### 2018-01-25 Manuscript1.Rmd
New files

---
title: "wholePlotTrajectories_manuscript"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

POST-DISTURBANCE DIVERSITY TRAJECTORIES IN A NEOTROPICAL FOREST

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2018-01-25 Manuscript1.Rmd
Save 1

@@
 
-## R Markdown
+# POST-DISTURBANCE DIVERSITY TRAJECTORIES IN A NEOTROPICAL FOREST
+\
+\
+\
 
-POST-DISTURBANCE DIVERSITY TRAJECTORIES IN A NEOTROPICAL FOREST
+## Summary
+\
+\
+\
+\
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+## Introduction
+\
+\
+\
+\
+
+## Material and Methods
+\
+\
+
+### Study site
+\
+\
+
+### Inventories protocol and dataset correction
+
+\
+\
+
+### Composition and diversity metrics
+
+\
+\
+
+##Results
+
+Over time, 104423 trees and 697 botanical species were recorded.
+The disturbance proved to significantly impact for more than 30 years the functional and taxonomic composition and diversity of the stands.
+
+### Taxonomic and functional composition
+
+*Composition trajectories* _ Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in a two-dimensional space based on NMDS analysis.
+Classifications were based on either abundance-based Bray-Curtis (Fig.
+1) or incidence-based Jaccard dissimilarity (data not shown).
+
+Both taxonomic and functional composition were substantially affected by disturbance and positively correlated with its intensity.
+*The distance from the initial condition was significantly time-dependent (P<0.01).* The taxonomic composition remained significantly different from the 1989 inventory until 30 years for all disturbed plots (Fig.
+2a).
+The distancing was positively correlated woth disturbance intensity.
+From 20 years after disturbance though the distancing reduced, suggesting the beginning of a return toward the reference 1989 inventory (Fig.
+1a & 2a).
+The functional composition distancing from 1989 was also positively correlated with disturbance intensity and similarly stabilized and reduced for 5 of T1, T2 and T3 plots from 20 years after disturbance (Fig.
+2a).
+*The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).*
+
+```{r pressure, echo=FALSE,fig.cap="Fig. 2"}
+treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
+names(treatments)<-c("Control","T1","T2","T3")
+load("FunctionalComposition_forGraph")
+ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
+
+par(mfrow=c(1,2))
+
+Matrep<-lapply(Matrep,function(Rep){
+  Rep<-as.data.frame(Rep)
+  Rep$plot<-substr(rownames(Rep),start=1,stop=regexpr("_",rownames(Rep))-1)
+  Rep$year<-substr(rownames(Rep),start=regexpr("_",rownames(Rep))+1,
+                   stop=nchar(rownames(Rep)))
+  return(Rep)
+})
+
+Dist<-lapply(1:12,function(pl){
+  ret<-lapply(Matrep,function(rep){return(rep[which(rep[,"plot"]==pl),])})
+  ret<-lapply(ret,function(rep){return(rep[which(rep[,"year"]%in%ret[[1]][,"year"]),])})
+  ret<-do.call(rbind,lapply(ret,function(rep){
+    ret2<-apply(rep[,c("NMDS1","NMDS2")],1,function(li){
+    return(sqrt(sum((rep[1,c("NMDS1","NMDS2")]-li)^2)))})
+    names(ret2)<-rep[,"year"]
+    return(ret2)}))
+  ret<-do.call(rbind,lapply(c(0.025,0.5,0.975),function(quant){
+    return(apply(ret,2,function(col){return(quantile(col,probs=quant))}))}))
+  rownames(ret)<-c(0.025,0.5,0.975)
+  return(ret)})
+names(Dist)<-1:12
+Dist<-lapply(Dist,function(pl){
+  colnames(pl)<-as.numeric(colnames(pl))-1984;return(pl)})
+
+plot(colnames(Dist[[1]]),Dist[[1]][1,],type="n",xlab="",ylab="",
+     ylim=c(min(unlist(Dist)),max(unlist(Dist))),cex.axis=0.7)
+invisible(lapply(1:length(treatments),function(tr){
+  toplot<-Dist[treatments[[tr]]]
+  invisible(lapply(toplot,function(pl){
+    lines(colnames(pl),pl["0.5",],col=ColorsTr[[tr]],lwd=2)
+    polygon(c(colnames(pl),rev(colnames(pl))),c(pl["0.025",],rev(pl["0.975",])),
+            col=rgb(0,0,0,alpha=0.1),border=NA)
+  }))}))
+mtext("Euclidean distance from 1989 inventory",side=2,padj=0,line=2,cex=0.8)
+mtext("(a) Taxonomic composition",side=3,adj=0,line=0.5)
+
+```
 
@@
 
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
 Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2018-01-26 Manuscript1.Rmd
Article update

--- 2018-01-25 Manuscript1.Rmd
@@
 ---
-title: "wholePlotTrajectories_manuscript"
+title: "POST-DISTURBANCE DIVERSITY TRAJECTORIES IN A NEOTROPICAL FOREST"
 output: html_document
+fig_caption: yes
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
+
+library(shape)
+
+treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
+names(treatments)<-c("Control","T1","T2","T3")
+load("FunctionalComposition_forGraph");load("TaxoComposition_forGraphs")
+load("ReplacementTraj_ForGraphs");load("FunctionalTraj_ForGraphs")
+ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
+Colors<-c("Greens","Blues","Oranges","Reds")
+
 ```
 
-## R Markdown
+# Summary
+\
+\
+\
+\
 
-POST-DISTURBANCE DIVERSITY TRAJECTORIES IN A NEOTROPICAL FOREST
+# Introduction
+\
+\
+\
+\
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+# Material and Methods
+\
+\
 
-```{r cars}
-summary(cars)
+## Study site
+\
+\
+
+## Inventories protocol and dataset correction
+
+\
+\
+
+## Composition and diversity metrics
+
+\
+\
+
+#Results
+
+Over time, 104423 trees and 697 botanical species were recorded.
+The disturbance proved to significantly impact for more than 30 years the functional and taxonomic composition and diversity of the stands.
+
+## Taxonomic and functional composition
+
+*Composition trajectories* _ Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in a two-dimensional space based on NMDS analysis.
+Classifications were based on either abundance-based Bray-Curtis \autoref{Fig1} or incidence-based Jaccard dissimilarity (data not shown).
+
+Both taxonomic and functional composition were substantially affected by disturbance and positively correlated with its intensity.
+*The distance from the initial condition was significantly time-dependent (P<0.01).* The taxonomic composition remained significantly different from the 1989 inventory until 30 years for all disturbed plots \autoref{Fig1}.
+The distancing was positively correlated woth disturbance intensity.
+From 20 years after disturbance though the distancing reduced, suggesting the beginning of a return toward the reference 1989 inventory \autoref{Fig1},\autoref{Fig2}.
+The functional composition distancing from 1989 was also positively correlated with disturbance intensity and similarly stabilized and reduced for 5 of T1, T2 and T3 plots from 20 years after disturbance \autoref{Fig2}.
+*The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).*
+
+```{r Fig1, echo=FALSE,fig.cap="**Fig.
+1 _ **Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
+Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories.
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3)."}
+par(mfrow=c(1,2))
+
+matT<-as.data.frame(MatrepTaxo[[1]])
+ matT$plot<-substr(rownames(matT),start=1,stop=regexpr("_",rownames(matT))-1)
+ matT$year<-substr(rownames(matT),start=regexpr("_",rownames(matT))+1,
+                   stop=nchar(rownames(matT)))
+  colnames(matT)<-c("X","Y","plot","year")
+  plot(matT$X,matT$Y,type="n",xlab="",ylab="",cex.axis=0.7)
+  invisible(lapply(1:length(treatments),function(tr){
+    plots<-treatments[[tr]]
+    invisible(lapply(plots,function(pl){
+      toplot<-matT[which(matT$plot==pl),]
+      points(toplot[1,"X"],toplot[1,"Y"],col=ColorsTr[[tr]],pch=16,cex=1.7)
+      points(toplot[nrow(toplot),"X"],toplot[nrow(toplot),"Y"],col=ColorsTr[[tr]],pch=1,cex=1.7,lwd=2)
+
+      lines(toplot$X,toplot$Y,col=ColorsTr[[tr]],lty=2)
+      invisible(lapply(round(seq(2,nrow(toplot),length.out=5)),function(step){
+        Arrows(x0=toplot[step,"X"],y0=toplot[step,"Y"],
+               x1=toplot[step-1,"X"],y1=toplot[step-1,"Y"],
+               col=ColorsTr[tr],code=1,lwd=2,arr.length=0.2,arr.type="simple")}))
+    }))
+  }))
+
+mtext("NMDS 1",side=1,line=2,cex=0.8)
+mtext("Euclidean distance from 1989 inventory",side=2,padj=0,line=3,cex=0.9)
+mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)
+mtext("(a) Taxonomic composition",side=3,adj=0,line=0.5)
+
+mat<-as.data.frame(MatrepFun[[3]])
+mat$plot<-substr(rownames(mat),start=1,stop=regexpr("_",rownames(mat))-1)
+mat$year<-substr(rownames(mat),start=regexpr("_",rownames(mat))+1,
+                   stop=nchar(rownames(mat)))
+colnames(mat)<-c("X","Y","plot","year")
+plot(mat$X,mat$Y,type="n",xlab="",ylab="",cex.axis=0.7)
+invisible(lapply(1:length(treatments),function(tr){
+  plots<-treatments[[tr]]
+    invisible(lapply(plots,function(pl){
+      toplot<-mat[which(mat$plot==pl),]
+      points(toplot[1,"X"],toplot[1,"Y"],col=ColorsTr[[tr]],pch=16,cex=1.7)
+      points(toplot[nrow(toplot),"X"],toplot[nrow(toplot),"Y"],col=ColorsTr[[tr]],pch=1,cex=1.7,lwd=2)
+
+  lines(toplot$X,toplot$Y,col=ColorsTr[[tr]],lty=2)
+  invisible(lapply(round(seq(2,nrow(toplot),length.out=5)),function(step){
+      Arrows(x0=toplot[step,"X"],y0=toplot[step,"Y"],
+            x1=toplot[step-1,"X"],y1=toplot[step-1,"Y"],
+            col=ColorsTr[tr],code=1,lwd=2,arr.length=0.2,arr.type="simple")}))
+  }))
+}))
+mtext("(b) Functional composition",side=3,adj=0,line=0.5)
+mtext("NMDS 1",side=1,line=2,cex=0.8)
+mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)
+
 ```
 
-## Including Plots
+```{r Fig2, echo=FALSE,fig.cap="**Fig.
+2 _ ** Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of **(a)** taxonomic composition at genus level and **(b)** functional composition.
+Distance are abundance-based Bary-Curtis metric.
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
+The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.
+"}
+par(mfrow=c(1,2))
 
-You can also embed plots, for example:
+MatrepT<-lapply(MatrepTaxo,function(Rep){
+  Rep<-as.data.frame(Rep)
+  Rep$plot<-substr(rownames(Rep),start=1,stop=regexpr("_",rownames(Rep))-1)
+  Rep$year<-substr(rownames(Rep),start=regexpr("_",rownames(Rep))+1,
+                   stop=nchar(rownames(Rep)))
+  return(Rep)
+})
 
-```{r pressure, echo=FALSE}
-plot(pressure)
+DistT<-lapply(1:12,function(pl){
+  ret<-lapply(MatrepT,function(rep){return(rep[which(rep[,"plot"]==pl),])})
+  ret<-lapply(ret,function(rep){return(rep[which(rep[,"year"]%in%
+  ret[[which(unlist(lapply(ret,nrow))==min(unlist(lapply(ret,nrow))))[1]]][,"year"]),])})
+  ret<-do.call(rbind,lapply(ret,function(rep){
+    ret2<-apply(rep[,c("NMDS1","NMDS2")],1,function(li){
+      return(sqrt(sum((rep[1,c("NMDS1","NMDS2")]-li)^2)))})
+    names(ret2)<-rep[,"year"]
+    return(ret2)}))
+  ret<-do.call(rbind,lapply(c(0.025,0.5,0.975),function(quant){
+    return(apply(ret,2,function(col){return(quantile(col,probs=quant))}))}))
+  rownames(ret)<-c(0.025,0.5,0.975)
+  return(ret)})
+names(DistT)<-1:12
+DistT<-lapply(DistT,function(pl){
+  colnames(pl)<-as.numeric(colnames(pl))-1984;return(pl)})
+
+plot(colnames(DistT[[1]]),DistT[[1]][1,],type="n",xlab="",ylab="",
+     ylim=c(min(unlist(DistT)),max(unlist(DistT))),cex.axis=0.7)
+invisible(lapply(1:length(treatments),function(tr){
+  toplot<-DistT[treatments[[tr]]]
+  invisible(lapply(toplot,function(pl){
+    lines(colnames(pl),pl["0.5",],col=ColorsTr[[tr]],lwd=2)
+    polygon(c(colnames(pl),rev(colnames(pl))),c(pl["0.025",],rev(pl["0.975",])),
+            col=rgb(0,0,0,alpha=0.1),border=NA)
+  }))}))
+mtext("Euclidean distance from 1989 inventory",side=2,padj=0,line=2,cex=0.8)
+mtext("(a) Taxonomic composition",side=3,adj=0,line=0.5)
+
+MatrepFun<-lapply(MatrepFun,function(Rep){
+  Rep<-as.data.frame(Rep)
+  Rep$plot<-substr(rownames(Rep),start=1,stop=regexpr("_",rownames(Rep))-1)
+  Rep$year<-substr(rownames(Rep),start=regexpr("_",rownames(Rep))+1,
+                   stop=nchar(rownames(Rep)))
+  return(Rep)
+})
+
+Dist<-lapply(1:12,function(pl){
+  ret<-lapply(MatrepFun,function(rep){return(rep[which(rep[,"plot"]==pl),])})
+  ret<-lapply(ret,function(rep){return(rep[which(rep[,"year"]%in%ret[[1]][,"year"]),])})
+  ret<-do.call(rbind,lapply(ret,function(rep){
+    ret2<-apply(rep[,c("NMDS1","NMDS2")],1,function(li){
+    return(sqrt(sum((rep[1,c("NMDS1","NMDS2")]-li)^2)))})
+    names(ret2)<-rep[,"year"]
+    return(ret2)}))
+  ret<-do.call(rbind,lapply(c(0.025,0.5,0.975),function(quant){
+    return(apply(ret,2,function(col){return(quantile(col,probs=quant))}))}))
+  rownames(ret)<-c(0.025,0.5,0.975)
+  return(ret)})
+names(Dist)<-1:12
+Dist<-lapply(Dist,function(pl){
+  colnames(pl)<-as.numeric(colnames(pl))-1984;return(pl)})
+
+plot(colnames(Dist[[1]]),Dist[[1]][1,],type="n",xlab="",ylab="",
+     ylim=c(min(unlist(Dist)),max(unlist(Dist))),cex.axis=0.7)
+invisible(lapply(1:length(treatments),function(tr){
+  toplot<-Dist[treatments[[tr]]]
+  invisible(lapply(toplot,function(pl){
+    lines(colnames(pl),pl["0.5",],col=ColorsTr[[tr]],lwd=2)
+    polygon(c(colnames(pl),rev(colnames(pl))),c(pl["0.025",],rev(pl["0.975",])),
+            col=rgb(0,0,0,alpha=0.1),border=NA)
+  }))}))
+mtext("(b) Functional composition",side=3,adj=0,line=0.5)
+mtext("Years since disturbance",side=1,adj=1,line=2,cex=0.8)
+
 ```
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+*Traits community weighted means* _ To complete the functional analysis the trajectories of the community weighted means (CWM) measured for the 6 functional traits at plot level \autoref{Fig3}.
+**To be coming**
+
+```{r Fig3, echo=FALSE,fig.cap="**Fig.
+3 ** _ Trajectories of the CWM until 30 years after disturbance of 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness).
+The CWM were measured at plot level and averaged among plots of the same treatment (green for control, blue for T1,orange for T2 and red for T3).
+The average CWM per treatment are drawn (solid line) drawn along with the 0.025 and 0.975 percentile (gray envelope) observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
+```
+
+## Disturbance impact on diversity
+
+*Taxonomic diversity* _ Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1994 inventories \autoref{Fig4}.
+*For the three indices the OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and a covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01).*
+
+```{r Fig4, echo=FALSE,fig.cap="**Fig.
+4 ** _ Trajectories of the difference to reference inventory, 10 years after disturbance, until 30 years after disturbance for plots **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversity.
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
+The average value of the 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness) are drawn along with the 0.025 and 0.975 percentile observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
+par(mfrow=c(1,3),oma=c(1,1,1,1),no.readonly=TRUE)
+for(q in 1:3){     
+  Toplot<-lapply(Complete,function(tr){return(tr[,,,q])})
+  Toplot<-lapply(Toplot,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
+  Toplot<-lapply(Toplot,function(tr){
+    ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col-tr[,1,rep]}))})#
+    ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),dim(tr)[3]),
+               dimnames=list(rownames(ret[[1]]),colnames(ret[[1]]),c(0.025,0.5,0.975)))
+    return(ret)})
+  Toplot<-lapply(Toplot,function(tr){return(tr[,which(!is.na(Toplot[[2]][1,,"0.5"])),])})
+
+  plot(as.numeric(colnames(Toplot[[1]]))-1984,Toplot[[1]][1,,1],type="n",xaxt="n",
+       xlab="",ylab="",ylim=c(min(unlist(Toplot),na.rm=T),max(unlist(Toplot),na.rm=T)))
+  axis(1,at=as.character(seq(5,33,5)),labels=TRUE)  
+  mtext(paste("(",c("a","b","c")[q],") ",c("Richness","Shannon","Simpson")[q],sep=""),
+        line=1,side=3,adj=0)
+
+  invisible(lapply(1:4,function(t){  
+    toplot<-Toplot[[t]]
+
+     invisible(lapply(1:3,function(i){
+      absc<-as.numeric(colnames(Toplot[[t]]))-1984
+      lines(absc,Toplot[[t]][i,,"0.5"], col = ColorsTr[[t]],lty = 1,lwd=2)
+      polygon(c(absc,rev(absc)),c(Toplot[[t]][i,,"0.025"],rev(Toplot[[t]][i,,"0.975"])),
+              col=rgb(0,0,0,alpha=0.1),border=NA)
+     }))
+  }))
+}
+mtext("Years since disturbance",side=1,adj=1,cex=0.8,line=-2,outer=TRUE)
+mtext("Equivalent diversity",side=2,padj=1,cex=0.8,line=-1,outer=TRUE)
+
+```
+
+The trajectories of the Richness, Shannon and Simpson differences to reference inventories in 1994 were rather even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase depending on the disturbance intensity \autoref{Fig4}.
+The richness stabilized for all disturbed plots from 20 years after disturbance \autoref{Fig4}.
+The Shannon and Simpson diversity also stabilized 20 years after disturbance for T1 and T2 plots \autoref{Fig4} whereas for 2 out of 3 plots T3 the Shannon and Simpson diversity continued increasing until 30 years after disturbance, illustrating a significant increase in communities’ evenness.
+
+*Functional  diversity* _ The trajectories of plot 7 from the T1 disturbance treatment was discarded because of too high taxonomic uncertainty preventing correct analysis.
+While the control plots showed an even trajectory until 30 years after disturbance, all disturbed plots displayed an increase of the functional diversity positively correlated to the disturbance intensity \autoref{Fig5}.
+The functional diversity observed 30 years after disturbance remained higher for all disturbed plots compared to control ones, without clear distinction following the disturbance intensity.
+The slope of the initial diversity increase and the maximum diversity were both positively correlated with the disturbance intensity.
+*The OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and the covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01)*.
+
+```{r Fig5, echo=FALSE,fig.cap="**Fig.
+5 ** _ Trajectories of the plots Rao functional diversity from10 to 30 years after disturbance.
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
+The value of the 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness) are drawn along with the 0.025 and 0.975 percentile observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
+plot(as.numeric(colnames(Toplot[[1]])),Toplot[[1]][1,,"0.5"],type="n",xaxt="n",
+       xlab="",ylab="",ylim=c(min(unlist(Toplot),na.rm=T),max(unlist(Toplot),na.rm=T)))
+axis(1,at=as.character(seq(5,33,5)),labels=TRUE)  
+mtext("Rao diversity",3,adj=0,line=1) 
+mtext("Years since disturbance",side=1,adj=1,cex=0.8,line=2)
+mtext("Equivalent diversity",side=2,padj=1,cex=0.8,line=3)
+
+ Toplot[[2]]<-Toplot[[2]][c(1,3),,]
+
+invisible(lapply(1:length(Toplot),function(t){
+
+    toplot<-Toplot[[t]]
+
+      invisible(lapply(1:nrow(toplot),function(i){
+      lines(colnames(toplot),toplot[i,,"0.5"], col = ColorsTr[[t]],lty = 1,lwd=2)
+      polygon(c(colnames(toplot),rev(colnames(toplot))),c(toplot[i,,"0.975"],rev(toplot[i,,"0.025"])),
+       col=rgb(0,0,0,alpha=0.1),border=NA)
+    }))
+}))
+```

### 2018-01-26 Manuscript1.Rmd
Article update

@@
 
-## Study site
+### Study site
 \
@@
 
-## Inventories protocol and dataset correction
+### Inventories protocol and dataset correction
 
@@
 
-## Composition and diversity metrics
+### Composition and diversity metrics
 
@@
 
-#Results
+# Results
 
@@
 
-## Taxonomic and functional composition
+### Taxonomic and functional composition
 
@@
 
-## Disturbance impact on diversity
+### Disturbance impact on diversity
 
@@
 The value of the 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness) are drawn along with the 0.025 and 0.975 percentile observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
-plot(as.numeric(colnames(Toplot[[1]])),Toplot[[1]][1,,"0.5"],type="n",xaxt="n",
-       xlab="",ylab="",ylim=c(min(unlist(Toplot),na.rm=T),max(unlist(Toplot),na.rm=T)))
-axis(1,at=as.character(seq(5,33,5)),labels=TRUE)  
+plot(as.numeric(colnames(Toplot_FunTraj[[1]])),Toplot_FunTraj[[1]][1,,"0.5"],type="n",xaxt="n",
+       xlab="",ylab="",ylim=c(min(unlist(Toplot_FunTraj),na.rm=T),max(unlist(Toplot_FunTraj),na.rm=T)),cex.axis=0.7)
+axis(1,at=as.character(seq(5,33,5)),labels=TRUE,cex.axis=0.7)  
 mtext("Rao diversity",3,adj=0,line=1) 
@@
 
- Toplot[[2]]<-Toplot[[2]][c(1,3),,]
+invisible(lapply(1:length(Toplot_FunTraj),function(t){
 
-invisible(lapply(1:length(Toplot),function(t){
-
-    toplot<-Toplot[[t]]
+    toplot<-Toplot_FunTraj[[t]]
 
@@
 ```
+
+# Discussion
+-	The disturbance significantly impact the functional and taxonomic composition and diversity of the stands for more than 30 years 
+-	Important differences among plots of same treatments: possible impact of initial diversity composition, local abiotic parameters (topography, soils, surrounding stands).
+Impacts depend on the disturbance intensity, but further parameters may be involved.
+-	The shift in functional composition differs from the null model: biotic and abiotic filters are at stake but only tangible from 10/15 years after logging.
+Discussion with Cannon, 1998: they found no different process than rarefaction for 8 years after logging, but the time might have been too short-> divergence due to recruitment and not rarefaction after random species disappearance.
+Same conclusions otherwise.
+-	Disturbance involves new filters for species settling in the stands, probably according to their functional strategy.
+Disturbance generates a functional shift of communities toward acquisitive strategies.
+Ecosystems function rely on abundant species: changes in stands functioning is then expected.
+(Flores, 2006  for differential sapling success after disturbance).
+-	If the recovery is not achieved 30 years after disturbance, it is on the way to.
+Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories.
+-	To go further we should focus on recruitment which would also highlight the mechanisms underlying these shifts and dynamics and explicit the environmental filter.
+
+# Appendix I _ Scheme of the taxonomic uncertainty handle
+
+![**Fig.
+A1 _ ** Whenever the family or genus were recorded for a vernacular name, the pool of species sampled to replace the vernacular name was restricted to those pertaining to the same higher taxonomic level (i.e species of the same genus when the species is missing, species of the same family when the genus is missing).](BayesianSheme.png)

### 2018-01-26 Manuscript1.Rmd
Premiere version

@@
 ---
-title: "POST-DISTURBANCE DIVERSITY TRAJECTORIES IN A NEOTROPICAL FOREST"
-output: html_document
-fig_caption: yes
+title: "Post-Disturbance Diversity Trajectories in a Neotropical Forest"
+author:
+  - name: Prénom Nom
+    authsuperscript: 1*
+  - name: Deuxième Auteur
+    authsuperscript: 2
+affiliation:
+  - affsuperscript: 1
+    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
+    address: >
+      Campus Agronomique,
+      97310 Kourou,
+      France.
+  - affsuperscript: 2
+    dptuniv: Department of Ecology, University of Edimburgh
+    address: >
+      Street address,
+      Zip code,
+      Country.
+corrauthor:
+    email: prenom.nom@ecofog.gf
+    url: http://www.ecofog.gf/spip.php?article47
+abstract: >   #Pour avoir plusieurs lignes qui suivent. Commencer à 2 caractères
+  Résumé de l'article.
+keywords: [mot-clés, séparés par des virgules]
+journalinfo: Hal 00679993
+archive: DOI xxxx
+date: "`r format(Sys.time(), '%d %B %Y')`"
+preamble: >  # Pour coordonne les packages LateX (supplémentaires)
+  \usepackage{amsmath}
+bibliography: references.bib # .bib à activer seulement pour git et word format (pandoc fait la biblio)
+csl: 
+lang: english # french
+keywordlabel: Keywords
+corrauthorlabel: Corresponding author
+output:
+  bookdown::gitbook
+#   bookdown::pdf_book:
+#     base_format: EcoFoG::article
+#     keep_tex: TRUE
+#  bookdown::word_document2
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+<!-- Options de knitr et de R (librairies, ...) -->
 
-library(shape)
+```{r Options, echo=FALSE, message=FALSE}
+knitr::opts_chunk$set(cache=TRUE, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
+# Installation des packages si nécessaire et chargement
+Library <- function(Packages) {
+  InstallAndLoad <- function(Package) {
+    if (!Package %in% installed.packages()[, 1]) {install.packages(Package)}
+    require(Package, character.only = TRUE)
+  }
+  invisible(sapply(Packages, InstallAndLoad))
+}
+# Ajouter les packages nécessaires ici
+Library(c("shape"))
 
@@
 ```
-
-# Summary
-\
-\
-\
-\
 
@@
 
-```{r Fig1, echo=FALSE,fig.cap="**Fig.
-1 _ **Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
+```{r Fig1, echo=FALSE,fig.cap="Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
 Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories.
@@
 
-```{r Fig2, echo=FALSE,fig.cap="**Fig.
-2 _ ** Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of **(a)** taxonomic composition at genus level and **(b)** functional composition.
+```{r Fig2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of **(a)** taxonomic composition at genus level and **(b)** functional composition.
 Distance are abundance-based Bary-Curtis metric.
@@
 
-```{r Fig3, echo=FALSE,fig.cap="**Fig.
-3 ** _ Trajectories of the CWM until 30 years after disturbance of 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness).
+```{r Fig3, echo=FALSE,fig.cap="Trajectories of the CWM until 30 years after disturbance of 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness).
 The CWM were measured at plot level and averaged among plots of the same treatment (green for control, blue for T1,orange for T2 and red for T3).
@@
 
-```{r Fig4, echo=FALSE,fig.cap="**Fig.
-4 ** _ Trajectories of the difference to reference inventory, 10 years after disturbance, until 30 years after disturbance for plots **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversity.
+```{r Fig4, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory, 10 years after disturbance, until 30 years after disturbance for plots **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversity.
 Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
@@
 The trajectories of the Richness, Shannon and Simpson differences to reference inventories in 1994 were rather even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase depending on the disturbance intensity \autoref{Fig4}.
+
 The richness stabilized for all disturbed plots from 20 years after disturbance \autoref{Fig4}.

### 2018-01-26 Manuscript1.Rmd
ajout docs

@@
 #  bookdown::word_document2
+# output_dir pour articles?
 ---

### 2018-01-26 Manuscript1.Rmd
manuscript update

@@
 author:
-  - name: Prénom Nom
+  - name: Ariane MIRABEL
     authsuperscript: 1*
-  - name: Deuxième Auteur
+  - name: Eric Marcon
+    authsuperscript: 1
+  - name: Bruno Hérault
     authsuperscript: 2
@@
 corrauthor:
-    email: prenom.nom@ecofog.gf
+    email: ariane.mirabel@ecofog.gf
     url: http://www.ecofog.gf/spip.php?article47
@@
 # output_dir pour articles?
+#organiser le code, avec les script et DB adéquats. Mettre un code dans les figures pour indiquer le code d'où elle provient. ../ pour référer à des sous dossiers
 ---
@@
 # Introduction
-\
-\
-\
-\
+The large areas covered with tropical forests worldwide hold crucial commercial, social and cultural value. 
+They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate local climate and support nutrients, water and carbon cycles and ensure cultural and human well-being. 
+The demand for forests products and the shrink of unmanaged areas increased the human pressure on forests and currently threaten the maintenance of their structure, composition and function. 
+The increasing demand for forests products, land-use changes and other human disturbances along with substantial climatic changes make a crucial issue of understanding the response of forests dynamics to such disturbance regimes [@Gibson2011a; @Morales-Hidalgo2015].
+This has been addressed from the point of view of biomass production and carbon storage which resilience proved subjected to identified cyclic dynamics and therefore gives hope for the maintenance of forest ecosystem and their services.
+The response of trees biological diversity however has been sparsely studied, although this diversity is a major determinant of forests productivity, stability and functioning as it ensures interspecific complementarity, enhances the use of resources and the store of nutrients, and mitigates the prejudices of diseases and herbivory [@Tilman2014].
+Maintain the integrity of forests functioning thus requires highlighting the impacts and underlying mechanisms of disturbance on trees diversity and question the extent of their resilience. 
+Disturbance alter the biotic and abiotic parameters of the ecosystem in modifying the light availability, the heat and water fluxes, the plants productivity, etc. 
+Although those changes proved impacting the proportion and identity of species, negligible or even positive impacts were highlighted when considering the community as a whole, in accordance with the intermediate disturbance hypothesis (IDH).
+The IDH that predicts a maximized species diversity at intermediate level of disturbance, however, mainly rely on the study of species richness which gives limited or even misleading information on forests recovery and functioning.
+To have a detailed understanding of forests response to disturbance it is essential to identify the subsequent impact on species identity per se, which underlies any diversity change, as well as on richness, equitability and functional profile of the whole community.
+Such combination of composition and diversity in browsing all facets of stands assemblage would highlight their resilience to perturbation, and its determinant and underlying ecological processes.
+Disturbance survey would first assess species turnover and communities convergence through stands taxonomic composition and unveil environmental filters through functional composition that is based on the biological attributes of plants.
+The set of functional traits chosen for this functional analysis considers functional traits representative of species functional strategy, fitness, and ecology as it is the case for traits of the leaf or stem economics.
+Composition analysis would be continued with diversity analysis assessing for the whole community the richness and evenness of the taxonomic and functional abundance distribution.   
+Here we examined the long-term trajectory, in the sense of the evolution along time, of tropical forests taxonomic and functional composition and diversity. 
+We studied the regular inventories over 30 years of twelve Amazonian forest plots, either untouched or disturbed with three disturbance intensity . 
+The inventories were combined with one of the largest functional trait database for tropical trees.
+We aimed to (i) explicit the impact of disturbance on the long-term trajectory of tropical forest communities composition and diversity, (ii) question the long-term convergence of the composition of communities and (iii) determine their resilience or not, after disturbance.
 
 # Material and Methods
-\
-\
 
 ### Study site 
-\
-\
+Paracou station in French Guiana (5°18'N and 52°53'W) encompasses a lowland tropical rain forest, corresponds to a tropical wet climate with mean annual precipitation averaging 2980 mm.y-1 (30-y period) with a 3-months dry season (< 100 mm.months-1) from mid-August to mid-November, and a one-month dry season in (wagner AFM 2011).
+Elevation ranges between 5 and 50 m and mean annual temperature is 26°C.
+Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
+The experiment corresponds to a network of twelve 6.25ha plots where three logging, thinning and fuelwood treatments on an intensity gradient.
+Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots.
+The logging correspond to averages of 10 trees remove per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3).
+Treatments T2 and T3 included besides logging of 10trees/ha the thinning by poison girdling.
 
 ### Inventories protocol and dataset correction
+The study site corresponds to a tropical rain forest with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae. 
+In the twelve experimental plots all trees above 10 cm DBH are mapped and measured annually since 1984. 
+During the inventories trees are first identified with a vernacular name, assigned by the field team, and then with a scientific name, assigned by a botanist during regular botanical campaigns. 
+In 1984, specific vernacular names were given to 62 commercial or common species whereas two identifiers separating palm trees and trees were given to all other species. 
+Botanical campaigns to identify all trees at the species level started in 2003 and continued every 5 to 6 years. 
+In total, from 1984 to 2016, 848 species were identified in the whole dispositive. 
+The variability of botanical identification practices along time and among plots raised some methodological issues. 
+The first vernacular name attributed to trees, when not associated to a trustworthy botanical name thereafter, generate some taxonomic uncertainty. 
+To account for this taxonomic uncertainty in the composition and diversity metrics we used a Bayesian framework estimating the expectancy and variance of composition and diversity metrics. 
+The frameworks is based on a set of complete inventories extrapolated from the initial one through the replacement of vernacular name by botanical species.
+These replacement were made through a multinomial distribution M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name v and the species recorded in the inventory [*s~1~, s~2~, …, s~N~*].
+See appendix 1 and @Aubry-Kientz2013 Plos one  for the detailed methodology.
 
-\
-\
+The functional metrics used a set of 6 functional traits values representing leaf and wood economics.
+The trait database encompassed leaves thickness, toughness, total chlorophyll content, dry mass and specific leaf area (leaf area per unit dry mass; *SLA*),  and wood density and bark thickness. 
+Species trait values were assessed in sampling at each iteration one of the individuals measured for traits collection as part of the BRIDGE project (http://www.ecofog.gf/Bridge/index.htm), which allowed to account for intraspecific variability.
+This project was carried out in nine permanent plots in French Guiana, including two in Paracou. 
+Individuals missing trait values were filled using multivariate imputation by chained equation (mice), restricted to the samples pertaining to the next higher taxonomic level to account for the phylogenetic signal of the traits.
+The dataset comprised 257 botanical species. Whenever an inventoried species was not in the dataset a set of traits values was sampled among species of the same next higher taxonomic level. 
+All composition and diversity metrics correspond to the expectancy obtained after 50 iterations of the taxonomic uncertainty propagation framework and the missing trait values filling process. 
+Because of taxonomic uncertainty issues the taxonomic composition and diversity analysis were conducted at the genus level, knowing that diversity patterns remain consistent in this case.
 
@@
 
-\
-\
+The analysis of taxonomic and functional composition of plots assemblages was visualized in a two-dimensional ordination space based on non-metric dimensional scaling of their floristic or functional composition.
+Their evolution along time relative to initial condition was reported using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) similarity measures. 
+To determine whether plots composition converged to their initial condition we computed the distance to initial condition of the sampled years in the two-dimensional space.
+Additionally, the functional composition was analyzed through the trajectory along time of the community weighted means (CWM) for the 6 functional traits chosen.
+To determine whether communities’ assembly deviated from the expectation of a random process we compared the observed similarity patterns to those generated by 50 repetitions of stochastic null models.
+The null model for taxonomic composition randomly shuffled individuals among plots while preserving the overall species abundance and plots’ tree density. 
+The null model for functional composition produced random traits assemblages by permuting the set of functional traits values among species.
+
+Plots taxonomic diversity corresponded to species Richness and the Hill number translation of Shannon and Simpson indices.
+These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q).
+This order q grasps the balance between richness and evenness in the community as it determines the emphasis on common species in the diversity metric, with common species weighting more than rare ones when q increases.
+The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on the 6 functional traits specific values.
 

### 2018-01-27 Manuscript1.Rmd
Manuscript update

@@
 # Ajouter les packages nécessaires ici
-Library(c("shape"))
-
-treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
-names(treatments)<-c("Control","T1","T2","T3")
-load("FunctionalComposition_forGraph");load("TaxoComposition_forGraphs")
-load("ReplacementTraj_ForGraphs");load("FunctionalTraj_ForGraphs")
-ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
-Colors<-c("Greens","Blues","Oranges","Reds")
+source("DrawingTools.R")
 
@@
 Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3)."}
-par(mfrow=c(1,2))
+load("DB/FunctionalComposition_forGraph");
+load("DB/TaxoComposition_forGraphs")
 
-matT<-as.data.frame(MatrepTaxo[[1]])
- matT$plot<-substr(rownames(matT),start=1,stop=regexpr("_",rownames(matT))-1)
- matT$year<-substr(rownames(matT),start=regexpr("_",rownames(matT))+1,
-                   stop=nchar(rownames(matT)))
-  colnames(matT)<-c("X","Y","plot","year")
-  plot(matT$X,matT$Y,type="n",xlab="",ylab="",cex.axis=0.7)
-  invisible(lapply(1:length(treatments),function(tr){
-    plots<-treatments[[tr]]
-    invisible(lapply(plots,function(pl){
-      toplot<-matT[which(matT$plot==pl),]
-      points(toplot[1,"X"],toplot[1,"Y"],col=ColorsTr[[tr]],pch=16,cex=1.7)
-      points(toplot[nrow(toplot),"X"],toplot[nrow(toplot),"Y"],col=ColorsTr[[tr]],pch=1,cex=1.7,lwd=2)
-
-      lines(toplot$X,toplot$Y,col=ColorsTr[[tr]],lty=2)
-      invisible(lapply(round(seq(2,nrow(toplot),length.out=5)),function(step){
-        Arrows(x0=toplot[step,"X"],y0=toplot[step,"Y"],
-               x1=toplot[step-1,"X"],y1=toplot[step-1,"Y"],
-               col=ColorsTr[tr],code=1,lwd=2,arr.length=0.2,arr.type="simple")}))
-    }))
-  }))
-
-mtext("NMDS 1",side=1,line=2,cex=0.8)
-mtext("Euclidean distance from 1989 inventory",side=2,padj=0,line=3,cex=0.9)
-mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)
-mtext("(a) Taxonomic composition",side=3,adj=0,line=0.5)
-
-mat<-as.data.frame(MatrepFun[[3]])
-mat$plot<-substr(rownames(mat),start=1,stop=regexpr("_",rownames(mat))-1)
-mat$year<-substr(rownames(mat),start=regexpr("_",rownames(mat))+1,
-                   stop=nchar(rownames(mat)))
-colnames(mat)<-c("X","Y","plot","year")
-plot(mat$X,mat$Y,type="n",xlab="",ylab="",cex.axis=0.7)
-invisible(lapply(1:length(treatments),function(tr){
-  plots<-treatments[[tr]]
-    invisible(lapply(plots,function(pl){
-      toplot<-mat[which(mat$plot==pl),]
-      points(toplot[1,"X"],toplot[1,"Y"],col=ColorsTr[[tr]],pch=16,cex=1.7)
-      points(toplot[nrow(toplot),"X"],toplot[nrow(toplot),"Y"],col=ColorsTr[[tr]],pch=1,cex=1.7,lwd=2)
-
-  lines(toplot$X,toplot$Y,col=ColorsTr[[tr]],lty=2)
-  invisible(lapply(round(seq(2,nrow(toplot),length.out=5)),function(step){
-      Arrows(x0=toplot[step,"X"],y0=toplot[step,"Y"],
-            x1=toplot[step-1,"X"],y1=toplot[step-1,"Y"],
-            col=ColorsTr[tr],code=1,lwd=2,arr.length=0.2,arr.type="simple")}))
-  }))
-}))
-mtext("(b) Functional composition",side=3,adj=0,line=0.5)
-mtext("NMDS 1",side=1,line=2,cex=0.8)
-mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)
+TaxoCompo(MatrepTaxo)
+FunCompo(MatrepFun)
 
@@
 "}
-par(mfrow=c(1,2))
 
-MatrepT<-lapply(MatrepTaxo,function(Rep){
-  Rep<-as.data.frame(Rep)
-  Rep$plot<-substr(rownames(Rep),start=1,stop=regexpr("_",rownames(Rep))-1)
-  Rep$year<-substr(rownames(Rep),start=regexpr("_",rownames(Rep))+1,
-                   stop=nchar(rownames(Rep)))
-  return(Rep)
-})
-
-DistT<-lapply(1:12,function(pl){
-  ret<-lapply(MatrepT,function(rep){return(rep[which(rep[,"plot"]==pl),])})
-  ret<-lapply(ret,function(rep){return(rep[which(rep[,"year"]%in%
-  ret[[which(unlist(lapply(ret,nrow))==min(unlist(lapply(ret,nrow))))[1]]][,"year"]),])})
-  ret<-do.call(rbind,lapply(ret,function(rep){
-    ret2<-apply(rep[,c("NMDS1","NMDS2")],1,function(li){
-      return(sqrt(sum((rep[1,c("NMDS1","NMDS2")]-li)^2)))})
-    names(ret2)<-rep[,"year"]
-    return(ret2)}))
-  ret<-do.call(rbind,lapply(c(0.025,0.5,0.975),function(quant){
-    return(apply(ret,2,function(col){return(quantile(col,probs=quant))}))}))
-  rownames(ret)<-c(0.025,0.5,0.975)
-  return(ret)})
-names(DistT)<-1:12
-DistT<-lapply(DistT,function(pl){
-  colnames(pl)<-as.numeric(colnames(pl))-1984;return(pl)})
-
-plot(colnames(DistT[[1]]),DistT[[1]][1,],type="n",xlab="",ylab="",
-     ylim=c(min(unlist(DistT)),max(unlist(DistT))),cex.axis=0.7)
-invisible(lapply(1:length(treatments),function(tr){
-  toplot<-DistT[treatments[[tr]]]
-  invisible(lapply(toplot,function(pl){
-    lines(colnames(pl),pl["0.5",],col=ColorsTr[[tr]],lwd=2)
-    polygon(c(colnames(pl),rev(colnames(pl))),c(pl["0.025",],rev(pl["0.975",])),
-            col=rgb(0,0,0,alpha=0.1),border=NA)
-  }))}))
-mtext("Euclidean distance from 1989 inventory",side=2,padj=0,line=2,cex=0.8)
-mtext("(a) Taxonomic composition",side=3,adj=0,line=0.5)
-
-MatrepFun<-lapply(MatrepFun,function(Rep){
-  Rep<-as.data.frame(Rep)
-  Rep$plot<-substr(rownames(Rep),start=1,stop=regexpr("_",rownames(Rep))-1)
-  Rep$year<-substr(rownames(Rep),start=regexpr("_",rownames(Rep))+1,
-                   stop=nchar(rownames(Rep)))
-  return(Rep)
-})
-
-Dist<-lapply(1:12,function(pl){
-  ret<-lapply(MatrepFun,function(rep){return(rep[which(rep[,"plot"]==pl),])})
-  ret<-lapply(ret,function(rep){return(rep[which(rep[,"year"]%in%ret[[1]][,"year"]),])})
-  ret<-do.call(rbind,lapply(ret,function(rep){
-    ret2<-apply(rep[,c("NMDS1","NMDS2")],1,function(li){
-    return(sqrt(sum((rep[1,c("NMDS1","NMDS2")]-li)^2)))})
-    names(ret2)<-rep[,"year"]
-    return(ret2)}))
-  ret<-do.call(rbind,lapply(c(0.025,0.5,0.975),function(quant){
-    return(apply(ret,2,function(col){return(quantile(col,probs=quant))}))}))
-  rownames(ret)<-c(0.025,0.5,0.975)
-  return(ret)})
-names(Dist)<-1:12
-Dist<-lapply(Dist,function(pl){
-  colnames(pl)<-as.numeric(colnames(pl))-1984;return(pl)})
-
-plot(colnames(Dist[[1]]),Dist[[1]][1,],type="n",xlab="",ylab="",
-     ylim=c(min(unlist(Dist)),max(unlist(Dist))),cex.axis=0.7)
-invisible(lapply(1:length(treatments),function(tr){
-  toplot<-Dist[treatments[[tr]]]
-  invisible(lapply(toplot,function(pl){
-    lines(colnames(pl),pl["0.5",],col=ColorsTr[[tr]],lwd=2)
-    polygon(c(colnames(pl),rev(colnames(pl))),c(pl["0.025",],rev(pl["0.975",])),
-            col=rgb(0,0,0,alpha=0.1),border=NA)
-  }))}))
-mtext("(b) Functional composition",side=3,adj=0,line=0.5)
-mtext("Years since disturbance",side=1,adj=1,line=2,cex=0.8)
+TaxoDist(MatrepTaxo)
+FunDist(MatrepFun)
 
@@
 The average value of the 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness) are drawn along with the 0.025 and 0.975 percentile observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
-par(mfrow=c(1,3),oma=c(1,1,1,1),no.readonly=TRUE)
-for(q in 1:3){     
-  Toplot<-lapply(Complete,function(tr){return(tr[,,,q])})
-  Toplot<-lapply(Toplot,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
-  Toplot<-lapply(Toplot,function(tr){
-    ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col-tr[,1,rep]}))})#
-    ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),dim(tr)[3]),
-               dimnames=list(rownames(ret[[1]]),colnames(ret[[1]]),c(0.025,0.5,0.975)))
-    return(ret)})
-  Toplot<-lapply(Toplot,function(tr){return(tr[,which(!is.na(Toplot[[2]][1,,"0.5"])),])})
+load("DB/ReplacementTraj_ForGraphs")
 
-  plot(as.numeric(colnames(Toplot[[1]]))-1984,Toplot[[1]][1,,1],type="n",xaxt="n",
-       xlab="",ylab="",ylim=c(min(unlist(Toplot),na.rm=T),max(unlist(Toplot),na.rm=T)))
-  axis(1,at=as.character(seq(5,33,5)),labels=TRUE)  
-  mtext(paste("(",c("a","b","c")[q],") ",c("Richness","Shannon","Simpson")[q],sep=""),
-        line=1,side=3,adj=0)
-
-  invisible(lapply(1:4,function(t){  
-    toplot<-Toplot[[t]]
-
-     invisible(lapply(1:3,function(i){
-      absc<-as.numeric(colnames(Toplot[[t]]))-1984
-      lines(absc,Toplot[[t]][i,,"0.5"], col = ColorsTr[[t]],lty = 1,lwd=2)
-      polygon(c(absc,rev(absc)),c(Toplot[[t]][i,,"0.025"],rev(Toplot[[t]][i,,"0.975"])),
-              col=rgb(0,0,0,alpha=0.1),border=NA)
-     }))
-  }))
-}
-mtext("Years since disturbance",side=1,adj=1,cex=0.8,line=-2,outer=TRUE)
-mtext("Equivalent diversity",side=2,padj=1,cex=0.8,line=-1,outer=TRUE)
+TaxoTraj(CompleteTaxo)
 
@@
 
-```{r Fig5, echo=FALSE,fig.cap="**Fig.
-5 ** _ Trajectories of the plots Rao functional diversity from10 to 30 years after disturbance.
+```{r Fig5, echo=FALSE,fig.cap="Trajectories of the plots Rao functional diversity from10 to 30 years after disturbance.
 Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
 The value of the 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness) are drawn along with the 0.025 and 0.975 percentile observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
-plot(as.numeric(colnames(Toplot_FunTraj[[1]])),Toplot_FunTraj[[1]][1,,"0.5"],type="n",xaxt="n",
-       xlab="",ylab="",ylim=c(min(unlist(Toplot_FunTraj),na.rm=T),max(unlist(Toplot_FunTraj),na.rm=T)),cex.axis=0.7)
-axis(1,at=as.character(seq(5,33,5)),labels=TRUE,cex.axis=0.7)  
-mtext("Rao diversity",3,adj=0,line=1) 
-mtext("Years since disturbance",side=1,adj=1,cex=0.8,line=2)
-mtext("Equivalent diversity",side=2,padj=1,cex=0.8,line=3)
+load("DB/FunctionalTraj_ForGraphs")
 
-invisible(lapply(1:length(Toplot_FunTraj),function(t){
-
-    toplot<-Toplot_FunTraj[[t]]
-
-      invisible(lapply(1:nrow(toplot),function(i){
-      lines(colnames(toplot),toplot[i,,"0.5"], col = ColorsTr[[t]],lty = 1,lwd=2)
-      polygon(c(colnames(toplot),rev(colnames(toplot))),c(toplot[i,,"0.975"],rev(toplot[i,,"0.025"])),
-       col=rgb(0,0,0,alpha=0.1),border=NA)
-    }))
-}))
+FunTraj(CompleteFun)
 ```

### 2018-01-27 DrawingTools.R

@@ -1,0 +1,185 @@
+### Manuscript drawing functions
+Library(c("shape"))
+
+treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
+names(treatments)<-c("Control","T1","T2","T3")
+ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
+
+TaxoCompo<-function(Data_TaxoComp){
+par(mfrow=c(1,2))
+matT<-as.data.frame(Data_TaxoComp[[1]])
+matT$plot<-substr(rownames(matT),start=1,stop=regexpr("_",rownames(matT))-1)
+matT$year<-substr(rownames(matT),start=regexpr("_",rownames(matT))+1,
+                  stop=nchar(rownames(matT)))
+colnames(matT)<-c("X","Y","plot","year")
+plot(matT$X,matT$Y,type="n",xlab="",ylab="",cex.axis=0.7)
+invisible(lapply(1:length(treatments),function(tr){
+  plots<-treatments[[tr]]
+  invisible(lapply(plots,function(pl){
+    toplot<-matT[which(matT$plot==pl),]
+    points(toplot[1,"X"],toplot[1,"Y"],col=ColorsTr[[tr]],pch=16,cex=1.7)
+    points(toplot[nrow(toplot),"X"],toplot[nrow(toplot),"Y"],col=ColorsTr[[tr]],pch=1,cex=1.7,lwd=2)
+    
+    lines(toplot$X,toplot$Y,col=ColorsTr[[tr]],lty=2)
+    invisible(lapply(round(seq(2,nrow(toplot),length.out=5)),function(step){
+      Arrows(x0=toplot[step,"X"],y0=toplot[step,"Y"],
+             x1=toplot[step-1,"X"],y1=toplot[step-1,"Y"],
+             col=ColorsTr[tr],code=1,lwd=2,arr.length=0.2,arr.type="simple")}))
+  }))
+}))
+
+mtext("NMDS 1",side=1,line=2,cex=0.8)
+mtext("Euclidean distance from 1989 inventory",side=2,padj=0,line=3,cex=0.9)
+mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)
+mtext("(a) Taxonomic composition",side=3,adj=0,line=0.5)
+}
+
+FunCompo<-function(Data_FunComp){mat<-as.data.frame(Data_FunComp[[3]])
[…]

### 2018-01-27 Scripts&DB_description.Rmd
Script & DB loading

---
title: "Notice Scripts"
author: "Ariane"
date: "27 janvier 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Files in the *Scripts* file are annotated .R files corresponding to the production of database and subsequent plots of the whole plot article.
Files in the *DB* are the initial inventories and produced database.

## DB file

- *Alpha_Plots*: Alpha matrices for the Dirichlet sampling of the vernacular names botanical trials. 
It is a list of plot-specific matrices, following previous investigation suggesting that botanical associations should rely on local community inventories and history to minimize biases.

- *Paracou_R_Subdivided_ok*: Formatted Paracou inventories, it contains various matrices and lists that are not all used here. 
*LivingStand_all* is the lists used here, inventories are subdivided by inventoried year and by plot.

- *BotanyGenus*: Botanical correspondence tables to associate genuses to botanical names trialed from vernacular names.

- *BridgeOK.csv*: csv file of the Bridge dataset: 6 leaves and 2 stem traits.

- *DataLifeTraits.csv*: csv file for the Hmax and seed mass class.

- *TaxoComposition_ForGraphs*: Coordinates of inventories and plots in the 2 dimensions space from taxonomic composition NMDS. 
Lists the coordinates along the various iteration: for now on, only 5 repetitions (becaus run locally).

- *FunctionalComposition_forGraphs*: Coordinates of inventories and plots in the 2 dimensions space from functional composition NMDS. 
Lists the coordinates along the 50 iteration. Comes from the script: *FunctionalNMDS.R*

-*ReplacementTraj_ForGraphs*: List by treatment of plots Richness, Shannon and Simpson diversity along inventoried year. 
It gives the 0.025, 0.5 and 0.975 quantiles from 50 repetitions.
Comes from the script: *ReplacementTraj_ForGraphs.R*

-*FunctionalTraj_ForGraphs*: List by treatment of plots Rao diversity along inventoried year. 
It gives the 0.025, 0.5 and 0.975 quantiles from *only 5 for now* repetitions.

## Scripts file

- *Vernacular_handle.R*: General functions, to propagate taxonomic uncertainty.

- *TraitsMiceFilling.R*: General functions, gap-filling of the traits database.

- *Functional NMDS.R*: Functional composition analysis: iterative NMDS on plots functional inventories.

- *DivTaxo_trajectories.R*: Taxonomic Richness, Shannon and Simpson diversity trajectories per plot along inventoried years.

### 2018-01-27 Scripts/DivTaxo_trajectories.R

@@ -1,0 +1,76 @@
+load("DB/Paracou_R_Subdivided_ok")
+source("Vernacular_handle.R")
+
+# Treatments definition
+T0<-c(1,6,11);T1<-c(2,7,9);T2<-c(3,5,10);T3<-c(4,8,12)
+treatments<-list(T0,T1,T2,T3);names(treatments)<-c("T0","T1","T2","T3")
+
+# Years with identified problems of inventories
+Pby<-c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014")
+dates<-sort(names(LivingStand_all)[which(!names(LivingStand_all)%in%Pby)])
+
+# Alpha matrix for vernacular names association
+alphas_plot<-AlphaPlots(LivingStand_all)
+
+# Reference for botany (Genus/species association)
+load("DB/BotanyGenus")
+
+#trajectories drawn by treatment
+CompleteTaxo<-lapply(1:4,function(t){
+  
+  treat<-treatments[[t]]
+  Nrep<-2   # Number of iteration: replacement of vernacular names
+  print(t)
+  
+  Trajectories<-lapply(0:2,function(Q){
+    
+          Repet<-mclapply(1:Nrep,function(r){
+            
+    Plot_trajectory<-do.call(cbind,lapply(dates,function(y){
+      
+      # Select current year and current plots
+      year<-LivingStand_all[[which(names(LivingStand_all)==y)]] 
+      yearPlot<-year[which(names(year)%in%treat)] 
+      
+      if(length(yearPlot)!=0){
+        
+        Div<-lapply(1:length(yearPlot),function(pl){
+          
+        # Replace vernacular names and get corresponding genus
+        Div_plot<-as.data.frame(Replacement(yearPlot[[pl]],Alpha=alphas_plot[[names(yearPlot)[pl]]]))
+        colnames(Div_plot)<-"names"
+        Div_plot<-as.character(merge(Div_plot,RefBota,by.x="names",by.y="row.names",all.x=T)[,"Genre"])
+         
+        return(expq(bcTsallis(as.AbdVector(tapply(Div_plot,Div_plot,length)),q=Q,Correction = "Best"),q=Q))
+       })
+        
+        names(Div)<-names(yearPlot)
+        Div<-do.call(rbind,Div)
+        Div<-merge(Div,data.frame(treat),by.x="row.names",by.y="treat",all=T)
+        
[…]

### 2018-01-27 Scripts/FunctionalNMDS.R

@@ -1,0 +1,104 @@
+source("Scripts/Vernacular_handle.R")           # Script for vernacular names replacement
+source("Scripts/TraitsMiceFilling.R")                                   # Traits Gap-filling
+
+load("DB/Alpha_Plots")                  # Alpha vector for vernacular replacement
+load("DB/Paracou_R_Subdivided_ok")      # Paracou dataset, lists by inventoried year, sublists by plot, etc...
+dates<-sort(names(LivingStand_all))[-1] # dates considered
+
+Traits1<-read.csv("DB/BridgeOK.csv",sep=";",na.strings="")       # Brigde functional traits dataset
+Traits2<-read.csv("DB/DataLifeTraits.csv",sep=";",na.strings="") # Paracou15 seed and Hmax dataset
+TraitsName<-c("Hmax","S_mass","L_thickness","L_chloro",
+              "L_toughness","L_DryMass","SLA","WD","Bark_thick") # Considered traits list
+
+# Treatments definition
+T0<-c(1,6,11); T1<-c(2,7,9); T2<-c(3,5,10); T3<-c(4,8,12)
+treatments<-list(T0,T1,T2,T3)
+names(treatments)<-c("Control","T1","T2","T3")
+
+# list of all species recorded: to have global identical NMDS classification
+AllInv<-unique(unlist(lapply(LivingStand_all,function(yr){
+  return(unique(unlist(lapply(yr,function(pl){return(unique(pl[,"name"]))}))))})))
+AllInv<-as.data.frame(unique(AllInv),row.names=as.character(AllInv))
+
+Nrep<-50 # Iterations for the vernacular names replacement and traits gap filling
+
+MatrepFun<-lapply(1:Nrep,function(rep){
+  
+  Mat<-lapply(1:12,function(p){
+    
+    # For on plot (p),the species proba vector of successive inventories
+    Trajnmds<-lapply(dates,function(y){
+      
+      temp<-LivingStand_all[[which(names(LivingStand_all)==y)]]
+      if(!any(names(temp)==p)){return(NA)}
+      if(any(names(temp)==p)){
+        temp<-temp[[which(names(temp)==p)]]
+        temp<-temp[!duplicated(temp),]
[…]

### 2018-01-27 Scripts/TraitsMiceFilling.R

@@ -1,0 +1,100 @@
+library(mice)
+
+Traits_filling<-function(Traits1,Traits2){
+  Traits1<-Traits1[which(apply(Traits1,1,function(li){return(any(!is.na(li)))})),]
+  Traits2<-Traits2[which(apply(Traits2,1,function(li){return(any(!is.na(li)))})),]
+  
+## Traits correlation?
+Traits1["name"]<-as.factor(paste(Traits1$Genus,"_",Traits1$species,sep=""))
+Traits2["name"]<-sub(" ","_",Traits2[,"Name"])
+
+traits1<-Traits1[,c("Family","Genus","species","name","bar_code","thickness","SPAD","toughness","dry_mass","traits_surf_area","ind_surf_area",
+                    "sapwood_dens","moisture","bark_thick")]
+colnames(traits1)<-c("Family","Genus","species","name","bar_code","L_thickness","L_chloro","L_toughness","L_DryMass","LA","SLA","WD",
+                     "moisture","Bark_thick")
+traits2<-Traits2[,c("name","Hauteur","Masse")]
+colnames(traits2)<-c("name","Hmax","S_mass")
+
+traits<-merge(traits2,traits1,by="name")
+
+##SLA calc for empty values with LA available
+traits[which(!is.na(traits[,"LA"])),"SLA"]<-
+  traits[which(!is.na(traits[,"LA"])),"LA"]/traits[which(!is.na(traits[,"LA"])),"L_DryMass"]
+Seltraits<-c("Hmax","S_mass","L_thickness","L_chloro","L_toughness","L_DryMass","SLA","WD","moisture","Bark_thick")
+traits<-traits[,c("Family","Genus","name","bar_code",Seltraits)]
+
+##########################  Only some missing values: MICE regression to keep all available information
+### Non entire missing lines
+traitsPartial<-traits[which(apply(traits[,Seltraits],1,function(li){return(any(!is.na(li)))})),]
+traitsPartial<-traitsPartial[which(apply(traitsPartial[,Seltraits],1,function(li){return(any(is.na(li)))})),]
+traitsComp<-traits[which(apply(traits[,Seltraits],1,function(li){return(!any(is.na(li)))})),]
+traitsEmpty<-traits[which(apply(traits[,Seltraits],1,function(li){return(!any(!is.na(li)))})),]
+
[…]

### 2018-01-27 Scripts/Vernacular_handle.R

@@ -1,0 +1,119 @@
+#### Loadings
+library(vegan)
+library(reshape2)
+library(ade4)
+library(shape)
+library(cluster)
+library(mvtnorm)
+library(entropart)
+library(lattice)
+library(plotrix)
+library(gtools)
+library(RcppParallel)
+library(parallel)
+
+# Obtain the vernacular/botanical observed association frequency
+alpha_construct<-function(Plot){
+  mat_eff<-as.data.frame.matrix(with(Plot, xtabs(~name+nomPilote)))
+  mat_freq <- apply(mat_eff,2,function(v){v/sum(v)})
+  return(mat_freq)
+}
+
+# Built the plot-specific association matrix
+AlphaPlots<-function(Inv){
+  unlisted<-do.call(rbind,lapply(Inv,function(x){do.call(rbind,x)}))
+  ListedByPlot<-lapply(unique(unlisted[,"n_parcelle"]),function(y){
+    p<-unlisted[which(unlisted[,"n_parcelle"]==y),-which(colnames(unlisted)=="campagne")]
+    p<-p[which(!duplicated(p)),]
+    ap<-alpha_construct(p); return(ap)
+  })
+  names(ListedByPlot)<-unique(unlisted[,"n_parcelle"])
+  return(ListedByPlot)
+}
+
+# Listing all vernacular/genus and family associations 
+Correspondences<-function(Plot){
+  Corr<-Plot[order(Plot[,"Famille"],Plot[,"Genre"],Plot[,"Espece"]),]
+  Corr<-Corr[which(!duplicated(Corr)),]
+  return(Corr)
+}
+
+# Multinomial/dirichlet process, return a botanical name trialed in proba vector
+Dirichlet_draw<-function(V){
+  Vdir<-rdirichlet(1,V)
+  names(Vdir)<-names(V)
+  res <- rmultinom(1,1,Vdir)
+  res<-rownames(res)[which(res>0)]
+  return(res)
+}
+
+####### The function to randomly draw a botanical name replacing
+#the undertermined species. Use alpha matrix and botanic table
+tirage <- function(alpha,bota,name,eps=0.05){ 
+  
+  if(name["nomPilote"]=="-"){
+    trial<-sample(rownames(alpha),1)
+    while(grepl("Indet.",trial)){trial<-sample(rownames(alpha),1)}
+  }
+  
[…]

### 2018-01-28 Manuscript1.Rmd
Files and html update

@@
 ```{r Options, echo=FALSE, message=FALSE}
-knitr::opts_chunk$set(cache=TRUE, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
+knitr::opts_chunk$set(cache=TRUE, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '50%',fig.align="center")
 # Installation des packages si nécessaire et chargement
@@
 # Ajouter les packages nécessaires ici
-source("DrawingTools.R")
-
+source("Scripts/DrawingTools.R")
+source("Scripts/GithubPages.R")
 ```
@@
 
-```{r Fig1, echo=FALSE,fig.cap="Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
+```{r Fig1, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
 Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories.
@@
 
-```{r Fig2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of **(a)** taxonomic composition at genus level and **(b)** functional composition.
+```{r Fig2, out.width = '100%', fig.asp = 1.2/2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of **(a)** taxonomic composition at genus level and **(b)** functional composition.
 Distance are abundance-based Bary-Curtis metric.
@@
 
-```{r Fig4, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory, 10 years after disturbance, until 30 years after disturbance for plots **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversity.
+```{r Fig4, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory, 10 years after disturbance, until 30 years after disturbance for plots **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversity.
 Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
@@
 
-```{r Fig5, echo=FALSE,fig.cap="Trajectories of the plots Rao functional diversity from10 to 30 years after disturbance.
+```{r Fig5, fig.asp=1/1, echo=FALSE,fig.cap="Trajectories of the plots Rao functional diversity from10 to 30 years after disturbance.
 Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).

### 2018-01-28 DrawingTools.R

@@ -1,185 +1,0 @@
-### Manuscript drawing functions
-Library(c("shape"))
-
-treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
-names(treatments)<-c("Control","T1","T2","T3")
-ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
-
-TaxoCompo<-function(Data_TaxoComp){
-par(mfrow=c(1,2))
-matT<-as.data.frame(Data_TaxoComp[[1]])
-matT$plot<-substr(rownames(matT),start=1,stop=regexpr("_",rownames(matT))-1)
-matT$year<-substr(rownames(matT),start=regexpr("_",rownames(matT))+1,
-                  stop=nchar(rownames(matT)))
-colnames(matT)<-c("X","Y","plot","year")
-plot(matT$X,matT$Y,type="n",xlab="",ylab="",cex.axis=0.7)
-invisible(lapply(1:length(treatments),function(tr){
-  plots<-treatments[[tr]]
-  invisible(lapply(plots,function(pl){
-    toplot<-matT[which(matT$plot==pl),]
-    points(toplot[1,"X"],toplot[1,"Y"],col=ColorsTr[[tr]],pch=16,cex=1.7)
-    points(toplot[nrow(toplot),"X"],toplot[nrow(toplot),"Y"],col=ColorsTr[[tr]],pch=1,cex=1.7,lwd=2)
-    
-    lines(toplot$X,toplot$Y,col=ColorsTr[[tr]],lty=2)
-    invisible(lapply(round(seq(2,nrow(toplot),length.out=5)),function(step){
-      Arrows(x0=toplot[step,"X"],y0=toplot[step,"Y"],
-             x1=toplot[step-1,"X"],y1=toplot[step-1,"Y"],
-             col=ColorsTr[tr],code=1,lwd=2,arr.length=0.2,arr.type="simple")}))
-  }))
-}))
-
-mtext("NMDS 1",side=1,line=2,cex=0.8)
-mtext("Euclidean distance from 1989 inventory",side=2,padj=0,line=3,cex=0.9)
-mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)
-mtext("(a) Taxonomic composition",side=3,adj=0,line=0.5)
-}
-
-FunCompo<-function(Data_FunComp){mat<-as.data.frame(Data_FunComp[[3]])
[…]

### 2018-01-28 GithubPages.R

@@ -1,39 +1,0 @@
-### Make github pages (master branch / docs as web site)
-# The README.md file is copied to /docs. Make a link to the main html file in it
-# html files, including figures and libs are copied to /docs
-# pdf files are copied to /docs
-# Source this file before committing
-
-# Create the docs folder
-dir.create("docs")
-
-# Copy knitted html files
-htmlFiles <- list.files(pattern="*.html")
-if (length(htmlFiles) > 0)
-  file.rename(from=htmlFiles, to=paste("docs/", htmlFiles, sep=""))
-# Figures
-html_filesDir <- list.files(pattern="*_files")
-if (length(html_filesDir) > 0) {
-  sapply(paste("docs/", html_filesDir, sep=""), dir.create)
-  sapply(paste("docs/", html_filesDir, "/figure-html", sep=""), dir.create)
-  html_files <- list.files(html_filesDir, full.names = TRUE, recursive=TRUE)
-  if (length(html_files) > 0)
-    file.copy(from=html_files, to=paste("docs/", html_files, sep = ""), overwrite=TRUE)
-}
-#libs
-libsDirs <- list.dirs(path="libs", full.names=TRUE, recursive=TRUE)
-if (length(libsDirs) > 0) {
-  sapply(paste("docs/", libsDirs, sep = ""), dir.create)
-  libsFiles <- list.files("libs", full.names = TRUE, recursive=TRUE)
-  file.copy(from=libsFiles, to=paste("docs/", libsFiles, sep = ""), overwrite=TRUE)
-}
-
-# Copy knitted pdf files
-RmdFiles <- list.files(pattern="*.Rmd")
-# Change .Rmd files extension
-pdfFiles <- gsub(".Rmd", ".pdf", RmdFiles)
-if (length(pdfFiles) > 0)
-  file.rename(from=pdfFiles, to=paste("docs/", pdfFiles, sep=""))
-
-# Copy README.md to docs
-file.copy(from="README.md", to="docs/README.md")

### 2018-01-28 Manuscript1.Rmd
Files update

@@
 #  bookdown::word_document2
-# output_dir pour articles?
-#organiser le code, avec les script et DB adéquats. Mettre un code dans les figures pour indiquer le code d'où elle provient. ../ pour référer à des sous dossiers
+#Mettre un code dans les figures pour indiquer le code d'où elle provient. ../ pour référer à des sous dossiers
 ---
@@
 ```{r Options, echo=FALSE, message=FALSE}
-knitr::opts_chunk$set(cache=TRUE, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '50%',fig.align="center")
+knitr::opts_chunk$set(cache=TRUE, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
 # Installation des packages si nécessaire et chargement

### 2018-01-28 Scripts&DB_description.Rmd
Files and html update
Files update

@@
 - *TaxoComposition_ForGraphs*: Coordinates of inventories and plots in the 2 dimensions space from taxonomic composition NMDS. 
-Lists the coordinates along the various iteration: for now on, only 5 repetitions (becaus run locally).
+Lists the coordinates along the various iteration: for now on, only 5 repetitions (becaus run locally). **Figure 1 & 2 (a)**.
 
 - *FunctionalComposition_forGraphs*: Coordinates of inventories and plots in the 2 dimensions space from functional composition NMDS. 
-Lists the coordinates along the 50 iteration. Comes from the script: *FunctionalNMDS.R*
+Lists the coordinates along the 50 iteration. Comes from the script: *FunctionalNMDS.R*. **Figure 1 & 2 (b)**.
 
--*ReplacementTraj_ForGraphs*: List by treatment of plots Richness, Shannon and Simpson diversity along inventoried year. 
+- *ReplacementTraj_ForGraphs*: List by treatment of plots Richness, Shannon and Simpson diversity along inventoried year. 
 It gives the 0.025, 0.5 and 0.975 quantiles from 50 repetitions.
-Comes from the script: *ReplacementTraj_ForGraphs.R*
+Comes from the script: *ReplacementTraj_ForGraphs.R*. **Figure 3**.
 
--*FunctionalTraj_ForGraphs*: List by treatment of plots Rao diversity along inventoried year. 
-It gives the 0.025, 0.5 and 0.975 quantiles from *only 5 for now* repetitions.
+- *FunctionalTraj_ForGraphs*: List by treatment of plots Rao diversity along inventoried year. 
+It gives the 0.025, 0.5 and 0.975 quantiles from *only 5 for now* repetitions. **Figure 4**.
 
@@
 - *DivTaxo_trajectories.R*: Taxonomic Richness, Shannon and Simpson diversity trajectories per plot along inventoried years.
+
+- *DrawingTools.R*: To alleviate the *Manuscript* file I defined separately the functions to draw the plots from dataset.

### 2018-01-30 Manuscript1.Rmd
Manuscript update

@@
 ### Study site 
-Paracou station in French Guiana (5°18'N and 52°53'W) encompasses a lowland tropical rain forest, corresponds to a tropical wet climate with mean annual precipitation averaging 2980 mm.y-1 (30-y period) with a 3-months dry season (< 100 mm.months-1) from mid-August to mid-November, and a one-month dry season in (wagner AFM 2011).
+Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-months dry season (< 100 mm.months^-1^) from mid-August to mid-November, and a one-month dry season in March @Wagner2011.
 Elevation ranges between 5 and 50 m and mean annual temperature is 26°C. 
@@
 Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
-The logging correspond to averages of 10 trees remove per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3).
-Treatments T2 and T3 included besides logging of 10trees/ha the thinning by poison girdling.
+The logging correspond to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3).
+Treatments T2 and T3 included besides logging of 10trees/ha the thinning by poison girdling \ref{@Blanc2009}.
+
+Table 1: Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
+----------------------------------------------------------------------------------------------------------------------
+                      ***Timber***            ***Thinning***               ***Fuelwood***               *%AGB lost*
+-------------     --------------------     -------------------------     ------------------------     ---------------
+***Control***                                                                                              0
+
+***T1***            DBH ≥ 50 cm,
+                    commercial species,                                                                [12%-33%]
+                    ~ 10 trees/ha
+
+***T2***            DBH ≥ 50 cm,            DBH ≥ 40 cm,
+                    commercial species,    non-valuable species,                                        [33%-56%]
+                    ~ 10 trees/ha          ~ 30 trees/ha
+
+***T2***            DBH ≥ 50 cm,            DBH ≥ 50 cm,                   40 cm ≤ DBH ≤ 50 cm
+                    commercial species,    non-valuable species,           non-valuable species,        [33%-56%]
+                    ~ 10 trees/ha          ~ 15 trees/ha                   ~ 15 trees/ha
+----------------------------------------------------------------------------------------------------------------------
 
@@
 The study site corresponds to a tropical rain forest with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae. 
-In the twelve experimental plots all trees above 10 cm DBH are mapped and measured annually since 1984. 
-During the inventories trees are first identified with a vernacular name, assigned by the field team, and then with a scientific name, assigned by a botanist during regular botanical campaigns. 
+In the twelve experimental plots of the experiment, all trees above 10 cm DBH are mapped and measured annually since 1984. 
+During the inventories, trees are first identified with a vernacular name, assigned by the field team, and then with a scientific name, assigned by a botanist during regular botanical campaigns. 
 In 1984, specific vernacular names were given to 62 commercial or common species whereas two identifiers separating palm trees and trees were given to all other species. 
-Botanical campaigns to identify all trees at the species level started in 2003 and continued every 5 to 6 years. 
-In total, from 1984 to 2016, 848 species were identified in the whole dispositive. 
-The variability of botanical identification practices along time and among plots raised some methodological issues. 
-The first vernacular name attributed to trees, when not associated to a trustworthy botanical name thereafter, generate some taxonomic uncertainty. 
-To account for this taxonomic uncertainty in the composition and diversity metrics we used a Bayesian framework estimating the expectancy and variance of composition and diversity metrics. 
-The frameworks is based on a set of complete inventories extrapolated from the initial one through the replacement of vernacular name by botanical species.
-These replacement were made through a multinomial distribution M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name v and the species recorded in the inventory [*s~1~, s~2~, …, s~N~*].
-See appendix 1 and @Aubry-Kientz2013 Plos one  for the detailed methodology.
+The botanical campaigns to identify all trees at the species level started in 2003 and continued every 5 to 6 years.
+This inventory protocol rasied some mehodological issues given the variability of botanical identification practices along time and among plots. 
+Because vernacular name attributed to trees usually correspond to a variety of botanical species, some taxonomic uncertainty remains whenever a trustworthy botanical name cannot be given thereafter. 
+This taxonomic uncertainty was accounted for in the composition and diversity metrics throught a Bayesian framework estimating the expectancy and variance of composition and diversity metrics. 
+The uncertainty propagation is done by the extrapolation of complete inventories at genus level from real incomplete inventories, on the basis vernacular names replacement.
+The replacement of vernacular names uses a multinomial distribution M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
+See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.
 
-The functional metrics used a set of 6 functional traits values representing leaf and wood economics.
-The trait database encompassed leaves thickness, toughness, total chlorophyll content, dry mass and specific leaf area (leaf area per unit dry mass; *SLA*),  and wood density and bark thickness. 
-Species trait values were assessed in sampling at each iteration one of the individuals measured for traits collection as part of the BRIDGE project (http://www.ecofog.gf/Bridge/index.htm), which allowed to account for intraspecific variability.
-This project was carried out in nine permanent plots in French Guiana, including two in Paracou. 
-Individuals missing trait values were filled using multivariate imputation by chained equation (mice), restricted to the samples pertaining to the next higher taxonomic level to account for the phylogenetic signal of the traits.
-The dataset comprised 257 botanical species. Whenever an inventoried species was not in the dataset a set of traits values was sampled among species of the same next higher taxonomic level. 
+The functional metrics used a set of 6 functional traits values representing traits of the leaf and wood economics and life history traits.
+The trait database encompassed 4 traits of the leaves: leaves thickness (*L_thick*), toughness (*L_tough*), total chlorophyll content (*LCC*) and specific leaf area (leaf area per unit dry mass; *SLA*), 2 traits of the stem: wood specific gravity (*SWG*) and bark thickness (*B_thick)* and two life-history traits: maximum specific height (*Hmax*) and the seed mass (*Smass*).
+Traits database came from the BRIDGE project (http://www.ecofog.gf/Bridge/index.html) where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou.
+To account for intraspecific traits variability, trait values at species level were assessed in sampling one of the measured individuals for iteration rather than averaging at species level. 
+Missing trait values were filled using multivariate imputation by chained equation (mice), restricted to the samples pertaining to the next higher taxonomic level, in order to account for the phylogenetic signal of the functional traits.
+The dataset comprised 257 botanical species.
+Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level.
 All composition and diversity metrics correspond to the expectancy obtained after 50 iterations of the taxonomic uncertainty propagation framework and the missing trait values filling process.
-Because of taxonomic uncertainty issues the taxonomic composition and diversity analysis were conducted at the genus level, knowing that diversity patterns remain consistent in this case.
 
@@
 
-The analysis of taxonomic and functional composition of plots assemblages was visualized in a two-dimensional ordination space based on non-metric dimensional scaling of their floristic or functional composition.
-Their evolution along time relative to initial condition was reported using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) similarity measures. 
-To determine whether plots composition converged to their initial condition we computed the distance to initial condition of the sampled years in the two-dimensional space.
-Additionally, the functional composition was analyzed through the trajectory along time of the community weighted means (CWM) for the 6 functional traits chosen.
+The analysis of taxonomic and functional composition of plots along time was visualized in a two-dimensional ordination space based on non-metric dimensional scaling of succesive floristic or functional inventories.
+Plots evolution along time was reportedcomparatively to the inventories in 1989, 5 years after disturbance, which correspond to first inventory with a sufficient degree of uncertainty (<30% of undetermined trees).
+The distancing relative to the reference 1989 inventory was reported using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) similarity measures.
+The euclidean distance in the two-dimensional ordination space was used to visualize the distancing to reference inventory along time.
+The functional composition was besides completed by trajectory along time of the community weighted means (CWM) for the 8 leaf, stem and life history traits. 
 To determine whether communities’ assembly deviated from the expectation of a random process we compared the observed similarity patterns to those generated by 50 repetitions of stochastic null models.
@@
 These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q). 
-This order q grasps the balance between richness and evenness in the community as it determines the emphasis on common species in the diversity metric, with common species weighting more than rare ones when q increases.
+This order q grasps the balance between richness and evenness in the community as it determines the weight of common species in the diversity metric, with an growing emphasis on dominant species as q increases.
 The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on the 6 functional traits specific values.
@@
 
-Over time, 104423 trees and 697 botanical species were recorded. 
+Over time, 104423 trees and 696 botanical species were recorded. 
 The disturbance proved to significantly impact for more than 30 years the functional and taxonomic composition and diversity of the stands.
@@
 Both taxonomic and functional composition were substantially affected by disturbance and positively correlated with its intensity.
-*The distance from the initial condition was significantly time-dependent (P<0.01).* The taxonomic composition remained significantly different from the 1989 inventory until 30 years for all disturbed plots \autoref{Fig1}.
-The distancing was positively correlated woth disturbance intensity.
-From 20 years after disturbance though the distancing reduced, suggesting the beginning of a return toward the reference 1989 inventory \autoref{Fig1},\autoref{Fig2}.
-The functional composition distancing from 1989 was also positively correlated with disturbance intensity and similarly stabilized and reduced for 5 of T1, T2 and T3 plots from 20 years after disturbance \autoref{Fig2}.
+*The distance from the initial condition was significantly time-dependent (P<0.01).* 
+The taxonomic composition remained significantly different from the 1989 reference inventory until 30 years for all disturbed plots \autoref{Fig1}. 
+From 20 years after disturbance, though, the distancing decreased which suggested a return toward the reference 1989 inventory and its resilience to disturbance \autoref{Fig1},\autoref{Fig2}. 
+The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance \autoref{Fig2}.
 *The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).*
@@
 
-*Traits community weighted means* _ To complete the functional analysis the trajectories of the community weighted means (CWM) measured for the 6 functional traits at plot level \autoref{Fig3}.
+*Traits community weighted means* _ The trajectories of the community weighted means (CWM) of the 8 functional traits were measured at plot level \autoref{Fig3}.
 **To be coming**
@@
 
-The trajectories of the Richness, Shannon and Simpson differences to reference inventories in 1994 were rather even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase depending on the disturbance intensity \autoref{Fig4}.
-
+The evolution along time of the difference of Richness, Shannon and Simpson diversity compared to reference inventories in 1989 remained even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase, positively correlated with the disturbance intensity \autoref{Fig4}.
 The richness stabilized for all disturbed plots from 20 years after disturbance \autoref{Fig4}. 
@@
 
-*Functional  diversity* _ The trajectories of plot 7 from the T1 disturbance treatment was discarded because of too high taxonomic uncertainty preventing correct analysis. 
-While the control plots showed an even trajectory until 30 years after disturbance, all disturbed plots displayed an increase of the functional diversity positively correlated to the disturbance intensity \autoref{Fig5}.
-The functional diversity observed 30 years after disturbance remained higher for all disturbed plots compared to control ones, without clear distinction following the disturbance intensity. 
+*Functional  diversity* _ The evolution along time of the Rao functinal diversity compared to reference inventories in 1989 remained even all along the survey for the control plots. 
+All disturbed plots however showed an increase of the functional diversity that was positively correlated to the disturbance intensity \autoref{Fig5}.
+30 after disturbance the diversity of disturbed plots remained significantly higher than this of control plots, thus whenever the initial disturbance intensity.
 The slope of the initial diversity increase and the maximum diversity were both positively correlated with the disturbance intensity. 
 *The OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and the covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01)*.
+The trajectories of plot 7 from the T1 disturbance treatment was discarded because of too high taxonomic uncertainty preventing correct analysis.
 

### 2018-01-30 AppendixI.Rmd
Manuscript update

--- 2018-01-28 Manuscript1.Rmd
@@
 ---
-title: "Post-Disturbance Diversity Trajectories in a Neotropical Forest"
-author:
-  - name: Ariane MIRABEL
-    authsuperscript: 1*
-  - name: Eric Marcon
-    authsuperscript: 1
-  - name: Bruno Hérault
-    authsuperscript: 2
-affiliation:
-  - affsuperscript: 1
-    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
-    address: >
-      Campus Agronomique,
-      97310 Kourou,
-      France.
-  - affsuperscript: 2
-    dptuniv: Department of Ecology, University of Edimburgh
-    address: >
-      Street address,
-      Zip code,
-      Country.
-corrauthor:
-    email: ariane.mirabel@ecofog.gf
-    url: http://www.ecofog.gf/spip.php?article47
-abstract: >   #Pour avoir plusieurs lignes qui suivent. Commencer à 2 caractères
-  Résumé de l'article.
-keywords: [mot-clés, séparés par des virgules]
-journalinfo: Hal 00679993
-archive: DOI xxxx
-date: "`r format(Sys.time(), '%d %B %Y')`"
-preamble: >  # Pour coordonne les packages LateX (supplémentaires)
-  \usepackage{amsmath}
-bibliography: references.bib # .bib à activer seulement pour git et word format (pandoc fait la biblio)
-csl: 
-lang: english # french
-keywordlabel: Keywords
-corrauthorlabel: Corresponding author
-output:
-  bookdown::gitbook
-#   bookdown::pdf_book:
-#     base_format: EcoFoG::article
-#     keep_tex: TRUE
-#  bookdown::word_document2
-#Mettre un code dans les figures pour indiquer le code d'où elle provient. ../ pour référer à des sous dossiers
+title: "Appendix I _ Scheme of the taxonomic uncertainty handle"
+author: "Ariane"
+date: "30 janvier 2018"
+output: html_document
 ---
-
-<!-- Options de knitr et de R (librairies, ...) -->
-
-```{r Options, echo=FALSE, message=FALSE}
-knitr::opts_chunk$set(cache=TRUE, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
-# Installation des packages si nécessaire et chargement
-Library <- function(Packages) {
-  InstallAndLoad <- function(Package) {
-    if (!Package %in% installed.packages()[, 1]) {install.packages(Package)}
-    require(Package, character.only = TRUE)
-  }
-  invisible(sapply(Packages, InstallAndLoad))
-}
-# Ajouter les packages nécessaires ici
-source("Scripts/DrawingTools.R")
-source("Scripts/GithubPages.R")
-```
-
-# Introduction
-The large areas covered with tropical forests worldwide hold crucial commercial, social and cultural value. 
-They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate local climate and support nutrients, water and carbon cycles and ensure cultural and human well-being. 
-The demand for forests products and the shrink of unmanaged areas increased the human pressure on forests and currently threaten the maintenance of their structure, composition and function. 
-The increasing demand for forests products, land-use changes and other human disturbances along with substantial climatic changes make a crucial issue of understanding the response of forests dynamics to such disturbance regimes [@Gibson2011a; @Morales-Hidalgo2015].
-This has been addressed from the point of view of biomass production and carbon storage which resilience proved subjected to identified cyclic dynamics and therefore gives hope for the maintenance of forest ecosystem and their services.
-The response of trees biological diversity however has been sparsely studied, although this diversity is a major determinant of forests productivity, stability and functioning as it ensures interspecific complementarity, enhances the use of resources and the store of nutrients, and mitigates the prejudices of diseases and herbivory [@Tilman2014].
-Maintain the integrity of forests functioning thus requires highlighting the impacts and underlying mechanisms of disturbance on trees diversity and question the extent of their resilience. 
-Disturbance alter the biotic and abiotic parameters of the ecosystem in modifying the light availability, the heat and water fluxes, the plants productivity, etc. 
-Although those changes proved impacting the proportion and identity of species, negligible or even positive impacts were highlighted when considering the community as a whole, in accordance with the intermediate disturbance hypothesis (IDH).
-The IDH that predicts a maximized species diversity at intermediate level of disturbance, however, mainly rely on the study of species richness which gives limited or even misleading information on forests recovery and functioning.
-To have a detailed understanding of forests response to disturbance it is essential to identify the subsequent impact on species identity per se, which underlies any diversity change, as well as on richness, equitability and functional profile of the whole community.
-Such combination of composition and diversity in browsing all facets of stands assemblage would highlight their resilience to perturbation, and its determinant and underlying ecological processes.
-Disturbance survey would first assess species turnover and communities convergence through stands taxonomic composition and unveil environmental filters through functional composition that is based on the biological attributes of plants.
-The set of functional traits chosen for this functional analysis considers functional traits representative of species functional strategy, fitness, and ecology as it is the case for traits of the leaf or stem economics.
-Composition analysis would be continued with diversity analysis assessing for the whole community the richness and evenness of the taxonomic and functional abundance distribution.   
-Here we examined the long-term trajectory, in the sense of the evolution along time, of tropical forests taxonomic and functional composition and diversity. 
-We studied the regular inventories over 30 years of twelve Amazonian forest plots, either untouched or disturbed with three disturbance intensity . 
-The inventories were combined with one of the largest functional trait database for tropical trees.
-We aimed to (i) explicit the impact of disturbance on the long-term trajectory of tropical forest communities composition and diversity, (ii) question the long-term convergence of the composition of communities and (iii) determine their resilience or not, after disturbance.
-
-# Material and Methods
-
-### Study site 
-Paracou station in French Guiana (5°18'N and 52°53'W) encompasses a lowland tropical rain forest, corresponds to a tropical wet climate with mean annual precipitation averaging 2980 mm.y-1 (30-y period) with a 3-months dry season (< 100 mm.months-1) from mid-August to mid-November, and a one-month dry season in (wagner AFM 2011).
-Elevation ranges between 5 and 50 m and mean annual temperature is 26°C.
-Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
-The experiment corresponds to a network of twelve 6.25ha plots where three logging, thinning and fuelwood treatments on an intensity gradient.
-Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots.
-The logging correspond to averages of 10 trees remove per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3).
-Treatments T2 and T3 included besides logging of 10trees/ha the thinning by poison girdling.
-
-### Inventories protocol and dataset correction
-The study site corresponds to a tropical rain forest with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae. 
-In the twelve experimental plots all trees above 10 cm DBH are mapped and measured annually since 1984. 
-During the inventories trees are first identified with a vernacular name, assigned by the field team, and then with a scientific name, assigned by a botanist during regular botanical campaigns. 
-In 1984, specific vernacular names were given to 62 commercial or common species whereas two identifiers separating palm trees and trees were given to all other species. 
-Botanical campaigns to identify all trees at the species level started in 2003 and continued every 5 to 6 years. 
-In total, from 1984 to 2016, 848 species were identified in the whole dispositive. 
-The variability of botanical identification practices along time and among plots raised some methodological issues. 
-The first vernacular name attributed to trees, when not associated to a trustworthy botanical name thereafter, generate some taxonomic uncertainty. 
-To account for this taxonomic uncertainty in the composition and diversity metrics we used a Bayesian framework estimating the expectancy and variance of composition and diversity metrics. 
-The frameworks is based on a set of complete inventories extrapolated from the initial one through the replacement of vernacular name by botanical species.
-These replacement were made through a multinomial distribution M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name v and the species recorded in the inventory [*s~1~, s~2~, …, s~N~*].
-See appendix 1 and @Aubry-Kientz2013 Plos one  for the detailed methodology.
-
-The functional metrics used a set of 6 functional traits values representing leaf and wood economics.
-The trait database encompassed leaves thickness, toughness, total chlorophyll content, dry mass and specific leaf area (leaf area per unit dry mass; *SLA*),  and wood density and bark thickness. 
-Species trait values were assessed in sampling at each iteration one of the individuals measured for traits collection as part of the BRIDGE project (http://www.ecofog.gf/Bridge/index.htm), which allowed to account for intraspecific variability.
-This project was carried out in nine permanent plots in French Guiana, including two in Paracou. 
-Individuals missing trait values were filled using multivariate imputation by chained equation (mice), restricted to the samples pertaining to the next higher taxonomic level to account for the phylogenetic signal of the traits.
-The dataset comprised 257 botanical species. Whenever an inventoried species was not in the dataset a set of traits values was sampled among species of the same next higher taxonomic level. 
-All composition and diversity metrics correspond to the expectancy obtained after 50 iterations of the taxonomic uncertainty propagation framework and the missing trait values filling process. 
-Because of taxonomic uncertainty issues the taxonomic composition and diversity analysis were conducted at the genus level, knowing that diversity patterns remain consistent in this case.
-
-### Composition and diversity metrics
-
-The analysis of taxonomic and functional composition of plots assemblages was visualized in a two-dimensional ordination space based on non-metric dimensional scaling of their floristic or functional composition.
-Their evolution along time relative to initial condition was reported using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) similarity measures. 
-To determine whether plots composition converged to their initial condition we computed the distance to initial condition of the sampled years in the two-dimensional space.
-Additionally, the functional composition was analyzed through the trajectory along time of the community weighted means (CWM) for the 6 functional traits chosen.
-To determine whether communities’ assembly deviated from the expectation of a random process we compared the observed similarity patterns to those generated by 50 repetitions of stochastic null models.
-The null model for taxonomic composition randomly shuffled individuals among plots while preserving the overall species abundance and plots’ tree density. 
-The null model for functional composition produced random traits assemblages by permuting the set of functional traits values among species.
-
-Plots taxonomic diversity corresponded to species Richness and the Hill number translation of Shannon and Simpson indices.
-These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q).
-This order q grasps the balance between richness and evenness in the community as it determines the emphasis on common species in the diversity metric, with common species weighting more than rare ones when q increases.
-The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on the 6 functional traits specific values.
-
-# Results
-
-Over time, 104423 trees and 697 botanical species were recorded. 
-The disturbance proved to significantly impact for more than 30 years the functional and taxonomic composition and diversity of the stands.
-
-### Taxonomic and functional composition
-
-*Composition trajectories* _ Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in a two-dimensional space based on NMDS analysis.
-Classifications were based on either abundance-based Bray-Curtis \autoref{Fig1} or incidence-based Jaccard dissimilarity (data not shown).
-
-Both taxonomic and functional composition were substantially affected by disturbance and positively correlated with its intensity. 
-*The distance from the initial condition was significantly time-dependent (P<0.01).* The taxonomic composition remained significantly different from the 1989 inventory until 30 years for all disturbed plots \autoref{Fig1}.
-The distancing was positively correlated woth disturbance intensity.
-From 20 years after disturbance though the distancing reduced, suggesting the beginning of a return toward the reference 1989 inventory \autoref{Fig1},\autoref{Fig2}.
-The functional composition distancing from 1989 was also positively correlated with disturbance intensity and similarly stabilized and reduced for 5 of T1, T2 and T3 plots from 20 years after disturbance \autoref{Fig2}.
-*The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).*
-
-```{r Fig1, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
-Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3)."}
-load("DB/FunctionalComposition_forGraph");
-load("DB/TaxoComposition_forGraphs")
-
-TaxoCompo(MatrepTaxo)
-FunCompo(MatrepFun)
-
-```
-
-```{r Fig2, out.width = '100%', fig.asp = 1.2/2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of **(a)** taxonomic composition at genus level and **(b)** functional composition.
-Distance are abundance-based Bary-Curtis metric.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.
-"}
-
-TaxoDist(MatrepTaxo)
-FunDist(MatrepFun)
-
-```
-
-*Traits community weighted means* _ To complete the functional analysis the trajectories of the community weighted means (CWM) measured for the 6 functional traits at plot level \autoref{Fig3}.
-**To be coming**
-
-```{r Fig3, echo=FALSE,fig.cap="Trajectories of the CWM until 30 years after disturbance of 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness).
-The CWM were measured at plot level and averaged among plots of the same treatment (green for control, blue for T1,orange for T2 and red for T3).
-The average CWM per treatment are drawn (solid line) drawn along with the 0.025 and 0.975 percentile (gray envelope) observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
-```
-
-### Disturbance impact on diversity
-
-*Taxonomic diversity* _ Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1994 inventories \autoref{Fig4}. 
-*For the three indices the OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and a covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01).*
-
-```{r Fig4, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory, 10 years after disturbance, until 30 years after disturbance for plots **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversity.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The average value of the 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness) are drawn along with the 0.025 and 0.975 percentile observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
-load("DB/ReplacementTraj_ForGraphs")
-
-TaxoTraj(CompleteTaxo)
-
-```
-
-The trajectories of the Richness, Shannon and Simpson differences to reference inventories in 1994 were rather even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase depending on the disturbance intensity \autoref{Fig4}.
-
-The richness stabilized for all disturbed plots from 20 years after disturbance \autoref{Fig4}.
-The Shannon and Simpson diversity also stabilized 20 years after disturbance for T1 and T2 plots \autoref{Fig4} whereas for 2 out of 3 plots T3 the Shannon and Simpson diversity continued increasing until 30 years after disturbance, illustrating a significant increase in communities’ evenness.
-
-*Functional  diversity* _ The trajectories of plot 7 from the T1 disturbance treatment was discarded because of too high taxonomic uncertainty preventing correct analysis. 
-While the control plots showed an even trajectory until 30 years after disturbance, all disturbed plots displayed an increase of the functional diversity positively correlated to the disturbance intensity \autoref{Fig5}.
-The functional diversity observed 30 years after disturbance remained higher for all disturbed plots compared to control ones, without clear distinction following the disturbance intensity. 
-The slope of the initial diversity increase and the maximum diversity were both positively correlated with the disturbance intensity. 
-*The OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and the covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01)*.
-
-```{r Fig5, fig.asp=1/1, echo=FALSE,fig.cap="Trajectories of the plots Rao functional diversity from10 to 30 years after disturbance.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The value of the 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness) are drawn along with the 0.025 and 0.975 percentile observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
-load("DB/FunctionalTraj_ForGraphs")
-
-FunTraj(CompleteFun)
-```
-
-# Discussion
--	The disturbance significantly impact the functional and taxonomic composition and diversity of the stands for more than 30 years 
--	Important differences among plots of same treatments: possible impact of initial diversity composition, local abiotic parameters (topography, soils, surrounding stands).
-Impacts depend on the disturbance intensity, but further parameters may be involved.
--	The shift in functional composition differs from the null model: biotic and abiotic filters are at stake but only tangible from 10/15 years after logging.
-Discussion with Cannon, 1998: they found no different process than rarefaction for 8 years after logging, but the time might have been too short-> divergence due to recruitment and not rarefaction after random species disappearance.
-Same conclusions otherwise.
--	Disturbance involves new filters for species settling in the stands, probably according to their functional strategy.
-Disturbance generates a functional shift of communities toward acquisitive strategies.
-Ecosystems function rely on abundant species: changes in stands functioning is then expected.
-(Flores, 2006  for differential sapling success after disturbance).
--	If the recovery is not achieved 30 years after disturbance, it is on the way to.
-Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories.
--	To go further we should focus on recruitment which would also highlight the mechanisms underlying these shifts and dynamics and explicit the environmental filter.
-
-# Appendix I _ Scheme of the taxonomic uncertainty handle
 

### 2018-01-30 Manuscript1.Rmd
html files update
Manuscript update

@@
 Table 1: Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
+
 ----------------------------------------------------------------------------------------------------------------------
@@
 Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3)."}
-load("DB/FunctionalComposition_forGraph");
+load("DB/FunctionalComposition_forGraph")
 load("DB/TaxoComposition_forGraphs")
@@
 
-# Appendix I _ Scheme of the taxonomic uncertainty handle
+`r if (knitr:::is_html_output()) '# References {-}'`
 
-![**Fig.
-A1 _ ** Whenever the family or genus were recorded for a vernacular name, the pool of species sampled to replace the vernacular name was restricted to those pertaining to the same higher taxonomic level (i.e species of the same genus when the species is missing, species of the same family when the genus is missing).](BayesianSheme.png)
+---
+bibliography: ["references.bib"]
+biblio-style: "apalike"
+link-citations: true
+---

### 2018-02-06 Manuscript1.Rmd
Manuscript update

@@
 ### Study site 
-Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-months dry season (< 100 mm.months^-1^) from mid-August to mid-November, and a one-month dry season in March @Wagner2011.
+Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-months dry season (< 100 mm.months^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
 Elevation ranges between 5 and 50 m and mean annual temperature is 26°C. 
 Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
-The experiment corresponds to a network of twelve 6.25ha plots where three logging, thinning and fuelwood treatments on an intensity gradient.
+The experiment corresponds to a network of twelve 6.25ha plots that have undergone a gradient of three logging, thinning and fuelwood treatments.
 Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
-The logging correspond to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3).
-Treatments T2 and T3 included besides logging of 10trees/ha the thinning by poison girdling \ref{@Blanc2009}.
+The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3).
+Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Blanc2009].
 
 Table 1: Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
+---------------------------------------------------------------------------------------------------------------------------------
+                          ***Timber***                 ***Thinning***              ***Fuelwood***           **%AGB lost**
 
-----------------------------------------------------------------------------------------------------------------------
-                      ***Timber***            ***Thinning***               ***Fuelwood***               *%AGB lost*
--------------     --------------------     -------------------------     ------------------------     ---------------
+-----------------    ------------------------    ------------------------    ------------------------    ------------------------
     ***Control***                                                                                                   0
 
+-----------------    ------------------------    ------------------------    ------------------------    ------------------------
          ***T1***     DBH ≥ 50 cm,                                                                                               
@@
 
+-----------------    ------------------------    ------------------------    ------------------------    ------------------------
          ***T2***     DBH ≥ 50 cm,                DBH ≥ 40 cm,                                                                   
@@
 
-***T2***            DBH ≥ 50 cm,            DBH ≥ 50 cm,                   40 cm ≤ DBH ≤ 50 cm
+-----------------    ------------------------    ------------------------    ------------------------    ------------------------
+         ***T3***     DBH ≥ 50 cm,                DBH ≥ 50 cm,                 40 cm ≤ DBH ≤ 50 cm                               
                       commercial species,         non-valuable species,        non-valuable species,           [33%-56%]         
                       ~ 10 trees/ha               ~ 15 trees/ha                ~ 15 trees/ha
-----------------------------------------------------------------------------------------------------------------------
+
+---------------------------------------------------------------------------------------------------------------------------------
 
@@
 Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories.
+however this is the impact of a single disturbance event: the repetition may have different impact (@Olander2005 demonstrated the impact of logging on soils nutrient and compaction)
 -	To go further we should focus on recruitment which would also highlight the mechanisms underlying these shifts and dynamics and explicit the environmental filter.

### 2018-02-06 Saucisson.tex
Manuscript update

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Article EcoFoG
% Version 2.1 (23/10/2017)
%
% adapté de :
% Stylish Article
% LaTeX Template
% Version 1.0 (31/1/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Mathias Legrand (legrand.mathias@gmail.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[fleqn,10pt]{ArtEcoFoG} % Document font size and equations flushed left

\setcounter{tocdepth}{3} % Show only three levels in the table of contents section: sections, subsections and subsubsections

% Pandoc environments
\usepackage{framed}
\usepackage{fancyvrb}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, fontsize=\scriptsize} % Code R
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable,booktabs}
\usepackage{caption}
% These lines are needed to make table captions work with longtable:
\makeatletter
\def\fnum@table{\tablename~\thetable}
\makeatother
% longtable 2 columns
% https://tex.stackexchange.com/questions/161431/how-to-solve-longtable-is-not-in-1-column-mode-error
\makeatletter
\let\oldlt\longtable
\let\endoldlt\endlongtable
\def\longtable{\@ifnextchar[\longtable@i \longtable@ii}
\def\longtable@i[#1]{\begin{figure}[t]
\onecolumn
\begin{minipage}{0.5\textwidth}\scriptsize
\oldlt[#1]
}
\def\longtable@ii{\begin{figure}[t]
\onecolumn
\begin{minipage}{0.5\textwidth}\scriptsize
\oldlt
}
\def\endlongtable{\endoldlt
\end{minipage}
\twocolumn
\end{figure}}
\makeatother

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight0.8\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

% User-adder preamble
\usepackage{amsmath}

%----------------------------------------------------------------------------------------
%	ARTICLE INFORMATION
%----------------------------------------------------------------------------------------

\JournalInfo{Hal 00679993} % Journal information
\Archive{DOI xxxx} % Additional notes (e.g. copyright, DOI, review/research article)

\PaperTitle{Titre de l'article} % Article title

\Authors{
Prénom Nom\textsuperscript{1*}\\ Deuxième Auteur\textsuperscript{2}
} % Authors
\affiliation{
\textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles,
Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}Department of Ecology, University of Edimburgh\\ \hspace{1em} Street address, Zip code, Country.
}
\affiliation{*\textbf{Contact}: prenom.nom@ecofog.gf, http://www.ecofog.gf/spip.php?article47} % Corresponding author

\Keywords{mot-clés, séparés par des virgules} % Keywords - if you don't want any simply remove all the text between the curly brackets
\newcommand{\keywordname}{Mots-clés} % Defines the keywords heading name

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\Abstract{
Résumé de l'article.
}

%----------------------------------------------------------------------------------------

\begin{document}

\selectlanguage{french}

\flushbottom % Makes all text pages the same height

\maketitle % Print the title and abstract box

\tableofcontents % Print the contents section

\thispagestyle{empty} % Removes page numbering from the first page

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}\label{introduction}

Ce modèle permet la rédaction d'articles au format Markdown. Il produit
directement des articles bien formatés pour l'auto-archivage (dépôt sur
HAL par exemple) ou sous d'autres formats, par exemple HTML.

\section{R Markdown}\label{markdown}

Markdown est un langage très simple pour produire divers types de
documents: HTML, PDF, et Word entre autres. Sa documentation est
disponible sur le site de RStudio\footnote{https://github.com/citation-style-language/styles}.

Markdown est étendu par Bookdown \footnote{\url{https://bookdown.org/yihui/bookdown/}},
qui permet la rédaction de livres et une syntaxe plus efficace pour les
articles. Ce document est réalisé avec Markdown dans RStudio: knitr
traite le code Markdown, le passe à Pandoc pour sa transformaton en
Latex, enfin MikteX le compile en PDF.

\subsection{Intérêt}\label{interet}

Markdown est très simple à apprendre.

Markdown permet d'intégrer son code R pour un résultat
\emph{reproductible}.

Markdown permet de produire, sans réécrire le texte, un document dans
différents formats: article Latex ou Word par exemple.

\subsection{Comment faire}\label{comment-faire}

Dans RStudio, créer un nouveau document de type Document R Markdown.
L'assistant permet de choisir entre divers formats. Les plus
intéressants sont:

\begin{figure}
\centering
\includegraphics{NouveauRMarkdown.PNG}
\caption{\label{fig:nouveau}Nouveau document}
\end{figure}

\begin{itemize}
\item
  Document: rapport simple
\item
  Presentation: diaporama
\item
  From template: à partir de modèles installés par des packages. Les
  modèles du package EcoFoG sont affichés (voir figure
  \ref{fig:nouveau}): choisir Article EcoFoG.
\item
  Ecrire le document dans RStudio.
\item
  Cliquer sur le bouton \textbf{Knit} de RStudio génère le document au
  format demandé.
\end{itemize}

\section{Code}\label{code}

Les principales caractéristiques de Markdown sont résumées ici.

\subsection{Code R}\label{code-r}

Le code R est inclus dans des bouts de code (\emph{code chunks}):

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(cars)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   speed dist
## 1     4    2
## 2     4   10
## 3     7    4
## 4     7   22
## 5     8   16
## 6     9   10
\end{verbatim}

\subsection{Tableaux}\label{tableaux}

Les séparateurs horizontaux - et verticaux \textbar{} permettent de
dessiner un tableau. Les tableaux peuvent aussi être produits par du
code R, avec le package knitr:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{names}\NormalTok{(iris) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Longueur sépales"}\NormalTok{, }\StringTok{"Largeur"}\NormalTok{, }\StringTok{"Longueur pétales"}\NormalTok{, }\StringTok{"Largeur"}\NormalTok{, }\StringTok{"Espèce"}\NormalTok{)}
\NormalTok{knitr}\OperatorTok{::}\KeywordTok{kable}\NormalTok{(}\KeywordTok{head}\NormalTok{(iris), }\DataTypeTok{caption=}\StringTok{"Tableau créé par R"}\NormalTok{, }\DataTypeTok{longtable =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{booktabs =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[t]{rrrrl}
\caption{\label{tab:kable}Tableau créé par R}\\
\toprule
Longueur sépales & Largeur & Longueur pétales & Largeur & Espèce\\
\midrule
5.1 & 3.5 & 1.4 & 0.2 & setosa\\
4.9 & 3.0 & 1.4 & 0.2 & setosa\\
4.7 & 3.2 & 1.3 & 0.2 & setosa\\
4.6 & 3.1 & 1.5 & 0.2 & setosa\\
5.0 & 3.6 & 1.4 & 0.2 & setosa\\
5.4 & 3.9 & 1.7 & 0.4 & setosa\\
\bottomrule
\end{longtable}

Une légende est possible et le référencement aussi (tableau
\ref{tab:kable}). Attention, la largeur des tableaux est limitée à une
colonne de texte, et les retours à la ligne dans les cellules ne sont
pas pris en charge.

\subsection{Figures}\label{figures}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(pressure)}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\centering
\includegraphics{Saucisson_files/figure-latex/pressure-1.pdf}
\caption{\label{fig:pressure}Titre de la figure}
\end{figure}

Les figures peuvent être créées par le code R (figure
\ref{fig:pressure}). Avec Bookdown, une étiquette est associée à chaque
figure: son nom est \texttt{fig:xxx} où \texttt{xxx} est le nom du bout
de code R. Les renvois se fonct avec la commande
\texttt{\textbackslash{}@ref(fig:xxx)}.

Une figure peut utiliser toute la largeur de la page en ajoutant les
options suivantes dans l'entête du bout de code qui la génère:
\texttt{fig.env="figure*"}et \texttt{out.extra=""}.

Les figures existantes sont intégrées dans un bout de code par la
fonction \texttt{include\_graphics}, voir la figure \ref{fig:nouveau}.

\subsection{Listes}\label{listes}

Les listes sont indiquées par des *, + et - (trois niveaux
hiérarchiques) ou des nombres 1., i. et A. (listes numérotées).

\begin{itemize}
\item
  Liste

  \begin{itemize}
  \tightlist
  \item
    sous-liste
  \end{itemize}
\item
  deuxième élément
\item
  Suite de la liste
\end{itemize}

\subsection{Maths}\label{maths}

Les équations au format Latex peuvent être insérées en ligne, comme
\(A=\pi r^2\) ou isolées comme \[e^{i \pi} = -1.\]

Elles peuvent être numérotées (voir équation \eqref{eq:disque}) en
utlisant l'environnement \emph{equation} :

\begin{equation}
A = \pi r^2.
\label{eq:disque}
\end{equation}

\subsection{Références croisées}\label{references-croisees}

Les figures et tableaux ont une étiquette générée automatiquement,
identique au nom du bout de code préfixé par \texttt{fig:} et
\texttt{tab:}.

Pour les équations, l'étiquette est ajoutée manuellement par le code
\texttt{(\textbackslash{}\#eq:xxx)} avant la fin de l'équation.

Les sections peuvent recevoir une étiquette en terminant leur titre par
\texttt{\{\#yyy\}}.

Des signets peuvent aussi être placés librement dans le texte avec la
commande \texttt{(ref:zzz)}.

Dans tous les cas, l'appel à la référence est fait par la commande
\texttt{\textbackslash{}@ref(ref:zzz)}.

\subsection{Bibliographie}\label{bibliographie}

Les références bibliographiques incluses dans le fichier references.bib
peuvent être appelées dans le texte, entre parenthèses
\citep{Marcon2014c}, ou dans le texte, à la façon de
\citet{Marcon2014c}.

La bibliographie est traitée par Pandoc lors de la production de
documents Word ou HTML. Le sytle bibliographique peut être précisé, en
ajoutant la ligne

\begin{verbatim}
csl:nom_du_fichier.csl
\end{verbatim}

dans l'entête du document et en copiant le fichier de style \emph{.csl}
dans le dossier du projet. Plus d'un millier de styles sont disponibles
\footnote{https://github.com/citation-style-language/styles}.

Pour les documents PDF, la bibliographie est gérée par \LaTeX: c'est
pourquoi l'extension \emph{.bib} du fichier doit être supprimée dans la
déclaration de la base bibliographique. Le style est inclus dans le
modèle EcoFoG: c'est celui de \emph{Methods in Ecology and Evolution}.
Il ne peut pas être changé, pour assurer l'homogénéité des documents
produits.

Pour préparer la soumission d'un manuscrit à une revue, il faudra ouvrir
le fichier \emph{.tex} intermédiaire produit par Pandoc et copier le
contenu de l'environnement \{document\} dans le modèle proposé par la
revue, qui se chargera du formatage.

\section{Types de document}\label{types-de-document}

Ce modèle est prévu pour fonctionner avec le modèle Article d'EcoFoG en
LaTeX et produire des documents au format PDF. En changeant l'entête du
document, un fichier HTML est produit: décommenter

\begin{verbatim}
 bookdown::gitbook
\end{verbatim}

et commenter

\begin{verbatim}
bookdown::pdf_book:
\end{verbatim}

et les deux lignes suivantes qui en dépendent; compléter le nom du
fichier \emph{.bib} avec son extension sur la ligne
\texttt{bibliography} (LaTeX attend le nom du fichier sans extension
contrairement aux autres générateurs).

Pour produire un document Word, procéder de la même façon que pour un
fichier HTML en décommentant

\begin{verbatim}
bookdown::word_document2.
\end{verbatim}

Le document Word peut ensuite être modifié pour respecter les
instructions aux auteurs des revues : double interlignage, police, etc.

\subsection{Document HTML/PDF}\label{document-htmlpdf}

Pour un usage quotidien: le modèle HTML simple à utiliser, son format de
sortie est propre correct. Le résultat est bien plus clair que du code R
commenté.

\subsection{Document Word}\label{document-word}

Son contenu peut être mis en forme ou copié dans un modèle. Les styles
de texte standard sont ``First Paragraph'' et ``Corps de de texte''.

L'intérêt du format Word est de produire un manuscrit pour les revues
qui ne supportent pas \LaTeX. Le style bibliographique de la revue est
très probablement disponible au format \emph{.csl}, ce qui permet de
minimiser la préparation manuelle.

\subsection{Présentation Beamer}\label{presentation-beamer}

Utiliser le modèle \emph{Présentation EcoFoG} proposé par le package.

Le document produit est un diaporama Beamer respectant la charte
graphique 2018.

\subsection{Autres Modèles}\label{autres-modeles}

Le package rticle fournit des modèles d'articles (PLOS, PNAS, etc.). Le
package xaringan fournit un modèle de présentation HTML 5.

Le modèle Ouvrage du package EcoFoG permet d'écrire des livres.

La dernière ligne du modèle (bout de code R) doit être conservée pour
afficher le titre \emph{References} (à traduire éventuellement dans la
langue du document) au format HTML. Le titre de niveau 1
\emph{Références} doit être ajouté manuellement aux fichiers Word.

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\bibliographystyle{mee}
\bibliography{references}

%----------------------------------------------------------------------------------------

\end{document}

### 2018-02-06 Manuscript1.Rmd
Manuscript update

@@
 source("Scripts/DrawingTools.R")
-source("Scripts/GithubPages.R")
+#source("Scripts/GithubPages.R")
 ```

### 2018-02-06 Manuscript1.Rmd
test
set up affiliations

@@
   - name: Bruno Hérault
-    authsuperscript: 2
+    authsuperscript: 2 3
 affiliation:
@@
   - affsuperscript: 2
-    dptuniv: Department of Ecology, University of Edimburgh
+    dptuniv: Cirad, Univ montpellier, UR Forests & Societies.
     address: >
-      Street address,
-      Zip code,
-      Country.
+      Montpellier,
+      France.
+  - affsuperscript: 3
+    dptuniv: INPHB, Institut National Polytechnique Félix Houphouet-Boigny
+    address: >
+      Yamoussoukro,
+      Ivory Coast.
 corrauthor:
@@
 # Introduction
-The large areas covered with tropical forests worldwide hold crucial commercial, social and cultural value. 
+The large areas covered with tropical forests worldwide hold crucial economic, social and cultural value. 
 They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate local climate and support nutrients, water and carbon cycles and ensure cultural and human well-being. 

### 2018-02-06 Manuscript1.Rmd
.

@@
 ---
-title: "Post-Disturbance Diversity Trajectories in a Neotropical Forest"
+title: "Post-Disturbance Tree Community Trajectories in a Neotropical Forest"
 author:
@@
 The large areas covered with tropical forests worldwide hold crucial economic, social and cultural value. 
-They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate local climate and support nutrients, water and carbon cycles and ensure cultural and human well-being. 
-The demand for forests products and the shrink of unmanaged areas increased the human pressure on forests and currently threaten the maintenance of their structure, composition and function. 
+They provide not only wood but also multiple non-timber forest products, shelter a diversified fauna, regulate local climate and support carbon, water and many nutrient cycles and ensure cultural and human well-being.
+The increasing demand for forests products and the resulting shrink of unmanaged areas increase the pressure on the remaining intact forest landscapes as well as threaten the maintenance of their structure, composition and functionning in the long term.
 The increasing demand for forests products, land-use changes and other human disturbances along with substantial climatic changes make a crucial issue of understanding the response of forests dynamics to such disturbance regimes [@Gibson2011a; @Morales-Hidalgo2015].

### 2018-02-07 Manuscript1.Rmd
Files update

@@
 ```{r Options, echo=FALSE, message=FALSE}
-knitr::opts_chunk$set(cache=TRUE, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
+knitr::opts_chunk$set(cache=F, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
 # Installation des packages si nécessaire et chargement

### 2018-02-08 Manuscript1.Rmd
Manuscript update

@@
 # Discussion
--	The disturbance significantly impact the functional and taxonomic composition and diversity of the stands for more than 30 years 
+-	The disturbance significantly impact the functional and taxonomic composition and diversity of the stands for more than 30 years. 
+< Question Hubbel vs. Molino: IDH or not, if the composition is affected the changes are oriented then the IDH is valid. >
 -	Important differences among plots of same treatments: possible impact of initial diversity composition, local abiotic parameters (topography, soils, surrounding stands).

### 2018-02-09 WholePlotTrajectories.Rmd
html update

--- 2018-02-08 Manuscript1.Rmd
@@
 The response of trees biological diversity however has been sparsely studied, although this diversity is a major determinant of forests productivity, stability and functioning as it ensures interspecific complementarity, enhances the use of resources and the store of nutrients, and mitigates the prejudices of diseases and herbivory [@Tilman2014].
-Maintain the integrity of forests functioning thus requires highlighting the impacts and underlying mechanisms of disturbance on trees diversity and question the extent of their resilience. 
-Disturbance alter the biotic and abiotic parameters of the ecosystem in modifying the light availability, the heat and water fluxes, the plants productivity, etc. 
+Maintain the integrity of forests functioning thus requires highlighting the impacts and underlying mechanisms of disturbance on trees diversity and question the extent of forests resilience.
+
+Disturbance is defined as rapid changes in the amount and allocation of resources such as variation in the availability of light, the fluxes of heat and water, etc.
+It influences the environment and competition among species and combineds with environmental filters that select the pool of species fitting the environmental context to shape the composition and structure of communities in the long term: understanding ecosystems requires understanding their disturbance history [@White2001].
 Although those changes proved impacting the proportion and identity of species, negligible or even positive impacts were highlighted when considering the community as a whole, in accordance with the intermediate disturbance hypothesis (IDH).
 The IDH that predicts a maximized species diversity at intermediate level of disturbance, however, mainly rely on the study of species richness which gives limited or even misleading information on forests recovery and functioning.
-To have a detailed understanding of forests response to disturbance it is essential to identify the subsequent impact on species identity per se, which underlies any diversity change, as well as on richness, equitability and functional profile of the whole community.
+The mechanisms under which disturbance shapes communities diversity and structure involve the foster of certain species either stachastically, through the recruitment and dispersal limitation [@Hubbell1999] or determinisitcally through the selection of suited ecological types.
+Such deterministic selection is though to be driven by the trade-off between rapid growth and resources acquisition, therefore "acquisitive strategies" and slow growth with scarce resources [].
+A way to disentangle the mechanisms of disturbance dynamics would then to identify the shifts in communities functional types definde by the acquisitive or conservative functional traits and by the reproductive strategy through seed and early stage traits [@TerSteege2001;@Reich2014;@Herault2011].
+Relevant functional types would be defined with the combination of resource acquisition traits: in that respect leaf and stem traits proved relevant in a large litterature [@Chave2009b;@Wright2004],and life history traits revealing the growth and reproductive strategy [@Westoby1998;@Herault2011]
+To have a detailed understanding of forests response to disturbance it is essential to identify the subsequent impact on species identity per se, which underlies any diversity change, as well as on richness, equitability and functional characteristics of the whole community.
+
 Such combination of composition and diversity in browsing all facets of stands assemblage would highlight their resilience to perturbation, and its determinant and underlying ecological processes.
@@
 # Discussion
+
 -	The disturbance significantly impact the functional and taxonomic composition and diversity of the stands for more than 30 years. 
-< Question Hubbel vs. Molino: IDH or not, if the composition is affected the changes are oriented then the IDH is valid. >
+> Question Hubbel vs. Molino: IDH or not, if the composition is affected the changes are oriented then the IDH is valid.
+> Recovery of the composition, as in storm regime [@Burslem2000] the recovery is probably due to the coexistence of species.
+It is consistent with the seed characteristics that have high mass and then short dispersion distance
+
 -	Important differences among plots of same treatments: possible impact of initial diversity composition, local abiotic parameters (topography, soils, surrounding stands).
 Impacts depend on the disturbance intensity, but further parameters may be involved.
+
 -	The shift in functional composition differs from the null model: biotic and abiotic filters are at stake but only tangible from 10/15 years after logging. 
-Discussion with Cannon, 1998: they found no different process than rarefaction for 8 years after logging, but the time might have been too short-> divergence due to recruitment and not rarefaction after random species disappearance.
+> Discussion with Cannon, 1998: they found no different process than rarefaction for 8 years after logging, but the time might have been too short -> divergence due to recruitment and not rarefaction after random species disappearance.
 Same conclusions otherwise.
+> hubbel stop the observation 13 years after disturbance, which is very short [@Hubbell1999;@Sheil2003], and according to our observatoin this is not the peak period yet
+
 -	Disturbance involves new filters for species settling in the stands, probably according to their functional strategy.
@@
 Ecosystems function rely on abundant species: changes in stands functioning is then expected.
-(Flores, 2006  for differential sapling success after disturbance).
+> (Flores, 2006  for differential sapling success after disturbance)
+> typical pattern of the IDH [@Sheil2003] with an initial rise of the diversity followed by a decline through compoetitive domination.
+IDH is a regression to previous successional states and the more intense the logging the further we go back.The overall diversit increases because the IDH is patchy
+> Acquisitive like functional traits associated with higher diversity and disturbance dynamics [@TerSteege2001].
+Increase in functional diversity correlated with shifts toward more disturbance-like functional traits.
+
 -	If the recovery is not achieved 30 years after disturbance, it is on the way to.
@@
 however this is the impact of a single disturbance event: the repetition may have different impact (@Olander2005 demonstrated the impact of logging on soils nutrient and compaction)
+> the IDH involves a unimodal response to disturbance [@Sheil2003] and the functional strategy follows this pattern, however the composition still didn't recover,neither did the taxo diversity.
+Functioning recovering faster than taxo?
+
 -	To go further we should focus on recruitment which would also highlight the mechanisms underlying these shifts and dynamics and explicit the environmental filter.
+> Discussion about the composition impact: do we regenerate SB and species from old-growth forest (see [@Sheil2003] for discussion).
 

### 2018-02-09 WholePlotTrajectories.Rmd
html update

@@
 The mechanisms under which disturbance shapes communities diversity and structure involve the foster of certain species either stachastically, through the recruitment and dispersal limitation [@Hubbell1999] or determinisitcally through the selection of suited ecological types.
-Such deterministic selection is though to be driven by the trade-off between rapid growth and resources acquisition, therefore "acquisitive strategies" and slow growth with scarce resources [].
+Such deterministic selection is though to be driven by the trade-off between rapid growth and resources acquisition, therefore "acquisitive strategies" and slow growth with scarce resources [@Sheil2003].
 A way to disentangle the mechanisms of disturbance dynamics would then to identify the shifts in communities functional types definde by the acquisitive or conservative functional traits and by the reproductive strategy through seed and early stage traits [@TerSteege2001;@Reich2014;@Herault2011].
@@
 > Question Hubbel vs. Molino: IDH or not, if the composition is affected the changes are oriented then the IDH is valid.
-> Recovery of the composition, as in storm regime [@Burslem2000] the recovery is probably due to the coexistence of species.
-It is consistent with the seed characteristics that have high mass and then short dispersion distance
+> Recovery of the composition, probably not much external arrivals (seed mass characteristics as in storm regime [@Burslem2000] the recovery is probably due to the coexistence of species.
+Sometimes recovery is only due to resprouting and regrowth of species in place [@Zimmerman1994]
+> Higher seed mass = more sensitive to fragmentation because of the dispersers ecology [@Cramer2007]
+> Hubbell: identical species abundance repartition so the recruitment is random. This is the natural regime: we also proved that.The disturbance changes the natural disturbance regime?
 
@@
 Impacts depend on the disturbance intensity, but further parameters may be involved.
+> Composition can change the environmental recovery: resprouting and seedling dynamics are highly dependant of taxonomic composition [@Paciorek2000]
+> The recovery dynamics largely depend on the past disturbance regime also [@Burselm2000]
 

### 2018-02-15 Manuscript1.Rmd
Manuscript & Graph update

@@
 The response of trees biological diversity however has been sparsely studied, although this diversity is a major determinant of forests productivity, stability and functioning as it ensures interspecific complementarity, enhances the use of resources and the store of nutrients, and mitigates the prejudices of diseases and herbivory [@Tilman2014].
-Maintain the integrity of forests functioning thus requires highlighting the impacts and underlying mechanisms of disturbance on trees diversity and question the extent of their resilience. 
-Disturbance alter the biotic and abiotic parameters of the ecosystem in modifying the light availability, the heat and water fluxes, the plants productivity, etc. 
+Maintain the integrity of forests functioning thus requires highlighting the impacts and underlying mechanisms of disturbance on trees diversity and question the extent of forests resilience.
+
+Disturbance is defined as rapid changes in the amount and allocation of resources such as variation in the availability of light, the fluxes of heat and water, etc. 
+It influences the environment and competition among species and combineds with environmental filters that select the pool of species fitting the environmental context to shape the composition and structure of communities in the long term: understanding ecosystems requires understanding their disturbance history [@White2001].
 Although those changes proved impacting the proportion and identity of species, negligible or even positive impacts were highlighted when considering the community as a whole, in accordance with the intermediate disturbance hypothesis (IDH).
 The IDH that predicts a maximized species diversity at intermediate level of disturbance, however, mainly rely on the study of species richness which gives limited or even misleading information on forests recovery and functioning.
-To have a detailed understanding of forests response to disturbance it is essential to identify the subsequent impact on species identity per se, which underlies any diversity change, as well as on richness, equitability and functional profile of the whole community.
+The mechanisms under which disturbance shapes communities diversity and structure involve the foster of certain species either stachastically, through the recruitment and dispersal limitation [@Hubbell1999] or determinisitcally through the selection of suited ecological types.
+Such deterministic selection is though to be driven by the trade-off between rapid growth and resources acquisition, therefore "acquisitive strategies" and slow growth with scarce resources [].
+A way to disentangle the mechanisms of disturbance dynamics would then to identify the shifts in communities functional types definde by the acquisitive or conservative functional traits and by the reproductive strategy through seed and early stage traits [@TerSteege2001;@Reich2014;@Herault2011].
+Relevant functional types would be defined with the combination of resource acquisition traits: in that respect leaf and stem traits proved relevant in a large litterature [@Chave2009b;@Wright2004],and life history traits revealing the growth and reproductive strategy [@Westoby1998;@Herault2011]
+Beyond diversity indices it is essential to analyse changes in taxonomic composition so that the impact on species per se and their particular impact on ecosystem can be followed [@╔Lavorel2002;@Bellwood;@Baraloto2012] along with the ecological rules shaping communities assembly and abundance distribution [@Magurran??].
+
 Such combination of composition and diversity in browsing all facets of stands assemblage would highlight their resilience to perturbation, and its determinant and underlying ecological processes.
@@
 Ecosystems function rely on abundant species: changes in stands functioning is then expected.
-(Flores, 2006  for differential sapling success after disturbance).
+> (Flores, 2006  for differential sapling success after disturbance)
+> typical pattern of the IDH [@Sheil2003] with an initial rise of the diversity followed by a decline through compoetitive domination.
+IDH is a regression to previous successional states and the more intense the logging the further we go back.The overall diversit increases because the IDH is patchy
+> Acquisitive like functional traits associated with higher diversity and disturbance dynamics [@TerSteege2001].
+Increase in functional diversity correlated with shifts toward more disturbance-like functional traits.
+
+- Functional and taxonomic trajectories don't follow the same trajectories.
+Functional redundancy [biblio to be found] and contrasting taxonomic and functional response to disturbane[@Baraloto2012 and ref therein].
+Ecosystem functioning is altered differently by disturbance than communities composition.
+Such difference could be explained by competitive exclusion at stake in control plots and not after disturbance.
+Baraloto showed that evenness of functional diversity was affected: resilience may be altered for future disturbance
+
 -	If the recovery is not achieved 30 years after disturbance, it is on the way to.

### 2018-02-15 WholePlotTrajectories.Rmd
Manuscript & Graph update

@@
     authsuperscript: 1*
-  - name: Eric Marcon
+  - name: Eric MARCON
     authsuperscript: 1
-  - name: Bruno Hérault
+  - name: Bruno HERAULT
     authsuperscript: 2 3
@@
 
+The long-term survey of tropical forest plots after disturbance demonstrated a significant impact on stands composition and abundance distribution lasting for more than 30 years.
+Disturbance induced environmental filters favoring acquisitive functional types and enhancing communities evenness.
+Plots response was highly variable and might depend on its initial conditions.
+
 -	The disturbance significantly impact the functional and taxonomic composition and diversity of the stands for more than 30 years. 

### 2018-02-15 Scripts/DrawingTools.R

@@ -166,7 +166,8 @@ for(q in 1:3){
 mtext("Years since disturbance",side=1,adj=1,cex=0.8,line=-2,outer=TRUE)
 mtext("Equivalent diversity",side=2,padj=1,cex=0.8,line=-1,outer=TRUE)}
 
-FunTraj<-function(CompFun){plot(as.numeric(colnames(CompFun[[1]])),CompFun[[1]][1,,"0.5"],type="n",xaxt="n",
+FunTraj<-function(CompFun){
+  plot(as.numeric(colnames(CompFun[[1]])),CompFun[[1]][1,,"0.5"],type="n",xaxt="n",
      xlab="",ylab="",ylim=c(min(unlist(CompFun),na.rm=T),max(unlist(CompFun),na.rm=T)),cex.axis=0.7)
 axis(1,at=as.character(seq(5,33,5)),labels=TRUE,cex.axis=0.7)  
 mtext("Rao diversity",3,adj=0,line=1,cex=1.5) 

### 2018-02-16 Manuscript1.Rmd
Manuscript & graph update

--- 2018-02-07 Manuscript1.Rmd
@@
 ```{r Options, echo=FALSE, message=FALSE}
-knitr::opts_chunk$set(cache=F, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
+knitr::opts_chunk$set(cache=T, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
 # Installation des packages si nécessaire et chargement
@@
 The response of trees biological diversity however has been sparsely studied, although this diversity is a major determinant of forests productivity, stability and functioning as it ensures interspecific complementarity, enhances the use of resources and the store of nutrients, and mitigates the prejudices of diseases and herbivory [@Tilman2014].
-Maintain the integrity of forests functioning thus requires highlighting the impacts and underlying mechanisms of disturbance on trees diversity and question the extent of their resilience. 
-Disturbance alter the biotic and abiotic parameters of the ecosystem in modifying the light availability, the heat and water fluxes, the plants productivity, etc. 
+Maintain the integrity of forests functioning thus requires highlighting the impacts and underlying mechanisms of disturbance on trees diversity and question the extent of forests resilience.
+
+Disturbance is defined as rapid changes in the amount and allocation of resources such as variation in the availability of light, the fluxes of heat and water, etc. 
+It influences the environment and competition among species and combineds with environmental filters that select the pool of species fitting the environmental context to shape the composition and structure of communities in the long term: understanding ecosystems requires understanding their disturbance history [@White2001].
 Although those changes proved impacting the proportion and identity of species, negligible or even positive impacts were highlighted when considering the community as a whole, in accordance with the intermediate disturbance hypothesis (IDH).
 The IDH that predicts a maximized species diversity at intermediate level of disturbance, however, mainly rely on the study of species richness which gives limited or even misleading information on forests recovery and functioning.
-To have a detailed understanding of forests response to disturbance it is essential to identify the subsequent impact on species identity per se, which underlies any diversity change, as well as on richness, equitability and functional profile of the whole community.
+The mechanisms under which disturbance shapes communities diversity and structure involve the foster of certain species either stochastically, through the recruitment and dispersal limitation [@Hubbell1999] or determinisitcally through the selection of suited ecological types.
+Disturbance indeed in the context of old growth, stable forests where competitive exclusion favors best competitor species causes a shift toward disturbance resistant species with rapid growth and efficient resources acquisition [@Haddad2008].
+This is illustrated by a shift from slow growth with scarce resources and high competitive ability, "conservative strategies", to fast growth with prolific resourcesn, "acquisitive strategies".
+A way to disentangle the mechanisms of disturbance dynamics would then to identify the shifts in communities functional types definde by the acquisitive or conservative functional traits and by the reproductive strategy through seed and early stage traits [@TerSteege2001;@Reich2014;@Herault2011].
+Relevant functional types would be defined with the combination of resource acquisition traits: in that respect leaf and stem traits proved relevant in a large litterature [@Chave2009b;@Wright2004],and life history traits revealing the growth and reproductive strategy [@Westoby1998;@Herault2011]
+Beyond diversity indices it is essential to analyse changes in taxonomic composition so that the impact on species per se and their particular impact on ecosystem can be followed [@╔Lavorel2002;@Bellwood2006;@Baraloto2012a] along with the ecological rules shaping communities assembly and abundance distribution [@Magurran1988].
+
 Such combination of composition and diversity in browsing all facets of stands assemblage would highlight their resilience to perturbation, and its determinant and underlying ecological processes.
@@
 
-```{r Fig3, echo=FALSE,fig.cap="Trajectories of the CWM until 30 years after disturbance of 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness).
+```{r Fig3, echo=FALSE, fig.cap="Trajectories of the CWM until 30 years after disturbance of 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness).
 The CWM were measured at plot level and averaged among plots of the same treatment (green for control, blue for T1,orange for T2 and red for T3).
 The average CWM per treatment are drawn (solid line) drawn along with the 0.025 and 0.975 percentile (gray envelope) observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
+
 ```
 
+```{r Fig4, echo=FALSE, out.width='100%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions until 30 years after disturbance.
+Plot treatments are indicated with line colors in green for control, blue for T1, orange for T2 and red for T3.
+The proportion reported are the median values (solid line) and the 0.025 and 0.975 percentile (gray envelope) observed for 50 iteration of the taxonomic uncertainty propagation.
+only species which seed mass values were reported are considered (*i.e* 414 out of 696 species recorded in total)."}
+load("DB/Smass_ClassProp")
+
+SeedMassProp(Smass)
+
+```
 ### Disturbance impact on diversity
 
-*Taxonomic diversity* _ Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1994 inventories \autoref{Fig4}. 
+*Taxonomic diversity* _ Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1994 inventories \autoref{Fig5}. 
 *For the three indices the OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and a covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01).*
 
-```{r Fig4, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory, 10 years after disturbance, until 30 years after disturbance for plots **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversity.
+```{r Fig5, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory, 10 years after disturbance, until 30 years after disturbance for plots **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversity.
 Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
@@
 
-The evolution along time of the difference of Richness, Shannon and Simpson diversity compared to reference inventories in 1989 remained even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase, positively correlated with the disturbance intensity \autoref{Fig4}.
-The richness stabilized for all disturbed plots from 20 years after disturbance \autoref{Fig4}. 
-The Shannon and Simpson diversity also stabilized 20 years after disturbance for T1 and T2 plots \autoref{Fig4} whereas for 2 out of 3 plots T3 the Shannon and Simpson diversity continued increasing until 30 years after disturbance, illustrating a significant increase in communities’ evenness.
+The evolution along time of the difference of Richness, Shannon and Simpson diversity compared to reference inventories in 1989 remained even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase, positively correlated with the disturbance intensity \autoref{Fig5}.
+The richness stabilized for all disturbed plots from 20 years after disturbance \autoref{Fig5}. 
+The Shannon and Simpson diversity also stabilized 20 years after disturbance for T1 and T2 plots \autoref{Fig5} whereas for 2 out of 3 plots T3 the Shannon and Simpson diversity continued increasing until 30 years after disturbance, illustrating a significant increase in communities’ evenness.
 
 *Functional  diversity* _ The evolution along time of the Rao functinal diversity compared to reference inventories in 1989 remained even all along the survey for the control plots. 
-All disturbed plots however showed an increase of the functional diversity that was positively correlated to the disturbance intensity \autoref{Fig5}.
+All disturbed plots however showed an increase of the functional diversity that was positively correlated to the disturbance intensity \autoref{Fig6}.
 30 after disturbance the diversity of disturbed plots remained significantly higher than this of control plots, thus whenever the initial disturbance intensity.
@@
 
-```{r Fig5, fig.asp=1/1, echo=FALSE,fig.cap="Trajectories of the plots Rao functional diversity from10 to 30 years after disturbance.
+```{r Fig6, fig.asp=1/1, echo=FALSE,fig.cap="Trajectories of the plots Rao functional diversity from10 to 30 years after disturbance.
 Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
@@
 # Discussion
--	The disturbance significantly impact the functional and taxonomic composition and diversity of the stands for more than 30 years 
+
+Our study underlined the significant but contrasting impact of disturbance on tropical forests taxonomic and functional diversity.
+Specifically we unveiled an important decoupling between taxonomic evenness and functional characteristic of the communities, and by extension ecosystem functioning.
+This highlighted the functional overlap between species in tropical forests,typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.
+
+In line with the assumtions of the IDH our 30 years survey confirmed the significant increase of communities evenness and functional richness after disturbance.
+As it was demonstrated for the conditions 10 years after disturbance [@Molino2001] the directional changes in communities composition reflected environmental filtering after disturbance which would support the IDH as demonstrated in [@Molino2001] rather than pure stochastic dynamics as suggested in [@Hubbell1998].
+The trajectories shaping a unimodal response to disturbance suggested the recovery of communities initial state as it was already suggested by the previous studies from the same site 10 years [@Molino2001] and 20 years [@Baraloto2012a] after disturbance.
+Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories. 
+This recovery corroborate previous post-disturbance dynamics observed in a Bornean rainforest [@Cannon1998] where at large scale the richness was preserved although it was diminished locally or in Solomon islands where stands demonstrated a fast recovery and identical composition [@Burselm2000].
+
+Although recovery seemed on its way the trajectories revealed significant changes in the ecological rules shaping communities over the 30 years of survey.
+Disturbance is thought to shift exclusive competition favoring competitive species to environmental filtering selecting acquisitive and disturbance resistant species [@TerSteege2001;@Flores2006;@Haddad2008]
+The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
+In particular this is demonstrated by the evolution of seed mass classes proportions revealing a clear increase in small seeded and then large dispersive species.
+Faster than for taxonomic composition and diversity though, functional characteristic returned to their initial values and the values observed for control plots.
+This suggested some functional overlap in the pre-disturbance condition (*i.e* the usual tropical forests conditions). 
+Such overlap and maybe functional versatility was already observed in highly diverse ecosystems [@Bellwood2006] and sugegsted as one mechanisms for high diversity maintenance.
+Decoupled taxonomic and functional trajectories questions to what extent communities recovery is completed: ecosystem functioning might be quickly recoverd, and taxonomic composition following suit although along for much longer, but the resilience of the community itself might be altered as functional overlap themselves would be lost.
+Along with probably other environmental modifications (like soils nutrient cycling and compaction [@Olander2005]) with alteration of communities resilience makes us expect different trajectories after subsequent disturbance.
+As disturbance seemed to update communities toward more disturbance resistant ones [@Haddad2008] it is possible that eventually the disturbance regie observed has set another permanent condition, which would only be more diverse if good competitors, poor disturbance resitant survive.
+> (Flores, 2006  for differential sapling success after disturbance)
+
 -	Important differences among plots of same treatments: possible impact of initial diversity composition, local abiotic parameters (topography, soils, surrounding stands).
 Impacts depend on the disturbance intensity, but further parameters may be involved.
--	The shift in functional composition differs from the null model: biotic and abiotic filters are at stake but only tangible from 10/15 years after logging.
-Discussion with Cannon, 1998: they found no different process than rarefaction for 8 years after logging, but the time might have been too short-> divergence due to recruitment and not rarefaction after random species disappearance.
-Same conclusions otherwise.
--	Disturbance involves new filters for species settling in the stands, probably according to their functional strategy.
-Disturbance generates a functional shift of communities toward acquisitive strategies.
-Ecosystems function rely on abundant species: changes in stands functioning is then expected.
-(Flores, 2006  for differential sapling success after disturbance).
--	If the recovery is not achieved 30 years after disturbance, it is on the way to.
-Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories.
-however this is the impact of a single disturbance event: the repetition may have different impact (@Olander2005 demonstrated the impact of logging on soils nutrient and compaction)
--	To go further we should focus on recruitment which would also highlight the mechanisms underlying these shifts and dynamics and explicit the environmental filter.
+
+-	To go beyond whole community diversity and highlight the ecological mechanisms behind communities assembly and repond to the question of their resilience we should focus on recruitment.
+It would also refine the role of environmental filters shaping communities assembly.
 

### 2018-02-16 Scripts/DrawingTools.R

@@ -184,3 +184,22 @@ invisible(lapply(1:length(CompFun),function(t){
             col=rgb(0,0,0,alpha=0.1),border=NA)
   }))
 }))}
+
+SeedMassProp<-function(SeedMass){
+  par(mfrow=c(1,5),mar=c(1,1,2,1),oma=c(2,1,4,1),no.readonly = T)
+  invisible(lapply(1:5,function(clas){
+    Smass_class<-lapply(SeedMass,function(pl){return(t(pl[clas,,]))})
+    plot(colnames(Smass_class[[1]]),Smass_class[[1]]["0.5",], ylim=c(min(unlist(Smass_class)),max(unlist(Smass_class))),
+         type="n",xlab="years",ylab="")
+    mtext(paste("class",clas),3,cex=1,line=1)
+    invisible(lapply(1:4,function(tr){
+      invisible(lapply(treatments[[tr]],function(plo){
+        lines(colnames(Smass_class[[plo]]),Smass_class[[plo]]["0.5",],col=ColorsTr[tr],lwd=2)
+        polygon(c(colnames(Smass_class[[plo]]),rev(colnames(Smass_class[[plo]]))),
+                c(Smass_class[[plo]]["0.05",],rev(Smass_class[[plo]]["0.975",])),
+                col=rgb(0,0,0,alpha=0.1),border=NA)}))
+    }))
+  }))
+mtext("Seed mass, class proportions",line=2,adj=0,outer=TRUE)
+mtext("Years since disturbance",side=1,line=1,adj=1,cex=0.9,outer=TRUE)
+}

### 2018-02-19 WholePlotTrajectories.Rmd
Files update

--- 2018-02-09 WholePlotTrajectories.Rmd
@@
 The IDH that predicts a maximized species diversity at intermediate level of disturbance, however, mainly rely on the study of species richness which gives limited or even misleading information on forests recovery and functioning.
-The mechanisms under which disturbance shapes communities diversity and structure involve the foster of certain species either stachastically, through the recruitment and dispersal limitation [@Hubbell1999] or determinisitcally through the selection of suited ecological types.
-Such deterministic selection is though to be driven by the trade-off between rapid growth and resources acquisition, therefore "acquisitive strategies" and slow growth with scarce resources [].
+The mechanisms under which disturbance shapes communities diversity and structure involve the foster of certain species either stochastically, through the recruitment and dispersal limitation [@Hubbell1999] or determinisitcally through the selection of suited ecological types.
+Disturbance indeed in the context of old growth, stable forests where competitive exclusion favors best competitor species causes a shift toward disturbance resistant species with rapid growth and efficient resources acquisition [@Haddad2008].
+This is illustrated by a shift from slow growth with scarce resources and high competitive ability, "conservative strategies", to fast growth with prolific resourcesn, "acquisitive strategies".
 A way to disentangle the mechanisms of disturbance dynamics would then to identify the shifts in communities functional types definde by the acquisitive or conservative functional traits and by the reproductive strategy through seed and early stage traits [@TerSteege2001;@Reich2014;@Herault2011].
 Relevant functional types would be defined with the combination of resource acquisition traits: in that respect leaf and stem traits proved relevant in a large litterature [@Chave2009b;@Wright2004],and life history traits revealing the growth and reproductive strategy [@Westoby1998;@Herault2011]
-To have a detailed understanding of forests response to disturbance it is essential to identify the subsequent impact on species identity per se, which underlies any diversity change, as well as on richness, equitability and functional characteristics of the whole community.
+Beyond diversity indices it is essential to analyse changes in taxonomic composition so that the impact on species per se and their particular impact on ecosystem can be followed [@╔Lavorel2002;@Bellwood2006;@Baraloto2012a] along with the ecological rules shaping communities assembly and abundance distribution [@Magurran1988].
 
@@
 
--	The disturbance significantly impact the functional and taxonomic composition and diversity of the stands for more than 30 years. 
-> Question Hubbel vs. Molino: IDH or not, if the composition is affected the changes are oriented then the IDH is valid.
-> Recovery of the composition, as in storm regime [@Burslem2000] the recovery is probably due to the coexistence of species.
-It is consistent with the seed characteristics that have high mass and then short dispersion distance
+Our study underlined the significant but contrasting impact of disturbance on tropical forests taxonomic and functional diversity.
+Specifically we unveiled an important decoupling between taxonomic evenness and functional characteristic of the communities, and by extension ecosystem functioning.
+This highlighted the functional overlap between species in tropical forests,typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.
+
+In line with the assumtions of the IDH our 30 years survey confirmed the significant increase of communities evenness and functional richness after disturbance.
+As it was demonstrated for the conditions 10 years after disturbance [@Molino2001] the directional changes in communities composition reflected environmental filtering after disturbance which would support the IDH as demonstrated in [@Molino2001] rather than pure stochastic dynamics as suggested in [@Hubbell1998].
+The trajectories shaping a unimodal response to disturbance suggested the recovery of communities initial state as it was already suggested by the previous studies from the same site 10 years [@Molino2001] and 20 years [@Baraloto2012a] after disturbance.
+Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories. 
+This recovery corroborate previous post-disturbance dynamics observed in a Bornean rainforest [@Cannon1998] where at large scale the richness was preserved although it was diminished locally or in Solomon islands where stands demonstrated a fast recovery and identical composition [@Burselm2000].
+
+Although recovery seemed on its way the trajectories revealed significant changes in the ecological rules shaping communities over the 30 years of survey.
+Disturbance is thought to shift exclusive competition favoring competitive species to environmental filtering selecting acquisitive and disturbance resistant species [@TerSteege2001;@Flores2006;@Haddad2008]
+The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
+In particular this is demonstrated by the evolution of seed mass classes proportions revealing a clear increase in small seeded and then large dispersive species.
+Faster than for taxonomic composition and diversity though, functional characteristic returned to their initial values and the values observed for control plots.
+This suggested some functional overlap in the pre-disturbance condition (*i.e* the usual tropical forests conditions). 
+Such overlap and maybe functional versatility was already observed in highly diverse ecosystems [@Bellwood2006] and sugegsted as one mechanisms for high diversity maintenance.
+Decoupled taxonomic and functional trajectories questions to what extent communities recovery is completed: ecosystem functioning might be quickly recoverd, and taxonomic composition following suit although along for much longer, but the resilience of the community itself might be altered as functional overlap themselves would be lost.
+Along with probably other environmental modifications (like soils nutrient cycling and compaction [@Olander2005]) with alteration of communities resilience makes us expect different trajectories after subsequent disturbance.
+As disturbance seemed to update communities toward more disturbance resistant ones [@Haddad2008] it is possible that eventually the disturbance regie observed has set another permanent condition, which would only be more diverse if good competitors, poor disturbance resitant survive.
+> (Flores, 2006  for differential sapling success after disturbance)
 
@@
 
--	The shift in functional composition differs from the null model: biotic and abiotic filters are at stake but only tangible from 10/15 years after logging. 
-> Discussion with Cannon, 1998: they found no different process than rarefaction for 8 years after logging, but the time might have been too short -> divergence due to recruitment and not rarefaction after random species disappearance.
-Same conclusions otherwise.
-> hubbel stop the observation 13 years after disturbance, which is very short [@Hubbell1999;@Sheil2003], and according to our observatoin this is not the peak period yet
-
--	Disturbance involves new filters for species settling in the stands, probably according to their functional strategy.
-Disturbance generates a functional shift of communities toward acquisitive strategies.
-Ecosystems function rely on abundant species: changes in stands functioning is then expected.
-> (Flores, 2006  for differential sapling success after disturbance)
-> typical pattern of the IDH [@Sheil2003] with an initial rise of the diversity followed by a decline through compoetitive domination.
-IDH is a regression to previous successional states and the more intense the logging the further we go back.The overall diversit increases because the IDH is patchy
-> Acquisitive like functional traits associated with higher diversity and disturbance dynamics [@TerSteege2001].
-Increase in functional diversity correlated with shifts toward more disturbance-like functional traits.
-
--	If the recovery is not achieved 30 years after disturbance, it is on the way to.
-Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories.
-however this is the impact of a single disturbance event: the repetition may have different impact (@Olander2005 demonstrated the impact of logging on soils nutrient and compaction)
-> the IDH involves a unimodal response to disturbance [@Sheil2003] and the functional strategy follows this pattern, however the composition still didn't recover,neither did the taxo diversity.
-Functioning recovering faster than taxo?
-
--	To go further we should focus on recruitment which would also highlight the mechanisms underlying these shifts and dynamics and explicit the environmental filter.
-> Discussion about the composition impact: do we regenerate SB and species from old-growth forest (see [@Sheil2003] for discussion).
+-	To go beyond whole community diversity and highlight the ecological mechanisms behind communities assembly and repond to the question of their resilience we should focus on recruitment.
+It would also refine the role of environmental filters shaping communities assembly.
 

### 2018-02-19 WholePlotTrajectories.Rmd
html update

@@
 
-Our study underlined the significant but contrasting impact of disturbance on tropical forests taxonomic and functional diversity.
-Specifically we unveiled an important decoupling between taxonomic evenness and functional characteristic of the communities, and by extension ecosystem functioning.
+Our study underlined the significant impact of disturbance on tropical forests communities but a contrasting response of taxonomic and functional diversity.
+Specifically,  it unveiled the decoupling between trajectories of taxonomic evenness and their functional diversity and dominant characteristics.
+This decoupling confirms the independence for tropical forests between the processes grasped through the studied functional traits (primary productivity, water and nutrient cycling, etc.) and communities composition and evenness as it was suggested .
 This highlighted the functional overlap between species in tropical forests,typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.

### 2018-02-19 WholePlotTrajectories.Rmd
Manuscript update

@@
 Specifically,  it unveiled the decoupling between trajectories of taxonomic evenness and their functional diversity and dominant characteristics.
-This decoupling confirms the independence for tropical forests between the processes grasped through the studied functional traits (primary productivity, water and nutrient cycling, etc.) and communities composition and evenness as it was suggested .
+This decoupling confirms the "vegetation quantity effect" suggesting that communities functional characteristics are determined by dominant species rather that subtle changes in biodiversity [@reftobefound].
+This would result in a faster recovery of ecosystem processes rates than in communities ocmposition and diversity as it mas largely demonstrated for grasslands [@Tilman1997; @Mokany2008; @Mouillot2011 *in Lohbeck2015*] and recently proved for tropical forests [@Lohbeck2015].
 This highlighted the functional overlap between species in tropical forests,typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.
+Ecosystem processes graped through the set of functional traits retained here however do not assess communities resilience and future dynamics.
 
@@
 Along with probably other environmental modifications (like soils nutrient cycling and compaction [@Olander2005]) with alteration of communities resilience makes us expect different trajectories after subsequent disturbance.
-As disturbance seemed to update communities toward more disturbance resistant ones [@Haddad2008] it is possible that eventually the disturbance regie observed has set another permanent condition, which would only be more diverse if good competitors, poor disturbance resitant survive.
-> (Flores, 2006  for differential sapling success after disturbance)
+As disturbance seemed to update communities toward more disturbance resistant ones [@Haddad2008] it is possible that eventually the disturbance regime observed has set another permanent condition, which would only be more diverse if good competitors, poor disturbance resitant survive.
+We assume that such post-disturbance forest community might have the same functioning, as suggested by the functional trajectories, but composition displays a poor recovery, especially regarding species contingent to undisturbed forests or different life forms as lianas or epiphytes that have not been monitored here [@Martin2013].
+Although forests functioning and processes seemed likely to recover after the range of disturbance studied, the hampered recovery of communities composition might jeopardize the conservation of some species, therefore threatening the global biodiversity [@Gardner2007].
+Additionally to threats on particular species conservation the consequences of biodiversity changes on hyper-diverse ecosystems might reveal dramatic, specifically in impairing ecosystems resilience and resistance to further disturbance [@Trenbath1999;@Elmqvist2003;@Diaz2006].
 

### 2018-02-19 WholePlotTrajectories.Rmd
Manuscript update

--- 2018-02-19 WholePlotTrajectories.Rmd
@@
 ```{r Options, echo=FALSE, message=FALSE}
-knitr::opts_chunk$set(cache=T, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
+knitr::opts_chunk$set(cache=F, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
 # Installation des packages si nécessaire et chargement
@@
 Relevant functional types would be defined with the combination of resource acquisition traits: in that respect leaf and stem traits proved relevant in a large litterature [@Chave2009b;@Wright2004],and life history traits revealing the growth and reproductive strategy [@Westoby1998;@Herault2011]
-Beyond diversity indices it is essential to analyse changes in taxonomic composition so that the impact on species per se and their particular impact on ecosystem can be followed [@╔Lavorel2002;@Bellwood2006;@Baraloto2012a] along with the ecological rules shaping communities assembly and abundance distribution [@Magurran1988].
+Beyond diversity indices it is essential to analyse changes in taxonomic composition so that the impact on species per se and their particular impact on ecosystem can be followed [@Lavorel2002;@Bellwood2006;@Baraloto2012a] along with the ecological rules shaping communities assembly and abundance distribution [@Magurran1988].
+> See Garnier 2001a for the description of traits functional impact
+
+Since the functional traits chosen showed low intraspecific compared to interspecific variability [@Garnier2001a;@Westoby2006a] it its valid to assess the functional trajectories of communities through the functional values observed or the species assembly.
 
@@
 Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
-The experiment corresponds to a network of twelve 6.25ha plots that have undergone a gradient of three logging, thinning and fuelwood treatments.
+The experiment corresponds to a network of twelve 6.25ha plots that have undergone a gradient of three logging, thinning and fuelwood treatments \@ref(tab:Tab1).
 Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
@@
 
-Table 1: Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
----------------------------------------------------------------------------------------------------------------------------------
+Table : (\#tab:Tab1) Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
+
+------------------------------------------------------------------------------------------------------------------------
                           ***Timber***              ***Thinning***          ***Fuelwood***           **%AGB lost**
 
------------------    ------------------------    ------------------------    ------------------------    ------------------------
+------------------  ------------------------ ------------------------- ------------------------ ------------------------
     ***Control***                                                                                          0
 
------------------    ------------------------    ------------------------    ------------------------    ------------------------
          ***T1***    DBH ≥ 50 cm,                                                                                       
@@
 
------------------    ------------------------    ------------------------    ------------------------    ------------------------
          ***T2***    DBH ≥ 50 cm,              DBH ≥ 40 cm,                                                                 
@@
 
------------------    ------------------------    ------------------------    ------------------------    ------------------------
          ***T3***    DBH ≥ 50 cm,              DBH ≥ 50 cm,               40 cm ≤ DBH ≤ 50 cm                               
-                      commercial species,         non-valuable species,        non-valuable species,           [33%-56%]         
+                     commercial species,       non-valuable species,      non-valuable species,         [35%-56%]         
                      ~ 10 trees/ha             ~ 15 trees/ha              ~ 15 trees/ha                                     
-
----------------------------------------------------------------------------------------------------------------------------------
+------------------------------------------------------------------------------------------------------------------------
 
 ### Inventories protocol and dataset correction
+
 The study site corresponds to a tropical rain forest with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae. 
@@
 The functional metrics used a set of 6 functional traits values representing traits of the leaf and wood economics and life history traits.
-The trait database encompassed 4 traits of the leaves: leaves thickness (*L_thick*), toughness (*L_tough*), total chlorophyll content (*LCC*) and specific leaf area (leaf area per unit dry mass; *SLA*), 2 traits of the stem: wood specific gravity (*SWG*) and bark thickness (*B_thick)* and two life-history traits: maximum specific height (*Hmax*) and the seed mass (*Smass*).
+The trait database encompassed 4 traits of the leaves: leaves thickness, toughness, total chlorophyll content and specific leaf area (leaf area per unit dry mass), 2 traits of the stem: wood specific gravity and bark thickness and two life-history traits: maximum specific height and the seed mass.
 Traits database came from the BRIDGE project (http://www.ecofog.gf/Bridge/index.html) where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou.
@@
 *Composition trajectories* _ Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in a two-dimensional space based on NMDS analysis.
-Classifications were based on either abundance-based Bray-Curtis \autoref{Fig1} or incidence-based Jaccard dissimilarity (data not shown).
+Classifications were based on either abundance-based Bray-Curtis \@ref(fig:Fig1) or incidence-based Jaccard dissimilarity (data not shown).
 
@@
 *The distance from the initial condition was significantly time-dependent (P<0.01).* 
-The taxonomic composition remained significantly different from the 1989 reference inventory until 30 years for all disturbed plots \autoref{Fig1}. 
-From 20 years after disturbance, though, the distancing decreased which suggested a return toward the reference 1989 inventory and its resilience to disturbance \autoref{Fig1},\autoref{Fig2}. 
-The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance \autoref{Fig2}.
+The taxonomic composition remained significantly different from the 1989 reference inventory until 30 years for all disturbed plots \@ref(fig:Fig1). 
+From 20 years after disturbance, though, the distancing decreased which suggested a return toward the reference 1989 inventory and its resilience to disturbance \@ref(fig:Fig1),\@ref(fig:Fig2). 
+The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance \@ref(fig:Fig2).
 *The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).*
@@
 
-*Traits community weighted means* _ The trajectories of the community weighted means (CWM) of the 8 functional traits were measured at plot level \autoref{Fig3}.
+*Traits community weighted means* _ The trajectories of the community weighted means (CWM) of the 8 functional traits were measured at plot level \@ref(fig:Fig3).
 **To be coming**
@@
 ```
+
 ### Disturbance impact on diversity
 
-*Taxonomic diversity* _ Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1994 inventories \autoref{Fig5}. 
+*Taxonomic diversity* _ Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1994 inventories \@ref(fig:Fig5). 
 *For the three indices the OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and a covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01).*
@@
 
-The evolution along time of the difference of Richness, Shannon and Simpson diversity compared to reference inventories in 1989 remained even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase, positively correlated with the disturbance intensity \autoref{Fig5}.
-The richness stabilized for all disturbed plots from 20 years after disturbance \autoref{Fig5}. 
-The Shannon and Simpson diversity also stabilized 20 years after disturbance for T1 and T2 plots \autoref{Fig5} whereas for 2 out of 3 plots T3 the Shannon and Simpson diversity continued increasing until 30 years after disturbance, illustrating a significant increase in communities’ evenness.
+The evolution along time of the difference of Richness, Shannon and Simpson diversity compared to reference inventories in 1989 remained even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase, positively correlated with the disturbance intensity \@ref(fig:Fig5).
+The richness stabilized for all disturbed plots from 20 years after disturbance \@ref(fig:Fig5). 
+The Shannon and Simpson diversity also stabilized 20 years after disturbance for T1 and T2 plots \@ref(fig:Fig5) whereas for 2 out of 3 plots T3 the Shannon and Simpson diversity continued increasing until 30 years after disturbance, illustrating a significant increase in communities’ evenness.
 
 *Functional  diversity* _ The evolution along time of the Rao functinal diversity compared to reference inventories in 1989 remained even all along the survey for the control plots. 
-All disturbed plots however showed an increase of the functional diversity that was positively correlated to the disturbance intensity \autoref{Fig6}.
+All disturbed plots however showed an increase of the functional diversity that was positively correlated to the disturbance intensity \@ref(fig:Fig6).
 30 after disturbance the diversity of disturbed plots remained significantly higher than this of control plots, thus whenever the initial disturbance intensity.
@@
 
-Our study underlined the significant but contrasting impact of disturbance on tropical forests taxonomic and functional diversity.
-Specifically we unveiled an important decoupling between taxonomic evenness and functional characteristic of the communities, and by extension ecosystem functioning.
+Our study showed the significant impact of disturbance on tropical forests communities.
+The subsequent diversity trajectories confirmed the intermediate disturbance hypothesis debated for tropical forests through their correlation with disturbance intensity.
+Besides it revealed the contrasting response of taxonomic and functional characteristics, specifically the decoupling between communities taxonomic evenness and their functional diversity and dominant functional traits values.
+The long-term disturbance trajectories observed highlighted the unachieved but consistent recovery of communities assembly for the lowest disturbance intensity but questioned it after higher disturbance.
+
+In line with the "vegetation quantity effect" suggesting that communities functional characteristics depends on dominant species rather that subtle changes in biodiversity [@reftobefound].
+This would result in a faster recovery of ecosystem processes rates than in communities ocmposition and diversity as it mas largely demonstrated for grasslands [@Tilman1997; @Mouillot2011] and recently proved for tropical forests [@Lohbeck2015].
 This highlighted the functional overlap between species in tropical forests,typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.
+Ecosystem processes graped through the set of functional traits retained here however do not assess communities resilience and future dynamics.
 
-In line with the assumtions of the IDH our 30 years survey confirmed the significant increase of communities evenness and functional richness after disturbance.
-As it was demonstrated for the conditions 10 years after disturbance [@Molino2001] the directional changes in communities composition reflected environmental filtering after disturbance which would support the IDH as demonstrated in [@Molino2001] rather than pure stochastic dynamics as suggested in [@Hubbell1998].
-The trajectories shaping a unimodal response to disturbance suggested the recovery of communities initial state as it was already suggested by the previous studies from the same site 10 years [@Molino2001] and 20 years [@Baraloto2012a] after disturbance.
-Only one plot displayed a still increasing Simpson diversity, and all plots showed a stabilization or a decrease of the composition distancing from 1989 inventories. 
-This recovery corroborate previous post-disturbance dynamics observed in a Bornean rainforest [@Cannon1998] where at large scale the richness was preserved although it was diminished locally or in Solomon islands where stands demonstrated a fast recovery and identical composition [@Burselm2000].
+In line with the assumtions of the IDH the observed diversity trajectories confirmed the increase of communities evenness and functional richness after disturbance.
+The consistency of the IDH has been debated for the higly diverse tropical forests, some observing the importance of environmental filters to shape communities assembly and others arguing for a genuine stochastic recovery [@Hubbell1999;@Molino2001].
+Consistently to the trajectories drawn here, a previous study of the same experiment revealed the directional changes in communities composition reflecting environmental filtering after disturbance [@Molino2001].
+Disturbance regimes brought a unimodal response depending on disturbance intensity as suggested by the outcome of 10 years and 20 years post-disturbance analysis[@Molino2001;@Baraloto2012a].
+Consistently with dynamicspreviously observed in a Bornean rainforest [@Cannon1998] or in Solomon islands [@Burslem2000], the unimodal response of diversities to disturbance suggested the maintenance of diversity levels, although for one of the intensively disturbed plot the community evenness was still increasing \@ref(fig:Fig5).
+What contrasted in our study though is that taxonomic composition did clearly recover \@ref(fig:Fig3), which may imply a long-term, if not irreversible, change in communities after disturbance.
 
-Although recovery seemed on its way the trajectories revealed significant changes in the ecological rules shaping communities over the 30 years of survey.
-Disturbance is thought to shift exclusive competition favoring competitive species to environmental filtering selecting acquisitive and disturbance resistant species [@TerSteege2001;@Flores2006;@Haddad2008]
+Regarding the functionl trajectories and dominnant trait values of disturbed communities \@ref(fig:Fig3), disturbance appeared to favor acquisitive strategies, *i.e* species with high SLA, low wood density, small seeds, etc [@Westoby1998;@Wright2004;@Reich2014].
 The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
-In particular this is demonstrated by the evolution of seed mass classes proportions revealing a clear increase in small seeded and then large dispersive species.
-Faster than for taxonomic composition and diversity though, functional characteristic returned to their initial values and the values observed for control plots.
-This suggested some functional overlap in the pre-disturbance condition (*i.e* the usual tropical forests conditions). 
-Such overlap and maybe functional versatility was already observed in highly diverse ecosystems [@Bellwood2006] and sugegsted as one mechanisms for high diversity maintenance.
-Decoupled taxonomic and functional trajectories questions to what extent communities recovery is completed: ecosystem functioning might be quickly recoverd, and taxonomic composition following suit although along for much longer, but the resilience of the community itself might be altered as functional overlap themselves would be lost.
-Along with probably other environmental modifications (like soils nutrient cycling and compaction [@Olander2005]) with alteration of communities resilience makes us expect different trajectories after subsequent disturbance.
-As disturbance seemed to update communities toward more disturbance resistant ones [@Haddad2008] it is possible that eventually the disturbance regie observed has set another permanent condition, which would only be more diverse if good competitors, poor disturbance resitant survive.
-> (Flores, 2006  for differential sapling success after disturbance)
+This means a shift from exclusive competition favoring competitive species before disturbance to environmental filtering selecting acquisitive and disturbance resistant species.
+In particular, the evolution seed mass classes proportions \@ref(fig:Fig4) revealed a clear increase in small seeded, therefore large dispersive, species [@TerSteege2001;@Flores2006;@Haddad2008].
+
+The diversity and composition functional trajectories showed a decoupling between functional and taxonomic after the most intense disturbance (treatments T2 and T3).
+Plots functional characteristic returned to their initial values, similar to those of control plots, faster than their taxonomic counterparts \@ref(fig:Fig2;fig:Fig3;fig:Fig4).
+As functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2006] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered.
+Such decoupling suggested the consistency of functional overlap and maybe functional versatility, implying a variety of functioning from the same characteristic, in the pre-disturbance condition.
+Functional overlap has been observed in highly diverse ecosystems [@Bellwood2006] and suggested as one mechanisms for high diversity maintenance.
+Knowing the decoupling between taxonomic and functional trajectories, this functional overlap may not be recovering after disturbance and therefore radically modify communities resilience and resistance [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+Given the observed lack of composition recovery, in favor of more disturbance resistant species [@Haddad2008], the likely decrease in functional overlap, and the probable changes in different life like lianas or epiphytes [@Martin2013] and environmental conditions like soils nutrient cycling and compaction [@Olander2005], the entirety of communities recovery after intense disturbance is questioned.
+In this case, post-disturbance communities might have the same functioning, as suggested by the observed functional recovery, but composition would be permanently altered in the case of most intense disturbance and species contingent to undisturbed forests in particular would be impaired.
+This threatens the global biodiversity along with the risk to loose cornerstone species and trigger unexpected consequences [@Diaz2005;@Gardner2007].
+
+To confirm this long-term communities changes and highlight the ecological rules shaping communities assembly, we suggest a focus on recruitment dynamics, that would follow the diversity and functional characteristic of recruited communities and, by extension, of the forest to be.
 
@@
 Impacts depend on the disturbance intensity, but further parameters may be involved.
-
--	To go beyond whole community diversity and highlight the ecological mechanisms behind communities assembly and repond to the question of their resilience we should focus on recruitment.
-It would also refine the role of environmental filters shaping communities assembly.
 

### 2018-02-19 Scripts/DrawingTools.R

@@ -29,7 +29,6 @@ invisible(lapply(1:length(treatments),function(tr){
 }))
 
 mtext("NMDS 1",side=1,line=2,cex=0.8)
-mtext("Euclidean distance from 1989 inventory",side=2,padj=0,line=3,cex=0.9)
 mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)
 mtext("(a) Taxonomic composition",side=3,adj=0,line=0.5)
 }
@@ -135,7 +134,7 @@ FunDist<-function(Data_FunComp){
   mtext("Years since disturbance",side=1,adj=1,line=2,cex=0.8)
 }
 
-TaxoTraj<-function(CompTaxo){par(mfrow=c(1,3),oma=c(1,1,1,1),no.readonly=TRUE)
+TaxoTraj<-function(CompTaxo){par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
 for(q in 1:3){     
   Toplot<-lapply(CompTaxo,function(tr){return(tr[,,,q])})
   Toplot<-lapply(Toplot,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
@@ -164,7 +163,7 @@ for(q in 1:3){
   }))
 }
 mtext("Years since disturbance",side=1,adj=1,cex=0.8,line=-2,outer=TRUE)
-mtext("Equivalent diversity",side=2,padj=1,cex=0.8,line=-1,outer=TRUE)}
+mtext("Equivalent diversity",side=2,padj=1,cex=0.8,line=1.5,outer=TRUE)}
 
 FunTraj<-function(CompFun){
   plot(as.numeric(colnames(CompFun[[1]])),CompFun[[1]][1,,"0.5"],type="n",xaxt="n",
@@ -200,6 +199,6 @@ SeedMassProp<-function(SeedMass){
                 col=rgb(0,0,0,alpha=0.1),border=NA)}))
     }))
   }))
-mtext("Seed mass, class proportions",line=2,adj=0,outer=TRUE)
-mtext("Years since disturbance",side=1,line=1,adj=1,cex=0.9,outer=TRUE)
+mtext("Proportion of seed mass classes",line=2,adj=0,outer=TRUE,cex=0.9)
+mtext("Years since disturbance",side=1,line=1.2,adj=1,cex=0.9,outer=TRUE)
 }

### 2018-02-20 WholePlotTrajectories.Rmd
Manuscript & Graphs update

--- 2018-02-19 WholePlotTrajectories.Rmd
@@
 # Introduction
+
 The large areas covered with tropical forests worldwide hold crucial economic, social and cultural value. 
-They provide not only wood but also multiple non-timber forest products, shelter a diversified fauna, regulate local climate and support carbon, water and many nutrient cycles and ensure cultural and human well-being.
-The increasing demand for forests products and the resulting shrink of unmanaged areas increase the pressure on the remaining intact forest landscapes as well as threaten the maintenance of their structure, composition and functionning in the long term.
-The increasing demand for forests products, land-use changes and other human disturbances along with substantial climatic changes make a crucial issue of understanding the response of forests dynamics to such disturbance regimes [@Gibson2011a; @Morales-Hidalgo2015].
-This has been addressed from the point of view of biomass production and carbon storage which resilience proved subjected to identified cyclic dynamics and therefore gives hope for the maintenance of forest ecosystem and their services.
-The response of trees biological diversity however has been sparsely studied, although this diversity is a major determinant of forests productivity, stability and functioning as it ensures interspecific complementarity, enhances the use of resources and the store of nutrients, and mitigates the prejudices of diseases and herbivory [@Tilman2014].
-Maintain the integrity of forests functioning thus requires highlighting the impacts and underlying mechanisms of disturbance on trees diversity and question the extent of forests resilience.
+They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate local climate and support carbon, water and nutrient cycles, and ensure cultural and human well-being. 
+The increasing demand for forests products reducing unmanaged areas along with climatic changes heightened the pressure on remaining forests [@Gibson2011a; @Morales-Hidalgo2015]. 
+Communities structure, composition and functioning are thus at stake, as well as the dynamics that drive them in the long term [@Anderson-Teixeira2013;@Sist2015].
 
-Disturbance is defined as rapid changes in the amount and allocation of resources such as variation in the availability of light, the fluxes of heat and water, etc. 
-It influences the environment and competition among species and combineds with environmental filters that select the pool of species fitting the environmental context to shape the composition and structure of communities in the long term: understanding ecosystems requires understanding their disturbance history [@White2001].
-Although those changes proved impacting the proportion and identity of species, negligible or even positive impacts were highlighted when considering the community as a whole, in accordance with the intermediate disturbance hypothesis (IDH).
-The IDH that predicts a maximized species diversity at intermediate level of disturbance, however, mainly rely on the study of species richness which gives limited or even misleading information on forests recovery and functioning.
-The mechanisms under which disturbance shapes communities diversity and structure involve the foster of certain species either stochastically, through the recruitment and dispersal limitation [@Hubbell1999] or determinisitcally through the selection of suited ecological types.
-Disturbance indeed in the context of old growth, stable forests where competitive exclusion favors best competitor species causes a shift toward disturbance resistant species with rapid growth and efficient resources acquisition [@Haddad2008].
-This is illustrated by a shift from slow growth with scarce resources and high competitive ability, "conservative strategies", to fast growth with prolific resourcesn, "acquisitive strategies".
-A way to disentangle the mechanisms of disturbance dynamics would then to identify the shifts in communities functional types definde by the acquisitive or conservative functional traits and by the reproductive strategy through seed and early stage traits [@TerSteege2001;@Reich2014;@Herault2011].
-Relevant functional types would be defined with the combination of resource acquisition traits: in that respect leaf and stem traits proved relevant in a large litterature [@Chave2009b;@Wright2004],and life history traits revealing the growth and reproductive strategy [@Westoby1998;@Herault2011]
-Beyond diversity indices it is essential to analyse changes in taxonomic composition so that the impact on species per se and their particular impact on ecosystem can be followed [@Lavorel2002;@Bellwood2006;@Baraloto2012a] along with the ecological rules shaping communities assembly and abundance distribution [@Magurran1988].
+Tropical forests are subject to a wide range of disturbance, defined as rapid changes in the amount and allocation of resources such as light, heat and water fluxes, etc.
+Those institute the intrinsinc dynamical state of forest communities in changing environmental conditions and interspecific relations which make the understanding of communities response to disturbance a cornerstone for ecology, modelling and management [@White2001;@Chazdon2003a].
+Disturbance recovery have been largely studied regarding structural parameters as aboveground biomass, tree height or stem density which display rapid and measurable changes related to ecosystems functioning [@Denslow2000;@Blanc2009;@Rutishauser2016].
+Species composition and diversity also are surely impacted by the environmental changes after disturbance and the subsequent modification of species selection and competition.
+The diversity and composition of communities are crucial for their functioning and dynamics as they determine their productivity, stability and functioning through interspecific complementarity which enhances the use of resources and the storing of nutrients, and mitigates the prejudices of diseases and herbivory [@Tilman2014].
+
+Although inescapable for the understanding of ecological rules shaping tropical forest communities, the long-term response of communities composition and diversity to disturbance remains unclear.
+Until now, short-term dynamics demonstrated negligible or even positive impacts on communities diversity in accordance with the intermediate disturbance hypothesis (IDH) [@Molino2001;@Kariuki2006a;@Berry2008a].
+The IDH predicting a maximized species diversity at intermediate level of disturbance, however, was mainly tested through species richness which gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+Communities response should encompass taxonomic composition and evenness which only can assess conservation issues and ecological rules shaping abundance distribution[@Magurran1988;@Lavorel2002;@Bellwood2006].
+Beyond taxonomic metrics, communities recovery should also encompass functional diversity and composition metrics [@Moretti2009;@Baraloto2012a].
+Functional traits, the species biological attributes that relate to individuals fitness and ecosystem properties [@Violle2007b;@Scheiter2013], were largely adopted and integrated within a workable framework to assess forests dynamics and functioning [@Diaz2005;@Villeger2008a].
+Numerous studies revealed some major functional traits determining species ecology and performance that are thus relevant to reveal the ecological rules shaping species assembly: post-disturbance trajectories in tropical rainforests for example, are assumed to be driven by environmental filters that foster disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
+This would be reflected by a shift from "conservative" species displaying long-lived, dense tissues growing slowly but with scarce resources, to "acquisitive" species with fast-growing, light tissues allowing fast resources acquisition [@TerSteege2001;@Reich2014;@Herault2011].
+In this perspective a combination of resource acquisition traits (leaf area,density and chlorophyll content and stem specific gravity and bark thickness) and of life history traits revealing the growth and reproductive strategy (seed mass and maximum height) already proved relevant [@Chave2009b;@Wright2004;@Westoby1998].
 > See Garnier 2001a for the description of traits functional impact
 
-Since the functional traits chosen showed low intraspecific compared to interspecific variability [@Garnier2001a;@Westoby2006a] it its valid to assess the functional trajectories of communities through the functional values observed or the species assembly.
+>Since the functional traits chosen showed low intraspecific compared to interspecific variability >[@Garnier2001a;@Westoby2006a] it its valid to assess the functional trajectories of communities through the >functional values observed or the species assembly.
 
@@
 *Traits community weighted means* _ The trajectories of the community weighted means (CWM) of the 8 functional traits were measured at plot level \@ref(fig:Fig3).
-**To be coming**
 
-```{r Fig3, echo=FALSE, fig.cap="Trajectories of the CWM until 30 years after disturbance of 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness).
+```{r Fig3, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the CWM until 30 years after disturbance of 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness).
 The CWM were measured at plot level and averaged among plots of the same treatment (green for control, blue for T1,orange for T2 and red for T3).
 The average CWM per treatment are drawn (solid line) drawn along with the 0.025 and 0.975 percentile (gray envelope) observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
+load("DB/CWM")
+
+CWMdraw(CWM)
 
@@
 Disturbance regimes brought a unimodal response depending on disturbance intensity as suggested by the outcome of 10 years and 20 years post-disturbance analysis[@Molino2001;@Baraloto2012a].
-Consistently with dynamicspreviously observed in a Bornean rainforest [@Cannon1998] or in Solomon islands [@Burslem2000], the unimodal response of diversities to disturbance suggested the maintenance of diversity levels, although for one of the intensively disturbed plot the community evenness was still increasing \@ref(fig:Fig5).
-What contrasted in our study though is that taxonomic composition did clearly recover \@ref(fig:Fig3), which may imply a long-term, if not irreversible, change in communities after disturbance.
+As observed after logging in a Boranean rainforest [@Cannon1998] or after cyclones in the Solomon islands [@Burslem2000], communities diversity showed a unimodal response to disturbance and a maintenance of diversity levels, except for one of the intensively disturbed community which evenness was still increasing \@ref(fig:Fig5).
+Despite the recovery of communities diversity, though, taxonomic composition was clearly altered \@ref(fig:Fig3) implying a long-term, if not permenant, composition change after disturbance as observed in East African tropical forests [@Bonnell2011].
 
-Regarding the functionl trajectories and dominnant trait values of disturbed communities \@ref(fig:Fig3), disturbance appeared to favor acquisitive strategies, *i.e* species with high SLA, low wood density, small seeds, etc [@Westoby1998;@Wright2004;@Reich2014].
+Regarding the functionl trajectories and dominant trait values of disturbed communities \@ref(fig:Fig3), disturbance appeared to favor acquisitive strategies, *i.e* species with high SLA, low wood density, small seeds, etc [@Westoby1998;@Wright2004;@Reich2014].
 The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
@@
 Knowing the decoupling between taxonomic and functional trajectories, this functional overlap may not be recovering after disturbance and therefore radically modify communities resilience and resistance [@Trenbath1999;@Elmqvist2003;@Diaz2005].
-Given the observed lack of composition recovery, in favor of more disturbance resistant species [@Haddad2008], the likely decrease in functional overlap, and the probable changes in different life like lianas or epiphytes [@Martin2013] and environmental conditions like soils nutrient cycling and compaction [@Olander2005], the entirety of communities recovery after intense disturbance is questioned.
-In this case, post-disturbance communities might have the same functioning, as suggested by the observed functional recovery, but composition would be permanently altered in the case of most intense disturbance and species contingent to undisturbed forests in particular would be impaired.
-This threatens the global biodiversity along with the risk to loose cornerstone species and trigger unexpected consequences [@Diaz2005;@Gardner2007].
+Given the impaired functional overlap, the persitant compositional changes regarding disturbance resistant species [@Haddad2008] and probably lianas or epiphytes [@Martin2013], and the modification of environmental conditions like soils nutrient cycling and compaction [@Olander2005] the entirety of communities recovery after intense disturbance is deeply questioned [@Chazdon2003a].
+New conditions would not only be longer lasting but self-maintained as tied to disturbance regime [@Burslem2000].
+Although communities recovered the core of their initial functioning, composition seemed permanently altered after intense disturbance.
+Specificaly, this would impair species contingent to undisturbed forests, threatening their maintenance, and run the risk of cornerstone species loss and unexpected ecological consequences [@Diaz2005;@Gardner2007].
 

### 2018-02-20 Scripts/DrawingTools.R

@@ -184,6 +184,27 @@ invisible(lapply(1:length(CompFun),function(t){
   }))
 }))}
 
+CWMdraw<-function(Cwm){
+  par(mfrow=c(2,4),mar=c(2,2,3,1),oma=c(2,1,2,1),no.readonly = T)
+invisible(lapply(colnames(Cwm[[1]]),function(trait){
+  Toplot<-lapply(Cwm,function(pl){return(t(pl[,trait,]))})
+  plot(colnames(Toplot[[1]]),Toplot[[1]]["0.5",], ylim=c(min(unlist(Toplot)),max(unlist(Toplot))),type="n",xlab="years",ylab="")
+  mtext(trait,3,cex=0.8,adj=0,line=0.5)
+  
+  invisible(lapply(1:4,function(tr){
+    toplot<-Toplot[which(names(Toplot)%in%treatments[[tr]])]
+    invisible(lapply(toplot,function(plo){
+      lines(colnames(plo),plo["0.5",],col=ColorsTr[tr],lwd=2)
+      polygon(c(colnames(plo),rev(colnames(plo))),c(plo["0.025",],rev(plo["0.975",])),
+              col=rgb(0,0,0,alpha=0.1),border=NA)
+      
+    }))
+  }))
+}))
+mtext("Community Weighted Means",line=0.5,adj=0,outer=TRUE,cex=1.1)
+mtext("Years since disturbance",side=1,line=1.2,adj=1,cex=0.9,outer=TRUE)
+}
+
 SeedMassProp<-function(SeedMass){
   par(mfrow=c(1,5),mar=c(1,1,2,1),oma=c(2,1,4,1),no.readonly = T)
   invisible(lapply(1:5,function(clas){

### 2018-02-21 WholePlotTrajectories.Rmd
Introduction update

@@
 This would be reflected by a shift from "conservative" species displaying long-lived, dense tissues growing slowly but with scarce resources, to "acquisitive" species with fast-growing, light tissues allowing fast resources acquisition [@TerSteege2001;@Reich2014;@Herault2011].
-In this perspective a combination of resource acquisition traits (leaf area,density and chlorophyll content and stem specific gravity and bark thickness) and of life history traits revealing the growth and reproductive strategy (seed mass and maximum height) already proved relevant [@Chave2009b;@Wright2004;@Westoby1998].
+In this perspective a combination of resource acquisition traits (leaf area,density and chlorophyll content and stem specific gravity and bark thickness) and of life history traits revealing the growth and reproductive strategy (seed mass and maximum height) already proved relevant [@Wright2004;@Westoby2006a;@Chave2009b].
 > See Garnier 2001a for the description of traits functional impact
@@
 
-Such combination of composition and diversity in browsing all facets of stands assemblage would highlight their resilience to perturbation, and its determinant and underlying ecological processes.
-Disturbance survey would first assess species turnover and communities convergence through stands taxonomic composition and unveil environmental filters through functional composition that is based on the biological attributes of plants.
-The set of functional traits chosen for this functional analysis considers functional traits representative of species functional strategy, fitness, and ecology as it is the case for traits of the leaf or stem economics.
-Composition analysis would be continued with diversity analysis assessing for the whole community the richness and evenness of the taxonomic and functional abundance distribution.   
-Here we examined the long-term trajectory, in the sense of the evolution along time, of tropical forests taxonomic and functional composition and diversity. 
-We studied the regular inventories over 30 years of twelve Amazonian forest plots, either untouched or disturbed with three disturbance intensity . 
-The inventories were combined with one of the largest functional trait database for tropical trees.
-We aimed to (i) explicit the impact of disturbance on the long-term trajectory of tropical forest communities composition and diversity, (ii) question the long-term convergence of the composition of communities and (iii) determine their resilience or not, after disturbance.
+Here we investigated over 30 years the response of 75ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
+We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw communities taxonomic and functional composition and diversity trajectories after disturbance.
+Specifically, we (i) tested the validity of the IDH arguing for enhanced diversity after moderate disturbance and (ii) confirmed the changes in ecological rules shaping communities, from competitive exclusion between "conservative", slow-growing tree species to environmental filtering of "acquisitive", fasting-growing ones.
+Eventually we examined the different trajectories to (iii) resolve the resilience of tropical forests and question their future dynamics regarding their state 30 years after disturbance.
 
@@
 In line with the "vegetation quantity effect" suggesting that communities functional characteristics depends on dominant species rather that subtle changes in biodiversity [@reftobefound].
-This would result in a faster recovery of ecosystem processes rates than in communities ocmposition and diversity as it mas largely demonstrated for grasslands [@Tilman1997; @Mouillot2011] and recently proved for tropical forests [@Lohbeck2015].
-This highlighted the functional overlap between species in tropical forests,typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.
-Ecosystem processes graped through the set of functional traits retained here however do not assess communities resilience and future dynamics.
+This would result in a faster recovery of ecosystem processes rates than composition and diversity as demonstrated for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
+This highlighted the functional overlap between species in tropical forests, typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.
+Ecosystem processes grasped through the set of functional traits retained here however do not assess communities resilience and future dynamics.
 
@@
 Plots functional characteristic returned to their initial values, similar to those of control plots, faster than their taxonomic counterparts \@ref(fig:Fig2;fig:Fig3;fig:Fig4).
-As functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2006] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered.
+As functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered [@Guariguata2001].
 Such decoupling suggested the consistency of functional overlap and maybe functional versatility, implying a variety of functioning from the same characteristic, in the pre-disturbance condition.

### 2018-02-21 WholePlotTrajectories.Rmd
Introduction upadte

@@
 The large areas covered with tropical forests worldwide hold crucial economic, social and cultural value. 
-They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate local climate and support carbon, water and nutrient cycles, and ensure cultural and human well-being. 
-The increasing demand for forests products reducing unmanaged areas along with climatic changes heightened the pressure on remaining forests [@Gibson2011a; @Morales-Hidalgo2015]. 
-Communities structure, composition and functioning are thus at stake, as well as the dynamics that drive them in the long term [@Anderson-Teixeira2013;@Sist2015].
+They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate local climate, support carbon, water and nutrient cycles, and ensure cultural and human well-being. 
+The simultaneous increase of forests products demand and substantial climatic changes heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threaten the maintenance of communities structure, composition and functioning and the dynamics shaping them [@Anderson-Teixeira2013;@Sist2015].
 
-Tropical forests are subject to a wide range of disturbance, defined as rapid changes in the amount and allocation of resources such as light, heat and water fluxes, etc.
-Those institute the intrinsinc dynamical state of forest communities in changing environmental conditions and interspecific relations which make the understanding of communities response to disturbance a cornerstone for ecology, modelling and management [@White2001;@Chazdon2003a].
+In tropical forests, a constant range of disturbances determines the ecological rules that shape the communities, through the reallocation of resources (as light, heat and water fluxes), changes in interspecific relations, etc.
+The response of communities to disturbance is then the cornerstone of tropical forests ecology, modelling and management [@White2001;@Chazdon2003a].
 Disturbance recovery have been largely studied regarding structural parameters as aboveground biomass, tree height or stem density which display rapid and measurable changes related to ecosystems functioning [@Denslow2000;@Blanc2009;@Rutishauser2016].

### 2018-02-21 Scripts/DrawingTools.R

@@ -134,7 +134,8 @@ FunDist<-function(Data_FunComp){
   mtext("Years since disturbance",side=1,adj=1,line=2,cex=0.8)
 }
 
-TaxoTraj<-function(CompTaxo){par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
+TaxoTraj<-function(CompTaxo){
+  par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
 for(q in 1:3){     
   Toplot<-lapply(CompTaxo,function(tr){return(tr[,,,q])})
   Toplot<-lapply(Toplot,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})

### 2018-02-21 WholePlotTrajectories.Rmd
Manuscript upadte

@@
 The response of communities to disturbance is then the cornerstone of tropical forests ecology, modelling and management [@White2001;@Chazdon2003a].
-Disturbance recovery have been largely studied regarding structural parameters as aboveground biomass, tree height or stem density which display rapid and measurable changes related to ecosystems functioning [@Denslow2000;@Blanc2009;@Rutishauser2016].
-Species composition and diversity also are surely impacted by the environmental changes after disturbance and the subsequent modification of species selection and competition.
-The diversity and composition of communities are crucial for their functioning and dynamics as they determine their productivity, stability and functioning through interspecific complementarity which enhances the use of resources and the storing of nutrients, and mitigates the prejudices of diseases and herbivory [@Tilman2014].
+Disturbance recovery have been largely studied through structural parameters as aboveground biomass, tree height or stem density that all reflect rapid and measurable changes of ecosystems functioning asses crucial issues about the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
+In this perspective tree species diversity is inevitable as it determines ecosystems productivity, stability and functioning through the interspecific complementarity that enhances resources use, nutrients storing, and hazards mitigation [@Tilman2014].
+Reveal diversity dynamics is besides all the more important that disturbance, modifying the environment and interspecific relations, suerly impacts trees composition and diversity.
 
-Although inescapable for the understanding of ecological rules shaping tropical forest communities, the long-term response of communities composition and diversity to disturbance remains unclear.
-Until now, short-term dynamics demonstrated negligible or even positive impacts on communities diversity in accordance with the intermediate disturbance hypothesis (IDH) [@Molino2001;@Kariuki2006a;@Berry2008a].
-The IDH predicting a maximized species diversity at intermediate level of disturbance, however, was mainly tested through species richness which gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-Communities response should encompass taxonomic composition and evenness which only can assess conservation issues and ecological rules shaping abundance distribution[@Magurran1988;@Lavorel2002;@Bellwood2006].
-Beyond taxonomic metrics, communities recovery should also encompass functional diversity and composition metrics [@Moretti2009;@Baraloto2012a].
-Functional traits, the species biological attributes that relate to individuals fitness and ecosystem properties [@Violle2007b;@Scheiter2013], were largely adopted and integrated within a workable framework to assess forests dynamics and functioning [@Diaz2005;@Villeger2008a].
-Numerous studies revealed some major functional traits determining species ecology and performance that are thus relevant to reveal the ecological rules shaping species assembly: post-disturbance trajectories in tropical rainforests for example, are assumed to be driven by environmental filters that foster disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
-This would be reflected by a shift from "conservative" species displaying long-lived, dense tissues growing slowly but with scarce resources, to "acquisitive" species with fast-growing, light tissues allowing fast resources acquisition [@TerSteege2001;@Reich2014;@Herault2011].
-In this perspective a combination of resource acquisition traits (leaf area,density and chlorophyll content and stem specific gravity and bark thickness) and of life history traits revealing the growth and reproductive strategy (seed mass and maximum height) already proved relevant [@Wright2004;@Westoby2006a;@Chave2009b].
-> See Garnier 2001a for the description of traits functional impact
-
->Since the functional traits chosen showed low intraspecific compared to interspecific variability >[@Garnier2001a;@Westoby2006a] it its valid to assess the functional trajectories of communities through the >functional values observed or the species assembly.
+Until now, although, short-term dynamics demonstrated negligible or even positive impacts on communities diversity in accordance with the intermediate disturbance hypothesis (IDH) [@Molino2001;@Kariuki2006a;@Berry2008a] the long-term response of communities composition and diversity to disturbance remains unclear.
+The IDH predicting a maximized species diversity at intermediate level of disturbance have besides mainly been tested through species richness, which gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+Studies of communities response to disturbance should encompass taxonomic composition and evenness, which only can assess some conservation issues and the ecological rules that determine communities abundance distribution [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Beyond taxonomic metrics, communities recovery should also encompass functional diversity metrics [@Moretti2009;@Baraloto2012a] through functional traits, that are species biological attributes relating to individuals fitness and ecosystem properties [@Violle2007b;@Scheiter2013].
+Functional traits were largely adopted and integrated within a workable framework to assess forests dynamics and functioning [@Diaz2005;@Villeger2008a] and a vast litterature revealed some major functional traits determinant for species ecology and performance.
+These major traits proved relevant to assess the ecological rules shaping species assembly: post-disturbance trajectories in tropical rainforests, for example, are assumed to be driven by environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
+Communities dynamics would then be reflected by a shift from "conservative" species displaying long-lived, dense tissues growing slowly but with scarce resources, to "acquisitive" species with fast-growing, light tissues allowing fast resources acquisition [@TerSteege2001;@Reich2014;@Herault2011].
+In this perspective a combination of resource acquisition traits (leaf area,density and chlorophyll content and stem specific gravity and bark thickness) and life history traits of growth and reproduction (seed mass and maximum height) already proved relevant [@Wright2004;@Westoby2006a;@Chave2009b].
 
@@
 
-The study site corresponds to a tropical rain forest with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae. 
-In the twelve experimental plots of the experiment, all trees above 10 cm DBH are mapped and measured annually since 1984. 
-During the inventories, trees are first identified with a vernacular name, assigned by the field team, and then with a scientific name, assigned by a botanist during regular botanical campaigns. 
-In 1984, specific vernacular names were given to 62 commercial or common species whereas two identifiers separating palm trees and trees were given to all other species. 
-The botanical campaigns to identify all trees at the species level started in 2003 and continued every 5 to 6 years.
-This inventory protocol rasied some mehodological issues given the variability of botanical identification practices along time and among plots. 
-Because vernacular name attributed to trees usually correspond to a variety of botanical species, some taxonomic uncertainty remains whenever a trustworthy botanical name cannot be given thereafter. 
-This taxonomic uncertainty was accounted for in the composition and diversity metrics throught a Bayesian framework estimating the expectancy and variance of composition and diversity metrics. 
-The uncertainty propagation is done by the extrapolation of complete inventories at genus level from real incomplete inventories, on the basis vernacular names replacement.
-The replacement of vernacular names uses a multinomial distribution M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
+The study site corresponds to a tropical rainforest with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical families. 
+In the twelve experimental plots of the experiment, all trees above 10 cm DBH were mapped and measured annually since 1984. 
+During inventories, trees were first identified with a vernacular name assigned by the field team, and afterward with a scientific name assigned by a botanist during regular botanical campaigns. 
+In 1984, specific vernacular names were given to 62 commercial or common species whereas other less common species were identified under two identifiers only separating trees and palm trees. 
+The botanical campaigns carried every 5 to 6 years to identify all trees at the species level only started in 2003 and identification practices varied among plots and successive campaigns. 
+This raised methodological issues as vernacular names usually correspond to different botanical species, resulting in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics throught a Bayesian framework.
+The uncertainty propagation was done by the replenishment of inventories completed at genus level from real incomplete ones on the basisof vernacular/botanical names association.
+Vernacular names were replaced through multinomial trials (M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) ) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
 See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.
 
-The functional metrics used a set of 6 functional traits values representing traits of the leaf and wood economics and life history traits.
-The trait database encompassed 4 traits of the leaves: leaves thickness, toughness, total chlorophyll content and specific leaf area (leaf area per unit dry mass), 2 traits of the stem: wood specific gravity and bark thickness and two life-history traits: maximum specific height and the seed mass.
+The functional diversity metrics used a dataset for 6 functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, theleaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass).
 Traits database came from the BRIDGE project (http://www.ecofog.gf/Bridge/index.html) where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou.
@@
 
-In line with the "vegetation quantity effect" suggesting that communities functional characteristics depends on dominant species rather that subtle changes in biodiversity [@reftobefound].
+In line with the "vegetation quantity effect" suggesting that communities functional characteristics depends on dominant species rather that subtle changes in biodiversity [@Grime1998].
 This would result in a faster recovery of ecosystem processes rates than composition and diversity as demonstrated for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].

### 2018-02-21 WholePlotTrajectories.Rmd
Manuscript update

@@
 ### Study site
+
 Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-months dry season (< 100 mm.months^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
@@
 Vernacular names were replaced through multinomial trials (M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) ) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
+To avoid remaining identification caveats and consider complete inventories, the simulated inventories were then reported at genus.
 See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.
 
-The functional diversity metrics used a dataset for 6 functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, theleaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass).
+The functional diversity metrics used a dataset for 6 functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, the leaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass).
 Traits database came from the BRIDGE project (http://www.ecofog.gf/Bridge/index.html) where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou.
-To account for intraspecific traits variability, trait values at species level were assessed in sampling one of the measured individuals for iteration rather than averaging at species level. 
-Missing trait values were filled using multivariate imputation by chained equation (mice), restricted to the samples pertaining to the next higher taxonomic level, in order to account for the phylogenetic signal of the functional traits.
-The dataset comprised 257 botanical species.
+Missing trait values were filled using multivariate imputation by chained equation (mice) restricted to samples pertaining to the next higher taxonomic level, in order to account for the phylogenetic signal of the functional traits.
+The dataset comprised 294 botanical species pertaining to 157 botanical genus. 
 Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level. 
-All composition and diversity metrics correspond to the expectancy obtained after 50 iterations of the taxonomic uncertainty propagation framework and the missing trait values filling process.
+All composition and diversity metrics corresponded to the expectancy obtained after 50 iterations of the taxonomic uncertainty propagation framework and of the filling process of missing trait values.
 
@@
 The analysis of taxonomic and functional composition of plots along time was visualized in a two-dimensional ordination space based on non-metric dimensional scaling of succesive floristic or functional inventories.
-Plots evolution along time was reportedcomparatively to the inventories in 1989, 5 years after disturbance, which correspond to first inventory with a sufficient degree of uncertainty (<30% of undetermined trees).
-The distancing relative to the reference 1989 inventory was reported using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) similarity measures.
-The euclidean distance in the two-dimensional ordination space was used to visualize the distancing to reference inventory along time.
-The functional composition was besides completed by trajectory along time of the community weighted means (CWM) for the 8 leaf, stem and life history traits. 
+Plots trajectories along time was reported comparatively to the inventories in 1989, 5 years after disturbance, which corresponded to first inventory with a sufficient degree of uncertainty (<30% of undetermined trees).
+The inventories dissimilarity compared to the reference 1989 inventory was reported using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) similarity measures.
+The trajectory of inventories along time was visualized with the euclidean distance in the two-dimensional ordination space to the 1989 inventory.
+The functional trajectories were also visualized with the community weighted means (CWM) of the leaf and stem and data life traits. 
 To determine whether communities’ assembly deviated from the expectation of a random process we compared the observed similarity patterns to those generated by 50 repetitions of stochastic null models.
@@
 This order q grasps the balance between richness and evenness in the community as it determines the weight of common species in the diversity metric, with an growing emphasis on dominant species as q increases.
-The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on the 6 functional traits specific values.
+The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on all functional and data life traits.
 

### 2018-02-22 WholePlotTrajectories.Rmd
Manuscript update

@@
 Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level. 
+As seed mass information correspond to a classification into mass classes, no data filling process was applied so analysis were performed considering the 414 botanical species of the seed mass dataset.
 All composition and diversity metrics corresponded to the expectancy obtained after 50 iterations of the taxonomic uncertainty propagation framework and of the filling process of missing trait values.
@@
 
+To counter the remaining taxonomic uncertainty (see previously) plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
 The analysis of taxonomic and functional composition of plots along time was visualized in a two-dimensional ordination space based on non-metric dimensional scaling of succesive floristic or functional inventories.
@@
 The trajectory of inventories along time was visualized with the euclidean distance in the two-dimensional ordination space to the 1989 inventory.
-The functional trajectories were also visualized with the community weighted means (CWM) of the leaf and stem and data life traits. 
+The functional trajectories of the leaf and stem and data life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004].
+Species seed mass correspond to 5 classes of increasing mass, the seed mass trajectories were reported by the proportion of each class recorded in the inventories.
 To determine whether communities’ assembly deviated from the expectation of a random process we compared the observed similarity patterns to those generated by 50 repetitions of stochastic null models.
@@
 
-Over time, 104423 trees and 696 botanical species were recorded. 
-The disturbance proved to significantly impact for more than 30 years the functional and taxonomic composition and diversity of the stands.
-
 ### Taxonomic and functional composition
 
-*Composition trajectories* _ Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in a two-dimensional space based on NMDS analysis.
-Classifications were based on either abundance-based Bray-Curtis \@ref(fig:Fig1) or incidence-based Jaccard dissimilarity (data not shown).
+*Composition trajectories*
 
-Both taxonomic and functional composition were substantially affected by disturbance and positively correlated with its intensity.
+Over time, 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
+Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in the two-dimensional space from NMDS analysis based on flora inventories and the corresponding functional composition.
+Classifications were performed using either abundance-based Bray-Curtis \@ref(fig:Fig1) or incidence-based Jaccard dissimilarity (data not shown).
+
+Both taxonomic and functional composition were substantially affected by disturbance, and the maximum dissimilarity to 1984 reference inventory composition was positively correlated with the disturbance intensity.
+Until 30 years after disturbance, all disturbed plots taxonomic composition remained significantly dissimilar to that of the 1989 reference inventory composition \@ref(fig:Fig1).
+All plots, though, displayed a unimodal trajectory with a the return toward the taxonomic and functional composition of the 1989 reference inventory suggesting a shift towards a cyclic regime \@ref(fig:Fig2).
+For taxonomic composition, the maximum dissimilarity to 1984 inventory composition was reach around 20 years after disturbance for the T2 and T3 plots and for one T1 plot while control and other T1 plots continued to increase \@ref(fig:Fig2.a).
+For functional composition, the maximum dissimilarity to 1984 inventory composition was reach from 15 to 20 years after disturbance for the T2 and T3 plots while T1 and control plots continued to increase \@ref(fig:Fig2.b).
+The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance \@ref(fig:Fig2).
+
 *The distance from the initial condition was significantly time-dependent (P<0.01).* 
-The taxonomic composition remained significantly different from the 1989 reference inventory until 30 years for all disturbed plots \@ref(fig:Fig1). 
-From 20 years after disturbance, though, the distancing decreased which suggested a return toward the reference 1989 inventory and its resilience to disturbance \@ref(fig:Fig1),\@ref(fig:Fig2). 
-The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance \@ref(fig:Fig2).
 *The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).*
@@
 
-*Traits community weighted means* _ The trajectories of the community weighted means (CWM) of the 8 functional traits were measured at plot level \@ref(fig:Fig3).
+*Traits community weighted means*
 
-```{r Fig3, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the CWM until 30 years after disturbance of 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness).
-The CWM were measured at plot level and averaged among plots of the same treatment (green for control, blue for T1,orange for T2 and red for T3).
-The average CWM per treatment are drawn (solid line) drawn along with the 0.025 and 0.975 percentile (gray envelope) observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
+For all plots the trajectories of the community weighted means (CWM) were drawn for the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area,wood specific gravity and barck thickness and seed mass and maximum adult height) \@ref(fig:Fig3).
+
+>To compensate the intrinsinc difference among plots the trajectories drawn correspond to the difference in value between the reference inventory in >1984 (5 year after disturbance) and the successive years inventoried.
+
+Except for leaves cholorphyll content, which displayed higher difference between plots than among treatments which may be due to the completeness of dataset, all CWM trajectories corresponded to significant changes after disturbance.
+All functional traits and seed mass proportions displayed a unimodal trajectories but with different times at maximum and different values at the end of the experiment (30 years after disturbance). 
+30 years after disturbance, the weighted means of communities specific maximum height at adult stage (Hmax), leaf toughness (L_toughness)and wood specific gravity (WD) remained significantly lower than their initial value and than these of the control plots \@ref(fig:Fig3).
+The weighted means of bark thickness (Bark_thick) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (SLA) had almost recovered its initial value and this of the undisturbed plots at the end of the experiment.
+
+```{r Fig3, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, *L_thickness*, chlorophyll content, *L_chloro*, toughness, *L_toughness* and specific area, *SLA*), 2 stem traits (wood specific gravity, *WD*, and bark thickness, *Bark-thick*) and one data life trait (Specific maximum height at adult stage,*Hmax*).
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3."}
 load("DB/CWM")
@@
 
-```{r Fig4, echo=FALSE, out.width='100%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions until 30 years after disturbance.
-Plot treatments are indicated with line colors in green for control, blue for T1, orange for T2 and red for T3.
-The proportion reported are the median values (solid line) and the 0.025 and 0.975 percentile (gray envelope) observed for 50 iteration of the taxonomic uncertainty propagation.
-only species which seed mass values were reported are considered (*i.e* 414 out of 696 species recorded in total)."}
+```{r Fig4, echo=FALSE, out.width='100%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+No gap filling process was applied in this case.
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3."}
 load("DB/Smass_ClassProp")
@@
 
-*Taxonomic diversity* _ Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1994 inventories \@ref(fig:Fig5). 
+*Taxonomic diversity*
+
+Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1989 inventories (5 years after disturbance) \@ref(fig:Fig5). 
 *For the three indices the OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and a covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01).*
 
-```{r Fig5, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory, 10 years after disturbance, until 30 years after disturbance for plots **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversity.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The average value of the 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness) are drawn along with the 0.025 and 0.975 percentile observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
+```{r Fig5, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3."}
 load("DB/ReplacementTraj_ForGraphs")
@@
 
-The evolution along time of the difference of Richness, Shannon and Simpson diversity compared to reference inventories in 1989 remained even for 30 years after disturbance for the control plots whereas disturbed plots showed significant increase, positively correlated with the disturbance intensity \@ref(fig:Fig5).
-The richness stabilized for all disturbed plots from 20 years after disturbance \@ref(fig:Fig5). 
-The Shannon and Simpson diversity also stabilized 20 years after disturbance for T1 and T2 plots \@ref(fig:Fig5) whereas for 2 out of 3 plots T3 the Shannon and Simpson diversity continued increasing until 30 years after disturbance, illustrating a significant increase in communities’ evenness.
+While for undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity \@ref(fig:Fig5).
+The Richness was less affected by the initial disturbance except for the higher intensity (T3 plots) \@ref(fig:Fig5). 
+The Shannon and Simpson diversities however first significantly increased, illustrating an increasing evenness, before reaching a maximum around 15 years after disturbance for all plots except one from T1 treatment and one from T2 treatment \@ref(fig:Fig5) .
 
-*Functional  diversity* _ The evolution along time of the Rao functinal diversity compared to reference inventories in 1989 remained even all along the survey for the control plots. 
-All disturbed plots however showed an increase of the functional diversity that was positively correlated to the disturbance intensity \@ref(fig:Fig6).
-30 after disturbance the diversity of disturbed plots remained significantly higher than this of control plots, thus whenever the initial disturbance intensity.
-The slope of the initial diversity increase and the maximum diversity were both positively correlated with the disturbance intensity. 
+*Functional  diversity*
+
+For all undisturbed plots the Rao diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity \@ref(fig:Fig6).
+Although a first increase of plots functional diversity that was all the more important than initial disturbance was intense, 30 after disturbance the diversity was similar among disturbed plots and significantly higher than this of control plots.
 *The OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and the covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01)*.
-The trajectories of plot 7 from the T1 disturbance treatment was discarded because of too high taxonomic uncertainty preventing correct analysis.
 
-```{r Fig6, fig.asp=1/1, echo=FALSE,fig.cap="Trajectories of the plots Rao functional diversity from10 to 30 years after disturbance.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The value of the 4 leaf traits (Leaf thickness, chlorophyll content, toughness and specific area) and the 2 stem traits (Wood density and barck thickness) are drawn along with the 0.025 and 0.975 percentile observed for 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes."}
+```{r Fig6, fig.asp=1/1, echo=FALSE,fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes."}
 load("DB/FunctionalTraj_ForGraphs")
@@
 FunTraj(CompleteFun)
+
 ```

### 2018-02-22 Scripts/DrawingTools.R

@@ -168,22 +168,22 @@ mtext("Equivalent diversity",side=2,padj=1,cex=0.8,line=1.5,outer=TRUE)}
 
 FunTraj<-function(CompFun){
   plot(as.numeric(colnames(CompFun[[1]])),CompFun[[1]][1,,"0.5"],type="n",xaxt="n",
-     xlab="",ylab="",ylim=c(min(unlist(CompFun),na.rm=T),max(unlist(CompFun),na.rm=T)),cex.axis=0.7)
-axis(1,at=as.character(seq(5,33,5)),labels=TRUE,cex.axis=0.7)  
-mtext("Rao diversity",3,adj=0,line=1,cex=1.5) 
-mtext("Years since disturbance",side=1,adj=1,line=2)
-mtext("Equivalent diversity",side=2,padj=1,line=3)
-
-invisible(lapply(1:length(CompFun),function(t){  
+       xlab="",ylab="",ylim=c(min(unlist(CompFun),na.rm=T),max(unlist(CompFun),na.rm=T)),cex.axis=0.7)
+  axis(1,at=as.character(seq(5,33,5)),labels=TRUE,cex.axis=0.7)  
+  mtext("Rao diversity",3,adj=0,line=1,cex=1.5) 
+  mtext("Years since disturbance",side=1,adj=1,line=2)
+  mtext("Equivalent diversity",side=2,padj=1,line=3)
   
-  toplot<-CompFun[[t]]
-  
-  invisible(lapply(1:nrow(toplot),function(i){
-    lines(colnames(toplot),toplot[i,,"0.5"], col = ColorsTr[[t]],lty = 1,lwd=2)
-    polygon(c(colnames(toplot),rev(colnames(toplot))),c(toplot[i,,"0.975"],rev(toplot[i,,"0.025"])),
-            col=rgb(0,0,0,alpha=0.1),border=NA)
-  }))
-}))}
+  invisible(lapply(1:length(CompFun),function(t){  
+    
+    toplot<-CompFun[[t]]
+    
+    invisible(lapply(1:nrow(toplot),function(i){
+      lines(colnames(toplot),toplot[i,,"0.5"], col = ColorsTr[[t]],lty = 1,lwd=2)
+      polygon(c(colnames(toplot),rev(colnames(toplot))),c(toplot[i,,"0.975"],rev(toplot[i,,"0.025"])),
[…]

### 2018-02-23 WholePlotTrajectories.Rmd
Manuscript update & DB remove (.gitignore)

@@
 All functional traits and seed mass proportions displayed a unimodal trajectories but with different times at maximum and different values at the end of the experiment (30 years after disturbance). 
-30 years after disturbance, the weighted means of communities specific maximum height at adult stage (Hmax), leaf toughness (L_toughness)and wood specific gravity (WD) remained significantly lower than their initial value and than these of the control plots \@ref(fig:Fig3).
-The weighted means of bark thickness (Bark_thick) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (SLA) had almost recovered its initial value and this of the undisturbed plots at the end of the experiment.
+30 years after disturbance, the weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value and than these of the control plots \@ref(fig:Fig3).
+The weighted means of bark thickness (*Bark_thick*) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value and this of the undisturbed plots at the end of the experiment.
 

### 2018-02-23 Scripts/DrawingTools.R

@@ -203,7 +203,7 @@ invisible(lapply(colnames(Cwm[[1]]),function(trait){
   }))
 }))
 mtext("Community Weighted Means",line=0.5,adj=0,outer=TRUE,cex=1.1)
-mtext("Years since disturbance",side=1,line=1.2,adj=1,cex=0.9,outer=TRUE)
+mtext("Years since disturbance",side=1,line=1.2,adj=0.65,cex=0.9,outer=TRUE)
 }
 
 SeedMassProp<-function(SeedMass){

### 2018-02-23 WholePlotTrajectories.Rmd
Manuscript update

@@
 The trajectory of inventories along time was visualized with the euclidean distance in the two-dimensional ordination space to the 1989 inventory.
-The functional trajectories of the leaf and stem and data life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004].
+The functional trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004].
 Species seed mass correspond to 5 classes of increasing mass, the seed mass trajectories were reported by the proportion of each class recorded in the inventories.
@@
 
-Plots taxonomic diversity corresponded to species Richness and the Hill number translation of Shannon and Simpson indices. 
-These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q). 
-This order q grasps the balance between richness and evenness in the community as it determines the weight of common species in the diversity metric, with an growing emphasis on dominant species as q increases.
-The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on all functional and data life traits.
+The taxonomic diversity was assessed through Richness and the Hill number translation of Shannon and Simpson indices [@Hill1973]. 
+To tackle the unequal number of recruited trees among treatments the indices bias corrected estimator were used, following [@chao2015estimating;@Marcon2015b].
+These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q), which proved well suited for diversity studies [@Patil1982;@Tothmeresz1995].
+The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on all functional and life traits.
 
@@
 
-```{r Fig3, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, *L_thickness*, chlorophyll content, *L_chloro*, toughness, *L_toughness* and specific area, *SLA*), 2 stem traits (wood specific gravity, *WD*, and bark thickness, *Bark-thick*) and one data life trait (Specific maximum height at adult stage,*Hmax*).
+```{r Fig3, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, *L_thickness*, chlorophyll content, *L_chloro*, toughness, *L_toughness* and specific area, *SLA*), 2 stem traits (wood specific gravity, *WD*, and bark thickness, *Bark-thick*) and one life trait (Specific maximum height at adult stage,*Hmax*).
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.

### 2018-02-26 WholePlotTrajectories.Rmd
Manuscript update

@@
 They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate local climate, support carbon, water and nutrient cycles, and ensure cultural and human well-being. 
-The simultaneous increase of forests products demand and substantial climatic changes heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threaten the maintenance of communities structure, composition and functioning and the dynamics shaping them [@Anderson-Teixeira2013;@Sist2015].
+The simultaneous increase of forests products demand and substantial climatic changes heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threaten the maintenance of communities structure, composition and functioning and the dynamics that shape them [@Anderson-Teixeira2013;@Sist2015].
 
-In tropical forests, a constant range of disturbances determines the ecological rules that shape the communities, through the reallocation of resources (as light, heat and water fluxes), changes in interspecific relations, etc.
+In tropical forests, a constant range of disturbance determines the ecological rules shaping the communities through the reallocation of resources (as light, heat and water fluxes), changes in interspecific relations, etc.
 The response of communities to disturbance is then the cornerstone of tropical forests ecology, modelling and management [@White2001;@Chazdon2003a].
-Disturbance recovery have been largely studied through structural parameters as aboveground biomass, tree height or stem density that all reflect rapid and measurable changes of ecosystems functioning asses crucial issues about the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
+Disturbance recovery have been largely studied through structural parameters as aboveground biomass, tree height or stem density, which all reflect rapid and measurable changes of ecosystems functioning asses crucial issues about the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
 In this perspective tree species diversity is inevitable as it determines ecosystems productivity, stability and functioning through the interspecific complementarity that enhances resources use, nutrients storing, and hazards mitigation [@Tilman2014].
-Reveal diversity dynamics is besides all the more important that disturbance, modifying the environment and interspecific relations, suerly impacts trees composition and diversity.
+Reveal diversity dynamics is besides all the more important that disturbance, modifies the environment and interspecific relations and surely impacts trees composition and diversity [@CazzollaGatti2014].
 
@@
 Although communities recovered the core of their initial functioning, composition seemed permanently altered after intense disturbance.
-Specificaly, this would impair species contingent to undisturbed forests, threatening their maintenance, and run the risk of cornerstone species loss and unexpected ecological consequences [@Diaz2005;@Gardner2007].
+Specificaly, this would impair species contingent to undisturbed forests, threatening their maintenance, and run the risk to loose cornerstone species and trigger unexpected ecological consequences [@Diaz2005;@Gardner2007].
 

### 2018-03-12 WholePlotTrajectories.Rmd
Corrections Eric

@@
 keywords: [mot-clés, séparés par des virgules]
-journalinfo: Hal 00679993
+journalinfo: Hal xxx
 archive: DOI xxxx
 date: "`r format(Sys.time(), '%d %B %Y')`"
-preamble: >  # Pour coordonne les packages LateX (supplémentaires)
-  \usepackage{amsmath}
-bibliography: references.bib # .bib à activer seulement pour git et word format (pandoc fait la biblio)
-csl: 
+preamble: >
+  \usepackage{textcomp}
+  \DeclareUnicodeCharacter{B0}{\textdegree}
+  \hyphenation{sa-plings}
+  \usepackage{longtable,tabu}
+bibliography: references.bib
 lang: english # french
@@
 output:
-  bookdown::gitbook
-#   bookdown::pdf_book:
-#     base_format: EcoFoG::article
-#     keep_tex: TRUE
-#  bookdown::word_document2
-#Mettre un code dans les figures pour indiquer le code d'où elle provient. ../ pour référer à des sous dossiers
+  bookdown::pdf_book:
+    base_format: EcoFoG::article
+    keep_tex: TRUE
+  bookdown::gitbook: default
+  bookdown::word_document2: default
 ---
@@
 ```{r Options, echo=FALSE, message=FALSE}
-knitr::opts_chunk$set(cache=F, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), out.width = '40%',fig.align="center")
+knitr::opts_chunk$set(cache=F, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
 # Installation des packages si nécessaire et chargement
@@
 # Ajouter les packages nécessaires ici
+Library("kableExtra")
 source("Scripts/DrawingTools.R")
@@
 In this perspective tree species diversity is inevitable as it determines ecosystems productivity, stability and functioning through the interspecific complementarity that enhances resources use, nutrients storing, and hazards mitigation [@Tilman2014].
-Reveal diversity dynamics is besides all the more important that disturbance, modifies the environment and interspecific relations and surely impacts trees composition and diversity [@CazzollaGatti2014].
+Revealing diversity dynamics is besides all the more important that disturbance, modifies the environment and interspecific relations and surely impacts trees composition and diversity [@CazzollaGatti2014].
 
-Until now, although, short-term dynamics demonstrated negligible or even positive impacts on communities diversity in accordance with the intermediate disturbance hypothesis (IDH) [@Molino2001;@Kariuki2006a;@Berry2008a] the long-term response of communities composition and diversity to disturbance remains unclear.
-The IDH predicting a maximized species diversity at intermediate level of disturbance have besides mainly been tested through species richness, which gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+Until now, although, short-term dynamics demonstrated negligible or even positive impacts on communities diversity in accordance with the intermediate disturbance hypothesis (IDH) [@Molino2001;@Kariuki2006a;@Berry2008a].
+The long-term response of communities composition and diversity to disturbance remains unclear.
+The IDH predicting a maximized species diversity at intermediate level of disturbance has besides mainly been tested through species richness, which gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
 Studies of communities response to disturbance should encompass taxonomic composition and evenness, which only can assess some conservation issues and the ecological rules that determine communities abundance distribution [@Magurran1988;@Lavorel2002;@Bellwood2006].
@@
 Communities dynamics would then be reflected by a shift from "conservative" species displaying long-lived, dense tissues growing slowly but with scarce resources, to "acquisitive" species with fast-growing, light tissues allowing fast resources acquisition [@TerSteege2001;@Reich2014;@Herault2011].
-In this perspective a combination of resource acquisition traits (leaf area,density and chlorophyll content and stem specific gravity and bark thickness) and life history traits of growth and reproduction (seed mass and maximum height) already proved relevant [@Wright2004;@Westoby2006a;@Chave2009b].
+In this perspective a combination of resource acquisition traits (leaf area, density and chlorophyll content and stem specific gravity and bark thickness) and life history traits of growth and reproduction (seed mass and maximum height) already proved relevant [@Wright2004;@Westoby2006a;@Chave2009b].
 
-Here we investigated over 30 years the response of 75ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
+Here we investigated over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
 We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw communities taxonomic and functional composition and diversity trajectories after disturbance.
@@
 
-### Study site
+## Study site
 
@@
 Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
-The experiment corresponds to a network of twelve 6.25ha plots that have undergone a gradient of three logging, thinning and fuelwood treatments \@ref(tab:Tab1).
+The experiment corresponds to a network of twelve 6.25ha plots that have undergone a gradient of three logging, thinning and fuelwood treatments (Table \@ref(tab:Tab1)).
 Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
@@
 
-Table : (\#tab:Tab1) Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
+```{r Tab1, echo=FALSE}
+Intervention <- data.frame(
+  Treatment=c("Control", "T1", "T2", "T3"),
+  Timber=c("", 
+           "DBH $\\geq$ 50 cm, commercial species, ~ 10 trees/ha", 
+           "DBH $\\geq$ 50 cm, commercial species, ~ 10 trees/ha",
+           "DBH $\\geq$ 50 cm, commercial species, ~ 10 trees/ha"),
+  Thinning=c("", 
+             "", 
+             "DBH $\\geq$ 40 cm, non-valuable species, ~ 30 trees/ha", 
+             "DBH $\\geq$ 50 cm, non-valuable species, ~ 15 trees/ha  "),
+  Fuelwood=c("", 
+             "", 
+             "", 
+             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-valuable species, ~ 15 trees/ha"),
+  AGBLost=c(0, "$[12\\%-33\\%]$", "$[33\\%-56\\%]$", "$[35\\%-56\\%]$")
+)
+names(Intervention)[5] <- "\\%AGB lost"
+knitr::kable(Intervention, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.", longtable = TRUE, booktabs = TRUE, escape = FALSE) %>%
+  kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
+```
 
-------------------------------------------------------------------------------------------------------------------------
-                          ***Timber***              ***Thinning***          ***Fuelwood***           **%AGB lost**
-
-------------------  ------------------------ ------------------------- ------------------------ ------------------------
-    ***Control***                                                                                          0
-
-         ***T1***    DBH ≥ 50 cm,                                                                                       
-                     commercial species,                                                                [12%-33%]         
-                     ~ 10 trees/ha
-
-         ***T2***    DBH ≥ 50 cm,              DBH ≥ 40 cm,                                                                 
-                     commercial species,       non-valuable species,                                    [33%-56%]         
-                     ~ 10 trees/ha             ~ 30 trees/ha
-
-         ***T3***    DBH ≥ 50 cm,              DBH ≥ 50 cm,               40 cm ≤ DBH ≤ 50 cm                               
-                     commercial species,       non-valuable species,      non-valuable species,         [35%-56%]         
-                     ~ 10 trees/ha             ~ 15 trees/ha              ~ 15 trees/ha                                     
-------------------------------------------------------------------------------------------------------------------------
-
-### Inventories protocol and dataset correction
+## Inventories protocol and dataset correction
 
@@
 The uncertainty propagation was done by the replenishment of inventories completed at genus level from real incomplete ones on the basisof vernacular/botanical names association. 
-Vernacular names were replaced through multinomial trials (M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) ) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
+Vernacular names were replaced through multinomial trials (M~v~([*s~1~, s~2~, …, s~N~*] ,[*$\alpha$~1~, $\alpha$~2~,…, $\alpha$~N~*]) ) based on the observed association probability [*$\alpha$~1~, $\alpha$~2~,…, $\alpha$~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
 To avoid remaining identification caveats and consider complete inventories, the simulated inventories were then reported at genus.
@@
 The functional diversity metrics used a dataset for 6 functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, the leaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass).
-Traits database came from the BRIDGE project (http://www.ecofog.gf/Bridge/index.html) where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou.
+The trait database came from the BRIDGE project [^1] where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou. 
 Missing trait values were filled using multivariate imputation by chained equation (mice) restricted to samples pertaining to the next higher taxonomic level, in order to account for the phylogenetic signal of the functional traits.
-The dataset comprised 294 botanical species pertaining to 157 botanical genus. 
+The dataset comprised 294 botanical species pertaining to 157 botanical genera. 
 Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level. 
-As seed mass information correspond to a classification into mass classes, no data filling process was applied so analysis were performed considering the 414 botanical species of the seed mass dataset.
-All composition and diversity metrics corresponded to the expectancy obtained after 50 iterations of the taxonomic uncertainty propagation framework and of the filling process of missing trait values.
+As seed mass information corresponds to a classification into mass classes, no data filling process was applied so analysis were performed considering the 414 botanical species of the seed mass dataset.
+All composition and diversity metrics corresponded to the average obtained after 50 iterations of the taxonomic uncertainty propagation framework and of the filling process of missing trait values.
 
-### Composition and diversity metrics
+[^1]: http://www.ecofog.gf/Bridge/index.html
 
-To counter the remaining taxonomic uncertainty (see previously) plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
+## Composition and diversity metrics
+
+To counter the remaining taxonomic uncertainty plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
 The analysis of taxonomic and functional composition of plots along time was visualized in a two-dimensional ordination space based on non-metric dimensional scaling of succesive floristic or functional inventories.
@@
 
-### Taxonomic and functional composition
+## Taxonomic and functional composition
 
-*Composition trajectories*
+### Composition trajectories
 
@@
 Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in the two-dimensional space from NMDS analysis based on flora inventories and the corresponding functional composition.
-Classifications were performed using either abundance-based Bray-Curtis \@ref(fig:Fig1) or incidence-based Jaccard dissimilarity (data not shown).
+Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:Fig1)) or incidence-based Jaccard dissimilarity (data not shown).
 
 Both taxonomic and functional composition were substantially affected by disturbance, and the maximum dissimilarity to 1984 reference inventory composition was positively correlated with the disturbance intensity.
-Until 30 years after disturbance, all disturbed plots taxonomic composition remained significantly dissimilar to that of the 1989 reference inventory composition \@ref(fig:Fig1).
-All plots, though, displayed a unimodal trajectory with a the return toward the taxonomic and functional composition of the 1989 reference inventory suggesting a shift towards a cyclic regime \@ref(fig:Fig2).
-For taxonomic composition, the maximum dissimilarity to 1984 inventory composition was reach around 20 years after disturbance for the T2 and T3 plots and for one T1 plot while control and other T1 plots continued to increase \@ref(fig:Fig2.a).
-For functional composition, the maximum dissimilarity to 1984 inventory composition was reach from 15 to 20 years after disturbance for the T2 and T3 plots while T1 and control plots continued to increase \@ref(fig:Fig2.b).
-The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance \@ref(fig:Fig2).
+Until 30 years after disturbance, all disturbed plots taxonomic composition remained significantly dissimilar to that of the 1989 reference inventory composition (Figure \@ref(fig:Fig1)).
+All plots, though, displayed a unimodal trajectory with a the return toward the taxonomic and functional composition of the 1989 reference inventory suggesting a shift towards a cyclic regime (Figure \@ref(fig:Fig2)).
+For taxonomic composition, the maximum dissimilarity to 1984 inventory composition was reach around 20 years after disturbance for the T2 and T3 plots and for one T1 plot while control and other T1 plots continued to increase (Figure \@ref(fig:Fig2).a).
+For functional composition, the maximum dissimilarity to 1984 inventory composition was reach from 15 to 20 years after disturbance for the T2 and T3 plots while T1 and control plots continued to increase (Figure \@ref(fig:Fig2).b).
+The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance (Figure \@ref(fig:Fig2)).
 
@@
 Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3)."}
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).", fig.env="figure*"}
 load("DB/FunctionalComposition_forGraph")
@@
 
-*Traits community weighted means*
+### Traits community weighted means
 
-For all plots the trajectories of the community weighted means (CWM) were drawn for the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area,wood specific gravity and barck thickness and seed mass and maximum adult height) \@ref(fig:Fig3).
+For all plots the trajectories of the community weighted means (CWM) were drawn for the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area,wood specific gravity and barck thickness and seed mass and maximum adult height) (Figure \@ref(fig:Fig3)).
 
@@
 
-Except for leaves cholorphyll content, which displayed higher difference between plots than among treatments which may be due to the completeness of dataset, all CWM trajectories corresponded to significant changes after disturbance.
-All functional traits and seed mass proportions displayed a unimodal trajectories but with different times at maximum and different values at the end of the experiment (30 years after disturbance). 
-30 years after disturbance, the weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value and than these of the control plots \@ref(fig:Fig3).
+Except for leaf cholorphyll content, which displayed higher difference between plots than among treatments which may be due to the completeness of dataset, all CWM trajectories corresponded to significant changes after disturbance.
+All functional traits and seed mass proportions displayed a unimodal trajectories but with different times at maximum and different values 30 years after disturbance. 
+The weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value and than these of the control plots (Figure \@ref(fig:Fig3)).
 The weighted means of bark thickness (*Bark_thick*) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value and this of the undisturbed plots at the end of the experiment.
@@
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3."}
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
 load("DB/CWM")
@@
 No gap filling process was applied in this case.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3."}
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
 load("DB/Smass_ClassProp")
@@
 
-### Disturbance impact on diversity
+## Disturbance impact on diversity
 
-*Taxonomic diversity*
+### Taxonomic diversity
 
-Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1989 inventories (5 years after disturbance) \@ref(fig:Fig5). 
+Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1989 inventories (5 years after disturbance) (Figure \@ref(fig:Fig5)). 
 *For the three indices the OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and a covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01).*
@@
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3."}
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
 load("DB/ReplacementTraj_ForGraphs")
@@
 
-While for undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity \@ref(fig:Fig5).
-The Richness was less affected by the initial disturbance except for the higher intensity (T3 plots) \@ref(fig:Fig5). 
-The Shannon and Simpson diversities however first significantly increased, illustrating an increasing evenness, before reaching a maximum around 15 years after disturbance for all plots except one from T1 treatment and one from T2 treatment \@ref(fig:Fig5) .
+While for undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity (Figure \@ref(fig:Fig5)).
+The Richness was less affected by the initial disturbance except for the higher intensity (T3 plots) (Figure \@ref(fig:Fig5)). 
+The Shannon and Simpson diversities however first significantly increased, illustrating an increasing evenness, before reaching a maximum around 15 years after disturbance for all plots except one from T1 treatment and one from T2 treatment (Figure \@ref(fig:Fig5)).
 
-*Functional  diversity*
+### Functional diversity
 
-For all undisturbed plots the Rao diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity \@ref(fig:Fig6).
+For all undisturbed plots the Rao diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity (Figure \@ref(fig:Fig6)).
 Although a first increase of plots functional diversity that was all the more important than initial disturbance was intense, 30 after disturbance the diversity was similar among disturbed plots and significantly higher than this of control plots.
@@
 
-```{r Fig6, fig.asp=1/1, echo=FALSE,fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
+```{r Fig6, echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 Disturbance regimes brought a unimodal response depending on disturbance intensity as suggested by the outcome of 10 years and 20 years post-disturbance analysis[@Molino2001;@Baraloto2012a].
-As observed after logging in a Boranean rainforest [@Cannon1998] or after cyclones in the Solomon islands [@Burslem2000], communities diversity showed a unimodal response to disturbance and a maintenance of diversity levels, except for one of the intensively disturbed community which evenness was still increasing \@ref(fig:Fig5).
-Despite the recovery of communities diversity, though, taxonomic composition was clearly altered \@ref(fig:Fig3) implying a long-term, if not permenant, composition change after disturbance as observed in East African tropical forests [@Bonnell2011].
+As observed after logging in a Boranean rainforest [@Cannon1998] or after cyclones in the Solomon islands [@Burslem2000], communities diversity showed a unimodal response to disturbance and a maintenance of diversity levels, except for one of the intensively disturbed community which evenness was still increasing (Figure \@ref(fig:Fig5)).
+Despite the recovery of communities diversity, though, taxonomic composition was clearly altered (Figure \@ref(fig:Fig3)) implying a long-term, if not permenant, composition change after disturbance as observed in East African tropical forests [@Bonnell2011].
 
-Regarding the functionl trajectories and dominant trait values of disturbed communities \@ref(fig:Fig3), disturbance appeared to favor acquisitive strategies, *i.e* species with high SLA, low wood density, small seeds, etc [@Westoby1998;@Wright2004;@Reich2014].
+Regarding the functional trajectories and dominant trait values of disturbed communities (Figure \@ref(fig:Fig3)), disturbance appeared to favor acquisitive strategies, *i.e* species with high SLA, low wood density, small seeds, etc [@Westoby1998;@Wright2004;@Reich2014].
 The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
 This means a shift from exclusive competition favoring competitive species before disturbance to environmental filtering selecting acquisitive and disturbance resistant species.
-In particular, the evolution seed mass classes proportions \@ref(fig:Fig4) revealed a clear increase in small seeded, therefore large dispersive, species [@TerSteege2001;@Flores2006;@Haddad2008].
+In particular, the evolution seed mass classes proportions (Figure \@ref(fig:Fig4)) revealed a clear increase in small seeded, therefore large dispersive, species [@TerSteege2001;@Flores2006;@Haddad2008].
 
 The diversity and composition functional trajectories showed a decoupling between functional and taxonomic after the most intense disturbance (treatments T2 and T3).
-Plots functional characteristic returned to their initial values, similar to those of control plots, faster than their taxonomic counterparts \@ref(fig:Fig2;fig:Fig3;fig:Fig4).
+Plots functional characteristic returned to their initial values, similar to those of control plots, faster than their taxonomic counterparts (Figures \@ref(fig:Fig2) to \@ref(fig:Fig4)).
 As functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered [@Guariguata2001].
@@
 
--	Important differences among plots of same treatments: possible impact of initial diversity composition, local abiotic parameters (topography, soils, surrounding stands).
+>	Important differences among plots of same treatments: possible impact of initial diversity composition, local abiotic parameters (topography, soils, surrounding stands).
 Impacts depend on the disturbance intensity, but further parameters may be involved.
@@
 `r if (knitr:::is_html_output()) '# References {-}'`
-
----
-bibliography: ["references.bib"]
-biblio-style: "apalike"
-link-citations: true
----

### 2018-03-13 WholePlotTrajectories.Rmd
Latests update before merge

--- 2018-02-26 WholePlotTrajectories.Rmd
@@
 
+```{r kable}
+textExp<- c(" ","Timber", "Thinning", "Fuelwood", "\\%AGB lost", 
+         "T1","DBH ≥ 50 cm,commercial species,~ 10 trees/ha"," "," ","[12\\%-33\\%]",
+         "T2","DBH ≥ 50 cm,commercial species,~ 10 trees/ha","DBH ≥ 40 cm,non-valuable species,~ 30 trees/ha",
+            " ","[33\\%-56\\%]",
+         "T3","DBH ≥ 50 cm,commercial species,~ 10 trees/ha","DBH ≥ 50 cm,non-valuable species,~ 15 trees/ha",
+         "40 cm ≤ DBH ≤ 50 cm,non-valuable species,~ 15 trees/ha","[35\\%-56\\%]",)
+knitr::kable(textExp, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou", booktabs = TRUE)
+```
+
 ### Inventories protocol and dataset correction
@@
 The uncertainty propagation was done by the replenishment of inventories completed at genus level from real incomplete ones on the basisof vernacular/botanical names association. 
-Vernacular names were replaced through multinomial trials (M~v~([*s~1~, s~2~, …, s~N~*] ,[*α~1~, α~2~,…, α~N~*]) ) based on the observed association probability [*α~1~, α~2~,…, α~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
+Vernacular names were replaced through multinomial trials (M~v~([*s~1~, s~2~, …, s~N~*] ,[*\alpha~1~, \alpha~2~,…, \alpha~N~*]) ) based on the observed association probability [*\alpha~1~, \alpha~2~,…, \alpha~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
 To avoid remaining identification caveats and consider complete inventories, the simulated inventories were then reported at genus.
@@
 The functional diversity metrics used a dataset for 6 functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, the leaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass).
-Traits database came from the BRIDGE project (http://www.ecofog.gf/Bridge/index.html) where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou.
-Missing trait values were filled using multivariate imputation by chained equation (mice) restricted to samples pertaining to the next higher taxonomic level, in order to account for the phylogenetic signal of the functional traits.
+Traits database came from the BRIDGE project [^1] where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou. 
+Missing trait values were filled using multivariate imputation by chained equation (mice) restricted to samples pertaining to the next higher taxonomic level, in order to account for the phylogenetic signal of functional traits.
 The dataset comprised 294 botanical species pertaining to 157 botanical genus. 
@@
 All composition and diversity metrics corresponded to the expectancy obtained after 50 iterations of the taxonomic uncertainty propagation framework and of the filling process of missing trait values.
+
+[^1] http://www.ecofog.gf/Bridge/
 
@@
 The taxonomic diversity was assessed through Richness and the Hill number translation of Shannon and Simpson indices [@Hill1973]. 
-To tackle the unequal number of recruited trees among treatments the indices bias corrected estimator were used, following [@chao2015estimating;@Marcon2015b].
+To tackle the unequal number of recruited trees among treatments the indices bias corrected estimator were used, following [@Chao2015;@Marcon2015b].
 These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q), which proved well suited for diversity studies [@Patil1982;@Tothmeresz1995].
@@
 
-*The distance from the initial condition was significantly time-dependent (P<0.01).* 
-*The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).*
+>The distance from the initial condition was significantly time-dependent (P<0.01).
+>The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).
 
@@
 In line with the assumtions of the IDH the observed diversity trajectories confirmed the increase of communities evenness and functional richness after disturbance.
-The consistency of the IDH has been debated for the higly diverse tropical forests, some observing the importance of environmental filters to shape communities assembly and others arguing for a genuine stochastic recovery [@Hubbell1999;@Molino2001].
-Consistently to the trajectories drawn here, a previous study of the same experiment revealed the directional changes in communities composition reflecting environmental filtering after disturbance [@Molino2001].
-Disturbance regimes brought a unimodal response depending on disturbance intensity as suggested by the outcome of 10 years and 20 years post-disturbance analysis[@Molino2001;@Baraloto2012a].
+The consistency of the IDH has been debated for the higly diverse tropical forests, some observing the importance of environmental filters to shape communities assembly and others arguing for a genuine stochastic recovery.
+Consistently to the trajectories drawn here, a previous study of the same experiment revealed the directional changes in communities composition reflecting environmental filtering after disturbance.
+Disturbance regimes brought a unimodal response depending on disturbance intensity as suggested by the outcome of 10 years and 20 years post-disturbance analysis[@Hubbell1999;@Molino2001;@Baraloto2012a].
 As observed after logging in a Boranean rainforest [@Cannon1998] or after cyclones in the Solomon islands [@Burslem2000], communities diversity showed a unimodal response to disturbance and a maintenance of diversity levels, except for one of the intensively disturbed community which evenness was still increasing \@ref(fig:Fig5).
@@
 Although communities recovered the core of their initial functioning, composition seemed permanently altered after intense disturbance.
-Specificaly, this would impair species contingent to undisturbed forests, threatening their maintenance, and run the risk to loose cornerstone species and trigger unexpected ecological consequences [@Diaz2005;@Gardner2007].
+Specificaly, this would impair species contingent to undisturbed forests, threatening their maintenance, and run the risk to loose cornerstone species and trigger unexpected ecological consequences [@Jones1994;@Diaz2005;@Gardner2007].
 

### 2018-03-13 WholePlotTrajectories.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Article EcoFoG
% Version 2.1 (23/10/2017)
%
% adapté de :
% Stylish Article
% LaTeX Template
% Version 1.0 (31/1/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Mathias Legrand (legrand.mathias@gmail.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[fleqn,10pt]{ArtEcoFoG} % Document font size and equations flushed left

\setcounter{tocdepth}{3} % Show only three levels in the table of contents section: sections, subsections and subsubsections

% Pandoc environments
\usepackage{framed}
\usepackage{fancyvrb}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, fontsize=\scriptsize} % Code R
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable,booktabs}
\usepackage{caption}
% These lines are needed to make table captions work with longtable:
\makeatletter
\def\fnum@table{\tablename~\thetable}
\makeatother
% longtable 2 columns
% https://tex.stackexchange.com/questions/161431/how-to-solve-longtable-is-not-in-1-column-mode-error
\makeatletter
\let\oldlt\longtable
\let\endoldlt\endlongtable
\def\longtable{\@ifnextchar[\longtable@i \longtable@ii}
\def\longtable@i[#1]{\begin{figure}[t]
\onecolumn
\begin{minipage}{0.5\textwidth}\scriptsize
\oldlt[#1]
}
\def\longtable@ii{\begin{figure}[t]
\onecolumn
\begin{minipage}{0.5\textwidth}\scriptsize
\oldlt
}
\def\endlongtable{\endoldlt
\end{minipage}
\twocolumn
\end{figure}}
\makeatother

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight0.8\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

% User-adder preamble
\usepackage{textcomp} \DeclareUnicodeCharacter{B0}{\textdegree}
\hyphenation{sa-plings} \usepackage{longtable,tabu}

%----------------------------------------------------------------------------------------
%	ARTICLE INFORMATION
%----------------------------------------------------------------------------------------

\JournalInfo{Hal 00679993} % Journal information
\Archive{DOI xxxx} % Additional notes (e.g. copyright, DOI, review/research article)

\PaperTitle{Post-Disturbance Tree Community Trajectories in a Neotropical Forest} % Article title

\Authors{
Ariane MIRABEL\textsuperscript{1*}\\ Eric Marcon\textsuperscript{1}\\ Bruno Hérault\textsuperscript{2 3}
} % Authors
\affiliation{
\textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles,
Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}Cirad, Univ montpellier, UR Forests \& Societies.\\ \hspace{1em} Montpellier, France.\\\textsuperscript{3}INPHB, Institut National Polytechnique Félix Houphouet-Boigny\\ \hspace{1em} Yamoussoukro, Ivory Coast.
}
\affiliation{*\textbf{Corresponding author}: ariane.mirabel@ecofog.gf, http://www.ecofog.gf/spip.php?article47} % Corresponding author

\Keywords{mot-clés, séparés par des virgules} % Keywords - if you don't want any simply remove all the text between the curly brackets
\newcommand{\keywordname}{Keywords} % Defines the keywords heading name

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\Abstract{
Résumé de l'article.
}

%----------------------------------------------------------------------------------------

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\selectlanguage{english}

\flushbottom % Makes all text pages the same height

\maketitle % Print the title and abstract box

\tableofcontents % Print the contents section

\thispagestyle{empty} % Removes page numbering from the first page

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}\label{introduction}

The large areas covered with tropical forests worldwide hold crucial
economic, social and cultural value. They provide wood and multiple
non-timber forest products, shelter a diversified fauna, regulate local
climate, support carbon, water and nutrient cycles, and ensure cultural
and human well-being. The simultaneous increase of forests products
demand and substantial climatic changes heightened the pressure on the
remaining forests \citep{Gibson2011a, Morales-Hidalgo2015} and threaten
the maintenance of communities structure, composition and functioning
and the dynamics that shape them
\citep{Anderson-Teixeira2013, Sist2015}.

In tropical forests, a constant range of disturbance determines the
ecological rules shaping the communities through the reallocation of
resources (as light, heat and water fluxes), changes in interspecific
relations, etc. The response of communities to disturbance is then the
cornerstone of tropical forests ecology, modelling and management
\citep{White2001, Chazdon2003a}. Disturbance recovery have been largely
studied through structural parameters as aboveground biomass, tree
height or stem density, which all reflect rapid and measurable changes
of ecosystems functioning asses crucial issues about the maintenance of
ecosystems processes and services
\citep{Denslow2000, Blanc2009, Rutishauser2016}. In this perspective
tree species diversity is inevitable as it determines ecosystems
productivity, stability and functioning through the interspecific
complementarity that enhances resources use, nutrients storing, and
hazards mitigation \citep{Tilman2014}. Revealing diversity dynamics is
besides all the more important that disturbance, modifies the
environment and interspecific relations and surely impacts trees
composition and diversity \citep{CazzollaGatti2014}.

Until now, although, short-term dynamics demonstrated negligible or even
positive impacts on communities diversity in accordance with the
intermediate disturbance hypothesis (IDH)
\citep{Molino2001, Kariuki2006a, Berry2008a}. The long-term response of
communities composition and diversity to disturbance remains unclear.
The IDH predicting a maximized species diversity at intermediate level
of disturbance has besides mainly been tested through species richness,
which gives limited or misleading information on forests recovery and
functioning \citep{Martin2015, Chaudhary2016}. Studies of communities
response to disturbance should encompass taxonomic composition and
evenness, which only can assess some conservation issues and the
ecological rules that determine communities abundance distribution
\citep{Magurran1988, Lavorel2002, Bellwood2006}. Beyond taxonomic
metrics, communities recovery should also encompass functional diversity
metrics \citep{Moretti2009, Baraloto2012a} through functional traits,
that are species biological attributes relating to individuals fitness
and ecosystem properties \citep{Violle2007b, Scheiter2013}. Functional
traits were largely adopted and integrated within a workable framework
to assess forests dynamics and functioning
\citep{Diaz2005, Villeger2008a} and a vast litterature revealed some
major functional traits determinant for species ecology and performance.
These major traits proved relevant to assess the ecological rules
shaping species assembly: post-disturbance trajectories in tropical
rainforests, for example, are assumed to be driven by environmental
filters fostering disturbance resistant species with rapid growth and
efficient resources acquisition \citep{Molino2001, Haddad2008}.
Communities dynamics would then be reflected by a shift from
``conservative'' species displaying long-lived, dense tissues growing
slowly but with scarce resources, to ``acquisitive'' species with
fast-growing, light tissues allowing fast resources acquisition
\citep{TerSteege2001, Reich2014, Herault2011}. In this perspective a
combination of resource acquisition traits (leaf area, density and
chlorophyll content and stem specific gravity and bark thickness) and
life history traits of growth and reproduction (seed mass and maximum
height) already proved relevant
\citep{Wright2004, Westoby2006a, Chave2009b}.

Here we investigated over 30 years the response of 75 ha of forests
plots set up on a gradient of disturbance intensity, from 10 to 60\% of
ecosystem biomass removed. We made use of a large functional traits
database browsing major leaf, stem and seed traits and species maximum
height to draw communities taxonomic and functional composition and
diversity trajectories after disturbance. Specifically, we (i) tested
the validity of the IDH arguing for enhanced diversity after moderate
disturbance and (ii) confirmed the changes in ecological rules shaping
communities, from competitive exclusion between ``conservative'',
slow-growing tree species to environmental filtering of ``acquisitive'',
fasting-growing ones. Eventually we examined the different trajectories
to (iii) resolve the resilience of tropical forests and question their
future dynamics regarding their state 30 years after disturbance.

\section{Material and Methods}\label{material-and-methods}

\subsection{Study site}\label{study-site}

Paracou station in French Guiana (5°18'N and 52°53'W) is located in a
lowland tropical rain forest corresponding to a tropical wet climate
with mean annual precipitation averaging 2980 mm.y\textsuperscript{-1}
(30-y period) with a 3-months dry season (\textless{} 100
mm.months\textsuperscript{-1}) from mid-August to mid-November, and a
one-month dry season in March \citep{Wagner2011}. Elevation ranges
between 5 and 50 m and mean annual temperature is 26°C. Soils correspond
to thin acrisols over a layer of transformed saprolite with low
permeability generating lateral drainage during heavy rains. The
experiment corresponds to a network of twelve 6.25ha plots that have
undergone a gradient of three logging, thinning and fuelwood treatments
(Table \ref{tab:Tab1}). Disturbance treatments were attributed according
to a randomized plot design with three replicate blocks of four plots.
The disturbance corresponds to averages of 10 trees removed per hectare
with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1),
32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40
cm DBH for treatment 3 (T3). Treatments T2 and T3 besides included the
thinning of trees by poison girdling \citep{Blanc2009}.

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.}\\
\toprule
Treatment & Timber & Thinning & Fuelwood & \%AGB lost\\
\midrule
Control &  &  &  & 0\\
T1 & DBH $\geq$ 50 cm, commercial species, ~ 10 trees/ha &  &  & $[12\%-33\%]$\\
T2 & DBH $\geq$ 50 cm, commercial species, ~ 10 trees/ha & DBH $\geq$ 40 cm, non-valuable species, ~ 30 trees/ha &  & $[33\%-56\%]$\\
T3 & DBH $\geq$ 50 cm, commercial species, ~ 10 trees/ha & DBH $\geq$ 50 cm, non-valuable species, ~ 15 trees/ha & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species, ~ 15 trees/ha & $[35\%-56\%]$\\
\bottomrule
\end{longtabu}

\subsection{Inventories protocol and dataset
correction}\label{inventories-protocol-and-dataset-correction}

The study site corresponds to a tropical rainforest with a dominance of
Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical
families. In the twelve experimental plots of the experiment, all trees
above 10 cm DBH were mapped and measured annually since 1984. During
inventories, trees were first identified with a vernacular name assigned
by the field team, and afterward with a scientific name assigned by a
botanist during regular botanical campaigns. In 1984, specific
vernacular names were given to 62 commercial or common species whereas
other less common species were identified under two identifiers only
separating trees and palm trees. The botanical campaigns carried every 5
to 6 years to identify all trees at the species level only started in
2003 and identification practices varied among plots and successive
campaigns. This raised methodological issues as vernacular names usually
correspond to different botanical species, resulting in significant
taxonomic uncertainties that had to be propagated to composition and
diversity metrics throught a Bayesian framework. The uncertainty
propagation was done by the replenishment of inventories completed at
genus level from real incomplete ones on the basisof
vernacular/botanical names association. Vernacular names were replaced
through multinomial trials
(M\textsubscript{v}({[}\emph{s\textsubscript{1}, s\textsubscript{2},
\ldots{}, s\textsubscript{N}}{]} ,{[}\emph{\(\alpha\)\textsubscript{1},
\(\alpha\)\textsubscript{2},\ldots{}, \(\alpha\)\textsubscript{N}}{]}) )
based on the observed association probability
{[}\emph{\(\alpha\)\textsubscript{1},
\(\alpha\)\textsubscript{2},\ldots{}, \(\alpha\)\textsubscript{N}}{]}
between each vernacular name \emph{v} and the species
{[}\emph{s\textsubscript{1}, s\textsubscript{2}, \ldots{},
s\textsubscript{N}}{]} recorded in the inventory. To avoid remaining
identification caveats and consider complete inventories, the simulated
inventories were then reported at genus. See appendix 1 and
\citet{Aubry-Kientz2013} for the detailed methodology.

The functional diversity metrics used a dataset for 6 functional traits
representing leaf economics (leaves thickness, toughness, total
chlorophyll content and specific leaf area, the leaf area per unit dry
mass) and wood economics (wood specific gravity and bark thickness), and
life history traits (maximum specific height and seed mass). The trait
database came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/index.html}
where traits values were assessed from a selection of individuals
located in nine permanent plots in French Guiana, including two in
Paracou. Missing trait values were filled using multivariate imputation
by chained equation (mice) restricted to samples pertaining to the next
higher taxonomic level, in order to account for the phylogenetic signal
of the functional traits. The dataset comprised 294 botanical species
pertaining to 157 botanical genera. Whenever a species inventoried was
not in the dataset, it was attributed a set of traits values randomly
sampled among species of the same next higher taxonomic level. As seed
mass information corresponds to a classification into mass classes, no
data filling process was applied so analysis were performed considering
the 414 botanical species of the seed mass dataset. All composition and
diversity metrics corresponded to the average obtained after 50
iterations of the taxonomic uncertainty propagation framework and of the
filling process of missing trait values.

\subsection{Composition and diversity
metrics}\label{composition-and-diversity-metrics}

To counter the remaining taxonomic uncertainty plots taxonomic
composition and diversity were analysed at the genus level, \emph{i.e.}
referring to the genus of observed or trialed botanical names. The
analysis of taxonomic and functional composition of plots along time was
visualized in a two-dimensional ordination space based on non-metric
dimensional scaling of succesive floristic or functional inventories.
Plots trajectories along time was reported comparatively to the
inventories in 1989, 5 years after disturbance, which corresponded to
first inventory with a sufficient degree of uncertainty (\textless{}30\%
of undetermined trees). The inventories dissimilarity compared to the
reference 1989 inventory was reported using occurrence-based (Jaccard)
and abundance-based (Bray-Curtis) similarity measures. The trajectory of
inventories along time was visualized with the euclidean distance in the
two-dimensional ordination space to the 1989 inventory. The functional
trajectories of the leaf and stem and life traits were also visualized
with the community weighted means (CWM), representing the average trait
value in a community weighted by relative abundance of the species
carrying each value \citep{Diaz2007, Garnier2004}. Species seed mass
correspond to 5 classes of increasing mass, the seed mass trajectories
were reported by the proportion of each class recorded in the
inventories. To determine whether communities' assembly deviated from
the expectation of a random process we compared the observed similarity
patterns to those generated by 50 repetitions of stochastic null models.
The null model for taxonomic composition randomly shuffled individuals
among plots while preserving the overall species abundance and plots'
tree density. The null model for functional composition produced random
traits assemblages by permuting the set of functional traits values
among species.

The taxonomic diversity was assessed through Richness and the Hill
number translation of Shannon and Simpson indices \citep{Hill1973}. To
tackle the unequal number of recruited trees among treatments the
indices bias corrected estimator were used, following
\citep{chao2015estimating, Marcon2015b}. These three indices belong to
the set of HCDT or generalized entropy, respectively corresponding to
the 0, 1 and 2 order of diversity (q), which proved well suited for
diversity studies \citep{Patil1982, Tothmeresz1995}. The functional
diversity was reported using the Rao index of quadratic entropy which
combines species abundance distribution and average pairwise
dissimilarity based on all functional and life traits.

\section{Results}\label{results}

\subsection{Taxonomic and functional
composition}\label{taxonomic-and-functional-composition}

\subsubsection{Composition trajectories}\label{composition-trajectories}

Over time, 828388 trees and 591 botanical species pertaining to 223
genus and 64 botanical families were recorded. Trajectories of taxonomic
and functional composition after disturbance were examined through the
ordination of successive inventories from 1989 (5 years after
disturbance) in the two-dimensional space from NMDS analysis based on
flora inventories and the corresponding functional composition.
Classifications were performed using either abundance-based Bray-Curtis
(Figure \ref{fig:Fig1}) or incidence-based Jaccard dissimilarity (data
not shown).

Both taxonomic and functional composition were substantially affected by
disturbance, and the maximum dissimilarity to 1984 reference inventory
composition was positively correlated with the disturbance intensity.
Until 30 years after disturbance, all disturbed plots taxonomic
composition remained significantly dissimilar to that of the 1989
reference inventory composition (Figure \ref{fig:Fig1}). All plots,
though, displayed a unimodal trajectory with a the return toward the
taxonomic and functional composition of the 1989 reference inventory
suggesting a shift towards a cyclic regime (Figure \ref{fig:Fig2}). For
taxonomic composition, the maximum dissimilarity to 1984 inventory
composition was reach around 20 years after disturbance for the T2 and
T3 plots and for one T1 plot while control and other T1 plots continued
to increase (Figure \ref{fig:Fig2}.a). For functional composition, the
maximum dissimilarity to 1984 inventory composition was reach from 15 to
20 years after disturbance for the T2 and T3 plots while T1 and control
plots continued to increase (Figure \ref{fig:Fig2}.b). The functional
composition distancing from the reference inventory was also positively
correlated with disturbance intensity and similarly stabilized or
reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance
(Figure \ref{fig:Fig2}).

\emph{The distance from the initial condition was significantly
time-dependent (P\textless{}0.01).} \emph{The coordinates of the
functional traits were measured in the two-dimensional ordination space
that mapped plots evolution along time (graphs not showed) which
revealed that the trajectories of disturbed plots along time headed
toward acquisitive functional strategies (from high SWG to high SLA and
chlorophyll content).}

\begin{figure*}

{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig1-1}

}

\caption{Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance. Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).}\label{fig:Fig1}
\end{figure*}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig2-1}

}

\caption{Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of **(a)** taxonomic composition at genus level and **(b)** functional composition. Distance are abundance-based Bary-Curtis metric. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3). The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes. }\label{fig:Fig2}
\end{figure}

\subsubsection{Traits community weighted
means}\label{traits-community-weighted-means}

For all plots the trajectories of the community weighted means (CWM)
were drawn for the 8 functional and life history traits (Leaf thickness,
chlorophyll content, toughness and specific area,wood specific gravity
and barck thickness and seed mass and maximum adult height) (Figure
\ref{fig:Fig3}).

\begin{quote}
To compensate the intrinsinc difference among plots the trajectories
drawn correspond to the difference in value between the reference
inventory in \textgreater{}1984 (5 year after disturbance) and the
successive years inventoried.
\end{quote}

Except for leaf cholorphyll content, which displayed higher difference
between plots than among treatments which may be due to the completeness
of dataset, all CWM trajectories corresponded to significant changes
after disturbance. All functional traits and seed mass proportions
displayed a unimodal trajectories but with different times at maximum
and different values 30 years after disturbance. The weighted means of
communities specific maximum height at adult stage (\emph{Hmax}), leaf
toughness (\emph{L\_toughness}) and wood specific gravity (\emph{WD})
remained significantly lower than their initial value and than these of
the control plots (Figure \ref{fig:Fig3}). The weighted means of bark
thickness (\emph{Bark\_thick}) similarly remained substantilly higher
than initially for all disturbed plots while the specfic leaf area
(\emph{SLA}) had almost recovered its initial value and this of the
undisturbed plots at the end of the experiment.

\begin{figure*}

{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig3-1}

}

\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, *L_thickness*, chlorophyll content, *L_chloro*, toughness, *L_toughness* and specific area, *SLA*), 2 stem traits (wood specific gravity, *WD*, and bark thickness, *Bark-thick*) and one life trait (Specific maximum height at adult stage,*Hmax*). Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig3}
\end{figure*}

\begin{figure*}

{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig4-1}

}

\caption{Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. No gap filling process was applied in this case. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig4}
\end{figure*}

\subsection{Disturbance impact on
diversity}\label{disturbance-impact-on-diversity}

\subsubsection{Taxonomic diversity}\label{taxonomic-diversity}

Trajectories of Richness, Shannon and Simpson taxonomic diversity were
examined in relation to the 1989 inventories (5 years after disturbance)
(Figure \ref{fig:Fig5}). \emph{For the three indices the OLS regression
confirmed the time-dependency of the diversity variations
(P\textless{}0.01) and a covariance analysis confirmed the significant
effect of the disturbance treatment (P\textless{}0.01).}

\begin{figure*}

{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig5-1}

}

\caption{Trajectories of the difference to reference inventory over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig5}
\end{figure*}

While for undisturbed plots the Richness, Shannon and Simpson diversity
remained comparable to the values in 1989, the trajectories of all
disturbed plots showed significant increase positively correlated with
the disturbance intensity (Figure \ref{fig:Fig5}). The Richness was less
affected by the initial disturbance except for the higher intensity (T3
plots) (Figure \ref{fig:Fig5}). The Shannon and Simpson diversities
however first significantly increased, illustrating an increasing
evenness, before reaching a maximum around 15 years after disturbance
for all plots except one from T1 treatment and one from T2 treatment
(Figure \ref{fig:Fig5}).

\subsubsection{Functional diversity}\label{functional-diversity}

For all undisturbed plots the Rao diversity remained comparable to the
values in 1989, the trajectories of all disturbed plots showed
significant increase positively correlated with the disturbance
intensity (Figure \ref{fig:Fig6}). Although a first increase of plots
functional diversity that was all the more important than initial
disturbance was intense, 30 after disturbance the diversity was similar
among disturbed plots and significantly higher than this of control
plots. \emph{The OLS regression confirmed the time-dependency of the
diversity variations (P\textless{}0.01) and the covariance analysis
confirmed the significant effect of the disturbance treatment
(P\textless{}0.01)}.

\begin{figure}

{\centering \includegraphics{WholePlotTrajectories_files/figure-latex/Fig6-1}

}

\caption{Trajectories of the Rao functional diversity over 30 years after disturbance. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:Fig6}
\end{figure}

\section{Discussion}\label{discussion}

Our study showed the significant impact of disturbance on tropical
forests communities. The subsequent diversity trajectories confirmed the
intermediate disturbance hypothesis debated for tropical forests through
their correlation with disturbance intensity. Besides it revealed the
contrasting response of taxonomic and functional characteristics,
specifically the decoupling between communities taxonomic evenness and
their functional diversity and dominant functional traits values. The
long-term disturbance trajectories observed highlighted the unachieved
but consistent recovery of communities assembly for the lowest
disturbance intensity but questioned it after higher disturbance.

In line with the ``vegetation quantity effect'' suggesting that
communities functional characteristics depends on dominant species
rather that subtle changes in biodiversity \citep{Grime1998}. This would
result in a faster recovery of ecosystem processes rates than
composition and diversity as demonstrated for grasslands
\citep{Tilman1997, Mouillot2011} and more recently for tropical forests
\citep{Lohbeck2015, Guariguata2001}. This highlighted the functional
overlap between species in tropical forests, typical of their huge
biodiversity, and intricately questioned the different facets of
communities resilience. Ecosystem processes grasped through the set of
functional traits retained here however do not assess communities
resilience and future dynamics.

In line with the assumtions of the IDH the observed diversity
trajectories confirmed the increase of communities evenness and
functional richness after disturbance. The consistency of the IDH has
been debated for the higly diverse tropical forests, some observing the
importance of environmental filters to shape communities assembly and
others arguing for a genuine stochastic recovery
\citep{Hubbell1999, Molino2001}. Consistently to the trajectories drawn
here, a previous study of the same experiment revealed the directional
changes in communities composition reflecting environmental filtering
after disturbance \citep{Molino2001}. Disturbance regimes brought a
unimodal response depending on disturbance intensity as suggested by the
outcome of 10 years and 20 years post-disturbance
analysis\citep{Molino2001, Baraloto2012a}. As observed after logging in
a Boranean rainforest \citep{Cannon1998} or after cyclones in the
Solomon islands \citep{Burslem2000}, communities diversity showed a
unimodal response to disturbance and a maintenance of diversity levels,
except for one of the intensively disturbed community which evenness was
still increasing (Figure \ref{fig:Fig5}). Despite the recovery of
communities diversity, though, taxonomic composition was clearly altered
(Figure \ref{fig:Fig3}) implying a long-term, if not permenant,
composition change after disturbance as observed in East African
tropical forests \citep{Bonnell2011}.

Regarding the functional trajectories and dominant trait values of
disturbed communities (Figure \ref{fig:Fig3}), disturbance appeared to
favor acquisitive strategies, \emph{i.e} species with high SLA, low wood
density, small seeds, etc \citep{Westoby1998, Wright2004, Reich2014}.
The comparison of observed trajectories to a null model with random
traits set shuffling showed an oriented functional shift after
disturbance which suggest additional biotic and abiotic filters for the
10 to 15 years after disturbance. This means a shift from exclusive
competition favoring competitive species before disturbance to
environmental filtering selecting acquisitive and disturbance resistant
species. In particular, the evolution seed mass classes proportions
(Figure \ref{fig:Fig4}) revealed a clear increase in small seeded,
therefore large dispersive, species
\citep{TerSteege2001, Flores2006, Haddad2008}.

The diversity and composition functional trajectories showed a
decoupling between functional and taxonomic after the most intense
disturbance (treatments T2 and T3). Plots functional characteristic
returned to their initial values, similar to those of control plots,
faster than their taxonomic counterparts (Figures \ref{fig:Fig2} to
\ref{fig:Fig4}). As functional traits are the most direct link between
biodiversity and ecosystem functioning \citep{Diaz2005} the recovery of
communities functional characteristics would go along the recovery of
ecosystems processes although taxonomic composition remain altered
\citep{Guariguata2001}. Such decoupling suggested the consistency of
functional overlap and maybe functional versatility, implying a variety
of functioning from the same characteristic, in the pre-disturbance
condition. Functional overlap has been observed in highly diverse
ecosystems \citep{Bellwood2006} and suggested as one mechanisms for high
diversity maintenance. Knowing the decoupling between taxonomic and
functional trajectories, this functional overlap may not be recovering
after disturbance and therefore radically modify communities resilience
and resistance \citep{Trenbath1999, Elmqvist2003, Diaz2005}. Given the
impaired functional overlap, the persitant compositional changes
regarding disturbance resistant species \citep{Haddad2008} and probably
lianas or epiphytes \citep{Martin2013}, and the modification of
environmental conditions like soils nutrient cycling and compaction
\citep{Olander2005} the entirety of communities recovery after intense
disturbance is deeply questioned \citep{Chazdon2003a}. New conditions
would not only be longer lasting but self-maintained as tied to
disturbance regime \citep{Burslem2000}. Although communities recovered
the core of their initial functioning, composition seemed permanently
altered after intense disturbance. Specificaly, this would impair
species contingent to undisturbed forests, threatening their
maintenance, and run the risk to loose cornerstone species and trigger
unexpected ecological consequences \citep{Diaz2005, Gardner2007}.

To confirm this long-term communities changes and highlight the
ecological rules shaping communities assembly, we suggest a focus on
recruitment dynamics, that would follow the diversity and functional
characteristic of recruited communities and, by extension, of the forest
to be.

\begin{quote}
Important differences among plots of same treatments: possible impact of
initial diversity composition, local abiotic parameters (topography,
soils, surrounding stands). Impacts depend on the disturbance intensity,
but further parameters may be involved.
\end{quote}

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\bibliographystyle{mee}
\makeatletter
% The filename has .bib extension the must be eliminated
\filename@parse{references.bib}
% parse stores the file name in base. Extension starts at the first dot, so don't use dots in file names.
\bibliography{\filename@base}
\makeatother

%----------------------------------------------------------------------------------------

\end{document}

### 2018-04-05 WholePlotTrajectories.tex
Scripts

@@
 
-\JournalInfo{Hal 00679993} % Journal information
+\JournalInfo{Hal xxx} % Journal information
 \Archive{DOI xxxx} % Additional notes (e.g. copyright, DOI, review/research article)
@@
 %----------------------------------------------------------------------------------------
+
+\begin{verbatim}
+## Warning: package 'kableExtra' was built under R version 3.3.3
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: package 'shape' was built under R version 3.3.3
+\end{verbatim}
 
@@
 \subsection{Inventories protocol and dataset
-correction}\label{inventories-protocol-and-dataset-correction}
+collection}\label{inventories-protocol-and-dataset-collection}
 
@@
 genus level from real incomplete ones on the basisof
-vernacular/botanical names association. Vernacular names were replaced
-through multinomial trials
-(M\textsubscript{v}({[}\emph{s\textsubscript{1}, s\textsubscript{2},
-\ldots{}, s\textsubscript{N}}{]} ,{[}\emph{\(\alpha\)\textsubscript{1},
-\(\alpha\)\textsubscript{2},\ldots{}, \(\alpha\)\textsubscript{N}}{]}) )
+vernacular/botanical names association.
+
+Vernacular names were replaced through multinomial trials
+\(M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)\)
 based on the observed association probability
-{[}\emph{\(\alpha\)\textsubscript{1},
-\(\alpha\)\textsubscript{2},\ldots{}, \(\alpha\)\textsubscript{N}}{]}
-between each vernacular name \emph{v} and the species
-{[}\emph{s\textsubscript{1}, s\textsubscript{2}, \ldots{},
-s\textsubscript{N}}{]} recorded in the inventory. To avoid remaining
-identification caveats and consider complete inventories, the simulated
-inventories were then reported at genus. See appendix 1 and
-\citet{Aubry-Kientz2013} for the detailed methodology.
+\(\big[\alpha_1, \alpha_2,…, \alpha_3\big]\) between each vernacular
+name \emph{v} and the species \(\big[s_1, s_2, …, s_N\big]\) recorded in
+the inventory.
+
+To avoid remaining identification caveats and consider complete
+inventories, the simulated inventories were then reported at genus. See
+appendix 1 and \citet{Aubry-Kientz2013} for the detailed methodology.
 
@@
 mass) and wood economics (wood specific gravity and bark thickness), and
-life history traits (maximum specific height and seed mass). The trait
-database came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/index.html}
-where traits values were assessed from a selection of individuals
-located in nine permanent plots in French Guiana, including two in
-Paracou. Missing trait values were filled using multivariate imputation
-by chained equation (mice) restricted to samples pertaining to the next
-higher taxonomic level, in order to account for the phylogenetic signal
-of the functional traits. The dataset comprised 294 botanical species
-pertaining to 157 botanical genera. Whenever a species inventoried was
-not in the dataset, it was attributed a set of traits values randomly
-sampled among species of the same next higher taxonomic level. As seed
-mass information corresponds to a classification into mass classes, no
-data filling process was applied so analysis were performed considering
-the 414 botanical species of the seed mass dataset. All composition and
-diversity metrics corresponded to the average obtained after 50
-iterations of the taxonomic uncertainty propagation framework and of the
-filling process of missing trait values.
+life history traits (maximum specific height and seed mass).
+
+The trait database came from the BRIDGE project {[}\^{}1{]} where traits
+values were assessed from a selection of individuals located in nine
+permanent plots in French Guiana, including two in Paracou. Missing
+trait values were filled using multivariate imputation by chained
+equation (mice) restricted to samples pertaining to the next higher
+taxonomic level, in order to account for the phylogenetic signal of the
+functional traits. The dataset comprised 294 botanical species
+pertaining to 157 botanical genera.
+
+Whenever a species inventoried was not in the dataset, it was attributed
+a set of traits values randomly sampled among species of the same next
+higher taxonomic level. As seed mass information corresponds to a
+classification into mass classes, no data filling process was applied so
+analysis were performed considering the 414 botanical species of the
+seed mass dataset. All composition and diversity metrics corresponded to
+the average obtained after 50 iterations of the taxonomic uncertainty
+propagation framework and of the filling process of missing trait
+values.
+
+{[}\^{}1{]} http://www.ecofog.gf/Bridge/
 
@@
 indices bias corrected estimator were used, following
-\citep{chao2015estimating, Marcon2015b}. These three indices belong to
-the set of HCDT or generalized entropy, respectively corresponding to
-the 0, 1 and 2 order of diversity (q), which proved well suited for
-diversity studies \citep{Patil1982, Tothmeresz1995}. The functional
-diversity was reported using the Rao index of quadratic entropy which
-combines species abundance distribution and average pairwise
-dissimilarity based on all functional and life traits.
+\citep{Chao2015, Marcon2015b}. These three indices belong to the set of
+HCDT or generalized entropy, respectively corresponding to the 0, 1 and
+2 order of diversity (q), which proved well suited for diversity studies
+\citep{Patil1982, Tothmeresz1995}. The functional diversity was reported
+using the Rao index of quadratic entropy which combines species
+abundance distribution and average pairwise dissimilarity based on all
+functional and life traits.
 
@@
 
-\emph{The distance from the initial condition was significantly
-time-dependent (P\textless{}0.01).} \emph{The coordinates of the
-functional traits were measured in the two-dimensional ordination space
-that mapped plots evolution along time (graphs not showed) which
-revealed that the trajectories of disturbed plots along time headed
-toward acquisitive functional strategies (from high SWG to high SLA and
-chlorophyll content).}
+\begin{quote}
+The distance from the initial condition was significantly time-dependent
+(P\textless{}0.01). The coordinates of the functional traits were
+measured in the two-dimensional ordination space that mapped plots
+evolution along time (graphs not showed) which revealed that the
+trajectories of disturbed plots along time headed toward acquisitive
+functional strategies (from high SWG to high SLA and chlorophyll
+content).
+\end{quote}
 
@@
 trajectories confirmed the increase of communities evenness and
-functional richness after disturbance. The consistency of the IDH has
-been debated for the higly diverse tropical forests, some observing the
-importance of environmental filters to shape communities assembly and
-others arguing for a genuine stochastic recovery
-\citep{Hubbell1999, Molino2001}. Consistently to the trajectories drawn
-here, a previous study of the same experiment revealed the directional
-changes in communities composition reflecting environmental filtering
-after disturbance \citep{Molino2001}. Disturbance regimes brought a
-unimodal response depending on disturbance intensity as suggested by the
-outcome of 10 years and 20 years post-disturbance
-analysis\citep{Molino2001, Baraloto2012a}. As observed after logging in
-a Boranean rainforest \citep{Cannon1998} or after cyclones in the
-Solomon islands \citep{Burslem2000}, communities diversity showed a
-unimodal response to disturbance and a maintenance of diversity levels,
-except for one of the intensively disturbed community which evenness was
-still increasing (Figure \ref{fig:Fig5}). Despite the recovery of
-communities diversity, though, taxonomic composition was clearly altered
-(Figure \ref{fig:Fig3}) implying a long-term, if not permenant,
+functional richness after disturbance.
+\textless{}\textless{}\textless{}\textless{}\textless{}\textless{}\textless{}
+HEAD The consistency of the IDH has been debated for the higly diverse
+tropical forests, some observing the importance of environmental filters
+to shape communities assembly and others arguing for a genuine
+stochastic recovery. Consistently to the trajectories drawn here, a
+previous study of the same experiment revealed the directional changes
+in communities composition reflecting environmental filtering after
+disturbance. Disturbance regimes brought a unimodal response depending
+on disturbance intensity as suggested by the outcome of 10 years and 20
+years post-disturbance
+analysis\citep{Hubbell1999, Molino2001, Baraloto2012a}. As observed
+after logging in a Boranean rainforest \citep{Cannon1998} or after
+cyclones in the Solomon islands \citep{Burslem2000}, communities
+diversity showed a unimodal response to disturbance and a maintenance of
+diversity levels, except for one of the intensively disturbed community
+which evenness was still increasing \ref{fig:Fig5}. Despite the recovery
+of communities diversity, though, taxonomic composition was clearly
+altered \ref{fig:Fig3} implying a long-term, if not permenant,
 composition change after disturbance as observed in East African
@@
 maintenance, and run the risk to loose cornerstone species and trigger
-unexpected ecological consequences \citep{Diaz2005, Gardner2007}.
+unexpected ecological consequences
+\citep{Jones1994, Diaz2005, Gardner2007}.
 
@@
 \bibliographystyle{mee}
-\makeatletter
-% The filename has .bib extension the must be eliminated
-\filename@parse{references.bib}
-% parse stores the file name in base. Extension starts at the first dot, so don't use dots in file names.
-\bibliography{\filename@base}
-\makeatother
+\bibliography{references.bib}
 

### 2018-04-05 Scripts/DivTaxo_trajectories.R

@@ -41,7 +41,7 @@ CompleteTaxo<-lapply(1:4,function(t){
         colnames(Div_plot)<-"names"
         Div_plot<-as.character(merge(Div_plot,RefBota,by.x="names",by.y="row.names",all.x=T)[,"Genre"])
          
-        return(expq(bcTsallis(as.AbdVector(tapply(Div_plot,Div_plot,length)),q=Q,Correction = "Best"),q=Q))
+        return(expq(bcTsallis(as.AbdVector(tapply(Div_plot,Div_plot,length)),q=Q,Correction = "None"),q=Q))
        })
         
         names(Div)<-names(yearPlot)

### 2018-04-05 Scripts/GithubPages.R

@@ -5,7 +5,7 @@
 # Source this file before committing
 
 # Create the docs folder
-dir.create("docs")
+if(!dir.exists("docs")){dir.create("docs")}
 
 # Copy knitted html files
 htmlFiles <- list.files(pattern="*.html")
@@ -14,9 +14,9 @@ if (length(htmlFiles) > 0)
 # Figures
 html_filesDir <- list.files(pattern="*_files")
 if (length(html_filesDir) > 0) {
-  sapply(paste("docs/", html_filesDir, sep=""), dir.create)
-  sapply(paste("docs/", html_filesDir, "/figure-html", sep=""), dir.create)
-  html_files <- list.files(html_filesDir, full.names = TRUE, recursive=TRUE)
+  sapply(paste("docs/", html_filesDir, sep=""), function(dir){if(!dir.exists(dir)){dir.create(dir)}})
+  sapply(paste("docs/", html_filesDir, "/figure-html", sep=""), function(dir){if(!dir.exists(dir)){dir.create(dir)}})
+  html_files <- list.files(path=paste(html_filesDir, "/figure-html", sep=""), full.names = T, recursive=TRUE)
   if (length(html_files) > 0)
     file.copy(from=html_files, to=paste("docs/", html_files, sep = ""), overwrite=TRUE)
 }
@@ -36,4 +36,6 @@ if (length(pdfFiles) > 0)
   file.rename(from=pdfFiles, to=paste("docs/", pdfFiles, sep=""))
 
 # Copy README.md to docs
-file.copy(from="README.md", to="docs/README.md")
+file.copy(from="README.md", to="docs/README.md", overwrite=TRUE)
+
+

### 2018-04-05 WholePlotTrajectories.Rmd
Biblio

--- 2018-03-12 WholePlotTrajectories.Rmd
@@
   \DeclareUnicodeCharacter{B0}{\textdegree}
-  \hyphenation{sa-plings}
-  \usepackage{longtable,tabu}
+  \usepackage{tabu}
+  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
+  \hyphenation{bio-di-ver-si-ty sap-lings}
 bibliography: references.bib
-lang: english # french
+lang: english
 keywordlabel: Keywords
@@
 
-<!-- Options de knitr et de R (librairies, ...) -->
-
 ```{r Options, echo=FALSE, message=FALSE}
@@
   InstallAndLoad <- function(Package) {
-    if (!Package %in% installed.packages()[, 1]) {install.packages(Package)}
+    if (!Package %in% installed.packages()[, 1]) {install.packages(Package, repos="https://cran.rstudio.com/")}
     require(Package, character.only = TRUE)
@@
   Timber=c("", 
-           "DBH $\\geq$ 50 cm, commercial species, ~ 10 trees/ha", 
-           "DBH $\\geq$ 50 cm, commercial species, ~ 10 trees/ha",
-           "DBH $\\geq$ 50 cm, commercial species, ~ 10 trees/ha"),
+           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha", 
+           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha",
+           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha"),
   Thinning=c("", 
              "", 
-             "DBH $\\geq$ 40 cm, non-valuable species, ~ 30 trees/ha", 
-             "DBH $\\geq$ 50 cm, non-valuable species, ~ 15 trees/ha  "),
+             "DBH $\\geq$ 40 cm, non-valuable species, $\\approx$ 30 trees/ha", 
+             "DBH $\\geq$ 50 cm, non-valuable species, $\\approx$ 15 trees/ha  "),
   Fuelwood=c("", 
@@
              "", 
-             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-valuable species, ~ 15 trees/ha"),
+             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-valuable species, $\\approx$ 15 trees/ha"),
   AGBLost=c(0, "$[12\\%-33\\%]$", "$[33\\%-56\\%]$", "$[35\\%-56\\%]$")
@@
 names(Intervention)[5] <- "\\%AGB lost"
-knitr::kable(Intervention, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.", longtable = TRUE, booktabs = TRUE, escape = FALSE) %>%
+knitr::kable(Intervention, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.", longtable = FALSE, booktabs = TRUE, escape = FALSE) %>%
   kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
@@
 
-## Inventories protocol and dataset correction
+## Inventories protocol and dataset collection
 
@@
 The uncertainty propagation was done by the replenishment of inventories completed at genus level from real incomplete ones on the basisof vernacular/botanical names association.
-Vernacular names were replaced through multinomial trials (M~v~([*s~1~, s~2~, …, s~N~*] ,[*$\alpha$~1~, $\alpha$~2~,…, $\alpha$~N~*]) ) based on the observed association probability [*$\alpha$~1~, $\alpha$~2~,…, $\alpha$~N~*] between each vernacular name *v* and the species [*s~1~, s~2~, …, s~N~*] recorded in the inventory.
+
+Vernacular names were replaced through multinomial trials $M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)$ based on the observed association probability $\big[\alpha_1, \alpha_2,…, \alpha_3\big]$ between each vernacular name *v* and the species $\big[s_1, s_2, …, s_N\big]$ recorded in the inventory.
+
 To avoid remaining identification caveats and consider complete inventories, the simulated inventories were then reported at genus.
@@
 The functional diversity metrics used a dataset for 6 functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, the leaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass).
+
 The trait database came from the BRIDGE project [^1] where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou. 
@@
 The dataset comprised 294 botanical species pertaining to 157 botanical genera.
+
 Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level. 
@@
 
-[^1]: http://www.ecofog.gf/Bridge/index.html
+[^1] http://www.ecofog.gf/Bridge/
 
@@
 
-```{r Fig6, echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
+```{r Fig6, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 In line with the assumtions of the IDH the observed diversity trajectories confirmed the increase of communities evenness and functional richness after disturbance.
-The consistency of the IDH has been debated for the higly diverse tropical forests, some observing the importance of environmental filters to shape communities assembly and others arguing for a genuine stochastic recovery [@Hubbell1999;@Molino2001].
-Consistently to the trajectories drawn here, a previous study of the same experiment revealed the directional changes in communities composition reflecting environmental filtering after disturbance [@Molino2001].
-Disturbance regimes brought a unimodal response depending on disturbance intensity as suggested by the outcome of 10 years and 20 years post-disturbance analysis[@Molino2001;@Baraloto2012a].
-As observed after logging in a Boranean rainforest [@Cannon1998] or after cyclones in the Solomon islands [@Burslem2000], communities diversity showed a unimodal response to disturbance and a maintenance of diversity levels, except for one of the intensively disturbed community which evenness was still increasing (Figure \@ref(fig:Fig5)).
-Despite the recovery of communities diversity, though, taxonomic composition was clearly altered (Figure \@ref(fig:Fig3)) implying a long-term, if not permenant, composition change after disturbance as observed in East African tropical forests [@Bonnell2011].
+The consistency of the IDH has been debated for the higly diverse tropical forests, some observing the importance of environmental filters to shape communities assembly and others arguing for a genuine stochastic recovery.
+Consistently to the trajectories drawn here, a previous study of the same experiment revealed the directional changes in communities composition reflecting environmental filtering after disturbance.
+Disturbance regimes brought a unimodal response depending on disturbance intensity as suggested by the outcome of 10 years and 20 years post-disturbance analysis[@Hubbell1999;@Molino2001;@Baraloto2012a].
+As observed after logging in a Boranean rainforest [@Cannon1998] or after cyclones in the Solomon islands [@Burslem2000], communities diversity showed a unimodal response to disturbance and a maintenance of diversity levels, except for one of the intensively disturbed community which evenness was still increasing \@ref(fig:Fig5).
+Despite the recovery of communities diversity, though, taxonomic composition was clearly altered \@ref(fig:Fig3) implying a long-term, if not permenant, composition change after disturbance as observed in East African tropical forests [@Bonnell2011].
 

### 2018-04-05 WholePlotTrajectories.tex
Biblio

@@
 \usepackage{textcomp} \DeclareUnicodeCharacter{B0}{\textdegree}
-\hyphenation{sa-plings} \usepackage{longtable,tabu}
+\usepackage{tabu}
+\renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
+\hyphenation{bio-di-ver-si-ty sap-lings}
 
@@
 
-\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
-\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.}\\
+\begin{table}
+
+\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.}
+\centering
+\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
 \toprule
@@
 Control &  &  &  & 0\\
-T1 & DBH $\geq$ 50 cm, commercial species, ~ 10 trees/ha &  &  & $[12\%-33\%]$\\
-T2 & DBH $\geq$ 50 cm, commercial species, ~ 10 trees/ha & DBH $\geq$ 40 cm, non-valuable species, ~ 30 trees/ha &  & $[33\%-56\%]$\\
-T3 & DBH $\geq$ 50 cm, commercial species, ~ 10 trees/ha & DBH $\geq$ 50 cm, non-valuable species, ~ 15 trees/ha & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species, ~ 15 trees/ha & $[35\%-56\%]$\\
+T1 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha &  &  & $[12\%-33\%]$\\
+T2 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30 trees/ha &  & $[33\%-56\%]$\\
+T3 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & $[35\%-56\%]$\\
 \bottomrule
-\end{longtabu}
+\end{tabu}
+\end{table}
 

### 2018-04-13 WholePlotTrajectories.Rmd
Scripts & Discussion

@@
 
-In line with the "vegetation quantity effect" suggesting that communities functional characteristics depends on dominant species rather that subtle changes in biodiversity [@Grime1998].
-This would result in a faster recovery of ecosystem processes rates than composition and diversity as demonstrated for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
+After disturbance, communities functional characteristics followed a similar but delayed trajectory as this of taxonomic characteristics which difference to initial state remained higher after 30 years.
+A faster recovery of communities processes and functioning than of composition and diversity has demonstrated for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
+This divergence  the "vegetation quantity effect" suggesting that communities functional characteristics depends on dominant species rather that subtle changes in biodiversity [@Grime1998]. 
 This highlighted the functional overlap between species in tropical forests, typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.

### 2018-04-13 Scripts/CWM.R

@@ -1,0 +1,270 @@
+library(vegan)
+library(reshape2)
+library(ade4)
+source("Scripts/Vernacular_handle.R")
+load("DB/Alpha_Plots")
+load("DB/Paracou_R_Subdivided_ok")
+
+Traits1<-read.csv("DB/BridgeOK.csv",sep=";",na.strings="")
+Traits2<-read.csv("DB/DataLifeTraits.csv",sep=";",na.strings="")
+source("Scripts/TraitsMiceFilling.R")
+
+T0<-c(1,6,11)
+T1<-c(2,7,9)
+T2<-c(3,5,10)
+T3<-c(4,8,12)
+
+treatments<-list(T0,T1,T2,T3)
+names(treatments)<-c("Control","T1","T2","T3")
+
+TraitsName<-c("Hmax","L_thickness","L_chloro","L_toughness","SLA","WD","Bark_thick")
+
+ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")#"chartreuse3",
+
+InventorySp<-do.call(rbind,lapply(LivingStand_all,function(yr){
+  ret<-do.call(rbind,yr)[,c("Famille","Genre","name")]
+  ret<-ret[which(!grepl("Indet.",ret[,"name"])),]
+  return(ret[which(!duplicated(ret)),])}))
+InventorySp<-InventorySp[which(!duplicated(InventorySp)),]
+
+dates<-sort(names(LivingStand_all))[-1]
+
+Nrep<-3
+CWM<-lapply(1:12,function(p){
+  matplot<-lapply(1:Nrep,function(rep){
+    Trajnmds<-lapply(dates,function(y){
+      temp<-LivingStand_all[[which(names(LivingStand_all)==y)]]
+      if(!any(names(temp)==p)){return(NA)}
+      if(any(names(temp)==p)){
+        temp<-temp[[which(names(temp)==p)]]
+        temp<-temp[!duplicated(temp),]
+        #temprep<-temp[which(temp[,"name"]!="Indet."),"name"]
+        temprep<-Replacement(temp,Alpha=alpha_construct(temp))
+        #temprep<-as.data.frame(Replacement(temp,Alpha=alpha_construct(temp)));colnames(temprep)<-"name"
+        #temp<-temp[,c("Genre","name")];temp<-as.data.frame(temp[which(!duplicated(temp)),],col.names="name")
[…]

### 2018-04-13 Scripts/GithubPages.R

@@ -23,8 +23,8 @@ if (length(html_filesDir) > 0) {
 #libs
 libsDirs <- list.dirs(path="libs", full.names=TRUE, recursive=TRUE)
 if (length(libsDirs) > 0) {
-  sapply(paste("docs/", libsDirs, sep = ""), dir.create)
-  libsFiles <- list.files("libs", full.names = TRUE, recursive=TRUE)
+  sapply(paste("docs/", libsDirs, sep = ""), function(dir){if(!dir.exists(dir)){dir.create(dir)}})
+  libsFiles <- list.files(path=libsDirs, full.names = TRUE, recursive=TRUE)
   file.copy(from=libsFiles, to=paste("docs/", libsFiles, sep = ""), overwrite=TRUE)
 }
 

### 2018-04-13 Scripts/TaxonomicNMDS.R

@@ -1,0 +1,65 @@
+library(vegan)
+library(reshape2)
+library(ade4)
+library(cluster)
+source("Vernacular_handle.R")
+load("DB/Alpha_Plots")
+load("DB/Paracou_R_Subdivided_ok")
+
+T0<-c(1,6,11);T1<-c(2,7,9);T2<-c(3,5,10);T3<-c(4,8,12)
+treatments<-list(T0,T1,T2,T3);names(treatments)<-c("Control","T1","T2","T3")
+
+ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
+
+dates<-sort(names(LivingStand_all))[-1]
+dates<-dates[which(!dates%in%c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014"))]
+
+AllInv<-do.call(rbind,lapply(LivingStand_all,function(yr){
+  Ret<-do.call(rbind,lapply(yr,function(pl){
+    ret<-pl[,c("Genre","name")]
+    return(ret[which(!duplicated(ret)),])}))
+  return(Ret[which(!duplicated(Ret)),])}))
+AllInv<-AllInv[which(!duplicated(AllInv)),]
+AllInv<-as.data.frame(unique(AllInv),row.names=as.character(AllInv[,"name"]))
+
+Nrep<-50
+
+Matrep<-lapply(1:Nrep,function(rep){
+  Mat<-lapply(1:12,function(p){
+    Trajnmds<-lapply(dates,function(y){
+      temp<-LivingStand_all[[which(names(LivingStand_all)==y)]]
+      if(!any(names(temp)==p)){return(NA)}
+      if(any(names(temp)==p)){
+        temp<-temp[[which(names(temp)==p)]]
+        temp<-temp[!duplicated(temp),]
+        temp<-as.data.frame(Replacement(temp,Alpha=alphas_plot[[which(names(alphas_plot)==p)]]))
+        colnames(temp)<-"name"
+        temp<-as.character(merge(AllInv,temp,by.x="row.names",by.y="name",all.y=TRUE)[,"Genre"])
+        return(as.ProbaVector(tapply(temp,temp,length)))}})
+    names(Trajnmds)<-dates
+    Ok<-names(Trajnmds)[which(unlist(lapply(Trajnmds,function(yr){return(any(!is.na(yr)))})))]
[…]

### 2018-04-13 WholePlotTrajectories.Rmd
Discussion

@@
 
+## Validation of the IDH
+
+The enhanced communities evenness and functional richness after disturbance confirmed the IDH for all treatment intensity.
+The divergence from the initial state however was very variable among plots, specifically after intense disturbance, which meant a strong dependency of plots recovery to initial composition or to abiotic conditions.
+All plots, though, followed the same unimodal trajectory which confirmed both the resilience of forests taxonomic and functional composition and the convergence of communities towards initial states [@Hubbell1999;@Molino2001;@Baraloto2012a].
+Functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered [@Guariguata2001].
+Disturbance proved to favor species with acquisitive strategies, displaying efficient resource acquisition (high SLA) but short-lived tissues (low wood density, leaf toughness and thickness) and with a dispresive strategy (low seed mass) [@Westoby1998;@Wright2004;@Reich2014].
+#The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
+#This means a shift from exclusive competition favoring competitive species before disturbance to environmental filtering selecting acquisitive and disturbance resistant species.
+In particular, the evolution seed mass classes proportions revealed a clear increase in small seeded, therefore large dispersive, species [@TerSteege2001;@Flores2006;@Haddad2008].
+
+## Functional redundancy of disturbed ecosystems
+
 After disturbance, communities functional characteristics followed a similar but delayed trajectory as this of taxonomic characteristics which difference to initial state remained higher after 30 years.
-A faster recovery of communities processes and functioning than of composition and diversity has demonstrated for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
-This divergence  the "vegetation quantity effect" suggesting that communities functional characteristics depends on dominant species rather that subtle changes in biodiversity [@Grime1998]. 
-This highlighted the functional overlap between species in tropical forests, typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.
-Ecosystem processes grasped through the set of functional traits retained here however do not assess communities resilience and future dynamics.
+This confirms conclusions drawn for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
+Assuming the "vegetation quantity effect", suggesting that communities functional characteristics are defined  by dominant species rather that subtle changes in biodiversity [@Grime1998], this delay between functional and taxonomic diversities meant the rapid recovery of dominant species while this of rare ones would be much longer.
+Besides, it meant that species missing to plots taxonomic recovery would be functionally redundant with those already in place.
+The functional overlap between species, typical of the huge biodiversity of tropical forests [@Bellwood2006], is then not trivially recovered after disturbance and as a determinant of forests resilience this questions forests resilience to multiple disturbance [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
-In line with the assumtions of the IDH the observed diversity trajectories confirmed the increase of communities evenness and functional richness after disturbance.
-The consistency of the IDH has been debated for the higly diverse tropical forests, some observing the importance of environmental filters to shape communities assembly and others arguing for a genuine stochastic recovery.
-Consistently to the trajectories drawn here, a previous study of the same experiment revealed the directional changes in communities composition reflecting environmental filtering after disturbance.
-Disturbance regimes brought a unimodal response depending on disturbance intensity as suggested by the outcome of 10 years and 20 years post-disturbance analysis[@Hubbell1999;@Molino2001;@Baraloto2012a].
-As observed after logging in a Boranean rainforest [@Cannon1998] or after cyclones in the Solomon islands [@Burslem2000], communities diversity showed a unimodal response to disturbance and a maintenance of diversity levels, except for one of the intensively disturbed community which evenness was still increasing \@ref(fig:Fig5).
-Despite the recovery of communities diversity, though, taxonomic composition was clearly altered \@ref(fig:Fig3) implying a long-term, if not permenant, composition change after disturbance as observed in East African tropical forests [@Bonnell2011].
-
-Regarding the functional trajectories and dominant trait values of disturbed communities (Figure \@ref(fig:Fig3)), disturbance appeared to favor acquisitive strategies, *i.e* species with high SLA, low wood density, small seeds, etc [@Westoby1998;@Wright2004;@Reich2014].
-The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
-This means a shift from exclusive competition favoring competitive species before disturbance to environmental filtering selecting acquisitive and disturbance resistant species.
-In particular, the evolution seed mass classes proportions (Figure \@ref(fig:Fig4)) revealed a clear increase in small seeded, therefore large dispersive, species [@TerSteege2001;@Flores2006;@Haddad2008].
-
-The diversity and composition functional trajectories showed a decoupling between functional and taxonomic after the most intense disturbance (treatments T2 and T3).
-Plots functional characteristic returned to their initial values, similar to those of control plots, faster than their taxonomic counterparts (Figures \@ref(fig:Fig2) to \@ref(fig:Fig4)).
-As functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered [@Guariguata2001].
-Such decoupling suggested the consistency of functional overlap and maybe functional versatility, implying a variety of functioning from the same characteristic, in the pre-disturbance condition.
-Functional overlap has been observed in highly diverse ecosystems [@Bellwood2006] and suggested as one mechanisms for high diversity maintenance.
-Knowing the decoupling between taxonomic and functional trajectories, this functional overlap may not be recovering after disturbance and therefore radically modify communities resilience and resistance [@Trenbath1999;@Elmqvist2003;@Diaz2005].
-Given the impaired functional overlap, the persitant compositional changes regarding disturbance resistant species [@Haddad2008] and probably lianas or epiphytes [@Martin2013], and the modification of environmental conditions like soils nutrient cycling and compaction [@Olander2005] the entirety of communities recovery after intense disturbance is deeply questioned [@Chazdon2003a].
+Besides such impact on functional redundancy along with persitant compositional changes regarding disturbance resistant species [@Haddad2008] and probably lianas or epiphytes [@Martin2013], and the modification of environmental conditions like soils nutrient cycling and compaction [@Olander2005] forest resilience remains under question [@Chazdon2003a].
 New conditions would not only be longer lasting but self-maintained as tied to disturbance regime [@Burslem2000].
-Although communities recovered the core of their initial functioning, composition seemed permanently altered after intense disturbance.
 Specificaly, this would impair species contingent to undisturbed forests, threatening their maintenance, and run the risk to loose cornerstone species and trigger unexpected ecological consequences [@Jones1994;@Diaz2005;@Gardner2007].
@@
 
->	Important differences among plots of same treatments: possible impact of initial diversity composition, local abiotic parameters (topography, soils, surrounding stands).
-Impacts depend on the disturbance intensity, but further parameters may be involved.
-
 `r if (knitr:::is_html_output()) '# References {-}'`

### 2018-04-13 Scripts/CWM.R

@@ -29,8 +29,8 @@ InventorySp<-InventorySp[which(!duplicated(InventorySp)),]
 
 dates<-sort(names(LivingStand_all))[-1]
 
-Nrep<-3
-CWM<-lapply(1:12,function(p){
+Nrep<-2
+CWM_genus<-lapply(1:12,function(p){
   matplot<-lapply(1:Nrep,function(rep){
     Trajnmds<-lapply(dates,function(y){
       temp<-LivingStand_all[[which(names(LivingStand_all)==y)]]
@@ -39,10 +39,10 @@ CWM<-lapply(1:12,function(p){
         temp<-temp[[which(names(temp)==p)]]
         temp<-temp[!duplicated(temp),]
         #temprep<-temp[which(temp[,"name"]!="Indet."),"name"]
-        temprep<-Replacement(temp,Alpha=alpha_construct(temp))
-        #temprep<-as.data.frame(Replacement(temp,Alpha=alpha_construct(temp)));colnames(temprep)<-"name"
-        #temp<-temp[,c("Genre","name")];temp<-as.data.frame(temp[which(!duplicated(temp)),],col.names="name")
-        #temprep<-merge(temprep,temp,by="name",all.x=T);temprep<-as.character(temprep[,"Genre"])
+        #temprep<-Replacement(temp,Alpha=alpha_construct(temp))
+        temprep<-as.data.frame(Replacement(temp,Alpha=alpha_construct(temp)));colnames(temprep)<-"name"
+        temp<-temp[,c("Genre","name")];temp<-as.data.frame(temp[which(!duplicated(temp)),],col.names="name")
+        temprep<-merge(temprep,temp,by="name",all.x=T);temprep<-as.character(temprep[,"Genre"])
         return(as.ProbaVector(tapply(temprep,temprep,length)))}})
     
     names(Trajnmds)<-dates
@@ -63,19 +63,49 @@ CWM<-lapply(1:12,function(p){
       return(ret[3])}))
     colnames(Trajnmds)<-namesOk
     
-    traits_filled<-Traits_filling(Traits1,Traits2,InventorySp)
-    traits_filled<-aggregate(traits_filled[,TraitsName],list(traits_filled$name),median)
+    #traits_filled<-Traits_filling(Traits1,Traits2,InventorySp)
[…]

### 2018-04-13 WholePlotTrajectories.Rmd
Files

@@
 
->The distance from the initial condition was significantly time-dependent (P<0.01).
->The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) which revealed that the trajectories of disturbed plots along time headed toward acquisitive functional strategies (from high SWG to high SLA and chlorophyll content).
+The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) revealed a trajectory of disturbed plots toward acquisitive functional strategies (from high WD to high SLA and chlorophyll content).
 
@@
 No gap filling process was applied in this case.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
 load("DB/Smass_ClassProp")

### 2018-04-13 WholePlotTrajectories.tex
Files

@@
 
-\begin{quote}
-The distance from the initial condition was significantly time-dependent
-(P\textless{}0.01). The coordinates of the functional traits were
-measured in the two-dimensional ordination space that mapped plots
-evolution along time (graphs not showed) which revealed that the
-trajectories of disturbed plots along time headed toward acquisitive
-functional strategies (from high SWG to high SLA and chlorophyll
-content).
-\end{quote}
+The coordinates of the functional traits were measured in the
+two-dimensional ordination space that mapped plots evolution along time
+(graphs not showed) revealed a trajectory of disturbed plots toward
+acquisitive functional strategies (from high WD to high SLA and
+chlorophyll content).
 
@@
 No gap filling process was applied in this case.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig4}
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig4}
 \end{figure*}
@@
 
-{\centering \includegraphics{WholePlotTrajectories_files/figure-latex/Fig6-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig6-1}
 
@@
 
-In line with the ``vegetation quantity effect'' suggesting that
-communities functional characteristics depends on dominant species
-rather that subtle changes in biodiversity \citep{Grime1998}. This would
-result in a faster recovery of ecosystem processes rates than
-composition and diversity as demonstrated for grasslands
-\citep{Tilman1997, Mouillot2011} and more recently for tropical forests
-\citep{Lohbeck2015, Guariguata2001}. This highlighted the functional
-overlap between species in tropical forests, typical of their huge
-biodiversity, and intricately questioned the different facets of
-communities resilience. Ecosystem processes grasped through the set of
-functional traits retained here however do not assess communities
-resilience and future dynamics.
+\subsection{Validation of the IDH}\label{validation-of-the-idh}
 
-In line with the assumtions of the IDH the observed diversity
-trajectories confirmed the increase of communities evenness and
-functional richness after disturbance.
-\textless{}\textless{}\textless{}\textless{}\textless{}\textless{}\textless{}
-HEAD The consistency of the IDH has been debated for the higly diverse
-tropical forests, some observing the importance of environmental filters
-to shape communities assembly and others arguing for a genuine
-stochastic recovery. Consistently to the trajectories drawn here, a
-previous study of the same experiment revealed the directional changes
-in communities composition reflecting environmental filtering after
-disturbance. Disturbance regimes brought a unimodal response depending
-on disturbance intensity as suggested by the outcome of 10 years and 20
-years post-disturbance
-analysis\citep{Hubbell1999, Molino2001, Baraloto2012a}. As observed
-after logging in a Boranean rainforest \citep{Cannon1998} or after
-cyclones in the Solomon islands \citep{Burslem2000}, communities
-diversity showed a unimodal response to disturbance and a maintenance of
-diversity levels, except for one of the intensively disturbed community
-which evenness was still increasing \ref{fig:Fig5}. Despite the recovery
-of communities diversity, though, taxonomic composition was clearly
-altered \ref{fig:Fig3} implying a long-term, if not permenant,
-composition change after disturbance as observed in East African
-tropical forests \citep{Bonnell2011}.
-
-Regarding the functional trajectories and dominant trait values of
-disturbed communities (Figure \ref{fig:Fig3}), disturbance appeared to
-favor acquisitive strategies, \emph{i.e} species with high SLA, low wood
-density, small seeds, etc \citep{Westoby1998, Wright2004, Reich2014}.
-The comparison of observed trajectories to a null model with random
-traits set shuffling showed an oriented functional shift after
-disturbance which suggest additional biotic and abiotic filters for the
-10 to 15 years after disturbance. This means a shift from exclusive
-competition favoring competitive species before disturbance to
-environmental filtering selecting acquisitive and disturbance resistant
-species. In particular, the evolution seed mass classes proportions
-(Figure \ref{fig:Fig4}) revealed a clear increase in small seeded,
-therefore large dispersive, species
+The enhanced communities evenness and functional richness after
+disturbance confirmed the IDH for all treatment intensity. The
+divergence from the initial state however was very variable among plots,
+specifically after intense disturbance, which meant a strong dependency
+of plots recovery to initial composition or to abiotic conditions. All
+plots, though, followed the same unimodal trajectory which confirmed
+both the resilience of forests taxonomic and functional composition and
+the convergence of communities towards initial states
+\citep{Hubbell1999, Molino2001, Baraloto2012a}. Functional traits are
+the most direct link between biodiversity and ecosystem functioning
+\citep{Diaz2005} the recovery of communities functional characteristics
+would go along the recovery of ecosystems processes although taxonomic
+composition remain altered \citep{Guariguata2001}. Disturbance proved to
+favor species with acquisitive strategies, displaying efficient resource
+acquisition (high SLA) but short-lived tissues (low wood density, leaf
+toughness and thickness) and with a dispresive strategy (low seed mass)
+\citep{Westoby1998, Wright2004, Reich2014}. \#The comparison of observed
+trajectories to a null model with random traits set shuffling showed an
+oriented functional shift after disturbance which suggest additional
+biotic and abiotic filters for the 10 to 15 years after disturbance.
+\#This means a shift from exclusive competition favoring competitive
+species before disturbance to environmental filtering selecting
+acquisitive and disturbance resistant species. In particular, the
+evolution seed mass classes proportions revealed a clear increase in
+small seeded, therefore large dispersive, species
 \citep{TerSteege2001, Flores2006, Haddad2008}.
 
-The diversity and composition functional trajectories showed a
-decoupling between functional and taxonomic after the most intense
-disturbance (treatments T2 and T3). Plots functional characteristic
-returned to their initial values, similar to those of control plots,
-faster than their taxonomic counterparts (Figures \ref{fig:Fig2} to
-\ref{fig:Fig4}). As functional traits are the most direct link between
-biodiversity and ecosystem functioning \citep{Diaz2005} the recovery of
-communities functional characteristics would go along the recovery of
-ecosystems processes although taxonomic composition remain altered
-\citep{Guariguata2001}. Such decoupling suggested the consistency of
-functional overlap and maybe functional versatility, implying a variety
-of functioning from the same characteristic, in the pre-disturbance
-condition. Functional overlap has been observed in highly diverse
-ecosystems \citep{Bellwood2006} and suggested as one mechanisms for high
-diversity maintenance. Knowing the decoupling between taxonomic and
-functional trajectories, this functional overlap may not be recovering
-after disturbance and therefore radically modify communities resilience
-and resistance \citep{Trenbath1999, Elmqvist2003, Diaz2005}. Given the
-impaired functional overlap, the persitant compositional changes
-regarding disturbance resistant species \citep{Haddad2008} and probably
-lianas or epiphytes \citep{Martin2013}, and the modification of
-environmental conditions like soils nutrient cycling and compaction
-\citep{Olander2005} the entirety of communities recovery after intense
-disturbance is deeply questioned \citep{Chazdon2003a}. New conditions
-would not only be longer lasting but self-maintained as tied to
-disturbance regime \citep{Burslem2000}. Although communities recovered
-the core of their initial functioning, composition seemed permanently
-altered after intense disturbance. Specificaly, this would impair
-species contingent to undisturbed forests, threatening their
-maintenance, and run the risk to loose cornerstone species and trigger
-unexpected ecological consequences
-\citep{Jones1994, Diaz2005, Gardner2007}.
+\subsection{Functional redundancy of disturbed
+ecosystems}\label{functional-redundancy-of-disturbed-ecosystems}
+
+After disturbance, communities functional characteristics followed a
+similar but delayed trajectory as this of taxonomic characteristics
+which difference to initial state remained higher after 30 years. This
+confirms conclusions drawn for grasslands
+\citep{Tilman1997, Mouillot2011} and more recently for tropical forests
+\citep{Lohbeck2015, Guariguata2001}. Assuming the ``vegetation quantity
+effect'', suggesting that communities functional characteristics are
+defined by dominant species rather that subtle changes in biodiversity
+\citep{Grime1998}, this delay between functional and taxonomic
+diversities meant the rapid recovery of dominant species while this of
+rare ones would be much longer. Besides, it meant that species missing
+to plots taxonomic recovery would be functionally redundant with those
+already in place. The functional overlap between species, typical of the
+huge biodiversity of tropical forests \citep{Bellwood2006}, is then not
+trivially recovered after disturbance and as a determinant of forests
+resilience this questions forests resilience to multiple disturbance
+\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+
+Besides such impact on functional redundancy along with persitant
+compositional changes regarding disturbance resistant species
+\citep{Haddad2008} and probably lianas or epiphytes \citep{Martin2013},
+and the modification of environmental conditions like soils nutrient
+cycling and compaction \citep{Olander2005} forest resilience remains
+under question \citep{Chazdon2003a}. New conditions would not only be
+longer lasting but self-maintained as tied to disturbance regime
+\citep{Burslem2000}. Specificaly, this would impair species contingent
+to undisturbed forests, threatening their maintenance, and run the risk
+to loose cornerstone species and trigger unexpected ecological
+consequences \citep{Jones1994, Diaz2005, Gardner2007}.
 
@@
 to be.
-
-\begin{quote}
-Important differences among plots of same treatments: possible impact of
-initial diversity composition, local abiotic parameters (topography,
-soils, surrounding stands). Impacts depend on the disturbance intensity,
-but further parameters may be involved.
-\end{quote}
 

### 2018-04-13 Scripts/CWM.R

@@ -30,7 +30,7 @@ InventorySp<-InventorySp[which(!duplicated(InventorySp)),]
 dates<-sort(names(LivingStand_all))[-1]
 
 Nrep<-2
-CWM_genus<-lapply(1:12,function(p){
+CWM<-lapply(1:12,function(p){
   matplot<-lapply(1:Nrep,function(rep){
     Trajnmds<-lapply(dates,function(y){
       temp<-LivingStand_all[[which(names(LivingStand_all)==y)]]
@@ -39,10 +39,10 @@ CWM_genus<-lapply(1:12,function(p){
         temp<-temp[[which(names(temp)==p)]]
         temp<-temp[!duplicated(temp),]
         #temprep<-temp[which(temp[,"name"]!="Indet."),"name"]
-        #temprep<-Replacement(temp,Alpha=alpha_construct(temp))
-        temprep<-as.data.frame(Replacement(temp,Alpha=alpha_construct(temp)));colnames(temprep)<-"name"
-        temp<-temp[,c("Genre","name")];temp<-as.data.frame(temp[which(!duplicated(temp)),],col.names="name")
-        temprep<-merge(temprep,temp,by="name",all.x=T);temprep<-as.character(temprep[,"Genre"])
+        temprep<-Replacement(temp,Alpha=alpha_construct(temp))
+        #temprep<-as.data.frame(Replacement(temp,Alpha=alpha_construct(temp)));colnames(temprep)<-"name"
+        #temp<-temp[,c("Genre","name")];temp<-as.data.frame(temp[which(!duplicated(temp)),],col.names="name")
+        #temprep<-merge(temprep,temp,by="name",all.x=T);temprep<-as.character(temprep[,"Genre"])
         return(as.ProbaVector(tapply(temprep,temprep,length)))}})
     
     names(Trajnmds)<-dates
@@ -64,7 +64,7 @@ CWM_genus<-lapply(1:12,function(p){
     colnames(Trajnmds)<-namesOk
     
     #traits_filled<-Traits_filling(Traits1,Traits2,InventorySp)
-    traits_filled<-aggregate(Traits_filled[,TraitsName],list(traits_filled$Genus),median)
+    traits_filled<-aggregate(Traits_filled[,TraitsName],list(traits_filled$name),median)
[…]

### 2018-04-16 WholePlotTrajectories.Rmd
Discussion

@@
 Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
-load("DB/ReplacementTraj_ForGraphs")
+load("DB/ReplacementTraj_ForGraphsN")
 
@@
 
-## Validation of the IDH
+## A validation of the intermediate disturbance hypothesis
 
-The enhanced communities evenness and functional richness after disturbance confirmed the IDH for all treatment intensity.
+The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging (@Cannon1998;@Baraloto2010).
+Only the most disturbed plots had not fully achieved the ongoing recovery of genus richness, although all had already reached equivalent levels as those of some control plots.
+Such long time richness recovery had already been observed in several logging experiments, which also highlighted the role of random disturbance damage in the richness decrease (@deAvila2015;@Hu2018).
+
+The evenness of all disturbed plots and the richness of low disturbance plots on the other hand mainly followed an asymptotic growth with a sharp increase until 15 years after disturbance.
+Such increased evenness is explained by a much more homogeneous species distribution after disturbance.
+Along with the IDH, disturbance then enhanced the growth and survival of previously infrequent species, reorganizing the typical high dominance structure of hyperdiverse mature forests.
+This increase in taxonomic diversity was accompanied by an increased taxonomic dissimilarity to initial state along time, translating a a taxonomic turnover benefiting to pioneers and light demanding species, as illustrated by the communities functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
+In line with the IDH, the post-disturbance changes in abiotic environment and competitive pressure then well favored pioneers outcompeting other species when abiotic resources are abundant but otherwise excluded in mature forests by long-lived, resistant and shade tolerant species.
+The trajectories of proportions among seed mass classes besides revealed the enhanced growth of small seeded, large dispersive, species and therefore the importance of species reproductive strategy which further supported the role of species dispersion and demographic strategy for the post-disturbance dynamics, as suggested by the IDH [@TerSteege2001;@Flores2006;@Haddad2008].
+
+## On the recovery of disturbed communities
+
 The divergence from the initial state however was very variable among plots, specifically after intense disturbance, which meant a strong dependency of plots recovery to initial composition or to abiotic conditions.
@@
 Functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered [@Guariguata2001].
-Disturbance proved to favor species with acquisitive strategies, displaying efficient resource acquisition (high SLA) but short-lived tissues (low wood density, leaf toughness and thickness) and with a dispresive strategy (low seed mass) [@Westoby1998;@Wright2004;@Reich2014].
 #The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
 #This means a shift from exclusive competition favoring competitive species before disturbance to environmental filtering selecting acquisitive and disturbance resistant species.
-In particular, the evolution seed mass classes proportions revealed a clear increase in small seeded, therefore large dispersive, species [@TerSteege2001;@Flores2006;@Haddad2008].
 

### 2018-04-16 WholePlotTrajectories.tex
Discussion

@@
 
-\subsection{Validation of the IDH}\label{validation-of-the-idh}
+\subsection{A validation of the intermediate disturbance
+hypothesis}\label{a-validation-of-the-intermediate-disturbance-hypothesis}
 
-The enhanced communities evenness and functional richness after
-disturbance confirmed the IDH for all treatment intensity. The
-divergence from the initial state however was very variable among plots,
-specifically after intense disturbance, which meant a strong dependency
-of plots recovery to initial composition or to abiotic conditions. All
-plots, though, followed the same unimodal trajectory which confirmed
-both the resilience of forests taxonomic and functional composition and
-the convergence of communities towards initial states
+The monitoring 30 years of disturbed forest communities confirmed the
+limited direct impact of low disturbance on species richness as it was
+found for impact survey of selective logging
+(\citet{Cannon1998};\citet{Baraloto2010}). Only the most disturbed plots
+had not fully achieved the ongoing recovery of genus richness, although
+all had already reached equivalent levels as those of some control
+plots. Such long time richness recovery had already been observed in
+several logging experiments, which also highlighted the role of random
+disturbance damage in the richness decrease
+(\citet{deAvila2015};\citet{Hu2018}).
+
+The evenness of all disturbed plots and the richness of low disturbance
+plots on the other hand mainly followed an asymptotic growth with a
+sharp increase until 15 years after disturbance. Such increased evenness
+is explained by a much more homogeneous species distribution after
+disturbance. Along with the IDH, disturbance then enhanced the growth
+and survival of previously infrequent species, reorganizing the typical
+high dominance structure of hyperdiverse mature forests. This increase
+in taxonomic diversity was accompanied by an increased taxonomic
+dissimilarity to initial state along time, translating a a taxonomic
+turnover benefiting to pioneers and light demanding species, as
+illustrated by the communities functional shifts towards
+resource-acquisitive strategies (sharp increase in the SLA, leaf
+thickness and bark thickness and decrease in wood density, leaf
+toughness and maximum height)
+\citep{Westoby1998, Wright2004, Reich2014}. In line with the IDH, the
+post-disturbance changes in abiotic environment and competitive pressure
+then well favored pioneers outcompeting other species when abiotic
+resources are abundant but otherwise excluded in mature forests by
+long-lived, resistant and shade tolerant species. The trajectories of
+proportions among seed mass classes besides revealed the enhanced growth
+of small seeded, large dispersive, species and therefore the importance
+of species reproductive strategy which further supported the role of
+species dispersion and demographic strategy for the post-disturbance
+dynamics, as suggested by the IDH
+\citep{TerSteege2001, Flores2006, Haddad2008}.
+
+\subsection{On the recovery of disturbed
+communities}\label{on-the-recovery-of-disturbed-communities}
+
+The divergence from the initial state however was very variable among
+plots, specifically after intense disturbance, which meant a strong
+dependency of plots recovery to initial composition or to abiotic
+conditions. All plots, though, followed the same unimodal trajectory
+which confirmed both the resilience of forests taxonomic and functional
+composition and the convergence of communities towards initial states
 \citep{Hubbell1999, Molino2001, Baraloto2012a}. Functional traits are
@@
 would go along the recovery of ecosystems processes although taxonomic
-composition remain altered \citep{Guariguata2001}. Disturbance proved to
-favor species with acquisitive strategies, displaying efficient resource
-acquisition (high SLA) but short-lived tissues (low wood density, leaf
-toughness and thickness) and with a dispresive strategy (low seed mass)
-\citep{Westoby1998, Wright2004, Reich2014}. \#The comparison of observed
-trajectories to a null model with random traits set shuffling showed an
-oriented functional shift after disturbance which suggest additional
-biotic and abiotic filters for the 10 to 15 years after disturbance.
-\#This means a shift from exclusive competition favoring competitive
-species before disturbance to environmental filtering selecting
-acquisitive and disturbance resistant species. In particular, the
-evolution seed mass classes proportions revealed a clear increase in
-small seeded, therefore large dispersive, species
-\citep{TerSteege2001, Flores2006, Haddad2008}.
+composition remain altered \citep{Guariguata2001}. \#The comparison of
+observed trajectories to a null model with random traits set shuffling
+showed an oriented functional shift after disturbance which suggest
+additional biotic and abiotic filters for the 10 to 15 years after
+disturbance. \#This means a shift from exclusive competition favoring
+competitive species before disturbance to environmental filtering
+selecting acquisitive and disturbance resistant species.
 

### 2018-04-16 Scripts/DivTaxo_trajectories.R

@@ -1,12 +1,12 @@
 load("DB/Paracou_R_Subdivided_ok")
-source("Vernacular_handle.R")
+source("Scripts/Vernacular_handle.R")
 
 # Treatments definition
 T0<-c(1,6,11);T1<-c(2,7,9);T2<-c(3,5,10);T3<-c(4,8,12)
 treatments<-list(T0,T1,T2,T3);names(treatments)<-c("T0","T1","T2","T3")
 
 # Years with identified problems of inventories
-Pby<-c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014")
+Pby<-c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014","2016","2017")
 dates<-sort(names(LivingStand_all)[which(!names(LivingStand_all)%in%Pby)])
 
 # Alpha matrix for vernacular names association
@@ -15,11 +15,16 @@ alphas_plot<-AlphaPlots(LivingStand_all)
 # Reference for botany (Genus/species association)
 load("DB/BotanyGenus")
 
+# Moving average function
+smooth<-function(mat,larg){return(do.call(cbind,lapply(1:ncol(mat),function(step){
+  range<-max(1,step-larg):min(ncol(mat),step+larg)
+  rowSums(mat[,range])/length(range)})))}
+
 #trajectories drawn by treatment
 CompleteTaxo<-lapply(1:4,function(t){
   
   treat<-treatments[[t]]
-  Nrep<-2   # Number of iteration: replacement of vernacular names
+  Nrep<-50   # Number of iteration: replacement of vernacular names
   print(t)
   
   Trajectories<-lapply(0:2,function(Q){
@@ -56,8 +61,10 @@ CompleteTaxo<-lapply(1:4,function(t){
     
     rownames(Plot_trajectory)<-treat
     colnames(Plot_trajectory)<-dates
+    Plot_trajectory<-Plot_trajectory[,which(!apply(Plot_trajectory,2,anyNA))]
     
-    return(Plot_trajectory)})
+    return(smooth(Plot_trajectory,2)) # moving average, path=2
+})
           
   Repet<-array(unlist(Repet),dim=c(length(treat),length(dates),Nrep),dimnames=list(treat,dates,1:Nrep))
    

### 2018-04-16 Scripts/DrawingTools.R

@@ -195,7 +195,7 @@ invisible(lapply(colnames(Cwm[[1]]),function(trait){
   invisible(lapply(1:4,function(tr){
     toplot<-Toplot[which(names(Toplot)%in%treatments[[tr]])]
     invisible(lapply(toplot,function(plo){
-      lines(colnames(plo),plo["0.5",],col=ColorsTr[tr],lwd=2)
+      lines(colnames(plo),plo["0.5",],col=ColorsTr[tr],lwd=1.5)
       polygon(c(colnames(plo),rev(colnames(plo))),c(plo["0.025",],rev(plo["0.975",])),
               col=rgb(0,0,0,alpha=0.1),border=NA)
       

### 2018-04-16 WholePlotTrajectories.Rmd
Discussion

@@
 
-Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined in relation to the 1989 inventories (5 years after disturbance) (Figure \@ref(fig:Fig5)). 
-*For the three indices the OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and a covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01).*
+Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined at genus level in relation to the 1989 inventories (5 years after disturbance) (Figure \@ref(fig:Fig5)).
 
@@
 Although a first increase of plots functional diversity that was all the more important than initial disturbance was intense, 30 after disturbance the diversity was similar among disturbed plots and significantly higher than this of control plots.
-*The OLS regression confirmed the time-dependency of the diversity variations (P<0.01) and the covariance analysis confirmed the significant effect of the disturbance treatment (P<0.01)*.
 
@@
 
-The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging (@Cannon1998;@Baraloto2010).
+The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging (@Cannon1998;@Baraloto2012).
 Only the most disturbed plots had not fully achieved the ongoing recovery of genus richness, although all had already reached equivalent levels as those of some control plots.
@@
 Such increased evenness is explained by a much more homogeneous species distribution after disturbance.
-Along with the IDH, disturbance then enhanced the growth and survival of previously infrequent species, reorganizing the typical high dominance structure of hyperdiverse mature forests.
+Communities composition after disturbance is either due to the old, pre-disturbance survivors or to the recruited trees.
+The composition of old survivors proved to mirror the initial community [@Herault2018], the observed composition turnover would stem from an enhanced the growth and survival of previously infrequent species which, along with the IDH, reorganizes the typical high dominance structure of hyperdiverse mature forests.
 This increase in taxonomic diversity was accompanied by an increased taxonomic dissimilarity to initial state along time, translating a a taxonomic turnover benefiting to pioneers and light demanding species, as illustrated by the communities functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
-In line with the IDH, the post-disturbance changes in abiotic environment and competitive pressure then well favored pioneers outcompeting other species when abiotic resources are abundant but otherwise excluded in mature forests by long-lived, resistant and shade tolerant species.
+Post-disturbance changes in abiotic environment and competitive pressure then well favored pioneers outcompeting other species when abiotic resources are abundant but otherwise excluded in mature forests by long-lived, resistant and shade tolerant species.
 The trajectories of proportions among seed mass classes besides revealed the enhanced growth of small seeded, large dispersive, species and therefore the importance of species reproductive strategy which further supported the role of species dispersion and demographic strategy for the post-disturbance dynamics, as suggested by the IDH [@TerSteege2001;@Flores2006;@Haddad2008].
@@
 
-The divergence from the initial state however was very variable among plots, specifically after intense disturbance, which meant a strong dependency of plots recovery to initial composition or to abiotic conditions.
-All plots, though, followed the same unimodal trajectory which confirmed both the resilience of forests taxonomic and functional composition and the convergence of communities towards initial states [@Hubbell1999;@Molino2001;@Baraloto2012a].
-Functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered [@Guariguata2001].
+Plots functional diversity had recovered initial values for almost all plots 30 years after disturbance. 
+It followed a hump-back trajectory that was mirror by all leaf, stem and life history traits, to the exception of leaf chlorophyll content.
+This recovery of communities functional characteristics would go along the recovery of ecosystems processes [@Guariguata2001], as functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005].
+Similar hump-back trajectories were followed by communities taxonomic and functional composition which argued for the convergence of communities composition maintaining the initial difference after the recovery cycle [@Hubbell1999;@Molino2001;@Baraloto2012a].
+Besides, desipte a homogeneous trend among plots, the trajectories proved quite variable among plots, specifically after intense disturbance.
+This meant a strong dependency of plots recovery time and trajectory to initial composition or abiotic conditions.
+
 #The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
@@
 
-After disturbance, communities functional characteristics followed a similar but delayed trajectory as this of taxonomic characteristics which difference to initial state remained higher after 30 years.
-This confirms conclusions drawn for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
-Assuming the "vegetation quantity effect", suggesting that communities functional characteristics are defined  by dominant species rather that subtle changes in biodiversity [@Grime1998], this delay between functional and taxonomic diversities meant the rapid recovery of dominant species while this of rare ones would be much longer.
-Besides, it meant that species missing to plots taxonomic recovery would be functionally redundant with those already in place.
+Composition trajectories, though, translated marked differences in communities composition as the functional diversity had already recovered.
+Such delay between functional and taxonomic dynamics, already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
+Assuming the "vegetation quantity effect", suggesting that communities functional characteristics are defined  by dominant species rather that subtle changes in biodiversity [@Grime1998], this delay meant a faster recovery of dominant species while this of infrequent ones would be much longer.
+These infrequent species missing to plots taxonomic recovery would consequently be functionally redundant with those already in place.
 The functional overlap between species, typical of the huge biodiversity of tropical forests [@Bellwood2006], is then not trivially recovered after disturbance and as a determinant of forests resilience this questions forests resilience to multiple disturbance [@Trenbath1999;@Elmqvist2003;@Diaz2005].

### 2018-04-17 WholePlotTrajectories.Rmd
Discussion

@@
 
-```{r Options, echo=FALSE, message=FALSE}
-knitr::opts_chunk$set(cache=F, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
+```{r Options, echo=FALSE, message=FALSE, warning=F}
+knitr::opts_chunk$set(cache=F, warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
 # Installation des packages si nécessaire et chargement
@@
 
-The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging (@Cannon1998;@Baraloto2012).
+The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging [@Cannon1998;@Baraloto2012].
 Only the most disturbed plots had not fully achieved the ongoing recovery of genus richness, although all had already reached equivalent levels as those of some control plots.
-Such long time richness recovery had already been observed in several logging experiments, which also highlighted the role of random disturbance damage in the richness decrease (@deAvila2015;@Hu2018).
+Such long time richness recovery had already been observed in several logging experiments, which also highlighted the role of random disturbance damage in the richness decrease [@deAvila2015;@Hu2018].
 

### 2018-04-17 WholePlotTrajectories.tex
Discussion

@@
 %----------------------------------------------------------------------------------------
-
-\begin{verbatim}
-## Warning: package 'kableExtra' was built under R version 3.3.3
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: package 'shape' was built under R version 3.3.3
-\end{verbatim}
 
@@
 Trajectories of Richness, Shannon and Simpson taxonomic diversity were
-examined in relation to the 1989 inventories (5 years after disturbance)
-(Figure \ref{fig:Fig5}). \emph{For the three indices the OLS regression
-confirmed the time-dependency of the diversity variations
-(P\textless{}0.01) and a covariance analysis confirmed the significant
-effect of the disturbance treatment (P\textless{}0.01).}
+examined at genus level in relation to the 1989 inventories (5 years
+after disturbance) (Figure \ref{fig:Fig5}).
 
@@
 among disturbed plots and significantly higher than this of control
-plots. \emph{The OLS regression confirmed the time-dependency of the
-diversity variations (P\textless{}0.01) and the covariance analysis
-confirmed the significant effect of the disturbance treatment
-(P\textless{}0.01)}.
+plots.
 
@@
 found for impact survey of selective logging
-(\citet{Cannon1998};\citet{Baraloto2010}). Only the most disturbed plots
-had not fully achieved the ongoing recovery of genus richness, although
-all had already reached equivalent levels as those of some control
-plots. Such long time richness recovery had already been observed in
-several logging experiments, which also highlighted the role of random
-disturbance damage in the richness decrease
-(\citet{deAvila2015};\citet{Hu2018}).
+\citep{Cannon1998, Baraloto2012}. Only the most disturbed plots had not
+fully achieved the ongoing recovery of genus richness, although all had
+already reached equivalent levels as those of some control plots. Such
+long time richness recovery had already been observed in several logging
+experiments, which also highlighted the role of random disturbance
+damage in the richness decrease \citep{deAvila2015, Hu2018}.
 
@@
 is explained by a much more homogeneous species distribution after
-disturbance. Along with the IDH, disturbance then enhanced the growth
-and survival of previously infrequent species, reorganizing the typical
-high dominance structure of hyperdiverse mature forests. This increase
-in taxonomic diversity was accompanied by an increased taxonomic
-dissimilarity to initial state along time, translating a a taxonomic
-turnover benefiting to pioneers and light demanding species, as
-illustrated by the communities functional shifts towards
-resource-acquisitive strategies (sharp increase in the SLA, leaf
-thickness and bark thickness and decrease in wood density, leaf
-toughness and maximum height)
-\citep{Westoby1998, Wright2004, Reich2014}. In line with the IDH, the
-post-disturbance changes in abiotic environment and competitive pressure
-then well favored pioneers outcompeting other species when abiotic
-resources are abundant but otherwise excluded in mature forests by
-long-lived, resistant and shade tolerant species. The trajectories of
-proportions among seed mass classes besides revealed the enhanced growth
-of small seeded, large dispersive, species and therefore the importance
-of species reproductive strategy which further supported the role of
-species dispersion and demographic strategy for the post-disturbance
-dynamics, as suggested by the IDH
-\citep{TerSteege2001, Flores2006, Haddad2008}.
+disturbance. Communities composition after disturbance is either due to
+the old, pre-disturbance survivors or to the recruited trees. The
+composition of old survivors proved to mirror the initial community
+\citep{Herault2018}, the observed composition turnover would stem from
+an enhanced the growth and survival of previously infrequent species
+which, along with the IDH, reorganizes the typical high dominance
+structure of hyperdiverse mature forests. This increase in taxonomic
+diversity was accompanied by an increased taxonomic dissimilarity to
+initial state along time, translating a a taxonomic turnover benefiting
+to pioneers and light demanding species, as illustrated by the
+communities functional shifts towards resource-acquisitive strategies
+(sharp increase in the SLA, leaf thickness and bark thickness and
+decrease in wood density, leaf toughness and maximum height)
+\citep{Westoby1998, Wright2004, Reich2014}. Post-disturbance changes in
+abiotic environment and competitive pressure then well favored pioneers
+outcompeting other species when abiotic resources are abundant but
+otherwise excluded in mature forests by long-lived, resistant and shade
+tolerant species. The trajectories of proportions among seed mass
+classes besides revealed the enhanced growth of small seeded, large
+dispersive, species and therefore the importance of species reproductive
+strategy which further supported the role of species dispersion and
+demographic strategy for the post-disturbance dynamics, as suggested by
+the IDH \citep{TerSteege2001, Flores2006, Haddad2008}.
 
@@
 
-The divergence from the initial state however was very variable among
-plots, specifically after intense disturbance, which meant a strong
-dependency of plots recovery to initial composition or to abiotic
-conditions. All plots, though, followed the same unimodal trajectory
-which confirmed both the resilience of forests taxonomic and functional
-composition and the convergence of communities towards initial states
-\citep{Hubbell1999, Molino2001, Baraloto2012a}. Functional traits are
-the most direct link between biodiversity and ecosystem functioning
-\citep{Diaz2005} the recovery of communities functional characteristics
-would go along the recovery of ecosystems processes although taxonomic
-composition remain altered \citep{Guariguata2001}. \#The comparison of
-observed trajectories to a null model with random traits set shuffling
-showed an oriented functional shift after disturbance which suggest
-additional biotic and abiotic filters for the 10 to 15 years after
-disturbance. \#This means a shift from exclusive competition favoring
+Plots functional diversity had recovered initial values for almost all
+plots 30 years after disturbance. It followed a hump-back trajectory
+that was mirror by all leaf, stem and life history traits, to the
+exception of leaf chlorophyll content. This recovery of communities
+functional characteristics would go along the recovery of ecosystems
+processes \citep{Guariguata2001}, as functional traits are the most
+direct link between biodiversity and ecosystem functioning
+\citep{Diaz2005}. Similar hump-back trajectories were followed by
+communities taxonomic and functional composition which argued for the
+convergence of communities composition maintaining the initial
+difference after the recovery cycle
+\citep{Hubbell1999, Molino2001, Baraloto2012a}. Besides, desipte a
+homogeneous trend among plots, the trajectories proved quite variable
+among plots, specifically after intense disturbance. This meant a strong
+dependency of plots recovery time and trajectory to initial composition
+or abiotic conditions.
+
+\section{The comparison of observed trajectories to a null model with
+random traits set shuffling showed an oriented functional shift after
+disturbance which suggest additional biotic and abiotic filters for the
+10 to 15 years after
+disturbance.}\label{the-comparison-of-observed-trajectories-to-a-null-model-with-random-traits-set-shuffling-showed-an-oriented-functional-shift-after-disturbance-which-suggest-additional-biotic-and-abiotic-filters-for-the-10-to-15-years-after-disturbance.}
+
+\section{This means a shift from exclusive competition favoring
 competitive species before disturbance to environmental filtering
-selecting acquisitive and disturbance resistant species.
+selecting acquisitive and disturbance resistant
+species.}\label{this-means-a-shift-from-exclusive-competition-favoring-competitive-species-before-disturbance-to-environmental-filtering-selecting-acquisitive-and-disturbance-resistant-species.}
 
@@
 
-After disturbance, communities functional characteristics followed a
-similar but delayed trajectory as this of taxonomic characteristics
-which difference to initial state remained higher after 30 years. This
-confirms conclusions drawn for grasslands
-\citep{Tilman1997, Mouillot2011} and more recently for tropical forests
-\citep{Lohbeck2015, Guariguata2001}. Assuming the ``vegetation quantity
-effect'', suggesting that communities functional characteristics are
-defined by dominant species rather that subtle changes in biodiversity
-\citep{Grime1998}, this delay between functional and taxonomic
-diversities meant the rapid recovery of dominant species while this of
-rare ones would be much longer. Besides, it meant that species missing
-to plots taxonomic recovery would be functionally redundant with those
-already in place. The functional overlap between species, typical of the
-huge biodiversity of tropical forests \citep{Bellwood2006}, is then not
+Composition trajectories, though, translated marked differences in
+communities composition as the functional diversity had already
+recovered. Such delay between functional and taxonomic dynamics, already
+observed for grasslands \citep{Tilman1997, Mouillot2011} and more
+recently for tropical forests \citep{Lohbeck2015, Guariguata2001}.
+Assuming the ``vegetation quantity effect'', suggesting that communities
+functional characteristics are defined by dominant species rather that
+subtle changes in biodiversity \citep{Grime1998}, this delay meant a
+faster recovery of dominant species while this of infrequent ones would
+be much longer. These infrequent species missing to plots taxonomic
+recovery would consequently be functionally redundant with those already
+in place. The functional overlap between species, typical of the huge
+biodiversity of tropical forests \citep{Bellwood2006}, is then not
 trivially recovered after disturbance and as a determinant of forests

### 2018-04-17 Scripts/Stats.R

@@ -1,0 +1,41 @@
+load("DB/Turnover_toInit")
+treats<-cbind(c(1,6,11,2,7,9,3,5,10,4,8,12),rep(0:3,each=3))
+treats<-treats[order(treats[,1]),]
+colnames(treats)<-c("plot","treat"); rownames(treats)<-treats[,"plot"]
+
+Maxs<-cbind(apply(Turn,1,max),treats[,"treat"])
+colnames(Maxs)<-c("Max","treat")
+
+cor(Maxs[,"Max"],Maxs[,"treat"],method="spearman")
+
+##################################################"
+
+Traits<-read.csv("DB/BridgeOK.csv",sep=";",na.strings="")
+Traits["name"]<-as.factor(paste(Traits$Genus,"_",Traits$species,sep=""))
+traits<-Traits1[,c("Family","Genus","species","name","bar_code","thickness","SPAD","toughness","dry_mass","traits_surf_area","ind_surf_area",
+                   "sapwood_dens","moisture","bark_thick")]
+colnames(traits)<-c("Family","Genus","species","name","bar_code","L_thickness","L_chloro","L_toughness","L_DryMass","LA","SLA","WD",
+                    "moisture","Bark_thick")
+##SLA calc for empty values with LA available
+traits[which(!is.na(traits[,"LA"])),"SLA"]<-
+  traits[which(!is.na(traits[,"LA"])),"LA"]/traits[which(!is.na(traits[,"LA"])),"L_DryMass"]
+
+traitsName<-c("L_thickness","L_chloro","L_toughness","L_DryMass","LA","SLA","WD","moisture","Bark_thick")
+
+allSd<-apply(traits[,traitsName],2,function(x){return(sd(x, na.rm=T))})
+
+SpSd<-aggregate(traits[,traitsName],list(traits$name),function(x){return(sd(x, na.rm=T))})
+SpSd<-apply(SpSd[,traitsName],2,function(x){return(sd(x, na.rm=T))})
+
+GenSd<-aggregate(traits[,traitsName],list(traits$Genus),function(x){return(sd(x, na.rm=T))})
+GenSd<-apply(GenSd[,traitsName],2,function(x){return(sd(x, na.rm=T))})
+
+Traits2<-read.csv("DB/DataLifeTraits.csv",sep=";",na.strings="")
+Traits2<-Traits2[which(apply(Traits2,1,function(li){return(any(!is.na(li)))})),]
[…]

### 2018-04-17 WholePlotTrajectories.Rmd
Discussion

@@
 
-The evenness of all disturbed plots and the richness of low disturbance plots on the other hand mainly followed an asymptotic growth with a sharp increase until 15 years after disturbance.
+The evenness of all disturbed plots and the richness of low disturbance plots on the other hand mainly followed an asymptotic growth sharply increasing until 15 years after disturbance.
 Such increased evenness is explained by a much more homogeneous species distribution after disturbance.
@@
 
-Plots functional diversity had recovered initial values for almost all plots 30 years after disturbance. 
-It followed a hump-back trajectory that was mirror by all leaf, stem and life history traits, to the exception of leaf chlorophyll content.
-This recovery of communities functional characteristics would go along the recovery of ecosystems processes [@Guariguata2001], as functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005].
-Similar hump-back trajectories were followed by communities taxonomic and functional composition which argued for the convergence of communities composition maintaining the initial difference after the recovery cycle [@Hubbell1999;@Molino2001;@Baraloto2012a].
-Besides, desipte a homogeneous trend among plots, the trajectories proved quite variable among plots, specifically after intense disturbance.
-This meant a strong dependency of plots recovery time and trajectory to initial composition or abiotic conditions.
-
-#The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
-#This means a shift from exclusive competition favoring competitive species before disturbance to environmental filtering selecting acquisitive and disturbance resistant species.
+The functional diversity followed a hump-back trajectory which maximum was positively correlated to the disturbance intensity ($\rho_{spearman}=$) and after 30 years, all disturbed plots had recovered diversities close to their initial values.
+Similar hump-back trajectories were followed by all leaf and stem functional traits and life history traits, to the exception of leaf chlorophyll content.
+Although the return to initial values was still ongoing the recovery of functional diversity and average communities trait values was undeniable.
+This translated the recovery of ecosystems processes [@Guariguata2001], as functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005].
+Same recovery trajectories were followed by taxonomic and functional composition which argued for communities convergence and the maintenance of plots initial difference [@Hubbell1999;@Molino2001;@Baraloto2012a].
+These compositon difference, and probably also some abiotic parameters, besides strongly determined the maximum and time path of the trajectories which, despite an homogeneous trend  were quite variable among plots.
 
@@
 
-Composition trajectories, though, translated marked differences in communities composition as the functional diversity had already recovered.
-Such delay between functional and taxonomic dynamics, already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
-Assuming the "vegetation quantity effect", suggesting that communities functional characteristics are defined  by dominant species rather that subtle changes in biodiversity [@Grime1998], this delay meant a faster recovery of dominant species while this of infrequent ones would be much longer.
-These infrequent species missing to plots taxonomic recovery would consequently be functionally redundant with those already in place.
-The functional overlap between species, typical of the huge biodiversity of tropical forests [@Bellwood2006], is then not trivially recovered after disturbance and as a determinant of forests resilience this questions forests resilience to multiple disturbance [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+Despite the consistent recovery of initial diversity and composition there was a time lag between the communities taxonomic and functional characteristics.
+Taxonomic composition had longer time path and while communities had recovered their functional diveristy, their taxonomic composition and evenness remained higher than before disturbance.
+This delay between functional and taxonomic dynamics was already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
+Communities functional diversity rely, according to the "vegetation quantity effect" [@Grime1998], on the functional strategy of dominant species.
+Communities functional trajectory is then first driven by the increase of dominant species diversity and evenness following disturbance and then by the emergence of recruitment of species progressively resembling the old pre-disturbance community which decreased the diversity.
+Dominant species then restored the functional diversity and the dominant functional type of the community although the functional and taxonomic composition remained altered and the infrequent species of initial communities are still missing to plots recovery.
+The functional traits trajectories suggest that those species would match the dominant functional type therefore increasing communities functional redundancy.
+The functional overlap between species is typical of the huge biodiversity of tropical forests [@Bellwood2006], and proved far from recovery 03 years after disturbance.
+It is though a determinant of forests resilience and should be accounted for [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
-Besides such impact on functional redundancy along with persitant compositional changes regarding disturbance resistant species [@Haddad2008] and probably lianas or epiphytes [@Martin2013], and the modification of environmental conditions like soils nutrient cycling and compaction [@Olander2005] forest resilience remains under question [@Chazdon2003a].
+Long-term alteration of functional redundancy besides came along with persistent compositional changes which most probably impact disturbance resistant species [@Haddad2008] and probably lianas or epiphytes [@Martin2013], and the modification of environmental conditions like soils nutrient cycling and compaction [@Olander2005] forest resilience remains under question [@Chazdon2003a].
 New conditions would not only be longer lasting but self-maintained as tied to disturbance regime [@Burslem2000].
@@
 
-To confirm this long-term communities changes and highlight the ecological rules shaping communities assembly, we suggest a focus on recruitment dynamics, that would follow the diversity and functional characteristic of recruited communities and, by extension, of the forest to be.
-
 `r if (knitr:::is_html_output()) '# References {-}'`

### 2018-04-17 Scripts/Stats.R

@@ -1,4 +1,7 @@
-load("DB/Turnover_toInit")
+load("DB/ReplacementTraj_ForGraphs")
+
+CompleteTaxo
+
 treats<-cbind(c(1,6,11,2,7,9,3,5,10,4,8,12),rep(0:3,each=3))
 treats<-treats[order(treats[,1]),]
 colnames(treats)<-c("plot","treat"); rownames(treats)<-treats[,"plot"]

### 2018-04-17 WholePlotTrajectories.Rmd
Intro & discussion

@@
 The large areas covered with tropical forests worldwide hold crucial economic, social and cultural value. 
-They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate local climate, support carbon, water and nutrient cycles, and ensure cultural and human well-being. 
-The simultaneous increase of forests products demand and substantial climatic changes heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threaten the maintenance of communities structure, composition and functioning and the dynamics that shape them [@Anderson-Teixeira2013;@Sist2015].
+They provide wood and multiple non-timber forest products, shelter for a diversified fauna, regulate the local climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being.
+The simultaneous increase of forests products demand and substantial climatic changes heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threatened the maintenance of communities structure, composition and functioning and their underlying dynamics [@Anderson-Teixeira2013;@Sist2015].
 
-In tropical forests, a constant range of disturbance determines the ecological rules shaping the communities through the reallocation of resources (as light, heat and water fluxes), changes in interspecific relations, etc.
-The response of communities to disturbance is then the cornerstone of tropical forests ecology, modelling and management [@White2001;@Chazdon2003a].
-Disturbance recovery have been largely studied through structural parameters as aboveground biomass, tree height or stem density, which all reflect rapid and measurable changes of ecosystems functioning asses crucial issues about the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
-In this perspective tree species diversity is inevitable as it determines ecosystems productivity, stability and functioning through the interspecific complementarity that enhances resources use, nutrients storing, and hazards mitigation [@Tilman2014].
-Revealing diversity dynamics is besides all the more important that disturbance, modifies the environment and interspecific relations and surely impacts trees composition and diversity [@CazzollaGatti2014].
+In tropical forests communities are shaped by a constant range of disturbance that changes the abiotic environment, as the light, heat and water fluxes, and the biotic interaction and competitive pressure.
+The cornerstone of tropical forests ecology is then to understand their response to disturbance and the corresponding mechanisms [@White2001;@Chazdon2003a].
+Forests response to disturbance has been largely studied through structural parameters rapid and convenient to measure, as aboveground biomass, tree height or stem density.
+These structural parameters have then been sucessfully modeled and assessed the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
+However the response of tree species diversity although it is determinant of ecosystems productivity, stability and functioning [@Tilman2014] and it would be most probably impacted by the biotic and abiotic changes after disturbance [@CazzollaGatti2014].
 
-Until now, although, short-term dynamics demonstrated negligible or even positive impacts on communities diversity in accordance with the intermediate disturbance hypothesis (IDH) [@Molino2001;@Kariuki2006a;@Berry2008a].
-The long-term response of communities composition and diversity to disturbance remains unclear.
-The IDH predicting a maximized species diversity at intermediate level of disturbance has besides mainly been tested through species richness, which gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-Studies of communities response to disturbance should encompass taxonomic composition and evenness, which only can assess some conservation issues and the ecological rules that determine communities abundance distribution [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Beyond taxonomic metrics, communities recovery should also encompass functional diversity metrics [@Moretti2009;@Baraloto2012a] through functional traits, that are species biological attributes relating to individuals fitness and ecosystem properties [@Violle2007b;@Scheiter2013].
-Functional traits were largely adopted and integrated within a workable framework to assess forests dynamics and functioning [@Diaz2005;@Villeger2008a] and a vast litterature revealed some major functional traits determinant for species ecology and performance.
-These major traits proved relevant to assess the ecological rules shaping species assembly: post-disturbance trajectories in tropical rainforests, for example, are assumed to be driven by environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
-Communities dynamics would then be reflected by a shift from "conservative" species displaying long-lived, dense tissues growing slowly but with scarce resources, to "acquisitive" species with fast-growing, light tissues allowing fast resources acquisition [@TerSteege2001;@Reich2014;@Herault2011].
-In this perspective a combination of resource acquisition traits (leaf area, density and chlorophyll content and stem specific gravity and bark thickness) and life history traits of growth and reproduction (seed mass and maximum height) already proved relevant [@Wright2004;@Westoby2006a;@Chave2009b].
+In the short-term, diversity dynamics demonstrated negligible or even positive impacts of disturbance on communities diversity, and have been formalized by the intermediate disturbance hypothesis (IDH) that predicts a maximized species diversity at intermediate level of disturbance[@Molino2001;@Kariuki2006a;@Berry2008a].
+The IDH however has scarcely been tested in the long term and mainly relied on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+Communities composition and evenness are crucial to resolve conservation issues and assess the ecological rules shaping communities abundance distribution [@Magurran1988;@Lavorel2002;@Bellwood2006].
+A promosing perspective is also the functional approach accounting for species biological attributes related to individuals fitness and ecosystem properties [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+Major functional traits revealing species ecology and performance were largely adopted through a workable framework assessing forests dynamics and functioning [@Diaz2005;@Villeger2008a].
+This approach for example highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
+This was translated by the shift from "conservative" species with long-lived, dense tissues growing slowly but with scarce resources, to "acquisitive" fast-growing species with light tissues and efficient resources use [@TerSteege2001;@Reich2014;@Herault2011].
+Such shift would be detected with a combination of resource acquisition funcitonal traits (leaf area, density and chlorophyll content and stem specific gravity and bark thickness) and growth and reproduction life history traits (seed mass and maximum height) already proved relevant [@Wright2004;@Westoby2006a;@Chave2009b].
+A functional approach combined to the analysis of taxonomic diversity and composition is then a reliable way to validate the IDH in the long term and from communities evenness and functinal perspective.
+A full interpretation of communities response would be completed with communities taxonomic and functional composition that would elucidate the determinism of forest recovery and the maintenance of intrinsinc differences.
+Those development would be as much information to understand the mechanisms of forests response to diversity and help conservation and management strategies.
 
@@
 
->To compensate the intrinsinc difference among plots the trajectories drawn correspond to the difference in value between the reference inventory in >1984 (5 year after disturbance) and the successive years inventoried.
+To compensate the intrinsinc difference among plots the trajectories drawn correspond to the difference in value between the reference inventory in 1984 (5 year after disturbance) and the successive years inventoried.
 
@@
 Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
-load("DB/ReplacementTraj_ForGraphsN")
+load("DB/ReplacementTraj_ForGraphs")
 

### 2018-04-17 WholePlotTrajectories.tex
Intro & discussion

@@
 economic, social and cultural value. They provide wood and multiple
-non-timber forest products, shelter a diversified fauna, regulate local
-climate, support carbon, water and nutrient cycles, and ensure cultural
-and human well-being. The simultaneous increase of forests products
-demand and substantial climatic changes heightened the pressure on the
-remaining forests \citep{Gibson2011a, Morales-Hidalgo2015} and threaten
-the maintenance of communities structure, composition and functioning
-and the dynamics that shape them
-\citep{Anderson-Teixeira2013, Sist2015}.
+non-timber forest products, shelter for a diversified fauna, regulate
+the local climate, support the carbon, water and nutrient cycles, and
+ensure cultural and human well-being. The simultaneous increase of
+forests products demand and substantial climatic changes heightened the
+pressure on the remaining forests
+\citep{Gibson2011a, Morales-Hidalgo2015} and threatened the maintenance
+of communities structure, composition and functioning and their
+underlying dynamics \citep{Anderson-Teixeira2013, Sist2015}.
 
-In tropical forests, a constant range of disturbance determines the
-ecological rules shaping the communities through the reallocation of
-resources (as light, heat and water fluxes), changes in interspecific
-relations, etc. The response of communities to disturbance is then the
-cornerstone of tropical forests ecology, modelling and management
-\citep{White2001, Chazdon2003a}. Disturbance recovery have been largely
-studied through structural parameters as aboveground biomass, tree
-height or stem density, which all reflect rapid and measurable changes
-of ecosystems functioning asses crucial issues about the maintenance of
-ecosystems processes and services
-\citep{Denslow2000, Blanc2009, Rutishauser2016}. In this perspective
-tree species diversity is inevitable as it determines ecosystems
-productivity, stability and functioning through the interspecific
-complementarity that enhances resources use, nutrients storing, and
-hazards mitigation \citep{Tilman2014}. Revealing diversity dynamics is
-besides all the more important that disturbance, modifies the
-environment and interspecific relations and surely impacts trees
-composition and diversity \citep{CazzollaGatti2014}.
+In tropical forests communities are shaped by a constant range of
+disturbance that changes the abiotic environment, as the light, heat and
+water fluxes, and the biotic interaction and competitive pressure. The
+cornerstone of tropical forests ecology is then to understand their
+response to disturbance and the corresponding mechanisms
+\citep{White2001, Chazdon2003a}. Forests response to disturbance has
+been largely studied through structural parameters rapid and convenient
+to measure, as aboveground biomass, tree height or stem density. These
+structural parameters have then been sucessfully modeled and assessed
+the maintenance of ecosystems processes and services
+\citep{Denslow2000, Blanc2009, Rutishauser2016}. However the response of
+tree species diversity although it is determinant of ecosystems
+productivity, stability and functioning \citep{Tilman2014} and it would
+be most probably impacted by the biotic and abiotic changes after
+disturbance \citep{CazzollaGatti2014}.
 
-Until now, although, short-term dynamics demonstrated negligible or even
-positive impacts on communities diversity in accordance with the
-intermediate disturbance hypothesis (IDH)
-\citep{Molino2001, Kariuki2006a, Berry2008a}. The long-term response of
-communities composition and diversity to disturbance remains unclear.
-The IDH predicting a maximized species diversity at intermediate level
-of disturbance has besides mainly been tested through species richness,
-which gives limited or misleading information on forests recovery and
-functioning \citep{Martin2015, Chaudhary2016}. Studies of communities
-response to disturbance should encompass taxonomic composition and
-evenness, which only can assess some conservation issues and the
-ecological rules that determine communities abundance distribution
-\citep{Magurran1988, Lavorel2002, Bellwood2006}. Beyond taxonomic
-metrics, communities recovery should also encompass functional diversity
-metrics \citep{Moretti2009, Baraloto2012a} through functional traits,
-that are species biological attributes relating to individuals fitness
-and ecosystem properties \citep{Violle2007b, Scheiter2013}. Functional
-traits were largely adopted and integrated within a workable framework
-to assess forests dynamics and functioning
-\citep{Diaz2005, Villeger2008a} and a vast litterature revealed some
-major functional traits determinant for species ecology and performance.
-These major traits proved relevant to assess the ecological rules
-shaping species assembly: post-disturbance trajectories in tropical
-rainforests, for example, are assumed to be driven by environmental
-filters fostering disturbance resistant species with rapid growth and
-efficient resources acquisition \citep{Molino2001, Haddad2008}.
-Communities dynamics would then be reflected by a shift from
-``conservative'' species displaying long-lived, dense tissues growing
-slowly but with scarce resources, to ``acquisitive'' species with
-fast-growing, light tissues allowing fast resources acquisition
-\citep{TerSteege2001, Reich2014, Herault2011}. In this perspective a
-combination of resource acquisition traits (leaf area, density and
-chlorophyll content and stem specific gravity and bark thickness) and
-life history traits of growth and reproduction (seed mass and maximum
-height) already proved relevant
-\citep{Wright2004, Westoby2006a, Chave2009b}.
+In the short-term, diversity dynamics demonstrated negligible or even
+positive impacts of disturbance on communities diversity, and have been
+formalized by the intermediate disturbance hypothesis (IDH) that
+predicts a maximized species diversity at intermediate level of
+disturbance\citep{Molino2001, Kariuki2006a, Berry2008a}. The IDH however
+has scarcely been tested in the long term and mainly relied on the
+analysis of species richness that gives limited or misleading
+information on forests recovery and functioning
+\citep{Martin2015, Chaudhary2016}. Communities composition and evenness
+are crucial to resolve conservation issues and assess the ecological
+rules shaping communities abundance distribution
+\citep{Magurran1988, Lavorel2002, Bellwood2006}. A promosing perspective
+is also the functional approach accounting for species biological
+attributes related to individuals fitness and ecosystem properties
+\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. Major
+functional traits revealing species ecology and performance were largely
+adopted through a workable framework assessing forests dynamics and
+functioning \citep{Diaz2005, Villeger2008a}. This approach for example
+highlighted in tropical rainforests the environmental filters fostering
+disturbance resistant species with rapid growth and efficient resources
+acquisition \citep{Molino2001, Haddad2008}. This was translated by the
+shift from ``conservative'' species with long-lived, dense tissues
+growing slowly but with scarce resources, to ``acquisitive''
+fast-growing species with light tissues and efficient resources use
+\citep{TerSteege2001, Reich2014, Herault2011}. Such shift would be
+detected with a combination of resource acquisition funcitonal traits
+(leaf area, density and chlorophyll content and stem specific gravity
+and bark thickness) and growth and reproduction life history traits
+(seed mass and maximum height) already proved relevant
+\citep{Wright2004, Westoby2006a, Chave2009b}. A functional approach
+combined to the analysis of taxonomic diversity and composition is then
+a reliable way to validate the IDH in the long term and from communities
+evenness and functinal perspective. A full interpretation of communities
+response would be completed with communities taxonomic and functional
+composition that would elucidate the determinism of forest recovery and
+the maintenance of intrinsinc differences. Those development would be as
+much information to understand the mechanisms of forests response to
+diversity and help conservation and management strategies.
 
@@
 
-\begin{quote}
 To compensate the intrinsinc difference among plots the trajectories
 drawn correspond to the difference in value between the reference
-inventory in \textgreater{}1984 (5 year after disturbance) and the
-successive years inventoried.
-\end{quote}
+inventory in 1984 (5 year after disturbance) and the successive years
+inventoried.
 
@@
 The evenness of all disturbed plots and the richness of low disturbance
-plots on the other hand mainly followed an asymptotic growth with a
-sharp increase until 15 years after disturbance. Such increased evenness
-is explained by a much more homogeneous species distribution after
+plots on the other hand mainly followed an asymptotic growth sharply
+increasing until 15 years after disturbance. Such increased evenness is
+explained by a much more homogeneous species distribution after
 disturbance. Communities composition after disturbance is either due to
@@
 
-Plots functional diversity had recovered initial values for almost all
-plots 30 years after disturbance. It followed a hump-back trajectory
-that was mirror by all leaf, stem and life history traits, to the
-exception of leaf chlorophyll content. This recovery of communities
-functional characteristics would go along the recovery of ecosystems
-processes \citep{Guariguata2001}, as functional traits are the most
-direct link between biodiversity and ecosystem functioning
-\citep{Diaz2005}. Similar hump-back trajectories were followed by
-communities taxonomic and functional composition which argued for the
-convergence of communities composition maintaining the initial
-difference after the recovery cycle
-\citep{Hubbell1999, Molino2001, Baraloto2012a}. Besides, desipte a
-homogeneous trend among plots, the trajectories proved quite variable
-among plots, specifically after intense disturbance. This meant a strong
-dependency of plots recovery time and trajectory to initial composition
-or abiotic conditions.
-
-\section{The comparison of observed trajectories to a null model with
-random traits set shuffling showed an oriented functional shift after
-disturbance which suggest additional biotic and abiotic filters for the
-10 to 15 years after
-disturbance.}\label{the-comparison-of-observed-trajectories-to-a-null-model-with-random-traits-set-shuffling-showed-an-oriented-functional-shift-after-disturbance-which-suggest-additional-biotic-and-abiotic-filters-for-the-10-to-15-years-after-disturbance.}
-
-\section{This means a shift from exclusive competition favoring
-competitive species before disturbance to environmental filtering
-selecting acquisitive and disturbance resistant
-species.}\label{this-means-a-shift-from-exclusive-competition-favoring-competitive-species-before-disturbance-to-environmental-filtering-selecting-acquisitive-and-disturbance-resistant-species.}
+The functional diversity followed a hump-back trajectory which maximum
+was positively correlated to the disturbance intensity
+(\(\rho_{spearman}=\)) and after 30 years, all disturbed plots had
+recovered diversities close to their initial values. Similar hump-back
+trajectories were followed by all leaf and stem functional traits and
+life history traits, to the exception of leaf chlorophyll content.
+Although the return to initial values was still ongoing the recovery of
+functional diversity and average communities trait values was
+undeniable. This translated the recovery of ecosystems processes
+\citep{Guariguata2001}, as functional traits are the most direct link
+between biodiversity and ecosystem functioning \citep{Diaz2005}. Same
+recovery trajectories were followed by taxonomic and functional
+composition which argued for communities convergence and the maintenance
+of plots initial difference
+\citep{Hubbell1999, Molino2001, Baraloto2012a}. These compositon
+difference, and probably also some abiotic parameters, besides strongly
+determined the maximum and time path of the trajectories which, despite
+an homogeneous trend were quite variable among plots.
 
@@
 
-Composition trajectories, though, translated marked differences in
-communities composition as the functional diversity had already
-recovered. Such delay between functional and taxonomic dynamics, already
-observed for grasslands \citep{Tilman1997, Mouillot2011} and more
-recently for tropical forests \citep{Lohbeck2015, Guariguata2001}.
-Assuming the ``vegetation quantity effect'', suggesting that communities
-functional characteristics are defined by dominant species rather that
-subtle changes in biodiversity \citep{Grime1998}, this delay meant a
-faster recovery of dominant species while this of infrequent ones would
-be much longer. These infrequent species missing to plots taxonomic
-recovery would consequently be functionally redundant with those already
-in place. The functional overlap between species, typical of the huge
-biodiversity of tropical forests \citep{Bellwood2006}, is then not
-trivially recovered after disturbance and as a determinant of forests
-resilience this questions forests resilience to multiple disturbance
+Despite the consistent recovery of initial diversity and composition
+there was a time lag between the communities taxonomic and functional
+characteristics. Taxonomic composition had longer time path and while
+communities had recovered their functional diveristy, their taxonomic
+composition and evenness remained higher than before disturbance. This
+delay between functional and taxonomic dynamics was already observed for
+grasslands \citep{Tilman1997, Mouillot2011} and more recently for
+tropical forests \citep{Lohbeck2015, Guariguata2001}. Communities
+functional diversity rely, according to the ``vegetation quantity
+effect'' \citep{Grime1998}, on the functional strategy of dominant
+species. Communities functional trajectory is then first driven by the
+increase of dominant species diversity and evenness following
+disturbance and then by the emergence of recruitment of species
+progressively resembling the old pre-disturbance community which
+decreased the diversity. Dominant species then restored the functional
+diversity and the dominant functional type of the community although the
+functional and taxonomic composition remained altered and the infrequent
+species of initial communities are still missing to plots recovery. The
+functional traits trajectories suggest that those species would match
+the dominant functional type therefore increasing communities functional
+redundancy. The functional overlap between species is typical of the
+huge biodiversity of tropical forests \citep{Bellwood2006}, and proved
+far from recovery 03 years after disturbance. It is though a determinant
+of forests resilience and should be accounted for
 \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-Besides such impact on functional redundancy along with persitant
-compositional changes regarding disturbance resistant species
-\citep{Haddad2008} and probably lianas or epiphytes \citep{Martin2013},
-and the modification of environmental conditions like soils nutrient
-cycling and compaction \citep{Olander2005} forest resilience remains
-under question \citep{Chazdon2003a}. New conditions would not only be
-longer lasting but self-maintained as tied to disturbance regime
-\citep{Burslem2000}. Specificaly, this would impair species contingent
-to undisturbed forests, threatening their maintenance, and run the risk
-to loose cornerstone species and trigger unexpected ecological
-consequences \citep{Jones1994, Diaz2005, Gardner2007}.
-
-To confirm this long-term communities changes and highlight the
-ecological rules shaping communities assembly, we suggest a focus on
-recruitment dynamics, that would follow the diversity and functional
-characteristic of recruited communities and, by extension, of the forest
-to be.
+Long-term alteration of functional redundancy besides came along with
+persistent compositional changes which most probably impact disturbance
+resistant species \citep{Haddad2008} and probably lianas or epiphytes
+\citep{Martin2013}, and the modification of environmental conditions
+like soils nutrient cycling and compaction \citep{Olander2005} forest
+resilience remains under question \citep{Chazdon2003a}. New conditions
+would not only be longer lasting but self-maintained as tied to
+disturbance regime \citep{Burslem2000}. Specificaly, this would impair
+species contingent to undisturbed forests, threatening their
+maintenance, and run the risk to loose cornerstone species and trigger
+unexpected ecological consequences
+\citep{Jones1994, Diaz2005, Gardner2007}.
 

### 2018-04-17 Scripts/DivTaxo_trajectories.R

@@ -81,3 +81,4 @@ CompleteTaxo<-lapply(1:4,function(t){
 })
 
 save(CompleteTaxo,file = "DB/ReplacementTraj_ForGraphs")
+

### 2018-04-19 WholePlotTrajectories.Rmd
Figures

@@
 The taxonomic diversity was assessed through Richness and the Hill number translation of Shannon and Simpson indices [@Hill1973]. 
-To tackle the unequal number of recruited trees among treatments the indices bias corrected estimator were used, following [@Chao2015;@Marcon2015b].
+>>To tackle the unequal number of recruited trees among treatments the indices bias corrected estimator were used, following [@Chao2015;@Marcon2015b].
+
 These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q), which proved well suited for diversity studies [@Patil1982;@Tothmeresz1995].
@@
 
-```{r Fig5, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
+```{r Fig5, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
 load("DB/ReplacementTraj_ForGraphs")
@@
 
-While for undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity (Figure \@ref(fig:Fig5)).
+While for undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity ($\rho_{spearman}^{Shannon}=0.87$, and $\rho_{spearman}^{Simpson}=0.88$) (Figure \@ref(fig:Fig5)).
 The Richness was less affected by the initial disturbance except for the higher intensity (T3 plots) (Figure \@ref(fig:Fig5)). 

### 2018-04-19 WholePlotTrajectories.tex
Figures

@@
 The taxonomic diversity was assessed through Richness and the Hill
-number translation of Shannon and Simpson indices \citep{Hill1973}. To
-tackle the unequal number of recruited trees among treatments the
-indices bias corrected estimator were used, following
-\citep{Chao2015, Marcon2015b}. These three indices belong to the set of
-HCDT or generalized entropy, respectively corresponding to the 0, 1 and
-2 order of diversity (q), which proved well suited for diversity studies
+number translation of Shannon and Simpson indices \citep{Hill1973}.
+\textgreater{}\textgreater{}To tackle the unequal number of recruited
+trees among treatments the indices bias corrected estimator were used,
+following \citep{Chao2015, Marcon2015b}.
+
+These three indices belong to the set of HCDT or generalized entropy,
+respectively corresponding to the 0, 1 and 2 order of diversity (q),
+which proved well suited for diversity studies
 \citep{Patil1982, Tothmeresz1995}. The functional diversity was reported
@@
 
-\caption{Trajectories of the difference to reference inventory over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
+\caption{Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig5}
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig5}
 \end{figure*}
@@
 disturbed plots showed significant increase positively correlated with
-the disturbance intensity (Figure \ref{fig:Fig5}). The Richness was less
-affected by the initial disturbance except for the higher intensity (T3
-plots) (Figure \ref{fig:Fig5}). The Shannon and Simpson diversities
-however first significantly increased, illustrating an increasing
-evenness, before reaching a maximum around 15 years after disturbance
-for all plots except one from T1 treatment and one from T2 treatment
-(Figure \ref{fig:Fig5}).
+the disturbance intensity (\(\rho_{spearman}^{Shannon}=0.87\), and
+\(\rho_{spearman}^{Simpson}=0.88\)) (Figure \ref{fig:Fig5}). The
+Richness was less affected by the initial disturbance except for the
+higher intensity (T3 plots) (Figure \ref{fig:Fig5}). The Shannon and
+Simpson diversities however first significantly increased, illustrating
+an increasing evenness, before reaching a maximum around 15 years after
+disturbance for all plots except one from T1 treatment and one from T2
+treatment (Figure \ref{fig:Fig5}).
 

### 2018-04-19 Scripts/DrawingTools.R

@@ -136,33 +136,34 @@ FunDist<-function(Data_FunComp){
 
 TaxoTraj<-function(CompTaxo){
   par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
-for(q in 1:3){     
-  Toplot<-lapply(CompTaxo,function(tr){return(tr[,,,q])})
-  Toplot<-lapply(Toplot,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
-  Toplot<-lapply(Toplot,function(tr){
-    ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col-tr[,1,rep]}))})#
-    ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),dim(tr)[3]),
-               dimnames=list(rownames(ret[[1]]),colnames(ret[[1]]),c(0.025,0.5,0.975)))
-    return(ret)})
-  Toplot<-lapply(Toplot,function(tr){return(tr[,which(!is.na(Toplot[[2]][1,,"0.5"])),])})
-  
-  plot(as.numeric(colnames(Toplot[[1]]))-1984,Toplot[[1]][1,,1],type="n",xaxt="n",
-       xlab="",ylab="",ylim=c(min(unlist(Toplot),na.rm=T),max(unlist(Toplot),na.rm=T)))
-  axis(1,at=as.character(seq(5,33,5)),labels=TRUE)  
-  mtext(paste("(",c("a","b","c")[q],") ",c("Richness","Shannon","Simpson")[q],sep=""),
-        line=1,side=3,adj=0)
-  
-  invisible(lapply(1:4,function(t){  
-    toplot<-Toplot[[t]]
+  for(q in 1:3){     
+    Toplot<-lapply(CompTaxo,function(tr){return(tr[,,,q])})
+    Toplot<-lapply(Toplot,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
+    Toplot<-lapply(Toplot,function(tr){
+      ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col-tr[,1,rep]}))})#
+      ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),
[…]

### 2018-04-19 Scripts/Stats.R

@@ -1,15 +1,34 @@
 load("DB/ReplacementTraj_ForGraphs")
 
-CompleteTaxo
+CompTaxo<-CompleteTaxo
 
 treats<-cbind(c(1,6,11,2,7,9,3,5,10,4,8,12),rep(0:3,each=3))
 treats<-treats[order(treats[,1]),]
 colnames(treats)<-c("plot","treat"); rownames(treats)<-treats[,"plot"]
 
-Maxs<-cbind(apply(Turn,1,max),treats[,"treat"])
-colnames(Maxs)<-c("Max","treat")
+for(q in 1:3){     
+  Toplot<-lapply(CompTaxo,function(tr){return(tr[,,,q])})
+  Toplot<-lapply(Toplot,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
+  Toplot<-lapply(Toplot,function(tr){
+    ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col-tr[,1,rep]}))})#
+    ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),
+               dimnames=list(rownames(ret[[1]]),as.numeric(colnames(ret[[1]]))-1984,1:length(ret)))
+    return(apply(apply(ret,c(1,2),median),1,max))
+    })
+  Ret<-as.data.frame(unlist(Toplot))
+  Ret$plot<-rownames(Ret)
+  Ret<-Ret[order(as.numeric(Ret$plot)),]
+  Ret<-merge(Ret,treats,by="plot")
+  assign(c("Richness","Shannon","Simpson")[q],Ret)
+}
 
-cor(Maxs[,"Max"],Maxs[,"treat"],method="spearman")
+colnames(Richness)<-c("plot","Max","treat")
+cor(Richness[,"Max"],Richness[,"treat"],method="spearman")
+colnames(Shannon)<-c("plot","Max","treat")
+cor(Shannon[,"Max"],Shannon[,"treat"],method="spearman")
+colnames(Simpson)<-c("plot","Max","treat")
+cor(Simpson[,"Max"],Simpson[,"treat"],method="spearman")
+
 
 ##################################################"
 

### 2018-04-19 AppendixI.Rmd
Figures captions

@@
 ---
-title: "Appendix I _ Scheme of the taxonomic uncertainty handle"
-author: "Ariane"
-date: "30 janvier 2018"
-output: html_document
+title: "Appendix I"
+author:
+  - name: Ariane MIRABEL
+    authsuperscript: 1*
+  - name: Eric MARCON
+    authsuperscript: 1
+  - name: Bruno HERAULT
+    authsuperscript: 2
+affiliation:
+  - affsuperscript: 1
+    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
+    address: >
+      Campus Agronomique,
+      97310 Kourou,
+      France.
+  - affsuperscript: 2
+    dptuniv: INPHB (Institut National Ploytechnique Félix Houphoüet Boigny) 
+    address: Yamoussoukro, Ivory Coast
+corrauthor:
+    email: ariane.mirabel@ecofog.gf
+    url: http://www.ecofog.gf/spip.php?article47
+journalinfo: Hal 00679993
+date: "`r format(Sys.time(), '%d %B %Y')`"
+preamble: >
+  \usepackage{textcomp}
+  \DeclareUnicodeCharacter{B0}{\textdegree}
+  \usepackage{tabu}
+  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
+  \hyphenation{bio-di-ver-si-ty sap-lings}
+lang: english
+output: 
+  bookdown::pdf_book:
+    base_format: EcoFoG::article
+    keep_tex: TRUE
+  bookdown::gitbook: default
+  bookdown::word_document2: default
 ---
 
-![**Fig.
-A1 _ ** Whenever the family or genus were recorded for a vernacular name, the pool of species sampled to replace the vernacular name was restricted to those pertaining to the same higher taxonomic level (i.e species of the same genus when the species is missing, species of the same family when the genus is missing).](BayesianSheme.png)
+```{r Options, echo=FALSE, message=FALSE, warning=F}
+knitr::opts_chunk$set(cache=F, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40),fig.align="center")
+# Installation des packages si nécessaire et chargement
+Library <- function(Packages) {
+  InstallAndLoad <- function(Package) {
+    if (!Package %in% installed.packages()[, 1]) {install.packages(Package, repos="https://cran.rstudio.com/")}
+    require(Package, character.only = TRUE)
+  }
+  invisible(sapply(Packages, InstallAndLoad))
+}
+#source("Scripts/GithubPages.R")
+source("Scripts/DrawingTools.R")
+```
+
+The following information are available for this article:
+
+**Figure A1**: Whenever the family or genus were recorded for a vernacular name, the pool of species sampled to replace the vernacular name was restricted to those pertaining to the same higher taxonomic level (i.e species of the same genus when the species is missing, species of the same family when the genus is missing) \@ref(fig:FigBayesianScheme).
+
+**Figure A2**: Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+No gap filling process was applied in this case.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3 \@ref(fig:FigSeedMassClasses).
+
+```{r FigBayesianScheme,fig.cap="Scheme of the Bayesian process propagating the taxonomic uncertainty"}
+knitr::include_graphics("ExternalFig/BayesianSheme.png")
+```
+
+```{r FigSeedMassClasses, echo=FALSE, out.width='100%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+No gap filling process was applied in this case.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
+load("DB/Smass_ClassProp")
+
+SeedMassProp(Smass)
+
+```

### 2018-04-19 AppendixI.tex
Figures captions

--- 2018-04-17 WholePlotTrajectories.tex
@@
 
-\JournalInfo{Hal xxx} % Journal information
-\Archive{DOI xxxx} % Additional notes (e.g. copyright, DOI, review/research article)
+\JournalInfo{Hal 00679993} % Journal information
+\Archive{} % Additional notes (e.g. copyright, DOI, review/research article)
 
-\PaperTitle{Post-Disturbance Tree Community Trajectories in a Neotropical Forest} % Article title
+\PaperTitle{Appendix I} % Article title
 
 \Authors{
-Ariane MIRABEL\textsuperscript{1*}\\ Eric Marcon\textsuperscript{1}\\ Bruno Hérault\textsuperscript{2 3}
+Ariane MIRABEL\textsuperscript{1*}\\ Eric MARCON\textsuperscript{1}\\ Bruno HERAULT\textsuperscript{2}
 } % Authors
@@
 \textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles,
-Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}Cirad, Univ montpellier, UR Forests \& Societies.\\ \hspace{1em} Montpellier, France.\\\textsuperscript{3}INPHB, Institut National Polytechnique Félix Houphouet-Boigny\\ \hspace{1em} Yamoussoukro, Ivory Coast.
+Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}INPHB (Institut National Ploytechnique Félix Houphoüet Boigny)\\ \hspace{1em} Yamoussoukro, Ivory Coast
 }
-\affiliation{*\textbf{Corresponding author}: ariane.mirabel@ecofog.gf, http://www.ecofog.gf/spip.php?article47} % Corresponding author
+\affiliation{*\textbf{}: ariane.mirabel@ecofog.gf, http://www.ecofog.gf/spip.php?article47} % Corresponding author
 
-\Keywords{mot-clés, séparés par des virgules} % Keywords - if you don't want any simply remove all the text between the curly brackets
-\newcommand{\keywordname}{Keywords} % Defines the keywords heading name
+\Keywords{} % Keywords - if you don't want any simply remove all the text between the curly brackets
+\newcommand{\keywordname}{} % Defines the keywords heading name
 
@@
 \Abstract{
-Résumé de l'article.
+
 }
@@
 
-\usepackage{amsthm}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{proposition}{Proposition}[section]
-\theoremstyle{definition}
-\newtheorem{example}{Example}[section]
-\theoremstyle{definition}
-\newtheorem{exercise}{Exercise}[section]
-\theoremstyle{remark}
-\newtheorem*{remark}{Remark}
-\newtheorem*{solution}{Solution}
 \begin{document}
@@
 
-\section{Introduction}\label{introduction}
+The following information are available for this article:
 
-The large areas covered with tropical forests worldwide hold crucial
-economic, social and cultural value. They provide wood and multiple
-non-timber forest products, shelter for a diversified fauna, regulate
-the local climate, support the carbon, water and nutrient cycles, and
-ensure cultural and human well-being. The simultaneous increase of
-forests products demand and substantial climatic changes heightened the
-pressure on the remaining forests
-\citep{Gibson2011a, Morales-Hidalgo2015} and threatened the maintenance
-of communities structure, composition and functioning and their
-underlying dynamics \citep{Anderson-Teixeira2013, Sist2015}.
+\textbf{Figure A1}: Whenever the family or genus were recorded for a
+vernacular name, the pool of species sampled to replace the vernacular
+name was restricted to those pertaining to the same higher taxonomic
+level (i.e species of the same genus when the species is missing,
+species of the same family when the genus is missing).
 
-In tropical forests communities are shaped by a constant range of
-disturbance that changes the abiotic environment, as the light, heat and
-water fluxes, and the biotic interaction and competitive pressure. The
-cornerstone of tropical forests ecology is then to understand their
-response to disturbance and the corresponding mechanisms
-\citep{White2001, Chazdon2003a}. Forests response to disturbance has
-been largely studied through structural parameters rapid and convenient
-to measure, as aboveground biomass, tree height or stem density. These
-structural parameters have then been sucessfully modeled and assessed
-the maintenance of ecosystems processes and services
-\citep{Denslow2000, Blanc2009, Rutishauser2016}. However the response of
-tree species diversity although it is determinant of ecosystems
-productivity, stability and functioning \citep{Tilman2014} and it would
-be most probably impacted by the biotic and abiotic changes after
-disturbance \citep{CazzollaGatti2014}.
+\textbf{Figure A2}: Trajectories of seed mass classes proportions over
+30 years after disturbance, corresponding to the median (solid line) and
+0.025 and 0.975 percentile (gray envelope) observed after 50 iteration
+of the taxonomic uncertainty propagation. No gap filling process was
+applied in this case. Initial treatments are represented by solid lines
+colors with green for control, blue for T1,orange for T2 and red for T3.
 
-In the short-term, diversity dynamics demonstrated negligible or even
-positive impacts of disturbance on communities diversity, and have been
-formalized by the intermediate disturbance hypothesis (IDH) that
-predicts a maximized species diversity at intermediate level of
-disturbance\citep{Molino2001, Kariuki2006a, Berry2008a}. The IDH however
-has scarcely been tested in the long term and mainly relied on the
-analysis of species richness that gives limited or misleading
-information on forests recovery and functioning
-\citep{Martin2015, Chaudhary2016}. Communities composition and evenness
-are crucial to resolve conservation issues and assess the ecological
-rules shaping communities abundance distribution
-\citep{Magurran1988, Lavorel2002, Bellwood2006}. A promosing perspective
-is also the functional approach accounting for species biological
-attributes related to individuals fitness and ecosystem properties
-\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. Major
-functional traits revealing species ecology and performance were largely
-adopted through a workable framework assessing forests dynamics and
-functioning \citep{Diaz2005, Villeger2008a}. This approach for example
-highlighted in tropical rainforests the environmental filters fostering
-disturbance resistant species with rapid growth and efficient resources
-acquisition \citep{Molino2001, Haddad2008}. This was translated by the
-shift from ``conservative'' species with long-lived, dense tissues
-growing slowly but with scarce resources, to ``acquisitive''
-fast-growing species with light tissues and efficient resources use
-\citep{TerSteege2001, Reich2014, Herault2011}. Such shift would be
-detected with a combination of resource acquisition funcitonal traits
-(leaf area, density and chlorophyll content and stem specific gravity
-and bark thickness) and growth and reproduction life history traits
-(seed mass and maximum height) already proved relevant
-\citep{Wright2004, Westoby2006a, Chave2009b}. A functional approach
-combined to the analysis of taxonomic diversity and composition is then
-a reliable way to validate the IDH in the long term and from communities
-evenness and functinal perspective. A full interpretation of communities
-response would be completed with communities taxonomic and functional
-composition that would elucidate the determinism of forest recovery and
-the maintenance of intrinsinc differences. Those development would be as
-much information to understand the mechanisms of forests response to
-diversity and help conservation and management strategies.
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{knitr}\OperatorTok{::}\KeywordTok{include_graphics}\NormalTok{(}\StringTok{"ExternalFig/BayesianSheme.png"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-Here we investigated over 30 years the response of 75 ha of forests
-plots set up on a gradient of disturbance intensity, from 10 to 60\% of
-ecosystem biomass removed. We made use of a large functional traits
-database browsing major leaf, stem and seed traits and species maximum
-height to draw communities taxonomic and functional composition and
-diversity trajectories after disturbance. Specifically, we (i) tested
-the validity of the IDH arguing for enhanced diversity after moderate
-disturbance and (ii) confirmed the changes in ecological rules shaping
-communities, from competitive exclusion between ``conservative'',
-slow-growing tree species to environmental filtering of ``acquisitive'',
-fasting-growing ones. Eventually we examined the different trajectories
-to (iii) resolve the resilience of tropical forests and question their
-future dynamics regarding their state 30 years after disturbance.
+\begin{figure}
 
-\section{Material and Methods}\label{material-and-methods}
+{\centering \includegraphics{ExternalFig/BayesianSheme}
 
-\subsection{Study site}\label{study-site}
+}
 
-Paracou station in French Guiana (5°18'N and 52°53'W) is located in a
-lowland tropical rain forest corresponding to a tropical wet climate
-with mean annual precipitation averaging 2980 mm.y\textsuperscript{-1}
-(30-y period) with a 3-months dry season (\textless{} 100
-mm.months\textsuperscript{-1}) from mid-August to mid-November, and a
-one-month dry season in March \citep{Wagner2011}. Elevation ranges
-between 5 and 50 m and mean annual temperature is 26°C. Soils correspond
-to thin acrisols over a layer of transformed saprolite with low
-permeability generating lateral drainage during heavy rains. The
-experiment corresponds to a network of twelve 6.25ha plots that have
-undergone a gradient of three logging, thinning and fuelwood treatments
-(Table \ref{tab:Tab1}). Disturbance treatments were attributed according
-to a randomized plot design with three replicate blocks of four plots.
-The disturbance corresponds to averages of 10 trees removed per hectare
-with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1),
-32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40
-cm DBH for treatment 3 (T3). Treatments T2 and T3 besides included the
-thinning of trees by poison girdling \citep{Blanc2009}.
-
-\begin{table}
-
-\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.}
-\centering
-\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
-\toprule
-Treatment & Timber & Thinning & Fuelwood & \%AGB lost\\
-\midrule
-Control &  &  &  & 0\\
-T1 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha &  &  & $[12\%-33\%]$\\
-T2 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30 trees/ha &  & $[33\%-56\%]$\\
-T3 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & $[35\%-56\%]$\\
-\bottomrule
-\end{tabu}
-\end{table}
-
-\subsection{Inventories protocol and dataset
-collection}\label{inventories-protocol-and-dataset-collection}
-
-The study site corresponds to a tropical rainforest with a dominance of
-Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical
-families. In the twelve experimental plots of the experiment, all trees
-above 10 cm DBH were mapped and measured annually since 1984. During
-inventories, trees were first identified with a vernacular name assigned
-by the field team, and afterward with a scientific name assigned by a
-botanist during regular botanical campaigns. In 1984, specific
-vernacular names were given to 62 commercial or common species whereas
-other less common species were identified under two identifiers only
-separating trees and palm trees. The botanical campaigns carried every 5
-to 6 years to identify all trees at the species level only started in
-2003 and identification practices varied among plots and successive
-campaigns. This raised methodological issues as vernacular names usually
-correspond to different botanical species, resulting in significant
-taxonomic uncertainties that had to be propagated to composition and
-diversity metrics throught a Bayesian framework. The uncertainty
-propagation was done by the replenishment of inventories completed at
-genus level from real incomplete ones on the basisof
-vernacular/botanical names association.
-
-Vernacular names were replaced through multinomial trials
-\(M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)\)
-based on the observed association probability
-\(\big[\alpha_1, \alpha_2,…, \alpha_3\big]\) between each vernacular
-name \emph{v} and the species \(\big[s_1, s_2, …, s_N\big]\) recorded in
-the inventory.
-
-To avoid remaining identification caveats and consider complete
-inventories, the simulated inventories were then reported at genus. See
-appendix 1 and \citet{Aubry-Kientz2013} for the detailed methodology.
-
-The functional diversity metrics used a dataset for 6 functional traits
-representing leaf economics (leaves thickness, toughness, total
-chlorophyll content and specific leaf area, the leaf area per unit dry
-mass) and wood economics (wood specific gravity and bark thickness), and
-life history traits (maximum specific height and seed mass).
-
-The trait database came from the BRIDGE project {[}\^{}1{]} where traits
-values were assessed from a selection of individuals located in nine
-permanent plots in French Guiana, including two in Paracou. Missing
-trait values were filled using multivariate imputation by chained
-equation (mice) restricted to samples pertaining to the next higher
-taxonomic level, in order to account for the phylogenetic signal of the
-functional traits. The dataset comprised 294 botanical species
-pertaining to 157 botanical genera.
-
-Whenever a species inventoried was not in the dataset, it was attributed
-a set of traits values randomly sampled among species of the same next
-higher taxonomic level. As seed mass information corresponds to a
-classification into mass classes, no data filling process was applied so
-analysis were performed considering the 414 botanical species of the
-seed mass dataset. All composition and diversity metrics corresponded to
-the average obtained after 50 iterations of the taxonomic uncertainty
-propagation framework and of the filling process of missing trait
-values.
-
-{[}\^{}1{]} http://www.ecofog.gf/Bridge/
-
-\subsection{Composition and diversity
-metrics}\label{composition-and-diversity-metrics}
-
-To counter the remaining taxonomic uncertainty plots taxonomic
-composition and diversity were analysed at the genus level, \emph{i.e.}
-referring to the genus of observed or trialed botanical names. The
-analysis of taxonomic and functional composition of plots along time was
-visualized in a two-dimensional ordination space based on non-metric
-dimensional scaling of succesive floristic or functional inventories.
-Plots trajectories along time was reported comparatively to the
-inventories in 1989, 5 years after disturbance, which corresponded to
-first inventory with a sufficient degree of uncertainty (\textless{}30\%
-of undetermined trees). The inventories dissimilarity compared to the
-reference 1989 inventory was reported using occurrence-based (Jaccard)
-and abundance-based (Bray-Curtis) similarity measures. The trajectory of
-inventories along time was visualized with the euclidean distance in the
-two-dimensional ordination space to the 1989 inventory. The functional
-trajectories of the leaf and stem and life traits were also visualized
-with the community weighted means (CWM), representing the average trait
-value in a community weighted by relative abundance of the species
-carrying each value \citep{Diaz2007, Garnier2004}. Species seed mass
-correspond to 5 classes of increasing mass, the seed mass trajectories
-were reported by the proportion of each class recorded in the
-inventories. To determine whether communities' assembly deviated from
-the expectation of a random process we compared the observed similarity
-patterns to those generated by 50 repetitions of stochastic null models.
-The null model for taxonomic composition randomly shuffled individuals
-among plots while preserving the overall species abundance and plots'
-tree density. The null model for functional composition produced random
-traits assemblages by permuting the set of functional traits values
-among species.
-
-The taxonomic diversity was assessed through Richness and the Hill
-number translation of Shannon and Simpson indices \citep{Hill1973}. To
-tackle the unequal number of recruited trees among treatments the
-indices bias corrected estimator were used, following
-\citep{Chao2015, Marcon2015b}. These three indices belong to the set of
-HCDT or generalized entropy, respectively corresponding to the 0, 1 and
-2 order of diversity (q), which proved well suited for diversity studies
-\citep{Patil1982, Tothmeresz1995}. The functional diversity was reported
-using the Rao index of quadratic entropy which combines species
-abundance distribution and average pairwise dissimilarity based on all
-functional and life traits.
-
-\section{Results}\label{results}
-
-\subsection{Taxonomic and functional
-composition}\label{taxonomic-and-functional-composition}
-
-\subsubsection{Composition trajectories}\label{composition-trajectories}
-
-Over time, 828388 trees and 591 botanical species pertaining to 223
-genus and 64 botanical families were recorded. Trajectories of taxonomic
-and functional composition after disturbance were examined through the
-ordination of successive inventories from 1989 (5 years after
-disturbance) in the two-dimensional space from NMDS analysis based on
-flora inventories and the corresponding functional composition.
-Classifications were performed using either abundance-based Bray-Curtis
-(Figure \ref{fig:Fig1}) or incidence-based Jaccard dissimilarity (data
-not shown).
-
-Both taxonomic and functional composition were substantially affected by
-disturbance, and the maximum dissimilarity to 1984 reference inventory
-composition was positively correlated with the disturbance intensity.
-Until 30 years after disturbance, all disturbed plots taxonomic
-composition remained significantly dissimilar to that of the 1989
-reference inventory composition (Figure \ref{fig:Fig1}). All plots,
-though, displayed a unimodal trajectory with a the return toward the
-taxonomic and functional composition of the 1989 reference inventory
-suggesting a shift towards a cyclic regime (Figure \ref{fig:Fig2}). For
-taxonomic composition, the maximum dissimilarity to 1984 inventory
-composition was reach around 20 years after disturbance for the T2 and
-T3 plots and for one T1 plot while control and other T1 plots continued
-to increase (Figure \ref{fig:Fig2}.a). For functional composition, the
-maximum dissimilarity to 1984 inventory composition was reach from 15 to
-20 years after disturbance for the T2 and T3 plots while T1 and control
-plots continued to increase (Figure \ref{fig:Fig2}.b). The functional
-composition distancing from the reference inventory was also positively
-correlated with disturbance intensity and similarly stabilized or
-reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance
-(Figure \ref{fig:Fig2}).
-
-The coordinates of the functional traits were measured in the
-two-dimensional ordination space that mapped plots evolution along time
-(graphs not showed) revealed a trajectory of disturbed plots toward
-acquisitive functional strategies (from high WD to high SLA and
-chlorophyll content).
+\caption{Scheme of the Bayesian process propagating the taxonomic uncertainty}\label{fig:FigA1}
+\end{figure}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig1-1}
-
-}
-
-\caption{Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
-Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).}\label{fig:Fig1}
-\end{figure*}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig2-1}
-
-}
-
-\caption{Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of **(a)** taxonomic composition at genus level and **(b)** functional composition.
-Distance are abundance-based Bary-Curtis metric.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.
-}\label{fig:Fig2}
-\end{figure}
-
-\subsubsection{Traits community weighted
-means}\label{traits-community-weighted-means}
-
-For all plots the trajectories of the community weighted means (CWM)
-were drawn for the 8 functional and life history traits (Leaf thickness,
-chlorophyll content, toughness and specific area,wood specific gravity
-and barck thickness and seed mass and maximum adult height) (Figure
-\ref{fig:Fig3}).
-
-To compensate the intrinsinc difference among plots the trajectories
-drawn correspond to the difference in value between the reference
-inventory in 1984 (5 year after disturbance) and the successive years
-inventoried.
-
-Except for leaf cholorphyll content, which displayed higher difference
-between plots than among treatments which may be due to the completeness
-of dataset, all CWM trajectories corresponded to significant changes
-after disturbance. All functional traits and seed mass proportions
-displayed a unimodal trajectories but with different times at maximum
-and different values 30 years after disturbance. The weighted means of
-communities specific maximum height at adult stage (\emph{Hmax}), leaf
-toughness (\emph{L\_toughness}) and wood specific gravity (\emph{WD})
-remained significantly lower than their initial value and than these of
-the control plots (Figure \ref{fig:Fig3}). The weighted means of bark
-thickness (\emph{Bark\_thick}) similarly remained substantilly higher
-than initially for all disturbed plots while the specfic leaf area
-(\emph{SLA}) had almost recovered its initial value and this of the
-undisturbed plots at the end of the experiment.
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig3-1}
-
-}
-
-\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, *L_thickness*, chlorophyll content, *L_chloro*, toughness, *L_toughness* and specific area, *SLA*), 2 stem traits (wood specific gravity, *WD*, and bark thickness, *Bark-thick*) and one life trait (Specific maximum height at adult stage,*Hmax*).
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig3}
-\end{figure*}
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig4-1}
+{\centering \includegraphics[width=1\linewidth]{AppendixI_files/figure-latex/FigA2-1}
 
@@
 No gap filling process was applied in this case.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig4}
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:FigA2}
 \end{figure*}
-
-\subsection{Disturbance impact on
-diversity}\label{disturbance-impact-on-diversity}
-
-\subsubsection{Taxonomic diversity}\label{taxonomic-diversity}
-
-Trajectories of Richness, Shannon and Simpson taxonomic diversity were
-examined at genus level in relation to the 1989 inventories (5 years
-after disturbance) (Figure \ref{fig:Fig5}).
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig5-1}
-
-}
-
-\caption{Trajectories of the difference to reference inventory over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig5}
-\end{figure*}
-
-While for undisturbed plots the Richness, Shannon and Simpson diversity
-remained comparable to the values in 1989, the trajectories of all
-disturbed plots showed significant increase positively correlated with
-the disturbance intensity (Figure \ref{fig:Fig5}). The Richness was less
-affected by the initial disturbance except for the higher intensity (T3
-plots) (Figure \ref{fig:Fig5}). The Shannon and Simpson diversities
-however first significantly increased, illustrating an increasing
-evenness, before reaching a maximum around 15 years after disturbance
-for all plots except one from T1 treatment and one from T2 treatment
-(Figure \ref{fig:Fig5}).
-
-\subsubsection{Functional diversity}\label{functional-diversity}
-
-For all undisturbed plots the Rao diversity remained comparable to the
-values in 1989, the trajectories of all disturbed plots showed
-significant increase positively correlated with the disturbance
-intensity (Figure \ref{fig:Fig6}). Although a first increase of plots
-functional diversity that was all the more important than initial
-disturbance was intense, 30 after disturbance the diversity was similar
-among disturbed plots and significantly higher than this of control
-plots.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig6-1}
-
-}
-
-\caption{Trajectories of the Rao functional diversity over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:Fig6}
-\end{figure}
-
-\section{Discussion}\label{discussion}
-
-Our study showed the significant impact of disturbance on tropical
-forests communities. The subsequent diversity trajectories confirmed the
-intermediate disturbance hypothesis debated for tropical forests through
-their correlation with disturbance intensity. Besides it revealed the
-contrasting response of taxonomic and functional characteristics,
-specifically the decoupling between communities taxonomic evenness and
-their functional diversity and dominant functional traits values. The
-long-term disturbance trajectories observed highlighted the unachieved
-but consistent recovery of communities assembly for the lowest
-disturbance intensity but questioned it after higher disturbance.
-
-\subsection{A validation of the intermediate disturbance
-hypothesis}\label{a-validation-of-the-intermediate-disturbance-hypothesis}
-
-The monitoring 30 years of disturbed forest communities confirmed the
-limited direct impact of low disturbance on species richness as it was
-found for impact survey of selective logging
-\citep{Cannon1998, Baraloto2012}. Only the most disturbed plots had not
-fully achieved the ongoing recovery of genus richness, although all had
-already reached equivalent levels as those of some control plots. Such
-long time richness recovery had already been observed in several logging
-experiments, which also highlighted the role of random disturbance
-damage in the richness decrease \citep{deAvila2015, Hu2018}.
-
-The evenness of all disturbed plots and the richness of low disturbance
-plots on the other hand mainly followed an asymptotic growth sharply
-increasing until 15 years after disturbance. Such increased evenness is
-explained by a much more homogeneous species distribution after
-disturbance. Communities composition after disturbance is either due to
-the old, pre-disturbance survivors or to the recruited trees. The
-composition of old survivors proved to mirror the initial community
-\citep{Herault2018}, the observed composition turnover would stem from
-an enhanced the growth and survival of previously infrequent species
-which, along with the IDH, reorganizes the typical high dominance
-structure of hyperdiverse mature forests. This increase in taxonomic
-diversity was accompanied by an increased taxonomic dissimilarity to
-initial state along time, translating a a taxonomic turnover benefiting
-to pioneers and light demanding species, as illustrated by the
-communities functional shifts towards resource-acquisitive strategies
-(sharp increase in the SLA, leaf thickness and bark thickness and
-decrease in wood density, leaf toughness and maximum height)
-\citep{Westoby1998, Wright2004, Reich2014}. Post-disturbance changes in
-abiotic environment and competitive pressure then well favored pioneers
-outcompeting other species when abiotic resources are abundant but
-otherwise excluded in mature forests by long-lived, resistant and shade
-tolerant species. The trajectories of proportions among seed mass
-classes besides revealed the enhanced growth of small seeded, large
-dispersive, species and therefore the importance of species reproductive
-strategy which further supported the role of species dispersion and
-demographic strategy for the post-disturbance dynamics, as suggested by
-the IDH \citep{TerSteege2001, Flores2006, Haddad2008}.
-
-\subsection{On the recovery of disturbed
-communities}\label{on-the-recovery-of-disturbed-communities}
-
-The functional diversity followed a hump-back trajectory which maximum
-was positively correlated to the disturbance intensity
-(\(\rho_{spearman}=\)) and after 30 years, all disturbed plots had
-recovered diversities close to their initial values. Similar hump-back
-trajectories were followed by all leaf and stem functional traits and
-life history traits, to the exception of leaf chlorophyll content.
-Although the return to initial values was still ongoing the recovery of
-functional diversity and average communities trait values was
-undeniable. This translated the recovery of ecosystems processes
-\citep{Guariguata2001}, as functional traits are the most direct link
-between biodiversity and ecosystem functioning \citep{Diaz2005}. Same
-recovery trajectories were followed by taxonomic and functional
-composition which argued for communities convergence and the maintenance
-of plots initial difference
-\citep{Hubbell1999, Molino2001, Baraloto2012a}. These compositon
-difference, and probably also some abiotic parameters, besides strongly
-determined the maximum and time path of the trajectories which, despite
-an homogeneous trend were quite variable among plots.
-
-\subsection{Functional redundancy of disturbed
-ecosystems}\label{functional-redundancy-of-disturbed-ecosystems}
-
-Despite the consistent recovery of initial diversity and composition
-there was a time lag between the communities taxonomic and functional
-characteristics. Taxonomic composition had longer time path and while
-communities had recovered their functional diveristy, their taxonomic
-composition and evenness remained higher than before disturbance. This
-delay between functional and taxonomic dynamics was already observed for
-grasslands \citep{Tilman1997, Mouillot2011} and more recently for
-tropical forests \citep{Lohbeck2015, Guariguata2001}. Communities
-functional diversity rely, according to the ``vegetation quantity
-effect'' \citep{Grime1998}, on the functional strategy of dominant
-species. Communities functional trajectory is then first driven by the
-increase of dominant species diversity and evenness following
-disturbance and then by the emergence of recruitment of species
-progressively resembling the old pre-disturbance community which
-decreased the diversity. Dominant species then restored the functional
-diversity and the dominant functional type of the community although the
-functional and taxonomic composition remained altered and the infrequent
-species of initial communities are still missing to plots recovery. The
-functional traits trajectories suggest that those species would match
-the dominant functional type therefore increasing communities functional
-redundancy. The functional overlap between species is typical of the
-huge biodiversity of tropical forests \citep{Bellwood2006}, and proved
-far from recovery 03 years after disturbance. It is though a determinant
-of forests resilience and should be accounted for
-\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
-
-Long-term alteration of functional redundancy besides came along with
-persistent compositional changes which most probably impact disturbance
-resistant species \citep{Haddad2008} and probably lianas or epiphytes
-\citep{Martin2013}, and the modification of environmental conditions
-like soils nutrient cycling and compaction \citep{Olander2005} forest
-resilience remains under question \citep{Chazdon2003a}. New conditions
-would not only be longer lasting but self-maintained as tied to
-disturbance regime \citep{Burslem2000}. Specificaly, this would impair
-species contingent to undisturbed forests, threatening their
-maintenance, and run the risk to loose cornerstone species and trigger
-unexpected ecological consequences
-\citep{Jones1994, Diaz2005, Gardner2007}.
 
@@
 \bibliographystyle{mee}
-\bibliography{references.bib}
+\bibliography{}
 

### 2018-04-19 WholePlotTrajectories.Rmd
Figures captions

@@
 The functional trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004].
-Species seed mass correspond to 5 classes of increasing mass, the seed mass trajectories were reported by the proportion of each class recorded in the inventories.
-To determine whether communities’ assembly deviated from the expectation of a random process we compared the observed similarity patterns to those generated by 50 repetitions of stochastic null models.
-The null model for taxonomic composition randomly shuffled individuals among plots while preserving the overall species abundance and plots’ tree density. 
-The null model for functional composition produced random traits assemblages by permuting the set of functional traits values among species.
+Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were  therefore reported as the proportion of each class in the inventories.
 
@@
 Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in the two-dimensional space from NMDS analysis based on flora inventories and the corresponding functional composition.
-Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:Fig1)) or incidence-based Jaccard dissimilarity (data not shown).
+Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity (data not shown).
 
 Both taxonomic and functional composition were substantially affected by disturbance, and the maximum dissimilarity to 1984 reference inventory composition was positively correlated with the disturbance intensity.
-Until 30 years after disturbance, all disturbed plots taxonomic composition remained significantly dissimilar to that of the 1989 reference inventory composition (Figure \@ref(fig:Fig1)).
-All plots, though, displayed a unimodal trajectory with a the return toward the taxonomic and functional composition of the 1989 reference inventory suggesting a shift towards a cyclic regime (Figure \@ref(fig:Fig2)).
-For taxonomic composition, the maximum dissimilarity to 1984 inventory composition was reach around 20 years after disturbance for the T2 and T3 plots and for one T1 plot while control and other T1 plots continued to increase (Figure \@ref(fig:Fig2).a).
-For functional composition, the maximum dissimilarity to 1984 inventory composition was reach from 15 to 20 years after disturbance for the T2 and T3 plots while T1 and control plots continued to increase (Figure \@ref(fig:Fig2).b).
-The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance (Figure \@ref(fig:Fig2)).
+Until 30 years after disturbance, all disturbed plots taxonomic composition remained significantly dissimilar to that of the 1989 reference inventory composition (Figure \@ref(fig:NMDSplans)).
+All plots, though, displayed a unimodal trajectory with a the return toward the taxonomic and functional composition of the 1989 reference inventory suggesting a shift towards a cyclic regime (Figure \@ref(fig:NMDSdivergence)).
+For taxonomic composition, the maximum dissimilarity to 1984 inventory composition was reach around 20 years after disturbance for the T2 and T3 plots and for one T1 plot while control and other T1 plots continued to increase (Figure \@ref(fig:NMDSdivergence).a).
+For functional composition, the maximum dissimilarity to 1984 inventory composition was reach from 15 to 20 years after disturbance for the T2 and T3 plots while T1 and control plots continued to increase (Figure \@ref(fig:NMDSdivergence).b).
+The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance (Figure \@ref(fig:NMDSdivergence)).
 
@@
 
-```{r Fig1, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
+(ref:Bolda) **(a)**
+
+(ref:Boldb) **(b)**
+
+(ref:Boldc) **(c)**
+
+```{r NMDSplans, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of (ref:Bolda) flora composition and (ref:Boldb) functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
 Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories.
@@
 
-```{r Fig2, out.width = '100%', fig.asp = 1.2/2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of **(a)** taxonomic composition at genus level and **(b)** functional composition.
+```{r NMDSdivergence, out.width = '100%', fig.asp = 1.2/2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of v taxonomic composition at genus level and (ref:Boldb) functional composition.
 Distance are abundance-based Bary-Curtis metric.
@@
 
-For all plots the trajectories of the community weighted means (CWM) were drawn for the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area,wood specific gravity and barck thickness and seed mass and maximum adult height) (Figure \@ref(fig:Fig3)).
+For all plots the trajectories of the community weighted means (CWM) were drawn for the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area,wood specific gravity and barck thickness and seed mass and maximum adult height) (Figure \@ref(fig:CWM)).
 
@@
 All functional traits and seed mass proportions displayed a unimodal trajectories but with different times at maximum and different values 30 years after disturbance. 
-The weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value and than these of the control plots (Figure \@ref(fig:Fig3)).
+The weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value and than these of the control plots (Figure \@ref(fig:CWM)).
 The weighted means of bark thickness (*Bark_thick*) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value and this of the undisturbed plots at the end of the experiment.
 
-```{r Fig3, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, *L_thickness*, chlorophyll content, *L_chloro*, toughness, *L_toughness* and specific area, *SLA*), 2 stem traits (wood specific gravity, *WD*, and bark thickness, *Bark-thick*) and one life trait (Specific maximum height at adult stage,*Hmax*).
+(ref:capThickness) *L_thickness*
+
+(ref:capChloro) *L_chloro*
+
+(ref:capThoughness) *L_toughness*
+
+(ref:capSLA) *SLA*
+
+(ref:capWD) *WD*
+
+(ref:capBarkThick) *Bark-thick*
+
+(ref:capHmax) *Hmax*
+
+```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWD), and bark thickness, (ref:capBarkThick)) and one life trait (Specific maximum height at adult stage, (ref:capHmax)).
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
@@
 
-```{r Fig4, echo=FALSE, out.width='100%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-No gap filling process was applied in this case.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
-load("DB/Smass_ClassProp")
-
-SeedMassProp(Smass)
-
-```
-
 ## Disturbance impact on diversity
@@
 
-Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined at genus level in relation to the 1989 inventories (5 years after disturbance) (Figure \@ref(fig:Fig5)).
+Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined at genus level in relation to the 1989 inventories (5 years after disturbance) (Figure \@ref(fig:DivTaxo)).
 
-```{r Fig5, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
+```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 
-While for undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity ($\rho_{spearman}^{Shannon}=0.87$, and $\rho_{spearman}^{Simpson}=0.88$) (Figure \@ref(fig:Fig5)).
-The Richness was less affected by the initial disturbance except for the higher intensity (T3 plots) (Figure \@ref(fig:Fig5)). 
-The Shannon and Simpson diversities however first significantly increased, illustrating an increasing evenness, before reaching a maximum around 15 years after disturbance for all plots except one from T1 treatment and one from T2 treatment (Figure \@ref(fig:Fig5)).
+While for undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity ($\rho_{spearman}^{Shannon}=0.87$, and $\rho_{spearman}^{Simpson}=0.88$) (Figure \@ref(fig:DivTaxo)).
+The Richness was less affected by the initial disturbance except for the higher intensity (T3 plots) (Figure \@ref(fig:DivTaxo)). 
+The Shannon and Simpson diversities however first significantly increased, illustrating an increasing evenness, before reaching a maximum around 15 years after disturbance for all plots except one from T1 treatment and one from T2 treatment (Figure \@ref(fig:DivTaxo)).
 
@@
 
-For all undisturbed plots the Rao diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity (Figure \@ref(fig:Fig6)).
+For all undisturbed plots the Rao diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity (Figure \@ref(fig:DivFun)).
 Although a first increase of plots functional diversity that was all the more important than initial disturbance was intense, 30 after disturbance the diversity was similar among disturbed plots and significantly higher than this of control plots.
 
-```{r Fig6, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
+```{r DivFun, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 
-The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging [@Cannon1998;@Baraloto2012].
+The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging [@Cannon1998;@Baraloto2012a].
 Only the most disturbed plots had not fully achieved the ongoing recovery of genus richness, although all had already reached equivalent levels as those of some control plots.

### 2018-04-19 WholePlotTrajectories.tex
Figures captions

@@
 carrying each value \citep{Diaz2007, Garnier2004}. Species seed mass
-correspond to 5 classes of increasing mass, the seed mass trajectories
-were reported by the proportion of each class recorded in the
-inventories. To determine whether communities' assembly deviated from
-the expectation of a random process we compared the observed similarity
-patterns to those generated by 50 repetitions of stochastic null models.
-The null model for taxonomic composition randomly shuffled individuals
-among plots while preserving the overall species abundance and plots'
-tree density. The null model for functional composition produced random
-traits assemblages by permuting the set of functional traits values
-among species.
+corresponded to 5 classes of increasing mass, seed mass trajectories
+were therefore reported as the proportion of each class in the
+inventories.
 
@@
 Classifications were performed using either abundance-based Bray-Curtis
-(Figure \ref{fig:Fig1}) or incidence-based Jaccard dissimilarity (data
-not shown).
+(Figure \ref{fig:NMDSplans}) or incidence-based Jaccard dissimilarity
+(data not shown).
 
@@
 composition remained significantly dissimilar to that of the 1989
-reference inventory composition (Figure \ref{fig:Fig1}). All plots,
+reference inventory composition (Figure \ref{fig:NMDSplans}). All plots,
 though, displayed a unimodal trajectory with a the return toward the
 taxonomic and functional composition of the 1989 reference inventory
-suggesting a shift towards a cyclic regime (Figure \ref{fig:Fig2}). For
-taxonomic composition, the maximum dissimilarity to 1984 inventory
-composition was reach around 20 years after disturbance for the T2 and
-T3 plots and for one T1 plot while control and other T1 plots continued
-to increase (Figure \ref{fig:Fig2}.a). For functional composition, the
-maximum dissimilarity to 1984 inventory composition was reach from 15 to
-20 years after disturbance for the T2 and T3 plots while T1 and control
-plots continued to increase (Figure \ref{fig:Fig2}.b). The functional
-composition distancing from the reference inventory was also positively
-correlated with disturbance intensity and similarly stabilized or
-reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance
-(Figure \ref{fig:Fig2}).
+suggesting a shift towards a cyclic regime (Figure
+\ref{fig:NMDSdivergence}). For taxonomic composition, the maximum
+dissimilarity to 1984 inventory composition was reach around 20 years
+after disturbance for the T2 and T3 plots and for one T1 plot while
+control and other T1 plots continued to increase (Figure
+\ref{fig:NMDSdivergence}.a). For functional composition, the maximum
+dissimilarity to 1984 inventory composition was reach from 15 to 20
+years after disturbance for the T2 and T3 plots while T1 and control
+plots continued to increase (Figure \ref{fig:NMDSdivergence}.b). The
+functional composition distancing from the reference inventory was also
+positively correlated with disturbance intensity and similarly
+stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after
+disturbance (Figure \ref{fig:NMDSdivergence}).
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig1-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/NMDSplans-1}
 
@@
 
-\caption{Trajectories of the plots in terms of **(a)** flora composition and **(b)** functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
+\caption{Trajectories of the plots in terms of \textbf{(a)} flora composition and \textbf{(b)} functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
 Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).}\label{fig:Fig1}
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).}\label{fig:NMDSplans}
 \end{figure*}
@@
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig2-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/NMDSdivergence-1}
 
@@
 
-\caption{Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of **(a)** taxonomic composition at genus level and **(b)** functional composition.
+\caption{Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of v taxonomic composition at genus level and \textbf{(b)} functional composition.
 Distance are abundance-based Bary-Curtis metric.
@@
 The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.
-}\label{fig:Fig2}
+}\label{fig:NMDSdivergence}
 \end{figure}
@@
 and barck thickness and seed mass and maximum adult height) (Figure
-\ref{fig:Fig3}).
+\ref{fig:CWM}).
 
@@
 remained significantly lower than their initial value and than these of
-the control plots (Figure \ref{fig:Fig3}). The weighted means of bark
+the control plots (Figure \ref{fig:CWM}). The weighted means of bark
 thickness (\emph{Bark\_thick}) similarly remained substantilly higher
@@
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig3-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/CWM-1}
 
@@
 
-\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, *L_thickness*, chlorophyll content, *L_chloro*, toughness, *L_toughness* and specific area, *SLA*), 2 stem traits (wood specific gravity, *WD*, and bark thickness, *Bark-thick*) and one life trait (Specific maximum height at adult stage,*Hmax*).
+\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life trait (Specific maximum height at adult stage, \emph{Hmax}).
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig3}
-\end{figure*}
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig4-1}
-
-}
-
-\caption{Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-No gap filling process was applied in this case.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig4}
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.}\label{fig:CWM}
 \end{figure*}
@@
 examined at genus level in relation to the 1989 inventories (5 years
-after disturbance) (Figure \ref{fig:Fig5}).
+after disturbance) (Figure \ref{fig:DivTaxo}).
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig5-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivTaxo-1}
 
@@
 
-\caption{Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
+\caption{Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities \textbf{(a)} Richness, \textbf{(b)} Shannon and \textbf{(c)} Simpson diversities.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:Fig5}
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:DivTaxo}
 \end{figure*}
@@
 the disturbance intensity (\(\rho_{spearman}^{Shannon}=0.87\), and
-\(\rho_{spearman}^{Simpson}=0.88\)) (Figure \ref{fig:Fig5}). The
+\(\rho_{spearman}^{Simpson}=0.88\)) (Figure \ref{fig:DivTaxo}). The
 Richness was less affected by the initial disturbance except for the
-higher intensity (T3 plots) (Figure \ref{fig:Fig5}). The Shannon and
+higher intensity (T3 plots) (Figure \ref{fig:DivTaxo}). The Shannon and
 Simpson diversities however first significantly increased, illustrating
@@
 disturbance for all plots except one from T1 treatment and one from T2
-treatment (Figure \ref{fig:Fig5}).
+treatment (Figure \ref{fig:DivTaxo}).
 
@@
 significant increase positively correlated with the disturbance
-intensity (Figure \ref{fig:Fig6}). Although a first increase of plots
+intensity (Figure \ref{fig:DivFun}). Although a first increase of plots
 functional diversity that was all the more important than initial
@@
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/Fig6-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivFun-1}
 
@@
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:Fig6}
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:DivFun}
 \end{figure}
@@
 found for impact survey of selective logging
-\citep{Cannon1998, Baraloto2012}. Only the most disturbed plots had not
+\citep{Cannon1998, Baraloto2012a}. Only the most disturbed plots had not
 fully achieved the ongoing recovery of genus richness, although all had

### 2018-04-19 Scripts/DivFunctional_trajectories.R

@@ -1,0 +1,89 @@
+library(RColorBrewer)
+library(mice)
+library(ade4)
+library(cluster)
+source("Scripts/Vernacular_handle.R")
+#source("NMDS/TraitsDB.R")
+source("Scripts/TraitsMiceFilling.R")
+load("DB/Paracou_R_Subdivided_ok")
+load("DB/Alpha_Plots")
+
+###################################   Data
+
+Traits1<-read.csv("DB/BridgeOK.csv",sep=";",na.strings="")
+Traits2<-read.csv("DB/DataLifeTraits.csv",sep=";",na.strings="")
+TraitsName<-c("Hmax","L_thickness","L_chloro","L_toughness","L_DryMass","SLA","WD","Bark_thick")
+
+InventorySp<-do.call(rbind,lapply(LivingStand_all,function(yr){
+  ret<-do.call(rbind,yr)[,c("Famille","Genre","name")]
+  ret<-ret[which(!grepl("Indet.",ret[,"name"])),]
+  return(ret[which(!duplicated(ret)),])}))
+InventorySp<-InventorySp[which(!duplicated(InventorySp)),]
+
+#Plots of the different treatments
+T0<-c(1,6,11);T1<-c(2,7,9);T2<-c(3,5,10);T3<-c(4,8,12)
+treatments<-list(T0,T1,T2,T3);names(treatments)<-c("T0","T1","T2","T3")
+
+Pby<-c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014","2016","2017")
+dates<-sort(names(LivingStand_all)[which(!names(LivingStand_all)%in%Pby)])
+
+# Moving average function
+smooth<-function(mat,larg){return(do.call(cbind,lapply(1:ncol(mat),function(step){
+  range<-max(1,step-larg):min(ncol(mat),step+larg)
+  rowSums(mat[,range])/length(range)})))}
+
+
+CompleteFun<-lapply(1:4,function(t){
+  
+  treat<-treatments[[t]]
+  Nrep<-50
+  
+   Mplot<-mclapply(1:Nrep,function(r){
+      
+      VernDiv_living<-do.call(cbind,lapply(dates,function(y){
+        x1<-LivingStand_all[[which(names(LivingStand_all)==y)]] 
+        
+        M<-x1[which(names(x1)%in%treat)] 
+        
+        if(length(M)!=0){ #If there are surveys for these year and treatment
+          
[…]

### 2018-04-19 Scripts/TraitsMiceFillingOld.R

@@ -1,100 +1,0 @@
-library(mice)
-
-Traits_filling<-function(Traits1,Traits2){
-  Traits1<-Traits1[which(apply(Traits1,1,function(li){return(any(!is.na(li)))})),]
-  Traits2<-Traits2[which(apply(Traits2,1,function(li){return(any(!is.na(li)))})),]
-  
-## Traits correlation?
-Traits1["name"]<-as.factor(paste(Traits1$Genus,"_",Traits1$species,sep=""))
-Traits2["name"]<-sub(" ","_",Traits2[,"Name"])
-
-traits1<-Traits1[,c("Family","Genus","species","name","bar_code","thickness","SPAD","toughness","dry_mass","traits_surf_area","ind_surf_area",
-                    "sapwood_dens","moisture","bark_thick")]
-colnames(traits1)<-c("Family","Genus","species","name","bar_code","L_thickness","L_chloro","L_toughness","L_DryMass","LA","SLA","WD",
-                     "moisture","Bark_thick")
-traits2<-Traits2[,c("name","Hauteur","Masse")]
-colnames(traits2)<-c("name","Hmax","S_mass")
-
-traits<-merge(traits2,traits1,by="name")
-
-##SLA calc for empty values with LA available
-traits[which(!is.na(traits[,"LA"])),"SLA"]<-
-  traits[which(!is.na(traits[,"LA"])),"LA"]/traits[which(!is.na(traits[,"LA"])),"L_DryMass"]
-Seltraits<-c("Hmax","S_mass","L_thickness","L_chloro","L_toughness","L_DryMass","SLA","WD","moisture","Bark_thick")
-traits<-traits[,c("Family","Genus","name","bar_code",Seltraits)]
-
-##########################  Only some missing values: MICE regression to keep all available information
-### Non entire missing lines
-traitsPartial<-traits[which(apply(traits[,Seltraits],1,function(li){return(any(!is.na(li)))})),]
-traitsPartial<-traitsPartial[which(apply(traitsPartial[,Seltraits],1,function(li){return(any(is.na(li)))})),]
-traitsComp<-traits[which(apply(traits[,Seltraits],1,function(li){return(!any(is.na(li)))})),]
-traitsEmpty<-traits[which(apply(traits[,Seltraits],1,function(li){return(!any(!is.na(li)))})),]
-
[…]

### 2018-04-19 Scripts/Vernacular_handleOld.R

@@ -1,119 +1,0 @@
-#### Loadings
-library(vegan)
-library(reshape2)
-library(ade4)
-library(shape)
-library(cluster)
-library(mvtnorm)
-library(entropart)
-library(lattice)
-library(plotrix)
-library(gtools)
-library(RcppParallel)
-library(parallel)
-
-# Obtain the vernacular/botanical observed association frequency
-alpha_construct<-function(Plot){
-  mat_eff<-as.data.frame.matrix(with(Plot, xtabs(~name+nomPilote)))
-  mat_freq <- apply(mat_eff,2,function(v){v/sum(v)})
-  return(mat_freq)
-}
-
-# Built the plot-specific association matrix
-AlphaPlots<-function(Inv){
-  unlisted<-do.call(rbind,lapply(Inv,function(x){do.call(rbind,x)}))
-  ListedByPlot<-lapply(unique(unlisted[,"n_parcelle"]),function(y){
-    p<-unlisted[which(unlisted[,"n_parcelle"]==y),-which(colnames(unlisted)=="campagne")]
-    p<-p[which(!duplicated(p)),]
-    ap<-alpha_construct(p); return(ap)
-  })
-  names(ListedByPlot)<-unique(unlisted[,"n_parcelle"])
-  return(ListedByPlot)
-}
-
-# Listing all vernacular/genus and family associations 
-Correspondences<-function(Plot){
-  Corr<-Plot[order(Plot[,"Famille"],Plot[,"Genre"],Plot[,"Espece"]),]
-  Corr<-Corr[which(!duplicated(Corr)),]
-  return(Corr)
-}
-
-# Multinomial/dirichlet process, return a botanical name trialed in proba vector
-Dirichlet_draw<-function(V){
-  Vdir<-rdirichlet(1,V)
-  names(Vdir)<-names(V)
-  res <- rmultinom(1,1,Vdir)
-  res<-rownames(res)[which(res>0)]
-  return(res)
-}
-
-####### The function to randomly draw a botanical name replacing
-#the undertermined species. Use alpha matrix and botanic table
-tirage <- function(alpha,bota,name,eps=0.05){ 
-  
-  if(name["nomPilote"]=="-"){
-    trial<-sample(rownames(alpha),1)
-    while(grepl("Indet.",trial)){trial<-sample(rownames(alpha),1)}
-  }
-  
[…]

### 2018-04-19 Scripts/CWM.R
Graphslegend

@@ -17,7 +17,7 @@ T3<-c(4,8,12)
 treatments<-list(T0,T1,T2,T3)
 names(treatments)<-c("Control","T1","T2","T3")
 
-TraitsName<-c("Hmax","L_thickness","L_chloro","L_toughness","SLA","WD","Bark_thick")
+TraitsName<-c("L_thickness","L_chloro","L_toughness","SLA","WD","Bark_thick","Hmax")
 
 ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")#"chartreuse3",
 

### 2018-04-19 Scripts/DrawingTools.R

@@ -191,7 +191,7 @@ CWMdraw<-function(Cwm){
 invisible(lapply(colnames(Cwm[[1]]),function(trait){
   Toplot<-lapply(Cwm,function(pl){return(t(pl[,trait,]))})
   plot(colnames(Toplot[[1]]),Toplot[[1]]["0.5",], ylim=c(min(unlist(Toplot)),max(unlist(Toplot))),type="n",xlab="years",ylab="")
-  mtext(trait,3,cex=0.8,adj=0,line=0.5)
+  #mtext(trait,3,cex=0.8,adj=0,line=0.5)
   
   invisible(lapply(1:4,function(tr){
     toplot<-Toplot[which(names(Toplot)%in%treatments[[tr]])]
@@ -206,6 +206,25 @@ invisible(lapply(colnames(Cwm[[1]]),function(trait){
 mtext("Community Weighted Means",line=0.5,adj=0,outer=TRUE,cex=1.1)
 mtext("Years since disturbance",side=1,line=1.2,adj=0.65,cex=0.9,outer=TRUE)
 }
+
+legendCWM<-function(){
+  mtext("Leaf thickness\n",at=0.13,line=-2.5,outer=TRUE,cex=0.9)
+  mtext(expression(paste(mu, "m",sep = "")),at=0.08,line=-3,outer=TRUE,cex=0.9)
+  mtext("Leaf cholophyll content\n",at=0.4,line=-2.5,outer=TRUE,cex=0.9)
+  mtext(expression(paste("g.",mm^-2,sep = "")),at=0.34,line=-3,outer=TRUE,cex=0.9)
+  mtext("Leaf toughness\n",at=0.64,line=-2.5,outer=TRUE,cex=0.9)
+  mtext("N",at=0.56,line=-3,outer=TRUE,cex=0.9)
+  mtext("SLA\n",at=0.88,line=-2.5,outer=TRUE,cex=0.9)
+  mtext(expression(paste(mm^2,".",mg^-1,sep = "")),at=0.84,line=-3,outer=TRUE,cex=0.9)
+  
+  mtext("WD\n",at=0.13,line=-17.8,outer=TRUE,cex=0.9)
+  mtext(expression(paste("g.",cm^-3,sep = "")),at=0.08,line=-18,outer=TRUE,cex=0.9)
[…]

### 2018-04-19 WholePlotTrajectories.Rmd
Graphslegend
Webpages

@@
 output:
+  bookdown::gitbook: default
   bookdown::pdf_book:
@@
     keep_tex: TRUE
-  bookdown::gitbook: default
   bookdown::word_document2: default
 ---
+
+(ref:Bolda) **(a)**
+
+(ref:Boldb) **(b)**
+
+(ref:Boldc) **(c)**
+
+(ref:capThickness) *L_thickness*
+
+(ref:capChloro) *L_chloro*
+
+(ref:capThoughness) *L_toughness*
+
+(ref:capSLA) *SLA*
+
+(ref:capWD) *WD*
+
+(ref:capBarkThick) *Bark-thick*
+
+(ref:capHmax) *Hmax*
 
@@
 
-(ref:Bolda) **(a)**
-
-(ref:Boldb) **(b)**
-
-(ref:Boldc) **(c)**
-
 ```{r NMDSplans, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of (ref:Bolda) flora composition and (ref:Boldb) functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
@@
 
-(ref:capThickness) *L_thickness*
-
-(ref:capChloro) *L_chloro*
-
-(ref:capThoughness) *L_toughness*
-
-(ref:capSLA) *SLA*
-
-(ref:capWD) *WD*
-
-(ref:capBarkThick) *Bark-thick*
-
-(ref:capHmax) *Hmax*
-
 ```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWD), and bark thickness, (ref:capBarkThick)) and one life trait (Specific maximum height at adult stage, (ref:capHmax)).
@@
 CWMdraw(CWM)
+legendCWM()
 
@@
 
-While for undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity ($\rho_{spearman}^{Shannon}=0.87$, and $\rho_{spearman}^{Simpson}=0.88$) (Figure \@ref(fig:DivTaxo)).
-The Richness was less affected by the initial disturbance except for the higher intensity (T3 plots) (Figure \@ref(fig:DivTaxo)). 
-The Shannon and Simpson diversities however first significantly increased, illustrating an increasing evenness, before reaching a maximum around 15 years after disturbance for all plots except one from T1 treatment and one from T2 treatment (Figure \@ref(fig:DivTaxo)).
+For undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the initial values 5 years after disturbance.
+After disturbance, the richness increased after low disturbance intensity, with a maximum increase of 14 botanical genuses (for plot 3 after treatment 2), but followed a unimodal decrease with a return to initial values after intense disturbance.
+The taxonomic evenness (Shannon and Simpson diversities), however, significantly increased after all disturbance regime.
+The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum reached around 20 years and positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.92$, and $\rho_{spearman}^{Simpson}=0.97$).
+Only two T3 plots, plots 8 and 12, remained increasing 30 years after disturbance \@ref(fig:DivTaxo)), suggesting a similar be much delayed unimodal trajectory.
 
@@
 
-For all undisturbed plots the Rao diversity remained comparable to the values in 1989, the trajectories of all disturbed plots showed significant increase positively correlated with the disturbance intensity (Figure \@ref(fig:DivFun)).
-Although a first increase of plots functional diversity that was all the more important than initial disturbance was intense, 30 after disturbance the diversity was similar among disturbed plots and significantly higher than this of control plots.
+For all undisturbed plots the Rao diversity remained comparable to the initial values, 5 years after disturbance (1989 inventories).
+The trajectories of all disturbed plots followed a unimodal trajectory with a maximum positively correlated to disturbance intensity ($\rho_{spearman}$) (Figure \@ref(fig:DivFun)). 
+Thirty years after disturbance all plots, whenever the initial disturbance intensity, regained diversity values similar to their initial avule and to those of control plots.
 
@@
 # Discussion
-
-Our study showed the significant impact of disturbance on tropical forests communities.
-The subsequent diversity trajectories confirmed the intermediate disturbance hypothesis debated for tropical forests through their correlation with disturbance intensity.
-Besides it revealed the contrasting response of taxonomic and functional characteristics, specifically the decoupling between communities taxonomic evenness and their functional diversity and dominant functional traits values.
-The long-term disturbance trajectories observed highlighted the unachieved but consistent recovery of communities assembly for the lowest disturbance intensity but questioned it after higher disturbance.
 
@@
 
+# Conclusions
+
+Our study showed the significant impact of disturbance on tropical forests communities.
+The subsequent diversity trajectories confirmed the intermediate disturbance hypothesis debated for tropical forests through their correlation with disturbance intensity.
+Besides it revealed the contrasting response of taxonomic and functional characteristics, specifically the decoupling between communities taxonomic evenness and their functional diversity and dominant functional traits values.
+The long-term disturbance trajectories observed highlighted the unachieved but consistent recovery of communities assembly for the lowest disturbance intensity but questioned it after higher disturbance.
+
 `r if (knitr:::is_html_output()) '# References {-}'`

### 2018-04-19 WholePlotTrajectories.tex
Graphslegend
Webpages

@@
 
-While for undisturbed plots the Richness, Shannon and Simpson diversity
-remained comparable to the values in 1989, the trajectories of all
-disturbed plots showed significant increase positively correlated with
-the disturbance intensity (\(\rho_{spearman}^{Shannon}=0.87\), and
-\(\rho_{spearman}^{Simpson}=0.88\)) (Figure \ref{fig:DivTaxo}). The
-Richness was less affected by the initial disturbance except for the
-higher intensity (T3 plots) (Figure \ref{fig:DivTaxo}). The Shannon and
-Simpson diversities however first significantly increased, illustrating
-an increasing evenness, before reaching a maximum around 15 years after
-disturbance for all plots except one from T1 treatment and one from T2
-treatment (Figure \ref{fig:DivTaxo}).
+For undisturbed plots the Richness, Shannon and Simpson diversity
+remained comparable to the initial values 5 years after disturbance.
+After disturbance, the richness increased after low disturbance
+intensity, with a maximum increase of 14 botanical genuses (for plot 3
+after treatment 2), but followed a unimodal decrease with a return to
+initial values after intense disturbance. The taxonomic evenness
+(Shannon and Simpson diversities), however, significantly increased
+after all disturbance regime. The evenness followed a unimodal
+trajectory with a just beginning return towards initial values and a
+maximum reached around 20 years and positively correlated to the
+disturbance intensity (\(\rho_{spearman}^{Shannon}=0.92\), and
+\(\rho_{spearman}^{Simpson}=0.97\)). Only two T3 plots, plots 8 and 12,
+remained increasing 30 years after disturbance \ref{fig:DivTaxo}),
+suggesting a similar be much delayed unimodal trajectory.
 
@@
 For all undisturbed plots the Rao diversity remained comparable to the
-values in 1989, the trajectories of all disturbed plots showed
-significant increase positively correlated with the disturbance
-intensity (Figure \ref{fig:DivFun}). Although a first increase of plots
-functional diversity that was all the more important than initial
-disturbance was intense, 30 after disturbance the diversity was similar
-among disturbed plots and significantly higher than this of control
-plots.
+initial values, 5 years after disturbance (1989 inventories). The
+trajectories of all disturbed plots followed a unimodal trajectory with
+a maximum positively correlated to disturbance intensity
+(\(\rho_{spearman}\)) (Figure \ref{fig:DivFun}). Thirty years after
+disturbance all plots, whenever the initial disturbance intensity,
+regained diversity values similar to their initial avule and to those of
+control plots.
 
@@
 \section{Discussion}\label{discussion}
-
-Our study showed the significant impact of disturbance on tropical
-forests communities. The subsequent diversity trajectories confirmed the
-intermediate disturbance hypothesis debated for tropical forests through
-their correlation with disturbance intensity. Besides it revealed the
-contrasting response of taxonomic and functional characteristics,
-specifically the decoupling between communities taxonomic evenness and
-their functional diversity and dominant functional traits values. The
-long-term disturbance trajectories observed highlighted the unachieved
-but consistent recovery of communities assembly for the lowest
-disturbance intensity but questioned it after higher disturbance.
 
@@
 
+\section{Conclusions}\label{conclusions}
+
+Our study showed the significant impact of disturbance on tropical
+forests communities. The subsequent diversity trajectories confirmed the
+intermediate disturbance hypothesis debated for tropical forests through
+their correlation with disturbance intensity. Besides it revealed the
+contrasting response of taxonomic and functional characteristics,
+specifically the decoupling between communities taxonomic evenness and
+their functional diversity and dominant functional traits values. The
+long-term disturbance trajectories observed highlighted the unachieved
+but consistent recovery of communities assembly for the lowest
+disturbance intensity but questioned it after higher disturbance.
+
 %----------------------------------------------------------------------------------------

### 2018-04-19 Scripts/Stats.R

@@ -25,39 +25,33 @@ for(q in 1:3){
 colnames(Richness)<-c("plot","Max","treat")
 cor(Richness[,"Max"],Richness[,"treat"],method="spearman")
 colnames(Shannon)<-c("plot","Max","treat")
-cor(Shannon[,"Max"],Shannon[,"treat"],method="spearman")
+cor(Shannon[which(!Shannon$plot%in%c(8:12)),"Max"],Shannon[which(!Shannon$plot%in%c(8:12)),"treat"],method="spearman")
 colnames(Simpson)<-c("plot","Max","treat")
-cor(Simpson[,"Max"],Simpson[,"treat"],method="spearman")
+cor(Simpson[which(!Simpson$plot%in%c(8:12)),"Max"],Simpson[which(!Simpson$plot%in%c(8:12)),"treat"],method="spearman")
 
 
-##################################################"
+######
+# maximums of diversity indices
 
-Traits<-read.csv("DB/BridgeOK.csv",sep=";",na.strings="")
-Traits["name"]<-as.factor(paste(Traits$Genus,"_",Traits$species,sep=""))
-traits<-Traits1[,c("Family","Genus","species","name","bar_code","thickness","SPAD","toughness","dry_mass","traits_surf_area","ind_surf_area",
-                   "sapwood_dens","moisture","bark_thick")]
-colnames(traits)<-c("Family","Genus","species","name","bar_code","L_thickness","L_chloro","L_toughness","L_DryMass","LA","SLA","WD",
-                    "moisture","Bark_thick")
-##SLA calc for empty values with LA available
-traits[which(!is.na(traits[,"LA"])),"SLA"]<-
-  traits[which(!is.na(traits[,"LA"])),"LA"]/traits[which(!is.na(traits[,"LA"])),"L_DryMass"]
+load("DB/ReplacementTraj_ForGraphs")
+CompTaxo<-CompleteTaxo
 
-traitsName<-c("L_thickness","L_chloro","L_toughness","L_DryMass","LA","SLA","WD","moisture","Bark_thick")
+q<-3
+Toplot<-lapply(CompTaxo,function(tr){return(tr[,,,q])})
+Toplot<-lapply(Toplot,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
+Toplot<-lapply(Toplot,function(tr){
[…]

### 2018-04-19 WholePlotTrajectories.Rmd
Intro

@@
 
-In the short-term, diversity dynamics demonstrated negligible or even positive impacts of disturbance on communities diversity, and have been formalized by the intermediate disturbance hypothesis (IDH) that predicts a maximized species diversity at intermediate level of disturbance[@Molino2001;@Kariuki2006a;@Berry2008a].
-The IDH however has scarcely been tested in the long term and mainly relied on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-Communities composition and evenness are crucial to resolve conservation issues and assess the ecological rules shaping communities abundance distribution [@Magurran1988;@Lavorel2002;@Bellwood2006].
-A promosing perspective is also the functional approach accounting for species biological attributes related to individuals fitness and ecosystem properties [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-Major functional traits revealing species ecology and performance were largely adopted through a workable framework assessing forests dynamics and functioning [@Diaz2005;@Villeger2008a].
-This approach for example highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
-This was translated by the shift from "conservative" species with long-lived, dense tissues growing slowly but with scarce resources, to "acquisitive" fast-growing species with light tissues and efficient resources use [@TerSteege2001;@Reich2014;@Herault2011].
-Such shift would be detected with a combination of resource acquisition funcitonal traits (leaf area, density and chlorophyll content and stem specific gravity and bark thickness) and growth and reproduction life history traits (seed mass and maximum height) already proved relevant [@Wright2004;@Westoby2006a;@Chave2009b].
-A functional approach combined to the analysis of taxonomic diversity and composition is then a reliable way to validate the IDH in the long term and from communities evenness and functinal perspective.
-A full interpretation of communities response would be completed with communities taxonomic and functional composition that would elucidate the determinism of forest recovery and the maintenance of intrinsinc differences.
-Those development would be as much information to understand the mechanisms of forests response to diversity and help conservation and management strategies.
+In the short-term diversity dynamics demonstrated negligible or even positive impacts of disturbance on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
+However, validations of the IDH in the long term remain scarce and mainly rely on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+More releveant monitoring would encompass communities composition, that are crucial for conservation issues, and evenness, that revealing ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Furthermore, the functional approach accounting for species biological attributes would be insightful as it reveals species fitness and ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+In that respect major functional traits related to species ecology and performance were largely adopted through a relevant framework [@Diaz2005;@Villeger2008a].
+The functonial trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
+This translated by shifts from "conservative" slow-growing species dealing with scarce resources, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
+It was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
+A proper monitoring of communities response should therefore encompass taxonomic and functional diversity and composition measures to test the validity of the IDH in the long term for tropical forests, and clarify the resilience of communities evenness, composition,and also functioning.
+The trajectories followed by all these facets would highlight the role of deterministic processes, like competitive exclusion or abiotic selection, and the communities' convergence maintaining intrinsinc differences in diversity and composition,which is as much insgights for future adaptive conservation strategies [@Adler2007].
 
 Here we investigated over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
-We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw communities taxonomic and functional composition and diversity trajectories after disturbance.
-Specifically, we (i) tested the validity of the IDH arguing for enhanced diversity after moderate disturbance and (ii) confirmed the changes in ecological rules shaping communities, from competitive exclusion between "conservative", slow-growing tree species to environmental filtering of "acquisitive", fasting-growing ones.
-Eventually we examined the different trajectories to (iii) resolve the resilience of tropical forests and question their future dynamics regarding their state 30 years after disturbance.
+We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
+Specifically, we (i) tested the validity of the IDH in the long term for tropical hiperdiverse forest and highlighted the ecological rules shaping their response to disturbance, (ii) clarified the different facets of communities resilience in terms of communities composition, diversity and functioning (iii) questioned the completeness of communities recovery given the altered functional redundancy.
+>>revoir la fin
 

### 2018-04-19 WholePlotTrajectories.tex
Intro

@@
 
-In the short-term, diversity dynamics demonstrated negligible or even
-positive impacts of disturbance on communities diversity, and have been
-formalized by the intermediate disturbance hypothesis (IDH) that
-predicts a maximized species diversity at intermediate level of
-disturbance\citep{Molino2001, Kariuki2006a, Berry2008a}. The IDH however
-has scarcely been tested in the long term and mainly relied on the
-analysis of species richness that gives limited or misleading
-information on forests recovery and functioning
-\citep{Martin2015, Chaudhary2016}. Communities composition and evenness
-are crucial to resolve conservation issues and assess the ecological
-rules shaping communities abundance distribution
-\citep{Magurran1988, Lavorel2002, Bellwood2006}. A promosing perspective
-is also the functional approach accounting for species biological
-attributes related to individuals fitness and ecosystem properties
-\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. Major
-functional traits revealing species ecology and performance were largely
-adopted through a workable framework assessing forests dynamics and
-functioning \citep{Diaz2005, Villeger2008a}. This approach for example
-highlighted in tropical rainforests the environmental filters fostering
-disturbance resistant species with rapid growth and efficient resources
-acquisition \citep{Molino2001, Haddad2008}. This was translated by the
-shift from ``conservative'' species with long-lived, dense tissues
-growing slowly but with scarce resources, to ``acquisitive''
-fast-growing species with light tissues and efficient resources use
-\citep{TerSteege2001, Reich2014, Herault2011}. Such shift would be
-detected with a combination of resource acquisition funcitonal traits
-(leaf area, density and chlorophyll content and stem specific gravity
-and bark thickness) and growth and reproduction life history traits
-(seed mass and maximum height) already proved relevant
-\citep{Wright2004, Westoby2006a, Chave2009b}. A functional approach
-combined to the analysis of taxonomic diversity and composition is then
-a reliable way to validate the IDH in the long term and from communities
-evenness and functinal perspective. A full interpretation of communities
-response would be completed with communities taxonomic and functional
-composition that would elucidate the determinism of forest recovery and
-the maintenance of intrinsinc differences. Those development would be as
-much information to understand the mechanisms of forests response to
-diversity and help conservation and management strategies.
+In the short-term diversity dynamics demonstrated negligible or even
+positive impacts of disturbance on communities diversity, which have
+been formalized by the intermediate disturbance hypothesis (IDH) stating
+a maximized species diversity at intermediate disturbance intensity
+\citep{Molino2001, Kariuki2006a, Berry2008a}. However, validations of
+the IDH in the long term remain scarce and mainly rely on the analysis
+of species richness that gives limited or misleading information on
+forests recovery and functioning \citep{Martin2015, Chaudhary2016}. More
+releveant monitoring would encompass communities composition, that are
+crucial for conservation issues, and evenness, that revealing ecological
+rules underlying communities' structure
+\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, the
+functional approach accounting for species biological attributes would
+be insightful as it reveals species fitness and ecosystems functioning
+\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
+respect major functional traits related to species ecology and
+performance were largely adopted through a relevant framework
+\citep{Diaz2005, Villeger2008a}. The functonial trait-based approach,
+for example, highlighted in tropical rainforests the environmental
+filters fostering disturbance resistant species with rapid growth and
+efficient resources acquisition \citep{Molino2001, Haddad2008}. This
+translated by shifts from ``conservative'' slow-growing species dealing
+with scarce resources, to ``acquisitive'' fast-growing species with
+rapid and efficient use of abundant resources
+\citep{TerSteege2001, Reich2014, Herault2011}. It was mirrored by shifts
+in key functional traits related to resource acquisition (leaf area,
+density and chlorophyll content, and stem specific gravity and bark
+thickness), tree growth and reproduction life history traits (seed mass
+and maximum height) \citep{Wright2004, Westoby2006a, Chave2009b}. A
+proper monitoring of communities response should therefore encompass
+taxonomic and functional diversity and composition measures to test the
+validity of the IDH in the long term for tropical forests, and clarify
+the resilience of communities evenness, composition,and also
+functioning. The trajectories followed by all these facets would
+highlight the role of deterministic processes, like competitive
+exclusion or abiotic selection, and the communities' convergence
+maintaining intrinsinc differences in diversity and composition,which is
+as much insgights for future adaptive conservation strategies
+\citep{Adler2007}.
 
@@
 database browsing major leaf, stem and seed traits and species maximum
-height to draw communities taxonomic and functional composition and
-diversity trajectories after disturbance. Specifically, we (i) tested
-the validity of the IDH arguing for enhanced diversity after moderate
-disturbance and (ii) confirmed the changes in ecological rules shaping
-communities, from competitive exclusion between ``conservative'',
-slow-growing tree species to environmental filtering of ``acquisitive'',
-fasting-growing ones. Eventually we examined the different trajectories
-to (iii) resolve the resilience of tropical forests and question their
-future dynamics regarding their state 30 years after disturbance.
+height to draw the trajectories over time of communities taxonomic and
+functional composition and diversity. Specifically, we (i) tested the
+validity of the IDH in the long term for tropical hiperdiverse forest
+and highlighted the ecological rules shaping their response to
+disturbance, (ii) clarified the different facets of communities
+resilience in terms of communities composition, diversity and
+functioning (iii) questioned the completeness of communities recovery
+given the altered functional redundancy.
+\textgreater{}\textgreater{}revoir la fin
 

### 2018-04-20 AppendixI.Rmd
Grphs

@@
 
+# Appendix figures
+
 The following information are available for this article:
@@
 
-```{r FigBayesianScheme,fig.cap="Scheme of the Bayesian process propagating the taxonomic uncertainty"}
+**Figure A3**: Trajectories of the Rao functional diversity over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.
+\@ref(fig:DivFunAppendix).
+
+\newpage
+
+# Figure A1
+
+```{r FigBayesianScheme,out.width='100%',fig.cap="Scheme of the Bayesian process propagating the taxonomic uncertainty",echo=F, fig.env="figure*"}
 knitr::include_graphics("ExternalFig/BayesianSheme.png")
@@
 
-```{r FigSeedMassClasses, echo=FALSE, out.width='100%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+# Figure A2
+
+```{r FigSeedMassClasses, echo=FALSE, out.width='80%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
 No gap filling process was applied in this case.
@@
 ```
+
+# Figure A3
+
+```{r DivFunAppendix, out.width = '60%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.", fig.env="figure*"}
+load("DB/FunctionalTraj_ForGraphsnew")
+
+FunTraj(CompleteFun,remove=F)
+
+```

### 2018-04-20 AppendixI.tex
Grphs

@@
 
+\section{Appendix figures}\label{appendix-figures}
+
 The following information are available for this article:
@@
 level (i.e species of the same genus when the species is missing,
-species of the same family when the genus is missing).
+species of the same family when the genus is missing)
+\ref{fig:FigBayesianScheme}.
 
@@
 applied in this case. Initial treatments are represented by solid lines
-colors with green for control, blue for T1,orange for T2 and red for T3.
+colors with green for control, blue for T1,orange for T2 and red for T3
+\ref{fig:FigSeedMassClasses}.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{knitr}\OperatorTok{::}\KeywordTok{include_graphics}\NormalTok{(}\StringTok{"ExternalFig/BayesianSheme.png"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+\textbf{Figure A3}: Trajectories of the Rao functional diversity over 30
+years after disturbance. Trajectories correspond to the median (solid
+line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
+iteration of the taxonomic uncertainty propagation. Initial treatments
+are represented by solid lines colorswith green for control, blue for
+T1,orange for T2 and red for T3.and the missing trait value filling
+processes. \ref{fig:DivFunAppendix}.
 
-\begin{figure}
+\newpage
 
-{\centering \includegraphics{ExternalFig/BayesianSheme}
+\section{Figure A1}\label{figure-a1}
+
+\begin{figure*}
+
+{\centering \includegraphics[width=1\linewidth]{ExternalFig/BayesianSheme}
 
@@
 
-\caption{Scheme of the Bayesian process propagating the taxonomic uncertainty}\label{fig:FigA1}
-\end{figure}
+\caption{Scheme of the Bayesian process propagating the taxonomic uncertainty}\label{fig:FigBayesianScheme}
+\end{figure*}
+
+\section{Figure A2}\label{figure-a2}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{AppendixI_files/figure-latex/FigA2-1}
+{\centering \includegraphics[width=0.8\linewidth]{AppendixI_files/figure-latex/FigSeedMassClasses-1}
 
@@
 No gap filling process was applied in this case.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:FigA2}
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:FigSeedMassClasses}
+\end{figure*}
+
+\section{Figure A3}\label{figure-a3}
+
+\begin{figure*}
+
+{\centering \includegraphics[width=0.6\linewidth]{AppendixI_files/figure-latex/DivFunAppendix-1}
+
+}
+
+\caption{Trajectories of the Rao functional diversity over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:DivFunAppendix}
 \end{figure*}

### 2018-04-20 WholePlotTrajectories.Rmd
Grphs

@@
 The trajectories of all disturbed plots followed a unimodal trajectory with a maximum positively correlated to disturbance intensity ($\rho_{spearman}$) (Figure \@ref(fig:DivFun)). 
+The plot 7 from treatment 1 was removed from the graphical representation for better readability, see appendix for full graph.
 Thirty years after disturbance all plots, whenever the initial disturbance intensity, regained diversity values similar to their initial avule and to those of control plots.
@@
 Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes."}
-load("DB/FunctionalTraj_ForGraphs")
+load("DB/FunctionalTraj_ForGraphsnew")
 
-FunTraj(CompleteFun)
+FunTraj(CompleteFun,remove=T)
 
@@
 Only the most disturbed plots had not fully achieved the ongoing recovery of genus richness, although all had already reached equivalent levels as those of some control plots.
-Such long time richness recovery had already been observed in several logging experiments, which also highlighted the role of random disturbance damage in the richness decrease [@deAvila2015;@Hu2018].
+Such long time richness recovery had already been observed in several logging experiments, which also highlighted the role of random disturbance damage in the richness decrease [@DeAvila2015;@Hu2018].
 

### 2018-04-20 WholePlotTrajectories.tex
Grphs

@@
 a maximum positively correlated to disturbance intensity
-(\(\rho_{spearman}\)) (Figure \ref{fig:DivFun}). Thirty years after
-disturbance all plots, whenever the initial disturbance intensity,
-regained diversity values similar to their initial avule and to those of
-control plots.
+(\(\rho_{spearman}\)) (Figure \ref{fig:DivFun}). The plot 7 from
+treatment 1 was removed from the graphical representation for better
+readability, see appendix for full graph. Thirty years after disturbance
+all plots, whenever the initial disturbance intensity, regained
+diversity values similar to their initial avule and to those of control
+plots.
 
@@
 experiments, which also highlighted the role of random disturbance
-damage in the richness decrease \citep{deAvila2015, Hu2018}.
+damage in the richness decrease \citep{DeAvila2015, Hu2018}.
 

### 2018-04-20 Scripts/DivFunctional_trajectories.R

@@ -86,4 +86,5 @@ CompleteFun<-lapply(1:4,function(t){
                dimnames=list(treat,dates,1:Nrep)))
 })
 
-save(CompleteFun,file="FunctionalTraj_ForGraphs")
+save(CompleteFun,file="DB/FunctionalTraj_ForGraphs")
+

### 2018-04-20 Scripts/DrawingTools.R

@@ -5,6 +5,10 @@ treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
 names(treatments)<-c("Control","T1","T2","T3")
 ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
 
+smooth<-function(mat,larg){return(do.call(cbind,lapply(1:ncol(mat),function(step){
+  range<-max(1,step-larg):min(ncol(mat),step+larg)
+  rowSums(mat[,range])/length(range)})))}
+
 TaxoCompo<-function(Data_TaxoComp){
 par(mfrow=c(1,2))
 matT<-as.data.frame(Data_TaxoComp[[1]])
@@ -167,7 +171,17 @@ TaxoTraj<-function(CompTaxo){
 mtext("Years since disturbance",side=1,adj=1,cex=0.8,line=-2,outer=TRUE)
 mtext("Equivalent diversity",side=2,padj=1,cex=0.8,line=1.5,outer=TRUE)}
 
-FunTraj<-function(CompFun){
+FunTraj<-function(CompFun,remove=TRUE){
+  if(remove){CompFun[[2]]<-CompFun[[2]][which(rownames(CompFun[[2]])!=7),,]}
+  
+  CompFun<-lapply(CompFun,function(tr){
+    ret<-tr[,which(colnames(tr)>=1989),]
+    #ret<-lapply(1:dim(ret)[3],function(rep){return(apply(ret[,,rep],2,function(col){col<-col-ret[,1,rep]}))})
+    #ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),dimnames=list(rownames(ret[[1]]),colnames(ret[[1]]),1:length(ret)))
+    ret<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(ret,c(1,2),function(x){return(quantile(x,probs=quant))}))})
+    return(array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),3),
+                 dimnames=list(rownames(ret[[1]]),as.numeric(colnames(ret[[1]]))-1984,c(0.025,0.5,0.975))))})
+  
   plot(as.numeric(colnames(CompFun[[1]])),CompFun[[1]][1,,"0.5"],type="n",xaxt="n",
[…]

### 2018-04-20 Scripts/Stats.R

@@ -52,6 +52,68 @@ MaxRich<-apply(do.call(rbind,lapply(Toplot,function(tr){return(tr[,,"0.5"])})),1
 MaxRich[order(MaxRich,decreasing = T)][1:2]
 
 
+######################r
+### The plot 7 funcitonal diversity
+
+tab<-do.call(rbind,lapply(1:12,function(pl){
+  P<-lapply(LivingStand_all,function(yr){if(any(names(yr)==pl)){return(yr[[which(names(yr)==pl)]])}})
+P<-P[which(!names(P)%in%Pby)]
+#lapply(P,nrow)
+
+Rao<-do.call(rbind,lapply(P,function(yr){
+  
+  yr<-Replacement(yr,Alpha=alphas_plot[[which(names(alphas_plot)==pl)]])
+      
+      #traits_filled<-Traits_filling(Traits1,Traits2,InventorySp)
+      #Traits_filled<-aggregate(Traits_filled[,TraitsName],list(Traits_filled$name),median)
+      #rownames(Traits_filled)<-Traits_filled[,1];Traits_filled<-Traits_filled[,TraitsName]
+      tra<-Traits_filled[which(rownames(Traits_filled)%in%yr),];tra<-tra[order(rownames(tra)),]
+      
+      return(apply(tra,2,sd))
+      
+      yr<-yr[which(yr%in%rownames(tra))]
+      #return(length(which(!yr%in%rownames(tra))))
+      
+      #dissim<-as.matrix(daisy(tra,metric="gower"))
+      #dissim <- 1 - dissim/max(dissim)
+      
+      #return(expq(Hqz(as.AbdVector(tapply(yr,yr,length)), q=2, Z=dissim,Correction="None"),q=2))
+    }))
+Rao<-apply(Rao,2,mean)
+}))
+rownames(tab)<-1:12
+
+PCA<-dudi.pca(tab,scale=T,scan=F)
+
+windows()
+plot(PCA$li[,"Axis1"],PCA$li[,"Axis2"],type="n")
+text(PCA$li[,"Axis1"],PCA$li[,"Axis2"],labels=rownames(PCA$li))
 
 
+P<-lapply(LivingStand_all,function(yr){if(any(names(yr)==pl)){return(yr[[which(names(yr)==pl)]])}})
+P<-P[which(!names(P)%in%Pby)]
+#lapply(P,nrow)
+
+pl<-7
[…]

### 2018-04-20 WholePlotTrajectories.Rmd
Graphs

@@
 output:
-  bookdown::gitbook: default
   bookdown::pdf_book:
@@
     keep_tex: TRUE
+  bookdown::gitbook: default
   bookdown::word_document2: default
@@
 
-```{r NMDSdivergence, out.width = '100%', fig.asp = 1.2/2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of v taxonomic composition at genus level and (ref:Boldb) functional composition.
+```{r NMDSdivergence, out.width = '80%', fig.asp = 1.2/2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of v taxonomic composition at genus level and (ref:Boldb) functional composition.
 Distance are abundance-based Bary-Curtis metric.
 Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.
-"}
+The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.", fig.env="figure*"}
 
@@
 
-```{r DivFun, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
+```{r DivFun, out.width = '80%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes."}
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.", fig.env="figure*"}
 load("DB/FunctionalTraj_ForGraphsnew")

### 2018-04-20 WholePlotTrajectories.tex
Graphs

@@
 
-\begin{figure}
+\begin{figure*}
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/NMDSdivergence-1}
+{\centering \includegraphics[width=0.8\linewidth]{WholePlotTrajectories_files/figure-latex/NMDSdivergence-1}
 
@@
 Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.
-}\label{fig:NMDSdivergence}
-\end{figure}
+The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.}\label{fig:NMDSdivergence}
+\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure*}
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivFun-1}
+{\centering \includegraphics[width=0.8\linewidth]{WholePlotTrajectories_files/figure-latex/DivFun-1}
 
@@
 Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:DivFun}
-\end{figure}
+\end{figure*}
 

### 2018-04-20 AppendixI.Rmd
results

@@
 
-**Figure A2**: Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+**Figure A2**: Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of v taxonomic composition at genus level and (ref:Boldb) functional composition.
+Distance are abundance-based Bary-Curtis metric.
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
+The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.
+
+**Figure A3**: Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
 No gap filling process was applied in this case.
@@
 
-**Figure A3**: Trajectories of the Rao functional diversity over 30 years after disturbance.
+**Figure A4**: Trajectories of the Rao functional diversity over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 
-# Figure A1
+# **Figure A1**: Bayesian process sheme
 
@@
 
-# Figure A2
+# **Figure A2**: Dissimilarity to initial inventory of plots taxonomic and functional composition
+
+```{r NMDSdivergence, out.width = '80%', fig.asp = 1.2/2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of v taxonomic composition at genus level and (ref:Boldb) functional composition.
+Distance are abundance-based Bary-Curtis metric.
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
+The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.", fig.env="figure*"}
+load("DB/FunctionalComposition_forGraph")
+load("DB/TaxoComposition_forGraphs")
+
+TaxoDist(MatrepTaxo)
+FunDist(MatrepFun)
+
+```
+
+# **Figure A3**: Proportion of seed mass classes along time
 
@@
 
-# Figure A3
+# **Figure A4**: Rao diversity trajectories with all plots display
 

### 2018-04-20 AppendixI.tex
results

@@
 
-\textbf{Figure A2}: Trajectories of seed mass classes proportions over
+\textbf{Figure A2}: Trajectories of the distance to initial condition of
+the 30 sampled years in the two-dimensional space from the NMDS of v
+taxonomic composition at genus level and (ref:Boldb) functional
+composition. Distance are abundance-based Bary-Curtis metric. Line
+colors represent the disturbance treatment (green for control, blue for
+T1,orange for T2 and red for T3). The 0.025 and 0.975 percentile
+correspond to the variance observed for 50 iteration of the taxonomic
+uncertainty propagation and functional trait filling processes.
+
+\textbf{Figure A3}: Trajectories of seed mass classes proportions over
 30 years after disturbance, corresponding to the median (solid line) and
@@
 
-\textbf{Figure A3}: Trajectories of the Rao functional diversity over 30
+\textbf{Figure A4}: Trajectories of the Rao functional diversity over 30
 years after disturbance. Trajectories correspond to the median (solid
@@
 
-\section{Figure A1}\label{figure-a1}
+\section{\texorpdfstring{\textbf{Figure A1}: Bayesian process
+sheme}{Figure A1: Bayesian process sheme}}\label{figure-a1-bayesian-process-sheme}
 
@@
 
-\section{Figure A2}\label{figure-a2}
+\section{\texorpdfstring{\textbf{Figure A2}: Dissimilarity to initial
+inventory of plots taxonomic and functional
+composition}{Figure A2: Dissimilarity to initial inventory of plots taxonomic and functional composition}}\label{figure-a2-dissimilarity-to-initial-inventory-of-plots-taxonomic-and-functional-composition}
+
+\begin{figure*}
+
+{\centering \includegraphics[width=0.8\linewidth]{AppendixI_files/figure-latex/NMDSdivergence-1}
+
+}
+
+\caption{Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of v taxonomic composition at genus level and (ref:Boldb) functional composition.
+Distance are abundance-based Bary-Curtis metric.
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
+The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.}\label{fig:NMDSdivergence}
+\end{figure*}
+
+\section{\texorpdfstring{\textbf{Figure A3}: Proportion of seed mass
+classes along
+time}{Figure A3: Proportion of seed mass classes along time}}\label{figure-a3-proportion-of-seed-mass-classes-along-time}
 
@@
 
-\section{Figure A3}\label{figure-a3}
+\section{\texorpdfstring{\textbf{Figure A4}: Rao diversity trajectories
+with all plots
+display}{Figure A4: Rao diversity trajectories with all plots display}}\label{figure-a4-rao-diversity-trajectories-with-all-plots-display}
 

### 2018-04-20 WholePlotTrajectories.tex
results

@@
 composition and diversity were analysed at the genus level, \emph{i.e.}
-referring to the genus of observed or trialed botanical names. The
-analysis of taxonomic and functional composition of plots along time was
-visualized in a two-dimensional ordination space based on non-metric
-dimensional scaling of succesive floristic or functional inventories.
-Plots trajectories along time was reported comparatively to the
-inventories in 1989, 5 years after disturbance, which corresponded to
-first inventory with a sufficient degree of uncertainty (\textless{}30\%
-of undetermined trees). The inventories dissimilarity compared to the
-reference 1989 inventory was reported using occurrence-based (Jaccard)
-and abundance-based (Bray-Curtis) similarity measures. The trajectory of
+referring to the genus of observed or trialed botanical names.
+Variations in plots taxonomic and functional composition after
+disturbance was visualized by their trajectories in a two-dimensional
+ordination space over 30 years. Two NMDS were conducted, either from
+taxonomic flora inventories or from plots functional composition based
+on the 8 leaf, stem and life history traits. Plots trajectories along
+time was reported comparatively to the inventories in 1989, 5 years
+after disturbance, which corresponded to first inventory with a
+sufficient degree of uncertainty (\textless{}30\% of undetermined
+trees). The inventories dissimilarity compared to the reference 1989
+inventory was reported using occurrence-based (Jaccard) and
+abundance-based (Bray-Curtis) similarity measures. The trajectory of
 inventories along time was visualized with the euclidean distance in the
-two-dimensional ordination space to the 1989 inventory. The functional
-trajectories of the leaf and stem and life traits were also visualized
-with the community weighted means (CWM), representing the average trait
-value in a community weighted by relative abundance of the species
-carrying each value \citep{Diaz2007, Garnier2004}. Species seed mass
-corresponded to 5 classes of increasing mass, seed mass trajectories
-were therefore reported as the proportion of each class in the
-inventories.
+two-dimensional ordination space to the 1989 inventory. The trajectories
+of the leaf and stem and life traits were also visualized with the
+community weighted means (CWM), representing the average trait value in
+a community weighted by relative abundance of the species carrying each
+value \citep{Diaz2007, Garnier2004}. To compensate the intrinsinc
+difference among plots the trajectories corresponded to the differences
+along time with the reference inventory in 1984 (5 year after
+disturbance). Species seed mass corresponded to 5 classes of increasing
+mass, seed mass trajectories were therefore reported as the proportion
+of each class in the inventories.
 
@@
 Over time, 828388 trees and 591 botanical species pertaining to 223
-genus and 64 botanical families were recorded. Trajectories of taxonomic
-and functional composition after disturbance were examined through the
-ordination of successive inventories from 1989 (5 years after
-disturbance) in the two-dimensional space from NMDS analysis based on
-flora inventories and the corresponding functional composition.
+genus and 64 botanical families were recorded. Taxonomic and functional
+plots trajectories were examined through the ordination in a two
+dimensional ordiantion space of successive inventories from 1989 (5
+years after disturbance) to 2015 (31 years after disturbance).
 Classifications were performed using either abundance-based Bray-Curtis
-(Figure \ref{fig:NMDSplans}) or incidence-based Jaccard dissimilarity
-(data not shown).
+(Figure \ref{fig:NMDSplans}) or incidence-based Jaccard dissimilarity,
+this last having given similar results (data not shown).
 
-Both taxonomic and functional composition were substantially affected by
-disturbance, and the maximum dissimilarity to 1984 reference inventory
-composition was positively correlated with the disturbance intensity.
-Until 30 years after disturbance, all disturbed plots taxonomic
-composition remained significantly dissimilar to that of the 1989
-reference inventory composition (Figure \ref{fig:NMDSplans}). All plots,
-though, displayed a unimodal trajectory with a the return toward the
-taxonomic and functional composition of the 1989 reference inventory
-suggesting a shift towards a cyclic regime (Figure
-\ref{fig:NMDSdivergence}). For taxonomic composition, the maximum
-dissimilarity to 1984 inventory composition was reach around 20 years
-after disturbance for the T2 and T3 plots and for one T1 plot while
-control and other T1 plots continued to increase (Figure
-\ref{fig:NMDSdivergence}.a). For functional composition, the maximum
-dissimilarity to 1984 inventory composition was reach from 15 to 20
-years after disturbance for the T2 and T3 plots while T1 and control
-plots continued to increase (Figure \ref{fig:NMDSdivergence}.b). The
-functional composition distancing from the reference inventory was also
-positively correlated with disturbance intensity and similarly
-stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after
-disturbance (Figure \ref{fig:NMDSdivergence}).
-
-The coordinates of the functional traits were measured in the
-two-dimensional ordination space that mapped plots evolution along time
-(graphs not showed) revealed a trajectory of disturbed plots toward
-acquisitive functional strategies (from high WD to high SLA and
-chlorophyll content).
+Both taxonomic and functional composition of disturbed plots followed
+consistent trajectories over time, translating compositional changes
+while control plots remained stable. The addition of functional traits
+in the two dimensional space revealed a shift of plots composition
+towards more acquisitive functional strategies (disturbed plots changed
+from high average WD to high average SLA and chlorophyll content). The
+dissimilarity of the successive inventories for disturbed plots compared
+to their respective reference inventory in 1989 followed unimodal
+trajectories (AppendixI, Figure A2). The maximum dissimilarity of plot
+strajectories was positively correlated to the disturbance intensity for
+taxonomic (\(\rho_{spearman}^{taxonomic}=0.91\)) as well as functional
+composition (\(\rho_{spearman}^{taxonomic}=0.96\)). The time at maximum
+dissimilarity for taxonomic composition was reached around 26 years
+after disturbance while it was 22 years for functional composition. All
+trajectories returned towards the initial composition and translated
+cyclic compositional changes as visualized in figure (Figure
+\ref{fig:NMDSplans}).
 
@@
 
-\begin{figure*}
-
-{\centering \includegraphics[width=0.8\linewidth]{WholePlotTrajectories_files/figure-latex/NMDSdivergence-1}
-
-}
-
-\caption{Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of v taxonomic composition at genus level and \textbf{(b)} functional composition.
-Distance are abundance-based Bary-Curtis metric.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.}\label{fig:NMDSdivergence}
-\end{figure*}
-
 \subsubsection{Traits community weighted
@@
 
-For all plots the trajectories of the community weighted means (CWM)
-were drawn for the 8 functional and life history traits (Leaf thickness,
-chlorophyll content, toughness and specific area,wood specific gravity
-and barck thickness and seed mass and maximum adult height) (Figure
+The changes observed in plots functional composition went hand to hand
+with consistent trajectories of the 8 functional and life history traits
+(Leaf thickness, chlorophyll content, toughness and specific area,wood
+specific gravity and barck thickness and seed mass and maximum adult
+height) visaualized with plots community weighted means (CWM) (Figure
 \ref{fig:CWM}).
-
-To compensate the intrinsinc difference among plots the trajectories
-drawn correspond to the difference in value between the reference
-inventory in 1984 (5 year after disturbance) and the successive years
-inventoried.
 
@@
 maximum reached around 20 years and positively correlated to the
-disturbance intensity (\(\rho_{spearman}^{Shannon}=0.92\), and
-\(\rho_{spearman}^{Simpson}=0.97\)). Only two T3 plots, plots 8 and 12,
+disturbance intensity (\(\rho_{spearman}^{Shannon}=0.86\), and
+\(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots, plots 8 and 12,
 remained increasing 30 years after disturbance \ref{fig:DivTaxo}),

### 2018-04-20 Scripts/DrawingTools.R

@@ -214,7 +214,7 @@ invisible(lapply(colnames(Cwm[[1]]),function(trait){
     invisible(lapply(toplot,function(plo){
       lines(colnames(plo),plo["0.5",],col=ColorsTr[tr],lwd=1.5)
       polygon(c(colnames(plo),rev(colnames(plo))),c(plo["0.025",],rev(plo["0.975",])),
-              col=rgb(0,0,0,alpha=0.1),border=NA)
+              col=rgb(0,0,0,alpha=0.05),border=NA)
       
     }))
   }))

### 2018-04-20 Scripts/Stats.R

@@ -1,3 +1,47 @@
+load("DB/TaxoComposition_ForGraphs")
+
+Data_TaxoComp<-MatrepTaxo
+
+MatrepT<-lapply(Data_TaxoComp,function(Rep){
+  Rep<-as.data.frame(Rep)
+  Rep$plot<-substr(rownames(Rep),start=1,stop=regexpr("_",rownames(Rep))-1)
+  Rep$year<-substr(rownames(Rep),start=regexpr("_",rownames(Rep))+1,
+                   stop=nchar(rownames(Rep)))
+  return(Rep)
+})
+
+DistT<-as.data.frame(unlist(lapply(1:12,function(pl){
+  ret<-lapply(MatrepT,function(rep){return(rep[which(rep[,"plot"]==pl),])})
+  ret<-lapply(ret,function(rep){return(rep[which(rep[,"year"]%in%
+                              ret[[which(unlist(lapply(ret,nrow))==min(unlist(lapply(ret,nrow))))[1]]][,"year"]),])})
+  ret<-do.call(rbind,lapply(ret,function(rep){
+    ret2<-apply(rep[,c("NMDS1","NMDS2")],1,function(li){
+      return(sqrt(sum((rep[1,c("NMDS1","NMDS2")]-li)^2)))})
+    names(ret2)<-rep[,"year"]
+    return(ret2)}))
+  ret<-apply(ret,2,median)
+  return(names(ret)[which(ret==max(ret))])})))
+treats<-cbind(c(1,6,11,2,7,9,3,5,10,4,8,12),rep(0:3,each=3))
+DistT<-cbind(DistT,rownames(DistT),treats[order(treats[,1]),2])
+colnames(DistT)<-c("Max","Plot","treat")
+cor(DistT[,"Max"],DistT[,"treat"],method="spearman")
+
+TimeMax<-unlist(lapply(1:12,function(pl){
+  ret<-lapply(MatrepT,function(rep){return(rep[which(rep[,"plot"]==pl),])})
+  ret<-lapply(ret,function(rep){return(rep[which(rep[,"year"]%in%
+                                                   ret[[which(unlist(lapply(ret,nrow))==min(unlist(lapply(ret,nrow))))[1]]][,"year"]),])})
+  ret<-do.call(rbind,lapply(ret,function(rep){
+    ret2<-apply(rep[,c("NMDS1","NMDS2")],1,function(li){
[…]

### 2018-04-20 WholePlotTrajectories.Rmd
html

@@
     keep_tex: TRUE
-  bookdown::gitbook: default
   bookdown::word_document2: default
@@
 To counter the remaining taxonomic uncertainty plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
-The analysis of taxonomic and functional composition of plots along time was visualized in a two-dimensional ordination space based on non-metric dimensional scaling of succesive floristic or functional inventories.
+Variations in plots taxonomic and functional composition after disturbance was visualized by their trajectories in a two-dimensional ordination space over 30 years.
+Two NMDS were conducted, either from taxonomic flora inventories or from plots functional composition based on the 8 leaf, stem and life history traits. 
 Plots trajectories along time was reported comparatively to the inventories in 1989, 5 years after disturbance, which corresponded to first inventory with a sufficient degree of uncertainty (<30% of undetermined trees).
@@
 The trajectory of inventories along time was visualized with the euclidean distance in the two-dimensional ordination space to the 1989 inventory.
-The functional trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004].
+The trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004].
+To compensate the intrinsinc difference among plots the trajectories corresponded to the differences along time with the reference inventory in 1984 (5 year after disturbance).
 Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were  therefore reported as the proportion of each class in the inventories.
@@
 Over time, 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
-Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in the two-dimensional space from NMDS analysis based on flora inventories and the corresponding functional composition.
-Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity (data not shown).
+Taxonomic and functional plots trajectories were examined through the ordination in a two dimensional ordiantion space of successive inventories from 1989 (5 years after disturbance) to 2015 (31 years after disturbance).
+Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity, this last having given similar results (data not shown).
 
-Both taxonomic and functional composition were substantially affected by disturbance, and the maximum dissimilarity to 1984 reference inventory composition was positively correlated with the disturbance intensity.
-Until 30 years after disturbance, all disturbed plots taxonomic composition remained significantly dissimilar to that of the 1989 reference inventory composition (Figure \@ref(fig:NMDSplans)).
-All plots, though, displayed a unimodal trajectory with a the return toward the taxonomic and functional composition of the 1989 reference inventory suggesting a shift towards a cyclic regime (Figure \@ref(fig:NMDSdivergence)).
-For taxonomic composition, the maximum dissimilarity to 1984 inventory composition was reach around 20 years after disturbance for the T2 and T3 plots and for one T1 plot while control and other T1 plots continued to increase (Figure \@ref(fig:NMDSdivergence).a).
-For functional composition, the maximum dissimilarity to 1984 inventory composition was reach from 15 to 20 years after disturbance for the T2 and T3 plots while T1 and control plots continued to increase (Figure \@ref(fig:NMDSdivergence).b).
-The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance (Figure \@ref(fig:NMDSdivergence)).
-
-The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) revealed a trajectory of disturbed plots toward acquisitive functional strategies (from high WD to high SLA and chlorophyll content).
+Both taxonomic and functional composition of disturbed plots followed consistent trajectories over time, translating compositional changes while control plots remained stable.
+The addition of functional traits in the two dimensional space revealed a shift of plots composition towards more acquisitive functional strategies (disturbed plots changed from high average WD to high average SLA and chlorophyll content).
+The dissimilarity of the successive inventories for disturbed plots compared to their respective reference inventory in 1989 followed unimodal trajectories (AppendixI, Figure A2).
+The maximum dissimilarity of plot strajectories was positively correlated to the disturbance intensity for taxonomic ($\rho_{spearman}^{taxonomic}=0.91$) as well as functional composition ($\rho_{spearman}^{taxonomic}=0.96$).
+The time at maximum dissimilarity for taxonomic composition was reached around 26 years after disturbance while it was 22 years for functional composition. 
+All trajectories returned towards the initial composition and translated cyclic compositional changes as visualized in figure (Figure \@ref(fig:NMDSplans)).
 
@@
 
-```{r NMDSdivergence, out.width = '80%', fig.asp = 1.2/2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of v taxonomic composition at genus level and (ref:Boldb) functional composition.
-Distance are abundance-based Bary-Curtis metric.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.", fig.env="figure*"}
-
-TaxoDist(MatrepTaxo)
-FunDist(MatrepFun)
-
-```
-
 ### Traits community weighted means
 
-For all plots the trajectories of the community weighted means (CWM) were drawn for the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area,wood specific gravity and barck thickness and seed mass and maximum adult height) (Figure \@ref(fig:CWM)).
-
-To compensate the intrinsinc difference among plots the trajectories drawn correspond to the difference in value between the reference inventory in 1984 (5 year after disturbance) and the successive years inventoried.
+The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area,wood specific gravity and barck thickness and seed mass and maximum adult height) visaualized with plots community weighted means (CWM) (Figure \@ref(fig:CWM)).
 
@@
 The taxonomic evenness (Shannon and Simpson diversities), however, significantly increased after all disturbance regime.
-The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum reached around 20 years and positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.92$, and $\rho_{spearman}^{Simpson}=0.97$).
+The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum reached around 20 years and positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
 Only two T3 plots, plots 8 and 12, remained increasing 30 years after disturbance \@ref(fig:DivTaxo)), suggesting a similar be much delayed unimodal trajectory.

### 2018-04-20 WholePlotTrajectories.Rmd
results

@@
 
-## Taxonomic and functional composition
+## Disturbance impacts on communities composition and average functional traits values
 
@@
 
-Except for leaf cholorphyll content, which displayed higher difference between plots than among treatments which may be due to the completeness of dataset, all CWM trajectories corresponded to significant changes after disturbance.
-All functional traits and seed mass proportions displayed a unimodal trajectories but with different times at maximum and different values 30 years after disturbance. 
+Except for leaf cholorphyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions displayed a unimodal trajectories.
 The weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value and than these of the control plots (Figure \@ref(fig:CWM)).
 The weighted means of bark thickness (*Bark_thick*) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value and this of the undisturbed plots at the end of the experiment.
+Once again the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_thickness}=0.67$,$\rho_{spearman}^{L_chloro}=0.45$,$\rho_{spearman}^{L_toughness}=-0.43$,$\rho_{spearman}^{SLA}=0.93$,$\rho_{spearman}^{WD}=-0.78$,$\rho_{spearman}^{Bark-thickness}=0.88$,$\rho_{spearman}^{Hmax}=-0.48$).
 
@@
 
-## Disturbance impact on diversity
+## Trajectories of communities diversity after disturbance
 
@@
 
-For undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the initial values 5 years after disturbance.
-After disturbance, the richness increased after low disturbance intensity, with a maximum increase of 14 botanical genuses (for plot 3 after treatment 2), but followed a unimodal decrease with a return to initial values after intense disturbance.
-The taxonomic evenness (Shannon and Simpson diversities), however, significantly increased after all disturbance regime.
+For undisturbed plots the Richness, Shannon and Simpson diversity remained stable over the 30 monitored years.
+After disturbance the richness increased when the disturbance intensity was low, with a maximum increase of 14 botanical genuses (plot 3 from treatment 2).
+When the disturbance was intense however, plot staxonomic richness followed a unimodal decrease with a return to initial values after intense disturbance.
+The taxonomic evenness (Shannon and Simpson diversities) significantly increased whenever the disturbance intensity.
 The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum reached around 20 years and positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
@@
 
-For all undisturbed plots the Rao diversity remained comparable to the initial values, 5 years after disturbance (1989 inventories).
-The trajectories of all disturbed plots followed a unimodal trajectory with a maximum positively correlated to disturbance intensity ($\rho_{spearman}$) (Figure \@ref(fig:DivFun)). 
-The plot 7 from treatment 1 was removed from the graphical representation for better readability, see appendix for full graph.
+The functional diversity trajectories of plots based on the 8 leaf, stem and life history traits was examined through the Rao diversity.
+For all undisturbed plots the functional diversity remained comparable to the reference values 5 years after disturbance (1989 inventories).
+For all disturbed plots the functional trajectories followed a unimodal trajectory with a maximum positively correlated to disturbance intensity ($\rho_{spearman}=XX$) (Figure \@ref(fig:DivFun)). 
+The plot 7 from treatment 1 that displayed a constantly outlying diversity was removed from the graphical representation for better readability, see appendix for full graph.
 Thirty years after disturbance all plots, whenever the initial disturbance intensity, regained diversity values similar to their initial avule and to those of control plots.
@@
 
-FunTraj(CompleteFun,remove=T)
+FunTraj(CompleteFun,remove=F)
 

### 2018-04-20 WholePlotTrajectories.tex
results

@@
 
-\subsection{Taxonomic and functional
-composition}\label{taxonomic-and-functional-composition}
+\subsection{Disturbance impacts on communities composition and average
+functional traits
+values}\label{disturbance-impacts-on-communities-composition-and-average-functional-traits-values}
 
@@
 
-Except for leaf cholorphyll content, which displayed higher difference
-between plots than among treatments which may be due to the completeness
-of dataset, all CWM trajectories corresponded to significant changes
-after disturbance. All functional traits and seed mass proportions
-displayed a unimodal trajectories but with different times at maximum
-and different values 30 years after disturbance. The weighted means of
-communities specific maximum height at adult stage (\emph{Hmax}), leaf
-toughness (\emph{L\_toughness}) and wood specific gravity (\emph{WD})
-remained significantly lower than their initial value and than these of
-the control plots (Figure \ref{fig:CWM}). The weighted means of bark
-thickness (\emph{Bark\_thick}) similarly remained substantilly higher
-than initially for all disturbed plots while the specfic leaf area
-(\emph{SLA}) had almost recovered its initial value and this of the
-undisturbed plots at the end of the experiment.
+Except for leaf cholorphyll content, which continued to increase for
+some T3 and T2 plots 30 years after disturbance, all traits and seed
+mass proportions displayed a unimodal trajectories. The weighted means
+of communities specific maximum height at adult stage (\emph{Hmax}),
+leaf toughness (\emph{L\_toughness}) and wood specific gravity
+(\emph{WD}) remained significantly lower than their initial value and
+than these of the control plots (Figure \ref{fig:CWM}). The weighted
+means of bark thickness (\emph{Bark\_thick}) similarly remained
+substantilly higher than initially for all disturbed plots while the
+specfic leaf area (\emph{SLA}) had almost recovered its initial value
+and this of the undisturbed plots at the end of the experiment. Once
+again the maximum difference to initial state was correlated to the
+disturbance intensity
+(\(\rho_{spearman}^{L_thickness}=0.67\),\(\rho_{spearman}^{L_chloro}=0.45\),\(\rho_{spearman}^{L_toughness}=-0.43\),\(\rho_{spearman}^{SLA}=0.93\),\(\rho_{spearman}^{WD}=-0.78\),\(\rho_{spearman}^{Bark-thickness}=0.88\),\(\rho_{spearman}^{Hmax}=-0.48\)).
 
@@
 
-\subsection{Disturbance impact on
-diversity}\label{disturbance-impact-on-diversity}
+\subsection{Trajectories of communities diversity after
+disturbance}\label{trajectories-of-communities-diversity-after-disturbance}
 
@@
 For undisturbed plots the Richness, Shannon and Simpson diversity
-remained comparable to the initial values 5 years after disturbance.
-After disturbance, the richness increased after low disturbance
-intensity, with a maximum increase of 14 botanical genuses (for plot 3
-after treatment 2), but followed a unimodal decrease with a return to
-initial values after intense disturbance. The taxonomic evenness
-(Shannon and Simpson diversities), however, significantly increased
-after all disturbance regime. The evenness followed a unimodal
-trajectory with a just beginning return towards initial values and a
-maximum reached around 20 years and positively correlated to the
-disturbance intensity (\(\rho_{spearman}^{Shannon}=0.86\), and
+remained stable over the 30 monitored years. After disturbance the
+richness increased when the disturbance intensity was low, with a
+maximum increase of 14 botanical genuses (plot 3 from treatment 2). When
+the disturbance was intense however, plot staxonomic richness followed a
+unimodal decrease with a return to initial values after intense
+disturbance. The taxonomic evenness (Shannon and Simpson diversities)
+significantly increased whenever the disturbance intensity. The evenness
+followed a unimodal trajectory with a just beginning return towards
+initial values and a maximum reached around 20 years and positively
+correlated to the disturbance intensity
+(\(\rho_{spearman}^{Shannon}=0.86\), and
 \(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots, plots 8 and 12,
@@
 
-For all undisturbed plots the Rao diversity remained comparable to the
-initial values, 5 years after disturbance (1989 inventories). The
-trajectories of all disturbed plots followed a unimodal trajectory with
-a maximum positively correlated to disturbance intensity
-(\(\rho_{spearman}\)) (Figure \ref{fig:DivFun}). The plot 7 from
-treatment 1 was removed from the graphical representation for better
-readability, see appendix for full graph. Thirty years after disturbance
-all plots, whenever the initial disturbance intensity, regained
-diversity values similar to their initial avule and to those of control
-plots.
+The functional diversity trajectories of plots based on the 8 leaf, stem
+and life history traits was examined through the Rao diversity. For all
+undisturbed plots the functional diversity remained comparable to the
+reference values 5 years after disturbance (1989 inventories). For all
+disturbed plots the functional trajectories followed a unimodal
+trajectory with a maximum positively correlated to disturbance intensity
+(\(\rho_{spearman}=XX\)) (Figure \ref{fig:DivFun}). The plot 7 from
+treatment 1 that displayed a constantly outlying diversity was removed
+from the graphical representation for better readability, see appendix
+for full graph. Thirty years after disturbance all plots, whenever the
+initial disturbance intensity, regained diversity values similar to
+their initial avule and to those of control plots.
 

### 2018-04-20 Scripts/DrawingTools.R

@@ -176,8 +176,8 @@ FunTraj<-function(CompFun,remove=TRUE){
   
   CompFun<-lapply(CompFun,function(tr){
     ret<-tr[,which(colnames(tr)>=1989),]
-    #ret<-lapply(1:dim(ret)[3],function(rep){return(apply(ret[,,rep],2,function(col){col<-col-ret[,1,rep]}))})
-    #ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),dimnames=list(rownames(ret[[1]]),colnames(ret[[1]]),1:length(ret)))
+    ret<-lapply(1:dim(ret)[3],function(rep){return(apply(ret[,,rep],2,function(col){col<-col-ret[,1,rep]}))})
+    ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),dimnames=list(rownames(ret[[1]]),colnames(ret[[1]]),1:length(ret)))
     ret<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(ret,c(1,2),function(x){return(quantile(x,probs=quant))}))})
     return(array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),3),
                  dimnames=list(rownames(ret[[1]]),as.numeric(colnames(ret[[1]]))-1984,c(0.025,0.5,0.975))))})

### 2018-04-23 WholePlotTrajectories.Rmd
FunctionalGraph

@@
 The taxonomic diversity was assessed through Richness and the Hill number translation of Shannon and Simpson indices [@Hill1973].
->>To tackle the unequal number of recruited trees among treatments the indices bias corrected estimator were used, following [@Chao2015;@Marcon2015b].
 
@@
 Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).", fig.env="figure*"}
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).", fig.env="figure"}
 load("DB/FunctionalComposition_forGraph")
@@
 For all undisturbed plots the functional diversity remained comparable to the reference values 5 years after disturbance (1989 inventories).
-For all disturbed plots the functional trajectories followed a unimodal trajectory with a maximum positively correlated to disturbance intensity ($\rho_{spearman}=XX$) (Figure \@ref(fig:DivFun)). 
+For all disturbed plots the functional trajectories followed a unimodal trajectory with a maximum positively correlated to disturbance intensity ($\rho_{spearman}=0.73$) (Figure \@ref(fig:DivFun)). 
 The plot 7 from treatment 1 that displayed a constantly outlying diversity was removed from the graphical representation for better readability, see appendix for full graph.
@@
 
-```{r DivFun, out.width = '80%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
+```{r DivFun, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.", fig.env="figure*"}
-load("DB/FunctionalTraj_ForGraphsnew")
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.", fig.env="figure"}
+load("DB/FunctionalTraj_ForGraphs")
 
-FunTraj(CompleteFun,remove=F)
+FunTraj(CompleteFun,remove=T)
 

### 2018-04-23 Scripts/DrawingTools.R

@@ -10,7 +10,7 @@ smooth<-function(mat,larg){return(do.call(cbind,lapply(1:ncol(mat),function(step
   rowSums(mat[,range])/length(range)})))}
 
 TaxoCompo<-function(Data_TaxoComp){
-par(mfrow=c(1,2))
+par(mfrow=c(1,2),mar=c(3,3,2,1))
 matT<-as.data.frame(Data_TaxoComp[[1]])
 matT$plot<-substr(rownames(matT),start=1,stop=regexpr("_",rownames(matT))-1)
 matT$year<-substr(rownames(matT),start=regexpr("_",rownames(matT))+1,
@@ -176,8 +176,8 @@ FunTraj<-function(CompFun,remove=TRUE){
   
   CompFun<-lapply(CompFun,function(tr){
     ret<-tr[,which(colnames(tr)>=1989),]
-    ret<-lapply(1:dim(ret)[3],function(rep){return(apply(ret[,,rep],2,function(col){col<-col-ret[,1,rep]}))})
-    ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),dimnames=list(rownames(ret[[1]]),colnames(ret[[1]]),1:length(ret)))
+    #ret<-lapply(1:dim(ret)[3],function(rep){return(apply(ret[,,rep],2,function(col){col<-col-ret[,1,rep]}))})
+    #ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),dimnames=list(rownames(ret[[1]]),colnames(ret[[1]]),1:length(ret)))
     ret<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(ret,c(1,2),function(x){return(quantile(x,probs=quant))}))})
     return(array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),3),
                  dimnames=list(rownames(ret[[1]]),as.numeric(colnames(ret[[1]]))-1984,c(0.025,0.5,0.975))))})

### 2018-04-23 Scripts/Stats.R

@@ -187,3 +187,21 @@ Rao<-do.call(rbind,lapply(P,function(yr){
   #return(expq(Hqz(as.AbdVector(tapply(yr,yr,length)), q=2, Z=dissim,Correction="None"),q=2))
 }))
 
+###### Functional spearman
+load("DB/FunctionalTraj_ForGraphs")
+
+CompFun<-CompleteFun
+
+#if(remove){CompFun[[2]]<-CompFun[[2]][which(rownames(CompFun[[2]])!=7),,]}
+
+CompFun<-do.call(c,lapply(CompFun,function(tr){
+  ret<-tr[,which(colnames(tr)>=1989),]
+  ret<-apply(ret,c(1,2),median)
+  return(apply(ret,1,max))}))
+
+CompFun<-cbind(CompFun,names(CompFun),rep(0:3,each=3))
+colnames(CompFun)<-c("Max","Plot","treat")
+cor(as.numeric(CompFun[,"Max"]),as.numeric(CompFun[,"treat"]),method="spearman")
+
+
+

### 2018-04-23 WholePlotTrajectories.Rmd
html pages

@@
 The weighted means of bark thickness (*Bark_thick*) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value and this of the undisturbed plots at the end of the experiment.
-Once again the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_thickness}=0.67$,$\rho_{spearman}^{L_chloro}=0.45$,$\rho_{spearman}^{L_toughness}=-0.43$,$\rho_{spearman}^{SLA}=0.93$,$\rho_{spearman}^{WD}=-0.78$,$\rho_{spearman}^{Bark-thickness}=0.88$,$\rho_{spearman}^{Hmax}=-0.48$).
+Once again the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_thickness}=0.67$, $\rho_{spearman}^{L_chloro}=0.45$, $\rho_{spearman}^{L_toughness}=-0.43$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WD}=-0.78$, $\rho_{spearman}^{Bark-thickness}=0.88$, $\rho_{spearman}^{Hmax}=-0.48$).
 
@@
 
-The functional diversity followed a hump-back trajectory which maximum was positively correlated to the disturbance intensity ($\rho_{spearman}=$) and after 30 years, all disturbed plots had recovered diversities close to their initial values.
+The functional diversity followed a hump-back trajectory which maximum was positively correlated to the disturbance intensity and after 30 years, all disturbed plots had recovered diversities close to their initial values.
 Similar hump-back trajectories were followed by all leaf and stem functional traits and life history traits, to the exception of leaf chlorophyll content.

### 2018-04-24 WholePlotTrajectories.Rmd
discussion

@@
 
-The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging [@Cannon1998;@Baraloto2012a].
-Only the most disturbed plots had not fully achieved the ongoing recovery of genus richness, although all had already reached equivalent levels as those of some control plots.
-Such long time richness recovery had already been observed in several logging experiments, which also highlighted the role of random disturbance damage in the richness decrease [@DeAvila2015;@Hu2018].
+The monitoring of disturbed forest communities confirmed a limited negative impact of disturbance on species richness, as observed on several post logging surveys [@Cannon1998;@Baraloto2012a].
+Thirty years after disturbance, the most disturbed plots had recovered their initial genus richness, similar to this of control plots.
+After low intensity plots the genus richness substantially increased for low intensity plots, reaching a gain of almost 12 genuses for some plots.
 

### 2018-04-24 Scripts/DrawingTools.R

@@ -144,7 +144,7 @@ TaxoTraj<-function(CompTaxo){
     Toplot<-lapply(CompTaxo,function(tr){return(tr[,,,q])})
     Toplot<-lapply(Toplot,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
     Toplot<-lapply(Toplot,function(tr){
-      ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col-tr[,1,rep]}))})#
+      ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col}))})#-tr[,1,rep]
       ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),
                  dimnames=list(rownames(ret[[1]]),as.numeric(colnames(ret[[1]]))-1984,1:length(ret)))
       ret<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(ret,c(1,2),function(x){return(quantile(x,probs=quant))}))})

### 2018-04-24 WholePlotTrajectories.Rmd
Discussion

@@
 The monitoring of disturbed forest communities confirmed a limited negative impact of disturbance on species richness, as observed on several post logging surveys [@Cannon1998;@Baraloto2012a].
-Thirty years after disturbance, the most disturbed plots had recovered their initial genus richness, similar to this of control plots.
-After low intensity plots the genus richness substantially increased for low intensity plots, reaching a gain of almost 12 genuses for some plots.
+Thirty years after disturbance, the genus richness was restored to initial plot values after high disturbance intensity while it substantially increased after low disturbance intensity, reaching a gain of almost 12 genuses for some plots.
 
-The evenness of all disturbed plots and the richness of low disturbance plots on the other hand mainly followed an asymptotic growth sharply increasing until 15 years after disturbance.
-Such increased evenness is explained by a much more homogeneous species distribution after disturbance.
-Communities composition after disturbance is either due to the old, pre-disturbance survivors or to the recruited trees.
-The composition of old survivors proved to mirror the initial community [@Herault2018], the observed composition turnover would stem from an enhanced the growth and survival of previously infrequent species which, along with the IDH, reorganizes the typical high dominance structure of hyperdiverse mature forests.
-This increase in taxonomic diversity was accompanied by an increased taxonomic dissimilarity to initial state along time, translating a a taxonomic turnover benefiting to pioneers and light demanding species, as illustrated by the communities functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
-Post-disturbance changes in abiotic environment and competitive pressure then well favored pioneers outcompeting other species when abiotic resources are abundant but otherwise excluded in mature forests by long-lived, resistant and shade tolerant species.
-The trajectories of proportions among seed mass classes besides revealed the enhanced growth of small seeded, large dispersive, species and therefore the importance of species reproductive strategy which further supported the role of species dispersion and demographic strategy for the post-disturbance dynamics, as suggested by the IDH [@TerSteege2001;@Flores2006;@Haddad2008].
+Both richness and evenness after disturbance followed asymptotic trajectories, sharply increasing for 15 years before stabilizing at higher values than those of control plots.
+An increase of communities evenness stems from a higher homogeneity of species distribution.
+After disturbance, communities are made of the old, pre-disturbance survivors and the newly recruited trees: changes in composition and abundance distribution are to be found in the recruitment processes and in the pre-disturbance survivors mortality.
+The composition of old survivors proved to mirror the initial community [@Herault2018], which leads to support a species turnover among recruited trees with enhanced growth and survival of previously infrequent species.
+Indeed, the increase in taxonomic diversity was accompanied by an increase of taxonomic dissimilarity with plots initial state and a functional shift towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
+Disturbance then causes a reorganization of the typical high dominance structure of hyperdiverse mature forests after disturbance, benefiting to pioneers and light demanding species. 
+Likely, the changes in abiotic environment and competitive pressure favored pioneers which outcompete others in non limiting resources but are excluded in mature forests by long-lived, more resistant and shade tolerant species.
+
+As stated by the IDH, communities dynamics after disturbance then rely on species functional strategy and corresponding ability to fill the environmental niches made available by disturbance.
+Recruited species then mix with pre-disturbance ones from which they differ and constitue a community all the more diversified that the disturbance is intense [@Molino2001].
 
@@
 
-The functional diversity followed a hump-back trajectory which maximum was positively correlated to the disturbance intensity and after 30 years, all disturbed plots had recovered diversities close to their initial values.
-Similar hump-back trajectories were followed by all leaf and stem functional traits and life history traits, to the exception of leaf chlorophyll content.
-Although the return to initial values was still ongoing the recovery of functional diversity and average communities trait values was undeniable.
-This translated the recovery of ecosystems processes [@Guariguata2001], as functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005].
-Same recovery trajectories were followed by taxonomic and functional composition which argued for communities convergence and the maintenance of plots initial difference [@Hubbell1999;@Molino2001;@Baraloto2012a].
-These compositon difference, and probably also some abiotic parameters, besides strongly determined the maximum and time path of the trajectories which, despite an homogeneous trend  were quite variable among plots.
+The trajectories of almost all leaf and stem functional traits, life history traits,and functional diversity followed a hump-back trajectory which maximum were positively correlated to the disturbance intensity, and were recovering towards initial values after 30 years that trajectory.
+Functoinal traits and communities functional diversity are the most direct link between biodiversity and ecosystem functioning [@Diaz2005]: these trajectories reflected a consistent recovery in the long term of ecosystem processes [@Guariguata2001].
+
+Same hump-back trajectories were followed by the composition of plots species and functoinal composition.
+This similarly meant a recovery of the pre-disturbance state and the maintenance of the initial differences among plots which supports the idea of a convergence of communities composition [@Hubbell1999;@Molino2001;@Baraloto2012a].
+Although plots followed homogeneous trends these initial differences in compositon difference, and probably also some abiotic parameters, strongly determined the maximum and time path of the trajectories [@Anderson2007]•.
 
@@
 
-Despite the consistent recovery of initial diversity and composition there was a time lag between the communities taxonomic and functional characteristics.
-Taxonomic composition had longer time path and while communities had recovered their functional diveristy, their taxonomic composition and evenness remained higher than before disturbance.
+Despite the consistent recovery of initial diversity and composition there was a time lag between the trajectories followed by taxonomic and functional characteristics.
+Taxonomic composition had longer time path and while communities had recovered their functional characteristics, their taxonomic composition and evenness remained altered by disturbance.
 This delay between functional and taxonomic dynamics was already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
-Communities functional diversity rely, according to the "vegetation quantity effect" [@Grime1998], on the functional strategy of dominant species.
-Communities functional trajectory is then first driven by the increase of dominant species diversity and evenness following disturbance and then by the emergence of recruitment of species progressively resembling the old pre-disturbance community which decreased the diversity.
-Dominant species then restored the functional diversity and the dominant functional type of the community although the functional and taxonomic composition remained altered and the infrequent species of initial communities are still missing to plots recovery.
-The functional traits trajectories suggest that those species would match the dominant functional type therefore increasing communities functional redundancy.
-The functional overlap between species is typical of the huge biodiversity of tropical forests [@Bellwood2006], and proved far from recovery 03 years after disturbance.
-It is though a determinant of forests resilience and should be accounted for [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+The functional diversity of communities rely on the functional strategy of dominant species, according to the "vegetation quantity effect" [@Grime1998].
+The functional trajectories of communities are then driven first by the increase of diversity and evenness of dominant species and then by the recruitment of species resembling the old pre-disturbance community which reduced the overall diversity.
+Although dominant species restoredthe functional pre-disturbance characteristics, communities composition remained altered and the infrequent species of initial communities remain missing to achieve plots recovery.
+According to the funcitonal traits trajectories, those infrequent species would match the dominant functional type and would participate to the functional redundancy.
+The functional redundancy, thas is the functional overlap between species that is typical of the huge biodiversity of tropical forests [@Bellwood2006], therefore was not restored 30 years after disturbance and this is majr to consider as it defines forests' resilience [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
-Long-term alteration of functional redundancy besides came along with persistent compositional changes which most probably impact disturbance resistant species [@Haddad2008] and probably lianas or epiphytes [@Martin2013], and the modification of environmental conditions like soils nutrient cycling and compaction [@Olander2005] forest resilience remains under question [@Chazdon2003a].
+Besides the long-term alteration of functional redundancy, the persistent compositional changes which might have favored disturbance resistant species [@Haddad2008], lianas or epiphytes [@Martin2013] and would probably have come along with environmental change, like in the soils nutrient cycling and compaction [@Olander2005].
+These presistent changes highly question forest's resilience[@Chazdon2003a].
 New conditions would not only be longer lasting but self-maintained as tied to disturbance regime [@Burslem2000].
@@
 Our study showed the significant impact of disturbance on tropical forests communities.
-The subsequent diversity trajectories confirmed the intermediate disturbance hypothesis debated for tropical forests through their correlation with disturbance intensity.
-Besides it revealed the contrasting response of taxonomic and functional characteristics, specifically the decoupling between communities taxonomic evenness and their functional diversity and dominant functional traits values.
-The long-term disturbance trajectories observed highlighted the unachieved but consistent recovery of communities assembly for the lowest disturbance intensity but questioned it after higher disturbance.
+It validated the consistency of the IDH in the long term and revealed the contrasting response of taxonomic and functional characteristics with the decoupling between communities taxonomic evenness and their functional diversity and th dominant functional strategies.
+Communities recovery remained unachieved but consistent recovery of communities assembly for the lowest disturbance intensity but questioned it after higher disturbance.
 

### 2018-05-02 WholePlotTrajectories.Rmd
Discussions
Table

@@
 In tropical forests communities are shaped by a constant range of disturbance that changes the abiotic environment, as the light, heat and water fluxes, and the biotic interaction and competitive pressure.
-The cornerstone of tropical forests ecology is then to understand their response to disturbance and the corresponding mechanisms [@White2001;@Chazdon2003a].
+The cornerstone of tropical forests ecology is then to understand their response to disturbance and its mechanisms [@White2001;@Chazdon2003a].
 Forests response to disturbance has been largely studied through structural parameters rapid and convenient to measure, as aboveground biomass, tree height or stem density.
@@
 
-```{r Tab1, echo=FALSE}
+```{r Tab1, echo=FALSE,is.as=T}
 Intervention <- data.frame(

### 2018-05-02 WholePlotTrajectories.tex
FunctionalGraph
Table

@@
 cornerstone of tropical forests ecology is then to understand their
-response to disturbance and the corresponding mechanisms
+response to disturbance and its mechanisms
 \citep{White2001, Chazdon2003a}. Forests response to disturbance has
@@
 number translation of Shannon and Simpson indices \citep{Hill1973}.
-\textgreater{}\textgreater{}To tackle the unequal number of recruited
-trees among treatments the indices bias corrected estimator were used,
-following \citep{Chao2015, Marcon2015b}.
 
@@
 
-\begin{figure*}
+\begin{figure}
 
@@
 Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).}\label{fig:NMDSplans}
-\end{figure*}
+\end{figure}
 
@@
 again the maximum difference to initial state was correlated to the
-disturbance intensity
-(\(\rho_{spearman}^{L_thickness}=0.67\),\(\rho_{spearman}^{L_chloro}=0.45\),\(\rho_{spearman}^{L_toughness}=-0.43\),\(\rho_{spearman}^{SLA}=0.93\),\(\rho_{spearman}^{WD}=-0.78\),\(\rho_{spearman}^{Bark-thickness}=0.88\),\(\rho_{spearman}^{Hmax}=-0.48\)).
+disturbance intensity (\(\rho_{spearman}^{L_thickness}=0.67\),
+\(\rho_{spearman}^{L_chloro}=0.45\),
+\(\rho_{spearman}^{L_toughness}=-0.43\), \(\rho_{spearman}^{SLA}=0.93\),
+\(\rho_{spearman}^{WD}=-0.78\),
+\(\rho_{spearman}^{Bark-thickness}=0.88\),
+\(\rho_{spearman}^{Hmax}=-0.48\)).
 
@@
 trajectory with a maximum positively correlated to disturbance intensity
-(\(\rho_{spearman}=XX\)) (Figure \ref{fig:DivFun}). The plot 7 from
+(\(\rho_{spearman}=0.73\)) (Figure \ref{fig:DivFun}). The plot 7 from
 treatment 1 that displayed a constantly outlying diversity was removed
@@
 
-\begin{figure*}
+\begin{figure}
 
-{\centering \includegraphics[width=0.8\linewidth]{WholePlotTrajectories_files/figure-latex/DivFun-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivFun-1}
 
@@
 Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:DivFun}
-\end{figure*}
+\end{figure}
 
@@
 
-The monitoring 30 years of disturbed forest communities confirmed the
-limited direct impact of low disturbance on species richness as it was
-found for impact survey of selective logging
-\citep{Cannon1998, Baraloto2012a}. Only the most disturbed plots had not
-fully achieved the ongoing recovery of genus richness, although all had
-already reached equivalent levels as those of some control plots. Such
-long time richness recovery had already been observed in several logging
-experiments, which also highlighted the role of random disturbance
-damage in the richness decrease \citep{DeAvila2015, Hu2018}.
+The monitoring of disturbed forest communities confirmed a limited
+negative impact of disturbance on species richness, as observed on
+several post logging surveys \citep{Cannon1998, Baraloto2012a}. Thirty
+years after disturbance, the genus richness was restored to initial plot
+values after high disturbance intensity while it substantially increased
+after low disturbance intensity, reaching a gain of almost 12 genuses
+for some plots.
 
-The evenness of all disturbed plots and the richness of low disturbance
-plots on the other hand mainly followed an asymptotic growth sharply
-increasing until 15 years after disturbance. Such increased evenness is
-explained by a much more homogeneous species distribution after
-disturbance. Communities composition after disturbance is either due to
-the old, pre-disturbance survivors or to the recruited trees. The
-composition of old survivors proved to mirror the initial community
-\citep{Herault2018}, the observed composition turnover would stem from
-an enhanced the growth and survival of previously infrequent species
-which, along with the IDH, reorganizes the typical high dominance
-structure of hyperdiverse mature forests. This increase in taxonomic
-diversity was accompanied by an increased taxonomic dissimilarity to
-initial state along time, translating a a taxonomic turnover benefiting
-to pioneers and light demanding species, as illustrated by the
-communities functional shifts towards resource-acquisitive strategies
-(sharp increase in the SLA, leaf thickness and bark thickness and
-decrease in wood density, leaf toughness and maximum height)
-\citep{Westoby1998, Wright2004, Reich2014}. Post-disturbance changes in
-abiotic environment and competitive pressure then well favored pioneers
-outcompeting other species when abiotic resources are abundant but
-otherwise excluded in mature forests by long-lived, resistant and shade
-tolerant species. The trajectories of proportions among seed mass
-classes besides revealed the enhanced growth of small seeded, large
-dispersive, species and therefore the importance of species reproductive
-strategy which further supported the role of species dispersion and
-demographic strategy for the post-disturbance dynamics, as suggested by
-the IDH \citep{TerSteege2001, Flores2006, Haddad2008}.
+Both richness and evenness after disturbance followed asymptotic
+trajectories, sharply increasing for 15 years before stabilizing at
+higher values than those of control plots. An increase of communities
+evenness stems from a higher homogeneity of species distribution. After
+disturbance, communities are made of the old, pre-disturbance survivors
+and the newly recruited trees: changes in composition and abundance
+distribution are to be found in the recruitment processes and in the
+pre-disturbance survivors mortality. The composition of old survivors
+proved to mirror the initial community \citep{Herault2018}, which leads
+to support a species turnover among recruited trees with enhanced growth
+and survival of previously infrequent species. Indeed, the increase in
+taxonomic diversity was accompanied by an increase of taxonomic
+dissimilarity with plots initial state and a functional shift towards
+resource-acquisitive strategies (sharp increase in the SLA, leaf
+thickness and bark thickness and decrease in wood density, leaf
+toughness and maximum height)
+\citep{Westoby1998, Wright2004, Reich2014}. Disturbance then causes a
+reorganization of the typical high dominance structure of hyperdiverse
+mature forests after disturbance, benefiting to pioneers and light
+demanding species. Likely, the changes in abiotic environment and
+competitive pressure favored pioneers which outcompete others in non
+limiting resources but are excluded in mature forests by long-lived,
+more resistant and shade tolerant species.
+
+As stated by the IDH, communities dynamics after disturbance then rely
+on species functional strategy and corresponding ability to fill the
+environmental niches made available by disturbance. Recruited species
+then mix with pre-disturbance ones from which they differ and constitue
+a community all the more diversified that the disturbance is intense
+\citep{Molino2001}.
 
@@
 
-The functional diversity followed a hump-back trajectory which maximum
-was positively correlated to the disturbance intensity
-(\(\rho_{spearman}=\)) and after 30 years, all disturbed plots had
-recovered diversities close to their initial values. Similar hump-back
-trajectories were followed by all leaf and stem functional traits and
-life history traits, to the exception of leaf chlorophyll content.
-Although the return to initial values was still ongoing the recovery of
-functional diversity and average communities trait values was
-undeniable. This translated the recovery of ecosystems processes
-\citep{Guariguata2001}, as functional traits are the most direct link
-between biodiversity and ecosystem functioning \citep{Diaz2005}. Same
-recovery trajectories were followed by taxonomic and functional
-composition which argued for communities convergence and the maintenance
-of plots initial difference
-\citep{Hubbell1999, Molino2001, Baraloto2012a}. These compositon
-difference, and probably also some abiotic parameters, besides strongly
-determined the maximum and time path of the trajectories which, despite
-an homogeneous trend were quite variable among plots.
+The trajectories of almost all leaf and stem functional traits, life
+history traits,and functional diversity followed a hump-back trajectory
+which maximum were positively correlated to the disturbance intensity,
+and were recovering towards initial values after 30 years that
+trajectory. Functoinal traits and communities functional diversity are
+the most direct link between biodiversity and ecosystem functioning
+\citep{Diaz2005}: these trajectories reflected a consistent recovery in
+the long term of ecosystem processes \citep{Guariguata2001}.
+
+Same hump-back trajectories were followed by the composition of plots
+species and functoinal composition. This similarly meant a recovery of
+the pre-disturbance state and the maintenance of the initial differences
+among plots which supports the idea of a convergence of communities
+composition \citep{Hubbell1999, Molino2001, Baraloto2012a}. Although
+plots followed homogeneous trends these initial differences in
+compositon difference, and probably also some abiotic parameters,
+strongly determined the maximum and time path of the trajectories
+\citep{Anderson2007}•.
 
@@
 Despite the consistent recovery of initial diversity and composition
-there was a time lag between the communities taxonomic and functional
-characteristics. Taxonomic composition had longer time path and while
-communities had recovered their functional diveristy, their taxonomic
-composition and evenness remained higher than before disturbance. This
-delay between functional and taxonomic dynamics was already observed for
-grasslands \citep{Tilman1997, Mouillot2011} and more recently for
-tropical forests \citep{Lohbeck2015, Guariguata2001}. Communities
-functional diversity rely, according to the ``vegetation quantity
-effect'' \citep{Grime1998}, on the functional strategy of dominant
-species. Communities functional trajectory is then first driven by the
-increase of dominant species diversity and evenness following
-disturbance and then by the emergence of recruitment of species
-progressively resembling the old pre-disturbance community which
-decreased the diversity. Dominant species then restored the functional
-diversity and the dominant functional type of the community although the
-functional and taxonomic composition remained altered and the infrequent
-species of initial communities are still missing to plots recovery. The
-functional traits trajectories suggest that those species would match
-the dominant functional type therefore increasing communities functional
-redundancy. The functional overlap between species is typical of the
-huge biodiversity of tropical forests \citep{Bellwood2006}, and proved
-far from recovery 03 years after disturbance. It is though a determinant
-of forests resilience and should be accounted for
-\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+there was a time lag between the trajectories followed by taxonomic and
+functional characteristics. Taxonomic composition had longer time path
+and while communities had recovered their functional characteristics,
+their taxonomic composition and evenness remained altered by
+disturbance. This delay between functional and taxonomic dynamics was
+already observed for grasslands \citep{Tilman1997, Mouillot2011} and
+more recently for tropical forests \citep{Lohbeck2015, Guariguata2001}.
+The functional diversity of communities rely on the functional strategy
+of dominant species, according to the ``vegetation quantity effect''
+\citep{Grime1998}. The functional trajectories of communities are then
+driven first by the increase of diversity and evenness of dominant
+species and then by the recruitment of species resembling the old
+pre-disturbance community which reduced the overall diversity. Although
+dominant species restoredthe functional pre-disturbance characteristics,
+communities composition remained altered and the infrequent species of
+initial communities remain missing to achieve plots recovery. According
+to the funcitonal traits trajectories, those infrequent species would
+match the dominant functional type and would participate to the
+functional redundancy. The functional redundancy, thas is the functional
+overlap between species that is typical of the huge biodiversity of
+tropical forests \citep{Bellwood2006}, therefore was not restored 30
+years after disturbance and this is majr to consider as it defines
+forests' resilience \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-Long-term alteration of functional redundancy besides came along with
-persistent compositional changes which most probably impact disturbance
-resistant species \citep{Haddad2008} and probably lianas or epiphytes
-\citep{Martin2013}, and the modification of environmental conditions
-like soils nutrient cycling and compaction \citep{Olander2005} forest
-resilience remains under question \citep{Chazdon2003a}. New conditions
-would not only be longer lasting but self-maintained as tied to
-disturbance regime \citep{Burslem2000}. Specificaly, this would impair
-species contingent to undisturbed forests, threatening their
-maintenance, and run the risk to loose cornerstone species and trigger
-unexpected ecological consequences
-\citep{Jones1994, Diaz2005, Gardner2007}.
+Besides the long-term alteration of functional redundancy, the
+persistent compositional changes which might have favored disturbance
+resistant species \citep{Haddad2008}, lianas or epiphytes
+\citep{Martin2013} and would probably have come along with environmental
+change, like in the soils nutrient cycling and compaction
+\citep{Olander2005}. These presistent changes highly question forest's
+resilience\citep{Chazdon2003a}. New conditions would not only be longer
+lasting but self-maintained as tied to disturbance regime
+\citep{Burslem2000}. Specificaly, this would impair species contingent
+to undisturbed forests, threatening their maintenance, and run the risk
+to loose cornerstone species and trigger unexpected ecological
+consequences \citep{Jones1994, Diaz2005, Gardner2007}.
 
@@
 Our study showed the significant impact of disturbance on tropical
-forests communities. The subsequent diversity trajectories confirmed the
-intermediate disturbance hypothesis debated for tropical forests through
-their correlation with disturbance intensity. Besides it revealed the
-contrasting response of taxonomic and functional characteristics,
-specifically the decoupling between communities taxonomic evenness and
-their functional diversity and dominant functional traits values. The
-long-term disturbance trajectories observed highlighted the unachieved
-but consistent recovery of communities assembly for the lowest
-disturbance intensity but questioned it after higher disturbance.
+forests communities. It validated the consistency of the IDH in the long
+term and revealed the contrasting response of taxonomic and functional
+characteristics with the decoupling between communities taxonomic
+evenness and their functional diversity and th dominant functional
+strategies. Communities recovery remained unachieved but consistent
+recovery of communities assembly for the lowest disturbance intensity
+but questioned it after higher disturbance.
 

### 2018-05-03 WholePlotTrajectories.Rmd
small corrections

@@
 
-In tropical forests communities are shaped by a constant range of disturbance that changes the abiotic environment, as the light, heat and water fluxes, and the biotic interaction and competitive pressure.
+In tropical forests communities are shaped by a constant range of disturbance that changes the abiotic environment, as the light, the heat and water fluxes, and the biotic interaction and competitive pressure.
 The cornerstone of tropical forests ecology is then to understand their response to disturbance and its mechanisms [@White2001;@Chazdon2003a].
-Forests response to disturbance has been largely studied through structural parameters rapid and convenient to measure, as aboveground biomass, tree height or stem density.
-These structural parameters have then been sucessfully modeled and assessed the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
-However the response of tree species diversity although it is determinant of ecosystems productivity, stability and functioning [@Tilman2014] and it would be most probably impacted by the biotic and abiotic changes after disturbance [@CazzollaGatti2014].
+Forests response has been largely studied through structural parameters rapid and convenient to measure, as aboveground biomass, tree height or stem density.
+These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
+However the response of tree species diversity, which is determinant of ecosystems productivity, stability and functioning [@Tilman2014] and would be most probably impacted by post-disturbance environmental changes [@CazzollaGatti2014], remains unclear.
 
@@
 However, validations of the IDH in the long term remain scarce and mainly rely on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-More releveant monitoring would encompass communities composition, that are crucial for conservation issues, and evenness, that revealing ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
+More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and evenness, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
 Furthermore, the functional approach accounting for species biological attributes would be insightful as it reveals species fitness and ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
 In that respect major functional traits related to species ecology and performance were largely adopted through a relevant framework [@Diaz2005;@Villeger2008a].
-The functonial trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
+The functional trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
 This translated by shifts from "conservative" slow-growing species dealing with scarce resources, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
 It was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
-A proper monitoring of communities response should therefore encompass taxonomic and functional diversity and composition measures to test the validity of the IDH in the long term for tropical forests, and clarify the resilience of communities evenness, composition,and also functioning.
+A proper monitoring of communities response should therefore encompass taxonomic and functional diversity and composition measures to test the validity of the IDH in the long term for tropical forests, and clarify the resilience of communities evenness, composition, and also functioning.
 The trajectories followed by all these facets would highlight the role of deterministic processes, like competitive exclusion or abiotic selection, and the communities' convergence maintaining intrinsinc differences in diversity and composition,which is as much insgights for future adaptive conservation strategies [@Adler2007].
@@
 
-Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-months dry season (< 100 mm.months^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
+Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
 Elevation ranges between 5 and 50 m and mean annual temperature is 26°C. 
@@
 The botanical campaigns carried every 5 to 6 years to identify all trees at the species level only started in 2003 and identification practices varied among plots and successive campaigns. 
-This raised methodological issues as vernacular names usually correspond to different botanical species, resulting in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics throught a Bayesian framework.
-The uncertainty propagation was done by the replenishment of inventories completed at genus level from real incomplete ones on the basisof vernacular/botanical names association.
+This raised methodological issues as vernacular names usually correspond to different botanical species, resulting in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics through a Bayesian framework.
+The uncertainty propagation was done by the replenishment of inventories completed at genus level from real incomplete ones on the basis of vernacular/botanical names association.
 
 Vernacular names were replaced through multinomial trials $M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)$ based on the observed association probability $\big[\alpha_1, \alpha_2,…, \alpha_3\big]$ between each vernacular name *v* and the species $\big[s_1, s_2, …, s_N\big]$ recorded in the inventory.
-
-To avoid remaining identification caveats and consider complete inventories, the simulated inventories were then reported at genus.
 See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.
@@
 
-The trait database came from the BRIDGE project [^1] where traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou. 
+The trait database came from the BRIDGE project [^1] where trait values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou. 
 Missing trait values were filled using multivariate imputation by chained equation (mice) restricted to samples pertaining to the next higher taxonomic level, in order to account for the phylogenetic signal of the functional traits.
@@
 
-[^1] http://www.ecofog.gf/Bridge/
+[^1]: http://www.ecofog.gf/Bridge/
 
@@
 
-To counter the remaining taxonomic uncertainty plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
+To counter the remaining taxonomic uncertainty, plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
 Variations in plots taxonomic and functional composition after disturbance was visualized by their trajectories in a two-dimensional ordination space over 30 years.
@@
 The trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004].
-To compensate the intrinsinc difference among plots the trajectories corresponded to the differences along time with the reference inventory in 1984 (5 year after disturbance).
+To compensate the intrinsinc difference among plots the trajectories corresponded to the differences along time with the reference inventory in 1989.
 Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were  therefore reported as the proportion of each class in the inventories.
@@
 Over time, 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
-Taxonomic and functional plots trajectories were examined through the ordination in a two dimensional ordiantion space of successive inventories from 1989 (5 years after disturbance) to 2015 (31 years after disturbance).
+Taxonomic and functional plots trajectories were examined through the trajecotries in a two dimensional ordination space of successive inventories from 1989 (5 years after disturbance) to 2015 (31 years after disturbance).
 Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity, this last having given similar results (data not shown).
@@
 
-The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area,wood specific gravity and barck thickness and seed mass and maximum adult height) visaualized with plots community weighted means (CWM) (Figure \@ref(fig:CWM)).
+The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area, wood specific gravity and bark thickness and seed mass and maximum adult height) visualized with plots community weighted means (CWM) (Figure \@ref(fig:CWM)).
 
@@
 The weighted means of bark thickness (*Bark_thick*) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value and this of the undisturbed plots at the end of the experiment.
-Once again the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_thickness}=0.67$, $\rho_{spearman}^{L_chloro}=0.45$, $\rho_{spearman}^{L_toughness}=-0.43$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WD}=-0.78$, $\rho_{spearman}^{Bark-thickness}=0.88$, $\rho_{spearman}^{Hmax}=-0.48$).
+Once again the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.67$, $\rho_{spearman}^{L_{chloro}}=0.45$, $\rho_{spearman}^{L_{toughness}}=-0.43$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WD}=-0.78$, $\rho_{spearman}^{Bark-thickness}=0.88$, $\rho_{spearman}^{Hmax}=-0.48$).
 
@@
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
 load("DB/CWM")
@@
 
-For undisturbed plots the Richness, Shannon and Simpson diversity remained stable over the 30 monitored years.
-After disturbance the richness increased when the disturbance intensity was low, with a maximum increase of 14 botanical genuses (plot 3 from treatment 2).
-When the disturbance was intense however, plot staxonomic richness followed a unimodal decrease with a return to initial values after intense disturbance.
-The taxonomic evenness (Shannon and Simpson diversities) significantly increased whenever the disturbance intensity.
+For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 monitored years.
+After disturbance the richness increased when the disturbance intensity was low, with a maximum increase of 14 botanical genera (plot 3 from treatment 2).
+When the disturbance was intense however, plots taxonomic richness followed a unimodal decrease with a return to initial values after intense disturbance.
+The taxonomic evenness (Shannon and Simpson diversities) significantly increased irrespective of the disturbance intensity.
 The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum reached around 20 years and positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
@@
 The plot 7 from treatment 1 that displayed a constantly outlying diversity was removed from the graphical representation for better readability, see appendix for full graph.
-Thirty years after disturbance all plots, whenever the initial disturbance intensity, regained diversity values similar to their initial avule and to those of control plots.
+Thirty years after disturbance all plots, whenever the initial disturbance intensity, regained diversity values similar to their initial value and to those of control plots.
 
 ```{r DivFun, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.", fig.env="figure"}
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure"}
 load("DB/FunctionalTraj_ForGraphs")
@@
 The monitoring of disturbed forest communities confirmed a limited negative impact of disturbance on species richness, as observed on several post logging surveys [@Cannon1998;@Baraloto2012a].
-Thirty years after disturbance, the genus richness was restored to initial plot values after high disturbance intensity while it substantially increased after low disturbance intensity, reaching a gain of almost 12 genuses for some plots.
+Thirty years after disturbance, the genus richness was restored to initial plot values after high disturbance intensity while it substantially increased after low disturbance intensity, reaching a gain of almost 12 genera for some plots.
 
@@
 
-The trajectories of almost all leaf and stem functional traits, life history traits,and functional diversity followed a hump-back trajectory which maximum were positively correlated to the disturbance intensity, and were recovering towards initial values after 30 years that trajectory.
-Functoinal traits and communities functional diversity are the most direct link between biodiversity and ecosystem functioning [@Diaz2005]: these trajectories reflected a consistent recovery in the long term of ecosystem processes [@Guariguata2001].
+The trajectories of almost all leaf and stem functional traits, life history traits, and functional diversity followed hump-back trajectories whose maximum were positively correlated to the disturbance intensity, and were recovering towards initial values after 30 years.
+Functional traits and communities functional diversity are the most direct link between biodiversity and ecosystem functioning [@Diaz2005]: these trajectories reflected a consistent recovery in the long term of ecosystem processes [@Guariguata2001].
 
@@
 Taxonomic composition had longer time path and while communities had recovered their functional characteristics, their taxonomic composition and evenness remained altered by disturbance.
+
 This delay between functional and taxonomic dynamics was already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
@@
 The functional trajectories of communities are then driven first by the increase of diversity and evenness of dominant species and then by the recruitment of species resembling the old pre-disturbance community which reduced the overall diversity.
-Although dominant species restoredthe functional pre-disturbance characteristics, communities composition remained altered and the infrequent species of initial communities remain missing to achieve plots recovery.
-According to the funcitonal traits trajectories, those infrequent species would match the dominant functional type and would participate to the functional redundancy.
-The functional redundancy, thas is the functional overlap between species that is typical of the huge biodiversity of tropical forests [@Bellwood2006], therefore was not restored 30 years after disturbance and this is majr to consider as it defines forests' resilience [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+Although dominant species restored the functional pre-disturbance characteristics, communities composition remained altered and the infrequent species of initial communities remain missing to achieve plots recovery.
+According to the functional traits trajectories, those infrequent species would match the dominant functional type and would participate to the functional redundancy.
+The functional redundancy, thas is the functional overlap between species that is typical of the huge biodiversity of tropical forests [@Bellwood2006], therefore was not restored 30 years after disturbance and this is major to consider as it defines forests' resilience [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
-Besides the long-term alteration of functional redundancy, the persistent compositional changes which might have favored disturbance resistant species [@Haddad2008], lianas or epiphytes [@Martin2013] and would probably have come along with environmental change, like in the soils nutrient cycling and compaction [@Olander2005].
-These presistent changes highly question forest's resilience[@Chazdon2003a].
+Besides the long-term alteration of functional redundancy, there was probably persistent compositional changes favoring disturbance resistant species [@Haddad2008], lianas or epiphytes [@Martin2013] and environmental changes, like in the soils nutrient cycling and compaction [@Olander2005].
+These presistent changes highly question forest's resilience [@Chazdon2003a].
 New conditions would not only be longer lasting but self-maintained as tied to disturbance regime [@Burslem2000].

### 2018-05-03 WholePlotTrajectories.tex
small corrections

@@
 In tropical forests communities are shaped by a constant range of
-disturbance that changes the abiotic environment, as the light, heat and
-water fluxes, and the biotic interaction and competitive pressure. The
-cornerstone of tropical forests ecology is then to understand their
+disturbance that changes the abiotic environment, as the light, the heat
+and water fluxes, and the biotic interaction and competitive pressure.
+The cornerstone of tropical forests ecology is then to understand their
 response to disturbance and its mechanisms
-\citep{White2001, Chazdon2003a}. Forests response to disturbance has
-been largely studied through structural parameters rapid and convenient
-to measure, as aboveground biomass, tree height or stem density. These
-structural parameters have then been sucessfully modeled and assessed
-the maintenance of ecosystems processes and services
+\citep{White2001, Chazdon2003a}. Forests response has been largely
+studied through structural parameters rapid and convenient to measure,
+as aboveground biomass, tree height or stem density. These structural
+parameters have then been sucessfully modeled and allowed to assess the
+maintenance of ecosystems processes and services
 \citep{Denslow2000, Blanc2009, Rutishauser2016}. However the response of
-tree species diversity although it is determinant of ecosystems
-productivity, stability and functioning \citep{Tilman2014} and it would
-be most probably impacted by the biotic and abiotic changes after
-disturbance \citep{CazzollaGatti2014}.
+tree species diversity, which is determinant of ecosystems productivity,
+stability and functioning \citep{Tilman2014} and would be most probably
+impacted by post-disturbance environmental changes
+\citep{CazzollaGatti2014}, remains unclear.
 
@@
 forests recovery and functioning \citep{Martin2015, Chaudhary2016}. More
-releveant monitoring would encompass communities composition, that are
-crucial for conservation issues, and evenness, that revealing ecological
+releveant monitoring would encompass communities composition, that is
+crucial for conservation issues, and evenness, that reveals ecological
 rules underlying communities' structure
@@
 performance were largely adopted through a relevant framework
-\citep{Diaz2005, Villeger2008a}. The functonial trait-based approach,
+\citep{Diaz2005, Villeger2008a}. The functional trait-based approach,
 for example, highlighted in tropical rainforests the environmental
@@
 validity of the IDH in the long term for tropical forests, and clarify
-the resilience of communities evenness, composition,and also
+the resilience of communities evenness, composition, and also
 functioning. The trajectories followed by all these facets would
@@
 with mean annual precipitation averaging 2980 mm.y\textsuperscript{-1}
-(30-y period) with a 3-months dry season (\textless{} 100
-mm.months\textsuperscript{-1}) from mid-August to mid-November, and a
+(30-y period) with a 3-month dry season (\textless{} 100
+mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
 one-month dry season in March \citep{Wagner2011}. Elevation ranges
@@
 taxonomic uncertainties that had to be propagated to composition and
-diversity metrics throught a Bayesian framework. The uncertainty
+diversity metrics through a Bayesian framework. The uncertainty
 propagation was done by the replenishment of inventories completed at
-genus level from real incomplete ones on the basisof
+genus level from real incomplete ones on the basis of
 vernacular/botanical names association.
@@
 name \emph{v} and the species \(\big[s_1, s_2, …, s_N\big]\) recorded in
-the inventory.
-
-To avoid remaining identification caveats and consider complete
-inventories, the simulated inventories were then reported at genus. See
-appendix 1 and \citet{Aubry-Kientz2013} for the detailed methodology.
+the inventory. See appendix 1 and \citet{Aubry-Kientz2013} for the
+detailed methodology.
 
@@
 
-The trait database came from the BRIDGE project {[}\^{}1{]} where traits
-values were assessed from a selection of individuals located in nine
-permanent plots in French Guiana, including two in Paracou. Missing
-trait values were filled using multivariate imputation by chained
-equation (mice) restricted to samples pertaining to the next higher
-taxonomic level, in order to account for the phylogenetic signal of the
-functional traits. The dataset comprised 294 botanical species
+The trait database came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}
+where trait values were assessed from a selection of individuals located
+in nine permanent plots in French Guiana, including two in Paracou.
+Missing trait values were filled using multivariate imputation by
+chained equation (mice) restricted to samples pertaining to the next
+higher taxonomic level, in order to account for the phylogenetic signal
+of the functional traits. The dataset comprised 294 botanical species
 pertaining to 157 botanical genera.
@@
 values.
-
-{[}\^{}1{]} http://www.ecofog.gf/Bridge/
 
@@
 \bibliographystyle{mee}
-\bibliography{references.bib}
+\makeatletter
+% The filename has .bib extension the must be eliminated
+\filename@parse{references.bib}
+% parse stores the file name in base. Extension starts at the first dot, so don't use dots in file names.
+\bibliography{\filename@base}
+\makeatother
 

### 2018-05-03 WholePlotTrajectories.Rmd
discussion

@@
 
-Both richness and evenness after disturbance followed asymptotic trajectories, sharply increasing for 15 years before stabilizing at higher values than those of control plots.
+Both richness and evenness followed asymptotic trajectories after disturbance, sharply increasing for 15 years before stabilizing at higher values than those of control plots.
 An increase of communities evenness stems from a higher homogeneity of species distribution.
 After disturbance, communities are made of the old, pre-disturbance survivors and the newly recruited trees: changes in composition and abundance distribution are to be found in the recruitment processes and in the pre-disturbance survivors mortality.
-The composition of old survivors proved to mirror the initial community [@Herault2018], which leads to support a species turnover among recruited trees with enhanced growth and survival of previously infrequent species.
+Because the composition of old survivors proved to mirror the initial community [@Herault2018], a specific turnover was expected among recruited trees with enhanced growth and survival of previously infrequent species.
 Indeed, the increase in taxonomic diversity was accompanied by an increase of taxonomic dissimilarity with plots initial state and a functional shift towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
@@
 
-As stated by the IDH, communities dynamics after disturbance then rely on species functional strategy and corresponding ability to fill the environmental niches made available by disturbance.
-Recruited species then mix with pre-disturbance ones from which they differ and constitue a community all the more diversified that the disturbance is intense [@Molino2001].
+As stated by the IDH, communities dynamics after disturbance relied on species functional strategy and corresponding ability to fill the environmental niches made available by disturbance.
+Recruited species then mixed with pre-disturbance ones, from which they differed, and constituted a community all the more diversified that the disturbance was intense [@Molino2001].
 
@@
 
-The trajectories of almost all leaf and stem functional traits, life history traits, and functional diversity followed hump-back trajectories whose maximum were positively correlated to the disturbance intensity, and were recovering towards initial values after 30 years.
-Functional traits and communities functional diversity are the most direct link between biodiversity and ecosystem functioning [@Diaz2005]: these trajectories reflected a consistent recovery in the long term of ecosystem processes [@Guariguata2001].
+All diversity and composition trajectories followed similar hump-back trajectories returning towards their initial values.
+Communities trajectories therefore illustrated their taxonomic and function resilience.
 
-Same hump-back trajectories were followed by the composition of plots species and functoinal composition.
-This similarly meant a recovery of the pre-disturbance state and the maintenance of the initial differences among plots which supports the idea of a convergence of communities composition [@Hubbell1999;@Molino2001;@Baraloto2012a].
-Although plots followed homogeneous trends these initial differences in compositon difference, and probably also some abiotic parameters, strongly determined the maximum and time path of the trajectories [@Anderson2007]•.
+Because functional traits and functional diversity are the most direct link between biodiversity and ecosystem functioning [@Diaz2005], their resilience meant a consistent recovery of ecosystem processes in the long term [@Guariguata2001].
+The resilience of the taxonomic diversity and composition similarly translated  the recovery of communities such as before disturbance.
+This resilience meant the maintenance of communities initial differences in composition and supported the idea of their convergence towards a determined state [@Hubbell1999;@Molino2001;@Baraloto2012a].
+If all plots followed similar hump-back trajectories, though, some significant differences emerged.
+To name but one, the taxonomic evenness of one plot displayed an extensive increase until 30 years after disturbance.
+The differences trajectories maxima and time path, along with the fact intrinsinc communities differences are maintained, suggested that the trajectories depend on the initial plots composition or on some abiotic parameters [@Anderson2007].
 
@@
 
-Despite the consistent recovery of initial diversity and composition there was a time lag between the trajectories followed by taxonomic and functional characteristics.
-Taxonomic composition had longer time path and while communities had recovered their functional characteristics, their taxonomic composition and evenness remained altered by disturbance.
+Despite the recovery of pre-disturbance state there was a time lag between the trajectories followed by communities taxonomic and functional characteristics.
+The trajectory of taxonomic composition and evenness had a longer time-path and remained altered while communities functional composition and diversity had already recovered.
 
-This delay between functional and taxonomic dynamics was already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
-The functional diversity of communities rely on the functional strategy of dominant species, according to the "vegetation quantity effect" [@Grime1998].
-The functional trajectories of communities are then driven first by the increase of diversity and evenness of dominant species and then by the recruitment of species resembling the old pre-disturbance community which reduced the overall diversity.
-Although dominant species restored the functional pre-disturbance characteristics, communities composition remained altered and the infrequent species of initial communities remain missing to achieve plots recovery.
-According to the functional traits trajectories, those infrequent species would match the dominant functional type and would participate to the functional redundancy.
-The functional redundancy, thas is the functional overlap between species that is typical of the huge biodiversity of tropical forests [@Bellwood2006], therefore was not restored 30 years after disturbance and this is major to consider as it defines forests' resilience [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+Such delay between functional and taxonomic dynamics was already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
+According to the "vegetation quantity effect" [@Grime1998], the functional diversity of communities rely on the functional strategy of dominant species.
+Communities functional trajectories were then driven first by the increase of the diversity and evenness of dominant species restoring the pre-disturbance functional characteristics.
+Then, the functional trajectories were driven by the recruitment of species resembling the old pre-disturbance community, which reduced the overall diversity.
+At that time, the taxonomic composition and diversity remained altered: the initially infrequent, functionally redundant species, therefore remained missing.
+The functional redundancy, the functional overlap between species that is typical of the huge biodiversity of tropical forests [@Bellwood2006], was then not restored 30 years after disturbance and this is major to consider as it defines forests' resilience [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
@@
 
-Our study showed the significant impact of disturbance on tropical forests communities.
-It validated the consistency of the IDH in the long term and revealed the contrasting response of taxonomic and functional characteristics with the decoupling between communities taxonomic evenness and their functional diversity and th dominant functional strategies.
-Communities recovery remained unachieved but consistent recovery of communities assembly for the lowest disturbance intensity but questioned it after higher disturbance.
+Our study showed the significant impact of disturbance on tropical forests communities and validated the consistency of the IDH in the long term.
+It revealed the contrasting response of communities taxonomic and functional characteristics, with persisting impacts on the species abundance distibution while the functional diversity and dominant functional strategies were restored.
+Communities recovery therefore remained unachieved but consistent for the range of disturbance studied here.
+The length of the recovery, however, severly questioned the sustainability of intense selective logging and advocated felling cycle much longer than 30 years [@Gourlet-Fleury2005].
 

### 2018-05-03 WholePlotTrajectories.Rmd
Discussion

@@
 Such delay between functional and taxonomic dynamics was already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
-According to the "vegetation quantity effect" [@Grime1998], the functional diversity of communities rely on the functional strategy of dominant species.
-Communities functional trajectories were then driven first by the increase of the diversity and evenness of dominant species restoring the pre-disturbance functional characteristics.
+According to the "vegetation quantity effect" [@Grime1998], the functional diversity of communities rely on the dominant species.
+Communities functional trajectories were then driven first by the increase in diversity and evenness of dominant species, which restored the pre-disturbance functional characteristics.
 Then, the functional trajectories were driven by the recruitment of species resembling the old pre-disturbance community, which reduced the overall diversity.
-At that time, the taxonomic composition and diversity remained altered: the initially infrequent, functionally redundant species, therefore remained missing.
+At that time, the taxonomic composition and diversity remained altered: the still missing species were then the initially infrequent and functionally redundant ones.
 The functional redundancy, the functional overlap between species that is typical of the huge biodiversity of tropical forests [@Bellwood2006], was then not restored 30 years after disturbance and this is major to consider as it defines forests' resilience [@Trenbath1999;@Elmqvist2003;@Diaz2005].

### 2018-05-04 WholePlotTrajectories.Rmd
Update WholePlotTrajectories.Rmd

@@
 The large areas covered with tropical forests worldwide hold crucial economic, social and cultural value. 
-They provide wood and multiple non-timber forest products, shelter for a diversified fauna, regulate the local climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The simultaneous increase of forests products demand and substantial climatic changes heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threatened the maintenance of communities structure, composition and functioning and their underlying dynamics [@Anderson-Teixeira2013;@Sist2015].
+They provide wood and multiple non-timber forest products, shelter for a diversified fauna, regulate the local and global climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being.
+The simultaneous increase of forests products demand and substantial climatic changes has heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threatened the maintenance of communities structure, composition and functioning and their underlying dynamics in space and time [@Anderson-Teixeira2013;@Sist2015].
 
-In tropical forests communities are shaped by a constant range of disturbance that changes the abiotic environment, as the light, the heat and water fluxes, and the biotic interaction and competitive pressure.
+In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events that changes the abiotic environment, as the light, the heat and water fluxes, and the biotic interaction and competitive pressure *(Goulamoussene Biogeosciences)*.
 The cornerstone of tropical forests ecology is then to understand their response to disturbance and its mechanisms [@White2001;@Chazdon2003a].

### 2018-05-04 WholePlotTrajectories.Rmd
intro OK

@@
 In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events that changes the abiotic environment, as the light, the heat and water fluxes, and the biotic interaction and competitive pressure *(Goulamoussene Biogeosciences)*.
-The cornerstone of tropical forests ecology is then to understand their response to disturbance and its mechanisms [@White2001;@Chazdon2003a].
-Forests response has been largely studied through structural parameters rapid and convenient to measure, as aboveground biomass, tree height or stem density.
-These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
-However the response of tree species diversity, which is determinant of ecosystems productivity, stability and functioning [@Tilman2014] and would be most probably impacted by post-disturbance environmental changes [@CazzollaGatti2014], remains unclear.
+The cornerstone of tropical forests ecology is then to understand the underlyings of ecosystem response to disturbance, its mechanisms and drivers [@White2001;@Chazdon2003a].
+For now, this has been largely studied through structural parameters rapid and convenient to measure, as aboveground biomass, tree height or stem density *(e.g.
+PiponiotElife, RutishauserBiotropica)*.
+These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Herault2018].
+However the response of tree species diversity, which is determinant of ecosystems productivity, stability and functioning [@Tilman2014, Liang2016Science] and would be most probably impacted by post-disturbance environmental changes [@BaralotoJAE], remains unclear.
 
 In the short-term diversity dynamics demonstrated negligible or even positive impacts of disturbance on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
-However, validations of the IDH in the long term remain scarce and mainly rely on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and evenness, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, the functional approach accounting for species biological attributes would be insightful as it reveals species fitness and ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-In that respect major functional traits related to species ecology and performance were largely adopted through a relevant framework [@Diaz2005;@Villeger2008a].
-The functional trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
-This translated by shifts from "conservative" slow-growing species dealing with scarce resources, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
-It was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
+However, validations of the IDH in the long term remain scarce and mainly rely on the analysis of the rough species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and diversity profiles, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Furthermore, the functional approach accounting for differenes in species biological attributes and functioning would be insightful as it reveals the species role in the post-disturbance trajectory of ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+In that respect major functional traits related to species ecology and performance were largely adopted through an integrative framework [@Diaz2016Nature, @Kunstler2016Nature].
+In disturbed tropical rainforests, this framework translates into the fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008] with a shift from "conservative" slow-growing species dealing with scarce resources that dominate before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011] and was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
+
 A proper monitoring of communities response should therefore encompass taxonomic and functional diversity and composition measures to test the validity of the IDH in the long term for tropical forests, and clarify the resilience of communities evenness, composition, and also functioning.
-The trajectories followed by all these facets would highlight the role of deterministic processes, like competitive exclusion or abiotic selection, and the communities' convergence maintaining intrinsinc differences in diversity and composition,which is as much insgights for future adaptive conservation strategies [@Adler2007].
-
+The trajectories followed by all these facets would highlight the role of deterministic processes, like competitive exclusion or abiotic selection, and the communities' convergence maintaining intrinsinc differences in diversity and composition, which is as much insights for future adaptive management strategies [@Adler2007].
 Here we investigated over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
 We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
-Specifically, we (i) tested the validity of the IDH in the long term for tropical hiperdiverse forest and highlighted the ecological rules shaping their response to disturbance, (ii) clarified the different facets of communities resilience in terms of communities composition, diversity and functioning (iii) questioned the completeness of communities recovery given the altered functional redundancy.
->>revoir la fin
+Specifically, we (i) tested the validity of the IDH in the long term for tropical hyperdiverse forest and highlighted the ecological rules shaping their response to disturbance, (ii) clarified the different facets of communities resilience in terms of communities composition, diversity and functioning *????(iii) questioned the completeness of communities recovery given the altered functional redundancy.*
 

### 2018-05-04 AppendixI.Rmd
appendix & graphs

@@
 Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.", fig.env="figure*"}
-load("DB/FunctionalTraj_ForGraphsnew")
+load("DB/FunctionalTraj_ForGraphs")
 

### 2018-05-04 WholePlotTrajectories.Rmd
appendix & graphs

--- 2018-05-03 WholePlotTrajectories.Rmd
@@
              "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-valuable species, $\\approx$ 15 trees/ha"),
-  AGBLost=c(0, "$[12\\%-33\\%]$", "$[33\\%-56\\%]$", "$[35\\%-56\\%]$")
+  AGBLost=c(0, "$[12\\%-33\\%]$", "$[33\\%-56\\%]$", "$[35\\%-66\\%]$")
 )
@@
 
-```{r DivFun, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
+```{r DivFun, out.width = '100%', echo=FALSE, fig.cap="Upper panels, Trajectories of the Simpson taxonomic diversity and Rao functional diversity over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure"}
-load("DB/FunctionalTraj_ForGraphs")
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.
+Lower panels, Relationship between the initial %AGB removed and the values of Simpson (c) and Rao (d) diversities at three times after disturbance.
+Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.", fig.env="figure"}
+load("DB/SimpsonDraw");load("DB/FunctionalTraj_ForGraphs")
+load("DB/SimpsonIDH");load("DB/RaoIDH")
 
-FunTraj(CompleteFun,remove=T)
+par(mfrow=c(2,2),mar=c(3,3,2,1),oma=c(2,1,2,3),no.readonly = T)
+plotDiv(SimpsonTaxo)
+mtext("Simpson diversity",side=3,adj=0,line=1)
+plotDiv(CompleteFun,remove=TRUE)
+mtext("Rao diversity",side=3,adj=0,line=1)
+mtext("Years since disturbance",side=1,line=2.2,adj=1)
+legend("right",inset=c(-0.28,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
+
+plotIDH(Sim)
+plotIDH(Rao)
+mtext("initial %AGB lost",side=1,line=2.2,adj=1)
+legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
 

### 2018-05-04 Scripts/DrawingTools.R

@@ -4,6 +4,9 @@ Library(c("shape"))
 treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
 names(treatments)<-c("Control","T1","T2","T3")
 ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
+treatAGB<-c(0,18,45,50)
+colyear<-c("darkgoldenrod1","darkorange2","darkred")
+time<-c("1995","2005","2015")
 
 smooth<-function(mat,larg){return(do.call(cbind,lapply(1:ncol(mat),function(step){
   range<-max(1,step-larg):min(ncol(mat),step+larg)
@@ -138,6 +141,109 @@ FunDist<-function(Data_FunComp){
   mtext("Years since disturbance",side=1,adj=1,line=2,cex=0.8)
 }
 
+plotIDH<-function(Data){
+  plot(treatAGB,treatAGB,type="n",xlab="",ylab="",
+       ylim=c(min(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))),
+              max(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])})))))
+  invisible(lapply(1:3,function(T){
+    toplot<-lapply(Data, function(tr){return(tr[,time[T],"0.5"])})
+    toplot<-do.call(rbind,lapply(1:length(toplot),function(tr){return(cbind(treatAGB[tr],toplot[[tr]]))}))
+    toplot<-apply(toplot,2,as.numeric)
+    points(toplot[,1],toplot[,2],col=colyear[T])
+    Lm<-lm(toplot[,2]~toplot[,1])
+    abline(a=Lm$coefficients[1],b=Lm$coefficients[2],col=colyear[T],lwd=2.5)
+  }))
+}
+
+plotDiv<-function(Data,remove=FALSE){
+  
+  if(remove){Data[[2]]<-Data[[2]][which(rownames(Data[[2]])!=7),,]}
+  
+  Toplot<-lapply(Data,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
+  Toplot<-lapply(Toplot,function(tr){
+    ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col}))})#-tr[,1,rep]
[…]

### 2018-05-04 Scripts/Stats.R

@@ -1,3 +1,41 @@
+load("DB/ReplacementTraj_ForGraphs")
+load("DB/FunctionalTraj_ForGraphs")
+
+SimpsonTaxo<-lapply(CompleteTaxo,function(tr){return(tr[,,,"Simpson"])})
+save(SimpsonTaxo,file="DB/SimpsonDraw")
+
+Sim<-lapply(CompTaxo,function(tr){return(tr[,c("1995","2005","2015"),])})
+Sim<-lapply(Sim, function(tr){
+  Ret<-lapply(c(0.025,0.5,0.975),function(quant){
+  return(apply(tr,c(1,2),function(x){return(quantile(x,probs=quant))}))})
+  Ret<-array(unlist(Ret),dim=c(nrow(Ret[[1]]),ncol(Ret[[1]]),3),
+             dimnames=list(rownames(Ret[[1]]),colnames(Ret[[1]]),c(0.025,0.5,0.975)))})
+save(Sim,file="DB/SimpsonIDH")
+
+Rao<-lapply(CompFun, function(tr){
+  Ret<-lapply(c(0.025,0.5,0.975),function(quant){
+    return(apply(tr[,c("1995","2005","2015"),],c(1,2),function(x){return(quantile(x,probs=quant))}))})
+  Ret<-array(unlist(Ret),dim=c(nrow(Ret[[1]]),ncol(Ret[[1]]),3),
+             dimnames=list(rownames(Ret[[1]]),colnames(Ret[[1]]),c(0.025,0.5,0.975)))})
+save(Rao,file="DB/RaoIDH")
+
+windows()
+par(mfrow=c(2,2),mar=c(3,3,2,1),oma=c(2,1,2,3),no.readonly = T)
+plotDiv(SimpsonTaxo)
+mtext("Simpson diversity",side=3,adj=0,line=1)
+plotDiv(CompleteFun,remove=TRUE)
+mtext("Rao diversity",side=3,adj=0,line=1)
+mtext("Years since disturbance",side=1,line=2.2,adj=1)
+legend("right",inset=c(-0.28,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
+
+plotIDH(Sim)
+plotIDH(Rao)
+mtext("initial %AGB lost",side=1,line=2.2,adj=1)
[…]

### 2018-05-09 WholePlotTrajectories.Rmd
Redundancy etc

@@
 
+```{r RedFun, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure"}
+load("DB/Redundancy")
+
+RedundancyPlot(RedundancyTraj)
+mtext("Functional Redundancy",side=3,adj=0,line=1.5)
+mtext("Sum of abundance-weighted Trait Probability Density",side=3,adj=0,cex=0.8,line=0.5)
+mtext("Years since disturbance",side=1,line=2.2,adj=1)
+
+```
+
 # Discussion

### 2018-05-09 WholePlotTrajectories.tex
Redundancy etc

@@
 T2 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30 trees/ha &  & $[33\%-56\%]$\\
-T3 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & $[35\%-56\%]$\\
+T3 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & $[35\%-66\%]$\\
 \bottomrule
@@
 
-To counter the remaining taxonomic uncertainty plots taxonomic
+To counter the remaining taxonomic uncertainty, plots taxonomic
 composition and diversity were analysed at the genus level, \emph{i.e.}
@@
 difference among plots the trajectories corresponded to the differences
-along time with the reference inventory in 1984 (5 year after
-disturbance). Species seed mass corresponded to 5 classes of increasing
-mass, seed mass trajectories were therefore reported as the proportion
-of each class in the inventories.
+along time with the reference inventory in 1989. Species seed mass
+corresponded to 5 classes of increasing mass, seed mass trajectories
+were therefore reported as the proportion of each class in the
+inventories.
 
@@
 genus and 64 botanical families were recorded. Taxonomic and functional
-plots trajectories were examined through the ordination in a two
-dimensional ordiantion space of successive inventories from 1989 (5
+plots trajectories were examined through the trajecotries in a two
+dimensional ordination space of successive inventories from 1989 (5
 years after disturbance) to 2015 (31 years after disturbance).
@@
 with consistent trajectories of the 8 functional and life history traits
-(Leaf thickness, chlorophyll content, toughness and specific area,wood
-specific gravity and barck thickness and seed mass and maximum adult
-height) visaualized with plots community weighted means (CWM) (Figure
+(Leaf thickness, chlorophyll content, toughness and specific area, wood
+specific gravity and bark thickness and seed mass and maximum adult
+height) visualized with plots community weighted means (CWM) (Figure
 \ref{fig:CWM}).
@@
 again the maximum difference to initial state was correlated to the
-disturbance intensity (\(\rho_{spearman}^{L_thickness}=0.67\),
-\(\rho_{spearman}^{L_chloro}=0.45\),
-\(\rho_{spearman}^{L_toughness}=-0.43\), \(\rho_{spearman}^{SLA}=0.93\),
-\(\rho_{spearman}^{WD}=-0.78\),
+disturbance intensity (\(\rho_{spearman}^{L_{thickness}}=0.67\),
+\(\rho_{spearman}^{L_{chloro}}=0.45\),
+\(\rho_{spearman}^{L_{toughness}}=-0.43\),
+\(\rho_{spearman}^{SLA}=0.93\), \(\rho_{spearman}^{WD}=-0.78\),
 \(\rho_{spearman}^{Bark-thickness}=0.88\),
@@
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.}\label{fig:CWM}
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:CWM}
 \end{figure*}
@@
 
-For undisturbed plots the Richness, Shannon and Simpson diversity
+For undisturbed plots the Richness, Shannon and Simpson diversities
 remained stable over the 30 monitored years. After disturbance the
 richness increased when the disturbance intensity was low, with a
-maximum increase of 14 botanical genuses (plot 3 from treatment 2). When
-the disturbance was intense however, plot staxonomic richness followed a
+maximum increase of 14 botanical genera (plot 3 from treatment 2). When
+the disturbance was intense however, plots taxonomic richness followed a
 unimodal decrease with a return to initial values after intense
 disturbance. The taxonomic evenness (Shannon and Simpson diversities)
-significantly increased whenever the disturbance intensity. The evenness
-followed a unimodal trajectory with a just beginning return towards
-initial values and a maximum reached around 20 years and positively
-correlated to the disturbance intensity
+significantly increased irrespective of the disturbance intensity. The
+evenness followed a unimodal trajectory with a just beginning return
+towards initial values and a maximum reached around 20 years and
+positively correlated to the disturbance intensity
 (\(\rho_{spearman}^{Shannon}=0.86\), and
@@
 initial disturbance intensity, regained diversity values similar to
-their initial avule and to those of control plots.
+their initial value and to those of control plots.
+
+\textbackslash{}begin\{figure\}
+
+\{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivFun-1}
+
+\}
+
+\textbackslash{}caption\{Upper panels, Trajectories of the Simpson
+taxonomic diversity and Rao functional diversity over 30 years after
+disturbance. Trajectories correspond to the median (solid line) and
+0.025 and 0.975 percentile (gray envelope) observed after 50 iteration
+of the taxonomic uncertainty propagation and the missing trait value
+filling processes. Initial treatments are represented by solid lines
+colors with green for control, blue for T1,orange for T2 and red for T3.
+Lower panels, Relationship between the initial \%AGB removed and the
+values of Simpson (c) and Rao (d) diversities at three times after
+disturbance. Solid lines colors represent the time, 10 years (yellow),
+20 years (orange) and 30 years (brown) after
+disturbance.\}\label{fig:DivFun} \textbackslash{}end\{figure\}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivFun-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/RedFun-1}
 
@@
 
-\caption{Trajectories of the Rao functional diversity over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:DivFun}
+\caption{Trajectories of the functional redundancy over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:RedFun}
 \end{figure}
@@
 values after high disturbance intensity while it substantially increased
-after low disturbance intensity, reaching a gain of almost 12 genuses
-for some plots.
+after low disturbance intensity, reaching a gain of almost 12 genera for
+some plots.
 
-Both richness and evenness after disturbance followed asymptotic
-trajectories, sharply increasing for 15 years before stabilizing at
+Both richness and evenness followed asymptotic trajectories after
+disturbance, sharply increasing for 15 years before stabilizing at
 higher values than those of control plots. An increase of communities
@@
 distribution are to be found in the recruitment processes and in the
-pre-disturbance survivors mortality. The composition of old survivors
-proved to mirror the initial community \citep{Herault2018}, which leads
-to support a species turnover among recruited trees with enhanced growth
-and survival of previously infrequent species. Indeed, the increase in
-taxonomic diversity was accompanied by an increase of taxonomic
-dissimilarity with plots initial state and a functional shift towards
-resource-acquisitive strategies (sharp increase in the SLA, leaf
+pre-disturbance survivors mortality. Because the composition of old
+survivors proved to mirror the initial community \citep{Herault2018}, a
+specific turnover was expected among recruited trees with enhanced
+growth and survival of previously infrequent species. Indeed, the
+increase in taxonomic diversity was accompanied by an increase of
+taxonomic dissimilarity with plots initial state and a functional shift
+towards resource-acquisitive strategies (sharp increase in the SLA, leaf
 thickness and bark thickness and decrease in wood density, leaf
@@
 
-As stated by the IDH, communities dynamics after disturbance then rely
-on species functional strategy and corresponding ability to fill the
+As stated by the IDH, communities dynamics after disturbance relied on
+species functional strategy and corresponding ability to fill the
 environmental niches made available by disturbance. Recruited species
-then mix with pre-disturbance ones from which they differ and constitue
-a community all the more diversified that the disturbance is intense
-\citep{Molino2001}.
+then mixed with pre-disturbance ones, from which they differed, and
+constituted a community all the more diversified that the disturbance
+was intense \citep{Molino2001}.
 
@@
 
-The trajectories of almost all leaf and stem functional traits, life
-history traits,and functional diversity followed a hump-back trajectory
-which maximum were positively correlated to the disturbance intensity,
-and were recovering towards initial values after 30 years that
-trajectory. Functoinal traits and communities functional diversity are
-the most direct link between biodiversity and ecosystem functioning
-\citep{Diaz2005}: these trajectories reflected a consistent recovery in
-the long term of ecosystem processes \citep{Guariguata2001}.
+All diversity and composition trajectories followed similar hump-back
+trajectories returning towards their initial values. Communities
+trajectories therefore illustrated their taxonomic and function
+resilience.
 
-Same hump-back trajectories were followed by the composition of plots
-species and functoinal composition. This similarly meant a recovery of
-the pre-disturbance state and the maintenance of the initial differences
-among plots which supports the idea of a convergence of communities
-composition \citep{Hubbell1999, Molino2001, Baraloto2012a}. Although
-plots followed homogeneous trends these initial differences in
-compositon difference, and probably also some abiotic parameters,
-strongly determined the maximum and time path of the trajectories
-\citep{Anderson2007}•.
+Because functional traits and functional diversity are the most direct
+link between biodiversity and ecosystem functioning \citep{Diaz2005},
+their resilience meant a consistent recovery of ecosystem processes in
+the long term \citep{Guariguata2001}. The resilience of the taxonomic
+diversity and composition similarly translated the recovery of
+communities such as before disturbance. This resilience meant the
+maintenance of communities initial differences in composition and
+supported the idea of their convergence towards a determined state
+\citep{Hubbell1999, Molino2001, Baraloto2012a}. If all plots followed
+similar hump-back trajectories, though, some significant differences
+emerged. To name but one, the taxonomic evenness of one plot displayed
+an extensive increase until 30 years after disturbance. The differences
+trajectories maxima and time path, along with the fact intrinsinc
+communities differences are maintained, suggested that the trajectories
+depend on the initial plots composition or on some abiotic parameters
+\citep{Anderson2007}.
 
@@
 
-Despite the consistent recovery of initial diversity and composition
-there was a time lag between the trajectories followed by taxonomic and
-functional characteristics. Taxonomic composition had longer time path
-and while communities had recovered their functional characteristics,
-their taxonomic composition and evenness remained altered by
-disturbance. This delay between functional and taxonomic dynamics was
-already observed for grasslands \citep{Tilman1997, Mouillot2011} and
-more recently for tropical forests \citep{Lohbeck2015, Guariguata2001}.
-The functional diversity of communities rely on the functional strategy
-of dominant species, according to the ``vegetation quantity effect''
-\citep{Grime1998}. The functional trajectories of communities are then
-driven first by the increase of diversity and evenness of dominant
-species and then by the recruitment of species resembling the old
-pre-disturbance community which reduced the overall diversity. Although
-dominant species restoredthe functional pre-disturbance characteristics,
-communities composition remained altered and the infrequent species of
-initial communities remain missing to achieve plots recovery. According
-to the funcitonal traits trajectories, those infrequent species would
-match the dominant functional type and would participate to the
-functional redundancy. The functional redundancy, thas is the functional
+Despite the recovery of pre-disturbance state there was a time lag
+between the trajectories followed by communities taxonomic and
+functional characteristics. The trajectory of taxonomic composition and
+evenness had a longer time-path and remained altered while communities
+functional composition and diversity had already recovered.
+
+Such delay between functional and taxonomic dynamics was already
+observed for grasslands \citep{Tilman1997, Mouillot2011} and more
+recently for tropical forests \citep{Lohbeck2015, Guariguata2001}.
+According to the ``vegetation quantity effect'' \citep{Grime1998}, the
+functional diversity of communities rely on the dominant species.
+Communities functional trajectories were then driven first by the
+increase in diversity and evenness of dominant species, which restored
+the pre-disturbance functional characteristics. Then, the functional
+trajectories were driven by the recruitment of species resembling the
+old pre-disturbance community, which reduced the overall diversity. At
+that time, the taxonomic composition and diversity remained altered: the
+still missing species were then the initially infrequent and
+functionally redundant ones. The functional redundancy, the functional
 overlap between species that is typical of the huge biodiversity of
-tropical forests \citep{Bellwood2006}, therefore was not restored 30
-years after disturbance and this is majr to consider as it defines
-forests' resilience \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+tropical forests \citep{Bellwood2006}, was then not restored 30 years
+after disturbance and this is major to consider as it defines forests'
+resilience \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-Besides the long-term alteration of functional redundancy, the
-persistent compositional changes which might have favored disturbance
-resistant species \citep{Haddad2008}, lianas or epiphytes
-\citep{Martin2013} and would probably have come along with environmental
-change, like in the soils nutrient cycling and compaction
+Besides the long-term alteration of functional redundancy, there was
+probably persistent compositional changes favoring disturbance resistant
+species \citep{Haddad2008}, lianas or epiphytes \citep{Martin2013} and
+environmental changes, like in the soils nutrient cycling and compaction
 \citep{Olander2005}. These presistent changes highly question forest's
-resilience\citep{Chazdon2003a}. New conditions would not only be longer
+resilience \citep{Chazdon2003a}. New conditions would not only be longer
 lasting but self-maintained as tied to disturbance regime
@@
 Our study showed the significant impact of disturbance on tropical
-forests communities. It validated the consistency of the IDH in the long
-term and revealed the contrasting response of taxonomic and functional
-characteristics with the decoupling between communities taxonomic
-evenness and their functional diversity and th dominant functional
-strategies. Communities recovery remained unachieved but consistent
-recovery of communities assembly for the lowest disturbance intensity
-but questioned it after higher disturbance.
+forests communities and validated the consistency of the IDH in the long
+term. It revealed the contrasting response of communities taxonomic and
+functional characteristics, with persisting impacts on the species
+abundance distibution while the functional diversity and dominant
+functional strategies were restored. Communities recovery therefore
+remained unachieved but consistent for the range of disturbance studied
+here. The length of the recovery, however, severly questioned the
+sustainability of intense selective logging and advocated felling cycle
+much longer than 30 years \citep{Gourlet-Fleury2005}.
 

### 2018-05-09 Scripts/AGBloss.R

@@ -1,0 +1,182 @@
+#### get data damage for Tortue and Manaré, using Guyafor
+#############################################################################
+
+#  This calculation uses the methods of Camilia but with the allometric equation fo Rejou-Mechain 2017 for AGB
+# and the tarif de cubage de l'Est for vol.ext
+# The methods calculate the ABG before logging and for all inventory after logging 
+# then dAGB = AGB initiale - AGB minimum i nthe 4 year after logging
+# AGB damage = dAGB - AGB extracted
+# uses a C content in AGB of 0.5
+# and average wd bellow
+
+### open libray and set values
+library(data.table)
+library(BIOMASS)
+
+
+###connection odbc a Guyafor sur serveur SQL
+library("RODBC")
+Connex=odbcConnect(dsn="Guyafor")
+####s?lection des donn?es
+req1<-"SELECT 
+dbo.TtGuyaforShiny.NomForet AS Forest, 
+dbo.TtGuyaforShiny.n_parcelle AS Plot, 
+dbo.TtGuyaforShiny.Surface AS PlotSurface, 
+dbo.TtGuyaforShiny.n_carre AS SubPlot, 
+dbo.TtGuyaforShiny.n_arbre AS TreeFieldNum, 
+dbo.TtGuyaforShiny.i_arbre AS idTree, 
+dbo.TtGuyaforShiny.X AS Xfield, 
+dbo.TtGuyaforShiny.Y AS Yfield, 
+dbo.TtGuyaforShiny.Xutm, 
+dbo.TtGuyaforShiny.Yutm, 
+dbo.TtGuyaforShiny.UTMZone, 
+dbo.TtGuyaforShiny.Lat, 
+dbo.TtGuyaforShiny.Lon, 
+dbo.TtGuyaforShiny.Famille AS Family, 
+dbo.TtGuyaforShiny.Genre AS Genus, 
+dbo.TtGuyaforShiny.Espece AS Species, 
+dbo.TtGuyaforShiny.SourceBota AS BotaSource, 
+dbo.TtGuyaforShiny.indSurete AS BotaCertainty, 
+dbo.TtGuyaforShiny.n_essence, 
+dbo.TtGuyaforShiny.idPilote AS idVern, 
+dbo.TtGuyaforShiny.nomPilote AS VernName, 
+dbo.TtGuyaforShiny.Commerciale AS CommercialSp, 
[…]

### 2018-05-09 Scripts/DrawingTools.R

@@ -179,7 +179,7 @@ plotDiv<-function(Data,remove=FALSE){
       absc<-colnames(Toplot[[t]])
       lines(absc,Toplot[[t]][i,,"0.5"], col = ColorsTr[[t]],lty = 1,lwd=2)
       polygon(c(absc,rev(absc)),c(Toplot[[t]][i,,"0.025"],rev(Toplot[[t]][i,,"0.975"])),
-              col=rgb(0,0,0,alpha=0.1),border=NA)
+              col=rgb(0,0,0,alpha=0.05),border=NA)
     }))
   }))
 }
@@ -241,6 +241,27 @@ SeedMassProp<-function(SeedMass){
 mtext("Proportion of seed mass classes",line=2,adj=0,outer=TRUE,cex=0.9)
 mtext("Years since disturbance",side=1,line=1.2,adj=1,cex=0.9,outer=TRUE)
 }
+
+RedundancyPlot<-function(Red){
+  abs<-as.numeric(colnames(Red[[1]]))-1984
+  Red<-lapply(Red,function(rep){return(smooth(rep,2))})
+  Red<-lapply(Red,function(rep){return(apply(rep,2,function(col){return(col-rep[,1])}))})
+  Red<-array(unlist(Red),dim=c(12,ncol(Red[[1]]),length(Red)),
+             dimnames=list(1:12,colnames(Red[[1]]),1:length(Red)))
+  Red<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(Red,c(1,2),function(rep){return(quantile(rep,probs=quant))}))})
+  Red<-array(unlist(Red),dim=c(12,ncol(Red[[1]]),3),
+             dimnames=list(1:12,colnames(Red[[1]]),c(0.025,0.5,0.975)))
+  #Red<-Red[,which(colnames(Red)>=1989),]
+  
+  plot(abs,Red[1,,1],type='n',ylim=c(min(Red),max(Red)),xlab="",ylab="")
+  invisible(lapply(1:4,function(tr){
+    toplot<-Red[which(rownames(Red)%in%treatments[[tr]]),,"0.5"]
+    apply(toplot,1,function(li){
+      lines(abs,li,col=ColorsTr[tr],lwd=2)
+    })
+  }))
+}
+
 
 
 ### Previous graphs

### 2018-05-09 Scripts/Functionalredundancy.R

@@ -1,0 +1,190 @@
+
+source("Scripts/Vernacular_handle.R")           # Script for vernacular names replacement
+source("Scripts/TraitsMiceFilling.R")                                   # Traits Gap-filling
+
+load("DB/Alpha_Plots")                  # Alpha vector for vernacular replacement
+load("DB/Paracou_R_Subdivided_ok")      # Paracou dataset, lists by inventoried year, sublists by plot, etc...
+dates<-sort(names(LivingStand_all))[-1] # dates considered
+
+Traits1<-read.csv("DB/BridgeOK.csv",sep=";",na.strings="")       # Brigde functional traits dataset
+Traits2<-read.csv("DB/DataLifeTraits.csv",sep=";",na.strings="") # Paracou15 seed and Hmax dataset
+TraitsName<-c("L_thickness","L_chloro","L_toughness","L_DryMass","SLA","WD","Bark_thick","Hmax") # Considered traits list
+
+### Whole community mapping, individuals: NMDS ou ACP?
+
+traits_filled<-Traits_filling(Traits1,Traits2,InventorySp)
+BigNmds<-as.matrix(daisy(scale(traits_filled[,TraitsName]),metric="gower"))
+BigNmds<-scores(metaMDS(BigNmds,distance="bray"),display="sites")
+BigNmds<-merge(BigNmds,traits_filled[,c("Family","Genus","name" )],by="row.names")[,c("Family","Genus","name","NMDS1","NMDS2")]
+save(BigNmds,file="DB/BigNmds")
+
+ACP<-dudi.pca(scale(traits_filled[,TraitsName]))
+Bigacp<-merge(ACP$li,traits_filled[,c("Family","Genus","name" )],by="row.names")[,c("Family","Genus","name","Axis1","Axis2")]
+
+plot(Bigacp[,"Axis1"], Bigacp[,"Axis2"],type="n")
+text(Bigacp[,"Axis1"], Bigacp[,"Axis2"],labels=Bigacp[,"name"])
+
+### Species TPD
+
+xgen<-c(min(Bigacp[,"Axis1"]),max(Bigacp[,"Axis1"]))
+ygen<-c(min(Bigacp[,"Axis2"]),max(Bigacp[,"Axis2"]))
+  
[…]

### 2018-05-09 WholePlotTrajectories.Rmd
M&M

@@
 The taxonomic diversity was assessed through Richness and the Hill number translation of Shannon and Simpson indices [@Hill1973].
-
 These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q), which proved well suited for diversity studies [@Patil1982;@Tothmeresz1995].
 The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on all functional and life traits.
+
+The functional redundancy was measured as the overlap among species in communities' functional space.
+The samples of the trait database were first located with a PCA analysis in a two-dimensional functional space summarizing the 8 functional traits considered.
+Then, multivariate kernel density estimator associated with individual trees were summed to give the distribution of traits probabilites of each species.
+For each community the trait probability distributions of corresponding species were combined and weighted by species abundance. 
+Eventually communities functional redundancy was measured as the sum of overlap between species weighted functional density.
 

### 2018-05-09 WholePlotTrajectories.Rmd
Ref
Graphs, precise AGB loss

@@
 
-The functional redundancy was measured as the overlap among species in communities' functional space.
+The functional redundancy was measured as the overlap among species in communities' functional space [@Carmona2016].
 The samples of the trait database were first located with a PCA analysis in a two-dimensional functional space summarizing the 8 functional traits considered.
@@
 load("DB/SimpsonDraw");load("DB/FunctionalTraj_ForGraphs")
-load("DB/SimpsonIDH");load("DB/RaoIDH")
+load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")
 
@@
 
-plotIDH(Sim)
-plotIDH(Rao)
+plotIDH(Data=Sim,AgbLoss=AGBloss)
+plotIDH(Data=Rao,AgbLoss=AGBloss)
 mtext("initial %AGB lost",side=1,line=2.2,adj=1)

### 2018-05-09 WholePlotTrajectories.tex
Graphs, precise AGB loss

@@
 number translation of Shannon and Simpson indices \citep{Hill1973}.
-
 These three indices belong to the set of HCDT or generalized entropy,
@@
 functional and life traits.
+
+The functional redundancy was measured as the overlap among species in
+communities' functional space \citep{Carmona2016}. The samples of the
+trait database were first located with a PCA analysis in a
+two-dimensional functional space summarizing the 8 functional traits
+considered. Then, multivariate kernel density estimator associated with
+individual trees were summed to give the distribution of traits
+probabilites of each species. For each community the trait probability
+distributions of corresponding species were combined and weighted by
+species abundance. Eventually communities functional redundancy was
+measured as the sum of overlap between species weighted functional
+density.
 

### 2018-05-09 Scripts/AGBloss.R

@@ -1,3 +1,8 @@
+load("DB/AGBlostDB")
+
+init<-AllYears[which(AllYears[,"campagne"]==1984),]
+logg<-AllYears[which(AllYears[,"campagne"]==1988),]
+
 #### get data damage for Tortue and Manaré, using Guyafor
 #############################################################################
 
@@ -13,170 +18,29 @@
 library(data.table)
 library(BIOMASS)
 
-
-###connection odbc a Guyafor sur serveur SQL
-library("RODBC")
-Connex=odbcConnect(dsn="Guyafor")
-####s?lection des donn?es
-req1<-"SELECT 
-dbo.TtGuyaforShiny.NomForet AS Forest, 
-dbo.TtGuyaforShiny.n_parcelle AS Plot, 
-dbo.TtGuyaforShiny.Surface AS PlotSurface, 
-dbo.TtGuyaforShiny.n_carre AS SubPlot, 
-dbo.TtGuyaforShiny.n_arbre AS TreeFieldNum, 
-dbo.TtGuyaforShiny.i_arbre AS idTree, 
-dbo.TtGuyaforShiny.X AS Xfield, 
-dbo.TtGuyaforShiny.Y AS Yfield, 
-dbo.TtGuyaforShiny.Xutm, 
-dbo.TtGuyaforShiny.Yutm, 
-dbo.TtGuyaforShiny.UTMZone, 
-dbo.TtGuyaforShiny.Lat, 
-dbo.TtGuyaforShiny.Lon, 
-dbo.TtGuyaforShiny.Famille AS Family, 
-dbo.TtGuyaforShiny.Genre AS Genus, 
-dbo.TtGuyaforShiny.Espece AS Species, 
-dbo.TtGuyaforShiny.SourceBota AS BotaSource, 
-dbo.TtGuyaforShiny.indSurete AS BotaCertainty, 
-dbo.TtGuyaforShiny.n_essence, 
-dbo.TtGuyaforShiny.idPilote AS idVern, 
-dbo.TtGuyaforShiny.nomPilote AS VernName, 
-dbo.TtGuyaforShiny.Commerciale AS CommercialSp, 
-dbo.TtGuyaforShiny.Densite AS WoodDensity, 
-dbo.TtGuyaforShiny.campagne AS CensusYear, 
-dbo.TtGuyaforShiny.DateMesure AS CensusDate, 
[…]

### 2018-05-09 Scripts/DrawingTools.R

@@ -4,7 +4,6 @@ Library(c("shape"))
 treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
 names(treatments)<-c("Control","T1","T2","T3")
 ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
-treatAGB<-c(0,18,45,50)
 colyear<-c("darkgoldenrod1","darkorange2","darkred")
 time<-c("1995","2005","2015")
 
@@ -141,16 +140,15 @@ FunDist<-function(Data_FunComp){
   mtext("Years since disturbance",side=1,adj=1,line=2,cex=0.8)
 }
 
-plotIDH<-function(Data){
-  plot(treatAGB,treatAGB,type="n",xlab="",ylab="",
+plotIDH<-function(Data,AgbLoss){
+  plot(AgbLoss[,"AGB"],AgbLoss[,"AGB"],type="n",xlab="",ylab="",
        ylim=c(min(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))),
               max(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])})))))
   invisible(lapply(1:3,function(T){
-    toplot<-lapply(Data, function(tr){return(tr[,time[T],"0.5"])})
-    toplot<-do.call(rbind,lapply(1:length(toplot),function(tr){return(cbind(treatAGB[tr],toplot[[tr]]))}))
-    toplot<-apply(toplot,2,as.numeric)
-    points(toplot[,1],toplot[,2],col=colyear[T])
-    Lm<-lm(toplot[,2]~toplot[,1])
+    toplot<-unlist(lapply(Data, function(tr){return(tr[,time[T],"0.5"])}))
+    abs<-AgbLoss[which(AgbLoss[,"plot"]%in%names(toplot)),"AGB"]
+    points(abs,toplot,col=colyear[T])
+    Lm<-lm(toplot~abs)
     abline(a=Lm$coefficients[1],b=Lm$coefficients[2],col=colyear[T],lwd=2.5)
   }))
 }

### 2018-05-10 WholePlotTrajectories.Rmd
figures formatting

@@
   \usepackage{tabu}
+  \usepackage{caption}
+  \captionsetup{justification = justified}
   \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
@@
 (ref:Boldc) **(c)**
+
+(ref:Boldd) **(d)**
 
@@
 
-```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
+```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 
-```{r DivFun, out.width = '100%', echo=FALSE, fig.cap="Upper panels, Trajectories of the Simpson taxonomic diversity and Rao functional diversity over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+```{r IDHplot,  echo=FALSE, fig.cap="Upper panels, Trajectories of the Simpson taxonomic diversity (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
 Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.
-Lower panels, Relationship between the initial %AGB removed and the values of Simpson (c) and Rao (d) diversities at three times after disturbance.
-Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.", fig.env="figure"}
+Lower panels, Relationship between the initial \\%AGB removed and the values of Simpson (ref:Boldc) and Rao (ref:Boldd) diversities at three times after disturbance.
+Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.", fig.env="figure*"}
+
 load("DB/SimpsonDraw");load("DB/FunctionalTraj_ForGraphs")
@@
 
+The functional redundancy was measured as the sum of weighted overlap among species in the communities.
+Species functional space were defined in the main two-dimensional space of a PCA analysis based on all leaf and stem and maximum height traits (see appendix I for more details). 
+Plots functoinal redundancy was followed in comparison to their initial value right after logging so that only communities dynamics are followed apart from the impact of logging itself.
+While the functional redundancy of control plots remained stable along the 30 years it systematically decreased after disturbance (\@ref(fig:RedFun)).
+The decrease of functional redundancy was comparable between all disturbed plots but started later for the most disturbed plots.
+Plots recovery did not seemed correlated to the initial disturbance, for five out of nine disturbed plots the initial functional redundancy was had recovered while it was still decreasing or stabilized at lower values for the fou other plots.
+
 ```{r RedFun, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy over 30 years after disturbance.
@@
 
+par(mar=c(5,4,5,2))
 RedundancyPlot(RedundancyTraj)

### 2018-05-10 WholePlotTrajectories.tex
figures formatting

@@
 \usepackage{textcomp} \DeclareUnicodeCharacter{B0}{\textdegree}
-\usepackage{tabu}
+\usepackage{tabu} \usepackage{caption}
+\captionsetup{justification = justified}
 \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
@@
 
-\textbackslash{}begin\{figure\}
+\begin{figure*}
 
-\{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivFun-1}
+{\centering \includegraphics{WholePlotTrajectories_files/figure-latex/IDHplot-1}
 
-\}
+}
 
-\textbackslash{}caption\{Upper panels, Trajectories of the Simpson
-taxonomic diversity and Rao functional diversity over 30 years after
-disturbance. Trajectories correspond to the median (solid line) and
-0.025 and 0.975 percentile (gray envelope) observed after 50 iteration
-of the taxonomic uncertainty propagation and the missing trait value
-filling processes. Initial treatments are represented by solid lines
-colors with green for control, blue for T1,orange for T2 and red for T3.
-Lower panels, Relationship between the initial \%AGB removed and the
-values of Simpson (c) and Rao (d) diversities at three times after
-disturbance. Solid lines colors represent the time, 10 years (yellow),
-20 years (orange) and 30 years (brown) after
-disturbance.\}\label{fig:DivFun} \textbackslash{}end\{figure\}
+\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.
+Lower panels, Relationship between the initial \%AGB removed and the values of Simpson \textbf{(c)} and Rao \textbf{(d)} diversities at three times after disturbance.
+Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.}\label{fig:IDHplot}
+\end{figure*}
+
+The functional redundancy was measured as the sum of weighted overlap
+among species in the communities. Species functional space were defined
+in the main two-dimensional space of a PCA analysis based on all leaf
+and stem and maximum height traits (see appendix I for more details).
+Plots functoinal redundancy was followed in comparison to their initial
+value right after logging so that only communities dynamics are followed
+apart from the impact of logging itself. While the functional redundancy
+of control plots remained stable along the 30 years it systematically
+decreased after disturbance (\ref{fig:RedFun}). The decrease of
+functional redundancy was comparable between all disturbed plots but
+started later for the most disturbed plots. Plots recovery did not
+seemed correlated to the initial disturbance, for five out of nine
+disturbed plots the initial functional redundancy was had recovered
+while it was still decreasing or stabilized at lower values for the fou
+other plots.
 

### 2018-05-10 WholePlotTrajectories.Rmd

@@
 The large areas covered with tropical forests worldwide hold crucial economic, social and cultural value. 
-They provide wood and multiple non-timber forest products, shelter for a diversified fauna, regulate the local climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The simultaneous increase of forests products demand and substantial climatic changes heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threatened the maintenance of communities structure, composition and functioning and their underlying dynamics [@Anderson-Teixeira2013;@Sist2015].
+They provide wood and multiple non-timber forest products, shelter for a diversified fauna, regulate the local and global climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being.
+The simultaneous increase of forests products demand and substantial climatic changes has heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threatened the maintenance of communities structure, composition and functioning and their underlying dynamics in space and time [@Anderson-Teixeira2013;@Sist2015].
 
-In tropical forests communities are shaped by a constant range of disturbance that changes the abiotic environment, as the light, the heat and water fluxes, and the biotic interaction and competitive pressure.
-The cornerstone of tropical forests ecology is then to understand their response to disturbance and its mechanisms [@White2001;@Chazdon2003a].
-Forests response has been largely studied through structural parameters rapid and convenient to measure, as aboveground biomass, tree height or stem density.
-These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
-However the response of tree species diversity, which is determinant of ecosystems productivity, stability and functioning [@Tilman2014] and would be most probably impacted by post-disturbance environmental changes [@CazzollaGatti2014], remains unclear.
+In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events that changes the abiotic environment, as the light, the heat and water fluxes, and the biotic interaction and competitive pressure *(Goulamoussene Biogeosciences)*.
+The cornerstone of tropical forests ecology is then to understand the underlyings of ecosystem response to disturbance, its mechanisms and drivers [@White2001;@Chazdon2003a].
+For now, this has been largely studied through structural parameters rapid and convenient to measure, as aboveground biomass, tree height or stem density *(e.g.
+PiponiotElife, RutishauserBiotropica)*.
+These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Herault2018].
+However the response of tree species diversity, which is determinant of ecosystems productivity, stability and functioning [@Tilman2014, Liang2016Science] and would be most probably impacted by post-disturbance environmental changes [@BaralotoJAE], remains unclear.
 
 In the short-term diversity dynamics demonstrated negligible or even positive impacts of disturbance on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
-However, validations of the IDH in the long term remain scarce and mainly rely on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and evenness, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, the functional approach accounting for species biological attributes would be insightful as it reveals species fitness and ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-In that respect major functional traits related to species ecology and performance were largely adopted through a relevant framework [@Diaz2005;@Villeger2008a].
-The functional trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
-This translated by shifts from "conservative" slow-growing species dealing with scarce resources, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
-It was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
+However, validations of the IDH in the long term remain scarce and mainly rely on the analysis of the rough species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and diversity profiles, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Furthermore, the functional approach accounting for differenes in species biological attributes and functioning would be insightful as it reveals the species role in the post-disturbance trajectory of ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+In that respect major functional traits related to species ecology and performance were largely adopted through an integrative framework [@Diaz2016Nature, @Kunstler2016Nature].
+In disturbed tropical rainforests, this framework translates into the fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008] with a shift from "conservative" slow-growing species dealing with scarce resources that dominate before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011] and was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
+
 A proper monitoring of communities response should therefore encompass taxonomic and functional diversity and composition measures to test the validity of the IDH in the long term for tropical forests, and clarify the resilience of communities evenness, composition, and also functioning.
-The trajectories followed by all these facets would highlight the role of deterministic processes, like competitive exclusion or abiotic selection, and the communities' convergence maintaining intrinsinc differences in diversity and composition,which is as much insgights for future adaptive conservation strategies [@Adler2007].
-
+The trajectories followed by all these facets would highlight the role of deterministic processes, like competitive exclusion or abiotic selection, and the communities' convergence maintaining intrinsinc differences in diversity and composition, which is as much insights for future adaptive management strategies [@Adler2007].
 Here we investigated over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
 We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
-Specifically, we (i) tested the validity of the IDH in the long term for tropical hiperdiverse forest and highlighted the ecological rules shaping their response to disturbance, (ii) clarified the different facets of communities resilience in terms of communities composition, diversity and functioning (iii) questioned the completeness of communities recovery given the altered functional redundancy.
->>revoir la fin
+Specifically, we (i) tested the validity of the IDH in the long term for tropical hyperdiverse forest and highlighted the ecological rules shaping their response to disturbance, (ii) clarified the different facets of communities resilience in terms of communities composition, diversity and functioning *????(iii) questioned the completeness of communities recovery given the altered functional redundancy.*
 

### 2018-05-11 WholePlotTrajectories.Rmd
redundancy script update

--- 2018-05-10 WholePlotTrajectories.Rmd
@@
 The large areas covered with tropical forests worldwide hold crucial economic, social and cultural value. 
-They provide wood and multiple non-timber forest products, shelter for a diversified fauna, regulate the local climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The simultaneous increase of forests products demand and substantial climatic changes heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threatened the maintenance of communities structure, composition and functioning and their underlying dynamics [@Anderson-Teixeira2013;@Sist2015].
+They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being.
+The simultaneous increase of forests products demand and substantial climatic changes currently heighten the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015].
+It threatens the maintenance of communities structure, composition and functioning and the underlying dynamics [@Anderson-Teixeira2013;@Sist2015].
 
-In tropical forests communities are shaped by a constant range of disturbance that changes the abiotic environment, as the light, the heat and water fluxes, and the biotic interaction and competitive pressure.
-The cornerstone of tropical forests ecology is then to understand their response to disturbance and its mechanisms [@White2001;@Chazdon2003a].
-Forests response has been largely studied through structural parameters rapid and convenient to measure, as aboveground biomass, tree height or stem density.
-These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
-However the response of tree species diversity, which is determinant of ecosystems productivity, stability and functioning [@Tilman2014] and would be most probably impacted by post-disturbance environmental changes [@CazzollaGatti2014], remains unclear.
+In tropical forests communities are shaped by a constant range of disturbance changing the abiotic environment, through the fluxes of light, heat and water, and the bioticenvironment, through the interaction and competition among species.
+The cornerstone of tropical forests ecology is to understand their response to disturbance and its mechanisms [@White2001;@Chazdon2003a].
+Forests response has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density.
+These structural parameters have then been sucessfully modeled which allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
+However the response of tree species diversity remains unclear.
+It is however crucial as it determines ecosystems productivity, stability and functioning [@Tilman2014] and would be most probably impacted by the environmental changes due to disturbance [@CazzollaGatti2014].
 
-In the short-term diversity dynamics demonstrated negligible or even positive impacts of disturbance on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
-However, validations of the IDH in the long term remain scarce and mainly rely on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+In the short-term disturbance demonstrated negligible or even positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
+In the long term however the application of the IDH has been scarcely tested.
+The analysis besides mainly relied on species richness that gives limited or even misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
 More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and evenness, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
@@
 The samples of the trait database were first located with a PCA analysis in a two-dimensional functional space summarizing the 8 functional traits considered.
+We choose to summarize species functional space through a PCA because functional traits considered were already somehow redundant and allows not to overweight the role of one trait.
 Then, multivariate kernel density estimator associated with individual trees were summed to give the distribution of traits probabilites of each species.
@@
 Eventually communities functional redundancy was measured as the sum of overlap between species weighted functional density.
+Communities functional redundancy is expressed as the average number of species that could be removed from the community without reducing the functional space.
 
@@
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure"}
-load("DB/Redundancy")
+load("DB/Redundancy2")
 

### 2018-05-11 Scripts/Functionalredundancy.R

@@ -1,4 +1,4 @@
-
+library("MASS")
 source("Scripts/Vernacular_handle.R")           # Script for vernacular names replacement
 source("Scripts/TraitsMiceFilling.R")                                   # Traits Gap-filling
 
@@ -10,9 +10,16 @@ Traits1<-read.csv("DB/BridgeOK.csv",sep=";",na.strings="")       # Brigde functi
 Traits2<-read.csv("DB/DataLifeTraits.csv",sep=";",na.strings="") # Paracou15 seed and Hmax dataset
 TraitsName<-c("L_thickness","L_chloro","L_toughness","L_DryMass","SLA","WD","Bark_thick","Hmax") # Considered traits list
 
-### Whole community mapping, individuals: NMDS ou ACP?
+InventorySp<-do.call(rbind,lapply(LivingStand_all,function(yr){
+  ret<-do.call(rbind,yr)[,c("Famille","Genre","name")]
+  ret<-ret[which(!grepl("Indet.",ret[,"name"])),]
+  return(ret[which(!duplicated(ret)),])}))
+InventorySp<-InventorySp[which(!duplicated(InventorySp)),]
 
 traits_filled<-Traits_filling(Traits1,Traits2,InventorySp)
+
+### Whole community mapping, individuals: NMDS ou ACP?
+
 BigNmds<-as.matrix(daisy(scale(traits_filled[,TraitsName]),metric="gower"))
 BigNmds<-scores(metaMDS(BigNmds,distance="bray"),display="sites")
 BigNmds<-merge(BigNmds,traits_filled[,c("Family","Genus","name" )],by="row.names")[,c("Family","Genus","name","NMDS1","NMDS2")]
@@ -73,10 +80,38 @@ smooth<-function(mat,larg){return(do.call(cbind,lapply(1:ncol(mat),function(step
   range<-max(1,step-larg):min(ncol(mat),step+larg)
   rowSums(mat[,range])/length(range)})))}
 
-Nrep<-10
-Mat<-lapply(1:Nrep,function(rep){
+RefSpaces<-lapply(1:12,function(p){   
+  temp<-LivingStand_all[[which(names(LivingStand_all)==1989)]]
+  temp<-temp[[which(names(temp)==p)]]
+  temp<-temp[!duplicated(temp),]
[…]

### 2018-05-11 WholePlotTrajectories.Rmd

@@
 The simultaneous increase of forests products demand and substantial climatic changes currently heighten the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015].
-It threatens the maintenance of communities structure, composition and functioning and the underlying dynamics [@Anderson-Teixeira2013;@Sist2015].
+It threatens the maintenance of communities structure, composition and functioning and the underlying dynamics  in space and time [@Anderson-Teixeira2013;@Sist2015].
 
-In tropical forests communities are shaped by a constant range of disturbance changing the abiotic environment, through the fluxes of light, heat and water, and the bioticenvironment, through the interaction and competition among species.
-The cornerstone of tropical forests ecology is to understand their response to disturbance and its mechanisms [@White2001;@Chazdon2003a].
-Forests response has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density.
-These structural parameters have then been sucessfully modeled which allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
-However the response of tree species diversity remains unclear.
-It is however crucial as it determines ecosystems productivity, stability and functioning [@Tilman2014] and would be most probably impacted by the environmental changes due to disturbance [@CazzollaGatti2014].
+In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events that changes the abiotic environment, through the fluxes of light, heat and water, and the biotic interaction and competitive pressure *(Goulamoussene Biogeosciences)*.
+The cornerstone of tropical forests ecology is then to understand the underlyings of ecosystem response to disturbance, its mechanisms and drivers [@White2001;@Chazdon2003a]. 
+For now, this has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density *(e.g.
+PiponiotElife, RutishauserBiotropica)*.
+These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016]. 
+However the response of tree species diversity remains unclear, also it is determinant of ecosystems productivity, stability and functioning [@Tilman2014, Liang2016Science] and would be most probably impacted by post-disturbance environmental changes [@BaralotoJAE.
 
 In the short-term disturbance demonstrated negligible or even positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
-In the long term however the application of the IDH has been scarcely tested.
-The analysis besides mainly relied on species richness that gives limited or even misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and evenness, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, the functional approach accounting for species biological attributes would be insightful as it reveals species fitness and ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-In that respect major functional traits related to species ecology and performance were largely adopted through a relevant framework [@Diaz2005;@Villeger2008a].
+Validations of the IDH in the long term, however, remain scarce and mainly rely on the analysis of the rough species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and diversity profiles, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Furthermore, the functional approach accounting for differenes species biological attributes would be insightful as it reveals the species role in the post-disturbance trajectory of ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+In that respect major functional traits related to species ecology and performance were largely adopted through an integrative framework [@Diaz2005;@Villeger2008a].
 The functional trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
-This translated by shifts from "conservative" slow-growing species dealing with scarce resources, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
+In disturbed tropical rainforests, this framework translates into the fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008] with a shift from "conservative" slow-growing species dealing with scarce resources that dominate before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
 It was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
+
 A proper monitoring of communities response should therefore encompass taxonomic and functional diversity and composition measures to test the validity of the IDH in the long term for tropical forests, and clarify the resilience of communities evenness, composition, and also functioning.
@@
 We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
-Specifically, we (i) tested the validity of the IDH in the long term for tropical hiperdiverse forest and highlighted the ecological rules shaping their response to disturbance, (ii) clarified the different facets of communities resilience in terms of communities composition, diversity and functioning (iii) questioned the completeness of communities recovery given the altered functional redundancy.
->>revoir la fin
+Specifically, we (i) tested the validity of the IDH in the long term for tropical hyperdiverse forest and highlighted the ecological rules shaping their response to disturbance, (ii) clarified the different facets of communities resilience in terms of communities composition, diversity and functioning *????(iii) questioned the completeness of communities recovery given the altered functional redundancy.*
 

### 2018-05-11 WholePlotTrajectories.Rmd
Redundancy plot

@@
 These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016]. 
-However the response of tree species diversity remains unclear, also it is determinant of ecosystems productivity, stability and functioning [@Tilman2014, Liang2016Science] and would be most probably impacted by post-disturbance environmental changes [@BaralotoJAE.
+However the response of tree species diversity remains unclear, also it is determinant of ecosystems productivity, stability and functioning [@Tilman2014, Liang2016Science] and would be most probably impacted by post-disturbance environmental changes [@BaralotoJAE].
 
@@
 For all undisturbed plots the functional diversity remained comparable to the reference values 5 years after disturbance (1989 inventories).
-For all disturbed plots the functional trajectories followed a unimodal trajectory with a maximum positively correlated to disturbance intensity ($\rho_{spearman}=0.73$) (Figure \@ref(fig:DivFun)). 
+For all disturbed plots the functional trajectories followed a unimodal trajectory with a maximum positively correlated to disturbance intensity ($\rho_{spearman}=0.73$) (Figure \@ref(fig:IDHplot)). 
 The plot 7 from treatment 1 that displayed a constantly outlying diversity was removed from the graphical representation for better readability, see appendix for full graph.
@@
 Species functional space were defined in the main two-dimensional space of a PCA analysis based on all leaf and stem and maximum height traits (see appendix I for more details). 
-Plots functoinal redundancy was followed in comparison to their initial value right after logging so that only communities dynamics are followed apart from the impact of logging itself.
-While the functional redundancy of control plots remained stable along the 30 years it systematically decreased after disturbance (\@ref(fig:RedFun)).
-The decrease of functional redundancy was comparable between all disturbed plots but started later for the most disturbed plots.
-Plots recovery did not seemed correlated to the initial disturbance, for five out of nine disturbed plots the initial functional redundancy was had recovered while it was still decreasing or stabilized at lower values for the fou other plots.
+Plots functional redundancy compared to the initial value after logging so that only communities dynamics are followed apart from the impact of logging itself.
+Communities functional redundancy remained stable in control plots but after disturbance the redundancy trajectories were quite variable (See appendix I).
+Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place and then increased to edge, recover or even exceed the initial value.
+Considering the functional redundancy restricted to the functional space of the initial inventory, though, all disturbed plots followed similar decreasing humped-shaped trajectories (\@ref(fig:RedFun)).
+The maximum redundancy loss observed was positively correlated with the disturbance intensity (XX spearman to be measured) and none of the disturbed plots had recovered the functional redundancy within their initial functional space.
 
@@
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure"}
-load("DB/Redundancy2")
+load("DB/Redundancy_restricted")
 
 par(mar=c(5,4,5,2))
-RedundancyPlot(RedundancyTraj)
+RedundancyPlot(RedundancyTraj_restricted)
 mtext("Functional Redundancy",side=3,adj=0,line=1.5)
-mtext("Sum of abundance-weighted Trait Probability Density",side=3,adj=0,cex=0.8,line=0.5)
+mtext("Restricted to initial functional space",side=3,adj=0,cex=0.8,line=0.5)
 mtext("Years since disturbance",side=1,line=2.2,adj=1)
@@
 Besides the long-term alteration of functional redundancy, there was probably persistent compositional changes favoring disturbance resistant species [@Haddad2008], lianas or epiphytes [@Martin2013] and environmental changes, like in the soils nutrient cycling and compaction [@Olander2005].
-These presistent changes highly question forest's resilience [@Chazdon2003a].
+These persistent changes highly question forest's resilience [@Chazdon2003a].
 New conditions would not only be longer lasting but self-maintained as tied to disturbance regime [@Burslem2000].

### 2018-05-11 WholePlotTrajectories.tex
Redundancy plot

@@
 economic, social and cultural value. They provide wood and multiple
-non-timber forest products, shelter for a diversified fauna, regulate
-the local climate, support the carbon, water and nutrient cycles, and
-ensure cultural and human well-being. The simultaneous increase of
-forests products demand and substantial climatic changes heightened the
+non-timber forest products, shelter a diversified fauna, regulate the
+local climate, support the carbon, water and nutrient cycles, and ensure
+cultural and human well-being. The simultaneous increase of forests
+products demand and substantial climatic changes currently heighten the
 pressure on the remaining forests
-\citep{Gibson2011a, Morales-Hidalgo2015} and threatened the maintenance
-of communities structure, composition and functioning and their
-underlying dynamics \citep{Anderson-Teixeira2013, Sist2015}.
+\citep{Gibson2011a, Morales-Hidalgo2015}. It threatens the maintenance
+of communities structure, composition and functioning and the underlying
+dynamics in space and time \citep{Anderson-Teixeira2013, Sist2015}.
 
-In tropical forests communities are shaped by a constant range of
-disturbance that changes the abiotic environment, as the light, the heat
-and water fluxes, and the biotic interaction and competitive pressure.
-The cornerstone of tropical forests ecology is then to understand their
-response to disturbance and its mechanisms
-\citep{White2001, Chazdon2003a}. Forests response has been largely
-studied through structural parameters rapid and convenient to measure,
-as aboveground biomass, tree height or stem density. These structural
+In tropical forest, ecological communities are contantly re-shaped by
+the natural disturbance events that changes the abiotic environment,
+through the fluxes of light, heat and water, and the biotic interaction
+and competitive pressure \emph{(Goulamoussene Biogeosciences)}. The
+cornerstone of tropical forests ecology is then to understand the
+underlyings of ecosystem response to disturbance, its mechanisms and
+drivers \citep{White2001, Chazdon2003a}. For now, this has been largely
+studied through structural parameters, rapid and convenient to measure,
+as aboveground biomass, tree height or stem density
+\emph{(e.g.~PiponiotElife, RutishauserBiotropica)}. These structural
 parameters have then been sucessfully modeled and allowed to assess the
@@
 \citep{Denslow2000, Blanc2009, Rutishauser2016}. However the response of
-tree species diversity, which is determinant of ecosystems productivity,
-stability and functioning \citep{Tilman2014} and would be most probably
-impacted by post-disturbance environmental changes
-\citep{CazzollaGatti2014}, remains unclear.
+tree species diversity remains unclear, also it is determinant of
+ecosystems productivity, stability and functioning
+\citep[Liang2016Science]{Tilman2014} and would be most probably impacted
+by post-disturbance environmental changes \citep{BaralotoJAE}.
 
-In the short-term diversity dynamics demonstrated negligible or even
-positive impacts of disturbance on communities diversity, which have
-been formalized by the intermediate disturbance hypothesis (IDH) stating
-a maximized species diversity at intermediate disturbance intensity
-\citep{Molino2001, Kariuki2006a, Berry2008a}. However, validations of
-the IDH in the long term remain scarce and mainly rely on the analysis
-of species richness that gives limited or misleading information on
-forests recovery and functioning \citep{Martin2015, Chaudhary2016}. More
-releveant monitoring would encompass communities composition, that is
-crucial for conservation issues, and evenness, that reveals ecological
-rules underlying communities' structure
+In the short-term disturbance demonstrated negligible or even positive
+impacts on communities diversity, which have been formalized by the
+intermediate disturbance hypothesis (IDH) stating a maximized species
+diversity at intermediate disturbance intensity
+\citep{Molino2001, Kariuki2006a, Berry2008a}. Validations of the IDH in
+the long term, however, remain scarce and mainly rely on the analysis of
+the rough species richness that gives limited or misleading information
+on forests recovery and functioning \citep{Martin2015, Chaudhary2016}.
+More releveant monitoring would encompass communities composition, that
+is crucial for conservation issues, and diversity profiles, that reveals
+ecological rules underlying communities' structure
 \citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, the
-functional approach accounting for species biological attributes would
-be insightful as it reveals species fitness and ecosystems functioning
+functional approach accounting for differenes species biological
+attributes would be insightful as it reveals the species role in the
+post-disturbance trajectory of ecosystems functioning
 \citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
 respect major functional traits related to species ecology and
-performance were largely adopted through a relevant framework
+performance were largely adopted through an integrative framework
 \citep{Diaz2005, Villeger2008a}. The functional trait-based approach,
@@
 filters fostering disturbance resistant species with rapid growth and
-efficient resources acquisition \citep{Molino2001, Haddad2008}. This
-translated by shifts from ``conservative'' slow-growing species dealing
-with scarce resources, to ``acquisitive'' fast-growing species with
-rapid and efficient use of abundant resources
+efficient resources acquisition \citep{Molino2001, Haddad2008}. In
+disturbed tropical rainforests, this framework translates into the
+fostering disturbance resistant species with rapid growth and efficient
+resources acquisition \citep{Molino2001, Haddad2008} with a shift from
+``conservative'' slow-growing species dealing with scarce resources that
+dominate before disturbance, to ``acquisitive'' fast-growing species
+with rapid and efficient use of abundant resources
 \citep{TerSteege2001, Reich2014, Herault2011}. It was mirrored by shifts
@@
 thickness), tree growth and reproduction life history traits (seed mass
-and maximum height) \citep{Wright2004, Westoby2006a, Chave2009b}. A
-proper monitoring of communities response should therefore encompass
+and maximum height) \citep{Wright2004, Westoby2006a, Chave2009b}.
+
+A proper monitoring of communities response should therefore encompass
 taxonomic and functional diversity and composition measures to test the
@@
 functional composition and diversity. Specifically, we (i) tested the
-validity of the IDH in the long term for tropical hiperdiverse forest
+validity of the IDH in the long term for tropical hyperdiverse forest
 and highlighted the ecological rules shaping their response to
@@
 resilience in terms of communities composition, diversity and
-functioning (iii) questioned the completeness of communities recovery
-given the altered functional redundancy.
-\textgreater{}\textgreater{}revoir la fin
+functioning \emph{????(iii) questioned the completeness of communities
+recovery given the altered functional redundancy.}
 
@@
 two-dimensional functional space summarizing the 8 functional traits
-considered. Then, multivariate kernel density estimator associated with
-individual trees were summed to give the distribution of traits
-probabilites of each species. For each community the trait probability
-distributions of corresponding species were combined and weighted by
-species abundance. Eventually communities functional redundancy was
-measured as the sum of overlap between species weighted functional
-density.
+considered. We choose to summarize species functional space through a
+PCA because functional traits considered were already somehow redundant
+and allows not to overweight the role of one trait. Then, multivariate
+kernel density estimator associated with individual trees were summed to
+give the distribution of traits probabilites of each species. For each
+community the trait probability distributions of corresponding species
+were combined and weighted by species abundance. Eventually communities
+functional redundancy was measured as the sum of overlap between species
+weighted functional density. Communities functional redundancy is
+expressed as the average number of species that could be removed from
+the community without reducing the functional space.
 
@@
 and stem and maximum height traits (see appendix I for more details).
-Plots functoinal redundancy was followed in comparison to their initial
-value right after logging so that only communities dynamics are followed
-apart from the impact of logging itself. While the functional redundancy
-of control plots remained stable along the 30 years it systematically
-decreased after disturbance (\ref{fig:RedFun}). The decrease of
-functional redundancy was comparable between all disturbed plots but
-started later for the most disturbed plots. Plots recovery did not
-seemed correlated to the initial disturbance, for five out of nine
-disturbed plots the initial functional redundancy was had recovered
-while it was still decreasing or stabilized at lower values for the fou
-other plots.
+Plots functional redundancy compared to the initial value after logging
+so that only communities dynamics are followed apart from the impact of
+logging itself. Communities functional redundancy remained stable in
+control plots but after disturbance the redundancy trajectories were
+quite variable (See appendix I). Globally after most intense disturbance
+(plots T2 and T3) communities redundancy decreased at first place and
+then increased to edge, recover or even exceed the initial value.
+Considering the functional redundancy restricted to the functional space
+of the initial inventory, though, all disturbed plots followed similar
+decreasing humped-shaped trajectories (\ref{fig:RedFun}). The maximum
+redundancy loss observed was positively correlated with the disturbance
+intensity (XX spearman to be measured) and none of the disturbed plots
+had recovered the functional redundancy within their initial functional
+space.
 
@@
 environmental changes, like in the soils nutrient cycling and compaction
-\citep{Olander2005}. These presistent changes highly question forest's
+\citep{Olander2005}. These persistent changes highly question forest's
 resilience \citep{Chazdon2003a}. New conditions would not only be longer

### 2018-05-11 Scripts/DrawingTools.R

@@ -241,26 +241,31 @@ mtext("Years since disturbance",side=1,line=1.2,adj=1,cex=0.9,outer=TRUE)
 }
 
 RedundancyPlot<-function(Red){
-  abs<-as.numeric(colnames(Red[[1]]))-1984
-  Red<-lapply(Red,function(rep){return(smooth(rep,2))})
-  Red<-lapply(Red,function(rep){return(apply(rep,2,function(col){return(col-rep[,1])}))})
+  Red<-lapply(Red,function(rep){
+    ret<-smooth(rep,2)
+   colnames(ret)<-colnames(rep)
+   ret<-apply(ret[,which(as.numeric(colnames(ret))>="1989")],2,function(col){return(col-ret[,"1989"])})
+   colnames(ret)<-as.numeric(colnames(ret))-1984
+   return(ret)})
+  
   Red<-array(unlist(Red),dim=c(12,ncol(Red[[1]]),length(Red)),
              dimnames=list(1:12,colnames(Red[[1]]),1:length(Red)))
-  Red<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(Red,c(1,2),function(rep){return(quantile(rep,probs=quant))}))})
+  Red<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(Red,c(1,2),function(rep){
+    return(quantile(rep,probs=quant))}))})
   Red<-array(unlist(Red),dim=c(12,ncol(Red[[1]]),3),
              dimnames=list(1:12,colnames(Red[[1]]),c(0.025,0.5,0.975)))
-  #Red<-Red[,which(colnames(Red)>=1989),]
   
-  plot(abs,Red[1,,1],type='n',ylim=c(min(Red),max(Red)),xlab="",ylab="")
+  plot(colnames(Red),Red[1,,1],type='n',ylim=c(min(Red),max(Red)),xlab="",ylab="")
+  
   invisible(lapply(1:4,function(tr){
-    toplot<-Red[which(rownames(Red)%in%treatments[[tr]]),,"0.5"]
+    toplot<-Red[which(rownames(Red)%in%treatments[[tr]]),,]
     apply(toplot,1,function(li){
-      lines(abs,li,col=ColorsTr[tr],lwd=2)
-    })
-  }))
+      lines(colnames(Red),li[,"0.5"],col=ColorsTr[tr],lwd=2)
+      polygon(c(colnames(Red),rev(colnames(Red))),
[…]

### 2018-05-11 Scripts/Functionalredundancy.R

@@ -4,7 +4,8 @@ source("Scripts/TraitsMiceFilling.R")                                   # Traits
 
 load("DB/Alpha_Plots")                  # Alpha vector for vernacular replacement
 load("DB/Paracou_R_Subdivided_ok")      # Paracou dataset, lists by inventoried year, sublists by plot, etc...
-dates<-sort(names(LivingStand_all))[-1] # dates considered
+dates<-sort(names(LivingStand_all)) # dates considered
+dates<-dates[which(!dates%in%c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014","2016","2017"))]
 
 Traits1<-read.csv("DB/BridgeOK.csv",sep=";",na.strings="")       # Brigde functional traits dataset
 Traits2<-read.csv("DB/DataLifeTraits.csv",sep=";",na.strings="") # Paracou15 seed and Hmax dataset
@@ -55,13 +56,15 @@ SpTdp<-lapply(unique(Bigacp[,"name"]),function(Sp){
   f<-kde2d(acp[,"Axis1"], acp[,"Axis2"], n = 100,lims=c(xgen,ygen),
                h = c(width.SJ(acp[,"Axis1"],method=method1),
                      width.SJ(acp[,"Axis2"],method=method2)) )
-  return(f$z)
+  z<-f$z
+  colnames(z)<-1:100;rownames(z)<-1:100
   }
   if(nrow(acp)==1){
     z<-outer(seq(xgen[1],xgen[2],length=100),seq(ygen[1],ygen[2],length=100), 
               function(x,y) dnorm(x,acp[,"Axis1"],MeanSd)*dnorm(y,acp[,"Axis2"],MeanSd))
-    return(z)
+    colnames(z)<-1:100;rownames(z)<-1:100
   }
+  return(z)
 })
 names(SpTdp)<-unique(Bigacp[,"name"])
 
@@ -72,16 +75,52 @@ T0<-c(1,6,11); T1<-c(2,7,9); T2<-c(3,5,10); T3<-c(4,8,12)
 treatments<-list(T0,T1,T2,T3)
 names(treatments)<-c("Control","T1","T2","T3")
 
-Pby<-c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014","2016","2017")
-
[…]

### 2018-05-11 WholePlotTrajectories.Rmd
code bidouille

@@
 
-In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events that changes the abiotic environment, through the fluxes of light, heat and water, and the biotic interaction and competitive pressure *(Goulamoussene Biogeosciences)*.
+In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events that changes the abiotic environment, through the fluxes of light, heat and water, and the biotic interaction and competitive pressure [@Goulamoussene2017].
 The cornerstone of tropical forests ecology is then to understand the underlyings of ecosystem response to disturbance, its mechanisms and drivers [@White2001;@Chazdon2003a]. 
-For now, this has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density *(e.g.
-PiponiotElife, RutishauserBiotropica)*.
-These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016]. 
-However the response of tree species diversity remains unclear, also it is determinant of ecosystems productivity, stability and functioning [@Tilman2014, Liang2016Science] and would be most probably impacted by post-disturbance environmental changes [@BaralotoJAE].
+For now, this has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
+These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009]. 
+However the response of tree species diversity remains unclear, also it is determinant of ecosystems productivity, stability and functioning [@Tilman2014, @Liang2016] and would be most probably impacted by post-disturbance environmental changes [@Baraloto2012a].
 
@@
 The functional trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
-In disturbed tropical rainforests, this framework translates into the fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008] with a shift from "conservative" slow-growing species dealing with scarce resources that dominate before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
+In disturbed tropical rainforests, this framework translates into the fostering of disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
+It results in a shift from "conservative" slow-growing species dealing with scarce resources that dominate before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
 It was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].

### 2018-05-11 Scripts/Functionalredundancy.R

@@ -1,6 +1,6 @@
 library("MASS")
 source("Scripts/Vernacular_handle.R")           # Script for vernacular names replacement
-source("Scripts/TraitsMiceFilling.R")                                   # Traits Gap-filling
+source("Scripts/TraitsMiceFilling.R")            # Traits Gap-filling
 
 load("DB/Alpha_Plots")                  # Alpha vector for vernacular replacement
 load("DB/Paracou_R_Subdivided_ok")      # Paracou dataset, lists by inventoried year, sublists by plot, etc...
@@ -17,59 +17,6 @@ InventorySp<-do.call(rbind,lapply(LivingStand_all,function(yr){
   return(ret[which(!duplicated(ret)),])}))
 InventorySp<-InventorySp[which(!duplicated(InventorySp)),]
 
-traits_filled<-Traits_filling(Traits1,Traits2,InventorySp)
-
-### Whole community mapping, individuals: NMDS ou ACP?
-
-BigNmds<-as.matrix(daisy(scale(traits_filled[,TraitsName]),metric="gower"))
-BigNmds<-scores(metaMDS(BigNmds,distance="bray"),display="sites")
-BigNmds<-merge(BigNmds,traits_filled[,c("Family","Genus","name" )],by="row.names")[,c("Family","Genus","name","NMDS1","NMDS2")]
-save(BigNmds,file="DB/BigNmds")
-
-ACP<-dudi.pca(scale(traits_filled[,TraitsName]))
-Bigacp<-merge(ACP$li,traits_filled[,c("Family","Genus","name" )],by="row.names")[,c("Family","Genus","name","Axis1","Axis2")]
-
-plot(Bigacp[,"Axis1"], Bigacp[,"Axis2"],type="n")
-text(Bigacp[,"Axis1"], Bigacp[,"Axis2"],labels=Bigacp[,"name"])
-
-### Species TPD
-
-xgen<-c(min(Bigacp[,"Axis1"]),max(Bigacp[,"Axis1"]))
-ygen<-c(min(Bigacp[,"Axis2"]),max(Bigacp[,"Axis2"]))
-  
-# For the species with one individual: reproduce a gaussian kernel,
-# centered in the individual location and with sd like the mean sd of all species
[…]

### 2018-05-11 Scripts/TraitsMiceFilling.R

@@ -68,8 +68,9 @@ traitsPartial_gen<-traitsPartial_gen[which(unlist(lapply(traitsPartial_gen,nrow)
 
 # gap filling à proporement parler, à chaque élément de la liste de sgenre, on applique la fonction mice
 traitsPartial_gen<-lapply(traitsPartial_gen,function(sub){
-  ret<-complete(mice(sub[,Seltraits],printFlag=F))
-  return(cbind(sub[,c("Family","Genus","name","bar_code")],ret))})
+  data<-rbind(traitsComp[which(traitsComp[,"Genus"]==unique(sub[,"Genus"])),],sub)
+  ret<-complete(mice(data[,Seltraits],printFlag=F))
+  return(cbind(sub[,c("Family","Genus","name","bar_code")],ret[(nrow(ret)-nrow(sub)+1):nrow(ret),]))})
 
 # J'avais de soucis avec les NA persistants pour la Wood Density, 
 #si le problème persiste je ramène ces espèces à la famille et je refais l'estimation
@@ -87,8 +88,9 @@ traitsPartial_fam<-traitsPartial_fam[which(unlist(lapply(traitsPartial_fam,nrow)
 
 # Gap filling des groupes par famille
 traitsPartial_fam<-lapply(traitsPartial_fam,function(sub){
-  ret<-complete(mice(sub[,Seltraits],printFlag=F))
-  return(cbind(sub[,c("Family","Genus","name","bar_code")],ret))})
+  data<-rbind(traitsComp[which(traitsComp[,"Family"]==unique(sub[,"Family"])),],sub)
+  ret<-complete(mice(data[,Seltraits],printFlag=F))
+  return(cbind(sub[,c("Family","Genus","name","bar_code")],ret[(nrow(ret)-nrow(sub)+1):nrow(ret),]))})
 
 # Pour les espèce squi n'avaient toujours pas assez d'individus dans leur famille pour permettre l'estimation
 idScarce<-as.character(traitsPartial_all[,"bar_code"],

### 2018-05-14 AppendixI.Rmd
Appendix, PCA graphs

--- 2018-04-20 AppendixI.Rmd
@@
 
+(ref:Bolda) **(a)**
+
+(ref:Boldb) **(b)**
+
+(ref:Boldc) **(c)**
+
 ```{r Options, echo=FALSE, message=FALSE, warning=F}
@@
 
+**Figure A4**: Trajectories of the Rao functional diversity over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.
+\@ref(fig:PCA_graphs).
+
+**Figure A5**: Trajectories of the Rao functional diversity over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.
+\@ref(fig:RedFun).
+
 \newpage
 
-# Figure A1
+# **Figure A1**: Bayesian process sheme
 
@@
 
-# Figure A2
+# **Figure A2**: Proportion of seed mass classes along time
 
@@
 
-# Figure A3
+# **Figure A3**: Rao diversity trajectories with all plots display
 
@@
 ```
+
+# **Figure A4**: Functional PCA
+
+```{r PCA_graphs, out.width = '100%', echo=FALSE, fig.cap="Summarizing plots of the functional PCA analysis defining communities functional space.
+(ref:Bolda) Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
+(ref:Boldb) bow plot of the PCA eignevalues in percentage.
+(ref:Boldc) Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).", fig.env="figure*"}
+load("DB/FunctionalPCA")
+
+plotPCA(DataACP)
+
+```
+
+# **Figure A5**: Overall functional redundancy
+
+```{r RedFun, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
+load("DB/Redundancy")
+
+par(mar=c(5,4,5,2))
+RedundancyPlot(RedundancyTraj)
+mtext("Functional Redundancy",side=3,adj=0,line=1.5)
+mtext("Restricted to initial functional space",side=3,adj=0,cex=0.8,line=0.5)
+mtext("Years since disturbance",side=1,line=2.2,adj=1)
+
+```

### 2018-05-14 AppendixI.tex
appendix & graphs
Appendix, PCA graphs

--- 2018-04-20 AppendixI.tex
@@
 \JournalInfo{Hal 00679993} % Journal information
-\Archive{} % Additional notes (e.g. copyright, DOI, review/research article)
+\Archive{\ }
 
@@
 
+\textbf{Figure A4}: Trajectories of the Rao functional diversity over 30
+years after disturbance. Trajectories correspond to the median (solid
+line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
+iteration of the taxonomic uncertainty propagation. Initial treatments
+are represented by solid lines colorswith green for control, blue for
+T1,orange for T2 and red for T3.and the missing trait value filling
+processes. @ref(fig:PCA\_graphs).
+
+\textbf{Figure A5}: Trajectories of the Rao functional diversity over 30
+years after disturbance. Trajectories correspond to the median (solid
+line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
+iteration of the taxonomic uncertainty propagation. Initial treatments
+are represented by solid lines colorswith green for control, blue for
+T1,orange for T2 and red for T3.and the missing trait value filling
+processes. \ref{fig:RedFun}.
+
 \newpage
 
-\section{Figure A1}\label{figure-a1}
+\section{\texorpdfstring{\textbf{Figure A1}: Bayesian process
+sheme}{Figure A1: Bayesian process sheme}}\label{figure-a1-bayesian-process-sheme}
 
@@
 
-\section{Figure A2}\label{figure-a2}
+\section{\texorpdfstring{\textbf{Figure A2}: Proportion of seed mass
+classes along
+time}{Figure A2: Proportion of seed mass classes along time}}\label{figure-a2-proportion-of-seed-mass-classes-along-time}
 
@@
 
-\section{Figure A3}\label{figure-a3}
+\section{\texorpdfstring{\textbf{Figure A3}: Rao diversity trajectories
+with all plots
+display}{Figure A3: Rao diversity trajectories with all plots display}}\label{figure-a3-rao-diversity-trajectories-with-all-plots-display}
 
@@
 
+\section{\texorpdfstring{\textbf{Figure A4}: Functional
+PCA}{Figure A4: Functional PCA}}\label{figure-a4-functional-pca}
+
+\begin{figure*}
+
+{\centering \includegraphics[width=1\linewidth]{AppendixI_files/figure-latex/PCA_graphs-1}
+
+}
+
+\caption{Summarizing plots of the functional PCA analysis defining communities functional space.
+\textbf{(a)} Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
+\textbf{(b)} bow plot of the PCA eignevalues in percentage.
+\textbf{(c)} Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).}(\#fig:PCA_graphs)
+\end{figure*}
+
+\section{\texorpdfstring{\textbf{Figure A5}: Overall functional
+redundancy}{Figure A5: Overall functional redundancy}}\label{figure-a5-overall-functional-redundancy}
+
+\begin{figure*}
+
+{\centering \includegraphics[width=1\linewidth]{AppendixI_files/figure-latex/RedFun-1}
+
+}
+
+\caption{Trajectories of the functional redundancy over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:RedFun}
+\end{figure*}
+
 %----------------------------------------------------------------------------------------
@@
 \bibliographystyle{mee}
-\bibliography{}
+\makeatletter
+% The filename has .bib extension the must be eliminated
+\filename@parse{}
+% parse stores the file name in base. Extension starts at the first dot, so don't use dots in file names.
+\bibliography{\filename@base}
+\makeatother
 

### 2018-05-14 WholePlotTrajectories.Rmd
Appendix, PCA graphs

@@
 
-```{r NMDSplans, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of (ref:Bolda) flora composition and (ref:Boldb) functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
-Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).", fig.env="figure"}
-load("DB/FunctionalComposition_forGraph")
-load("DB/TaxoComposition_forGraphs")
+```{r NMDSplans, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) regarding the 6 leaf and stem functional traits, the maximum allometric height and seed mass class.
+Plots trajectories are first represented in the two-dimensional space from the NMDS performed for the 30 years after disturbance based on Bray-Curtis dissimilarity measures between successive inventories (Upper panels (ref:Bolda) and (ref:Boldb)).
+Then the lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years.
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
+The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.", fig.env="figure*"}
+load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")
+load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")
 
+par(mfrow=c(2,2))
 TaxoCompo(MatrepTaxo)
 FunCompo(MatrepFun)
+TaxoDist(MatrepTaxo)
+FunDist(MatrepFun)
 
@@
 legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
-
 ```
 
-The functional redundancy was measured as the sum of weighted overlap among species in the communities.
-Species functional space were defined in the main two-dimensional space of a PCA analysis based on all leaf and stem and maximum height traits (see appendix I for more details). 
+The functional redundancy was measured as the sum of weighted overlap among species in communities' functional space.
+Functional space were defined in the main two-dimensional space of a PCA analysis based on all leaf and stem and maximum height traits (see appendix I for PCA details). 
 Plots functional redundancy compared to the initial value after logging so that only communities dynamics are followed apart from the impact of logging itself.
@@
 Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place and then increased to edge, recover or even exceed the initial value.
-Considering the functional redundancy restricted to the functional space of the initial inventory, though, all disturbed plots followed similar decreasing humped-shaped trajectories (\@ref(fig:RedFun)).
+Considering the functional redundancy restricted to the functional space of the initial inventory, though, all disturbed plots followed similar decreasing humped-shaped trajectories (\@ref(fig:RedFun_rest)).
 The maximum redundancy loss observed was positively correlated with the disturbance intensity (XX spearman to be measured) and none of the disturbed plots had recovered the functional redundancy within their initial functional space.
 
-```{r RedFun, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure"}
+```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial cpùùunities functional space over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure"}
 load("DB/Redundancy_restricted")

### 2018-05-14 WholePlotTrajectories.tex
Appendix, PCA graphs

@@
 through the fluxes of light, heat and water, and the biotic interaction
-and competitive pressure \emph{(Goulamoussene Biogeosciences)}. The
-cornerstone of tropical forests ecology is then to understand the
-underlyings of ecosystem response to disturbance, its mechanisms and
-drivers \citep{White2001, Chazdon2003a}. For now, this has been largely
-studied through structural parameters, rapid and convenient to measure,
-as aboveground biomass, tree height or stem density
-\emph{(e.g.~PiponiotElife, RutishauserBiotropica)}. These structural
-parameters have then been sucessfully modeled and allowed to assess the
-maintenance of ecosystems processes and services
-\citep{Denslow2000, Blanc2009, Rutishauser2016}. However the response of
-tree species diversity remains unclear, also it is determinant of
-ecosystems productivity, stability and functioning
-\citep[Liang2016Science]{Tilman2014} and would be most probably impacted
-by post-disturbance environmental changes \citep{BaralotoJAE}.
+and competitive pressure \citep{Goulamoussene2017}. The cornerstone of
+tropical forests ecology is then to understand the underlyings of
+ecosystem response to disturbance, its mechanisms and drivers
+\citep{White2001, Chazdon2003a}. For now, this has been largely studied
+through structural parameters, rapid and convenient to measure, as
+aboveground biomass, tree height or stem density
+\citep{Piponiot2016, Rutishauser2016}. These structural parameters have
+then been sucessfully modeled and allowed to assess the maintenance of
+ecosystems processes and services \citep{Denslow2000, Blanc2009}.
+However the response of tree species diversity remains unclear, also it
+is determinant of ecosystems productivity, stability and functioning
+\citep[\citet{Liang2016}]{Tilman2014} and would be most probably
+impacted by post-disturbance environmental changes
+\citep{Baraloto2012a}.
 
@@
 disturbed tropical rainforests, this framework translates into the
-fostering disturbance resistant species with rapid growth and efficient
-resources acquisition \citep{Molino2001, Haddad2008} with a shift from
-``conservative'' slow-growing species dealing with scarce resources that
-dominate before disturbance, to ``acquisitive'' fast-growing species
-with rapid and efficient use of abundant resources
-\citep{TerSteege2001, Reich2014, Herault2011}. It was mirrored by shifts
-in key functional traits related to resource acquisition (leaf area,
-density and chlorophyll content, and stem specific gravity and bark
-thickness), tree growth and reproduction life history traits (seed mass
-and maximum height) \citep{Wright2004, Westoby2006a, Chave2009b}.
+fostering of disturbance resistant species with rapid growth and
+efficient resources acquisition \citep{Molino2001, Haddad2008}. It
+results in a shift from ``conservative'' slow-growing species dealing
+with scarce resources that dominate before disturbance, to
+``acquisitive'' fast-growing species with rapid and efficient use of
+abundant resources \citep{TerSteege2001, Reich2014, Herault2011}. It was
+mirrored by shifts in key functional traits related to resource
+acquisition (leaf area, density and chlorophyll content, and stem
+specific gravity and bark thickness), tree growth and reproduction life
+history traits (seed mass and maximum height)
+\citep{Wright2004, Westoby2006a, Chave2009b}.
 
@@
 
-\begin{figure}
+\begin{figure*}
 
@@
 
-\caption{Trajectories of the plots in terms of \textbf{(a)} flora composition and \textbf{(b)} functional composition regarding the 6 leaf and stem functional traits,the maximum allometric height and seed mass class in the two-dimensional space from the NMDS performed for the 30 years after disturbance.
-Distance matrix for NMDS were computed from the Bray-curtis dissimilarity between successive inventories.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).}\label{fig:NMDSplans}
-\end{figure}
+\caption{Trajectories of the plots in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) regarding the 6 leaf and stem functional traits, the maximum allometric height and seed mass class.
+Plots trajectories are first represented in the two-dimensional space from the NMDS performed for the 30 years after disturbance based on Bray-Curtis dissimilarity measures between successive inventories (Upper panels \textbf{(a)} and \textbf{(b)}).
+Then the lower panels (\textbf{(c)} and \textbf{(d)}) represent the euclidean distance to initial condition along the 30 sampled years.
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
+The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.}\label{fig:NMDSplans}
+\end{figure*}
 
@@
 trajectory with a maximum positively correlated to disturbance intensity
-(\(\rho_{spearman}=0.73\)) (Figure \ref{fig:DivFun}). The plot 7 from
+(\(\rho_{spearman}=0.73\)) (Figure \ref{fig:IDHplot}). The plot 7 from
 treatment 1 that displayed a constantly outlying diversity was removed
@@
 The functional redundancy was measured as the sum of weighted overlap
-among species in the communities. Species functional space were defined
-in the main two-dimensional space of a PCA analysis based on all leaf
-and stem and maximum height traits (see appendix I for more details).
-Plots functional redundancy compared to the initial value after logging
-so that only communities dynamics are followed apart from the impact of
-logging itself. Communities functional redundancy remained stable in
-control plots but after disturbance the redundancy trajectories were
-quite variable (See appendix I). Globally after most intense disturbance
-(plots T2 and T3) communities redundancy decreased at first place and
-then increased to edge, recover or even exceed the initial value.
-Considering the functional redundancy restricted to the functional space
-of the initial inventory, though, all disturbed plots followed similar
-decreasing humped-shaped trajectories (\ref{fig:RedFun}). The maximum
-redundancy loss observed was positively correlated with the disturbance
-intensity (XX spearman to be measured) and none of the disturbed plots
-had recovered the functional redundancy within their initial functional
-space.
+among species in communities' functional space. Functional space were
+defined in the main two-dimensional space of a PCA analysis based on all
+leaf and stem and maximum height traits (see appendix I for PCA
+details). Plots functional redundancy compared to the initial value
+after logging so that only communities dynamics are followed apart from
+the impact of logging itself. Communities functional redundancy remained
+stable in control plots but after disturbance the redundancy
+trajectories were quite variable (See appendix I). Globally after most
+intense disturbance (plots T2 and T3) communities redundancy decreased
+at first place and then increased to edge, recover or even exceed the
+initial value. Considering the functional redundancy restricted to the
+functional space of the initial inventory, though, all disturbed plots
+followed similar decreasing humped-shaped trajectories
+(@ref(fig:RedFun\_rest)). The maximum redundancy loss observed was
+positively correlated with the disturbance intensity (XX spearman to be
+measured) and none of the disturbed plots had recovered the functional
+redundancy within their initial functional space.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/RedFun-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/RedFun_rest-1}
 
@@
 
-\caption{Trajectories of the functional redundancy over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:RedFun}
+\caption{Trajectories of the functional redundancy within the initial cpùùunities functional space over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}(\#fig:RedFun_rest)
 \end{figure}

### 2018-05-14 Scripts/Functionalredundancy.R

@@ -9,7 +9,7 @@ dates<-dates[which(!dates%in%c("1996","1998","2000","2002","2004","2006","2008",
 
 Traits1<-read.csv("DB/BridgeOK.csv",sep=";",na.strings="")       # Brigde functional traits dataset
 Traits2<-read.csv("DB/DataLifeTraits.csv",sep=";",na.strings="") # Paracou15 seed and Hmax dataset
-TraitsName<-c("L_thickness","L_chloro","L_toughness","L_DryMass","SLA","WD","Bark_thick","Hmax") # Considered traits list
+TraitsName<-c("L_thickness","L_chloro","L_toughness","SLA","WD","Bark_thick","Hmax") # Considered traits list
 
 InventorySp<-do.call(rbind,lapply(LivingStand_all,function(yr){
   ret<-do.call(rbind,yr)[,c("Famille","Genre","name")]
@@ -43,7 +43,7 @@ RedundancyTraj<-lapply(1:Nrep,function(rep){
   
   ACP<-dudi.pca(scale(traits_filled[,TraitsName]),scannf=FALSE,nf=2)
   Bigacp<-merge(ACP$li,traits_filled[,c("Family","Genus","name" )],by="row.names")[,c("Family","Genus","name","Axis1","Axis2")]
-  
+
   xgen<-c(min(Bigacp[,"Axis1"]),max(Bigacp[,"Axis1"]))
   ygen<-c(min(Bigacp[,"Axis2"]),max(Bigacp[,"Axis2"]))
   
@@ -176,89 +176,29 @@ save(RedundancyTraj_restricted,file="DB/Redundancy_restricted")
 
 
 #####################################################
-load("DB/Redundancy")
+traits_filled<-Traits_filling(Traits1,Traits2,InventorySp)
+
+ACP<-dudi.pca(scale(traits_filled[,TraitsName]),scannf=FALSE,nf=2)
+ACP_Indiv<-merge(ACP$li,traits_filled[,c("Family","Genus","name" )],by="row.names")[,c("Family","Genus","name","Axis1","Axis2")]
+ACP_traits<-ACP$co
+Eigen<-as.vector(ACP$eig*100/sum(ACP$eig))
+DataACP<-list(ACP_Indiv,ACP_traits,Eigen)
+names(DataACP)<-c("Indiv","Traits","Eigen")
+
[…]

### 2018-05-14 WholePlotTrajectories.Rmd
web pages

@@
 
-```{r NMDSplans, out.width = '100%', fig.asp = 1.2/2, echo=FALSE, fig.cap="Trajectories of the plots in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) regarding the 6 leaf and stem functional traits, the maximum allometric height and seed mass class.
+```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the plots in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) regarding the 6 leaf and stem functional traits, the maximum allometric height and seed mass class.
 Plots trajectories are first represented in the two-dimensional space from the NMDS performed for the 30 years after disturbance based on Bray-Curtis dissimilarity measures between successive inventories (Upper panels (ref:Bolda) and (ref:Boldb)).
@@
 
-The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area, wood specific gravity and bark thickness and seed mass and maximum adult height) visualized with plots community weighted means (CWM) (Figure \@ref(fig:CWM)).
+The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits visualized through plots community weighted means (CWM) (Figure \@ref(fig:CWM)) of the leaves thickness, chlorophyll content, toughness and specific area, wood specific gravity, bark thickness, seed mass and maximum adult height.
 
@@
 
-### Taxonomic diversity
+### Communities taxonomic and functional diversity
 
@@
 
-For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 monitored years.
-After disturbance the richness increased when the disturbance intensity was low, with a maximum increase of 14 botanical genera (plot 3 from treatment 2).
-When the disturbance was intense however, plots taxonomic richness followed a unimodal decrease with a return to initial values after intense disturbance.
-The taxonomic evenness (Shannon and Simpson diversities) significantly increased irrespective of the disturbance intensity.
-The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum reached around 20 years and positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
-Only two T3 plots, plots 8 and 12, remained increasing 30 years after disturbance \@ref(fig:DivTaxo)), suggesting a similar be much delayed unimodal trajectory.
+For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 years monitored.
+After low disturbance intensity the richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
+After intense disturbance however, the taxonomic richness followed a unimodal decrease with a return to pre-disturbance values.
+The taxonomic evenness (Shannon and Simpson diversities) increased after disturbance, whenever its intensity.
+The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum value, reached around 20 years, positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
+Only two T3 plots (plots 8 and 12) were still increasing 30 years after disturbance \@ref(fig:DivTaxo)), suggesting a similar be much delayed unimodal trajectory.
 
-### Functional diversity
+Communities functional diversity trajectories were examined through the Rao diversity based on the 7 leaf, stem and life history traits (the seed mass was excluded as a qualitative variable).
+The plot 7 from treatment 1 that displayed a constantly outlying diversity was removed from the graphical representation for better readability, see figure A3 in appendix for full graph.
+The functional diversity of all undisturbed communities remained stable along the 30 years monitored.
+For all disturbed communities the functional diversity followed unimodal trajectories with a marked return towards the initial state (5 years after disturbance) with values comparable to those of control plots.
 
-The functional diversity trajectories of plots based on the 8 leaf, stem and life history traits was examined through the Rao diversity.
-For all undisturbed plots the functional diversity remained comparable to the reference values 5 years after disturbance (1989 inventories).
-For all disturbed plots the functional trajectories followed a unimodal trajectory with a maximum positively correlated to disturbance intensity ($\rho_{spearman}=0.73$) (Figure \@ref(fig:IDHplot)). 
-The plot 7 from treatment 1 that displayed a constantly outlying diversity was removed from the graphical representation for better readability, see appendix for full graph.
-Thirty years after disturbance all plots, whenever the initial disturbance intensity, regained diversity values similar to their initial value and to those of control plots.
+The impact of disturbance was examined specifically at 10, 20 and 30 years after disturbance through the linear correlation between the initial percentage of AGB removed and the value of communities Simpson and Rao diversities (both diversities of order 2).
 
-```{r IDHplot,  echo=FALSE, fig.cap="Upper panels, Trajectories of the Simpson taxonomic diversity (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Upper panels, Trajectories of the Simpson taxonomic diversity (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
 Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.
-Lower panels, Relationship between the initial \\%AGB removed and the values of Simpson (ref:Boldc) and Rao (ref:Boldd) diversities at three times after disturbance.
+Lower panels, Relationship between the initial \\%AGB removed and the median values of Simpson (ref:Boldc) and Rao (ref:Boldd) diversities at three times after disturbance.
 Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.", fig.env="figure*"}
-
 load("DB/SimpsonDraw");load("DB/FunctionalTraj_ForGraphs")
@@
 legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
+
 ```

### 2018-05-14 WholePlotTrajectories.tex
web pages

@@
 with consistent trajectories of the 8 functional and life history traits
-(Leaf thickness, chlorophyll content, toughness and specific area, wood
-specific gravity and bark thickness and seed mass and maximum adult
-height) visualized with plots community weighted means (CWM) (Figure
-\ref{fig:CWM}).
+visualized through plots community weighted means (CWM) (Figure
+\ref{fig:CWM}) of the leaves thickness, chlorophyll content, toughness
+and specific area, wood specific gravity, bark thickness, seed mass and
+maximum adult height.
 
@@
 
-\subsubsection{Taxonomic diversity}\label{taxonomic-diversity}
+\subsubsection{Communities taxonomic and functional
+diversity}\label{communities-taxonomic-and-functional-diversity}
 
@@
 For undisturbed plots the Richness, Shannon and Simpson diversities
-remained stable over the 30 monitored years. After disturbance the
-richness increased when the disturbance intensity was low, with a
-maximum increase of 14 botanical genera (plot 3 from treatment 2). When
-the disturbance was intense however, plots taxonomic richness followed a
-unimodal decrease with a return to initial values after intense
-disturbance. The taxonomic evenness (Shannon and Simpson diversities)
-significantly increased irrespective of the disturbance intensity. The
-evenness followed a unimodal trajectory with a just beginning return
-towards initial values and a maximum reached around 20 years and
-positively correlated to the disturbance intensity
+remained stable over the 30 years monitored. After low disturbance
+intensity the richness increased, reaching a maximum gain of 14
+botanical genera (plot 3 from treatment 2). After intense disturbance
+however, the taxonomic richness followed a unimodal decrease with a
+return to pre-disturbance values. The taxonomic evenness (Shannon and
+Simpson diversities) increased after disturbance, whenever its
+intensity. The evenness followed a unimodal trajectory with a just
+beginning return towards initial values and a maximum value, reached
+around 20 years, positively correlated to the disturbance intensity
 (\(\rho_{spearman}^{Shannon}=0.86\), and
-\(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots, plots 8 and 12,
-remained increasing 30 years after disturbance \ref{fig:DivTaxo}),
+\(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots (plots 8 and 12)
+were still increasing 30 years after disturbance \ref{fig:DivTaxo}),
 suggesting a similar be much delayed unimodal trajectory.
 
-\subsubsection{Functional diversity}\label{functional-diversity}
+Communities functional diversity trajectories were examined through the
+Rao diversity based on the 7 leaf, stem and life history traits (the
+seed mass was excluded as a qualitative variable). The plot 7 from
+treatment 1 that displayed a constantly outlying diversity was removed
+from the graphical representation for better readability, see figure A3
+in appendix for full graph. The functional diversity of all undisturbed
+communities remained stable along the 30 years monitored. For all
+disturbed communities the functional diversity followed unimodal
+trajectories with a marked return towards the initial state (5 years
+after disturbance) with values comparable to those of control plots.
 
-The functional diversity trajectories of plots based on the 8 leaf, stem
-and life history traits was examined through the Rao diversity. For all
-undisturbed plots the functional diversity remained comparable to the
-reference values 5 years after disturbance (1989 inventories). For all
-disturbed plots the functional trajectories followed a unimodal
-trajectory with a maximum positively correlated to disturbance intensity
-(\(\rho_{spearman}=0.73\)) (Figure \ref{fig:IDHplot}). The plot 7 from
-treatment 1 that displayed a constantly outlying diversity was removed
-from the graphical representation for better readability, see appendix
-for full graph. Thirty years after disturbance all plots, whenever the
-initial disturbance intensity, regained diversity values similar to
-their initial value and to those of control plots.
+The impact of disturbance was examined more accurately at 10, 20 and 30
+years after disturbance through the linear correlation between the
+initial percentage of AGB removed and the value of communities Simpson
+and Rao diversities (both diversities of order 2).
 
@@
 
-{\centering \includegraphics{WholePlotTrajectories_files/figure-latex/IDHplot-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/IDHplot-1}
 
@@
 Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.
-Lower panels, Relationship between the initial \%AGB removed and the values of Simpson \textbf{(c)} and Rao \textbf{(d)} diversities at three times after disturbance.
+Lower panels, Relationship between the initial \%AGB removed and the median values of Simpson \textbf{(c)} and Rao \textbf{(d)} diversities at three times after disturbance.
 Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.}\label{fig:IDHplot}

### 2018-05-14 Scripts/DrawingTools.R

@@ -141,16 +141,21 @@ FunDist<-function(Data_FunComp){
 }
 
 plotIDH<-function(Data,AgbLoss){
+  Ylim<-c(min(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))),
+          max(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))))
   plot(AgbLoss[,"AGB"],AgbLoss[,"AGB"],type="n",xlab="",ylab="",
-       ylim=c(min(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))),
-              max(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])})))))
-  invisible(lapply(1:3,function(T){
-    toplot<-unlist(lapply(Data, function(tr){return(tr[,time[T],"0.5"])}))
+       ylim=Ylim)
+  leg<-lapply(1:3,function(Ti){
+    toplot<-unlist(lapply(Data, function(tr){return(tr[,time[Ti],"0.5"])}))
     abs<-AgbLoss[which(AgbLoss[,"plot"]%in%names(toplot)),"AGB"]
-    points(abs,toplot,col=colyear[T])
+    points(abs,toplot,col=colyear[Ti],pch=20)
     Lm<-lm(toplot~abs)
-    abline(a=Lm$coefficients[1],b=Lm$coefficients[2],col=colyear[T],lwd=2.5)
-  }))
+    abline(a=Lm$coefficients[1],b=Lm$coefficients[2],col=colyear[Ti],lwd=2.5)
+    return(round(summary(Lm)$adj.r.squared,2))
+   #legend(round(summary(Lm)$adj.r.squared,2),x=min(AgbLoss),y=seq(Ylim[2],Ylim[1],length.out=10)[Ti],
+    #       bty="n",lty=1,col=colyear[Ti],lwd=2.5,cex=0.8)
+  })
+  legend("topleft",legend=unlist(leg),bty="n",lty=1,col=colyear,lwd=2.5,cex=0.8,title=expression(paste('R'^2,'adjusted')))
 }
 
 plotDiv<-function(Data,remove=FALSE){

### 2018-05-14 Scripts/Stats.R

@@ -34,6 +34,13 @@ mtext("initial %AGB lost",side=1,line=2.2,adj=1)
 legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
 
 
+load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")
+time<-c("1995","2005","2015")
+colyear<-c("darkgoldenrod1","darkorange2","darkred")
+Data<-Rao
+AgbLoss<-AGBloss
+#plotIDH...
+
 ##########################################
 ## Rho Spearman
 load("DB/TaxoComposition_ForGraphs")
@@ -161,7 +168,7 @@ MaxRich[order(MaxRich,decreasing = T)][1:2]
 
 
 ######################r
-### The plot 7 funcitonal diversity
+### The plot 7 functional diversity
 
 tab<-do.call(rbind,lapply(1:12,function(pl){
   P<-lapply(LivingStand_all,function(yr){if(any(names(yr)==pl)){return(yr[[which(names(yr)==pl)]])}})

### 2018-05-14 WholePlotTrajectories.Rmd
results

@@
 Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Blanc2009]. 
+The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance (ref to be found).
+Biomass measurements were performed with the BIOMASS package R package [@Biomass2018].
 
@@
 The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum value, reached around 20 years, positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
-Only two T3 plots (plots 8 and 12) were still increasing 30 years after disturbance \@ref(fig:DivTaxo)), suggesting a similar be much delayed unimodal trajectory.
+Only two T3 plots (plots 8 and 12) were still increasing 30 years after disturbance \@ref(fig:DivTaxo), suggesting a similar be much delayed unimodal trajectory.
 
@@
 
-The impact of disturbance was examined specifically at 10, 20 and 30 years after disturbance through the linear correlation between the initial percentage of AGB removed and the value of communities Simpson and Rao diversities (both diversities of order 2).
+The impact of disturbance was examined specifically at three times after disturbance (10, 20 and 30 years) through the linear correlation between the disturbance intensity (intial %AGB removed) and the value of communities Simpson and Rao diversities (diversities of order 2) \@ref(fig:IDHplot).
+The Simpson diversity was only weakly related to the disturbance intensity ($R^2<0.25$) and only from 20 years after disturbance.
+The Rao diversity was more strongly correlated to the disturbance intensity ($0.55<R^2<0.75$) for the three times studied and the impact of disturbance was observed 20 years after disturbance.
 
@@
 plotIDH(Data=Rao,AgbLoss=AGBloss)
-mtext("initial %AGB lost",side=1,line=2.2,adj=1)
+mtext("initial AGB lost (%)",side=1,line=2.2,adj=1)
 legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")

### 2018-05-14 WholePlotTrajectories.tex
results

@@
 cm DBH for treatment 3 (T3). Treatments T2 and T3 besides included the
-thinning of trees by poison girdling \citep{Blanc2009}.
+thinning of trees by poison girdling \citep{Blanc2009}. The disturbance
+intensity was measured as the percentage of aboveground biomass (\%AGB)
+lost between the first inventory in 1984 and five years after
+disturbance (ref to be found). Biomass measurements were performed with
+the BIOMASS package R package \citep{Biomass2018}.
 
@@
 \(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots (plots 8 and 12)
-were still increasing 30 years after disturbance \ref{fig:DivTaxo}),
+were still increasing 30 years after disturbance \ref{fig:DivTaxo},
 suggesting a similar be much delayed unimodal trajectory.
@@
 
-The impact of disturbance was examined more accurately at 10, 20 and 30
-years after disturbance through the linear correlation between the
-initial percentage of AGB removed and the value of communities Simpson
-and Rao diversities (both diversities of order 2).
+The impact of disturbance was examined specifically at three times after
+disturbance (10, 20 and 30 years) through the linear correlation between
+the disturbance intensity (intial \%AGB removed) and the value of
+communities Simpson and Rao diversities (diversities of order 2)
+\ref{fig:IDHplot}. The Simpson diversity was only weakly related to the
+disturbance intensity (\(R^2<0.25\)) and only from 20 years after
+disturbance. The Rao diversity was more strongly correlated to the
+disturbance intensity (\(0.55<R^2<0.75\)) for the three times studied
+and the impact of disturbance was observed 20 years after disturbance.
 

### 2018-05-15 AppendixI.Rmd
Results & Appendix

@@
 output: 
+  bookdown::gitbook: default
   bookdown::pdf_book:
@@
 
-**Figure A3**: Trajectories of the Rao functional diversity over 30 years after disturbance.
+**Figure A3**: Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.
-\@ref(fig:DivFunAppendix).
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3 \@ref(fig:DivTaxo).
 
@@
 Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.
+\@ref(fig:DivFun).
+
+**Figure A5**: Summarizing plots of the functional PCA analysis defining communities functional space.
+**(a)** Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
+**(b)** bow plot of the PCA eignevalues in percentage.
+**(c)** Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).
 \@ref(fig:PCA_graphs).
 
-**Figure A5**: Trajectories of the Rao functional diversity over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.
+**Figure A6**: Trajectories of the functional redundancy over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.
 \@ref(fig:RedFun).
@@
 
-# **Figure A3**: Rao diversity trajectories with all plots display
+# **Figure A3**: Richness, Shannon and Simpson taxonomic diversity trajectories
 
-```{r DivFunAppendix, out.width = '60%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
+```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
+load("DB/ReplacementTraj_ForGraphs")
+
+TaxoTraj(CompleteTaxo)
+
+```
+
+# **Figure A4**: Rao diversity trajectories with all plots display
+
+```{r DivFun, out.width = '60%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 
-# **Figure A4**: Functional PCA
+# **Figure A5**: Functional PCA
 
@@
 
-# **Figure A5**: Overall functional redundancy
+# **Figure A6**: Overall functional redundancy
 

### 2018-05-15 AppendixI.tex
Results & Appendix

@@
 
-\textbf{Figure A3}: Trajectories of the Rao functional diversity over 30
-years after disturbance. Trajectories correspond to the median (solid
-line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
-iteration of the taxonomic uncertainty propagation. Initial treatments
-are represented by solid lines colorswith green for control, blue for
-T1,orange for T2 and red for T3.and the missing trait value filling
-processes. \ref{fig:DivFunAppendix}.
+\textbf{Figure A3}: Trajectories of the difference to the 1989
+inventories (5 years after disturbance) over 30 years after disturbance
+of plots communities \textbf{(a)} Richness, \textbf{(b)} Shannon and
+\textbf{(c)} Simpson diversities. Trajectories correspond to the median
+(solid line) and 0.025 and 0.975 percentile (gray envelope) observed
+after 50 iteration of the taxonomic uncertainty propagation. Initial
+treatments are represented by solid lines colors with green for control,
+blue for T1,orange for T2 and red for T3 \ref{fig:DivTaxo}.
 
@@
 T1,orange for T2 and red for T3.and the missing trait value filling
-processes. @ref(fig:PCA\_graphs).
+processes. \ref{fig:DivFun}.
 
-\textbf{Figure A5}: Trajectories of the Rao functional diversity over 30
+\textbf{Figure A5}: Summarizing plots of the functional PCA analysis
+defining communities functional space. \textbf{(a)} Plot of the samples
+from the BRIDGE functional database in the two dimensional main plan.
+\textbf{(b)} bow plot of the PCA eignevalues in percentage. \textbf{(c)}
+Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem
+traits, and maximum species height). @ref(fig:PCA\_graphs).
+
+\textbf{Figure A6}: Trajectories of the functional redundancy over 30
 years after disturbance. Trajectories correspond to the median (solid
 line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
-iteration of the taxonomic uncertainty propagation. Initial treatments
-are represented by solid lines colorswith green for control, blue for
-T1,orange for T2 and red for T3.and the missing trait value filling
-processes. \ref{fig:RedFun}.
+iteration of the taxonomic uncertainty propagation and the missing trait
+value filling processes. Initial treatments are represented by solid
+lines colors with green for control, blue for T1,orange for T2 and red
+for T3. \ref{fig:RedFun}.
 
@@
 
-\section{\texorpdfstring{\textbf{Figure A3}: Rao diversity trajectories
-with all plots
-display}{Figure A3: Rao diversity trajectories with all plots display}}\label{figure-a3-rao-diversity-trajectories-with-all-plots-display}
+\section{\texorpdfstring{\textbf{Figure A3}: Richness, Shannon and
+Simpson taxonomic diversity
+trajectories}{Figure A3: Richness, Shannon and Simpson taxonomic diversity trajectories}}\label{figure-a3-richness-shannon-and-simpson-taxonomic-diversity-trajectories}
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{AppendixI_files/figure-latex/DivFunAppendix-1}
+{\centering \includegraphics[width=1\linewidth]{AppendixI_files/figure-latex/DivTaxo-1}
+
+}
+
+\caption{Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities \textbf{(a)} Richness, \textbf{(b)} Shannon and \textbf{(c)} Simpson diversities.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:DivTaxo}
+\end{figure*}
+
+\section{\texorpdfstring{\textbf{Figure A4}: Rao diversity trajectories
+with all plots
+display}{Figure A4: Rao diversity trajectories with all plots display}}\label{figure-a4-rao-diversity-trajectories-with-all-plots-display}
+
+\begin{figure*}
+
+{\centering \includegraphics[width=0.6\linewidth]{AppendixI_files/figure-latex/DivFun-1}
 
@@
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:DivFunAppendix}
+Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:DivFun}
 \end{figure*}
 
-\section{\texorpdfstring{\textbf{Figure A4}: Functional
-PCA}{Figure A4: Functional PCA}}\label{figure-a4-functional-pca}
+\section{\texorpdfstring{\textbf{Figure A5}: Functional
+PCA}{Figure A5: Functional PCA}}\label{figure-a5-functional-pca}
 
@@
 
-\section{\texorpdfstring{\textbf{Figure A5}: Overall functional
-redundancy}{Figure A5: Overall functional redundancy}}\label{figure-a5-overall-functional-redundancy}
+\section{\texorpdfstring{\textbf{Figure A6}: Overall functional
+redundancy}{Figure A6: Overall functional redundancy}}\label{figure-a6-overall-functional-redundancy}
 

### 2018-05-15 WholePlotTrajectories.Rmd
Results & Appendix

@@
 To counter the remaining taxonomic uncertainty, plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
-Variations in plots taxonomic and functional composition after disturbance was visualized by their trajectories in a two-dimensional ordination space over 30 years.
-Two NMDS were conducted, either from taxonomic flora inventories or from plots functional composition based on the 8 leaf, stem and life history traits. 
-Plots trajectories along time was reported comparatively to the inventories in 1989, 5 years after disturbance, which corresponded to first inventory with a sufficient degree of uncertainty (<30% of undetermined trees).
-The inventories dissimilarity compared to the reference 1989 inventory was reported using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) similarity measures. 
-The trajectory of inventories along time was visualized with the euclidean distance in the two-dimensional ordination space to the 1989 inventory.
+Variations in plots taxonomic and functional composition after disturbance was visualized by their trajectories over 30 years in a two-dimensional ordination space.
+Two NMDS were conducted, either from taxonomic flora inventories or from plots functional composition based on the 7 leaf, stem and life history traits (without seed mass classes).
+In both cases the NMDS were performed using occurrence-based (Jaccard) as well as abundance-based (Bray-Curtis) dissimilarity measures. 
+Trajectories along time were reported through the euclidean distance of successive inventories to the reference inventories in 1989, 5 years after disturbance, when the uncertainty degree did not exceed 30% of undetermined trees.
 The trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004].
@@
 
-## Disturbance impacts on communities composition and average functional traits values
+## Communities Diversity
 
-### Composition trajectories
-
-Over time, 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
-Taxonomic and functional plots trajectories were examined through the trajecotries in a two dimensional ordination space of successive inventories from 1989 (5 years after disturbance) to 2015 (31 years after disturbance).
-Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity, this last having given similar results (data not shown).
-
-Both taxonomic and functional composition of disturbed plots followed consistent trajectories over time, translating compositional changes while control plots remained stable.
-The addition of functional traits in the two dimensional space revealed a shift of plots composition towards more acquisitive functional strategies (disturbed plots changed from high average WD to high average SLA and chlorophyll content).
-The dissimilarity of the successive inventories for disturbed plots compared to their respective reference inventory in 1989 followed unimodal trajectories (AppendixI, Figure A2).
-The maximum dissimilarity of plot strajectories was positively correlated to the disturbance intensity for taxonomic ($\rho_{spearman}^{taxonomic}=0.91$) as well as functional composition ($\rho_{spearman}^{taxonomic}=0.96$).
-The time at maximum dissimilarity for taxonomic composition was reached around 26 years after disturbance while it was 22 years for functional composition. 
-All trajectories returned towards the initial composition and translated cyclic compositional changes as visualized in figure (Figure \@ref(fig:NMDSplans)).
-
-```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the plots in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) regarding the 6 leaf and stem functional traits, the maximum allometric height and seed mass class.
-Plots trajectories are first represented in the two-dimensional space from the NMDS performed for the 30 years after disturbance based on Bray-Curtis dissimilarity measures between successive inventories (Upper panels (ref:Bolda) and (ref:Boldb)).
-Then the lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.", fig.env="figure*"}
-load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")
-load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")
-
-par(mfrow=c(2,2))
-TaxoCompo(MatrepTaxo)
-FunCompo(MatrepFun)
-TaxoDist(MatrepTaxo)
-FunDist(MatrepFun)
-
-```
-
-### Traits community weighted means
-
-The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits visualized through plots community weighted means (CWM) (Figure \@ref(fig:CWM)) of the leaves thickness, chlorophyll content, toughness and specific area, wood specific gravity, bark thickness, seed mass and maximum adult height.
-
-Except for leaf cholorphyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions displayed a unimodal trajectories.
-The weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value and than these of the control plots (Figure \@ref(fig:CWM)).
-The weighted means of bark thickness (*Bark_thick*) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value and this of the undisturbed plots at the end of the experiment.
-Once again the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.67$, $\rho_{spearman}^{L_{chloro}}=0.45$, $\rho_{spearman}^{L_{toughness}}=-0.43$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WD}=-0.78$, $\rho_{spearman}^{Bark-thickness}=0.88$, $\rho_{spearman}^{Hmax}=-0.48$).
-
-```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWD), and bark thickness, (ref:capBarkThick)) and one life trait (Specific maximum height at adult stage, (ref:capHmax)).
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
-load("DB/CWM")
-
-CWMdraw(CWM)
-legendCWM()
-
-```
-
-## Trajectories of communities diversity after disturbance
-
-### Communities taxonomic and functional diversity
-
-Trajectories of Richness, Shannon and Simpson taxonomic diversity were examined at genus level in relation to the 1989 inventories (5 years after disturbance) (Figure \@ref(fig:DivTaxo)).
-
-```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
-load("DB/ReplacementTraj_ForGraphs")
-
-TaxoTraj(CompleteTaxo)
-
-```
-
+Communities taxonomic diversity trajectories were examined through the Richness, Shannon and Simpson diversities at genus level, in relation to the 1989 inventories (5 years after disturbance) (See annexe I).
 For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 years monitored.
 After low disturbance intensity the richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
-After intense disturbance however, the taxonomic richness followed a unimodal decrease with a return to pre-disturbance values.
-The taxonomic evenness (Shannon and Simpson diversities) increased after disturbance, whenever its intensity.
-The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum value, reached around 20 years, positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
-Only two T3 plots (plots 8 and 12) were still increasing 30 years after disturbance \@ref(fig:DivTaxo), suggesting a similar be much delayed unimodal trajectory.
+After intense disturbance, plots taxonomic richness followed unimodal trajectories first decrease for around ten years and then increasing to return to pre-disturbance values.
+For all disturbed plots the taxonomic evenness (Shannon and Simpson diversities) increased, following unimodal trajectories with a just beginning return towards initial values after 30 years.
+The maximum evenness, reached after around 20 years, was positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
+Only two T3 plots (plots 8 and 12) were still increasing 30 years after disturbance \@ref(fig:IDHplot), suggesting similar but delayed trajectories.
 
 Communities functional diversity trajectories were examined through the Rao diversity based on the 7 leaf, stem and life history traits (the seed mass was excluded as a qualitative variable).
-The plot 7 from treatment 1 that displayed a constantly outlying diversity was removed from the graphical representation for better readability, see figure A3 in appendix for full graph.
-The functional diversity of all undisturbed communities remained stable along the 30 years monitored.
-For all disturbed communities the functional diversity followed unimodal trajectories with a marked return towards the initial state (5 years after disturbance) with values comparable to those of control plots.
+The plot 7 from treatment 1 displayed a constantly outlying diversity and was removed from the graphical representation for better readability (see appendix for full graphs).
+The functional diversity of all undisturbed communities remained stable along the 30 years.
+For all disturbed communities the functional diversity followed unimodal trajectories with a return towards the initial state, comparable to the control plots.
 
-The impact of disturbance was examined specifically at three times after disturbance (10, 20 and 30 years) through the linear correlation between the disturbance intensity (intial %AGB removed) and the value of communities Simpson and Rao diversities (diversities of order 2) \@ref(fig:IDHplot).
-The Simpson diversity was only weakly related to the disturbance intensity ($R^2<0.25$) and only from 20 years after disturbance.
-The Rao diversity was more strongly correlated to the disturbance intensity ($0.55<R^2<0.75$) for the three times studied and the impact of disturbance was observed 20 years after disturbance.
+The impact of disturbance was examined specifically through the linear correlation between the disturbance intensity (intial %AGB removed) and the Simpson and Rao diversities (diversities of order 2) at three focus times after disturbance (10, 20 and 30 years) \@ref(fig:IDHplot).
+The Simpson diversity was weakly related to the disturbance intensity ($R^2<0.25$) and only from 20 years after disturbance.
+The Rao diversity was more strongly correlated to the disturbance intensity ($0.55<R^2<0.75$) for the three times studied.
+The impact of disturbance, translated by the slope of the linear correlation, was the highest 20 years after disturbance.
 
@@
 
-The functional redundancy was measured as the sum of weighted overlap among species in communities' functional space.
-Functional space were defined in the main two-dimensional space of a PCA analysis based on all leaf and stem and maximum height traits (see appendix I for PCA details). 
-Plots functional redundancy compared to the initial value after logging so that only communities dynamics are followed apart from the impact of logging itself.
+## Communities composition
+
+### Taxonomic and functional trajectories
+
+In the inventories from 1989 (5 years after disturbance) to 2015 (31 years after disturbance), 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
+The trajectories of taxonomic and functional composition were visualized in a two dimensional ordination space mapping the successive inventories according to their flora and the corresponding traits.
+Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity.
+Both metrics gave similar results, so only analysis using Bray-Curtis dissimilarity will be discussed here.
+
+While both taxonomic and functional composition remained stable in undistrubed communities, after disturbed they followed consistent trajectories over time that revealed significant compositional changes.
+Considering the mapping of functional traits in the same dimensional space, the post-disturbance trajectories corresponded to shifts towards more acquisitive functional strategies: disturbed communities changed from high average WD to high average SLA and chlorophyll content.
+For disturbed communities the distance of successive inventories to the reference inventory in 1989 followed unimodal trajectories (AppendixI, Figure A2).
+The maximum of communities dissimilarity to their initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.91$ and $\rho_{spearman}^{functional}=0.96$ respectively).
+The time at maximum dissimilarity was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition. 
+All trajectories followed cyclic composition changes with a return towards the initial composition (Figure \@ref(fig:NMDSplans)).
+
+```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the plots in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) regarding the 6 leaf and stem functional traits, the maximum allometric height and seed mass class.
+Plots trajectories are first represented in the two-dimensional space from the NMDS performed for the 30 years after disturbance based on Bray-Curtis dissimilarity measures between successive inventories (Upper panels (ref:Bolda) and (ref:Boldb)).
+Then the lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years.
+Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
+The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.", fig.env="figure*"}
+load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")
+load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")
+
+par(mfrow=c(2,2))
+TaxoCompo(MatrepTaxo)
+FunCompo(MatrepFun)
+TaxoDist(MatrepTaxo)
+FunDist(MatrepFun)
+
+```
+
+### Traits community weighted means (CWM)
+
+The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits visualized through plots community weighted means (CWM) (Figure \@ref(fig:CWM)) of the leaves thickness, chlorophyll content, toughness and specific area, wood specific gravity, bark thickness, seed mass and maximum adult height.
+
+Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots, all traits and seed mass proportions followed a unimodal trajectories and either stabilized or returned towards their initial values 30 years after disturbance.
+At that time the weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value and than these of the control plots (Figure \@ref(fig:CWM)).
+The weighted means of bark thickness (*Bark_thick*) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value, similar as those of undisturbed plots.
+For all traits the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.67$, $\rho_{spearman}^{L_{chloro}}=0.45$, $\rho_{spearman}^{L_{toughness}}=-0.43$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WD}=-0.78$, $\rho_{spearman}^{Bark-thickness}=0.88$, $\rho_{spearman}^{Hmax}=-0.48$).
+
+```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWD), and bark thickness, (ref:capBarkThick)) and one life trait (Specific maximum height at adult stage, (ref:capHmax)).
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
+load("DB/CWM")
+
+CWMdraw(CWM)
+legendCWM()
+
+```
+
+### Functional redundancy
+
+Communities functional redundancy was measured as the sum of weighted overlap among species in communities' functional space.
+Functional spaces were defined in the main two-dimensional space of a PCA analysis summarizing the 7 leaf, stem, and maximum height traits (see appendix I for PCA details). 
 Communities functional redundancy remained stable in control plots but after disturbance the redundancy trajectories were quite variable (See appendix I).
-Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place and then increased to edge, recover or even exceed the initial value.
-Considering the functional redundancy restricted to the functional space of the initial inventory, though, all disturbed plots followed similar decreasing humped-shaped trajectories (\@ref(fig:RedFun_rest)).
-The maximum redundancy loss observed was positively correlated with the disturbance intensity (XX spearman to be measured) and none of the disturbed plots had recovered the functional redundancy within their initial functional space.
+Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place before increasing to edge, recover or exceed the initial value.
+
+Considering the functional redundancy restricted to the functional space of the initial inventory, all disturbed plots followed similar decreasing humped-shaped trajectories (\@ref(fig:RedFun_rest)).
+The maximum redundancy loss observed was positively correlated with the disturbance intensity (XX spearman to be measured) and none of the disturbed plots had recovered the functional redundancy within the initial functional space.
 

### 2018-05-15 Scripts/Functionalredundancy.R

@@ -1,4 +1,4 @@
-library("MASS")
+library("MASS");library("plotrix")
 source("Scripts/Vernacular_handle.R")           # Script for vernacular names replacement
 source("Scripts/TraitsMiceFilling.R")            # Traits Gap-filling
 
@@ -180,6 +180,7 @@ traits_filled<-Traits_filling(Traits1,Traits2,InventorySp)
 
 ACP<-dudi.pca(scale(traits_filled[,TraitsName]),scannf=FALSE,nf=2)
 ACP_Indiv<-merge(ACP$li,traits_filled[,c("Family","Genus","name" )],by="row.names")[,c("Family","Genus","name","Axis1","Axis2")]
+ACP_Indiv[which(ACP_Indiv[,"name"]=="Lecythis_corrugata subsp. corrugata"),"name"]<-"Lecythis_corrugata"
 ACP_traits<-ACP$co
 Eigen<-as.vector(ACP$eig*100/sum(ACP$eig))
 DataACP<-list(ACP_Indiv,ACP_traits,Eigen)

### 2018-05-15 AppendixI.Rmd
Discussion

@@
 output: 
-  bookdown::gitbook: default
   bookdown::pdf_book:

### 2018-05-15 WholePlotTrajectories.Rmd
Discussion

@@
 
+## Decoupled recovery of communities taxonomic and functional characteristics
+
+Both communities taxonomic and functional diversity and composition proved resilient, following similar humped-back trajectories with a return towards initial values.
+The resilience of communities functional characteristics, the most direct link between biodiversity and ecosystem functioning [@Diaz2005], meant a consistent recovery of ecosystem processes in the long term [@Guariguata2001].
+The resilience of communities taxonomic characteristics, in turn, meant the maintenance of their initial differences in composition and structure.
+It suggested that communities response to disturbance were somehow constraint to converge towards determined compositions [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
+
+Although both communities taxonomic and functional characteristics proved resilient and followed similar humped-back trajectories, the taxonomic recovery systematically lagged behind the corresponding functional dynamics.
+Such delay between functional and taxonomic dynamics has already been observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
+According to the "vegetation quantity effect" [@Grime1998] functional trajectories rely on the pool of dominant species, which diversity and evenness were enhanced after disturbance, and which rapidly restored the functional diversity and composition.
+However, communities evenness remained high so infrequent species still missed to the taxonomic recovery that mechanically lagged behind, all the more so that unrecovered species would be functionally redundant and probably undergo competition processes.
+
 ## A validation of the intermediate disturbance hypothesis
 
-The monitoring of disturbed forest communities confirmed a limited negative impact of disturbance on species richness, as observed on several post logging surveys [@Cannon1998;@Baraloto2012a].
-Thirty years after disturbance, the genus richness was restored to initial plot values after high disturbance intensity while it substantially increased after low disturbance intensity, reaching a gain of almost 12 genera for some plots.
+Validating the IDH, communities trajectories confirmed the diversity increase after disturbance through an enhanced growth of otherwise less favored species.
+This was however completely clear regarding the functional diversity but much more blurry for communities taxonomic characteristics.
 
-Both richness and evenness followed asymptotic trajectories after disturbance, sharply increasing for 15 years before stabilizing at higher values than those of control plots.
-An increase of communities evenness stems from a higher homogeneity of species distribution.
-After disturbance, communities are made of the old, pre-disturbance survivors and the newly recruited trees: changes in composition and abundance distribution are to be found in the recruitment processes and in the pre-disturbance survivors mortality.
-Because the composition of old survivors proved to mirror the initial community [@Herault2018], a specific turnover was expected among recruited trees with enhanced growth and survival of previously infrequent species.
-Indeed, the increase in taxonomic diversity was accompanied by an increase of taxonomic dissimilarity with plots initial state and a functional shift towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
-Disturbance then causes a reorganization of the typical high dominance structure of hyperdiverse mature forests after disturbance, benefiting to pioneers and light demanding species. 
+The taxonomic richness was weakly or negatively impacted by intense disturbance, as observed on several post logging surveys [@Cannon1998; @Baraloto2012a], while it substantially increased after low intensity disturbance.
+Disturbance also enhanced taxonomic evenness, but its intensity was only weakly and lately correlated to communities eveness (*i.e.* Simpson diversity).
+Contrastingly, disturbance intensity consistently predicted the significant increase of communities functional diversity for 30 years.
+Disturbance impacts involve specific turnover within communities, either among pre-disturbance survivors or among newly recruited trees.
+As the composition of old-growth survivors proved to mirror initial communities [@Herault2018], disturbance would impact trees recruitment through the enhanced growth and survival of previously infrequent species and functional characteristics.
+Consistently, disturbance resulted in increasing taxonomic dissimilarity compared to pre-disturbance communities and functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
+Disturbance then caused a reorganization of the typical high dominance structure of hyperdiverse mature forests after disturbance, benefiting to pioneers and light demanding species. 
 Likely, the changes in abiotic environment and competitive pressure favored pioneers which outcompete others in non limiting resources but are excluded in mature forests by long-lived, more resistant and shade tolerant species.
-
-As stated by the IDH, communities dynamics after disturbance relied on species functional strategy and corresponding ability to fill the environmental niches made available by disturbance.
+Therefore, consistently with communities dynamics after disturbance relied on species functional strategy and corresponding ability to fill the environmental niches made available by disturbance.
 Recruited species then mixed with pre-disturbance ones, from which they differed, and constituted a community all the more diversified that the disturbance was intense [@Molino2001].
 
-## On the recovery of disturbed communities
+## Functional redundancy of disturbed communities
 
-All diversity and composition trajectories followed similar hump-back trajectories returning towards their initial values.
-Communities trajectories therefore illustrated their taxonomic and function resilience.
+both the lag between taxonomic and functional recovery and the much better consistency of the IDH when the functional diversity is considered rely on the central role of functional redundancy.
+The functional redundancy that is the overlap between species, typical of the huge biodiversity of tropical forests [@Bellwood2006] and determinant for communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005], was not recovered 30 years after disturbance.
 
-Because functional traits and functional diversity are the most direct link between biodiversity and ecosystem functioning [@Diaz2005], their resilience meant a consistent recovery of ecosystem processes in the long term [@Guariguata2001].
-The resilience of the taxonomic diversity and composition similarly translated  the recovery of communities such as before disturbance.
-This resilience meant the maintenance of communities initial differences in composition and supported the idea of their convergence towards a determined state [@Hubbell1999;@Molino2001;@Baraloto2012a].
-If all plots followed similar hump-back trajectories, though, some significant differences emerged.
-To name but one, the taxonomic evenness of one plot displayed an extensive increase until 30 years after disturbance.
-The differences trajectories maxima and time path, along with the fact intrinsinc communities differences are maintained, suggested that the trajectories depend on the initial plots composition or on some abiotic parameters [@Anderson2007].
-
-## Functional redundancy of disturbed ecosystems
-
-Despite the recovery of pre-disturbance state there was a time lag between the trajectories followed by communities taxonomic and functional characteristics.
-The trajectory of taxonomic composition and evenness had a longer time-path and remained altered while communities functional composition and diversity had already recovered.
-
-Such delay between functional and taxonomic dynamics was already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
-According to the "vegetation quantity effect" [@Grime1998], the functional diversity of communities rely on the dominant species.
-Communities functional trajectories were then driven first by the increase in diversity and evenness of dominant species, which restored the pre-disturbance functional characteristics.
-Then, the functional trajectories were driven by the recruitment of species resembling the old pre-disturbance community, which reduced the overall diversity.
-At that time, the taxonomic composition and diversity remained altered: the still missing species were then the initially infrequent and functionally redundant ones.
-The functional redundancy, the functional overlap between species that is typical of the huge biodiversity of tropical forests [@Bellwood2006], was then not restored 30 years after disturbance and this is major to consider as it defines forests' resilience [@Trenbath1999;@Elmqvist2003;@Diaz2005].
-
+Disturbance globally reduced the overall functional redundancy of communities but this restored rapidly without consistent rtajectories along time.
+The initial functional space corresponding to undistrubed 
 Besides the long-term alteration of functional redundancy, there was probably persistent compositional changes favoring disturbance resistant species [@Haddad2008], lianas or epiphytes [@Martin2013] and environmental changes, like in the soils nutrient cycling and compaction [@Olander2005].

### 2018-05-15 WholePlotTrajectories.tex
Discussion

--- 2018-05-14 WholePlotTrajectories.tex
@@
 Variations in plots taxonomic and functional composition after
-disturbance was visualized by their trajectories in a two-dimensional
-ordination space over 30 years. Two NMDS were conducted, either from
+disturbance was visualized by their trajectories over 30 years in a
+two-dimensional ordination space. Two NMDS were conducted, either from
 taxonomic flora inventories or from plots functional composition based
-on the 8 leaf, stem and life history traits. Plots trajectories along
-time was reported comparatively to the inventories in 1989, 5 years
-after disturbance, which corresponded to first inventory with a
-sufficient degree of uncertainty (\textless{}30\% of undetermined
-trees). The inventories dissimilarity compared to the reference 1989
-inventory was reported using occurrence-based (Jaccard) and
-abundance-based (Bray-Curtis) similarity measures. The trajectory of
-inventories along time was visualized with the euclidean distance in the
-two-dimensional ordination space to the 1989 inventory. The trajectories
-of the leaf and stem and life traits were also visualized with the
-community weighted means (CWM), representing the average trait value in
-a community weighted by relative abundance of the species carrying each
-value \citep{Diaz2007, Garnier2004}. To compensate the intrinsinc
-difference among plots the trajectories corresponded to the differences
-along time with the reference inventory in 1989. Species seed mass
-corresponded to 5 classes of increasing mass, seed mass trajectories
-were therefore reported as the proportion of each class in the
-inventories.
+on the 7 leaf, stem and life history traits (without seed mass classes).
+In both cases the NMDS were performed using occurrence-based (Jaccard)
+as well as abundance-based (Bray-Curtis) dissimilarity measures.
+Trajectories along time were reported through the euclidean distance of
+successive inventories to the reference inventories in 1989, 5 years
+after disturbance, when the uncertainty degree did not exceed 30\% of
+undetermined trees. The trajectories of the leaf and stem and life
+traits were also visualized with the community weighted means (CWM),
+representing the average trait value in a community weighted by relative
+abundance of the species carrying each value
+\citep{Diaz2007, Garnier2004}. To compensate the intrinsinc difference
+among plots the trajectories corresponded to the differences along time
+with the reference inventory in 1989. Species seed mass corresponded to
+5 classes of increasing mass, seed mass trajectories were therefore
+reported as the proportion of each class in the inventories.
 
@@
 
-\subsection{Disturbance impacts on communities composition and average
-functional traits
-values}\label{disturbance-impacts-on-communities-composition-and-average-functional-traits-values}
+\subsection{Communities Diversity}\label{communities-diversity}
 
-\subsubsection{Composition trajectories}\label{composition-trajectories}
+Communities taxonomic diversity trajectories were examined through the
+Richness, Shannon and Simpson diversities at genus level, in relation to
+the 1989 inventories (5 years after disturbance) (See annexe I). For
+undisturbed plots the Richness, Shannon and Simpson diversities remained
+stable over the 30 years monitored. After low disturbance intensity the
+richness increased, reaching a maximum gain of 14 botanical genera (plot
+3 from treatment 2). After intense disturbance, plots taxonomic richness
+followed unimodal trajectories first decrease for around ten years and
+then increasing to return to pre-disturbance values. For all disturbed
+plots the taxonomic evenness (Shannon and Simpson diversities)
+increased, following unimodal trajectories with a just beginning return
+towards initial values after 30 years. The maximum evenness, reached
+after around 20 years, was positively correlated to the disturbance
+intensity (\(\rho_{spearman}^{Shannon}=0.86\), and
+\(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots (plots 8 and 12)
+were still increasing 30 years after disturbance \ref{fig:IDHplot},
+suggesting similar but delayed trajectories.
 
-Over time, 828388 trees and 591 botanical species pertaining to 223
-genus and 64 botanical families were recorded. Taxonomic and functional
-plots trajectories were examined through the trajecotries in a two
-dimensional ordination space of successive inventories from 1989 (5
-years after disturbance) to 2015 (31 years after disturbance).
-Classifications were performed using either abundance-based Bray-Curtis
-(Figure \ref{fig:NMDSplans}) or incidence-based Jaccard dissimilarity,
-this last having given similar results (data not shown).
+Communities functional diversity trajectories were examined through the
+Rao diversity based on the 7 leaf, stem and life history traits (the
+seed mass was excluded as a qualitative variable). The plot 7 from
+treatment 1 displayed a constantly outlying diversity and was removed
+from the graphical representation for better readability (see appendix
+for full graphs). The functional diversity of all undisturbed
+communities remained stable along the 30 years. For all disturbed
+communities the functional diversity followed unimodal trajectories with
+a return towards the initial state, comparable to the control plots.
 
-Both taxonomic and functional composition of disturbed plots followed
-consistent trajectories over time, translating compositional changes
-while control plots remained stable. The addition of functional traits
-in the two dimensional space revealed a shift of plots composition
-towards more acquisitive functional strategies (disturbed plots changed
-from high average WD to high average SLA and chlorophyll content). The
-dissimilarity of the successive inventories for disturbed plots compared
-to their respective reference inventory in 1989 followed unimodal
-trajectories (AppendixI, Figure A2). The maximum dissimilarity of plot
-strajectories was positively correlated to the disturbance intensity for
-taxonomic (\(\rho_{spearman}^{taxonomic}=0.91\)) as well as functional
-composition (\(\rho_{spearman}^{taxonomic}=0.96\)). The time at maximum
-dissimilarity for taxonomic composition was reached around 26 years
-after disturbance while it was 22 years for functional composition. All
-trajectories returned towards the initial composition and translated
-cyclic compositional changes as visualized in figure (Figure
-\ref{fig:NMDSplans}).
+The impact of disturbance was examined specifically through the linear
+correlation between the disturbance intensity (intial \%AGB removed) and
+the Simpson and Rao diversities (diversities of order 2) at three focus
+times after disturbance (10, 20 and 30 years) \ref{fig:IDHplot}. The
+Simpson diversity was weakly related to the disturbance intensity
+(\(R^2<0.25\)) and only from 20 years after disturbance. The Rao
+diversity was more strongly correlated to the disturbance intensity
+(\(0.55<R^2<0.75\)) for the three times studied. The impact of
+disturbance, translated by the slope of the linear correlation, was the
+highest 20 years after disturbance.
+
+\begin{figure*}
+
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/IDHplot-1}
+
+}
+
+\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.
+Lower panels, Relationship between the initial \%AGB removed and the median values of Simpson \textbf{(c)} and Rao \textbf{(d)} diversities at three times after disturbance.
+Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.}\label{fig:IDHplot}
+\end{figure*}
+
+\subsection{Communities composition}\label{communities-composition}
+
+\subsubsection{Taxonomic and functional
+trajectories}\label{taxonomic-and-functional-trajectories}
+
+In the inventories from 1989 (5 years after disturbance) to 2015 (31
+years after disturbance), 828388 trees and 591 botanical species
+pertaining to 223 genus and 64 botanical families were recorded. The
+trajectories of taxonomic and functional composition were visualized in
+a two dimensional ordination space mapping the successive inventories
+according to their flora and the corresponding traits. Classifications
+were performed using either abundance-based Bray-Curtis (Figure
+\ref{fig:NMDSplans}) or incidence-based Jaccard dissimilarity. Both
+metrics gave similar results, so only analysis using Bray-Curtis
+dissimilarity will be discussed here.
+
+While both taxonomic and functional composition remained stable in
+undistrubed communities, after disturbed they followed consistent
+trajectories over time that revealed significant compositional changes.
+Considering the mapping of functional traits in the same dimensional
+space, the post-disturbance trajectories corresponded to shifts towards
+more acquisitive functional strategies: disturbed communities changed
+from high average WD to high average SLA and chlorophyll content. For
+disturbed communities the distance of successive inventories to the
+reference inventory in 1989 followed unimodal trajectories (AppendixI,
+Figure A2). The maximum of communities dissimilarity to their initial
+state was positively correlated to the disturbance intensity for both
+taxonomic and functional composition
+(\(\rho_{spearman}^{taxonomic}=0.91\) and
+\(\rho_{spearman}^{functional}=0.96\) respectively). The time at maximum
+dissimilarity was reached around 26 years after disturbance for
+taxonomic composition and 22 years for functional composition. All
+trajectories followed cyclic composition changes with a return towards
+the initial composition (Figure \ref{fig:NMDSplans}).
 
@@
 
-\subsubsection{Traits community weighted
-means}\label{traits-community-weighted-means}
+\subsubsection{Traits community weighted means
+(CWM)}\label{traits-community-weighted-means-cwm}
 
@@
 
-Except for leaf cholorphyll content, which continued to increase for
-some T3 and T2 plots 30 years after disturbance, all traits and seed
-mass proportions displayed a unimodal trajectories. The weighted means
-of communities specific maximum height at adult stage (\emph{Hmax}),
-leaf toughness (\emph{L\_toughness}) and wood specific gravity
-(\emph{WD}) remained significantly lower than their initial value and
-than these of the control plots (Figure \ref{fig:CWM}). The weighted
-means of bark thickness (\emph{Bark\_thick}) similarly remained
+Except for leaf chlorophyll content, which continued to increase for
+some T3 and T2 plots, all traits and seed mass proportions followed a
+unimodal trajectories and either stabilized or returned towards their
+initial values 30 years after disturbance. At that time the weighted
+means of communities specific maximum height at adult stage
+(\emph{Hmax}), leaf toughness (\emph{L\_toughness}) and wood specific
+gravity (\emph{WD}) remained significantly lower than their initial
+value and than these of the control plots (Figure \ref{fig:CWM}). The
+weighted means of bark thickness (\emph{Bark\_thick}) similarly remained
 substantilly higher than initially for all disturbed plots while the
-specfic leaf area (\emph{SLA}) had almost recovered its initial value
-and this of the undisturbed plots at the end of the experiment. Once
-again the maximum difference to initial state was correlated to the
-disturbance intensity (\(\rho_{spearman}^{L_{thickness}}=0.67\),
+specfic leaf area (\emph{SLA}) had almost recovered its initial value,
+similar as those of undisturbed plots. For all traits the maximum
+difference to initial state was correlated to the disturbance intensity
+(\(\rho_{spearman}^{L_{thickness}}=0.67\),
 \(\rho_{spearman}^{L_{chloro}}=0.45\),
@@
 
-\subsection{Trajectories of communities diversity after
-disturbance}\label{trajectories-of-communities-diversity-after-disturbance}
+\subsubsection{Functional redundancy}\label{functional-redundancy}
 
-\subsubsection{Communities taxonomic and functional
-diversity}\label{communities-taxonomic-and-functional-diversity}
+Communities functional redundancy was measured as the sum of weighted
+overlap among species in communities' functional space. Functional
+spaces were defined in the main two-dimensional space of a PCA analysis
+summarizing the 7 leaf, stem, and maximum height traits (see appendix I
+for PCA details). Communities functional redundancy remained stable in
+control plots but after disturbance the redundancy trajectories were
+quite variable (See appendix I). Globally after most intense disturbance
+(plots T2 and T3) communities redundancy decreased at first place before
+increasing to edge, recover or exceed the initial value.
 
-Trajectories of Richness, Shannon and Simpson taxonomic diversity were
-examined at genus level in relation to the 1989 inventories (5 years
-after disturbance) (Figure \ref{fig:DivTaxo}).
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivTaxo-1}
-
-}
-
-\caption{Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities \textbf{(a)} Richness, \textbf{(b)} Shannon and \textbf{(c)} Simpson diversities.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:DivTaxo}
-\end{figure*}
-
-For undisturbed plots the Richness, Shannon and Simpson diversities
-remained stable over the 30 years monitored. After low disturbance
-intensity the richness increased, reaching a maximum gain of 14
-botanical genera (plot 3 from treatment 2). After intense disturbance
-however, the taxonomic richness followed a unimodal decrease with a
-return to pre-disturbance values. The taxonomic evenness (Shannon and
-Simpson diversities) increased after disturbance, whenever its
-intensity. The evenness followed a unimodal trajectory with a just
-beginning return towards initial values and a maximum value, reached
-around 20 years, positively correlated to the disturbance intensity
-(\(\rho_{spearman}^{Shannon}=0.86\), and
-\(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots (plots 8 and 12)
-were still increasing 30 years after disturbance \ref{fig:DivTaxo}),
-suggesting a similar be much delayed unimodal trajectory.
-
-Communities functional diversity trajectories were examined through the
-Rao diversity based on the 7 leaf, stem and life history traits (the
-seed mass was excluded as a qualitative variable). The plot 7 from
-treatment 1 that displayed a constantly outlying diversity was removed
-from the graphical representation for better readability, see figure A3
-in appendix for full graph. The functional diversity of all undisturbed
-communities remained stable along the 30 years monitored. For all
-disturbed communities the functional diversity followed unimodal
-trajectories with a marked return towards the initial state (5 years
-after disturbance) with values comparable to those of control plots.
-
-The impact of disturbance was examined more accurately at 10, 20 and 30
-years after disturbance through the linear correlation between the
-initial percentage of AGB removed and the value of communities Simpson
-and Rao diversities (both diversities of order 2).
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/IDHplot-1}
-
-}
-
-\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.
-Lower panels, Relationship between the initial \%AGB removed and the median values of Simpson \textbf{(c)} and Rao \textbf{(d)} diversities at three times after disturbance.
-Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.}\label{fig:IDHplot}
-\end{figure*}
-
-The functional redundancy was measured as the sum of weighted overlap
-among species in communities' functional space. Functional space were
-defined in the main two-dimensional space of a PCA analysis based on all
-leaf and stem and maximum height traits (see appendix I for PCA
-details). Plots functional redundancy compared to the initial value
-after logging so that only communities dynamics are followed apart from
-the impact of logging itself. Communities functional redundancy remained
-stable in control plots but after disturbance the redundancy
-trajectories were quite variable (See appendix I). Globally after most
-intense disturbance (plots T2 and T3) communities redundancy decreased
-at first place and then increased to edge, recover or even exceed the
-initial value. Considering the functional redundancy restricted to the
-functional space of the initial inventory, though, all disturbed plots
-followed similar decreasing humped-shaped trajectories
-(@ref(fig:RedFun\_rest)). The maximum redundancy loss observed was
-positively correlated with the disturbance intensity (XX spearman to be
-measured) and none of the disturbed plots had recovered the functional
-redundancy within their initial functional space.
+Considering the functional redundancy restricted to the functional space
+of the initial inventory, all disturbed plots followed similar
+decreasing humped-shaped trajectories (@ref(fig:RedFun\_rest)). The
+maximum redundancy loss observed was positively correlated with the
+disturbance intensity (XX spearman to be measured) and none of the
+disturbed plots had recovered the functional redundancy within the
+initial functional space.
 
@@
 
+\subsection{Decoupled recovery of communities taxonomic and functional
+characteristics}\label{decoupled-recovery-of-communities-taxonomic-and-functional-characteristics}
+
+Both communities taxonomic and functional diversity and composition
+proved resilient, following similar humped-back trajectories with a
+return towards initial values. The resilience of communities functional
+characteristics, the most direct link between biodiversity and ecosystem
+functioning \citep{Diaz2005}, meant a consistent recovery of ecosystem
+processes in the long term \citep{Guariguata2001}. The resilience of
+communities taxonomic characteristics, in turn, meant the maintenance of
+their initial differences in composition and structure. It suggested
+that communities response to disturbance were somehow constraint to
+converge towards determined compositions
+\citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.
+
+Although both communities taxonomic and functional characteristics
+proved resilient and followed similar humped-back trajectories, the
+taxonomic recovery systematically lagged behind the corresponding
+functional dynamics. Such delay between functional and taxonomic
+dynamics has already been observed for grasslands
+\citep{Tilman1997, Mouillot2011} and more recently for tropical forests
+\citep{Lohbeck2015, Guariguata2001}. According to the ``vegetation
+quantity effect'' \citep{Grime1998}, functional trajectories rely on the
+pool of dominant species which diversity and evenness were enhanced
+after disturbance and which rapidly restored the functional diversity
+and composition. However, infrequent species that would further lower
+the functional diversity and communities evenness still missed to the
+recovery of taxonomic characteristics. Taxonomic recovery then
+mechanically lagged behind, all the more so that unrecovered species are
+functionally redundant and would probably undergo competitive pressures.
+
 \subsection{A validation of the intermediate disturbance
@@
 
-The monitoring of disturbed forest communities confirmed a limited
-negative impact of disturbance on species richness, as observed on
-several post logging surveys \citep{Cannon1998, Baraloto2012a}. Thirty
-years after disturbance, the genus richness was restored to initial plot
-values after high disturbance intensity while it substantially increased
-after low disturbance intensity, reaching a gain of almost 12 genera for
-some plots.
+The monitoring of disturbed forest communities confirmed a limited or
+negative impact of intense disturbance on species richness, as observed
+on several post logging surveys \citep{Cannon1998, Baraloto2012a}, while
+. Thirty years after disturbance, the genus richness was restored to
+initial plot values after high disturbance intensity while it
+substantially increased after low disturbance intensity, reaching a gain
+of almost 12 genera for some plots.
 
@@
 
-\subsection{On the recovery of disturbed
-communities}\label{on-the-recovery-of-disturbed-communities}
-
-All diversity and composition trajectories followed similar hump-back
-trajectories returning towards their initial values. Communities
-trajectories therefore illustrated their taxonomic and function
-resilience.
-
-Because functional traits and functional diversity are the most direct
-link between biodiversity and ecosystem functioning \citep{Diaz2005},
-their resilience meant a consistent recovery of ecosystem processes in
-the long term \citep{Guariguata2001}. The resilience of the taxonomic
-diversity and composition similarly translated the recovery of
-communities such as before disturbance. This resilience meant the
-maintenance of communities initial differences in composition and
-supported the idea of their convergence towards a determined state
-\citep{Hubbell1999, Molino2001, Baraloto2012a}. If all plots followed
-similar hump-back trajectories, though, some significant differences
-emerged. To name but one, the taxonomic evenness of one plot displayed
-an extensive increase until 30 years after disturbance. The differences
-trajectories maxima and time path, along with the fact intrinsinc
-communities differences are maintained, suggested that the trajectories
-depend on the initial plots composition or on some abiotic parameters
-\citep{Anderson2007}.
-
 \subsection{Functional redundancy of disturbed
@@
 
-Despite the recovery of pre-disturbance state there was a time lag
-between the trajectories followed by communities taxonomic and
-functional characteristics. The trajectory of taxonomic composition and
-evenness had a longer time-path and remained altered while communities
-functional composition and diversity had already recovered.
-
-Such delay between functional and taxonomic dynamics was already
-observed for grasslands \citep{Tilman1997, Mouillot2011} and more
-recently for tropical forests \citep{Lohbeck2015, Guariguata2001}.
-According to the ``vegetation quantity effect'' \citep{Grime1998}, the
-functional diversity of communities rely on the dominant species.
-Communities functional trajectories were then driven first by the
-increase in diversity and evenness of dominant species, which restored
-the pre-disturbance functional characteristics. Then, the functional
-trajectories were driven by the recruitment of species resembling the
-old pre-disturbance community, which reduced the overall diversity. At
-that time, the taxonomic composition and diversity remained altered: the
-still missing species were then the initially infrequent and
-functionally redundant ones. The functional redundancy, the functional
-overlap between species that is typical of the huge biodiversity of
-tropical forests \citep{Bellwood2006}, was then not restored 30 years
-after disturbance and this is major to consider as it defines forests'
-resilience \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+The functional redundancy, the functional overlap between species that
+is typical of the huge biodiversity of tropical forests
+\citep{Bellwood2006}, was then not restored 30 years after disturbance
+and this is major to consider as it defines forests' resilience
+\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 

### 2018-05-16 WholePlotTrajectories.Rmd
Discussion

@@
 
-both the lag between taxonomic and functional recovery and the much better consistency of the IDH when the functional diversity is considered rely on the central role of functional redundancy.
-The functional redundancy that is the overlap between species, typical of the huge biodiversity of tropical forests [@Bellwood2006] and determinant for communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005], was not recovered 30 years after disturbance.
+Both the lag between taxonomic and functional recovery and the middling consistency of the IDH regarding communities taxonomic diversity entail a central role of functional redundancy, *i.e.* the overlap between species functional spaces.
+Functional redundancy is typical of the huge biodiversity of tropical forests [@Bellwood2006] and is determinant of communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005], and it seemed not to have recovered 30 years after disturbance.
 
-Disturbance globally reduced the overall functional redundancy of communities but this restored rapidly without consistent rtajectories along time.
-The initial functional space corresponding to undistrubed 
-Besides the long-term alteration of functional redundancy, there was probably persistent compositional changes favoring disturbance resistant species [@Haddad2008], lianas or epiphytes [@Martin2013] and environmental changes, like in the soils nutrient cycling and compaction [@Olander2005].
-These persistent changes highly question forest's resilience [@Chazdon2003a].
-New conditions would not only be longer lasting but self-maintained as tied to disturbance regime [@Burslem2000].
+Although disturbance would be mitigated by communities functional redundancy, we expect the redundancy itself to be reduced or re-organized within the functional space.
+Globally, disturbed communities did not follow consistent trajectories after disturbance and 30 years after disturbance, irrespective of the intensity, disturbed communities displayed lower, similar or higher functional redundancy.
+The redundancy within the functional space of the initial community however clearly followed humped-back trajectories which decrease was determined by the disturbance intensity and which return to initial state was not observed for any disturbed community.
+Not only did the functional redundancy and hence the resilience of pre-disturbance community decrease, but also did the ones of disturbance-specific communities increase.
+It means higher chances to have long lasting or self-maintained compositional changes towards disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013] and thus highly question forest's resilience [@Chazdon2003a].
 Specificaly, this would impair species contingent to undisturbed forests, threatening their maintenance, and run the risk to loose cornerstone species and trigger unexpected ecological consequences [@Jones1994;@Diaz2005;@Gardner2007].

### 2018-05-16 WholePlotTrajectories.Rmd
Conclusion

@@
 Although both communities taxonomic and functional characteristics proved resilient and followed similar humped-back trajectories, the taxonomic recovery systematically lagged behind the corresponding functional dynamics.
-Such delay between functional and taxonomic dynamics has already been observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
+Such delay between functional and taxonomic dynamics has already been observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015; @Guariguata2001].
 According to the "vegetation quantity effect" [@Grime1998] functional trajectories rely on the pool of dominant species, which diversity and evenness were enhanced after disturbance, and which rapidly restored the functional diversity and composition.
@@
 As the composition of old-growth survivors proved to mirror initial communities [@Herault2018], disturbance would impact trees recruitment through the enhanced growth and survival of previously infrequent species and functional characteristics.
-Consistently, disturbance resulted in increasing taxonomic dissimilarity compared to pre-disturbance communities and functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
+Consistently, disturbance resulted in increasing taxonomic dissimilarity compared to pre-disturbance communities and functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
 Disturbance then caused a reorganization of the typical high dominance structure of hyperdiverse mature forests after disturbance, benefiting to pioneers and light demanding species. 
@@
 
-## Functional redundancy of disturbed communities
+## The functional redundancu and the resilience of communities
 
@@
 
-Our study showed the significant impact of disturbance on tropical forests communities and validated the consistency of the IDH in the long term.
-It revealed the contrasting response of communities taxonomic and functional characteristics, with persisting impacts on the species abundance distibution while the functional diversity and dominant functional strategies were restored.
-Communities recovery therefore remained unachieved but consistent for the range of disturbance studied here.
-The length of the recovery, however, severly questioned the sustainability of intense selective logging and advocated felling cycle much longer than 30 years [@Gourlet-Fleury2005].
+our study disentangled tropical forests functional and taxonomic decoupled response to disturbance, with a rapid recovery of communities functioning but slower and more variable taxonomic dynamics.
+Consistently with the IDH functional trajectories were constrained by environmental pressures favoring pioneers and light-demanding species after disturbance, but taxonomic dynamics were more stochastic as they involved infrequent speciesfunctoinally redundant with the dominants and thus under competitive pressure.
+While communities functioning rapidly recovered, taxonomic trajectories, although consistent and converging towards full recovery, were much slower and entailed persisting alteration of communities functional redundancy.
+the trajectories therefore suggested a potential sustainability of tropical forests in the face of quite intense disturbance but only when followed by long recovery periods, longer than 30 years [@Gourlet-Fleury2005].
+This study besides highlighted the central role of recruitment processes underlying the IDH mechanisms and holding the whole community response to disturbance.
 

### 2018-05-16 AppendixI.Rmd
Discussion

@@
 Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
-load("DB/Redundancy")
+load("DB/Redundancy_sd0.2")
 
@@
 RedundancyPlot(RedundancyTraj)
-mtext("Functional Redundancy",side=3,adj=0,line=1.5)
-mtext("Restricted to initial functional space",side=3,adj=0,cex=0.8,line=0.5)
+mtext("Functional Redundancy",side=3,adj=0,line=0.5)
 mtext("Years since disturbance",side=1,line=2.2,adj=1)

### 2018-05-16 WholePlotTrajectories.Rmd
Discussion

@@
 
-In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events that changes the abiotic environment, through the fluxes of light, heat and water, and the biotic interaction and competitive pressure [@Goulamoussene2017].
-The cornerstone of tropical forests ecology is then to understand the underlyings of ecosystem response to disturbance, its mechanisms and drivers [@White2001;@Chazdon2003a]. 
+In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events changing both abiotic environment, through the fluxes of light, heat and water, and biotic interactions and competitive pressure [@Goulamoussene2017].
+The cornerstone of tropical forests ecology is to understand the underlyings of ecosystem response to disturbance, its mechanisms and drivers [@White2001;@Chazdon2003a]. 
 For now, this has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
-These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009]. 
-However the response of tree species diversity remains unclear, also it is determinant of ecosystems productivity, stability and functioning [@Tilman2014, @Liang2016] and would be most probably impacted by post-disturbance environmental changes [@Baraloto2012a].
+Consistent models assessed these structural parameters and highlighted the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009]. 
+The response of tree diversity, however, remains unclear albeit it determines ecosystems productivity, stability and functioning [@Tilman2014, @Liang2016] and would be most probably impacted by post-disturbance environmental changes [@Baraloto2012a].
 
@@
 More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and diversity profiles, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, the functional approach accounting for differenes species biological attributes would be insightful as it reveals the species role in the post-disturbance trajectory of ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-In that respect major functional traits related to species ecology and performance were largely adopted through an integrative framework [@Diaz2005;@Villeger2008a].
-The functional trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
+Furthermore, account for species biological attributes and role the ecosystem through a functional approach assesses communities functioning and corresponding mechanisms [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+In that respect major functional traits related to species ecology and performance were largely adopted [@Diaz2005;@Villeger2008a] highlighting, for example, the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition in tropical rainforests [@Molino2001;@Haddad2008].
 In disturbed tropical rainforests, this framework translates into the fostering of disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
@@
 
-A proper monitoring of communities response should therefore encompass taxonomic and functional diversity and composition measures to test the validity of the IDH in the long term for tropical forests, and clarify the resilience of communities evenness, composition, and also functioning.
-The trajectories followed by all these facets would highlight the role of deterministic processes, like competitive exclusion or abiotic selection, and the communities' convergence maintaining intrinsinc differences in diversity and composition,which is as much insgights for future adaptive conservation strategies [@Adler2007].
+Functional and taxonomic trajectories relate to distinct characteristics and community processes, and might therefore be decoupled as already observed in tropical forests [@Lohbeck2015; @Guariguata2001].
+The combination of taxonomic and functional characteristics is translated by the functional redundancy, measuring the amount of species sharing same trait values [@Carmona2016].
+Functional redundancy is a major determinant of communities resilience as high redundancy, like in highly diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning [@Trenbath1999; @Elmqvist2003; @Diaz2005].
+It is therefore a key element as a link between taxonomic and functional diversities and as a determinant of communities resilience.
 
-Here we investigated over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
+Grasping all facets of communities response to disturbance then requires to disentangle the taxonomic and functional trajectories in diversity and composition.
+These trajectories would highlight communities resilience and the drivers of their response to disturbance, clarifying the tenants of the IDH in the long term for tropical forests, that are as much insights for future adaptive conservation strategies [@Adler2007].
+Here we monitored over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
 We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
-Specifically, we (i) tested the validity of the IDH in the long term for tropical hyperdiverse forest and highlighted the ecological rules shaping their response to disturbance, (ii) clarified the different facets of communities resilience in terms of communities composition, diversity and functioning *????(iii) questioned the completeness of communities recovery given the altered functional redundancy.*
+Specifically, we (i) highlighted the determinant of communities functoinal and taxonomic response to disturbance, (ii) tested the validity of the IDH in the long term for tropical forest and (iii) questioned the completeness of communities recovery regarding their functional redundancy.
 
@@
 
-Both the lag between taxonomic and functional recovery and the middling consistency of the IDH regarding communities taxonomic diversity entail a central role of functional redundancy, *i.e.* the overlap between species functional spaces.
-Functional redundancy is typical of the huge biodiversity of tropical forests [@Bellwood2006] and is determinant of communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005], and it seemed not to have recovered 30 years after disturbance.
+Both the lag between taxonomic and functional recovery and the middling consistency of the IDH regarding communities taxonomic diversity entail a central role of functional redundancy.
+Functional redundancy, determinant of communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005], seemed not to have recovered 30 years after disturbance.
 

### 2018-05-16 Scripts/Functionalredundancy.R

@@ -104,6 +104,56 @@ save(RedundancyTraj,file="DB/Redundancy")
 
 
 ## Redundancy restricted to the initial community functional space
+
+Nrep<-2
+
+# Pb species, "Pouteria_sp.42CAY-ATDN", "Inga_nobilis"
+Sptdp<-lapply(1:Nrep,function(rep){
+  traits_filled<-Traits_filling(Traits1,Traits2,InventorySp)
+  
+  ACP<-dudi.pca(scale(traits_filled[,TraitsName]),scannf=FALSE,nf=2)
+  Bigacp<-merge(ACP$li,traits_filled[,c("Family","Genus","name" )],by="row.names")[,c("Family","Genus","name","Axis1","Axis2")]
+  
+  xgen<-c(min(Bigacp[,"Axis1"]),max(Bigacp[,"Axis1"]))
+  ygen<-c(min(Bigacp[,"Axis2"]),max(Bigacp[,"Axis2"]))
+  
+  sptdp<-lapply(unique(Bigacp[,"name"]),function(Sp){
+    #sptdp<-lapply(c("Pouteria_sp.42CAY-ATDN", "Inga_nobilis"),function(Sp){
+    
+    acp<-Bigacp[which(Bigacp[,"name"]==Sp),4:5]
+    if(nrow(acp)!=1){
+      method1<-method2<-"ste"
+      if(tryCatch(width.SJ(acp[,"Axis1"]), error=function(e) "error")=="error")
+      {method1<-"dpi"}
+      if(tryCatch(width.SJ(acp[,"Axis2"]), error=function(e) "error")=="error")
+      {method2<-"dpi"}
+      f<-kde2d(acp[,"Axis1"], acp[,"Axis2"], n = 100,lims=c(xgen,ygen),
+               h = c(width.SJ(acp[,"Axis1"],method=method1),
+                     width.SJ(acp[,"Axis2"],method=method2)) )
+      z<-f$z
+      colnames(z)<-1:100;rownames(z)<-1:100
+    }
+    if(nrow(acp)==1){
+      z<-outer(seq(xgen[1],xgen[2],length=100),seq(ygen[1],ygen[2],length=100), 
+               function(x,y) dnorm(x,acp[,"Axis1"],MeanSd)*dnorm(y,acp[,"Axis2"],MeanSd))
+      colnames(z)<-1:100;rownames(z)<-1:100
+    }
[…]

### 2018-05-16 Scripts/Stats.R

@@ -42,7 +42,7 @@ AgbLoss<-AGBloss
 #plotIDH...
 
 ##########################################
-## Rho Spearman
+## Rho Spearman, Taxo
 load("DB/TaxoComposition_ForGraphs")
 
 Data_TaxoComp<-MatrepTaxo
@@ -83,6 +83,12 @@ TimeMax<-unlist(lapply(1:12,function(pl){
   ret<-apply(ret,2,median)
   return(names(ret)[which(ret==max(ret))])}))
 mean(as.numeric(TimeMax))-1984
+
+##########################################
+## Rho Spearman, redundancy
+load("DB/Redundancy_restricted")
+RedundancyTraj_restricted
+
 
 ###############
 ## CWM distance

### 2018-05-17 AppendixI.Rmd
Appendix, redundancy

--- 2018-05-15 AppendixI.Rmd
@@
 
-**Figure A2**: Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+**Figure A2**: Scheme of the redundancy measurement process
+
+**Figure A3**: Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
 No gap filling process was applied in this case.
@@
 
-**Figure A3**: Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
+**Figure A4**: Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 
-**Figure A4**: Trajectories of the Rao functional diversity over 30 years after disturbance.
+**Figure A5**: Trajectories of the Rao functional diversity over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 
-**Figure A5**: Summarizing plots of the functional PCA analysis defining communities functional space.
+**Figure A6**: Summarizing plots of the functional PCA analysis defining communities functional space.
 **(a)** Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
@@
 
-**Figure A6**: Trajectories of the functional redundancy over 30 years after disturbance.
+**Figure A7**: Trajectories of the functional redundancy over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
@@
 
-# **Figure A2**: Proportion of seed mass classes along time
+# **Figure A2**: Redundancy metric
+
+```{r RedMM,out.width='100%',fig.cap="Scheme of the redundancy measurement process",echo=F, fig.env="figure*"}
+knitr::include_graphics("ExternalFig/RedundancyMetric.png")
+```
+
+# **Figure A3**: Proportion of seed mass classes along time
 
@@
 
-# **Figure A3**: Richness, Shannon and Simpson taxonomic diversity trajectories
+# **Figure A4**: Richness, Shannon and Simpson taxonomic diversity trajectories
 
@@
 
-# **Figure A4**: Rao diversity trajectories with all plots display
+# **Figure A5**: Rao diversity trajectories with all plots display
 
@@
 
-# **Figure A5**: Functional PCA
+# **Figure A6**: Functional PCA
 
@@
 
-# **Figure A6**: Overall functional redundancy
+# **Figure A7**: Overall functional redundancy
 

### 2018-05-17 AppendixI.tex
Appendix, redundancy

@@
 
-\textbf{Figure A2}: Trajectories of seed mass classes proportions over
+\textbf{Figure A2}: Scheme of the redundancy measurement process
+
+\textbf{Figure A3}: Trajectories of seed mass classes proportions over
 30 years after disturbance, corresponding to the median (solid line) and
@@
 
-\textbf{Figure A3}: Trajectories of the difference to the 1989
+\textbf{Figure A4}: Trajectories of the difference to the 1989
 inventories (5 years after disturbance) over 30 years after disturbance
@@
 
-\textbf{Figure A4}: Trajectories of the Rao functional diversity over 30
+\textbf{Figure A5}: Trajectories of the Rao functional diversity over 30
 years after disturbance. Trajectories correspond to the median (solid
@@
 
-\textbf{Figure A5}: Summarizing plots of the functional PCA analysis
+\textbf{Figure A6}: Summarizing plots of the functional PCA analysis
 defining communities functional space. \textbf{(a)} Plot of the samples
@@
 
-\textbf{Figure A6}: Trajectories of the functional redundancy over 30
+\textbf{Figure A7}: Trajectories of the functional redundancy over 30
 years after disturbance. Trajectories correspond to the median (solid
@@
 
-\section{\texorpdfstring{\textbf{Figure A2}: Proportion of seed mass
+\section{\texorpdfstring{\textbf{Figure A2}: Redundancy
+metric}{Figure A2: Redundancy metric}}\label{figure-a2-redundancy-metric}
+
+\begin{figure*}
+
+{\centering \includegraphics[width=1\linewidth]{ExternalFig/RedundancyMetric}
+
+}
+
+\caption{Scheme of the redundancy measurement process}\label{fig:RedMM}
+\end{figure*}
+
+\section{\texorpdfstring{\textbf{Figure A3}: Proportion of seed mass
 classes along
-time}{Figure A2: Proportion of seed mass classes along time}}\label{figure-a2-proportion-of-seed-mass-classes-along-time}
+time}{Figure A3: Proportion of seed mass classes along time}}\label{figure-a3-proportion-of-seed-mass-classes-along-time}
 
@@
 
-\section{\texorpdfstring{\textbf{Figure A3}: Richness, Shannon and
+\section{\texorpdfstring{\textbf{Figure A4}: Richness, Shannon and
 Simpson taxonomic diversity
-trajectories}{Figure A3: Richness, Shannon and Simpson taxonomic diversity trajectories}}\label{figure-a3-richness-shannon-and-simpson-taxonomic-diversity-trajectories}
+trajectories}{Figure A4: Richness, Shannon and Simpson taxonomic diversity trajectories}}\label{figure-a4-richness-shannon-and-simpson-taxonomic-diversity-trajectories}
 
@@
 
-\section{\texorpdfstring{\textbf{Figure A4}: Rao diversity trajectories
+\section{\texorpdfstring{\textbf{Figure A5}: Rao diversity trajectories
 with all plots
-display}{Figure A4: Rao diversity trajectories with all plots display}}\label{figure-a4-rao-diversity-trajectories-with-all-plots-display}
+display}{Figure A5: Rao diversity trajectories with all plots display}}\label{figure-a5-rao-diversity-trajectories-with-all-plots-display}
 
@@
 
-\section{\texorpdfstring{\textbf{Figure A5}: Functional
-PCA}{Figure A5: Functional PCA}}\label{figure-a5-functional-pca}
+\section{\texorpdfstring{\textbf{Figure A6}: Functional
+PCA}{Figure A6: Functional PCA}}\label{figure-a6-functional-pca}
 
@@
 
-\section{\texorpdfstring{\textbf{Figure A6}: Overall functional
-redundancy}{Figure A6: Overall functional redundancy}}\label{figure-a6-overall-functional-redundancy}
+\section{\texorpdfstring{\textbf{Figure A7}: Overall functional
+redundancy}{Figure A7: Overall functional redundancy}}\label{figure-a7-overall-functional-redundancy}
 

### 2018-05-17 WholePlotTrajectories.Rmd
Appendix, redundancy

@@
 
-The functional redundancy was measured as the overlap among species in communities' functional space [@Carmona2016].
-The samples of the trait database were first located with a PCA analysis in a two-dimensional functional space summarizing the 8 functional traits considered.
-We choose to summarize species functional space through a PCA because functional traits considered were already somehow redundant and allows not to overweight the role of one trait.
-Then, multivariate kernel density estimator associated with individual trees were summed to give the distribution of traits probabilites of each species.
-For each community the trait probability distributions of corresponding species were combined and weighted by species abundance. 
-Eventually communities functional redundancy was measured as the sum of overlap between species weighted functional density.
-Communities functional redundancy is expressed as the average number of species that could be removed from the community without reducing the functional space.
+The functional redundancy is measured as the overlap among species in communities' functional space [@Carmona2016].
+First, the samples of the trait database are mapped in a 2-dimensional plan with a PCA analysis.
+Then, multivariate kernel density estimator associated with individual trees were give species traits probability distribution (TDP).
+TDP, weighted by species abundance, are summed to give communities TDP: the functional redundancy is the sum of TDP overlaps,  corresponding to the average number of species that could be removed from without reducing the functional space (see appendix I for a more comprehensive sheme).
 

### 2018-05-17 WholePlotTrajectories.tex
Appendix, redundancy

@@
 In tropical forest, ecological communities are contantly re-shaped by
-the natural disturbance events that changes the abiotic environment,
-through the fluxes of light, heat and water, and the biotic interaction
-and competitive pressure \citep{Goulamoussene2017}. The cornerstone of
-tropical forests ecology is then to understand the underlyings of
-ecosystem response to disturbance, its mechanisms and drivers
+the natural disturbance events changing both abiotic environment,
+through the fluxes of light, heat and water, and biotic interactions and
+competitive pressure \citep{Goulamoussene2017}. The cornerstone of
+tropical forests ecology is to understand the underlyings of ecosystem
+response to disturbance, its mechanisms and drivers
 \citep{White2001, Chazdon2003a}. For now, this has been largely studied
@@
 aboveground biomass, tree height or stem density
-\citep{Piponiot2016, Rutishauser2016}. These structural parameters have
-then been sucessfully modeled and allowed to assess the maintenance of
-ecosystems processes and services \citep{Denslow2000, Blanc2009}.
-However the response of tree species diversity remains unclear, also it
-is determinant of ecosystems productivity, stability and functioning
+\citep{Piponiot2016, Rutishauser2016}. Consistent models assessed these
+structural parameters and highlighted the maintenance of ecosystems
+processes and services \citep{Denslow2000, Blanc2009}. The response of
+tree diversity, however, remains unclear albeit it determines ecosystems
+productivity, stability and functioning
 \citep[\citet{Liang2016}]{Tilman2014} and would be most probably
@@
 ecological rules underlying communities' structure
-\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, the
-functional approach accounting for differenes species biological
-attributes would be insightful as it reveals the species role in the
-post-disturbance trajectory of ecosystems functioning
+\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, account
+for species biological attributes and role the ecosystem through a
+functional approach assesses communities functioning and corresponding
+mechanisms
 \citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
 respect major functional traits related to species ecology and
-performance were largely adopted through an integrative framework
-\citep{Diaz2005, Villeger2008a}. The functional trait-based approach,
-for example, highlighted in tropical rainforests the environmental
-filters fostering disturbance resistant species with rapid growth and
-efficient resources acquisition \citep{Molino2001, Haddad2008}. In
+performance were largely adopted \citep{Diaz2005, Villeger2008a}
+highlighting, for example, the environmental filters fostering
+disturbance resistant species with rapid growth and efficient resources
+acquisition in tropical rainforests \citep{Molino2001, Haddad2008}. In
 disturbed tropical rainforests, this framework translates into the
@@
 
-A proper monitoring of communities response should therefore encompass
-taxonomic and functional diversity and composition measures to test the
-validity of the IDH in the long term for tropical forests, and clarify
-the resilience of communities evenness, composition, and also
-functioning. The trajectories followed by all these facets would
-highlight the role of deterministic processes, like competitive
-exclusion or abiotic selection, and the communities' convergence
-maintaining intrinsinc differences in diversity and composition,which is
-as much insgights for future adaptive conservation strategies
-\citep{Adler2007}.
+Functional and taxonomic trajectories relate to distinct characteristics
+and community processes, and might therefore be decoupled as already
+observed in tropical forests \citep{Lohbeck2015, Guariguata2001}. The
+combination of taxonomic and functional characteristics is translated by
+the functional redundancy, measuring the amount of species sharing same
+trait values \citep{Carmona2016}. Functional redundancy is a major
+determinant of communities resilience as high redundancy, like in highly
+diverse tropical forests \citep{Bellwood2006}, mitigates the impacts of
+species removal on ecosystem functioning
+\citep{Trenbath1999, Elmqvist2003, Diaz2005}. It is therefore a key
+element as a link between taxonomic and functional diversities and as a
+determinant of communities resilience.
 
-Here we investigated over 30 years the response of 75 ha of forests
-plots set up on a gradient of disturbance intensity, from 10 to 60\% of
-ecosystem biomass removed. We made use of a large functional traits
-database browsing major leaf, stem and seed traits and species maximum
-height to draw the trajectories over time of communities taxonomic and
-functional composition and diversity. Specifically, we (i) tested the
-validity of the IDH in the long term for tropical hyperdiverse forest
-and highlighted the ecological rules shaping their response to
-disturbance, (ii) clarified the different facets of communities
-resilience in terms of communities composition, diversity and
-functioning \emph{????(iii) questioned the completeness of communities
-recovery given the altered functional redundancy.}
+Grasping all facets of communities response to disturbance then requires
+to disentangle the taxonomic and functional trajectories in diversity
+and composition. These trajectories would highlight communities
+resilience and the drivers of their response to disturbance, clarifying
+the tenants of the IDH in the long term for tropical forests, that are
+as much insights for future adaptive conservation strategies
+\citep{Adler2007}. Here we monitored over 30 years the response of 75 ha
+of forests plots set up on a gradient of disturbance intensity, from 10
+to 60\% of ecosystem biomass removed. We made use of a large functional
+traits database browsing major leaf, stem and seed traits and species
+maximum height to draw the trajectories over time of communities
+taxonomic and functional composition and diversity. Specifically, we (i)
+highlighted the determinant of communities functoinal and taxonomic
+response to disturbance, (ii) tested the validity of the IDH in the long
+term for tropical forest and (iii) questioned the completeness of
+communities recovery regarding their functional redundancy.
 
@@
 
-The functional redundancy was measured as the overlap among species in
-communities' functional space \citep{Carmona2016}. The samples of the
-trait database were first located with a PCA analysis in a
-two-dimensional functional space summarizing the 8 functional traits
-considered. We choose to summarize species functional space through a
-PCA because functional traits considered were already somehow redundant
-and allows not to overweight the role of one trait. Then, multivariate
-kernel density estimator associated with individual trees were summed to
-give the distribution of traits probabilites of each species. For each
-community the trait probability distributions of corresponding species
-were combined and weighted by species abundance. Eventually communities
-functional redundancy was measured as the sum of overlap between species
-weighted functional density. Communities functional redundancy is
-expressed as the average number of species that could be removed from
-the community without reducing the functional space.
+The functional redundancy is measured as the overlap among species in
+communities' functional space \citep{Carmona2016}. First, the samples of
+the trait database are mapped in a 2-dimensional plan with a PCA
+analysis. Then, multivariate kernel density estimator associated with
+individual trees were give species traits probability distribution
+(TDP). TDP, weighted by species abundance, are summed to give
+communities TDP: the functional redundancy is the sum of TDP overlaps,
+corresponding to the average number of species that could be removed
+from without reducing the functional space (see appendix I for a more
+comprehensive sheme).
 
@@
 \citep{Lohbeck2015, Guariguata2001}. According to the ``vegetation
-quantity effect'' \citep{Grime1998}, functional trajectories rely on the
-pool of dominant species which diversity and evenness were enhanced
-after disturbance and which rapidly restored the functional diversity
-and composition. However, infrequent species that would further lower
-the functional diversity and communities evenness still missed to the
-recovery of taxonomic characteristics. Taxonomic recovery then
-mechanically lagged behind, all the more so that unrecovered species are
-functionally redundant and would probably undergo competitive pressures.
+quantity effect'' \citep{Grime1998} functional trajectories rely on the
+pool of dominant species, which diversity and evenness were enhanced
+after disturbance, and which rapidly restored the functional diversity
+and composition. However, communities evenness remained high so
+infrequent species still missed to the taxonomic recovery that
+mechanically lagged behind, all the more so that unrecovered species
+would be functionally redundant and probably undergo competition
+processes.
 
@@
 
-The monitoring of disturbed forest communities confirmed a limited or
-negative impact of intense disturbance on species richness, as observed
-on several post logging surveys \citep{Cannon1998, Baraloto2012a}, while
-. Thirty years after disturbance, the genus richness was restored to
-initial plot values after high disturbance intensity while it
-substantially increased after low disturbance intensity, reaching a gain
-of almost 12 genera for some plots.
+Validating the IDH, communities trajectories confirmed the diversity
+increase after disturbance through an enhanced growth of otherwise less
+favored species. This was however completely clear regarding the
+functional diversity but much more blurry for communities taxonomic
+characteristics.
 
-Both richness and evenness followed asymptotic trajectories after
-disturbance, sharply increasing for 15 years before stabilizing at
-higher values than those of control plots. An increase of communities
-evenness stems from a higher homogeneity of species distribution. After
-disturbance, communities are made of the old, pre-disturbance survivors
-and the newly recruited trees: changes in composition and abundance
-distribution are to be found in the recruitment processes and in the
-pre-disturbance survivors mortality. Because the composition of old
-survivors proved to mirror the initial community \citep{Herault2018}, a
-specific turnover was expected among recruited trees with enhanced
-growth and survival of previously infrequent species. Indeed, the
-increase in taxonomic diversity was accompanied by an increase of
-taxonomic dissimilarity with plots initial state and a functional shift
-towards resource-acquisitive strategies (sharp increase in the SLA, leaf
-thickness and bark thickness and decrease in wood density, leaf
-toughness and maximum height)
-\citep{Westoby1998, Wright2004, Reich2014}. Disturbance then causes a
+The taxonomic richness was weakly or negatively impacted by intense
+disturbance, as observed on several post logging surveys
+\citep{Cannon1998, Baraloto2012a}, while it substantially increased
+after low intensity disturbance. Disturbance also enhanced taxonomic
+evenness, but its intensity was only weakly and lately correlated to
+communities eveness (\emph{i.e.} Simpson diversity). Contrastingly,
+disturbance intensity consistently predicted the significant increase of
+communities functional diversity for 30 years. Disturbance impacts
+involve specific turnover within communities, either among
+pre-disturbance survivors or among newly recruited trees. As the
+composition of old-growth survivors proved to mirror initial communities
+\citep{Herault2018}, disturbance would impact trees recruitment through
+the enhanced growth and survival of previously infrequent species and
+functional characteristics. Consistently, disturbance resulted in
+increasing taxonomic dissimilarity compared to pre-disturbance
+communities and functional shifts towards resource-acquisitive
+strategies (sharp increase in the SLA, leaf thickness and bark thickness
+and decrease in wood density, leaf toughness and maximum height)
+\citep{Westoby1998, Wright2004, Reich2014}. Disturbance then caused a
 reorganization of the typical high dominance structure of hyperdiverse
@@
 limiting resources but are excluded in mature forests by long-lived,
-more resistant and shade tolerant species.
+more resistant and shade tolerant species. Therefore, consistently with
+communities dynamics after disturbance relied on species functional
+strategy and corresponding ability to fill the environmental niches made
+available by disturbance. Recruited species then mixed with
+pre-disturbance ones, from which they differed, and constituted a
+community all the more diversified that the disturbance was intense
+\citep{Molino2001}.
 
-As stated by the IDH, communities dynamics after disturbance relied on
-species functional strategy and corresponding ability to fill the
-environmental niches made available by disturbance. Recruited species
-then mixed with pre-disturbance ones, from which they differed, and
-constituted a community all the more diversified that the disturbance
-was intense \citep{Molino2001}.
+\subsection{The functional redundancu and the resilience of
+communities}\label{the-functional-redundancu-and-the-resilience-of-communities}
 
-\subsection{Functional redundancy of disturbed
-ecosystems}\label{functional-redundancy-of-disturbed-ecosystems}
+Both the lag between taxonomic and functional recovery and the middling
+consistency of the IDH regarding communities taxonomic diversity entail
+a central role of functional redundancy. Functional redundancy,
+determinant of communities resilience
+\citep{Trenbath1999, Elmqvist2003, Diaz2005}, seemed not to have
+recovered 30 years after disturbance.
 
-The functional redundancy, the functional overlap between species that
-is typical of the huge biodiversity of tropical forests
-\citep{Bellwood2006}, was then not restored 30 years after disturbance
-and this is major to consider as it defines forests' resilience
-\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
-
-Besides the long-term alteration of functional redundancy, there was
-probably persistent compositional changes favoring disturbance resistant
-species \citep{Haddad2008}, lianas or epiphytes \citep{Martin2013} and
-environmental changes, like in the soils nutrient cycling and compaction
-\citep{Olander2005}. These persistent changes highly question forest's
-resilience \citep{Chazdon2003a}. New conditions would not only be longer
-lasting but self-maintained as tied to disturbance regime
-\citep{Burslem2000}. Specificaly, this would impair species contingent
-to undisturbed forests, threatening their maintenance, and run the risk
-to loose cornerstone species and trigger unexpected ecological
-consequences \citep{Jones1994, Diaz2005, Gardner2007}.
+Although disturbance would be mitigated by communities functional
+redundancy, we expect the redundancy itself to be reduced or
+re-organized within the functional space. Globally, disturbed
+communities did not follow consistent trajectories after disturbance and
+30 years after disturbance, irrespective of the intensity, disturbed
+communities displayed lower, similar or higher functional redundancy.
+The redundancy within the functional space of the initial community
+however clearly followed humped-back trajectories which decrease was
+determined by the disturbance intensity and which return to initial
+state was not observed for any disturbed community. Not only did the
+functional redundancy and hence the resilience of pre-disturbance
+community decrease, but also did the ones of disturbance-specific
+communities increase. It means higher chances to have long lasting or
+self-maintained compositional changes towards disturbance resistant
+species, lianas or epiphytes \citep{Haddad2008, Burslem2000, Martin2013}
+and thus highly question forest's resilience \citep{Chazdon2003a}.
+Specificaly, this would impair species contingent to undisturbed
+forests, threatening their maintenance, and run the risk to loose
+cornerstone species and trigger unexpected ecological consequences
+\citep{Jones1994, Diaz2005, Gardner2007}.
 
@@
 
-Our study showed the significant impact of disturbance on tropical
-forests communities and validated the consistency of the IDH in the long
-term. It revealed the contrasting response of communities taxonomic and
-functional characteristics, with persisting impacts on the species
-abundance distibution while the functional diversity and dominant
-functional strategies were restored. Communities recovery therefore
-remained unachieved but consistent for the range of disturbance studied
-here. The length of the recovery, however, severly questioned the
-sustainability of intense selective logging and advocated felling cycle
-much longer than 30 years \citep{Gourlet-Fleury2005}.
+our study disentangled tropical forests functional and taxonomic
+decoupled response to disturbance, with a rapid recovery of communities
+functioning but slower and more variable taxonomic dynamics.
+Consistently with the IDH functional trajectories were constrained by
+environmental pressures favoring pioneers and light-demanding species
+after disturbance, but taxonomic dynamics were more stochastic as they
+involved infrequent speciesfunctoinally redundant with the dominants and
+thus under competitive pressure. While communities functioning rapidly
+recovered, taxonomic trajectories, although consistent and converging
+towards full recovery, were much slower and entailed persisting
+alteration of communities functional redundancy. the trajectories
+therefore suggested a potential sustainability of tropical forests in
+the face of quite intense disturbance but only when followed by long
+recovery periods, longer than 30 years \citep{Gourlet-Fleury2005}. This
+study besides highlighted the central role of recruitment processes
+underlying the IDH mechanisms and holding the whole community response
+to disturbance.
 

### 2018-05-18 WholePlotTrajectories.Rmd
Intro

@@
 
-The large areas covered with tropical forests worldwide hold crucial economic, social and cultural value. 
+The large areas covered with tropical forests worldwide hold crucial economic, social and cultural values. 
 They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The simultaneous increase of forests products demand and substantial climatic changes currently heighten the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015].
-It threatens the maintenance of communities structure, composition and functioning and the underlying dynamics in space and time [@Anderson-Teixeira2013;@Sist2015].
+The simultaneous increase of forests products demand and the substantial climatic changes currently heighten the pressure on remaining forests [@Gibson2011a; @Morales-Hidalgo2015], threatening the maintenance of communities structure, composition and functioning and their dynamics in space and time [@Anderson-Teixeira2013;@Sist2015].
 
-In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events changing both abiotic environment, through the fluxes of light, heat and water, and biotic interactions and competitive pressure [@Goulamoussene2017].
-The cornerstone of tropical forests ecology is to understand the underlyings of ecosystem response to disturbance, its mechanisms and drivers [@White2001;@Chazdon2003a]. 
+In tropical forest, ecological communities are constantly re-shaped by the natural disturbance events changing both abiotic environment, through the fluxes of light, heat and water, and biotic interactions and competitive pressure [@Goulamoussene2017].
+The cornerstone of tropical forests ecology is to understand the mechanisms and the determinants of ecosystems response to disturbance [@White2001;@Chazdon2003a]. 
 For now, this has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
-Consistent models assessed these structural parameters and highlighted the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009]. 
-The response of tree diversity, however, remains unclear albeit it determines ecosystems productivity, stability and functioning [@Tilman2014, @Liang2016] and would be most probably impacted by post-disturbance environmental changes [@Baraloto2012a].
+These structural parameters were thereafter consistently modeled and gave important insights into the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009]. 
+However the response of forests diversity in tree species remains unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014, @Liang2016] and would be most probably impacted by the changes following disturbance [@Baraloto2012a].
 
 In the short-term disturbance demonstrated negligible or even positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
-Validations of the IDH in the long term, however, remain scarce and mainly rely on the analysis of the rough species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and diversity profiles, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, account for species biological attributes and role the ecosystem through a functional approach assesses communities functioning and corresponding mechanisms [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-In that respect major functional traits related to species ecology and performance were largely adopted [@Diaz2005;@Villeger2008a] highlighting, for example, the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition in tropical rainforests [@Molino2001;@Haddad2008].
-In disturbed tropical rainforests, this framework translates into the fostering of disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
-It results in a shift from "conservative" slow-growing species dealing with scarce resources that dominate before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
-It was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
+Still, validations of the IDH remain scarce in the long term and mainly rely on species richness analyses that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+More releveant monitoring would encompass communities composition, crucial for conservation issues, and complete diversity profiles encompassing species evenness in addition to their richness, that reveals ecological rules shaping communities [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Furthermore, account for species biological attributes and role in the ecosystem is essential to understand the correlations between ecosystems biodiversity, functioning and environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+In that respect functional approaches based on major traits related to species ecology and performance was largely adopted [@Diaz2005;@Villeger2008a], for example highlighting in tropical rainforests the environmental filters fostering fast growing species with efficient resources acquisition after disturbance [@Molino2001;@Haddad2008].
+Post-disturbance trajectories then corresponded to a shift from "conservative" slow-growing species dealing with scarce resources dominating before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
+This shift translated into consistent trajectories for key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction and life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
 
-Functional and taxonomic trajectories relate to distinct characteristics and community processes, and might therefore be decoupled as already observed in tropical forests [@Lohbeck2015; @Guariguata2001].
-The combination of taxonomic and functional characteristics is translated by the functional redundancy, measuring the amount of species sharing same trait values [@Carmona2016].
-Functional redundancy is a major determinant of communities resilience as high redundancy, like in highly diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning [@Trenbath1999; @Elmqvist2003; @Diaz2005].
-It is therefore a key element as a link between taxonomic and functional diversities and as a determinant of communities resilience.
+Functional and taxonomic trajectories are both essentiall to fully assess communities response to disturbance.
+Both relate to distinct ecological rules and processes and might then be decoupled as it was observed already in tropical forests [@Lohbeck2015; @Guariguata2001].
+Although taxonomic and functional diversity are complementary, they are combine into communities functional redundancy that measures the amount of species sharing same trait values [@Carmona2016].
+Functional redundancy then explicits the link between otherwise decoupled taxonomic and functional diversities and is central for communities description.
+Besides, functional redundancy determines communities resilience as high redundancy, like in highly diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning [@Trenbath1999; @Elmqvist2003; @Diaz2005].
 
-Grasping all facets of communities response to disturbance then requires to disentangle the taxonomic and functional trajectories in diversity and composition.
-These trajectories would highlight communities resilience and the drivers of their response to disturbance, clarifying the tenants of the IDH in the long term for tropical forests, that are as much insights for future adaptive conservation strategies [@Adler2007].
+Grasp all facets of communities response to disturbance amount to disentangle the taxonomic and functional trajectories in diversity and composition.
+These trajectories would highlight the ecological rules constraining, or not, communities dynamics towards a recovery of initial composition, diversity and functioning.
+They would therefore clarify the tenants in the long term of the debated Intermediate Disturbance Hypothesis for tropical forests, which would be as much insights for future adaptive conservation strategies [@Adler2007].
 Here we monitored over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
 We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
-Specifically, we (i) highlighted the determinant of communities functoinal and taxonomic response to disturbance, (ii) tested the validity of the IDH in the long term for tropical forest and (iii) questioned the completeness of communities recovery regarding their functional redundancy.
+Specifically, we (i) questioned the coupling between taxonomic and functional response to disturbance and identified the underlying assembly processes, which allowed to (ii) clarify the validity of the IDH in the long term for tropical forest and (iii) question the completeness of communities recovery regarding their functional redundancy.
 

### 2018-05-18 AppendixI.Rmd
Results & appendix

@@
 }
+
+Library("kableExtra")
 #source("Scripts/GithubPages.R")
@@
 
-# **Figure A6**: Functional PCA
+# **Figure A6**: Simpson and Rao Linera regrassion summary
+
+```{r LinReg, echo=FALSE,is.as=T}
+LmSummary <- data.frame(
+  Diversity=c(rep("Simpson",2), rep("Rao",2)),
+  Metrics=rep(c("R^2^","p~value~"),times=2),
+  Time1=c("-0.08","0.7",
+            "0.58","4e-3"),
+  Time2=c("0.25","6e-2",
+            "0.58","4e-3"),
+  Time3=c("0.24","6e-2",
+            "0.58","4e-3")
+)
+names(LmSummary) <- c("Diversity","Metrics","10 years", "20 years", "30 years")
+knitr::kable(LmSummary, "html", caption="Linear models information", longtable = FALSE, booktabs = TRUE, escape = FALSE) %>%
+  collapse_rows(columns=1) %>%
+  kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
+```
+
+# **Figure A7**: Functional PCA
 
@@
 
-# **Figure A7**: Overall functional redundancy
+# **Figure A8**: Overall functional redundancy
 

### 2018-05-18 WholePlotTrajectories.Rmd
Results & appendix

@@
 
-Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest corresponding to a tropical wet climate with mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) with a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
-Elevation ranges between 5 and 50 m and mean annual temperature is 26°C. 
-Soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
-The experiment corresponds to a network of twelve 6.25ha plots that have undergone a gradient of three logging, thinning and fuelwood treatments (Table \@ref(tab:Tab1)).
+Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26°C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
+Elevation ranges between 5 and 50 m and soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
+
+The experiment is a network of twelve 6.25ha plots that underwent a gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)).
 Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
@@
 Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Blanc2009]. 
-The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance (ref to be found).
-Biomass measurements were performed with the BIOMASS package R package [@Biomass2018].
+The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance (ref to be found) measured with the BIOMASS R package [@Biomass2018].
 
@@
 
-## Inventories protocol and dataset collection
+## Inventories protocol and dataset collection {#protocols}
 
 The study site corresponds to a tropical rainforest with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical families. 
-In the twelve experimental plots of the experiment, all trees above 10 cm DBH were mapped and measured annually since 1984. 
-During inventories, trees were first identified with a vernacular name assigned by the field team, and afterward with a scientific name assigned by a botanist during regular botanical campaigns. 
-In 1984, specific vernacular names were given to 62 commercial or common species whereas other less common species were identified under two identifiers only separating trees and palm trees. 
-The botanical campaigns carried every 5 to 6 years to identify all trees at the species level only started in 2003 and identification practices varied among plots and successive campaigns. 
-This raised methodological issues as vernacular names usually correspond to different botanical species, resulting in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics through a Bayesian framework.
-The uncertainty propagation was done by the replenishment of inventories completed at genus level from real incomplete ones on the basis of vernacular/botanical names association.
+In the twelve experimental plots of the experiment, all trees above 10 cm DBH are mapped and measured annually since 1984. 
+Trees are first identified during inventories with a vernacular name assigned by the field team, and afterward with a scientific name assigned by a botanist during regular botanical campaigns. 
+In 1984, specific vernacular names are given to 62 commercial or common species whereas more infrequent ones were identified under general identifiers only distinguiching trees and palm trees. 
+From 2003 botanical campaigns were conducted every 5 to 6 years to identify all trees at the species level but identification practices still varied among plots and campaigns.
 
-Vernacular names were replaced through multinomial trials $M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)$ based on the observed association probability $\big[\alpha_1, \alpha_2,…, \alpha_3\big]$ between each vernacular name *v* and the species $\big[s_1, s_2, …, s_N\big]$ recorded in the inventory.
+This variability of protocols raised methodological issues as vernacular names usually correspond to different botanical species. 
+It resulted in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics.
+The uncertainty propagation was done through a Bayesian framework reconstituting complete inventories at genus level from real incomplete ones on the basis of vernacular/botanical names association. 
+Vernacular names were replaced through multinomial trials $M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)$ based on the association probability $\big[\alpha_1, \alpha_2,…, \alpha_3\big]$ observed across all inventories between each vernacular name *v* and the species $\big[s_1, s_2, …, s_N\big]$.
 See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.
 
-The functional diversity metrics used a dataset for 6 functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, the leaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass).
+The functional approach used a dataset of 6 functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, the leaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass).
+The trait database came from the BRIDGE project [^1] where trait values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 botanical species pertaining to 157 botanical genera.
+Missing trait values were filled using multivariate imputation by chained equation (mice) from the mice R package [@Mice2011].
+Imputations were restricted within genus, or family when samples wre too scarce, in order to account for the phylogenetic signal of the functional traits. 
+Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level (same genus or family). 
+As seed mass information corresponds to a classification into mass classes, no data filling process was applied and analysis were restricted to the 414 botanical species recorded.
 
-The trait database came from the BRIDGE project [^1] where trait values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou. 
-Missing trait values were filled using multivariate imputation by chained equation (mice) restricted to samples pertaining to the next higher taxonomic level, in order to account for the phylogenetic signal of the functional traits.
-The dataset comprised 294 botanical species pertaining to 157 botanical genera.
-
-Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level. 
-As seed mass information corresponds to a classification into mass classes, no data filling process was applied so analysis were performed considering the 414 botanical species of the seed mass dataset.
 All composition and diversity metrics corresponded to the average obtained after 50 iterations of the taxonomic uncertainty propagation framework and of the filling process of missing trait values.
@@
 
-To counter the remaining taxonomic uncertainty, plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
-Variations in plots taxonomic and functional composition after disturbance was visualized by their trajectories over 30 years in a two-dimensional ordination space.
-Two NMDS were conducted, either from taxonomic flora inventories or from plots functional composition based on the 7 leaf, stem and life history traits (without seed mass classes).
-In both cases the NMDS were performed using occurrence-based (Jaccard) as well as abundance-based (Bray-Curtis) dissimilarity measures. 
-Trajectories along time were reported through the euclidean distance of successive inventories to the reference inventories in 1989, 5 years after disturbance, when the uncertainty degree did not exceed 30% of undetermined trees.
+To counter taxonomic uncertainties due to the variability of botanical identification protocols (see [#protocols]), the taxonomic composition and diversity analysis were conducted at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
+Trajectories of communities taxonomic and functional variations in composition after disturbance were followed in a two-dimensional ordination space the 30 years monitored.
+Two NMDS were conducted to map either taxonomic flora inventories or communities functional composition based on the 7 leaf, stem and life history traits (without seed mass classes).
+In both cases the NMDS were performed using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) dissimilarity measures. 
+Trajectories along time in the plan were reported through the euclidean distance of successive inventories to the reference inventories in 1989, 5 years after disturbance, when the uncertainty degree did not exceed 30% of undetermined trees.
 The trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004].
@@
 
-The functional redundancy is measured as the overlap among species in communities' functional space [@Carmona2016].
-First, the samples of the trait database are mapped in a 2-dimensional plan with a PCA analysis.
-Then, multivariate kernel density estimator associated with individual trees were give species traits probability distribution (TDP).
-TDP, weighted by species abundance, are summed to give communities TDP: the functional redundancy is the sum of TDP overlaps,  corresponding to the average number of species that could be removed from without reducing the functional space (see appendix I for a more comprehensive sheme).
+The impacts of initial disturbance were first tested with the spearman rank correlation between the extremum of taxonomic and functional metrics reached over the 30 years and the initial %AGB removed.
+Then they were analysed through the linear correlations between Simpson and Rao diversities and the initial %AGB removed at 10, 20 and 30 years after disturbance.
+
+The functional redundancy was measured as the overlap among species in communities' functional space [@Carmona2016]. 
+The samples of the trait database were first mapped in a 2-dimensional plan from a PCA analysis. 
+Then, multivariate kernel density estimator associated with individual trees returned species traits probability distribution (TDP).
+Species TDP weighted by species abundance were eventually summed for each community: the functional redundancy was the sum of TDPs overlap, expressed as the average number of species that could be removed from without reducing the functional space (see appendix I for a more comprehensive sheme).
 
@@
 
+In the inventories from 1989 (5 years after disturbance) to 2015 (31 years after disturbance), 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
 Communities taxonomic diversity trajectories were examined through the Richness, Shannon and Simpson diversities at genus level, in relation to the 1989 inventories (5 years after disturbance) (See annexe I).
 For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 years monitored.
-After low disturbance intensity the richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
-After intense disturbance, plots taxonomic richness followed unimodal trajectories first decrease for around ten years and then increasing to return to pre-disturbance values.
-For all disturbed plots the taxonomic evenness (Shannon and Simpson diversities) increased, following unimodal trajectories with a just beginning return towards initial values after 30 years.
-The maximum evenness, reached after around 20 years, was positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
-Only two T3 plots (plots 8 and 12) were still increasing 30 years after disturbance \@ref(fig:IDHplot), suggesting similar but delayed trajectories.
+In disturbed communities the taxonomic richness increased after low disturbance intensity, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2) while it followed unimodal trajectories after intense disturbance, decreasing for ten years before recovering pre-disturbance values.
+In all disturbed plots the taxonomic evenness (Shannon and Simpson diversities) increased, following unimodal trajectories with a maximum, reached after around 20 years, positively correlated to the disturbance treatment ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
+Return towards initial evenness values was beginning after 30 years except for two T3 plots (plots 8 and 12) which evenness still increased, suggesting similar but delayed trajectories \@ref(fig:IDHplot).
 
-Communities functional diversity trajectories were examined through the Rao diversity based on the 7 leaf, stem and life history traits (the seed mass was excluded as a qualitative variable).
+Trajectories of communities functional diversity were examined through the Rao diversity based on the 7 leaf, stem and life history traits (to the exception of seed mass).
 The plot 7 from treatment 1 displayed a constantly outlying diversity and was removed from the graphical representation for better readability (see appendix for full graphs).
-The functional diversity of all undisturbed communities remained stable along the 30 years.
-For all disturbed communities the functional diversity followed unimodal trajectories with a return towards the initial state, comparable to the control plots.
+In undisturbed plots the functional diversity remained stable along the 30 years while in disturbed plots it followed unimodal trajectories with a return towards initial values that strated around 20 years after disturbance.
 
-The impact of disturbance was examined specifically through the linear correlation between the disturbance intensity (intial %AGB removed) and the Simpson and Rao diversities (diversities of order 2) at three focus times after disturbance (10, 20 and 30 years) \@ref(fig:IDHplot).
-The Simpson diversity was weakly related to the disturbance intensity ($R^2<0.25$) and only from 20 years after disturbance.
-The Rao diversity was more strongly correlated to the disturbance intensity ($0.55<R^2<0.75$) for the three times studied.
-The impact of disturbance, translated by the slope of the linear correlation, was the highest 20 years after disturbance.
+The impact of disturbance was examined specifically through the linear correlation between the intial %AGB removed and the Simpson and Rao diversities (diversities of order 2) after 10, 20 and 30 years \@ref(fig:IDHplot).
+The correlation with disturbance intensity was weak for the Simpson diversity ($R^2<0.25$) and only valid from 20 years after disturbance but it was much stronger for the Rao diversity ($0.55<R^2<0.75$) for all the time studied.
+Slope of linear correlations, reflecting the impact of disturbance, was the highest 20 years after disturbance.
 
@@
 
-In the inventories from 1989 (5 years after disturbance) to 2015 (31 years after disturbance), 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
-The trajectories of taxonomic and functional composition were visualized in a two dimensional ordination space mapping the successive inventories according to their flora and the corresponding traits.
-Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity.
-Both metrics gave similar results, so only analysis using Bray-Curtis dissimilarity will be discussed here.
+The trajectories of taxonomic and functional composition were visualized in a two dimensional ordination space mapping the successive inventories according to their flora and corresponding traits.
+Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity, both giving similar results only analysis using Bray-Curtis dissimilarity are discussed.
 
-While both taxonomic and functional composition remained stable in undistrubed communities, after disturbed they followed consistent trajectories over time that revealed significant compositional changes.
-Considering the mapping of functional traits in the same dimensional space, the post-disturbance trajectories corresponded to shifts towards more acquisitive functional strategies: disturbed communities changed from high average WD to high average SLA and chlorophyll content.
-For disturbed communities the distance of successive inventories to the reference inventory in 1989 followed unimodal trajectories (AppendixI, Figure A2).
-The maximum of communities dissimilarity to their initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.91$ and $\rho_{spearman}^{functional}=0.96$ respectively).
-The time at maximum dissimilarity was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition. 
-All trajectories followed cyclic composition changes with a return towards the initial composition (Figure \@ref(fig:NMDSplans)).
+While both taxonomic and functional composition remained stable in undistrubed communities, they followed consistent trajectories over time after disturbance which revealed significant compositional changes.
+According to the mapping of functional traits (see appendix I) these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WD to high average SLA and chlorophyll content.
+For disturbed communities the distance of successive inventories to the 1989 reference inventory followed unimodal trajectories translating cyclic compositional changes with a recovery of the initial composition (Figure \@ref(fig:NMDSplans)).
+The maximum dissimilarity with the initial state was positively correlated to the disturbance treatment for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.91$ and $\rho_{spearman}^{functional}=0.96$ respectively) and the time at maximum was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition.
 
@@
 
-The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits visualized through plots community weighted means (CWM) (Figure \@ref(fig:CWM)) of the leaves thickness, chlorophyll content, toughness and specific area, wood specific gravity, bark thickness, seed mass and maximum adult height.
+The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits visualized with the trajectories of community weighted means (CWM) of leaves economics (leaves thickness, chlorophyll content, toughness and specific area), wood economics (wood specific gravity, bark thickness), and life history traits (seed mass and maximum adult height) (Figure \@ref(fig:CWM)).
 
-Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots, all traits and seed mass proportions followed a unimodal trajectories and either stabilized or returned towards their initial values 30 years after disturbance.
-At that time the weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value and than these of the control plots (Figure \@ref(fig:CWM)).
-The weighted means of bark thickness (*Bark_thick*) similarly remained substantilly higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value, similar as those of undisturbed plots.
+Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions followed unimodal trajectories either stabilizing or returning towards their initial values.
+Thirty years after disturbance the weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value (Figure \@ref(fig:CWM)).
+The weighted means of bark thickness (*Bark_thick*) similarly remained substantially higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value.
 For all traits the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.67$, $\rho_{spearman}^{L_{chloro}}=0.45$, $\rho_{spearman}^{L_{toughness}}=-0.43$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WD}=-0.78$, $\rho_{spearman}^{Bark-thickness}=0.88$, $\rho_{spearman}^{Hmax}=-0.48$).
@@
 
-Communities functional redundancy was measured as the sum of weighted overlap among species in communities' functional space.
-Functional spaces were defined in the main two-dimensional space of a PCA analysis summarizing the 7 leaf, stem, and maximum height traits (see appendix I for PCA details). 
-Communities functional redundancy remained stable in control plots but after disturbance the redundancy trajectories were quite variable (See appendix I).
+Communities functional redundancy was measured as the sum within communities the species weighted functional overlap based on the 7 leaf, stem, and maximum height traits (see appendix I for PCA details).
+Communities functional redundancy remained stable in control plots but after disturbance the redundancy trajectories were quite variable (See appendix I) and apparently independently of the initial disturbance.
 Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place before increasing to edge, recover or exceed the initial value.
@@
 Considering the functional redundancy restricted to the functional space of the initial inventory, all disturbed plots followed similar decreasing humped-shaped trajectories (\@ref(fig:RedFun_rest)).
-The maximum redundancy loss observed was positively correlated with the disturbance intensity (XX spearman to be measured) and none of the disturbed plots had recovered the functional redundancy within the initial functional space.
+The maximum redundancy loss was positively correlated with the disturbance intensity (XX spearman to be measured) and the initial value had not recovered for any disturbed communities.
 

### 2018-05-18 WholePlotTrajectories.tex
Results & appendix

@@
 The large areas covered with tropical forests worldwide hold crucial
-economic, social and cultural value. They provide wood and multiple
+economic, social and cultural values. They provide wood and multiple
 non-timber forest products, shelter a diversified fauna, regulate the
@@
 cultural and human well-being. The simultaneous increase of forests
-products demand and substantial climatic changes currently heighten the
-pressure on the remaining forests
-\citep{Gibson2011a, Morales-Hidalgo2015}. It threatens the maintenance
-of communities structure, composition and functioning and the underlying
-dynamics in space and time \citep{Anderson-Teixeira2013, Sist2015}.
+products demand and the substantial climatic changes currently heighten
+the pressure on remaining forests
+\citep{Gibson2011a, Morales-Hidalgo2015}, threatening the maintenance of
+communities structure, composition and functioning and their dynamics in
+space and time \citep{Anderson-Teixeira2013, Sist2015}.
 
-In tropical forest, ecological communities are contantly re-shaped by
+In tropical forest, ecological communities are constantly re-shaped by
 the natural disturbance events changing both abiotic environment,
@@
 competitive pressure \citep{Goulamoussene2017}. The cornerstone of
-tropical forests ecology is to understand the underlyings of ecosystem
-response to disturbance, its mechanisms and drivers
+tropical forests ecology is to understand the mechanisms and the
+determinants of ecosystems response to disturbance
 \citep{White2001, Chazdon2003a}. For now, this has been largely studied
@@
 aboveground biomass, tree height or stem density
-\citep{Piponiot2016, Rutishauser2016}. Consistent models assessed these
-structural parameters and highlighted the maintenance of ecosystems
-processes and services \citep{Denslow2000, Blanc2009}. The response of
-tree diversity, however, remains unclear albeit it determines ecosystems
-productivity, stability and functioning
+\citep{Piponiot2016, Rutishauser2016}. These structural parameters were
+thereafter consistently modeled and gave important insights into the
+maintenance of ecosystems processes and services
+\citep{Denslow2000, Blanc2009}. However the response of forests
+diversity in tree species remains unclear, albeit it determines the
+productivity, stability and functioning of ecosystems
 \citep[\citet{Liang2016}]{Tilman2014} and would be most probably
-impacted by post-disturbance environmental changes
-\citep{Baraloto2012a}.
+impacted by the changes following disturbance \citep{Baraloto2012a}.
 
@@
 diversity at intermediate disturbance intensity
-\citep{Molino2001, Kariuki2006a, Berry2008a}. Validations of the IDH in
-the long term, however, remain scarce and mainly rely on the analysis of
-the rough species richness that gives limited or misleading information
-on forests recovery and functioning \citep{Martin2015, Chaudhary2016}.
-More releveant monitoring would encompass communities composition, that
-is crucial for conservation issues, and diversity profiles, that reveals
-ecological rules underlying communities' structure
+\citep{Molino2001, Kariuki2006a, Berry2008a}. Still, validations of the
+IDH remain scarce in the long term and mainly rely on species richness
+analyses that gives limited or misleading information on forests
+recovery and functioning \citep{Martin2015, Chaudhary2016}. More
+releveant monitoring would encompass communities composition, crucial
+for conservation issues, and complete diversity profiles encompassing
+species evenness in addition to their richness, that reveals ecological
+rules shaping communities
 \citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, account
-for species biological attributes and role the ecosystem through a
-functional approach assesses communities functioning and corresponding
-mechanisms
+for species biological attributes and role in the ecosystem is essential
+to understand the correlations between ecosystems biodiversity,
+functioning and environmental constraints
 \citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
-respect major functional traits related to species ecology and
-performance were largely adopted \citep{Diaz2005, Villeger2008a}
-highlighting, for example, the environmental filters fostering
-disturbance resistant species with rapid growth and efficient resources
-acquisition in tropical rainforests \citep{Molino2001, Haddad2008}. In
-disturbed tropical rainforests, this framework translates into the
-fostering of disturbance resistant species with rapid growth and
-efficient resources acquisition \citep{Molino2001, Haddad2008}. It
-results in a shift from ``conservative'' slow-growing species dealing
-with scarce resources that dominate before disturbance, to
+respect functional approaches based on major traits related to species
+ecology and performance was largely adopted
+\citep{Diaz2005, Villeger2008a}, for example highlighting in tropical
+rainforests the environmental filters fostering fast growing species
+with efficient resources acquisition after disturbance
+\citep{Molino2001, Haddad2008}. Post-disturbance trajectories then
+corresponded to a shift from ``conservative'' slow-growing species
+dealing with scarce resources dominating before disturbance, to
 ``acquisitive'' fast-growing species with rapid and efficient use of
-abundant resources \citep{TerSteege2001, Reich2014, Herault2011}. It was
-mirrored by shifts in key functional traits related to resource
-acquisition (leaf area, density and chlorophyll content, and stem
-specific gravity and bark thickness), tree growth and reproduction life
-history traits (seed mass and maximum height)
+abundant resources \citep{TerSteege2001, Reich2014, Herault2011}. This
+shift translated into consistent trajectories for key functional traits
+related to resource acquisition (leaf area, density and chlorophyll
+content, and stem specific gravity and bark thickness), tree growth and
+reproduction and life history traits (seed mass and maximum height)
 \citep{Wright2004, Westoby2006a, Chave2009b}.
 
-Functional and taxonomic trajectories relate to distinct characteristics
-and community processes, and might therefore be decoupled as already
-observed in tropical forests \citep{Lohbeck2015, Guariguata2001}. The
-combination of taxonomic and functional characteristics is translated by
-the functional redundancy, measuring the amount of species sharing same
-trait values \citep{Carmona2016}. Functional redundancy is a major
-determinant of communities resilience as high redundancy, like in highly
-diverse tropical forests \citep{Bellwood2006}, mitigates the impacts of
-species removal on ecosystem functioning
-\citep{Trenbath1999, Elmqvist2003, Diaz2005}. It is therefore a key
-element as a link between taxonomic and functional diversities and as a
-determinant of communities resilience.
+Functional and taxonomic trajectories are both essentiall to fully
+assess communities response to disturbance. Both relate to distinct
+ecological rules and processes and might then be decoupled as it was
+observed already in tropical forests
+\citep{Lohbeck2015, Guariguata2001}. Although taxonomic and functional
+diversity are complementary, they are combine into communities
+functional redundancy that measures the amount of species sharing same
+trait values \citep{Carmona2016}. Functional redundancy then explicits
+the link between otherwise decoupled taxonomic and functional
+diversities and is central for communities description. Besides,
+functional redundancy determines communities resilience as high
+redundancy, like in highly diverse tropical forests
+\citep{Bellwood2006}, mitigates the impacts of species removal on
+ecosystem functioning \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-Grasping all facets of communities response to disturbance then requires
-to disentangle the taxonomic and functional trajectories in diversity
-and composition. These trajectories would highlight communities
-resilience and the drivers of their response to disturbance, clarifying
-the tenants of the IDH in the long term for tropical forests, that are
-as much insights for future adaptive conservation strategies
-\citep{Adler2007}. Here we monitored over 30 years the response of 75 ha
-of forests plots set up on a gradient of disturbance intensity, from 10
-to 60\% of ecosystem biomass removed. We made use of a large functional
-traits database browsing major leaf, stem and seed traits and species
-maximum height to draw the trajectories over time of communities
-taxonomic and functional composition and diversity. Specifically, we (i)
-highlighted the determinant of communities functoinal and taxonomic
-response to disturbance, (ii) tested the validity of the IDH in the long
-term for tropical forest and (iii) questioned the completeness of
-communities recovery regarding their functional redundancy.
+Grasp all facets of communities response to disturbance amount to
+disentangle the taxonomic and functional trajectories in diversity and
+composition. These trajectories would highlight the ecological rules
+constraining, or not, communities dynamics towards a recovery of initial
+composition, diversity and functioning. They would therefore clarify the
+tenants in the long term of the debated Intermediate Disturbance
+Hypothesis for tropical forests, which would be as much insights for
+future adaptive conservation strategies \citep{Adler2007}. Here we
+monitored over 30 years the response of 75 ha of forests plots set up on
+a gradient of disturbance intensity, from 10 to 60\% of ecosystem
+biomass removed. We made use of a large functional traits database
+browsing major leaf, stem and seed traits and species maximum height to
+draw the trajectories over time of communities taxonomic and functional
+composition and diversity. Specifically, we (i) questioned the coupling
+between taxonomic and functional response to disturbance and identified
+the underlying assembly processes, which allowed to (ii) clarify the
+validity of the IDH in the long term for tropical forest and (iii)
+question the completeness of communities recovery regarding their
+functional redundancy.
 
@@
 Paracou station in French Guiana (5°18'N and 52°53'W) is located in a
-lowland tropical rain forest corresponding to a tropical wet climate
-with mean annual precipitation averaging 2980 mm.y\textsuperscript{-1}
-(30-y period) with a 3-month dry season (\textless{} 100
-mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
-one-month dry season in March \citep{Wagner2011}. Elevation ranges
-between 5 and 50 m and mean annual temperature is 26°C. Soils correspond
-to thin acrisols over a layer of transformed saprolite with low
-permeability generating lateral drainage during heavy rains. The
-experiment corresponds to a network of twelve 6.25ha plots that have
-undergone a gradient of three logging, thinning and fuelwood treatments
+lowland tropical rain forest in a tropical wet climate with mean annual
+temperature of 26°C, mean annual precipitation averaging 2980
+mm.y\textsuperscript{-1} (30-y period) and a 3-month dry season
+(\textless{} 100 mm.month\textsuperscript{-1}) from mid-August to
+mid-November, and a one-month dry season in March \citep{Wagner2011}.
+Elevation ranges between 5 and 50 m and soils correspond to thin
+acrisols over a layer of transformed saprolite with low permeability
+generating lateral drainage during heavy rains.
+
+The experiment is a network of twelve 6.25ha plots that underwent a
+gradient of three logging, thinning and fuelwood cutting treatments
 (Table \ref{tab:Tab1}). Disturbance treatments were attributed according
@@
 lost between the first inventory in 1984 and five years after
-disturbance (ref to be found). Biomass measurements were performed with
-the BIOMASS package R package \citep{Biomass2018}.
+disturbance (ref to be found) measured with the BIOMASS R package
+\citep{Biomass2018}.
 
@@
 \subsection{Inventories protocol and dataset
-collection}\label{inventories-protocol-and-dataset-collection}
+collection}\label{protocols}
 
@@
 families. In the twelve experimental plots of the experiment, all trees
-above 10 cm DBH were mapped and measured annually since 1984. During
-inventories, trees were first identified with a vernacular name assigned
-by the field team, and afterward with a scientific name assigned by a
+above 10 cm DBH are mapped and measured annually since 1984. Trees are
+first identified during inventories with a vernacular name assigned by
+the field team, and afterward with a scientific name assigned by a
 botanist during regular botanical campaigns. In 1984, specific
-vernacular names were given to 62 commercial or common species whereas
-other less common species were identified under two identifiers only
-separating trees and palm trees. The botanical campaigns carried every 5
-to 6 years to identify all trees at the species level only started in
-2003 and identification practices varied among plots and successive
-campaigns. This raised methodological issues as vernacular names usually
-correspond to different botanical species, resulting in significant
-taxonomic uncertainties that had to be propagated to composition and
-diversity metrics through a Bayesian framework. The uncertainty
-propagation was done by the replenishment of inventories completed at
+vernacular names are given to 62 commercial or common species whereas
+more infrequent ones were identified under general identifiers only
+distinguiching trees and palm trees. From 2003 botanical campaigns were
+conducted every 5 to 6 years to identify all trees at the species level
+but identification practices still varied among plots and campaigns.
+
+This variability of protocols raised methodological issues as vernacular
+names usually correspond to different botanical species. It resulted in
+significant taxonomic uncertainties that had to be propagated to
+composition and diversity metrics. The uncertainty propagation was done
+through a Bayesian framework reconstituting complete inventories at
 genus level from real incomplete ones on the basis of
-vernacular/botanical names association.
+vernacular/botanical names association. Vernacular names were replaced
+through multinomial trials
+\(M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)\)
+based on the association probability
+\(\big[\alpha_1, \alpha_2,…, \alpha_3\big]\) observed across all
+inventories between each vernacular name \emph{v} and the species
+\(\big[s_1, s_2, …, s_N\big]\). See appendix 1 and
+\citet{Aubry-Kientz2013} for the detailed methodology.
 
-Vernacular names were replaced through multinomial trials
-\(M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)\)
-based on the observed association probability
-\(\big[\alpha_1, \alpha_2,…, \alpha_3\big]\) between each vernacular
-name \emph{v} and the species \(\big[s_1, s_2, …, s_N\big]\) recorded in
-the inventory. See appendix 1 and \citet{Aubry-Kientz2013} for the
-detailed methodology.
-
-The functional diversity metrics used a dataset for 6 functional traits
+The functional approach used a dataset of 6 functional traits
 representing leaf economics (leaves thickness, toughness, total
@@
 mass) and wood economics (wood specific gravity and bark thickness), and
-life history traits (maximum specific height and seed mass).
+life history traits (maximum specific height and seed mass). The trait
+database came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}
+where trait values were assessed from a selection of individuals located
+in nine permanent plots in French Guiana, including two in Paracou, and
+comprised 294 botanical species pertaining to 157 botanical genera.
+Missing trait values were filled using multivariate imputation by
+chained equation (mice) from the mice R package \citep{Mice2011}.
+Imputations were restricted within genus, or family when samples wre too
+scarce, in order to account for the phylogenetic signal of the
+functional traits. Whenever a species inventoried was not in the
+dataset, it was attributed a set of traits values randomly sampled among
+species of the same next higher taxonomic level (same genus or family).
+As seed mass information corresponds to a classification into mass
+classes, no data filling process was applied and analysis were
+restricted to the 414 botanical species recorded.
 
-The trait database came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}
-where trait values were assessed from a selection of individuals located
-in nine permanent plots in French Guiana, including two in Paracou.
-Missing trait values were filled using multivariate imputation by
-chained equation (mice) restricted to samples pertaining to the next
-higher taxonomic level, in order to account for the phylogenetic signal
-of the functional traits. The dataset comprised 294 botanical species
-pertaining to 157 botanical genera.
-
-Whenever a species inventoried was not in the dataset, it was attributed
-a set of traits values randomly sampled among species of the same next
-higher taxonomic level. As seed mass information corresponds to a
-classification into mass classes, no data filling process was applied so
-analysis were performed considering the 414 botanical species of the
-seed mass dataset. All composition and diversity metrics corresponded to
-the average obtained after 50 iterations of the taxonomic uncertainty
-propagation framework and of the filling process of missing trait
-values.
+All composition and diversity metrics corresponded to the average
+obtained after 50 iterations of the taxonomic uncertainty propagation
+framework and of the filling process of missing trait values.
 
@@
 
-To counter the remaining taxonomic uncertainty, plots taxonomic
-composition and diversity were analysed at the genus level, \emph{i.e.}
-referring to the genus of observed or trialed botanical names.
-Variations in plots taxonomic and functional composition after
-disturbance was visualized by their trajectories over 30 years in a
-two-dimensional ordination space. Two NMDS were conducted, either from
-taxonomic flora inventories or from plots functional composition based
-on the 7 leaf, stem and life history traits (without seed mass classes).
-In both cases the NMDS were performed using occurrence-based (Jaccard)
-as well as abundance-based (Bray-Curtis) dissimilarity measures.
-Trajectories along time were reported through the euclidean distance of
-successive inventories to the reference inventories in 1989, 5 years
-after disturbance, when the uncertainty degree did not exceed 30\% of
-undetermined trees. The trajectories of the leaf and stem and life
-traits were also visualized with the community weighted means (CWM),
-representing the average trait value in a community weighted by relative
-abundance of the species carrying each value
+To counter taxonomic uncertainties due to the variability of botanical
+identification protocols (see {[}\#protocols{]}), the taxonomic
+composition and diversity analysis were conducted at the genus level,
+\emph{i.e.} referring to the genus of observed or trialed botanical
+names. Trajectories of communities taxonomic and functional variations
+in composition after disturbance were followed in a two-dimensional
+ordination space the 30 years monitored. Two NMDS were conducted to map
+either taxonomic flora inventories or communities functional composition
+based on the 7 leaf, stem and life history traits (without seed mass
+classes). In both cases the NMDS were performed using occurrence-based
+(Jaccard) and abundance-based (Bray-Curtis) dissimilarity measures.
+Trajectories along time in the plan were reported through the euclidean
+distance of successive inventories to the reference inventories in 1989,
+5 years after disturbance, when the uncertainty degree did not exceed
+30\% of undetermined trees. The trajectories of the leaf and stem and
+life traits were also visualized with the community weighted means
+(CWM), representing the average trait value in a community weighted by
+relative abundance of the species carrying each value
 \citep{Diaz2007, Garnier2004}. To compensate the intrinsinc difference
@@
 
-The functional redundancy is measured as the overlap among species in
-communities' functional space \citep{Carmona2016}. First, the samples of
-the trait database are mapped in a 2-dimensional plan with a PCA
+The impacts of initial disturbance were first tested with the spearman
+rank correlation between the extremum of taxonomic and functional
+metrics reached over the 30 years and the initial \%AGB removed. Then
+they were analysed through the linear correlations between Simpson and
+Rao diversities and the initial \%AGB removed at 10, 20 and 30 years
+after disturbance.
+
+The functional redundancy was measured as the overlap among species in
+communities' functional space \citep{Carmona2016}. The samples of the
+trait database were first mapped in a 2-dimensional plan from a PCA
 analysis. Then, multivariate kernel density estimator associated with
-individual trees were give species traits probability distribution
-(TDP). TDP, weighted by species abundance, are summed to give
-communities TDP: the functional redundancy is the sum of TDP overlaps,
-corresponding to the average number of species that could be removed
-from without reducing the functional space (see appendix I for a more
+individual trees returned species traits probability distribution (TDP).
+Species TDP weighted by species abundance were eventually summed for
+each community: the functional redundancy was the sum of TDPs overlap,
+expressed as the average number of species that could be removed from
+without reducing the functional space (see appendix I for a more
 comprehensive sheme).
@@
 
+In the inventories from 1989 (5 years after disturbance) to 2015 (31
+years after disturbance), 828388 trees and 591 botanical species
+pertaining to 223 genus and 64 botanical families were recorded.
 Communities taxonomic diversity trajectories were examined through the
@@
 undisturbed plots the Richness, Shannon and Simpson diversities remained
-stable over the 30 years monitored. After low disturbance intensity the
-richness increased, reaching a maximum gain of 14 botanical genera (plot
-3 from treatment 2). After intense disturbance, plots taxonomic richness
-followed unimodal trajectories first decrease for around ten years and
-then increasing to return to pre-disturbance values. For all disturbed
+stable over the 30 years monitored. In disturbed communities the
+taxonomic richness increased after low disturbance intensity, reaching a
+maximum gain of 14 botanical genera (plot 3 from treatment 2) while it
+followed unimodal trajectories after intense disturbance, decreasing for
+ten years before recovering pre-disturbance values. In all disturbed
 plots the taxonomic evenness (Shannon and Simpson diversities)
-increased, following unimodal trajectories with a just beginning return
-towards initial values after 30 years. The maximum evenness, reached
-after around 20 years, was positively correlated to the disturbance
-intensity (\(\rho_{spearman}^{Shannon}=0.86\), and
-\(\rho_{spearman}^{Simpson}=0.89\)). Only two T3 plots (plots 8 and 12)
-were still increasing 30 years after disturbance \ref{fig:IDHplot},
-suggesting similar but delayed trajectories.
+increased, following unimodal trajectories with a maximum, reached after
+around 20 years, positively correlated to the disturbance treatment
+(\(\rho_{spearman}^{Shannon}=0.86\), and
+\(\rho_{spearman}^{Simpson}=0.89\)). Return towards initial evenness
+values was beginning after 30 years except for two T3 plots (plots 8 and
+12) which evenness still increased, suggesting similar but delayed
+trajectories \ref{fig:IDHplot}.
 
-Communities functional diversity trajectories were examined through the
-Rao diversity based on the 7 leaf, stem and life history traits (the
-seed mass was excluded as a qualitative variable). The plot 7 from
-treatment 1 displayed a constantly outlying diversity and was removed
-from the graphical representation for better readability (see appendix
-for full graphs). The functional diversity of all undisturbed
-communities remained stable along the 30 years. For all disturbed
-communities the functional diversity followed unimodal trajectories with
-a return towards the initial state, comparable to the control plots.
+Trajectories of communities functional diversity were examined through
+the Rao diversity based on the 7 leaf, stem and life history traits (to
+the exception of seed mass). The plot 7 from treatment 1 displayed a
+constantly outlying diversity and was removed from the graphical
+representation for better readability (see appendix for full graphs). In
+undisturbed plots the functional diversity remained stable along the 30
+years while in disturbed plots it followed unimodal trajectories with a
+return towards initial values that strated around 20 years after
+disturbance.
 
 The impact of disturbance was examined specifically through the linear
-correlation between the disturbance intensity (intial \%AGB removed) and
-the Simpson and Rao diversities (diversities of order 2) at three focus
-times after disturbance (10, 20 and 30 years) \ref{fig:IDHplot}. The
-Simpson diversity was weakly related to the disturbance intensity
-(\(R^2<0.25\)) and only from 20 years after disturbance. The Rao
-diversity was more strongly correlated to the disturbance intensity
-(\(0.55<R^2<0.75\)) for the three times studied. The impact of
-disturbance, translated by the slope of the linear correlation, was the
-highest 20 years after disturbance.
+correlation between the intial \%AGB removed and the Simpson and Rao
+diversities (diversities of order 2) after 10, 20 and 30 years
+\ref{fig:IDHplot}. The correlation with disturbance intensity was weak
+for the Simpson diversity (\(R^2<0.25\)) and only valid from 20 years
+after disturbance but it was much stronger for the Rao diversity
+(\(0.55<R^2<0.75\)) for all the time studied. Slope of linear
+correlations, reflecting the impact of disturbance, was the highest 20
+years after disturbance.
 
@@
 
-In the inventories from 1989 (5 years after disturbance) to 2015 (31
-years after disturbance), 828388 trees and 591 botanical species
-pertaining to 223 genus and 64 botanical families were recorded. The
-trajectories of taxonomic and functional composition were visualized in
-a two dimensional ordination space mapping the successive inventories
-according to their flora and the corresponding traits. Classifications
-were performed using either abundance-based Bray-Curtis (Figure
-\ref{fig:NMDSplans}) or incidence-based Jaccard dissimilarity. Both
-metrics gave similar results, so only analysis using Bray-Curtis
-dissimilarity will be discussed here.
+The trajectories of taxonomic and functional composition were visualized
+in a two dimensional ordination space mapping the successive inventories
+according to their flora and corresponding traits. Classifications were
+performed using either abundance-based Bray-Curtis (Figure
+\ref{fig:NMDSplans}) or incidence-based Jaccard dissimilarity, both
+giving similar results only analysis using Bray-Curtis dissimilarity are
+discussed.
 
 While both taxonomic and functional composition remained stable in
-undistrubed communities, after disturbed they followed consistent
-trajectories over time that revealed significant compositional changes.
-Considering the mapping of functional traits in the same dimensional
-space, the post-disturbance trajectories corresponded to shifts towards
-more acquisitive functional strategies: disturbed communities changed
-from high average WD to high average SLA and chlorophyll content. For
-disturbed communities the distance of successive inventories to the
-reference inventory in 1989 followed unimodal trajectories (AppendixI,
-Figure A2). The maximum of communities dissimilarity to their initial
-state was positively correlated to the disturbance intensity for both
-taxonomic and functional composition
-(\(\rho_{spearman}^{taxonomic}=0.91\) and
-\(\rho_{spearman}^{functional}=0.96\) respectively). The time at maximum
-dissimilarity was reached around 26 years after disturbance for
-taxonomic composition and 22 years for functional composition. All
-trajectories followed cyclic composition changes with a return towards
-the initial composition (Figure \ref{fig:NMDSplans}).
+undistrubed communities, they followed consistent trajectories over time
+after disturbance which revealed significant compositional changes.
+According to the mapping of functional traits (see appendix I) these
+compositional changes corresponded to shifts towards species with more
+acquisitive functional strategies, from communities with high average WD
+to high average SLA and chlorophyll content. For disturbed communities
+the distance of successive inventories to the 1989 reference inventory
+followed unimodal trajectories translating cyclic compositional changes
+with a recovery of the initial composition (Figure \ref{fig:NMDSplans}).
+The maximum dissimilarity with the initial state was positively
+correlated to the disturbance treatment for both taxonomic and
+functional composition (\(\rho_{spearman}^{taxonomic}=0.91\) and
+\(\rho_{spearman}^{functional}=0.96\) respectively) and the time at
+maximum was reached around 26 years after disturbance for taxonomic
+composition and 22 years for functional composition.
 
@@
 with consistent trajectories of the 8 functional and life history traits
-visualized through plots community weighted means (CWM) (Figure
-\ref{fig:CWM}) of the leaves thickness, chlorophyll content, toughness
-and specific area, wood specific gravity, bark thickness, seed mass and
-maximum adult height.
+visualized with the trajectories of community weighted means (CWM) of
+leaves economics (leaves thickness, chlorophyll content, toughness and
+specific area), wood economics (wood specific gravity, bark thickness),
+and life history traits (seed mass and maximum adult height) (Figure
+\ref{fig:CWM}).
 
 Except for leaf chlorophyll content, which continued to increase for
-some T3 and T2 plots, all traits and seed mass proportions followed a
-unimodal trajectories and either stabilized or returned towards their
-initial values 30 years after disturbance. At that time the weighted
-means of communities specific maximum height at adult stage
+some T3 and T2 plots 30 years after disturbance, all traits and seed
+mass proportions followed unimodal trajectories either stabilizing or
+returning towards their initial values. Thirty years after disturbance
+the weighted means of communities specific maximum height at adult stage
 (\emph{Hmax}), leaf toughness (\emph{L\_toughness}) and wood specific
 gravity (\emph{WD}) remained significantly lower than their initial
-value and than these of the control plots (Figure \ref{fig:CWM}). The
-weighted means of bark thickness (\emph{Bark\_thick}) similarly remained
-substantilly higher than initially for all disturbed plots while the
-specfic leaf area (\emph{SLA}) had almost recovered its initial value,
-similar as those of undisturbed plots. For all traits the maximum
-difference to initial state was correlated to the disturbance intensity
-(\(\rho_{spearman}^{L_{thickness}}=0.67\),
+value (Figure \ref{fig:CWM}). The weighted means of bark thickness
+(\emph{Bark\_thick}) similarly remained substantially higher than
+initially for all disturbed plots while the specfic leaf area
+(\emph{SLA}) had almost recovered its initial value. For all traits the
+maximum difference to initial state was correlated to the disturbance
+intensity (\(\rho_{spearman}^{L_{thickness}}=0.67\),
 \(\rho_{spearman}^{L_{chloro}}=0.45\),
@@
 
-Communities functional redundancy was measured as the sum of weighted
-overlap among species in communities' functional space. Functional
-spaces were defined in the main two-dimensional space of a PCA analysis
-summarizing the 7 leaf, stem, and maximum height traits (see appendix I
-for PCA details). Communities functional redundancy remained stable in
-control plots but after disturbance the redundancy trajectories were
-quite variable (See appendix I). Globally after most intense disturbance
-(plots T2 and T3) communities redundancy decreased at first place before
-increasing to edge, recover or exceed the initial value.
+Communities functional redundancy was measured as the sum within
+communities the species weighted functional overlap based on the 7 leaf,
+stem, and maximum height traits (see appendix I for PCA details).
+Communities functional redundancy remained stable in control plots but
+after disturbance the redundancy trajectories were quite variable (See
+appendix I) and apparently independently of the initial disturbance.
+Globally after most intense disturbance (plots T2 and T3) communities
+redundancy decreased at first place before increasing to edge, recover
+or exceed the initial value.
 
@@
 decreasing humped-shaped trajectories (@ref(fig:RedFun\_rest)). The
-maximum redundancy loss observed was positively correlated with the
-disturbance intensity (XX spearman to be measured) and none of the
-disturbed plots had recovered the functional redundancy within the
-initial functional space.
+maximum redundancy loss was positively correlated with the disturbance
+intensity (XX spearman to be measured) and the initial value had not
+recovered for any disturbed communities.
 

### 2018-05-18 Scripts/Stats.R

@@ -37,9 +37,29 @@ legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=
 load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")
 time<-c("1995","2005","2015")
 colyear<-c("darkgoldenrod1","darkorange2","darkred")
-Data<-Rao
+Data<-Sim
 AgbLoss<-AGBloss
-#plotIDH...
+
+Ylim<-c(min(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))),
+          max(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))))
+plot(AgbLoss[,"AGB"],AgbLoss[,"AGB"],type="n",xlab="",ylab="",
+       ylim=Ylim)
+leg<-lapply(1:3,function(Ti){
+    toplot<-unlist(lapply(Data, function(tr){return(tr[,time[Ti],"0.5"])}))
+    abs<-AgbLoss[which(AgbLoss[,"plot"]%in%names(toplot)),"AGB"]
+    points(abs,toplot,col=colyear[Ti],pch=20)
+    Lm<-lm(toplot~abs)
+    #summary(Lm)
+    #anova(Lm)
+    #par(mfrow=c(2,2));plot(Lm)
+    abline(a=Lm$coefficients[1],b=Lm$coefficients[2],col=colyear[Ti],lwd=2.5)
+    return(round(summary(Lm)$adj.r.squared,2))
+    #legend(round(summary(Lm)$adj.r.squared,2),x=min(AgbLoss),y=seq(Ylim[2],Ylim[1],length.out=10)[Ti],
+    #       bty="n",lty=1,col=colyear[Ti],lwd=2.5,cex=0.8)
+  })
+  legend("topleft",legend=unlist(leg),bty="n",lty=1,col=colyear,lwd=2.5,cex=0.8,title=expression(paste('R'^2,'adjusted')))
+
+
 
 ##########################################
 ## Rho Spearman, Taxo

### 2018-05-25 WholePlotTrajectories.Rmd
web pages
Graphs

@@
 Furthermore, account for species biological attributes and role in the ecosystem is essential to understand the correlations between ecosystems biodiversity, functioning and environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-In that respect functional approaches based on major traits related to species ecology and performance was largely adopted [@Diaz2005;@Villeger2008a], for example highlighting in tropical rainforests the environmental filters fostering fast growing species with efficient resources acquisition after disturbance [@Molino2001;@Haddad2008].
+In that respect functional approaches based on major traits related to species ecology and performance was largely adopted [@Diaz2005;@Villeger2008a], for example highlighting in tropical rainforests the environmental filters fostering fast growing species with efficient resources acquisition after disturbance [@Molino2001;@Haddad2008;@Ruger2009].
 Post-disturbance trajectories then corresponded to a shift from "conservative" slow-growing species dealing with scarce resources dominating before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
@@
 The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3).
-Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Blanc2009]. 
+Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Schmitt1989;@Blanc2009]. 
 The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance (ref to be found) measured with the BIOMASS R package [@Biomass2018].
@@
 The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.", fig.env="figure*"}
-load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")
-load("DB/FunctionalComposition_forGraph");load("DB/TaxoComposition_forGraphs")
+load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
+load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
 

### 2018-05-25 Scripts/DrawingTools.R

@@ -39,7 +39,8 @@ mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)
 mtext("(a) Taxonomic composition",side=3,adj=0,line=0.5)
 }
 
-FunCompo<-function(Data_FunComp){mat<-as.data.frame(Data_FunComp[[3]])
+FunCompo<-function(Data_FunComp){
+  mat<-as.data.frame(Data_FunComp[[8]])
 mat$plot<-substr(rownames(mat),start=1,stop=regexpr("_",rownames(mat))-1)
 mat$year<-substr(rownames(mat),start=regexpr("_",rownames(mat))+1,
                  stop=nchar(rownames(mat)))
@@ -61,7 +62,8 @@ invisible(lapply(1:length(treatments),function(tr){
 }))
 mtext("(b) Functional composition",side=3,adj=0,line=0.5)
 mtext("NMDS 1",side=1,line=2,cex=0.8)
-mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)}
+mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)    
+}
 
 TaxoDist<-function(Data_TaxoComp){
 MatrepT<-lapply(Data_TaxoComp,function(Rep){
@@ -249,7 +251,7 @@ RedundancyPlot<-function(Red){
   Red<-lapply(Red,function(rep){
     ret<-smooth(rep,2)
    colnames(ret)<-colnames(rep)
-   ret<-apply(ret[,which(as.numeric(colnames(ret))>="1989")],2,function(col){return(col)})#-ret[,"1989"]
+   ret<-apply(ret[,which(as.numeric(colnames(ret))>="1989")],2,function(col){return(col-ret[,"1989"])})#
    colnames(ret)<-as.numeric(colnames(ret))-1984
    return(ret)})
   

### 2018-05-25 WholePlotTrajectories.Rmd
Graphs

@@
 Library("kableExtra")
+
 source("Scripts/DrawingTools.R")
@@
 The impact of disturbance was examined specifically through the linear correlation between the intial %AGB removed and the Simpson and Rao diversities (diversities of order 2) after 10, 20 and 30 years \@ref(fig:IDHplot).
-The correlation with disturbance intensity was weak for the Simpson diversity ($R^2<0.25$) and only valid from 20 years after disturbance but it was much stronger for the Rao diversity ($0.55<R^2<0.75$) for all the time studied.
+The correlation with disturbance intensity was weak for the Simpson diversity ($R^2<0.25$) and only valid from 20 years after disturbance but it was much stronger for the Rao diversity ($0.60<R^2<0.75$) for all the time studied.
 Slope of linear correlations, reflecting the impact of disturbance, was the highest 20 years after disturbance.

### 2018-05-25 Scripts/CWM.R

@@ -298,3 +298,7 @@ invisible(lapply(colnames(CWM[[1]]),function(trait){
   
 }))
 
+load("DB/CWMnew")
+
+
+

### 2018-05-25 Scripts/Stats.R

@@ -12,6 +12,8 @@ Sim<-lapply(Sim, function(tr){
              dimnames=list(rownames(Ret[[1]]),colnames(Ret[[1]]),c(0.025,0.5,0.975)))})
 save(Sim,file="DB/SimpsonIDH")
 
+CompFun<-CompleteFun
+CompFun[[2]]<-CompFun[[2]][which(rownames(CompFun[[2]])!=7),,]
 Rao<-lapply(CompFun, function(tr){
   Ret<-lapply(c(0.025,0.5,0.975),function(quant){
     return(apply(tr[,c("1995","2005","2015"),],c(1,2),function(x){return(quantile(x,probs=quant))}))})
@@ -37,7 +39,7 @@ legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=
 load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")
 time<-c("1995","2005","2015")
 colyear<-c("darkgoldenrod1","darkorange2","darkred")
-Data<-Sim
+Data<-Rao
 AgbLoss<-AGBloss
 
 Ylim<-c(min(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))),

### 2018-05-31 WholePlotTrajectories.Rmd
intro

@@
 
-The large areas covered with tropical forests worldwide hold crucial economic, social and cultural values. 
+The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
 They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The simultaneous increase of forests products demand and the substantial climatic changes currently heighten the pressure on remaining forests [@Gibson2011a; @Morales-Hidalgo2015], threatening the maintenance of communities structure, composition and functioning and their dynamics in space and time [@Anderson-Teixeira2013;@Sist2015].
+The simultaneous increased demand in forests products and the substantial climatic changes currently heighten the pressure on remaining forests [@Gibson2011a; @Morales-Hidalgo2015], threatening the maintenance of communities structure, composition and functioning and their dynamics in space and time [@Anderson-Teixeira2013;@Sist2015].
 
-In tropical forest, ecological communities are constantly re-shaped by the natural disturbance events changing both abiotic environment, through the fluxes of light, heat and water, and biotic interactions and competitive pressure [@Goulamoussene2017].
+In tropical forest, ecological communities are constantly re-shaped by natural disturbance events that change both abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and biotic interactions like competition among species [@Herault2018].
 The cornerstone of tropical forests ecology is to understand the mechanisms and the determinants of ecosystems response to disturbance [@White2001;@Chazdon2003a]. 
 For now, this has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
-These structural parameters were thereafter consistently modeled and gave important insights into the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009]. 
+These structural parameters were thereafter sucessfully modeled, giving important insights into the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009]. 
 However the response of forests diversity in tree species remains unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014, @Liang2016] and would be most probably impacted by the changes following disturbance [@Baraloto2012a].
@@
 In the short-term disturbance demonstrated negligible or even positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
-Still, validations of the IDH remain scarce in the long term and mainly rely on species richness analyses that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-More releveant monitoring would encompass communities composition, crucial for conservation issues, and complete diversity profiles encompassing species evenness in addition to their richness, that reveals ecological rules shaping communities [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, account for species biological attributes and role in the ecosystem is essential to understand the correlations between ecosystems biodiversity, functioning and environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-In that respect functional approaches based on major traits related to species ecology and performance was largely adopted [@Diaz2005;@Villeger2008a], for example highlighting in tropical rainforests the environmental filters fostering fast growing species with efficient resources acquisition after disturbance [@Molino2001;@Haddad2008;@Ruger2009].
-Post-disturbance trajectories then corresponded to a shift from "conservative" slow-growing species dealing with scarce resources dominating before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
+Still, validations of the IDH remain scarce in the long term and mainly rely on species richness analyses that give limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+More releveant monitoring would encompass communities composition, crucial for conservation issues, and complete diversity profiles integrating both species richness and evenness, to reveal the ecological rules structuring communities [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Furthermore, the functional approach that accounts for species biological attributes and role in the ecosystem would be essential to understand the correlations between ecosystems biodiversity, functioning and environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+The functional trait-based approach focused on major traits related to species ecology and performance was sucessfully adopted [@Diaz2005;@Villeger2008a], for example in highlighting deterministic processes in tropical rainforests that foster fast growing species with efficient resources acquisition after disturbance [@Molino2001;@Haddad2008;@Ruger2009].
+Communities then shifted from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
 This shift translated into consistent trajectories for key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction and life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
 
-Functional and taxonomic trajectories are both essentiall to fully assess communities response to disturbance.
-Both relate to distinct ecological rules and processes and might then be decoupled as it was observed already in tropical forests [@Lohbeck2015; @Guariguata2001].
-Although taxonomic and functional diversity are complementary, they are combine into communities functional redundancy that measures the amount of species sharing same trait values [@Carmona2016].
-Functional redundancy then explicits the link between otherwise decoupled taxonomic and functional diversities and is central for communities description.
-Besides, functional redundancy determines communities resilience as high redundancy, like in highly diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning [@Trenbath1999; @Elmqvist2003; @Diaz2005].
+A combination of taxonomic and functional approaches are essential to fully assess communities response to disturbance and differences in their respective trajectories are insightful of underlying mechanisms [@Lohbeck2015;@Guariguata2001].
+Although taxonomic and functional diversity are complementary and sometimes decouples, they are combined in the measure of communities functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
+Functional redundancy is determinant of communities resilience as high redundancy, like in highly diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning [@Trenbath1999; @Elmqvist2003; @Diaz2005].
 
-Grasp all facets of communities response to disturbance amount to disentangle the taxonomic and functional trajectories in diversity and composition.
-These trajectories would highlight the ecological rules constraining, or not, communities dynamics towards a recovery of initial composition, diversity and functioning.
-They would therefore clarify the tenants in the long term of the debated Intermediate Disturbance Hypothesis for tropical forests, which would be as much insights for future adaptive conservation strategies [@Adler2007].
+Grasp all facets of communities response to disturbance comes to examine the taxonomic and functional trajectories in diversity and composition.
+These trajectories would highlight the ecological rules constraining or not communities dynamics towards the recovery of initial composition, diversity and functioning.
+They would therefore clarify the tenants in the long term of the debated Intermediate Disturbance Hypothesis for tropical forests, a cruicial point for future adaptive conservation strategies [@Adler2007].
 Here we monitored over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
-We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
+We made use of a large functional traits database browsing major leaf, stem and seed functional traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
 Specifically, we (i) questioned the coupling between taxonomic and functional response to disturbance and identified the underlying assembly processes, which allowed to (ii) clarify the validity of the IDH in the long term for tropical forest and (iii) question the completeness of communities recovery regarding their functional redundancy.

### 2018-05-31 WholePlotTrajectories.Rmd
Discussion

@@
 Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
-The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40 cm DBH for treatment 3 (T3).
+The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3).
 Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Schmitt1989;@Blanc2009]. 
@@
 The taxonomic diversity was assessed through Richness and the Hill number translation of Shannon and Simpson indices [@Hill1973].
-These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q), which proved well suited for diversity studies [@Patil1982;@Tothmeresz1995].
+These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995].
 The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on all functional and life traits.
@@
 
-## Decoupled recovery of communities taxonomic and functional characteristics
+## Decoupled taxonomic and functional trajectories
 
-Both communities taxonomic and functional diversity and composition proved resilient, following similar humped-back trajectories with a return towards initial values.
-The resilience of communities functional characteristics, the most direct link between biodiversity and ecosystem functioning [@Diaz2005], meant a consistent recovery of ecosystem processes in the long term [@Guariguata2001].
-The resilience of communities taxonomic characteristics, in turn, meant the maintenance of their initial differences in composition and structure.
-It suggested that communities response to disturbance were somehow constraint to converge towards determined compositions [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
+Both communities taxonomic and functional diversity and composition proved resilient, following similar humped-back trajectories returning towards initial values.
+The resilience of communities functional structure, that is the most direct link between biodiversity and ecosystem functioning [@Diaz2005], meant a consistent recovery of ecosystem processes in the long term [@Guariguata2001].
+The resilience of communities taxonomy meant the maintenance of their initial differences in composition and diversity, suggesting the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003].
+It besides suggested that recovery trajectories depend on the initial composition eventually restored that would the the asymptot of the humped-shaped dynamics [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
 
-Although both communities taxonomic and functional characteristics proved resilient and followed similar humped-back trajectories, the taxonomic recovery systematically lagged behind the corresponding functional dynamics.
-Such delay between functional and taxonomic dynamics has already been observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015; @Guariguata2001].
-According to the "vegetation quantity effect" [@Grime1998] functional trajectories rely on the pool of dominant species, which diversity and evenness were enhanced after disturbance, and which rapidly restored the functional diversity and composition.
-However, communities evenness remained high so infrequent species still missed to the taxonomic recovery that mechanically lagged behind, all the more so that unrecovered species would be functionally redundant and probably undergo competition processes.
+Although both communities taxonomic and functional characteristics proved resilient and followed similar humped-back trajectories, the taxonomic recovery systematically lagged behind the functional.
+The decoupling between functional and taxonomic dynamics has already been observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015; @Guariguata2001].
+According to the "vegetation quantity effect" [@Grime1998] functional trajectories rely on the pool of dominant species, which diversity and evenness were enhanced after disturbance, and which rapidly recovered their initial functional structure.
+However, communities evenness remained high so infrequent species still missed to the taxonomic recovery that proved all the more longer to recover that unrecovered species would be functionally redundant and probably underwent competition processes.
 
@@
 Validating the IDH, communities trajectories confirmed the diversity increase after disturbance through an enhanced growth of otherwise less favored species.
-This was however completely clear regarding the functional diversity but much more blurry for communities taxonomic characteristics.
+This was strong regarding the functional diversity but more blurry regarding their taxonomic structure.
 
 The taxonomic richness was weakly or negatively impacted by intense disturbance, as observed on several post logging surveys [@Cannon1998; @Baraloto2012a], while it substantially increased after low intensity disturbance.
-Disturbance also enhanced taxonomic evenness, but its intensity was only weakly and lately correlated to communities eveness (*i.e.* Simpson diversity).
+Disturbance also enhanced taxonomic evenness, but the disturbance intensity was only weakly and lately correlated to communities eveness (*i.e.* Simpson diversity).
 Contrastingly, disturbance intensity consistently predicted the significant increase of communities functional diversity for 30 years.
-Disturbance impacts involve specific turnover within communities, either among pre-disturbance survivors or among newly recruited trees.
-As the composition of old-growth survivors proved to mirror initial communities [@Herault2018], disturbance would impact trees recruitment through the enhanced growth and survival of previously infrequent species and functional characteristics.
-Consistently, disturbance resulted in increasing taxonomic dissimilarity compared to pre-disturbance communities and functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
+Disturbance entailed specific turnover within communities, either among pre-disturbance survivors or among newly recruited trees.
+As the composition of old-growth survivors proved to mirror initial communities [@Herault2018], disturbance likely impacted trees recruitment through the enhanced growth and survival of previously infrequent species and functional types.
+Consistently, disturbance resulted in increasing taxonomic dissimilarity compared to pre-disturbance communities and significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
 Disturbance then caused a reorganization of the typical high dominance structure of hyperdiverse mature forests after disturbance, benefiting to pioneers and light demanding species. 
-Likely, the changes in abiotic environment and competitive pressure favored pioneers which outcompete others in non limiting resources but are excluded in mature forests by long-lived, more resistant and shade tolerant species.
-Therefore, consistently with communities dynamics after disturbance relied on species functional strategy and corresponding ability to fill the environmental niches made available by disturbance.
-Recruited species then mixed with pre-disturbance ones, from which they differed, and constituted a community all the more diversified that the disturbance was intense [@Molino2001].
+Likely, the changes in abiotic environment and competitive pressure favored pioneers which outcompete other species in non limiting resources but are excluded in mature forests through the cmpetition with long-lived, more resistant and shade bearers.
+Therefore, consistently with the IDH the trajectories after disturbance relied on the environmental niches made available by disturbance filled by species from a restrictied functional range.
+Recruited species therefore differed from pre-disturbance ones, constituting a community all the more diversified that the disturbance was intense [@Molino2001].
 
-## The functional redundancu and the resilience of communities
+## The functional redundancy, key of communities resilience
 
-Both the lag between taxonomic and functional recovery and the middling consistency of the IDH regarding communities taxonomic diversity entail a central role of functional redundancy.
+Both the lag between taxonomic and functional recovery and the middling consistency of the IDH regarding communities taxonomic diversity made central the role of functional redundancy.
 Functional redundancy, determinant of communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005], seemed not to have recovered 30 years after disturbance.
 
-Although disturbance would be mitigated by communities functional redundancy, we expect the redundancy itself to be reduced or re-organized within the functional space.
-Globally, disturbed communities did not follow consistent trajectories after disturbance and 30 years after disturbance, irrespective of the intensity, disturbed communities displayed lower, similar or higher functional redundancy.
-The redundancy within the functional space of the initial community however clearly followed humped-back trajectories which decrease was determined by the disturbance intensity and which return to initial state was not observed for any disturbed community.
-Not only did the functional redundancy and hence the resilience of pre-disturbance community decrease, but also did the ones of disturbance-specific communities increase.
-It means higher chances to have long lasting or self-maintained compositional changes towards disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013] and thus highly question forest's resilience [@Chazdon2003a].
-Specificaly, this would impair species contingent to undisturbed forests, threatening their maintenance, and run the risk to loose cornerstone species and trigger unexpected ecological consequences [@Jones1994;@Diaz2005;@Gardner2007].
+Communities with high functional redundancy are expected to be able to lose species with a limited impact on their functional structure [@Carmona2016].
+Functional redundancy therefore mitigates the impacts of disturbance, but we expect the redundancy itself to be reduced or re-organized within the functional space.
+Although the overall functional redundancy did not follow consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped-back trajectories.
+This restricted redundancy decreased according to the disturbance intensity before returning towards the initial value but had not recovered for any disturbed community after 30 years.
+The recovery of the redundancy within the initial functional space explained the hampered recovery of infrequent species specific to old-growth forests that would be functionally redundant and therefore underwent competitive pressures limiting species functional similarity [@Mayfield2010].
+The unachieved humped-shape redundancy trajectory meant a decrease of the resilience of pre-disturbance communities and an increase in the resilience of disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species).
+After disturbance the chances were higher to have long lasting or self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013], which highly questioned the completeness of the resilience [@Chazdon2003a].
+Specificaly, species contingent to undisturbed forests might be impaired, threatening their maintenance, the risk to loose cornerstone species and trigger unexpected ecological consequences would be higher [@Jones1994;@Diaz2005;@Gardner2007].
 
@@
 
-our study disentangled tropical forests functional and taxonomic decoupled response to disturbance, with a rapid recovery of communities functioning but slower and more variable taxonomic dynamics.
-Consistently with the IDH functional trajectories were constrained by environmental pressures favoring pioneers and light-demanding species after disturbance, but taxonomic dynamics were more stochastic as they involved infrequent speciesfunctoinally redundant with the dominants and thus under competitive pressure.
-While communities functioning rapidly recovered, taxonomic trajectories, although consistent and converging towards full recovery, were much slower and entailed persisting alteration of communities functional redundancy.
-the trajectories therefore suggested a potential sustainability of tropical forests in the face of quite intense disturbance but only when followed by long recovery periods, longer than 30 years [@Gourlet-Fleury2005].
+our study defined the decoupled functional and taxonomic trajectories of tropical rainforests after disturbance.
+These trajectories shed light on their rapid functional recovery but slower and more variable taxonomic dynamics.
+Consistently with the IDH, functional trajectories were constrained by determined processes favoring pioneers and light-demanding species after disturbance.
+The taxonomic dynamics however proved more stochastic due to the recovery of infrequent, functionally redundant species hampered by limiting similarity processes.
+Although communities functioning rapidly recovered, their taxonomic structured and functional redundancy remained altered for several decades.
+The trajectories therefore suggested a potential sustainability of tropical forests in the face of quite intense disturbance but only when followed by long recovery periods, longer than 30 years [@Gourlet-Fleury2005].
 This study besides highlighted the central role of recruitment processes underlying the IDH mechanisms and holding the whole community response to disturbance.

### 2018-06-01 WholePlotTrajectories.Rmd
Abstract tentative

@@
     url: http://www.ecofog.gf/spip.php?article47
-abstract: >   #Pour avoir plusieurs lignes qui suivent. Commencer à 2 caractères
-  Résumé de l'article.
-keywords: [mot-clés, séparés par des virgules]
+abstract: >   
+    Currently the increasing threats on tropical forests and the global changing climate challenge the maintenance of forests biodiversity structure, functioning and dynamics.
+    Communities diversity structure is assumed to rely on a constant regime of disturbance and to peak at intermediate level of disturbance. 
+    In tropical forests however both this intermediate disturbance hypothesis (IDH) and the underlying processes remain debated.
+    The ecological processes driving trajectories after disturbance leave a signature on communities taxonomic or functional structure which differences give major insights to validate the IDH and identify the ecological processes involved.
+    The completeness of forests resilience regarding all taxonomic and functional facets besdides remain to be clarified.
+
+  Here we examined in a neotropical forest the trajectories over 30 years after a gradient of logging and thinning disturbance, focusing on communities taxonomic and functional composition and diversity.
+  Specifically we analysed the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 key functional traits assessing species ecology.
+
+  We highlighted a decoupling between functional and taxonomic trajectories, translating a fast recovery of communities functioning while their taxonomic recovery was slowed by the recruitment of infrequent species specific to mature forests.
+  The IDH consistently predicted communities functional trajectories but poorly reflected the taxonomic ones, blurried by changes in the functional redundancy entailing competitive exclusion.
+  Although communities consistently recovered their functioning after 30 years, initially infrequent species proved long to recover which impacted the taxonomic structure and functoinal redundancy over several decades.
+  Our results acknowledged the need of decades-long recovery cycles to restore the resilience of the community and prevent species loss in the long term.
+
+keywords: [Taxonomic and Functional Biodiversity, Neotropical Forests, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience]
 journalinfo: Hal xxx
@@
 
-Although both communities taxonomic and functional characteristics proved resilient and followed similar humped-back trajectories, the taxonomic recovery systematically lagged behind the functional.
+Although both communities taxonomic and functional characteristics proved resilient and followed similar humped-back trajectories, the taxonomic recovery systematically lagged behind the functional one.
 The decoupling between functional and taxonomic dynamics has already been observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015; @Guariguata2001].
@@
 The taxonomic dynamics however proved more stochastic due to the recovery of infrequent, functionally redundant species hampered by limiting similarity processes.
-Although communities functioning rapidly recovered, their taxonomic structured and functional redundancy remained altered for several decades.
-The trajectories therefore suggested a potential sustainability of tropical forests in the face of quite intense disturbance but only when followed by long recovery periods, longer than 30 years [@Gourlet-Fleury2005].
-This study besides highlighted the central role of recruitment processes underlying the IDH mechanisms and holding the whole community response to disturbance.
+Although communities functioning rapidly recovered, their taxonomic structure and functional redundancy remained altered for several decades.
+The trajectories therefore demonstrated a decades-long resilience of tropical forests in the face of quite intense disturbance [@Gourlet-Fleury2005].
+Within the mechanisms of this resilience the recruitment processes proved central, in accordance with the IDH, and would deserved closer focus to clarify their determinisms and ecological rules.
+The disturbance range however stayed within the spectrum of selective logging, with a forest cover remaining all along the experiment, and the response mechanisms would probably be much different after harder disturbance.
 

### 2018-06-04 WholePlotTrajectories.Rmd
abstract

@@
 abstract: >   
-    Currently the increasing threats on tropical forests and the global changing climate challenge the maintenance of forests biodiversity structure, functioning and dynamics.
-    Communities diversity structure is assumed to rely on a constant regime of disturbance and to peak at intermediate level of disturbance. 
-    In tropical forests however both this intermediate disturbance hypothesis (IDH) and the underlying processes remain debated.
-    The ecological processes driving trajectories after disturbance leave a signature on communities taxonomic or functional structure which differences give major insights to validate the IDH and identify the ecological processes involved.
-    The completeness of forests resilience regarding all taxonomic and functional facets besdides remain to be clarified.
+    Understand the maintenance of forests biodiversity structure, functioning and dynamics is urgent to anticipate their fate in the global changing context.
+    Communities diversity structure is assumed to rely on a constant regime of disturbance and peak at intermediate intensity.
+    For tropical forests this intermediate disturbance hypothesis (IDH) remain debated as well as the completeness of forests resilience in all taxonomic and functoinal facets.
 
-  Here we examined in a neotropical forest the trajectories over 30 years after a gradient of logging and thinning disturbance, focusing on communities taxonomic and functional composition and diversity.
+  To disentangle the ecological processes driving forests dynamics we examined the the trajectories over 30 years after a gradient of logging and thinning disturbance, focusing on communities taxonomic and functional composition and diversity.
   Specifically we analysed the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 key functional traits assessing species ecology.
@@
 
-Considering the functional redundancy restricted to the functional space of the initial inventory, all disturbed plots followed similar decreasing humped-shaped trajectories (\@ref(fig:RedFun_rest)).
+Considering the functional redundancy restricted to the functional space of the initial inventory, all disturbed plots followed similar decreasing humped shaped trajectories (\@ref(fig:RedFun_rest)).
 The maximum redundancy loss was positively correlated with the disturbance intensity (XX spearman to be measured) and the initial value had not recovered for any disturbed communities.
@@
 
-Both communities taxonomic and functional diversity and composition proved resilient, following similar humped-back trajectories returning towards initial values.
+Both communities taxonomic and functional diversity and composition proved resilient, following similar humped shaped trajectories returning towards initial values.
 The resilience of communities functional structure, that is the most direct link between biodiversity and ecosystem functioning [@Diaz2005], meant a consistent recovery of ecosystem processes in the long term [@Guariguata2001].
 The resilience of communities taxonomy meant the maintenance of their initial differences in composition and diversity, suggesting the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003].
-It besides suggested that recovery trajectories depend on the initial composition eventually restored that would the the asymptot of the humped-shaped dynamics [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
+It besides suggested that recovery trajectories depend on the initial composition eventually restored that would the the asymptot of the humped shaped dynamics [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
 
-Although both communities taxonomic and functional characteristics proved resilient and followed similar humped-back trajectories, the taxonomic recovery systematically lagged behind the functional one.
+Although both communities taxonomic and functional characteristics proved resilient and followed similar humped shaped trajectories, the taxonomic recovery systematically lagged behind the functional one.
 The decoupling between functional and taxonomic dynamics has already been observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015; @Guariguata2001].
@@
 Functional redundancy therefore mitigates the impacts of disturbance, but we expect the redundancy itself to be reduced or re-organized within the functional space.
-Although the overall functional redundancy did not follow consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped-back trajectories.
+Although the overall functional redundancy did not follow consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
 This restricted redundancy decreased according to the disturbance intensity before returning towards the initial value but had not recovered for any disturbed community after 30 years.
 The recovery of the redundancy within the initial functional space explained the hampered recovery of infrequent species specific to old-growth forests that would be functionally redundant and therefore underwent competitive pressures limiting species functional similarity [@Mayfield2010].
-The unachieved humped-shape redundancy trajectory meant a decrease of the resilience of pre-disturbance communities and an increase in the resilience of disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species).
+The unachieved humped shaped redundancy trajectory meant a decrease of the resilience of pre-disturbance communities and an increase in the resilience of disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species).
 After disturbance the chances were higher to have long lasting or self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013], which highly questioned the completeness of the resilience [@Chazdon2003a].

### 2018-06-04 WholePlotTrajectories.Rmd
pdf

@@
     Communities diversity structure is assumed to rely on a constant regime of disturbance and peak at intermediate intensity.
-    For tropical forests this intermediate disturbance hypothesis (IDH) remain debated as well as the completeness of forests resilience in all taxonomic and functoinal facets.
+    For tropical forests this intermediate disturbance hypothesis (IDH) remain debated as well as the completeness of forests resilience in all taxonomic and functional facets.
 

### 2018-06-04 WholePlotTrajectories.tex
Graphs
pdf

@@
 
-\Keywords{mot-clés, séparés par des virgules} % Keywords - if you don't want any simply remove all the text between the curly brackets
+\Keywords{Taxonomic and Functional Biodiversity, Neotropical Forests, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience} % Keywords - if you don't want any simply remove all the text between the curly brackets
 \newcommand{\keywordname}{Keywords} % Defines the keywords heading name
@@
 \Abstract{
-Résumé de l'article.
+Understand the maintenance of forests biodiversity structure,
+functioning and dynamics is urgent to anticipate their fate in the
+global changing context. Communities diversity structure is assumed to
+rely on a constant regime of disturbance and peak at intermediate
+intensity. For tropical forests this intermediate disturbance hypothesis
+(IDH) remain debated as well as the completeness of forests resilience
+in all taxonomic and functional facets.
 }
@@
 The large areas covered with tropical forests worldwide hold crucial
-economic, social and cultural values. They provide wood and multiple
-non-timber forest products, shelter a diversified fauna, regulate the
-local climate, support the carbon, water and nutrient cycles, and ensure
-cultural and human well-being. The simultaneous increase of forests
-products demand and the substantial climatic changes currently heighten
-the pressure on remaining forests
+environmental, economic and social values. They provide wood and
+multiple non-timber forest products, shelter a diversified fauna,
+regulate the local climate, support the carbon, water and nutrient
+cycles, and ensure cultural and human well-being. The simultaneous
+increased demand in forests products and the substantial climatic
+changes currently heighten the pressure on remaining forests
 \citep{Gibson2011a, Morales-Hidalgo2015}, threatening the maintenance of
@@
 In tropical forest, ecological communities are constantly re-shaped by
-the natural disturbance events changing both abiotic environment,
-through the fluxes of light, heat and water, and biotic interactions and
-competitive pressure \citep{Goulamoussene2017}. The cornerstone of
-tropical forests ecology is to understand the mechanisms and the
-determinants of ecosystems response to disturbance
+natural disturbance events that change both abiotic environment, through
+the fluxes of light, heat and water \citep{Goulamoussene2017}, and
+biotic interactions like competition among species \citep{Herault2018}.
+The cornerstone of tropical forests ecology is to understand the
+mechanisms and the determinants of ecosystems response to disturbance
 \citep{White2001, Chazdon2003a}. For now, this has been largely studied
@@
 \citep{Piponiot2016, Rutishauser2016}. These structural parameters were
-thereafter consistently modeled and gave important insights into the
+thereafter sucessfully modeled, giving important insights into the
 maintenance of ecosystems processes and services
@@
 IDH remain scarce in the long term and mainly rely on species richness
-analyses that gives limited or misleading information on forests
-recovery and functioning \citep{Martin2015, Chaudhary2016}. More
-releveant monitoring would encompass communities composition, crucial
-for conservation issues, and complete diversity profiles encompassing
-species evenness in addition to their richness, that reveals ecological
-rules shaping communities
-\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, account
-for species biological attributes and role in the ecosystem is essential
-to understand the correlations between ecosystems biodiversity,
-functioning and environmental constraints
-\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
-respect functional approaches based on major traits related to species
-ecology and performance was largely adopted
-\citep{Diaz2005, Villeger2008a}, for example highlighting in tropical
-rainforests the environmental filters fostering fast growing species
-with efficient resources acquisition after disturbance
-\citep{Molino2001, Haddad2008}. Post-disturbance trajectories then
-corresponded to a shift from ``conservative'' slow-growing species
-dealing with scarce resources dominating before disturbance, to
-``acquisitive'' fast-growing species with rapid and efficient use of
-abundant resources \citep{TerSteege2001, Reich2014, Herault2011}. This
-shift translated into consistent trajectories for key functional traits
-related to resource acquisition (leaf area, density and chlorophyll
-content, and stem specific gravity and bark thickness), tree growth and
-reproduction and life history traits (seed mass and maximum height)
+analyses that give limited or misleading information on forests recovery
+and functioning \citep{Martin2015, Chaudhary2016}. More releveant
+monitoring would encompass communities composition, crucial for
+conservation issues, and complete diversity profiles integrating both
+species richness and evenness, to reveal the ecological rules
+structuring communities \citep{Magurran1988, Lavorel2002, Bellwood2006}.
+Furthermore, the functional approach that accounts for species
+biological attributes and role in the ecosystem would be essential to
+understand the correlations between ecosystems biodiversity, functioning
+and environmental constraints
+\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. The
+functional trait-based approach focused on major traits related to
+species ecology and performance was sucessfully adopted
+\citep{Diaz2005, Villeger2008a}, for example in highlighting
+deterministic processes in tropical rainforests that foster fast growing
+species with efficient resources acquisition after disturbance
+\citep{Molino2001, Haddad2008, Ruger2009}. Communities then shifted from
+a dominance of ``conservative'' slow-growing species dealing with scarce
+resources to ``acquisitive'' fast-growing species with rapid and
+efficient use of abundant resources
+\citep{TerSteege2001, Reich2014, Herault2011}. This shift translated
+into consistent trajectories for key functional traits related to
+resource acquisition (leaf area, density and chlorophyll content, and
+stem specific gravity and bark thickness), tree growth and reproduction
+and life history traits (seed mass and maximum height)
 \citep{Wright2004, Westoby2006a, Chave2009b}.
 
-Functional and taxonomic trajectories are both essentiall to fully
-assess communities response to disturbance. Both relate to distinct
-ecological rules and processes and might then be decoupled as it was
-observed already in tropical forests
+A combination of taxonomic and functional approaches are essential to
+fully assess communities response to disturbance and differences in
+their respective trajectories are insightful of underlying mechanisms
 \citep{Lohbeck2015, Guariguata2001}. Although taxonomic and functional
-diversity are complementary, they are combine into communities
-functional redundancy that measures the amount of species sharing same
-trait values \citep{Carmona2016}. Functional redundancy then explicits
-the link between otherwise decoupled taxonomic and functional
-diversities and is central for communities description. Besides,
-functional redundancy determines communities resilience as high
+diversity are complementary and sometimes decouples, they are combined
+in the measure of communities functional redundancy that quantifies the
+amount of shared trait values among species \citep{Carmona2016}.
+Functional redundancy is determinant of communities resilience as high
 redundancy, like in highly diverse tropical forests
@@
 
-Grasp all facets of communities response to disturbance amount to
-disentangle the taxonomic and functional trajectories in diversity and
-composition. These trajectories would highlight the ecological rules
-constraining, or not, communities dynamics towards a recovery of initial
-composition, diversity and functioning. They would therefore clarify the
-tenants in the long term of the debated Intermediate Disturbance
-Hypothesis for tropical forests, which would be as much insights for
-future adaptive conservation strategies \citep{Adler2007}. Here we
-monitored over 30 years the response of 75 ha of forests plots set up on
-a gradient of disturbance intensity, from 10 to 60\% of ecosystem
-biomass removed. We made use of a large functional traits database
-browsing major leaf, stem and seed traits and species maximum height to
-draw the trajectories over time of communities taxonomic and functional
-composition and diversity. Specifically, we (i) questioned the coupling
-between taxonomic and functional response to disturbance and identified
-the underlying assembly processes, which allowed to (ii) clarify the
+Grasp all facets of communities response to disturbance comes to examine
+the taxonomic and functional trajectories in diversity and composition.
+These trajectories would highlight the ecological rules constraining or
+not communities dynamics towards the recovery of initial composition,
+diversity and functioning. They would therefore clarify the tenants in
+the long term of the debated Intermediate Disturbance Hypothesis for
+tropical forests, a cruicial point for future adaptive conservation
+strategies \citep{Adler2007}. Here we monitored over 30 years the
+response of 75 ha of forests plots set up on a gradient of disturbance
+intensity, from 10 to 60\% of ecosystem biomass removed. We made use of
+a large functional traits database browsing major leaf, stem and seed
+functional traits and species maximum height to draw the trajectories
+over time of communities taxonomic and functional composition and
+diversity. Specifically, we (i) questioned the coupling between
+taxonomic and functional response to disturbance and identified the
+underlying assembly processes, which allowed to (ii) clarify the
 validity of the IDH in the long term for tropical forest and (iii)
@@
 with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1),
-32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees above 40
-cm DBH for treatment 3 (T3). Treatments T2 and T3 besides included the
-thinning of trees by poison girdling \citep{Blanc2009}. The disturbance
-intensity was measured as the percentage of aboveground biomass (\%AGB)
-lost between the first inventory in 1984 and five years after
-disturbance (ref to be found) measured with the BIOMASS R package
+32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above
+40 cm DBH for treatment 3 (T3). Treatments T2 and T3 besides included
+the thinning of trees by poison girdling \citep{Schmitt1989, Blanc2009}.
+The disturbance intensity was measured as the percentage of aboveground
+biomass (\%AGB) lost between the first inventory in 1984 and five years
+after disturbance (ref to be found) measured with the BIOMASS R package
 \citep{Biomass2018}.
@@
 respectively corresponding to the 0, 1 and 2 order of diversity (q),
-which proved well suited for diversity studies
-\citep{Patil1982, Tothmeresz1995}. The functional diversity was reported
-using the Rao index of quadratic entropy which combines species
-abundance distribution and average pairwise dissimilarity based on all
-functional and life traits.
+recomended for diversity studies \citep{Patil1982, Tothmeresz1995}. The
+functional diversity was reported using the Rao index of quadratic
+entropy which combines species abundance distribution and average
+pairwise dissimilarity based on all functional and life traits.
 
@@
 after disturbance but it was much stronger for the Rao diversity
-(\(0.55<R^2<0.75\)) for all the time studied. Slope of linear
+(\(0.60<R^2<0.75\)) for all the time studied. Slope of linear
 correlations, reflecting the impact of disturbance, was the highest 20
@@
 of the initial inventory, all disturbed plots followed similar
-decreasing humped-shaped trajectories (@ref(fig:RedFun\_rest)). The
+decreasing humped shaped trajectories (@ref(fig:RedFun\_rest)). The
 maximum redundancy loss was positively correlated with the disturbance
@@
 
-\subsection{Decoupled recovery of communities taxonomic and functional
-characteristics}\label{decoupled-recovery-of-communities-taxonomic-and-functional-characteristics}
+\subsection{Decoupled taxonomic and functional
+trajectories}\label{decoupled-taxonomic-and-functional-trajectories}
 
 Both communities taxonomic and functional diversity and composition
-proved resilient, following similar humped-back trajectories with a
-return towards initial values. The resilience of communities functional
-characteristics, the most direct link between biodiversity and ecosystem
-functioning \citep{Diaz2005}, meant a consistent recovery of ecosystem
-processes in the long term \citep{Guariguata2001}. The resilience of
-communities taxonomic characteristics, in turn, meant the maintenance of
-their initial differences in composition and structure. It suggested
-that communities response to disturbance were somehow constraint to
-converge towards determined compositions
+proved resilient, following similar humped shaped trajectories returning
+towards initial values. The resilience of communities functional
+structure, that is the most direct link between biodiversity and
+ecosystem functioning \citep{Diaz2005}, meant a consistent recovery of
+ecosystem processes in the long term \citep{Guariguata2001}. The
+resilience of communities taxonomy meant the maintenance of their
+initial differences in composition and diversity, suggesting the
+existence of multiple stable equilibria as assumed for highly diverse
+and productive ecosystems \citep{Chase2003}. It besides suggested that
+recovery trajectories depend on the initial composition eventually
+restored that would the the asymptot of the humped shaped dynamics
 \citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.
@@
 Although both communities taxonomic and functional characteristics
-proved resilient and followed similar humped-back trajectories, the
-taxonomic recovery systematically lagged behind the corresponding
-functional dynamics. Such delay between functional and taxonomic
-dynamics has already been observed for grasslands
-\citep{Tilman1997, Mouillot2011} and more recently for tropical forests
-\citep{Lohbeck2015, Guariguata2001}. According to the ``vegetation
-quantity effect'' \citep{Grime1998} functional trajectories rely on the
-pool of dominant species, which diversity and evenness were enhanced
-after disturbance, and which rapidly restored the functional diversity
-and composition. However, communities evenness remained high so
-infrequent species still missed to the taxonomic recovery that
-mechanically lagged behind, all the more so that unrecovered species
-would be functionally redundant and probably undergo competition
-processes.
+proved resilient and followed similar humped shaped trajectories, the
+taxonomic recovery systematically lagged behind the functional one. The
+decoupling between functional and taxonomic dynamics has already been
+observed for grasslands \citep{Tilman1997, Mouillot2011} and more
+recently for tropical forests \citep{Lohbeck2015, Guariguata2001}.
+According to the ``vegetation quantity effect'' \citep{Grime1998}
+functional trajectories rely on the pool of dominant species, which
+diversity and evenness were enhanced after disturbance, and which
+rapidly recovered their initial functional structure. However,
+communities evenness remained high so infrequent species still missed to
+the taxonomic recovery that proved all the more longer to recover that
+unrecovered species would be functionally redundant and probably
+underwent competition processes.
 
@@
 increase after disturbance through an enhanced growth of otherwise less
-favored species. This was however completely clear regarding the
-functional diversity but much more blurry for communities taxonomic
-characteristics.
+favored species. This was strong regarding the functional diversity but
+more blurry regarding their taxonomic structure.
 
@@
 after low intensity disturbance. Disturbance also enhanced taxonomic
-evenness, but its intensity was only weakly and lately correlated to
-communities eveness (\emph{i.e.} Simpson diversity). Contrastingly,
-disturbance intensity consistently predicted the significant increase of
-communities functional diversity for 30 years. Disturbance impacts
-involve specific turnover within communities, either among
+evenness, but the disturbance intensity was only weakly and lately
+correlated to communities eveness (\emph{i.e.} Simpson diversity).
+Contrastingly, disturbance intensity consistently predicted the
+significant increase of communities functional diversity for 30 years.
+Disturbance entailed specific turnover within communities, either among
 pre-disturbance survivors or among newly recruited trees. As the
 composition of old-growth survivors proved to mirror initial communities
-\citep{Herault2018}, disturbance would impact trees recruitment through
-the enhanced growth and survival of previously infrequent species and
-functional characteristics. Consistently, disturbance resulted in
+\citep{Herault2018}, disturbance likely impacted trees recruitment
+through the enhanced growth and survival of previously infrequent
+species and functional types. Consistently, disturbance resulted in
 increasing taxonomic dissimilarity compared to pre-disturbance
-communities and functional shifts towards resource-acquisitive
-strategies (sharp increase in the SLA, leaf thickness and bark thickness
-and decrease in wood density, leaf toughness and maximum height)
+communities and significant functional shifts towards
+resource-acquisitive strategies (sharp increase in the SLA, leaf
+thickness and bark thickness and decrease in wood density, leaf
+toughness and maximum height)
 \citep{Westoby1998, Wright2004, Reich2014}. Disturbance then caused a
@@
 demanding species. Likely, the changes in abiotic environment and
-competitive pressure favored pioneers which outcompete others in non
-limiting resources but are excluded in mature forests by long-lived,
-more resistant and shade tolerant species. Therefore, consistently with
-communities dynamics after disturbance relied on species functional
-strategy and corresponding ability to fill the environmental niches made
-available by disturbance. Recruited species then mixed with
-pre-disturbance ones, from which they differed, and constituted a
-community all the more diversified that the disturbance was intense
-\citep{Molino2001}.
+competitive pressure favored pioneers which outcompete other species in
+non limiting resources but are excluded in mature forests through the
+cmpetition with long-lived, more resistant and shade bearers. Therefore,
+consistently with the IDH the trajectories after disturbance relied on
+the environmental niches made available by disturbance filled by species
+from a restrictied functional range. Recruited species therefore
+differed from pre-disturbance ones, constituting a community all the
+more diversified that the disturbance was intense \citep{Molino2001}.
 
-\subsection{The functional redundancu and the resilience of
-communities}\label{the-functional-redundancu-and-the-resilience-of-communities}
+\subsection{The functional redundancy, key of communities
+resilience}\label{the-functional-redundancy-key-of-communities-resilience}
 
 Both the lag between taxonomic and functional recovery and the middling
-consistency of the IDH regarding communities taxonomic diversity entail
-a central role of functional redundancy. Functional redundancy,
+consistency of the IDH regarding communities taxonomic diversity made
+central the role of functional redundancy. Functional redundancy,
 determinant of communities resilience
@@
 
-Although disturbance would be mitigated by communities functional
-redundancy, we expect the redundancy itself to be reduced or
-re-organized within the functional space. Globally, disturbed
-communities did not follow consistent trajectories after disturbance and
-30 years after disturbance, irrespective of the intensity, disturbed
-communities displayed lower, similar or higher functional redundancy.
-The redundancy within the functional space of the initial community
-however clearly followed humped-back trajectories which decrease was
-determined by the disturbance intensity and which return to initial
-state was not observed for any disturbed community. Not only did the
-functional redundancy and hence the resilience of pre-disturbance
-community decrease, but also did the ones of disturbance-specific
-communities increase. It means higher chances to have long lasting or
-self-maintained compositional changes towards disturbance resistant
-species, lianas or epiphytes \citep{Haddad2008, Burslem2000, Martin2013}
-and thus highly question forest's resilience \citep{Chazdon2003a}.
-Specificaly, this would impair species contingent to undisturbed
-forests, threatening their maintenance, and run the risk to loose
-cornerstone species and trigger unexpected ecological consequences
+Communities with high functional redundancy are expected to be able to
+lose species with a limited impact on their functional structure
+\citep{Carmona2016}. Functional redundancy therefore mitigates the
+impacts of disturbance, but we expect the redundancy itself to be
+reduced or re-organized within the functional space. Although the
+overall functional redundancy did not follow consistent trajectories
+after disturbance, the redundancy restricted to the functional space of
+the initial community clearly followed humped shaped trajectories. This
+restricted redundancy decreased according to the disturbance intensity
+before returning towards the initial value but had not recovered for any
+disturbed community after 30 years. The recovery of the redundancy
+within the initial functional space explained the hampered recovery of
+infrequent species specific to old-growth forests that would be
+functionally redundant and therefore underwent competitive pressures
+limiting species functional similarity \citep{Mayfield2010}. The
+unachieved humped shaped redundancy trajectory meant a decrease of the
+resilience of pre-disturbance communities and an increase in the
+resilience of disturbance-specific communities (corresponding to lower
+richness and higher dominance of pioneer species). After disturbance the
+chances were higher to have long lasting or self-maintained
+compositional changes in favor of disturbance resistant species, lianas
+or epiphytes \citep{Haddad2008, Burslem2000, Martin2013}, which highly
+questioned the completeness of the resilience \citep{Chazdon2003a}.
+Specificaly, species contingent to undisturbed forests might be
+impaired, threatening their maintenance, the risk to loose cornerstone
+species and trigger unexpected ecological consequences would be higher
 \citep{Jones1994, Diaz2005, Gardner2007}.
@@
 
-our study disentangled tropical forests functional and taxonomic
-decoupled response to disturbance, with a rapid recovery of communities
-functioning but slower and more variable taxonomic dynamics.
-Consistently with the IDH functional trajectories were constrained by
-environmental pressures favoring pioneers and light-demanding species
-after disturbance, but taxonomic dynamics were more stochastic as they
-involved infrequent speciesfunctoinally redundant with the dominants and
-thus under competitive pressure. While communities functioning rapidly
-recovered, taxonomic trajectories, although consistent and converging
-towards full recovery, were much slower and entailed persisting
-alteration of communities functional redundancy. the trajectories
-therefore suggested a potential sustainability of tropical forests in
-the face of quite intense disturbance but only when followed by long
-recovery periods, longer than 30 years \citep{Gourlet-Fleury2005}. This
-study besides highlighted the central role of recruitment processes
-underlying the IDH mechanisms and holding the whole community response
-to disturbance.
+our study defined the decoupled functional and taxonomic trajectories of
+tropical rainforests after disturbance. These trajectories shed light on
+their rapid functional recovery but slower and more variable taxonomic
+dynamics. Consistently with the IDH, functional trajectories were
+constrained by determined processes favoring pioneers and
+light-demanding species after disturbance. The taxonomic dynamics
+however proved more stochastic due to the recovery of infrequent,
+functionally redundant species hampered by limiting similarity
+processes. Although communities functioning rapidly recovered, their
+taxonomic structure and functional redundancy remained altered for
+several decades. The trajectories therefore demonstrated a decades-long
+resilience of tropical forests in the face of quite intense disturbance
+\citep{Gourlet-Fleury2005}. Within the mechanisms of this resilience the
+recruitment processes proved central, in accordance with the IDH, and
+would deserved closer focus to clarify their determinisms and ecological
+rules. The disturbance range however stayed within the spectrum of
+selective logging, with a forest cover remaining all along the
+experiment, and the response mechanisms would probably be much different
+after harder disturbance.
 

### 2018-06-04 WholePlotTrajectories.Rmd
pdf

--- 2018-06-04 WholePlotTrajectories.Rmd
@@
     Specifically we analysed the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 key functional traits assessing species ecology.
-
     We highlighted a decoupling between functional and taxonomic trajectories, translating a fast recovery of communities functioning while their taxonomic recovery was slowed by the recruitment of infrequent species specific to mature forests.

### 2018-06-04 WholePlotTrajectories.tex
pdf

@@
 (IDH) remain debated as well as the completeness of forests resilience
-in all taxonomic and functional facets.
+in all taxonomic and functoinal facets. To disentangle the ecological
+processes driving forests dynamics we examined the the trajectories over
+30 years after a gradient of logging and thinning disturbance, focusing
+on communities taxonomic and functional composition and diversity.
+Specifically we analysed the trajectories of communities taxonomic
+richness and evennes and functional composition, diversity and
+redundancy based on 7 key functional traits assessing species ecology.
+We highlighted a decoupling between functional and taxonomic
+trajectories, translating a fast recovery of communities functioning
+while their taxonomic recovery was slowed by the recruitment of
+infrequent species specific to mature forests. The IDH consistently
+predicted communities functional trajectories but poorly reflected the
+taxonomic ones, blurried by changes in the functional redundancy
+entailing competitive exclusion. Although communities consistently
+recovered their functioning after 30 years, initially infrequent species
+proved long to recover which impacted the taxonomic structure and
+functoinal redundancy over several decades. Our results acknowledged the
+need of decades-long recovery cycles to restore the resilience of the
+community and prevent species loss in the long term.
 }

### 2018-06-05 WholePlotTrajectories.Rmd
relecture

--- 2018-06-04 WholePlotTrajectories.Rmd
@@
 The resilience of communities functional structure, that is the most direct link between biodiversity and ecosystem functioning [@Diaz2005], meant a consistent recovery of ecosystem processes in the long term [@Guariguata2001].
-The resilience of communities taxonomy meant the maintenance of their initial differences in composition and diversity, suggesting the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003].
-It besides suggested that recovery trajectories depend on the initial composition eventually restored that would the the asymptot of the humped shaped dynamics [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
+The resilience of communities taxonomy meant the maintenance of their initial differences in composition and diversity, suggesting the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003], and the dependency of recovery trajectories on the initial composition to be recovered that would the the asymptot of the humped shaped dynamics [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
 
@@
 The decoupling between functional and taxonomic dynamics has already been observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015; @Guariguata2001].
-According to the "vegetation quantity effect" [@Grime1998] functional trajectories rely on the pool of dominant species, which diversity and evenness were enhanced after disturbance, and which rapidly recovered their initial functional structure.
-However, communities evenness remained high so infrequent species still missed to the taxonomic recovery that proved all the more longer to recover that unrecovered species would be functionally redundant and probably underwent competition processes.
+According to the "vegetation quantity effect" [@Grime1998] functional trajectories rely on the pool of dominant species, which diversity and evenness increased after disturbance but rapidly recovered their initial functional structure.
+Although the pool of dominant species recovered quickly, communities evenness remained high so infrequent species still missed to the taxonomic recovery.
+Taxonomic recovery was then all the more longer that unrecovered species would be functionally redundant and probably underwent selective competition processes.
 
@@
 
-Validating the IDH, communities trajectories confirmed the diversity increase after disturbance through an enhanced growth of otherwise less favored species.
-This was strong regarding the functional diversity but more blurry regarding their taxonomic structure.
+In accordance with the IDH the trajectories confirmed the increase of communities diversity after disturbance with the favored growth of otherwise less favored species.
+However, although it was clear for the functional diversity this tendency was more blurry regarding communities taxonomic structure.
 
-The taxonomic richness was weakly or negatively impacted by intense disturbance, as observed on several post logging surveys [@Cannon1998; @Baraloto2012a], while it substantially increased after low intensity disturbance.
-Disturbance also enhanced taxonomic evenness, but the disturbance intensity was only weakly and lately correlated to communities eveness (*i.e.* Simpson diversity).
-Contrastingly, disturbance intensity consistently predicted the significant increase of communities functional diversity for 30 years.
-Disturbance entailed specific turnover within communities, either among pre-disturbance survivors or among newly recruited trees.
-As the composition of old-growth survivors proved to mirror initial communities [@Herault2018], disturbance likely impacted trees recruitment through the enhanced growth and survival of previously infrequent species and functional types.
-Consistently, disturbance resulted in increasing taxonomic dissimilarity compared to pre-disturbance communities and significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
-Disturbance then caused a reorganization of the typical high dominance structure of hyperdiverse mature forests after disturbance, benefiting to pioneers and light demanding species. 
-Likely, the changes in abiotic environment and competitive pressure favored pioneers which outcompete other species in non limiting resources but are excluded in mature forests through the cmpetition with long-lived, more resistant and shade bearers.
-Therefore, consistently with the IDH the trajectories after disturbance relied on the environmental niches made available by disturbance filled by species from a restrictied functional range.
-Recruited species therefore differed from pre-disturbance ones, constituting a community all the more diversified that the disturbance was intense [@Molino2001].
+The taxonomic richness was enhanced after low disturbance but it was weakly or negatively impacted by intense disturbance, as observed on several post logging surveys [@Cannon1998; @Baraloto2012a].
+The taxonomic evenness also increased after disturbance, but the correlation between the disturbance intensity and the diversity increase remained weak and valid only after 20 years (*i.e.* Simpson diversity).
+
+Contrastingly the functional diversity increase after disturbance was determined by the disturbance intensity all along the 30 years monitored.
+We already demonstrated that disturbance primarily entailed turnover of specific within communities, either among pre-disturbance survivors or among newly recruited trees.
+It wasdemonstrated that the composition of old-growth survivors proved to mirror initial communities [@Herault2018], so disturbance rather impacted trees recruitment with the enhanced growth and survival of previously infrequent species and functional types.
+Consistently disturbance resulted in an increase of the taxonomic dissimilarity between communities and their pre-disturbance composition, and in significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
+Disturbance entailed a reorganization of the typical high dominance structure of hyperdiverse mature forests benefiting to pioneers and light demanding species: the changes in abiotic environment and competitive pressure favored pioneers which outcompete other species in non limiting resources but were excluded in mature forests by long-lived, more resistant and shade tolerant species.
+Corroborating the IDH, the trajectories after disturbance relied on the environmental niches made available and filled by species from a restrictied functional range.
+Recruited species therefore differed from pre-disturbance ones, constituting a community all the more diverse that the disturbance was intense [@Molino2001].
 
@@
 
-Both the lag between taxonomic and functional recovery and the middling consistency of the IDH regarding communities taxonomic diversity made central the role of functional redundancy.
-Functional redundancy, determinant of communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005], seemed not to have recovered 30 years after disturbance.
+Both the lag between taxonomic and functional recovery and the middling consistency of the IDH regarding communities taxonomic diversity were explained by the incomplete recovery of communities functional redundancy, which besides implies that communities resilience remained altered 30 years after disturbance [@Trenbath1999; @Elmqvist2003; @Diaz2005].
 
-Communities with high functional redundancy are expected to be able to lose species with a limited impact on their functional structure [@Carmona2016].
-Functional redundancy therefore mitigates the impacts of disturbance, but we expect the redundancy itself to be reduced or re-organized within the functional space.
-Although the overall functional redundancy did not follow consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
-This restricted redundancy decreased according to the disturbance intensity before returning towards the initial value but had not recovered for any disturbed community after 30 years.
-The recovery of the redundancy within the initial functional space explained the hampered recovery of infrequent species specific to old-growth forests that would be functionally redundant and therefore underwent competitive pressures limiting species functional similarity [@Mayfield2010].
-The unachieved humped shaped redundancy trajectory meant a decrease of the resilience of pre-disturbance communities and an increase in the resilience of disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species).
-After disturbance the chances were higher to have long lasting or self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013], which highly questioned the completeness of the resilience [@Chazdon2003a].
-Specificaly, species contingent to undisturbed forests might be impaired, threatening their maintenance, the risk to loose cornerstone species and trigger unexpected ecological consequences would be higher [@Jones1994;@Diaz2005;@Gardner2007].
+Functional redundancy mitigates the loss of species on communities functional structure [@Carmona2016], and therefore reduces the impact on disturbance.
+The functional redundancy itself however might be reduced or re-organized within the functional space  and alter communities resilience.
+Although the functional redundancy considered in the whole commnities functional space did not follow consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
+This restricted redundancy first decreased, all the more so that the disturbance was intense, and then returned towards the initial value but this had not recovered for any disturbed community after 30 years.
+After disturbance communities then differently occupied the functional space, as it was suggested by the shifts observed for the functional diversity and traits trajectories, before recovering the specific highly redundant initial functional space.
+This increasing redundancy along the trajectory explained the hampered recovery of infrequent species specific to old-growth forests, as they were likely functionally redundant and therefore underwent competitive pressures limiting species functional similarity [@Mayfield2010].
+The unachieved recovery of the functional redundancy meant a decreased resilience of pre-disturbance communities and an increased of disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species).
+After disturbance the chances were therefore higher to have long lasting or self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013].
+The recovery was not complete until the functional redundancy and resilience of the initial communities remained altered.
+Specificaly, such incomplete recovery would threaten the maintenance of species contingent to undisturbed forests and increase the risk of cornerstone species loss and unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
 
@@
 
-our study defined the decoupled functional and taxonomic trajectories of tropical rainforests after disturbance.
-These trajectories shed light on their rapid functional recovery but slower and more variable taxonomic dynamics.
-Consistently with the IDH, functional trajectories were constrained by determined processes favoring pioneers and light-demanding species after disturbance.
-The taxonomic dynamics however proved more stochastic due to the recovery of infrequent, functionally redundant species hampered by limiting similarity processes.
-Although communities functioning rapidly recovered, their taxonomic structure and functional redundancy remained altered for several decades.
-The trajectories therefore demonstrated a decades-long resilience of tropical forests in the face of quite intense disturbance [@Gourlet-Fleury2005].
-Within the mechanisms of this resilience the recruitment processes proved central, in accordance with the IDH, and would deserved closer focus to clarify their determinisms and ecological rules.
+Our study defined the decoupled functional and taxonomic trajectories of tropical rainforests after disturbancethat demonstrated a rapid functional recovery but a slower and more variable taxonomic one.
+Consistently with the IDH, functional trajectories were driven by deterministic processes favoring pioneers and light-demanding species after disturbance.
+The following  functional shifts entailed a re-organization of communities functional redundancy in the functional space that proved long to recover and explained the more variable and longer taxonomic trajectories as shade tolerant species of old-growth forests faced functional redundancy and limiting similarity processes.
+The resilience of tropical forests then proved long in the face of quite intense disturbance and did not preclude the settling of a persistent disturbanc-specific community [@Gourlet-Fleury2005].
+Within the mechanisms of this resilience the recruitment processes proved central, in accordance with the IDH, and would deserve closer focus to specify their ecological drivers.
 The disturbance range however stayed within the spectrum of selective logging, with a forest cover remaining all along the experiment, and the response mechanisms would probably be much different after harder disturbance.

### 2018-06-05 WholePlotTrajectories.tex
relecture

@@
 existence of multiple stable equilibria as assumed for highly diverse
-and productive ecosystems \citep{Chase2003}. It besides suggested that
-recovery trajectories depend on the initial composition eventually
-restored that would the the asymptot of the humped shaped dynamics
+and productive ecosystems \citep{Chase2003}, and the dependency of
+recovery trajectories on the initial composition to be recovered that
+would the the asymptot of the humped shaped dynamics
 \citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.
@@
 functional trajectories rely on the pool of dominant species, which
-diversity and evenness were enhanced after disturbance, and which
-rapidly recovered their initial functional structure. However,
-communities evenness remained high so infrequent species still missed to
-the taxonomic recovery that proved all the more longer to recover that
-unrecovered species would be functionally redundant and probably
-underwent competition processes.
+diversity and evenness increased after disturbance but rapidly recovered
+their initial functional structure. Although the pool of dominant
+species recovered quickly, communities evenness remained high so
+infrequent species still missed to the taxonomic recovery. Taxonomic
+recovery was then all the more longer that unrecovered species would be
+functionally redundant and probably underwent selective competition
+processes.
 
@@
 
-Validating the IDH, communities trajectories confirmed the diversity
-increase after disturbance through an enhanced growth of otherwise less
-favored species. This was strong regarding the functional diversity but
-more blurry regarding their taxonomic structure.
+In accordance with the IDH the trajectories confirmed the increase of
+communities diversity after disturbance with the favored growth of
+otherwise less favored species. However, although it was clear for the
+functional diversity this tendency was more blurry regarding communities
+taxonomic structure.
 
-The taxonomic richness was weakly or negatively impacted by intense
-disturbance, as observed on several post logging surveys
-\citep{Cannon1998, Baraloto2012a}, while it substantially increased
-after low intensity disturbance. Disturbance also enhanced taxonomic
-evenness, but the disturbance intensity was only weakly and lately
-correlated to communities eveness (\emph{i.e.} Simpson diversity).
-Contrastingly, disturbance intensity consistently predicted the
-significant increase of communities functional diversity for 30 years.
-Disturbance entailed specific turnover within communities, either among
-pre-disturbance survivors or among newly recruited trees. As the
+The taxonomic richness was enhanced after low disturbance but it was
+weakly or negatively impacted by intense disturbance, as observed on
+several post logging surveys \citep{Cannon1998, Baraloto2012a}. The
+taxonomic evenness also increased after disturbance, but the correlation
+between the disturbance intensity and the diversity increase remained
+weak and valid only after 20 years (\emph{i.e.} Simpson diversity).
+
+Contrastingly the functional diversity increase after disturbance was
+determined by the disturbance intensity all along the 30 years
+monitored. We already demonstrated that disturbance primarily entailed
+turnover of specific within communities, either among pre-disturbance
+survivors or among newly recruited trees. It wasdemonstrated that the
 composition of old-growth survivors proved to mirror initial communities
-\citep{Herault2018}, disturbance likely impacted trees recruitment
-through the enhanced growth and survival of previously infrequent
-species and functional types. Consistently, disturbance resulted in
-increasing taxonomic dissimilarity compared to pre-disturbance
-communities and significant functional shifts towards
-resource-acquisitive strategies (sharp increase in the SLA, leaf
+\citep{Herault2018}, so disturbance rather impacted trees recruitment
+with the enhanced growth and survival of previously infrequent species
+and functional types. Consistently disturbance resulted in an increase
+of the taxonomic dissimilarity between communities and their
+pre-disturbance composition, and in significant functional shifts
+towards resource-acquisitive strategies (sharp increase in the SLA, leaf
 thickness and bark thickness and decrease in wood density, leaf
 toughness and maximum height)
-\citep{Westoby1998, Wright2004, Reich2014}. Disturbance then caused a
+\citep{Westoby1998, Wright2004, Reich2014}. Disturbance entailed a
 reorganization of the typical high dominance structure of hyperdiverse
-mature forests after disturbance, benefiting to pioneers and light
-demanding species. Likely, the changes in abiotic environment and
-competitive pressure favored pioneers which outcompete other species in
-non limiting resources but are excluded in mature forests through the
-cmpetition with long-lived, more resistant and shade bearers. Therefore,
-consistently with the IDH the trajectories after disturbance relied on
-the environmental niches made available by disturbance filled by species
-from a restrictied functional range. Recruited species therefore
-differed from pre-disturbance ones, constituting a community all the
-more diversified that the disturbance was intense \citep{Molino2001}.
+mature forests benefiting to pioneers and light demanding species: the
+changes in abiotic environment and competitive pressure favored pioneers
+which outcompete other species in non limiting resources but were
+excluded in mature forests by long-lived, more resistant and shade
+tolerant species. Corroborating the IDH, the trajectories after
+disturbance relied on the environmental niches made available and filled
+by species from a restrictied functional range. Recruited species
+therefore differed from pre-disturbance ones, constituting a community
+all the more diverse that the disturbance was intense
+\citep{Molino2001}.
 
@@
 Both the lag between taxonomic and functional recovery and the middling
-consistency of the IDH regarding communities taxonomic diversity made
-central the role of functional redundancy. Functional redundancy,
-determinant of communities resilience
-\citep{Trenbath1999, Elmqvist2003, Diaz2005}, seemed not to have
-recovered 30 years after disturbance.
+consistency of the IDH regarding communities taxonomic diversity were
+explained by the incomplete recovery of communities functional
+redundancy, which besides implies that communities resilience remained
+altered 30 years after disturbance
+\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-Communities with high functional redundancy are expected to be able to
-lose species with a limited impact on their functional structure
-\citep{Carmona2016}. Functional redundancy therefore mitigates the
-impacts of disturbance, but we expect the redundancy itself to be
-reduced or re-organized within the functional space. Although the
-overall functional redundancy did not follow consistent trajectories
-after disturbance, the redundancy restricted to the functional space of
-the initial community clearly followed humped shaped trajectories. This
-restricted redundancy decreased according to the disturbance intensity
-before returning towards the initial value but had not recovered for any
-disturbed community after 30 years. The recovery of the redundancy
-within the initial functional space explained the hampered recovery of
-infrequent species specific to old-growth forests that would be
+Functional redundancy mitigates the loss of species on communities
+functional structure \citep{Carmona2016}, and therefore reduces the
+impact on disturbance. The functional redundancy itself however might be
+reduced or re-organized within the functional space and alter
+communities resilience. Although the functional redundancy considered in
+the whole commnities functional space did not follow consistent
+trajectories after disturbance, the redundancy restricted to the
+functional space of the initial community clearly followed humped shaped
+trajectories. This restricted redundancy first decreased, all the more
+so that the disturbance was intense, and then returned towards the
+initial value but this had not recovered for any disturbed community
+after 30 years. After disturbance communities then differently occupied
+the functional space, as it was suggested by the shifts observed for the
+functional diversity and traits trajectories, before recovering the
+specific highly redundant initial functional space. This increasing
+redundancy along the trajectory explained the hampered recovery of
+infrequent species specific to old-growth forests, as they were likely
 functionally redundant and therefore underwent competitive pressures
 limiting species functional similarity \citep{Mayfield2010}. The
-unachieved humped shaped redundancy trajectory meant a decrease of the
-resilience of pre-disturbance communities and an increase in the
-resilience of disturbance-specific communities (corresponding to lower
-richness and higher dominance of pioneer species). After disturbance the
-chances were higher to have long lasting or self-maintained
-compositional changes in favor of disturbance resistant species, lianas
-or epiphytes \citep{Haddad2008, Burslem2000, Martin2013}, which highly
-questioned the completeness of the resilience \citep{Chazdon2003a}.
-Specificaly, species contingent to undisturbed forests might be
-impaired, threatening their maintenance, the risk to loose cornerstone
-species and trigger unexpected ecological consequences would be higher
-\citep{Jones1994, Diaz2005, Gardner2007}.
+unachieved recovery of the functional redundancy meant a decreased
+resilience of pre-disturbance communities and an increased of
+disturbance-specific communities (corresponding to lower richness and
+higher dominance of pioneer species). After disturbance the chances were
+therefore higher to have long lasting or self-maintained compositional
+changes in favor of disturbance resistant species, lianas or epiphytes
+\citep{Haddad2008, Burslem2000, Martin2013}. The recovery was not
+complete until the functional redundancy and resilience of the initial
+communities remained altered. Specificaly, such incomplete recovery
+would threaten the maintenance of species contingent to undisturbed
+forests and increase the risk of cornerstone species loss and unexpected
+ecological consequences
+\citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.
 
@@
 
-our study defined the decoupled functional and taxonomic trajectories of
-tropical rainforests after disturbance. These trajectories shed light on
-their rapid functional recovery but slower and more variable taxonomic
-dynamics. Consistently with the IDH, functional trajectories were
-constrained by determined processes favoring pioneers and
-light-demanding species after disturbance. The taxonomic dynamics
-however proved more stochastic due to the recovery of infrequent,
-functionally redundant species hampered by limiting similarity
-processes. Although communities functioning rapidly recovered, their
-taxonomic structure and functional redundancy remained altered for
-several decades. The trajectories therefore demonstrated a decades-long
-resilience of tropical forests in the face of quite intense disturbance
-\citep{Gourlet-Fleury2005}. Within the mechanisms of this resilience the
-recruitment processes proved central, in accordance with the IDH, and
-would deserved closer focus to clarify their determinisms and ecological
-rules. The disturbance range however stayed within the spectrum of
-selective logging, with a forest cover remaining all along the
-experiment, and the response mechanisms would probably be much different
-after harder disturbance.
+Our study defined the decoupled functional and taxonomic trajectories of
+tropical rainforests after disturbancethat demonstrated a rapid
+functional recovery but a slower and more variable taxonomic one.
+Consistently with the IDH, functional trajectories were driven by
+deterministic processes favoring pioneers and light-demanding species
+after disturbance. The following functional shifts entailed a
+re-organization of communities functional redundancy in the functional
+space that proved long to recover and explained the more variable and
+longer taxonomic trajectories as shade tolerant species of old-growth
+forests faced functional redundancy and limiting similarity processes.
+The resilience of tropical forests then proved long in the face of quite
+intense disturbance and did not preclude the settling of a persistent
+disturbanc-specific community \citep{Gourlet-Fleury2005}. Within the
+mechanisms of this resilience the recruitment processes proved central,
+in accordance with the IDH, and would deserve closer focus to specify
+their ecological drivers. The disturbance range however stayed within
+the spectrum of selective logging, with a forest cover remaining all
+along the experiment, and the response mechanisms would probably be much
+different after harder disturbance.
 

### 2018-06-06 WholePlotTrajectories.Rmd
relecture

@@
 
-    To disentangle the ecological processes driving forests dynamics we examined the the trajectories over 30 years after a gradient of logging and thinning disturbance, focusing on communities taxonomic and functional composition and diversity.
+    To disentangle the ecological processes driving forests dynamics we examined the trajectories over 30 years after a gradient of logging and thinning disturbance, focusing on communities taxonomic and functional composition and diversity.
     Specifically we analysed the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 key functional traits assessing species ecology.
@@
 Communities then shifted from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
-This shift translated into consistent trajectories for key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction and life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
+This shift translated into consistent trajectories for key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
 

### 2018-06-07 WholePlotTrajectories.Rmd
coquilles

@@
 A combination of taxonomic and functional approaches are essential to fully assess communities response to disturbance and differences in their respective trajectories are insightful of underlying mechanisms [@Lohbeck2015;@Guariguata2001].
-Although taxonomic and functional diversity are complementary and sometimes decouples, they are combined in the measure of communities functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
+Although taxonomic and functional diversity are complementary and sometimes decoupled, they are combined in the measure of communities functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
 Functional redundancy is determinant of communities resilience as high redundancy, like in highly diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning [@Trenbath1999; @Elmqvist2003; @Diaz2005].
@@
 We already demonstrated that disturbance primarily entailed turnover of specific within communities, either among pre-disturbance survivors or among newly recruited trees.
-It wasdemonstrated that the composition of old-growth survivors proved to mirror initial communities [@Herault2018], so disturbance rather impacted trees recruitment with the enhanced growth and survival of previously infrequent species and functional types.
+It was demonstrated that the composition of old-growth survivors proved to mirror initial communities [@Herault2018], so disturbance rather impacted trees recruitment with the enhanced growth and survival of previously infrequent species and functional types.
 Consistently disturbance resulted in an increase of the taxonomic dissimilarity between communities and their pre-disturbance composition, and in significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].

### 2018-06-12 WholePlotTrajectories.Rmd
discussion

@@
 ```{r Options, echo=FALSE, message=FALSE, warning=F}
-knitr::opts_chunk$set(cache=F, warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
+knitr::opts_chunk$set(cache=T, warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
 # Installation des packages si nécessaire et chargement
@@
 
-## A validation of the intermediate disturbance hypothesis
+## The extent of the intermediate disturbance hypothesis
 
-In accordance with the IDH the trajectories confirmed the increase of communities diversity after disturbance with the favored growth of otherwise less favored species.
-However, although it was clear for the functional diversity this tendency was more blurry regarding communities taxonomic structure.
+The examination of taxonomic and functional trajectories suggested that communities functional structure and dominant species pool relied upon deterministic processes following the IDH while the pool of rare species relied upon more stochastic processes.
 
-The taxonomic richness was enhanced after low disturbance but it was weakly or negatively impacted by intense disturbance, as observed on several post logging surveys [@Cannon1998; @Baraloto2012a].
-The taxonomic evenness also increased after disturbance, but the correlation between the disturbance intensity and the diversity increase remained weak and valid only after 20 years (*i.e.* Simpson diversity).
+The disturbance intensity did not predict the taxonomic richness, as already observed in the Paracou station [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998], nor did it strongly correlate with the taxonomic evenness (*i.e.* Simpson diversity) although the it was enhanced after disturbance.
 
-Contrastingly the functional diversity increase after disturbance was determined by the disturbance intensity all along the 30 years monitored.
-We already demonstrated that disturbance primarily entailed turnover of specific within communities, either among pre-disturbance survivors or among newly recruited trees.
-It was demonstrated that the composition of old-growth survivors proved to mirror initial communities [@Herault2018], so disturbance rather impacted trees recruitment with the enhanced growth and survival of previously infrequent species and functional types.
-Consistently disturbance resulted in an increase of the taxonomic dissimilarity between communities and their pre-disturbance composition, and in significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
+Contrastingly, the disturbance intensity predicted the increase in functional diversity along the 30 years monitored after disturbance.
+The increase in functional diversity paralleled the important specific turnover and significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
+As the old-growth, pre-disturbance survivor trees proved to mirror initial communities [@Herault2018] the functional trajectories and taxonomic turnover rather relied on the pool of newly recruited trees driven by the enhanced growth and survival of previously infrequent species and functional types.
 Disturbance entailed a reorganization of the typical high dominance structure of hyperdiverse mature forests benefiting to pioneers and light demanding species: the changes in abiotic environment and competitive pressure favored pioneers which outcompete other species in non limiting resources but were excluded in mature forests by long-lived, more resistant and shade tolerant species.
-Corroborating the IDH, the trajectories after disturbance relied on the environmental niches made available and filled by species from a restrictied functional range.
-Recruited species therefore differed from pre-disturbance ones, constituting a community all the more diverse that the disturbance was intense [@Molino2001].
+Disturbance trajectories therefore relied on the environmental niches made available and filled by species from a restricted functional range that became dominant and structured the functionalcharacteristics of the community [@Grime1998].
+The intermediate disturbance hypothesis predicted communities functional diversity and pool of dominant species [@Molino2001], but rare species relied on different processes and disprove the IDh for the whole taxonomic community [@Hubbell2001;@Chave2004].
 
@@
 
-Both the lag between taxonomic and functional recovery and the middling consistency of the IDH regarding communities taxonomic diversity were explained by the incomplete recovery of communities functional redundancy, which besides implies that communities resilience remained altered 30 years after disturbance [@Trenbath1999; @Elmqvist2003; @Diaz2005].
+The functional and taxonomic decoupling and the disprove of the IDh for communities taxonomic structure were explained were explained by persisting changes in communities functoinal redundancy that implied communities incomplete resilience [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
@@
 The functional redundancy itself however might be reduced or re-organized within the functional space  and alter communities resilience.
-Although the functional redundancy considered in the whole commnities functional space did not follow consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
-This restricted redundancy first decreased, all the more so that the disturbance was intense, and then returned towards the initial value but this had not recovered for any disturbed community after 30 years.
-After disturbance communities then differently occupied the functional space, as it was suggested by the shifts observed for the functional diversity and traits trajectories, before recovering the specific highly redundant initial functional space.
-This increasing redundancy along the trajectory explained the hampered recovery of infrequent species specific to old-growth forests, as they were likely functionally redundant and therefore underwent competitive pressures limiting species functional similarity [@Mayfield2010].
+Although the functional redundancy considered in the whole communities functional space did not follow consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
+This restricted redundancy first decreased, all the more so that the disturbance was intense, before returning towards the initial value but this had not recovered for any disturbed community after 30 years.
+After disturbance communities differently occupied the functional space following traits trajectories and the shifts in functional composition, before recovering the initial functional space and its high redundancy.
+This recovery phase was slow as it depend on infrequent, functionally redundant species specific to old-growth forests that rely upon different ecological processes, stochastic or deterministic competitive pressures limiting species functional similarity [@Chave2004;@Mayfield2010].
+
 The unachieved recovery of the functional redundancy meant a decreased resilience of pre-disturbance communities and an increased of disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species).
-After disturbance the chances were therefore higher to have long lasting or self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013].
-The recovery was not complete until the functional redundancy and resilience of the initial communities remained altered.
-Specificaly, such incomplete recovery would threaten the maintenance of species contingent to undisturbed forests and increase the risk of cornerstone species loss and unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
+After disturbance, chances were higher to observe long lasting or self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013].
+The recovery remained incomplete until the recovery of the initial functional redundancy and resilience, and this specifically threatened the maintenance of species contingent to undisturbed forests and increased the risk of loosing cornerstone species and having unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
 

### 2018-06-12 WholePlotTrajectories.tex
discussion

@@
 in all taxonomic and functoinal facets. To disentangle the ecological
-processes driving forests dynamics we examined the the trajectories over
-30 years after a gradient of logging and thinning disturbance, focusing
-on communities taxonomic and functional composition and diversity.
+processes driving forests dynamics we examined the trajectories over 30
+years after a gradient of logging and thinning disturbance, focusing on
+communities taxonomic and functional composition and diversity.
 Specifically we analysed the trajectories of communities taxonomic
@@
 stem specific gravity and bark thickness), tree growth and reproduction
-and life history traits (seed mass and maximum height)
-\citep{Wright2004, Westoby2006a, Chave2009b}.
+life history traits (seed mass and maximum height)
+\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}.
 
@@
 \citep{Lohbeck2015, Guariguata2001}. Although taxonomic and functional
-diversity are complementary and sometimes decouples, they are combined
+diversity are complementary and sometimes decoupled, they are combined
 in the measure of communities functional redundancy that quantifies the
@@
 
-\subsection{A validation of the intermediate disturbance
-hypothesis}\label{a-validation-of-the-intermediate-disturbance-hypothesis}
+\subsection{The extent of the intermediate disturbance
+hypothesis}\label{the-extent-of-the-intermediate-disturbance-hypothesis}
 
-In accordance with the IDH the trajectories confirmed the increase of
-communities diversity after disturbance with the favored growth of
-otherwise less favored species. However, although it was clear for the
-functional diversity this tendency was more blurry regarding communities
-taxonomic structure.
+The examination of taxonomic and functional trajectories suggested that
+communities functional structure and dominant species pool relied upon
+deterministic processes following the IDH while the pool of rare species
+relied upon more stochastic processes.
 
-The taxonomic richness was enhanced after low disturbance but it was
-weakly or negatively impacted by intense disturbance, as observed on
-several post logging surveys \citep{Cannon1998, Baraloto2012a}. The
-taxonomic evenness also increased after disturbance, but the correlation
-between the disturbance intensity and the diversity increase remained
-weak and valid only after 20 years (\emph{i.e.} Simpson diversity).
+The disturbance intensity did not predict the taxonomic richness, as
+already observed in the Paracou station \citep{Baraloto2012a} and in
+Bornean tropical forests \citep{Cannon1998}, nor did it strongly
+correlate with the taxonomic evenness (\emph{i.e.} Simpson diversity)
+although the it was enhanced after disturbance.
 
-Contrastingly the functional diversity increase after disturbance was
-determined by the disturbance intensity all along the 30 years
-monitored. We already demonstrated that disturbance primarily entailed
-turnover of specific within communities, either among pre-disturbance
-survivors or among newly recruited trees. It wasdemonstrated that the
-composition of old-growth survivors proved to mirror initial communities
-\citep{Herault2018}, so disturbance rather impacted trees recruitment
-with the enhanced growth and survival of previously infrequent species
-and functional types. Consistently disturbance resulted in an increase
-of the taxonomic dissimilarity between communities and their
-pre-disturbance composition, and in significant functional shifts
-towards resource-acquisitive strategies (sharp increase in the SLA, leaf
-thickness and bark thickness and decrease in wood density, leaf
-toughness and maximum height)
-\citep{Westoby1998, Wright2004, Reich2014}. Disturbance entailed a
-reorganization of the typical high dominance structure of hyperdiverse
-mature forests benefiting to pioneers and light demanding species: the
-changes in abiotic environment and competitive pressure favored pioneers
-which outcompete other species in non limiting resources but were
-excluded in mature forests by long-lived, more resistant and shade
-tolerant species. Corroborating the IDH, the trajectories after
-disturbance relied on the environmental niches made available and filled
-by species from a restrictied functional range. Recruited species
-therefore differed from pre-disturbance ones, constituting a community
-all the more diverse that the disturbance was intense
-\citep{Molino2001}.
+Contrastingly, the disturbance intensity predicted the increase in
+functional diversity along the 30 years monitored after disturbance. The
+increase in functional diversity paralleled the important specific
+turnover and significant functional shifts towards resource-acquisitive
+strategies (sharp increase in the SLA, leaf thickness and bark thickness
+and decrease in wood density, leaf toughness and maximum height)
+\citep{Westoby1998, Wright2004, Reich2014}. As the old-growth,
+pre-disturbance survivor trees proved to mirror initial communities
+\citep{Herault2018} the functional trajectories and taxonomic turnover
+rather relied on the pool of newly recruited trees driven by the
+enhanced growth and survival of previously infrequent species and
+functional types. Disturbance entailed a reorganization of the typical
+high dominance structure of hyperdiverse mature forests benefiting to
+pioneers and light demanding species: the changes in abiotic environment
+and competitive pressure favored pioneers which outcompete other species
+in non limiting resources but were excluded in mature forests by
+long-lived, more resistant and shade tolerant species. Disturbance
+trajectories therefore relied on the environmental niches made available
+and filled by species from a restricted functional range that became
+dominant and structured the functionalcharacteristics of the community
+\citep{Grime1998}. The intermediate disturbance hypothesis predicted
+communities functional diversity and pool of dominant species
+\citep{Molino2001}, but rare species relied on different processes and
+disprove the IDh for the whole taxonomic community
+\citep{Hubbell2001, Chave2004}.
 
@@
 
-Both the lag between taxonomic and functional recovery and the middling
-consistency of the IDH regarding communities taxonomic diversity were
-explained by the incomplete recovery of communities functional
-redundancy, which besides implies that communities resilience remained
-altered 30 years after disturbance
+The functional and taxonomic decoupling and the disprove of the IDh for
+communities taxonomic structure were explained were explained by
+persisting changes in communities functoinal redundancy that implied
+communities incomplete resilience
 \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
@@
 communities resilience. Although the functional redundancy considered in
-the whole commnities functional space did not follow consistent
+the whole communities functional space did not follow consistent
 trajectories after disturbance, the redundancy restricted to the
@@
 trajectories. This restricted redundancy first decreased, all the more
-so that the disturbance was intense, and then returned towards the
+so that the disturbance was intense, before returning towards the
 initial value but this had not recovered for any disturbed community
-after 30 years. After disturbance communities then differently occupied
-the functional space, as it was suggested by the shifts observed for the
-functional diversity and traits trajectories, before recovering the
-specific highly redundant initial functional space. This increasing
-redundancy along the trajectory explained the hampered recovery of
-infrequent species specific to old-growth forests, as they were likely
-functionally redundant and therefore underwent competitive pressures
-limiting species functional similarity \citep{Mayfield2010}. The
-unachieved recovery of the functional redundancy meant a decreased
+after 30 years. After disturbance communities differently occupied the
+functional space following traits trajectories and the shifts in
+functional composition, before recovering the initial functional space
+and its high redundancy. This recovery phase was slow as it depend on
+infrequent, functionally redundant species specific to old-growth
+forests that rely upon different ecological processes, stochastic or
+deterministic competitive pressures limiting species functional
+similarity \citep{Chave2004, Mayfield2010}.
+
+The unachieved recovery of the functional redundancy meant a decreased
 resilience of pre-disturbance communities and an increased of
 disturbance-specific communities (corresponding to lower richness and
-higher dominance of pioneer species). After disturbance the chances were
-therefore higher to have long lasting or self-maintained compositional
-changes in favor of disturbance resistant species, lianas or epiphytes
-\citep{Haddad2008, Burslem2000, Martin2013}. The recovery was not
-complete until the functional redundancy and resilience of the initial
-communities remained altered. Specificaly, such incomplete recovery
-would threaten the maintenance of species contingent to undisturbed
-forests and increase the risk of cornerstone species loss and unexpected
-ecological consequences
+higher dominance of pioneer species). After disturbance, chances were
+higher to observe long lasting or self-maintained compositional changes
+in favor of disturbance resistant species, lianas or epiphytes
+\citep{Haddad2008, Burslem2000, Martin2013}. The recovery remained
+incomplete until the recovery of the initial functional redundancy and
+resilience, and this specifically threatened the maintenance of species
+contingent to undisturbed forests and increased the risk of loosing
+cornerstone species and having unexpected ecological consequences
 \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.

### 2018-06-13 WholePlotTrajectories.Rmd
relecture

@@
 
-A combination of taxonomic and functional approaches are essential to fully assess communities response to disturbance and differences in their respective trajectories are insightful of underlying mechanisms [@Lohbeck2015;@Guariguata2001].
-Although taxonomic and functional diversity are complementary and sometimes decoupled, they are combined in the measure of communities functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
-Functional redundancy is determinant of communities resilience as high redundancy, like in highly diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning [@Trenbath1999; @Elmqvist2003; @Diaz2005].
+A combination of taxonomic and functional approaches are essential to fully assess communities response to disturbance and the differences in their trajectories are insightful of assembly processes [@Lohbeck2015;@Guariguata2001].
+The decoupling between functional and taxonomic trajectories is assessed through communities functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
+Functional redundancy, which is high in the very diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning [@Trenbath1999; @Elmqvist2003; @Diaz2005] and thus determines communities resilience.
+High functional redundancy besides increases the functional overlap among species, increasing their similarity, and would foster neutral stochastic processes [@Gravel2006].
 

### 2018-06-15 WholePlotTrajectories.tex
relecture

@@
 A combination of taxonomic and functional approaches are essential to
-fully assess communities response to disturbance and differences in
-their respective trajectories are insightful of underlying mechanisms
-\citep{Lohbeck2015, Guariguata2001}. Although taxonomic and functional
-diversity are complementary and sometimes decoupled, they are combined
-in the measure of communities functional redundancy that quantifies the
-amount of shared trait values among species \citep{Carmona2016}.
-Functional redundancy is determinant of communities resilience as high
-redundancy, like in highly diverse tropical forests
-\citep{Bellwood2006}, mitigates the impacts of species removal on
-ecosystem functioning \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+fully assess communities response to disturbance and the differences in
+their trajectories are insightful of assembly processes
+\citep{Lohbeck2015, Guariguata2001}. The decoupling between functional
+and taxonomic trajectories is assessed through communities functional
+redundancy that quantifies the amount of shared trait values among
+species \citep{Carmona2016}. Functional redundancy, which is high in the
+very diverse tropical forests \citep{Bellwood2006}, mitigates the
+impacts of species removal on ecosystem functioning
+\citep{Trenbath1999, Elmqvist2003, Diaz2005} and thus determines
+communities resilience. High functional redundancy besides increases the
+functional overlap among species, increasing their similarity, and would
+foster neutral stochastic processes \citep{Gravel2006}.
 

### 2018-06-18 WholePlotTrajectories.Rmd
relecture discussion

--- 2018-06-12 WholePlotTrajectories.Rmd
@@
     Understand the maintenance of forests biodiversity structure, functioning and dynamics is urgent to anticipate their fate in the global changing context.
-    Communities diversity structure is assumed to rely on a constant regime of disturbance and peak at intermediate intensity.
-    For tropical forests this intermediate disturbance hypothesis (IDH) remain debated as well as the completeness of forests resilience in all taxonomic and functoinal facets.
+    Communities diversity are assumed to be shaped by constant regime of disturbance yielding a diversity peak at intermediate intensity, but for tropical forests this intermediate disturbance hypothesis (IDH) remains debated as well as the completeness their resilience in all taxonomic and functional facets after disturbance.
 
-    To disentangle the ecological processes driving forests dynamics we examined the trajectories over 30 years after a gradient of logging and thinning disturbance, focusing on communities taxonomic and functional composition and diversity.
-    Specifically we analysed the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 key functional traits assessing species ecology.
-    We highlighted a decoupling between functional and taxonomic trajectories, translating a fast recovery of communities functioning while their taxonomic recovery was slowed by the recruitment of infrequent species specific to mature forests.
-    The IDH consistently predicted communities functional trajectories but poorly reflected the taxonomic ones, blurried by changes in the functional redundancy entailing competitive exclusion.
-    Although communities consistently recovered their functioning after 30 years, initially infrequent species proved long to recover which impacted the taxonomic structure and functoinal redundancy over several decades.
-    Our results acknowledged the need of decades-long recovery cycles to restore the resilience of the community and prevent species loss in the long term.
+    To disentangle the ecological processes driving communities response to disturbance we analysed the taxonomic and functional trajectories following a logging and thinning disturbance gradient.
+    Specifically we examined over 30 years the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 leaf, stem and life history traits assessing species ecology.
+    Trajectories highlighted communities taxonomic resilience, maintaining their initial differences in composition, and functional resilience through fast functional trajectory common to all communities in the functional space.
+    The IDH consistently predicted communities functional trajectories according to the disturbance intensity, but poorly represented their taxonomic trajectories, blurried by the alteration of the functional redundancy and the recruitment of infrequent, shade tolerant species slowed by competitive exclusion processes.
+
+    Although communities functioning had recovered after 30 years their taxonomic diversity and composition in infrequent species and their functional redundancy remained altered.
+    these results acknowledged the need of decades-long recovery cycles to ensure communities complete resilience, and questioned communities maintenance after repeated disturbance.
 
@@
   \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
-  \hyphenation{bio-di-ver-si-ty sap-lings}
+  \hyphenation{bio-di-ver-si-ty sap-lings re-or-ga-ni-za-tion}
 bibliography: references.bib
@@
 The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
-They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The simultaneous increased demand in forests products and the substantial climatic changes currently heighten the pressure on remaining forests [@Gibson2011a; @Morales-Hidalgo2015], threatening the maintenance of communities structure, composition and functioning and their dynamics in space and time [@Anderson-Teixeira2013;@Sist2015].
+They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local climate, and the carbon, water and nutrient cycles, and ensure cultural and human well-being.
+The growing demand in forests products together with current climatic changes increases the pressure on remaining forests [@Gibson2011a; @Morales-Hidalgo2015]and threatens the maintenance and dynamics in space and time of communities structure, composition and functioning [@Anderson-Teixeira2013;@Sist2015].
 
-In tropical forest, ecological communities are constantly re-shaped by natural disturbance events that change both abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and biotic interactions like competition among species [@Herault2018].
-The cornerstone of tropical forests ecology is to understand the mechanisms and the determinants of ecosystems response to disturbance [@White2001;@Chazdon2003a]. 
+In tropical forest, ecological communities are constantly re-shaped by natural disturbance events that change both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions like competition among species [@Herault2018].
+The cornerstone of tropical forests ecology is to understand the processes and the determinants of ecosystems response to disturbance [@White2001;@Chazdon2003a]. 
 For now, this has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
 These structural parameters were thereafter sucessfully modeled, giving important insights into the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009]. 
-However the response of forests diversity in tree species remains unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014, @Liang2016] and would be most probably impacted by the changes following disturbance [@Baraloto2012a].
+However the response of forests diversity in tree species remains unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014, @Liang2016] and would be most probably impacted by the changes induced by disturbance [@Baraloto2012a].
 
 In the short-term disturbance demonstrated negligible or even positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
-Still, validations of the IDH remain scarce in the long term and mainly rely on species richness analyses that give limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-More releveant monitoring would encompass communities composition, crucial for conservation issues, and complete diversity profiles integrating both species richness and evenness, to reveal the ecological rules structuring communities [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, the functional approach that accounts for species biological attributes and role in the ecosystem would be essential to understand the correlations between ecosystems biodiversity, functioning and environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-The functional trait-based approach focused on major traits related to species ecology and performance was sucessfully adopted [@Diaz2005;@Villeger2008a], for example in highlighting deterministic processes in tropical rainforests that foster fast growing species with efficient resources acquisition after disturbance [@Molino2001;@Haddad2008;@Ruger2009].
-Communities then shifted from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
-This shift translated into consistent trajectories for key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
+Still, validations of the IDH remain scarce in the long term and mainly rely on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+More complete analysis would besides encompass communities explicit composition that is crucial for conservation issues, and communities evenness to reveal the ecological rules structuring communities [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Furthermore, a functional approach accoutning for species biological attributes and role in the ecosystem directly links ecosystems biodiversity, functioning and environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+The functional trait-based approach focused on major traits related to species ecology and performance was sucessfully adopted [@Diaz2005;@Villeger2008a], for example to reveal in tropical rainforests the deterministic processes that foster after disturbance the fast growing species with efficient resources acquisition [@Molino2001;@Haddad2008;@Ruger2009].
+This basic deterministic process entails a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
+This shift is translated into consistent trajectories of key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
+Besides communities functional composition and diversity, a key aspect of communties response to disturbance is the changes in functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
+High functional redundancy, as in the very diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning and determines communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005].
+High functional redundancy also increases the functional overlap among species and would foster neutral stochastic processes [@Gravel2006].
 
-A combination of taxonomic and functional approaches are essential to fully assess communities response to disturbance and differences in their respective trajectories are insightful of underlying mechanisms [@Lohbeck2015;@Guariguata2001].
-Although taxonomic and functional diversity are complementary and sometimes decoupled, they are combined in the measure of communities functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
-Functional redundancy is determinant of communities resilience as high redundancy, like in highly diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning [@Trenbath1999; @Elmqvist2003; @Diaz2005].
-
-Grasp all facets of communities response to disturbance comes to examine the taxonomic and functional trajectories in diversity and composition.
-These trajectories would highlight the ecological rules constraining or not communities dynamics towards the recovery of initial composition, diversity and functioning.
-They would therefore clarify the tenants in the long term of the debated Intermediate Disturbance Hypothesis for tropical forests, a cruicial point for future adaptive conservation strategies [@Adler2007].
-Here we monitored over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
+To grasp all facets of communities response to disturbance we examined here the taxonomic and functional trajectories in diversity and composition [@Lohbeck2015;@Guariguata2001].
+These trajectories would highlight the recovery of communities initial composition, diversity and functioning and the underlying ecological processes.
+They would clarify the tenants in the long term of the Intermediate Disturbance Hypothesis debated in tropical forests, and provide indications for future adaptive conservation strategies [@Adler2007].
+We monitored over 30 years the response of 75 ha of neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
 We made use of a large functional traits database browsing major leaf, stem and seed functional traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
-Specifically, we (i) questioned the coupling between taxonomic and functional response to disturbance and identified the underlying assembly processes, which allowed to (ii) clarify the validity of the IDH in the long term for tropical forest and (iii) question the completeness of communities recovery regarding their functional redundancy.
+Specifically, we (i) questioned the coupling between taxonomic and functional response to disturbance and identified the underlying assembly processes, (ii) clarified the validity of the IDH in the long term for tropical forest and (iii) questioned the completeness of communities recovery regarding communities functional redundancy.
 
@@
 
-The experiment is a network of twelve 6.25ha plots that underwent a gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)).
-Disturbance treatments were attributed according to a randomized plot design with three replicate blocks of four plots. 
+The experiment is a network of twelve 6.25ha plots that underwent a gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) attributed according to a randomized plot design with three replicate blocks of four plots.
 The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3).
@@
 Trees are first identified during inventories with a vernacular name assigned by the field team, and afterward with a scientific name assigned by a botanist during regular botanical campaigns. 
-In 1984, specific vernacular names are given to 62 commercial or common species whereas more infrequent ones were identified under general identifiers only distinguiching trees and palm trees. 
+In 1984, specific vernacular names are given to 62 commercial or common species whereas more infrequent ones were identified under general identifiers only distinguishing trees and palm trees. 
 From 2003 botanical campaigns were conducted every 5 to 6 years to identify all trees at the species level but identification practices still varied among plots and campaigns.
@@
 
-Both communities taxonomic and functional diversity and composition proved resilient, following similar humped shaped trajectories returning towards initial values.
-The resilience of communities functional structure, that is the most direct link between biodiversity and ecosystem functioning [@Diaz2005], meant a consistent recovery of ecosystem processes in the long term [@Guariguata2001].
-The resilience of communities taxonomy meant the maintenance of their initial differences in composition and diversity, suggesting the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003], and the dependency of recovery trajectories on the initial composition to be recovered that would the the asymptot of the humped shaped dynamics [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
+Both communities taxonomic and functional diversities and composition proved resilient, following similar humped shaped trajectories returning towards initial values.
+Communities functional structure is the most direct link between biodiversity and ecosystem functioning [@Diaz2005] and its resilience meant the recovery of ecosystem processes in the long term [@Guariguata2001].
+The resilience of communities taxonomy in turn meant the maintenance of their initial differences in species composition and diversity.
+The maintenance of initial differences suggested the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003] and the dependency of recovery trajectories on the initial composition to be recovered [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
 
-Although both communities taxonomic and functional characteristics proved resilient and followed similar humped shaped trajectories, the taxonomic recovery systematically lagged behind the functional one.
-The decoupling between functional and taxonomic dynamics has already been observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015; @Guariguata2001].
-According to the "vegetation quantity effect" [@Grime1998] functional trajectories rely on the pool of dominant species, which diversity and evenness increased after disturbance but rapidly recovered their initial functional structure.
-Although the pool of dominant species recovered quickly, communities evenness remained high so infrequent species still missed to the taxonomic recovery.
-Taxonomic recovery was then all the more longer that unrecovered species would be functionally redundant and probably underwent selective competition processes.
+Although communities taxonomic and functional trajectories were similar the taxonomic recovery lagged behind the functional one, confirming a decoupling between functional and taxonomic dynamics already observed for grasslands [@Tilman1997; @Mouillot2011] and tropical forests [@Lohbeck2015; @Guariguata2001].
+According to the "vegetation quantity effect" [@Grime1998] the functional trajectories rely on the pool of dominant species, which diversity and evenness increased and rapidly recovered after disturbance.
+At the same time communities taxonomic evenness remained high, revealing the unachieved recovery of infrequent species.
 
@@
 
-The functional and taxonomic decoupling and the disprove of the IDh for communities taxonomic structure were explained were explained by persisting changes in communities functoinal redundancy that implied communities incomplete resilience [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+The functional and taxonomic decoupling and the disprove of the IDH for communities taxonomic structure were explained by persisting rearranging of communities functional redundancy that implied an incomplete resilience of communities [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
-Functional redundancy mitigates the loss of species on communities functional structure [@Carmona2016], and therefore reduces the impact on disturbance.
-The functional redundancy itself however might be reduced or re-organized within the functional space  and alter communities resilience.
-Although the functional redundancy considered in the whole communities functional space did not follow consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
-This restricted redundancy first decreased, all the more so that the disturbance was intense, before returning towards the initial value but this had not recovered for any disturbed community after 30 years.
-After disturbance communities differently occupied the functional space following traits trajectories and the shifts in functional composition, before recovering the initial functional space and its high redundancy.
-This recovery phase was slow as it depend on infrequent, functionally redundant species specific to old-growth forests that rely upon different ecological processes, stochastic or deterministic competitive pressures limiting species functional similarity [@Chave2004;@Mayfield2010].
-
-The unachieved recovery of the functional redundancy meant a decreased resilience of pre-disturbance communities and an increased of disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species).
-After disturbance, chances were higher to observe long lasting or self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013].
-The recovery remained incomplete until the recovery of the initial functional redundancy and resilience, and this specifically threatened the maintenance of species contingent to undisturbed forests and increased the risk of loosing cornerstone species and having unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
+Although the functional redundancy considered in the functional space of the whole communities did not display consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
+After disturbance communities differently occupied the functional space, following the observed shifts in traits trajectories and functional composition, and the redundancy in the initial functional space decreased according to the disturbance intensity.
+After 30 years the recovery of the initial redundancy structure was consistent but remained unachieved.
+It depended on infrequent species specific to old-growth forests that were functionally and therefore underwent competitive exclusion limiting functional similarity [@Chave2004;@Mayfield2010].
+The persistent alteration of functional redundancy meant a lower resilience of pre-disturbance communities and an higher resilience disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species) so chances to observe self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes were increased after disturbance [@Haddad2008; @Burslem2000; @Martin2013].
+Communities resilience depended on the functional redundancy structure that proved long to recover.
+This decades-long recovery specifically increased the risks of loosing cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
 
@@
 
-Our study defined the decoupled functional and taxonomic trajectories of tropical rainforests after disturbancethat demonstrated a rapid functional recovery but a slower and more variable taxonomic one.
-Consistently with the IDH, functional trajectories were driven by deterministic processes favoring pioneers and light-demanding species after disturbance.
-The following  functional shifts entailed a re-organization of communities functional redundancy in the functional space that proved long to recover and explained the more variable and longer taxonomic trajectories as shade tolerant species of old-growth forests faced functional redundancy and limiting similarity processes.
-The resilience of tropical forests then proved long in the face of quite intense disturbance and did not preclude the settling of a persistent disturbanc-specific community [@Gourlet-Fleury2005].
-Within the mechanisms of this resilience the recruitment processes proved central, in accordance with the IDH, and would deserve closer focus to specify their ecological drivers.
+Our study defined the decoupled functional and taxonomic trajectories of tropical rainforests after disturbance demonstrating a rapid functional recovery but a slower and more variable taxonomic one.
+Consistently with the IDH, functional trajectories were driven after disturbance by deterministic processes favoring pioneers and light-demanding species.
+The following functional shifts entailed a re-organization of communities functional redundancy that proved long to recover and explained the more variable and longer taxonomic trajectories as shade tolerant species of old-growth forests faced functional redundancy and limiting similarity processes.
+The resilience of tropical forests then proved long in the face of quite intense disturbance and did not preclude the settling of a persistent disturbance-specific community [@Gourlet-Fleury2005].
+As the trajectories highlighted the recruitment processes proved central for communities response to disturbance and closer focus on demographical drivers of communities response would clarify the fate of the future forests.
 The disturbance range however stayed within the spectrum of selective logging, with a forest cover remaining all along the experiment, and the response mechanisms would probably be much different after harder disturbance.

### 2018-06-18 WholePlotTrajectories.tex
relecture discussion

--- 2018-06-12 WholePlotTrajectories.tex
@@
 functioning and dynamics is urgent to anticipate their fate in the
-global changing context. Communities diversity structure is assumed to
-rely on a constant regime of disturbance and peak at intermediate
-intensity. For tropical forests this intermediate disturbance hypothesis
-(IDH) remain debated as well as the completeness of forests resilience
-in all taxonomic and functoinal facets. To disentangle the ecological
-processes driving forests dynamics we examined the trajectories over 30
-years after a gradient of logging and thinning disturbance, focusing on
-communities taxonomic and functional composition and diversity.
-Specifically we analysed the trajectories of communities taxonomic
-richness and evennes and functional composition, diversity and
-redundancy based on 7 key functional traits assessing species ecology.
-We highlighted a decoupling between functional and taxonomic
-trajectories, translating a fast recovery of communities functioning
-while their taxonomic recovery was slowed by the recruitment of
-infrequent species specific to mature forests. The IDH consistently
-predicted communities functional trajectories but poorly reflected the
-taxonomic ones, blurried by changes in the functional redundancy
-entailing competitive exclusion. Although communities consistently
-recovered their functioning after 30 years, initially infrequent species
-proved long to recover which impacted the taxonomic structure and
-functoinal redundancy over several decades. Our results acknowledged the
-need of decades-long recovery cycles to restore the resilience of the
-community and prevent species loss in the long term.
+global changing context. Communities diversity are assumed to be shaped
+by constant regime of disturbance yielding a diversity peak at
+intermediate intensity, but for tropical forests this intermediate
+disturbance hypothesis (IDH) remains debated as well as the completeness
+their resilience in all taxonomic and functional facets after
+disturbance. To disentangle the ecological processes driving communities
+response to disturbance we analysed the taxonomic and functional
+trajectories following a logging and thinning disturbance gradient.
+Specifically we examined over 30 years the trajectories of communities
+taxonomic richness and evennes and functional composition, diversity and
+redundancy based on 7 leaf, stem and life history traits assessing
+species ecology. Trajectories highlighted communities taxonomic
+resilience, maintaining their initial differences in composition, and
+functional resilience through fast functional trajectory common to all
+communities in the functional space. The IDH consistently predicted
+communities functional trajectories according to the disturbance
+intensity, but poorly represented their taxonomic trajectories, blurried
+by the alteration of the functional redundancy and the recruitment of
+infrequent, shade tolerant species slowed by competitive exclusion
+processes. Although communities functioning had recovered after 30 years
+their taxonomic diversity and composition in infrequent species and
+their functional redundancy remained altered. these results acknowledged
+the need of decades-long recovery cycles to ensure communities complete
+resilience, and questioned communities maintenance after repeated
+disturbance.
 }
@@
 multiple non-timber forest products, shelter a diversified fauna,
-regulate the local climate, support the carbon, water and nutrient
-cycles, and ensure cultural and human well-being. The simultaneous
-increased demand in forests products and the substantial climatic
-changes currently heighten the pressure on remaining forests
-\citep{Gibson2011a, Morales-Hidalgo2015}, threatening the maintenance of
-communities structure, composition and functioning and their dynamics in
-space and time \citep{Anderson-Teixeira2013, Sist2015}.
+regulate the local climate, and the carbon, water and nutrient cycles,
+and ensure cultural and human well-being. The growing demand in forests
+products together with current climatic changes increases the pressure
+on remaining forests \citep{Gibson2011a, Morales-Hidalgo2015}and
+threatens the maintenance and dynamics in space and time of communities
+structure, composition and functioning
+\citep{Anderson-Teixeira2013, Sist2015}.
 
 In tropical forest, ecological communities are constantly re-shaped by
-natural disturbance events that change both abiotic environment, through
-the fluxes of light, heat and water \citep{Goulamoussene2017}, and
-biotic interactions like competition among species \citep{Herault2018}.
-The cornerstone of tropical forests ecology is to understand the
-mechanisms and the determinants of ecosystems response to disturbance
-\citep{White2001, Chazdon2003a}. For now, this has been largely studied
-through structural parameters, rapid and convenient to measure, as
-aboveground biomass, tree height or stem density
+natural disturbance events that change both the abiotic environment,
+through the fluxes of light, heat and water \citep{Goulamoussene2017},
+and the biotic interactions like competition among species
+\citep{Herault2018}. The cornerstone of tropical forests ecology is to
+understand the processes and the determinants of ecosystems response to
+disturbance \citep{White2001, Chazdon2003a}. For now, this has been
+largely studied through structural parameters, rapid and convenient to
+measure, as aboveground biomass, tree height or stem density
 \citep{Piponiot2016, Rutishauser2016}. These structural parameters were
@@
 \citep[\citet{Liang2016}]{Tilman2014} and would be most probably
-impacted by the changes following disturbance \citep{Baraloto2012a}.
+impacted by the changes induced by disturbance \citep{Baraloto2012a}.
 
@@
 \citep{Molino2001, Kariuki2006a, Berry2008a}. Still, validations of the
-IDH remain scarce in the long term and mainly rely on species richness
-analyses that give limited or misleading information on forests recovery
-and functioning \citep{Martin2015, Chaudhary2016}. More releveant
-monitoring would encompass communities composition, crucial for
-conservation issues, and complete diversity profiles integrating both
-species richness and evenness, to reveal the ecological rules
-structuring communities \citep{Magurran1988, Lavorel2002, Bellwood2006}.
-Furthermore, the functional approach that accounts for species
-biological attributes and role in the ecosystem would be essential to
-understand the correlations between ecosystems biodiversity, functioning
-and environmental constraints
+IDH remain scarce in the long term and mainly rely on the analysis of
+species richness that gives limited or misleading information on forests
+recovery and functioning \citep{Martin2015, Chaudhary2016}. More
+complete analysis would besides encompass communities explicit
+composition that is crucial for conservation issues, and communities
+evenness to reveal the ecological rules structuring communities
+\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
+functional approach accoutning for species biological attributes and
+role in the ecosystem directly links ecosystems biodiversity,
+functioning and environmental constraints
 \citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. The
@@
 species ecology and performance was sucessfully adopted
-\citep{Diaz2005, Villeger2008a}, for example in highlighting
-deterministic processes in tropical rainforests that foster fast growing
-species with efficient resources acquisition after disturbance
-\citep{Molino2001, Haddad2008, Ruger2009}. Communities then shifted from
-a dominance of ``conservative'' slow-growing species dealing with scarce
-resources to ``acquisitive'' fast-growing species with rapid and
-efficient use of abundant resources
-\citep{TerSteege2001, Reich2014, Herault2011}. This shift translated
-into consistent trajectories for key functional traits related to
+\citep{Diaz2005, Villeger2008a}, for example to reveal in tropical
+rainforests the deterministic processes that foster after disturbance
+the fast growing species with efficient resources acquisition
+\citep{Molino2001, Haddad2008, Ruger2009}. This basic deterministic
+process entails a functional shift from a dominance of ``conservative''
+slow-growing species dealing with scarce resources to ``acquisitive''
+fast-growing species with rapid and efficient use of abundant resources
+\citep{TerSteege2001, Reich2014, Herault2011}. This shift is translated
+into consistent trajectories of key functional traits related to
 resource acquisition (leaf area, density and chlorophyll content, and
@@
 life history traits (seed mass and maximum height)
-\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}.
+\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Besides
+communities functional composition and diversity, a key aspect of
+communties response to disturbance is the changes in functional
+redundancy that quantifies the amount of shared trait values among
+species \citep{Carmona2016}. High functional redundancy, as in the very
+diverse tropical forests \citep{Bellwood2006}, mitigates the impacts of
+species removal on ecosystem functioning and determines communities
+resilience \citep{Trenbath1999, Elmqvist2003, Diaz2005}. High functional
+redundancy also increases the functional overlap among species and would
+foster neutral stochastic processes \citep{Gravel2006}.
 
-A combination of taxonomic and functional approaches are essential to
-fully assess communities response to disturbance and differences in
-their respective trajectories are insightful of underlying mechanisms
-\citep{Lohbeck2015, Guariguata2001}. Although taxonomic and functional
-diversity are complementary and sometimes decoupled, they are combined
-in the measure of communities functional redundancy that quantifies the
-amount of shared trait values among species \citep{Carmona2016}.
-Functional redundancy is determinant of communities resilience as high
-redundancy, like in highly diverse tropical forests
-\citep{Bellwood2006}, mitigates the impacts of species removal on
-ecosystem functioning \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
-
-Grasp all facets of communities response to disturbance comes to examine
-the taxonomic and functional trajectories in diversity and composition.
-These trajectories would highlight the ecological rules constraining or
-not communities dynamics towards the recovery of initial composition,
-diversity and functioning. They would therefore clarify the tenants in
-the long term of the debated Intermediate Disturbance Hypothesis for
-tropical forests, a cruicial point for future adaptive conservation
-strategies \citep{Adler2007}. Here we monitored over 30 years the
-response of 75 ha of forests plots set up on a gradient of disturbance
-intensity, from 10 to 60\% of ecosystem biomass removed. We made use of
-a large functional traits database browsing major leaf, stem and seed
-functional traits and species maximum height to draw the trajectories
-over time of communities taxonomic and functional composition and
-diversity. Specifically, we (i) questioned the coupling between
-taxonomic and functional response to disturbance and identified the
-underlying assembly processes, which allowed to (ii) clarify the
-validity of the IDH in the long term for tropical forest and (iii)
-question the completeness of communities recovery regarding their
-functional redundancy.
+To grasp all facets of communities response to disturbance we examined
+here the taxonomic and functional trajectories in diversity and
+composition \citep{Lohbeck2015, Guariguata2001}. These trajectories
+would highlight the recovery of communities initial composition,
+diversity and functioning and the underlying ecological processes. They
+would clarify the tenants in the long term of the Intermediate
+Disturbance Hypothesis debated in tropical forests, and provide
+indications for future adaptive conservation strategies
+\citep{Adler2007}. We monitored over 30 years the response of 75 ha of
+neotropical forest plots set up on a gradient of disturbance intensity,
+from 10 to 60\% of ecosystem biomass removed. We made use of a large
+functional traits database browsing major leaf, stem and seed functional
+traits and species maximum height to draw the trajectories over time of
+communities taxonomic and functional composition and diversity.
+Specifically, we (i) questioned the coupling between taxonomic and
+functional response to disturbance and identified the underlying
+assembly processes, (ii) clarified the validity of the IDH in the long
+term for tropical forest and (iii) questioned the completeness of
+communities recovery regarding communities functional redundancy.
 
@@
 gradient of three logging, thinning and fuelwood cutting treatments
-(Table \ref{tab:Tab1}). Disturbance treatments were attributed according
-to a randomized plot design with three replicate blocks of four plots.
-The disturbance corresponds to averages of 10 trees removed per hectare
-with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1),
-32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above
-40 cm DBH for treatment 3 (T3). Treatments T2 and T3 besides included
-the thinning of trees by poison girdling \citep{Schmitt1989, Blanc2009}.
-The disturbance intensity was measured as the percentage of aboveground
-biomass (\%AGB) lost between the first inventory in 1984 and five years
-after disturbance (ref to be found) measured with the BIOMASS R package
+(Table \ref{tab:Tab1}) attributed according to a randomized plot design
+with three replicate blocks of four plots. The disturbance corresponds
+to averages of 10 trees removed per hectare with a diameter at 1.3 m
+height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm
+DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3
+(T3). Treatments T2 and T3 besides included the thinning of trees by
+poison girdling \citep{Schmitt1989, Blanc2009}. The disturbance
+intensity was measured as the percentage of aboveground biomass (\%AGB)
+lost between the first inventory in 1984 and five years after
+disturbance (ref to be found) measured with the BIOMASS R package
 \citep{Biomass2018}.
@@
 more infrequent ones were identified under general identifiers only
-distinguiching trees and palm trees. From 2003 botanical campaigns were
+distinguishing trees and palm trees. From 2003 botanical campaigns were
 conducted every 5 to 6 years to identify all trees at the species level
@@
 
-Both communities taxonomic and functional diversity and composition
+Both communities taxonomic and functional diversities and composition
 proved resilient, following similar humped shaped trajectories returning
-towards initial values. The resilience of communities functional
-structure, that is the most direct link between biodiversity and
-ecosystem functioning \citep{Diaz2005}, meant a consistent recovery of
-ecosystem processes in the long term \citep{Guariguata2001}. The
-resilience of communities taxonomy meant the maintenance of their
-initial differences in composition and diversity, suggesting the
-existence of multiple stable equilibria as assumed for highly diverse
-and productive ecosystems \citep{Chase2003}, and the dependency of
-recovery trajectories on the initial composition to be recovered that
-would the the asymptot of the humped shaped dynamics
+towards initial values. Communities functional structure is the most
+direct link between biodiversity and ecosystem functioning
+\citep{Diaz2005} and its resilience meant the recovery of ecosystem
+processes in the long term \citep{Guariguata2001}. The resilience of
+communities taxonomy in turn meant the maintenance of their initial
+differences in species composition and diversity. The maintenance of
+initial differences suggested the existence of multiple stable
+equilibria as assumed for highly diverse and productive ecosystems
+\citep{Chase2003} and the dependency of recovery trajectories on the
+initial composition to be recovered
 \citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.
 
-Although both communities taxonomic and functional characteristics
-proved resilient and followed similar humped shaped trajectories, the
-taxonomic recovery systematically lagged behind the functional one. The
-decoupling between functional and taxonomic dynamics has already been
-observed for grasslands \citep{Tilman1997, Mouillot2011} and more
-recently for tropical forests \citep{Lohbeck2015, Guariguata2001}.
-According to the ``vegetation quantity effect'' \citep{Grime1998}
-functional trajectories rely on the pool of dominant species, which
-diversity and evenness increased after disturbance but rapidly recovered
-their initial functional structure. Although the pool of dominant
-species recovered quickly, communities evenness remained high so
-infrequent species still missed to the taxonomic recovery. Taxonomic
-recovery was then all the more longer that unrecovered species would be
-functionally redundant and probably underwent selective competition
-processes.
+Although communities taxonomic and functional trajectories were similar
+the taxonomic recovery lagged behind the functional one, confirming a
+decoupling between functional and taxonomic dynamics already observed
+for grasslands \citep{Tilman1997, Mouillot2011} and tropical forests
+\citep{Lohbeck2015, Guariguata2001}. According to the ``vegetation
+quantity effect'' \citep{Grime1998} the functional trajectories rely on
+the pool of dominant species, which diversity and evenness increased and
+rapidly recovered after disturbance. At the same time communities
+taxonomic evenness remained high, revealing the unachieved recovery of
+infrequent species.
 
@@
 
-The functional and taxonomic decoupling and the disprove of the IDh for
-communities taxonomic structure were explained were explained by
-persisting changes in communities functoinal redundancy that implied
-communities incomplete resilience
-\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+The functional and taxonomic decoupling and the disprove of the IDH for
+communities taxonomic structure were explained by persisting rearranging
+of communities functional redundancy that implied an incomplete
+resilience of communities \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-Functional redundancy mitigates the loss of species on communities
-functional structure \citep{Carmona2016}, and therefore reduces the
-impact on disturbance. The functional redundancy itself however might be
-reduced or re-organized within the functional space and alter
-communities resilience. Although the functional redundancy considered in
-the whole communities functional space did not follow consistent
-trajectories after disturbance, the redundancy restricted to the
-functional space of the initial community clearly followed humped shaped
-trajectories. This restricted redundancy first decreased, all the more
-so that the disturbance was intense, before returning towards the
-initial value but this had not recovered for any disturbed community
-after 30 years. After disturbance communities differently occupied the
-functional space following traits trajectories and the shifts in
-functional composition, before recovering the initial functional space
-and its high redundancy. This recovery phase was slow as it depend on
-infrequent, functionally redundant species specific to old-growth
-forests that rely upon different ecological processes, stochastic or
-deterministic competitive pressures limiting species functional
-similarity \citep{Chave2004, Mayfield2010}.
-
-The unachieved recovery of the functional redundancy meant a decreased
-resilience of pre-disturbance communities and an increased of
-disturbance-specific communities (corresponding to lower richness and
-higher dominance of pioneer species). After disturbance, chances were
-higher to observe long lasting or self-maintained compositional changes
-in favor of disturbance resistant species, lianas or epiphytes
-\citep{Haddad2008, Burslem2000, Martin2013}. The recovery remained
-incomplete until the recovery of the initial functional redundancy and
-resilience, and this specifically threatened the maintenance of species
-contingent to undisturbed forests and increased the risk of loosing
-cornerstone species and having unexpected ecological consequences
+Although the functional redundancy considered in the functional space of
+the whole communities did not display consistent trajectories after
+disturbance, the redundancy restricted to the functional space of the
+initial community clearly followed humped shaped trajectories. After
+disturbance communities differently occupied the functional space,
+following the observed shifts in traits trajectories and functional
+composition, and the redundancy in the initial functional space
+decreased according to the disturbance intensity. After 30 years the
+recovery of the initial redundancy structure was consistent but remained
+unachieved. It depended on infrequent species specific to old-growth
+forests that were functionally and therefore underwent competitive
+exclusion limiting functional similarity
+\citep{Chave2004, Mayfield2010}. The persistent alteration of functional
+redundancy meant a lower resilience of pre-disturbance communities and
+an higher resilience disturbance-specific communities (corresponding to
+lower richness and higher dominance of pioneer species) so chances to
+observe self-maintained compositional changes in favor of disturbance
+resistant species, lianas or epiphytes were increased after disturbance
+\citep{Haddad2008, Burslem2000, Martin2013}. Communities resilience
+depended on the functional redundancy structure that proved long to
+recover. This decades-long recovery specifically increased the risks of
+loosing cornerstone specie, with unexpected ecological consequences
 \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.
@@
 Our study defined the decoupled functional and taxonomic trajectories of
-tropical rainforests after disturbancethat demonstrated a rapid
-functional recovery but a slower and more variable taxonomic one.
-Consistently with the IDH, functional trajectories were driven by
-deterministic processes favoring pioneers and light-demanding species
-after disturbance. The following functional shifts entailed a
-re-organization of communities functional redundancy in the functional
-space that proved long to recover and explained the more variable and
-longer taxonomic trajectories as shade tolerant species of old-growth
-forests faced functional redundancy and limiting similarity processes.
-The resilience of tropical forests then proved long in the face of quite
-intense disturbance and did not preclude the settling of a persistent
-disturbanc-specific community \citep{Gourlet-Fleury2005}. Within the
-mechanisms of this resilience the recruitment processes proved central,
-in accordance with the IDH, and would deserve closer focus to specify
-their ecological drivers. The disturbance range however stayed within
-the spectrum of selective logging, with a forest cover remaining all
-along the experiment, and the response mechanisms would probably be much
-different after harder disturbance.
+tropical rainforests after disturbance demonstrating a rapid functional
+recovery but a slower and more variable taxonomic one. Consistently with
+the IDH, functional trajectories were driven after disturbance by
+deterministic processes favoring pioneers and light-demanding species.
+The following functional shifts entailed a re-organization of
+communities functional redundancy that proved long to recover and
+explained the more variable and longer taxonomic trajectories as shade
+tolerant species of old-growth forests faced functional redundancy and
+limiting similarity processes. The resilience of tropical forests then
+proved long in the face of quite intense disturbance and did not
+preclude the settling of a persistent disturbance-specific community
+\citep{Gourlet-Fleury2005}. As the trajectories highlighted the
+recruitment processes proved central for communities response to
+disturbance and closer focus on demographical drivers of communities
+response would clarify the fate of the future forests. The disturbance
+range however stayed within the spectrum of selective logging, with a
+forest cover remaining all along the experiment, and the response
+mechanisms would probably be much different after harder disturbance.
 

### 2018-06-19 WholePlotTrajectories.Rmd
relecture

@@
 
-```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWD), and bark thickness, (ref:capBarkThick)) and one life trait (Specific maximum height at adult stage, (ref:capHmax)).
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWD), and bark thickness, (ref:capBarkThick)) and one life history trait (Specific maximum height at adult stage, (ref:capHmax)).
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values.
 Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
@@
 
-Communities functional redundancy was measured as the sum within communities the species weighted functional overlap based on the 7 leaf, stem, and maximum height traits (see appendix I for PCA details).
+Communities functional redundancy was measured as the sum within communities of species functional overlap weighted by their abundance.
+The functional redundancy was measured in the functional space defined by  7 leaf, stem, and maximum height traits (see appendix I for PCA details). 
 Communities functional redundancy remained stable in control plots but after disturbance the redundancy trajectories were quite variable (See appendix I) and apparently independently of the initial disturbance.
@@
 
-```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial cpùùunities functional space over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial communities functional space over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values.
 Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure"}
@@
 
-The examination of taxonomic and functional trajectories suggested that communities functional structure and dominant species pool relied upon deterministic processes following the IDH while the pool of rare species relied upon more stochastic processes.
+The taxonomic and functional trajectories validated the IDH as determinant of communities functional structure and of the dominant species pool, while contrastingly the pool of rare species relied upon more stochastic processes.
 
-The disturbance intensity did not predict the taxonomic richness, as already observed in the Paracou station [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998], nor did it strongly correlate with the taxonomic evenness (*i.e.* Simpson diversity) although the it was enhanced after disturbance.
+The disturbance intensity poorly predicted the taxonomic richness, as already observed in the Paracou station [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998], was only weakly correlated with the increase in taxonomic evenness (*i.e.* Simpson diversity).
 
-Contrastingly, the disturbance intensity predicted the increase in functional diversity along the 30 years monitored after disturbance.
-The increase in functional diversity paralleled the important specific turnover and significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
-As the old-growth, pre-disturbance survivor trees proved to mirror initial communities [@Herault2018] the functional trajectories and taxonomic turnover rather relied on the pool of newly recruited trees driven by the enhanced growth and survival of previously infrequent species and functional types.
-Disturbance entailed a reorganization of the typical high dominance structure of hyperdiverse mature forests benefiting to pioneers and light demanding species: the changes in abiotic environment and competitive pressure favored pioneers which outcompete other species in non limiting resources but were excluded in mature forests by long-lived, more resistant and shade tolerant species.
-Disturbance trajectories therefore relied on the environmental niches made available and filled by species from a restricted functional range that became dominant and structured the functionalcharacteristics of the community [@Grime1998].
-The intermediate disturbance hypothesis predicted communities functional diversity and pool of dominant species [@Molino2001], but rare species relied on different processes and disprove the IDh for the whole taxonomic community [@Hubbell2001;@Chave2004].
+Contrastingly, the disturbance intensity consistently predicted the increase in functional diversity after disturbance along the 30 years.
+The functional diversity increase paralleled the significant specific turnover and functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
+As the old-growth, pre-disturbance survivor trees proved mirroring the initial communities [@Herault2018] the functional trajectories and taxonomic turnover rather relied upon the pool of newly recruited trees driven by the enhanced growth and survival of previously infrequent species and functional types.
+Disturbance entailed a reorganization, benefiting to pioneers and light demanding species, of the typical high dominance structure of hyperdiverse mature forests.
+The changes in abiotic environment and competitive pressure favored pioneers which outcompeted other species in non limiting light resources but were excluded in mature forests by long-lived, more resistant and shade tolerant species.
+Disturbance trajectories therefore relied upon the environmental niches made available and filled by species from a restricted functional range that became dominant and structured the functional characteristics of the community [@Grime1998].
+The intermediate disturbance hypothesis was suitabe to predict communities functional diversity and pool of dominant species [@Molino2001], but it poorly represented the whole taxonomic structure that was hampered by the slow recovery of rare species [@Hubbell2001;@Chave2004].
 
@@
 
-The functional and taxonomic decoupling and the disprove of the IDH for communities taxonomic structure were explained by persisting rearranging of communities functional redundancy that implied an incomplete resilience of communities [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+The functional and taxonomic decoupling and the rebuttal of the IDH for communities taxonomic structure were explained by a reorganization of the functional redundancy within the functional traits space that implied the incomplete resilience of communities [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 

### 2018-06-19 Scripts/Stats.R

@@ -108,8 +108,49 @@ mean(as.numeric(TimeMax))-1984
 
 ##########################################
 ## Rho Spearman, redundancy
-load("DB/Redundancy_restricted")
-RedundancyTraj_restricted
+
+smooth<-function(mat,larg){return(do.call(cbind,lapply(1:ncol(mat),function(step){
+  range<-max(1,step-larg):min(ncol(mat),step+larg)
+  rowSums(mat[,range])/length(range)})))}
+treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
+names(treatments)<-c("Control","T1","T2","T3")
+ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
+colyear<-c("darkgoldenrod1","darkorange2","darkred")
+time<-c("1995","2005","2015")
+
+load("DB/Redundancy_sd0.7")
+RedundancyTraj_restricted->Red
+
+RedundancyPlot<-function(Red){
+  Red<-lapply(Red,function(rep){
+  ret<-smooth(rep,2)
+  #ret<-rep
+  colnames(ret)<-colnames(rep)
+  ret<-apply(ret[,which(as.numeric(colnames(ret))>="1989")],2,function(col){return(col-ret[,"1989"])})#
+  colnames(ret)<-as.numeric(colnames(ret))-1984
+  return(ret)})
+
+Red<-array(unlist(Red),dim=c(12,ncol(Red[[1]]),length(Red)),
+           dimnames=list(1:12,colnames(Red[[1]]),1:length(Red)))
+Red<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(Red,c(1,2),function(rep){
+  return(quantile(rep,probs=quant))}))})
+Red<-array(unlist(Red),dim=c(12,ncol(Red[[1]]),3),
+           dimnames=list(1:12,colnames(Red[[1]]),c(0.025,0.5,0.975)))
+
+
+plot(colnames(Red),Red[1,,1],type='n',ylim=c(min(Red),max(Red)),xlab="",ylab="")
+  
+  invisible(lapply(1:4,function(tr){
+    toplot<-Red[which(rownames(Red)%in%treatments[[tr]]),,]
+    apply(toplot,1,function(li){
+      lines(colnames(Red),li[,"0.5"],col=ColorsTr[tr],lwd=2)
+      polygon(c(colnames(Red),rev(colnames(Red))),
[…]

### 2018-06-19 WholePlotTrajectories.Rmd
citatoin check

@@
   \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
-  \hyphenation{bio-di-ver-si-ty sap-lings re-or-ga-ni-za-tion}
+  \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion}
 bibliography: references.bib
@@
 ```{r Options, echo=FALSE, message=FALSE, warning=F}
-knitr::opts_chunk$set(cache=T, warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
+knitr::opts_chunk$set(cache=F, warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
 # Installation des packages si nécessaire et chargement
@@
 The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3).
-Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Schmitt1989;@Blanc2009]. 
+Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Schmitt1990;@Blanc2009]. 
 The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance (ref to be found) measured with the BIOMASS R package [@Biomass2018].
@@
 The uncertainty propagation was done through a Bayesian framework reconstituting complete inventories at genus level from real incomplete ones on the basis of vernacular/botanical names association. 
-Vernacular names were replaced through multinomial trials $M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)$ based on the association probability $\big[\alpha_1, \alpha_2,…, \alpha_3\big]$ observed across all inventories between each vernacular name *v* and the species $\big[s_1, s_2, …, s_N\big]$.
+Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,…, \alpha_3\big]$ observed across all inventories between each vernacular name *v* and the species $\big[s_1, s_2, …, s_N\big]$:
+
+\begin{align}
+M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big) \nonumber
+\end{align}
+
 See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.
@@
 
-## Communities composition
+## Communities Composition
 
@@
 Considering the functional redundancy restricted to the functional space of the initial inventory, all disturbed plots followed similar decreasing humped shaped trajectories (\@ref(fig:RedFun_rest)).
-The maximum redundancy loss was positively correlated with the disturbance intensity (XX spearman to be measured) and the initial value had not recovered for any disturbed communities.
+The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{spearman}=0.50$) and the initial value had not recovered for any disturbed communities.
 
-```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial communities functional space over 30 years after disturbance.
+```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the difference from initial state of functional redundancy within the initial functional space over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values.

### 2018-06-19 WholePlotTrajectories.tex
citatoin check

@@
 \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
-\hyphenation{bio-di-ver-si-ty sap-lings}
+\hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion}
 
@@
 (T3). Treatments T2 and T3 besides included the thinning of trees by
-poison girdling \citep{Schmitt1989, Blanc2009}. The disturbance
+poison girdling \citep{Schmitt1990, Blanc2009}. The disturbance
 intensity was measured as the percentage of aboveground biomass (\%AGB)
@@
 vernacular/botanical names association. Vernacular names were replaced
-through multinomial trials
-\(M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)\)
-based on the association probability
+through multinomial trials based on the association probability
 \(\big[\alpha_1, \alpha_2,…, \alpha_3\big]\) observed across all
 inventories between each vernacular name \emph{v} and the species
-\(\big[s_1, s_2, …, s_N\big]\). See appendix 1 and
-\citet{Aubry-Kientz2013} for the detailed methodology.
+\(\big[s_1, s_2, …, s_N\big]\):
+
+\begin{align}
+M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big) \nonumber
+\end{align}
+
+See appendix 1 and \citet{Aubry-Kientz2013} for the detailed
+methodology.
 
@@
 
-\subsection{Communities composition}\label{communities-composition}
+\subsection{Communities Composition}\label{communities-composition}
 
@@
 
-\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life trait (Specific maximum height at adult stage, \emph{Hmax}).
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}).
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values.
 Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:CWM}
@@
 Communities functional redundancy was measured as the sum within
-communities the species weighted functional overlap based on the 7 leaf,
-stem, and maximum height traits (see appendix I for PCA details).
-Communities functional redundancy remained stable in control plots but
-after disturbance the redundancy trajectories were quite variable (See
-appendix I) and apparently independently of the initial disturbance.
-Globally after most intense disturbance (plots T2 and T3) communities
-redundancy decreased at first place before increasing to edge, recover
-or exceed the initial value.
+communities of species functional overlap weighted by their abundance.
+The functional redundancy was measured in the functional space defined
+by 7 leaf, stem, and maximum height traits (see appendix I for PCA
+details). Communities functional redundancy remained stable in control
+plots but after disturbance the redundancy trajectories were quite
+variable (See appendix I) and apparently independently of the initial
+disturbance. Globally after most intense disturbance (plots T2 and T3)
+communities redundancy decreased at first place before increasing to
+edge, recover or exceed the initial value.
 
@@
 maximum redundancy loss was positively correlated with the disturbance
-intensity (XX spearman to be measured) and the initial value had not
+intensity (\(\rho_{spearman}=0.50\)) and the initial value had not
 recovered for any disturbed communities.
@@
 
-\caption{Trajectories of the functional redundancy within the initial cpùùunities functional space over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
+\caption{Trajectories of the difference from initial state of functional redundancy within the initial functional space over 30 years after disturbance.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values.
 Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}(\#fig:RedFun_rest)
@@
 
-The examination of taxonomic and functional trajectories suggested that
-communities functional structure and dominant species pool relied upon
-deterministic processes following the IDH while the pool of rare species
-relied upon more stochastic processes.
+The taxonomic and functional trajectories validated the IDH as
+determinant of communities functional structure and of the dominant
+species pool, while contrastingly the pool of rare species relied upon
+more stochastic processes.
 
-The disturbance intensity did not predict the taxonomic richness, as
+The disturbance intensity poorly predicted the taxonomic richness, as
 already observed in the Paracou station \citep{Baraloto2012a} and in
-Bornean tropical forests \citep{Cannon1998}, nor did it strongly
-correlate with the taxonomic evenness (\emph{i.e.} Simpson diversity)
-although the it was enhanced after disturbance.
+Bornean tropical forests \citep{Cannon1998}, was only weakly correlated
+with the increase in taxonomic evenness (\emph{i.e.} Simpson diversity).
 
-Contrastingly, the disturbance intensity predicted the increase in
-functional diversity along the 30 years monitored after disturbance. The
-increase in functional diversity paralleled the important specific
-turnover and significant functional shifts towards resource-acquisitive
-strategies (sharp increase in the SLA, leaf thickness and bark thickness
-and decrease in wood density, leaf toughness and maximum height)
+Contrastingly, the disturbance intensity consistently predicted the
+increase in functional diversity after disturbance along the 30 years.
+The functional diversity increase paralleled the significant specific
+turnover and functional shifts towards resource-acquisitive strategies
+(sharp increase in the SLA, leaf thickness and bark thickness and
+decrease in wood density, leaf toughness and maximum height)
 \citep{Westoby1998, Wright2004, Reich2014}. As the old-growth,
-pre-disturbance survivor trees proved to mirror initial communities
+pre-disturbance survivor trees proved mirroring the initial communities
 \citep{Herault2018} the functional trajectories and taxonomic turnover
-rather relied on the pool of newly recruited trees driven by the
+rather relied upon the pool of newly recruited trees driven by the
 enhanced growth and survival of previously infrequent species and
-functional types. Disturbance entailed a reorganization of the typical
-high dominance structure of hyperdiverse mature forests benefiting to
-pioneers and light demanding species: the changes in abiotic environment
-and competitive pressure favored pioneers which outcompete other species
-in non limiting resources but were excluded in mature forests by
-long-lived, more resistant and shade tolerant species. Disturbance
-trajectories therefore relied on the environmental niches made available
-and filled by species from a restricted functional range that became
-dominant and structured the functionalcharacteristics of the community
-\citep{Grime1998}. The intermediate disturbance hypothesis predicted
-communities functional diversity and pool of dominant species
-\citep{Molino2001}, but rare species relied on different processes and
-disprove the IDh for the whole taxonomic community
-\citep{Hubbell2001, Chave2004}.
+functional types. Disturbance entailed a reorganization, benefiting to
+pioneers and light demanding species, of the typical high dominance
+structure of hyperdiverse mature forests. The changes in abiotic
+environment and competitive pressure favored pioneers which outcompeted
+other species in non limiting light resources but were excluded in
+mature forests by long-lived, more resistant and shade tolerant species.
+Disturbance trajectories therefore relied upon the environmental niches
+made available and filled by species from a restricted functional range
+that became dominant and structured the functional characteristics of
+the community \citep{Grime1998}. The intermediate disturbance hypothesis
+was suitabe to predict communities functional diversity and pool of
+dominant species \citep{Molino2001}, but it poorly represented the whole
+taxonomic structure that was hampered by the slow recovery of rare
+species \citep{Hubbell2001, Chave2004}.
 
@@
 
-The functional and taxonomic decoupling and the disprove of the IDH for
-communities taxonomic structure were explained by persisting rearranging
-of communities functional redundancy that implied an incomplete
-resilience of communities \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+The functional and taxonomic decoupling and the rebuttal of the IDH for
+communities taxonomic structure were explained by a reorganization of
+the functional redundancy within the functional traits space that
+implied the incomplete resilience of communities
+\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 

### 2018-06-19 Scripts/Stats.R

@@ -88,6 +88,7 @@ DistT<-as.data.frame(unlist(lapply(1:12,function(pl){
     return(ret2)}))
   ret<-apply(ret,2,median)
   return(names(ret)[which(ret==max(ret))])})))
+
 treats<-cbind(c(1,6,11,2,7,9,3,5,10,4,8,12),rep(0:3,each=3))
 DistT<-cbind(DistT,rownames(DistT),treats[order(treats[,1]),2])
 colnames(DistT)<-c("Max","Plot","treat")
@@ -118,39 +119,18 @@ ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
 colyear<-c("darkgoldenrod1","darkorange2","darkred")
 time<-c("1995","2005","2015")
 
-load("DB/Redundancy_sd0.7")
+load("DB/Redundancy_restricted_sd0.7")
 RedundancyTraj_restricted->Red
 
-RedundancyPlot<-function(Red){
-  Red<-lapply(Red,function(rep){
-  ret<-smooth(rep,2)
-  #ret<-rep
-  colnames(ret)<-colnames(rep)
-  ret<-apply(ret[,which(as.numeric(colnames(ret))>="1989")],2,function(col){return(col-ret[,"1989"])})#
-  colnames(ret)<-as.numeric(colnames(ret))-1984
-  return(ret)})
+Maxred<-unlist(lapply(1:12,function(pl){
+  ret<-do.call(rbind,lapply(Red,function(iter){return(iter[pl,])}))
+  return(max(apply(ret,2,median)))
+}))
 
-Red<-array(unlist(Red),dim=c(12,ncol(Red[[1]]),length(Red)),
-           dimnames=list(1:12,colnames(Red[[1]]),1:length(Red)))
-Red<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(Red,c(1,2),function(rep){
-  return(quantile(rep,probs=quant))}))})
-Red<-array(unlist(Red),dim=c(12,ncol(Red[[1]]),3),
-           dimnames=list(1:12,colnames(Red[[1]]),c(0.025,0.5,0.975)))
-
-
-plot(colnames(Red),Red[1,,1],type='n',ylim=c(min(Red),max(Red)),xlab="",ylab="")
-  
-  invisible(lapply(1:4,function(tr){
-    toplot<-Red[which(rownames(Red)%in%treatments[[tr]]),,]
[…]

### 2018-06-22 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

@@
 \Abstract{
-Understand the maintenance of forests biodiversity structure,
+Understanding the ecological rules underlying the maintenance of tropical forest biodiversity, structure,
 functioning and dynamics is urgent to anticipate their fate in the

### 2018-06-22 WholePlotTrajectories.Rmd
manuscript

@@
 
-In tropical forest, ecological communities are constantly re-shaped by natural disturbance events that change both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions like competition among species [@Herault2018].
+In tropical forest, ecological communities are constantly re-shaped by natural disturbance events that change both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions like competition among species [@Chesson2000;@Herault2018].
 The cornerstone of tropical forests ecology is to understand the processes and the determinants of ecosystems response to disturbance [@White2001;@Chazdon2003a]. 

### 2018-06-22 WholePlotTrajectories.Rmd
abstract

@@
 abstract: >   
-    Understand the maintenance of forests biodiversity structure, functioning and dynamics is urgent to anticipate their fate in the global changing context.
+    Understand the ecological rules underlying the maintenance of tropical forests biodiversity, structure, functioning and dynamics is urgent to anticipate their fate in the global changing context.
     Communities diversity are assumed to be shaped by constant regime of disturbance yielding a diversity peak at intermediate intensity, but for tropical forests this intermediate disturbance hypothesis (IDH) remains debated as well as the completeness their resilience in all taxonomic and functional facets after disturbance.

### 2018-06-22 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

--- 2018-06-19 WholePlotTrajectories.tex
@@
 \Abstract{
-Understand the maintenance of forests biodiversity structure,
+Understanding the ecological rules underlying the maintenance of tropical forest
 functioning and dynamics is urgent to anticipate their fate in the
-global changing context. Communities diversity are assumed to be shaped
+global changing context. The huge diversity if tropical forests is often assumed to be shaped
 by constant regime of disturbance yielding a diversity peak at
 intermediate intensity, but for tropical forests this intermediate
-disturbance hypothesis (IDH) remains debated as well as the completeness
+disturbance hypothesis (IDH) remains debated ????as well as the completeness
 their resilience in all taxonomic and functional facets after
-disturbance. To disentangle the ecological processes driving communities
+disturbance????. To disentangle the ecological processes driving communities
 response to disturbance we analysed the taxonomic and functional
-trajectories following a logging and thinning disturbance gradient.
-Specifically we examined over 30 years the trajectories of communities
+diversity trajectories following a logging and thinning disturbance gradient.
+Specifically we examined over 30 years the trajectories of
 taxonomic richness and evennes and functional composition, diversity and
-redundancy based on 7 leaf, stem and life history traits assessing
-species ecology. Trajectories highlighted communities taxonomic
+redundancy based on 7 leaf, stem and life history traits. Trajectories highlighted communities taxonomic
 resilience, maintaining their initial differences in composition, and
@@
 products together with current climatic changes increases the pressure
-on remaining forests \citep{Gibson2011a, Morales-Hidalgo2015}and
+on remaining forests \citep{Gibson2011a, Morales-Hidalgo2015} and
 threatens the maintenance and dynamics in space and time of communities
-structure, composition and functioning
-\citep{Anderson-Teixeira2013, Sist2015}.
+structure, composition and functioning \citep{Anderson-Teixeira2013, Sist2015}.
 
@@
 and the biotic interactions like competition among species
-\citep{Herault2018}. The cornerstone of tropical forests ecology is to
-understand the processes and the determinants of ecosystems response to
+\citep{Herault2018}. One of the cornerstone of tropical forest ecology is to
+understand the processes and drivers of ecosystems response to
 disturbance \citep{White2001, Chazdon2003a}. For now, this has been
-largely studied through structural parameters, rapid and convenient to
+largely studied through forest structural parameters, rapid and convenient to
 measure, as aboveground biomass, tree height or stem density
@@
 thereafter sucessfully modeled, giving important insights into the
-maintenance of ecosystems processes and services
-\citep{Denslow2000, Blanc2009}. However the response of forests
-diversity in tree species remains unclear, albeit it determines the
+recovery of ecosystems processes and services
+\citep{Herault2018}. However the response of forests
+diversity is still unclear, albeit it determines the
 productivity, stability and functioning of ecosystems
@@
 
-In the short-term disturbance demonstrated negligible or even positive
+In the short-term, moderate disturbance may lead to positive
 impacts on communities diversity, which have been formalized by the
@@
 recovery and functioning \citep{Martin2015, Chaudhary2016}. More
-complete analysis would besides encompass communities explicit
-composition that is crucial for conservation issues, and communities
-evenness to reveal the ecological rules structuring communities
+complete analysis would besides encompass communities
+composition that is crucial for conservation issues, and abundance distribution
+to reveal the ecological rules structuring communities
 \citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
-functional approach accoutning for species biological attributes and
+functional approach accounting for species biological attributes and
 role in the ecosystem directly links ecosystems biodiversity,
@@
 \citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Besides
-communities functional composition and diversity, a key aspect of
-communties response to disturbance is the changes in functional
+communities functional composition and diversity, a last key aspect of
+communities response to disturbance is the changes in functional
 redundancy that quantifies the amount of shared trait values among

### 2018-06-23 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

@@
 lost between the first inventory in 1984 and five years after
-disturbance (ref to be found) measured with the BIOMASS R package
+disturbance \citep{Piponiot2016} measured with the BIOMASS R package
 \citep{Biomass2018}.

### 2018-06-23 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

--- 2018-06-22 WholePlotTrajectories.tex
@@
 gradient of three logging, thinning and fuelwood cutting treatments
-(Table \ref{tab:Tab1}) attributed according to a randomized plot design
+(Table \ref{tab:Tab1}) according to a randomized plot design
 with three replicate blocks of four plots. The disturbance corresponds
@@
 lost between the first inventory in 1984 and five years after
-disturbance (ref to be found) measured with the BIOMASS R package
+disturbance \citep{Piponiot2016} estimated with the BIOMASS R package
 \citep{Biomass2018}.
@@
 
-The study site corresponds to a tropical rainforest with a dominance of
+The study site corresponds to a tropical rainforest typical if the Guiana Shield with a dominance of
 Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical
 families. In the twelve experimental plots of the experiment, all trees
-above 10 cm DBH are mapped and measured annually since 1984. Trees are
+above 10 cm DBH have been mapped and measured annually since 1984. Trees are
 first identified during inventories with a vernacular name assigned by
-the field team, and afterward with a scientific name assigned by a
-botanist during regular botanical campaigns. In 1984, specific
+the forest worker team, and afterward with a scientific name assigned by
+botanists during regular botanical campaigns. In 1984, specific
 vernacular names are given to 62 commercial or common species whereas
 more infrequent ones were identified under general identifiers only
-distinguishing trees and palm trees. From 2003 botanical campaigns were
+distinguishing trees and palm trees. From 2003, botanical campaigns have been
 conducted every 5 to 6 years to identify all trees at the species level
-but identification practices still varied among plots and campaigns.
+but identification levels still varied among plots and campaigns.
 
@@
 
-The functional approach used a dataset of 6 functional traits
+Six functional traits
 representing leaf economics (leaves thickness, toughness, total
@@
 mass) and wood economics (wood specific gravity and bark thickness), and
-life history traits (maximum specific height and seed mass). The trait
-database came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}
-where trait values were assessed from a selection of individuals located
+life history traits (maximum specific height and seed mass) came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}. Trait values were assessed from a selection of individuals located
 in nine permanent plots in French Guiana, including two in Paracou, and
@@
 Missing trait values were filled using multivariate imputation by
-chained equation (mice) from the mice R package \citep{Mice2011}.
-Imputations were restricted within genus, or family when samples wre too
+chained equation \citep{Mice2011}.
+Imputations were restricted within genus, or family when samples were too
 scarce, in order to account for the phylogenetic signal of the
@@
 
-All composition and diversity metrics corresponded to the average
+All composition and diversity metrics were
 obtained after 50 iterations of the taxonomic uncertainty propagation
@@
 To counter taxonomic uncertainties due to the variability of botanical
-identification protocols (see {[}\#protocols{]}), the taxonomic
+identification levels (in space) and protocols (in time) (see {[}\#protocols{]}), the taxonomic
 composition and diversity analysis were conducted at the genus level,

### 2018-06-23 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

@@
 in composition after disturbance were followed in a two-dimensional
-ordination space the 30 years monitored. Two NMDS were conducted to map
-either taxonomic flora inventories or communities functional composition
-based on the 7 leaf, stem and life history traits (without seed mass
-classes). In both cases the NMDS were performed using occurrence-based
+NMDS ordination space the 30 years monitored. In both cases the NMDS were performed using occurrence-based
 (Jaccard) and abundance-based (Bray-Curtis) dissimilarity measures.
@@
 distance of successive inventories to the reference inventories in 1989,
-5 years after disturbance, when the uncertainty degree did not exceed
-30\% of undetermined trees. The trajectories of the leaf and stem and
-life traits were also visualized with the community weighted means
-(CWM), representing the average trait value in a community weighted by
-relative abundance of the species carrying each value
-\citep{Diaz2007, Garnier2004}. To compensate the intrinsinc difference
-among plots the trajectories corresponded to the differences along time
-with the reference inventory in 1989. Species seed mass corresponded to
+i.e. 2-3 years after disturbance, when the uncertainty degree did not exceed
+30\% of undetermined trees. To compensate the intrinsinc difference
+of initial position among plots, the trajectories corresponded to the euclidian differences along time
+with the reference position in 1989. Univariate trajectories of the leaf, stem and
+life-history traits were inspected with the community weighted means
+(CWM)
+\citep{Diaz2007, Garnier2004}. Species seed mass corresponded to
 5 classes of increasing mass, seed mass trajectories were therefore
@@
 respectively corresponding to the 0, 1 and 2 order of diversity (q),
-recomended for diversity studies \citep{Patil1982, Tothmeresz1995}. The
+recomended for diversity studies \citep{Patil1982, Tothmeresz1995, Marcon2015}. The
 functional diversity was reported using the Rao index of quadratic
@@
 The functional redundancy was measured as the overlap among species in
-communities' functional space \citep{Carmona2016}. The samples of the
+community functional space \citep{Carmona2016}. The samples of the
 trait database were first mapped in a 2-dimensional plan from a PCA

### 2018-06-23 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

@@
 
-In the inventories from 1989 (5 years after disturbance) to 2015 (31
-years after disturbance), 828388 trees and 591 botanical species
+From 1989 (2-3 years after disturbance) to 2015 (28-29
+years after disturbance), 828388 individual trees and 591 botanical species
 pertaining to 223 genus and 64 botanical families were recorded.
-Communities taxonomic diversity trajectories were examined through the
-Richness, Shannon and Simpson diversities at genus level, in relation to
-the 1989 inventories (5 years after disturbance) (See annexe I). For
-undisturbed plots the Richness, Shannon and Simpson diversities remained
+For undisturbed plots Richness, Shannon and Simpson diversities remained
 stable over the 30 years monitored. In disturbed communities the
@@
 followed unimodal trajectories after intense disturbance, decreasing for
-ten years before recovering pre-disturbance values. In all disturbed
+ten years before recovering to pre-disturbance values. In all disturbed
 plots the taxonomic evenness (Shannon and Simpson diversities)
@@
 giving similar results only analysis using Bray-Curtis dissimilarity are
-discussed.
+presented.
 
 While both taxonomic and functional composition remained stable in
-undistrubed communities, they followed consistent trajectories over time
+undisturbed communities, they followed consistent trajectories over time
 after disturbance which revealed significant compositional changes.
@@
 The changes observed in plots functional composition went hand to hand
-with consistent trajectories of the 8 functional and life history traits
+with consistent trajectories of the 8 functional traits
 visualized with the trajectories of community weighted means (CWM) of
@@
 
-Communities functional redundancy was measured as the sum within
-communities of species functional overlap weighted by their abundance.
-The functional redundancy was measured in the functional space defined
-by 7 leaf, stem, and maximum height traits (see appendix I for PCA
-details). Communities functional redundancy remained stable in control
+Communities functional redundancy remained stable in control
 plots but after disturbance the redundancy trajectories were quite

### 2018-06-24 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

@@
 pertaining to 223 genus and 64 botanical families were recorded.
-For undisturbed plots Richness, Shannon and Simpson diversities remained
+For undisturbed plots Richness and taxonomic evenness (Shannon and Simpson diversities) remained
 stable over the 30 years monitored. In disturbed communities the
@@
 
-\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.
-Lower panels, Relationship between the initial \%AGB removed and the median values of Simpson \textbf{(c)} and Rao \textbf{(d)} diversities at three times after disturbance.
-Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.}\label{fig:IDHplot}
+\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance.
+Colors are treaments: green (control), blue (T1), orange (T2), red (T3) with sahded areas the credibility intervals.
+Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities, 10, 20 and 30 years after disturbance}\label{fig:IDHplot}
 \end{figure*}
@@
 
-The trajectories of taxonomic and functional composition were visualized
-in a two dimensional ordination space mapping the successive inventories
-according to their flora and corresponding traits. Classifications were
-performed using either abundance-based Bray-Curtis (Figure
-\ref{fig:NMDSplans}) or incidence-based Jaccard dissimilarity, both
-giving similar results only analysis using Bray-Curtis dissimilarity are
-presented.
-
 While both taxonomic and functional composition remained stable in
-undisturbed communities, they followed consistent trajectories over time
-after disturbance which revealed significant compositional changes.
-According to the mapping of functional traits (see appendix I) these
+undisturbed communities (ref to fig), they followed consistent trajectories over time
+after disturbance. According to the mapping of functional traits (see appendix I) these
 compositional changes corresponded to shifts towards species with more
@@
 
-\caption{Trajectories of the plots in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) regarding the 6 leaf and stem functional traits, the maximum allometric height and seed mass class.
-Plots trajectories are first represented in the two-dimensional space from the NMDS performed for the 30 years after disturbance based on Bray-Curtis dissimilarity measures between successive inventories (Upper panels \textbf{(a)} and \textbf{(b)}).
-Then the lower panels (\textbf{(c)} and \textbf{(d)}) represent the euclidean distance to initial condition along the 30 sampled years.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.}\label{fig:NMDSplans}
+\caption{Plot trajectories in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) in a the two-dimensional NMDS space.
+Lower panels (\textbf{(c)} and \textbf{(d)}) represent the euclidean distance to initial condition along the 30 sampled years.
+Colors are disturbance treatments (green for control, blue for T1,orange for T2 and red for T3) with shaded areas, the credibility intervals}\label{fig:NMDSplans}
 \end{figure*}
@@
 
-The changes observed in plots functional composition went hand to hand
-with consistent trajectories of the 8 functional traits
-visualized with the trajectories of community weighted means (CWM) of
-leaves economics (leaves thickness, chlorophyll content, toughness and
-specific area), wood economics (wood specific gravity, bark thickness),
-and life history traits (seed mass and maximum adult height) (Figure
-\ref{fig:CWM}).
-
-Except for leaf chlorophyll content, which continued to increase for
+Changes in functional composition trajectories went hand to hand
+with consistent trajectories of the 8 functional traits (Figure
+\ref{fig:CWM}). Except for leaf chlorophyll content, which continued to increase for
 some T3 and T2 plots 30 years after disturbance, all traits and seed
@@
 \caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}).
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:CWM}
+Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
 \end{figure*}
@@
 communities redundancy decreased at first place before increasing to
-edge, recover or exceed the initial value.
+edge, recover or exceed the initial value. *I would suggest to consider only the Functional redundancy restricted to the functional space in the paper*
 
@@
 
-\caption{Trajectories of the difference from initial state of functional redundancy within the initial functional space over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}(\#fig:RedFun_rest)
+\caption{Trajectories of functional redundancy within the initial functional space over 30 years after disturbance.
+Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}(\#fig:RedFun_rest)
 \end{figure}

### 2018-06-24 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

@@
 Both communities taxonomic and functional diversities and composition
-proved resilient, following similar humped shaped trajectories returning
-towards initial values. Communities functional structure is the most
+proved resilient, following similar humped-shaped trajectories starting to return
+towards initial values 30 years after disturbance. Communities functional signature is the most
 direct link between biodiversity and ecosystem functioning
@@
 processes in the long term \citep{Guariguata2001}. The resilience of
-communities taxonomy in turn meant the maintenance of their initial
-differences in species composition and diversity. The maintenance of
-initial differences suggested the existence of multiple stable
+communities taxonomy, despite the multiple NMDS starting points, suggest the long-term maintenance of the initial
+differences in species composition and diversity, in turn suggesting the existence of multiple stable
 equilibria as assumed for highly diverse and productive ecosystems

### 2018-06-24 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

@@
 
-The taxonomic and functional trajectories validated the IDH as
-determinant of communities functional structure and of the dominant
-species pool, while contrastingly the pool of rare species relied upon
-more stochastic processes.
-
-The disturbance intensity poorly predicted the taxonomic richness, as
-already observed in the Paracou station \citep{Baraloto2012a} and in
-Bornean tropical forests \citep{Cannon1998}, was only weakly correlated
-with the increase in taxonomic evenness (\emph{i.e.} Simpson diversity).
-
-Contrastingly, the disturbance intensity consistently predicted the
+Disturbance intensity poorly predicted the taxonomic richness regardless of the period studied, as
+already observed in the Guiana Shield \citep{Baraloto2012a} and in
+Bornean tropical forests \citep{Cannon1998}. Contrastingly, the disturbance intensity consistently predicted the
 increase in functional diversity after disturbance along the 30 years.
-The functional diversity increase paralleled the significant specific
-turnover and functional shifts towards resource-acquisitive strategies
+The functional diversity increase paralleled the significant functional shifts towards resource-acquisitive strategies
 (sharp increase in the SLA, leaf thickness and bark thickness and
@@
 mature forests by long-lived, more resistant and shade tolerant species.
+
+* Not clear from here*
 Disturbance trajectories therefore relied upon the environmental niches

### 2018-06-24 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

@@
 
-The functional and taxonomic decoupling and the rebuttal of the IDH for
+The functional and taxonomic decoupling as well as the rebuttal of the IDH for
 communities taxonomic structure were explained by a reorganization of
 the functional redundancy within the functional traits space that
-implied the incomplete resilience of communities
+implied the incomplete resilience (???) of communities
 \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-Although the functional redundancy considered in the functional space of
-the whole communities did not display consistent trajectories after
-disturbance, the redundancy restricted to the functional space of the
+The functional redundancy restricted to the functional space of the
 initial community clearly followed humped shaped trajectories. After
@@
 recovery of the initial redundancy structure was consistent but remained
-unachieved. It depended on infrequent species specific to old-growth
+unachieved. ?????-> It depended on infrequent species specific to old-growth
 forests that were functionally and therefore underwent competitive
 exclusion limiting functional similarity
-\citep{Chave2004, Mayfield2010}. The persistent alteration of functional
+\citep{Chave2004, Mayfield2010}. ?????->The persistent alteration of functional
 redundancy meant a lower resilience of pre-disturbance communities and
@@
 resistant species, lianas or epiphytes were increased after disturbance
-\citep{Haddad2008, Burslem2000, Martin2013}. Communities resilience
+\citep{Haddad2008, Burslem2000, Martin2013}. ?? Not so long no? -> Communities resilience
 depended on the functional redundancy structure that proved long to
 recover. This decades-long recovery specifically increased the risks of
-loosing cornerstone specie, with unexpected ecological consequences
+loosing cornerstone (????) species, with unexpected ecological consequences
 \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.
@@
 the IDH, functional trajectories were driven after disturbance by
-deterministic processes favoring pioneers and light-demanding species.
-The following functional shifts entailed a re-organization of
+enrichment of the system with pioneers and light-demanding species.
+????-> The ????following functional shifts entailed a re-organization of
 communities functional redundancy that proved long to recover and
@@
 disturbance and closer focus on demographical drivers of communities
-response would clarify the fate of the future forests. The disturbance
+response would clarify the fate of the future forests. ????-> The disturbance
 range however stayed within the spectrum of selective logging, with a

### 2018-06-25 WholePlotTrajectories.Rmd
relecture

@@
 They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local climate, and the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The growing demand in forests products together with current climatic changes increases the pressure on remaining forests [@Gibson2011a; @Morales-Hidalgo2015]and threatens the maintenance and dynamics in space and time of communities structure, composition and functioning [@Anderson-Teixeira2013;@Sist2015].
+The growing demand in forests products together with current climatic changes increases the pressure on remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threatens the maintenance and dynamics in space and time of communities structure, composition and functioning [@Anderson-Teixeira2013;@Sist2015].
 

### 2018-06-26 WholePlotTrajectories.Rmd
comments bubs

@@
 abstract: >   
-    Understand the ecological rules underlying the maintenance of tropical forests biodiversity, structure, functioning and dynamics is urgent to anticipate their fate in the global changing context.
-    Communities diversity are assumed to be shaped by constant regime of disturbance yielding a diversity peak at intermediate intensity, but for tropical forests this intermediate disturbance hypothesis (IDH) remains debated as well as the completeness their resilience in all taxonomic and functional facets after disturbance.
+    Understanding the ecological rules underlying the maintenance of tropical forests biodiversity, structure, functioning and dynamics is urgent to anticipate their fate in the global changing context.
+    The huge diversity of tropical forests is often assumed to be shaped by constant regime of disturbance yielding a diversity peak at intermediate intensity, but for tropical forests this intermediate disturbance hypothesis (IDH) remains debated, and this controversy also questions the extent of communities resilience regarding their functional and taxonomic facets is also to be clarified.
 
-    To disentangle the ecological processes driving communities response to disturbance we analysed the taxonomic and functional trajectories following a logging and thinning disturbance gradient.
-    Specifically we examined over 30 years the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 leaf, stem and life history traits assessing species ecology.
+    To disentangle the ecological processes driving communities response to disturbance we analysed the taxonomic and functional diversity trajectories following a logging and thinning disturbance gradient.
+    Specifically we examined over 30 years the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 leaf, stem and life history traits.
     Trajectories highlighted communities taxonomic resilience, maintaining their initial differences in composition, and functional resilience through fast functional trajectory common to all communities in the functional space.
@@
 In tropical forest, ecological communities are constantly re-shaped by natural disturbance events that change both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions like competition among species [@Chesson2000;@Herault2018].
-The cornerstone of tropical forests ecology is to understand the processes and the determinants of ecosystems response to disturbance [@White2001;@Chazdon2003a]. 
-For now, this has been largely studied through structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
-These structural parameters were thereafter sucessfully modeled, giving important insights into the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009]. 
-However the response of forests diversity in tree species remains unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014, @Liang2016] and would be most probably impacted by the changes induced by disturbance [@Baraloto2012a].
+One of the cornerstone of tropical forest ecology is to understand the processes and drivers of ecosystems response to disturbance [@White2001;@Chazdon2003a]. 
+For now, this has been largely studied through forest structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
+These structural parameters were thereafter sucessfully modeled, giving important insights into the recovery of ecosystems processes and services [@Herault2018]. 
+However the response of forests diversity is still unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014, @Liang2016] and would be most probably impacted by the changes induced by disturbance [@Baraloto2012a].
 
-In the short-term disturbance demonstrated negligible or even positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
+In the short-term, moderate disturbance may lead to positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
 Still, validations of the IDH remain scarce in the long term and mainly rely on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-More complete analysis would besides encompass communities explicit composition that is crucial for conservation issues, and communities evenness to reveal the ecological rules structuring communities [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, a functional approach accoutning for species biological attributes and role in the ecosystem directly links ecosystems biodiversity, functioning and environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+More complete analysis would besides encompass communities composition that is crucial for conservation issues, and abundance distribution to reveal the ecological rules structuring communities [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Furthermore, a functional approach accounting for species biological attributes and role in the ecosystem directly links ecosystems biodiversity, functioning and environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
 The functional trait-based approach focused on major traits related to species ecology and performance was sucessfully adopted [@Diaz2005;@Villeger2008a], for example to reveal in tropical rainforests the deterministic processes that foster after disturbance the fast growing species with efficient resources acquisition [@Molino2001;@Haddad2008;@Ruger2009].
@@
 This shift is translated into consistent trajectories of key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
-Besides communities functional composition and diversity, a key aspect of communties response to disturbance is the changes in functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
+Besides communities functional composition and diversity, a last key aspect of communities response to disturbance is the changes in functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
 High functional redundancy, as in the very diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning and determines communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005].
@@
 
-The experiment is a network of twelve 6.25ha plots that underwent a gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) attributed according to a randomized plot design with three replicate blocks of four plots.
+The experiment is a network of twelve 6.25ha plots that underwent a gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) according to a randomized plot design with three replicate blocks of four plots.
 The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3).
 Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Schmitt1990;@Blanc2009]. 
-The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance (ref to be found) measured with the BIOMASS R package [@Biomass2018].
-
-```{r Tab1, echo=FALSE,is.as=T}
-Intervention <- data.frame(
-  Treatment=c("Control", "T1", "T2", "T3"),
-  Timber=c("", 
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha", 
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha",
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha"),
-  Thinning=c("", 
-             "", 
-             "DBH $\\geq$ 40 cm, non-valuable species, $\\approx$ 30 trees/ha", 
-             "DBH $\\geq$ 50 cm, non-valuable species, $\\approx$ 15 trees/ha  "),
-  Fuelwood=c("", 
-             "", 
-             "", 
-             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-valuable species, $\\approx$ 15 trees/ha"),
-  AGBLost=c(0, "$[12\\%-33\\%]$", "$[33\\%-56\\%]$", "$[35\\%-66\\%]$")
-)
-names(Intervention)[5] <- "\\%AGB lost"
-knitr::kable(Intervention, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.", longtable = FALSE, booktabs = TRUE, escape = FALSE) %>%
-kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
-```
+The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Biomass2018].
 
@@
 
-The study site corresponds to a tropical rainforest with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical families. 
-In the twelve experimental plots of the experiment, all trees above 10 cm DBH are mapped and measured annually since 1984. 
-Trees are first identified during inventories with a vernacular name assigned by the field team, and afterward with a scientific name assigned by a botanist during regular botanical campaigns. 
+The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical families. 
+In the twelve experimental plots of the experiment, all trees above 10 cm DBH have been mapped and measured annually since 1984. 
+Trees are first identified during inventories with a vernacular name assigned by the forest worker team, and afterward with a scientific name assigned by botanists during regular botanical campaigns. 
 In 1984, specific vernacular names are given to 62 commercial or common species whereas more infrequent ones were identified under general identifiers only distinguishing trees and palm trees. 
-From 2003 botanical campaigns were conducted every 5 to 6 years to identify all trees at the species level but identification practices still varied among plots and campaigns.
+From 2003, botanical campaigns have been conducted every 5 to 6 years to identify all trees at the species level but identification levels still varied among plots and campaigns.
 
@@
 
-The functional approach used a dataset of 6 functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, the leaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass).
-The trait database came from the BRIDGE project [^1] where trait values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 botanical species pertaining to 157 botanical genera.
-Missing trait values were filled using multivariate imputation by chained equation (mice) from the mice R package [@Mice2011].
-Imputations were restricted within genus, or family when samples wre too scarce, in order to account for the phylogenetic signal of the functional traits. 
+Six functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, the leaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass) came from the BRIDGE project [^1].
+Traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 botanical species pertaining to 157 botanical genera.
+Missing trait values were filled using multivariate imputation by chained equation [@Mice2011].
+Imputations were restricted within genus, or family when samples were too scarce, in order to account for the phylogenetic signal of the functional traits. 
 Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level (same genus or family). 
@@
 
-All composition and diversity metrics corresponded to the average obtained after 50 iterations of the taxonomic uncertainty propagation framework and of the filling process of missing trait values.
+All composition and diversity metrics were obtained after 50 iterations of the taxonomic uncertainty propagation framework and of the filling process of missing trait values.
 
@@
 
-To counter taxonomic uncertainties due to the variability of botanical identification protocols (see [#protocols]), the taxonomic composition and diversity analysis were conducted at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
-Trajectories of communities taxonomic and functional variations in composition after disturbance were followed in a two-dimensional ordination space the 30 years monitored.
+To counter taxonomic uncertainties due to the variability of botanical identification levels (in space) and protocols (in time), the taxonomic composition and diversity analysis were conducted at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
+Trajectories of communities taxonomic and functional variations in composition after disturbance were followed in a two-dimensional NMDS ordination space of the 30 years monitored.
 Two NMDS were conducted to map either taxonomic flora inventories or communities functional composition based on the 7 leaf, stem and life history traits (without seed mass classes).
 In both cases the NMDS were performed using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) dissimilarity measures. 
-Trajectories along time in the plan were reported through the euclidean distance of successive inventories to the reference inventories in 1989, 5 years after disturbance, when the uncertainty degree did not exceed 30% of undetermined trees.
-The trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004].
-To compensate the intrinsinc difference among plots the trajectories corresponded to the differences along time with the reference inventory in 1989.
+Trajectories along time in the plan were reported through the euclidean distance of successive inventories to the reference inventories in 1989, *i.e.* 2-3 years after disturbance, when the uncertainty degree did not exceed 30% of undetermined trees.
+To compensate the intrinsinc difference among plots, the trajectories corresponded to the euclidean differences along time with the reference inventory in 1989.
+Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007;@Garnier2004]. 
 Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were therefore reported as the proportion of each class in the inventories.
@@
 The taxonomic diversity was assessed through Richness and the Hill number translation of Shannon and Simpson indices [@Hill1973].
-These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995].
+These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995;@Marcon2015].
 The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on all functional and life traits.
@@
 
-The functional redundancy was measured as the overlap among species in communities' functional space [@Carmona2016]. 
+The functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
 The samples of the trait database were first mapped in a 2-dimensional plan from a PCA analysis. 
@@
 
-In the inventories from 1989 (5 years after disturbance) to 2015 (31 years after disturbance), 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
-Communities taxonomic diversity trajectories were examined through the Richness, Shannon and Simpson diversities at genus level, in relation to the 1989 inventories (5 years after disturbance) (See annexe I).
-For undisturbed plots the Richness, Shannon and Simpson diversities remained stable over the 30 years monitored.
-In disturbed communities the taxonomic richness increased after low disturbance intensity, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2) while it followed unimodal trajectories after intense disturbance, decreasing for ten years before recovering pre-disturbance values.
+From 1989 (2-3 years after disturbance) to 2015 (28-29 years after disturbance), 828388 individual trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
+For undisturbed plots Richness and taxonomic evenness (Shannon and Simpson diversities) remained stable over the 30 years monitored.
+In disturbed communities the taxonomic richness increased after low disturbance intensity, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2) while it followed unimodal trajectories after intense disturbance, decreasing for ten years before recovering to pre-disturbance values.
 In all disturbed plots the taxonomic evenness (Shannon and Simpson diversities) increased, following unimodal trajectories with a maximum, reached after around 20 years, positively correlated to the disturbance treatment ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
@@
 
-```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Upper panels, Trajectories of the Simpson taxonomic diversity (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance, corresponding to the median and 0.025 and 0.975 percentile observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.
-Lower panels, Relationship between the initial \\%AGB removed and the median values of Simpson (ref:Boldc) and Rao (ref:Boldd) diversities at three times after disturbance.
-Solid lines colors represent the time, 10 years (yellow), 20 years (orange) and 30 years (brown) after disturbance.", fig.env="figure*"}
+```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Upper panels, Trajectories of the Simpson taxonomic diversity (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.
+Lower panels, Relationship between the initial \\%AGB removed and Simpson (ref:Boldc) and Rao (ref:Boldd) diversities 10, 20 and 30 years after disturbance.", fig.env="figure*"}
 load("DB/SimpsonDraw");load("DB/FunctionalTraj_ForGraphs")
@@
 
-The trajectories of taxonomic and functional composition were visualized in a two dimensional ordination space mapping the successive inventories according to their flora and corresponding traits.
-Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity, both giving similar results only analysis using Bray-Curtis dissimilarity are discussed.
-
-While both taxonomic and functional composition remained stable in undistrubed communities, they followed consistent trajectories over time after disturbance which revealed significant compositional changes.
+While both taxonomic and functional composition remained stable in undisturbed communities  (Figure \@ref(fig:NMDSplans)), they followed consistent trajectories over time after disturbance which revealed significant compositional changes.
 According to the mapping of functional traits (see appendix I) these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WD to high average SLA and chlorophyll content.
@@
 
-```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the plots in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) regarding the 6 leaf and stem functional traits, the maximum allometric height and seed mass class.
-Plots trajectories are first represented in the two-dimensional space from the NMDS performed for the 30 years after disturbance based on Bray-Curtis dissimilarity measures between successive inventories (Upper panels (ref:Bolda) and (ref:Boldb)).
-Then the lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years.
-Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3).
-The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.", fig.env="figure*"}
+```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS space.
+Lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals", fig.env="figure*"}
 load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
@@
 
-The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits visualized with the trajectories of community weighted means (CWM) of leaves economics (leaves thickness, chlorophyll content, toughness and specific area), wood economics (wood specific gravity, bark thickness), and life history traits (seed mass and maximum adult height) (Figure \@ref(fig:CWM)).
-
+Changes in functional composition trajectories went hand to hand with consistent trajectories of the 8 functional traits (Figure \@ref(fig:CWM)).
 Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions followed unimodal trajectories either stabilizing or returning towards their initial values.
@@
 ```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWD), and bark thickness, (ref:capBarkThick)) and one life history trait (Specific maximum height at adult stage, (ref:capHmax)).
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure*"}
 load("DB/CWM")
@@
 
-Communities functional redundancy was measured as the sum within communities of species functional overlap weighted by their abundance.
-The functional redundancy was measured in the functional space defined by  7 leaf, stem, and maximum height traits (see appendix I for PCA details). 
-Communities functional redundancy remained stable in control plots but after disturbance the redundancy trajectories were quite variable (See appendix I) and apparently independently of the initial disturbance.
+Communities functional remained stable in control plots but after disturbance the redundancy trajectories were quite variable (See appendix I) and apparently independently of the initial disturbance.
 Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place before increasing to edge, recover or exceed the initial value.
+> restrict to initial functional space
 
@@
 
-```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the difference from initial state of functional redundancy within the initial functional space over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iterations of the taxonomic uncertainty propagation and the filling processes of missing trait values.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure"}
+```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="TTrajectories of functional redundancy within the initial functional space over 30 years after disturbance.
+Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
 load("DB/Redundancy_restricted")
@@
 
-Both communities taxonomic and functional diversities and composition proved resilient, following similar humped shaped trajectories returning towards initial values.
-Communities functional structure is the most direct link between biodiversity and ecosystem functioning [@Diaz2005] and its resilience meant the recovery of ecosystem processes in the long term [@Guariguata2001].
-The resilience of communities taxonomy in turn meant the maintenance of their initial differences in species composition and diversity.
-The maintenance of initial differences suggested the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003] and the dependency of recovery trajectories on the initial composition to be recovered [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
+Both communities taxonomic and functional diversities and composition proved resilient, following similar humped-shaped trajectories starting to return towards initial values 30 years after disturbance.
+Communities functional signature is the most direct link between biodiversity and ecosystem functioning [@Diaz2005] and its resilience meant the recovery of ecosystem processes in the long term [@Guariguata2001].
+The resilience of communities taxonomy, despite the multiple NMDS starting points, suggest the long term maintenance of the initial differences in species composition and diversity.
+This maintenance in turn suggest the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003] and the dependency of recovery trajectories on the initial composition to be recovered [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
 
@@
 
-The taxonomic and functional trajectories validated the IDH as determinant of communities functional structure and of the dominant species pool, while contrastingly the pool of rare species relied upon more stochastic processes.
-
-The disturbance intensity poorly predicted the taxonomic richness, as already observed in the Paracou station [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998], was only weakly correlated with the increase in taxonomic evenness (*i.e.* Simpson diversity).
-
+Disturbance intensity poorly predicted the taxonomic richness regadless of the period studied, as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
 Contrastingly, the disturbance intensity consistently predicted the increase in functional diversity after disturbance along the 30 years.
-The functional diversity increase paralleled the significant specific turnover and functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
+The functional diversity increase paralleled the significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
 As the old-growth, pre-disturbance survivor trees proved mirroring the initial communities [@Herault2018] the functional trajectories and taxonomic turnover rather relied upon the pool of newly recruited trees driven by the enhanced growth and survival of previously infrequent species and functional types.
@@
 The changes in abiotic environment and competitive pressure favored pioneers which outcompeted other species in non limiting light resources but were excluded in mature forests by long-lived, more resistant and shade tolerant species.
+
+>> Not clear
 Disturbance trajectories therefore relied upon the environmental niches made available and filled by species from a restricted functional range that became dominant and structured the functional characteristics of the community [@Grime1998].
@@
 
-The functional and taxonomic decoupling and the rebuttal of the IDH for communities taxonomic structure were explained by a reorganization of the functional redundancy within the functional traits space that implied the incomplete resilience of communities [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+The functional and taxonomic decoupling as well as the rebuttal of the IDH for communities taxonomic structure were explained by a reorganization of the functional redundancy within the functional traits space that implied the
+>> incomplete resilience of communities [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
-Although the functional redundancy considered in the functional space of the whole communities did not display consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
+The functional redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
 After disturbance communities differently occupied the functional space, following the observed shifts in traits trajectories and functional composition, and the redundancy in the initial functional space decreased according to the disturbance intensity.
 After 30 years the recovery of the initial redundancy structure was consistent but remained unachieved.
+>> unclear
 It depended on infrequent species specific to old-growth forests that were functionally and therefore underwent competitive exclusion limiting functional similarity [@Chave2004;@Mayfield2010].
+
+>>unclear
 The persistent alteration of functional redundancy meant a lower resilience of pre-disturbance communities and an higher resilience disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species) so chances to observe self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes were increased after disturbance [@Haddad2008; @Burslem2000; @Martin2013].
+
+>>not so long
 Communities resilience depended on the functional redundancy structure that proved long to recover.
+
+>>unclear (cornerstone?)
 This decades-long recovery specifically increased the risks of loosing cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
@@
 Our study defined the decoupled functional and taxonomic trajectories of tropical rainforests after disturbance demonstrating a rapid functional recovery but a slower and more variable taxonomic one.
-Consistently with the IDH, functional trajectories were driven after disturbance by deterministic processes favoring pioneers and light-demanding species.
+Consistently with the IDH, functional trajectories were driven after disturbance by enrichment of the system with pioneers and light-demanding species.
+
+>>unclear
 The following functional shifts entailed a re-organization of communities functional redundancy that proved long to recover and explained the more variable and longer taxonomic trajectories as shade tolerant species of old-growth forests faced functional redundancy and limiting similarity processes.
+
 The resilience of tropical forests then proved long in the face of quite intense disturbance and did not preclude the settling of a persistent disturbance-specific community [@Gourlet-Fleury2005].
 As the trajectories highlighted the recruitment processes proved central for communities response to disturbance and closer focus on demographical drivers of communities response would clarify the fate of the future forests.
+
+>>unclear
 The disturbance range however stayed within the spectrum of selective logging, with a forest cover remaining all along the experiment, and the response mechanisms would probably be much different after harder disturbance.

### 2018-06-26 WholePlotTrajectories.tex
comments bubs

@@
 \Abstract{
-Understanding the ecological rules underlying the maintenance of tropical forest
-functioning and dynamics is urgent to anticipate their fate in the
-global changing context. The huge diversity if tropical forests is often assumed to be shaped
-by constant regime of disturbance yielding a diversity peak at
-intermediate intensity, but for tropical forests this intermediate
-disturbance hypothesis (IDH) remains debated ????as well as the completeness
-their resilience in all taxonomic and functional facets after
-disturbance????. To disentangle the ecological processes driving communities
-response to disturbance we analysed the taxonomic and functional
-diversity trajectories following a logging and thinning disturbance gradient.
-Specifically we examined over 30 years the trajectories of
-taxonomic richness and evennes and functional composition, diversity and
-redundancy based on 7 leaf, stem and life history traits. Trajectories highlighted communities taxonomic
+Understanding the ecological rules underlying the maintenance of
+tropical forests biodiversity, structure, functioning and dynamics is
+urgent to anticipate their fate in the global changing context. The huge
+diversity of tropical forests is often assumed to be shaped by constant
+regime of disturbance yielding a diversity peak at intermediate
+intensity, but for tropical forests this intermediate disturbance
+hypothesis (IDH) remains debated, and this controversy also questions
+the extent of communities resilience regarding their functional and
+taxonomic facets is also to be clarified. To disentangle the ecological
+processes driving communities response to disturbance we analysed the
+taxonomic and functional diversity trajectories following a logging and
+thinning disturbance gradient. Specifically we examined over 30 years
+the trajectories of communities taxonomic richness and evennes and
+functional composition, diversity and redundancy based on 7 leaf, stem
+and life history traits. Trajectories highlighted communities taxonomic
 resilience, maintaining their initial differences in composition, and
@@
 threatens the maintenance and dynamics in space and time of communities
-structure, composition and functioning \citep{Anderson-Teixeira2013, Sist2015}.
+structure, composition and functioning
+\citep{Anderson-Teixeira2013, Sist2015}.
 
@@
 and the biotic interactions like competition among species
-\citep{Herault2018}. One of the cornerstone of tropical forest ecology is to
-understand the processes and drivers of ecosystems response to
-disturbance \citep{White2001, Chazdon2003a}. For now, this has been
-largely studied through forest structural parameters, rapid and convenient to
-measure, as aboveground biomass, tree height or stem density
-\citep{Piponiot2016, Rutishauser2016}. These structural parameters were
-thereafter sucessfully modeled, giving important insights into the
-recovery of ecosystems processes and services
-\citep{Herault2018}. However the response of forests
-diversity is still unclear, albeit it determines the
-productivity, stability and functioning of ecosystems
-\citep[\citet{Liang2016}]{Tilman2014} and would be most probably
-impacted by the changes induced by disturbance \citep{Baraloto2012a}.
+\citep{Chesson2000, Herault2018}. One of the cornerstone of tropical
+forest ecology is to understand the processes and drivers of ecosystems
+response to disturbance \citep{White2001, Chazdon2003a}. For now, this
+has been largely studied through forest structural parameters, rapid and
+convenient to measure, as aboveground biomass, tree height or stem
+density \citep{Piponiot2016, Rutishauser2016}. These structural
+parameters were thereafter sucessfully modeled, giving important
+insights into the recovery of ecosystems processes and services
+\citep{Herault2018}. However the response of forests diversity is still
+unclear, albeit it determines the productivity, stability and
+functioning of ecosystems \citep[\citet{Liang2016}]{Tilman2014} and
+would be most probably impacted by the changes induced by disturbance
+\citep{Baraloto2012a}.
 
-In the short-term, moderate disturbance may lead to positive
-impacts on communities diversity, which have been formalized by the
-intermediate disturbance hypothesis (IDH) stating a maximized species
-diversity at intermediate disturbance intensity
+In the short-term, moderate disturbance may lead to positive impacts on
+communities diversity, which have been formalized by the intermediate
+disturbance hypothesis (IDH) stating a maximized species diversity at
+intermediate disturbance intensity
 \citep{Molino2001, Kariuki2006a, Berry2008a}. Still, validations of the
@@
 recovery and functioning \citep{Martin2015, Chaudhary2016}. More
-complete analysis would besides encompass communities
-composition that is crucial for conservation issues, and abundance distribution
-to reveal the ecological rules structuring communities
+complete analysis would besides encompass communities composition that
+is crucial for conservation issues, and abundance distribution to reveal
+the ecological rules structuring communities
 \citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
@@
 gradient of three logging, thinning and fuelwood cutting treatments
-(Table \ref{tab:Tab1}) according to a randomized plot design
-with three replicate blocks of four plots. The disturbance corresponds
-to averages of 10 trees removed per hectare with a diameter at 1.3 m
-height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm
-DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3
-(T3). Treatments T2 and T3 besides included the thinning of trees by
-poison girdling \citep{Schmitt1990, Blanc2009}. The disturbance
-intensity was measured as the percentage of aboveground biomass (\%AGB)
-lost between the first inventory in 1984 and five years after
-disturbance \citep{Piponiot2016} estimated with the BIOMASS R package
+(Table \ref{tab:Tab1}) according to a randomized plot design with three
+replicate blocks of four plots. The disturbance corresponds to averages
+of 10 trees removed per hectare with a diameter at 1.3 m height (DBH)
+above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for
+treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3).
+Treatments T2 and T3 besides included the thinning of trees by poison
+girdling \citep{Schmitt1990, Blanc2009}. The disturbance intensity was
+measured as the percentage of aboveground biomass (\%AGB) lost between
+the first inventory in 1984 and five years after disturbance
+\citep{Piponiot2016} estimated with the BIOMASS R package
 \citep{Biomass2018}.
-
-\begin{table}
-
-\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.}
-\centering
-\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
-\toprule
-Treatment & Timber & Thinning & Fuelwood & \%AGB lost\\
-\midrule
-Control &  &  &  & 0\\
-T1 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha &  &  & $[12\%-33\%]$\\
-T2 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30 trees/ha &  & $[33\%-56\%]$\\
-T3 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & $[35\%-66\%]$\\
-\bottomrule
-\end{tabu}
-\end{table}
 
@@
 
-The study site corresponds to a tropical rainforest typical if the Guiana Shield with a dominance of
-Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical
-families. In the twelve experimental plots of the experiment, all trees
-above 10 cm DBH have been mapped and measured annually since 1984. Trees are
-first identified during inventories with a vernacular name assigned by
-the forest worker team, and afterward with a scientific name assigned by
-botanists during regular botanical campaigns. In 1984, specific
-vernacular names are given to 62 commercial or common species whereas
-more infrequent ones were identified under general identifiers only
-distinguishing trees and palm trees. From 2003, botanical campaigns have been
-conducted every 5 to 6 years to identify all trees at the species level
-but identification levels still varied among plots and campaigns.
+The study site corresponds to a tropical rainforest typical of the
+Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae,
+Lecythidaceae and Sapotaceae botanical families. In the twelve
+experimental plots of the experiment, all trees above 10 cm DBH have
+been mapped and measured annually since 1984. Trees are first identified
+during inventories with a vernacular name assigned by the forest worker
+team, and afterward with a scientific name assigned by botanists during
+regular botanical campaigns. In 1984, specific vernacular names are
+given to 62 commercial or common species whereas more infrequent ones
+were identified under general identifiers only distinguishing trees and
+palm trees. From 2003, botanical campaigns have been conducted every 5
+to 6 years to identify all trees at the species level but identification
+levels still varied among plots and campaigns.
 
@@
 
-Six functional traits
-representing leaf economics (leaves thickness, toughness, total
-chlorophyll content and specific leaf area, the leaf area per unit dry
-mass) and wood economics (wood specific gravity and bark thickness), and
-life history traits (maximum specific height and seed mass) came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}. Trait values were assessed from a selection of individuals located
-in nine permanent plots in French Guiana, including two in Paracou, and
+Six functional traits representing leaf economics (leaves thickness,
+toughness, total chlorophyll content and specific leaf area, the leaf
+area per unit dry mass) and wood economics (wood specific gravity and
+bark thickness), and life history traits (maximum specific height and
+seed mass) came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}.
+Traits values were assessed from a selection of individuals located in
+nine permanent plots in French Guiana, including two in Paracou, and
 comprised 294 botanical species pertaining to 157 botanical genera.
 Missing trait values were filled using multivariate imputation by
-chained equation \citep{Mice2011}.
-Imputations were restricted within genus, or family when samples were too
-scarce, in order to account for the phylogenetic signal of the
-functional traits. Whenever a species inventoried was not in the
-dataset, it was attributed a set of traits values randomly sampled among
-species of the same next higher taxonomic level (same genus or family).
-As seed mass information corresponds to a classification into mass
-classes, no data filling process was applied and analysis were
-restricted to the 414 botanical species recorded.
+chained equation \citep{Mice2011}. Imputations were restricted within
+genus, or family when samples were too scarce, in order to account for
+the phylogenetic signal of the functional traits. Whenever a species
+inventoried was not in the dataset, it was attributed a set of traits
+values randomly sampled among species of the same next higher taxonomic
+level (same genus or family). As seed mass information corresponds to a
+classification into mass classes, no data filling process was applied
+and analysis were restricted to the 414 botanical species recorded.
 
-All composition and diversity metrics were
-obtained after 50 iterations of the taxonomic uncertainty propagation
-framework and of the filling process of missing trait values.
+All composition and diversity metrics were obtained after 50 iterations
+of the taxonomic uncertainty propagation framework and of the filling
+process of missing trait values.
 
@@
 To counter taxonomic uncertainties due to the variability of botanical
-identification levels (in space) and protocols (in time) (see {[}\#protocols{]}), the taxonomic
+identification levels (in space) and protocols (in time), the taxonomic
 composition and diversity analysis were conducted at the genus level,
@@
 names. Trajectories of communities taxonomic and functional variations
-in composition after disturbance were followed in a two-dimensional
-NMDS ordination space the 30 years monitored. In both cases the NMDS were performed using occurrence-based
-(Jaccard) and abundance-based (Bray-Curtis) dissimilarity measures.
-Trajectories along time in the plan were reported through the euclidean
-distance of successive inventories to the reference inventories in 1989,
-i.e. 2-3 years after disturbance, when the uncertainty degree did not exceed
-30\% of undetermined trees. To compensate the intrinsinc difference
-of initial position among plots, the trajectories corresponded to the euclidian differences along time
-with the reference position in 1989. Univariate trajectories of the leaf, stem and
-life-history traits were inspected with the community weighted means
-(CWM)
-\citep{Diaz2007, Garnier2004}. Species seed mass corresponded to
-5 classes of increasing mass, seed mass trajectories were therefore
-reported as the proportion of each class in the inventories.
+in composition after disturbance were followed in a two-dimensional NMDS
+ordination space of the 30 years monitored. Two NMDS were conducted to
+map either taxonomic flora inventories or communities functional
+composition based on the 7 leaf, stem and life history traits (without
+seed mass classes). In both cases the NMDS were performed using
+occurrence-based (Jaccard) and abundance-based (Bray-Curtis)
+dissimilarity measures. Trajectories along time in the plan were
+reported through the euclidean distance of successive inventories to the
+reference inventories in 1989, \emph{i.e.} 2-3 years after disturbance,
+when the uncertainty degree did not exceed 30\% of undetermined trees.
+To compensate the intrinsinc difference among plots, the trajectories
+corresponded to the euclidean differences along time with the reference
+inventory in 1989. Univariate trajectories of the leaf, stem and
+life-history traits were also visualized with the community weighted
+means (CWM) \citep{Diaz2007, Garnier2004}. Species seed mass
+corresponded to 5 classes of increasing mass, seed mass trajectories
+were therefore reported as the proportion of each class in the
+inventories.
 
@@
 respectively corresponding to the 0, 1 and 2 order of diversity (q),
-recomended for diversity studies \citep{Patil1982, Tothmeresz1995, Marcon2015}. The
-functional diversity was reported using the Rao index of quadratic
-entropy which combines species abundance distribution and average
-pairwise dissimilarity based on all functional and life traits.
+recomended for diversity studies
+\citep{Patil1982, Tothmeresz1995, Marcon2015}. The functional diversity
+was reported using the Rao index of quadratic entropy which combines
+species abundance distribution and average pairwise dissimilarity based
+on all functional and life traits.
 
@@
 The functional redundancy was measured as the overlap among species in
-community functional space \citep{Carmona2016}. The samples of the
-trait database were first mapped in a 2-dimensional plan from a PCA
-analysis. Then, multivariate kernel density estimator associated with
-individual trees returned species traits probability distribution (TDP).
-Species TDP weighted by species abundance were eventually summed for
-each community: the functional redundancy was the sum of TDPs overlap,
+community functional space \citep{Carmona2016}. The samples of the trait
+database were first mapped in a 2-dimensional plan from a PCA analysis.
+Then, multivariate kernel density estimator associated with individual
+trees returned species traits probability distribution (TDP). Species
+TDP weighted by species abundance were eventually summed for each
+community: the functional redundancy was the sum of TDPs overlap,
 expressed as the average number of species that could be removed from
@@
 
-From 1989 (2-3 years after disturbance) to 2015 (28-29
-years after disturbance), 828388 individual trees and 591 botanical species
-pertaining to 223 genus and 64 botanical families were recorded.
-For undisturbed plots Richness and taxonomic evenness (Shannon and Simpson diversities) remained
-stable over the 30 years monitored. In disturbed communities the
-taxonomic richness increased after low disturbance intensity, reaching a
-maximum gain of 14 botanical genera (plot 3 from treatment 2) while it
-followed unimodal trajectories after intense disturbance, decreasing for
-ten years before recovering to pre-disturbance values. In all disturbed
-plots the taxonomic evenness (Shannon and Simpson diversities)
-increased, following unimodal trajectories with a maximum, reached after
-around 20 years, positively correlated to the disturbance treatment
+From 1989 (2-3 years after disturbance) to 2015 (28-29 years after
+disturbance), 828388 individual trees and 591 botanical species
+pertaining to 223 genus and 64 botanical families were recorded. For
+undisturbed plots Richness and taxonomic evenness (Shannon and Simpson
+diversities) remained stable over the 30 years monitored. In disturbed
+communities the taxonomic richness increased after low disturbance
+intensity, reaching a maximum gain of 14 botanical genera (plot 3 from
+treatment 2) while it followed unimodal trajectories after intense
+disturbance, decreasing for ten years before recovering to
+pre-disturbance values. In all disturbed plots the taxonomic evenness
+(Shannon and Simpson diversities) increased, following unimodal
+trajectories with a maximum, reached after around 20 years, positively
+correlated to the disturbance treatment
 (\(\rho_{spearman}^{Shannon}=0.86\), and
@@
 \caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance.
-Colors are treaments: green (control), blue (T1), orange (T2), red (T3) with sahded areas the credibility intervals.
-Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities, 10, 20 and 30 years after disturbance}\label{fig:IDHplot}
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.
+Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities 10, 20 and 30 years after disturbance.}\label{fig:IDHplot}
 \end{figure*}
@@
 While both taxonomic and functional composition remained stable in
-undisturbed communities (ref to fig), they followed consistent trajectories over time
-after disturbance. According to the mapping of functional traits (see appendix I) these
-compositional changes corresponded to shifts towards species with more
-acquisitive functional strategies, from communities with high average WD
-to high average SLA and chlorophyll content. For disturbed communities
-the distance of successive inventories to the 1989 reference inventory
-followed unimodal trajectories translating cyclic compositional changes
-with a recovery of the initial composition (Figure \ref{fig:NMDSplans}).
-The maximum dissimilarity with the initial state was positively
-correlated to the disturbance treatment for both taxonomic and
-functional composition (\(\rho_{spearman}^{taxonomic}=0.91\) and
+undisturbed communities (Figure \ref{fig:NMDSplans}), they followed
+consistent trajectories over time after disturbance which revealed
+significant compositional changes. According to the mapping of
+functional traits (see appendix I) these compositional changes
+corresponded to shifts towards species with more acquisitive functional
+strategies, from communities with high average WD to high average SLA
+and chlorophyll content. For disturbed communities the distance of
+successive inventories to the 1989 reference inventory followed unimodal
+trajectories translating cyclic compositional changes with a recovery of
+the initial composition (Figure \ref{fig:NMDSplans}). The maximum
+dissimilarity with the initial state was positively correlated to the
+disturbance treatment for both taxonomic and functional composition
+(\(\rho_{spearman}^{taxonomic}=0.91\) and
 \(\rho_{spearman}^{functional}=0.96\) respectively) and the time at
@@
 
-\caption{Plot trajectories in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) in a the two-dimensional NMDS space.
+\caption{Plot trajectories in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS space.
 Lower panels (\textbf{(c)} and \textbf{(d)}) represent the euclidean distance to initial condition along the 30 sampled years.
-Colors are disturbance treatments (green for control, blue for T1,orange for T2 and red for T3) with shaded areas, the credibility intervals}\label{fig:NMDSplans}
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals}\label{fig:NMDSplans}
 \end{figure*}
@@
 
-Changes in functional composition trajectories went hand to hand
-with consistent trajectories of the 8 functional traits (Figure
-\ref{fig:CWM}). Except for leaf chlorophyll content, which continued to increase for
-some T3 and T2 plots 30 years after disturbance, all traits and seed
-mass proportions followed unimodal trajectories either stabilizing or
-returning towards their initial values. Thirty years after disturbance
-the weighted means of communities specific maximum height at adult stage
-(\emph{Hmax}), leaf toughness (\emph{L\_toughness}) and wood specific
-gravity (\emph{WD}) remained significantly lower than their initial
-value (Figure \ref{fig:CWM}). The weighted means of bark thickness
-(\emph{Bark\_thick}) similarly remained substantially higher than
-initially for all disturbed plots while the specfic leaf area
-(\emph{SLA}) had almost recovered its initial value. For all traits the
-maximum difference to initial state was correlated to the disturbance
-intensity (\(\rho_{spearman}^{L_{thickness}}=0.67\),
+Changes in functional composition trajectories went hand to hand with
+consistent trajectories of the 8 functional traits (Figure
+\ref{fig:CWM}). Except for leaf chlorophyll content, which continued to
+increase for some T3 and T2 plots 30 years after disturbance, all traits
+and seed mass proportions followed unimodal trajectories either
+stabilizing or returning towards their initial values. Thirty years
+after disturbance the weighted means of communities specific maximum
+height at adult stage (\emph{Hmax}), leaf toughness
+(\emph{L\_toughness}) and wood specific gravity (\emph{WD}) remained
+significantly lower than their initial value (Figure \ref{fig:CWM}). The
+weighted means of bark thickness (\emph{Bark\_thick}) similarly remained
+substantially higher than initially for all disturbed plots while the
+specfic leaf area (\emph{SLA}) had almost recovered its initial value.
+For all traits the maximum difference to initial state was correlated to
+the disturbance intensity (\(\rho_{spearman}^{L_{thickness}}=0.67\),
 \(\rho_{spearman}^{L_{chloro}}=0.45\),
@@
 \caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}).
-Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
 \end{figure*}
@@
 
-Communities functional redundancy remained stable in control
-plots but after disturbance the redundancy trajectories were quite
-variable (See appendix I) and apparently independently of the initial
-disturbance. Globally after most intense disturbance (plots T2 and T3)
-communities redundancy decreased at first place before increasing to
-edge, recover or exceed the initial value. *I would suggest to consider only the Functional redundancy restricted to the functional space in the paper*
+Communities functional remained stable in control plots but after
+disturbance the redundancy trajectories were quite variable (See
+appendix I) and apparently independently of the initial disturbance.
+Globally after most intense disturbance (plots T2 and T3) communities
+redundancy decreased at first place before increasing to edge, recover
+or exceed the initial value. \textgreater{} restrict to initial
+functional space
 
@@
 
-\caption{Trajectories of functional redundancy within the initial functional space over 30 years after disturbance.
+\caption{TTrajectories of functional redundancy within the initial functional space over 30 years after disturbance.
 Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}(\#fig:RedFun_rest)
@@
 Both communities taxonomic and functional diversities and composition
-proved resilient, following similar humped-shaped trajectories starting to return
-towards initial values 30 years after disturbance. Communities functional signature is the most
-direct link between biodiversity and ecosystem functioning
-\citep{Diaz2005} and its resilience meant the recovery of ecosystem
-processes in the long term \citep{Guariguata2001}. The resilience of
-communities taxonomy, despite the multiple NMDS starting points, suggest the long-term maintenance of the initial
-differences in species composition and diversity, in turn suggesting the existence of multiple stable
-equilibria as assumed for highly diverse and productive ecosystems
-\citep{Chase2003} and the dependency of recovery trajectories on the
-initial composition to be recovered
-\citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.
+proved resilient, following similar humped-shaped trajectories starting
+to return towards initial values 30 years after disturbance. Communities
+functional signature is the most direct link between biodiversity and
+ecosystem functioning \citep{Diaz2005} and its resilience meant the
+recovery of ecosystem processes in the long term \citep{Guariguata2001}.
+The resilience of communities taxonomy, despite the multiple NMDS
+starting points, suggest the long term maintenance of the initial
+differences in species composition and diversity. This maintenance in
+turn suggest the existence of multiple stable equilibria as assumed for
+highly diverse and productive ecosystems \citep{Chase2003} and the
+dependency of recovery trajectories on the initial composition to be
+recovered \citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.
 
@@
 
-Disturbance intensity poorly predicted the taxonomic richness regardless of the period studied, as
-already observed in the Guiana Shield \citep{Baraloto2012a} and in
-Bornean tropical forests \citep{Cannon1998}. Contrastingly, the disturbance intensity consistently predicted the
-increase in functional diversity after disturbance along the 30 years.
-The functional diversity increase paralleled the significant functional shifts towards resource-acquisitive strategies
-(sharp increase in the SLA, leaf thickness and bark thickness and
-decrease in wood density, leaf toughness and maximum height)
+Disturbance intensity poorly predicted the taxonomic richness regadless
+of the period studied, as already observed in the Guiana Shield
+\citep{Baraloto2012a} and in Bornean tropical forests
+\citep{Cannon1998}. Contrastingly, the disturbance intensity
+consistently predicted the increase in functional diversity after
+disturbance along the 30 years. The functional diversity increase
+paralleled the significant functional shifts towards
+resource-acquisitive strategies (sharp increase in the SLA, leaf
+thickness and bark thickness and decrease in wood density, leaf
+toughness and maximum height)
 \citep{Westoby1998, Wright2004, Reich2014}. As the old-growth,
@@
 
-* Not clear from here*
-Disturbance trajectories therefore relied upon the environmental niches
-made available and filled by species from a restricted functional range
-that became dominant and structured the functional characteristics of
-the community \citep{Grime1998}. The intermediate disturbance hypothesis
-was suitabe to predict communities functional diversity and pool of
-dominant species \citep{Molino2001}, but it poorly represented the whole
-taxonomic structure that was hampered by the slow recovery of rare
-species \citep{Hubbell2001, Chave2004}.
+\begin{quote}
+\begin{quote}
+Not clear Disturbance trajectories therefore relied upon the
+environmental niches made available and filled by species from a
+restricted functional range that became dominant and structured the
+functional characteristics of the community \citep{Grime1998}. The
+intermediate disturbance hypothesis was suitabe to predict communities
+functional diversity and pool of dominant species \citep{Molino2001},
+but it poorly represented the whole taxonomic structure that was
+hampered by the slow recovery of rare species
+\citep{Hubbell2001, Chave2004}.
+\end{quote}
+\end{quote}
 
@@
 
-The functional and taxonomic decoupling as well as the rebuttal of the IDH for
-communities taxonomic structure were explained by a reorganization of
-the functional redundancy within the functional traits space that
-implied the incomplete resilience (???) of communities
-\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+The functional and taxonomic decoupling as well as the rebuttal of the
+IDH for communities taxonomic structure were explained by a
+reorganization of the functional redundancy within the functional traits
+space that implied the \textgreater{}\textgreater{} incomplete
+resilience of communities \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
@@
 recovery of the initial redundancy structure was consistent but remained
-unachieved. ?????-> It depended on infrequent species specific to old-growth
-forests that were functionally and therefore underwent competitive
-exclusion limiting functional similarity
-\citep{Chave2004, Mayfield2010}. ?????->The persistent alteration of functional
-redundancy meant a lower resilience of pre-disturbance communities and
-an higher resilience disturbance-specific communities (corresponding to
-lower richness and higher dominance of pioneer species) so chances to
-observe self-maintained compositional changes in favor of disturbance
-resistant species, lianas or epiphytes were increased after disturbance
-\citep{Haddad2008, Burslem2000, Martin2013}. ?? Not so long no? -> Communities resilience
-depended on the functional redundancy structure that proved long to
-recover. This decades-long recovery specifically increased the risks of
-loosing cornerstone (????) species, with unexpected ecological consequences
-\citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.
+unachieved. \textgreater{}\textgreater{} unclear It depended on
+infrequent species specific to old-growth forests that were functionally
+and therefore underwent competitive exclusion limiting functional
+similarity \citep{Chave2004, Mayfield2010}.
+
+\begin{quote}
+\begin{quote}
+unclear The persistent alteration of functional redundancy meant a lower
+resilience of pre-disturbance communities and an higher resilience
+disturbance-specific communities (corresponding to lower richness and
+higher dominance of pioneer species) so chances to observe
+self-maintained compositional changes in favor of disturbance resistant
+species, lianas or epiphytes were increased after disturbance
+\citep{Haddad2008, Burslem2000, Martin2013}.
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+not so long Communities resilience depended on the functional redundancy
+structure that proved long to recover.
+\end{quote}
+\end{quote}
+
+\begin{quote}
+\begin{quote}
+unclear (cornerstone?) This decades-long recovery specifically increased
+the risks of loosing cornerstone specie, with unexpected ecological
+consequences \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.
+\end{quote}
+\end{quote}
 
@@
 enrichment of the system with pioneers and light-demanding species.
-????-> The ????following functional shifts entailed a re-organization of
+
+\begin{quote}
+\begin{quote}
+unclear The following functional shifts entailed a re-organization of
 communities functional redundancy that proved long to recover and
@@
 tolerant species of old-growth forests faced functional redundancy and
-limiting similarity processes. The resilience of tropical forests then
-proved long in the face of quite intense disturbance and did not
-preclude the settling of a persistent disturbance-specific community
-\citep{Gourlet-Fleury2005}. As the trajectories highlighted the
-recruitment processes proved central for communities response to
-disturbance and closer focus on demographical drivers of communities
-response would clarify the fate of the future forests. ????-> The disturbance
-range however stayed within the spectrum of selective logging, with a
-forest cover remaining all along the experiment, and the response
-mechanisms would probably be much different after harder disturbance.
+limiting similarity processes.
+\end{quote}
+\end{quote}
+
+The resilience of tropical forests then proved long in the face of quite
+intense disturbance and did not preclude the settling of a persistent
+disturbance-specific community \citep{Gourlet-Fleury2005}. As the
+trajectories highlighted the recruitment processes proved central for
+communities response to disturbance and closer focus on demographical
+drivers of communities response would clarify the fate of the future
+forests.
+
+\begin{quote}
+\begin{quote}
+unclear The disturbance range however stayed within the spectrum of
+selective logging, with a forest cover remaining all along the
+experiment, and the response mechanisms would probably be much different
+after harder disturbance.
+\end{quote}
+\end{quote}
 

### 2018-06-28 Scripts/density_fromEric.Rmd
new graphs

---
title: "Comparaison de distributions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Données simulées

Deux distributions D1 : normale centrée réduite et normale de moyenne 0.5 et écart-type 2. 100 valeurs pour chacune. Correspondent à la distribution des traits à deux dates.

```{r distributions}
D1 <- rnorm(100)
D2 <- rnorm(100, mean = .5, sd = 2)
```

## Estimation de la densité de probabilité

### Recherche de la bande passante optimale

```{r bp}
# Comparaison entre les bp de chaque distribution et de l'ensemble des données
bw.SJ(D1)
bw.SJ(D2)
bw.SJ(c(D1, D2))
# On retient la bp de la distribution de référence, ici D1
OptimalBW <- bw.SJ(D1)
```

### Estimation des densités

Avec la même bande passante

```{r density}
dD1 <- density(D1, bw=OptimalBW)
dD2 <- density(D2, bw=OptimalBW)
# figure
plot(dD1)
lines(dD2, col="red")
```

## Distance entre les distributions

### Valeurs de x

Les valeurs de x sont différentes entre les deux distributions de densité. On complète chaque distribution par les valeurs de x de l'autre.

```{r approx}
# Estimation de la distribution 1 aux x de la diostrbution 2
adD1 <- approx(dD1$x, dD1$y, xout=dD2$x)
# Les NA sont des zéros (au-delà des bornes d'origine)
adD1$y[is.na(adD1$y)] <- 0
# Estimation de la distribution 2 aux x de la diostrbution 1
adD2 <- approx(dD2$x, dD2$y, xout=dD1$x)
# Les NA sont des zéros (au-delà des bornes d'origine)
adD2$y[is.na(adD2$y)] <- 0
# Fusion. Toutes les valeurs de x
x <- c(dD1$x, dD2$x)
# y de la distribution 1
y1 <- c(dD1$y, adD1$y)
# y de la distribution 2. Attention à l'ordre
y2 <- c(adD2$y, dD2$y)
# Tri par x croissant.
sortedXindex <- sort(x, index.return=TRUE)$ix
y1 <- y1[sortedXindex]
y2 <- y2[sortedXindex]
x <- x[sortedXindex]
# Figure: la même qu'avant mais avec toutes les valeurs de x et y
plot(x, y1, type="l")
lines(x, y2, col="red")
```

### Valeurs de x

Calcul de la surface commune aux deux courbes.
La distance entre les deux courbes, normalisé, est 1 - deux fois la surface commune / la surface totale.

```{r distance}
library(sfsmisc)
# calculate intersection densities
w <- pmin(y1, y2)
# integrate areas under curves
total <- integrate.xy(x, y1) + integrate.xy(x, y2)
intersection <- integrate.xy(x, w)
# compute non-overlap coefficient
1 -  2*intersection/total
# Figure : surface commune
plot(x, y1, type="l")
lines(x, y2, col="red")
polygon(c(x, rev(x)), c(w, rep(0, length(x))), col = "orange")
```

### 2018-06-28 WholePlotTrajectories.Rmd
new graphs

@@
 Two NMDS were conducted to map either taxonomic flora inventories or communities functional composition based on the 7 leaf, stem and life history traits (without seed mass classes).
-In both cases the NMDS were performed using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) dissimilarity measures. 
-Trajectories along time in the plan were reported through the euclidean distance of successive inventories to the reference inventories in 1989, *i.e.* 2-3 years after disturbance, when the uncertainty degree did not exceed 30% of undetermined trees.
+In both cases the NMDS were performed using abundance-based (Bray-Curtis) dissimilarity measures. 
+Trajectories along time were reported through the distance between successive inventories and the reference inventories in 1989, *i.e.* 2-3 years after disturbance, when the uncertainty degree did not exceed 30% of undetermined trees.
+Taxonomic trajectories were measured through the euclidean distance between inventories in the multivariate space of all inventories species.
+Functional trajectories were measured through the sum of euclidean distances among traits density distributions between inventories.
 To compensate the intrinsinc difference among plots, the trajectories corresponded to the euclidean differences along time with the reference inventory in 1989.
@@
 load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
-load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
+load("DB/FunDistance_ForGraphs");load("DB/TaxoDistance_ForGraphs")
 
@@
 FunCompo(MatrepFun)
-TaxoDist(MatrepTaxo)
-FunDist(MatrepFun)
+EuclidDist(TaxoEuclid)
+EuclidDist(FunEuclid)
 

### 2018-06-28 WholePlotTrajectories.tex
new graphs

@@
 seed mass classes). In both cases the NMDS were performed using
-occurrence-based (Jaccard) and abundance-based (Bray-Curtis)
-dissimilarity measures. Trajectories along time in the plan were
-reported through the euclidean distance of successive inventories to the
-reference inventories in 1989, \emph{i.e.} 2-3 years after disturbance,
-when the uncertainty degree did not exceed 30\% of undetermined trees.
-To compensate the intrinsinc difference among plots, the trajectories
-corresponded to the euclidean differences along time with the reference
-inventory in 1989. Univariate trajectories of the leaf, stem and
-life-history traits were also visualized with the community weighted
-means (CWM) \citep{Diaz2007, Garnier2004}. Species seed mass
+abundance-based (Bray-Curtis) dissimilarity measures. Trajectories along
+time were reported through the distance between successive inventories
+and the reference inventories in 1989, \emph{i.e.} 2-3 years after
+disturbance, when the uncertainty degree did not exceed 30\% of
+undetermined trees. Taxonomic trajectories were measured through the
+euclidean distance between inventories in the multivariate space of all
+inventories species. Functional trajectories were measured through the
+sum of euclidean distances among traits density distributions between
+inventories. To compensate the intrinsinc difference among plots, the
+trajectories corresponded to the euclidean differences along time with
+the reference inventory in 1989. Univariate trajectories of the leaf,
+stem and life-history traits were also visualized with the community
+weighted means (CWM) \citep{Diaz2007, Garnier2004}. Species seed mass
 corresponded to 5 classes of increasing mass, seed mass trajectories

### 2018-06-28 Scripts/DrawingTools.R

@@ -65,81 +65,23 @@ mtext("NMDS 1",side=1,line=2,cex=0.8)
 mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)    
 }
 
-TaxoDist<-function(Data_TaxoComp){
-MatrepT<-lapply(Data_TaxoComp,function(Rep){
-  Rep<-as.data.frame(Rep)
-  Rep$plot<-substr(rownames(Rep),start=1,stop=regexpr("_",rownames(Rep))-1)
-  Rep$year<-substr(rownames(Rep),start=regexpr("_",rownames(Rep))+1,
-                   stop=nchar(rownames(Rep)))
-  return(Rep)
-})
+EuclidDist<-function(Taxo_euclid){
+  ret<-lapply(c(0.025,0.5,0.975),function(quant){
+  return(apply(Taxo_euclid,c(1,2),function(col){return(quantile(col,probs=quant))}))})
+names(ret)<-c(0.025,0.5,0.975)
 
-DistT<-lapply(1:12,function(pl){
-  ret<-lapply(MatrepT,function(rep){return(rep[which(rep[,"plot"]==pl),])})
-  ret<-lapply(ret,function(rep){return(rep[which(rep[,"year"]%in%
-                                                   ret[[which(unlist(lapply(ret,nrow))==min(unlist(lapply(ret,nrow))))[1]]][,"year"]),])})
-  ret<-do.call(rbind,lapply(ret,function(rep){
-    ret2<-apply(rep[,c("NMDS1","NMDS2")],1,function(li){
-      return(sqrt(sum((rep[1,c("NMDS1","NMDS2")]-li)^2)))})
-    names(ret2)<-rep[,"year"]
-    return(ret2)}))
-  ret<-do.call(rbind,lapply(c(0.025,0.5,0.975),function(quant){
-    return(apply(ret,2,function(col){return(quantile(col,probs=quant))}))}))
-  rownames(ret)<-c(0.025,0.5,0.975)
-  return(ret)})
-names(DistT)<-1:12
-DistT<-lapply(DistT,function(pl){
-  colnames(pl)<-as.numeric(colnames(pl))-1984;return(pl)})
-
-plot(colnames(DistT[[1]]),DistT[[1]][1,],type="n",xlab="",ylab="",
-     ylim=c(min(unlist(DistT)),max(unlist(DistT))),cex.axis=0.7)
+plot(colnames(ret[[2]]),ret[[2]][1,],type="n",xlab="",ylab="",
[…]

### 2018-06-28 Scripts/FunctionalNMDS.R

@@ -4,10 +4,12 @@ source("Scripts/TraitsMiceFilling.R")                                   # Traits
 load("DB/Alpha_Plots")                  # Alpha vector for vernacular replacement
 load("DB/Paracou_R_Subdivided_ok")      # Paracou dataset, lists by inventoried year, sublists by plot, etc...
 dates<-sort(names(LivingStand_all))[-1] # dates considered
+dates<-dates[which(!dates%in%c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014","2016"))]
+
 
 Traits1<-read.csv("DB/BridgeOK.csv",sep=";",na.strings="")       # Brigde functional traits dataset
 Traits2<-read.csv("DB/DataLifeTraits.csv",sep=";",na.strings="") # Paracou15 seed and Hmax dataset
-TraitsName<-c("Hmax","S_mass","L_thickness","L_chloro",
+TraitsName<-c("Hmax","L_thickness","L_chloro",
               "L_toughness","L_DryMass","SLA","WD","Bark_thick") # Considered traits list
 
 # Treatments definition
@@ -19,6 +21,12 @@ names(treatments)<-c("Control","T1","T2","T3")
 AllInv<-unique(unlist(lapply(LivingStand_all,function(yr){
   return(unique(unlist(lapply(yr,function(pl){return(unique(pl[,"name"]))}))))})))
 AllInv<-as.data.frame(unique(AllInv),row.names=as.character(AllInv))
+
+InventorySp<-do.call(rbind,lapply(LivingStand_all,function(yr){
+  ret<-do.call(rbind,yr)[,c("Famille","Genre","name")]
+  ret<-ret[which(!grepl("Indet.",ret[,"name"])),]
+  return(ret[which(!duplicated(ret)),])}))
+InventorySp<-InventorySp[which(!duplicated(InventorySp)),]
 
 Nrep<-50 # Iterations for the vernacular names replacement and traits gap filling
 
@@ -102,3 +110,133 @@ Mat2<-as.matrix(daisy(Mat2,metric="gower"))
 })
 
 save(MatrepFun,file = "DB/FunctionalComposition_forGraph")
+
+
+###########################################################
+## traits density distance
+
+library(sfsmisc)
+
[…]

### 2018-06-28 Scripts/TaxonomicNMDS.R

@@ -2,7 +2,7 @@ library(vegan)
 library(reshape2)
 library(ade4)
 library(cluster)
-source("Vernacular_handle.R")
+source("Scripts/Vernacular_handle.R")
 load("DB/Alpha_Plots")
 load("DB/Paracou_R_Subdivided_ok")
 
@@ -12,7 +12,7 @@ treatments<-list(T0,T1,T2,T3);names(treatments)<-c("Control","T1","T2","T3")
 ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
 
 dates<-sort(names(LivingStand_all))[-1]
-dates<-dates[which(!dates%in%c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014"))]
+dates<-dates[which(!dates%in%c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014","2016"))]
 
 AllInv<-do.call(rbind,lapply(LivingStand_all,function(yr){
   Ret<-do.call(rbind,lapply(yr,function(pl){
@@ -43,16 +43,18 @@ Matrep<-lapply(1:Nrep,function(rep){
     Ok<-names(which(unlist(lapply(2:length(Trajnmds),function(step){return(Ntrees[step]-Ntrees[step-1])}))<=-90))
     Trajnmds<-Trajnmds[which(!names(Trajnmds)%in%Ok)]
     namesOk<-names(Trajnmds)
+    ref<-as.data.frame(Trajnmds[["1985"]])
     Trajnmds<-do.call(cbind,lapply(Trajnmds,function(yr){
       All<-as.data.frame(AllInv[, "Genre"]);colnames(All)<-"Genre"
       ret<-merge(All,yr,by.x="Genre",by.y="row.names",all.x=TRUE)
-      ret<-ret[which(!duplicated(ret)),]
+      ret<-merge(ret,ref,by.x="Genre",by.y="row.names",all.x=TRUE)
+      ret<-ret[which(!duplicated(ret)),2:3]
       ret[is.na(ret)]<-0
-      return(ret[2])}))
-    colnames(Trajnmds)<-paste(p,"_",namesOk,sep="")
+      return(dist(t(ret)))}))
+    colnames(Trajnmds)<-namesOk
     return(Trajnmds)})
     
-Mat<-do.call(cbind,Mat)
+Mat<-do.call(rbind,Mat)
 Mat<-t(Mat[which(rowSums(Mat)>0),])
 
[…]

### 2018-06-28 Scripts/TraitsMiceFilling.R

@@ -93,6 +93,7 @@ traitsPartial_fam<-lapply(traitsPartial_fam,function(sub){
   return(cbind(sub[,c("Family","Genus","name","bar_code")],ret[(nrow(ret)-nrow(sub)+1):nrow(ret),]))})
 
 # Pour les espèce squi n'avaient toujours pas assez d'individus dans leur famille pour permettre l'estimation
+if(any(unlist(lapply(traitsPartial_fam,function(sub){anyNA(sub[,"WD"])})))){
 idScarce<-as.character(traitsPartial_all[,"bar_code"],
    do.call(rbind,traitsPartial_fam[which(unlist(
      lapply(traitsPartial_fam,function(sub){anyNA(sub[,"WD"])})))])[,"bar_code"])
@@ -102,9 +103,13 @@ comScarce<-cbind(traits[,c("Family","Genus","name","bar_code")],
 comScarce<-comScarce[which(comScarce[,"bar_code"]%in%idScarce),]
 traitsPartial_fam<-do.call(rbind,
                     traitsPartial_fam[which(!unlist(lapply(traitsPartial_fam,function(sub){anyNA(sub[,"WD"])})))])
-
+traitsPartial_fam<-rbind(traitsPartial_fam,comScarce)
+}
+if(!any(unlist(lapply(traitsPartial_fam,function(sub){anyNA(sub[,"WD"])})))){
+traitsPartial_fam<-do.call(rbind,traitsPartial_fam[which(!unlist(lapply(traitsPartial_fam,function(sub){anyNA(sub[,"WD"])})))])
+}
 # On rassemble tout
-traitsPartial_filled<-rbind(traitsPartial_gen,traitsPartial_fam,comScarce)
+traitsPartial_filled<-rbind(traitsPartial_gen,traitsPartial_fam)
 
 # Pour les espèces dont on a aucune valeur de traits, 
 # on tire un ensemble de traits complet correspondant à l'une des espcèces de l'inventaire.

### 2018-06-30 Scripts/DrawingTools.R
graphs, html & pdf

@@ -110,7 +110,7 @@ plotDiv<-function(Data,remove=FALSE){
   Toplot<-lapply(Toplot,function(tr){
     ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col}))})#-tr[,1,rep]
     ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),
-               dimnames=list(rownames(ret[[1]]),as.numeric(colnames(ret[[1]]))-1984,1:length(ret)))
+               dimnames=list(rownames(ret[[1]]),as.numeric(colnames(ret[[1]]))-1986,1:length(ret)))
     ret<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(ret,c(1,2),function(x){return(quantile(x,probs=quant))}))})
     return(array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),3),
                  dimnames=list(rownames(ret[[1]]),colnames(ret[[1]]),c(0.025,0.5,0.975))))})
@@ -194,7 +194,7 @@ RedundancyPlot<-function(Red){
     ret<-smooth(rep,2)
    colnames(ret)<-colnames(rep)
    ret<-apply(ret[,which(as.numeric(colnames(ret))>="1989")],2,function(col){return(col-ret[,"1989"])})#
-   colnames(ret)<-as.numeric(colnames(ret))-1984
+   colnames(ret)<-as.numeric(colnames(ret))-1986
    return(ret)})
   
   Red<-array(unlist(Red),dim=c(12,ncol(Red[[1]]),length(Red)),
@@ -260,7 +260,7 @@ TaxoTraj<-function(CompTaxo){
     Toplot<-lapply(Toplot,function(tr){
       ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col-tr[,1,rep]}))})#
       ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),
-                 dimnames=list(rownames(ret[[1]]),as.numeric(colnames(ret[[1]]))-1984,1:length(ret)))
+                 dimnames=list(rownames(ret[[1]]),as.numeric(colnames(ret[[1]]))-1986,1:length(ret)))
[…]

### 2018-06-30 Scripts/FunctionalNMDS.R

@@ -123,7 +123,7 @@ rownames(traits_filled)<-traits_filled[,1];traits_filled<-traits_filled[,TraitsN
 
 Nrep<-3
 
-MatrepFun<-lapply(1:Nrep,function(rep){
+FunEuclid<-lapply(1:Nrep,function(rep){
   
 Mat<-do.call(rbind,lapply(1:12,function(p){
     
@@ -210,33 +210,5 @@ rownames(Mat)<-1:12
 return(Mat)
 })
 
-
-Matrep<-Mat
-
-
-Matrep<-array(unlist(Matrep),dim=c(nrow(Matrep[[1]]),ncol(Matrep[[1]]),length(Matrep)),
-                  dimnames=list(1:nrow(Matrep[[1]]),colnames(Matrep[[1]]),1:length(Matrep)))
-
-ret<-lapply(c(0.025,0.5,0.975),function(quant){
-    return(apply(Matrep,c(1,2),function(col){return(quantile(col,probs=quant))}))})
-  names(ret)<-c(0.025,0.5,0.975)
-  
-  plot(colnames(ret[[2]]),ret[[2]][1,],type="n",xlab="",ylab="",
-       ylim=c(min(unlist(ret)),max(unlist(ret))),cex.axis=0.7)
-  invisible(lapply(1:length(treatments),function(tr){
-    Toplot<-lapply(treatments[[tr]],function(plo){return(do.call(rbind,lapply(ret,function(quant){return(quant[plo,])})))})
-    invisible(lapply(Toplot,function(plo){
-      lines(colnames(plo),plo["0.5",],col=ColorsTr[[tr]],lwd=2)
-      polygon(c(colnames(plo),rev(colnames(plo))),c(plo["0.025",],rev(plo["0.975",])),
-              col=rgb(0,0,0,alpha=0.1),border=NA)
-    }))
-  }))
-  mtext("Distance from 1989 inventory",side=2,padj=0,line=2,cex=0.8)
-  mtext("(c)",side=3,adj=0,line=0.5)
-
-save(FunEuclid,file="DB/FunDistance_ForGraphs")
-windows()
-par(mfrow=c(1,2))
-EuclidDist(FunEuclid)
-EuclidDist(TaxoEuclid)
-
+FunEuclid<-array(unlist(FunEuclid),dim=c(nrow(FunEuclid[[1]]),ncol(FunEuclid[[1]]),length(FunEuclid)),
[…]

### 2018-06-30 WholePlotTrajectories.Rmd
relecture

@@
 
-Communities functional remained stable in control plots but after disturbance the redundancy trajectories were quite variable (See appendix I) and apparently independently of the initial disturbance.
+In control plots the functional redundancy remained stable but in undisturbed plots it was quite variable and independent of the disturbance intensity (See appendix I).
 Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place before increasing to edge, recover or exceed the initial value.
-> restrict to initial functional space
 
@@
 
-## Decoupled taxonomic and functional trajectories
+>> Si on met que richesse et evenness, pour functional c'est de l'evenness (Rao = order2)
 
-Both communities taxonomic and functional diversities and composition proved resilient, following similar humped-shaped trajectories starting to return towards initial values 30 years after disturbance.
-Communities functional signature is the most direct link between biodiversity and ecosystem functioning [@Diaz2005] and its resilience meant the recovery of ecosystem processes in the long term [@Guariguata2001].
-The resilience of communities taxonomy, despite the multiple NMDS starting points, suggest the long term maintenance of the initial differences in species composition and diversity.
-This maintenance in turn suggest the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003] and the dependency of recovery trajectories on the initial composition to be recovered [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
+## Communities cyclic recovery
 
-Although communities taxonomic and functional trajectories were similar the taxonomic recovery lagged behind the functional one, confirming a decoupling between functional and taxonomic dynamics already observed for grasslands [@Tilman1997; @Mouillot2011] and tropical forests [@Lohbeck2015; @Guariguata2001].
-According to the "vegetation quantity effect" [@Grime1998] the functional trajectories rely on the pool of dominant species, which diversity and evenness increased and rapidly recovered after disturbance.
-At the same time communities taxonomic evenness remained high, revealing the unachieved recovery of infrequent species.
+Both taxonomic and functional composition proved resilient and followed similar humped-shaped trajectories starting to return towards initial values 30 years after disturbance.
 
-## The extent of the intermediate disturbance hypothesis
+The taxonomic NMDS revealed differences in the composition of initial communities that proved maintained throughout the recovery trajectories after disturbance.
+More than commonly though, communities trajectories after disturbance was specific to the initial composition and converged towards its recovery [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
+The resilience of commnities taxonomy suggested the existence of multiple stable equilibria, corresponding here to communities initial composition, as assumed for highly diverse and productive ecosystems [@Chase2003].
 
-Disturbance intensity poorly predicted the taxonomic richness regadless of the period studied, as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
-Contrastingly, the disturbance intensity consistently predicted the increase in functional diversity after disturbance along the 30 years.
+Contrastingly communities functional composition and diversity proved more homogeneous among communities and communities functional trajectories were similar in the functional space.
+Communities functional characteristics are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] so their resilience meant the recovery of ecosystem processes in the long term [@Guariguata2001].
+
+## Another view of the intermediate disturbance hypothesis
+
+The initial disturbance intensity determined the maximum functional diversity increase, confirming the IDh within this range of disturbance.
+Regarding communities communities along time, the humped-shaped trajectories of the functional diversity following disturbance demonstrated the positive impact of disturbance that enhanced the diversity before a return to initial lower diversity levels.
+In accordance with the IDH disturbance entailed ecological processes different than in undisturbed communities and a temporal variability correlated to the intensity of the disturbance.
 The functional diversity increase paralleled the significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
 As the old-growth, pre-disturbance survivor trees proved mirroring the initial communities [@Herault2018] the functional trajectories and taxonomic turnover rather relied upon the pool of newly recruited trees driven by the enhanced growth and survival of previously infrequent species and functional types.
-Disturbance entailed a reorganization, benefiting to pioneers and light demanding species, of the typical high dominance structure of hyperdiverse mature forests.
 The changes in abiotic environment and competitive pressure favored pioneers which outcompeted other species in non limiting light resources but were excluded in mature forests by long-lived, more resistant and shade tolerant species.
-
 >> Not clear
 Disturbance trajectories therefore relied upon the environmental niches made available and filled by species from a restricted functional range that became dominant and structured the functional characteristics of the community [@Grime1998].
+
+In contrast the trajectories of taxonomic richness and evenness were decoupled from the disturbance intensity, questioning the consistency of the IDH to predict communities taxonomic diversity, as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+
+In the most disturbed plots the taxonomic richness did not exceed this of the initial communities but the taxonomic evenness significantly increased, revealing the loss of most infrequent species.
+Dominant species fastly recovered and restored the communities functoinal characteristics that mostly rely on dominant species, according to the "vegetation quantity effect" [@Grime1998], while the recovery of more infrequent species was hampered.
+>> Not clear
 The intermediate disturbance hypothesis was suitabe to predict communities functional diversity and pool of dominant species [@Molino2001], but it poorly represented the whole taxonomic structure that was hampered by the slow recovery of rare species [@Hubbell2001;@Chave2004].
@@
 
-The functional and taxonomic decoupling as well as the rebuttal of the IDH for communities taxonomic structure were explained by a reorganization of the functional redundancy within the functional traits space that implied the
->> incomplete resilience of communities [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+The recovery of the most infrequent species was hampered by the competition among species, in accordance with the lottery recruitment [@Busing2002] and entailed long term recovery cycles [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
-The functional redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
+The progressive increase of the functional redundancy within the initial functional space, that had first sharply decreased, explains the lottery recruitment that first allow an easy taxonomic recovery and then a progressive slow down.
+functional redundancy restricted to the functional space of the initial community clearly followed.
 After disturbance communities differently occupied the functional space, following the observed shifts in traits trajectories and functional composition, and the redundancy in the initial functional space decreased according to the disturbance intensity.
 After 30 years the recovery of the initial redundancy structure was consistent but remained unachieved.
->> unclear
-It depended on infrequent species specific to old-growth forests that were functionally and therefore underwent competitive exclusion limiting functional similarity [@Chave2004;@Mayfield2010].
 
@@
 
->>not so long
-Communities resilience depended on the functional redundancy structure that proved long to recover.
-
 >>unclear (cornerstone?)
 This decades-long recovery specifically increased the risks of loosing cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
+Indeed the flora-faune relationship and the functional traits studied are not exhaustive and key function might be lost. 
+specifically for the fauna-flora relationships.
 

### 2018-06-30 WholePlotTrajectories.tex
relecture

@@
 
-Communities functional remained stable in control plots but after
-disturbance the redundancy trajectories were quite variable (See
-appendix I) and apparently independently of the initial disturbance.
-Globally after most intense disturbance (plots T2 and T3) communities
-redundancy decreased at first place before increasing to edge, recover
-or exceed the initial value. \textgreater{} restrict to initial
-functional space
+In control plots the functional redundancy remained stable but in
+undisturbed plots it was quite variable and independent of the
+disturbance intensity (See appendix I). Globally after most intense
+disturbance (plots T2 and T3) communities redundancy decreased at first
+place before increasing to edge, recover or exceed the initial value.
 
@@
 
-\subsection{Decoupled taxonomic and functional
-trajectories}\label{decoupled-taxonomic-and-functional-trajectories}
+\begin{quote}
+\begin{quote}
+Si on met que richesse et evenness, pour functional c'est de l'evenness
+(Rao = order2)
+\end{quote}
+\end{quote}
 
-Both communities taxonomic and functional diversities and composition
-proved resilient, following similar humped-shaped trajectories starting
-to return towards initial values 30 years after disturbance. Communities
-functional signature is the most direct link between biodiversity and
-ecosystem functioning \citep{Diaz2005} and its resilience meant the
+\subsection{Communities cyclic
+recovery}\label{communities-cyclic-recovery}
+
+Both taxonomic and functional composition proved resilient and followed
+similar humped-shaped trajectories starting to return towards initial
+values 30 years after disturbance.
+
+The taxonomic NMDS revealed differences in the composition of initial
+communities that proved maintained throughout the recovery trajectories
+after disturbance. More than commonly though, communities trajectories
+after disturbance was specific to the initial composition and converged
+towards its recovery
+\citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}. The
+resilience of commnities taxonomy suggested the existence of multiple
+stable equilibria, corresponding here to communities initial
+composition, as assumed for highly diverse and productive ecosystems
+\citep{Chase2003}.
+
+Contrastingly communities functional composition and diversity proved
+more homogeneous among communities and communities functional
+trajectories were similar in the functional space. Communities
+functional characteristics are the most direct link between biodiversity
+and ecosystem functioning \citep{Diaz2005} so their resilience meant the
 recovery of ecosystem processes in the long term \citep{Guariguata2001}.
-The resilience of communities taxonomy, despite the multiple NMDS
-starting points, suggest the long term maintenance of the initial
-differences in species composition and diversity. This maintenance in
-turn suggest the existence of multiple stable equilibria as assumed for
-highly diverse and productive ecosystems \citep{Chase2003} and the
-dependency of recovery trajectories on the initial composition to be
-recovered \citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.
 
-Although communities taxonomic and functional trajectories were similar
-the taxonomic recovery lagged behind the functional one, confirming a
-decoupling between functional and taxonomic dynamics already observed
-for grasslands \citep{Tilman1997, Mouillot2011} and tropical forests
-\citep{Lohbeck2015, Guariguata2001}. According to the ``vegetation
-quantity effect'' \citep{Grime1998} the functional trajectories rely on
-the pool of dominant species, which diversity and evenness increased and
-rapidly recovered after disturbance. At the same time communities
-taxonomic evenness remained high, revealing the unachieved recovery of
-infrequent species.
+\subsection{Another view of the intermediate disturbance
+hypothesis}\label{another-view-of-the-intermediate-disturbance-hypothesis}
 
-\subsection{The extent of the intermediate disturbance
-hypothesis}\label{the-extent-of-the-intermediate-disturbance-hypothesis}
-
-Disturbance intensity poorly predicted the taxonomic richness regadless
-of the period studied, as already observed in the Guiana Shield
-\citep{Baraloto2012a} and in Bornean tropical forests
-\citep{Cannon1998}. Contrastingly, the disturbance intensity
-consistently predicted the increase in functional diversity after
-disturbance along the 30 years. The functional diversity increase
-paralleled the significant functional shifts towards
+The initial disturbance intensity determined the maximum functional
+diversity increase, confirming the IDh within this range of disturbance.
+Regarding communities communities along time, the humped-shaped
+trajectories of the functional diversity following disturbance
+demonstrated the positive impact of disturbance that enhanced the
+diversity before a return to initial lower diversity levels. In
+accordance with the IDH disturbance entailed ecological processes
+different than in undisturbed communities and a temporal variability
+correlated to the intensity of the disturbance. The functional diversity
+increase paralleled the significant functional shifts towards
 resource-acquisitive strategies (sharp increase in the SLA, leaf
@@
 enhanced growth and survival of previously infrequent species and
-functional types. Disturbance entailed a reorganization, benefiting to
-pioneers and light demanding species, of the typical high dominance
-structure of hyperdiverse mature forests. The changes in abiotic
-environment and competitive pressure favored pioneers which outcompeted
-other species in non limiting light resources but were excluded in
-mature forests by long-lived, more resistant and shade tolerant species.
+functional types. The changes in abiotic environment and competitive
+pressure favored pioneers which outcompeted other species in non
+limiting light resources but were excluded in mature forests by
+long-lived, more resistant and shade tolerant species.
+\textgreater{}\textgreater{} Not clear Disturbance trajectories
+therefore relied upon the environmental niches made available and filled
+by species from a restricted functional range that became dominant and
+structured the functional characteristics of the community
+\citep{Grime1998}.
 
-\begin{quote}
-\begin{quote}
-Not clear Disturbance trajectories therefore relied upon the
-environmental niches made available and filled by species from a
-restricted functional range that became dominant and structured the
-functional characteristics of the community \citep{Grime1998}. The
-intermediate disturbance hypothesis was suitabe to predict communities
-functional diversity and pool of dominant species \citep{Molino2001},
-but it poorly represented the whole taxonomic structure that was
-hampered by the slow recovery of rare species
+In contrast the trajectories of taxonomic richness and evenness were
+decoupled from the disturbance intensity, questioning the consistency of
+the IDH to predict communities taxonomic diversity, as already observed
+in the Guiana Shield \citep{Baraloto2012a} and in Bornean tropical
+forests \citep{Cannon1998}.
+
+In the most disturbed plots the taxonomic richness did not exceed this
+of the initial communities but the taxonomic evenness significantly
+increased, revealing the loss of most infrequent species. Dominant
+species fastly recovered and restored the communities functoinal
+characteristics that mostly rely on dominant species, according to the
+``vegetation quantity effect'' \citep{Grime1998}, while the recovery of
+more infrequent species was hampered. \textgreater{}\textgreater{} Not
+clear The intermediate disturbance hypothesis was suitabe to predict
+communities functional diversity and pool of dominant species
+\citep{Molino2001}, but it poorly represented the whole taxonomic
+structure that was hampered by the slow recovery of rare species
 \citep{Hubbell2001, Chave2004}.
-\end{quote}
-\end{quote}
 
@@
 
-The functional and taxonomic decoupling as well as the rebuttal of the
-IDH for communities taxonomic structure were explained by a
-reorganization of the functional redundancy within the functional traits
-space that implied the \textgreater{}\textgreater{} incomplete
-resilience of communities \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+The recovery of the most infrequent species was hampered by the
+competition among species, in accordance with the lottery recruitment
+\citep{Busing2002} and entailed long term recovery cycles
+\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-The functional redundancy restricted to the functional space of the
-initial community clearly followed humped shaped trajectories. After
+The progressive increase of the functional redundancy within the initial
+functional space, that had first sharply decreased, explains the lottery
+recruitment that first allow an easy taxonomic recovery and then a
+progressive slow down. functional redundancy restricted to the
+functional space of the initial community clearly followed. After
 disturbance communities differently occupied the functional space,
@@
 recovery of the initial redundancy structure was consistent but remained
-unachieved. \textgreater{}\textgreater{} unclear It depended on
-infrequent species specific to old-growth forests that were functionally
-and therefore underwent competitive exclusion limiting functional
-similarity \citep{Chave2004, Mayfield2010}.
+unachieved.
 
@@
 \begin{quote}
-not so long Communities resilience depended on the functional redundancy
-structure that proved long to recover.
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
 unclear (cornerstone?) This decades-long recovery specifically increased
@@
 consequences \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.
+Indeed the flora-faune relationship and the functional traits studied
+are not exhaustive and key function might be lost. specifically for the
+fauna-flora relationships.
 \end{quote}

### 2018-06-30 Scripts/DrawingTools.R

@@ -66,12 +66,15 @@ mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)
 }
 
 EuclidDist<-function(Distances){
+  colnames(Distances)<-as.numeric(colnames(Distances))-1986
   ret<-lapply(c(0.025,0.5,0.975),function(quant){
-  return(apply(Distances,c(1,2),function(col){return(quantile(col,probs=quant))}))})
+  return(apply(Distances[,which(as.numeric(colnames(Distances))>=3),],c(1,2),function(col){
+    return(quantile(col,probs=quant))}))})
 names(ret)<-c(0.025,0.5,0.975)
 
-plot(colnames(ret[[2]]),ret[[2]][1,],type="n",xlab="",ylab="",
+plot(colnames(ret[[2]]),ret[[2]][1,],type="n",xlab="",ylab="",xaxt="n",
      ylim=c(min(unlist(ret)),max(unlist(ret))),cex.axis=0.7)
+axis(1,at=as.numeric(colnames(ret[[1]])),labels=TRUE)  
 invisible(lapply(1:length(treatments),function(tr){
   Toplot<-lapply(treatments[[tr]],function(plo){return(do.call(rbind,lapply(ret,function(quant){return(quant[plo,])})))})
   invisible(lapply(Toplot,function(plo){
@@ -80,7 +83,7 @@ invisible(lapply(1:length(treatments),function(tr){
             col=rgb(0,0,0,alpha=0.1),border=NA)
   }))
 }))
-mtext("Distance from 1989 inventory",side=2,padj=0,line=2,cex=0.8)
+mtext("Euclidean distance from 1985 inventory",side=2,padj=0,line=2,cex=0.8)
 mtext("(c)",side=3,adj=0,line=0.5)
 }
 
@@ -107,6 +110,7 @@ plotDiv<-function(Data,remove=FALSE){
   if(remove){Data[[2]]<-Data[[2]][which(rownames(Data[[2]])!=7),,]}
   
   Toplot<-lapply(Data,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
+  absc<-as.numeric(colnames(Toplot[[1]]))-1986
   Toplot<-lapply(Toplot,function(tr){
[…]

### 2018-07-01 WholePlotTrajectories.Rmd
discussion

@@
 
-Both taxonomic and functional composition proved resilient and followed similar humped-shaped trajectories starting to return towards initial values 30 years after disturbance.
+Communities taxonomic and functional composition and functional evenness proved resilient, following similar humped-shaped trajectories starting to return towards pre-disturbance composition and functioning after 30 years.
 
-The taxonomic NMDS revealed differences in the composition of initial communities that proved maintained throughout the recovery trajectories after disturbance.
-More than commonly though, communities trajectories after disturbance was specific to the initial composition and converged towards its recovery [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
-The resilience of commnities taxonomy suggested the existence of multiple stable equilibria, corresponding here to communities initial composition, as assumed for highly diverse and productive ecosystems [@Chase2003].
+The taxonomic differences among communities, marked before disturbance by the distinct starting points of the NMDS, were maintained throughout communities recovery trajectories.
+More than commonly thought post-disturbance trajectories depended on communities initial composition, that partly determined the pool of rectuitable species and that constrained the converge towards the recovery of the initial composition [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
+The resilience of communities taxonomy suggested the existence of multiple stable equilibria, here the pre-disturbance communities, as assumed for highly diverse and productive ecosystems [@Chase2003].
 
-Contrastingly communities functional composition and diversity proved more homogeneous among communities and communities functional trajectories were similar in the functional space.
-Communities functional characteristics are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] so their resilience meant the recovery of ecosystem processes in the long term [@Guariguata2001].
+Communities functional composition followed similar trajectories that were not so distinct in the functional space.
+The trajectories of functional evenness as well were similar among communities 
+Communities functional characteristics are the most direct link between biodiversity and ecosystem functioning and their resilience here meant the recovery of ecosystem processes and services ona cycle of several decades [@Diaz2005;@Guariguata2001].
 
@@
 
-The initial disturbance intensity determined the maximum functional diversity increase, confirming the IDh within this range of disturbance.
-Regarding communities communities along time, the humped-shaped trajectories of the functional diversity following disturbance demonstrated the positive impact of disturbance that enhanced the diversity before a return to initial lower diversity levels.
-In accordance with the IDH disturbance entailed ecological processes different than in undisturbed communities and a temporal variability correlated to the intensity of the disturbance.
-The functional diversity increase paralleled the significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
-As the old-growth, pre-disturbance survivor trees proved mirroring the initial communities [@Herault2018] the functional trajectories and taxonomic turnover rather relied upon the pool of newly recruited trees driven by the enhanced growth and survival of previously infrequent species and functional types.
-The changes in abiotic environment and competitive pressure favored pioneers which outcompeted other species in non limiting light resources but were excluded in mature forests by long-lived, more resistant and shade tolerant species.
->> Not clear
-Disturbance trajectories therefore relied upon the environmental niches made available and filled by species from a restricted functional range that became dominant and structured the functional characteristics of the community [@Grime1998].
+The IDH consistently predicted communities functional response to disturbance and the functional trajectories in time highlighted the underlying ecological mechanisms, but the theory was diproved regarding communities taxonomic richness and evenness.
 
-In contrast the trajectories of taxonomic richness and evenness were decoupled from the disturbance intensity, questioning the consistency of the IDH to predict communities taxonomic diversity, as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+With the studied range of intensity, disturbance enhanced communities functional evenness but did not have clear impact on taxonomic richness or evenness. 
+The IDH proved a good predictor of the disturbance impact on communities functioning but not on their taxonomic richness or evenness.
 
-In the most disturbed plots the taxonomic richness did not exceed this of the initial communities but the taxonomic evenness significantly increased, revealing the loss of most infrequent species.
-Dominant species fastly recovered and restored the communities functoinal characteristics that mostly rely on dominant species, according to the "vegetation quantity effect" [@Grime1998], while the recovery of more infrequent species was hampered.
->> Not clear
-The intermediate disturbance hypothesis was suitabe to predict communities functional diversity and pool of dominant species [@Molino2001], but it poorly represented the whole taxonomic structure that was hampered by the slow recovery of rare species [@Hubbell2001;@Chave2004].
+The humped-shaped trajectories of communities along time demonstrated the enhanced functional evennes and changes in functional composition following disturbance before a return towards initial, lower values.
+In accordance with the IDH, disturbance entailed significant temporal variability with the emergence of ecological processes different from those rueling undisturbed forest.
+The increase in functional evenness went with significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
+Because the pre-disturbance survivor trees mirror the initial communities [@Herault2018], the functional changes relied upon recruited trees and the enhanced growth and survival of previously infrequent species and functional types.
+Pioneers outcompeted other species in the post-disturbance environment of non limiting resources and alleviated competition.
+Before disturbance and over communities recovery time though, these species would be excluded by long-lived, more resistant and shade tolerant species.
 
-## The functional redundancy, key of communities resilience
+In contrast the taxonomic richness and evenness were decoupled from the disturbance intensity, questioning the consistency of the IDH as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+The recovery of communities functional characteristics, that mostly rely on dominant species according to the "vegetation quantity effect" [@Grime1998], suggested a similar recovery of the dominant species [@Molino2001].
+Still, the taxonomic trajectories were blurried by the hampered recovery of infrequent species that entailed after disturbance a decrease in the taxonomic richness and an increase of the evenness [@Hubbell2001;@Chave2004].
+The disturbance therefore porrly predicted commuities taxonomic richnessa and evennes that rely upon different ecological processes.
 
-The recovery of the most infrequent species was hampered by the competition among species, in accordance with the lottery recruitment [@Busing2002] and entailed long term recovery cycles [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+## The functional redundancy, key of the taxonomic resilience
 
-The progressive increase of the functional redundancy within the initial functional space, that had first sharply decreased, explains the lottery recruitment that first allow an easy taxonomic recovery and then a progressive slow down.
-functional redundancy restricted to the functional space of the initial community clearly followed.
-After disturbance communities differently occupied the functional space, following the observed shifts in traits trajectories and functional composition, and the redundancy in the initial functional space decreased according to the disturbance intensity.
-After 30 years the recovery of the initial redundancy structure was consistent but remained unachieved.
+The recovery of the most infrequent species followed the lottery recruitment rules [@Busing2002] and was hampered by the increasing competition among species whih entailed long term recovery cycles [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
->>unclear
-The persistent alteration of functional redundancy meant a lower resilience of pre-disturbance communities and an higher resilience disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species) so chances to observe self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes were increased after disturbance [@Haddad2008; @Burslem2000; @Martin2013].
-
->>unclear (cornerstone?)
-This decades-long recovery specifically increased the risks of loosing cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
-Indeed the flora-faune relationship and the functional traits studied are not exhaustive and key function might be lost. 
-specifically for the fauna-flora relationships.
+The rapid decrease of the functional redundancy within the initial functional space, correlated with the disturbance intensity, illustrated the space and resources made available after disturbance.
+Following the lottery recruitment, communities taxonomic richness increased firsthand along the enhanced recruitment rate but decreased thereafter, following the emergence of exclusive competition. 
+After 30 years the recovery of the initial redundancy structure was underway but remained unachieved.
+This alteration of the functional redundancy meant a lower resilience of pre-disturbance communities and higher chances to see the persistence of disturbance-specific communities, with less species and more pioneers [@Haddad2008; @Burslem2000; @Martin2013].
+Besides the hampered recovery of rare species increased the risks to loose cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
+Apart from the functional characteristics considered here, infrequent species mighthave unique functions i nthe ecosystem or be a key for some fauna.
 
@@
 
-Our study defined the decoupled functional and taxonomic trajectories of tropical rainforests after disturbance demonstrating a rapid functional recovery but a slower and more variable taxonomic one.
-Consistently with the IDH, functional trajectories were driven after disturbance by enrichment of the system with pioneers and light-demanding species.
-
->>unclear
-The following functional shifts entailed a re-organization of communities functional redundancy that proved long to recover and explained the more variable and longer taxonomic trajectories as shade tolerant species of old-growth forests faced functional redundancy and limiting similarity processes.
-
-The resilience of tropical forests then proved long in the face of quite intense disturbance and did not preclude the settling of a persistent disturbance-specific community [@Gourlet-Fleury2005].
+Our study revealed communities cyclic recovery after disturbance entailing the resilience of their functioning and taxonomic composition with the maintenance of initial differences among communities.
+Communities functional evenness was enhanced for  20 years after disturbance through the enrichment of the communities with pioneers and light-demanding species,in accordance with the IDH.
+The IDH, though, poorly predicted the disturbance impact on communities taxonomic richness and evenness that were blurried by the emergence of competitive exclusion along time.
+The resilience of tropical forests proved consistent although spread over several decades. 
+Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Gourlet-Fleury2005].
 As the trajectories highlighted the recruitment processes proved central for communities response to disturbance and closer focus on demographical drivers of communities response would clarify the fate of the future forests.
 
->>unclear
-The disturbance range however stayed within the spectrum of selective logging, with a forest cover remaining all along the experiment, and the response mechanisms would probably be much different after harder disturbance.
-
 `r if (knitr:::is_html_output()) '# References {-}'`

### 2018-07-01 AppendixI.Rmd
intro

@@
 
-**Figure A4**: Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3 \@ref(fig:DivTaxo).
-
-**Figure A5**: Trajectories of the Rao functional diversity over 30 years after disturbance.
+**Figure A4**: Trajectories of the Rao functional diversity over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 
-**Figure A6**: Summarizing plots of the functional PCA analysis defining communities functional space.
+**Figure A5**: Summarizing plots of the functional PCA analysis defining communities functional space.
 **(a)** Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
@@
 
-**Figure A7**: Trajectories of the functional redundancy over 30 years after disturbance.
+**Figure A6**: Trajectories of the functional redundancy over 30 years after disturbance.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
@@
 
-# **Figure A4**: Richness, Shannon and Simpson taxonomic diversity trajectories
-
-```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
-load("DB/ReplacementTraj_ForGraphs")
-
-TaxoTraj(CompleteTaxo)
-
-```
-
-# **Figure A5**: Rao diversity trajectories with all plots display
+# **Figure A4**: Rao diversity trajectories with all plots display
 
@@
 
-# **Figure A6**: Simpson and Rao Linera regrassion summary
+# **Figure A5**: Simpson and Rao Linera regrassion summary
 
@@
 
-# **Figure A7**: Functional PCA
+# **Figure A6**: Functional PCA
 
@@
 
-# **Figure A8**: Overall functional redundancy
+# **Figure A7**: Overall functional redundancy
 

### 2018-07-01 WholePlotTrajectories.Rmd
intro

@@
     Specifically we examined over 30 years the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 leaf, stem and life history traits.
-    Trajectories highlighted communities taxonomic resilience, maintaining their initial differences in composition, and functional resilience through fast functional trajectory common to all communities in the functional space.
-    The IDH consistently predicted communities functional trajectories according to the disturbance intensity, but poorly represented their taxonomic trajectories, blurried by the alteration of the functional redundancy and the recruitment of infrequent, shade tolerant species slowed by competitive exclusion processes.
+    Trajectories highlighted the cyclic recovery of communities taxonomic composition, maintaining the pre-disturbance differences, as well as functional composition and diversity.
+    The IDH predicted communities functional response to disturbance that resulted from competition processes determined by the disturbance intensity.
+    The IDH however poorly represented the taxonomic response that relied upon lottery recruitment and recruitment limitation which hampered the recovery of infrequent, shade tolerant species.
 
-    Although communities functioning had recovered after 30 years their taxonomic diversity and composition in infrequent species and their functional redundancy remained altered.
-    these results acknowledged the need of decades-long recovery cycles to ensure communities complete resilience, and questioned communities maintenance after repeated disturbance.
+    After disturbance communities functioning followed a cyclic recovery and the initial composition differences among plots were maintained.
+    Although consistent the recovery of communities functioning and taxonomic composition and diversity remained unachieved after 30 years which acknowledged the need of decades-long recovery cycles to ensure a complete recovery, and questioned communities sustainability after repeated disturbance.
 
@@
 
-In tropical forest, ecological communities are constantly re-shaped by natural disturbance events that change both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions like competition among species [@Chesson2000;@Herault2018].
+In tropical forests, ecological communities are constantly re-shaped by natural disturbance events that change both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions like competition among species [@Chesson2000;@Herault2018].
 One of the cornerstone of tropical forest ecology is to understand the processes and drivers of ecosystems response to disturbance [@White2001;@Chazdon2003a]. 
@@
 However the response of forests diversity is still unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014, @Liang2016] and would be most probably impacted by the changes induced by disturbance [@Baraloto2012a].
+In the short-term, moderate disturbance may lead to positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity when the disturbance regime is moderate but significant [@Molino2001;@Kariuki2006a;@Berry2008a].
+Along time the IDH would translate into a variety of post-disturbance trajectories shaped by the different ecological processes at stake.
 
-In the short-term, moderate disturbance may lead to positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a].
-Still, validations of the IDH remain scarce in the long term and mainly rely on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-More complete analysis would besides encompass communities composition that is crucial for conservation issues, and abundance distribution to reveal the ecological rules structuring communities [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, a functional approach accounting for species biological attributes and role in the ecosystem directly links ecosystems biodiversity, functioning and environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-The functional trait-based approach focused on major traits related to species ecology and performance was sucessfully adopted [@Diaz2005;@Villeger2008a], for example to reveal in tropical rainforests the deterministic processes that foster after disturbance the fast growing species with efficient resources acquisition [@Molino2001;@Haddad2008;@Ruger2009].
-This basic deterministic process entails a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
+Validations of the IDH though remain scarce in the long term and mainly rely on the analysis of species richness.
+Species richness though gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016]: more complete analysis would encompass communities composition that is crucial for conservation issues, and abundance distribution that reveals the ecological processes involved [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Furthermore, a functional approach accounting for species biological attributes and assessing their role in the ecosystem would directly link communities diversity to the ecosystem functioning and to its environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+In that respect, the functional trait-based approach that focused on major traits related to species ecology and performance was sucessfully adopted [@Diaz2005;@Villeger2008a].
+For example the functoinal approach revealed in tropical rainforests the deterministic processes fostering fast growing species with efficient resources acquisition after disturbance [@Molino2001;@Haddad2008;@Ruger2009].
+Disturbance therefore entails a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
 This shift is translated into consistent trajectories of key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
-Besides communities functional composition and diversity, a last key aspect of communities response to disturbance is the changes in functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
-High functional redundancy, as in the very diverse tropical forests [@Bellwood2006], mitigates the impacts of species removal on ecosystem functioning and determines communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005].
+Eventually a complete overview of communities response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
+High functional redundancy like in the very diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities [@Trenbath1999; @Elmqvist2003; @Diaz2005].
 High functional redundancy also increases the functional overlap among species and would foster neutral stochastic processes [@Gravel2006].
 
-To grasp all facets of communities response to disturbance we examined here the taxonomic and functional trajectories in diversity and composition [@Lohbeck2015;@Guariguata2001].
-These trajectories would highlight the recovery of communities initial composition, diversity and functioning and the underlying ecological processes.
-They would clarify the tenants in the long term of the Intermediate Disturbance Hypothesis debated in tropical forests, and provide indications for future adaptive conservation strategies [@Adler2007].
+To grasp all facets of communities response to disturbance we examined here the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy [@Lohbeck2015;@Guariguata2001].
+These trajectories would highlight the recovery of communities initial characteristics and the underlying ecological processes, and provide indications for future adaptive conservation strategies [@Adler2007].
+They would clarify the tenants of the Intermediate Disturbance Hypothesis in the long term, debated in tropical forests, and explicit the variety of post-disturbance trajectories defined by the disturbance intensity.
 We monitored over 30 years the response of 75 ha of neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
-We made use of a large functional traits database browsing major leaf, stem and seed functional traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
-Specifically, we (i) questioned the coupling between taxonomic and functional response to disturbance and identified the underlying assembly processes, (ii) clarified the validity of the IDH in the long term for tropical forest and (iii) questioned the completeness of communities recovery regarding communities functional redundancy.
+We made use of a large functional traits database browsing major leaf, stem and seed functional traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition, richness, evenness and redundancy.
+Specifically, we (i) questioned the recovery of communities taxonomic and functional characteristics and identify the underlying assembly processes, (ii) clarified the validity of the IDH in the long term for tropical forest and elucidated its declination into different trajectories, and (iii) questioned the length of complete communities recovery.
 
@@
 
-## Inventories protocol and dataset collection {#protocols}
+## Inventories protocol and dataset collection
 
@@
 Trajectories along time were reported through the distance between successive inventories and the reference inventories in 1989, *i.e.* 2-3 years after disturbance, when the uncertainty degree did not exceed 30% of undetermined trees.
-Taxonomic trajectories were measured through the euclidean distance between inventories in the multivariate space of all inventories species.
-Functional trajectories were measured through the sum of euclidean distances among traits density distributions between inventories.
+for the taxonomix composition trajectories were measured by the euclidean distance between inventories in the multivariate space of all inventories species.
+For the functoinal composition trajectories were measured by the sum for all traits of the euclidean distances between the traits density distributions of successive inventories.
 To compensate the intrinsinc difference among plots, the trajectories corresponded to the euclidean differences along time with the reference inventory in 1989.
@@
 
-The taxonomic diversity was assessed through Richness and the Hill number translation of Shannon and Simpson indices [@Hill1973].
-These three indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0, 1 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995;@Marcon2015].
+The taxonomic diversity was reported through species richness and the evenness, *i.e* the Hill number translation of the Simpson index [@Hill1973].
+These two indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995;@Marcon2015].
 The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on all functional and life traits.
 
-The impacts of initial disturbance were first tested with the spearman rank correlation between the extremum of taxonomic and functional metrics reached over the 30 years and the initial %AGB removed.
-Then they were analysed through the linear correlations between Simpson and Rao diversities and the initial %AGB removed at 10, 20 and 30 years after disturbance.
+The impacts of initial disturbance were tested with the spearman rank correlation between the extremum of taxonomic and functional metrics reached over the 30 years and the initial %AGB removed. 
+The were besides analysed through the linear correlations between Simpson and Rao diversities and the initial %AGB removed at 10, 20 and 30 years after disturbance.
 
@@
 
-## Communities Diversity
+## Communities richness and evenness
 
 From 1989 (2-3 years after disturbance) to 2015 (28-29 years after disturbance), 828388 individual trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
-For undisturbed plots Richness and taxonomic evenness (Shannon and Simpson diversities) remained stable over the 30 years monitored.
-In disturbed communities the taxonomic richness increased after low disturbance intensity, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2) while it followed unimodal trajectories after intense disturbance, decreasing for ten years before recovering to pre-disturbance values.
-In all disturbed plots the taxonomic evenness (Shannon and Simpson diversities) increased, following unimodal trajectories with a maximum, reached after around 20 years, positively correlated to the disturbance treatment ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
-Return towards initial evenness values was beginning after 30 years except for two T3 plots (plots 8 and 12) which evenness still increased, suggesting similar but delayed trajectories \@ref(fig:IDHplot).
+For undisturbed plots taxonomic Richness and Evenness (Simpson diversity) remained stable over the 30 years monitored.
+In disturbed communities after low disturbance intensity the taxonomic richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
+After intense disturbance the taxonomic richness followed unimodal trajectories, decreasing for ten years after disturbance before recovering to pre-disturbance values.
+In all disturbed plots the taxonomic evenness (Simpson diversity) followed unimodal trajectories with first an increase until a maximum reached after around 20 years that was positively correlated to the disturbance treatment ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
+The return towards initial values of communities evenness was beginning after 30 years except for two T3 plots (plots 8 and 12) which evenness still increased, suggesting similar but delayed trajectories \@ref(fig:DivTaxo).
 
-Trajectories of communities functional diversity were examined through the Rao diversity based on the 7 leaf, stem and life history traits (to the exception of seed mass).
-The plot 7 from treatment 1 displayed a constantly outlying diversity and was removed from the graphical representation for better readability (see appendix for full graphs).
-In undisturbed plots the functional diversity remained stable along the 30 years while in disturbed plots it followed unimodal trajectories with a return towards initial values that strated around 20 years after disturbance.
+```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
+load("DB/ReplacementTraj_ForGraphs")
+
+TaxoTraj(CompleteTaxo)
+
+```
+
+The plot 7 from treatment 1 displayed a constantly outlying functional evenness and was removed from the graphical representation for better readability (see appendix for full graphs).
+In undisturbed plots the functional evenness remained stable along the 30 years while in disturbed plots it followed unimodal trajectories with a return towards initial values starting around 20 years after disturbance.
+
+```{r FunDiv, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
+load("DB/FunctionalTraj_ForGraphs")
+
+plotDiv(CompleteFun,remove=TRUE)
+mtext("Rao diversity",side=3,adj=0,line=1)
+mtext("Years since disturbance",side=1,line=2.2,adj=1)
+legend("right",inset=c(-0.28,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
+```
 
@@
 
-```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Upper panels, Trajectories of the Simpson taxonomic diversity (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance.
+```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance.
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.
@@
 
-par(mfrow=c(2,2),mar=c(3,3,2,1),oma=c(2,1,2,3),no.readonly = T)
+par(mfrow=c(1,2),mar=c(3,3,2,1),oma=c(2,1,2,3),no.readonly = T)
 plotDiv(SimpsonTaxo)
@@
 
->> Si on met que richesse et evenness, pour functional c'est de l'evenness (Rao = order2)
-
 ## Communities cyclic recovery
@@
 The taxonomic differences among communities, marked before disturbance by the distinct starting points of the NMDS, were maintained throughout communities recovery trajectories.
-More than commonly thought post-disturbance trajectories depended on communities initial composition, that partly determined the pool of rectuitable species and that constrained the converge towards the recovery of the initial composition [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
+More than commonly thought post-disturbance trajectories depended on communities initial composition, that partly determined the pool of recruitable species and that constrained the converge towards the recovery of the initial composition [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
 The resilience of communities taxonomy suggested the existence of multiple stable equilibria, here the pre-disturbance communities, as assumed for highly diverse and productive ecosystems [@Chase2003].
@@
 
-The humped-shaped trajectories of communities along time demonstrated the enhanced functional evennes and changes in functional composition following disturbance before a return towards initial, lower values.
-In accordance with the IDH, disturbance entailed significant temporal variability with the emergence of ecological processes different from those rueling undisturbed forest.
+The humped-shaped trajectories of communities along time illustrated the processes enhancing the functional evennes and the changes in functional composition following disturbance before communities return towards initial composition and lower evenness.
+Disturbance entailed significant variability in the communities through the emergence of ecological processes different from thoses involvemnt before disturbance, in accordance with the IDH.
 The increase in functional evenness went with significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
 Because the pre-disturbance survivor trees mirror the initial communities [@Herault2018], the functional changes relied upon recruited trees and the enhanced growth and survival of previously infrequent species and functional types.
-Pioneers outcompeted other species in the post-disturbance environment of non limiting resources and alleviated competition.
+Disturbance vacated environmental niches of high light, space and nutrient availability, filled by competitive pioneers that became dominant and determined the functional characteristics of the community [@Grime1998].
 Before disturbance and over communities recovery time though, these species would be excluded by long-lived, more resistant and shade tolerant species.
@@
 
-The recovery of the most infrequent species followed the lottery recruitment rules [@Busing2002] and was hampered by the increasing competition among species whih entailed long term recovery cycles [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+The recovery of the most infrequent species followed the lottery recruitment rules [@Busing2002] and was hampered by the increasing competition among species, implying long term recovery [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
-The rapid decrease of the functional redundancy within the initial functional space, correlated with the disturbance intensity, illustrated the space and resources made available after disturbance.
-Following the lottery recruitment, communities taxonomic richness increased firsthand along the enhanced recruitment rate but decreased thereafter, following the emergence of exclusive competition. 
-After 30 years the recovery of the initial redundancy structure was underway but remained unachieved.
+Following disturbance the space and resources made available corresponded to a rapid decrease of the functional redundancy within the initial functional space.
+The high light, space and nutrient niches were rapidly filled by species, mainly the most dominant and frequent ones.
+Thereafter competitive exclusion emerged following the filling of the niches and limited the recruitment of infrequent species [@Busing2002].
+The recovery of communities taxonomic richness and evenness relied upon the recovery of the initial functional redudancy.
+Although underway 30 years after disturbance the recovery remained unachieved and after 30 years the recovery it was underway but remained unachieved.
 This alteration of the functional redundancy meant a lower resilience of pre-disturbance communities and higher chances to see the persistence of disturbance-specific communities, with less species and more pioneers [@Haddad2008; @Burslem2000; @Martin2013].
-Besides the hampered recovery of rare species increased the risks to loose cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
+Besides the slowed recovery of rare species increased the risks to loose cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
 Apart from the functional characteristics considered here, infrequent species mighthave unique functions i nthe ecosystem or be a key for some fauna.
@@
 
-Our study revealed communities cyclic recovery after disturbance entailing the resilience of their functioning and taxonomic composition with the maintenance of initial differences among communities.
+Our study revealed communities cyclic recovery after disturbance allowing the resilience of their functioning and taxonomic composition with the maintenance of initial differences among communities.
 Communities functional evenness was enhanced for  20 years after disturbance through the enrichment of the communities with pioneers and light-demanding species,in accordance with the IDH.

### 2018-07-01 WholePlotTrajectories.tex
intro

@@
 To grasp all facets of communities response to disturbance we examined
-here the taxonomic and functional trajectories in diversity and
-composition \citep{Lohbeck2015, Guariguata2001}. These trajectories
-would highlight the recovery of communities initial composition,
-diversity and functioning and the underlying ecological processes. They
-would clarify the tenants in the long term of the Intermediate
-Disturbance Hypothesis debated in tropical forests, and provide
-indications for future adaptive conservation strategies
-\citep{Adler2007}. We monitored over 30 years the response of 75 ha of
-neotropical forest plots set up on a gradient of disturbance intensity,
-from 10 to 60\% of ecosystem biomass removed. We made use of a large
-functional traits database browsing major leaf, stem and seed functional
-traits and species maximum height to draw the trajectories over time of
-communities taxonomic and functional composition and diversity.
-Specifically, we (i) questioned the coupling between taxonomic and
-functional response to disturbance and identified the underlying
-assembly processes, (ii) clarified the validity of the IDH in the long
-term for tropical forest and (iii) questioned the completeness of
-communities recovery regarding communities functional redundancy.
+here the taxonomic and functional trajectories in terms of richness,
+evennes and composition \citep{Lohbeck2015, Guariguata2001}. These
+trajectories would highlight the recovery of communities initial
+characteristics and the underlying ecological processes. They would
+clarify the tenants in the long term of the Intermediate Disturbance
+Hypothesis debated in tropical forests, and provide indications for
+future adaptive conservation strategies \citep{Adler2007}. We monitored
+over 30 years the response of 75 ha of neotropical forest plots set up
+on a gradient of disturbance intensity, from 10 to 60\% of ecosystem
+biomass removed. We made use of a large functional traits database
+browsing major leaf, stem and seed functional traits and species maximum
+height to draw the trajectories over time of communities taxonomic and
+functional composition,richness and evenness. Specifically, we (i)
+questioned the coupling between taxonomic and functional response to
+disturbance and identified the underlying assembly processes, (ii)
+clarified the validity of the IDH in the long term for tropical forest
+and (iii) questioned the completeness of communities recovery regarding
+communities functional redundancy.
 
@@
 \subsection{Inventories protocol and dataset
-collection}\label{protocols}
+collection}\label{inventories-protocol-and-dataset-collection}
 
@@
 disturbance, when the uncertainty degree did not exceed 30\% of
-undetermined trees. Taxonomic trajectories were measured through the
-euclidean distance between inventories in the multivariate space of all
-inventories species. Functional trajectories were measured through the
-sum of euclidean distances among traits density distributions between
-inventories. To compensate the intrinsinc difference among plots, the
-trajectories corresponded to the euclidean differences along time with
-the reference inventory in 1989. Univariate trajectories of the leaf,
-stem and life-history traits were also visualized with the community
-weighted means (CWM) \citep{Diaz2007, Garnier2004}. Species seed mass
-corresponded to 5 classes of increasing mass, seed mass trajectories
-were therefore reported as the proportion of each class in the
-inventories.
+undetermined trees. for the taxonomix composition trajectories were
+measured by the euclidean distance between inventories in the
+multivariate space of all inventories species. For the functoinal
+composition trajectories were measured by the sum for all traits of the
+euclidean distances between the traits density distributions of
+successive inventories. To compensate the intrinsinc difference among
+plots, the trajectories corresponded to the euclidean differences along
+time with the reference inventory in 1989. Univariate trajectories of
+the leaf, stem and life-history traits were also visualized with the
+community weighted means (CWM) \citep{Diaz2007, Garnier2004}. Species
+seed mass corresponded to 5 classes of increasing mass, seed mass
+trajectories were therefore reported as the proportion of each class in
+the inventories.
 
-The taxonomic diversity was assessed through Richness and the Hill
-number translation of Shannon and Simpson indices \citep{Hill1973}.
-These three indices belong to the set of HCDT or generalized entropy,
-respectively corresponding to the 0, 1 and 2 order of diversity (q),
-recomended for diversity studies
+The taxonomic diversity was reported through species richness and the
+evenness, \emph{i.e} the Hill number translation of the Simpson index
+\citep{Hill1973}. These two indices belong to the set of HCDT or
+generalized entropy, respectively corresponding to the 0 and 2 order of
+diversity (q), recomended for diversity studies
 \citep{Patil1982, Tothmeresz1995, Marcon2015}. The functional diversity
@@
 
-The impacts of initial disturbance were first tested with the spearman
-rank correlation between the extremum of taxonomic and functional
-metrics reached over the 30 years and the initial \%AGB removed. Then
-they were analysed through the linear correlations between Simpson and
-Rao diversities and the initial \%AGB removed at 10, 20 and 30 years
-after disturbance.
+The impacts of initial disturbance were tested with the spearman rank
+correlation between the extremum of taxonomic and functional metrics
+reached over the 30 years and the initial \%AGB removed. The were
+besides analysed through the linear correlations between Simpson and Rao
+diversities and the initial \%AGB removed at 10, 20 and 30 years after
+disturbance.
 
@@
 
-\subsection{Communities Diversity}\label{communities-diversity}
+\subsection{Communities richness and
+evenness}\label{communities-richness-and-evenness}
 
@@
 pertaining to 223 genus and 64 botanical families were recorded. For
-undisturbed plots Richness and taxonomic evenness (Shannon and Simpson
-diversities) remained stable over the 30 years monitored. In disturbed
-communities the taxonomic richness increased after low disturbance
-intensity, reaching a maximum gain of 14 botanical genera (plot 3 from
-treatment 2) while it followed unimodal trajectories after intense
-disturbance, decreasing for ten years before recovering to
-pre-disturbance values. In all disturbed plots the taxonomic evenness
-(Shannon and Simpson diversities) increased, following unimodal
-trajectories with a maximum, reached after around 20 years, positively
-correlated to the disturbance treatment
+undisturbed plots taxonomic Richness and Evenness (Simpson diversity)
+remained stable over the 30 years monitored. In disturbed communities
+after low disturbance intensity the taxonomic richness increased,
+reaching a maximum gain of 14 botanical genera (plot 3 from treatment
+2). After intense disturbance the taxonomic richness followed unimodal
+trajectories, decreasing for ten years after disturbance before
+recovering to pre-disturbance values. In all disturbed plots the
+taxonomic evenness (Simpson diversity) followed unimodal trajectories
+with first an increase until a maximum reached after around 20 years
+that was positively correlated to the disturbance treatment
 (\(\rho_{spearman}^{Shannon}=0.86\), and
-\(\rho_{spearman}^{Simpson}=0.89\)). Return towards initial evenness
-values was beginning after 30 years except for two T3 plots (plots 8 and
-12) which evenness still increased, suggesting similar but delayed
-trajectories \ref{fig:IDHplot}.
+\(\rho_{spearman}^{Simpson}=0.89\)). The return towards initial values
+of communities evenness was beginning after 30 years except for two T3
+plots (plots 8 and 12) which evenness still increased, suggesting
+similar but delayed trajectories \ref{fig:DivTaxo}.
 
-Trajectories of communities functional diversity were examined through
-the Rao diversity based on the 7 leaf, stem and life history traits (to
-the exception of seed mass). The plot 7 from treatment 1 displayed a
-constantly outlying diversity and was removed from the graphical
-representation for better readability (see appendix for full graphs). In
-undisturbed plots the functional diversity remained stable along the 30
-years while in disturbed plots it followed unimodal trajectories with a
-return towards initial values that strated around 20 years after
-disturbance.
+\begin{figure*}
+
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivTaxo-1}
+
+}
+
+\caption{Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities \textbf{(a)} Richness, \textbf{(b)} Shannon and \textbf{(c)} Simpson diversities.
+Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:DivTaxo}
+\end{figure*}
+
+The plot 7 from treatment 1 displayed a constantly outlying functoinal
+evenness and was removed from the graphical representation for better
+readability (see appendix for full graphs). In undisturbed plots the
+functional evenness remained stable along the 30 years while in
+disturbed plots it followed unimodal trajectories with a return towards
+initial values starting around 20 years after disturbance.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/FunDiv-1}
+
+}
+
+\caption{Trajectories of the Rao functional diversity over 30 years after disturbance.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:FunDiv}
+\end{figure}
 
@@
 
-\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance.
+\caption{Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance.
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.
-Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities 10, 20 and 30 years after disturbance.}\label{fig:IDHplot}
+Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities 10, 20 and 30 years after disturbance.}\label{fig:IDHplot1}
+\end{figure*}\begin{figure*}
+
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/IDHplot-2}
+
+}
+
+\caption{Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.
+Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities 10, 20 and 30 years after disturbance.}\label{fig:IDHplot2}
 \end{figure*}
@@
 
-Both taxonomic and functional composition proved resilient and followed
-similar humped-shaped trajectories starting to return towards initial
-values 30 years after disturbance.
+Communities taxonomic and functional composition and functional evenness
+proved resilient, following similar humped-shaped trajectories starting
+to return towards pre-disturbance composition and functioning after 30
+years.
 
-The taxonomic NMDS revealed differences in the composition of initial
-communities that proved maintained throughout the recovery trajectories
-after disturbance. More than commonly though, communities trajectories
-after disturbance was specific to the initial composition and converged
-towards its recovery
+The taxonomic differences among communities, marked before disturbance
+by the distinct starting points of the NMDS, were maintained throughout
+communities recovery trajectories. More than commonly thought
+post-disturbance trajectories depended on communities initial
+composition, that partly determined the pool of rectuitable species and
+that constrained the converge towards the recovery of the initial
+composition
 \citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}. The
-resilience of commnities taxonomy suggested the existence of multiple
-stable equilibria, corresponding here to communities initial
-composition, as assumed for highly diverse and productive ecosystems
-\citep{Chase2003}.
+resilience of communities taxonomy suggested the existence of multiple
+stable equilibria, here the pre-disturbance communities, as assumed for
+highly diverse and productive ecosystems \citep{Chase2003}.
 
-Contrastingly communities functional composition and diversity proved
-more homogeneous among communities and communities functional
-trajectories were similar in the functional space. Communities
+Communities functional composition followed similar trajectories that
+were not so distinct in the functional space. The trajectories of
+functional evenness as well were similar among communities Communities
 functional characteristics are the most direct link between biodiversity
-and ecosystem functioning \citep{Diaz2005} so their resilience meant the
-recovery of ecosystem processes in the long term \citep{Guariguata2001}.
+and ecosystem functioning and their resilience here meant the recovery
+of ecosystem processes and services ona cycle of several decades
+\citep{Diaz2005, Guariguata2001}.
 
@@
 
-The initial disturbance intensity determined the maximum functional
-diversity increase, confirming the IDh within this range of disturbance.
-Regarding communities communities along time, the humped-shaped
-trajectories of the functional diversity following disturbance
-demonstrated the positive impact of disturbance that enhanced the
-diversity before a return to initial lower diversity levels. In
-accordance with the IDH disturbance entailed ecological processes
-different than in undisturbed communities and a temporal variability
-correlated to the intensity of the disturbance. The functional diversity
-increase paralleled the significant functional shifts towards
-resource-acquisitive strategies (sharp increase in the SLA, leaf
-thickness and bark thickness and decrease in wood density, leaf
-toughness and maximum height)
-\citep{Westoby1998, Wright2004, Reich2014}. As the old-growth,
-pre-disturbance survivor trees proved mirroring the initial communities
-\citep{Herault2018} the functional trajectories and taxonomic turnover
-rather relied upon the pool of newly recruited trees driven by the
-enhanced growth and survival of previously infrequent species and
-functional types. The changes in abiotic environment and competitive
-pressure favored pioneers which outcompeted other species in non
-limiting light resources but were excluded in mature forests by
-long-lived, more resistant and shade tolerant species.
-\textgreater{}\textgreater{} Not clear Disturbance trajectories
-therefore relied upon the environmental niches made available and filled
-by species from a restricted functional range that became dominant and
-structured the functional characteristics of the community
-\citep{Grime1998}.
+The IDH consistently predicted communities functional response to
+disturbance and the functional trajectories in time highlighted the
+underlying ecological mechanisms, but the theory was diproved regarding
+communities taxonomic richness and evenness.
 
-In contrast the trajectories of taxonomic richness and evenness were
-decoupled from the disturbance intensity, questioning the consistency of
-the IDH to predict communities taxonomic diversity, as already observed
-in the Guiana Shield \citep{Baraloto2012a} and in Bornean tropical
-forests \citep{Cannon1998}.
+With the studied range of intensity, disturbance enhanced communities
+functional evenness but did not have clear impact on taxonomic richness
+or evenness. The IDH proved a good predictor of the disturbance impact
+on communities functioning but not on their taxonomic richness or
+evenness.
 
-In the most disturbed plots the taxonomic richness did not exceed this
-of the initial communities but the taxonomic evenness significantly
-increased, revealing the loss of most infrequent species. Dominant
-species fastly recovered and restored the communities functoinal
-characteristics that mostly rely on dominant species, according to the
-``vegetation quantity effect'' \citep{Grime1998}, while the recovery of
-more infrequent species was hampered. \textgreater{}\textgreater{} Not
-clear The intermediate disturbance hypothesis was suitabe to predict
-communities functional diversity and pool of dominant species
-\citep{Molino2001}, but it poorly represented the whole taxonomic
-structure that was hampered by the slow recovery of rare species
-\citep{Hubbell2001, Chave2004}.
+The humped-shaped trajectories of communities along time demonstrated
+the enhanced functional evennes and changes in functional composition
+following disturbance before a return towards initial, lower values. In
+accordance with the IDH, disturbance entailed significant temporal
+variability with the emergence of ecological processes different from
+those rueling undisturbed forest. The increase in functional evenness
+went with significant functional shifts towards resource-acquisitive
+strategies (sharp increase in the SLA, leaf thickness and bark thickness
+and decrease in wood density, leaf toughness and maximum height)
+\citep{Westoby1998, Wright2004, Reich2014}. Because the pre-disturbance
+survivor trees mirror the initial communities \citep{Herault2018}, the
+functional changes relied upon recruited trees and the enhanced growth
+and survival of previously infrequent species and functional types.
+Pioneers outcompeted other species in the post-disturbance environment
+of non limiting resources and alleviated competition. Before disturbance
+and over communities recovery time though, these species would be
+excluded by long-lived, more resistant and shade tolerant species.
 
-\subsection{The functional redundancy, key of communities
-resilience}\label{the-functional-redundancy-key-of-communities-resilience}
+In contrast the taxonomic richness and evenness were decoupled from the
+disturbance intensity, questioning the consistency of the IDH as already
+observed in the Guiana Shield \citep{Baraloto2012a} and in Bornean
+tropical forests \citep{Cannon1998}. The recovery of communities
+functional characteristics, that mostly rely on dominant species
+according to the ``vegetation quantity effect'' \citep{Grime1998},
+suggested a similar recovery of the dominant species \citep{Molino2001}.
+Still, the taxonomic trajectories were blurried by the hampered recovery
+of infrequent species that entailed after disturbance a decrease in the
+taxonomic richness and an increase of the evenness
+\citep{Hubbell2001, Chave2004}. The disturbance therefore porrly
+predicted commuities taxonomic richnessa and evennes that rely upon
+different ecological processes.
 
-The recovery of the most infrequent species was hampered by the
-competition among species, in accordance with the lottery recruitment
-\citep{Busing2002} and entailed long term recovery cycles
+\subsection{The functional redundancy, key of the taxonomic
+resilience}\label{the-functional-redundancy-key-of-the-taxonomic-resilience}
+
+The recovery of the most infrequent species followed the lottery
+recruitment rules \citep{Busing2002} and was hampered by the increasing
+competition among species whih entailed long term recovery cycles
 \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-The progressive increase of the functional redundancy within the initial
-functional space, that had first sharply decreased, explains the lottery
-recruitment that first allow an easy taxonomic recovery and then a
-progressive slow down. functional redundancy restricted to the
-functional space of the initial community clearly followed. After
-disturbance communities differently occupied the functional space,
-following the observed shifts in traits trajectories and functional
-composition, and the redundancy in the initial functional space
-decreased according to the disturbance intensity. After 30 years the
-recovery of the initial redundancy structure was consistent but remained
-unachieved.
-
-\begin{quote}
-\begin{quote}
-unclear The persistent alteration of functional redundancy meant a lower
-resilience of pre-disturbance communities and an higher resilience
-disturbance-specific communities (corresponding to lower richness and
-higher dominance of pioneer species) so chances to observe
-self-maintained compositional changes in favor of disturbance resistant
-species, lianas or epiphytes were increased after disturbance
-\citep{Haddad2008, Burslem2000, Martin2013}.
-\end{quote}
-\end{quote}
-
-\begin{quote}
-\begin{quote}
-unclear (cornerstone?) This decades-long recovery specifically increased
-the risks of loosing cornerstone specie, with unexpected ecological
-consequences \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.
-Indeed the flora-faune relationship and the functional traits studied
-are not exhaustive and key function might be lost. specifically for the
-fauna-flora relationships.
-\end{quote}
-\end{quote}
+The rapid decrease of the functional redundancy within the initial
+functional space, correlated with the disturbance intensity, illustrated
+the space and resources made available after disturbance. Following the
+lottery recruitment, communities taxonomic richness increased firsthand
+along the enhanced recruitment rate but decreased thereafter, following
+the emergence of exclusive competition. After 30 years the recovery of
+the initial redundancy structure was underway but remained unachieved.
+This alteration of the functional redundancy meant a lower resilience of
+pre-disturbance communities and higher chances to see the persistence of
+disturbance-specific communities, with less species and more pioneers
+\citep{Haddad2008, Burslem2000, Martin2013}. Besides the hampered
+recovery of rare species increased the risks to loose cornerstone
+specie, with unexpected ecological consequences
+\citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}. Apart from the
+functional characteristics considered here, infrequent species mighthave
+unique functions i nthe ecosystem or be a key for some fauna.
 
@@
 
-Our study defined the decoupled functional and taxonomic trajectories of
-tropical rainforests after disturbance demonstrating a rapid functional
-recovery but a slower and more variable taxonomic one. Consistently with
-the IDH, functional trajectories were driven after disturbance by
-enrichment of the system with pioneers and light-demanding species.
-
-\begin{quote}
-\begin{quote}
-unclear The following functional shifts entailed a re-organization of
-communities functional redundancy that proved long to recover and
-explained the more variable and longer taxonomic trajectories as shade
-tolerant species of old-growth forests faced functional redundancy and
-limiting similarity processes.
-\end{quote}
-\end{quote}
-
-The resilience of tropical forests then proved long in the face of quite
-intense disturbance and did not preclude the settling of a persistent
-disturbance-specific community \citep{Gourlet-Fleury2005}. As the
-trajectories highlighted the recruitment processes proved central for
-communities response to disturbance and closer focus on demographical
-drivers of communities response would clarify the fate of the future
-forests.
-
-\begin{quote}
-\begin{quote}
-unclear The disturbance range however stayed within the spectrum of
-selective logging, with a forest cover remaining all along the
-experiment, and the response mechanisms would probably be much different
-after harder disturbance.
-\end{quote}
-\end{quote}
+Our study revealed communities cyclic recovery after disturbance
+entailing the resilience of their functioning and taxonomic composition
+with the maintenance of initial differences among communities.
+Communities functional evenness was enhanced for 20 years after
+disturbance through the enrichment of the communities with pioneers and
+light-demanding species,in accordance with the IDH. The IDH, though,
+poorly predicted the disturbance impact on communities taxonomic
+richness and evenness that were blurried by the emergence of competitive
+exclusion along time. The resilience of tropical forests proved
+consistent although spread over several decades. Still, the disturbance
+impact on communities redundancy cautioned against the risks of
+infrequent species loss and the persistence of disturbance-specific
+communities \citep{Gourlet-Fleury2005}. As the trajectories highlighted
+the recruitment processes proved central for communities response to
+disturbance and closer focus on demographical drivers of communities
+response would clarify the fate of the future forests.
 

### 2018-07-01 Scripts/DrawingTools.R

@@ -87,6 +87,39 @@ mtext("Euclidean distance from 1985 inventory",side=2,padj=0,line=2,cex=0.8)
 mtext("(c)",side=3,adj=0,line=0.5)
 }
 
+TaxoTraj<-function(CompTaxo){
+  par(mfrow=c(1,2),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
+  for(q in c(1,3)){     
+    Toplot<-lapply(CompTaxo,function(tr){return(tr[,,,q])})
+    Toplot<-lapply(Toplot,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
+    Toplot<-lapply(Toplot,function(tr){
+      ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col-tr[,1,rep]}))})#
+      ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),
+                 dimnames=list(rownames(ret[[1]]),as.numeric(colnames(ret[[1]]))-1986,1:length(ret)))
+      ret<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(ret,c(1,2),function(x){return(quantile(x,probs=quant))}))})
+      return(array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),3),
+                   dimnames=list(rownames(ret[[1]]),colnames(ret[[1]]),c(0.025,0.5,0.975))))})
+    
+    plot(colnames(Toplot[[1]]),Toplot[[1]][1,,1],type="n",xaxt="n",
+         xlab="",ylab="",ylim=c(min(unlist(Toplot),na.rm=T),max(unlist(Toplot),na.rm=T)))
+    axis(1,at=as.character(seq(5,33,5)),labels=TRUE)  
+    mtext(paste("(",c("a","b","c")[q],") ",c("Richness","Shannon","Simpson")[q],sep=""),
+          line=1,side=3,adj=0)
+    
+    invisible(lapply(1:4,function(t){  
+      toplot<-Toplot[[t]]
+      
+      invisible(lapply(1:3,function(i){
+        absc<-colnames(Toplot[[t]])
+        lines(absc,Toplot[[t]][i,,"0.5"], col = ColorsTr[[t]],lty = 1,lwd=2)
[…]

### 2018-07-02 WholePlotTrajectories.tex
maj

@@
 functional composition, diversity and redundancy based on 7 leaf, stem
-and life history traits. Trajectories highlighted communities taxonomic
-resilience, maintaining their initial differences in composition, and
-functional resilience through fast functional trajectory common to all
-communities in the functional space. The IDH consistently predicted
-communities functional trajectories according to the disturbance
-intensity, but poorly represented their taxonomic trajectories, blurried
-by the alteration of the functional redundancy and the recruitment of
-infrequent, shade tolerant species slowed by competitive exclusion
-processes. Although communities functioning had recovered after 30 years
-their taxonomic diversity and composition in infrequent species and
-their functional redundancy remained altered. these results acknowledged
-the need of decades-long recovery cycles to ensure communities complete
-resilience, and questioned communities maintenance after repeated
-disturbance.
+and life history traits. Trajectories highlighted the cyclic recovery of
+communities taxonomic composition, maintaining the pre-disturbance
+differences, as well as functional composition and diversity. The IDH
+predicted communities functional response to disturbance that resulted
+from competition processes determined by the disturbance intensity. The
+IDH however poorly represented the taxonomic response that relied upon
+lottery recruitment and recruitment limitation which hampered the
+recovery of infrequent, shade tolerant species. After disturbance
+communities functioning followed a cyclic recovery and the initial
+composition differences among plots were maintained. Although consistent
+the recovery of communities functioning and taxonomic composition and
+diversity remained unachieved after 30 years which acknowledged the need
+of decades-long recovery cycles to ensure a complete recovery, and
+questioned communities sustainability after repeated disturbance.
 }
@@
 
-In tropical forest, ecological communities are constantly re-shaped by
+In tropical forests, ecological communities are constantly re-shaped by
 natural disturbance events that change both the abiotic environment,
@@
 would be most probably impacted by the changes induced by disturbance
-\citep{Baraloto2012a}.
+\citep{Baraloto2012a}. In the short-term, moderate disturbance may lead
+to positive impacts on communities diversity, which have been formalized
+by the intermediate disturbance hypothesis (IDH) stating a maximized
+species diversity when the disturbance regime is moderate but
+significant \citep{Molino2001, Kariuki2006a, Berry2008a}. Along time the
+IDH would translate into a variety of post-disturbance trajectories
+shaped by the different ecological processes at stake.
 
-In the short-term, moderate disturbance may lead to positive impacts on
-communities diversity, which have been formalized by the intermediate
-disturbance hypothesis (IDH) stating a maximized species diversity at
-intermediate disturbance intensity
-\citep{Molino2001, Kariuki2006a, Berry2008a}. Still, validations of the
-IDH remain scarce in the long term and mainly rely on the analysis of
-species richness that gives limited or misleading information on forests
-recovery and functioning \citep{Martin2015, Chaudhary2016}. More
-complete analysis would besides encompass communities composition that
-is crucial for conservation issues, and abundance distribution to reveal
-the ecological rules structuring communities
-\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
+Validations of the IDH though remain scarce in the long term and mainly
+rely on the analysis of species richness. Species richness though gives
+limited or misleading information on forests recovery and functioning
+\citep{Martin2015, Chaudhary2016}: more complete analysis would
+encompass communities composition that is crucial for conservation
+issues, and abundance distribution that reveals the ecological processes
+involved \citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
 functional approach accounting for species biological attributes and
-role in the ecosystem directly links ecosystems biodiversity,
-functioning and environmental constraints
-\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. The
-functional trait-based approach focused on major traits related to
-species ecology and performance was sucessfully adopted
-\citep{Diaz2005, Villeger2008a}, for example to reveal in tropical
-rainforests the deterministic processes that foster after disturbance
-the fast growing species with efficient resources acquisition
-\citep{Molino2001, Haddad2008, Ruger2009}. This basic deterministic
-process entails a functional shift from a dominance of ``conservative''
-slow-growing species dealing with scarce resources to ``acquisitive''
-fast-growing species with rapid and efficient use of abundant resources
-\citep{TerSteege2001, Reich2014, Herault2011}. This shift is translated
-into consistent trajectories of key functional traits related to
-resource acquisition (leaf area, density and chlorophyll content, and
-stem specific gravity and bark thickness), tree growth and reproduction
-life history traits (seed mass and maximum height)
-\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Besides
-communities functional composition and diversity, a last key aspect of
-communities response to disturbance is the changes in functional
-redundancy that quantifies the amount of shared trait values among
-species \citep{Carmona2016}. High functional redundancy, as in the very
-diverse tropical forests \citep{Bellwood2006}, mitigates the impacts of
-species removal on ecosystem functioning and determines communities
-resilience \citep{Trenbath1999, Elmqvist2003, Diaz2005}. High functional
-redundancy also increases the functional overlap among species and would
-foster neutral stochastic processes \citep{Gravel2006}.
+assessing their role in the ecosystem would directly link communities
+diversity to the ecosystem functioning and to its environmental
+constraints
+\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
+respect, the functional trait-based approach that focused on major
+traits related to species ecology and performance was sucessfully
+adopted \citep{Diaz2005, Villeger2008a}. For example the functoinal
+approach revealed in tropical rainforests the deterministic processes
+fostering fast growing species with efficient resources acquisition
+after disturbance \citep{Molino2001, Haddad2008, Ruger2009}. Disturbance
+therefore entails a functional shift from a dominance of
+``conservative'' slow-growing species dealing with scarce resources to
+``acquisitive'' fast-growing species with rapid and efficient use of
+abundant resources \citep{TerSteege2001, Reich2014, Herault2011}. This
+shift is translated into consistent trajectories of key functional
+traits related to resource acquisition (leaf area, density and
+chlorophyll content, and stem specific gravity and bark thickness), tree
+growth and reproduction life history traits (seed mass and maximum
+height) \citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}.
+Eventually a complete overview of communities response to disturbance
+would encompass the changes in functional redundancy, that quantifies
+the amount of shared trait values among species \citep{Carmona2016}.
+High functional redundancy like in the very diverse tropical forests
+\citep{Bellwood2006} mitigates the impacts of species removal on
+ecosystem functioning and determines the resilience of communities
+\citep{Trenbath1999, Elmqvist2003, Diaz2005}. High functional redundancy
+also increases the functional overlap among species and would foster
+neutral stochastic processes \citep{Gravel2006}.
 
@@
 here the taxonomic and functional trajectories in terms of richness,
-evennes and composition \citep{Lohbeck2015, Guariguata2001}. These
-trajectories would highlight the recovery of communities initial
-characteristics and the underlying ecological processes. They would
-clarify the tenants in the long term of the Intermediate Disturbance
-Hypothesis debated in tropical forests, and provide indications for
-future adaptive conservation strategies \citep{Adler2007}. We monitored
-over 30 years the response of 75 ha of neotropical forest plots set up
-on a gradient of disturbance intensity, from 10 to 60\% of ecosystem
-biomass removed. We made use of a large functional traits database
-browsing major leaf, stem and seed functional traits and species maximum
-height to draw the trajectories over time of communities taxonomic and
-functional composition,richness and evenness. Specifically, we (i)
-questioned the coupling between taxonomic and functional response to
-disturbance and identified the underlying assembly processes, (ii)
-clarified the validity of the IDH in the long term for tropical forest
-and (iii) questioned the completeness of communities recovery regarding
-communities functional redundancy.
+evenness, composition and redundancy
+\citep{Lohbeck2015, Guariguata2001}. These trajectories would highlight
+the recovery of communities initial characteristics and the underlying
+ecological processes, and provide indications for future adaptive
+conservation strategies \citep{Adler2007}. They would clarify the
+tenants of the Intermediate Disturbance Hypothesis in the long term,
+debated in tropical forests, and explicit the variety of
+post-disturbance trajectories defined by the disturbance intensity. We
+monitored over 30 years the response of 75 ha of neotropical forest
+plots set up on a gradient of disturbance intensity, from 10 to 60\% of
+ecosystem biomass removed. We made use of a large functional traits
+database browsing major leaf, stem and seed functional traits and
+species maximum height to draw the trajectories over time of communities
+taxonomic and functional composition, richness, evenness and redundancy.
+Specifically, we (i) questioned the recovery of communities taxonomic
+and functional characteristics and identify the underlying assembly
+processes, (ii) clarified the validity of the IDH in the long term for
+tropical forest and elucidated its declination into different
+trajectories, and (iii) questioned the length of complete communities
+recovery.
 
@@
 
-The plot 7 from treatment 1 displayed a constantly outlying functoinal
+The plot 7 from treatment 1 displayed a constantly outlying functional
 evenness and was removed from the graphical representation for better
@@
 
-\begin{quote}
-\begin{quote}
-Si on met que richesse et evenness, pour functional c'est de l'evenness
-(Rao = order2)
-\end{quote}
-\end{quote}
-
 \subsection{Communities cyclic
@@
 post-disturbance trajectories depended on communities initial
-composition, that partly determined the pool of rectuitable species and
+composition, that partly determined the pool of recruitable species and
 that constrained the converge towards the recovery of the initial
@@
 
-The humped-shaped trajectories of communities along time demonstrated
-the enhanced functional evennes and changes in functional composition
-following disturbance before a return towards initial, lower values. In
-accordance with the IDH, disturbance entailed significant temporal
-variability with the emergence of ecological processes different from
-those rueling undisturbed forest. The increase in functional evenness
-went with significant functional shifts towards resource-acquisitive
-strategies (sharp increase in the SLA, leaf thickness and bark thickness
-and decrease in wood density, leaf toughness and maximum height)
+The humped-shaped trajectories of communities along time illustrated the
+processes enhancing the functional evennes and the changes in functional
+composition following disturbance before communities return towards
+initial composition and lower evenness. Disturbance entailed significant
+variability in the communities through the emergence of ecological
+processes different from thoses involvemnt before disturbance, in
+accordance with the IDH. The increase in functional evenness went with
+significant functional shifts towards resource-acquisitive strategies
+(sharp increase in the SLA, leaf thickness and bark thickness and
+decrease in wood density, leaf toughness and maximum height)
 \citep{Westoby1998, Wright2004, Reich2014}. Because the pre-disturbance
@@
 and survival of previously infrequent species and functional types.
-Pioneers outcompeted other species in the post-disturbance environment
-of non limiting resources and alleviated competition. Before disturbance
-and over communities recovery time though, these species would be
-excluded by long-lived, more resistant and shade tolerant species.
+Disturbance vacated environmental niches of high light, space and
+nutrient availability, filled by competitive pioneers that became
+dominant and determined the functional characteristics of the community
+\citep{Grime1998}. Before disturbance and over communities recovery time
+though, these species would be excluded by long-lived, more resistant
+and shade tolerant species.
 
@@
 recruitment rules \citep{Busing2002} and was hampered by the increasing
-competition among species whih entailed long term recovery cycles
+competition among species, implying long term recovery
 \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-The rapid decrease of the functional redundancy within the initial
-functional space, correlated with the disturbance intensity, illustrated
-the space and resources made available after disturbance. Following the
-lottery recruitment, communities taxonomic richness increased firsthand
-along the enhanced recruitment rate but decreased thereafter, following
-the emergence of exclusive competition. After 30 years the recovery of
-the initial redundancy structure was underway but remained unachieved.
-This alteration of the functional redundancy meant a lower resilience of
-pre-disturbance communities and higher chances to see the persistence of
-disturbance-specific communities, with less species and more pioneers
-\citep{Haddad2008, Burslem2000, Martin2013}. Besides the hampered
-recovery of rare species increased the risks to loose cornerstone
+Following disturbance the space and resources made available
+corresponded to a rapid decrease of the functional redundancy within the
+initial functional space. The high light, space and nutrient niches were
+rapidly filled by species, mainly the most dominant and frequent ones.
+Thereafter competitive exclusion emerged following the filling of the
+niches and limited the recruitment of infrequent species
+\citep{Busing2002}. The recovery of communities taxonomic richness and
+evenness relied upon the recovery of the initial functional redudancy.
+Although underway 30 years after disturbance the recovery remained
+unachieved and after 30 years the recovery it was underway but remained
+unachieved. This alteration of the functional redundancy meant a lower
+resilience of pre-disturbance communities and higher chances to see the
+persistence of disturbance-specific communities, with less species and
+more pioneers \citep{Haddad2008, Burslem2000, Martin2013}. Besides the
+slowed recovery of rare species increased the risks to loose cornerstone
 specie, with unexpected ecological consequences
@@
 Our study revealed communities cyclic recovery after disturbance
-entailing the resilience of their functioning and taxonomic composition
+allowing the resilience of their functioning and taxonomic composition
 with the maintenance of initial differences among communities.

### 2018-07-03 Untitled.Rmd
graphs and relecture

--- 2018-07-02 WholePlotTrajectories.Rmd
@@
     authsuperscript: 1
-  - name: Bruno Hérault
+  - name: Bruno H??rault
     authsuperscript: 2 3
@@
   - affsuperscript: 1
-    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
+    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Universit?? des Antilles, Universit?? de Guyane.
     address: >
@@
   - affsuperscript: 3
-    dptuniv: INPHB, Institut National Polytechnique Félix Houphouet-Boigny
+    dptuniv: INPHB, Institut National Polytechnique F??lix Houphouet-Boigny
     address: >
@@
 abstract: >   
-    Understanding the ecological rules underlying the maintenance of tropical forests biodiversity, structure, functioning and dynamics is urgent to anticipate their fate in the global changing context.
-    The huge diversity of tropical forests is often assumed to be shaped by constant regime of disturbance yielding a diversity peak at intermediate intensity, but for tropical forests this intermediate disturbance hypothesis (IDH) remains debated, and this controversy also questions the extent of communities resilience regarding their functional and taxonomic facets is also to be clarified.
+    Understanding the ecological rules underlying the maintenance of tropical forests biodiversity, structure, functioning and dynamics is urgent to anticipate their fate in the global change context.
+    The huge diversity of tropical forests is often assumed to be regularly reshaped by natural disturbance yielding a diversity peak at intermediate intensity, but this intermediate disturbance hypothesis (IDH) remains debated, and this controversy also questions the extent of communities resilience regarding their functional and taxonomic facets.
 
-    To disentangle the ecological processes driving communities response to disturbance we analysed the taxonomic and functional diversity trajectories following a logging and thinning disturbance gradient.
-    Specifically we examined over 30 years the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 leaf, stem and life history traits.
-    Trajectories highlighted the cyclic recovery of communities taxonomic composition, maintaining the pre-disturbance differences, as well as functional composition and diversity.
-    The IDH predicted communities functional response to disturbance that resulted from competition processes determined by the disturbance intensity.
-    The IDH however poorly represented the taxonomic response that relied upon lottery recruitment and recruitment limitation which hampered the recovery of infrequent, shade tolerant species.
-
-    After disturbance communities functioning followed a cyclic recovery and the initial composition differences among plots were maintained.
-    Although consistent the recovery of communities functioning and taxonomic composition and diversity remained unachieved after 30 years which acknowledged the need of decades-long recovery cycles to ensure a complete recovery, and questioned communities sustainability after repeated disturbance.
+    To disentangle the ecological processes driving community response to disturbance, we analysed the taxonomic and functional diversity trajectories following a disturbance gradient.
+    Specifically, we examined, over 30 years, the functional and taxonomic community trajectories with regards to diversity, composition and redundancy.
+Functional trajectories were drawn based on 7 leaf, stem and life-history traits.
+    We highlighted the cyclic recovery of community taxonomic and functional composition.
+The pre-disturbance taxonomic differences were maintained over time while the functional composition trajectories were quite similar.
+    The IDH did predict communities functional diversity response while taxonomic diversity remained poorly sensitive to disturbance intensity.
+    Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years.
+This acknowledged the need of decades-long cycles with no disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbance.
 
@@
 knitr::opts_chunk$set(cache=F, warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
-# Installation des packages si nécessaire et chargement
+# Installation des packages si n??cessaire et chargement
 Library <- function(Packages) {
@@
 }
-# Ajouter les packages nécessaires ici
+# Ajouter les packages n??cessaires ici
 Library("kableExtra")
@@
 The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
-They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local climate, and the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The growing demand in forests products together with current climatic changes increases the pressure on remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threatens the maintenance and dynamics in space and time of communities structure, composition and functioning [@Anderson-Teixeira2013;@Sist2015].
+They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, the carbon, water and nutrient cycles, and ensure cultural and human well-being.
+The growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Gibson2011a; @Morales-Hidalgo2015] and threatens the maintenance and dynamics in space and time of communities structure, composition and functioning [@Anderson-Teixeira2013;@Sist2015].
 
-In tropical forests, ecological communities are constantly re-shaped by natural disturbance events that change both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions like competition among species [@Chesson2000;@Herault2018].
+In tropical forests, ecological communities are regularly re-shaped by natural disturbance events that change both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions such as competition among species [@Chesson2000;@Herault2018].
 One of the cornerstone of tropical forest ecology is to understand the processes and drivers of ecosystems response to disturbance [@White2001;@Chazdon2003a]. 
-For now, this has been largely studied through forest structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
-These structural parameters were thereafter sucessfully modeled, giving important insights into the recovery of ecosystems processes and services [@Herault2018]. 
-However the response of forests diversity is still unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014, @Liang2016] and would be most probably impacted by the changes induced by disturbance [@Baraloto2012a].
-In the short-term, moderate disturbance may lead to positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity when the disturbance regime is moderate but significant [@Molino2001;@Kariuki2006a;@Berry2008a].
-Along time the IDH would translate into a variety of post-disturbance trajectories shaped by the different ecological processes at stake.
+For now, this has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016] that are rapid and convenient to measure.
+These structural parameters have been sucessfully modeled, giving important insights into the recovery of ecosystem processes and services [@Herault2018]. 
+However the response of forests diversity and composition is still unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014; @Liang2016].
+In the short-term, moderate disturbance may lead to positive impacts on communities diversity, an idea formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity when disturbance intensity is not too high [@Molino2001;@Kariuki2006a;@Berry2008a].
 
-Validations of the IDH though remain scarce in the long term and mainly rely on the analysis of species richness.
-Species richness though gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016]: more complete analysis would encompass communities composition that is crucial for conservation issues, and abundance distribution that reveals the ecological processes involved [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, a functional approach accounting for species biological attributes and assessing their role in the ecosystem would directly link communities diversity to the ecosystem functioning and to its environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-In that respect, the functional trait-based approach that focused on major traits related to species ecology and performance was sucessfully adopted [@Diaz2005;@Villeger2008a].
-For example the functoinal approach revealed in tropical rainforests the deterministic processes fostering fast growing species with efficient resources acquisition after disturbance [@Molino2001;@Haddad2008;@Ruger2009].
-Disturbance therefore entails a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
-This shift is translated into consistent trajectories of key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
+Validations of the IDH though remain scarce in the long-term and mainly rely on the analysis of taxonomic richness [@Molino2001].
+Taxonomic richness, alone, gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+More ecological-meaningful analysis would couple richness to (i) evenness to reveal the underlying ecological processes and (ii) composition that is crucial for conservation issues [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Furthermore, a functional approach accounting for species biological attributes would directly link communities diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+In that respect, the functional trait-based approach that focused on major traits related to species ecology and thta mediated species performance in a given environment was sucessfully adopted [@Diaz2005;@Villeger2008a].
+For instance, the functional approach revealed in tropical rainforests the deterministic processes entailing, after disturbance, a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
+This shift is translated into consistent trajectories of key functional traits related to resource acquisition (leaf and stem traits) and life-history traits (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
 Eventually a complete overview of communities response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
-High functional redundancy like in the very diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities [@Trenbath1999; @Elmqvist2003; @Diaz2005].
-High functional redundancy also increases the functional overlap among species and would foster neutral stochastic processes [@Gravel2006].
+The high functional redundancy of hyperdiverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Trenbath1999; @Elmqvist2003; @Diaz2005].
 
-To grasp all facets of communities response to disturbance we examined here the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy [@Lohbeck2015;@Guariguata2001].
-These trajectories would highlight the recovery of communities initial characteristics and the underlying ecological processes, and provide indications for future adaptive conservation strategies [@Adler2007].
-They would clarify the tenants of the Intermediate Disturbance Hypothesis in the long term, debated in tropical forests, and explicit the variety of post-disturbance trajectories defined by the disturbance intensity.
-We monitored over 30 years the response of 75 ha of neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
-We made use of a large functional traits database browsing major leaf, stem and seed functional traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition, richness, evenness and redundancy.
+In this study, we monitored over 30 years the response of 75 ha of neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
+We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy [@Lohbeck2015;@Guariguata2001].
 Specifically, we (i) questioned the recovery of communities taxonomic and functional characteristics and identify the underlying assembly processes, (ii) clarified the validity of the IDH in the long term for tropical forest and elucidated its declination into different trajectories, and (iii) questioned the length of complete communities recovery.
@@
 
-Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26°C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
+Paracou station in French Guiana (5??18'N and 52??53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26??C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
 Elevation ranges between 5 and 50 m and soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
 
-The experiment is a network of twelve 6.25ha plots that underwent a gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) according to a randomized plot design with three replicate blocks of four plots.
+The experiment is a network of twelve 6.25ha plots that underwent a disturbance gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) according to a randomized plot design with three replicate blocks of four plots.
 The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3).
@@
 
-The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical families. 
-In the twelve experimental plots of the experiment, all trees above 10 cm DBH have been mapped and measured annually since 1984. 
-Trees are first identified during inventories with a vernacular name assigned by the forest worker team, and afterward with a scientific name assigned by botanists during regular botanical campaigns. 
+The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae. 
+In the twelve experimental plots, all trees above 10 cm DBH have been mapped and measured annually since 1984. 
+Trees are first identified with a vernacular name assigned by the forest worker team, and afterward with a scientific name assigned by botanists during regular botanical campaigns. 
 In 1984, specific vernacular names are given to 62 commercial or common species whereas more infrequent ones were identified under general identifiers only distinguishing trees and palm trees. 
@@
 
-This variability of protocols raised methodological issues as vernacular names usually correspond to different botanical species. 
+This variability of protocols in time raised methodological issues as vernacular names usually correspond to different botanical species. 
 It resulted in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics.
 The uncertainty propagation was done through a Bayesian framework reconstituting complete inventories at genus level from real incomplete ones on the basis of vernacular/botanical names association. 
-Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,…, \alpha_3\big]$ observed across all inventories between each vernacular name *v* and the species $\big[s_1, s_2, …, s_N\big]$:
+Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,???, \alpha_3\big]$ observed across all inventories between each vernacular name *v* and the species $\big[s_1, s_2, ???, s_N\big]$:
 
 \begin{align}
-M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big) \nonumber
+M_v\Big(\big[s_1, s_2, ???, s_N\big],\big[\alpha_1, \alpha_2,???, \alpha_3\big]\Big) \nonumber
 \end{align}
@@
 
-Six functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, the leaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass) came from the BRIDGE project [^1].
-Traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 botanical species pertaining to 157 botanical genera.
+Six functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area) and wood economics (wood specific gravity and bark thickness), and life-history traits (maximum specific height and seed mass) came from the BRIDGE project [^1].
+Trait values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 species pertaining to 157 genera.
 Missing trait values were filled using multivariate imputation by chained equation [@Mice2011].
-Imputations were restricted within genus, or family when samples were too scarce, in order to account for the phylogenetic signal of the functional traits. 
-Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level (same genus or family). 
-As seed mass information corresponds to a classification into mass classes, no data filling process was applied and analysis were restricted to the 414 botanical species recorded.
+Imputations were restricted within genus, or family when samples were too scarce, in order to account for the phylogenetic signal. 
+Whenever a species was not in the dataset, it was attributed a set of traits values randomly sampled among species of the next higher taxonomic level (same genus or family). 
+As seed mass information was classified into classes, no data filling process was applied and analyses were restricted to the 414 botanical species recorded.
 
-All composition and diversity metrics were obtained after 50 iterations of the taxonomic uncertainty propagation framework and of the filling process of missing trait values.
+All composition and diversity metrics were obtained after 50 iterations of the taxonomic uncertainty propagation framework.
 
@@
 
-To counter taxonomic uncertainties due to the variability of botanical identification levels (in space) and protocols (in time), the taxonomic composition and diversity analysis were conducted at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
-Trajectories of communities taxonomic and functional variations in composition after disturbance were followed in a two-dimensional NMDS ordination space of the 30 years monitored.
-Two NMDS were conducted to map either taxonomic flora inventories or communities functional composition based on the 7 leaf, stem and life history traits (without seed mass classes).
-In both cases the NMDS were performed using abundance-based (Bray-Curtis) dissimilarity measures. 
-Trajectories along time were reported through the distance between successive inventories and the reference inventories in 1989, *i.e.* 2-3 years after disturbance, when the uncertainty degree did not exceed 30% of undetermined trees.
-for the taxonomix composition trajectories were measured by the euclidean distance between inventories in the multivariate space of all inventories species.
-For the functoinal composition trajectories were measured by the sum for all traits of the euclidean distances between the traits density distributions of successive inventories.
-To compensate the intrinsinc difference among plots, the trajectories corresponded to the euclidean differences along time with the reference inventory in 1989.
+To counter taxonomic uncertainties due to the variability of botanical identification levels (in space) and protocols (in time), the taxonomic composition and diversity analysis were conducted at the genus level.
+Taxonomic and functional trajectories of community composition were followed in a two-dimensional NMDS ordination space.
+Two NMDS were conducted to map either taxonomic or functional composition, the later based on the 7 leaf, stem and life history traits (without seed mass classes).
+NMDS were performed using abundance-based (Bray-Curtis) dissimilarity measures.
+Trajectories along time were reported through the euclidean distance between the target inventories and the reference inventories in 1989, *i.e.* 2-3 years after disturbance. 
 Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007;@Garnier2004]. 
-Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were therefore reported as the proportion of each class in the inventories.
+Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were therefore reported as the proportion of each class in the inventories (Supp Mat XX).
 
@@
 These two indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995;@Marcon2015].
-The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on all functional and life traits.
+The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on all functional traits.
 
-The impacts of initial disturbance were tested with the spearman rank correlation between the extremum of taxonomic and functional metrics reached over the 30 years and the initial %AGB removed. 
+The impacts of initial disturbance were tested with the spearman rank correlation between the extrema of taxonomic and functional metrics reached over the 30 years and the initial %AGB removed.
 The were besides analysed through the linear correlations between Simpson and Rao diversities and the initial %AGB removed at 10, 20 and 30 years after disturbance.
@@
 The functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
-The samples of the trait database were first mapped in a 2-dimensional plan from a PCA analysis. 
+The samples of the trait database were first mapped in a 2-dimensional plan with a PCA analysis. 
 Then, multivariate kernel density estimator associated with individual trees returned species traits probability distribution (TDP).
@@
 
-## Communities richness and evenness
+## Community diversity: richness and evenness
 
-From 1989 (2-3 years after disturbance) to 2015 (28-29 years after disturbance), 828388 individual trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
-For undisturbed plots taxonomic Richness and Evenness (Simpson diversity) remained stable over the 30 years monitored.
-In disturbed communities after low disturbance intensity the taxonomic richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
-After intense disturbance the taxonomic richness followed unimodal trajectories, decreasing for ten years after disturbance before recovering to pre-disturbance values.
-In all disturbed plots the taxonomic evenness (Simpson diversity) followed unimodal trajectories with first an increase until a maximum reached after around 20 years that was positively correlated to the disturbance treatment ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
-The return towards initial values of communities evenness was beginning after 30 years except for two T3 plots (plots 8 and 12) which evenness still increased, suggesting similar but delayed trajectories \@ref(fig:DivTaxo).
+From 1989 (2 years after disturbance) to 2015 (28 years after disturbance), 828388 individual trees and 591 botanical species pertaining to 223 genus and 64 families were recorded. 
+For undisturbed plots, taxonomic Richness and Evenness remained stable over the 30 years monitored.
+In disturbed communities, after low disturbance intensity the taxonomic richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
+After intense disturbance the taxonomic richness followed a more complex trajectory, decreasing for ten years after disturbance before recovering to pre-disturbance values.
+In all disturbed plots the taxonomic evenness first increased until a maximum reached after around 20 years.
+This maximum was positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
+The evenness then stabilized except for two T3 plots (plots 8 and 12) for which evenness still increased.
 
-```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
+```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 
-The plot 7 from treatment 1 displayed a constantly outlying functional evenness and was removed from the graphical representation for better readability (see appendix for full graphs).
-In undisturbed plots the functional evenness remained stable along the 30 years while in disturbed plots it followed unimodal trajectories with a return towards initial values starting around 20 years after disturbance.
+The plot 7 from treatment 1 displayed a constantly outlying functional diversity and was removed from the graphical representation for better readability (see appendix for full graphs).
+In undisturbed plots the functional diversity remained stable along the 30 years.
+In disturbed plots, trajectories depend on the disturbnace intensity with, for low intensity, a low but long-lasting increase up to amximum reached after 20-25 years and, for, high intensity, a fast but short increase followed, after 10 years, by a slow decrease towards the inital values.
 
@@
 
-The impact of disturbance was examined specifically through the linear correlation between the intial %AGB removed and the Simpson and Rao diversities (diversities of order 2) after 10, 20 and 30 years \@ref(fig:IDHplot).
-The correlation with disturbance intensity was weak for the Simpson diversity ($R^2<0.25$) and only valid from 20 years after disturbance but it was much stronger for the Rao diversity ($0.60<R^2<0.75$) for all the time studied.
-Slope of linear correlations, reflecting the impact of disturbance, was the highest 20 years after disturbance.
+The impact of disturbance intensity was examined through the linear correlation between (i) the %AGB removed and (ii) taxonomic richness, evenness and functional diversity after 10, 20 and 30 years \@ref(fig:IDHplot).
+The correlation with disturbance intensity was weak for XXXXXXXXXX, for the taxonomic evenness ($R^2<0.25$) while highly significant for functional diversity ($0.60<R^2<0.75$) whatever the time since disturbance.
 
-```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance.
+```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="xxx to be changed xxx Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance.
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.
@@
 
-## Communities Composition
+## Community Composition
 
-### Taxonomic and functional trajectories
-
-While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed consistent trajectories over time after disturbance which revealed significant compositional changes.
-According to the mapping of functional traits (see appendix I) these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WD to high average SLA and chlorophyll content.
-For disturbed communities the distance of successive inventories to the 1989 reference inventory followed unimodal trajectories translating cyclic compositional changes with a recovery of the initial composition (Figure \@ref(fig:NMDSplans)).
-The maximum dissimilarity with the initial state was positively correlated to the disturbance treatment for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.91$ and $\rho_{spearman}^{functional}=0.96$ respectively) and the time at maximum was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition.
+While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-disturbance time.
+in disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WD to high average SLA and chlorophyll content (see appendix I).
+For functional composition, this translated into cyclic compositional changes with an unachieved recovery of the initial composition (Figure \@ref(fig:NMDSplans)).
+The maximum dissimilarity with the initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.91$ and $\rho_{spearman}^{functional}=0.96$ respectively).
+The maximum value was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition.
 
@@
 
-### Traits community weighted means (CWM)
+Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions followed unimodal trajectories either stabilizing or returning towards their initial values.
 
-Changes in functional composition trajectories went hand to hand with consistent trajectories of the 8 functional traits (Figure \@ref(fig:CWM)).
-Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions followed unimodal trajectories either stabilizing or returning towards their initial values.
-Thirty years after disturbance the weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value (Figure \@ref(fig:CWM)).
-The weighted means of bark thickness (*Bark_thick*) similarly remained substantially higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value.
-For all traits the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.67$, $\rho_{spearman}^{L_{chloro}}=0.45$, $\rho_{spearman}^{L_{toughness}}=-0.43$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WD}=-0.78$, $\rho_{spearman}^{Bark-thickness}=0.88$, $\rho_{spearman}^{Hmax}=-0.48$).
+Maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood density (*WD*) first decreased and then slightly increased but remained significantly lower than their initial value (Figure \@ref(fig:CWM)).
+On the other side, Bark thickness (*Bark_thick*) and specific leaf area (*SLA*) increased and while *Bark_thick* remained substantially high after 30 years, *SLA* had almost recovered its initial value.
+For all traits, the maximum difference to initial value was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.67$, $\rho_{spearman}^{L_{chloro}}=0.45$, $\rho_{spearman}^{L_{toughness}}=-0.43$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WD}=-0.78$, $\rho_{spearman}^{Bark-thickness}=0.88$, $\rho_{spearman}^{Hmax}=-0.48$).
 
@@
 
-### Functional redundancy
+## Functional redundancy
 
-In control plots the functional redundancy remained stable but in undisturbed plots it was quite variable and independent of the disturbance intensity (See appendix I).
-Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place before increasing to edge, recover or exceed the initial value.
+All disturbed plots had lower functional redundancy than control plots and followed similar hump-shaped trajectories (\@ref(fig:RedFun_rest)).
+The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{spearman}=0.50$) and the initial value had not recovered for any disturbed communities after 30 years.
 
-Considering the functional redundancy restricted to the functional space of the initial inventory, all disturbed plots followed similar decreasing humped shaped trajectories (\@ref(fig:RedFun_rest)).
-The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{spearman}=0.50$) and the initial value had not recovered for any disturbed communities.
-
-```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="TTrajectories of functional redundancy within the initial functional space over 30 years after disturbance.
+```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of functional redundancy within the initial functional space over 30 years after disturbance.
 Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
@@
 
-## Communities cyclic recovery
+## A cyclic recovery of community composition
 
-Communities taxonomic and functional composition and functional evenness proved resilient, following similar humped-shaped trajectories starting to return towards pre-disturbance composition and functioning after 30 years.
+Communities taxonomic and functional composition appeared resilient, following similar hump-shaped trajectories starting to return towards pre-disturbance composition after 30 years.
+The taxonomic differences among communities, marked before disturbance by the distinct starting points on the NMDS axis 2, were maintained throughout recovery trajectories.
+More than commonly thought, post-disturbance trajectories depended on community initial composition, that partly determined the pool of recruited species and that constrained the trajectories towards the initial composition.
+The high resilience of communities taxonomy revealed that species not belonging to the pre-disturbance community were hardly recruited because of the commonness of dispersal limitation among tropical tree species (Svenning, Wright JECOL 2005).
 
-The taxonomic differences among communities, marked before disturbance by the distinct starting points of the NMDS, were maintained throughout communities recovery trajectories.
-More than commonly thought post-disturbance trajectories depended on communities initial composition, that partly determined the pool of recruitable species and that constrained the converge towards the recovery of the initial composition [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
-The resilience of communities taxonomy suggested the existence of multiple stable equilibria, here the pre-disturbance communities, as assumed for highly diverse and productive ecosystems [@Chase2003].
-
-Communities functional composition followed similar trajectories that were not so distinct in the functional space.
-The trajectories of functional evenness as well were similar among communities 
-Communities functional characteristics are the most direct link between biodiversity and ecosystem functioning and their resilience here meant the recovery of ecosystem processes and services ona cycle of several decades [@Diaz2005;@Guariguata2001].
-
-## Another view of the intermediate disturbance hypothesis
-
-The IDH consistently predicted communities functional response to disturbance and the functional trajectories in time highlighted the underlying ecological mechanisms, but the theory was diproved regarding communities taxonomic richness and evenness.
-
-With the studied range of intensity, disturbance enhanced communities functional evenness but did not have clear impact on taxonomic richness or evenness. 
-The IDH proved a good predictor of the disturbance impact on communities functioning but not on their taxonomic richness or evenness.
-
-The humped-shaped trajectories of communities along time illustrated the processes enhancing the functional evennes and the changes in functional composition following disturbance before communities return towards initial composition and lower evenness.
-Disturbance entailed significant variability in the communities through the emergence of ecological processes different from thoses involvemnt before disturbance, in accordance with the IDH.
-The increase in functional evenness went with significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
+Community functional composition followed similar trajectories that were not so distinct in the functional space.......
 Because the pre-disturbance survivor trees mirror the initial communities [@Herault2018], the functional changes relied upon recruited trees and the enhanced growth and survival of previously infrequent species and functional types.
@@
 Before disturbance and over communities recovery time though, these species would be excluded by long-lived, more resistant and shade tolerant species.
+The recovery of communities functional characteristics, that mostly rely on dominant species according to the "vegetation quantity effect" [@Grime1998], suggested a similar recovery of the dominant species [@Molino2001].
+significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
 
+## Another perspective on the intermediate disturbance hypothesis
+
+1. IDH in Space (Disturbance gradient)
+The IDH consistently (predicted) the response of functional diversity to disturbance, but the theory was disproved regarding communities taxonomic richness and evenness.
 In contrast the taxonomic richness and evenness were decoupled from the disturbance intensity, questioning the consistency of the IDH as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
-The recovery of communities functional characteristics, that mostly rely on dominant species according to the "vegetation quantity effect" [@Grime1998], suggested a similar recovery of the dominant species [@Molino2001].
+
+2. IDH in Time
+The humped-shaped trajectories of functional diversity along time 
 Still, the taxonomic trajectories were blurried by the hampered recovery of infrequent species that entailed after disturbance a decrease in the taxonomic richness and an increase of the evenness [@Hubbell2001;@Chave2004].
-The disturbance therefore porrly predicted commuities taxonomic richnessa and evennes that rely upon different ecological processes.
 
@@
 
+Communities functional characteristics are the most direct link between biodiversity and ecosystem functioning and their resilience here meant the recovery of ecosystem processes and services on a cycle of several decades [@Diaz2005;@Guariguata2001].
+
 `r if (knitr:::is_html_output()) '# References {-}'`

### 2018-07-03 WholePlotTrajectories.tex
graphs and relecture

@@
 tropical forests biodiversity, structure, functioning and dynamics is
-urgent to anticipate their fate in the global changing context. The huge
-diversity of tropical forests is often assumed to be shaped by constant
-regime of disturbance yielding a diversity peak at intermediate
-intensity, but for tropical forests this intermediate disturbance
-hypothesis (IDH) remains debated, and this controversy also questions
-the extent of communities resilience regarding their functional and
-taxonomic facets is also to be clarified. To disentangle the ecological
-processes driving communities response to disturbance we analysed the
-taxonomic and functional diversity trajectories following a logging and
-thinning disturbance gradient. Specifically we examined over 30 years
-the trajectories of communities taxonomic richness and evennes and
-functional composition, diversity and redundancy based on 7 leaf, stem
-and life history traits. Trajectories highlighted the cyclic recovery of
-communities taxonomic composition, maintaining the pre-disturbance
-differences, as well as functional composition and diversity. The IDH
-predicted communities functional response to disturbance that resulted
-from competition processes determined by the disturbance intensity. The
-IDH however poorly represented the taxonomic response that relied upon
-lottery recruitment and recruitment limitation which hampered the
-recovery of infrequent, shade tolerant species. After disturbance
-communities functioning followed a cyclic recovery and the initial
-composition differences among plots were maintained. Although consistent
-the recovery of communities functioning and taxonomic composition and
-diversity remained unachieved after 30 years which acknowledged the need
-of decades-long recovery cycles to ensure a complete recovery, and
-questioned communities sustainability after repeated disturbance.
+urgent to anticipate their fate in the global change context. The huge
+diversity of tropical forests is often assumed to be regularly reshaped
+by natural disturbance yielding a diversity peak at intermediate
+intensity, but this intermediate disturbance hypothesis (IDH) remains
+debated, and this controversy also questions the extent of communities
+resilience regarding their functional and taxonomic facets. To
+disentangle the ecological processes driving community response to
+disturbance, we analysed the taxonomic and functional diversity
+trajectories following a disturbance gradient. Specifically, we
+examined, over 30 years, the functional and taxonomic community
+trajectories with regards to diversity, composition and redundancy.
+Functional trajectories were drawn based on 7 leaf, stem and
+life-history traits. We highlighted the cyclic recovery of community
+taxonomic and functional composition. The pre-disturbance taxonomic
+differences were maintained over time while the functional composition
+trajectories were quite similar. The IDH did predict communities
+functional diversity response while taxonomic diversity remained poorly
+sensitive to disturbance intensity. Although consistent, the recovery of
+community composition, diversity and redundancy remained unachieved
+after 30 years. This acknowledged the need of decades-long cycles with
+no disturbance to ensure a complete recovery, and questioned tropical
+forest community resilience after repeated disturbance.
 }
@@
 multiple non-timber forest products, shelter a diversified fauna,
-regulate the local climate, and the carbon, water and nutrient cycles,
-and ensure cultural and human well-being. The growing demand in forests
-products together with current climatic changes increases the pressure
-on remaining forests \citep{Gibson2011a, Morales-Hidalgo2015} and
-threatens the maintenance and dynamics in space and time of communities
-structure, composition and functioning
-\citep{Anderson-Teixeira2013, Sist2015}.
+regulate the local and regional climates, the carbon, water and nutrient
+cycles, and ensure cultural and human well-being. The growing demand in
+forests products together with current global changes increases the
+pressure on remaining natural forests
+\citep{Gibson2011a, Morales-Hidalgo2015} and threatens the maintenance
+and dynamics in space and time of communities structure, composition and
+functioning \citep{Anderson-Teixeira2013, Sist2015}.
 
-In tropical forests, ecological communities are constantly re-shaped by
+In tropical forests, ecological communities are regularly re-shaped by
 natural disturbance events that change both the abiotic environment,
 through the fluxes of light, heat and water \citep{Goulamoussene2017},
-and the biotic interactions like competition among species
+and the biotic interactions such as competition among species
 \citep{Chesson2000, Herault2018}. One of the cornerstone of tropical
@@
 response to disturbance \citep{White2001, Chazdon2003a}. For now, this
-has been largely studied through forest structural parameters, rapid and
-convenient to measure, as aboveground biomass, tree height or stem
-density \citep{Piponiot2016, Rutishauser2016}. These structural
-parameters were thereafter sucessfully modeled, giving important
-insights into the recovery of ecosystems processes and services
-\citep{Herault2018}. However the response of forests diversity is still
-unclear, albeit it determines the productivity, stability and
-functioning of ecosystems \citep[\citet{Liang2016}]{Tilman2014} and
-would be most probably impacted by the changes induced by disturbance
-\citep{Baraloto2012a}. In the short-term, moderate disturbance may lead
-to positive impacts on communities diversity, which have been formalized
-by the intermediate disturbance hypothesis (IDH) stating a maximized
-species diversity when the disturbance regime is moderate but
-significant \citep{Molino2001, Kariuki2006a, Berry2008a}. Along time the
-IDH would translate into a variety of post-disturbance trajectories
-shaped by the different ecological processes at stake.
+has been largely studied through forest structural parameters such as
+aboveground biomass, tree height or stem density
+\citep{Piponiot2016, Rutishauser2016} that are rapid and convenient to
+measure. These structural parameters have been sucessfully modeled,
+giving important insights into the recovery of ecosystem processes and
+services \citep{Herault2018}. However the response of forests diversity
+and composition is still unclear, albeit it determines the productivity,
+stability and functioning of ecosystems \citep{Tilman2014, Liang2016}.
+In the short-term, moderate disturbance may lead to positive impacts on
+communities diversity, an idea formalized by the intermediate
+disturbance hypothesis (IDH) stating a maximized species diversity when
+disturbance intensity is not too high
+\citep{Molino2001, Kariuki2006a, Berry2008a}.
 
-Validations of the IDH though remain scarce in the long term and mainly
-rely on the analysis of species richness. Species richness though gives
-limited or misleading information on forests recovery and functioning
-\citep{Martin2015, Chaudhary2016}: more complete analysis would
-encompass communities composition that is crucial for conservation
-issues, and abundance distribution that reveals the ecological processes
-involved \citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
-functional approach accounting for species biological attributes and
-assessing their role in the ecosystem would directly link communities
-diversity to the ecosystem functioning and to its environmental
-constraints
+Validations of the IDH though remain scarce in the long-term and mainly
+rely on the analysis of taxonomic richness \citep{Molino2001}. Taxonomic
+richness, alone, gives limited or misleading information on forests
+recovery and functioning \citep{Martin2015, Chaudhary2016}. More
+ecological-meaningful analysis would couple richness to (i) evenness to
+reveal the underlying ecological processes and (ii) composition that is
+crucial for conservation issues
+\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
+functional approach accounting for species biological attributes would
+directly link communities diversity, composition and redundancy to
+ecosystem functioning and to its environmental constraints
 \citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
 respect, the functional trait-based approach that focused on major
-traits related to species ecology and performance was sucessfully
-adopted \citep{Diaz2005, Villeger2008a}. For example the functoinal
-approach revealed in tropical rainforests the deterministic processes
-fostering fast growing species with efficient resources acquisition
-after disturbance \citep{Molino2001, Haddad2008, Ruger2009}. Disturbance
-therefore entails a functional shift from a dominance of
+traits related to species ecology and thta mediated species performance
+in a given environment was sucessfully adopted
+\citep{Diaz2005, Villeger2008a}. For instance, the functional approach
+revealed in tropical rainforests the deterministic processes entailing,
+after disturbance, a functional shift from a dominance of
 ``conservative'' slow-growing species dealing with scarce resources to
@@
 shift is translated into consistent trajectories of key functional
-traits related to resource acquisition (leaf area, density and
-chlorophyll content, and stem specific gravity and bark thickness), tree
-growth and reproduction life history traits (seed mass and maximum
-height) \citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}.
-Eventually a complete overview of communities response to disturbance
-would encompass the changes in functional redundancy, that quantifies
-the amount of shared trait values among species \citep{Carmona2016}.
-High functional redundancy like in the very diverse tropical forests
+traits related to resource acquisition (leaf and stem traits) and
+life-history traits (seed mass, maximum size)
+\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Eventually
+a complete overview of communities response to disturbance would
+encompass the changes in functional redundancy, that quantifies the
+amount of shared trait values among species \citep{Carmona2016}. The
+high functional redundancy of hyperdiverse tropical forests
 \citep{Bellwood2006} mitigates the impacts of species removal on
-ecosystem functioning and determines the resilience of communities
-\citep{Trenbath1999, Elmqvist2003, Diaz2005}. High functional redundancy
-also increases the functional overlap among species and would foster
-neutral stochastic processes \citep{Gravel2006}.
+ecosystem functioning and determines the resilience of communities after
+disturbance \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-To grasp all facets of communities response to disturbance we examined
-here the taxonomic and functional trajectories in terms of richness,
-evenness, composition and redundancy
-\citep{Lohbeck2015, Guariguata2001}. These trajectories would highlight
-the recovery of communities initial characteristics and the underlying
-ecological processes, and provide indications for future adaptive
-conservation strategies \citep{Adler2007}. They would clarify the
-tenants of the Intermediate Disturbance Hypothesis in the long term,
-debated in tropical forests, and explicit the variety of
-post-disturbance trajectories defined by the disturbance intensity. We
-monitored over 30 years the response of 75 ha of neotropical forest
-plots set up on a gradient of disturbance intensity, from 10 to 60\% of
-ecosystem biomass removed. We made use of a large functional traits
-database browsing major leaf, stem and seed functional traits and
-species maximum height to draw the trajectories over time of communities
-taxonomic and functional composition, richness, evenness and redundancy.
-Specifically, we (i) questioned the recovery of communities taxonomic
-and functional characteristics and identify the underlying assembly
-processes, (ii) clarified the validity of the IDH in the long term for
-tropical forest and elucidated its declination into different
-trajectories, and (iii) questioned the length of complete communities
-recovery.
+In this study, we monitored over 30 years the response of 75 ha of
+neotropical forest plots set up on a gradient of disturbance intensity,
+from 10 to 60\% of ecosystem biomass removed. We made use of a large
+functional traits database encompassing major leaf, stem and
+life-history traits in order to draw the taxonomic and functional
+trajectories in terms of richness, evenness, composition and redundancy
+\citep{Lohbeck2015, Guariguata2001}. Specifically, we (i) questioned the
+recovery of communities taxonomic and functional characteristics and
+identify the underlying assembly processes, (ii) clarified the validity
+of the IDH in the long term for tropical forest and elucidated its
+declination into different trajectories, and (iii) questioned the length
+of complete communities recovery.
 
@@
 
-Paracou station in French Guiana (5°18'N and 52°53'W) is located in a
+Paracou station in French Guiana (5??18'N and 52??53'W) is located in a
 lowland tropical rain forest in a tropical wet climate with mean annual
-temperature of 26°C, mean annual precipitation averaging 2980
+temperature of 26??C, mean annual precipitation averaging 2980
 mm.y\textsuperscript{-1} (30-y period) and a 3-month dry season
@@
 The experiment is a network of twelve 6.25ha plots that underwent a
-gradient of three logging, thinning and fuelwood cutting treatments
-(Table \ref{tab:Tab1}) according to a randomized plot design with three
-replicate blocks of four plots. The disturbance corresponds to averages
-of 10 trees removed per hectare with a diameter at 1.3 m height (DBH)
-above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for
-treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3).
-Treatments T2 and T3 besides included the thinning of trees by poison
-girdling \citep{Schmitt1990, Blanc2009}. The disturbance intensity was
-measured as the percentage of aboveground biomass (\%AGB) lost between
-the first inventory in 1984 and five years after disturbance
-\citep{Piponiot2016} estimated with the BIOMASS R package
+disturbance gradient of three logging, thinning and fuelwood cutting
+treatments (Table \ref{tab:Tab1}) according to a randomized plot design
+with three replicate blocks of four plots. The disturbance corresponds
+to averages of 10 trees removed per hectare with a diameter at 1.3 m
+height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm
+DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3
+(T3). Treatments T2 and T3 besides included the thinning of trees by
+poison girdling \citep{Schmitt1990, Blanc2009}. The disturbance
+intensity was measured as the percentage of aboveground biomass (\%AGB)
+lost between the first inventory in 1984 and five years after
+disturbance \citep{Piponiot2016} estimated with the BIOMASS R package
 \citep{Biomass2018}.
@@
 Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae,
-Lecythidaceae and Sapotaceae botanical families. In the twelve
-experimental plots of the experiment, all trees above 10 cm DBH have
-been mapped and measured annually since 1984. Trees are first identified
-during inventories with a vernacular name assigned by the forest worker
-team, and afterward with a scientific name assigned by botanists during
-regular botanical campaigns. In 1984, specific vernacular names are
-given to 62 commercial or common species whereas more infrequent ones
-were identified under general identifiers only distinguishing trees and
-palm trees. From 2003, botanical campaigns have been conducted every 5
-to 6 years to identify all trees at the species level but identification
-levels still varied among plots and campaigns.
+Lecythidaceae and Sapotaceae. In the twelve experimental plots, all
+trees above 10 cm DBH have been mapped and measured annually since 1984.
+Trees are first identified with a vernacular name assigned by the forest
+worker team, and afterward with a scientific name assigned by botanists
+during regular botanical campaigns. In 1984, specific vernacular names
+are given to 62 commercial or common species whereas more infrequent
+ones were identified under general identifiers only distinguishing trees
+and palm trees. From 2003, botanical campaigns have been conducted every
+5 to 6 years to identify all trees at the species level but
+identification levels still varied among plots and campaigns.
 
-This variability of protocols raised methodological issues as vernacular
-names usually correspond to different botanical species. It resulted in
-significant taxonomic uncertainties that had to be propagated to
-composition and diversity metrics. The uncertainty propagation was done
-through a Bayesian framework reconstituting complete inventories at
-genus level from real incomplete ones on the basis of
-vernacular/botanical names association. Vernacular names were replaced
-through multinomial trials based on the association probability
-\(\big[\alpha_1, \alpha_2,…, \alpha_3\big]\) observed across all
+This variability of protocols in time raised methodological issues as
+vernacular names usually correspond to different botanical species. It
+resulted in significant taxonomic uncertainties that had to be
+propagated to composition and diversity metrics. The uncertainty
+propagation was done through a Bayesian framework reconstituting
+complete inventories at genus level from real incomplete ones on the
+basis of vernacular/botanical names association. Vernacular names were
+replaced through multinomial trials based on the association probability
+\(\big[\alpha_1, \alpha_2,???, \alpha_3\big]\) observed across all
 inventories between each vernacular name \emph{v} and the species
-\(\big[s_1, s_2, …, s_N\big]\):
+\(\big[s_1, s_2, ???, s_N\big]\):
 
 \begin{align}
-M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big) \nonumber
+M_v\Big(\big[s_1, s_2, ???, s_N\big],\big[\alpha_1, \alpha_2,???, \alpha_3\big]\Big) \nonumber
 \end{align}
@@
 Six functional traits representing leaf economics (leaves thickness,
-toughness, total chlorophyll content and specific leaf area, the leaf
-area per unit dry mass) and wood economics (wood specific gravity and
-bark thickness), and life history traits (maximum specific height and
-seed mass) came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}.
-Traits values were assessed from a selection of individuals located in
-nine permanent plots in French Guiana, including two in Paracou, and
-comprised 294 botanical species pertaining to 157 botanical genera.
-Missing trait values were filled using multivariate imputation by
-chained equation \citep{Mice2011}. Imputations were restricted within
-genus, or family when samples were too scarce, in order to account for
-the phylogenetic signal of the functional traits. Whenever a species
-inventoried was not in the dataset, it was attributed a set of traits
-values randomly sampled among species of the same next higher taxonomic
-level (same genus or family). As seed mass information corresponds to a
-classification into mass classes, no data filling process was applied
-and analysis were restricted to the 414 botanical species recorded.
+toughness, total chlorophyll content and specific leaf area) and wood
+economics (wood specific gravity and bark thickness), and life-history
+traits (maximum specific height and seed mass) came from the BRIDGE
+project \footnote{http://www.ecofog.gf/Bridge/}. Trait values were
+assessed from a selection of individuals located in nine permanent plots
+in French Guiana, including two in Paracou, and comprised 294 species
+pertaining to 157 genera. Missing trait values were filled using
+multivariate imputation by chained equation \citep{Mice2011}.
+Imputations were restricted within genus, or family when samples were
+too scarce, in order to account for the phylogenetic signal. Whenever a
+species was not in the dataset, it was attributed a set of traits values
+randomly sampled among species of the next higher taxonomic level (same
+genus or family). As seed mass information was classified into classes,
+no data filling process was applied and analyses were restricted to the
+414 botanical species recorded.
 
 All composition and diversity metrics were obtained after 50 iterations
-of the taxonomic uncertainty propagation framework and of the filling
-process of missing trait values.
+of the taxonomic uncertainty propagation framework.
 
@@
 identification levels (in space) and protocols (in time), the taxonomic
-composition and diversity analysis were conducted at the genus level,
-\emph{i.e.} referring to the genus of observed or trialed botanical
-names. Trajectories of communities taxonomic and functional variations
-in composition after disturbance were followed in a two-dimensional NMDS
-ordination space of the 30 years monitored. Two NMDS were conducted to
-map either taxonomic flora inventories or communities functional
-composition based on the 7 leaf, stem and life history traits (without
-seed mass classes). In both cases the NMDS were performed using
-abundance-based (Bray-Curtis) dissimilarity measures. Trajectories along
-time were reported through the distance between successive inventories
-and the reference inventories in 1989, \emph{i.e.} 2-3 years after
-disturbance, when the uncertainty degree did not exceed 30\% of
-undetermined trees. for the taxonomix composition trajectories were
-measured by the euclidean distance between inventories in the
-multivariate space of all inventories species. For the functoinal
-composition trajectories were measured by the sum for all traits of the
-euclidean distances between the traits density distributions of
-successive inventories. To compensate the intrinsinc difference among
-plots, the trajectories corresponded to the euclidean differences along
-time with the reference inventory in 1989. Univariate trajectories of
-the leaf, stem and life-history traits were also visualized with the
-community weighted means (CWM) \citep{Diaz2007, Garnier2004}. Species
-seed mass corresponded to 5 classes of increasing mass, seed mass
-trajectories were therefore reported as the proportion of each class in
-the inventories.
+composition and diversity analysis were conducted at the genus level.
+Taxonomic and functional trajectories of community composition were
+followed in a two-dimensional NMDS ordination space. Two NMDS were
+conducted to map either taxonomic or functional composition, the later
+based on the 7 leaf, stem and life history traits (without seed mass
+classes). NMDS were performed using abundance-based (Bray-Curtis)
+dissimilarity measures. Trajectories along time were reported through
+the euclidean distance between the target inventories and the reference
+inventories in 1989, \emph{i.e.} 2-3 years after disturbance. Univariate
+trajectories of the leaf, stem and life-history traits were also
+visualized with the community weighted means (CWM)
+\citep{Diaz2007, Garnier2004}. Species seed mass corresponded to 5
+classes of increasing mass, seed mass trajectories were therefore
+reported as the proportion of each class in the inventories (Supp Mat
+XX).
 
@@
 species abundance distribution and average pairwise dissimilarity based
-on all functional and life traits.
+on all functional traits.
 
 The impacts of initial disturbance were tested with the spearman rank
-correlation between the extremum of taxonomic and functional metrics
+correlation between the extrema of taxonomic and functional metrics
 reached over the 30 years and the initial \%AGB removed. The were
@@
 community functional space \citep{Carmona2016}. The samples of the trait
-database were first mapped in a 2-dimensional plan from a PCA analysis.
+database were first mapped in a 2-dimensional plan with a PCA analysis.
 Then, multivariate kernel density estimator associated with individual
@@
 
-From 1989 (2-3 years after disturbance) to 2015 (28-29 years after
+From 1989 (2 years after disturbance) to 2015 (28 years after
 disturbance), 828388 individual trees and 591 botanical species
-pertaining to 223 genus and 64 botanical families were recorded. For
-undisturbed plots taxonomic Richness and Evenness (Simpson diversity)
-remained stable over the 30 years monitored. In disturbed communities
-after low disturbance intensity the taxonomic richness increased,
-reaching a maximum gain of 14 botanical genera (plot 3 from treatment
-2). After intense disturbance the taxonomic richness followed unimodal
-trajectories, decreasing for ten years after disturbance before
-recovering to pre-disturbance values. In all disturbed plots the
-taxonomic evenness (Simpson diversity) followed unimodal trajectories
-with first an increase until a maximum reached after around 20 years
-that was positively correlated to the disturbance treatment
+pertaining to 223 genus and 64 families were recorded. For undisturbed
+plots, taxonomic Richness and Evenness remained stable over the 30 years
+monitored. In disturbed communities, after low disturbance intensity the
+taxonomic richness increased, reaching a maximum gain of 14 botanical
+genera (plot 3 from treatment 2). After intense disturbance the
+taxonomic richness followed a more complex trajectory, decreasing for
+ten years after disturbance before recovering to pre-disturbance values.
+In all disturbed plots the taxonomic evenness first increased until a
+maximum reached after around 20 years. This maximum was positively
+correlated to the disturbance intensity
 (\(\rho_{spearman}^{Shannon}=0.86\), and
-\(\rho_{spearman}^{Simpson}=0.89\)). The return towards initial values
-of communities evenness was beginning after 30 years except for two T3
-plots (plots 8 and 12) which evenness still increased, suggesting
-similar but delayed trajectories \ref{fig:DivTaxo}.
+\(\rho_{spearman}^{Simpson}=0.89\)). The evenness then stabilized except
+for two T3 plots (plots 8 and 12) for which evenness still increased.
 
@@
 The plot 7 from treatment 1 displayed a constantly outlying functional
-evenness and was removed from the graphical representation for better
+diversity and was removed from the graphical representation for better
 readability (see appendix for full graphs). In undisturbed plots the
-functional evenness remained stable along the 30 years while in
-disturbed plots it followed unimodal trajectories with a return towards
-initial values starting around 20 years after disturbance.
+functional diversity remained stable along the 30 years. In disturbed
+plots, trajectories depend on the disturbnace intensity with, for low
+intensity, a low but long-lasting increase up to amximum reached after
+20-25 years and, for, high intensity, a fast but short increase
+followed, after 10 years, by a slow decrease towards the inital values.
 
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/FunDiv-1}
-
-}
-
-\caption{Trajectories of the Rao functional diversity over 30 years after disturbance.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:FunDiv}
-\end{figure}
-
-The impact of disturbance was examined specifically through the linear
-correlation between the intial \%AGB removed and the Simpson and Rao
-diversities (diversities of order 2) after 10, 20 and 30 years
+The impact of disturbance intensity was examined through the linear
+correlation between (i) the \%AGB removed and (ii) taxonomic richness,
+evenness and functional diversity after 10, 20 and 30 years
 \ref{fig:IDHplot}. The correlation with disturbance intensity was weak
-for the Simpson diversity (\(R^2<0.25\)) and only valid from 20 years
-after disturbance but it was much stronger for the Rao diversity
-(\(0.60<R^2<0.75\)) for all the time studied. Slope of linear
-correlations, reflecting the impact of disturbance, was the highest 20
-years after disturbance.
+for XXXXXXXXXX, for the taxonomic evenness (\(R^2<0.25\)) while highly
+significant for functional diversity (\(0.60<R^2<0.75\)) whatever the
+time since disturbance.
 
@@
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.
-Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities 10, 20 and 30 years after disturbance.}\label{fig:IDHplot1}
-\end{figure*}\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/IDHplot-2}
-
-}
-
-\caption{Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.
-Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities 10, 20 and 30 years after disturbance.}\label{fig:IDHplot2}
+Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities 10, 20 and 30 years after disturbance.}\label{fig:IDHplot}
 \end{figure*}
@@
 
-\subsubsection{Taxonomic and functional
-trajectories}\label{taxonomic-and-functional-trajectories}
-
 While both taxonomic and functional composition remained stable in
 undisturbed communities (Figure \ref{fig:NMDSplans}), they followed
-consistent trajectories over time after disturbance which revealed
-significant compositional changes. According to the mapping of
-functional traits (see appendix I) these compositional changes
-corresponded to shifts towards species with more acquisitive functional
-strategies, from communities with high average WD to high average SLA
-and chlorophyll content. For disturbed communities the distance of
-successive inventories to the 1989 reference inventory followed unimodal
-trajectories translating cyclic compositional changes with a recovery of
-the initial composition (Figure \ref{fig:NMDSplans}). The maximum
+marked and consistent trajectories over post-disturbance time. in
+disturbed communities, these compositional changes corresponded to
+shifts towards species with more acquisitive functional strategies, from
+communities with high average WD to high average SLA and chlorophyll
+content (see appendix I). For functional composition, this translated
+into cyclic compositional changes with an unachieved recovery of the
+initial composition (Figure \ref{fig:NMDSplans}). The maximum
 dissimilarity with the initial state was positively correlated to the
-disturbance treatment for both taxonomic and functional composition
+disturbance intensity for both taxonomic and functional composition
 (\(\rho_{spearman}^{taxonomic}=0.91\) and
-\(\rho_{spearman}^{functional}=0.96\) respectively) and the time at
-maximum was reached around 26 years after disturbance for taxonomic
-composition and 22 years for functional composition.
+\(\rho_{spearman}^{functional}=0.96\) respectively). The maximum value
+was reached around 26 years after disturbance for taxonomic composition
+and 22 years for functional composition.
 
@@
 
-\subsubsection{Traits community weighted means
-(CWM)}\label{traits-community-weighted-means-cwm}
+Except for leaf chlorophyll content, which continued to increase for
+some T3 and T2 plots 30 years after disturbance, all traits and seed
+mass proportions followed unimodal trajectories either stabilizing or
+returning towards their initial values.
 
-Changes in functional composition trajectories went hand to hand with
-consistent trajectories of the 8 functional traits (Figure
-\ref{fig:CWM}). Except for leaf chlorophyll content, which continued to
-increase for some T3 and T2 plots 30 years after disturbance, all traits
-and seed mass proportions followed unimodal trajectories either
-stabilizing or returning towards their initial values. Thirty years
-after disturbance the weighted means of communities specific maximum
-height at adult stage (\emph{Hmax}), leaf toughness
-(\emph{L\_toughness}) and wood specific gravity (\emph{WD}) remained
-significantly lower than their initial value (Figure \ref{fig:CWM}). The
-weighted means of bark thickness (\emph{Bark\_thick}) similarly remained
-substantially higher than initially for all disturbed plots while the
-specfic leaf area (\emph{SLA}) had almost recovered its initial value.
-For all traits the maximum difference to initial state was correlated to
-the disturbance intensity (\(\rho_{spearman}^{L_{thickness}}=0.67\),
+Maximum height at adult stage (\emph{Hmax}), leaf toughness
+(\emph{L\_toughness}) and wood density (\emph{WD}) first decreased and
+then slightly increased but remained significantly lower than their
+initial value (Figure \ref{fig:CWM}). On the other side, Bark thickness
+(\emph{Bark\_thick}) and specific leaf area (\emph{SLA}) increased and
+while \emph{Bark\_thick} remained substantially high after 30 years,
+\emph{SLA} had almost recovered its initial value. For all traits, the
+maximum difference to initial value was correlated to the disturbance
+intensity (\(\rho_{spearman}^{L_{thickness}}=0.67\),
 \(\rho_{spearman}^{L_{chloro}}=0.45\),
@@
 
-\subsubsection{Functional redundancy}\label{functional-redundancy}
+\subsection{Functional redundancy}\label{functional-redundancy}
 
-In control plots the functional redundancy remained stable but in
-undisturbed plots it was quite variable and independent of the
-disturbance intensity (See appendix I). Globally after most intense
-disturbance (plots T2 and T3) communities redundancy decreased at first
-place before increasing to edge, recover or exceed the initial value.
-
-Considering the functional redundancy restricted to the functional space
-of the initial inventory, all disturbed plots followed similar
-decreasing humped shaped trajectories (@ref(fig:RedFun\_rest)). The
-maximum redundancy loss was positively correlated with the disturbance
-intensity (\(\rho_{spearman}=0.50\)) and the initial value had not
-recovered for any disturbed communities.
+All disturbed plots had lower functional redundancy than control plots
+and followed similar hump-shaped trajectories (@ref(fig:RedFun\_rest)).
+The maximum redundancy loss was positively correlated with the
+disturbance intensity (\(\rho_{spearman}=0.50\)) and the initial value
+had not recovered for any disturbed communities after 30 years.
 
@@
 
-\subsection{Communities cyclic
-recovery}\label{communities-cyclic-recovery}
+\subsection{A cyclic recovery of community
+composition}\label{a-cyclic-recovery-of-community-composition}
 
-Communities taxonomic and functional composition and functional evenness
-proved resilient, following similar humped-shaped trajectories starting
-to return towards pre-disturbance composition and functioning after 30
-years.
+Communities taxonomic and functional composition appeared resilient,
+following similar hump-shaped trajectories starting to return towards
+pre-disturbance composition after 30 years. The taxonomic differences
+among communities, marked before disturbance by the distinct starting
+points on the NMDS axis 2, were maintained throughout recovery
+trajectories. More than commonly thought, post-disturbance trajectories
+depended on community initial composition, that partly determined the
+pool of recruited species and that constrained the trajectories towards
+the initial composition. The high resilience of communities taxonomy
+revealed that species not belonging to the pre-disturbance community
+were hardly recruited because of the commonness of dispersal limitation
+among tropical tree species (Svenning, Wright JECOL 2005).
 
-The taxonomic differences among communities, marked before disturbance
-by the distinct starting points of the NMDS, were maintained throughout
-communities recovery trajectories. More than commonly thought
-post-disturbance trajectories depended on communities initial
-composition, that partly determined the pool of recruitable species and
-that constrained the converge towards the recovery of the initial
-composition
-\citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}. The
-resilience of communities taxonomy suggested the existence of multiple
-stable equilibria, here the pre-disturbance communities, as assumed for
-highly diverse and productive ecosystems \citep{Chase2003}.
-
-Communities functional composition followed similar trajectories that
-were not so distinct in the functional space. The trajectories of
-functional evenness as well were similar among communities Communities
-functional characteristics are the most direct link between biodiversity
-and ecosystem functioning and their resilience here meant the recovery
-of ecosystem processes and services ona cycle of several decades
-\citep{Diaz2005, Guariguata2001}.
-
-\subsection{Another view of the intermediate disturbance
-hypothesis}\label{another-view-of-the-intermediate-disturbance-hypothesis}
-
-The IDH consistently predicted communities functional response to
-disturbance and the functional trajectories in time highlighted the
-underlying ecological mechanisms, but the theory was diproved regarding
-communities taxonomic richness and evenness.
-
-With the studied range of intensity, disturbance enhanced communities
-functional evenness but did not have clear impact on taxonomic richness
-or evenness. The IDH proved a good predictor of the disturbance impact
-on communities functioning but not on their taxonomic richness or
-evenness.
-
-The humped-shaped trajectories of communities along time illustrated the
-processes enhancing the functional evennes and the changes in functional
-composition following disturbance before communities return towards
-initial composition and lower evenness. Disturbance entailed significant
-variability in the communities through the emergence of ecological
-processes different from thoses involvemnt before disturbance, in
-accordance with the IDH. The increase in functional evenness went with
+Community functional composition followed similar trajectories that were
+not so distinct in the functional space\ldots{}\ldots{}. Because the
+pre-disturbance survivor trees mirror the initial communities
+\citep{Herault2018}, the functional changes relied upon recruited trees
+and the enhanced growth and survival of previously infrequent species
+and functional types. Disturbance vacated environmental niches of high
+light, space and nutrient availability, filled by competitive pioneers
+that became dominant and determined the functional characteristics of
+the community \citep{Grime1998}. Before disturbance and over communities
+recovery time though, these species would be excluded by long-lived,
+more resistant and shade tolerant species. The recovery of communities
+functional characteristics, that mostly rely on dominant species
+according to the ``vegetation quantity effect'' \citep{Grime1998},
+suggested a similar recovery of the dominant species \citep{Molino2001}.
 significant functional shifts towards resource-acquisitive strategies
@@
 decrease in wood density, leaf toughness and maximum height)
-\citep{Westoby1998, Wright2004, Reich2014}. Because the pre-disturbance
-survivor trees mirror the initial communities \citep{Herault2018}, the
-functional changes relied upon recruited trees and the enhanced growth
-and survival of previously infrequent species and functional types.
-Disturbance vacated environmental niches of high light, space and
-nutrient availability, filled by competitive pioneers that became
-dominant and determined the functional characteristics of the community
-\citep{Grime1998}. Before disturbance and over communities recovery time
-though, these species would be excluded by long-lived, more resistant
-and shade tolerant species.
+\citep{Westoby1998, Wright2004, Reich2014}.
 
-In contrast the taxonomic richness and evenness were decoupled from the
-disturbance intensity, questioning the consistency of the IDH as already
-observed in the Guiana Shield \citep{Baraloto2012a} and in Bornean
-tropical forests \citep{Cannon1998}. The recovery of communities
-functional characteristics, that mostly rely on dominant species
-according to the ``vegetation quantity effect'' \citep{Grime1998},
-suggested a similar recovery of the dominant species \citep{Molino2001}.
-Still, the taxonomic trajectories were blurried by the hampered recovery
-of infrequent species that entailed after disturbance a decrease in the
-taxonomic richness and an increase of the evenness
-\citep{Hubbell2001, Chave2004}. The disturbance therefore porrly
-predicted commuities taxonomic richnessa and evennes that rely upon
-different ecological processes.
+\subsection{Another perspective on the intermediate disturbance
+hypothesis}\label{another-perspective-on-the-intermediate-disturbance-hypothesis}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  IDH in Space (Disturbance gradient) The IDH consistently (predicted)
+  the response of functional diversity to disturbance, but the theory
+  was disproved regarding communities taxonomic richness and evenness.
+  The functional range of species that could establish after disturbance
+  was determined by the expanse of the environmental space made
+  available by disturbance.
+\end{enumerate}
+
+In contrast the IDH translated in a decrease of the taxonomic richness
+above an intensity threshold while the taxonomic evenness was decoupled
+from disturbance intensity, as already observed in the Guiana Shield
+\citep{Baraloto2012a} and in Bornean tropical forests
+\citep{Cannon1998}. The taxonomic richness and evenness of recruited
+species was fixed and at community scale the post-disturbance taxonomic
+characterisitcs depended on the combination of these recruits and the
+remaining pre-disturbance community.
+
+The intensity of the disturbance determined the balance between trees
+recruited in the pre-disturbance community and those recruited in
+disturbance-specific community, functionally more diverse than
+undisturbed forests but with lower taxonomic richness and higher
+evenness.
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  IDH in Time For the first years following disturbance community
+  functional diversity increased through the filling of the vacated
+  environmental space without competitive exclusion contraints. The
+  maximum functional diversity was reached at the same time so the rate
+  of space filling was similar among communities. The maximum was
+  followed by the emergence of competitive exclusion processes bringing
+  the functional diversity back to pre-disturbance levels and made the
+  disturbance hypothesis intermediate in time.
+\end{enumerate}
+
+For medium disturbance intensity enhancing the taxonomic richness and
+evennes, the taxonomic trajectories reached asymptots of high richness
+and high evenness decoupled from their functional trajectories. Species
+first established after disturbance remained in the community but became
+less dominant, replaced the abundant pre-disturbance species, but the
+recovery of rare species was hampered in the long-term
+\citep{Hubbell2001, Chave2004}.
 

### 2018-07-03 Scripts/DrawingTools.R

@@ -88,7 +88,6 @@ mtext("(c)",side=3,adj=0,line=0.5)
 }
 
 TaxoTraj<-function(CompTaxo){
-  par(mfrow=c(1,2),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
   for(q in c(1,3)){     
     Toplot<-lapply(CompTaxo,function(tr){return(tr[,,,q])})
     Toplot<-lapply(Toplot,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
@@ -103,7 +102,7 @@ TaxoTraj<-function(CompTaxo){
     plot(colnames(Toplot[[1]]),Toplot[[1]][1,,1],type="n",xaxt="n",
          xlab="",ylab="",ylim=c(min(unlist(Toplot),na.rm=T),max(unlist(Toplot),na.rm=T)))
     axis(1,at=as.character(seq(5,33,5)),labels=TRUE)  
-    mtext(paste("(",c("a","b","c")[q],") ",c("Richness","Shannon","Simpson")[q],sep=""),
+    mtext(paste("(",c("a","b","b")[q],") ",c("Taxonomic Richness","Shannon","Taxonomic Evenness")[q],sep=""),
           line=1,side=3,adj=0)
     
     invisible(lapply(1:4,function(t){  
@@ -117,7 +116,7 @@ TaxoTraj<-function(CompTaxo){
       }))
     }))
   }
-  mtext("Years since disturbance",side=1,adj=1,cex=0.8,line=-2,outer=TRUE)
+  #mtext("Years since disturbance",side=1,adj=1,cex=0.8,line=-2,outer=TRUE)
   mtext("Equivalent diversity",side=2,padj=1,cex=0.8,line=1.5,outer=TRUE)}
 
 plotIDH<-function(Data,AgbLoss){
@@ -136,7 +135,7 @@ plotIDH<-function(Data,AgbLoss){
     #       bty="n",lty=1,col=colyear[Ti],lwd=2.5,cex=0.8)
   })
   legend("topleft",legend=unlist(leg),bty="n",lty=1,col=colyear,lwd=2.5,cex=0.8,title=expression(paste('R'^2,'adjusted')))
-}
+  }
 
 plotDiv<-function(Data,remove=FALSE){
   

### 2018-07-03 Scripts/Stats.R

@@ -12,6 +12,15 @@ Sim<-lapply(Sim, function(tr){
              dimnames=list(rownames(Ret[[1]]),colnames(Ret[[1]]),c(0.025,0.5,0.975)))})
 save(Sim,file="DB/SimpsonIDH")
 
+RichTaxo<-lapply(CompleteTaxo,function(tr){return(tr[,,,"Richness"])})
+Rich<-lapply(RichTaxo,function(tr){return(tr[,c("1995","2005","2015"),])})
+Rich<-lapply(Rich, function(tr){
+  Ret<-lapply(c(0.025,0.5,0.975),function(quant){
+    return(apply(tr,c(1,2),function(x){return(quantile(x,probs=quant))}))})
+  Ret<-array(unlist(Ret),dim=c(nrow(Ret[[1]]),ncol(Ret[[1]]),3),
+             dimnames=list(rownames(Ret[[1]]),colnames(Ret[[1]]),c(0.025,0.5,0.975)))})
+save(Rich,file="DB/RichnessIDH")
+
 CompFun<-CompleteFun
 CompFun[[2]]<-CompFun[[2]][which(rownames(CompFun[[2]])!=7),,]
 Rao<-lapply(CompFun, function(tr){

### 2018-07-04 WholePlotTrajectories.tex
IDH graphs

@@
 In tropical forests, ecological communities are regularly re-shaped by
-natural disturbance events that change both the abiotic environment,
+natural disturbance events changing both the abiotic environment,
 through the fluxes of light, heat and water \citep{Goulamoussene2017},
@@
 recovery and functioning \citep{Martin2015, Chaudhary2016}. More
-ecological-meaningful analysis would couple richness to (i) evenness to
-reveal the underlying ecological processes and (ii) composition that is
-crucial for conservation issues
+ecological-meaningful analysis would couple richness with (i) evenness
+that would reveal underlying ecological processes and (ii) composition
+that is crucial for conservation issues
 \citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
@@
 \citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
-respect, the functional trait-based approach that focused on major
-traits related to species ecology and thta mediated species performance
-in a given environment was sucessfully adopted
-\citep{Diaz2005, Villeger2008a}. For instance, the functional approach
-revealed in tropical rainforests the deterministic processes entailing,
-after disturbance, a functional shift from a dominance of
-``conservative'' slow-growing species dealing with scarce resources to
-``acquisitive'' fast-growing species with rapid and efficient use of
-abundant resources \citep{TerSteege2001, Reich2014, Herault2011}. This
-shift is translated into consistent trajectories of key functional
-traits related to resource acquisition (leaf and stem traits) and
-life-history traits (seed mass, maximum size)
+respect, the functional trait-based approach that focus on major traits
+related to species ecology and mediate species performance in a given
+environment was sucessfully adopted \citep{Diaz2005, Villeger2008a}. For
+instance, the functional approach revealed in tropical rainforests the
+deterministic processes entailing, after disturbance, a functional shift
+from a dominance of ``conservative'' slow-growing species dealing with
+scarce resources to ``acquisitive'' fast-growing species with rapid and
+efficient use of abundant resources
+\citep{TerSteege2001, Reich2014, Herault2011}. This shift is translated
+into the trajectories of key functional traits related to resource
+acquisition (leaf and stem traits) and life-history traits (seed mass,
+maximum size)
 \citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Eventually
@@
 neotropical forest plots set up on a gradient of disturbance intensity,
-from 10 to 60\% of ecosystem biomass removed. We made use of a large
-functional traits database encompassing major leaf, stem and
-life-history traits in order to draw the taxonomic and functional
+from 10 to 60\% of ecosystem above-ground biomass (AGB) removed. We made
+use of a large functional traits database encompassing major leaf, stem
+and life-history traits in order to draw the taxonomic and functional
 trajectories in terms of richness, evenness, composition and redundancy
@@
 recovery of communities taxonomic and functional characteristics and
-identify the underlying assembly processes, (ii) clarified the validity
-of the IDH in the long term for tropical forest and elucidated its
-declination into different trajectories, and (iii) questioned the length
-of complete communities recovery.
+identified the underlying assembly processes, (ii) clarified the
+validity of the IDH in the long term for tropical forest and elucidated
+its translation into different trajectories, and (iii) questioned
+community recovery time.
 
@@
 
-Paracou station in French Guiana (5??18'N and 52??53'W) is located in a
-lowland tropical rain forest in a tropical wet climate with mean annual
-temperature of 26??C, mean annual precipitation averaging 2980
-mm.y\textsuperscript{-1} (30-y period) and a 3-month dry season
-(\textless{} 100 mm.month\textsuperscript{-1}) from mid-August to
-mid-November, and a one-month dry season in March \citep{Wagner2011}.
-Elevation ranges between 5 and 50 m and soils correspond to thin
-acrisols over a layer of transformed saprolite with low permeability
-generating lateral drainage during heavy rains.
+Paracou station in French Guiana (5\textdegree 18'N and
+52\textdegree 53'W) is located in a lowland tropical rain forest in a
+tropical wet climate with mean annual temperature of 26\textdegree C,
+mean annual precipitation averaging 2980 mm.y\textsuperscript{-1} (30-y
+period) and a 3-month dry season (\textless{} 100
+mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
+one-month dry season in March \citep{Wagner2011}. Elevation ranges
+between 5 and 50 m and soils correspond to thin acrisols over a layer of
+transformed saprolite with low permeability generating lateral drainage
+during heavy rains.
 
@@
 during regular botanical campaigns. In 1984, specific vernacular names
-are given to 62 commercial or common species whereas more infrequent
+were given to 62 commercial or common species whereas more infrequent
 ones were identified under general identifiers only distinguishing trees
@@
 replaced through multinomial trials based on the association probability
-\(\big[\alpha_1, \alpha_2,???, \alpha_3\big]\) observed across all
-inventories between each vernacular name \emph{v} and the species
-\(\big[s_1, s_2, ???, s_N\big]\):
+\(\big[\alpha_1, \alpha_2,..., \alpha_3\big]\) observed across all
+inventories between each vernacular name \emph{v} and all species
+\(\big[s_1, s_2,..., s_N\big]\):
 
 \begin{align}
-M_v\Big(\big[s_1, s_2, ???, s_N\big],\big[\alpha_1, \alpha_2,???, \alpha_3\big]\Big) \nonumber
+M_v\Big(\big[s_1, s_2,..., s_N\big],\big[\alpha_1, \alpha_2,..., \alpha_3\big]\Big) \nonumber
 \end{align}
@@
 in French Guiana, including two in Paracou, and comprised 294 species
-pertaining to 157 genera. Missing trait values were filled using
+pertaining to 157 genera. Missing trait values (10\%) were filled using
 multivariate imputation by chained equation \citep{Mice2011}.
-Imputations were restricted within genus, or family when samples were
-too scarce, in order to account for the phylogenetic signal. Whenever a
-species was not in the dataset, it was attributed a set of traits values
+Imputations were restricted within genus or family when samples were too
+scarce, in order to account for the phylogenetic signal. Whenever a
+species was not in the dataset, it was attributed a set of trait values
 randomly sampled among species of the next higher taxonomic level (same
@@
 All composition and diversity metrics were obtained after 50 iterations
-of the taxonomic uncertainty propagation framework.
+of the uncertainty propagation framework.
 
@@
 Taxonomic and functional trajectories of community composition were
-followed in a two-dimensional NMDS ordination space. Two NMDS were
-conducted to map either taxonomic or functional composition, the later
-based on the 7 leaf, stem and life history traits (without seed mass
-classes). NMDS were performed using abundance-based (Bray-Curtis)
-dissimilarity measures. Trajectories along time were reported through
-the euclidean distance between the target inventories and the reference
-inventories in 1989, \emph{i.e.} 2-3 years after disturbance. Univariate
-trajectories of the leaf, stem and life-history traits were also
-visualized with the community weighted means (CWM)
-\citep{Diaz2007, Garnier2004}. Species seed mass corresponded to 5
-classes of increasing mass, seed mass trajectories were therefore
-reported as the proportion of each class in the inventories (Supp Mat
-XX).
+followed in a two-dimensional NMDS ordination space. Two NMDS using
+abundance-based (Bray-Curtis) dissimilarity measures were conducted to
+map either taxonomic or functional composition, the later based on the 7
+leaf, stem and life history traits (without seed mass classes).
+Trajectories along time were reported through the euclidean distance
+between the target inventories and the reference inventories in 1989,
+\emph{i.e.} 2 years after disturbance. Univariate trajectories of the
+leaf, stem and life-history traits were also visualized with the
+community weighted means (CWM) \citep{Diaz2007, Garnier2004}. Species
+seed mass corresponded to 5 classes of increasing mass, seed mass
+trajectories were therefore reported as the proportion of each class in
+the inventories (Supp Mat XX).
 
-The taxonomic diversity was reported through species richness and the
+The taxonomic diversity was reported through species richness and
 evenness, \emph{i.e} the Hill number translation of the Simpson index
-\citep{Hill1973}. These two indices belong to the set of HCDT or
-generalized entropy, respectively corresponding to the 0 and 2 order of
-diversity (q), recomended for diversity studies
+\citep{Hill1973}. These indices belong to the set of HCDT or generalized
+entropy, respectively corresponding to the 0 and 2 order of diversity
+(q), recomended for diversity studies
 \citep{Patil1982, Tothmeresz1995, Marcon2015}. The functional diversity
@@
 species abundance distribution and average pairwise dissimilarity based
-on all functional traits.
+on species functional traits.
 
@@
 reached over the 30 years and the initial \%AGB removed. The were
-besides analysed through the linear correlations between Simpson and Rao
-diversities and the initial \%AGB removed at 10, 20 and 30 years after
-disturbance.
+besides analysed through the linear correlations between (i) taxonomic
+and functional richness, evenness and diversity and (ii) the initial
+\%AGB removed at 10, 20 and 30 years after disturbance.
 
@@
 TDP weighted by species abundance were eventually summed for each
-community: the functional redundancy was the sum of TDPs overlap,
+community. Community functional redundancy was the sum of TDPs overlap,
 expressed as the average number of species that could be removed from
@@
 \ref{fig:IDHplot}. The correlation with disturbance intensity was weak
-for XXXXXXXXXX, for the taxonomic evenness (\(R^2<0.25\)) while highly
-significant for functional diversity (\(0.60<R^2<0.75\)) whatever the
-time since disturbance.
+for te taxonomic richness (\(-0.09<R^2<-0.06\)), for the taxonomic
+evenness (\(R^2<0.25\)) while highly significant for functional
+diversity (\(0.60<R^2<0.75\)) whatever the time since disturbance.
 
@@
 depended on community initial composition, that partly determined the
-pool of recruited species and that constrained the trajectories towards
-the initial composition. The high resilience of communities taxonomy
+pool of recruited species and constrained the trajectories towards the
+initial composition. The high resilience of communities taxonomy
 revealed that species not belonging to the pre-disturbance community

### 2018-07-04 withBubs.Rmd
IDH graphs

--- 2018-07-03 Untitled.Rmd
@@
 
-## Community diversity: richness and evenness
+## Communities richness and evenness
 
@@
 
-```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
+```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
 Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
@@
 load("DB/ReplacementTraj_ForGraphs")
+load("DB/FunctionalTraj_ForGraphs")
 
+par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
 TaxoTraj(CompleteTaxo)
+plotDiv(CompleteFun,remove=TRUE)
+mtext("(c) Rao diversity",side=3,adj=0,line=1)
+mtext("Years since disturbance",side=1,line=2.2,adj=1)
+legend("right",inset=c(-0.34,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
 
@@
 
-```{r FunDiv, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
-load("DB/FunctionalTraj_ForGraphs")
-
-plotDiv(CompleteFun,remove=TRUE)
-mtext("Rao diversity",side=3,adj=0,line=1)
-mtext("Years since disturbance",side=1,line=2.2,adj=1)
-legend("right",inset=c(-0.28,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
-```
-
 The impact of disturbance intensity was examined through the linear correlation between (i) the %AGB removed and (ii) taxonomic richness, evenness and functional diversity after 10, 20 and 30 years \@ref(fig:IDHplot).
@@
 
-```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="xxx to be changed xxx Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance.
+```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/3,fig.cap="Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance.
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.
 Lower panels, Relationship between the initial \\%AGB removed and Simpson (ref:Boldc) and Rao (ref:Boldd) diversities 10, 20 and 30 years after disturbance.", fig.env="figure*"}
-load("DB/SimpsonDraw");load("DB/FunctionalTraj_ForGraphs")
-load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")
+load("DB/RichnessIDH");load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")
 
-par(mfrow=c(1,2),mar=c(3,3,2,1),oma=c(2,1,2,3),no.readonly = T)
-plotDiv(SimpsonTaxo)
-mtext("Simpson diversity",side=3,adj=0,line=1)
-plotDiv(CompleteFun,remove=TRUE)
-mtext("Rao diversity",side=3,adj=0,line=1)
-mtext("Years since disturbance",side=1,line=2.2,adj=1)
-legend("right",inset=c(-0.28,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
-
+par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
+plotIDH(Data=Rich,AgbLoss=AGBloss)
+mtext("(a) Taxonomic Richness",line=1,side=3,adj=0,cex=0.8)
 plotIDH(Data=Sim,AgbLoss=AGBloss)
+mtext("(b) Taxonomic Evenness",line=1,side=3,adj=0,cex=0.8)
 plotIDH(Data=Rao,AgbLoss=AGBloss)
+mtext("(c) Functional diversity",line=1,side=3,adj=0,cex=0.8)
 mtext("initial AGB lost (%)",side=1,line=2.2,adj=1)
-legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
+legend("right",inset=c(-0.34,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
 
@@
 
-## Community Composition
+## Communities Composition
 
@@
 
-```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of functional redundancy within the initial functional space over 30 years after disturbance.
+```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="TTrajectories of functional redundancy within the initial functional space over 30 years after disturbance.
 Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
@@
 
-1. IDH in Space (Disturbance gradient)
-The IDH consistently (predicted) the response of functional diversity to disturbance, but the theory was disproved regarding communities taxonomic richness and evenness.
-In contrast the taxonomic richness and evenness were decoupled from the disturbance intensity, questioning the consistency of the IDH as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+The IDH consistently predicted the response of taxonomic richness to disturbance.
+The IDH translated in first an increase of richness with diturbance intensity and then a decrease above an intensity threshold (20-25% of initial AGB lost).
+The taxonomic evenness was somewhat decoupled from disturbance intensity, as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+In the Guiana Shield, the pool of true pioneer species that recruited after disturbance is restricted to a few common general (e.g.
+Cecropia spp., Vismia spp.) and the post-disturbance richness and evenness, in the taxonomic space, will only depend on the relative proportion of these recruits and the surviving species.
+-> explain what happens at low, intermediate, high intensity.
+-> functionally diversity: pioneer specis are functionally highly different, we still do not how diverse they are. Their increase in abundance is enough to increase the overall FDiv.
 
 2. IDH in Time
-The humped-shaped trajectories of functional diversity along time 
-Still, the taxonomic trajectories were blurried by the hampered recovery of infrequent species that entailed after disturbance a decrease in the taxonomic richness and an increase of the evenness [@Hubbell2001;@Chave2004].
+a. first the drop and then gain. quite similar drop, highly different species gain.
+b. short-lived pioneers versus long-lived pioneers dynamics linked to habitat diversity
+c. porbably, long-lived come after short-lived for high disturbance while they come together for intermediate disturbance.
+d. link to fuctional diversity, the decoupling in time between high and low disturbance intensity.
 
-## The functional redundancy, key of the taxonomic resilience
+## The functional redundancy, key of the system resilience
 
-The recovery of the most infrequent species followed the lottery recruitment rules [@Busing2002] and was hampered by the increasing competition among species, implying long term recovery [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+1.
+drop in functional redundancy due to random species loss and replacement by pioneer species ( high light, space and nutrient niches) that do not belong to the initial functional space.
+the highest the disturbance, the highest the loss and replacement, the deepest the drop in FR.
+- first 15 years
+2. 2nd step: the system started losing pioneer species (less competitive) and replaced them by not-pioneer (either long-lived or end-succesional sp, more competitive), FR increased.
+3. end of the process: all functionally different species are lost and replaced by individuals belonging to species from the pre-dicturbance community.
+-> long-process.
+hard to see how long it may last.
+The recovery of the most infrequent species followed the lottery recruitment rules [@Busing2002] mplying long term recovery [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+Although underway 30 years after disturbance the recovery remained unachieved and after 30 years the recovery it was underway but remained unachieved.
 
-Following disturbance the space and resources made available corresponded to a rapid decrease of the functional redundancy within the initial functional space.
-The high light, space and nutrient niches were rapidly filled by species, mainly the most dominant and frequent ones.
-Thereafter competitive exclusion emerged following the filling of the niches and limited the recruitment of infrequent species [@Busing2002].
-The recovery of communities taxonomic richness and evenness relied upon the recovery of the initial functional redudancy.
-Although underway 30 years after disturbance the recovery remained unachieved and after 30 years the recovery it was underway but remained unachieved.
-This alteration of the functional redundancy meant a lower resilience of pre-disturbance communities and higher chances to see the persistence of disturbance-specific communities, with less species and more pioneers [@Haddad2008; @Burslem2000; @Martin2013].
+This alteration of the functional redundancy likely meant a lower resilience of pre-disturbance communities and higher chances to see the persistence of disturbance-specific communities, with less functionzl redundancy and more pioneers [@Haddad2008; @Burslem2000; @Martin2013].
 Besides the slowed recovery of rare species increased the risks to loose cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
-Apart from the functional characteristics considered here, infrequent species mighthave unique functions i nthe ecosystem or be a key for some fauna.
+Apart from the functional characteristics considered here, infrequent species might have unique functions i nthe ecosystem or be a key ressource or habitat for some fauna.
 
@@
 
-Our study revealed communities cyclic recovery after disturbance allowing the resilience of their functioning and taxonomic composition with the maintenance of initial differences among communities.
+[Our study revealed communities cyclic recovery after disturbance allowing the resilience of their functioning and taxonomic composition with the maintenance of initial differences among communities.
 Communities functional evenness was enhanced for  20 years after disturbance through the enrichment of the communities with pioneers and light-demanding species,in accordance with the IDH.
-The IDH, though, poorly predicted the disturbance impact on communities taxonomic richness and evenness that were blurried by the emergence of competitive exclusion along time.
+The IDH, though, poorly predicted the disturbance impact on communities taxonomic richness and evenness that were blurried by the emergence of competitive exclusion along time.]
 The resilience of tropical forests proved consistent although spread over several decades. 
 Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Gourlet-Fleury2005].
-As the trajectories highlighted the recruitment processes proved central for communities response to disturbance and closer focus on demographical drivers of communities response would clarify the fate of the future forests.
-
-Communities functional characteristics are the most direct link between biodiversity and ecosystem functioning and their resilience here meant the recovery of ecosystem processes and services on a cycle of several decades [@Diaz2005;@Guariguata2001].
+and we know that regularly disturbed communities are less efficient to recover after disturbance that undisturbed ones..
 

### 2018-07-04 Scripts/DivFunctional_trajectories.R

@@ -33,10 +33,10 @@ smooth<-function(mat,larg){return(do.call(cbind,lapply(1:ncol(mat),function(step
   rowSums(mat[,range])/length(range)})))}
 
 
-CompleteFun<-lapply(1:4,function(t){
+CompleteRichnessFun<-lapply(1:4,function(t){
   
   treat<-treatments[[t]]
-  Nrep<-50
+  Nrep<-5
   
    Mplot<-mclapply(1:Nrep,function(r){
       
@@ -62,7 +62,7 @@ CompleteFun<-lapply(1:4,function(t){
             dissim<-as.matrix(daisy(tra,metric="gower"))
             dissim <- 1 - dissim/max(dissim)
             
-            return(expq(Hqz(as.AbdVector(tapply(Mc,Mc,length)), q=2, Z=dissim,Correction="None"),q=2))
+            return(expq(Hqz(as.AbdVector(tapply(Mc,Mc,length)), q=0, Z=dissim,Correction="None"),q=0))
           })
       
           names(Mean)<-names(M)
@@ -86,5 +86,5 @@ CompleteFun<-lapply(1:4,function(t){
                dimnames=list(treat,dates,1:Nrep)))
 })
 
-save(CompleteFun,file="DB/FunctionalTraj_ForGraphs")
+save(CompleteRichnessFun,file="DB/FunctionalRichnessTraj_ForGraphs")
 

### 2018-07-04 Scripts/DrawingTools.R

@@ -129,13 +129,17 @@ plotIDH<-function(Data,AgbLoss){
     abs<-AgbLoss[which(AgbLoss[,"plot"]%in%names(toplot)),"AGB"]
     points(abs,toplot,col=colyear[Ti],pch=20)
     Lm<-lm(toplot~abs)
-    abline(a=Lm$coefficients[1],b=Lm$coefficients[2],col=colyear[Ti],lwd=2.5)
-    return(round(summary(Lm)$adj.r.squared,2))
-   #legend(round(summary(Lm)$adj.r.squared,2),x=min(AgbLoss),y=seq(Ylim[2],Ylim[1],length.out=10)[Ti],
-    #       bty="n",lty=1,col=colyear[Ti],lwd=2.5,cex=0.8)
+    Lm2<-lm(toplot~abs+I(abs^2))
+    if(abs(round(summary(Lm)$adj.r.squared,2))<abs(round(summary(Lm2)$adj.r.squared,2))){
+      abs_pred<-seq(min(abs),max(abs),length.out=100)
+      lines(sort(abs_pred),predict(Lm2,newdata=data.frame(abs=abs_pred)),col=colyear[Ti],lwd=2.5)
+      return(round(summary(Lm2)$adj.r.squared,2)) }
+    if(abs(round(summary(Lm)$adj.r.squared,2))>abs(round(summary(Lm2)$adj.r.squared,2))){
+      abline(a=Lm$coefficients[1],b=Lm$coefficients[2],col=colyear[Ti],lwd=2.5)
+      return(round(summary(Lm)$adj.r.squared,2)) }
   })
   legend("topleft",legend=unlist(leg),bty="n",lty=1,col=colyear,lwd=2.5,cex=0.8,title=expression(paste('R'^2,'adjusted')))
-  }
+}
 
 plotDiv<-function(Data,remove=FALSE){
   

### 2018-07-04 Scripts/Stats.R

@@ -45,13 +45,14 @@ mtext("initial %AGB lost",side=1,line=2.2,adj=1)
 legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
 
 
-load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")
+load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB");load("DB/RichnessIDH")
 time<-c("1995","2005","2015")
 colyear<-c("darkgoldenrod1","darkorange2","darkred")
-Data<-Rao
+Data<-Rich
 AgbLoss<-AGBloss
 
-Ylim<-c(min(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))),
+plotIDH<-function(Data,AgbLoss){
+  Ylim<-c(min(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))),
           max(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))))
 plot(AgbLoss[,"AGB"],AgbLoss[,"AGB"],type="n",xlab="",ylab="",
        ylim=Ylim)
@@ -60,17 +61,25 @@ leg<-lapply(1:3,function(Ti){
     abs<-AgbLoss[which(AgbLoss[,"plot"]%in%names(toplot)),"AGB"]
     points(abs,toplot,col=colyear[Ti],pch=20)
     Lm<-lm(toplot~abs)
+    Lm2<-lm(toplot~abs+I(abs^2))
     #summary(Lm)
     #anova(Lm)
     #par(mfrow=c(2,2));plot(Lm)
+    if(abs(round(summary(Lm)$adj.r.squared,2))<abs(round(summary(Lm2)$adj.r.squared,2))){
+    abs_pred<-seq(min(abs),max(abs),length.out=100)
+    lines(sort(abs_pred),predict(Lm2,newdata=data.frame(abs=abs_pred)),col=colyear[Ti],lwd=2.5)
+    return(round(summary(Lm2)$adj.r.squared,2)) }
+    if(abs(round(summary(Lm)$adj.r.squared,2))>abs(round(summary(Lm2)$adj.r.squared,2))){
     abline(a=Lm$coefficients[1],b=Lm$coefficients[2],col=colyear[Ti],lwd=2.5)
-    return(round(summary(Lm)$adj.r.squared,2))
[…]

### 2018-07-06 WholePlotTrajectories.Rmd
graphs, stats,etc

@@
 
-(ref:capWD) *WD*
+(ref:capWSG) *WSG*
 
@@
 The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
-They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local climate, and the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The growing demand in forests products together with current climatic changes increases the pressure on remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threatens the maintenance and dynamics in space and time of communities structure, composition and functioning [@Anderson-Teixeira2013;@Sist2015].
+They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, the carbon, water and nutrient cycles, and ensure cultural and human well-being.
+The growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Gibson2011a; @Morales-Hidalgo2015] and threatens the maintenance and dynamics in space and time of communities structure, composition and functioning [@Anderson-Teixeira2013;@Sist2015].
 
-In tropical forests, ecological communities are constantly re-shaped by natural disturbance events that change both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions like competition among species [@Chesson2000;@Herault2018].
+In tropical forests, ecological communities are regularly re-shaped by natural disturbance events changing both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions such as competition among species [@Chesson2000;@Herault2018].
 One of the cornerstone of tropical forest ecology is to understand the processes and drivers of ecosystems response to disturbance [@White2001;@Chazdon2003a]. 
-For now, this has been largely studied through forest structural parameters, rapid and convenient to measure, as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
-These structural parameters were thereafter sucessfully modeled, giving important insights into the recovery of ecosystems processes and services [@Herault2018]. 
-However the response of forests diversity is still unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014, @Liang2016] and would be most probably impacted by the changes induced by disturbance [@Baraloto2012a].
-In the short-term, moderate disturbance may lead to positive impacts on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity when the disturbance regime is moderate but significant [@Molino2001;@Kariuki2006a;@Berry2008a].
-Along time the IDH would translate into a variety of post-disturbance trajectories shaped by the different ecological processes at stake.
+For now, this has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016] that are rapid and convenient to measure.
+These structural parameters have been sucessfully modeled, giving important insights into the recovery of ecosystem processes and services [@Herault2018]. 
+However the response of forests diversity and composition is still unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014; @Liang2016].
+In the short-term, moderate disturbance may lead to positive impacts on communities diversity, an idea formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity when disturbance intensity is not too high [@Molino2001;@Kariuki2006a;@Berry2008a].
 
-Validations of the IDH though remain scarce in the long term and mainly rely on the analysis of species richness.
-Species richness though gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016]: more complete analysis would encompass communities composition that is crucial for conservation issues, and abundance distribution that reveals the ecological processes involved [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, a functional approach accounting for species biological attributes and assessing their role in the ecosystem would directly link communities diversity to the ecosystem functioning and to its environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-In that respect, the functional trait-based approach that focused on major traits related to species ecology and performance was sucessfully adopted [@Diaz2005;@Villeger2008a].
-For example the functoinal approach revealed in tropical rainforests the deterministic processes fostering fast growing species with efficient resources acquisition after disturbance [@Molino2001;@Haddad2008;@Ruger2009].
-Disturbance therefore entails a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
-This shift is translated into consistent trajectories of key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
+Validations of the IDH though remain scarce in the long-term and mainly rely on the analysis of taxonomic richness [@Molino2001].
+Taxonomic richness, alone, gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+More ecological-meaningful analysis would couple richness with (i) evenness that would reveal underlying ecological processes and (ii) composition that is crucial for conservation issues [@Magurran1988;@Lavorel2002;@Bellwood2006].
+Furthermore, a functional approach accounting for species biological attributes would directly link communities diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
+In that respect, the functional trait-based approach that focus on major traits related to species ecology and mediate species performance in a given environment was sucessfully adopted [@Diaz2005;@Villeger2008a].
+For instance, the functional approach revealed in tropical rainforests the deterministic processes entailing, after disturbance, a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
+This shift is translated into the trajectories of key functional traits related to resource acquisition (leaf and stem traits) and life-history traits (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
 Eventually a complete overview of communities response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
-High functional redundancy like in the very diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities [@Trenbath1999; @Elmqvist2003; @Diaz2005].
-High functional redundancy also increases the functional overlap among species and would foster neutral stochastic processes [@Gravel2006].
+The high functional redundancy of hyperdiverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Trenbath1999; @Elmqvist2003; @Diaz2005].
 
-To grasp all facets of communities response to disturbance we examined here the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy [@Lohbeck2015;@Guariguata2001].
-These trajectories would highlight the recovery of communities initial characteristics and the underlying ecological processes, and provide indications for future adaptive conservation strategies [@Adler2007].
-They would clarify the tenants of the Intermediate Disturbance Hypothesis in the long term, debated in tropical forests, and explicit the variety of post-disturbance trajectories defined by the disturbance intensity.
-We monitored over 30 years the response of 75 ha of neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
-We made use of a large functional traits database browsing major leaf, stem and seed functional traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition, richness, evenness and redundancy.
-Specifically, we (i) questioned the recovery of communities taxonomic and functional characteristics and identify the underlying assembly processes, (ii) clarified the validity of the IDH in the long term for tropical forest and elucidated its declination into different trajectories, and (iii) questioned the length of complete communities recovery.
+In this study, we monitored over 30 years the response of 75 ha of neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem above-ground biomass (AGB) loss. 
+We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy [@Lohbeck2015;@Guariguata2001].
+Specifically, we (i) questioned the recovery of communities taxonomic and functional characteristics and identified the underlying assembly processes, (ii) clarified the validity of the IDH in the long term for tropical forest and elucidated its translation into different trajectories, and (iii) questioned community recovery time.
 
@@
 
-Paracou station in French Guiana (5°18'N and 52°53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26°C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
+Paracou station in French Guiana (5\textdegree 18'N and 52\textdegree 53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26\textdegree C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
 Elevation ranges between 5 and 50 m and soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
 
-The experiment is a network of twelve 6.25ha plots that underwent a gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) according to a randomized plot design with three replicate blocks of four plots.
+The experiment is a network of twelve 6.25ha plots that underwent a disturbance gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) according to a randomized plot design with three replicate blocks of four plots.
 The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3).
@@
 
-The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical families. 
-In the twelve experimental plots of the experiment, all trees above 10 cm DBH have been mapped and measured annually since 1984. 
-Trees are first identified during inventories with a vernacular name assigned by the forest worker team, and afterward with a scientific name assigned by botanists during regular botanical campaigns. 
-In 1984, specific vernacular names are given to 62 commercial or common species whereas more infrequent ones were identified under general identifiers only distinguishing trees and palm trees. 
+The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae. 
+In the twelve experimental plots, all trees above 10 cm DBH have been mapped and measured annually since 1984. 
+Trees are first identified with a vernacular name assigned by the forest worker team, and afterward with a scientific name assigned by botanists during regular botanical campaigns. 
+In 1984, specific vernacular names were given to 62 commercial or common species whereas more infrequent ones were identified under general identifiers only distinguishing trees and palm trees. 
 From 2003, botanical campaigns have been conducted every 5 to 6 years to identify all trees at the species level but identification levels still varied among plots and campaigns.
 
-This variability of protocols raised methodological issues as vernacular names usually correspond to different botanical species. 
+This variability of protocols in time raised methodological issues as vernacular names usually correspond to different botanical species. 
 It resulted in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics.
 The uncertainty propagation was done through a Bayesian framework reconstituting complete inventories at genus level from real incomplete ones on the basis of vernacular/botanical names association. 
-Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,…, \alpha_3\big]$ observed across all inventories between each vernacular name *v* and the species $\big[s_1, s_2, …, s_N\big]$:
+Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,..., \alpha_3\big]$ observed across all inventories between each vernacular name *v* and all species $\big[s_1, s_2,..., s_N\big]$:
 
 \begin{align}
-M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big) \nonumber
+M_v\Big(\big[s_1, s_2,..., s_N\big],\big[\alpha_1, \alpha_2,..., \alpha_3\big]\Big) \nonumber
 \end{align}
@@
 
-Six functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area, the leaf area per unit dry mass) and wood economics (wood specific gravity and bark thickness), and life history traits (maximum specific height and seed mass) came from the BRIDGE project [^1].
-Traits values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 botanical species pertaining to 157 botanical genera.
-Missing trait values were filled using multivariate imputation by chained equation [@Mice2011].
-Imputations were restricted within genus, or family when samples were too scarce, in order to account for the phylogenetic signal of the functional traits. 
-Whenever a species inventoried was not in the dataset, it was attributed a set of traits values randomly sampled among species of the same next higher taxonomic level (same genus or family). 
-As seed mass information corresponds to a classification into mass classes, no data filling process was applied and analysis were restricted to the 414 botanical species recorded.
+Six functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area) and stem economics (wood specific gravity and bark thickness), and life-history traits (maximum specific height and seed mass) came from the BRIDGE project [^1].
+Trait values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 species pertaining to 157 genera.
+Missing trait values (10%) were filled using multivariate imputation by chained equation [@Mice2011].
+Imputations were restricted within genus or family when samples were too scarce, in order to account for the phylogenetic signal. 
+Whenever a species was not in the dataset, it was attributed a set of trait values randomly sampled among species of the next higher taxonomic level (same genus or family). 
+As seed mass information was classified into classes, no data filling process was applied and analyses were restricted to the 414 botanical species recorded.
 
-All composition and diversity metrics were obtained after 50 iterations of the taxonomic uncertainty propagation framework and of the filling process of missing trait values.
+All composition and diversity metrics were obtained after 50 iterations of the uncertainty propagation framework.
 
@@
 
-To counter taxonomic uncertainties due to the variability of botanical identification levels (in space) and protocols (in time), the taxonomic composition and diversity analysis were conducted at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
-Trajectories of communities taxonomic and functional variations in composition after disturbance were followed in a two-dimensional NMDS ordination space of the 30 years monitored.
-Two NMDS were conducted to map either taxonomic flora inventories or communities functional composition based on the 7 leaf, stem and life history traits (without seed mass classes).
-In both cases the NMDS were performed using abundance-based (Bray-Curtis) dissimilarity measures. 
-Trajectories along time were reported through the distance between successive inventories and the reference inventories in 1989, *i.e.* 2-3 years after disturbance, when the uncertainty degree did not exceed 30% of undetermined trees.
-for the taxonomix composition trajectories were measured by the euclidean distance between inventories in the multivariate space of all inventories species.
-For the functoinal composition trajectories were measured by the sum for all traits of the euclidean distances between the traits density distributions of successive inventories.
-To compensate the intrinsinc difference among plots, the trajectories corresponded to the euclidean differences along time with the reference inventory in 1989.
+To counter taxonomic uncertainties due to the variability of botanical identification levels (in space) and protocols (in time), the taxonomic composition and diversity analysis were conducted at the genus level.
+Taxonomic and functional trajectories of community composition were followed in a two-dimensional NMDS ordination space.
+Two NMDS using abundance-based (Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the later based on the 7 leaf, stem and life history traits (without seed mass classes).
+Trajectories along time were reported through the euclidean distance between the target inventories and the reference inventories in 1989, *i.e.* 2 years after disturbance. 
 Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007;@Garnier2004]. 
-Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were therefore reported as the proportion of each class in the inventories.
+Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were therefore reported as the proportion of each class in the inventories (Supp Mat XX).
 
-The taxonomic diversity was reported through species richness and the evenness, *i.e* the Hill number translation of the Simpson index [@Hill1973].
-These two indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995;@Marcon2015].
-The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on all functional and life traits.
+The taxonomic diversity was reported through species richness and evenness, *i.e* the Hill number translation of the Simpson index [@Hill1973].
+These indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995;@Marcon2015].
+The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on species functional traits.
 
-The impacts of initial disturbance were tested with the spearman rank correlation between the extremum of taxonomic and functional metrics reached over the 30 years and the initial %AGB removed. 
-The were besides analysed through the linear correlations between Simpson and Rao diversities and the initial %AGB removed at 10, 20 and 30 years after disturbance.
+The impacts of initial disturbance were tested with the spearman rank correlation between the extrema of taxonomic and functional metrics reached over the 30 years and the initial %AGB loss. 
+They were besides analysed through polynomial regression between (i) taxonomic and functional richness, evenness and diversity and (ii) the initial %AGB loss at 10, 20 and 30 years after disturbance.
 
 The functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
-The samples of the trait database were first mapped in a 2-dimensional plan from a PCA analysis. 
+The samples of the trait database were first mapped in a 2-dimensional plan with a PCA analysis. 
 Then, multivariate kernel density estimator associated with individual trees returned species traits probability distribution (TDP).
-Species TDP weighted by species abundance were eventually summed for each community: the functional redundancy was the sum of TDPs overlap, expressed as the average number of species that could be removed from without reducing the functional space (see appendix I for a more comprehensive sheme).
+Species TDP weighted by species abundance were eventually summed for each community.
+Community functional redundancy was the sum of TDPs overlap, expressed as the average number of species that could be removed from without reducing the functional space (see appendix I for a more comprehensive sheme).
 
@@
 
-From 1989 (2-3 years after disturbance) to 2015 (28-29 years after disturbance), 828388 individual trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
-For undisturbed plots taxonomic Richness and Evenness (Simpson diversity) remained stable over the 30 years monitored.
-In disturbed communities after low disturbance intensity the taxonomic richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
-After intense disturbance the taxonomic richness followed unimodal trajectories, decreasing for ten years after disturbance before recovering to pre-disturbance values.
-In all disturbed plots the taxonomic evenness (Simpson diversity) followed unimodal trajectories with first an increase until a maximum reached after around 20 years that was positively correlated to the disturbance treatment ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
-The return towards initial values of communities evenness was beginning after 30 years except for two T3 plots (plots 8 and 12) which evenness still increased, suggesting similar but delayed trajectories \@ref(fig:DivTaxo).
+From 1989 (2 years after disturbance) to 2015 (28 years after disturbance), 828 388 individual trees and 591 botanical species pertaining to 223 genus and 64 families were recorded. 
+For undisturbed plots, taxonomic Richness and Evenness remained stable over the 30 years monitored.
+In disturbed communities, after low disturbance intensity the taxonomic richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
+After intense disturbance the taxonomic richness followed a more complex trajectory, decreasing for ten years after disturbance before recovering to pre-disturbance values.
+The maximum richness loss or gain after disturbance was positively correlated to the disturbance intensity ($\rho_{spearman}^{Richness}=0.50$).
+In all disturbed plots the taxonomic evenness first increased until a maximum reached after around 20 years.
+This maximum was positively correlated to the disturbance intensity ($\rho_{spearman}^{Evenness}=0.77$).
+The evenness then stabilized except for two T3 plots (plots 8 and 12) for which evenness kept increasing.
 
-```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
+```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic (ref:Bolda) richness, (ref:Boldb), taxonomic evenness and (ref:Boldc) functional diversity.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
 load("DB/ReplacementTraj_ForGraphs")
+load("DB/FunctionalTraj_ForGraphs")
 
+par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
 TaxoTraj(CompleteTaxo)
+plotDiv(CompleteFun,remove=TRUE)
+mtext("(c) Functional diversity",side=3,adj=0,line=1)
+mtext("Years since disturbance",side=1,line=2.2,adj=1)
+legend("right",inset=c(-0.34,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
 
@@
 
-The plot 7 from treatment 1 displayed a constantly outlying functional evenness and was removed from the graphical representation for better readability (see appendix for full graphs).
-In undisturbed plots the functional evenness remained stable along the 30 years while in disturbed plots it followed unimodal trajectories with a return towards initial values starting around 20 years after disturbance.
+The plot 7 from treatment 1 displayed a constantly outlying functional diversity and was removed from the graphical representation for better readability (see appendix for full graphs).
+In undisturbed plots the functional diversity remained stable along the 30 years.
+In disturbed plots, trajectories depend on the disturbnace intensity with, for low intensity, a low but long-lasting increase up to amximum reached after 20-25 years and, for, high intensity, a fast but short increase followed, after 10 years, by a slow decrease towards the inital values.
 
-```{r FunDiv, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
-load("DB/FunctionalTraj_ForGraphs")
+The second-degree polynomial regressions between (i) the %AGB loss and (ii) taxonomic richness, evenness and functional diversity after 10, 20 and 30 years best predicted the hump-shaped curve of the disturbance impact along the disturbance intensity gradient \@ref(fig:IDHplot).
+Along the disturbance intensity gradient the disturbance impact was more markedly hump-shaped for the taxonomic richness than for the taxonomic evenness and the functional diversity, the latest being almost linear.
+The regression model better predicted the functional diversity ($0.60<R^2_{Functional}<0.81$) than the taxonomic richness and evenness ($0.21<R^2_{Richness}<0.4$, and $-0.15<R^2_{Evenness}<0.43$ respectively)
 
-plotDiv(CompleteFun,remove=TRUE)
-mtext("Rao diversity",side=3,adj=0,line=1)
-mtext("Years since disturbance",side=1,line=2.2,adj=1)
-legend("right",inset=c(-0.28,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
-```
+```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/3,fig.cap="Relationship between the initial \\%AGB loss and community taxonmic richness (ref:Bolda), taxonomic evenness (ref:Boldb) and functional diversity (ref:Boldc) at 10, 20 and 30 years after disturbance.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
+load("DB/RichnessIDH");load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")
 
-The impact of disturbance was examined specifically through the linear correlation between the intial %AGB removed and the Simpson and Rao diversities (diversities of order 2) after 10, 20 and 30 years \@ref(fig:IDHplot).
-The correlation with disturbance intensity was weak for the Simpson diversity ($R^2<0.25$) and only valid from 20 years after disturbance but it was much stronger for the Rao diversity ($0.60<R^2<0.75$) for all the time studied.
-Slope of linear correlations, reflecting the impact of disturbance, was the highest 20 years after disturbance.
-
-```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.
-Lower panels, Relationship between the initial \\%AGB removed and Simpson (ref:Boldc) and Rao (ref:Boldd) diversities 10, 20 and 30 years after disturbance.", fig.env="figure*"}
-load("DB/SimpsonDraw");load("DB/FunctionalTraj_ForGraphs")
-load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")
-
-par(mfrow=c(1,2),mar=c(3,3,2,1),oma=c(2,1,2,3),no.readonly = T)
-plotDiv(SimpsonTaxo)
-mtext("Simpson diversity",side=3,adj=0,line=1)
-plotDiv(CompleteFun,remove=TRUE)
-mtext("Rao diversity",side=3,adj=0,line=1)
-mtext("Years since disturbance",side=1,line=2.2,adj=1)
-legend("right",inset=c(-0.28,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
-
+par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
+plotIDH(Data=Rich,AgbLoss=AGBloss)
+mtext("(a) Taxonomic Richness",line=1,side=3,adj=0,cex=0.8)
 plotIDH(Data=Sim,AgbLoss=AGBloss)
+mtext("(b) Taxonomic Evenness",line=1,side=3,adj=0,cex=0.8)
 plotIDH(Data=Rao,AgbLoss=AGBloss)
-mtext("initial AGB lost (%)",side=1,line=2.2,adj=1)
-legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
+mtext("(c) Functional diversity",line=1,side=3,adj=0,cex=0.8)
+mtext("initial AGB loss (%)",side=1,line=2.2,adj=1)
+legend("right",inset=c(-0.34,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
 
@@
 
-### Taxonomic and functional trajectories
-
-While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed consistent trajectories over time after disturbance which revealed significant compositional changes.
-According to the mapping of functional traits (see appendix I) these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WD to high average SLA and chlorophyll content.
-For disturbed communities the distance of successive inventories to the 1989 reference inventory followed unimodal trajectories translating cyclic compositional changes with a recovery of the initial composition (Figure \@ref(fig:NMDSplans)).
-The maximum dissimilarity with the initial state was positively correlated to the disturbance treatment for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.91$ and $\rho_{spearman}^{functional}=0.96$ respectively) and the time at maximum was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition.
+While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-disturbance time.
+in disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WSG to high average SLA and chlorophyll content (see appendix I).
+For functional composition, this translated into cyclic compositional changes with an unachieved recovery of the initial composition (Figure \@ref(fig:NMDSplans)).
+The maximum dissimilarity with the initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.87$ and $\rho_{spearman}^{functional}=0.90$ respectively).
+The maximum value was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition.
 
@@
 
-### Traits community weighted means (CWM)
+Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions followed unimodal trajectories either stabilizing or returning towards their initial values.
 
-Changes in functional composition trajectories went hand to hand with consistent trajectories of the 8 functional traits (Figure \@ref(fig:CWM)).
-Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions followed unimodal trajectories either stabilizing or returning towards their initial values.
-Thirty years after disturbance the weighted means of communities specific maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WD*) remained significantly lower than their initial value (Figure \@ref(fig:CWM)).
-The weighted means of bark thickness (*Bark_thick*) similarly remained substantially higher than initially for all disturbed plots while the specfic leaf area (*SLA*) had almost recovered its initial value.
-For all traits the maximum difference to initial state was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.67$, $\rho_{spearman}^{L_{chloro}}=0.45$, $\rho_{spearman}^{L_{toughness}}=-0.43$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WD}=-0.78$, $\rho_{spearman}^{Bark-thickness}=0.88$, $\rho_{spearman}^{Hmax}=-0.48$).
+Maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WSG*) first decreased and then slightly increased but remained significantly lower than their initial value (Figure \@ref(fig:CWM)).
+On the other side, Bark thickness (*Bark_thick*) and specific leaf area (*SLA*) increased and while *Bark_thick* remained substantially high after 30 years, *SLA* had almost recovered its initial value.
+For all traits, the maximum difference to initial value was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.76$, $\rho_{spearman}^{L_{chloro}}=0.60$, $\rho_{spearman}^{L_{toughness}}=-0.53$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WSG}=-0.75$, $\rho_{spearman}^{Bark-thickness}=0.71$, $\rho_{spearman}^{Hmax}=-0.40$).
 
-```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWD), and bark thickness, (ref:capBarkThick)) and one life history trait (Specific maximum height at adult stage, (ref:capHmax)).
+```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWSG), and bark thickness, (ref:capBarkThick)) and one life history trait (Specific maximum height at adult stage, (ref:capHmax)).
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure*"}
@@
 
-### Functional redundancy
+## Functional redundancy
 
-In control plots the functional redundancy remained stable but in undisturbed plots it was quite variable and independent of the disturbance intensity (See appendix I).
-Globally after most intense disturbance (plots T2 and T3) communities redundancy decreased at first place before increasing to edge, recover or exceed the initial value.
+All disturbed plots had lower functional redundancy than control plots and followed similar hump-shaped trajectories (\@ref(fig:RedFun_rest)).
+The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{spearman}=0.47$) and the initial value had not recovered for any disturbed communities after 30 years.
 
-Considering the functional redundancy restricted to the functional space of the initial inventory, all disturbed plots followed similar decreasing humped shaped trajectories (\@ref(fig:RedFun_rest)).
-The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{spearman}=0.50$) and the initial value had not recovered for any disturbed communities.
-
-```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="TTrajectories of functional redundancy within the initial functional space over 30 years after disturbance.
+```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
 Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
@@
 
-par(mar=c(5,4,5,2))
+par(mar=c(3,2,3,2))
 RedundancyPlot(RedundancyTraj_restricted)
-mtext("Functional Redundancy",side=3,adj=0,line=1.5)
-mtext("Restricted to initial functional space",side=3,adj=0,cex=0.8,line=0.5)
-mtext("Years since disturbance",side=1,line=2.2,adj=1)
+mtext("Functional Redundancy",side=3,adj=0,line=1.5,cex=1.2)
+mtext("Restricted to initial functional space",side=3,adj=0,cex=0.9,line=0.5)
+mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=1.2)
 
@@
 
-## Communities cyclic recovery
+## A cyclic recovery of community composition
 
-Communities taxonomic and functional composition and functional evenness proved resilient, following similar humped-shaped trajectories starting to return towards pre-disturbance composition and functioning after 30 years.
+Communities taxonomic and functional composition appeared resilient, following similar hump-shaped trajectories starting to return towards pre-disturbance composition after 30 years. 
+The taxonomic differences among communities, marked before disturbance by the distinct starting points on the NMDS axis 2, were maintained throughout recovery trajectories. 
+More than commonly thought, post-disturbance trajectories depended on community initial composition, that partly determined the pool of recruited species and constrained the trajectories towards the initial composition.
+The high resilience of communities taxonomy revealed that species not belonging to the pre-disturbance community were hardly recruited because of the commonness of dispersal limitation among tropical tree species [@Svenning2005].
 
-The taxonomic differences among communities, marked before disturbance by the distinct starting points of the NMDS, were maintained throughout communities recovery trajectories.
-More than commonly thought post-disturbance trajectories depended on communities initial composition, that partly determined the pool of recruitable species and that constrained the converge towards the recovery of the initial composition [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].
-The resilience of communities taxonomy suggested the existence of multiple stable equilibria, here the pre-disturbance communities, as assumed for highly diverse and productive ecosystems [@Chase2003].
+Trajectories of community functional composition appeared similar in time and close in terms of functional composition.
+Because pre-disturbance surviving trees mirror the initial community [@Herault2018], changes in functional composition relied upon the recruitment of species or functional types infrequent or absent before disturbance.
+Competitive pioneers became dominant in filling the environmental niches of high availability of light, space and nutrient made available by the disturbance.
+It changed community functional composition in the same way for all communities, towards more resource-acquisitive strategies [@Grime1998;@Westoby1998; @Wright2004; @Reich2014].
+Later during the community recovery long-lived, more resistant and shade-tolerant species, would exclude those pioneers and restore the pre-disturbance composition.
 
-Communities functional composition followed similar trajectories that were not so distinct in the functional space.
-The trajectories of functional evenness as well were similar among communities 
-Communities functional characteristics are the most direct link between biodiversity and ecosystem functioning and their resilience here meant the recovery of ecosystem processes and services ona cycle of several decades [@Diaz2005;@Guariguata2001].
+## Another perspective on the intermediate disturbance hypothesis
 
-## Another view of the intermediate disturbance hypothesis
+Disturbance clearly enhanced community taxonomic richness until an intensity threshold (20-25% of initial AGB loss), while the taxonomic evenness was somewhat decoupled from the disturbance intensity, as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+The IDH well predicted the response of taxonomic richness to disturbance, and to a lesser extent the taxonomic evenness.
+Post-disturbance communities combine pre-disturbance surviving trees and recruited trees, either from the remaining pre-disturbance community or from a disturbance-specific community.
+In the Guiana Shield, the pool of true pioneer species specifically recruited after disturbance is restricted to a few common genera (e.g. Cecropia spp., Vismia spp.) 
+Below the intensity threshold the taxonomic richness of the pre-disturbance community is maintained and the recruitment of pioneers, previously infrequent or absent, increases both taxonomic richness and evenness.
+The disturbance intensity determined the relative importance of pioneers recruited, so the taxonomic richness and evennes increase all the more so that disturbance is intense.
+However, disturbance in the same time decreases the taxonomic richness of the pre-disturbance community, and beyond the intensity threshold the decrease exceeds the taxonomic enrichment from pioneers.
+The taxonomic evenness for its part is less sensitive to the loss of rare species and the disturbance impact is mitigated as pioneers balance the hyper-dominance of undisturbed tropical forests.
 
-The IDH consistently predicted communities functional response to disturbance and the functional trajectories in time highlighted the underlying ecological mechanisms, but the theory was diproved regarding communities taxonomic richness and evenness.
+Conversely the IDH was disproved for the functional diversity.
+Pioneer species are functionally highly different, and whenever the disturbance intensity their increased abundance suffice to increase the community functional diversity and evenness.
 
-With the studied range of intensity, disturbance enhanced communities functional evenness but did not have clear impact on taxonomic richness or evenness. 
-The IDH proved a good predictor of the disturbance impact on communities functioning but not on their taxonomic richness or evenness.
+Following disturbance the taxonomic richness droped at first place independently of the disturbance intensity, but then increased through a species gain that was dermined by the disturbance intensity.
+Below the intensity threshold the establishment of long-lived pioneers significantly enhanced both taxonomic and functional richness and evenness in the long-term.
+Beyond the intensity threshold the establishment of short-lived pioneers rapidly filling the environmental space made available did not increased the taxonomic richness and entailed a decrease in the functional diversity after 10 years.
+Likely, after intense disturbance the long-lived pioneers will succeed and post-disturbance trajectories will match those of intermediate disturbance, which explains the decoupling in time between high and low disturbance intensity trajectories.
 
-The humped-shaped trajectories of communities along time illustrated the processes enhancing the functional evennes and the changes in functional composition following disturbance before communities return towards initial composition and lower evenness.
-Disturbance entailed significant variability in the communities through the emergence of ecological processes different from thoses involvemnt before disturbance, in accordance with the IDH.
-The increase in functional evenness went with significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
-Because the pre-disturbance survivor trees mirror the initial communities [@Herault2018], the functional changes relied upon recruited trees and the enhanced growth and survival of previously infrequent species and functional types.
-Disturbance vacated environmental niches of high light, space and nutrient availability, filled by competitive pioneers that became dominant and determined the functional characteristics of the community [@Grime1998].
-Before disturbance and over communities recovery time though, these species would be excluded by long-lived, more resistant and shade tolerant species.
-
-In contrast the taxonomic richness and evenness were decoupled from the disturbance intensity, questioning the consistency of the IDH as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
-The recovery of communities functional characteristics, that mostly rely on dominant species according to the "vegetation quantity effect" [@Grime1998], suggested a similar recovery of the dominant species [@Molino2001].
-Still, the taxonomic trajectories were blurried by the hampered recovery of infrequent species that entailed after disturbance a decrease in the taxonomic richness and an increase of the evenness [@Hubbell2001;@Chave2004].
-The disturbance therefore porrly predicted commuities taxonomic richnessa and evennes that rely upon different ecological processes.
+Species first established after disturbance remained in the community but became less dominant, replaced the abundant pre-disturbance species, but the recovery of rare species was hampered in the long-term [@Hubbell2001;@Chave2004].
 

### 2018-07-06 WholePlotTrajectories.tex
graphs, stats,etc

@@
 Six functional traits representing leaf economics (leaves thickness,
-toughness, total chlorophyll content and specific leaf area) and wood
+toughness, total chlorophyll content and specific leaf area) and stem
 economics (wood specific gravity and bark thickness), and life-history
@@
 correlation between the extrema of taxonomic and functional metrics
-reached over the 30 years and the initial \%AGB removed. The were
-besides analysed through the linear correlations between (i) taxonomic
-and functional richness, evenness and diversity and (ii) the initial
-\%AGB removed at 10, 20 and 30 years after disturbance.
+reached over the 30 years and the initial \%AGB removed. They were
+besides analysed through polynomial regression between (i) taxonomic and
+functional richness, evenness and diversity and (ii) the initial \%AGB
+removed at 10, 20 and 30 years after disturbance.
 
@@
 From 1989 (2 years after disturbance) to 2015 (28 years after
-disturbance), 828388 individual trees and 591 botanical species
+disturbance), 828 388 individual trees and 591 botanical species
 pertaining to 223 genus and 64 families were recorded. For undisturbed
@@
 ten years after disturbance before recovering to pre-disturbance values.
-In all disturbed plots the taxonomic evenness first increased until a
-maximum reached after around 20 years. This maximum was positively
+The maximum richness loss or gain after disturbance was positively
 correlated to the disturbance intensity
-(\(\rho_{spearman}^{Shannon}=0.86\), and
-\(\rho_{spearman}^{Simpson}=0.89\)). The evenness then stabilized except
-for two T3 plots (plots 8 and 12) for which evenness still increased.
+(\(\rho_{spearman}^{Richness}=0.50\)). In all disturbed plots the
+taxonomic evenness first increased until a maximum reached after around
+20 years. This maximum was positively correlated to the disturbance
+intensity (\(\rho_{spearman}^{Evenness}=0.77\)). The evenness then
+stabilized except for two T3 plots (plots 8 and 12) for which evenness
+kept increasing.
 
@@
 
-\caption{Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities \textbf{(a)} Richness, \textbf{(b)} Shannon and \textbf{(c)} Simpson diversities.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:DivTaxo}
+\caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic \textbf{(a)} richness, \textbf{(b)}, taxonomic evenness and \textbf{(c)} functional diversity.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:DivTaxo}
 \end{figure*}
@@
 
-The impact of disturbance intensity was examined through the linear
-correlation between (i) the \%AGB removed and (ii) taxonomic richness,
-evenness and functional diversity after 10, 20 and 30 years
-\ref{fig:IDHplot}. The correlation with disturbance intensity was weak
-for te taxonomic richness (\(-0.09<R^2<-0.06\)), for the taxonomic
-evenness (\(R^2<0.25\)) while highly significant for functional
-diversity (\(0.60<R^2<0.75\)) whatever the time since disturbance.
+Among polynomial regressions between (i) the \%AGB removed and (ii)
+taxonomic richness, evenness and functional diversity after 10, 20 and
+30 years the second degree better predicted the hump-shaped curve of the
+disturbance impact in function of the intensity \ref{fig:IDHplot}. The
+correlation with disturbance intensity was weak for the taxonomic
+richness (\(R^2<0.37\)) and taxonomic evenness (\(R^2<0.43\)) but highly
+significant for functional diversity (\(0.60<R^2<0.81\)).
 
@@
 
-\caption{Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.
-Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities 10, 20 and 30 years after disturbance.}\label{fig:IDHplot}
+\caption{Relationship between the initial \%AGB removed and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)} and functional diversity \textbf{(c)} at 10, 20 and 30 years after disturbance.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:IDHplot}
 \end{figure*}
@@
 shifts towards species with more acquisitive functional strategies, from
-communities with high average WD to high average SLA and chlorophyll
+communities with high average WSG to high average SLA and chlorophyll
 content (see appendix I). For functional composition, this translated
@@
 disturbance intensity for both taxonomic and functional composition
-(\(\rho_{spearman}^{taxonomic}=0.91\) and
-\(\rho_{spearman}^{functional}=0.96\) respectively). The maximum value
+(\(\rho_{spearman}^{taxonomic}=0.87\) and
+\(\rho_{spearman}^{functional}=0.90\) respectively). The maximum value
 was reached around 26 years after disturbance for taxonomic composition
@@
 Maximum height at adult stage (\emph{Hmax}), leaf toughness
-(\emph{L\_toughness}) and wood density (\emph{WD}) first decreased and
-then slightly increased but remained significantly lower than their
-initial value (Figure \ref{fig:CWM}). On the other side, Bark thickness
-(\emph{Bark\_thick}) and specific leaf area (\emph{SLA}) increased and
-while \emph{Bark\_thick} remained substantially high after 30 years,
-\emph{SLA} had almost recovered its initial value. For all traits, the
-maximum difference to initial value was correlated to the disturbance
-intensity (\(\rho_{spearman}^{L_{thickness}}=0.67\),
-\(\rho_{spearman}^{L_{chloro}}=0.45\),
-\(\rho_{spearman}^{L_{toughness}}=-0.43\),
-\(\rho_{spearman}^{SLA}=0.93\), \(\rho_{spearman}^{WD}=-0.78\),
-\(\rho_{spearman}^{Bark-thickness}=0.88\),
-\(\rho_{spearman}^{Hmax}=-0.48\)).
+(\emph{L\_toughness}) and wood specific gravity (\emph{WSG}) first
+decreased and then slightly increased but remained significantly lower
+than their initial value (Figure \ref{fig:CWM}). On the other side, Bark
+thickness (\emph{Bark\_thick}) and specific leaf area (\emph{SLA})
+increased and while \emph{Bark\_thick} remained substantially high after
+30 years, \emph{SLA} had almost recovered its initial value. For all
+traits, the maximum difference to initial value was correlated to the
+disturbance intensity (\(\rho_{spearman}^{L_{thickness}}=0.76\),
+\(\rho_{spearman}^{L_{chloro}}=0.60\),
+\(\rho_{spearman}^{L_{toughness}}=-0.53\),
+\(\rho_{spearman}^{SLA}=0.93\), \(\rho_{spearman}^{WSG}=-0.75\),
+\(\rho_{spearman}^{Bark-thickness}=0.71\),
+\(\rho_{spearman}^{Hmax}=-0.40\)).
 
@@
 
-\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}).
+\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WSG}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}).
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
@@
 The maximum redundancy loss was positively correlated with the
-disturbance intensity (\(\rho_{spearman}=0.50\)) and the initial value
+disturbance intensity (\(\rho_{spearman}=0.47\)) and the initial value
 had not recovered for any disturbed communities after 30 years.
@@
 
-\caption{TTrajectories of functional redundancy within the initial functional space over 30 years after disturbance.
+\caption{Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
 Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}(\#fig:RedFun_rest)
@@
 (sharp increase in the SLA, leaf thickness and bark thickness and
-decrease in wood density, leaf toughness and maximum height)
+decrease in wood specific gravity, leaf toughness and maximum height)
 \citep{Westoby1998, Wright2004, Reich2014}.

### 2018-07-06 Scripts/DrawingTools.R

@@ -3,8 +3,8 @@ Library(c("shape","plotrix"))
 
 treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
 names(treatments)<-c("Control","T1","T2","T3")
-ColorsTr<-c("darkolivegreen2","deepskyblue2","darkorange1","red2")
-colyear<-c("darkgoldenrod1","darkorange2","darkred")
+ColorsTr<-c("darkolivegreen2","gold","orangered","darkred")
+colyear<-c("deepskyblue","cornflowerblue","darkslateblue")
 time<-c("1995","2005","2015")
 
 smooth<-function(mat,larg){return(do.call(cbind,lapply(1:ncol(mat),function(step){
@@ -203,7 +203,7 @@ legendCWM<-function(){
   mtext("SLA\n",at=0.88,line=-2.5,outer=TRUE,cex=0.9)
   mtext(expression(paste(mm^2,".",mg^-1,sep = "")),at=0.84,line=-3,outer=TRUE,cex=0.9)
   
-  mtext("WD\n",at=0.13,line=-17.8,outer=TRUE,cex=0.9)
+  mtext("WSG\n",at=0.13,line=-17.8,outer=TRUE,cex=0.9)
   mtext(expression(paste("g.",cm^-3,sep = "")),at=0.08,line=-18,outer=TRUE,cex=0.9)
   mtext("Bark thickness\n",at=0.4,line=-17.8,outer=TRUE,cex=0.9)
   mtext("mm",at=0.32,line=-18,outer=TRUE,cex=0.9)

### 2018-07-06 Scripts/Stats.R

@@ -83,35 +83,17 @@ plotIDH(Rao,AGBloss)
 
 ##########################################
 ## Rho Spearman, Taxo
-load("DB/TaxoComposition_ForGraphs")
+load("DB/FunDistance_ForGraphs");load("DB/TaxoDistance_ForGraphs");load("DB/LostAGB")
 
-Data_TaxoComp<-MatrepTaxo
+Dist<-TaxoEuclid
+ret<-apply(apply(Dist,c(1,2),median),1,function(li){return(max(abs(li)))})
+ret<-data.frame(cbind(ret,names(ret)))
+colnames(ret)<-c("Dist","plot")
+ret<-merge(ret,AGBloss,by="plot")
+ret<-apply(ret,2,as.numeric)
+cor(ret[,"Dist"],ret[,"AGB"],method="spearman")
 
-MatrepT<-lapply(Data_TaxoComp,function(Rep){
-  Rep<-as.data.frame(Rep)
-  Rep$plot<-substr(rownames(Rep),start=1,stop=regexpr("_",rownames(Rep))-1)
-  Rep$year<-substr(rownames(Rep),start=regexpr("_",rownames(Rep))+1,
-                   stop=nchar(rownames(Rep)))
-  return(Rep)
-})
-
-DistT<-as.data.frame(unlist(lapply(1:12,function(pl){
-  ret<-lapply(MatrepT,function(rep){return(rep[which(rep[,"plot"]==pl),])})
-  ret<-lapply(ret,function(rep){return(rep[which(rep[,"year"]%in%
-                              ret[[which(unlist(lapply(ret,nrow))==min(unlist(lapply(ret,nrow))))[1]]][,"year"]),])})
-  ret<-do.call(rbind,lapply(ret,function(rep){
-    ret2<-apply(rep[,c("NMDS1","NMDS2")],1,function(li){
-      return(sqrt(sum((rep[1,c("NMDS1","NMDS2")]-li)^2)))})
-    names(ret2)<-rep[,"year"]
-    return(ret2)}))
-  ret<-apply(ret,2,median)
-  return(names(ret)[which(ret==max(ret))])})))
-
-treats<-cbind(c(1,6,11,2,7,9,3,5,10,4,8,12),rep(0:3,each=3))
-DistT<-cbind(DistT,rownames(DistT),treats[order(treats[,1]),2])
-colnames(DistT)<-c("Max","Plot","treat")
[…]

### 2018-07-06 WholePlotTrajectories.Rmd
discuss

@@
 The taxonomic evenness for its part is less sensitive to the loss of rare species and the disturbance impact is mitigated as pioneers balance the hyper-dominance of undisturbed tropical forests.
+Conversely the IDH was disproved for the functional diversity.
+The pioneer species were functionally highly different from the pre-disturbance community composition, so their recruitment increased the community functional richness and eveness after disturbance whenever its intensity.
 
-Conversely the IDH was disproved for the functional diversity.
-Pioneer species are functionally highly different, and whenever the disturbance intensity their increased abundance suffice to increase the community functional diversity and evenness.
-
-Following disturbance the taxonomic richness droped at first place independently of the disturbance intensity, but then increased through a species gain that was dermined by the disturbance intensity.
-Below the intensity threshold the establishment of long-lived pioneers significantly enhanced both taxonomic and functional richness and evenness in the long-term.
-Beyond the intensity threshold the establishment of short-lived pioneers rapidly filling the environmental space made available did not increased the taxonomic richness and entailed a decrease in the functional diversity after 10 years.
-Likely, after intense disturbance the long-lived pioneers will succeed and post-disturbance trajectories will match those of intermediate disturbance, which explains the decoupling in time between high and low disturbance intensity trajectories.
-
-Species first established after disturbance remained in the community but became less dominant, replaced the abundant pre-disturbance species, but the recovery of rare species was hampered in the long-term [@Hubbell2001;@Chave2004].
+Community post-disturbance trajectories showed a first drop of the taxonomic richness, similar among communities, followed by a species gain determined by the disturbance intensity.
+Until an intensity threshold the gain in taxonomic richness increased with the disturbance intensity.
+This corresponded to the establishment of long-lived pioneers enhancing community taxonomic richness and evenness in the long term.
+Beyond the intensity threshold a few short-lived pioneers occupied the vacated environmental space, preventing the establishment of further species. 
+The establishment of short-lived pioneers, functionally very different from the pre-disturbance community, rapidly and significantly increased the functional diversity but it started decreasing after 10 years.
+Below the disturbance intensity threshold on the other hand the establishment of long-lived pionneers entailed a later and lower but long-lasting increase of the functional diversity.
+Likely the establishment of long-lasting pioneers will follow the decline of short-live pionneers and we expect the taxonomic and functoinal trajectories to catch up with those observed after intermediate disturbance.
 
@@
 
-The recovery of the most infrequent species followed the lottery recruitment rules [@Busing2002] and was hampered by the increasing competition among species, implying long term recovery [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+The functional redundancy within the pre-disturbance functional space decreased for 15 years because of the loss of species, and was not compensated by the functionally different pionneers.
+The depth of the redundancy decrease relied on the disturbance intensity that enhanced the initial decrease and the functional differences of pionneers from the pre-disturance composition.
+After 15 years the first established pioneers pionneers started to be replaced by more competitive long-lived pionneers or late-successional species resembling more the pre-disturbance functional composition and hence increasing the functional redundancy.
+This replacement though is stochastic and follows the lottery recruitment so because of increasing competition it is increasingly difficult recover pre-disturbance species,specifically the most infrequent ones, as the functional redundancy recovers [@Busing2002].
+The recovery time proved long as for some community it was just initiated after 30 years but it is thus difficult to estimate [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
-Following disturbance the space and resources made available corresponded to a rapid decrease of the functional redundancy within the initial functional space.
-The high light, space and nutrient niches were rapidly filled by species, mainly the most dominant and frequent ones.
-Thereafter competitive exclusion emerged following the filling of the niches and limited the recruitment of infrequent species [@Busing2002].
-The recovery of communities taxonomic richness and evenness relied upon the recovery of the initial functional redudancy.
-Although underway 30 years after disturbance the recovery remained unachieved and after 30 years the recovery it was underway but remained unachieved.
-This alteration of the functional redundancy meant a lower resilience of pre-disturbance communities and higher chances to see the persistence of disturbance-specific communities, with less species and more pioneers [@Haddad2008; @Burslem2000; @Martin2013].
+The long-term impact of disturbance on community functional redundancy meant a lower resilience of pre-disturbance communities and higher chances to see the persistence of disturbance-specific communities, with lower taxonomic richness and more pioneers [@Haddad2008; @Burslem2000; @Martin2013].
 Besides the slowed recovery of rare species increased the risks to loose cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].

### 2018-07-06 WholePlotTrajectories.tex
discuss

@@
 neotropical forest plots set up on a gradient of disturbance intensity,
-from 10 to 60\% of ecosystem above-ground biomass (AGB) removed. We made
+from 10 to 60\% of ecosystem above-ground biomass (AGB) loss. We made
 use of a large functional traits database encompassing major leaf, stem
@@
 correlation between the extrema of taxonomic and functional metrics
-reached over the 30 years and the initial \%AGB removed. They were
-besides analysed through polynomial regression between (i) taxonomic and
+reached over the 30 years and the initial \%AGB loss. They were besides
+analysed through polynomial regression between (i) taxonomic and
 functional richness, evenness and diversity and (ii) the initial \%AGB
-removed at 10, 20 and 30 years after disturbance.
+loss at 10, 20 and 30 years after disturbance.
 
@@
 
-Among polynomial regressions between (i) the \%AGB removed and (ii)
-taxonomic richness, evenness and functional diversity after 10, 20 and
-30 years the second degree better predicted the hump-shaped curve of the
-disturbance impact in function of the intensity \ref{fig:IDHplot}. The
-correlation with disturbance intensity was weak for the taxonomic
-richness (\(R^2<0.37\)) and taxonomic evenness (\(R^2<0.43\)) but highly
-significant for functional diversity (\(0.60<R^2<0.81\)).
+The second-degree polynomial regressions between (i) the \%AGB loss and
+(ii) taxonomic richness, evenness and functional diversity after 10, 20
+and 30 years best predicted the hump-shaped curve of the disturbance
+impact along the disturbance intensity gradient \ref{fig:IDHplot}. Along
+the disturbance intensity gradient the disturbance impact was more
+markedly hump-shaped for the taxonomic richness than for the taxonomic
+evenness and the functional diversity, the latest being almost linear.
+The regression model better predicted the functional diversity
+(\(0.60<R^2_{Functional}<0.81\)) than the taxonomic richness and
+evenness (\(0.21<R^2_{Richness}<0.4\), and \(-0.15<R^2_{Evenness}<0.43\)
+respectively)
 
@@
 
-\caption{Relationship between the initial \%AGB removed and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)} and functional diversity \textbf{(c)} at 10, 20 and 30 years after disturbance.
+\caption{Relationship between the initial \%AGB loss and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)} and functional diversity \textbf{(c)} at 10, 20 and 30 years after disturbance.
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:IDHplot}
@@
 were hardly recruited because of the commonness of dispersal limitation
-among tropical tree species (Svenning, Wright JECOL 2005).
+among tropical tree species \citep{Svenning2005}.
 
-Community functional composition followed similar trajectories that were
-not so distinct in the functional space\ldots{}\ldots{}. Because the
-pre-disturbance survivor trees mirror the initial communities
-\citep{Herault2018}, the functional changes relied upon recruited trees
-and the enhanced growth and survival of previously infrequent species
-and functional types. Disturbance vacated environmental niches of high
-light, space and nutrient availability, filled by competitive pioneers
-that became dominant and determined the functional characteristics of
-the community \citep{Grime1998}. Before disturbance and over communities
-recovery time though, these species would be excluded by long-lived,
-more resistant and shade tolerant species. The recovery of communities
-functional characteristics, that mostly rely on dominant species
-according to the ``vegetation quantity effect'' \citep{Grime1998},
-suggested a similar recovery of the dominant species \citep{Molino2001}.
-significant functional shifts towards resource-acquisitive strategies
-(sharp increase in the SLA, leaf thickness and bark thickness and
-decrease in wood specific gravity, leaf toughness and maximum height)
-\citep{Westoby1998, Wright2004, Reich2014}.
+Trajectories of community functional composition appeared similar in
+time and close in terms of functional composition. Because
+pre-disturbance surviving trees mirror the initial community
+\citep{Herault2018}, changes in functional composition relied upon the
+recruitment of species or functional types infrequent or absent before
+disturbance. Competitive pioneers became dominant in filling the
+environmental niches of high availability of light, space and nutrient
+made available by the disturbance. It changed community functional
+composition in the same way for all communities, towards more
+resource-acquisitive strategies
+\citep{Grime1998, Westoby1998, Wright2004, Reich2014}. Later during the
+community recovery long-lived, more resistant and shade-tolerant
+species, would exclude those pioneers and restore the pre-disturbance
+composition.
 
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  IDH in Space (Disturbance gradient) The IDH consistently (predicted)
-  the response of functional diversity to disturbance, but the theory
-  was disproved regarding communities taxonomic richness and evenness.
-  The functional range of species that could establish after disturbance
-  was determined by the expanse of the environmental space made
-  available by disturbance.
-\end{enumerate}
+Disturbance clearly enhanced community taxonomic richness until an
+intensity threshold (20-25\% of initial AGB loss), while the taxonomic
+evenness was somewhat decoupled from the disturbance intensity, as
+already observed in the Guiana Shield \citep{Baraloto2012a} and in
+Bornean tropical forests \citep{Cannon1998}. The IDH well predicted the
+response of taxonomic richness to disturbance, and to a lesser extent
+the taxonomic evenness. Post-disturbance communities combine
+pre-disturbance surviving trees and recruited trees, either from the
+remaining pre-disturbance community or from a disturbance-specific
+community. In the Guiana Shield, the pool of true pioneer species
+specifically recruited after disturbance is restricted to a few common
+genera (e.g.~Cecropia spp., Vismia spp.) Below the intensity threshold
+the taxonomic richness of the pre-disturbance community is maintained
+and the recruitment of pioneers, previously infrequent or absent,
+increases both taxonomic richness and evenness. The disturbance
+intensity determined the relative importance of pioneers recruited, so
+the taxonomic richness and evennes increase all the more so that
+disturbance is intense. However, disturbance in the same time decreases
+the taxonomic richness of the pre-disturbance community, and beyond the
+intensity threshold the decrease exceeds the taxonomic enrichment from
+pioneers. The taxonomic evenness for its part is less sensitive to the
+loss of rare species and the disturbance impact is mitigated as pioneers
+balance the hyper-dominance of undisturbed tropical forests. Conversely
+the IDH was disproved for the functional diversity. The pioneer species
+were functionally highly different from the pre-disturbance community
+composition, so their recruitment increased the community functional
+richness and eveness after disturbance whenever its intensity.
 
-In contrast the IDH translated in a decrease of the taxonomic richness
-above an intensity threshold while the taxonomic evenness was decoupled
-from disturbance intensity, as already observed in the Guiana Shield
-\citep{Baraloto2012a} and in Bornean tropical forests
-\citep{Cannon1998}. The taxonomic richness and evenness of recruited
-species was fixed and at community scale the post-disturbance taxonomic
-characterisitcs depended on the combination of these recruits and the
-remaining pre-disturbance community.
-
-The intensity of the disturbance determined the balance between trees
-recruited in the pre-disturbance community and those recruited in
-disturbance-specific community, functionally more diverse than
-undisturbed forests but with lower taxonomic richness and higher
-evenness.
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  IDH in Time For the first years following disturbance community
-  functional diversity increased through the filling of the vacated
-  environmental space without competitive exclusion contraints. The
-  maximum functional diversity was reached at the same time so the rate
-  of space filling was similar among communities. The maximum was
-  followed by the emergence of competitive exclusion processes bringing
-  the functional diversity back to pre-disturbance levels and made the
-  disturbance hypothesis intermediate in time.
-\end{enumerate}
-
-For medium disturbance intensity enhancing the taxonomic richness and
-evennes, the taxonomic trajectories reached asymptots of high richness
-and high evenness decoupled from their functional trajectories. Species
-first established after disturbance remained in the community but became
-less dominant, replaced the abundant pre-disturbance species, but the
-recovery of rare species was hampered in the long-term
-\citep{Hubbell2001, Chave2004}.
+Community post-disturbance trajectories showed a first drop of the
+taxonomic richness, similar among communities, followed by a species
+gain determined by the disturbance intensity. Until an intensity
+threshold the gain in taxonomic richness increased with the disturbance
+intensity. This corresponded to the establishment of long-lived pioneers
+enhancing community taxonomic richness and evenness in the long term.
+Beyond the intensity threshold a few short-lived pioneers occupied the
+vacated environmental space, preventing the establishment of further
+species. The establishment of short-lived pioneers, functionally very
+different from the pre-disturbance community, rapidly and significantly
+increased the functional diversity but it started decreasing after 10
+years. Below the disturbance intensity threshold on the other hand the
+establishment of long-lived pionneers entailed a later and lower but
+long-lasting increase of the functional diversity. Likely the
+establishment of long-lasting pioneers will follow the decline of
+short-live pionneers and we expect the taxonomic and functoinal
+trajectories to catch up with those observed after intermediate
+disturbance.
 
@@
 
-The recovery of the most infrequent species followed the lottery
-recruitment rules \citep{Busing2002} and was hampered by the increasing
-competition among species, implying long term recovery
-\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+The functional redundancy within the pre-disturbance functional space
+decreased for 15 years because of the loss of species, and was not
+compensated by the functionally different pionneers. The depth of the
+redundancy decrease relied on the disturbance intensity that enhanced
+the initial decrease and the functional differences of pionneers from
+the pre-disturance composition. After 15 years the first established
+pioneers pionneers started to be replaced by more competitive long-lived
+pionneers or late-successional species resembling more the
+pre-disturbance functional composition and hence increasing the
+functional redundancy. This replacement though is stochastic and follows
+the lottery recruitment so because of increasing competition it is
+increasingly difficult recover pre-disturbance species,specifically the
+most infrequent ones, as the functional redundancy recovers
+\citep{Busing2002}. The recovery time proved long as for some community
+it was just initiated after 30 years but it is thus difficult to
+estimate \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
-Following disturbance the space and resources made available
-corresponded to a rapid decrease of the functional redundancy within the
-initial functional space. The high light, space and nutrient niches were
-rapidly filled by species, mainly the most dominant and frequent ones.
-Thereafter competitive exclusion emerged following the filling of the
-niches and limited the recruitment of infrequent species
-\citep{Busing2002}. The recovery of communities taxonomic richness and
-evenness relied upon the recovery of the initial functional redudancy.
-Although underway 30 years after disturbance the recovery remained
-unachieved and after 30 years the recovery it was underway but remained
-unachieved. This alteration of the functional redundancy meant a lower
-resilience of pre-disturbance communities and higher chances to see the
-persistence of disturbance-specific communities, with less species and
-more pioneers \citep{Haddad2008, Burslem2000, Martin2013}. Besides the
-slowed recovery of rare species increased the risks to loose cornerstone
-specie, with unexpected ecological consequences
+The long-term impact of disturbance on community functional redundancy
+meant a lower resilience of pre-disturbance communities and higher
+chances to see the persistence of disturbance-specific communities, with
+lower taxonomic richness and more pioneers
+\citep{Haddad2008, Burslem2000, Martin2013}. Besides the slowed recovery
+of rare species increased the risks to loose cornerstone specie, with
+unexpected ecological consequences
 \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}. Apart from the

### 2018-07-09 WholePlotTrajectories.Rmd
graphs

--- 2018-07-06 WholePlotTrajectories.Rmd
@@
 These indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995;@Marcon2015].
-The functional diversity was reported using the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on species functional traits.
+The functional diversity was reported using the functional richness and evenness, *i.e* Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on species functional traits.
 
@@
 
-```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic (ref:Bolda) richness, (ref:Boldb), taxonomic evenness and (ref:Boldc) functional diversity.
+```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic (ref:Bolda) richness, (ref:Boldb), taxonomic evenness, (ref:Boldc) functional richness, and (ref:Boldd) functional evenness.
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
@@
 load("DB/FunctionalTraj_ForGraphs")
+load("DB/FunctionalRichnessTraj_ForGraphs")
 
-par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
+par(mfrow=c(1,4),mar=c(5,2,4,2),oma=c(1,1.7,1,2),no.readonly=TRUE)
 TaxoTraj(CompleteTaxo)
+plotDiv(CompleteRichFun,remove=TRUE)
+mtext("(c) Functional Richness",side=3,adj=0,line=1,cex=0.75)
 plotDiv(CompleteFun,remove=TRUE)
-mtext("(c) Functional diversity",side=3,adj=0,line=1)
+mtext("(d) Functional Evenness",side=3,adj=0,line=1,cex=0.75)
 mtext("Years since disturbance",side=1,line=2.2,adj=1)
@@
 
-The plot 7 from treatment 1 displayed a constantly outlying functional diversity and was removed from the graphical representation for better readability (see appendix for full graphs).
-In undisturbed plots the functional diversity remained stable along the 30 years.
-In disturbed plots, trajectories depend on the disturbnace intensity with, for low intensity, a low but long-lasting increase up to amximum reached after 20-25 years and, for, high intensity, a fast but short increase followed, after 10 years, by a slow decrease towards the inital values.
+The plot 7 from treatment 1 displayed constantly outlying functional richness and evenness and was removed from the graphical representation for better readability.
+In undisturbed plots both functional richness and evenness remained stable along the 30 years. 
+In disturbed plots, richness and evenness trajectories depended on the disturbance intensity with their maximum positively correlated to %AGB loss $\rho_{spearman}^{Richness}=0.76$ and $\rho_{spearman}^{Evenness}=0.60$.
+Functional richness and evenness displayed for low disturbance intensity a low but long-lasting increase up to maximum reached after 20-25 years and, and for high intensity, a fast but short increase followed after 10 years by a slow decrease towards the inital values.
 
-The second-degree polynomial regressions between (i) the %AGB loss and (ii) taxonomic richness, evenness and functional diversity after 10, 20 and 30 years best predicted the hump-shaped curve of the disturbance impact along the disturbance intensity gradient \@ref(fig:IDHplot).
+The second-degree polynomial regressions between (i) the %AGB loss and (ii) taxonomic and functional richness and evenness after 10, 20 and 30 years best predicted the hump-shaped curve of the disturbance impact along the disturbance intensity gradient \@ref(fig:IDHplot).
 Along the disturbance intensity gradient the disturbance impact was more markedly hump-shaped for the taxonomic richness than for the taxonomic evenness and the functional diversity, the latest being almost linear.
-The regression model better predicted the functional diversity ($0.60<R^2_{Functional}<0.81$) than the taxonomic richness and evenness ($0.21<R^2_{Richness}<0.4$, and $-0.15<R^2_{Evenness}<0.43$ respectively)
+The regression model better predicted the functional richness and evenness ($0.55<R^2_{Functional Richness}<0.72$, and $0.60<R^2_{Functional Evenness}<0.81$) than the taxonomic richness and evenness ($0.21<R^2_{Taxonomic Richness}<0.4$, and $-0.15<R^2_{Taxonomic Evenness}<0.43$ respectively)
 
-```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/3,fig.cap="Relationship between the initial \\%AGB loss and community taxonmic richness (ref:Bolda), taxonomic evenness (ref:Boldb) and functional diversity (ref:Boldc) at 10, 20 and 30 years after disturbance.
+```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/3,fig.cap="Relationship between the initial \\%AGB loss and community taxonmic richness (ref:Bolda), taxonomic evenness (ref:Boldb), functional richness (ref:Boldc),and functional evenness (ref:Boldd) at 10, 20 and 30 years after disturbance.
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
-load("DB/RichnessIDH");load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")
+load("DB/RichnessIDH");load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB");load("DB/FunRichIDH")
 
-par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
+par(mfrow=c(1,4),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
 plotIDH(Data=Rich,AgbLoss=AGBloss)
-mtext("(a) Taxonomic Richness",line=1,side=3,adj=0,cex=0.8)
+mtext("(a) Taxonomic Richness",line=1,side=3,adj=0,cex=0.75)
 plotIDH(Data=Sim,AgbLoss=AGBloss)
-mtext("(b) Taxonomic Evenness",line=1,side=3,adj=0,cex=0.8)
+mtext("(b) Taxonomic Evenness",line=1,side=3,adj=0,cex=0.75)
 plotIDH(Data=Rao,AgbLoss=AGBloss)
-mtext("(c) Functional diversity",line=1,side=3,adj=0,cex=0.8)
+mtext("(c) Functional richness",line=1,side=3,adj=0,cex=0.75)
+plotIDH(Data=FunRich,AgbLoss=AGBloss)
+mtext("(d) Functional evenness",line=1,side=3,adj=0,cex=0.75)
 mtext("initial AGB loss (%)",side=1,line=2.2,adj=1)
@@
 
-Disturbance clearly enhanced community taxonomic richness until an intensity threshold (20-25% of initial AGB loss), while the taxonomic evenness was somewhat decoupled from the disturbance intensity, as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
-The IDH well predicted the response of taxonomic richness to disturbance, and to a lesser extent the taxonomic evenness.
-Post-disturbance communities combine pre-disturbance surviving trees and recruited trees, either from the remaining pre-disturbance community or from a disturbance-specific community.
-In the Guiana Shield, the pool of true pioneer species specifically recruited after disturbance is restricted to a few common genera (e.g. Cecropia spp., Vismia spp.) 
-Below the intensity threshold the taxonomic richness of the pre-disturbance community is maintained and the recruitment of pioneers, previously infrequent or absent, increases both taxonomic richness and evenness.
-The disturbance intensity determined the relative importance of pioneers recruited, so the taxonomic richness and evennes increase all the more so that disturbance is intense.
-However, disturbance in the same time decreases the taxonomic richness of the pre-disturbance community, and beyond the intensity threshold the decrease exceeds the taxonomic enrichment from pioneers.
-The taxonomic evenness for its part is less sensitive to the loss of rare species and the disturbance impact is mitigated as pioneers balance the hyper-dominance of undisturbed tropical forests.
+Disturbance appeared to enhance community taxonomic richness until an intensity threshold (20-25% of initial AGB loss), while the taxonomic evenness was somewhat decoupled from the disturbance intensity as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+The IDH well predicted the disturbance impact on community taxonomic richness and to some extent on taxonomic evenness.
+The disturbance intensity determines the balance in the community between pre-disturbance surviving trees and the trees recruited afterward.
+In the Guiana Shield, the pool of true pioneer species specifically recruited after disturbance is restricted to a few common genera (e.g. Cecropia spp., Vismia spp.).
+Below the intensity threshold the size of the surviving community maintains its pre-disturbance taxonomic richness and the recruitment of pioneers, infrequent or absent before disturbance, increases both community taxonomic richness and evenness.
+Beyond the intensity threshold the taxonomic diversity of the surviving trees decreases and is not offset by the taxonomic enrichment of pioneers.
+The disturbance impact was similar but slighter for community taxonomic evenness, that is less sensitive to the loss of rare species and rather represents the increased dominance of pionneers that balanced the usual hyper-dominance of tropical forest.
+While below the intensity threshold pioneers species increase the diversity of dominant or common species, beyond the threshold they become in turn highly dominant.
 
-Conversely the IDH was disproved for the functional diversity.
-Pioneer species are functionally highly different, and whenever the disturbance intensity their increased abundance suffice to increase the community functional diversity and evenness.
+Conversely the IDH was disproved to predict the dsiturbance impact on community functional diversity.
+Pioneer species were functionally highly different from the pre-disturbance community composition, so their recruitment increased the community functional richness and eveness whenever its intensity.
 
-Following disturbance the taxonomic richness droped at first place independently of the disturbance intensity, but then increased through a species gain that was dermined by the disturbance intensity.
-Below the intensity threshold the establishment of long-lived pioneers significantly enhanced both taxonomic and functional richness and evenness in the long-term.
-Beyond the intensity threshold the establishment of short-lived pioneers rapidly filling the environmental space made available did not increased the taxonomic richness and entailed a decrease in the functional diversity after 10 years.
-Likely, after intense disturbance the long-lived pioneers will succeed and post-disturbance trajectories will match those of intermediate disturbance, which explains the decoupling in time between high and low disturbance intensity trajectories.
+Along time, the post-disturbance taxonomic richness trajectories showed first a similar drop following the species loss during disturbance, and then a species gain determined by the disturbance intensity.
+Until an intensity threshold the gain in taxonomic richness increased with the disturbance intensity through the establishment of long-lived pioneers enhancing community taxonomic richness and evenness in the long term.
+These long-lived pionneers were functionally quite different from the functional composition and their establishment entailed a progressive and long-lasting increase of the functional diversity.
+Beyond the intensity threshold a few short-lived pioneers occupied the vacated environmental space and prevented further species establishment. 
+These short-lived pioneers were functionally very different from the pre-disturbance community and rapidly and significantly increased the functional diversity.
+Already after 10 years, though, they started to decline and the functional diversity started to decrease.
+Likely the establishment of long-lasting pioneers will follow and we expect the taxonomic and functional trajectories to catch up with those observed after intermediate disturbance.
 
-Species first established after disturbance remained in the community but became less dominant, replaced the abundant pre-disturbance species, but the recovery of rare species was hampered in the long-term [@Hubbell2001;@Chave2004].
+## The functional redundancy, key of the community resilience
 
-## The functional redundancy, key of the taxonomic resilience
+For 15 years the species loss during disturbance, determined by the disturbance intensity, commensurately decreased the functional redundancy within the pre-disturbance functional space.
+At first place the pioneers recruited, functionally different from the pre-disturbance functional composition did not compensate for the reundancy decrease.
+Progressively though, first establish species were replaced by more competitive long-lived pionneers or late-successional species resembling more the pre-disturbance functional composition and restoring the functional redundancy.
+This replacement was stochastic and followed the lottery recruitment rules, that ease the recovery for the first recruited species but then increasingly hamper it with the emergence of interspecific competition due to increasing functional redundancy [@Busing2002].
+Along time the recovery of infrequent species is increasingly difficult, and the recovery of the functional redundancy that was just initiated after 30 years in some communities is difficult to estimate [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
-The recovery of the most infrequent species followed the lottery recruitment rules [@Busing2002] and was hampered by the increasing competition among species, implying long term recovery [@Trenbath1999;@Elmqvist2003;@Diaz2005].
-
-Following disturbance the space and resources made available corresponded to a rapid decrease of the functional redundancy within the initial functional space.
-The high light, space and nutrient niches were rapidly filled by species, mainly the most dominant and frequent ones.
-Thereafter competitive exclusion emerged following the filling of the niches and limited the recruitment of infrequent species [@Busing2002].
-The recovery of communities taxonomic richness and evenness relied upon the recovery of the initial functional redudancy.
-Although underway 30 years after disturbance the recovery remained unachieved and after 30 years the recovery it was underway but remained unachieved.
-This alteration of the functional redundancy meant a lower resilience of pre-disturbance communities and higher chances to see the persistence of disturbance-specific communities, with less species and more pioneers [@Haddad2008; @Burslem2000; @Martin2013].
-Besides the slowed recovery of rare species increased the risks to loose cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
-Apart from the functional characteristics considered here, infrequent species mighthave unique functions i nthe ecosystem or be a key for some fauna.
+The long-term impact of disturbance on community functional redundancy meant a lower resilience of pre-disturbance communities and higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008; @Burslem2000; @Martin2013].
+Besides the long-term recovery of infrequent species increases the risks to loose cornerstone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
+Apart from the functional characteristics considered here, infrequent species might indeed have unique functions in the ecosystem or be a key for some fauna.
 
@@
 
-Our study revealed communities cyclic recovery after disturbance allowing the resilience of their functioning and taxonomic composition with the maintenance of initial differences among communities.
+[Our study revealed communities cyclic recovery after disturbance allowing the resilience of their functioning and taxonomic composition with the maintenance of initial differences among communities.
 Communities functional evenness was enhanced for  20 years after disturbance through the enrichment of the communities with pioneers and light-demanding species,in accordance with the IDH.
-The IDH, though, poorly predicted the disturbance impact on communities taxonomic richness and evenness that were blurried by the emergence of competitive exclusion along time.
-The resilience of tropical forests proved consistent although spread over several decades. 
+The IDH, though, poorly predicted the disturbance impact on communities taxonomic richness and evenness that were blurried by the emergence of competitive exclusion along time.]
+The resilience of tropical forests proved consistent although several decades-long. 
 Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Gourlet-Fleury2005].

### 2018-07-09 WholePlotTrajectories.tex
graphs

@@
 \citep{Patil1982, Tothmeresz1995, Marcon2015}. The functional diversity
-was reported using the Rao index of quadratic entropy which combines
-species abundance distribution and average pairwise dissimilarity based
-on species functional traits.
+was reported using the functional richness and evenness, \emph{i.e} Rao
+index of quadratic entropy which combines species abundance distribution
+and average pairwise dissimilarity based on species functional traits.
 
@@
 
-\caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic \textbf{(a)} richness, \textbf{(b)}, taxonomic evenness and \textbf{(c)} functional diversity.
+\caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic \textbf{(a)} richness, \textbf{(b)}, taxonomic evenness, \textbf{(c)} functional richness, and \textbf{(d)} functional evenness.
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:DivTaxo}
@@
 
-The plot 7 from treatment 1 displayed a constantly outlying functional
-diversity and was removed from the graphical representation for better
-readability (see appendix for full graphs). In undisturbed plots the
-functional diversity remained stable along the 30 years. In disturbed
-plots, trajectories depend on the disturbnace intensity with, for low
-intensity, a low but long-lasting increase up to amximum reached after
-20-25 years and, for, high intensity, a fast but short increase
-followed, after 10 years, by a slow decrease towards the inital values.
+The plot 7 from treatment 1 displayed constantly outlying functional
+richness and evenness and was removed from the graphical representation
+for better readability. In undisturbed plots both functional richness
+and evenness remained stable along the 30 years. In disturbed plots,
+richness and evenness trajectories depended on the disturbance intensity
+with their maximum positively correlated to \%AGB loss
+\(\rho_{spearman}^{Richness}=0.76\) and
+\(\rho_{spearman}^{Evenness}=0.60\). Functional richness and evenness
+displayed for low disturbance intensity a low but long-lasting increase
+up to maximum reached after 20-25 years and, and for high intensity, a
+fast but short increase followed after 10 years by a slow decrease
+towards the inital values.
 
 The second-degree polynomial regressions between (i) the \%AGB loss and
-(ii) taxonomic richness, evenness and functional diversity after 10, 20
-and 30 years best predicted the hump-shaped curve of the disturbance
-impact along the disturbance intensity gradient \ref{fig:IDHplot}. Along
-the disturbance intensity gradient the disturbance impact was more
-markedly hump-shaped for the taxonomic richness than for the taxonomic
-evenness and the functional diversity, the latest being almost linear.
-The regression model better predicted the functional diversity
-(\(0.60<R^2_{Functional}<0.81\)) than the taxonomic richness and
-evenness (\(0.21<R^2_{Richness}<0.4\), and \(-0.15<R^2_{Evenness}<0.43\)
-respectively)
+(ii) taxonomic and functional richness and evenness after 10, 20 and 30
+years best predicted the hump-shaped curve of the disturbance impact
+along the disturbance intensity gradient \ref{fig:IDHplot}. Along the
+disturbance intensity gradient the disturbance impact was more markedly
+hump-shaped for the taxonomic richness than for the taxonomic evenness
+and the functional diversity, the latest being almost linear. The
+regression model better predicted the functional richness and evenness
+(\(0.55<R^2_{Functional Richness}<0.72\), and
+\(0.60<R^2_{Functional Evenness}<0.81\)) than the taxonomic richness and
+evenness (\(0.21<R^2_{Taxonomic Richness}<0.4\), and
+\(-0.15<R^2_{Taxonomic Evenness}<0.43\) respectively)
 
@@
 
-\caption{Relationship between the initial \%AGB loss and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)} and functional diversity \textbf{(c)} at 10, 20 and 30 years after disturbance.
+\caption{Relationship between the initial \%AGB loss and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)}, functional richness \textbf{(c)},and functional evenness \textbf{(d)} at 10, 20 and 30 years after disturbance.
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:IDHplot}
@@
 
-Disturbance clearly enhanced community taxonomic richness until an
+Disturbance appeared to enhance community taxonomic richness until an
 intensity threshold (20-25\% of initial AGB loss), while the taxonomic
-evenness was somewhat decoupled from the disturbance intensity, as
+evenness was somewhat decoupled from the disturbance intensity as
 already observed in the Guiana Shield \citep{Baraloto2012a} and in
 Bornean tropical forests \citep{Cannon1998}. The IDH well predicted the
-response of taxonomic richness to disturbance, and to a lesser extent
-the taxonomic evenness. Post-disturbance communities combine
-pre-disturbance surviving trees and recruited trees, either from the
-remaining pre-disturbance community or from a disturbance-specific
-community. In the Guiana Shield, the pool of true pioneer species
-specifically recruited after disturbance is restricted to a few common
-genera (e.g.~Cecropia spp., Vismia spp.) Below the intensity threshold
-the taxonomic richness of the pre-disturbance community is maintained
-and the recruitment of pioneers, previously infrequent or absent,
-increases both taxonomic richness and evenness. The disturbance
-intensity determined the relative importance of pioneers recruited, so
-the taxonomic richness and evennes increase all the more so that
-disturbance is intense. However, disturbance in the same time decreases
-the taxonomic richness of the pre-disturbance community, and beyond the
-intensity threshold the decrease exceeds the taxonomic enrichment from
-pioneers. The taxonomic evenness for its part is less sensitive to the
-loss of rare species and the disturbance impact is mitigated as pioneers
-balance the hyper-dominance of undisturbed tropical forests. Conversely
-the IDH was disproved for the functional diversity. The pioneer species
-were functionally highly different from the pre-disturbance community
-composition, so their recruitment increased the community functional
-richness and eveness after disturbance whenever its intensity.
+disturbance impact on community taxonomic richness and to some extent on
+taxonomic evenness. The disturbance intensity determines the balance in
+the community between pre-disturbance surviving trees and the trees
+recruited afterward. In the Guiana Shield, the pool of true pioneer
+species specifically recruited after disturbance is restricted to a few
+common genera (e.g.~Cecropia spp., Vismia spp.). Below the intensity
+threshold the size of the surviving community maintains its
+pre-disturbance taxonomic richness and the recruitment of pioneers,
+infrequent or absent before disturbance, increases both community
+taxonomic richness and evenness. Beyond the intensity threshold the
+taxonomic diversity of the surviving trees decreases and is not offset
+by the taxonomic enrichment of pioneers. The disturbance impact was
+similar but slighter for community taxonomic evenness, that is less
+sensitive to the loss of rare species and rather represents the
+increased dominance of pionneers that balanced the usual hyper-dominance
+of tropical forest. While below the intensity threshold pioneers species
+increase the diversity of dominant or common species, beyond the
+threshold they become in turn highly dominant.
 
-Community post-disturbance trajectories showed a first drop of the
-taxonomic richness, similar among communities, followed by a species
-gain determined by the disturbance intensity. Until an intensity
-threshold the gain in taxonomic richness increased with the disturbance
-intensity. This corresponded to the establishment of long-lived pioneers
+Conversely the IDH was disproved to predict the dsiturbance impact on
+community functional diversity. Pioneer species were functionally highly
+different from the pre-disturbance community composition, so their
+recruitment increased the community functional richness and eveness
+whenever its intensity.
+
+Along time, the post-disturbance taxonomic richness trajectories showed
+first a similar drop following the species loss during disturbance, and
+then a species gain determined by the disturbance intensity. Until an
+intensity threshold the gain in taxonomic richness increased with the
+disturbance intensity through the establishment of long-lived pioneers
 enhancing community taxonomic richness and evenness in the long term.
-Beyond the intensity threshold a few short-lived pioneers occupied the
-vacated environmental space, preventing the establishment of further
-species. The establishment of short-lived pioneers, functionally very
-different from the pre-disturbance community, rapidly and significantly
-increased the functional diversity but it started decreasing after 10
-years. Below the disturbance intensity threshold on the other hand the
-establishment of long-lived pionneers entailed a later and lower but
-long-lasting increase of the functional diversity. Likely the
-establishment of long-lasting pioneers will follow the decline of
-short-live pionneers and we expect the taxonomic and functoinal
-trajectories to catch up with those observed after intermediate
-disturbance.
+These long-lived pionneers were functionally quite different from the
+functional composition and their establishment entailed a progressive
+and long-lasting increase of the functional diversity. Beyond the
+intensity threshold a few short-lived pioneers occupied the vacated
+environmental space and prevented further species establishment. These
+short-lived pioneers were functionally very different from the
+pre-disturbance community and rapidly and significantly increased the
+functional diversity. Already after 10 years, though, they started to
+decline and the functional diversity started to decrease. Likely the
+establishment of long-lasting pioneers will follow and we expect the
+taxonomic and functional trajectories to catch up with those observed
+after intermediate disturbance.
 
-\subsection{The functional redundancy, key of the taxonomic
-resilience}\label{the-functional-redundancy-key-of-the-taxonomic-resilience}
+\subsection{The functional redundancy, key of the community
+resilience}\label{the-functional-redundancy-key-of-the-community-resilience}
 
-The functional redundancy within the pre-disturbance functional space
-decreased for 15 years because of the loss of species, and was not
-compensated by the functionally different pionneers. The depth of the
-redundancy decrease relied on the disturbance intensity that enhanced
-the initial decrease and the functional differences of pionneers from
-the pre-disturance composition. After 15 years the first established
-pioneers pionneers started to be replaced by more competitive long-lived
-pionneers or late-successional species resembling more the
-pre-disturbance functional composition and hence increasing the
-functional redundancy. This replacement though is stochastic and follows
-the lottery recruitment so because of increasing competition it is
-increasingly difficult recover pre-disturbance species,specifically the
-most infrequent ones, as the functional redundancy recovers
-\citep{Busing2002}. The recovery time proved long as for some community
-it was just initiated after 30 years but it is thus difficult to
-estimate \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+For 15 years the species loss during disturbance, determined by the
+disturbance intensity, commensurately decreased the functional
+redundancy within the pre-disturbance functional space. At first place
+the pioneers recruited, functionally different from the pre-disturbance
+functional composition did not compensate for the reundancy decrease.
+Progressively though, first establish species were replaced by more
+competitive long-lived pionneers or late-successional species resembling
+more the pre-disturbance functional composition and restoring the
+functional redundancy. This replacement was stochastic and followed the
+lottery recruitment rules, that ease the recovery for the first
+recruited species but then increasingly hamper it with the emergence of
+interspecific competition due to increasing functional redundancy
+\citep{Busing2002}. Along time the recovery of infrequent species is
+increasingly difficult, and the recovery of the functional redundancy
+that was just initiated after 30 years in some communities is difficult
+to estimate \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
@@
 meant a lower resilience of pre-disturbance communities and higher
-chances to see the persistence of disturbance-specific communities, with
-lower taxonomic richness and more pioneers
-\citep{Haddad2008, Burslem2000, Martin2013}. Besides the slowed recovery
-of rare species increased the risks to loose cornerstone specie, with
-unexpected ecological consequences
+chances to see the persistence of disturbance-specific species at the
+expense of late-successional ones
+\citep{Haddad2008, Burslem2000, Martin2013}. Besides the long-term
+recovery of infrequent species increases the risks to loose cornerstone
+species, with unexpected ecological consequences
 \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}. Apart from the
-functional characteristics considered here, infrequent species mighthave
-unique functions i nthe ecosystem or be a key for some fauna.
+functional characteristics considered here, infrequent species might
+indeed have unique functions in the ecosystem or be a key for some
+fauna.
 
@@
 
-Our study revealed communities cyclic recovery after disturbance
+{[}Our study revealed communities cyclic recovery after disturbance
 allowing the resilience of their functioning and taxonomic composition
@@
 richness and evenness that were blurried by the emergence of competitive
-exclusion along time. The resilience of tropical forests proved
-consistent although spread over several decades. Still, the disturbance
-impact on communities redundancy cautioned against the risks of
-infrequent species loss and the persistence of disturbance-specific
-communities \citep{Gourlet-Fleury2005}. As the trajectories highlighted
-the recruitment processes proved central for communities response to
+exclusion along time.{]} The resilience of tropical forests proved
+consistent although several decades-long. Still, the disturbance impact
+on communities redundancy cautioned against the risks of infrequent
+species loss and the persistence of disturbance-specific communities
+\citep{Gourlet-Fleury2005}. As the trajectories highlighted the
+recruitment processes proved central for communities response to
 disturbance and closer focus on demographical drivers of communities

### 2018-07-09 Scripts/DrawingTools.R

@@ -103,7 +103,7 @@ TaxoTraj<-function(CompTaxo){
          xlab="",ylab="",ylim=c(min(unlist(Toplot),na.rm=T),max(unlist(Toplot),na.rm=T)))
     axis(1,at=as.character(seq(5,33,5)),labels=TRUE)  
     mtext(paste("(",c("a","b","b")[q],") ",c("Taxonomic Richness","Shannon","Taxonomic Evenness")[q],sep=""),
-          line=1,side=3,adj=0)
+          line=1,side=3,adj=0,cex=0.8)
     
     invisible(lapply(1:4,function(t){  
       toplot<-Toplot[[t]]
@@ -146,7 +146,6 @@ plotDiv<-function(Data,remove=FALSE){
   if(remove){Data[[2]]<-Data[[2]][which(rownames(Data[[2]])!=7),,]}
   
   Toplot<-lapply(Data,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
-  absc<-as.numeric(colnames(Toplot[[1]]))-1986
   Toplot<-lapply(Toplot,function(tr){
     ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col}))})#-tr[,1,rep]
     ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),
@@ -157,7 +156,7 @@ plotDiv<-function(Data,remove=FALSE){
   
   plot(colnames(Toplot[[1]]),Toplot[[1]][1,,1],type="n",xaxt="n",
        xlab="",ylab="",ylim=c(min(unlist(Toplot),na.rm=T),max(unlist(Toplot),na.rm=T)))
-  axis(1,at=absc,labels=TRUE)  
+  axis(1,at=as.character(seq(5,33,5)),labels=TRUE)  
   
   invisible(lapply(1:4,function(t){  
     toplot<-Toplot[[t]]

### 2018-07-09 Scripts/Stats.R

@@ -1,5 +1,6 @@
 load("DB/ReplacementTraj_ForGraphs")
 load("DB/FunctionalTraj_ForGraphs")
+load("DB/FunctionalRichnessTraj_ForGraphs")
 
 SimpsonTaxo<-lapply(CompleteTaxo,function(tr){return(tr[,,,"Simpson"])})
 save(SimpsonTaxo,file="DB/SimpsonDraw")
@@ -21,14 +22,14 @@ Rich<-lapply(Rich, function(tr){
              dimnames=list(rownames(Ret[[1]]),colnames(Ret[[1]]),c(0.025,0.5,0.975)))})
 save(Rich,file="DB/RichnessIDH")
 
-CompFun<-CompleteFun
+CompFun<-CompleteRichFun
 CompFun[[2]]<-CompFun[[2]][which(rownames(CompFun[[2]])!=7),,]
-Rao<-lapply(CompFun, function(tr){
+FunRich<-lapply(CompFun, function(tr){
   Ret<-lapply(c(0.025,0.5,0.975),function(quant){
     return(apply(tr[,c("1995","2005","2015"),],c(1,2),function(x){return(quantile(x,probs=quant))}))})
   Ret<-array(unlist(Ret),dim=c(nrow(Ret[[1]]),ncol(Ret[[1]]),3),
              dimnames=list(rownames(Ret[[1]]),colnames(Ret[[1]]),c(0.025,0.5,0.975)))})
-save(Rao,file="DB/RaoIDH")
+save(FunRich,file="DB/FunRichIDH")
 
 windows()
 par(mfrow=c(2,2),mar=c(3,3,2,1),oma=c(2,1,2,3),no.readonly = T)
@@ -40,7 +41,7 @@ mtext("Years since disturbance",side=1,line=2.2,adj=1)
 legend("right",inset=c(-0.28,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
 
 plotIDH(Sim)
-plotIDH(Rao)
+plotIDH(FunRich)
 mtext("initial %AGB lost",side=1,line=2.2,adj=1)
 legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
 
[…]

### 2018-07-09 WholePlotTrajectories.Rmd
dicussion

@@
 Validations of the IDH though remain scarce in the long-term and mainly rely on the analysis of taxonomic richness [@Molino2001].
-Taxonomic richness, alone, gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
+Taxonomic richness alone, though, gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
 More ecological-meaningful analysis would couple richness with (i) evenness that would reveal underlying ecological processes and (ii) composition that is crucial for conservation issues [@Magurran1988;@Lavorel2002;@Bellwood2006].
@@
 These indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995;@Marcon2015].
-The functional diversity was reported using the functional richness and evenness, *i.e* Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on species functional traits.
+The functional diversity was reported using the functional richness and functional evenness, *i.e* Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on species functional traits.
 
 The impacts of initial disturbance were tested with the spearman rank correlation between the extrema of taxonomic and functional metrics reached over the 30 years and the initial %AGB loss. 
-They were besides analysed through polynomial regression between (i) taxonomic and functional richness, evenness and diversity and (ii) the initial %AGB loss at 10, 20 and 30 years after disturbance.
+They were besides analysed through polynomial regression between (i) taxonomic and functional richness and evenness and (ii) the initial %AGB loss at 10, 20 and 30 years after disturbance.
 
@@
 
-```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic (ref:Bolda) richness, (ref:Boldb), taxonomic evenness, (ref:Boldc) functional richness, and (ref:Boldd) functional evenness.
+```{r DivTaxo, out.width = '100%', echo=FALSE, fig.cap="Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic (ref:Bolda) richness, (ref:Boldb), taxonomic evenness, (ref:Boldc) functional richness, and (ref:Boldd) functional evenness.
 Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
@@
 
-par(mfrow=c(1,4),mar=c(5,2,4,2),oma=c(1,1.7,1,2),no.readonly=TRUE)
+par(mfrow=c(2,2),mar=c(2,2.5,2.5,1),oma=c(1,2,1,3),no.readonly = T)
 TaxoTraj(CompleteTaxo)
@@
 mtext("Years since disturbance",side=1,line=2.2,adj=1)
-legend("right",inset=c(-0.34,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
+legend("right",inset=c(-0.3,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
 
@@
 In undisturbed plots both functional richness and evenness remained stable along the 30 years. 
-In disturbed plots, richness and evenness trajectories depended on the disturbance intensity with their maximum positively correlated to %AGB loss $\rho_{spearman}^{Richness}=0.76$ and $\rho_{spearman}^{Evenness}=0.60$.
-Functional richness and evenness displayed for low disturbance intensity a low but long-lasting increase up to maximum reached after 20-25 years and, and for high intensity, a fast but short increase followed after 10 years by a slow decrease towards the inital values.
+In disturbed plots, functional richness and evenness trajectories depended on the disturbance intensity with their maximum positively correlated to %AGB loss $\rho_{spearman}^{Richness}=0.76$ and $\rho_{spearman}^{Evenness}=0.60$.
+Functional richness and evenness displayed for low disturbance intensity a low but long-lasting increase up to a maximum reached after 20-25 years, and for high intensity, a fast but short increase followed after 10 years by a slow decrease towards the inital values.
 
 The second-degree polynomial regressions between (i) the %AGB loss and (ii) taxonomic and functional richness and evenness after 10, 20 and 30 years best predicted the hump-shaped curve of the disturbance impact along the disturbance intensity gradient \@ref(fig:IDHplot).
-Along the disturbance intensity gradient the disturbance impact was more markedly hump-shaped for the taxonomic richness than for the taxonomic evenness and the functional diversity, the latest being almost linear.
+The relationship between the disturbance impact and its intensity was more markedly hump-shaped for the taxonomic richness than for the taxonomic evenness.
+For both functional richness and evenness the relationship was almost linear.
 The regression model better predicted the functional richness and evenness ($0.55<R^2_{Functional Richness}<0.72$, and $0.60<R^2_{Functional Evenness}<0.81$) than the taxonomic richness and evenness ($0.21<R^2_{Taxonomic Richness}<0.4$, and $-0.15<R^2_{Taxonomic Evenness}<0.43$ respectively)
@@
 
-par(mfrow=c(1,4),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
+par(mfrow=c(1,4),mar=c(5,2,4,2),oma=c(1,1.5,1,2),no.readonly=TRUE)
 plotIDH(Data=Rich,AgbLoss=AGBloss)
@@
 mtext("initial AGB loss (%)",side=1,line=2.2,adj=1)
-legend("right",inset=c(-0.34,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
+legend("right",inset=c(-0.37,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
 
@@
 While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-disturbance time.
-in disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WSG to high average SLA and chlorophyll content (see appendix I).
+In disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WSG to high average SLA and chlorophyll content (see appendix I).
 For functional composition, this translated into cyclic compositional changes with an unachieved recovery of the initial composition (Figure \@ref(fig:NMDSplans)).
@@
 
-par(mfrow=c(2,2))
+par(mfrow=c(2,2),mar=c(2,2.5,2.5,1),oma=c(1,2,1,3),no.readonly = T)
 TaxoCompo(MatrepTaxo)
+mtext("Taxonomic composition",side=3,line=1.8,cex=1.05)
 FunCompo(MatrepFun)
+mtext("Functional composition",side=3,line=1.8,cex=1.05)
 EuclidDist(TaxoEuclid)
+mtext("Euclidean distance\nfrom 1985 inventory",side=2,padj=0,line=2,cex=0.8)
+mtext("(c)",side=3,adj=0,line=0.5)
 EuclidDist(FunEuclid)
+mtext("(d)",side=3,adj=0,line=0.5)
+mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=0.8)
+legend("right",inset=c(-0.3,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
 
@@
 Communities taxonomic and functional composition appeared resilient, following similar hump-shaped trajectories starting to return towards pre-disturbance composition after 30 years.
+
 The taxonomic differences among communities, marked before disturbance by the distinct starting points on the NMDS axis 2, were maintained throughout recovery trajectories. 
@@
 
-Trajectories of community functional composition appeared similar in time and close in terms of functional composition.
-Because pre-disturbance surviving trees mirror the initial community [@Herault2018], changes in functional composition relied upon the recruitment of species or functional types infrequent or absent before disturbance.
-Competitive pioneers became dominant in filling the environmental niches of high availability of light, space and nutrient made available by the disturbance.
-It changed community functional composition in the same way for all communities, towards more resource-acquisitive strategies [@Grime1998;@Westoby1998; @Wright2004; @Reich2014].
-Later during the community recovery long-lived, more resistant and shade-tolerant species, would exclude those pioneers and restore the pre-disturbance composition.
+Conversely, disturbed communities followed functional trajecotories similar in terms of functional composition.
+As pre-disturbance surviving trees mirror the initial community [@Herault2018], changes in functional composition relied upon the recruitment of species or functional types that were infrequent or absent before disturbance.
+Competitive pioneers became dominant in filling the environmental niches of high availability of light, space and nutrient vacated by the disturbance.
+The recruitment of pioneers similarly changed community functional composition towards more resource-acquisitive strategies [@Grime1998;@Westoby1998; @Wright2004; @Reich2014].
+Thereafter, again in the same way for all disturbed communities, long-lived, more resistant and shade-tolerant species excluded the first established pioneers and started the recovery of pre-disturbance functional composition.
 
@@
 
-Disturbance appeared to enhance community taxonomic richness until an intensity threshold (20-25% of initial AGB loss), while the taxonomic evenness was somewhat decoupled from the disturbance intensity as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
-The IDH well predicted the disturbance impact on community taxonomic richness and to some extent on taxonomic evenness.
-The disturbance intensity determines the balance in the community between pre-disturbance surviving trees and the trees recruited afterward.
-In the Guiana Shield, the pool of true pioneer species specifically recruited after disturbance is restricted to a few common genera (e.g. Cecropia spp., Vismia spp.).
-Below the intensity threshold the size of the surviving community maintains its pre-disturbance taxonomic richness and the recruitment of pioneers, infrequent or absent before disturbance, increases both community taxonomic richness and evenness.
-Beyond the intensity threshold the taxonomic diversity of the surviving trees decreases and is not offset by the taxonomic enrichment of pioneers.
-The disturbance impact was similar but slighter for community taxonomic evenness, that is less sensitive to the loss of rare species and rather represents the increased dominance of pionneers that balanced the usual hyper-dominance of tropical forest.
-While below the intensity threshold pioneers species increase the diversity of dominant or common species, beyond the threshold they become in turn highly dominant.
+The IDH predicted well the disturbance impact on community taxonomic richness, enhanced until an intensity threshold (20-25% AGB loss), and to some extent on taxonomic evenness, somewhat decoupled from the disturbance intensity as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+The disturbance intensity determined the balance in the community between pre-disturbance surviving trees and those recruited afterward.
+The pool of true pioneer species specifically recruited after disturbance is restricted in the Guiana Shield to a few common genera (e.g. Cecropia spp., Vismia spp.).
+Below the intensity threshold the size of the surviving community maintained the pre-disturbance taxonomic richness while the recruitment of pioneers, infrequent or absent before disturbance, increased both community taxonomic richness and evenness.
+Beyond the intensity threshold, the disturbance decreased the taxonomic richness of surviving trees which was not offset by the enrichment of pioneers, so that the overall community taxonomic richness decreased according to the disturbance intensity.
+For community taxonomic evenness the disturbance impact was similar but slighter, as the evenness is less sensitive to the loss of rare species.
+Taxonomic evenness rather represented the increasing dominance of pionneers that balanced the usual hyper-dominance of tropical forest below the intensity threshold, thus increasing community overall evenness up to the intensity threshold beyond which pioneers became in turn highly dominant and decreased the overall evenness.
 
-Conversely the IDH was disproved to predict the dsiturbance impact on community functional diversity.
-Pioneer species were functionally highly different from the pre-disturbance community composition, so their recruitment increased the community functional richness and eveness whenever its intensity.
+Conversely the IDH was disproved regarding the disturbance impact on community functional diversity.
+Irrespective of the disturbance intensity the recruitment of pionneers, functionally highly different from the composition of pre-disturbance community, increased the community functional richness and eveness.
 
-Along time, the post-disturbance taxonomic richness trajectories showed first a similar drop following the species loss during disturbance, and then a species gain determined by the disturbance intensity.
-Until an intensity threshold the gain in taxonomic richness increased with the disturbance intensity through the establishment of long-lived pioneers enhancing community taxonomic richness and evenness in the long term.
-These long-lived pionneers were functionally quite different from the functional composition and their establishment entailed a progressive and long-lasting increase of the functional diversity.
-Beyond the intensity threshold a few short-lived pioneers occupied the vacated environmental space and prevented further species establishment. 
-These short-lived pioneers were functionally very different from the pre-disturbance community and rapidly and significantly increased the functional diversity.
-Already after 10 years, though, they started to decline and the functional diversity started to decrease.
-Likely the establishment of long-lasting pioneers will follow and we expect the taxonomic and functional trajectories to catch up with those observed after intermediate disturbance.
+Along time, taxonomic richness trajectories for all disturbed communities first dropped similarly following the loss of species during disturbance.
+Then, trajectories showed a species gain depending on the disturbance intensity.
+The species gain was all the more significant as the disturbance intensity increased, up to an intensity threshold, through the establishment of long-lived pioneers enhancing community taxonomic richness and evenness in the long term.
+These long-lived pionneers, functionally quite different from the functional composition, entailed in parallel a progressive and long-lasting increase of the functional diversity.
+Beyond an intensity threshold, though, a few short-lived pioneers occupied the vacated environmental space and prevented the establishment of other species. 
+These short-lived pioneers were functionally very different from the pre-disturbance community and entailed a rapid and significant increase of functional richness en evenness.
+Already after 10 years, though, short-lived pioneers started to decline and the functional richness and evenness decreased.
+Likely the establishment of long-lasting pioneers will follow this decrease and we expect the taxonomic and functional trajectories to catch up with those observed after intermediate disturbance.
 
-## The functional redundancy, key of the community resilience
+## The functional redundancy, key of community resilience
 
 For 15 years the species loss during disturbance, determined by the disturbance intensity, commensurately decreased the functional redundancy within the pre-disturbance functional space.
-At first place the pioneers recruited, functionally different from the pre-disturbance functional composition did not compensate for the reundancy decrease.
-Progressively though, first establish species were replaced by more competitive long-lived pionneers or late-successional species resembling more the pre-disturbance functional composition and restoring the functional redundancy.
-This replacement was stochastic and followed the lottery recruitment rules, that ease the recovery for the first recruited species but then increasingly hamper it with the emergence of interspecific competition due to increasing functional redundancy [@Busing2002].
-Along time the recovery of infrequent species is increasingly difficult, and the recovery of the functional redundancy that was just initiated after 30 years in some communities is difficult to estimate [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+The redundancy decrease was not compensated in the first place as the pionneers first recruited were functionally different from the pre-disturbance functional composition.
+Progressively though, first established species were replaced by more competitive long-lived pionneers or late-successional species resembling more the pre-disturbance functional composition and restoring the functional redundancy.
+This replacement was stochastic and followed the lottery recruitment rules, implying a recruitment eased for the first recruited species but then increasingly hampered by the emergence of interspecific competition [@Busing2002].
+Along time the recovery of infrequent species was increasingly difficult, so that the time for the full recovery of the functional redundancy, in some communities just initiated after 30 years, was difficult to estimate [@Trenbath1999;@Elmqvist2003;@Diaz2005].
 
-The long-term impact of disturbance on community functional redundancy meant a lower resilience of pre-disturbance communities and higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008; @Burslem2000; @Martin2013].
+The long-term impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008; @Burslem2000; @Martin2013].
 Besides the long-term recovery of infrequent species increases the risks to loose cornerstone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
@@
 
-[Our study revealed communities cyclic recovery after disturbance allowing the resilience of their functioning and taxonomic composition with the maintenance of initial differences among communities.
-Communities functional evenness was enhanced for  20 years after disturbance through the enrichment of the communities with pioneers and light-demanding species,in accordance with the IDH.
-The IDH, though, poorly predicted the disturbance impact on communities taxonomic richness and evenness that were blurried by the emergence of competitive exclusion along time.]
+Our study revealed the post-disturbance recovery of both taxonomic composition, maintaining the initial differences among communities, and functional composition, that was in contrast similar among communities.
+The IDH was validated for community taxonomic richness and, to some extent, taxonomic evenness but disproved regarding community functional richness and evennes that were enhanced for any disturbance intensity by the high functional differences of pioneers compared to late-successional functoinal composition.
+The IDH was translated in time by the recruitment, beyond an intensity threshold, of short-lived pionneers that prevented in the first times the establishment of more diverse long-lived pionneers, recruited otherwise below the intensity threshold.
 The resilience of tropical forests proved consistent although several decades-long. 

### 2018-07-09 WholePlotTrajectories.tex
dicussion

@@
 rely on the analysis of taxonomic richness \citep{Molino2001}. Taxonomic
-richness, alone, gives limited or misleading information on forests
-recovery and functioning \citep{Martin2015, Chaudhary2016}. More
+richness alone, though, gives limited or misleading information on
+forests recovery and functioning \citep{Martin2015, Chaudhary2016}. More
 ecological-meaningful analysis would couple richness with (i) evenness
@@
 \citep{Patil1982, Tothmeresz1995, Marcon2015}. The functional diversity
-was reported using the functional richness and evenness, \emph{i.e} Rao
-index of quadratic entropy which combines species abundance distribution
-and average pairwise dissimilarity based on species functional traits.
+was reported using the functional richness and functional evenness,
+\emph{i.e} Rao index of quadratic entropy which combines species
+abundance distribution and average pairwise dissimilarity based on
+species functional traits.
 
@@
 analysed through polynomial regression between (i) taxonomic and
-functional richness, evenness and diversity and (ii) the initial \%AGB
-loss at 10, 20 and 30 years after disturbance.
+functional richness and evenness and (ii) the initial \%AGB loss at 10,
+20 and 30 years after disturbance.
 
@@
 and evenness remained stable along the 30 years. In disturbed plots,
-richness and evenness trajectories depended on the disturbance intensity
-with their maximum positively correlated to \%AGB loss
-\(\rho_{spearman}^{Richness}=0.76\) and
+functional richness and evenness trajectories depended on the
+disturbance intensity with their maximum positively correlated to \%AGB
+loss \(\rho_{spearman}^{Richness}=0.76\) and
 \(\rho_{spearman}^{Evenness}=0.60\). Functional richness and evenness
 displayed for low disturbance intensity a low but long-lasting increase
-up to maximum reached after 20-25 years and, and for high intensity, a
+up to a maximum reached after 20-25 years, and for high intensity, a
 fast but short increase followed after 10 years by a slow decrease
@@
 years best predicted the hump-shaped curve of the disturbance impact
-along the disturbance intensity gradient \ref{fig:IDHplot}. Along the
-disturbance intensity gradient the disturbance impact was more markedly
-hump-shaped for the taxonomic richness than for the taxonomic evenness
-and the functional diversity, the latest being almost linear. The
-regression model better predicted the functional richness and evenness
-(\(0.55<R^2_{Functional Richness}<0.72\), and
+along the disturbance intensity gradient \ref{fig:IDHplot}. The
+relationship between the disturbance impact and its intensity was more
+markedly hump-shaped for the taxonomic richness than for the taxonomic
+evenness. For both functional richness and evenness the relationship was
+almost linear. The regression model better predicted the functional
+richness and evenness (\(0.55<R^2_{Functional Richness}<0.72\), and
 \(0.60<R^2_{Functional Evenness}<0.81\)) than the taxonomic richness and
@@
 undisturbed communities (Figure \ref{fig:NMDSplans}), they followed
-marked and consistent trajectories over post-disturbance time. in
+marked and consistent trajectories over post-disturbance time. In
 disturbed communities, these compositional changes corresponded to
@@
 following similar hump-shaped trajectories starting to return towards
-pre-disturbance composition after 30 years. The taxonomic differences
-among communities, marked before disturbance by the distinct starting
-points on the NMDS axis 2, were maintained throughout recovery
-trajectories. More than commonly thought, post-disturbance trajectories
-depended on community initial composition, that partly determined the
-pool of recruited species and constrained the trajectories towards the
-initial composition. The high resilience of communities taxonomy
-revealed that species not belonging to the pre-disturbance community
-were hardly recruited because of the commonness of dispersal limitation
-among tropical tree species \citep{Svenning2005}.
+pre-disturbance composition after 30 years.
 
-Trajectories of community functional composition appeared similar in
-time and close in terms of functional composition. Because
-pre-disturbance surviving trees mirror the initial community
-\citep{Herault2018}, changes in functional composition relied upon the
-recruitment of species or functional types infrequent or absent before
-disturbance. Competitive pioneers became dominant in filling the
-environmental niches of high availability of light, space and nutrient
-made available by the disturbance. It changed community functional
-composition in the same way for all communities, towards more
-resource-acquisitive strategies
-\citep{Grime1998, Westoby1998, Wright2004, Reich2014}. Later during the
-community recovery long-lived, more resistant and shade-tolerant
-species, would exclude those pioneers and restore the pre-disturbance
+The taxonomic differences among communities, marked before disturbance
+by the distinct starting points on the NMDS axis 2, were maintained
+throughout recovery trajectories. More than commonly thought,
+post-disturbance trajectories depended on community initial composition,
+that partly determined the pool of recruited species and constrained the
+trajectories towards the initial composition. The high resilience of
+communities taxonomy revealed that species not belonging to the
+pre-disturbance community were hardly recruited because of the
+commonness of dispersal limitation among tropical tree species
+\citep{Svenning2005}.
+
+Conversely, disturbed communities followed functional trajecotories
+similar in terms of functional composition. As pre-disturbance surviving
+trees mirror the initial community \citep{Herault2018}, changes in
+functional composition relied upon the recruitment of species or
+functional types that were infrequent or absent before disturbance.
+Competitive pioneers became dominant in filling the environmental niches
+of high availability of light, space and nutrient vacated by the
+disturbance. The recruitment of pioneers similarly changed community
+functional composition towards more resource-acquisitive strategies
+\citep{Grime1998, Westoby1998, Wright2004, Reich2014}. Thereafter, again
+in the same way for all disturbed communities, long-lived, more
+resistant and shade-tolerant species excluded the first established
+pioneers and started the recovery of pre-disturbance functional
 composition.
@@
 
-Disturbance appeared to enhance community taxonomic richness until an
-intensity threshold (20-25\% of initial AGB loss), while the taxonomic
-evenness was somewhat decoupled from the disturbance intensity as
-already observed in the Guiana Shield \citep{Baraloto2012a} and in
-Bornean tropical forests \citep{Cannon1998}. The IDH well predicted the
-disturbance impact on community taxonomic richness and to some extent on
-taxonomic evenness. The disturbance intensity determines the balance in
-the community between pre-disturbance surviving trees and the trees
-recruited afterward. In the Guiana Shield, the pool of true pioneer
-species specifically recruited after disturbance is restricted to a few
+The IDH predicted well the disturbance impact on community taxonomic
+richness, enhanced until an intensity threshold (20-25\% AGB loss), and
+to some extent on taxonomic evenness, somewhat decoupled from the
+disturbance intensity as already observed in the Guiana Shield
+\citep{Baraloto2012a} and in Bornean tropical forests
+\citep{Cannon1998}. The disturbance intensity determined the balance in
+the community between pre-disturbance surviving trees and those
+recruited afterward. The pool of true pioneer species specifically
+recruited after disturbance is restricted in the Guiana Shield to a few
 common genera (e.g.~Cecropia spp., Vismia spp.). Below the intensity
-threshold the size of the surviving community maintains its
-pre-disturbance taxonomic richness and the recruitment of pioneers,
-infrequent or absent before disturbance, increases both community
-taxonomic richness and evenness. Beyond the intensity threshold the
-taxonomic diversity of the surviving trees decreases and is not offset
-by the taxonomic enrichment of pioneers. The disturbance impact was
-similar but slighter for community taxonomic evenness, that is less
-sensitive to the loss of rare species and rather represents the
-increased dominance of pionneers that balanced the usual hyper-dominance
-of tropical forest. While below the intensity threshold pioneers species
-increase the diversity of dominant or common species, beyond the
-threshold they become in turn highly dominant.
+threshold the size of the surviving community maintained the
+pre-disturbance taxonomic richness while the recruitment of pioneers,
+infrequent or absent before disturbance, increased both community
+taxonomic richness and evenness. Beyond the intensity threshold, the
+disturbance decreased the taxonomic richness of surviving trees which
+was not offset by the enrichment of pioneers, so that the overall
+community taxonomic richness decreased according to the disturbance
+intensity. For community taxonomic evenness the disturbance impact was
+similar but slighter, as the evenness is less sensitive to the loss of
+rare species. Taxonomic evenness rather represented the increasing
+dominance of pionneers that balanced the usual hyper-dominance of
+tropical forest below the intensity threshold, thus increasing community
+overall evenness up to the intensity threshold beyond which pioneers
+became in turn highly dominant and decreased the overall evenness.
 
-Conversely the IDH was disproved to predict the dsiturbance impact on
-community functional diversity. Pioneer species were functionally highly
-different from the pre-disturbance community composition, so their
-recruitment increased the community functional richness and eveness
-whenever its intensity.
+Conversely the IDH was disproved regarding the disturbance impact on
+community functional diversity. Irrespective of the disturbance
+intensity the recruitment of pionneers, functionally highly different
+from the composition of pre-disturbance community, increased the
+community functional richness and eveness.
 
-Along time, the post-disturbance taxonomic richness trajectories showed
-first a similar drop following the species loss during disturbance, and
-then a species gain determined by the disturbance intensity. Until an
-intensity threshold the gain in taxonomic richness increased with the
-disturbance intensity through the establishment of long-lived pioneers
-enhancing community taxonomic richness and evenness in the long term.
-These long-lived pionneers were functionally quite different from the
-functional composition and their establishment entailed a progressive
-and long-lasting increase of the functional diversity. Beyond the
-intensity threshold a few short-lived pioneers occupied the vacated
-environmental space and prevented further species establishment. These
-short-lived pioneers were functionally very different from the
-pre-disturbance community and rapidly and significantly increased the
-functional diversity. Already after 10 years, though, they started to
-decline and the functional diversity started to decrease. Likely the
-establishment of long-lasting pioneers will follow and we expect the
-taxonomic and functional trajectories to catch up with those observed
-after intermediate disturbance.
+Along time, taxonomic richness trajectories for all disturbed
+communities first dropped similarly following the loss of species during
+disturbance. Then, trajectories showed a species gain depending on the
+disturbance intensity. The species gain was all the more significant as
+the disturbance intensity increased, up to an intensity threshold,
+through the establishment of long-lived pioneers enhancing community
+taxonomic richness and evenness in the long term. These long-lived
+pionneers, functionally quite different from the functional composition,
+entailed in parallel a progressive and long-lasting increase of the
+functional diversity. Beyond an intensity threshold, though, a few
+short-lived pioneers occupied the vacated environmental space and
+prevented the establishment of other species. These short-lived pioneers
+were functionally very different from the pre-disturbance community and
+entailed a rapid and significant increase of functional richness en
+evenness. Already after 10 years, though, short-lived pioneers started
+to decline and the functional richness and evenness decreased. Likely
+the establishment of long-lasting pioneers will follow this decrease and
+we expect the taxonomic and functional trajectories to catch up with
+those observed after intermediate disturbance.
 
-\subsection{The functional redundancy, key of the community
-resilience}\label{the-functional-redundancy-key-of-the-community-resilience}
+\subsection{The functional redundancy, key of community
+resilience}\label{the-functional-redundancy-key-of-community-resilience}
 
@@
 disturbance intensity, commensurately decreased the functional
-redundancy within the pre-disturbance functional space. At first place
-the pioneers recruited, functionally different from the pre-disturbance
-functional composition did not compensate for the reundancy decrease.
-Progressively though, first establish species were replaced by more
-competitive long-lived pionneers or late-successional species resembling
-more the pre-disturbance functional composition and restoring the
-functional redundancy. This replacement was stochastic and followed the
-lottery recruitment rules, that ease the recovery for the first
-recruited species but then increasingly hamper it with the emergence of
-interspecific competition due to increasing functional redundancy
-\citep{Busing2002}. Along time the recovery of infrequent species is
-increasingly difficult, and the recovery of the functional redundancy
-that was just initiated after 30 years in some communities is difficult
-to estimate \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+redundancy within the pre-disturbance functional space. The redundancy
+decrease was not compensated in the first place as the pionneers first
+recruited were functionally different from the pre-disturbance
+functional composition. Progressively though, first established species
+were replaced by more competitive long-lived pionneers or
+late-successional species resembling more the pre-disturbance functional
+composition and restoring the functional redundancy. This replacement
+was stochastic and followed the lottery recruitment rules, implying a
+recruitment eased for the first recruited species but then increasingly
+hampered by the emergence of interspecific competition
+\citep{Busing2002}. Along time the recovery of infrequent species was
+increasingly difficult, so that the time for the full recovery of the
+functional redundancy, in some communities just initiated after 30
+years, was difficult to estimate
+\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
 
 The long-term impact of disturbance on community functional redundancy
-meant a lower resilience of pre-disturbance communities and higher
+meant a lower resilience of the pre-disturbance communities, with higher
 chances to see the persistence of disturbance-specific species at the
@@
 
-{[}Our study revealed communities cyclic recovery after disturbance
-allowing the resilience of their functioning and taxonomic composition
-with the maintenance of initial differences among communities.
-Communities functional evenness was enhanced for 20 years after
-disturbance through the enrichment of the communities with pioneers and
-light-demanding species,in accordance with the IDH. The IDH, though,
-poorly predicted the disturbance impact on communities taxonomic
-richness and evenness that were blurried by the emergence of competitive
-exclusion along time.{]} The resilience of tropical forests proved
-consistent although several decades-long. Still, the disturbance impact
-on communities redundancy cautioned against the risks of infrequent
-species loss and the persistence of disturbance-specific communities
+Our study revealed the post-disturbance recovery of both taxonomic
+composition, maintaining the initial differences among communities, and
+functional composition, that was in contrast similar among communities.
+The IDH was validated for community taxonomic richness and, to some
+extent, taxonomic evenness but disproved regarding community functional
+richness and evennes that were enhanced for any disturbance intensity by
+the high functional differences of pioneers compared to
+late-successional functoinal composition. The IDH was translated in time
+by the recruitment, beyond an intensity threshold, of short-lived
+pionneers that prevented in the first times the establishment of more
+diverse long-lived pionneers, recruited otherwise below the intensity
+threshold. The resilience of tropical forests proved consistent although
+several decades-long. Still, the disturbance impact on communities
+redundancy cautioned against the risks of infrequent species loss and
+the persistence of disturbance-specific communities
 \citep{Gourlet-Fleury2005}. As the trajectories highlighted the

### 2018-07-09 WholePlotTrajectories.Rmd
relecture

@@
     Understanding the ecological rules underlying the maintenance of tropical forests biodiversity, structure, functioning and dynamics is urgent to anticipate their fate in the global change context.
-    The huge diversity of tropical forests is often assumed to be regularly reshaped by natural disturbance yielding a diversity peak at intermediate intensity, but this intermediate disturbance hypothesis (IDH) remains debated, and this controversy also questions the extent of communities resilience regarding their functional and taxonomic facets.
+    The huge diversity of tropical forests is often assumed to be regularly reshaped by natural disturbance yielding a diversity peak at intermediate intensity.
+This intermediate disturbance hypothesis (IDH),though, remains debated and the controversy questions the extent of communities' resilience regarding their functional and taxonomic facets.
 
-    To disentangle the ecological processes driving community response to disturbance, we analysed the taxonomic and functional diversity trajectories following a disturbance gradient.
-    Specifically, we examined, over 30 years, the functional and taxonomic community trajectories with regards to diversity, composition and redundancy.
+    To disentangle the ecological processes driving community response to disturbance, we analysed the diversity trajectories over 30 years following a disturbance gradient.
+    Specifically, we examined community functional and taxonomic trajectories with regards to diversity, composition and redundancy.
 Functional trajectories were drawn based on 7 leaf, stem and life-history traits.
     We highlighted the cyclic recovery of community taxonomic and functional composition.
-The pre-disturbance taxonomic differences were maintained over time while the functional composition trajectories were quite similar.
-    The IDH did predict communities functional diversity response while taxonomic diversity remained poorly sensitive to disturbance intensity.
+While pre-disturbance taxonomic differences were maintained over time, functional composition trajectories were quite similar among communities.
+    The IDH did predict communities taxonomic diversity response while functional diversity was enhanced whenever the disturbance intensity.
     Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years.
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy [@Lohbeck2015;@Guariguata2001].
-Specifically, we (i) questioned the recovery of communities taxonomic and functional characteristics and identified the underlying assembly processes, (ii) clarified the validity of the IDH in the long term for tropical forest and elucidated its translation into different trajectories, and (iii) questioned community recovery time.
+Specifically, we (i) elucidated community taxonomic and functional recovery and the determinism of underlying processes, (ii) clarified the validity of the IDH in the long term for tropical forest and its translation into different trajectories in time, and (iii) questioned community recovery time.
 
@@
 The taxonomic diversity was reported through species richness and evenness, *i.e* the Hill number translation of the Simpson index [@Hill1973].
-These indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995;@Marcon2015].
+These indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995;@Marcon2015b].
 The functional diversity was reported using the functional richness and functional evenness, *i.e* Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on species functional traits.
@@
 
+## Communities Composition
+
+From 1989 (2 years after disturbance) to 2015 (28 years after disturbance), 828 388 individual trees and 591 botanical species pertaining to 223 genus and 64 families were recorded.
+
+While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-disturbance time.
+In disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WSG to high average SLA and chlorophyll content (see appendix I).
+For functional composition, this translated into cyclic compositional changes with an unachieved recovery of the initial composition (Figure \@ref(fig:NMDSplans)).
+The maximum dissimilarity with the initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.87$ and $\rho_{spearman}^{functional}=0.90$ respectively).
+The maximum value was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition.
+
+```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS space.
+Lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals", fig.env="figure*"}
+load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
+load("DB/FunDistance_ForGraphs");load("DB/TaxoDistance_ForGraphs")
+
+par(mfrow=c(2,2),mar=c(2,2.5,2.5,1),oma=c(1,2,1,3),no.readonly = T)
+TaxoCompo(MatrepTaxo)
+mtext("Taxonomic composition",side=3,line=1.8,cex=1.05)
+FunCompo(MatrepFun)
+mtext("Functional composition",side=3,line=1.8,cex=1.05)
+EuclidDist(TaxoEuclid)
+mtext("Euclidean distance\nfrom 1985 inventory",side=2,padj=0,line=2,cex=0.8)
+mtext("(c)",side=3,adj=0,line=0.5)
+EuclidDist(FunEuclid)
+mtext("(d)",side=3,adj=0,line=0.5)
+mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=0.8)
+legend("right",inset=c(-0.35,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
+
+```
+
+Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions followed unimodal trajectories either stabilizing or returning towards their initial values.
+
+Maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WSG*) first decreased and then slightly increased but remained significantly lower than their initial value (Figure \@ref(fig:CWM)).
+On the other side, Bark thickness (*Bark_thick*) and specific leaf area (*SLA*) increased and while *Bark_thick* remained substantially high after 30 years, *SLA* had almost recovered its initial value.
+For all traits, the maximum difference to initial value was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.76$, $\rho_{spearman}^{L_{chloro}}=0.60$, $\rho_{spearman}^{L_{toughness}}=-0.53$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WSG}=-0.75$, $\rho_{spearman}^{Bark-thickness}=0.71$, $\rho_{spearman}^{Hmax}=-0.40$).
+
+```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWSG), and bark thickness, (ref:capBarkThick)) and one life history trait (Specific maximum height at adult stage, (ref:capHmax)).
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure*"}
+load("DB/CWM")
+
+CWMdraw(CWM)
+legendCWM()
+
+```
+
 ## Communities richness and evenness
 
-From 1989 (2 years after disturbance) to 2015 (28 years after disturbance), 828 388 individual trees and 591 botanical species pertaining to 223 genus and 64 families were recorded. 
 For undisturbed plots, taxonomic Richness and Evenness remained stable over the 30 years monitored.
@@
 
-## Communities Composition
-
-While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-disturbance time.
-In disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WSG to high average SLA and chlorophyll content (see appendix I).
-For functional composition, this translated into cyclic compositional changes with an unachieved recovery of the initial composition (Figure \@ref(fig:NMDSplans)).
-The maximum dissimilarity with the initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.87$ and $\rho_{spearman}^{functional}=0.90$ respectively).
-The maximum value was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition.
-
-```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS space.
-Lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals", fig.env="figure*"}
-load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
-load("DB/FunDistance_ForGraphs");load("DB/TaxoDistance_ForGraphs")
-
-par(mfrow=c(2,2),mar=c(2,2.5,2.5,1),oma=c(1,2,1,3),no.readonly = T)
-TaxoCompo(MatrepTaxo)
-mtext("Taxonomic composition",side=3,line=1.8,cex=1.05)
-FunCompo(MatrepFun)
-mtext("Functional composition",side=3,line=1.8,cex=1.05)
-EuclidDist(TaxoEuclid)
-mtext("Euclidean distance\nfrom 1985 inventory",side=2,padj=0,line=2,cex=0.8)
-mtext("(c)",side=3,adj=0,line=0.5)
-EuclidDist(FunEuclid)
-mtext("(d)",side=3,adj=0,line=0.5)
-mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=0.8)
-legend("right",inset=c(-0.3,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
-
-```
-
-Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions followed unimodal trajectories either stabilizing or returning towards their initial values.
-
-Maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WSG*) first decreased and then slightly increased but remained significantly lower than their initial value (Figure \@ref(fig:CWM)).
-On the other side, Bark thickness (*Bark_thick*) and specific leaf area (*SLA*) increased and while *Bark_thick* remained substantially high after 30 years, *SLA* had almost recovered its initial value.
-For all traits, the maximum difference to initial value was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.76$, $\rho_{spearman}^{L_{chloro}}=0.60$, $\rho_{spearman}^{L_{toughness}}=-0.53$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WSG}=-0.75$, $\rho_{spearman}^{Bark-thickness}=0.71$, $\rho_{spearman}^{Hmax}=-0.40$).
-
-```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWSG), and bark thickness, (ref:capBarkThick)) and one life history trait (Specific maximum height at adult stage, (ref:capHmax)).
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure*"}
-load("DB/CWM")
-
-CWMdraw(CWM)
-legendCWM()
-
-```
-
 ## Functional redundancy
@@
 
-Conversely, disturbed communities followed functional trajecotories similar in terms of functional composition.
+Conversely, disturbed communities followed functional trajectories similar in terms of functional composition.
 As pre-disturbance surviving trees mirror the initial community [@Herault2018], changes in functional composition relied upon the recruitment of species or functional types that were infrequent or absent before disturbance.
 Competitive pioneers became dominant in filling the environmental niches of high availability of light, space and nutrient vacated by the disturbance.
-The recruitment of pioneers similarly changed community functional composition towards more resource-acquisitive strategies [@Grime1998;@Westoby1998; @Wright2004; @Reich2014].
-Thereafter, again in the same way for all disturbed communities, long-lived, more resistant and shade-tolerant species excluded the first established pioneers and started the recovery of pre-disturbance functional composition.
+The recruitment of pioneers changed community functional composition in the same way for all disturbance intensity towards more resource-acquisitive strategies, moving community functional composition right along the first axis in Figure \@ref(fig:NMDSplans) [@Grime1998;@Westoby1998; @Wright2004; @Reich2014].
+Thereafter long-lived, more resistant and shade-tolerant species excluded the first established pioneers and started the recovery of pre-disturbance functional composition, moving similarly community functional composition left along the first axis and upward along the second axis in Figure \@ref(fig:NMDSplans) .
+
+These trajectories provided empirical support to the hypothesis that community assembly is both deterministic and historically convergent at different levels of community organization.
+Deterministic, trait-based processes drove community convergence in functional composition, while at the same time dispersal limitation maintained their divergence in taxonomic composition [@Fukami2005].
 
@@
 The disturbance intensity determined the balance in the community between pre-disturbance surviving trees and those recruited afterward.
-The pool of true pioneer species specifically recruited after disturbance is restricted in the Guiana Shield to a few common genera (e.g. Cecropia spp., Vismia spp.).
-Below the intensity threshold the size of the surviving community maintained the pre-disturbance taxonomic richness while the recruitment of pioneers, infrequent or absent before disturbance, increased both community taxonomic richness and evenness.
-Beyond the intensity threshold, the disturbance decreased the taxonomic richness of surviving trees which was not offset by the enrichment of pioneers, so that the overall community taxonomic richness decreased according to the disturbance intensity.
+The pool of true pioneer species specifically recruited after disturbance is restricted in the Guiana Shield to a few common genera (e.g. Cecropia spp., Vismia spp.) [@Zalamea2012;@Guitet2018].
+Below the intensity threshold the size of the surviving community maintained the pre-disturbance high taxonomic richness while the recruitment of pioneers, infrequent or absent before disturbance, increased both community taxonomic richness and evenness.
+Beyond the intensity threshold, the disturbance decreased the taxonomic richness of surviving trees which was not offset by the enrichment of pioneers, so that the overall community taxonomic richness decreased according to the disturbance intensity [@Molino2001].
 For community taxonomic evenness the disturbance impact was similar but slighter, as the evenness is less sensitive to the loss of rare species.
-Taxonomic evenness rather represented the increasing dominance of pionneers that balanced the usual hyper-dominance of tropical forest below the intensity threshold, thus increasing community overall evenness up to the intensity threshold beyond which pioneers became in turn highly dominant and decreased the overall evenness.
+Taxonomic evenness rather represented the increasing dominance of pionneers that balanced the usual hyper-dominance of tropical forest below the intensity threshold, thus increasing community overall evenness up to the intensity threshold beyond which pioneers became in turn highly dominant and decreased the overall evenness [@Baraloto2012a].
 
-Conversely the IDH was disproved regarding the disturbance impact on community functional diversity.
-Irrespective of the disturbance intensity the recruitment of pionneers, functionally highly different from the composition of pre-disturbance community, increased the community functional richness and eveness.
+Conversely the IDH was disproved regarding the disturbance impact on community functional richness and evenness.
+Irrespective of the disturbance intensity the recruitment of pionneers, functionally highly different from the composition of pre-disturbance community, increased both community functional richness and eveness.
 
-Along time, taxonomic richness trajectories for all disturbed communities first dropped similarly following the loss of species during disturbance.
-Then, trajectories showed a species gain depending on the disturbance intensity.
-The species gain was all the more significant as the disturbance intensity increased, up to an intensity threshold, through the establishment of long-lived pioneers enhancing community taxonomic richness and evenness in the long term.
-These long-lived pionneers, functionally quite different from the functional composition, entailed in parallel a progressive and long-lasting increase of the functional diversity.
+Along time, taxonomic richness trajectories of all disturbed communities first dropped similarly, following the species loss due to disturbance, and then displayed a species gain depending on the disturbance intensity.
+Up to an intensity threshold, the species gain was all the more significant that the disturbance intensity increased, with the establishment of long-lived pioneers enhancing community taxonomic richness and evenness in the long term.
+These long-lived pionneers, functionally quite different from the functional composition, entailed as well a progressive and long-lasting increase of the functional richness and evenness [@Denslow1980;@Molino2001].
 Beyond an intensity threshold, though, a few short-lived pioneers occupied the vacated environmental space and prevented the establishment of other species. 
@@
 Already after 10 years, though, short-lived pioneers started to decline and the functional richness and evenness decreased.
-Likely the establishment of long-lasting pioneers will follow this decrease and we expect the taxonomic and functional trajectories to catch up with those observed after intermediate disturbance.
+Likely this decrease will be followed by the establishment of long-lasting pioneers, and by the time they recruit we expect the taxonomic and functional trajectories to catch up with those observed after intermediate disturbance [@Anderson2007;@Walker2009].
 
@@
 For 15 years the species loss during disturbance, determined by the disturbance intensity, commensurately decreased the functional redundancy within the pre-disturbance functional space.
-The redundancy decrease was not compensated in the first place as the pionneers first recruited were functionally different from the pre-disturbance functional composition.
+The redundancy decrease was not compensated in the first place as the first recruited pionneers were functionally different from the pre-disturbance functional composition.
 Progressively though, first established species were replaced by more competitive long-lived pionneers or late-successional species resembling more the pre-disturbance functional composition and restoring the functional redundancy.
@@
 
-Our study revealed the post-disturbance recovery of both taxonomic composition, maintaining the initial differences among communities, and functional composition, that was in contrast similar among communities.
+Our study revealed community recovery through the combination of deterministic processes driving their convergence in functional composition, and dispersal limitation maintaining their divergence in taxonomic composition.
 The IDH was validated for community taxonomic richness and, to some extent, taxonomic evenness but disproved regarding community functional richness and evennes that were enhanced for any disturbance intensity by the high functional differences of pioneers compared to late-successional functoinal composition.

### 2018-07-09 WholePlotTrajectories.tex
relecture

@@
 by natural disturbance yielding a diversity peak at intermediate
-intensity, but this intermediate disturbance hypothesis (IDH) remains
-debated, and this controversy also questions the extent of communities
+intensity. This intermediate disturbance hypothesis (IDH),though,
+remains debated and the controversy questions the extent of communities'
 resilience regarding their functional and taxonomic facets. To
 disentangle the ecological processes driving community response to
-disturbance, we analysed the taxonomic and functional diversity
-trajectories following a disturbance gradient. Specifically, we
-examined, over 30 years, the functional and taxonomic community
-trajectories with regards to diversity, composition and redundancy.
-Functional trajectories were drawn based on 7 leaf, stem and
-life-history traits. We highlighted the cyclic recovery of community
-taxonomic and functional composition. The pre-disturbance taxonomic
-differences were maintained over time while the functional composition
-trajectories were quite similar. The IDH did predict communities
-functional diversity response while taxonomic diversity remained poorly
-sensitive to disturbance intensity. Although consistent, the recovery of
-community composition, diversity and redundancy remained unachieved
-after 30 years. This acknowledged the need of decades-long cycles with
-no disturbance to ensure a complete recovery, and questioned tropical
-forest community resilience after repeated disturbance.
+disturbance, we analysed the diversity trajectories over 30 years
+following a disturbance gradient. Specifically, we examined community
+functional and taxonomic trajectories with regards to diversity,
+composition and redundancy. Functional trajectories were drawn based on
+7 leaf, stem and life-history traits. We highlighted the cyclic recovery
+of community taxonomic and functional composition. While pre-disturbance
+taxonomic differences were maintained over time, functional composition
+trajectories were quite similar among communities. The IDH did predict
+communities taxonomic diversity response while functional diversity was
+enhanced whenever the disturbance intensity. Although consistent, the
+recovery of community composition, diversity and redundancy remained
+unachieved after 30 years. This acknowledged the need of decades-long
+cycles with no disturbance to ensure a complete recovery, and questioned
+tropical forest community resilience after repeated disturbance.
 }
@@
 trajectories in terms of richness, evenness, composition and redundancy
-\citep{Lohbeck2015, Guariguata2001}. Specifically, we (i) questioned the
-recovery of communities taxonomic and functional characteristics and
-identified the underlying assembly processes, (ii) clarified the
-validity of the IDH in the long term for tropical forest and elucidated
-its translation into different trajectories, and (iii) questioned
-community recovery time.
+\citep{Lohbeck2015, Guariguata2001}. Specifically, we (i) elucidated
+community taxonomic and functional recovery and the determinism of
+underlying processes, (ii) clarified the validity of the IDH in the long
+term for tropical forest and its translation into different trajectories
+in time, and (iii) questioned community recovery time.
 
@@
 
-\subsection{Communities richness and
-evenness}\label{communities-richness-and-evenness}
+\subsection{Communities Composition}\label{communities-composition}
 
@@
 disturbance), 828 388 individual trees and 591 botanical species
-pertaining to 223 genus and 64 families were recorded. For undisturbed
-plots, taxonomic Richness and Evenness remained stable over the 30 years
-monitored. In disturbed communities, after low disturbance intensity the
-taxonomic richness increased, reaching a maximum gain of 14 botanical
-genera (plot 3 from treatment 2). After intense disturbance the
-taxonomic richness followed a more complex trajectory, decreasing for
-ten years after disturbance before recovering to pre-disturbance values.
-The maximum richness loss or gain after disturbance was positively
-correlated to the disturbance intensity
-(\(\rho_{spearman}^{Richness}=0.50\)). In all disturbed plots the
-taxonomic evenness first increased until a maximum reached after around
-20 years. This maximum was positively correlated to the disturbance
-intensity (\(\rho_{spearman}^{Evenness}=0.77\)). The evenness then
-stabilized except for two T3 plots (plots 8 and 12) for which evenness
-kept increasing.
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivTaxo-1}
-
-}
-
-\caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic \textbf{(a)} richness, \textbf{(b)}, taxonomic evenness, \textbf{(c)} functional richness, and \textbf{(d)} functional evenness.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:DivTaxo}
-\end{figure*}
-
-The plot 7 from treatment 1 displayed constantly outlying functional
-richness and evenness and was removed from the graphical representation
-for better readability. In undisturbed plots both functional richness
-and evenness remained stable along the 30 years. In disturbed plots,
-functional richness and evenness trajectories depended on the
-disturbance intensity with their maximum positively correlated to \%AGB
-loss \(\rho_{spearman}^{Richness}=0.76\) and
-\(\rho_{spearman}^{Evenness}=0.60\). Functional richness and evenness
-displayed for low disturbance intensity a low but long-lasting increase
-up to a maximum reached after 20-25 years, and for high intensity, a
-fast but short increase followed after 10 years by a slow decrease
-towards the inital values.
-
-The second-degree polynomial regressions between (i) the \%AGB loss and
-(ii) taxonomic and functional richness and evenness after 10, 20 and 30
-years best predicted the hump-shaped curve of the disturbance impact
-along the disturbance intensity gradient \ref{fig:IDHplot}. The
-relationship between the disturbance impact and its intensity was more
-markedly hump-shaped for the taxonomic richness than for the taxonomic
-evenness. For both functional richness and evenness the relationship was
-almost linear. The regression model better predicted the functional
-richness and evenness (\(0.55<R^2_{Functional Richness}<0.72\), and
-\(0.60<R^2_{Functional Evenness}<0.81\)) than the taxonomic richness and
-evenness (\(0.21<R^2_{Taxonomic Richness}<0.4\), and
-\(-0.15<R^2_{Taxonomic Evenness}<0.43\) respectively)
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/IDHplot-1}
-
-}
-
-\caption{Relationship between the initial \%AGB loss and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)}, functional richness \textbf{(c)},and functional evenness \textbf{(d)} at 10, 20 and 30 years after disturbance.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:IDHplot}
-\end{figure*}
-
-\subsection{Communities Composition}\label{communities-composition}
+pertaining to 223 genus and 64 families were recorded.
 
@@
 
+\subsection{Communities richness and
+evenness}\label{communities-richness-and-evenness}
+
+For undisturbed plots, taxonomic Richness and Evenness remained stable
+over the 30 years monitored. In disturbed communities, after low
+disturbance intensity the taxonomic richness increased, reaching a
+maximum gain of 14 botanical genera (plot 3 from treatment 2). After
+intense disturbance the taxonomic richness followed a more complex
+trajectory, decreasing for ten years after disturbance before recovering
+to pre-disturbance values. The maximum richness loss or gain after
+disturbance was positively correlated to the disturbance intensity
+(\(\rho_{spearman}^{Richness}=0.50\)). In all disturbed plots the
+taxonomic evenness first increased until a maximum reached after around
+20 years. This maximum was positively correlated to the disturbance
+intensity (\(\rho_{spearman}^{Evenness}=0.77\)). The evenness then
+stabilized except for two T3 plots (plots 8 and 12) for which evenness
+kept increasing.
+
+\begin{figure*}
+
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivTaxo-1}
+
+}
+
+\caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic \textbf{(a)} richness, \textbf{(b)}, taxonomic evenness, \textbf{(c)} functional richness, and \textbf{(d)} functional evenness.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:DivTaxo}
+\end{figure*}
+
+The plot 7 from treatment 1 displayed constantly outlying functional
+richness and evenness and was removed from the graphical representation
+for better readability. In undisturbed plots both functional richness
+and evenness remained stable along the 30 years. In disturbed plots,
+functional richness and evenness trajectories depended on the
+disturbance intensity with their maximum positively correlated to \%AGB
+loss \(\rho_{spearman}^{Richness}=0.76\) and
+\(\rho_{spearman}^{Evenness}=0.60\). Functional richness and evenness
+displayed for low disturbance intensity a low but long-lasting increase
+up to a maximum reached after 20-25 years, and for high intensity, a
+fast but short increase followed after 10 years by a slow decrease
+towards the inital values.
+
+The second-degree polynomial regressions between (i) the \%AGB loss and
+(ii) taxonomic and functional richness and evenness after 10, 20 and 30
+years best predicted the hump-shaped curve of the disturbance impact
+along the disturbance intensity gradient \ref{fig:IDHplot}. The
+relationship between the disturbance impact and its intensity was more
+markedly hump-shaped for the taxonomic richness than for the taxonomic
+evenness. For both functional richness and evenness the relationship was
+almost linear. The regression model better predicted the functional
+richness and evenness (\(0.55<R^2_{Functional Richness}<0.72\), and
+\(0.60<R^2_{Functional Evenness}<0.81\)) than the taxonomic richness and
+evenness (\(0.21<R^2_{Taxonomic Richness}<0.4\), and
+\(-0.15<R^2_{Taxonomic Evenness}<0.43\) respectively)
+
+\begin{figure*}
+
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/IDHplot-1}
+
+}
+
+\caption{Relationship between the initial \%AGB loss and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)}, functional richness \textbf{(c)},and functional evenness \textbf{(d)} at 10, 20 and 30 years after disturbance.
+Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:IDHplot}
+\end{figure*}
+
 \subsection{Functional redundancy}\label{functional-redundancy}
@@
 
-Conversely, disturbed communities followed functional trajecotories
+Conversely, disturbed communities followed functional trajectories
 similar in terms of functional composition. As pre-disturbance surviving
@@
 of high availability of light, space and nutrient vacated by the
-disturbance. The recruitment of pioneers similarly changed community
-functional composition towards more resource-acquisitive strategies
-\citep{Grime1998, Westoby1998, Wright2004, Reich2014}. Thereafter, again
-in the same way for all disturbed communities, long-lived, more
-resistant and shade-tolerant species excluded the first established
-pioneers and started the recovery of pre-disturbance functional
-composition.
+disturbance. The recruitment of pioneers changed community functional
+composition in the same way for all disturbance intensity towards more
+resource-acquisitive strategies, moving community functional composition
+right along the first axis in Figure \ref{fig:NMDSplans}
+\citep{Grime1998, Westoby1998, Wright2004, Reich2014}. Thereafter
+long-lived, more resistant and shade-tolerant species excluded the first
+established pioneers and started the recovery of pre-disturbance
+functional composition, moving similarly community functional
+composition left along the first axis and upward along the second axis
+in Figure \ref{fig:NMDSplans} .
+
+These trajectories provided empirical support to the hypothesis that
+community assembly is both deterministic and historically convergent at
+different levels of community organization. Deterministic, trait-based
+processes drove community convergence in functional composition, while
+at the same time dispersal limitation maintained their divergence in
+taxonomic composition \citep{Fukami2005}.
 
@@
 recruited after disturbance is restricted in the Guiana Shield to a few
-common genera (e.g.~Cecropia spp., Vismia spp.). Below the intensity
-threshold the size of the surviving community maintained the
-pre-disturbance taxonomic richness while the recruitment of pioneers,
-infrequent or absent before disturbance, increased both community
-taxonomic richness and evenness. Beyond the intensity threshold, the
-disturbance decreased the taxonomic richness of surviving trees which
-was not offset by the enrichment of pioneers, so that the overall
-community taxonomic richness decreased according to the disturbance
-intensity. For community taxonomic evenness the disturbance impact was
-similar but slighter, as the evenness is less sensitive to the loss of
-rare species. Taxonomic evenness rather represented the increasing
-dominance of pionneers that balanced the usual hyper-dominance of
-tropical forest below the intensity threshold, thus increasing community
-overall evenness up to the intensity threshold beyond which pioneers
-became in turn highly dominant and decreased the overall evenness.
+common genera (e.g.~Cecropia spp., Vismia spp.)
+\citep{Zalamea2012, Guitet2018}. Below the intensity threshold the size
+of the surviving community maintained the pre-disturbance high taxonomic
+richness while the recruitment of pioneers, infrequent or absent before
+disturbance, increased both community taxonomic richness and evenness.
+Beyond the intensity threshold, the disturbance decreased the taxonomic
+richness of surviving trees which was not offset by the enrichment of
+pioneers, so that the overall community taxonomic richness decreased
+according to the disturbance intensity \citep{Molino2001}. For community
+taxonomic evenness the disturbance impact was similar but slighter, as
+the evenness is less sensitive to the loss of rare species. Taxonomic
+evenness rather represented the increasing dominance of pionneers that
+balanced the usual hyper-dominance of tropical forest below the
+intensity threshold, thus increasing community overall evenness up to
+the intensity threshold beyond which pioneers became in turn highly
+dominant and decreased the overall evenness \citep{Baraloto2012a}.
 
 Conversely the IDH was disproved regarding the disturbance impact on
-community functional diversity. Irrespective of the disturbance
-intensity the recruitment of pionneers, functionally highly different
-from the composition of pre-disturbance community, increased the
-community functional richness and eveness.
+community functional richness and evenness. Irrespective of the
+disturbance intensity the recruitment of pionneers, functionally highly
+different from the composition of pre-disturbance community, increased
+both community functional richness and eveness.
 
-Along time, taxonomic richness trajectories for all disturbed
-communities first dropped similarly following the loss of species during
-disturbance. Then, trajectories showed a species gain depending on the
-disturbance intensity. The species gain was all the more significant as
-the disturbance intensity increased, up to an intensity threshold,
-through the establishment of long-lived pioneers enhancing community
-taxonomic richness and evenness in the long term. These long-lived
-pionneers, functionally quite different from the functional composition,
-entailed in parallel a progressive and long-lasting increase of the
-functional diversity. Beyond an intensity threshold, though, a few
-short-lived pioneers occupied the vacated environmental space and
-prevented the establishment of other species. These short-lived pioneers
-were functionally very different from the pre-disturbance community and
-entailed a rapid and significant increase of functional richness en
-evenness. Already after 10 years, though, short-lived pioneers started
-to decline and the functional richness and evenness decreased. Likely
-the establishment of long-lasting pioneers will follow this decrease and
-we expect the taxonomic and functional trajectories to catch up with
-those observed after intermediate disturbance.
+Along time, taxonomic richness trajectories of all disturbed communities
+first dropped similarly, following the species loss due to disturbance,
+and then displayed a species gain depending on the disturbance
+intensity. Up to an intensity threshold, the species gain was all the
+more significant that the disturbance intensity increased, with the
+establishment of long-lived pioneers enhancing community taxonomic
+richness and evenness in the long term. These long-lived pionneers,
+functionally quite different from the functional composition, entailed
+as well a progressive and long-lasting increase of the functional
+richness and evenness \citep{Denslow1980, Molino2001}. Beyond an
+intensity threshold, though, a few short-lived pioneers occupied the
+vacated environmental space and prevented the establishment of other
+species. These short-lived pioneers were functionally very different
+from the pre-disturbance community and entailed a rapid and significant
+increase of functional richness en evenness. Already after 10 years,
+though, short-lived pioneers started to decline and the functional
+richness and evenness decreased. Likely this decrease will be followed
+by the establishment of long-lasting pioneers, and by the time they
+recruit we expect the taxonomic and functional trajectories to catch up
+with those observed after intermediate disturbance
+\citep{Anderson2007, Walker2009}.
 
@@
 redundancy within the pre-disturbance functional space. The redundancy
-decrease was not compensated in the first place as the pionneers first
-recruited were functionally different from the pre-disturbance
+decrease was not compensated in the first place as the first recruited
+pionneers were functionally different from the pre-disturbance
 functional composition. Progressively though, first established species
@@
 
-Our study revealed the post-disturbance recovery of both taxonomic
-composition, maintaining the initial differences among communities, and
-functional composition, that was in contrast similar among communities.
-The IDH was validated for community taxonomic richness and, to some
-extent, taxonomic evenness but disproved regarding community functional
-richness and evennes that were enhanced for any disturbance intensity by
-the high functional differences of pioneers compared to
-late-successional functoinal composition. The IDH was translated in time
-by the recruitment, beyond an intensity threshold, of short-lived
-pionneers that prevented in the first times the establishment of more
-diverse long-lived pionneers, recruited otherwise below the intensity
-threshold. The resilience of tropical forests proved consistent although
-several decades-long. Still, the disturbance impact on communities
-redundancy cautioned against the risks of infrequent species loss and
-the persistence of disturbance-specific communities
+Our study revealed community recovery through the combination of
+deterministic processes driving their convergence in functional
+composition, and dispersal limitation maintaining their divergence in
+taxonomic composition. The IDH was validated for community taxonomic
+richness and, to some extent, taxonomic evenness but disproved regarding
+community functional richness and evennes that were enhanced for any
+disturbance intensity by the high functional differences of pioneers
+compared to late-successional functoinal composition. The IDH was
+translated in time by the recruitment, beyond an intensity threshold, of
+short-lived pionneers that prevented in the first times the
+establishment of more diverse long-lived pionneers, recruited otherwise
+below the intensity threshold. The resilience of tropical forests proved
+consistent although several decades-long. Still, the disturbance impact
+on communities redundancy cautioned against the risks of infrequent
+species loss and the persistence of disturbance-specific communities
 \citep{Gourlet-Fleury2005}. As the trajectories highlighted the

### 2018-07-10 WholePlotTrajectories.Rmd
coquille

@@
 Lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals", fig.env="figure*"}
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure*"}
 load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
@@
 ```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWSG), and bark thickness, (ref:capBarkThick)) and one life history trait (Specific maximum height at adult stage, (ref:capHmax)).
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure*"}
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure*"}
 load("DB/CWM")
@@
 ```{r DivTaxo, out.width = '100%', echo=FALSE, fig.cap="Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic (ref:Bolda) richness, (ref:Boldb), taxonomic evenness, (ref:Boldc) functional richness, and (ref:Boldd) functional evenness.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
 load("DB/ReplacementTraj_ForGraphs")
@@
 ```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/3,fig.cap="Relationship between the initial \\%AGB loss and community taxonmic richness (ref:Bolda), taxonomic evenness (ref:Boldb), functional richness (ref:Boldc),and functional evenness (ref:Boldd) at 10, 20 and 30 years after disturbance.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
 load("DB/RichnessIDH");load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB");load("DB/FunRichIDH")
@@
 ```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
-Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
+Colors are disturbance treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
 load("DB/Redundancy_restricted")

### 2018-07-11 WholePlotTrajectories.Rmd
biblio last reread

@@
     The huge diversity of tropical forests is often assumed to be regularly reshaped by natural disturbance yielding a diversity peak at intermediate intensity.
-This intermediate disturbance hypothesis (IDH),though, remains debated and the controversy questions the extent of communities' resilience regarding their functional and taxonomic facets.
+This intermediate disturbance hypothesis (IDH), though, remains debated and the controversy questions the extent of communities' resilience regarding their functional and taxonomic facets.
 
@@
   \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
-  \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion}
+  \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
 bibliography: references.bib
@@
 
+## abstract: 204 words / text without legends : 3659 words
+
 source("Scripts/DrawingTools.R")
@@
 They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Gibson2011a; @Morales-Hidalgo2015] and threatens the maintenance and dynamics in space and time of communities structure, composition and functioning [@Anderson-Teixeira2013;@Sist2015].
+The growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Morales-Hidalgo2015] and threatens the maintenance and dynamics in space and time of communities structure, composition and functioning [@Anderson-Teixeira2013].
 
 In tropical forests, ecological communities are regularly re-shaped by natural disturbance events changing both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions such as competition among species [@Chesson2000;@Herault2018].
-One of the cornerstone of tropical forest ecology is to understand the processes and drivers of ecosystems response to disturbance [@White2001;@Chazdon2003a]. 
+One of the cornerstone of tropical forest ecology is to understand the processes and drivers of ecosystems response to disturbance [@Chazdon2003a]. 
 For now, this has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016] that are rapid and convenient to measure.
@@
 However the response of forests diversity and composition is still unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014; @Liang2016].
-In the short-term, moderate disturbance may lead to positive impacts on communities diversity, an idea formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity when disturbance intensity is not too high [@Molino2001;@Kariuki2006a;@Berry2008a].
+In the short-term, moderate disturbance may lead to positive impacts on communities diversity, an idea formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity when disturbance intensity is not too high [@Molino2001;@Kariuki2006a].
 
 Validations of the IDH though remain scarce in the long-term and mainly rely on the analysis of taxonomic richness [@Molino2001].
-Taxonomic richness alone, though, gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
-More ecological-meaningful analysis would couple richness with (i) evenness that would reveal underlying ecological processes and (ii) composition that is crucial for conservation issues [@Magurran1988;@Lavorel2002;@Bellwood2006].
-Furthermore, a functional approach accounting for species biological attributes would directly link communities diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
-In that respect, the functional trait-based approach that focus on major traits related to species ecology and mediate species performance in a given environment was sucessfully adopted [@Diaz2005;@Villeger2008a].
-For instance, the functional approach revealed in tropical rainforests the deterministic processes entailing, after disturbance, a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
+Taxonomic richness alone, though, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016].
+More ecological-meaningful analysis would couple richness with (i) evenness that would reveal underlying ecological processes and (ii) composition that is crucial for conservation issues [@Lavorel2002;@Bellwood2006].
+Furthermore, a functional approach accounting for species biological attributes would directly link communities diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Baraloto2012a].
+In that respect, the functional trait-based approach that focus on major traits related to species ecology and mediate species performance in a given environment was sucessfully adopted [@Diaz2005].
+For instance, the functional approach revealed in tropical rainforests the deterministic processes entailing, after disturbance, a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Reich2014;@Herault2011].
 This shift is translated into the trajectories of key functional traits related to resource acquisition (leaf and stem traits) and life-history traits (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
 Eventually a complete overview of communities response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
-The high functional redundancy of hyperdiverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Trenbath1999; @Elmqvist2003; @Diaz2005].
+The high functional redundancy of hyperdiverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Elmqvist2003; @Diaz2005].
 
 In this study, we monitored over 30 years the response of 75 ha of neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem above-ground biomass (AGB) loss. 
-We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy [@Lohbeck2015;@Guariguata2001].
+We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
 Specifically, we (i) elucidated community taxonomic and functional recovery and the determinism of underlying processes, (ii) clarified the validity of the IDH in the long term for tropical forest and its translation into different trajectories in time, and (iii) questioned community recovery time.
@@
 The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3).
-Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Schmitt1990;@Blanc2009]. 
+Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Schmitt1990]. 
 The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Biomass2018].
+
+```{r Tab1, echo=FALSE,as.is=T}
+Intervention <- data.frame(
+  Treatment=c("Control", "T1", "T2", "T3"),
+  Timber=c("", 
+           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha", 
+           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha",
+           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha"),
+  Thinning=c("", 
+             "", 
+             "DBH $\\geq$ 40 cm, non-valuable species, $\\approx$ 30 trees/ha", 
+             "DBH $\\geq$ 50 cm, non-valuable species, $\\approx$ 15 trees/ha  "),
+  Fuelwood=c("", 
+             "", 
+             "", 
+             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-valuable species, $\\approx$ 15 trees/ha"),
+  AGBLost=c(0, "$[12\\%-33\\%]$", "$[33\\%-56\\%]$", "$[35\\%-56\\%]$")
+)
+names(Intervention)[5] <- "\\%AGB lost"
+knitr::kable(Intervention, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.", longtable = FALSE, booktabs = TRUE, escape = FALSE,format="latex") %>%
+  kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
+```
 
@@
 
-See appendix 1 and @Aubry-Kientz2013 for the detailed methodology.
+See Supplementary Materials -Figure S1 and @Aubry-Kientz2013 for the detailed methodology.
 
@@
 Trajectories along time were reported through the euclidean distance between the target inventories and the reference inventories in 1989, *i.e.* 2 years after disturbance. 
-Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007;@Garnier2004]. 
-Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were therefore reported as the proportion of each class in the inventories (Supp Mat XX).
+Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007]. 
+Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were therefore reported as the proportion of each class in the inventories (Supplementary materials).
 
 The taxonomic diversity was reported through species richness and evenness, *i.e* the Hill number translation of the Simpson index [@Hill1973].
-These indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recomended for diversity studies [@Patil1982;@Tothmeresz1995;@Marcon2015b].
+These indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recomended for diversity studies [@Marcon2015b]. 
 The functional diversity was reported using the functional richness and functional evenness, *i.e* Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on species functional traits.
@@
 Species TDP weighted by species abundance were eventually summed for each community.
-Community functional redundancy was the sum of TDPs overlap, expressed as the average number of species that could be removed from without reducing the functional space (see appendix I for a more comprehensive sheme).
+Community functional redundancy was the sum of TDPs overlap, expressed as the average number of species that could be removed from without reducing the functional space (Supp.
+Mat.
+- Figure S1 for a more comprehensive sheme).
 
@@
 For all traits, the maximum difference to initial value was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.76$, $\rho_{spearman}^{L_{chloro}}=0.60$, $\rho_{spearman}^{L_{toughness}}=-0.53$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WSG}=-0.75$, $\rho_{spearman}^{Bark-thickness}=0.71$, $\rho_{spearman}^{Hmax}=-0.40$).
+The proportions of the three lightest seed mass classes increased in all disturbed plots, and decreased after 30 years for the lightest class while it stabilized for the two other (Supp.
+Mat.
+- Figure S2).
 
@@
 Competitive pioneers became dominant in filling the environmental niches of high availability of light, space and nutrient vacated by the disturbance.
-The recruitment of pioneers changed community functional composition in the same way for all disturbance intensity towards more resource-acquisitive strategies, moving community functional composition right along the first axis in Figure \@ref(fig:NMDSplans) [@Grime1998;@Westoby1998; @Wright2004; @Reich2014].
-Thereafter long-lived, more resistant and shade-tolerant species excluded the first established pioneers and started the recovery of pre-disturbance functional composition, moving similarly community functional composition left along the first axis and upward along the second axis in Figure \@ref(fig:NMDSplans) .
+The recruitment of pioneers changed community functional composition in the same way for all disturbance intensity towards more resource-acquisitive strategies, moving community functional composition right along the first axis in Figure \@ref(fig:NMDSplans) [@Westoby1998; @Wright2004; @Reich2014].
+Thereafter long-lived, more resistant and shade-tolerant species excluded the first established pioneers and started the recovery of pre-disturbance functional composition, moving similarly community functional composition left along the first axis and upward along the second axis in Figure \@ref(fig:NMDSplans).
 
@@
 The disturbance intensity determined the balance in the community between pre-disturbance surviving trees and those recruited afterward.
-The pool of true pioneer species specifically recruited after disturbance is restricted in the Guiana Shield to a few common genera (e.g. Cecropia spp., Vismia spp.) [@Zalamea2012;@Guitet2018].
+The pool of true pioneer species specifically recruited after disturbance is restricted in the Guiana Shield to a few common genera (e.g. Cecropia spp., Vismia spp.) [@Guitet2018].
 Below the intensity threshold the size of the surviving community maintained the pre-disturbance high taxonomic richness while the recruitment of pioneers, infrequent or absent before disturbance, increased both community taxonomic richness and evenness.
@@
 Already after 10 years, though, short-lived pioneers started to decline and the functional richness and evenness decreased.
-Likely this decrease will be followed by the establishment of long-lasting pioneers, and by the time they recruit we expect the taxonomic and functional trajectories to catch up with those observed after intermediate disturbance [@Anderson2007;@Walker2009].
+Likely this decrease will be followed by the establishment of long-lasting pioneers, and by the time they recruit we expect the taxonomic and functional trajectories to catch up with those observed after intermediate disturbance [@Walker2009].
 
@@
 This replacement was stochastic and followed the lottery recruitment rules, implying a recruitment eased for the first recruited species but then increasingly hampered by the emergence of interspecific competition [@Busing2002].
-Along time the recovery of infrequent species was increasingly difficult, so that the time for the full recovery of the functional redundancy, in some communities just initiated after 30 years, was difficult to estimate [@Trenbath1999;@Elmqvist2003;@Diaz2005].
+Along time the recovery of infrequent species was increasingly difficult, so that the time for the full recovery of the functional redundancy, in some communities just initiated after 30 years, was difficult to estimate [@Elmqvist2003;@Diaz2005].
 
-The long-term impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008; @Burslem2000; @Martin2013].
+The long-term impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008].
 Besides the long-term recovery of infrequent species increases the risks to loose cornerstone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].

### 2018-07-11 WholePlotTrajectories.tex
pdf
biblio last reread

@@
 \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
-\hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion}
+\hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
 
@@
 by natural disturbance yielding a diversity peak at intermediate
-intensity. This intermediate disturbance hypothesis (IDH),though,
+intensity. This intermediate disturbance hypothesis (IDH), though,
 remains debated and the controversy questions the extent of communities'
@@
 forests products together with current global changes increases the
-pressure on remaining natural forests
-\citep{Gibson2011a, Morales-Hidalgo2015} and threatens the maintenance
-and dynamics in space and time of communities structure, composition and
-functioning \citep{Anderson-Teixeira2013, Sist2015}.
+pressure on remaining natural forests \citep{Morales-Hidalgo2015} and
+threatens the maintenance and dynamics in space and time of communities
+structure, composition and functioning \citep{Anderson-Teixeira2013}.
 
@@
 forest ecology is to understand the processes and drivers of ecosystems
-response to disturbance \citep{White2001, Chazdon2003a}. For now, this
-has been largely studied through forest structural parameters such as
-aboveground biomass, tree height or stem density
+response to disturbance \citep{Chazdon2003a}. For now, this has been
+largely studied through forest structural parameters such as aboveground
+biomass, tree height or stem density
 \citep{Piponiot2016, Rutishauser2016} that are rapid and convenient to
@@
 disturbance hypothesis (IDH) stating a maximized species diversity when
-disturbance intensity is not too high
-\citep{Molino2001, Kariuki2006a, Berry2008a}.
+disturbance intensity is not too high \citep{Molino2001, Kariuki2006a}.
 
@@
 richness alone, though, gives limited or misleading information on
-forests recovery and functioning \citep{Martin2015, Chaudhary2016}. More
+forests recovery and functioning \citep{Chaudhary2016}. More
 ecological-meaningful analysis would couple richness with (i) evenness
@@
 that is crucial for conservation issues
-\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
-functional approach accounting for species biological attributes would
-directly link communities diversity, composition and redundancy to
-ecosystem functioning and to its environmental constraints
-\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
-respect, the functional trait-based approach that focus on major traits
-related to species ecology and mediate species performance in a given
-environment was sucessfully adopted \citep{Diaz2005, Villeger2008a}. For
-instance, the functional approach revealed in tropical rainforests the
-deterministic processes entailing, after disturbance, a functional shift
-from a dominance of ``conservative'' slow-growing species dealing with
-scarce resources to ``acquisitive'' fast-growing species with rapid and
-efficient use of abundant resources
-\citep{TerSteege2001, Reich2014, Herault2011}. This shift is translated
-into the trajectories of key functional traits related to resource
-acquisition (leaf and stem traits) and life-history traits (seed mass,
-maximum size)
+\citep{Lavorel2002, Bellwood2006}. Furthermore, a functional approach
+accounting for species biological attributes would directly link
+communities diversity, composition and redundancy to ecosystem
+functioning and to its environmental constraints
+\citep{Violle2007b, Baraloto2012a}. In that respect, the functional
+trait-based approach that focus on major traits related to species
+ecology and mediate species performance in a given environment was
+sucessfully adopted \citep{Diaz2005}. For instance, the functional
+approach revealed in tropical rainforests the deterministic processes
+entailing, after disturbance, a functional shift from a dominance of
+``conservative'' slow-growing species dealing with scarce resources to
+``acquisitive'' fast-growing species with rapid and efficient use of
+abundant resources \citep{Reich2014, Herault2011}. This shift is
+translated into the trajectories of key functional traits related to
+resource acquisition (leaf and stem traits) and life-history traits
+(seed mass, maximum size)
 \citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Eventually
@@
 ecosystem functioning and determines the resilience of communities after
-disturbance \citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+disturbance \citep{Elmqvist2003, Diaz2005}.
 
@@
 and life-history traits in order to draw the taxonomic and functional
-trajectories in terms of richness, evenness, composition and redundancy
-\citep{Lohbeck2015, Guariguata2001}. Specifically, we (i) elucidated
-community taxonomic and functional recovery and the determinism of
-underlying processes, (ii) clarified the validity of the IDH in the long
-term for tropical forest and its translation into different trajectories
-in time, and (iii) questioned community recovery time.
+trajectories in terms of richness, evenness, composition and redundancy.
+Specifically, we (i) elucidated community taxonomic and functional
+recovery and the determinism of underlying processes, (ii) clarified the
+validity of the IDH in the long term for tropical forest and its
+translation into different trajectories in time, and (iii) questioned
+community recovery time.
 
@@
 (T3). Treatments T2 and T3 besides included the thinning of trees by
-poison girdling \citep{Schmitt1990, Blanc2009}. The disturbance
-intensity was measured as the percentage of aboveground biomass (\%AGB)
-lost between the first inventory in 1984 and five years after
-disturbance \citep{Piponiot2016} estimated with the BIOMASS R package
+poison girdling \citep{Schmitt1990}. The disturbance intensity was
+measured as the percentage of aboveground biomass (\%AGB) lost between
+the first inventory in 1984 and five years after disturbance
+\citep{Piponiot2016} estimated with the BIOMASS R package
 \citep{Biomass2018}.
+
+\begin{table}
+
+\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.}
+\centering
+\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
+\toprule
+Treatment & Timber & Thinning & Fuelwood & \%AGB lost\\
+\midrule
+Control &  &  &  & 0\\
+T1 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha &  &  & $[12\%-33\%]$\\
+T2 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30 trees/ha &  & $[33\%-56\%]$\\
+T3 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & $[35\%-56\%]$\\
+\bottomrule
+\end{tabu}
+\end{table}
 
@@
 
-See appendix 1 and \citet{Aubry-Kientz2013} for the detailed
-methodology.
+See Supplementary Materials -Figure S1 and \citet{Aubry-Kientz2013} for
+the detailed methodology.
 
@@
 leaf, stem and life-history traits were also visualized with the
-community weighted means (CWM) \citep{Diaz2007, Garnier2004}. Species
-seed mass corresponded to 5 classes of increasing mass, seed mass
-trajectories were therefore reported as the proportion of each class in
-the inventories (Supp Mat XX).
+community weighted means (CWM) \citep{Diaz2007}. Species seed mass
+corresponded to 5 classes of increasing mass, seed mass trajectories
+were therefore reported as the proportion of each class in the
+inventories (Supplementary materials).
 
@@
 entropy, respectively corresponding to the 0 and 2 order of diversity
-(q), recomended for diversity studies
-\citep{Patil1982, Tothmeresz1995, Marcon2015}. The functional diversity
-was reported using the functional richness and functional evenness,
-\emph{i.e} Rao index of quadratic entropy which combines species
-abundance distribution and average pairwise dissimilarity based on
-species functional traits.
+(q), recomended for diversity studies \citep{Marcon2015b}. The
+functional diversity was reported using the functional richness and
+functional evenness, \emph{i.e} Rao index of quadratic entropy which
+combines species abundance distribution and average pairwise
+dissimilarity based on species functional traits.
 
@@
 expressed as the average number of species that could be removed from
-without reducing the functional space (see appendix I for a more
+without reducing the functional space (Supp. Mat. - Figure S1 for a more
 comprehensive sheme).
@@
 Lower panels (\textbf{(c)} and \textbf{(d)}) represent the euclidean distance to initial condition along the 30 sampled years.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals}\label{fig:NMDSplans}
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:NMDSplans}
 \end{figure*}
@@
 \(\rho_{spearman}^{Bark-thickness}=0.71\),
-\(\rho_{spearman}^{Hmax}=-0.40\)).
+\(\rho_{spearman}^{Hmax}=-0.40\)). The proportions of the three lightest
+seed mass classes increased in all disturbed plots, and decreased after
+30 years for the lightest class while it stabilized for the two other
+(Supp. Mat. - Figure S2).
 
@@
 \caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WSG}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}).
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
 \end{figure*}
@@
 \caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic \textbf{(a)} richness, \textbf{(b)}, taxonomic evenness, \textbf{(c)} functional richness, and \textbf{(d)} functional evenness.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:DivTaxo}
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:DivTaxo}
 \end{figure*}
@@
 \caption{Relationship between the initial \%AGB loss and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)}, functional richness \textbf{(c)},and functional evenness \textbf{(d)} at 10, 20 and 30 years after disturbance.
-Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:IDHplot}
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:IDHplot}
 \end{figure*}
@@
 \caption{Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
-Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}(\#fig:RedFun_rest)
+Colors are disturbance treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}(\#fig:RedFun_rest)
 \end{figure}
@@
 right along the first axis in Figure \ref{fig:NMDSplans}
-\citep{Grime1998, Westoby1998, Wright2004, Reich2014}. Thereafter
-long-lived, more resistant and shade-tolerant species excluded the first
-established pioneers and started the recovery of pre-disturbance
-functional composition, moving similarly community functional
-composition left along the first axis and upward along the second axis
-in Figure \ref{fig:NMDSplans} .
+\citep{Westoby1998, Wright2004, Reich2014}. Thereafter long-lived, more
+resistant and shade-tolerant species excluded the first established
+pioneers and started the recovery of pre-disturbance functional
+composition, moving similarly community functional composition left
+along the first axis and upward along the second axis in Figure
+\ref{fig:NMDSplans}.
 
@@
 recruited after disturbance is restricted in the Guiana Shield to a few
-common genera (e.g.~Cecropia spp., Vismia spp.)
-\citep{Zalamea2012, Guitet2018}. Below the intensity threshold the size
-of the surviving community maintained the pre-disturbance high taxonomic
-richness while the recruitment of pioneers, infrequent or absent before
-disturbance, increased both community taxonomic richness and evenness.
-Beyond the intensity threshold, the disturbance decreased the taxonomic
-richness of surviving trees which was not offset by the enrichment of
-pioneers, so that the overall community taxonomic richness decreased
-according to the disturbance intensity \citep{Molino2001}. For community
-taxonomic evenness the disturbance impact was similar but slighter, as
-the evenness is less sensitive to the loss of rare species. Taxonomic
+common genera (e.g.~Cecropia spp., Vismia spp.) \citep{Guitet2018}.
+Below the intensity threshold the size of the surviving community
+maintained the pre-disturbance high taxonomic richness while the
+recruitment of pioneers, infrequent or absent before disturbance,
+increased both community taxonomic richness and evenness. Beyond the
+intensity threshold, the disturbance decreased the taxonomic richness of
+surviving trees which was not offset by the enrichment of pioneers, so
+that the overall community taxonomic richness decreased according to the
+disturbance intensity \citep{Molino2001}. For community taxonomic
+evenness the disturbance impact was similar but slighter, as the
+evenness is less sensitive to the loss of rare species. Taxonomic
 evenness rather represented the increasing dominance of pionneers that
@@
 recruit we expect the taxonomic and functional trajectories to catch up
-with those observed after intermediate disturbance
-\citep{Anderson2007, Walker2009}.
+with those observed after intermediate disturbance \citep{Walker2009}.
 
@@
 functional redundancy, in some communities just initiated after 30
-years, was difficult to estimate
-\citep{Trenbath1999, Elmqvist2003, Diaz2005}.
+years, was difficult to estimate \citep{Elmqvist2003, Diaz2005}.
 
@@
 chances to see the persistence of disturbance-specific species at the
-expense of late-successional ones
-\citep{Haddad2008, Burslem2000, Martin2013}. Besides the long-term
-recovery of infrequent species increases the risks to loose cornerstone
-species, with unexpected ecological consequences
+expense of late-successional ones \citep{Haddad2008}. Besides the
+long-term recovery of infrequent species increases the risks to loose
+cornerstone species, with unexpected ecological consequences
 \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}. Apart from the

### 2018-07-11 WholePlotTrajectories_SuppMat.Rmd
biblio last reread

--- 2018-07-01 AppendixI.Rmd
@@
 ---
-title: "Appendix I"
+title: "Supplementary materials"
 author:
@@
 
-**Figure A1**: Whenever the family or genus were recorded for a vernacular name, the pool of species sampled to replace the vernacular name was restricted to those pertaining to the same higher taxonomic level (i.e species of the same genus when the species is missing, species of the same family when the genus is missing) \@ref(fig:FigBayesianScheme).
+**Figure S1**: Whenever the family or genus were recorded for a vernacular name, the pool of species sampled to replace the vernacular name was restricted to those pertaining to the same higher taxonomic level (i.e species of the same genus when the species is missing, species of the same family when the genus is missing) \@ref(fig:FigBayesianScheme).
 
-**Figure A2**: Scheme of the redundancy measurement process
-
-**Figure A3**: Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+**Figure S2**: Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
 No gap filling process was applied in this case.
@@
 
-**Figure A4**: Trajectories of the Rao functional diversity over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.
-\@ref(fig:DivFun).
-
-**Figure A5**: Summarizing plots of the functional PCA analysis defining communities functional space.
-**(a)** Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
-**(b)** bow plot of the PCA eignevalues in percentage.
-**(c)** Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).
-\@ref(fig:PCA_graphs).
-
-**Figure A6**: Trajectories of the functional redundancy over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.
-\@ref(fig:RedFun).
-
 \newpage
 
-# **Figure A1**: Bayesian process sheme
+# **Figure S1**: Bayesian process sheme
 
@@
 
-# **Figure A2**: Redundancy metric
+# **Figure S2**: Proportion of seed mass classes along time
 
-```{r RedMM,out.width='100%',fig.cap="Scheme of the redundancy measurement process",echo=F, fig.env="figure*"}
-knitr::include_graphics("ExternalFig/RedundancyMetric.png")
-```
-
-# **Figure A3**: Proportion of seed mass classes along time
-
-```{r FigSeedMassClasses, echo=FALSE, out.width='80%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-No gap filling process was applied in this case.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
+```{r FigSeedMassClasses, echo=FALSE, out.width='80%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions over 30 years after disturbance.
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).", fig.env="figure*"}
 load("DB/Smass_ClassProp")
@@
 ```
-
-# **Figure A4**: Rao diversity trajectories with all plots display
-
-```{r DivFun, out.width = '60%', echo=FALSE, fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.", fig.env="figure*"}
-load("DB/FunctionalTraj_ForGraphs")
-
-FunTraj(CompleteFun,remove=F)
-
-```
-
-# **Figure A5**: Simpson and Rao Linera regrassion summary
-
-```{r LinReg, echo=FALSE,is.as=T}
-LmSummary <- data.frame(
-  Diversity=c(rep("Simpson",2), rep("Rao",2)),
-  Metrics=rep(c("R^2^","p~value~"),times=2),
-  Time1=c("-0.08","0.7",
-            "0.58","4e-3"),
-  Time2=c("0.25","6e-2",
-            "0.58","4e-3"),
-  Time3=c("0.24","6e-2",
-            "0.58","4e-3")
-)
-names(LmSummary) <- c("Diversity","Metrics","10 years", "20 years", "30 years")
-knitr::kable(LmSummary, "html", caption="Linear models information", longtable = FALSE, booktabs = TRUE, escape = FALSE) %>%
-  collapse_rows(columns=1) %>%
-  kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
-```
-
-# **Figure A6**: Functional PCA
-
-```{r PCA_graphs, out.width = '100%', echo=FALSE, fig.cap="Summarizing plots of the functional PCA analysis defining communities functional space.
-(ref:Bolda) Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
-(ref:Boldb) bow plot of the PCA eignevalues in percentage.
-(ref:Boldc) Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).", fig.env="figure*"}
-load("DB/FunctionalPCA")
-
-plotPCA(DataACP)
-
-```
-
-# **Figure A7**: Overall functional redundancy
-
-```{r RedFun, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
-load("DB/Redundancy")
-
-par(mar=c(5,4,5,2))
-RedundancyPlot(RedundancyTraj)
-mtext("Functional Redundancy",side=3,adj=0,line=0.5)
-mtext("Years since disturbance",side=1,line=2.2,adj=1)
-
-```

### 2018-07-11 WholePlotTrajectories_SuppMat.tex
biblio last reread

--- 2018-05-14 AppendixI.tex
@@
 
-\PaperTitle{Appendix I} % Article title
+\PaperTitle{Supplementary materials} % Article title
 
@@
 
-\textbf{Figure A1}: Whenever the family or genus were recorded for a
+\textbf{Figure S1}: Whenever the family or genus were recorded for a
 vernacular name, the pool of species sampled to replace the vernacular
@@
 
-\textbf{Figure A2}: Trajectories of seed mass classes proportions over
+\textbf{Figure S2}: Scheme of the redundancy measurement process
+
+\textbf{Figure S3}: Trajectories of seed mass classes proportions over
 30 years after disturbance, corresponding to the median (solid line) and
@@
 
-\textbf{Figure A3}: Trajectories of the Rao functional diversity over 30
-years after disturbance. Trajectories correspond to the median (solid
-line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
-iteration of the taxonomic uncertainty propagation. Initial treatments
-are represented by solid lines colorswith green for control, blue for
-T1,orange for T2 and red for T3.and the missing trait value filling
-processes. \ref{fig:DivFunAppendix}.
-
-\textbf{Figure A4}: Trajectories of the Rao functional diversity over 30
-years after disturbance. Trajectories correspond to the median (solid
-line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
-iteration of the taxonomic uncertainty propagation. Initial treatments
-are represented by solid lines colorswith green for control, blue for
-T1,orange for T2 and red for T3.and the missing trait value filling
-processes. @ref(fig:PCA\_graphs).
-
-\textbf{Figure A5}: Trajectories of the Rao functional diversity over 30
-years after disturbance. Trajectories correspond to the median (solid
-line) and 0.025 and 0.975 percentile (gray envelope) observed after 50
-iteration of the taxonomic uncertainty propagation. Initial treatments
-are represented by solid lines colorswith green for control, blue for
-T1,orange for T2 and red for T3.and the missing trait value filling
-processes. \ref{fig:RedFun}.
-
 \newpage
 
-\section{\texorpdfstring{\textbf{Figure A1}: Bayesian process
-sheme}{Figure A1: Bayesian process sheme}}\label{figure-a1-bayesian-process-sheme}
+\section{\texorpdfstring{\textbf{Figure S1}: Bayesian process
+sheme}{Figure S1: Bayesian process sheme}}\label{figure-s1-bayesian-process-sheme}
 
@@
 
-\section{\texorpdfstring{\textbf{Figure A2}: Proportion of seed mass
-classes along
-time}{Figure A2: Proportion of seed mass classes along time}}\label{figure-a2-proportion-of-seed-mass-classes-along-time}
+\section{\texorpdfstring{\textbf{Figure S2}: Redundancy
+metric}{Figure S2: Redundancy metric}}\label{figure-s2-redundancy-metric}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{AppendixI_files/figure-latex/FigSeedMassClasses-1}
+{\centering \includegraphics[width=1\linewidth]{ExternalFig/RedundancyMetric}
 
@@
 
-\caption{Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-No gap filling process was applied in this case.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:FigSeedMassClasses}
+\caption{Scheme of the redundancy measurement process}\label{fig:RedMM}
 \end{figure*}
 
-\section{\texorpdfstring{\textbf{Figure A3}: Rao diversity trajectories
-with all plots
-display}{Figure A3: Rao diversity trajectories with all plots display}}\label{figure-a3-rao-diversity-trajectories-with-all-plots-display}
+\section{\texorpdfstring{\textbf{Figure S3}: Proportion of seed mass
+classes along
+time}{Figure S3: Proportion of seed mass classes along time}}\label{figure-s3-proportion-of-seed-mass-classes-along-time}
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{AppendixI_files/figure-latex/DivFunAppendix-1}
+{\centering \includegraphics[width=0.8\linewidth]{WholePlotTrajectories_SuppMat_files/figure-latex/FigSeedMassClasses-1}
 
@@
 
-\caption{Trajectories of the Rao functional diversity over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
-Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.and the missing trait value filling processes.}\label{fig:DivFunAppendix}
-\end{figure*}
-
-\section{\texorpdfstring{\textbf{Figure A4}: Functional
-PCA}{Figure A4: Functional PCA}}\label{figure-a4-functional-pca}
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{AppendixI_files/figure-latex/PCA_graphs-1}
-
-}
-
-\caption{Summarizing plots of the functional PCA analysis defining communities functional space.
-\textbf{(a)} Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
-\textbf{(b)} bow plot of the PCA eignevalues in percentage.
-\textbf{(c)} Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).}(\#fig:PCA_graphs)
-\end{figure*}
-
-\section{\texorpdfstring{\textbf{Figure A5}: Overall functional
-redundancy}{Figure A5: Overall functional redundancy}}\label{figure-a5-overall-functional-redundancy}
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{AppendixI_files/figure-latex/RedFun-1}
-
-}
-
-\caption{Trajectories of the functional redundancy over 30 years after disturbance.
-Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation and the missing trait value filling processes.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.}\label{fig:RedFun}
+\caption{Trajectories of seed mass classes proportions over 30 years after disturbance. Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).}\label{fig:FigSeedMassClasses}
 \end{figure*}

### 2018-07-11 WholePlotTrajectories_SuppMat.Rmd
Supp Mat

@@
 
-# Appendix figures
-
 The following information are available for this article:
@@
 No gap filling process was applied in this case.
-Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3 \@ref(fig:FigSeedMassClasses).
-
-\newpage
-
-# **Figure S1**: Bayesian process sheme
+Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3 .
 
@@
 ```
-
-# **Figure S2**: Proportion of seed mass classes along time
 

### 2018-07-11 WholePlotTrajectories_SuppMat.tex
Supp Mat

@@
 
-\section{Appendix figures}\label{appendix-figures}
-
 The following information are available for this article:
@@
 
-\textbf{Figure S2}: Scheme of the redundancy measurement process
-
-\textbf{Figure S3}: Trajectories of seed mass classes proportions over
+\textbf{Figure S2}: Trajectories of seed mass classes proportions over
 30 years after disturbance, corresponding to the median (solid line) and
@@
 colors with green for control, blue for T1,orange for T2 and red for T3
-\ref{fig:FigSeedMassClasses}.
-
-\newpage
-
-\section{\texorpdfstring{\textbf{Figure S1}: Bayesian process
-sheme}{Figure S1: Bayesian process sheme}}\label{figure-s1-bayesian-process-sheme}
+.
 
@@
 \end{figure*}
-
-\section{\texorpdfstring{\textbf{Figure S2}: Redundancy
-metric}{Figure S2: Redundancy metric}}\label{figure-s2-redundancy-metric}
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{ExternalFig/RedundancyMetric}
-
-}
-
-\caption{Scheme of the redundancy measurement process}\label{fig:RedMM}
-\end{figure*}
-
-\section{\texorpdfstring{\textbf{Figure S3}: Proportion of seed mass
-classes along
-time}{Figure S3: Proportion of seed mass classes along time}}\label{figure-s3-proportion-of-seed-mass-classes-along-time}
 

### 2018-07-11 WholePlotTrajectories.Rmd
ref

@@
 
-## abstract: 204 words / text without legends : 3659 words
+## abstract: 204 words / text without legends : 3659 words / references: 68
 
@@
 The long-term impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008].
-Besides the long-term recovery of infrequent species increases the risks to loose cornerstone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
+Besides the long-term recovery of infrequent species increases the risks to loose cornerstone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
 Apart from the functional characteristics considered here, infrequent species might indeed have unique functions in the ecosystem or be a key for some fauna.

### 2018-07-11 WholePlotTrajectories.Rmd
graphs pimp

@@
 ```{r Options, echo=FALSE, message=FALSE, warning=F}
-knitr::opts_chunk$set(cache=F, warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
+knitr::opts_chunk$set(cache=F, echo=F,warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
 # Installation des packages si nécessaire et chargement
@@
 mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=0.8)
-legend("right",inset=c(-0.35,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
+legend("right",inset=c(-0.345,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
 
@@
 plotDiv(CompleteRichFun,remove=TRUE)
-mtext("(c) Functional Richness",side=3,adj=0,line=1,cex=0.75)
+mtext("(c) Functional Richness",side=3,adj=0,line=1,cex=0.9)
 plotDiv(CompleteFun,remove=TRUE)
-mtext("(d) Functional Evenness",side=3,adj=0,line=1,cex=0.75)
+mtext("(d) Functional Evenness",side=3,adj=0,line=1,cex=0.9)
 mtext("Years since disturbance",side=1,line=2.2,adj=1)
@@
 
-```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/3,fig.cap="Relationship between the initial \\%AGB loss and community taxonmic richness (ref:Bolda), taxonomic evenness (ref:Boldb), functional richness (ref:Boldc),and functional evenness (ref:Boldd) at 10, 20 and 30 years after disturbance.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
+```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/4,fig.cap="Relationship between the initial \\%AGB loss and community taxonmic richness (ref:Bolda), taxonomic evenness (ref:Boldb), functional richness (ref:Boldc),and functional evenness (ref:Boldd) at 10, 20 and 30 years after disturbance", fig.env="figure*"}
 load("DB/RichnessIDH");load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB");load("DB/FunRichIDH")
 
-par(mfrow=c(1,4),mar=c(5,2,4,2),oma=c(1,1.5,1,2),no.readonly=TRUE)
+par(mfrow=c(1,4),mar=c(2,2.5,3,1),oma=c(1,1,1,2.5),no.readonly=TRUE)
 plotIDH(Data=Rich,AgbLoss=AGBloss)
-mtext("(a) Taxonomic Richness",line=1,side=3,adj=0,cex=0.75)
+mtext("(a) Taxonomic\nRichness",line=1,side=3,adj=0,cex=0.85)
+mtext("Equivalent diversity",side=2,line=2.2,adj=1,cex=0.8)
 plotIDH(Data=Sim,AgbLoss=AGBloss)
-mtext("(b) Taxonomic Evenness",line=1,side=3,adj=0,cex=0.75)
+mtext("(b) Taxonomic\nEvenness",line=1,side=3,adj=0,cex=0.85)
 plotIDH(Data=Rao,AgbLoss=AGBloss)
-mtext("(c) Functional richness",line=1,side=3,adj=0,cex=0.75)
+mtext("(c) Functional\nRichness",line=1,side=3,adj=0,cex=0.85)
 plotIDH(Data=FunRich,AgbLoss=AGBloss)
-mtext("(d) Functional evenness",line=1,side=3,adj=0,cex=0.75)
-mtext("initial AGB loss (%)",side=1,line=2.2,adj=1)
-legend("right",inset=c(-0.37,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
+mtext("(d) Functional\nEvenness",line=1,side=3,adj=0,cex=0.85)
+mtext("initial AGB loss (%)",side=1,line=2.2,adj=1,cex=0.8)
+legend("right",inset=c(-0.5,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
 
@@
 
-All disturbed plots had lower functional redundancy than control plots and followed similar hump-shaped trajectories (\@ref(fig:RedFun_rest)).
+All disturbed plots had lower functional redundancy than control plots and followed similar hump-shaped trajectories (\@ref(fig:RedFunRest)).
 The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{spearman}=0.47$) and the initial value had not recovered for any disturbed communities after 30 years.
 
-```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
+```{r RedFunRest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
 Colors are disturbance treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
@@
 RedundancyPlot(RedundancyTraj_restricted)
-mtext("Functional Redundancy",side=3,adj=0,line=1.5,cex=1.2)
-mtext("Restricted to initial functional space",side=3,adj=0,cex=0.9,line=0.5)
-mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=1.2)
+mtext("Functional Redundancy",side=3,adj=0,line=1.5,cex=1.5)
+mtext("Restricted to initial functional space",side=3,adj=0,cex=1.2,line=0.5)
+mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=1.5)
 

### 2018-07-11 WholePlotTrajectories.tex
graphs pimp

@@
 
-\caption{Relationship between the initial \%AGB loss and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)}, functional richness \textbf{(c)},and functional evenness \textbf{(d)} at 10, 20 and 30 years after disturbance.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:IDHplot}
+\caption{Relationship between the initial \%AGB loss and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)}, functional richness \textbf{(c)},and functional evenness \textbf{(d)} at 10, 20 and 30 years after disturbance}\label{fig:IDHplot}
 \end{figure*}
@@
 All disturbed plots had lower functional redundancy than control plots
-and followed similar hump-shaped trajectories (@ref(fig:RedFun\_rest)).
+and followed similar hump-shaped trajectories (\ref{fig:RedFunRest}).
 The maximum redundancy loss was positively correlated with the
@@
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/RedFun_rest-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/RedFunRest-1}
 
@@
 \caption{Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
-Colors are disturbance treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}(\#fig:RedFun_rest)
+Colors are disturbance treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:RedFunRest}
 \end{figure}
@@
 cornerstone species, with unexpected ecological consequences
-\citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}. Apart from the
-functional characteristics considered here, infrequent species might
-indeed have unique functions in the ecosystem or be a key for some
-fauna.
+\citep{Jones1994, Chazdon2003a, Diaz2005}. Apart from the functional
+characteristics considered here, infrequent species might indeed have
+unique functions in the ecosystem or be a key for some fauna.
 

### 2018-07-11 Scripts/DrawingTools.R

@@ -101,7 +101,7 @@ TaxoTraj<-function(CompTaxo){
          xlab="",ylab="",ylim=c(min(unlist(Toplot),na.rm=T),max(unlist(Toplot),na.rm=T)))
     axis(1,at=as.character(seq(5,33,5)),labels=TRUE)  
     mtext(paste("(",c("a","b","b")[q],") ",c("Taxonomic Richness","Shannon","Taxonomic Evenness")[q],sep=""),
-          line=1,side=3,adj=0,cex=0.8)
+          line=1,side=3,adj=0,cex=0.9)
     
     invisible(lapply(1:4,function(t){  
       toplot<-Toplot[[t]]

### 2018-07-15 WholePlotTrajectories.Rmd
relecture

@@
 These structural parameters have been sucessfully modeled, giving important insights into the recovery of ecosystem processes and services [@Herault2018]. 
-However the response of forests diversity and composition is still unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014; @Liang2016].
+However the response of forests diversity and composition remains unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014; @Liang2016].
 In the short-term, moderate disturbance may lead to positive impacts on communities diversity, an idea formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity when disturbance intensity is not too high [@Molino2001;@Kariuki2006a].

### 2018-07-23 WholePlotTrajectories.Rmd
annex

@@
 The uncertainty propagation was done through a Bayesian framework reconstituting complete inventories at genus level from real incomplete ones on the basis of vernacular/botanical names association. 
-Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,..., \alpha_3\big]$ observed across all inventories between each vernacular name *v* and all species $\big[s_1, s_2,..., s_N\big]$:
+Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,..., \alpha_V\big]$ observed across all inventories between each vernacular name *v* and all species $\big[s_1, s_2,..., s_N\big]$:
 
 \begin{align}
-M_v\Big(\big[s_1, s_2,..., s_N\big],\big[\alpha_1, \alpha_2,..., \alpha_3\big]\Big) \nonumber
+M_v\Big(\big[s_1, s_2,..., s_N\big],\big[\alpha_1, \alpha_2,..., \alpha_V\big]\Big) \nonumber
 \end{align}
@@
 
-While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-disturbance time.
+While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-\break disturbance time.
 In disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WSG to high average SLA and chlorophyll content (see appendix I).
@@
 The long-term impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008].
-Besides the long-term recovery of infrequent species increases the risks to loose cornerstone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
+Besides, the long-term recovery of infrequent species increases the risks to loose cornerstone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
 Apart from the functional characteristics considered here, infrequent species might indeed have unique functions in the ecosystem or be a key for some fauna.
@@
 Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Gourlet-Fleury2005].
-As the trajectories highlighted the recruitment processes proved central for communities response to disturbance and closer focus on demographical drivers of communities response would clarify the fate of the future forests.
 

### 2018-07-23 WholePlotTrajectories_SuppMat.Rmd
annex

@@
 
+The biodiversity estimator is based on the association probability between vernacular names and botanical species. 
+The association probability is estimated by a Dirichlet trial based on the vernacular/botanical association frequency observed in the whole inventory.
+This association probability then allows generating theoretical inventories with complete botanical identification with the assignment of botanical species to trees with only a vernacular name.
+The repetition of the replacement process produces a set of theoretical inventories which diversities gives the probability distribution of the initial inventory diversity.
+
 ```{r FigSeedMassClasses, echo=FALSE, out.width='80%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions over 30 years after disturbance.

### 2018-07-23 WholePlotTrajectories_SuppMat.tex
annex

@@
 
+The biodiversity estimator is based on the association probability
+between vernacular names and botanical species. The association
+probability is estimated by a Dirichlet trial based on the
+vernacular/botanical association frequency observed in the whole
+inventory. This association probability then allows generating
+theoretical inventories with complete botanical identification with the
+assignment of botanical species to trees with only a vernacular name.
+The repetition of the replacement process produces a set of theoretical
+inventories which diversities gives the probability distribution of the
+initial inventory diversity.
+
 \begin{figure*}

### 2018-07-23 Scripts/DrawingTools.R

@@ -136,7 +136,7 @@ plotIDH<-function(Data,AgbLoss){
       abline(a=Lm$coefficients[1],b=Lm$coefficients[2],col=colyear[Ti],lwd=2.5)
       return(round(summary(Lm)$adj.r.squared,2)) }
   })
-  legend("topleft",legend=unlist(leg),bty="n",lty=1,col=colyear,lwd=2.5,cex=0.8,title=expression(paste('R'^2,'adjusted')))
+  legend("topleft",legend=unlist(leg),bty="n",lty=1,col=colyear,lwd=2.5,cex=0.8,title=expression(paste('adjusted','R'^2)))
 }
 
 plotDiv<-function(Data,remove=FALSE){

### 2018-07-23 WholePlotTrajectories.Rmd
coquilles

@@
 Besides, the long-term recovery of infrequent species increases the risks to loose cornerstone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
-Apart from the functional characteristics considered here, infrequent species might indeed have unique functions in the ecosystem or be a key for some fauna.
+Apart from the functional characteristics considered here, infrequent species might indeed have unique functions in the ecosystem or be a key for some fauna [@Schleuning2016].
 

### 2018-07-23 WholePlotTrajectories.tex
coquilles

@@
 services \citep{Herault2018}. However the response of forests diversity
-and composition is still unclear, albeit it determines the productivity,
+and composition remains unclear, albeit it determines the productivity,
 stability and functioning of ecosystems \citep{Tilman2014, Liang2016}.
@@
 replaced through multinomial trials based on the association probability
-\(\big[\alpha_1, \alpha_2,..., \alpha_3\big]\) observed across all
+\(\big[\alpha_1, \alpha_2,..., \alpha_V\big]\) observed across all
 inventories between each vernacular name \emph{v} and all species
@@
 \begin{align}
-M_v\Big(\big[s_1, s_2,..., s_N\big],\big[\alpha_1, \alpha_2,..., \alpha_3\big]\Big) \nonumber
+M_v\Big(\big[s_1, s_2,..., s_N\big],\big[\alpha_1, \alpha_2,..., \alpha_V\big]\Big) \nonumber
 \end{align}
@@
 undisturbed communities (Figure \ref{fig:NMDSplans}), they followed
-marked and consistent trajectories over post-disturbance time. In
+marked and consistent trajectories over post-\break disturbance time. In
 disturbed communities, these compositional changes corresponded to
@@
 chances to see the persistence of disturbance-specific species at the
-expense of late-successional ones \citep{Haddad2008}. Besides the
+expense of late-successional ones \citep{Haddad2008}. Besides, the
 long-term recovery of infrequent species increases the risks to loose
@@
 characteristics considered here, infrequent species might indeed have
-unique functions in the ecosystem or be a key for some fauna.
+unique functions in the ecosystem or be a key for some fauna
+\citep{Schleuning2016}.
 
@@
 species loss and the persistence of disturbance-specific communities
-\citep{Gourlet-Fleury2005}. As the trajectories highlighted the
-recruitment processes proved central for communities response to
-disturbance and closer focus on demographical drivers of communities
-response would clarify the fate of the future forests.
+\citep{Gourlet-Fleury2005}.
 

### 2018-07-23 Scripts/DrawingTools.R

@@ -136,7 +136,7 @@ plotIDH<-function(Data,AgbLoss){
       abline(a=Lm$coefficients[1],b=Lm$coefficients[2],col=colyear[Ti],lwd=2.5)
       return(round(summary(Lm)$adj.r.squared,2)) }
   })
-  legend("topleft",legend=unlist(leg),bty="n",lty=1,col=colyear,lwd=2.5,cex=0.8,title=expression(paste('adjusted','R'^2)))
+  legend("topleft",legend=unlist(leg),bty="n",lty=1,col=colyear,lwd=2.5,cex=0.8,title=expression(paste('Adjusted ','R'^2)))
 }
 
 plotDiv<-function(Data,remove=FALSE){

### 2018-07-30 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

@@
 remains debated and the controversy questions the extent of communities'
-resilience regarding their functional and taxonomic facets. To
+resilience regarding their taxonomic and functional facets. To
 disentangle the ecological processes driving community response to
-disturbance, we analysed the diversity trajectories over 30 years
-following a disturbance gradient. Specifically, we examined community
+disturbance, we analysed the tree community trajectories over 30 years
+following a disturbance gradient in a neotropical forest. Specifically, we examined community
 functional and taxonomic trajectories with regards to diversity,
@@
 communities taxonomic diversity response while functional diversity was
-enhanced whenever the disturbance intensity. Although consistent, the
+enhanced whatever the disturbance intensity. Although consistent, the
 recovery of community composition, diversity and redundancy remained
@@
 cycles with no disturbance to ensure a complete recovery, and questioned
-tropical forest community resilience after repeated disturbance.
+tropical forest community resilience after repeated disturbances.
 }
@@
 and the biotic interactions such as competition among species
-\citep{Chesson2000, Herault2018}. One of the cornerstone of tropical
+\citep{Chesson2000}. One of the cornerstone of tropical
 forest ecology is to understand the processes and drivers of ecosystems
@@
 ecological-meaningful analysis would couple richness with (i) evenness
-that would reveal underlying ecological processes and (ii) composition
+that would reveal the changes in the species abundance distribution and thus the underlying ecological processes and (ii) composition
 that is crucial for conservation issues
@@
 Specifically, we (i) elucidated community taxonomic and functional
-recovery and the determinism of underlying processes, (ii) clarified the
+recovery and the underlying ecological processes, (ii) clarified the
 validity of the IDH in the long term for tropical forest and its

### 2018-07-30 WholePlotTrajectories.tex
Update WholePlotTrajectories.tex

@@
 
-The taxonomic differences among communities, marked before disturbance
+The taxonomic differences among local communities, marked before disturbance
 by the distinct starting points on the NMDS axis 2, were maintained
@@
 Conversely, disturbed communities followed functional trajectories
-similar in terms of functional composition. As pre-disturbance surviving
+that are highly similar in terms of functional composition. As pre-disturbance surviving
 trees mirror the initial community \citep{Herault2018}, changes in
@@
 Competitive pioneers became dominant in filling the environmental niches
-of high availability of light, space and nutrient vacated by the
+of high availability of light, space and nutrients vacated by the
 disturbance. The recruitment of pioneers changed community functional
@@
 
-The IDH predicted well the disturbance impact on community taxonomic
+The IDH well predicted the disturbance impact on community taxonomic
 richness, enhanced until an intensity threshold (20-25\% AGB loss), and
@@
 evenness rather represented the increasing dominance of pionneers that
-balanced the usual hyper-dominance of tropical forest below the
+balanced the usual hyper-dominance of a few species in tropical forests below the
 intensity threshold, thus increasing community overall evenness up to
@@
 functional redundancy, in some communities just initiated after 30
-years, was difficult to estimate \citep{Elmqvist2003, Diaz2005}.
+years, was extremely difficult to estimate \citep{Elmqvist2003, Diaz2005}.
 
@@
 species loss and the persistence of disturbance-specific communities
-\citep{Gourlet-Fleury2005}.
+\citep{Herault2018}.
 

### 2018-07-30 WholePlotTrajectories.Rmd
details
Bubs correction

@@
     The huge diversity of tropical forests is often assumed to be regularly reshaped by natural disturbance yielding a diversity peak at intermediate intensity.
-This intermediate disturbance hypothesis (IDH), though, remains debated and the controversy questions the extent of communities' resilience regarding their functional and taxonomic facets.
+This intermediate disturbance hypothesis (IDH), though, remains debated and the controversy questions the extent of communities' resilience regarding their taxonomic and functional facets.
 
-    To disentangle the ecological processes driving community response to disturbance, we analysed the diversity trajectories over 30 years following a disturbance gradient.
+    To disentangle the ecological processes driving community response to disturbance, we analysed the tree community trajectories over 30 years following a disturbance gradient in a Neotropical forest.
     Specifically, we examined community functional and taxonomic trajectories with regards to diversity, composition and redundancy.
@@
 While pre-disturbance taxonomic differences were maintained over time, functional composition trajectories were quite similar among communities.
-    The IDH did predict communities taxonomic diversity response while functional diversity was enhanced whenever the disturbance intensity.
+    The IDH did predict communities taxonomic diversity response while functional diversity was enhanced whatever the disturbance intensity.
     Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years.
-This acknowledged the need of decades-long cycles with no disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbance.
+This acknowledged the need of decades-long cycles with no disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbances.
 
@@
 
-In tropical forests, ecological communities are regularly re-shaped by natural disturbance events changing both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions such as competition among species [@Chesson2000;@Herault2018].
+In tropical forests, ecological communities are regularly re-shaped by natural disturbance events changing both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions such as competition among species [@Chesson2000].
 One of the cornerstone of tropical forest ecology is to understand the processes and drivers of ecosystems response to disturbance [@Chazdon2003a]. 
@@
 Validations of the IDH though remain scarce in the long-term and mainly rely on the analysis of taxonomic richness [@Molino2001].
-Taxonomic richness alone, though, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016].
-More ecological-meaningful analysis would couple richness with (i) evenness that would reveal underlying ecological processes and (ii) composition that is crucial for conservation issues [@Lavorel2002;@Bellwood2006].
+Taxonomic richness alone, however, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016].
+More ecological-meaningful analysis would couple richness with (i) evenness, that would reveal the changes in the species abundance distribution and thus the underlying ecological processes, and (ii) composition that is crucial for conservation issues [@Lavorel2002;@Bellwood2006].
 Furthermore, a functional approach accounting for species biological attributes would directly link communities diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Baraloto2012a].
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, we (i) elucidated community taxonomic and functional recovery and the determinism of underlying processes, (ii) clarified the validity of the IDH in the long term for tropical forest and its translation into different trajectories in time, and (iii) questioned community recovery time.
+Specifically, we (i) elucidated community taxonomic and functional recovery and the underlying ecological processes, (ii) clarified the validity of the IDH in the long term for tropical forest and its translation into different trajectories in time, and (iii) questioned community recovery time.
 
@@
 
-The taxonomic differences among communities, marked before disturbance by the distinct starting points on the NMDS axis 2, were maintained throughout recovery trajectories. 
+The taxonomic differences among local communities, marked before disturbance by the distinct starting points on the NMDS axis 2, were maintained throughout recovery trajectories. 
 More than commonly thought, post-disturbance trajectories depended on community initial composition, that partly determined the pool of recruited species and constrained the trajectories towards the initial composition.
@@
 
-Conversely, disturbed communities followed functional trajectories similar in terms of functional composition.
+Conversely, disturbed communities followed functional trajectories that are highly similar in terms of functional composition.
 As pre-disturbance surviving trees mirror the initial community [@Herault2018], changes in functional composition relied upon the recruitment of species or functional types that were infrequent or absent before disturbance.
-Competitive pioneers became dominant in filling the environmental niches of high availability of light, space and nutrient vacated by the disturbance.
+Competitive pioneers became dominant in filling the environmental niches of high availability of light, space and nutrients vacated by the disturbance.
 The recruitment of pioneers changed community functional composition in the same way for all disturbance intensity towards more resource-acquisitive strategies, moving community functional composition right along the first axis in Figure \@ref(fig:NMDSplans) [@Westoby1998; @Wright2004; @Reich2014].
@@
 
-The IDH predicted well the disturbance impact on community taxonomic richness, enhanced until an intensity threshold (20-25% AGB loss), and to some extent on taxonomic evenness, somewhat decoupled from the disturbance intensity as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+The IDH well predicted well the disturbance impact on community taxonomic richness, enhanced until an intensity threshold (20-25% AGB loss), and to some extent on taxonomic evenness, somewhat decoupled from the disturbance intensity as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
 The disturbance intensity determined the balance in the community between pre-disturbance surviving trees and those recruited afterward.
@@
 For community taxonomic evenness the disturbance impact was similar but slighter, as the evenness is less sensitive to the loss of rare species.
-Taxonomic evenness rather represented the increasing dominance of pionneers that balanced the usual hyper-dominance of tropical forest below the intensity threshold, thus increasing community overall evenness up to the intensity threshold beyond which pioneers became in turn highly dominant and decreased the overall evenness [@Baraloto2012a].
+Taxonomic evenness rather represented the increasing dominance of pionneers that balanced the usual hyper-dominance of a few species in tropical forests below the intensity threshold, thus increasing community overall evenness up to the intensity threshold beyond which pioneers became in turn highly dominant and decreased the overall evenness [@Baraloto2012a].
 
@@
 This replacement was stochastic and followed the lottery recruitment rules, implying a recruitment eased for the first recruited species but then increasingly hampered by the emergence of interspecific competition [@Busing2002].
-Along time the recovery of infrequent species was increasingly difficult, so that the time for the full recovery of the functional redundancy, in some communities just initiated after 30 years, was difficult to estimate [@Elmqvist2003;@Diaz2005].
+Along time the recovery of infrequent species was increasingly slow, so that the time for the full recovery of the functional redundancy, in some communities just initiated after 30 years, was extremely difficult to estimate [@Elmqvist2003;@Diaz2005].
 
@@
 The resilience of tropical forests proved consistent although several decades-long. 
-Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Gourlet-Fleury2005].
+Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].
 

### 2018-07-30 WholePlotTrajectories.tex
Final version 1

@@
 disturbance, we analysed the tree community trajectories over 30 years
-following a disturbance gradient in a neotropical forest. Specifically, we examined community
-functional and taxonomic trajectories with regards to diversity,
-composition and redundancy. Functional trajectories were drawn based on
-7 leaf, stem and life-history traits. We highlighted the cyclic recovery
-of community taxonomic and functional composition. While pre-disturbance
-taxonomic differences were maintained over time, functional composition
-trajectories were quite similar among communities. The IDH did predict
-communities taxonomic diversity response while functional diversity was
-enhanced whatever the disturbance intensity. Although consistent, the
-recovery of community composition, diversity and redundancy remained
-unachieved after 30 years. This acknowledged the need of decades-long
-cycles with no disturbance to ensure a complete recovery, and questioned
-tropical forest community resilience after repeated disturbances.
+following a disturbance gradient in a Neotropical forest. Specifically,
+we examined community functional and taxonomic trajectories with regards
+to diversity, composition and redundancy. Functional trajectories were
+drawn based on 7 leaf, stem and life-history traits. We highlighted the
+cyclic recovery of community taxonomic and functional composition. While
+pre-disturbance taxonomic differences were maintained over time,
+functional composition trajectories were quite similar among
+communities. The IDH did predict communities taxonomic diversity
+response while functional diversity was enhanced whatever the
+disturbance intensity. Although consistent, the recovery of community
+composition, diversity and redundancy remained unachieved after 30
+years. This acknowledged the need of decades-long cycles with no
+disturbance to ensure a complete recovery, and questioned tropical
+forest community resilience after repeated disturbances.
 }
@@
 and the biotic interactions such as competition among species
-\citep{Chesson2000}. One of the cornerstone of tropical
-forest ecology is to understand the processes and drivers of ecosystems
-response to disturbance \citep{Chazdon2003a}. For now, this has been
-largely studied through forest structural parameters such as aboveground
-biomass, tree height or stem density
-\citep{Piponiot2016, Rutishauser2016} that are rapid and convenient to
-measure. These structural parameters have been sucessfully modeled,
-giving important insights into the recovery of ecosystem processes and
-services \citep{Herault2018}. However the response of forests diversity
-and composition remains unclear, albeit it determines the productivity,
-stability and functioning of ecosystems \citep{Tilman2014, Liang2016}.
-In the short-term, moderate disturbance may lead to positive impacts on
-communities diversity, an idea formalized by the intermediate
-disturbance hypothesis (IDH) stating a maximized species diversity when
-disturbance intensity is not too high \citep{Molino2001, Kariuki2006a}.
+\citep{Chesson2000}. One of the cornerstone of tropical forest ecology
+is to understand the processes and drivers of ecosystems response to
+disturbance \citep{Chazdon2003a}. For now, this has been largely studied
+through forest structural parameters such as aboveground biomass, tree
+height or stem density \citep{Piponiot2016, Rutishauser2016} that are
+rapid and convenient to measure. These structural parameters have been
+sucessfully modeled, giving important insights into the recovery of
+ecosystem processes and services \citep{Herault2018}. However the
+response of forests diversity and composition remains unclear, albeit it
+determines the productivity, stability and functioning of ecosystems
+\citep{Tilman2014, Liang2016}. In the short-term, moderate disturbance
+may lead to positive impacts on communities diversity, an idea
+formalized by the intermediate disturbance hypothesis (IDH) stating a
+maximized species diversity when disturbance intensity is not too high
+\citep{Molino2001, Kariuki2006a}.
 
@@
 rely on the analysis of taxonomic richness \citep{Molino2001}. Taxonomic
-richness alone, though, gives limited or misleading information on
+richness alone, however, gives limited or misleading information on
 forests recovery and functioning \citep{Chaudhary2016}. More
-ecological-meaningful analysis would couple richness with (i) evenness
-that would reveal the changes in the species abundance distribution and thus the underlying ecological processes and (ii) composition
-that is crucial for conservation issues
-\citep{Lavorel2002, Bellwood2006}. Furthermore, a functional approach
-accounting for species biological attributes would directly link
-communities diversity, composition and redundancy to ecosystem
-functioning and to its environmental constraints
+ecological-meaningful analysis would couple richness with (i) evenness,
+that would reveal the changes in the species abundance distribution and
+thus the underlying ecological processes, and (ii) composition that is
+crucial for conservation issues \citep{Lavorel2002, Bellwood2006}.
+Furthermore, a functional approach accounting for species biological
+attributes would directly link communities diversity, composition and
+redundancy to ecosystem functioning and to its environmental constraints
 \citep{Violle2007b, Baraloto2012a}. In that respect, the functional
@@
 
-The taxonomic differences among local communities, marked before disturbance
-by the distinct starting points on the NMDS axis 2, were maintained
-throughout recovery trajectories. More than commonly thought,
+The taxonomic differences among local communities, marked before
+disturbance by the distinct starting points on the NMDS axis 2, were
+maintained throughout recovery trajectories. More than commonly thought,
 post-disturbance trajectories depended on community initial composition,
@@
 
-Conversely, disturbed communities followed functional trajectories
-that are highly similar in terms of functional composition. As pre-disturbance surviving
-trees mirror the initial community \citep{Herault2018}, changes in
-functional composition relied upon the recruitment of species or
-functional types that were infrequent or absent before disturbance.
-Competitive pioneers became dominant in filling the environmental niches
-of high availability of light, space and nutrients vacated by the
-disturbance. The recruitment of pioneers changed community functional
-composition in the same way for all disturbance intensity towards more
-resource-acquisitive strategies, moving community functional composition
-right along the first axis in Figure \ref{fig:NMDSplans}
-\citep{Westoby1998, Wright2004, Reich2014}. Thereafter long-lived, more
-resistant and shade-tolerant species excluded the first established
-pioneers and started the recovery of pre-disturbance functional
-composition, moving similarly community functional composition left
-along the first axis and upward along the second axis in Figure
-\ref{fig:NMDSplans}.
+Conversely, disturbed communities followed functional trajectories that
+are highly similar in terms of functional composition. As
+pre-disturbance surviving trees mirror the initial community
+\citep{Herault2018}, changes in functional composition relied upon the
+recruitment of species or functional types that were infrequent or
+absent before disturbance. Competitive pioneers became dominant in
+filling the environmental niches of high availability of light, space
+and nutrients vacated by the disturbance. The recruitment of pioneers
+changed community functional composition in the same way for all
+disturbance intensity towards more resource-acquisitive strategies,
+moving community functional composition right along the first axis in
+Figure \ref{fig:NMDSplans} \citep{Westoby1998, Wright2004, Reich2014}.
+Thereafter long-lived, more resistant and shade-tolerant species
+excluded the first established pioneers and started the recovery of
+pre-disturbance functional composition, moving similarly community
+functional composition left along the first axis and upward along the
+second axis in Figure \ref{fig:NMDSplans}.
 
@@
 
-The IDH well predicted the disturbance impact on community taxonomic
-richness, enhanced until an intensity threshold (20-25\% AGB loss), and
-to some extent on taxonomic evenness, somewhat decoupled from the
-disturbance intensity as already observed in the Guiana Shield
+The IDH well predicted well the disturbance impact on community
+taxonomic richness, enhanced until an intensity threshold (20-25\% AGB
+loss), and to some extent on taxonomic evenness, somewhat decoupled from
+the disturbance intensity as already observed in the Guiana Shield
 \citep{Baraloto2012a} and in Bornean tropical forests
@@
 evenness rather represented the increasing dominance of pionneers that
-balanced the usual hyper-dominance of a few species in tropical forests below the
-intensity threshold, thus increasing community overall evenness up to
-the intensity threshold beyond which pioneers became in turn highly
-dominant and decreased the overall evenness \citep{Baraloto2012a}.
+balanced the usual hyper-dominance of a few species in tropical forests
+below the intensity threshold, thus increasing community overall
+evenness up to the intensity threshold beyond which pioneers became in
+turn highly dominant and decreased the overall evenness
+\citep{Baraloto2012a}.
 
@@
 \citep{Busing2002}. Along time the recovery of infrequent species was
-increasingly difficult, so that the time for the full recovery of the
+increasingly slow, so that the time for the full recovery of the
 functional redundancy, in some communities just initiated after 30
-years, was extremely difficult to estimate \citep{Elmqvist2003, Diaz2005}.
+years, was extremely difficult to estimate
+\citep{Elmqvist2003, Diaz2005}.
 
@@
 on communities redundancy cautioned against the risks of infrequent
-species loss and the persistence of disturbance-specific communities
-\citep{Herault2018}.
+species loss and the persistence of disturbance-specific communities.
 

### 2018-08-13 Bibliotemplate.Rmd
mise en page

--- 2018-07-30 WholePlotTrajectories.Rmd
@@
 ---
-title: "Post-Disturbance Tree Community Trajectories in a Neotropical Forest"
+title             : "The title"
+shorttitle        : "Title"
+
 author: 
-  - name: Ariane MIRABEL
-    authsuperscript: 1*
-  - name: Eric Marcon
-    authsuperscript: 1
-  - name: Bruno Hérault
-    authsuperscript: 2 3
+  - name          : "First Author"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       : "Postal address"
+    email         : "my@email.com"
+  - name          : "Ernst-August Doelle"
+    affiliation   : "1,2"
+
 affiliation:
-  - affsuperscript: 1
-    dptuniv: UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.
-    address: >
-      Campus Agronomique,
-      97310 Kourou,
-      France.
-  - affsuperscript: 2
-    dptuniv: Cirad, Univ montpellier, UR Forests & Societies.
-    address: >
-      Montpellier,
-      France.
-  - affsuperscript: 3
-    dptuniv: INPHB, Institut National Polytechnique Félix Houphouet-Boigny
-    address: >
-      Yamoussoukro,
-      Ivory Coast.
-corrauthor:
-    email: ariane.mirabel@ecofog.gf
-    url: http://www.ecofog.gf/spip.php?article47
-abstract: >   
-    Understanding the ecological rules underlying the maintenance of tropical forests biodiversity, structure, functioning and dynamics is urgent to anticipate their fate in the global change context.
-    The huge diversity of tropical forests is often assumed to be regularly reshaped by natural disturbance yielding a diversity peak at intermediate intensity.
-This intermediate disturbance hypothesis (IDH), though, remains debated and the controversy questions the extent of communities' resilience regarding their taxonomic and functional facets.
+  - id            : "1"
+    institution   : "Wilhelm-Wundt-University"
+  - id            : "2"
+    institution   : "Konstanz Business School"
 
-    To disentangle the ecological processes driving community response to disturbance, we analysed the tree community trajectories over 30 years following a disturbance gradient in a Neotropical forest.
-    Specifically, we examined community functional and taxonomic trajectories with regards to diversity, composition and redundancy.
-Functional trajectories were drawn based on 7 leaf, stem and life-history traits.
-    We highlighted the cyclic recovery of community taxonomic and functional composition.
-While pre-disturbance taxonomic differences were maintained over time, functional composition trajectories were quite similar among communities.
-    The IDH did predict communities taxonomic diversity response while functional diversity was enhanced whatever the disturbance intensity.
-    Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years.
-This acknowledged the need of decades-long cycles with no disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbances.
+author_note: |
+  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
 
-keywords: [Taxonomic and Functional Biodiversity, Neotropical Forests, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience]
-journalinfo: Hal xxx
-archive: DOI xxxx
-date: "`r format(Sys.time(), '%d %B %Y')`"
-preamble: >
-  \usepackage{textcomp}
-  \DeclareUnicodeCharacter{B0}{\textdegree}
-  \usepackage{tabu}
-  \usepackage{caption}
-  \captionsetup{justification = justified}
-  \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
-  \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
-bibliography: references.bib
-lang: english
-keywordlabel: Keywords
-corrauthorlabel: Corresponding author
-output:
-  bookdown::pdf_book:
-    base_format: EcoFoG::article
-    keep_tex: TRUE
-  bookdown::gitbook: default
-  bookdown::word_document2: default
+  Enter author note here.
+
+abstract: |
+  Enter abstract here. Each new line herein must be indented, like this line.
+
+keywords          : "keywords"
+wordcount         : "X"
+
+bibliography      : [references.bib]
+
+figsintext        : no
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+lineno            : yes
+mask              : no
+
+class             : "man"
+output            : papaja::apa6_word
 ---
 
-(ref:Bolda) **(a)**
+```{r load_packages, include = FALSE}
+library("papaja")
+```
 
-(ref:Boldb) **(b)**
-
-(ref:Boldc) **(c)**
-
-(ref:Boldd) **(d)**
-
-(ref:capThickness) *L_thickness*
-
-(ref:capChloro) *L_chloro*
-
-(ref:capThoughness) *L_toughness*
-
-(ref:capSLA) *SLA*
-
-(ref:capWSG) *WSG*
-
-(ref:capBarkThick) *Bark-thick*
-
-(ref:capHmax) *Hmax*
-
-```{r Options, echo=FALSE, message=FALSE, warning=F}
-knitr::opts_chunk$set(cache=F, echo=F,warning=F, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=40), fig.align="center")
-# Installation des packages si nécessaire et chargement
-Library <- function(Packages) {
-  InstallAndLoad <- function(Package) {
-    if (!Package %in% installed.packages()[, 1]) {install.packages(Package, repos="https://cran.rstudio.com/")}
-    require(Package, character.only = TRUE)
-  }
-  invisible(sapply(Packages, InstallAndLoad))
-}
-# Ajouter les packages nécessaires ici
-Library("kableExtra")
-
-## abstract: 204 words / text without legends : 3659 words / references: 68
-
-source("Scripts/DrawingTools.R")
-#source("Scripts/GithubPages.R")
+```{r analysis_preferences}
+# Seed for random number generation
+set.seed(42)
 ```
@@
 The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Biomass2018].
-
-```{r Tab1, echo=FALSE,as.is=T}
-Intervention <- data.frame(
-  Treatment=c("Control", "T1", "T2", "T3"),
-  Timber=c("", 
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha", 
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha",
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha"),
-  Thinning=c("", 
-             "", 
-             "DBH $\\geq$ 40 cm, non-valuable species, $\\approx$ 30 trees/ha", 
-             "DBH $\\geq$ 50 cm, non-valuable species, $\\approx$ 15 trees/ha  "),
-  Fuelwood=c("", 
-             "", 
-             "", 
-             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-valuable species, $\\approx$ 15 trees/ha"),
-  AGBLost=c(0, "$[12\\%-33\\%]$", "$[33\\%-56\\%]$", "$[35\\%-56\\%]$")
-)
-names(Intervention)[5] <- "\\%AGB lost"
-knitr::kable(Intervention, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.", longtable = FALSE, booktabs = TRUE, escape = FALSE,format="latex") %>%
-  kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
-```
 
@@
 
-```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS space.
-Lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure*"}
-load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
-load("DB/FunDistance_ForGraphs");load("DB/TaxoDistance_ForGraphs")
-
-par(mfrow=c(2,2),mar=c(2,2.5,2.5,1),oma=c(1,2,1,3),no.readonly = T)
-TaxoCompo(MatrepTaxo)
-mtext("Taxonomic composition",side=3,line=1.8,cex=1.05)
-FunCompo(MatrepFun)
-mtext("Functional composition",side=3,line=1.8,cex=1.05)
-EuclidDist(TaxoEuclid)
-mtext("Euclidean distance\nfrom 1985 inventory",side=2,padj=0,line=2,cex=0.8)
-mtext("(c)",side=3,adj=0,line=0.5)
-EuclidDist(FunEuclid)
-mtext("(d)",side=3,adj=0,line=0.5)
-mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=0.8)
-legend("right",inset=c(-0.345,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
-
-```
-
 Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions followed unimodal trajectories either stabilizing or returning towards their initial values.
@@
 - Figure S2).
-
-```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWSG), and bark thickness, (ref:capBarkThick)) and one life history trait (Specific maximum height at adult stage, (ref:capHmax)).
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure*"}
-load("DB/CWM")
-
-CWMdraw(CWM)
-legendCWM()
-
-```
 
@@
 
-```{r DivTaxo, out.width = '100%', echo=FALSE, fig.cap="Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic (ref:Bolda) richness, (ref:Boldb), taxonomic evenness, (ref:Boldc) functional richness, and (ref:Boldd) functional evenness.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
-load("DB/ReplacementTraj_ForGraphs")
-load("DB/FunctionalTraj_ForGraphs")
-load("DB/FunctionalRichnessTraj_ForGraphs")
-
-par(mfrow=c(2,2),mar=c(2,2.5,2.5,1),oma=c(1,2,1,3),no.readonly = T)
-TaxoTraj(CompleteTaxo)
-plotDiv(CompleteRichFun,remove=TRUE)
-mtext("(c) Functional Richness",side=3,adj=0,line=1,cex=0.9)
-plotDiv(CompleteFun,remove=TRUE)
-mtext("(d) Functional Evenness",side=3,adj=0,line=1,cex=0.9)
-mtext("Years since disturbance",side=1,line=2.2,adj=1)
-legend("right",inset=c(-0.3,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
-
-```
-
 The plot 7 from treatment 1 displayed constantly outlying functional richness and evenness and was removed from the graphical representation for better readability.
@@
 
-```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/4,fig.cap="Relationship between the initial \\%AGB loss and community taxonmic richness (ref:Bolda), taxonomic evenness (ref:Boldb), functional richness (ref:Boldc),and functional evenness (ref:Boldd) at 10, 20 and 30 years after disturbance", fig.env="figure*"}
-load("DB/RichnessIDH");load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB");load("DB/FunRichIDH")
-
-par(mfrow=c(1,4),mar=c(2,2.5,3,1),oma=c(1,1,1,2.5),no.readonly=TRUE)
-plotIDH(Data=Rich,AgbLoss=AGBloss)
-mtext("(a) Taxonomic\nRichness",line=1,side=3,adj=0,cex=0.85)
-mtext("Equivalent diversity",side=2,line=2.2,adj=1,cex=0.8)
-plotIDH(Data=Sim,AgbLoss=AGBloss)
-mtext("(b) Taxonomic\nEvenness",line=1,side=3,adj=0,cex=0.85)
-plotIDH(Data=Rao,AgbLoss=AGBloss)
-mtext("(c) Functional\nRichness",line=1,side=3,adj=0,cex=0.85)
-plotIDH(Data=FunRich,AgbLoss=AGBloss)
-mtext("(d) Functional\nEvenness",line=1,side=3,adj=0,cex=0.85)
-mtext("initial AGB loss (%)",side=1,line=2.2,adj=1,cex=0.8)
-legend("right",inset=c(-0.5,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
-
-```
-
 ## Functional redundancy
@@
 The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{spearman}=0.47$) and the initial value had not recovered for any disturbed communities after 30 years.
-
-```{r RedFunRest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
-Colors are disturbance treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
-load("DB/Redundancy_restricted")
-
-par(mar=c(3,2,3,2))
-RedundancyPlot(RedundancyTraj_restricted)
-mtext("Functional Redundancy",side=3,adj=0,line=1.5,cex=1.5)
-mtext("Restricted to initial functional space",side=3,adj=0,cex=1.2,line=0.5)
-mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=1.5)
-
-```
 
@@
 
-`r if (knitr:::is_html_output()) '# References {-}'`
+\newpage
+
+# References
+```{r create_r-references}
+r_refs(file = "r-references.bib")
+```
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+
+<div id = "refs"></div>
+\endgroup

### 2018-08-13 Bibliotemplate.tex
mise en page

--- 2018-07-30 WholePlotTrajectories.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Article EcoFoG
-% Version 2.1 (23/10/2017)
-%
-% adapté de :
-% Stylish Article
-% LaTeX Template
-% Version 1.0 (31/1/13)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Mathias Legrand (legrand.mathias@gmail.com)
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[man]{apa6}
 
-%----------------------------------------------------------------------------------------
-%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+    \usepackage{xltxtra,xunicode}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
 
-\documentclass[fleqn,10pt]{ArtEcoFoG} % Document font size and equations flushed left
+% Table formatting
+\usepackage{longtable, booktabs}
+\usepackage{lscape}
+% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
+\usepackage{multirow}		% Table styling
+\usepackage{tabularx}		% Control Column width
+\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
+\usepackage{threeparttablex}            % Lets threeparttable work with longtable
 
-\setcounter{tocdepth}{3} % Show only three levels in the table of contents section: sections, subsections and subsubsections
+% Create new environments so endfloat can handle them
+% \newenvironment{ltable}
+%   {\begin{landscape}\begin{center}\begin{threeparttable}}
+%   {\end{threeparttable}\end{center}\end{landscape}}
 
-% Pandoc environments
-\usepackage{framed}
-\usepackage{fancyvrb}
+\newenvironment{lltable}
+  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
+  {\end{ThreePartTable}\end{center}\end{landscape}}
+
+  \usepackage{ifthen} % Only add declarations when endfloat package is loaded
+  \ifthenelse{\equal{\string man}{\string man}}{%
+   \DeclareDelayedFloatFlavor{ThreePartTable}{table} % Make endfloat play with longtable
+   % \DeclareDelayedFloatFlavor{ltable}{table} % Make endfloat play with lscape
+   \DeclareDelayedFloatFlavor{lltable}{table} % Make endfloat play with lscape & longtable
+  }{}%
+
+% The following enables adjusting longtable caption width to table width
+% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
+\makeatletter
+\newcommand\LastLTentrywidth{1em}
+\newlength\longtablewidth
+\setlength{\longtablewidth}{1in}
+\newcommand\getlongtablewidth{%
+ \begingroup
+  \ifcsname LT@\roman{LT@tables}\endcsname
+  \global\longtablewidth=0pt
+  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
+  \@nameuse{LT@\roman{LT@tables}}%
+  \fi
+\endgroup}
+
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            pdfauthor={},
+            pdftitle={The title},
+            colorlinks=true,
+            citecolor=blue,
+            urlcolor=blue,
+            linkcolor=black,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+
+\setlength{\parindent}{0pt}
+%\setlength{\parskip}{0pt plus 0pt minus 0pt}
+
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+
+% Manuscript styling
+\captionsetup{font=singlespacing,justification=justified}
+\usepackage{csquotes}
+\usepackage{upgreek}
+
+ % Line numbering
+  \usepackage{lineno}
+  \linenumbers
+
+\usepackage{tikz} % Variable definition to generate author note
+
+% fix for \tightlist problem in pandoc 1.14
 \providecommand{\tightlist}{%
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, fontsize=\scriptsize} % Code R
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\ImportTok}[1]{{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
-\newcommand{\BuiltInTok}[1]{{#1}}
-\newcommand{\ExtensionTok}[1]{{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
-\newcommand{\RegionMarkerTok}[1]{{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
-\newcommand{\NormalTok}[1]{{#1}}
-\usepackage{longtable,booktabs}
-\usepackage{caption}
-% These lines are needed to make table captions work with longtable:
-\makeatletter
-\def\fnum@table{\tablename~\thetable}
-\makeatother
-% longtable 2 columns
-% https://tex.stackexchange.com/questions/161431/how-to-solve-longtable-is-not-in-1-column-mode-error
-\makeatletter
-\let\oldlt\longtable
-\let\endoldlt\endlongtable
-\def\longtable{\@ifnextchar[\longtable@i \longtable@ii}
-\def\longtable@i[#1]{\begin{figure}[t]
-\onecolumn
-\begin{minipage}{0.5\textwidth}\scriptsize
-\oldlt[#1]
-}
-\def\longtable@ii{\begin{figure}[t]
-\onecolumn
-\begin{minipage}{0.5\textwidth}\scriptsize
-\oldlt
-}
-\def\endlongtable{\endoldlt
-\end{minipage}
-\twocolumn
-\end{figure}}
-\makeatother
 
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight0.8\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Essential manuscript parts
+  \title{The title}
 
-% User-adder preamble
-\usepackage{textcomp} \DeclareUnicodeCharacter{B0}{\textdegree}
-\usepackage{tabu} \usepackage{caption}
-\captionsetup{justification = justified}
-\renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
-\hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
+  \shorttitle{Title}
 
-%----------------------------------------------------------------------------------------
-%	ARTICLE INFORMATION
-%----------------------------------------------------------------------------------------
+  \author{First Author\textsuperscript{1}~\& Ernst-August Doelle\textsuperscript{1,2}}
 
-\JournalInfo{Hal xxx} % Journal information
-\Archive{DOI xxxx} % Additional notes (e.g. copyright, DOI, review/research article)
+  % \def\affdep{{"", ""}}%
+  % \def\affcity{{"", ""}}%
 
-\PaperTitle{Post-Disturbance Tree Community Trajectories in a Neotropical Forest} % Article title
+  \affiliation{
+    \vspace{0.5cm}
+          \textsuperscript{1} Wilhelm-Wundt-University\\
+          \textsuperscript{2} Konstanz Business School  }
 
-\Authors{
-Ariane MIRABEL\textsuperscript{1*}\\ Eric Marcon\textsuperscript{1}\\ Bruno Hérault\textsuperscript{2 3}
-} % Authors
-\affiliation{
-\textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles,
-Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}Cirad, Univ montpellier, UR Forests \& Societies.\\ \hspace{1em} Montpellier, France.\\\textsuperscript{3}INPHB, Institut National Polytechnique Félix Houphouet-Boigny\\ \hspace{1em} Yamoussoukro, Ivory Coast.
-}
-\affiliation{*\textbf{Corresponding author}: ariane.mirabel@ecofog.gf, http://www.ecofog.gf/spip.php?article47} % Corresponding author
+  \authornote{
+    Add complete departmental affiliations for each author here. Each new
+    line herein must be indented, like this line.
 
-\Keywords{Taxonomic and Functional Biodiversity, Neotropical Forests, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience} % Keywords - if you don't want any simply remove all the text between the curly brackets
-\newcommand{\keywordname}{Keywords} % Defines the keywords heading name
+    Enter author note here.
 
-%----------------------------------------------------------------------------------------
-%	ABSTRACT
-%----------------------------------------------------------------------------------------
-
-\Abstract{
-Understanding the ecological rules underlying the maintenance of
-tropical forests biodiversity, structure, functioning and dynamics is
-urgent to anticipate their fate in the global change context. The huge
-diversity of tropical forests is often assumed to be regularly reshaped
-by natural disturbance yielding a diversity peak at intermediate
-intensity. This intermediate disturbance hypothesis (IDH), though,
-remains debated and the controversy questions the extent of communities'
-resilience regarding their taxonomic and functional facets. To
-disentangle the ecological processes driving community response to
-disturbance, we analysed the tree community trajectories over 30 years
-following a disturbance gradient in a Neotropical forest. Specifically,
-we examined community functional and taxonomic trajectories with regards
-to diversity, composition and redundancy. Functional trajectories were
-drawn based on 7 leaf, stem and life-history traits. We highlighted the
-cyclic recovery of community taxonomic and functional composition. While
-pre-disturbance taxonomic differences were maintained over time,
-functional composition trajectories were quite similar among
-communities. The IDH did predict communities taxonomic diversity
-response while functional diversity was enhanced whatever the
-disturbance intensity. Although consistent, the recovery of community
-composition, diversity and redundancy remained unachieved after 30
-years. This acknowledged the need of decades-long cycles with no
-disturbance to ensure a complete recovery, and questioned tropical
-forest community resilience after repeated disturbances.
+    Correspondence concerning this article should be addressed to First
+    Author, Postal address. E-mail:
+    \href{mailto:my@email.com}{\nolinkurl{my@email.com}}
   }
 
-%----------------------------------------------------------------------------------------
+  \abstract{Enter abstract here. Each new line herein must be indented, like this
+line.}
+  \keywords{keywords \\
 
+    \indent Word count: X
+  }
+
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}{Lemma}[section]
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{definition}
+\newtheorem{example}{Example}[section]
+\theoremstyle{definition}
+\newtheorem{exercise}{Exercise}[section]
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
+\newtheorem*{solution}{Solution}
 \begin{document}
 
-\selectlanguage{english}
+\maketitle
 
-\flushbottom % Makes all text pages the same height
-
-\maketitle % Print the title and abstract box
-
-\tableofcontents % Print the contents section
-
-\thispagestyle{empty} % Removes page numbering from the first page
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
+\setcounter{secnumdepth}{0}
 
@@
 forests products together with current global changes increases the
-pressure on remaining natural forests \citep{Morales-Hidalgo2015} and
-threatens the maintenance and dynamics in space and time of communities
-structure, composition and functioning \citep{Anderson-Teixeira2013}.
+pressure on remaining natural forests ({\textbf{???}}) and threatens the
+maintenance and dynamics in space and time of communities structure,
+composition and functioning ({\textbf{???}}).
 
@@
 natural disturbance events changing both the abiotic environment,
-through the fluxes of light, heat and water \citep{Goulamoussene2017},
-and the biotic interactions such as competition among species
-\citep{Chesson2000}. One of the cornerstone of tropical forest ecology
-is to understand the processes and drivers of ecosystems response to
-disturbance \citep{Chazdon2003a}. For now, this has been largely studied
-through forest structural parameters such as aboveground biomass, tree
-height or stem density \citep{Piponiot2016, Rutishauser2016} that are
-rapid and convenient to measure. These structural parameters have been
-sucessfully modeled, giving important insights into the recovery of
-ecosystem processes and services \citep{Herault2018}. However the
-response of forests diversity and composition remains unclear, albeit it
-determines the productivity, stability and functioning of ecosystems
-\citep{Tilman2014, Liang2016}. In the short-term, moderate disturbance
-may lead to positive impacts on communities diversity, an idea
-formalized by the intermediate disturbance hypothesis (IDH) stating a
-maximized species diversity when disturbance intensity is not too high
-\citep{Molino2001, Kariuki2006a}.
+through the fluxes of light, heat and water ({\textbf{???}}), and the
+biotic interactions such as competition among species ({\textbf{???}}).
+One of the cornerstone of tropical forest ecology is to understand the
+processes and drivers of ecosystems response to disturbance
+({\textbf{???}}). For now, this has been largely studied through forest
+structural parameters such as aboveground biomass, tree height or stem
+density ({\textbf{???}}, {\textbf{???}}) that are rapid and convenient
+to measure. These structural parameters have been sucessfully modeled,
+giving important insights into the recovery of ecosystem processes and
+services ({\textbf{???}}). However the response of forests diversity and
+composition remains unclear, albeit it determines the productivity,
+stability and functioning of ecosystems ({\textbf{???}},
+{\textbf{???}}). In the short-term, moderate disturbance may lead to
+positive impacts on communities diversity, an idea formalized by the
+intermediate disturbance hypothesis (IDH) stating a maximized species
+diversity when disturbance intensity is not too high ({\textbf{???}},
+{\textbf{???}}).
 
 Validations of the IDH though remain scarce in the long-term and mainly
-rely on the analysis of taxonomic richness \citep{Molino2001}. Taxonomic
+rely on the analysis of taxonomic richness ({\textbf{???}}). Taxonomic
 richness alone, however, gives limited or misleading information on
-forests recovery and functioning \citep{Chaudhary2016}. More
+forests recovery and functioning ({\textbf{???}}). More
 ecological-meaningful analysis would couple richness with (i) evenness,
@@
 thus the underlying ecological processes, and (ii) composition that is
-crucial for conservation issues \citep{Lavorel2002, Bellwood2006}.
+crucial for conservation issues ({\textbf{???}}, {\textbf{???}}).
 Furthermore, a functional approach accounting for species biological
@@
 redundancy to ecosystem functioning and to its environmental constraints
-\citep{Violle2007b, Baraloto2012a}. In that respect, the functional
+({\textbf{???}}, {\textbf{???}}). In that respect, the functional
 trait-based approach that focus on major traits related to species
 ecology and mediate species performance in a given environment was
-sucessfully adopted \citep{Diaz2005}. For instance, the functional
+sucessfully adopted ({\textbf{???}}). For instance, the functional
 approach revealed in tropical rainforests the deterministic processes
 entailing, after disturbance, a functional shift from a dominance of
-``conservative'' slow-growing species dealing with scarce resources to
-``acquisitive'' fast-growing species with rapid and efficient use of
-abundant resources \citep{Reich2014, Herault2011}. This shift is
-translated into the trajectories of key functional traits related to
-resource acquisition (leaf and stem traits) and life-history traits
-(seed mass, maximum size)
-\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Eventually
-a complete overview of communities response to disturbance would
-encompass the changes in functional redundancy, that quantifies the
-amount of shared trait values among species \citep{Carmona2016}. The
-high functional redundancy of hyperdiverse tropical forests
-\citep{Bellwood2006} mitigates the impacts of species removal on
-ecosystem functioning and determines the resilience of communities after
-disturbance \citep{Elmqvist2003, Diaz2005}.
+\enquote{conservative} slow-growing species dealing with scarce
+resources to \enquote{acquisitive} fast-growing species with rapid and
+efficient use of abundant resources ({\textbf{???}}, {\textbf{???}}).
+This shift is translated into the trajectories of key functional traits
+related to resource acquisition (leaf and stem traits) and life-history
+traits (seed mass, maximum size) ({\textbf{???}}, {\textbf{???}},
+{\textbf{???}}, {\textbf{???}}). Eventually a complete overview of
+communities response to disturbance would encompass the changes in
+functional redundancy, that quantifies the amount of shared trait values
+among species ({\textbf{???}}). The high functional redundancy of
+hyperdiverse tropical forests ({\textbf{???}}) mitigates the impacts of
+species removal on ecosystem functioning and determines the resilience
+of communities after disturbance ({\textbf{???}}, {\textbf{???}}).
 
@@
 mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
-one-month dry season in March \citep{Wagner2011}. Elevation ranges
-between 5 and 50 m and soils correspond to thin acrisols over a layer of
+one-month dry season in March ({\textbf{???}}). Elevation ranges between
+5 and 50 m and soils correspond to thin acrisols over a layer of
 transformed saprolite with low permeability generating lateral drainage
@@
 (T3). Treatments T2 and T3 besides included the thinning of trees by
-poison girdling \citep{Schmitt1990}. The disturbance intensity was
-measured as the percentage of aboveground biomass (\%AGB) lost between
-the first inventory in 1984 and five years after disturbance
-\citep{Piponiot2016} estimated with the BIOMASS R package
-\citep{Biomass2018}.
-
-\begin{table}
-
-\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.}
-\centering
-\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
-\toprule
-Treatment & Timber & Thinning & Fuelwood & \%AGB lost\\
-\midrule
-Control &  &  &  & 0\\
-T1 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha &  &  & $[12\%-33\%]$\\
-T2 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30 trees/ha &  & $[33\%-56\%]$\\
-T3 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & $[35\%-56\%]$\\
-\bottomrule
-\end{tabu}
-\end{table}
+poison girdling ({\textbf{???}}). The disturbance intensity was measured
+as the percentage of aboveground biomass (\%AGB) lost between the first
+inventory in 1984 and five years after disturbance ({\textbf{???}})
+estimated with the BIOMASS R package ({\textbf{???}}).
 
@@
 
-See Supplementary Materials -Figure S1 and \citet{Aubry-Kientz2013} for
-the detailed methodology.
+See Supplementary Materials -Figure S1 and ({\textbf{???}}) for the
+detailed methodology.
 
@@
 traits (maximum specific height and seed mass) came from the BRIDGE
-project \footnote{http://www.ecofog.gf/Bridge/}. Trait values were
+project.\footnote{\url{http://www.ecofog.gf/Bridge/}} Trait values were
 assessed from a selection of individuals located in nine permanent plots
@@
 pertaining to 157 genera. Missing trait values (10\%) were filled using
-multivariate imputation by chained equation \citep{Mice2011}.
+multivariate imputation by chained equation ({\textbf{???}}).
 Imputations were restricted within genus or family when samples were too
@@
 leaf, stem and life-history traits were also visualized with the
-community weighted means (CWM) \citep{Diaz2007}. Species seed mass
+community weighted means (CWM) ({\textbf{???}}). Species seed mass
 corresponded to 5 classes of increasing mass, seed mass trajectories
@@
 evenness, \emph{i.e} the Hill number translation of the Simpson index
-\citep{Hill1973}. These indices belong to the set of HCDT or generalized
+({\textbf{???}}). These indices belong to the set of HCDT or generalized
 entropy, respectively corresponding to the 0 and 2 order of diversity
-(q), recomended for diversity studies \citep{Marcon2015b}. The
-functional diversity was reported using the functional richness and
-functional evenness, \emph{i.e} Rao index of quadratic entropy which
-combines species abundance distribution and average pairwise
-dissimilarity based on species functional traits.
+(q), recomended for diversity studies ({\textbf{???}}). The functional
+diversity was reported using the functional richness and functional
+evenness, \emph{i.e} Rao index of quadratic entropy which combines
+species abundance distribution and average pairwise dissimilarity based
+on species functional traits.
 
@@
 The functional redundancy was measured as the overlap among species in
-community functional space \citep{Carmona2016}. The samples of the trait
+community functional space ({\textbf{???}}). The samples of the trait
 database were first mapped in a 2-dimensional plan with a PCA analysis.
@@
 
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/NMDSplans-1}
-
-}
-
-\caption{Plot trajectories in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS space.
-Lower panels (\textbf{(c)} and \textbf{(d)}) represent the euclidean distance to initial condition along the 30 sampled years.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:NMDSplans}
-\end{figure*}
-
 Except for leaf chlorophyll content, which continued to increase for
@@
 
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/CWM-1}
-
-}
-
-\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WSG}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}).
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
-\end{figure*}
-
 \subsection{Communities richness and
@@
 kept increasing.
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/DivTaxo-1}
-
-}
-
-\caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic \textbf{(a)} richness, \textbf{(b)}, taxonomic evenness, \textbf{(c)} functional richness, and \textbf{(d)} functional evenness.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:DivTaxo}
-\end{figure*}
 
@@
 
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/IDHplot-1}
-
-}
-
-\caption{Relationship between the initial \%AGB loss and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)}, functional richness \textbf{(c)},and functional evenness \textbf{(d)} at 10, 20 and 30 years after disturbance}\label{fig:IDHplot}
-\end{figure*}
-
 \subsection{Functional redundancy}\label{functional-redundancy}
@@
 had not recovered for any disturbed communities after 30 years.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/RedFunRest-1}
-
-}
-
-\caption{Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
-Colors are disturbance treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:RedFunRest}
-\end{figure}
 
@@
 commonness of dispersal limitation among tropical tree species
-\citep{Svenning2005}.
+({\textbf{???}}).
 
@@
 pre-disturbance surviving trees mirror the initial community
-\citep{Herault2018}, changes in functional composition relied upon the
+({\textbf{???}}), changes in functional composition relied upon the
 recruitment of species or functional types that were infrequent or
@@
 moving community functional composition right along the first axis in
-Figure \ref{fig:NMDSplans} \citep{Westoby1998, Wright2004, Reich2014}.
-Thereafter long-lived, more resistant and shade-tolerant species
-excluded the first established pioneers and started the recovery of
-pre-disturbance functional composition, moving similarly community
-functional composition left along the first axis and upward along the
-second axis in Figure \ref{fig:NMDSplans}.
+Figure \ref{fig:NMDSplans} ({\textbf{???}}, {\textbf{???}},
+{\textbf{???}}). Thereafter long-lived, more resistant and
+shade-tolerant species excluded the first established pioneers and
+started the recovery of pre-disturbance functional composition, moving
+similarly community functional composition left along the first axis and
+upward along the second axis in Figure \ref{fig:NMDSplans}.
 
@@
 at the same time dispersal limitation maintained their divergence in
-taxonomic composition \citep{Fukami2005}.
+taxonomic composition ({\textbf{???}}).
 
@@
 the disturbance intensity as already observed in the Guiana Shield
-\citep{Baraloto2012a} and in Bornean tropical forests
-\citep{Cannon1998}. The disturbance intensity determined the balance in
-the community between pre-disturbance surviving trees and those
-recruited afterward. The pool of true pioneer species specifically
-recruited after disturbance is restricted in the Guiana Shield to a few
-common genera (e.g.~Cecropia spp., Vismia spp.) \citep{Guitet2018}.
-Below the intensity threshold the size of the surviving community
-maintained the pre-disturbance high taxonomic richness while the
-recruitment of pioneers, infrequent or absent before disturbance,
-increased both community taxonomic richness and evenness. Beyond the
-intensity threshold, the disturbance decreased the taxonomic richness of
-surviving trees which was not offset by the enrichment of pioneers, so
-that the overall community taxonomic richness decreased according to the
-disturbance intensity \citep{Molino2001}. For community taxonomic
-evenness the disturbance impact was similar but slighter, as the
-evenness is less sensitive to the loss of rare species. Taxonomic
-evenness rather represented the increasing dominance of pionneers that
-balanced the usual hyper-dominance of a few species in tropical forests
-below the intensity threshold, thus increasing community overall
-evenness up to the intensity threshold beyond which pioneers became in
-turn highly dominant and decreased the overall evenness
-\citep{Baraloto2012a}.
+({\textbf{???}}) and in Bornean tropical forests ({\textbf{???}}). The
+disturbance intensity determined the balance in the community between
+pre-disturbance surviving trees and those recruited afterward. The pool
+of true pioneer species specifically recruited after disturbance is
+restricted in the Guiana Shield to a few common genera (e.g.~Cecropia
+spp., Vismia spp.) ({\textbf{???}}). Below the intensity threshold the
+size of the surviving community maintained the pre-disturbance high
+taxonomic richness while the recruitment of pioneers, infrequent or
+absent before disturbance, increased both community taxonomic richness
+and evenness. Beyond the intensity threshold, the disturbance decreased
+the taxonomic richness of surviving trees which was not offset by the
+enrichment of pioneers, so that the overall community taxonomic richness
+decreased according to the disturbance intensity ({\textbf{???}}). For
+community taxonomic evenness the disturbance impact was similar but
+slighter, as the evenness is less sensitive to the loss of rare species.
+Taxonomic evenness rather represented the increasing dominance of
+pionneers that balanced the usual hyper-dominance of a few species in
+tropical forests below the intensity threshold, thus increasing
+community overall evenness up to the intensity threshold beyond which
+pioneers became in turn highly dominant and decreased the overall
+evenness ({\textbf{???}}).
 
@@
 as well a progressive and long-lasting increase of the functional
-richness and evenness \citep{Denslow1980, Molino2001}. Beyond an
+richness and evenness ({\textbf{???}}, {\textbf{???}}). Beyond an
 intensity threshold, though, a few short-lived pioneers occupied the
@@
 recruit we expect the taxonomic and functional trajectories to catch up
-with those observed after intermediate disturbance \citep{Walker2009}.
+with those observed after intermediate disturbance ({\textbf{???}}).
 
@@
 recruitment eased for the first recruited species but then increasingly
-hampered by the emergence of interspecific competition
-\citep{Busing2002}. Along time the recovery of infrequent species was
-increasingly slow, so that the time for the full recovery of the
-functional redundancy, in some communities just initiated after 30
-years, was extremely difficult to estimate
-\citep{Elmqvist2003, Diaz2005}.
+hampered by the emergence of interspecific competition ({\textbf{???}}).
+Along time the recovery of infrequent species was increasingly slow, so
+that the time for the full recovery of the functional redundancy, in
+some communities just initiated after 30 years, was extremely difficult
+to estimate ({\textbf{???}}, {\textbf{???}}).
 
@@
 chances to see the persistence of disturbance-specific species at the
-expense of late-successional ones \citep{Haddad2008}. Besides, the
+expense of late-successional ones ({\textbf{???}}). Besides, the
 long-term recovery of infrequent species increases the risks to loose
 cornerstone species, with unexpected ecological consequences
-\citep{Jones1994, Chazdon2003a, Diaz2005}. Apart from the functional
-characteristics considered here, infrequent species might indeed have
-unique functions in the ecosystem or be a key for some fauna
-\citep{Schleuning2016}.
+({\textbf{???}}, {\textbf{???}}, {\textbf{???}}). Apart from the
+functional characteristics considered here, infrequent species might
+indeed have unique functions in the ecosystem or be a key for some fauna
+({\textbf{???}}).
 
@@
 on communities redundancy cautioned against the risks of infrequent
-species loss and the persistence of disturbance-specific communities.
+species loss and the persistence of disturbance-specific communities
+({\textbf{???}}).
 
-%----------------------------------------------------------------------------------------
-%	REFERENCE LIST
-%----------------------------------------------------------------------------------------
+\newpage
 
-\bibliographystyle{mee}
-\makeatletter
-% The filename has .bib extension the must be eliminated
-\filename@parse{references.bib}
-% parse stores the file name in base. Extension starts at the first dot, so don't use dots in file names.
-\bibliography{\filename@base}
-\makeatother
+\section{References}\label{references}
 
-%----------------------------------------------------------------------------------------
+\begingroup
+\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}
+
+\hypertarget{refs}{}
+
+\endgroup
 

### 2018-08-13 WholePlotTrajectories.tex
mise en page

@@
 
-\JournalInfo{Hal xxx} % Journal information
-\Archive{DOI xxxx} % Additional notes (e.g. copyright, DOI, review/research article)
+\JournalInfo{\ }
+\Archive{\ }
 
@@
 \textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles,
-Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}Cirad, Univ montpellier, UR Forests \& Societies.\\ \hspace{1em} Montpellier, France.\\\textsuperscript{3}INPHB, Institut National Polytechnique Félix Houphouet-Boigny\\ \hspace{1em} Yamoussoukro, Ivory Coast.
+Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}Cirad, Univ montpellier, UR Forests \& Societies.\\ \hspace{1em} Montpellier, France.\\\textsuperscript{3}INPHB, Institut National Polytechnique Félix Houphoüet-Boigny\\ \hspace{1em} Yamoussoukro, Ivory Coast.
 }
-\affiliation{*\textbf{Corresponding author}: ariane.mirabel@ecofog.gf, http://www.ecofog.gf/spip.php?article47} % Corresponding author
+\affiliation{*\textbf{Corresponding author}: ariane.mirabel@ecofog.gf, https://github.com/ArianeMirabel} % Corresponding author
 
-\Keywords{Taxonomic and Functional Biodiversity, Neotropical Forests, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience} % Keywords - if you don't want any simply remove all the text between the curly brackets
+\Keywords{Community Ecology, Determinants of Plant Community Diversity and Structure, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity} % Keywords - if you don't want any simply remove all the text between the curly brackets
 \newcommand{\keywordname}{Keywords} % Defines the keywords heading name
@@
 disentangle the ecological processes driving community response to
-disturbance, we analysed the tree community trajectories over 30 years
+disturbance, we analyzed the tree community trajectories over 30 years
 following a disturbance gradient in a Neotropical forest. Specifically,
@@
 rapid and convenient to measure. These structural parameters have been
-sucessfully modeled, giving important insights into the recovery of
+successfully modeled, giving important insights into the recovery of
 ecosystem processes and services \citep{Herault2018}. However the
@@
 ecology and mediate species performance in a given environment was
-sucessfully adopted \citep{Diaz2005}. For instance, the functional
+successfully adopted \citep{Diaz2005}. For instance, the functional
 approach revealed in tropical rainforests the deterministic processes
@@
 \citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Eventually
-a complete overview of communities response to disturbance would
-encompass the changes in functional redundancy, that quantifies the
-amount of shared trait values among species \citep{Carmona2016}. The
-high functional redundancy of hyperdiverse tropical forests
+a complete overview of community response to disturbance would encompass
+the changes in functional redundancy, that quantifies the amount of
+shared trait values among species \citep{Carmona2016}. The high
+functional redundancy of hyper-diverse tropical forests
 \citep{Bellwood2006} mitigates the impacts of species removal on
@@
 In this study, we monitored over 30 years the response of 75 ha of
-neotropical forest plots set up on a gradient of disturbance intensity,
+Neotropical forest plots set up on a gradient of disturbance intensity,
 from 10 to 60\% of ecosystem above-ground biomass (AGB) loss. We made
@@
 
-See Supplementary Materials -Figure S1 and \citet{Aubry-Kientz2013} for
+See Supplementary Materials -Fig. S1 and \citet{Aubry-Kientz2013} for
 the detailed methodology.
@@
 leaf, stem and life history traits (without seed mass classes).
-Trajectories along time were reported through the euclidean distance
+Trajectories along time were reported through the Euclidean distance
 between the target inventories and the reference inventories in 1989,
@@
 entropy, respectively corresponding to the 0 and 2 order of diversity
-(q), recomended for diversity studies \citep{Marcon2015b}. The
+(q), recommended for diversity studies \citep{Marcon2015b}. The
 functional diversity was reported using the functional richness and
@@
 reached over the 30 years and the initial \%AGB loss. They were besides
-analysed through polynomial regression between (i) taxonomic and
+analyzed through polynomial regression between (i) taxonomic and
 functional richness and evenness and (ii) the initial \%AGB loss at 10,
@@
 expressed as the average number of species that could be removed from
-without reducing the functional space (Supp. Mat. - Figure S1 for a more
-comprehensive sheme).
+without reducing the functional space (Supp. Mat. - Fig. S1 for a more
+comprehensive scheme).
 
@@
 While both taxonomic and functional composition remained stable in
-undisturbed communities (Figure \ref{fig:NMDSplans}), they followed
-marked and consistent trajectories over post-\break disturbance time. In
+undisturbed communities (Fig. \ref{fig:NMDSplans}), they followed marked
+and consistent trajectories over post-\break disturbance time. In
 disturbed communities, these compositional changes corresponded to
@@
 into cyclic compositional changes with an unachieved recovery of the
-initial composition (Figure \ref{fig:NMDSplans}). The maximum
+initial composition (Fig. \ref{fig:NMDSplans}). The maximum
 dissimilarity with the initial state was positively correlated to the
 disturbance intensity for both taxonomic and functional composition
-(\(\rho_{spearman}^{taxonomic}=0.87\) and
-\(\rho_{spearman}^{functional}=0.90\) respectively). The maximum value
+(\(\rho_{spearman}^{Taxonomic}=0.87\) and
+\(\rho_{spearman}^{Functional}=0.90\) respectively). The maximum value
 was reached around 26 years after disturbance for taxonomic composition
@@
 \caption{Plot trajectories in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS space.
-Lower panels (\textbf{(c)} and \textbf{(d)}) represent the euclidean distance to initial condition along the 30 sampled years.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:NMDSplans}
+Lower panels (\textbf{(c)} and \textbf{(d)}) represent the Euclidean distance to initial condition along the 30 sampled years.
+Shaded areas are the credibility intervals.}\label{fig:NMDSplans}
 \end{figure*}
@@
 
-Maximum height at adult stage (\emph{Hmax}), leaf toughness
-(\emph{L\_toughness}) and wood specific gravity (\emph{WSG}) first
-decreased and then slightly increased but remained significantly lower
-than their initial value (Figure \ref{fig:CWM}). On the other side, Bark
-thickness (\emph{Bark\_thick}) and specific leaf area (\emph{SLA})
-increased and while \emph{Bark\_thick} remained substantially high after
-30 years, \emph{SLA} had almost recovered its initial value. For all
-traits, the maximum difference to initial value was correlated to the
-disturbance intensity (\(\rho_{spearman}^{L_{thickness}}=0.76\),
-\(\rho_{spearman}^{L_{chloro}}=0.60\),
-\(\rho_{spearman}^{L_{toughness}}=-0.53\),
+Maximum height at adult stage (\emph{Hmax}), leaf toughness and wood
+specific gravity (\emph{WSG}) first decreased and then slightly
+increased but remained significantly lower than their initial value
+(Fig. \ref{fig:CWM}). On the other side, bark thickness and specific
+leaf area (\emph{SLA}) increased and while bark thickness remained
+substantially high after 30 years, \emph{SLA} had almost recovered its
+initial value. For all traits, the maximum difference to initial value
+was correlated to the disturbance intensity
+(\(\rho_{spearman}^{Leaf thickness}=0.76\),
+\(\rho_{spearman}^{Chlorophyll content}=0.60\),
+\(\rho_{spearman}^{Leaf toughness}=-0.53\),
 \(\rho_{spearman}^{SLA}=0.93\), \(\rho_{spearman}^{WSG}=-0.75\),
-\(\rho_{spearman}^{Bark-thickness}=0.71\),
+\(\rho_{spearman}^{Bark thickness}=0.71\),
 \(\rho_{spearman}^{Hmax}=-0.40\)). The proportions of the three lightest
@@
 30 years for the lightest class while it stabilized for the two other
-(Supp. Mat. - Figure S2).
+(Supp. Mat. - Fig. S2).
 
@@
 
-\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WSG}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}).
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
+\caption{Trajectories of the communities weighted means over 30 years after disturbance of four leaf traits (Leaf thickness, chlorophyll content, toughness, and specific area), two stem traits (wood specific gravity, and bark thickness) and one life history trait (Specific maximum height at adult stage).
+}\label{fig:CWM}
 \end{figure*}
@@
 \caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic \textbf{(a)} richness, \textbf{(b)}, taxonomic evenness, \textbf{(c)} functional richness, and \textbf{(d)} functional evenness.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals }\label{fig:DivTaxo}
+Shaded areas are the credibility intervals }\label{fig:DivTaxo}
 \end{figure*}
@@
 fast but short increase followed after 10 years by a slow decrease
-towards the inital values.
+towards the initial values.
 
-The second-degree polynomial regressions between (i) the \%AGB loss and
-(ii) taxonomic and functional richness and evenness after 10, 20 and 30
-years best predicted the hump-shaped curve of the disturbance impact
-along the disturbance intensity gradient \ref{fig:IDHplot}. The
+The second-degree polynomial regressions between (i) the percentage AGB
+loss and (ii) taxonomic and functional richness and evenness after 10,
+20 and 30 years best predicted the hump-shaped curve of the disturbance
+impact along the disturbance intensity gradient \ref{fig:IDHplot}. The
 relationship between the disturbance impact and its intensity was more
@@
 
-\caption{Relationship between the initial \%AGB loss and community taxonmic richness \textbf{(a)}, taxonomic evenness \textbf{(b)}, functional richness \textbf{(c)},and functional evenness \textbf{(d)} at 10, 20 and 30 years after disturbance}\label{fig:IDHplot}
+\caption{Relationship between the initial \%AGB loss and community taxonomic richness \textbf{(a)}, taxonomic evenness \textbf{(b)}, functional richness \textbf{(c)},and functional evenness \textbf{(d)} at 10, 20 and 30 years after disturbance}\label{fig:IDHplot}
 \end{figure*}
@@
 
-\caption{Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
-Colors are disturbance treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:RedFunRest}
+\caption{Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance. Shaded areas are the credibility intervals.}\label{fig:RedFunRest}
 \end{figure}
@@
 
-Communities taxonomic and functional composition appeared resilient,
+Community taxonomic and functional composition appeared resilient,
 following similar hump-shaped trajectories starting to return towards
@@
 trajectories towards the initial composition. The high resilience of
-communities taxonomy revealed that species not belonging to the
-pre-disturbance community were hardly recruited because of the
-commonness of dispersal limitation among tropical tree species
-\citep{Svenning2005}.
+community taxonomy, in the sense of the initial state recovery, revealed
+that species not belonging to the pre-disturbance community were hardly
+recruited because of the commonness of dispersal limitation among
+tropical tree species \citep{Svenning2005}.
 
@@
 moving community functional composition right along the first axis in
-Figure \ref{fig:NMDSplans} \citep{Westoby1998, Wright2004, Reich2014}.
+Fig. \ref{fig:NMDSplans} \citep{Westoby1998, Wright2004, Reich2014}.
 Thereafter long-lived, more resistant and shade-tolerant species
@@
 functional composition left along the first axis and upward along the
-second axis in Figure \ref{fig:NMDSplans}.
+second axis in Fig. \ref{fig:NMDSplans}.
 
@@
 evenness is less sensitive to the loss of rare species. Taxonomic
-evenness rather represented the increasing dominance of pionneers that
+evenness rather represented the increasing dominance of pioneers that
 balanced the usual hyper-dominance of a few species in tropical forests
@@
 community functional richness and evenness. Irrespective of the
-disturbance intensity the recruitment of pionneers, functionally highly
+disturbance intensity the recruitment of pioneers, functionally highly
 different from the composition of pre-disturbance community, increased
-both community functional richness and eveness.
+both community functional richness and evenness.
 
@@
 establishment of long-lived pioneers enhancing community taxonomic
-richness and evenness in the long term. These long-lived pionneers,
+richness and evenness in the long term. These long-lived pioneers,
 functionally quite different from the functional composition, entailed
@@
 from the pre-disturbance community and entailed a rapid and significant
-increase of functional richness en evenness. Already after 10 years,
+increase of functional richness and evenness. Already after 10 years,
 though, short-lived pioneers started to decline and the functional
@@
 decrease was not compensated in the first place as the first recruited
-pionneers were functionally different from the pre-disturbance
-functional composition. Progressively though, first established species
-were replaced by more competitive long-lived pionneers or
-late-successional species resembling more the pre-disturbance functional
-composition and restoring the functional redundancy. This replacement
-was stochastic and followed the lottery recruitment rules, implying a
-recruitment eased for the first recruited species but then increasingly
-hampered by the emergence of interspecific competition
-\citep{Busing2002}. Along time the recovery of infrequent species was
-increasingly slow, so that the time for the full recovery of the
-functional redundancy, in some communities just initiated after 30
-years, was extremely difficult to estimate
-\citep{Elmqvist2003, Diaz2005}.
+pioneers were functionally different from the pre-disturbance functional
+composition. Progressively though, first established species were
+replaced by more competitive long-lived pioneers or late-successional
+species resembling more the pre-disturbance functional composition and
+restoring the functional redundancy. This replacement was stochastic and
+followed the lottery recruitment rules, implying a recruitment eased for
+the first recruited species but then increasingly hampered by the
+emergence of interspecific competition \citep{Busing2002}. Along time
+the recovery of infrequent species was increasingly slow, so that the
+time for the full recovery of the functional redundancy, in some
+communities just initiated after 30 years, was extremely difficult to
+estimate \citep{Elmqvist2003, Diaz2005}.
 
@@
 richness and, to some extent, taxonomic evenness but disproved regarding
-community functional richness and evennes that were enhanced for any
+community functional richness and evenness that were enhanced for any
 disturbance intensity by the high functional differences of pioneers
-compared to late-successional functoinal composition. The IDH was
+compared to late-successional functional composition. The IDH was
 translated in time by the recruitment, beyond an intensity threshold, of
-short-lived pionneers that prevented in the first times the
-establishment of more diverse long-lived pionneers, recruited otherwise
-below the intensity threshold. The resilience of tropical forests proved
-consistent although several decades-long. Still, the disturbance impact
-on communities redundancy cautioned against the risks of infrequent
-species loss and the persistence of disturbance-specific communities.
+short-lived pioneers that prevented in the first times the establishment
+of more diverse long-lived pioneers, recruited otherwise below the
+intensity threshold. The resilience of tropical forests, in the sense of
+the pre-disturbance state recovery, proved tangible but requiring
+several decades. Still, the disturbance impact on communities redundancy
+cautioned against the risks of infrequent species loss and the
+persistence of disturbance-specific communities \citep{Herault2018}.
+
+\section{Acknowledgement}\label{acknowledgement}
+
+We are in debt with all technicians and colleagues who helped setting up
+the plots and collecting data over years. Without their precious work,
+this study would have not been possible and they may be warmly thanked
+here.
+
+\section{Author's contributions}\label{authors-contributions}
+
+AM, EM \& BH designed the study, developed the analysis framework,
+interpreted the results and wrote the manuscript. All authors gave final
+approval for publication.
+
+\section{Data availability}\label{data-availability}
+
+This article is based upon the dataset of the Paracou station, which is
+part of the Guyafor permanent plot network in French Guiana
+(Cirad-CNRS-ONF). The dataset is available upon request to the
+scientific director (https://paracou.cirad. fr).
 

### 2018-08-13 WholePlotTrajectories_SuppMat.Rmd
mise en page

@@
     email: ariane.mirabel@ecofog.gf
-    url: http://www.ecofog.gf/spip.php?article47
-journalinfo: Hal 00679993
-date: "`r format(Sys.time(), '%d %B %Y')`"
+    url: https://github.com/ArianeMirabel
 preamble: >
@@
 
-The following information are available for this article:
+\onecolumn
 
 **Figure S1**: Whenever the family or genus were recorded for a vernacular name, the pool of species sampled to replace the vernacular name was restricted to those pertaining to the same higher taxonomic level (i.e species of the same genus when the species is missing, species of the same family when the genus is missing) \@ref(fig:FigBayesianScheme).
+
+```{r FigBayesianScheme,out.width='100%',fig.cap="Scheme of the Bayesian process propagating the taxonomic uncertainty",echo=F, fig.env='figure',fig.pos='h'}
+knitr::include_graphics("ExternalFig/BayesianSheme.png")
+```
 
@@
 Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3 .
-
-```{r FigBayesianScheme,out.width='100%',fig.cap="Scheme of the Bayesian process propagating the taxonomic uncertainty",echo=F, fig.env="figure*"}
-knitr::include_graphics("ExternalFig/BayesianSheme.png")
-```
 
@@
 ```{r FigSeedMassClasses, echo=FALSE, out.width='80%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions over 30 years after disturbance.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).", fig.env="figure*"}
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).", fig.env='figure',fig.pos='h'}
 load("DB/Smass_ClassProp")
@@
 ```
+
+**Figure S3**: Functional PCA
+
+```{r PCA_graphs, out.width = '100%', echo=FALSE, fig.cap="Summarizing plots of the functional PCA analysis defining communities functional space.
+(ref:Bolda) Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
+(ref:Boldb) bow plot of the PCA eignevalues in percentage.
+(ref:Boldc) Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).", fig.env='figure',fig.pos='h'}
+load("DB/FunctionalPCA")
+
+plotPCA(DataACP)
+
+```

### 2018-08-13 WholePlotTrajectories_SuppMat.tex
mise en page

@@
 
-\JournalInfo{Hal 00679993} % Journal information
+\JournalInfo{\ }
 \Archive{\ }
@@
 }
-\affiliation{*\textbf{}: ariane.mirabel@ecofog.gf, http://www.ecofog.gf/spip.php?article47} % Corresponding author
+\affiliation{*\textbf{}: ariane.mirabel@ecofog.gf, https://github.com/ArianeMirabel} % Corresponding author
 
@@
 
+\onecolumn
+
 The following information are available for this article:
@@
 
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\linewidth]{ExternalFig/BayesianSheme}
+
+}
+
+\caption{Scheme of the Bayesian process propagating the taxonomic uncertainty}\label{fig:FigBayesianScheme}
+\end{figure}
+
 \textbf{Figure S2}: Trajectories of seed mass classes proportions over
@@
 .
-
-\begin{figure*}
-
-{\centering \includegraphics[width=1\linewidth]{ExternalFig/BayesianSheme}
-
-}
-
-\caption{Scheme of the Bayesian process propagating the taxonomic uncertainty}\label{fig:FigBayesianScheme}
-\end{figure*}
 
@@
 
-\begin{figure*}
+\begin{figure}[h]
 
@@
 \caption{Trajectories of seed mass classes proportions over 30 years after disturbance. Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).}\label{fig:FigSeedMassClasses}
-\end{figure*}
+\end{figure}
+
+\textbf{Figure S3}: Functional PCA
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_SuppMat_files/figure-latex/PCA_graphs-1}
+
+}
+
+\caption{Summarizing plots of the functional PCA analysis defining communities functional space.
+\textbf{(a)} Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
+\textbf{(b)} bow plot of the PCA eignevalues in percentage.
+\textbf{(c)} Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).}(\#fig:PCA_graphs)
+\end{figure}
 

### 2018-08-13 WholePlotTrajectories.Rmd
Final version 1
mise en page

@@
     authsuperscript: 1*
+  - name: Bruno Hérault
+    authsuperscript: 2
   - name: Eric Marcon
     authsuperscript: 1
-  - name: Bruno Hérault
-    authsuperscript: 2 3
 affiliation:
@@
   - affsuperscript: 2
-    dptuniv: Cirad, Univ montpellier, UR Forests & Societies.
-    address: >
-      Montpellier,
-      France.
-  - affsuperscript: 3
-    dptuniv: INPHB, Institut National Polytechnique Félix Houphouet-Boigny
+    dptuniv: INPHB, Institut National Polytechnique Félix Houphoüet-Boigny
     address: >
@@
     email: ariane.mirabel@ecofog.gf
-    url: http://www.ecofog.gf/spip.php?article47
+    url: https://github.com/ArianeMirabel
 abstract: >   
@@
  The huge diversity of tropical forests is often assumed to be regularly reshaped by natural disturbance yielding a diversity peak at intermediate intensity.
-This intermediate disturbance hypothesis (IDH), though, remains debated and the controversy questions the extent of communities' resilience regarding their taxonomic and functional facets.
-
-    To disentangle the ecological processes driving community response to disturbance, we analysed the tree community trajectories over 30 years following a disturbance gradient in a Neotropical forest.
+This intermediate disturbance hypothesis (IDH), though, remains debated and the controversy questions the extent of communities’ resilience regarding their taxonomic and functional facets.
+ To disentangle the ecological processes driving community response to disturbance, we analyzed the tree community trajectories over 30 years following a disturbance gradient in a Neotropical forest.
 Specifically, we examined community functional and taxonomic trajectories with regards to diversity, composition and redundancy.
@@
 This acknowledged the need of decades-long cycles with no disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbances.
-
-keywords: [Taxonomic and Functional Biodiversity, Neotropical Forests, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience]
-journalinfo: Hal xxx
-archive: DOI xxxx
-date: "`r format(Sys.time(), '%d %B %Y')`"
+keywords: [Community Ecology, Determinants of Plant Community Diversity and Structure, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
 preamble: >
@@
 output:
+  bookdown::gitbook: default
+  bookdown::word_document2: default
   bookdown::pdf_book:
@@
     keep_tex: TRUE
-  bookdown::gitbook: default
-  bookdown::word_document2: default
 ---
@@
 For now, this has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016] that are rapid and convenient to measure.
-These structural parameters have been sucessfully modeled, giving important insights into the recovery of ecosystem processes and services [@Herault2018]. 
+These structural parameters have been successfully modeled, giving important insights into the recovery of ecosystem processes and services [@Herault2018]. 
 However the response of forests diversity and composition remains unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014; @Liang2016].
@@
 Furthermore, a functional approach accounting for species biological attributes would directly link communities diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Baraloto2012a].
-In that respect, the functional trait-based approach that focus on major traits related to species ecology and mediate species performance in a given environment was sucessfully adopted [@Diaz2005].
+In that respect, the functional trait-based approach that focus on major traits related to species ecology and mediate species performance in a given environment was successfully adopted [@Diaz2005].
 For instance, the functional approach revealed in tropical rainforests the deterministic processes entailing, after disturbance, a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Reich2014;@Herault2011].
 This shift is translated into the trajectories of key functional traits related to resource acquisition (leaf and stem traits) and life-history traits (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
-Eventually a complete overview of communities response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
-The high functional redundancy of hyperdiverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Elmqvist2003; @Diaz2005].
+Eventually a complete overview of community response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
+The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Elmqvist2003; @Diaz2005].
 
-In this study, we monitored over 30 years the response of 75 ha of neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem above-ground biomass (AGB) loss. 
+In this study, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem above-ground biomass (AGB) loss. 
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
@@
   Treatment=c("Control", "T1", "T2", "T3"),
-  Timber=c("", 
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha", 
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha",
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10 trees/ha"),
-  Thinning=c("", 
-             "", 
-             "DBH $\\geq$ 40 cm, non-valuable species, $\\approx$ 30 trees/ha", 
-             "DBH $\\geq$ 50 cm, non-valuable species, $\\approx$ 15 trees/ha  "),
-  Fuelwood=c("", 
-             "", 
-             "", 
-             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-valuable species, $\\approx$ 15 trees/ha"),
-  AGBLost=c(0, "$[12\\%-33\\%]$", "$[33\\%-56\\%]$", "$[35\\%-56\\%]$")
+  Timber=c("-", 
+           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10   $trees.ha^{-1}$", 
+           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10  $trees.ha^{-1}$",
+           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10  $trees.ha^{-1}$"),
+  Thinning=c("-", 
+             "-", 
+             "DBH $\\geq$ 40 cm, non-valuable species, $\\approx$ 30   $trees.ha^{-1}$", 
+             "DBH $\\geq$ 50 cm, non-valuable species, $\\approx$ 15  $trees.ha^{-1}$  "),
+  Fuelwood=c("-", 
+             "-", 
+             "-", 
+             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-valuable species,\\ $\\approx$ 15 $trees.ha^{-1}$"),
+  AGBLost=c(0, "$[12-33]$", "$[33-56]$", "$[35-56]$")
 )
@@
 
-See Supplementary Materials -Figure S1 and @Aubry-Kientz2013 for the detailed methodology.
+See Supplementary Materials -Fig. S1 and @Aubry-Kientz2013 for the detailed methodology.
 
@@
 Two NMDS using abundance-based (Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the later based on the 7 leaf, stem and life history traits (without seed mass classes).
-Trajectories along time were reported through the euclidean distance between the target inventories and the reference inventories in 1989, *i.e.* 2 years after disturbance. 
+Trajectories along time were reported through the Euclidean distance between the target inventories and the reference inventories in 1989, *i.e.* 2 years after disturbance. 
 Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007]. 
@@
 The taxonomic diversity was reported through species richness and evenness, *i.e* the Hill number translation of the Simpson index [@Hill1973].
-These indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recomended for diversity studies [@Marcon2015b]. 
+These indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recommended for diversity studies [@Marcon2015b]. 
 The functional diversity was reported using the functional richness and functional evenness, *i.e* Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on species functional traits.
@@
 The impacts of initial disturbance were tested with the spearman rank correlation between the extrema of taxonomic and functional metrics reached over the 30 years and the initial %AGB loss. 
-They were besides analysed through polynomial regression between (i) taxonomic and functional richness and evenness and (ii) the initial %AGB loss at 10, 20 and 30 years after disturbance.
+They were besides analyzed through polynomial regression between (i) taxonomic and functional richness and evenness and (ii) the initial %AGB loss at 10, 20 and 30 years after disturbance.
 
@@
 Mat.
-- Figure S1 for a more comprehensive sheme).
+- Fig.
+S1 for a more comprehensive scheme).
 
@@
 
-While both taxonomic and functional composition remained stable in undisturbed communities (Figure \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-\break disturbance time.
+While both taxonomic and functional composition remained stable in undisturbed communities (Fig.
+\@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-\break disturbance time.
 In disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WSG to high average SLA and chlorophyll content (see appendix I).
-For functional composition, this translated into cyclic compositional changes with an unachieved recovery of the initial composition (Figure \@ref(fig:NMDSplans)).
-The maximum dissimilarity with the initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{taxonomic}=0.87$ and $\rho_{spearman}^{functional}=0.90$ respectively).
+For functional composition, this translated into cyclic compositional changes with an unachieved recovery of the initial composition (Fig.
+\@ref(fig:NMDSplans)).
+The maximum dissimilarity with the initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{Taxonomic}=0.87$ and $\rho_{spearman}^{Functional}=0.90$ respectively).
 The maximum value was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition.
@@
 ```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS space.
-Lower panels ((ref:Boldc) and (ref:Boldd)) represent the euclidean distance to initial condition along the 30 sampled years.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure*"}
+Lower panels ((ref:Boldc) and (ref:Boldd)) represent the Euclidean distance to initial condition along the 30 sampled years.
+Shaded areas are the credibility intervals.", fig.env="figure*"}
 load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
@@
 mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=0.8)
-legend("right",inset=c(-0.345,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
+legend("right",inset=c(-0.32,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
 
@@
 
-Maximum height at adult stage (*Hmax*), leaf toughness (*L_toughness*) and wood specific gravity (*WSG*) first decreased and then slightly increased but remained significantly lower than their initial value (Figure \@ref(fig:CWM)).
-On the other side, Bark thickness (*Bark_thick*) and specific leaf area (*SLA*) increased and while *Bark_thick* remained substantially high after 30 years, *SLA* had almost recovered its initial value.
-For all traits, the maximum difference to initial value was correlated to the disturbance intensity ($\rho_{spearman}^{L_{thickness}}=0.76$, $\rho_{spearman}^{L_{chloro}}=0.60$, $\rho_{spearman}^{L_{toughness}}=-0.53$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WSG}=-0.75$, $\rho_{spearman}^{Bark-thickness}=0.71$, $\rho_{spearman}^{Hmax}=-0.40$).
+Maximum height at adult stage (*Hmax*), leaf toughness and wood specific gravity (*WSG*) first decreased and then slightly increased but remained significantly lower than their initial value (Fig.
+\@ref(fig:CWM)).
+On the other side, bark thickness and specific leaf area (*SLA*) increased and while bark thickness remained substantially high after 30 years, *SLA* had almost recovered its initial value.
+For all traits, the maximum difference to initial value was correlated to the disturbance intensity ($\rho_{spearman}^{Leaf thickness}=0.76$, $\rho_{spearman}^{Chlorophyll content}=0.60$, $\rho_{spearman}^{Leaf toughness}=-0.53$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WSG}=-0.75$, $\rho_{spearman}^{Bark thickness}=0.71$, $\rho_{spearman}^{Hmax}=-0.40$).
 The proportions of the three lightest seed mass classes increased in all disturbed plots, and decreased after 30 years for the lightest class while it stabilized for the two other (Supp.
 Mat.
-- Figure S2).
+- Fig.
+S2).
 
-```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, (ref:capThickness), chlorophyll content, (ref:capChloro), toughness, (ref:capThoughness) and specific area, (ref:capSLA)), 2 stem traits (wood specific gravity, (ref:capWSG), and bark thickness, (ref:capBarkThick)) and one life history trait (Specific maximum height at adult stage, (ref:capHmax)).
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure*"}
+```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means over 30 years after disturbance of four leaf traits (Leaf thickness, chlorophyll content, toughness, and specific area), two stem traits (wood specific gravity, and bark thickness) and one life history trait (Specific maximum height at adult stage).
+", fig.env="figure*"}
 load("DB/CWM")
@@
 legendCWM()
-
+legend("right",inset=c(-1,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
 ```
@@
 ```{r DivTaxo, out.width = '100%', echo=FALSE, fig.cap="Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic (ref:Bolda) richness, (ref:Boldb), taxonomic evenness, (ref:Boldc) functional richness, and (ref:Boldd) functional evenness.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals ", fig.env="figure*"}
+Shaded areas are the credibility intervals ", fig.env="figure*"}
 load("DB/ReplacementTraj_ForGraphs")
@@
 
-par(mfrow=c(2,2),mar=c(2,2.5,2.5,1),oma=c(1,2,1,3),no.readonly = T)
+par(mfrow=c(2,2),mar=c(2,2,2.5,5),oma=c(1,2,1,1),no.readonly = T)
 TaxoTraj(CompleteTaxo)
@@
 mtext("Years since disturbance",side=1,line=2.2,adj=1)
-legend("right",inset=c(-0.3,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
+legend("right",xpd=NA,inset=-0.4,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
 
@@
 In disturbed plots, functional richness and evenness trajectories depended on the disturbance intensity with their maximum positively correlated to %AGB loss $\rho_{spearman}^{Richness}=0.76$ and $\rho_{spearman}^{Evenness}=0.60$.
-Functional richness and evenness displayed for low disturbance intensity a low but long-lasting increase up to a maximum reached after 20-25 years, and for high intensity, a fast but short increase followed after 10 years by a slow decrease towards the inital values.
+Functional richness and evenness displayed for low disturbance intensity a low but long-lasting increase up to a maximum reached after 20-25 years, and for high intensity, a fast but short increase followed after 10 years by a slow decrease towards the initial values.
 
-The second-degree polynomial regressions between (i) the %AGB loss and (ii) taxonomic and functional richness and evenness after 10, 20 and 30 years best predicted the hump-shaped curve of the disturbance impact along the disturbance intensity gradient \@ref(fig:IDHplot).
+The second-degree polynomial regressions between (i) the percentage AGB loss and (ii) taxonomic and functional richness and evenness after 10, 20 and 30 years best predicted the hump-shaped curve of the disturbance impact along the disturbance intensity gradient \@ref(fig:IDHplot).
 The relationship between the disturbance impact and its intensity was more markedly hump-shaped for the taxonomic richness than for the taxonomic evenness.
@@
 
-```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/4,fig.cap="Relationship between the initial \\%AGB loss and community taxonmic richness (ref:Bolda), taxonomic evenness (ref:Boldb), functional richness (ref:Boldc),and functional evenness (ref:Boldd) at 10, 20 and 30 years after disturbance", fig.env="figure*"}
+```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/4,fig.cap="Relationship between the initial \\%AGB loss and community taxonomic richness (ref:Bolda), taxonomic evenness (ref:Boldb), functional richness (ref:Boldc),and functional evenness (ref:Boldd) at 10, 20 and 30 years after disturbance", fig.env="figure*"}
 load("DB/RichnessIDH");load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB");load("DB/FunRichIDH")
 
-par(mfrow=c(1,4),mar=c(2,2.5,3,1),oma=c(1,1,1,2.5),no.readonly=TRUE)
+par(mfrow=c(1,4),mar=c(2,2.5,3,0),oma=c(1,1,1,4),no.readonly=TRUE)
 plotIDH(Data=Rich,AgbLoss=AGBloss)
@@
 mtext("initial AGB loss (%)",side=1,line=2.2,adj=1,cex=0.8)
-legend("right",inset=c(-0.5,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
+legend("right",inset=-0.5,xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
 
@@
 ```{r RedFunRest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
-Colors are disturbance treatments: green (control), yellow (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
+Shaded areas are the credibility intervals.", fig.env="figure"}
 load("DB/Redundancy_restricted")
@@
 
-Communities taxonomic and functional composition appeared resilient, following similar hump-shaped trajectories starting to return towards pre-disturbance composition after 30 years.
+Community taxonomic and functional composition appeared resilient, following similar hump-shaped trajectories starting to return towards pre-disturbance composition after 30 years.
 
@@
 More than commonly thought, post-disturbance trajectories depended on community initial composition, that partly determined the pool of recruited species and constrained the trajectories towards the initial composition.
-The high resilience of communities taxonomy revealed that species not belonging to the pre-disturbance community were hardly recruited because of the commonness of dispersal limitation among tropical tree species [@Svenning2005].
+The high resilience of community taxonomy, in the sense of the initial state recovery, revealed that species not belonging to the pre-disturbance community were hardly recruited because of the commonness of dispersal limitation among tropical tree species [@Svenning2005].
 
@@
 Competitive pioneers became dominant in filling the environmental niches of high availability of light, space and nutrients vacated by the disturbance.
-The recruitment of pioneers changed community functional composition in the same way for all disturbance intensity towards more resource-acquisitive strategies, moving community functional composition right along the first axis in Figure \@ref(fig:NMDSplans) [@Westoby1998; @Wright2004; @Reich2014].
-Thereafter long-lived, more resistant and shade-tolerant species excluded the first established pioneers and started the recovery of pre-disturbance functional composition, moving similarly community functional composition left along the first axis and upward along the second axis in Figure \@ref(fig:NMDSplans).
+The recruitment of pioneers changed community functional composition in the same way for all disturbance intensity towards more resource-acquisitive strategies, moving community functional composition right along the first axis in Fig.
+\@ref(fig:NMDSplans) [@Westoby1998; @Wright2004; @Reich2014].
+Thereafter long-lived, more resistant and shade-tolerant species excluded the first established pioneers and started the recovery of pre-disturbance functional composition, moving similarly community functional composition left along the first axis and upward along the second axis in Fig.
+\@ref(fig:NMDSplans).
 
@@
 For community taxonomic evenness the disturbance impact was similar but slighter, as the evenness is less sensitive to the loss of rare species.
-Taxonomic evenness rather represented the increasing dominance of pionneers that balanced the usual hyper-dominance of a few species in tropical forests below the intensity threshold, thus increasing community overall evenness up to the intensity threshold beyond which pioneers became in turn highly dominant and decreased the overall evenness [@Baraloto2012a].
+Taxonomic evenness rather represented the increasing dominance of pioneers that balanced the usual hyper-dominance of a few species in tropical forests below the intensity threshold, thus increasing community overall evenness up to the intensity threshold beyond which pioneers became in turn highly dominant and decreased the overall evenness [@Baraloto2012a].
 
 Conversely the IDH was disproved regarding the disturbance impact on community functional richness and evenness.
-Irrespective of the disturbance intensity the recruitment of pionneers, functionally highly different from the composition of pre-disturbance community, increased both community functional richness and eveness.
+Irrespective of the disturbance intensity the recruitment of pioneers, functionally highly different from the composition of pre-disturbance community, increased both community functional richness and evenness.
 
@@
 Up to an intensity threshold, the species gain was all the more significant that the disturbance intensity increased, with the establishment of long-lived pioneers enhancing community taxonomic richness and evenness in the long term.
-These long-lived pionneers, functionally quite different from the functional composition, entailed as well a progressive and long-lasting increase of the functional richness and evenness [@Denslow1980;@Molino2001].
+These long-lived pioneers, functionally quite different from the functional composition, entailed as well a progressive and long-lasting increase of the functional richness and evenness [@Denslow1980;@Molino2001].
 Beyond an intensity threshold, though, a few short-lived pioneers occupied the vacated environmental space and prevented the establishment of other species. 
-These short-lived pioneers were functionally very different from the pre-disturbance community and entailed a rapid and significant increase of functional richness en evenness.
+These short-lived pioneers were functionally very different from the pre-disturbance community and entailed a rapid and significant increase of functional richness and evenness.
 Already after 10 years, though, short-lived pioneers started to decline and the functional richness and evenness decreased.
@@
 For 15 years the species loss during disturbance, determined by the disturbance intensity, commensurately decreased the functional redundancy within the pre-disturbance functional space.
-The redundancy decrease was not compensated in the first place as the first recruited pionneers were functionally different from the pre-disturbance functional composition.
-Progressively though, first established species were replaced by more competitive long-lived pionneers or late-successional species resembling more the pre-disturbance functional composition and restoring the functional redundancy.
+The redundancy decrease was not compensated in the first place as the first recruited pioneers were functionally different from the pre-disturbance functional composition.
+Progressively though, first established species were replaced by more competitive long-lived pioneers or late-successional species resembling more the pre-disturbance functional composition and restoring the functional redundancy.
 This replacement was stochastic and followed the lottery recruitment rules, implying a recruitment eased for the first recruited species but then increasingly hampered by the emergence of interspecific competition [@Busing2002].
@@
 Our study revealed community recovery through the combination of deterministic processes driving their convergence in functional composition, and dispersal limitation maintaining their divergence in taxonomic composition.
-The IDH was validated for community taxonomic richness and, to some extent, taxonomic evenness but disproved regarding community functional richness and evennes that were enhanced for any disturbance intensity by the high functional differences of pioneers compared to late-successional functoinal composition.
-The IDH was translated in time by the recruitment, beyond an intensity threshold, of short-lived pionneers that prevented in the first times the establishment of more diverse long-lived pionneers, recruited otherwise below the intensity threshold.
-The resilience of tropical forests proved consistent although several decades-long. 
+The IDH was validated for community taxonomic richness and, to some extent, taxonomic evenness but disproved regarding community functional richness and evenness that were enhanced for any disturbance intensity by the high functional differences of pioneers compared to late-successional functional composition.
+The IDH was translated in time by the recruitment, beyond an intensity threshold, of short-lived pioneers that prevented in the first times the establishment of more diverse long-lived pioneers, recruited otherwise below the intensity threshold.
+The resilience of tropical forests, in the sense of the pre-disturbance state recovery, proved tangible but requiring several decades. 
 Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].
 
+# Acknowledgement
+
+We are in debt with all technicians and colleagues who helped setting up the plots and collecting data over years.
+Without their precious work, this study would have not been possible and they may be warmly thanked here.
+
+# Author's contributions
+
+AM, EM & BH designed the study, developed the analysis framework, interpreted the results and wrote the manuscript. All authors gave final approval for publication.
+
+# Data availability
+
+This article is based upon the dataset of the Paracou station, which is part of the Guyafor permanent plot network in French Guiana (Cirad-CNRS-ONF).
+The dataset is available upon request to the scientific director (https://paracou.cirad.
+fr).
+
 `r if (knitr:::is_html_output()) '# References {-}'`

### 2018-08-13 WholePlotTrajectories.tex
mise en page

@@
 \Authors{
-Ariane MIRABEL\textsuperscript{1*}\\ Eric Marcon\textsuperscript{1}\\ Bruno Hérault\textsuperscript{2 3}
+Ariane MIRABEL\textsuperscript{1*}\\ Bruno Hérault\textsuperscript{2}\\ Eric Marcon\textsuperscript{1}
 } % Authors
@@
 \textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles,
-Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}Cirad, Univ montpellier, UR Forests \& Societies.\\ \hspace{1em} Montpellier, France.\\\textsuperscript{3}INPHB, Institut National Polytechnique Félix Houphoüet-Boigny\\ \hspace{1em} Yamoussoukro, Ivory Coast.
+Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}INPHB, Institut National Polytechnique Félix Houphoüet-Boigny\\ \hspace{1em} Yamoussoukro, Ivory Coast.
 }
@@
 \midrule
-Control &  &  &  & 0\\
-T1 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha &  &  & $[12\%-33\%]$\\
-T2 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30 trees/ha &  & $[33\%-56\%]$\\
-T3 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & $[35\%-56\%]$\\
+Control & - & - & - & 0\\
+T1 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10   $trees.ha^{-1}$ & - & - & $[12-33]$\\
+T2 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30   $trees.ha^{-1}$ & - & $[33-56]$\\
+T3 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15  $trees.ha^{-1}$ & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species,\ $\approx$ 15 $trees.ha^{-1}$ & $[35-56]$\\
 \bottomrule

### 2018-10-18 WholePlotTrajectories.Rmd
Jecol review

@@
     authsuperscript: 1*
-  - name: Bruno Hérault
+  - name: Bruno Herault
     authsuperscript: 2
@@
     url: https://github.com/ArianeMirabel
-abstract: >   
+abstract: 
  Understanding the ecological rules underlying the maintenance of tropical forests biodiversity, structure, functioning and dynamics is urgent to anticipate their fate in the global change context.
  The huge diversity of tropical forests is often assumed to be regularly reshaped by natural disturbance yielding a diversity peak at intermediate intensity.
-This intermediate disturbance hypothesis (IDH), though, remains debated and the controversy questions the extent of communities’ resilience regarding their taxonomic and functional facets.
+This intermediate disturbance hypothesis (IDH), though, remains debated and the controversy questions the extent of community resilience regarding their taxonomic and functional facets.
  To disentangle the ecological processes driving community response to disturbance, we analyzed the tree community trajectories over 30 years following a disturbance gradient in a Neotropical forest.
@@
 While pre-disturbance taxonomic differences were maintained over time, functional composition trajectories were quite similar among communities.
-The IDH did predict communities taxonomic diversity response while functional diversity was enhanced whatever the disturbance intensity.
+The IDH did predict community taxonomic diversity response while functional diversity was enhanced whatever the disturbance intensity.
  Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years.
 This acknowledged the need of decades-long cycles with no disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbances.
-keywords: [Community Ecology, Determinants of Plant Community Diversity and Structure, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
+keywords: [Community Ecology, Community Diversity Determinants, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
 preamble: >
@@
 They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Morales-Hidalgo2015] and threatens the maintenance and dynamics in space and time of communities structure, composition and functioning [@Anderson-Teixeira2013].
+The growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Morales-Hidalgo2015] and threatens the maintenance and dynamics in space and time of community structure, composition and functioning [@Anderson-Teixeira2013].
 
@@
 However the response of forests diversity and composition remains unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014; @Liang2016].
-In the short-term, moderate disturbance may lead to positive impacts on communities diversity, an idea formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity when disturbance intensity is not too high [@Molino2001;@Kariuki2006a].
+In the short-term, moderate disturbance may lead to positive impacts on community diversity, an idea formalized by the intermediate disturbance hypothesis (IDH) which predicts a maximum species diversity when disturbance intensity is not too high [@Molino2001;@Kariuki2006a].
 
@@
 Taxonomic richness alone, however, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016].
-More ecological-meaningful analysis would couple richness with (i) evenness, that would reveal the changes in the species abundance distribution and thus the underlying ecological processes, and (ii) composition that is crucial for conservation issues [@Lavorel2002;@Bellwood2006].
-Furthermore, a functional approach accounting for species biological attributes would directly link communities diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Baraloto2012a].
+More ecological-meaningful analysis would couple richness with (i) evenness, which would reveal the changes in the species abundance distribution and thus the underlying ecological processes, and (ii) composition, which is crucial for conservation issues [@Lavorel2002;@Bellwood2006].
+Furthermore, a functional approach accounting for species biological attributes would directly link community diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Baraloto2012a].
 In that respect, the functional trait-based approach that focus on major traits related to species ecology and mediate species performance in a given environment was successfully adopted [@Diaz2005].
@@
 This shift is translated into the trajectories of key functional traits related to resource acquisition (leaf and stem traits) and life-history traits (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
-Eventually a complete overview of community response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
+Eventually, a complete overview of community response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
 The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Elmqvist2003; @Diaz2005].
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, we (i) elucidated community taxonomic and functional recovery and the underlying ecological processes, (ii) clarified the validity of the IDH in the long term for tropical forest and its translation into different trajectories in time, and (iii) questioned community recovery time.
+Specifically, we (i) elucidated community taxonomic and functional recovery and the underlying ecological processes, (ii) clarified the validity of the IDH in the long term for tropical forest and its translation into different trajectories in time, and (iii) examined community recovery time.
 
@@
 Paracou station in French Guiana (5\textdegree 18'N and 52\textdegree 53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26\textdegree C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
-Elevation ranges between 5 and 50 m and soils correspond to thin acrisols over a layer of transformed saprolite with low permeability generating lateral drainage during heavy rains.
+Elevation ranges between 5 and 50 m and soils correspond to thin acrisols over a layer of transformed saprolite with low permeability which produces lateral drainage during heavy rains.
 
-The experiment is a network of twelve 6.25ha plots that underwent a disturbance gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) according to a randomized plot design with three replicate blocks of four plots.
-The disturbance corresponds to averages of 10 trees removed per hectare with a diameter at 1.3 m height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3 (T3).
-Treatments T2 and T3 besides included the thinning of trees by poison girdling [@Schmitt1990]. 
-The disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Biomass2018].
+The experiment is a network of 12 6.25 ha plots that underwent a disturbance gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) according to a randomized plot design with three replicate blocks of four plots.
+The disturbance corresponds to an intensity gradient.
+For treatment 1 (T1) 10 trees of commercial species (of a diameter at 1.3 m height (DBH) equal or above 50 cm) were felled per hectare.
+For treatment 2 (T2) 10 trees/ha of commercial species (DBH  $\geq$ 50 cm) were felled and 30 trees/ha of non-valuable species (DBH $\geq$ 40 cm) were removed by poison girdling.
+For treatment 3 (T3) 10 trees/ha of commercial species (DBH $\geq$ 50 cm) were felled and 30 trees/ha of non-valuable species (15 with DBH $\geq$ 50 cm and 15 with DBH $\geq$ 40 cm) were removed by poison girdling [@Schmitt1990].
+Disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Biomass2018].
 
@@
 
-The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae. 
-In the twelve experimental plots, all trees above 10 cm DBH have been mapped and measured annually since 1984. 
+The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of *Fabaceae*, *Chrysobalanaceae*, *Lecythidaceae* and *Sapotaceae*. 
+In the 12 experimental plots, all trees above 10 cm DBH have been mapped and measured annually since 1984. 
 Trees are first identified with a vernacular name assigned by the forest worker team, and afterward with a scientific name assigned by botanists during regular botanical campaigns. 
-In 1984, specific vernacular names were given to 62 commercial or common species whereas more infrequent ones were identified under general identifiers only distinguishing trees and palm trees. 
+In 1984, specific vernacular names were given to 62 commercial or common species whereas more infrequent ones were identified under general identifiers only distinguishing trees and palms. 
 From 2003, botanical campaigns have been conducted every 5 to 6 years to identify all trees at the species level but identification levels still varied among plots and campaigns.
@@
 This variability of protocols in time raised methodological issues as vernacular names usually correspond to different botanical species. 
-It resulted in significant taxonomic uncertainties that had to be propagated to composition and diversity metrics.
+This resulted in significant taxonomic uncertainty that had to be propagated to composition and diversity metrics.
 The uncertainty propagation was done through a Bayesian framework reconstituting complete inventories at genus level from real incomplete ones on the basis of vernacular/botanical names association. 
@@
 
-Six functional traits representing leaf economics (leaves thickness, toughness, total chlorophyll content and specific leaf area) and stem economics (wood specific gravity and bark thickness), and life-history traits (maximum specific height and seed mass) came from the BRIDGE project [^1].
+Six functional traits representing leaf economics (leaf thickness, toughness, total chlorophyll content and specific leaf area) and stem economics (wood specific gravity and bark thickness), and life-history traits (maximum specific height and seed mass) came from the BRIDGE project [^1].
 Trait values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 species pertaining to 157 genera.
@@
 
-To counter taxonomic uncertainties due to the variability of botanical identification levels (in space) and protocols (in time), the taxonomic composition and diversity analysis were conducted at the genus level.
+Because of the variability in the precision of botanical identification efforts, we were constraint to conduct the taxonomic composition and diversity analysis at the genus level. 
 Taxonomic and functional trajectories of community composition were followed in a two-dimensional NMDS ordination space.
-Two NMDS using abundance-based (Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the later based on the 7 leaf, stem and life history traits (without seed mass classes).
+Two NMDS using abundance-based (Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the latter based on the seven leaf, stem and life history traits (without seed mass classes).
 Trajectories along time were reported through the Euclidean distance between the target inventories and the reference inventories in 1989, *i.e.* 2 years after disturbance. 
 Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007]. 
-Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were therefore reported as the proportion of each class in the inventories (Supplementary materials).
+Species seed mass were given in 5 mass classes.
+Seed mass trajectories were therefore reported as the proportion of each class in the inventories (Supplementary materials).
 
-The taxonomic diversity was reported through species richness and evenness, *i.e* the Hill number translation of the Simpson index [@Hill1973].
-These indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recommended for diversity studies [@Marcon2015b]. 
-The functional diversity was reported using the functional richness and functional evenness, *i.e* Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on species functional traits.
+The taxonomic diversity was reported through species richness and the Hill number translation of the Simpson index [@Hill1973].
+The comparison between richness and Simpson diversity assessed the variations in taxonomic evenness. 
+Both indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recommended for diversity studies [@Marcon2015b]. 
+The functional diversity was reported using the functional richness and functional evenness, *i.e* the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on species functional traits.
 
-The impacts of initial disturbance were tested with the spearman rank correlation between the extrema of taxonomic and functional metrics reached over the 30 years and the initial %AGB loss. 
+The impacts of initial disturbance were tested with the Spearman rank correlation between the extremes of taxonomic and functional metrics reached over the 30 years and the initial %AGB loss. 
 They were besides analyzed through polynomial regression between (i) taxonomic and functional richness and evenness and (ii) the initial %AGB loss at 10, 20 and 30 years after disturbance.
 
-The functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
-The samples of the trait database were first mapped in a 2-dimensional plan with a PCA analysis. 
-Then, multivariate kernel density estimator associated with individual trees returned species traits probability distribution (TDP).
-Species TDP weighted by species abundance were eventually summed for each community.
-Community functional redundancy was the sum of TDPs overlap, expressed as the average number of species that could be removed from without reducing the functional space (Supp.
-Mat.
-- Fig.
-S1 for a more comprehensive scheme).
+Functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
+First, the individuals of the trait database were mapped in the plane of the two first axes from a PCA analysis.
+Then, for each species, the traits probability density (TPD) were computed through multivariate kernel density estimators.
+Second, for each community, the functional traits probability density were computed as the sum of the TDP of the inventoried species weighted by species abundance.
+Third, the whole functional space was divided into a 100 x 100 grid, before counting for each cell the number of species which TPD is positive.
+The average count across cells, minus 1, gives the average number of species in the community that share the same traits, which is the Community Functional Redundancy.
 
@@
 
-## Communities Composition
+## Community Composition
 
-From 1989 (2 years after disturbance) to 2015 (28 years after disturbance), 828 388 individual trees and 591 botanical species pertaining to 223 genus and 64 families were recorded.
+From 1989 (2 years after disturbance) to 2015 (28 years after disturbance), 828 388 individual trees and 591 botanical species pertaining to 223 genera and 64 families were recorded.
 
@@
 \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-\break disturbance time.
-In disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high average WSG to high average SLA and chlorophyll content (see appendix I).
-For functional composition, this translated into cyclic compositional changes with an unachieved recovery of the initial composition (Fig.
+In disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high mean WSG to high mean SLA and chlorophyll content (see appendix I).
+For functional composition, this translated into cyclic compositional changes with an incomplete recovery of the initial composition (Fig.
 \@ref(fig:NMDSplans)).
-The maximum dissimilarity with the initial state was positively correlated to the disturbance intensity for both taxonomic and functional composition ($\rho_{spearman}^{Taxonomic}=0.87$ and $\rho_{spearman}^{Functional}=0.90$ respectively).
-The maximum value was reached around 26 years after disturbance for taxonomic composition and 22 years for functional composition.
+The maximum dissimilarity with the initial state was positively correlated with the disturbance intensity for both taxonomic and functional composition ($\rho_{Spearman}^{Taxonomic}=0.87$ and $\rho_{Spearman}^{Functional}=0.90$ respectively).
+The maximum dissimilarity with the initial was reached for taxonomic composition between 15 to 25 years, for most of the plots, and around 22 years for functional composition.
 
-```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of flora composition (left panels (ref:Bolda) and (ref:Boldc)) and functional composition (right panels (ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS space.
+```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of taxonomic composition ((ref:Bolda) and (ref:Boldc)) and functional composition ((ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS space.
 Lower panels ((ref:Boldc) and (ref:Boldd)) represent the Euclidean distance to initial condition along the 30 sampled years.
@@
 
-Except for leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance, all traits and seed mass proportions followed unimodal trajectories either stabilizing or returning towards their initial values.
+Community CWM average value of all traits and seed mass proportions followed unimodal trajectories, either stabilizing or returning towards their initial values, to the exception of leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance.
 
-Maximum height at adult stage (*Hmax*), leaf toughness and wood specific gravity (*WSG*) first decreased and then slightly increased but remained significantly lower than their initial value (Fig.
+Community CWM average value of Maximum height at adult stage (*Hmax*), leaf toughness and wood specific gravity (*WSG*) first decreased and then slightly increased but remained significantly lower than their initial value (Fig.
 \@ref(fig:CWM)).
-On the other side, bark thickness and specific leaf area (*SLA*) increased and while bark thickness remained substantially high after 30 years, *SLA* had almost recovered its initial value.
-For all traits, the maximum difference to initial value was correlated to the disturbance intensity ($\rho_{spearman}^{Leaf thickness}=0.76$, $\rho_{spearman}^{Chlorophyll content}=0.60$, $\rho_{spearman}^{Leaf toughness}=-0.53$, $\rho_{spearman}^{SLA}=0.93$, $\rho_{spearman}^{WSG}=-0.75$, $\rho_{spearman}^{Bark thickness}=0.71$, $\rho_{spearman}^{Hmax}=-0.40$).
+On the other side, bark thickness and specific leaf area (*SLA*) increased and while bark thickness remained substantially high after 30 years, *SLA* had almost recovered to its initial value.
+For all traits, the maximum difference to initial value was correlated to the disturbance intensity.
+Positive correlations were observed for Leaf thickness, chlorophyll content, SLA and bark thickness ($\rho_{Spearman}^{Leaf thickness}=0.76$, $\rho_{Spearman}^{Chlorophyll content}=0.60$, $\rho_{Spearman}^{SLA}=0.93$, $\rho_{Spearman}^{Bark thickness}=0.71$).
+Negative correlation was observed for Leaf toughness, WSG and Hmax ($\rho_{Spearman}^{Leaf toughness}=-0.53$, $\rho_{Spearman}^{WSG}=-0.75$, $\rho_{Spearman}^{Hmax}=-0.40$)
 The proportions of the three lightest seed mass classes increased in all disturbed plots, and decreased after 30 years for the lightest class while it stabilized for the two other (Supp.
@@
 
-```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the communities weighted means over 30 years after disturbance of four leaf traits (Leaf thickness, chlorophyll content, toughness, and specific area), two stem traits (wood specific gravity, and bark thickness) and one life history trait (Specific maximum height at adult stage).
+```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of community weighted means over 30 years after disturbance of four leaf traits (Leaf thickness, chlorophyll content, toughness, and specific area), two stem traits (wood specific gravity, and bark thickness) and one life history trait (Specific maximum height at adult stage).
 ", fig.env="figure*"}
@@
 
-## Communities richness and evenness
+## Community taxonomic and functional diversity
 
-For undisturbed plots, taxonomic Richness and Evenness remained stable over the 30 years monitored.
+For undisturbed plots, taxonomic richness and Simpson diversity remained stable over the 30 years of monitoring.
 In disturbed communities, after low disturbance intensity the taxonomic richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2).
 After intense disturbance the taxonomic richness followed a more complex trajectory, decreasing for ten years after disturbance before recovering to pre-disturbance values.
-The maximum richness loss or gain after disturbance was positively correlated to the disturbance intensity ($\rho_{spearman}^{Richness}=0.50$).
-In all disturbed plots the taxonomic evenness first increased until a maximum reached after around 20 years.
-This maximum was positively correlated to the disturbance intensity ($\rho_{spearman}^{Evenness}=0.77$).
-The evenness then stabilized except for two T3 plots (plots 8 and 12) for which evenness kept increasing.
+The maximum richness loss or gain after disturbance was positively correlated with the disturbance intensity ($\rho_{Spearman}^{Richness}=0.50$).
+In all disturbed plots the Simpson diversity first increased until a maximum reached after around 20 years.
+This maximum was positively correlated with the disturbance intensity ($\rho_{Spearman}^{Simpson}=0.77$).
+The Simpson diversity then stabilized except for two T3 plots (plots 8 and 12) for which it kept increasing (Fig. \@ref(fig:DivTaxo)).
 
-```{r DivTaxo, out.width = '100%', echo=FALSE, fig.cap="Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic (ref:Bolda) richness, (ref:Boldb), taxonomic evenness, (ref:Boldc) functional richness, and (ref:Boldd) functional evenness.
+```{r DivTaxo, out.width = '100%', echo=FALSE, fig.cap="Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic richness (ref:Bolda), Simpson diversity (ref:Boldb), functional richness (ref:Boldc), and Rao diversity (ref:Boldd).
 Shaded areas are the credibility intervals ", fig.env="figure*"}
@@
 
-The plot 7 from treatment 1 displayed constantly outlying functional richness and evenness and was removed from the graphical representation for better readability.
-In undisturbed plots both functional richness and evenness remained stable along the 30 years. 
-In disturbed plots, functional richness and evenness trajectories depended on the disturbance intensity with their maximum positively correlated to %AGB loss $\rho_{spearman}^{Richness}=0.76$ and $\rho_{spearman}^{Evenness}=0.60$.
-Functional richness and evenness displayed for low disturbance intensity a low but long-lasting increase up to a maximum reached after 20-25 years, and for high intensity, a fast but short increase followed after 10 years by a slow decrease towards the initial values.
+The plot 7 from treatment 1 displayed constantly outlying functional richness and Rao diversity and was removed from the graphical representation for better readability.
+In undisturbed plots both functional richness and Rao diversity remained stable along the 30 years. 
+In disturbed plots, both trajectories depended on the disturbance intensity, with their maximum being positively correlated to %AGB loss $\rho_{Spearman}^{Richness}=0.76$ and $\rho_{Spearman}^{Rao}=0.60$.
+Functional richness and Rao diversity displayed for low disturbance intensity a low but long-lasting increase up to a maximum reached after 20-25 years, and for high intensity, a fast but short increase followed after 10 years by a slow decrease towards the initial values.
 
-The second-degree polynomial regressions between (i) the percentage AGB loss and (ii) taxonomic and functional richness and evenness after 10, 20 and 30 years best predicted the hump-shaped curve of the disturbance impact along the disturbance intensity gradient \@ref(fig:IDHplot).
-The relationship between the disturbance impact and its intensity was more markedly hump-shaped for the taxonomic richness than for the taxonomic evenness.
-For both functional richness and evenness the relationship was almost linear.
-The regression model better predicted the functional richness and evenness ($0.55<R^2_{Functional Richness}<0.72$, and $0.60<R^2_{Functional Evenness}<0.81$) than the taxonomic richness and evenness ($0.21<R^2_{Taxonomic Richness}<0.4$, and $-0.15<R^2_{Taxonomic Evenness}<0.43$ respectively)
+The second-degree polynomial regressions between (i) the percentage AGB loss and (ii) taxonomic and functional diversity after 10, 20 and 30 years best predicted the hump-shaped curve of the disturbance impact along the disturbance intensity gradient \@ref(fig:IDHplot).
+The relationship between the disturbance impact and its intensity was more markedly hump-shaped for the taxonomic richness than for the Simpson diversity.
+For both functional richness and Rao diversity the relationship was almost linear.
+The regression model better predicted the functional richness and Rao diversity ($0.55<R^2_{Functional Richness}<0.72$, and $0.60<R^2_{Functional Rao}<0.81$) than the taxonomic richness and evenness ($0.21<R^2_{Taxonomic Richness}<0.4$, and $-0.15<R^2_{Taxonomic Simpson}<0.43$ respectively)
 
@@
 All disturbed plots had lower functional redundancy than control plots and followed similar hump-shaped trajectories (\@ref(fig:RedFunRest)).
-The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{spearman}=0.47$) and the initial value had not recovered for any disturbed communities after 30 years.
+The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{Spearman}=0.47$) and the recovery had not attained initial values for any disturbed communities after 30 years.
 
@@
 RedundancyPlot(RedundancyTraj_restricted)
-mtext("Functional Redundancy",side=3,adj=0,line=1.5,cex=1.5)
-mtext("Restricted to initial functional space",side=3,adj=0,cex=1.2,line=0.5)
 mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=1.5)
@@
 
-The taxonomic differences among local communities, marked before disturbance by the distinct starting points on the NMDS axis 2, were maintained throughout recovery trajectories. 
-More than commonly thought, post-disturbance trajectories depended on community initial composition, that partly determined the pool of recruited species and constrained the trajectories towards the initial composition.
-The high resilience of community taxonomy, in the sense of the initial state recovery, revealed that species not belonging to the pre-disturbance community were hardly recruited because of the commonness of dispersal limitation among tropical tree species [@Svenning2005].
+The taxonomic differences among local communities were already apparent before disturbance, as revealed by the distinct starting points on the NMDS axis 2. 
+These differences were maintained throughout recovery trajectories.  
+More than commonly thought, post-disturbance trajectories depended on community initial composition, which partly determined the pool of recruited species and constrained the trajectories towards the initial composition.
+The high resilience of community taxonomy, in the sense of the recovery of initial state, revealed that species not belonging to the pre-disturbance community were rarely recruited because of the commonness of dispersal limitation among tropical tree species [@Svenning2005].
 
 Conversely, disturbed communities followed functional trajectories that are highly similar in terms of functional composition.
-As pre-disturbance surviving trees mirror the initial community [@Herault2018], changes in functional composition relied upon the recruitment of species or functional types that were infrequent or absent before disturbance.
+As the composition of pre-disturbance surviving trees is representative of the initial community [@Herault2018], changes in functional composition relied upon the recruitment of species or functional types that were infrequent or absent before disturbance.
 Competitive pioneers became dominant in filling the environmental niches of high availability of light, space and nutrients vacated by the disturbance.
@@
 
-## Another perspective on the intermediate disturbance hypothesis
+## A new perspective on the intermediate disturbance hypothesis
 
-The IDH well predicted well the disturbance impact on community taxonomic richness, enhanced until an intensity threshold (20-25% AGB loss), and to some extent on taxonomic evenness, somewhat decoupled from the disturbance intensity as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
-The disturbance intensity determined the balance in the community between pre-disturbance surviving trees and those recruited afterward.
-The pool of true pioneer species specifically recruited after disturbance is restricted in the Guiana Shield to a few common genera (e.g. Cecropia spp., Vismia spp.) [@Guitet2018].
+The IDH predicted well the enhancement of taxonomic richness after disturbance, until an intensity threshold (20-25% AGB loss).
+It also explained the impact on the Simpson diversity and hence on community taxonomic evenness.
+Taxonomic evenness was somewhat decoupled from the disturbance intensity as already observed in the Guiana Shield  [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+Disturbance intensity determined the balance in the community between pre-disturbance surviving trees and those recruited afterward.
+The pool of true pioneer species specifically recruited after disturbance is restricted in the Guiana Shield to a few common genera (e.g. *Cecropia* spp., *Vismia* spp.) [@Guitet2018].
 Below the intensity threshold the size of the surviving community maintained the pre-disturbance high taxonomic richness while the recruitment of pioneers, infrequent or absent before disturbance, increased both community taxonomic richness and evenness.
 Beyond the intensity threshold, the disturbance decreased the taxonomic richness of surviving trees which was not offset by the enrichment of pioneers, so that the overall community taxonomic richness decreased according to the disturbance intensity [@Molino2001].
-For community taxonomic evenness the disturbance impact was similar but slighter, as the evenness is less sensitive to the loss of rare species.
-Taxonomic evenness rather represented the increasing dominance of pioneers that balanced the usual hyper-dominance of a few species in tropical forests below the intensity threshold, thus increasing community overall evenness up to the intensity threshold beyond which pioneers became in turn highly dominant and decreased the overall evenness [@Baraloto2012a].
+An intermediate disturbance hence increased the availability of resources and created opportunities for pioneers, without implying too important loss of shade-tolerant species and without preventing their maintenance in the community [@Bongers2009].
+For community taxonomic evenness the disturbance impact was similar but milder, as the evenness is less sensitive to the loss of rare species.
+Taxonomic evenness trajectories, below the intensity threshold, rather represented the increasing dominance of pioneers that balanced the usual hyper-dominance of a few species in tropical forests.
+Beyond the intensity threshold, however, pioneers became in turn highly dominant and decreased the overall evenness [@Baraloto2012a].
 
 Conversely the IDH was disproved regarding the disturbance impact on community functional richness and evenness.
-Irrespective of the disturbance intensity the recruitment of pioneers, functionally highly different from the composition of pre-disturbance community, increased both community functional richness and evenness.
+Irrespective of the disturbance intensity, both community functional richness and evenness increased following the recruitment of pioneers, functionally highly different from the composition of pre-disturbance community.
 
-Along time, taxonomic richness trajectories of all disturbed communities first dropped similarly, following the species loss due to disturbance, and then displayed a species gain depending on the disturbance intensity.
+Along time, taxonomic richness trajectories of all disturbed communities similarly dropped at first place, following the species loss due to disturbance, and then displayed a species gain depending on the disturbance intensity.
 Up to an intensity threshold, the species gain was all the more significant that the disturbance intensity increased, with the establishment of long-lived pioneers enhancing community taxonomic richness and evenness in the long term.
@@
 
-For 15 years the species loss during disturbance, determined by the disturbance intensity, commensurately decreased the functional redundancy within the pre-disturbance functional space.
-The redundancy decrease was not compensated in the first place as the first recruited pioneers were functionally different from the pre-disturbance functional composition.
+For 15 years the species loss during disturbance, determined by the disturbance intensity, significantly decreased the functional redundancy within the pre-disturbance functional space.
+The redundancy decrease was not compensated for in the first place as the first recruited pioneers were functionally different from the pre-disturbance functional composition.
 Progressively though, first established species were replaced by more competitive long-lived pioneers or late-successional species resembling more the pre-disturbance functional composition and restoring the functional redundancy.
 This replacement was stochastic and followed the lottery recruitment rules, implying a recruitment eased for the first recruited species but then increasingly hampered by the emergence of interspecific competition [@Busing2002].
-Along time the recovery of infrequent species was increasingly slow, so that the time for the full recovery of the functional redundancy, in some communities just initiated after 30 years, was extremely difficult to estimate [@Elmqvist2003;@Diaz2005].
+Along time the recovery of infrequent species was increasingly slow, so that the time to full recovery of the functional redundancy, in some communities just initiated after 30 years, was extremely difficult to estimate [@Elmqvist2003;@Diaz2005].
 
-The long-term impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008].
-Besides, the long-term recovery of infrequent species increases the risks to loose cornerstone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
+The mid-term impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008].
+Besides, the mid-term recovery of infrequent species increases the risks to loose keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
 Apart from the functional characteristics considered here, infrequent species might indeed have unique functions in the ecosystem or be a key for some fauna [@Schleuning2016].
@@
 Our study revealed community recovery through the combination of deterministic processes driving their convergence in functional composition, and dispersal limitation maintaining their divergence in taxonomic composition.
-The IDH was validated for community taxonomic richness and, to some extent, taxonomic evenness but disproved regarding community functional richness and evenness that were enhanced for any disturbance intensity by the high functional differences of pioneers compared to late-successional functional composition.
-The IDH was translated in time by the recruitment, beyond an intensity threshold, of short-lived pioneers that prevented in the first times the establishment of more diverse long-lived pioneers, recruited otherwise below the intensity threshold.
-The resilience of tropical forests, in the sense of the pre-disturbance state recovery, proved tangible but requiring several decades. 
-Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].
+The IDH was validated for community taxonomic richness and, to some extent, taxonomic evenness but disproved regarding community functional richness and evenness which were enhanced for any disturbance intensity by the high functional differences of pioneers compared to late-successional functional composition.
+The IDH was translated in time by the recruitment, beyond an intensity threshold, of short-lived pioneers that prevented in the first years after disturbance the establishment of more diverse long-lived pioneers, recruited otherwise below the intensity threshold.
+The resilience of tropical forests, defined in terms of recovery to pre-disturbance state, proved tangible but requiring several decades.
+Still, the disturbance impact on community redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities  [@Herault2018].
 

### 2018-10-18 WholePlotTrajectories.tex
Jecol review

@@
 \Authors{
-Ariane MIRABEL\textsuperscript{1*}\\ Bruno Hérault\textsuperscript{2}\\ Eric Marcon\textsuperscript{1}
+Ariane MIRABEL\textsuperscript{1*}\\ Bruno Herault\textsuperscript{2}\\ Eric Marcon\textsuperscript{1}
 } % Authors
@@
 
-\Keywords{Community Ecology, Determinants of Plant Community Diversity and Structure, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Long-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity} % Keywords - if you don't want any simply remove all the text between the curly brackets
+\Keywords{Community Ecology, Community Diversity Determinants, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity} % Keywords - if you don't want any simply remove all the text between the curly brackets
 \newcommand{\keywordname}{Keywords} % Defines the keywords heading name
@@
 intensity. This intermediate disturbance hypothesis (IDH), though,
-remains debated and the controversy questions the extent of communities'
+remains debated and the controversy questions the extent of community
 resilience regarding their taxonomic and functional facets. To
@@
 functional composition trajectories were quite similar among
-communities. The IDH did predict communities taxonomic diversity
-response while functional diversity was enhanced whatever the
-disturbance intensity. Although consistent, the recovery of community
-composition, diversity and redundancy remained unachieved after 30
-years. This acknowledged the need of decades-long cycles with no
-disturbance to ensure a complete recovery, and questioned tropical
-forest community resilience after repeated disturbances.
+communities. The IDH did predict community taxonomic diversity response
+while functional diversity was enhanced whatever the disturbance
+intensity. Although consistent, the recovery of community composition,
+diversity and redundancy remained unachieved after 30 years. This
+acknowledged the need of decades-long cycles with no disturbance to
+ensure a complete recovery, and questioned tropical forest community
+resilience after repeated disturbances.
 }
@@
 pressure on remaining natural forests \citep{Morales-Hidalgo2015} and
-threatens the maintenance and dynamics in space and time of communities
+threatens the maintenance and dynamics in space and time of community
 structure, composition and functioning \citep{Anderson-Teixeira2013}.
@@
 \citep{Tilman2014, Liang2016}. In the short-term, moderate disturbance
-may lead to positive impacts on communities diversity, an idea
-formalized by the intermediate disturbance hypothesis (IDH) stating a
-maximized species diversity when disturbance intensity is not too high
+may lead to positive impacts on community diversity, an idea formalized
+by the intermediate disturbance hypothesis (IDH) which predicts a
+maximum species diversity when disturbance intensity is not too high
 \citep{Molino2001, Kariuki2006a}.
@@
 ecological-meaningful analysis would couple richness with (i) evenness,
-that would reveal the changes in the species abundance distribution and
-thus the underlying ecological processes, and (ii) composition that is
+which would reveal the changes in the species abundance distribution and
+thus the underlying ecological processes, and (ii) composition, which is
 crucial for conservation issues \citep{Lavorel2002, Bellwood2006}.
 Furthermore, a functional approach accounting for species biological
-attributes would directly link communities diversity, composition and
+attributes would directly link community diversity, composition and
 redundancy to ecosystem functioning and to its environmental constraints
@@
 (seed mass, maximum size)
-\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Eventually
+\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Eventually,
 a complete overview of community response to disturbance would encompass
@@
 validity of the IDH in the long term for tropical forest and its
-translation into different trajectories in time, and (iii) questioned
+translation into different trajectories in time, and (iii) examined
 community recovery time.
@@
 between 5 and 50 m and soils correspond to thin acrisols over a layer of
-transformed saprolite with low permeability generating lateral drainage
-during heavy rains.
+transformed saprolite with low permeability which produces lateral
+drainage during heavy rains.
 
-The experiment is a network of twelve 6.25ha plots that underwent a
+The experiment is a network of 12 6.25 ha plots that underwent a
 disturbance gradient of three logging, thinning and fuelwood cutting
@@
 with three replicate blocks of four plots. The disturbance corresponds
-to averages of 10 trees removed per hectare with a diameter at 1.3 m
-height (DBH) above 50 cm for treatment 1 (T1), 32 trees/ha above 40 cm
-DBH for treatment 2 (T2) and 40 trees/ha above 40 cm DBH for treatment 3
-(T3). Treatments T2 and T3 besides included the thinning of trees by
-poison girdling \citep{Schmitt1990}. The disturbance intensity was
-measured as the percentage of aboveground biomass (\%AGB) lost between
-the first inventory in 1984 and five years after disturbance
-\citep{Piponiot2016} estimated with the BIOMASS R package
-\citep{Biomass2018}.
+to an intensity gradient. For treatment 1 (T1) 10 trees of commercial
+species (of a diameter at 1.3 m height (DBH) equal or above 50 cm) were
+felled per hectare. For treatment 2 (T2) 10 trees/ha of commercial
+species (DBH \(\geq\) 50 cm) were felled and 30 trees/ha of non-valuable
+species (DBH \(\geq\) 40 cm) were removed by poison girdling. For
+treatment 3 (T3) 10 trees/ha of commercial species (DBH \(\geq\) 50 cm)
+were felled and 30 trees/ha of non-valuable species (15 with DBH
+\(\geq\) 50 cm and 15 with DBH \(\geq\) 40 cm) were removed by poison
+girdling \citep{Schmitt1990}. Disturbance intensity was measured as the
+percentage of aboveground biomass (\%AGB) lost between the first
+inventory in 1984 and five years after disturbance \citep{Piponiot2016}
+estimated with the BIOMASS R package \citep{Biomass2018}.
 
@@
 The study site corresponds to a tropical rainforest typical of the
-Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae,
-Lecythidaceae and Sapotaceae. In the twelve experimental plots, all
-trees above 10 cm DBH have been mapped and measured annually since 1984.
-Trees are first identified with a vernacular name assigned by the forest
-worker team, and afterward with a scientific name assigned by botanists
-during regular botanical campaigns. In 1984, specific vernacular names
-were given to 62 commercial or common species whereas more infrequent
-ones were identified under general identifiers only distinguishing trees
-and palm trees. From 2003, botanical campaigns have been conducted every
-5 to 6 years to identify all trees at the species level but
-identification levels still varied among plots and campaigns.
+Guiana Shield with a dominance of \emph{Fabaceae},
+\emph{Chrysobalanaceae}, \emph{Lecythidaceae} and \emph{Sapotaceae}. In
+the 12 experimental plots, all trees above 10 cm DBH have been mapped
+and measured annually since 1984. Trees are first identified with a
+vernacular name assigned by the forest worker team, and afterward with a
+scientific name assigned by botanists during regular botanical
+campaigns. In 1984, specific vernacular names were given to 62
+commercial or common species whereas more infrequent ones were
+identified under general identifiers only distinguishing trees and
+palms. From 2003, botanical campaigns have been conducted every 5 to 6
+years to identify all trees at the species level but identification
+levels still varied among plots and campaigns.
 
 This variability of protocols in time raised methodological issues as
-vernacular names usually correspond to different botanical species. It
-resulted in significant taxonomic uncertainties that had to be
-propagated to composition and diversity metrics. The uncertainty
-propagation was done through a Bayesian framework reconstituting
-complete inventories at genus level from real incomplete ones on the
-basis of vernacular/botanical names association. Vernacular names were
-replaced through multinomial trials based on the association probability
+vernacular names usually correspond to different botanical species. This
+resulted in significant taxonomic uncertainty that had to be propagated
+to composition and diversity metrics. The uncertainty propagation was
+done through a Bayesian framework reconstituting complete inventories at
+genus level from real incomplete ones on the basis of
+vernacular/botanical names association. Vernacular names were replaced
+through multinomial trials based on the association probability
 \(\big[\alpha_1, \alpha_2,..., \alpha_V\big]\) observed across all
@@
 
-Six functional traits representing leaf economics (leaves thickness,
+Six functional traits representing leaf economics (leaf thickness,
 toughness, total chlorophyll content and specific leaf area) and stem
@@
 
-To counter taxonomic uncertainties due to the variability of botanical
-identification levels (in space) and protocols (in time), the taxonomic
-composition and diversity analysis were conducted at the genus level.
-Taxonomic and functional trajectories of community composition were
-followed in a two-dimensional NMDS ordination space. Two NMDS using
-abundance-based (Bray-Curtis) dissimilarity measures were conducted to
-map either taxonomic or functional composition, the later based on the 7
-leaf, stem and life history traits (without seed mass classes).
-Trajectories along time were reported through the Euclidean distance
-between the target inventories and the reference inventories in 1989,
-\emph{i.e.} 2 years after disturbance. Univariate trajectories of the
-leaf, stem and life-history traits were also visualized with the
-community weighted means (CWM) \citep{Diaz2007}. Species seed mass
-corresponded to 5 classes of increasing mass, seed mass trajectories
-were therefore reported as the proportion of each class in the
-inventories (Supplementary materials).
+Because of the variability in the precision of botanical identification
+efforts, we were constraint to conduct the taxonomic composition and
+diversity analysis at the genus level. Taxonomic and functional
+trajectories of community composition were followed in a two-dimensional
+NMDS ordination space. Two NMDS using abundance-based (Bray-Curtis)
+dissimilarity measures were conducted to map either taxonomic or
+functional composition, the latter based on the seven leaf, stem and
+life history traits (without seed mass classes). Trajectories along time
+were reported through the Euclidean distance between the target
+inventories and the reference inventories in 1989, \emph{i.e.} 2 years
+after disturbance. Univariate trajectories of the leaf, stem and
+life-history traits were also visualized with the community weighted
+means (CWM) \citep{Diaz2007}. Species seed mass were given in 5 mass
+classes. Seed mass trajectories were therefore reported as the
+proportion of each class in the inventories (Supplementary materials).
 
-The taxonomic diversity was reported through species richness and
-evenness, \emph{i.e} the Hill number translation of the Simpson index
-\citep{Hill1973}. These indices belong to the set of HCDT or generalized
-entropy, respectively corresponding to the 0 and 2 order of diversity
-(q), recommended for diversity studies \citep{Marcon2015b}. The
-functional diversity was reported using the functional richness and
-functional evenness, \emph{i.e} Rao index of quadratic entropy which
+The taxonomic diversity was reported through species richness and the
+Hill number translation of the Simpson index \citep{Hill1973}. The
+comparison between richness and Simpson diversity assessed the
+variations in taxonomic evenness. Both indices belong to the set of HCDT
+or generalized entropy, respectively corresponding to the 0 and 2 order
+of diversity (q), recommended for diversity studies \citep{Marcon2015b}.
+The functional diversity was reported using the functional richness and
+functional evenness, \emph{i.e} the Rao index of quadratic entropy which
 combines species abundance distribution and average pairwise
@@
 
-The impacts of initial disturbance were tested with the spearman rank
-correlation between the extrema of taxonomic and functional metrics
+The impacts of initial disturbance were tested with the Spearman rank
+correlation between the extremes of taxonomic and functional metrics
 reached over the 30 years and the initial \%AGB loss. They were besides
@@
 
-The functional redundancy was measured as the overlap among species in
-community functional space \citep{Carmona2016}. The samples of the trait
-database were first mapped in a 2-dimensional plan with a PCA analysis.
-Then, multivariate kernel density estimator associated with individual
-trees returned species traits probability distribution (TDP). Species
-TDP weighted by species abundance were eventually summed for each
-community. Community functional redundancy was the sum of TDPs overlap,
-expressed as the average number of species that could be removed from
-without reducing the functional space (Supp. Mat. - Fig. S1 for a more
-comprehensive scheme).
+Functional redundancy was measured as the overlap among species in
+community functional space \citep{Carmona2016}. First, the individuals
+of the trait database were mapped in the plane of the two first axes
+from a PCA analysis. Then, for each species, the traits probability
+density (TPD) were computed through multivariate kernel density
+estimators. Second, for each community, the functional traits
+probability density were computed as the sum of the TDP of the
+inventoried species weighted by species abundance. Third, the whole
+functional space was divided into a 100 x 100 grid, before counting for
+each cell the number of species which TPD is positive. The average count
+across cells, minus 1, gives the average number of species in the
+community that share the same traits, which is the Community Functional
+Redundancy.
 
@@
 
-\subsection{Communities Composition}\label{communities-composition}
+\subsection{Community Composition}\label{community-composition}
 
@@
 disturbance), 828 388 individual trees and 591 botanical species
-pertaining to 223 genus and 64 families were recorded.
+pertaining to 223 genera and 64 families were recorded.
 
@@
 shifts towards species with more acquisitive functional strategies, from
-communities with high average WSG to high average SLA and chlorophyll
-content (see appendix I). For functional composition, this translated
-into cyclic compositional changes with an unachieved recovery of the
-initial composition (Fig. \ref{fig:NMDSplans}). The maximum
-dissimilarity with the initial state was positively correlated to the
-disturbance intensity for both taxonomic and functional composition
-(\(\rho_{spearman}^{Taxonomic}=0.87\) and
-\(\rho_{spearman}^{Functional}=0.90\) respectively). The maximum value
-was reached around 26 years after disturbance for taxonomic composition
-and 22 years for functional composition.
+communities with high mean WSG to high mean SLA and chlorophyll content
+(see appendix I). For functional composition, this translated into
+cyclic compositional changes with an incomplete recovery of the initial
+composition (Fig. \ref{fig:NMDSplans}). The maximum dissimilarity with
+the initial state was positively correlated with the disturbance
+intensity for both taxonomic and functional composition
+(\(\rho_{Spearman}^{Taxonomic}=0.87\) and
+\(\rho_{Spearman}^{Functional}=0.90\) respectively). The maximum
+dissimilarity with the initial was reached for taxonomic composition
+between 15 to 25 years, for most of the plots, and around 22 years for
+functional composition.
 
@@
 
-\caption{Plot trajectories in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS space.
+\caption{Plot trajectories in terms of taxonomic composition (\textbf{(a)} and \textbf{(c)}) and functional composition (\textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS space.
 Lower panels (\textbf{(c)} and \textbf{(d)}) represent the Euclidean distance to initial condition along the 30 sampled years.
@@
 
-Except for leaf chlorophyll content, which continued to increase for
-some T3 and T2 plots 30 years after disturbance, all traits and seed
-mass proportions followed unimodal trajectories either stabilizing or
-returning towards their initial values.
+Community CWM average value of all traits and seed mass proportions
+followed unimodal trajectories, either stabilizing or returning towards
+their initial values, to the exception of leaf chlorophyll content,
+which continued to increase for some T3 and T2 plots 30 years after
+disturbance.
 
-Maximum height at adult stage (\emph{Hmax}), leaf toughness and wood
-specific gravity (\emph{WSG}) first decreased and then slightly
-increased but remained significantly lower than their initial value
-(Fig. \ref{fig:CWM}). On the other side, bark thickness and specific
-leaf area (\emph{SLA}) increased and while bark thickness remained
-substantially high after 30 years, \emph{SLA} had almost recovered its
-initial value. For all traits, the maximum difference to initial value
-was correlated to the disturbance intensity
-(\(\rho_{spearman}^{Leaf thickness}=0.76\),
-\(\rho_{spearman}^{Chlorophyll content}=0.60\),
-\(\rho_{spearman}^{Leaf toughness}=-0.53\),
-\(\rho_{spearman}^{SLA}=0.93\), \(\rho_{spearman}^{WSG}=-0.75\),
-\(\rho_{spearman}^{Bark thickness}=0.71\),
-\(\rho_{spearman}^{Hmax}=-0.40\)). The proportions of the three lightest
-seed mass classes increased in all disturbed plots, and decreased after
-30 years for the lightest class while it stabilized for the two other
-(Supp. Mat. - Fig. S2).
+Community CWM average value of Maximum height at adult stage
+(\emph{Hmax}), leaf toughness and wood specific gravity (\emph{WSG})
+first decreased and then slightly increased but remained significantly
+lower than their initial value (Fig. \ref{fig:CWM}). On the other side,
+bark thickness and specific leaf area (\emph{SLA}) increased and while
+bark thickness remained substantially high after 30 years, \emph{SLA}
+had almost recovered to its initial value. For all traits, the maximum
+difference to initial value was correlated to the disturbance intensity.
+Positive correlations were observed for Leaf thickness, chlorophyll
+content, SLA and bark thickness
+(\(\rho_{Spearman}^{Leaf thickness}=0.76\),
+\(\rho_{Spearman}^{Chlorophyll content}=0.60\),
+\(\rho_{Spearman}^{SLA}=0.93\),
+\(\rho_{Spearman}^{Bark thickness}=0.71\)). Negative correlation was
+observed for Leaf toughness, WSG and Hmax
+(\(\rho_{Spearman}^{Leaf toughness}=-0.53\),
+\(\rho_{Spearman}^{WSG}=-0.75\), \(\rho_{Spearman}^{Hmax}=-0.40\)) The
+proportions of the three lightest seed mass classes increased in all
+disturbed plots, and decreased after 30 years for the lightest class
+while it stabilized for the two other (Supp. Mat. - Fig. S2).
 
@@
 
-\caption{Trajectories of the communities weighted means over 30 years after disturbance of four leaf traits (Leaf thickness, chlorophyll content, toughness, and specific area), two stem traits (wood specific gravity, and bark thickness) and one life history trait (Specific maximum height at adult stage).
+\caption{Trajectories of community weighted means over 30 years after disturbance of four leaf traits (Leaf thickness, chlorophyll content, toughness, and specific area), two stem traits (wood specific gravity, and bark thickness) and one life history trait (Specific maximum height at adult stage).
 }\label{fig:CWM}
@@
 
-\subsection{Communities richness and
-evenness}\label{communities-richness-and-evenness}
+\subsection{Community taxonomic and functional
+diversity}\label{community-taxonomic-and-functional-diversity}
 
-For undisturbed plots, taxonomic Richness and Evenness remained stable
-over the 30 years monitored. In disturbed communities, after low
-disturbance intensity the taxonomic richness increased, reaching a
+For undisturbed plots, taxonomic richness and Simpson diversity remained
+stable over the 30 years of monitoring. In disturbed communities, after
+low disturbance intensity the taxonomic richness increased, reaching a
 maximum gain of 14 botanical genera (plot 3 from treatment 2). After
@@
 to pre-disturbance values. The maximum richness loss or gain after
-disturbance was positively correlated to the disturbance intensity
-(\(\rho_{spearman}^{Richness}=0.50\)). In all disturbed plots the
-taxonomic evenness first increased until a maximum reached after around
-20 years. This maximum was positively correlated to the disturbance
-intensity (\(\rho_{spearman}^{Evenness}=0.77\)). The evenness then
-stabilized except for two T3 plots (plots 8 and 12) for which evenness
-kept increasing.
+disturbance was positively correlated with the disturbance intensity
+(\(\rho_{Spearman}^{Richness}=0.50\)). In all disturbed plots the
+Simpson diversity first increased until a maximum reached after around
+20 years. This maximum was positively correlated with the disturbance
+intensity (\(\rho_{Spearman}^{Simpson}=0.77\)). The Simpson diversity
+then stabilized except for two T3 plots (plots 8 and 12) for which it
+kept increasing (Fig. \ref{fig:DivTaxo}).
 
@@
 
-\caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic \textbf{(a)} richness, \textbf{(b)}, taxonomic evenness, \textbf{(c)} functional richness, and \textbf{(d)} functional evenness.
+\caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic richness \textbf{(a)}, Simpson diversity \textbf{(b)}, functional richness \textbf{(c)}, and Rao diversity \textbf{(d)}.
 Shaded areas are the credibility intervals }\label{fig:DivTaxo}
@@
 The plot 7 from treatment 1 displayed constantly outlying functional
-richness and evenness and was removed from the graphical representation
-for better readability. In undisturbed plots both functional richness
-and evenness remained stable along the 30 years. In disturbed plots,
-functional richness and evenness trajectories depended on the
-disturbance intensity with their maximum positively correlated to \%AGB
-loss \(\rho_{spearman}^{Richness}=0.76\) and
-\(\rho_{spearman}^{Evenness}=0.60\). Functional richness and evenness
-displayed for low disturbance intensity a low but long-lasting increase
-up to a maximum reached after 20-25 years, and for high intensity, a
-fast but short increase followed after 10 years by a slow decrease
-towards the initial values.
+richness and Rao diversity and was removed from the graphical
+representation for better readability. In undisturbed plots both
+functional richness and Rao diversity remained stable along the 30
+years. In disturbed plots, both trajectories depended on the disturbance
+intensity, with their maximum being positively correlated to \%AGB loss
+\(\rho_{Spearman}^{Richness}=0.76\) and \(\rho_{Spearman}^{Rao}=0.60\).
+Functional richness and Rao diversity displayed for low disturbance
+intensity a low but long-lasting increase up to a maximum reached after
+20-25 years, and for high intensity, a fast but short increase followed
+after 10 years by a slow decrease towards the initial values.
 
 The second-degree polynomial regressions between (i) the percentage AGB
-loss and (ii) taxonomic and functional richness and evenness after 10,
-20 and 30 years best predicted the hump-shaped curve of the disturbance
-impact along the disturbance intensity gradient \ref{fig:IDHplot}. The
+loss and (ii) taxonomic and functional diversity after 10, 20 and 30
+years best predicted the hump-shaped curve of the disturbance impact
+along the disturbance intensity gradient \ref{fig:IDHplot}. The
 relationship between the disturbance impact and its intensity was more
-markedly hump-shaped for the taxonomic richness than for the taxonomic
-evenness. For both functional richness and evenness the relationship was
-almost linear. The regression model better predicted the functional
-richness and evenness (\(0.55<R^2_{Functional Richness}<0.72\), and
-\(0.60<R^2_{Functional Evenness}<0.81\)) than the taxonomic richness and
+markedly hump-shaped for the taxonomic richness than for the Simpson
+diversity. For both functional richness and Rao diversity the
+relationship was almost linear. The regression model better predicted
+the functional richness and Rao diversity
+(\(0.55<R^2_{Functional Richness}<0.72\), and
+\(0.60<R^2_{Functional Rao}<0.81\)) than the taxonomic richness and
 evenness (\(0.21<R^2_{Taxonomic Richness}<0.4\), and
-\(-0.15<R^2_{Taxonomic Evenness}<0.43\) respectively)
+\(-0.15<R^2_{Taxonomic Simpson}<0.43\) respectively)
 
@@
 The maximum redundancy loss was positively correlated with the
-disturbance intensity (\(\rho_{spearman}=0.47\)) and the initial value
-had not recovered for any disturbed communities after 30 years.
+disturbance intensity (\(\rho_{Spearman}=0.47\)) and the recovery had
+not attained initial values for any disturbed communities after 30
+years.
 
@@
 
-The taxonomic differences among local communities, marked before
-disturbance by the distinct starting points on the NMDS axis 2, were
-maintained throughout recovery trajectories. More than commonly thought,
-post-disturbance trajectories depended on community initial composition,
-that partly determined the pool of recruited species and constrained the
-trajectories towards the initial composition. The high resilience of
-community taxonomy, in the sense of the initial state recovery, revealed
-that species not belonging to the pre-disturbance community were hardly
-recruited because of the commonness of dispersal limitation among
-tropical tree species \citep{Svenning2005}.
+The taxonomic differences among local communities were already apparent
+before disturbance, as revealed by the distinct starting points on the
+NMDS axis 2. These differences were maintained throughout recovery
+trajectories.\\
+More than commonly thought, post-disturbance trajectories depended on
+community initial composition, which partly determined the pool of
+recruited species and constrained the trajectories towards the initial
+composition. The high resilience of community taxonomy, in the sense of
+the recovery of initial state, revealed that species not belonging to
+the pre-disturbance community were rarely recruited because of the
+commonness of dispersal limitation among tropical tree species
+\citep{Svenning2005}.
 
 Conversely, disturbed communities followed functional trajectories that
-are highly similar in terms of functional composition. As
-pre-disturbance surviving trees mirror the initial community
-\citep{Herault2018}, changes in functional composition relied upon the
-recruitment of species or functional types that were infrequent or
-absent before disturbance. Competitive pioneers became dominant in
-filling the environmental niches of high availability of light, space
-and nutrients vacated by the disturbance. The recruitment of pioneers
-changed community functional composition in the same way for all
-disturbance intensity towards more resource-acquisitive strategies,
+are highly similar in terms of functional composition. As the
+composition of pre-disturbance surviving trees is representative of the
+initial community \citep{Herault2018}, changes in functional composition
+relied upon the recruitment of species or functional types that were
+infrequent or absent before disturbance. Competitive pioneers became
+dominant in filling the environmental niches of high availability of
+light, space and nutrients vacated by the disturbance. The recruitment
+of pioneers changed community functional composition in the same way for
+all disturbance intensity towards more resource-acquisitive strategies,
 moving community functional composition right along the first axis in
@@
 
-\subsection{Another perspective on the intermediate disturbance
-hypothesis}\label{another-perspective-on-the-intermediate-disturbance-hypothesis}
+\subsection{A new perspective on the intermediate disturbance
+hypothesis}\label{a-new-perspective-on-the-intermediate-disturbance-hypothesis}
 
-The IDH well predicted well the disturbance impact on community
-taxonomic richness, enhanced until an intensity threshold (20-25\% AGB
-loss), and to some extent on taxonomic evenness, somewhat decoupled from
-the disturbance intensity as already observed in the Guiana Shield
+The IDH predicted well the enhancement of taxonomic richness after
+disturbance, until an intensity threshold (20-25\% AGB loss). It also
+explained the impact on the Simpson diversity and hence on community
+taxonomic evenness. Taxonomic evenness was somewhat decoupled from the
+disturbance intensity as already observed in the Guiana Shield
 \citep{Baraloto2012a} and in Bornean tropical forests
-\citep{Cannon1998}. The disturbance intensity determined the balance in
-the community between pre-disturbance surviving trees and those
-recruited afterward. The pool of true pioneer species specifically
-recruited after disturbance is restricted in the Guiana Shield to a few
-common genera (e.g.~Cecropia spp., Vismia spp.) \citep{Guitet2018}.
+\citep{Cannon1998}. Disturbance intensity determined the balance in the
+community between pre-disturbance surviving trees and those recruited
+afterward. The pool of true pioneer species specifically recruited after
+disturbance is restricted in the Guiana Shield to a few common genera
+(e.g. \emph{Cecropia} spp., \emph{Vismia} spp.) \citep{Guitet2018}.
 Below the intensity threshold the size of the surviving community
@@
 that the overall community taxonomic richness decreased according to the
-disturbance intensity \citep{Molino2001}. For community taxonomic
-evenness the disturbance impact was similar but slighter, as the
-evenness is less sensitive to the loss of rare species. Taxonomic
-evenness rather represented the increasing dominance of pioneers that
-balanced the usual hyper-dominance of a few species in tropical forests
-below the intensity threshold, thus increasing community overall
-evenness up to the intensity threshold beyond which pioneers became in
-turn highly dominant and decreased the overall evenness
+disturbance intensity \citep{Molino2001}. An intermediate disturbance
+hence increased the availability of resources and created opportunities
+for pioneers, without implying too important loss of shade-tolerant
+species and without preventing their maintenance in the community
+\citep{Bongers2009}. For community taxonomic evenness the disturbance
+impact was similar but milder, as the evenness is less sensitive to the
+loss of rare species. Taxonomic evenness trajectories, below the
+intensity threshold, rather represented the increasing dominance of
+pioneers that balanced the usual hyper-dominance of a few species in
+tropical forests. Beyond the intensity threshold, however, pioneers
+became in turn highly dominant and decreased the overall evenness
 \citep{Baraloto2012a}.
@@
 community functional richness and evenness. Irrespective of the
-disturbance intensity the recruitment of pioneers, functionally highly
-different from the composition of pre-disturbance community, increased
-both community functional richness and evenness.
+disturbance intensity, both community functional richness and evenness
+increased following the recruitment of pioneers, functionally highly
+different from the composition of pre-disturbance community.
 
 Along time, taxonomic richness trajectories of all disturbed communities
-first dropped similarly, following the species loss due to disturbance,
-and then displayed a species gain depending on the disturbance
-intensity. Up to an intensity threshold, the species gain was all the
-more significant that the disturbance intensity increased, with the
-establishment of long-lived pioneers enhancing community taxonomic
-richness and evenness in the long term. These long-lived pioneers,
-functionally quite different from the functional composition, entailed
-as well a progressive and long-lasting increase of the functional
-richness and evenness \citep{Denslow1980, Molino2001}. Beyond an
-intensity threshold, though, a few short-lived pioneers occupied the
+similarly dropped at first place, following the species loss due to
+disturbance, and then displayed a species gain depending on the
+disturbance intensity. Up to an intensity threshold, the species gain
+was all the more significant that the disturbance intensity increased,
+with the establishment of long-lived pioneers enhancing community
+taxonomic richness and evenness in the long term. These long-lived
+pioneers, functionally quite different from the functional composition,
+entailed as well a progressive and long-lasting increase of the
+functional richness and evenness \citep{Denslow1980, Molino2001}. Beyond
+an intensity threshold, though, a few short-lived pioneers occupied the
 vacated environmental space and prevented the establishment of other
@@
 For 15 years the species loss during disturbance, determined by the
-disturbance intensity, commensurately decreased the functional
-redundancy within the pre-disturbance functional space. The redundancy
-decrease was not compensated in the first place as the first recruited
-pioneers were functionally different from the pre-disturbance functional
+disturbance intensity, significantly decreased the functional redundancy
+within the pre-disturbance functional space. The redundancy decrease was
+not compensated for in the first place as the first recruited pioneers
+were functionally different from the pre-disturbance functional
 composition. Progressively though, first established species were
@@
 the recovery of infrequent species was increasingly slow, so that the
-time for the full recovery of the functional redundancy, in some
-communities just initiated after 30 years, was extremely difficult to
-estimate \citep{Elmqvist2003, Diaz2005}.
+time to full recovery of the functional redundancy, in some communities
+just initiated after 30 years, was extremely difficult to estimate
+\citep{Elmqvist2003, Diaz2005}.
 
-The long-term impact of disturbance on community functional redundancy
+The mid-term impact of disturbance on community functional redundancy
 meant a lower resilience of the pre-disturbance communities, with higher
@@
 expense of late-successional ones \citep{Haddad2008}. Besides, the
-long-term recovery of infrequent species increases the risks to loose
-cornerstone species, with unexpected ecological consequences
+mid-term recovery of infrequent species increases the risks to loose
+keystone species, with unexpected ecological consequences
 \citep{Jones1994, Chazdon2003a, Diaz2005}. Apart from the functional
@@
 richness and, to some extent, taxonomic evenness but disproved regarding
-community functional richness and evenness that were enhanced for any
+community functional richness and evenness which were enhanced for any
 disturbance intensity by the high functional differences of pioneers
@@
 translated in time by the recruitment, beyond an intensity threshold, of
-short-lived pioneers that prevented in the first times the establishment
-of more diverse long-lived pioneers, recruited otherwise below the
-intensity threshold. The resilience of tropical forests, in the sense of
-the pre-disturbance state recovery, proved tangible but requiring
-several decades. Still, the disturbance impact on communities redundancy
-cautioned against the risks of infrequent species loss and the
-persistence of disturbance-specific communities \citep{Herault2018}.
+short-lived pioneers that prevented in the first years after disturbance
+the establishment of more diverse long-lived pioneers, recruited
+otherwise below the intensity threshold. The resilience of tropical
+forests, defined in terms of recovery to pre-disturbance state, proved
+tangible but requiring several decades. Still, the disturbance impact on
+community redundancy cautioned against the risks of infrequent species
+loss and the persistence of disturbance-specific communities
+\citep{Herault2018}.
 

### 2018-10-18 Scripts/Stats.R

@@ -305,4 +305,19 @@ colnames(CompFun)<-c("Max","Plot","treat")
 cor(as.numeric(CompFun[,"Max"]),as.numeric(CompFun[,"treat"]),method="spearman")
 
 
+### traits dataset coverage:
 
+splist<-lapply(LivingStand_all,function(yr){
+  return(unlist(lapply(yr,function(plo){
+    freq<-as.character(plo[which(plo[,"name"]%in%traits[,"name"]),"name"])
+    freq<-sum(tapply(freq,freq,length))
+    return(freq/sum(tapply(as.character(plo[,"name"]),as.character(plo[,"name"]),length)))
+  })))
+})
+splist<-lapply(splist,function(yr){return(as.numeric(yr[which(names(yr)%in%1:12)]))})
+splist<-splist[which(unlist(lapply(splist,function(yr){length(yr)!=0})))]
+
+Pby<-c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014","2016","2017")
+splist<-unlist(splist[which(!names(splist)%in%Pby)])
+min(splist)
+mean(splist)

### 2018-10-18 WholePlotTrajectories.Rmd
graph

@@
 output:
-  bookdown::gitbook: default
-  bookdown::word_document2: default
   bookdown::pdf_book:
@@
     keep_tex: TRUE
+  bookdown::word_document2: default
+  bookdown::gitbook: default
 ---
@@
 RedundancyPlot(RedundancyTraj_restricted)
+mtext("Number of Species",side=3,line=1,adj=0,cex=1.5)
 mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=1.5)

### 2018-12-03 WholePlotTrajectories.Rmd
JEcol response

@@
 The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
-They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, the carbon, water and nutrient cycles, and ensure cultural and human well-being.
+They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, determine the cycles of carbon, water and nutrient, and ensure cultural and human well-being.
 The growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Morales-Hidalgo2015] and threatens the maintenance and dynamics in space and time of community structure, composition and functioning [@Anderson-Teixeira2013].
@@
 In tropical forests, ecological communities are regularly re-shaped by natural disturbance events changing both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions such as competition among species [@Chesson2000].
-One of the cornerstone of tropical forest ecology is to understand the processes and drivers of ecosystems response to disturbance [@Chazdon2003a]. 
-For now, this has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016] that are rapid and convenient to measure.
-These structural parameters have been successfully modeled, giving important insights into the recovery of ecosystem processes and services [@Herault2018]. 
-However the response of forests diversity and composition remains unclear, albeit it determines the productivity, stability and functioning of ecosystems [@Tilman2014; @Liang2016].
-In the short-term, moderate disturbance may lead to positive impacts on community diversity, an idea formalized by the intermediate disturbance hypothesis (IDH) which predicts a maximum species diversity when disturbance intensity is not too high [@Molino2001;@Kariuki2006a].
+One of the cornerstone of tropical forest ecology is to understand the processes and drivers of ecosystems response to such disturbance [@Chazdon2003a].
+This response has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
+Thereafter, models based on the observed trajectories after disturbance gave important insights into the recovery of ecosystem processes and services [@Herault2018]. 
+Regarding community diversity and composition, however, post-disturbance trajectories are not as thoroughly understood.
+Several studies and meta-analyses highlighted determined successions after disturbance, mainly driven by the recruitment of pioneer and heliophilous species [@Carreno-Rocabado2012;@Poorter2016].
+Community trajectories thus proved to rely on deterministic processes and post-disturbance responses would be predictable according to the disturbance intensity and the comunity diversity and composition.
+Still, observed trajectories in diversity and composition often deviate from predictions and the underlying ecological processes are debated [@Norden2015].
+Deeper understanding of tropical forest response to disturbance could be provided by long term monitoring, to complete the conclusions often based on single decades monitoring [@Carreno-Rocabado2012], dynamic trajectories to complete chronosequence studies [@Chazdon2007], and through a complete set of taxonomic and functional composition and diversity metrics to complete studies based only on taxonomic richness or functional composition [@Carreno-Rocabado2012,@Mouillot2013].
 
+Deterministic processes are thus recognized as important drivers of community trajectories.
+This is formalized by the intermediate disturbance hypothesis (IDH), predicting a miximum species diversity after moderate disturbance intensity [@Molino2001;@Kariuki2006a].
 Validations of the IDH though remain scarce in the long-term and mainly rely on the analysis of taxonomic richness [@Molino2001].
@@
 Because of the variability in the precision of botanical identification efforts, we were constraint to conduct the taxonomic composition and diversity analysis at the genus level. 
-Taxonomic and functional trajectories of community composition were followed in a two-dimensional NMDS ordination space.
+Taxonomic and functional trajectories of community composition were followed in a two-dimensional NMDS ordination plan.
 Two NMDS using abundance-based (Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the latter based on the seven leaf, stem and life history traits (without seed mass classes).
@@
 The taxonomic diversity was reported through species richness and the Hill number translation of the Simpson index [@Hill1973].
-The comparison between richness and Simpson diversity assessed the variations in taxonomic evenness. 
+These metrics allowed assessing the taxonomic richness as well as evenness, through the comparison between richness and Simpson diversity.
+R.esults will thus be discussed directly in terms of taxonomic richness and evenness.
 Both indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recommended for diversity studies [@Marcon2015b]. 
-The functional diversity was reported using the functional richness and functional evenness, *i.e* the Rao index of quadratic entropy which combines species abundance distribution and average pairwise dissimilarity based on species functional traits.
+The functional diversity was reported using the functional richness and functional evenness, *i.e* the Rao index of quadratic entropy.
+The Rao index combines species abundance distribution, and the average pairwise functional dissimilarity between species computed by the gower distance.
 
@@
 First, the individuals of the trait database were mapped in the plane of the two first axes from a PCA analysis.
-Then, for each species, the traits probability density (TPD) were computed through multivariate kernel density estimators.
-Second, for each community, the functional traits probability density were computed as the sum of the TDP of the inventoried species weighted by species abundance.
-Third, the whole functional space was divided into a 100 x 100 grid, before counting for each cell the number of species which TPD is positive.
-The average count across cells, minus 1, gives the average number of species in the community that share the same traits, which is the Community Functional Redundancy.
+Then, for each species, the traits probability density (TPD) were computed through two-dimension kernel density estimators.
+Second, for each community, the TDB weigthed by species abundance were summed accross the functional space.
+Third, the functional space was divided into a 100 x 100 grid and the number of species with a positive TDP was counted in each cell.
+The average count across cells minus 1 returnedthe Community Functional Redundancy, which was the average number of species in the community that share the same trait values.
 
@@
 
-```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of taxonomic composition ((ref:Bolda) and (ref:Boldc)) and functional composition ((ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS space.
+```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of taxonomic composition ((ref:Bolda) and (ref:Boldc)) and functional composition ((ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS plan.
 Lower panels ((ref:Boldc) and (ref:Boldd)) represent the Euclidean distance to initial condition along the 30 sampled years.
@@
 TaxoTraj(CompleteTaxo)
-plotDiv(CompleteRichFun,remove=TRUE)
+plotDiv(CompleteRichFun)
 mtext("(c) Functional Richness",side=3,adj=0,line=1,cex=0.9)
-plotDiv(CompleteFun,remove=TRUE)
+plotDiv(CompleteFun)
 mtext("(d) Functional Evenness",side=3,adj=0,line=1,cex=0.9)
@@
 For both functional richness and Rao diversity the relationship was almost linear.
-The regression model better predicted the functional richness and Rao diversity ($0.55<R^2_{Functional Richness}<0.72$, and $0.60<R^2_{Functional Rao}<0.81$) than the taxonomic richness and evenness ($0.21<R^2_{Taxonomic Richness}<0.4$, and $-0.15<R^2_{Taxonomic Simpson}<0.43$ respectively)
+The regression model better predicted the functional richness and Rao diversity ($0.55<R^2_{Functional Richness}<0.72$, and $0.60<R^2_{Functional Rao}<0.81$) than the taxonomic richness and evenness ($0.21<R^2_{Taxonomic Richness}<0.4$, and $-0.15<R^2_{Taxonomic Simpson}<0.43$ respectively).
 
@@
 More than commonly thought, post-disturbance trajectories depended on community initial composition, which partly determined the pool of recruited species and constrained the trajectories towards the initial composition.
-The high resilience of community taxonomy, in the sense of the recovery of initial state, revealed that species not belonging to the pre-disturbance community were rarely recruited because of the commonness of dispersal limitation among tropical tree species [@Svenning2005].
+The community taxonomy proved highly resilient, as it absorbed the disturbance and maintained commnity initial composition characteristics [@Folke2006].
+This high resilience suggested that species not belonging to the pre-disturbance community were rarely recruited because of the commonness of dispersal limitation among tropical tree species [@Svenning2005].
 
@@
 
-The IDH predicted well the enhancement of taxonomic richness after disturbance, until an intensity threshold (20-25% AGB loss).
-It also explained the impact on the Simpson diversity and hence on community taxonomic evenness.
-Taxonomic evenness was somewhat decoupled from the disturbance intensity as already observed in the Guiana Shield  [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
-Disturbance intensity determined the balance in the community between pre-disturbance surviving trees and those recruited afterward.
-The pool of true pioneer species specifically recruited after disturbance is restricted in the Guiana Shield to a few common genera (e.g. *Cecropia* spp., *Vismia* spp.) [@Guitet2018].
-Below the intensity threshold the size of the surviving community maintained the pre-disturbance high taxonomic richness while the recruitment of pioneers, infrequent or absent before disturbance, increased both community taxonomic richness and evenness.
-Beyond the intensity threshold, the disturbance decreased the taxonomic richness of surviving trees which was not offset by the enrichment of pioneers, so that the overall community taxonomic richness decreased according to the disturbance intensity [@Molino2001].
-An intermediate disturbance hence increased the availability of resources and created opportunities for pioneers, without implying too important loss of shade-tolerant species and without preventing their maintenance in the community [@Bongers2009].
-For community taxonomic evenness the disturbance impact was similar but milder, as the evenness is less sensitive to the loss of rare species.
-Taxonomic evenness trajectories, below the intensity threshold, rather represented the increasing dominance of pioneers that balanced the usual hyper-dominance of a few species in tropical forests.
+Community taxonomic richness increased after disturbance until an intensity threshold (20-25% AGB loss) above which the taxonomic richness decreased.
+The increase of taxonomic evenness after disturbance also depended on the disturbance intensity, although the correlation was weaker than for the taxonomic richness as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+In accordance with the IDH, then, community taxonomic trajectories depended on the disturbance intensity.
+They markedly changed above an intensity threshold, with a decrease of the richness and a persistent increase of the evenness.
+Disturbance intensity determined the balance in the community between surviving trees from pre-disturbance communities and trees recruited afterward.
+The pool of true pioneer species specifically recruited after disturbance is, in the Guiana Shield, restricted to a few common genera (e.g.
+*Cecropia* spp., *Vismia* spp.)[@Guitet2018].
+Below the intensity threshold the surviving community trees remained  numerous enough to maintain the pre-disturbance high taxonomic richness while the pioneers recruited, infrequent or absent before disturbance, increased both community taxonomic richness and evenness.
+Beyond the intensity threshold, disturbance decreased the taxonomic richness of surviving trees, and this decrease was not offset by the recruitment of pioneers.
+The overall community taxonomic richness therefore decreased according to the disturbance intensity [@Molino2001].
+An intermediate disturbance was then detected, defined by an intensity for which post-disturbance trajectories markedly changed.
+Such disturbance increased the availability of resources and created opportunities for pioneers, without implying too important loss of shade-tolerant species or preventing their maintenance in the community [@Bongers2009].
+For community taxonomic evenness, the disturbance impact was similar but milder.
+Because the evenness is less sensitive to the loss of rare species, below the intensity threshold the trajectories of taxonomic evenness rather represented the increasing dominance of pioneers that balancing the usual hyper-dominance of a few species.
 Beyond the intensity threshold, however, pioneers became in turn highly dominant and decreased the overall evenness [@Baraloto2012a].

### 2018-12-03 Scripts/AGBloss.R

@@ -1,4 +1,5 @@
 load("DB/AGBlostDB")
+load("DB/AllCirc")
 
 init<-AllYears[which(AllYears[,"campagne"]==1984),]
 logg<-AllYears[which(AllYears[,"campagne"]==1988),]
@@ -22,25 +23,30 @@ wd=0.736 # average Wodd density for Vol.ext
  #coord=c(5.18,52.53) paracou latitude and longitude
 AGBi<-unlist(lapply(1:12,function(plo){
 Plo<-init[which(init[,"n_parcelle"]==plo),]
-return(sum(computeAGB(D=Plo[,"circonf"],WD=rep(wd,nrow(Plo)),coord=c(mean(Plo[,"Lon"]),mean(Plo[,"Lat"])))))
+Carr<-unlist(lapply(unique(Plo[,"n_carre"]),function(car){
+  carr<-Plo[which(Plo[,"n_carre"]==car),]
+return(sum(computeAGB(D=carr[,"circ_corr"],WD=rep(wd,nrow(carr)),coord=c(mean(carr[,"Lon"]),mean(carr[,"Lat"])))))
 }))
-names(AGBi)<-1:12
+names(Carr)<-paste(plo,unique(Plo[,"n_carre"]),sep=".")
+return(Carr)
+}))
 
 AGBp<-unlist(lapply(1:12,function(plo){
   Plo<-logg[which(logg[,"n_parcelle"]==plo),]
-  return(sum(computeAGB(D=Plo[,"circonf"],WD=rep(wd,nrow(Plo)),coord=c(mean(Plo[,"Lon"]),mean(Plo[,"Lat"])))))
+  Carr<-unlist(lapply(unique(Plo[,"n_carre"]),function(car){
+    carr<-Plo[which(Plo[,"n_carre"]==car),]
+    return(sum(computeAGB(D=carr[,"circ_corr"],WD=rep(wd,nrow(carr)),coord=c(mean(carr[,"Lon"]),mean(carr[,"Lat"])))))
+  }))
+  names(Carr)<-paste(plo,unique(Plo[,"n_carre"]),sep=".")
+  return(Carr)
+  #return(cbind(rep(plo,length(unique(Plo[,"n_carre"]))),unique(Plo[,"n_carre"]),Carr))
 }))
-names(AGBp)<-1:12
+#colnames(AGBp)<-c("Plot","carre","AGB")
 
[…]

### 2018-12-03 Scripts/DivTaxo_trajectories_Carre.R

@@ -1,0 +1,87 @@
+load("DB/LivingStand_carre")
+source("Scripts/Vernacular_handle.R")
+
+# Treatments definition
+T0<-c(1,6,11);T1<-c(2,7,9);T2<-c(3,5,10);T3<-c(4,8,12)
+treatments<-list(T0,T1,T2,T3);names(treatments)<-c("T0","T1","T2","T3")
+
+# Years with identified problems of inventories
+Pby<-c("1996","1998","2000","2002","2004","2006","2008","2010","2012","2014","2016","2017")
+dates<-sort(names(LivingStand_all)[which(!names(LivingStand_all)%in%Pby)])
+
+# Alpha matrix for vernacular names association
+alphas_plot<-AlphaPlots(LivingStand_all)
+
+# Reference for botany (Genus/species association)
+load("DB/BotanyGenus")
+
+# Moving average function
+smooth<-function(mat,larg){return(do.call(cbind,lapply(1:ncol(mat),function(step){
+  range<-max(1,step-larg):min(ncol(mat),step+larg)
+  rowSums(mat[,range])/length(range)})))}
+
+#trajectories drawn by treatment
+CompleteTaxo<-lapply(1:4,function(t){
+  
+  treat<-treatments[[t]]
+  Nrep<-5   # Number of iteration: replacement of vernacular names
+  print(t)
+  
+  Trajectories<-lapply(c(0,2),function(Q){
+    
+          Repet<-mclapply(1:Nrep,function(r){
+            
+    Plot_trajectory<-do.call(cbind,lapply(dates,function(y){
+      
+      # Select current year and current plots
+      year<-LivingStand_all[[which(names(LivingStand_all)==y)]] 
+      yearPlot<-year[which(names(year)%in%treat)] 
+      
+      if(length(yearPlot)!=0){
+        
+        Div<-lapply(1:length(yearPlot),function(pl){
+          
+        # Replace vernacular names and get corresponding genus
+        Div_carre<-unlist(lapply(unique(yearPlot[[pl]][,"n_carre"]),function(car){
+          carr<-yearPlot[[pl]][which(yearPlot[[pl]][,"n_carre"]==car),]
+          carr<-as.data.frame(Replacement(carr,Alpha=alphas_plot[[names(yearPlot)[pl]]]))
+           colnames(carr)<-"names"
[…]

### 2018-12-03 Scripts/DrawingTools.R

@@ -2,7 +2,7 @@
 Library(c("shape","plotrix"))
 
 treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
-names(treatments)<-c("Control","T1","T2","T3")
+names(treatments)<-c("Control","Low","Intermediate","High")
 ColorsTr<-c("darkolivegreen2","gold","orangered","darkred")
 colyear<-c("deepskyblue","cornflowerblue","darkslateblue")
 time<-c("1995","2005","2015")
@@ -145,7 +145,7 @@ plotDiv<-function(Data,remove=FALSE){
   
   Toplot<-lapply(Data,function(toplot){return(toplot[,which(colnames(toplot)>=1989),])})
   Toplot<-lapply(Toplot,function(tr){
-    ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col}))})#-tr[,1,rep]
+    ret<-lapply(1:dim(tr)[3],function(rep){return(apply(tr[,,rep],2,function(col){col<-col-tr[,1,rep]}))})#
     ret<-array(unlist(ret),dim=c(nrow(ret[[1]]),ncol(ret[[1]]),length(ret)),
                dimnames=list(rownames(ret[[1]]),as.numeric(colnames(ret[[1]]))-1986,1:length(ret)))
     ret<-lapply(c(0.025,0.5,0.975),function(quant){return(apply(ret,c(1,2),function(x){return(quantile(x,probs=quant))}))})

### 2018-12-04 WholePlotTrajectories.Rmd
intro

@@
 They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, determine the cycles of carbon, water and nutrient, and ensure cultural and human well-being.
-The growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Morales-Hidalgo2015] and threatens the maintenance and dynamics in space and time of community structure, composition and functioning [@Anderson-Teixeira2013].
+The growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Morales-Hidalgo2015].
+In this context, the maintenance of tropical forest community structure, composition and functioning is increasingly threatened [@Anderson-Teixeira2013].
 
-In tropical forests, ecological communities are regularly re-shaped by natural disturbance events changing both the abiotic environment, through the fluxes of light, heat and water [@Goulamoussene2017], and the biotic interactions such as competition among species [@Chesson2000].
-One of the cornerstone of tropical forest ecology is to understand the processes and drivers of ecosystems response to such disturbance [@Chazdon2003a].
-This response has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
+Tropical forests are subject to natural disturbance events, specifically tree fall gaps, determining and maintaining the characteristics of ecosystems [@Schnitzer2001]. 
+Disturbances change both the biotic and abiotic environment, through modifications of the fluxes of light, heat and water [@Goulamoussene2017], and of the interactions among species such as competition [@Chesson2000].
+Understand the drivers of community response to disturbance and the underlying ecological processes is then crucial to anticipate the fate of tropical forests in the changing context [@Chazdon2003a].
+Community response has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
 Thereafter, models based on the observed trajectories after disturbance gave important insights into the recovery of ecosystem processes and services [@Herault2018]. 
 Regarding community diversity and composition, however, post-disturbance trajectories are not as thoroughly understood.
-Several studies and meta-analyses highlighted determined successions after disturbance, mainly driven by the recruitment of pioneer and heliophilous species [@Carreno-Rocabado2012;@Poorter2016].
-Community trajectories thus proved to rely on deterministic processes and post-disturbance responses would be predictable according to the disturbance intensity and the comunity diversity and composition.
-Still, observed trajectories in diversity and composition often deviate from predictions and the underlying ecological processes are debated [@Norden2015].
-Deeper understanding of tropical forest response to disturbance could be provided by long term monitoring, to complete the conclusions often based on single decades monitoring [@Carreno-Rocabado2012], dynamic trajectories to complete chronosequence studies [@Chazdon2007], and through a complete set of taxonomic and functional composition and diversity metrics to complete studies based only on taxonomic richness or functional composition [@Carreno-Rocabado2012,@Mouillot2013].
-
-Deterministic processes are thus recognized as important drivers of community trajectories.
-This is formalized by the intermediate disturbance hypothesis (IDH), predicting a miximum species diversity after moderate disturbance intensity [@Molino2001;@Kariuki2006a].
-Validations of the IDH though remain scarce in the long-term and mainly rely on the analysis of taxonomic richness [@Molino2001].
+It is recognized that determined succession, mainly driven by the recruitment of pioneer and heliophilous species, shape post-disturbance trajectories [@Carreno-Rocabado2012;@Poorter2016].
+Community trajectories would then rely on deterministic processes and would be predictable according to the disturbance intensity and to the characteristics of the pre-disturbance community.
+Yet, observed trajectories in diversity and composition often deviate from predictions.
+Besides, although the deterministic ecological processes at stake are identified, their relative importance compared to stochasticity remains debated [@Norden2015].
+This debate translates into the validation and the extent of the intermediate disturbance hypothesis (IDH).
+The IDH assumes the predominance of deterministic processes, and predicts a maximum species diversity after moderate disturbance intensity [@Molino2001;@Kariuki2006a].
+Validations of the IDH remain scarce in the long-term and often rely on the analysis of taxonomic richness [@Molino2001]. 
 Taxonomic richness alone, however, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016]. 
-More ecological-meaningful analysis would couple richness with (i) evenness, which would reveal the changes in the species abundance distribution and thus the underlying ecological processes, and (ii) composition, which is crucial for conservation issues [@Lavorel2002;@Bellwood2006].
+More ecological-meaningful analysis would couple richness with species evenness, which reveals the changes in the species abundance distribution, and (ii)species composition, which is crucial for conservation issues [@Lavorel2002;@Bellwood2006].
 Furthermore, a functional approach accounting for species biological attributes would directly link community diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Baraloto2012a].
@@
 The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Elmqvist2003; @Diaz2005].
+
+Improve the understanding of diversity and composition response to disturbance would then require to combine taxonomic and functional diversity and composition metrics that is not often provided [@Carreno-Rocabado2012,@Mouillot2013].
+Long term monitoring provided trajectories over several decades would allow to put in perspectives the existing studies that often correspond to chronosequence studies [@Chazdon2007] or cover 10 or 15 years after disturbance [@Carreno-Rocabado2012].
 

### 2018-12-04 WholePlotTrajectories.tex
intro

@@
 multiple non-timber forest products, shelter a diversified fauna,
-regulate the local and regional climates, the carbon, water and nutrient
-cycles, and ensure cultural and human well-being. The growing demand in
-forests products together with current global changes increases the
-pressure on remaining natural forests \citep{Morales-Hidalgo2015} and
-threatens the maintenance and dynamics in space and time of community
-structure, composition and functioning \citep{Anderson-Teixeira2013}.
+regulate the local and regional climates, determine the cycles of
+carbon, water and nutrient, and ensure cultural and human well-being.
+The growing demand in forests products together with current global
+changes increases the pressure on remaining natural forests
+\citep{Morales-Hidalgo2015}. In this context, the maintenance of
+tropical forest community structure, composition and functioning is
+increasingly threatened \citep{Anderson-Teixeira2013}.
 
-In tropical forests, ecological communities are regularly re-shaped by
-natural disturbance events changing both the abiotic environment,
-through the fluxes of light, heat and water \citep{Goulamoussene2017},
-and the biotic interactions such as competition among species
-\citep{Chesson2000}. One of the cornerstone of tropical forest ecology
-is to understand the processes and drivers of ecosystems response to
-disturbance \citep{Chazdon2003a}. For now, this has been largely studied
-through forest structural parameters such as aboveground biomass, tree
-height or stem density \citep{Piponiot2016, Rutishauser2016} that are
-rapid and convenient to measure. These structural parameters have been
-successfully modeled, giving important insights into the recovery of
-ecosystem processes and services \citep{Herault2018}. However the
-response of forests diversity and composition remains unclear, albeit it
-determines the productivity, stability and functioning of ecosystems
-\citep{Tilman2014, Liang2016}. In the short-term, moderate disturbance
-may lead to positive impacts on community diversity, an idea formalized
-by the intermediate disturbance hypothesis (IDH) which predicts a
-maximum species diversity when disturbance intensity is not too high
-\citep{Molino2001, Kariuki2006a}.
-
-Validations of the IDH though remain scarce in the long-term and mainly
-rely on the analysis of taxonomic richness \citep{Molino2001}. Taxonomic
-richness alone, however, gives limited or misleading information on
-forests recovery and functioning \citep{Chaudhary2016}. More
-ecological-meaningful analysis would couple richness with (i) evenness,
-which would reveal the changes in the species abundance distribution and
-thus the underlying ecological processes, and (ii) composition, which is
-crucial for conservation issues \citep{Lavorel2002, Bellwood2006}.
-Furthermore, a functional approach accounting for species biological
-attributes would directly link community diversity, composition and
-redundancy to ecosystem functioning and to its environmental constraints
+Tropical forests are subject to natural disturbance events, specifically
+tree fall gaps, determining and maintaining the characteristics of
+ecosystems \citep{Schnitzer2001}. Disturbances change both the biotic
+and abiotic environment, through modifications of the fluxes of light,
+heat and water \citep{Goulamoussene2017}, and of the interactions among
+species such as competition \citep{Chesson2000}. Understand the drivers
+of community response to disturbance and the underlying ecological
+processes is then crucial to anticipate the fate of tropical forests in
+the changing context \citep{Chazdon2003a}. Community response has been
+largely studied through forest structural parameters such as aboveground
+biomass, tree height or stem density
+\citep{Piponiot2016, Rutishauser2016}. Thereafter, models based on the
+observed trajectories after disturbance gave important insights into the
+recovery of ecosystem processes and services \citep{Herault2018}.
+Regarding community diversity and composition, however, post-disturbance
+trajectories are not as thoroughly understood. It is recognized that
+determined succession, mainly driven by the recruitment of pioneer and
+heliophilous species, shape post-disturbance trajectories
+\citep{Carreno-Rocabado2012, Poorter2016}. Community trajectories would
+then rely on deterministic processes and would be predictable according
+to the disturbance intensity and to the characteristics of the
+pre-disturbance community. Yet, observed trajectories in diversity and
+composition often deviate from predictions. Besides, although the
+deterministic ecological processes at stake are identified, their
+relative importance compared to stochasticity remains debated
+\citep{Norden2015}. This debate translates into the validation and the
+extent of the intermediate disturbance hypothesis (IDH). The IDH assumes
+the predominance of deterministic processes, and predicts a maximum
+species diversity after moderate disturbance intensity
+\citep{Molino2001, Kariuki2006a}. Validations of the IDH remain scarce
+in the long-term and often rely on the analysis of taxonomic richness
+\citep{Molino2001}. Taxonomic richness alone, however, gives limited or
+misleading information on forests recovery and functioning
+\citep{Chaudhary2016}. More ecological-meaningful analysis would couple
+richness with species evenness, which reveals the changes in the species
+abundance distribution, and (ii)species composition, which is crucial
+for conservation issues \citep{Lavorel2002, Bellwood2006}. Furthermore,
+a functional approach accounting for species biological attributes would
+directly link community diversity, composition and redundancy to
+ecosystem functioning and to its environmental constraints
 \citep{Violle2007b, Baraloto2012a}. In that respect, the functional
@@
 disturbance \citep{Elmqvist2003, Diaz2005}.
+
+Improve the understanding of diversity and composition response to
+disturbance would then require to combine taxonomic and functional
+diversity and composition metrics that is not often provided
+\citep[\citet{Mouillot2013}]{Carreno-Rocabado2012}. Long term monitoring
+provided trajectories over several decades would allow to put in
+perspectives the existing studies that often correspond to
+chronosequence studies \citep{Chazdon2007} or cover 10 or 15 years after
+disturbance \citep{Carreno-Rocabado2012}.
 
@@
 trajectories of community composition were followed in a two-dimensional
-NMDS ordination space. Two NMDS using abundance-based (Bray-Curtis)
+NMDS ordination plan. Two NMDS using abundance-based (Bray-Curtis)
 dissimilarity measures were conducted to map either taxonomic or
@@
 The taxonomic diversity was reported through species richness and the
-Hill number translation of the Simpson index \citep{Hill1973}. The
-comparison between richness and Simpson diversity assessed the
-variations in taxonomic evenness. Both indices belong to the set of HCDT
-or generalized entropy, respectively corresponding to the 0 and 2 order
-of diversity (q), recommended for diversity studies \citep{Marcon2015b}.
-The functional diversity was reported using the functional richness and
-functional evenness, \emph{i.e} the Rao index of quadratic entropy which
-combines species abundance distribution and average pairwise
-dissimilarity based on species functional traits.
+Hill number translation of the Simpson index \citep{Hill1973}. These
+metrics allowed assessing the taxonomic richness as well as evenness,
+through the comparison between richness and Simpson diversity. R.esults
+will thus be discussed directly in terms of taxonomic richness and
+evenness. Both indices belong to the set of HCDT or generalized entropy,
+respectively corresponding to the 0 and 2 order of diversity (q),
+recommended for diversity studies \citep{Marcon2015b}. The functional
+diversity was reported using the functional richness and functional
+evenness, \emph{i.e} the Rao index of quadratic entropy. The Rao index
+combines species abundance distribution, and the average pairwise
+functional dissimilarity between species computed by the gower distance.
 
@@
 from a PCA analysis. Then, for each species, the traits probability
-density (TPD) were computed through multivariate kernel density
-estimators. Second, for each community, the functional traits
-probability density were computed as the sum of the TDP of the
-inventoried species weighted by species abundance. Third, the whole
-functional space was divided into a 100 x 100 grid, before counting for
-each cell the number of species which TPD is positive. The average count
-across cells, minus 1, gives the average number of species in the
-community that share the same traits, which is the Community Functional
-Redundancy.
+density (TPD) were computed through two-dimension kernel density
+estimators. Second, for each community, the TDB weigthed by species
+abundance were summed accross the functional space. Third, the
+functional space was divided into a 100 x 100 grid and the number of
+species with a positive TDP was counted in each cell. The average count
+across cells minus 1 returnedthe Community Functional Redundancy, which
+was the average number of species in the community that share the same
+trait values.
 
@@
 
-\caption{Plot trajectories in terms of taxonomic composition (\textbf{(a)} and \textbf{(c)}) and functional composition (\textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS space.
+\caption{Plot trajectories in terms of taxonomic composition (\textbf{(a)} and \textbf{(c)}) and functional composition (\textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS plan.
 Lower panels (\textbf{(c)} and \textbf{(d)}) represent the Euclidean distance to initial condition along the 30 sampled years.
@@
 evenness (\(0.21<R^2_{Taxonomic Richness}<0.4\), and
-\(-0.15<R^2_{Taxonomic Simpson}<0.43\) respectively)
+\(-0.15<R^2_{Taxonomic Simpson}<0.43\) respectively).
 
@@
 recruited species and constrained the trajectories towards the initial
-composition. The high resilience of community taxonomy, in the sense of
-the recovery of initial state, revealed that species not belonging to
-the pre-disturbance community were rarely recruited because of the
-commonness of dispersal limitation among tropical tree species
-\citep{Svenning2005}.
+composition. The community taxonomy proved highly resilient, as it
+absorbed the disturbance and maintained commnity initial composition
+characteristics \citep{Folke2006}. This high resilience suggested that
+species not belonging to the pre-disturbance community were rarely
+recruited because of the commonness of dispersal limitation among
+tropical tree species \citep{Svenning2005}.
 
@@
 
-The IDH predicted well the enhancement of taxonomic richness after
-disturbance, until an intensity threshold (20-25\% AGB loss). It also
-explained the impact on the Simpson diversity and hence on community
-taxonomic evenness. Taxonomic evenness was somewhat decoupled from the
-disturbance intensity as already observed in the Guiana Shield
-\citep{Baraloto2012a} and in Bornean tropical forests
-\citep{Cannon1998}. Disturbance intensity determined the balance in the
-community between pre-disturbance surviving trees and those recruited
-afterward. The pool of true pioneer species specifically recruited after
-disturbance is restricted in the Guiana Shield to a few common genera
-(e.g. \emph{Cecropia} spp., \emph{Vismia} spp.) \citep{Guitet2018}.
-Below the intensity threshold the size of the surviving community
-maintained the pre-disturbance high taxonomic richness while the
-recruitment of pioneers, infrequent or absent before disturbance,
-increased both community taxonomic richness and evenness. Beyond the
-intensity threshold, the disturbance decreased the taxonomic richness of
-surviving trees which was not offset by the enrichment of pioneers, so
-that the overall community taxonomic richness decreased according to the
-disturbance intensity \citep{Molino2001}. An intermediate disturbance
-hence increased the availability of resources and created opportunities
-for pioneers, without implying too important loss of shade-tolerant
-species and without preventing their maintenance in the community
-\citep{Bongers2009}. For community taxonomic evenness the disturbance
-impact was similar but milder, as the evenness is less sensitive to the
-loss of rare species. Taxonomic evenness trajectories, below the
-intensity threshold, rather represented the increasing dominance of
-pioneers that balanced the usual hyper-dominance of a few species in
-tropical forests. Beyond the intensity threshold, however, pioneers
-became in turn highly dominant and decreased the overall evenness
-\citep{Baraloto2012a}.
+Community taxonomic richness increased after disturbance until an
+intensity threshold (20-25\% AGB loss) above which the taxonomic
+richness decreased. The increase of taxonomic evenness after disturbance
+also depended on the disturbance intensity, although the correlation was
+weaker than for the taxonomic richness as already observed in the Guiana
+Shield \citep{Baraloto2012a} and in Bornean tropical forests
+\citep{Cannon1998}. In accordance with the IDH, then, community
+taxonomic trajectories depended on the disturbance intensity. They
+markedly changed above an intensity threshold, with a decrease of the
+richness and a persistent increase of the evenness. Disturbance
+intensity determined the balance in the community between surviving
+trees from pre-disturbance communities and trees recruited afterward.
+The pool of true pioneer species specifically recruited after
+disturbance is, in the Guiana Shield, restricted to a few common genera
+(e.g. \emph{Cecropia} spp., \emph{Vismia} spp.)\citep{Guitet2018}. Below
+the intensity threshold the surviving community trees remained numerous
+enough to maintain the pre-disturbance high taxonomic richness while the
+pioneers recruited, infrequent or absent before disturbance, increased
+both community taxonomic richness and evenness. Beyond the intensity
+threshold, disturbance decreased the taxonomic richness of surviving
+trees, and this decrease was not offset by the recruitment of pioneers.
+The overall community taxonomic richness therefore decreased according
+to the disturbance intensity \citep{Molino2001}. An intermediate
+disturbance was then detected, defined by an intensity for which
+post-disturbance trajectories markedly changed. Such disturbance
+increased the availability of resources and created opportunities for
+pioneers, without implying too important loss of shade-tolerant species
+or preventing their maintenance in the community \citep{Bongers2009}.
+For community taxonomic evenness, the disturbance impact was similar but
+milder. Because the evenness is less sensitive to the loss of rare
+species, below the intensity threshold the trajectories of taxonomic
+evenness rather represented the increasing dominance of pioneers that
+balancing the usual hyper-dominance of a few species. Beyond the
+intensity threshold, however, pioneers became in turn highly dominant
+and decreased the overall evenness \citep{Baraloto2012a}.
 

### 2018-12-05 WholePlotTrajectories.Rmd
intro

@@
 
-The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
+Tropical forests cover large areas worldwide and represent crucial environmental, economic and social values. 
 They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, determine the cycles of carbon, water and nutrient, and ensure cultural and human well-being.
-The growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Morales-Hidalgo2015].
-In this context, the maintenance of tropical forest community structure, composition and functioning is increasingly threatened [@Anderson-Teixeira2013].
+However, the growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Morales-Hidalgo2015].
+This context specifically threatens the maintenance of community structure, composition and functioning [@Anderson-Teixeira2013].
+In tropical forests indeed, tree communities are shaped and maintained by a natural disturbance regime, specifically tree fall gaps[@Schnitzer2001].
+This disturbance regime is affected by the current changing context and anthropic activities [@Sist2015], so clarify community responses to disturbance and the underlying ecological processes is crucial to anticipate the fate of tropical forests [@Chazdon2003a].
 
-Tropical forests are subject to natural disturbance events, specifically tree fall gaps, determining and maintaining the characteristics of ecosystems [@Schnitzer2001]. 
-Disturbances change both the biotic and abiotic environment, through modifications of the fluxes of light, heat and water [@Goulamoussene2017], and of the interactions among species such as competition [@Chesson2000].
-Understand the drivers of community response to disturbance and the underlying ecological processes is then crucial to anticipate the fate of tropical forests in the changing context [@Chazdon2003a].
+Disturbances change both biotic and abiotic environment, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and of the interactions among species such as competition [@Chesson2000].
 Community response has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
@@
 Eventually, a complete overview of community response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
-The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Elmqvist2003; @Diaz2005].
+The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines community resilience after disturbance [@Elmqvist2003; @Diaz2005].
 
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, we (i) elucidated community taxonomic and functional recovery and the underlying ecological processes, (ii) clarified the validity of the IDH in the long term for tropical forest and its translation into different trajectories in time, and (iii) examined community recovery time.
+From post-disturbance trajectories (i) we determined the taxonomic and functional translation post-disturbance succession at community scale.
+(ii) we validated the importance of deterministic processes in community response to disturbance.
+We defined th interplay of deterministic and stochastic processes along time, which allowed discussing the IDH.
+(iii) examined community recovery time.
 

### 2018-12-05 WholePlotTrajectories.tex
intro

@@
 
-The large areas covered with tropical forests worldwide hold crucial
+Tropical forests cover large areas worldwide and represent crucial
 environmental, economic and social values. They provide wood and
@@
 carbon, water and nutrient, and ensure cultural and human well-being.
-The growing demand in forests products together with current global
-changes increases the pressure on remaining natural forests
-\citep{Morales-Hidalgo2015}. In this context, the maintenance of
-tropical forest community structure, composition and functioning is
-increasingly threatened \citep{Anderson-Teixeira2013}.
+However, the growing demand in forests products together with current
+global changes increases the pressure on remaining natural forests
+\citep{Morales-Hidalgo2015}. This context specifically threatens the
+maintenance of community structure, composition and functioning
+\citep{Anderson-Teixeira2013}. In tropical forests indeed, tree
+communities are shaped and maintained by a natural disturbance regime,
+specifically tree fall gaps\citep{Schnitzer2001}. This disturbance
+regime is affected by the current changing context and anthropic
+activities \citep{Sist2015}, so clarify community responses to
+disturbance and the underlying ecological processes is crucial to
+anticipate the fate of tropical forests \citep{Chazdon2003a}.
 
-Tropical forests are subject to natural disturbance events, specifically
-tree fall gaps, determining and maintaining the characteristics of
-ecosystems \citep{Schnitzer2001}. Disturbances change both the biotic
-and abiotic environment, through modifications of the fluxes of light,
-heat and water \citep{Goulamoussene2017}, and of the interactions among
-species such as competition \citep{Chesson2000}. Understand the drivers
-of community response to disturbance and the underlying ecological
-processes is then crucial to anticipate the fate of tropical forests in
-the changing context \citep{Chazdon2003a}. Community response has been
-largely studied through forest structural parameters such as aboveground
+Disturbances change both biotic and abiotic environment, through
+modifications in the fluxes of light, heat and water
+\citep{Goulamoussene2017}, and of the interactions among species such as
+competition \citep{Chesson2000}. Community response has been largely
+studied through forest structural parameters such as aboveground
 biomass, tree height or stem density
@@
 \citep{Bellwood2006} mitigates the impacts of species removal on
-ecosystem functioning and determines the resilience of communities after
+ecosystem functioning and determines community resilience after
 disturbance \citep{Elmqvist2003, Diaz2005}.
@@
 trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, we (i) elucidated community taxonomic and functional
-recovery and the underlying ecological processes, (ii) clarified the
-validity of the IDH in the long term for tropical forest and its
-translation into different trajectories in time, and (iii) examined
-community recovery time.
+From post-disturbance trajectories (i) we determined the taxonomic and
+functional translation post-disturbance succession at community scale.
+(ii) we validated the importance of deterministic processes in community
+response to disturbance. We defined th interplay of deterministic and
+stochastic processes along time, which allowed discussing the IDH. (iii)
+examined community recovery time.
 

### 2018-12-07 WholePlotTrajectories.Rmd
intro

@@
 
-Disturbances change both biotic and abiotic environment, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and of the interactions among species such as competition [@Chesson2000].
+Disturbances change both biotic and abiotic environment, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and modifications in the interactions among species [@Chesson2000].
 Community response has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
-Thereafter, models based on the observed trajectories after disturbance gave important insights into the recovery of ecosystem processes and services [@Herault2018]. 
+Thereafter, models based on the observed trajectories after disturbance gave important insights into community recovery and resilience [@Herault2018]. 
 Regarding community diversity and composition, however, post-disturbance trajectories are not as thoroughly understood.
-It is recognized that determined succession, mainly driven by the recruitment of pioneer and heliophilous species, shape post-disturbance trajectories [@Carreno-Rocabado2012;@Poorter2016].
-Community trajectories would then rely on deterministic processes and would be predictable according to the disturbance intensity and to the characteristics of the pre-disturbance community.
+It is recognized that a determined succession, mainly driven by the recruitment of pioneer and heliophilous species, shape community trajectories afte disturbance [@Carreno-Rocabado2012;@Poorter2016].
+Deterministic processes would then be major processes of community response to disturbance.
+This would make post-disturbance trajectories deterministic, and then predictable according to the disturbance intensity and to the characteristics of the pre-disturbance community.
 Yet, observed trajectories in diversity and composition often deviate from predictions.
-Besides, although the deterministic ecological processes at stake are identified, their relative importance compared to stochasticity remains debated [@Norden2015].
+Although the deterministic ecological processes at stake are identified, They interplay with stochastic processes and their relative importance remains debated [@Norden2015].
 This debate translates into the validation and the extent of the intermediate disturbance hypothesis (IDH).
 The IDH assumes the predominance of deterministic processes, and predicts a maximum species diversity after moderate disturbance intensity [@Molino2001;@Kariuki2006a].
-Validations of the IDH remain scarce in the long-term and often rely on the analysis of taxonomic richness [@Molino2001]. 
+VThe IDH has been scarcely validated in the long-term and often rely on the analysis of taxonomic richness [@Molino2001]. 
 Taxonomic richness alone, however, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016]. 

### 2018-12-14 WholePlotTrajectories.Rmd
review

@@
     keep_tex: TRUE
-  bookdown::word_document2: default
-  bookdown::gitbook: default
 ---
@@
 Regarding community diversity and composition, however, post-disturbance trajectories are not as thoroughly understood.
-It is recognized that a determined succession, mainly driven by the recruitment of pioneer and heliophilous species, shape community trajectories afte disturbance [@Carreno-Rocabado2012;@Poorter2016].
+It is recognized that a determined succession, mainly driven by the recruitment of pioneer and heliophilous species, shape community post-disturbance trajectories [@Carreno-Rocabado2012;@Poorter2016].
 Deterministic processes would then be major processes of community response to disturbance.
@@
 Yet, observed trajectories in diversity and composition often deviate from predictions.
-Although the deterministic ecological processes at stake are identified, They interplay with stochastic processes and their relative importance remains debated [@Norden2015].
+Although the deterministic ecological processes at stake are identified, their interplay with stochastic processes and their relative importance remains debated [@Norden2015].
 This debate translates into the validation and the extent of the intermediate disturbance hypothesis (IDH).
 The IDH assumes the predominance of deterministic processes, and predicts a maximum species diversity after moderate disturbance intensity [@Molino2001;@Kariuki2006a].
-VThe IDH has been scarcely validated in the long-term and often rely on the analysis of taxonomic richness [@Molino2001]. 
+The IDH has been scarcely validated in the long-term and often rely on the analysis of taxonomic richness [@Molino2001]. 
 Taxonomic richness alone, however, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016]. 
@@
 Intervention <- data.frame(
-  Treatment=c("Control", "T1", "T2", "T3"),
+  Treatment=c("Control", "T1, low", "T2, intermediate", "T3, high"),
   Timber=c("-", 
@@
 
-par(mfrow=c(2,2),mar=c(2,2.5,2.5,1),oma=c(1,2,1,3),no.readonly = T)
+par(mfrow=c(2,2),mar=c(2,2.5,2.5,1),oma=c(1,2,1,4),no.readonly = T)
 TaxoCompo(MatrepTaxo)
@@
 mtext("Years since disturbance",side=1,line=2.2,adj=1,cex=0.8)
-legend("right",inset=c(-0.32,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
+legend("right",inset=-0.4,xpd=NA,legend=c("Control","Low","Interm.","High"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment", cex=0.85)
 
@@
 legendCWM()
-legend("right",inset=c(-1,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
+legend("right",inset=c(-1,0),xpd=NA,legend=c("Control","Low","Intermediate","High"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
 ```
@@
 mtext("Years since disturbance",side=1,line=2.2,adj=1)
-legend("right",xpd=NA,inset=-0.4,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
+legend("right",xpd=NA,legend=c("Control","Low","Intermediate","High"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment", cex=0.85,inset=-0.57)
 
@@
 
-Community taxonomic and functional composition appeared resilient, following similar hump-shaped trajectories starting to return towards pre-disturbance composition after 30 years.
+The hump-shaped trajectories of community composition returned towards pre-disturbance state after 30 years, highlighting community resilience in terms of taxonomic and functional composition.
 
-The taxonomic differences among local communities were already apparent before disturbance, as revealed by the distinct starting points on the NMDS axis 2. 
-These differences were maintained throughout recovery trajectories.  
-More than commonly thought, post-disturbance trajectories depended on community initial composition, which partly determined the pool of recruited species and constrained the trajectories towards the initial composition.
-The community taxonomy proved highly resilient, as it absorbed the disturbance and maintained commnity initial composition characteristics [@Folke2006].
-This high resilience suggested that species not belonging to the pre-disturbance community were rarely recruited because of the commonness of dispersal limitation among tropical tree species [@Svenning2005].
+Before disturbance, there were already taxonomic differences among local communities, as revealed by the distinct starting points on the NMDS axis 2.
+These initial differences were maintained throughout the trajectories following disturbance, so community initial state influenced the trajectories following disturbance more than commonly thought.
+The initial composition indeed partly determined the pool of species recruited, and hence shaped the composition trajectories and drove them towards the recovery of the initial composition.
+This taxonomic recovery, although it was not fully achieved after 30 years, suggested the resilience of community taxonomic composition and the maintenance of community initial differences [@Folke2006].
+Species not belonging to the pre-disturbance community were then rarely recruited, probably because of the common dispersal limitations among tropical tree species [@Svenning2005].
 
-Conversely, disturbed communities followed functional trajectories that are highly similar in terms of functional composition.
+Functional composition trajectories were, conversely, highly similar among disturbed communities.
 As the composition of pre-disturbance surviving trees is representative of the initial community [@Herault2018], changes in functional composition relied upon the recruitment of species or functional types that were infrequent or absent before disturbance.
-Competitive pioneers became dominant in filling the environmental niches of high availability of light, space and nutrients vacated by the disturbance.
-The recruitment of pioneers changed community functional composition in the same way for all disturbance intensity towards more resource-acquisitive strategies, moving community functional composition right along the first axis in Fig.
-\@ref(fig:NMDSplans) [@Westoby1998; @Wright2004; @Reich2014].
-Thereafter long-lived, more resistant and shade-tolerant species excluded the first established pioneers and started the recovery of pre-disturbance functional composition, moving similarly community functional composition left along the first axis and upward along the second axis in Fig.
-\@ref(fig:NMDSplans).
+Competitive pioneers became dominant in filling the environmental niches vacated by the disturbance, with a high availability of light, space and nutrients.
+This recruitment of pioneers similarly changed the functional composition of all communities, whenever the disturbance intensity.
+The functional composition went towards more resource-acquisitive strategies, translated in the functional plan (Fig.
+\@ref(fig:NMDSplans)) by a displacement on the right along the first axis [@Westoby1998; @Wright2004; @Reich2014].
+Thereafter, the pioneers first recruited were progressively excluded by long-lived, more resistant and shade-tolerant species.
+The commpunity functional composition then returned towards more resource-conservative strategies, suggesting the recovery of the initial community composition and translated in the functional plan by a displacement left along the first axis and upward along the second axis (Fig.
+\@ref(fig:NMDSplans)).
 
-These trajectories provided empirical support to the hypothesis that community assembly is both deterministic and historically convergent at different levels of community organization.
-Deterministic, trait-based processes drove community convergence in functional composition, while at the same time dispersal limitation maintained their divergence in taxonomic composition [@Fukami2005].
+In the light of these taxonomic and functional composition trajectories, community response to disturbance proved simultaneously deterministic and stochastic at different levels of community organization.
+Taxonomic trajectories were shaped by dispersal limitations maintaining the stochastic differences among communities due to historical contingency. 
+Functional trajectories were shaped by trait-based processes resulting the the convergence of community functional composition [@Fukami2005].
 
@@
 
-Community taxonomic richness increased after disturbance until an intensity threshold (20-25% AGB loss) above which the taxonomic richness decreased.
-The increase of taxonomic evenness after disturbance also depended on the disturbance intensity, although the correlation was weaker than for the taxonomic richness as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
-In accordance with the IDH, then, community taxonomic trajectories depended on the disturbance intensity.
-They markedly changed above an intensity threshold, with a decrease of the richness and a persistent increase of the evenness.
-Disturbance intensity determined the balance in the community between surviving trees from pre-disturbance communities and trees recruited afterward.
-The pool of true pioneer species specifically recruited after disturbance is, in the Guiana Shield, restricted to a few common genera (e.g.
-*Cecropia* spp., *Vismia* spp.)[@Guitet2018].
-Below the intensity threshold the surviving community trees remained  numerous enough to maintain the pre-disturbance high taxonomic richness while the pioneers recruited, infrequent or absent before disturbance, increased both community taxonomic richness and evenness.
-Beyond the intensity threshold, disturbance decreased the taxonomic richness of surviving trees, and this decrease was not offset by the recruitment of pioneers.
-The overall community taxonomic richness therefore decreased according to the disturbance intensity [@Molino2001].
-An intermediate disturbance was then detected, defined by an intensity for which post-disturbance trajectories markedly changed.
-Such disturbance increased the availability of resources and created opportunities for pioneers, without implying too important loss of shade-tolerant species or preventing their maintenance in the community [@Bongers2009].
-For community taxonomic evenness, the disturbance impact was similar but milder.
-Because the evenness is less sensitive to the loss of rare species, below the intensity threshold the trajectories of taxonomic evenness rather represented the increasing dominance of pioneers that balancing the usual hyper-dominance of a few species.
-Beyond the intensity threshold, however, pioneers became in turn highly dominant and decreased the overall evenness [@Baraloto2012a].
+Community taxonomic richness and evenness trajectories both proved to depend on the intensity of disturbance and to markedly change above an intensity threshold, thus supporting the IDH.
+The disturbance intensity determined the balance in the community between pre-disturbance surviving trees, and trees recruited afterward. 
+Below an intensity of 20%-25% AGB loss, the trees surviving after disturbance remained numerous enough to maintain the high taxonomic richness of the pre-disturbance community.
+The recruitment of pionners, infrequent or absent before disturbance, then increased the community taxonomic richness positively correlated with the disturbance intensity [@Martin2015; @Chaudhary2016].
+Besides, the pioneers became increasingly dominant and temporarily balanced the usual hyper-dominance of a few species [@Baraloto2012a]. 
+The taxonomic evenness then followed hump-shaped trajectoriesas already observed in the Guiana shield [@Baraloto2012a] and in Borenan tropical forests [@Cannon1998], with first an increase positively correlated to the disturbance intensity, and then an return towards initial values.
+Above the intensity threshold, the taxonomic richness of surviving trees was too low to be offset by the recruitment of pioneers. 
+In the Guiana shield indeed, the pool of true pioneers specifically recruited after disturbance is restricted to a few common genera (e.g. *Cecropia* spp., *Vismia* spp.) [@Guitet2018].
+The taxonomic richness trajectories following intense disturbance then decreased, all the more so that the disturbance intensity was high [@Molino2001].
+At the same time, pioneers became persistently dominant and prevented a return towards the initial evenness and the recovery of hyper-dominant shade-tolerant species.
+The intensity threshold then marked two distinct post-disturbance trajectories, and a change in community response to disturbance. 
+In accordance with the IDH, an intense disturbance below the intensity threshold would then maximise community richness and evenness.
+Such intermediate disturbance would increase the availability of resources, creating opportunities for pioneers, while allowing a high diversity among surviving trees and the maintenance of shade-tolerant species [@Bongers2009].
 
-Conversely the IDH was disproved regarding the disturbance impact on community functional richness and evenness.
-Irrespective of the disturbance intensity, both community functional richness and evenness increased following the recruitment of pioneers, functionally highly different from the composition of pre-disturbance community.
+Regarding community functional response in contrast, the IDH seemed disproved.
+Whenever the disturbance intensity, both community functional richness and evenness increased, with the recruitment of pioneers that were functionally highly different from the species of the pre-disturbance community [@Denslow1980;@Molino2001].
+Above the intensity threshold however, for the most intense disturbance, community functional richness and evenness started to decrease after 15 to 20 years. 
+The vacated environmental space was then occupied by short-lived species which prevented the establishment of other species. 
+The decline of these short-lived pioneers decreased the functional richness and evenness, but we suggest that the establishment of long-lasting pioneers will follow and the taxonomic and functional trajectories will catch up with those observed for intermediate disturbance [@Walker2009].
 
-Along time, taxonomic richness trajectories of all disturbed communities similarly dropped at first place, following the species loss due to disturbance, and then displayed a species gain depending on the disturbance intensity.
-Up to an intensity threshold, the species gain was all the more significant that the disturbance intensity increased, with the establishment of long-lived pioneers enhancing community taxonomic richness and evenness in the long term.
-These long-lived pioneers, functionally quite different from the functional composition, entailed as well a progressive and long-lasting increase of the functional richness and evenness [@Denslow1980;@Molino2001].
-Beyond an intensity threshold, though, a few short-lived pioneers occupied the vacated environmental space and prevented the establishment of other species. 
-These short-lived pioneers were functionally very different from the pre-disturbance community and entailed a rapid and significant increase of functional richness and evenness.
-Already after 10 years, though, short-lived pioneers started to decline and the functional richness and evenness decreased.
-Likely this decrease will be followed by the establishment of long-lasting pioneers, and by the time they recruit we expect the taxonomic and functional trajectories to catch up with those observed after intermediate disturbance [@Walker2009].
+The IDH then translated into community post-disturbance taxonomic trajectories, with marked differences below and above an intensity threshold. 
+It was however not tangible regarding the functional trajectories, because of their decoupling with taxonomic trajectories.
 
@@
 
-For 15 years the species loss during disturbance, determined by the disturbance intensity, significantly decreased the functional redundancy within the pre-disturbance functional space.
-The redundancy decrease was not compensated for in the first place as the first recruited pioneers were functionally different from the pre-disturbance functional composition.
-Progressively though, first established species were replaced by more competitive long-lived pioneers or late-successional species resembling more the pre-disturbance functional composition and restoring the functional redundancy.
-This replacement was stochastic and followed the lottery recruitment rules, implying a recruitment eased for the first recruited species but then increasingly hampered by the emergence of interspecific competition [@Busing2002].
-Along time the recovery of infrequent species was increasingly slow, so that the time to full recovery of the functional redundancy, in some communities just initiated after 30 years, was extremely difficult to estimate [@Elmqvist2003;@Diaz2005].
+The decoupling between taxonomic and functional trajectories was explained by a decrease in the functional redundancy within the pre-disturbance functional space, due to the loss of species following disturbance.
+The recruitment of pioneers, functionally different from the pre-disturbance functional composition, did not compensate the decrease of functional redundancy in the first place.
+Progressively though, the functional redundancy was restored through the replacement of the first established species by more competitive, long-lived pioneers or late-successional species that were functionally closer to the pre-disturbance community.
+This replacement followed the lottery recruitment rules, implying a recruitment easy for the first recruited species but increasingly hampered by the emergence of interspecific competition [@Busing2002].
+The functional redundancy then relied upon stochastic processes and its recovery, depending on the recruitment of infrequent species constrained by the increasing competition, was increasingly slow and difficult to anticipate [@Elmqvist2003;@Diaz2005].
 
-The mid-term impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008].
-Besides, the mid-term recovery of infrequent species increases the risks to loose keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
+The impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008].
+Besides, the stochastic recovery of infrequent species increases the risks to lose keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
 Apart from the functional characteristics considered here, infrequent species might indeed have unique functions in the ecosystem or be a key for some fauna [@Schleuning2016].
@@
 
-Our study revealed community recovery through the combination of deterministic processes driving their convergence in functional composition, and dispersal limitation maintaining their divergence in taxonomic composition.
-The IDH was validated for community taxonomic richness and, to some extent, taxonomic evenness but disproved regarding community functional richness and evenness which were enhanced for any disturbance intensity by the high functional differences of pioneers compared to late-successional functional composition.
-The IDH was translated in time by the recruitment, beyond an intensity threshold, of short-lived pioneers that prevented in the first years after disturbance the establishment of more diverse long-lived pioneers, recruited otherwise below the intensity threshold.
+Our study highlighted the combination of deterministic and stochastic processes determining the succession of recruited species and shaping community response to disturbance. 
+The deterministic recruitment of pioneers shaped functional trajectories convergent in the functional space, while dispersal limitations shaped stochastic taxonomic trajectories maintaining community divergence in taxonomic composition.
+In accordance with the IDH, community taxonomic diversity was maximise for intermediate disturbance intensity, before reaching an intensity threshold beyond which the taxonomic richness decreased and the taxonomic evenness remained persistently higher.
+Conversely, functional trajectories proved decoupled from the taxonomic ones and did not follow the hypothesis of the IDH as no distinct trajectories emerged with increasing disturbance intensity. 
+The decoupling between taxonomic and functional disturbance was explained by a decrease of community functional redundancy mitigating the functoinal impact of disturbance.
 The resilience of tropical forests, defined in terms of recovery to pre-disturbance state, proved tangible but requiring several decades. 

### 2018-12-14 WholePlotTrajectories.tex
review

@@
 modifications in the fluxes of light, heat and water
-\citep{Goulamoussene2017}, and of the interactions among species such as
-competition \citep{Chesson2000}. Community response has been largely
-studied through forest structural parameters such as aboveground
-biomass, tree height or stem density
-\citep{Piponiot2016, Rutishauser2016}. Thereafter, models based on the
-observed trajectories after disturbance gave important insights into the
-recovery of ecosystem processes and services \citep{Herault2018}.
-Regarding community diversity and composition, however, post-disturbance
-trajectories are not as thoroughly understood. It is recognized that
-determined succession, mainly driven by the recruitment of pioneer and
-heliophilous species, shape post-disturbance trajectories
-\citep{Carreno-Rocabado2012, Poorter2016}. Community trajectories would
-then rely on deterministic processes and would be predictable according
-to the disturbance intensity and to the characteristics of the
-pre-disturbance community. Yet, observed trajectories in diversity and
-composition often deviate from predictions. Besides, although the
-deterministic ecological processes at stake are identified, their
-relative importance compared to stochasticity remains debated
-\citep{Norden2015}. This debate translates into the validation and the
-extent of the intermediate disturbance hypothesis (IDH). The IDH assumes
-the predominance of deterministic processes, and predicts a maximum
-species diversity after moderate disturbance intensity
-\citep{Molino2001, Kariuki2006a}. Validations of the IDH remain scarce
-in the long-term and often rely on the analysis of taxonomic richness
-\citep{Molino2001}. Taxonomic richness alone, however, gives limited or
-misleading information on forests recovery and functioning
-\citep{Chaudhary2016}. More ecological-meaningful analysis would couple
-richness with species evenness, which reveals the changes in the species
-abundance distribution, and (ii)species composition, which is crucial
-for conservation issues \citep{Lavorel2002, Bellwood2006}. Furthermore,
-a functional approach accounting for species biological attributes would
+\citep{Goulamoussene2017}, and modifications in the interactions among
+species \citep{Chesson2000}. Community response has been largely studied
+through forest structural parameters such as aboveground biomass, tree
+height or stem density \citep{Piponiot2016, Rutishauser2016}.
+Thereafter, models based on the observed trajectories after disturbance
+gave important insights into community recovery and resilience
+\citep{Herault2018}. Regarding community diversity and composition,
+however, post-disturbance trajectories are not as thoroughly understood.
+It is recognized that a determined succession, mainly driven by the
+recruitment of pioneer and heliophilous species, shape community
+post-disturbance trajectories \citep{Carreno-Rocabado2012, Poorter2016}.
+Deterministic processes would then be major processes of community
+response to disturbance. This would make post-disturbance trajectories
+deterministic, and then predictable according to the disturbance
+intensity and to the characteristics of the pre-disturbance community.
+Yet, observed trajectories in diversity and composition often deviate
+from predictions. Although the deterministic ecological processes at
+stake are identified, their interplay with stochastic processes and
+their relative importance remains debated \citep{Norden2015}. This
+debate translates into the validation and the extent of the intermediate
+disturbance hypothesis (IDH). The IDH assumes the predominance of
+deterministic processes, and predicts a maximum species diversity after
+moderate disturbance intensity \citep{Molino2001, Kariuki2006a}. The IDH
+has been scarcely validated in the long-term and often rely on the
+analysis of taxonomic richness \citep{Molino2001}. Taxonomic richness
+alone, however, gives limited or misleading information on forests
+recovery and functioning \citep{Chaudhary2016}. More
+ecological-meaningful analysis would couple richness with species
+evenness, which reveals the changes in the species abundance
+distribution, and (ii)species composition, which is crucial for
+conservation issues \citep{Lavorel2002, Bellwood2006}. Furthermore, a
+functional approach accounting for species biological attributes would
 directly link community diversity, composition and redundancy to
@@
 Control & - & - & - & 0\\
-T1 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10   $trees.ha^{-1}$ & - & - & $[12-33]$\\
-T2 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30   $trees.ha^{-1}$ & - & $[33-56]$\\
-T3 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15  $trees.ha^{-1}$ & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species,\ $\approx$ 15 $trees.ha^{-1}$ & $[35-56]$\\
+T1, low & DBH $\geq$ 50 cm, commercial species, $\approx$ 10   $trees.ha^{-1}$ & - & - & $[12-33]$\\
+T2, intermediate & DBH $\geq$ 50 cm, commercial species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30   $trees.ha^{-1}$ & - & $[33-56]$\\
+T3, high & DBH $\geq$ 50 cm, commercial species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15  $trees.ha^{-1}$ & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species,\ $\approx$ 15 $trees.ha^{-1}$ & $[35-56]$\\
 \bottomrule
@@
 
-Community taxonomic and functional composition appeared resilient,
-following similar hump-shaped trajectories starting to return towards
-pre-disturbance composition after 30 years.
+The hump-shaped trajectories of community composition returned towards
+pre-disturbance state after 30 years, highlighting community resilience
+in terms of taxonomic and functional composition.
 
-The taxonomic differences among local communities were already apparent
-before disturbance, as revealed by the distinct starting points on the
-NMDS axis 2. These differences were maintained throughout recovery
-trajectories.\\
-More than commonly thought, post-disturbance trajectories depended on
-community initial composition, which partly determined the pool of
-recruited species and constrained the trajectories towards the initial
-composition. The community taxonomy proved highly resilient, as it
-absorbed the disturbance and maintained commnity initial composition
-characteristics \citep{Folke2006}. This high resilience suggested that
-species not belonging to the pre-disturbance community were rarely
-recruited because of the commonness of dispersal limitation among
+Before disturbance, there were already taxonomic differences among local
+communities, as revealed by the distinct starting points on the NMDS
+axis 2. These initial differences were maintained throughout the
+trajectories following disturbance, so community initial state
+influenced the trajectories following disturbance more than commonly
+thought. The initial composition indeed partly determined the pool of
+species recruited, and hence shaped the composition trajectories and
+drove them towards the recovery of the initial composition. This
+taxonomic recovery, although it was not fully achieved after 30 years,
+suggested the resilience of community taxonomic composition and the
+maintenance of community initial differences \citep{Folke2006}. Species
+not belonging to the pre-disturbance community were then rarely
+recruited, probably because of the common dispersal limitations among
 tropical tree species \citep{Svenning2005}.
 
-Conversely, disturbed communities followed functional trajectories that
-are highly similar in terms of functional composition. As the
-composition of pre-disturbance surviving trees is representative of the
-initial community \citep{Herault2018}, changes in functional composition
-relied upon the recruitment of species or functional types that were
-infrequent or absent before disturbance. Competitive pioneers became
-dominant in filling the environmental niches of high availability of
-light, space and nutrients vacated by the disturbance. The recruitment
-of pioneers changed community functional composition in the same way for
-all disturbance intensity towards more resource-acquisitive strategies,
-moving community functional composition right along the first axis in
-Fig. \ref{fig:NMDSplans} \citep{Westoby1998, Wright2004, Reich2014}.
-Thereafter long-lived, more resistant and shade-tolerant species
-excluded the first established pioneers and started the recovery of
-pre-disturbance functional composition, moving similarly community
-functional composition left along the first axis and upward along the
-second axis in Fig. \ref{fig:NMDSplans}.
+Functional composition trajectories were, conversely, highly similar
+among disturbed communities. As the composition of pre-disturbance
+surviving trees is representative of the initial community
+\citep{Herault2018}, changes in functional composition relied upon the
+recruitment of species or functional types that were infrequent or
+absent before disturbance. Competitive pioneers became dominant in
+filling the environmental niches vacated by the disturbance, with a high
+availability of light, space and nutrients. This recruitment of pioneers
+similarly changed the functional composition of all communities,
+whenever the disturbance intensity. The functional composition went
+towards more resource-acquisitive strategies, translated in the
+functional plan (Fig. \ref{fig:NMDSplans}) by a displacement on the
+right along the first axis \citep{Westoby1998, Wright2004, Reich2014}.
+Thereafter, the pioneers first recruited were progressively excluded by
+long-lived, more resistant and shade-tolerant species. The commpunity
+functional composition then returned towards more resource-conservative
+strategies, suggesting the recovery of the initial community composition
+and translated in the functional plan by a displacement left along the
+first axis and upward along the second axis (Fig. \ref{fig:NMDSplans}).
 
-These trajectories provided empirical support to the hypothesis that
-community assembly is both deterministic and historically convergent at
-different levels of community organization. Deterministic, trait-based
-processes drove community convergence in functional composition, while
-at the same time dispersal limitation maintained their divergence in
-taxonomic composition \citep{Fukami2005}.
+In the light of these taxonomic and functional composition trajectories,
+community response to disturbance proved simultaneously deterministic
+and stochastic at different levels of community organization. Taxonomic
+trajectories were shaped by dispersal limitations maintaining the
+stochastic differences among communities due to historical contingency.
+Functional trajectories were shaped by trait-based processes resulting
+the the convergence of community functional composition
+\citep{Fukami2005}.
 
@@
 
-Community taxonomic richness increased after disturbance until an
-intensity threshold (20-25\% AGB loss) above which the taxonomic
-richness decreased. The increase of taxonomic evenness after disturbance
-also depended on the disturbance intensity, although the correlation was
-weaker than for the taxonomic richness as already observed in the Guiana
-Shield \citep{Baraloto2012a} and in Bornean tropical forests
-\citep{Cannon1998}. In accordance with the IDH, then, community
-taxonomic trajectories depended on the disturbance intensity. They
-markedly changed above an intensity threshold, with a decrease of the
-richness and a persistent increase of the evenness. Disturbance
-intensity determined the balance in the community between surviving
-trees from pre-disturbance communities and trees recruited afterward.
-The pool of true pioneer species specifically recruited after
-disturbance is, in the Guiana Shield, restricted to a few common genera
-(e.g. \emph{Cecropia} spp., \emph{Vismia} spp.)\citep{Guitet2018}. Below
-the intensity threshold the surviving community trees remained numerous
-enough to maintain the pre-disturbance high taxonomic richness while the
-pioneers recruited, infrequent or absent before disturbance, increased
-both community taxonomic richness and evenness. Beyond the intensity
-threshold, disturbance decreased the taxonomic richness of surviving
-trees, and this decrease was not offset by the recruitment of pioneers.
-The overall community taxonomic richness therefore decreased according
-to the disturbance intensity \citep{Molino2001}. An intermediate
-disturbance was then detected, defined by an intensity for which
-post-disturbance trajectories markedly changed. Such disturbance
-increased the availability of resources and created opportunities for
-pioneers, without implying too important loss of shade-tolerant species
-or preventing their maintenance in the community \citep{Bongers2009}.
-For community taxonomic evenness, the disturbance impact was similar but
-milder. Because the evenness is less sensitive to the loss of rare
-species, below the intensity threshold the trajectories of taxonomic
-evenness rather represented the increasing dominance of pioneers that
-balancing the usual hyper-dominance of a few species. Beyond the
-intensity threshold, however, pioneers became in turn highly dominant
-and decreased the overall evenness \citep{Baraloto2012a}.
+Community taxonomic richness and evenness trajectories both proved to
+depend on the intensity of disturbance and to markedly change above an
+intensity threshold, thus supporting the IDH. The disturbance intensity
+determined the balance in the community between pre-disturbance
+surviving trees, and trees recruited afterward. Below an intensity of
+20\%-25\% AGB loss, the trees surviving after disturbance remained
+numerous enough to maintain the high taxonomic richness of the
+pre-disturbance community. The recruitment of pionners, infrequent or
+absent before disturbance, then increased the community taxonomic
+richness positively correlated with the disturbance intensity
+\citep{Martin2015, Chaudhary2016}. Besides, the pioneers became
+increasingly dominant and temporarily balanced the usual hyper-dominance
+of a few species \citep{Baraloto2012a}. The taxonomic evenness then
+followed hump-shaped trajectoriesas already observed in the Guiana
+shield \citep{Baraloto2012a} and in Borenan tropical forests
+\citep{Cannon1998}, with first an increase positively correlated to the
+disturbance intensity, and then an return towards initial values. Above
+the intensity threshold, the taxonomic richness of surviving trees was
+too low to be offset by the recruitment of pioneers. In the Guiana
+shield indeed, the pool of true pioneers specifically recruited after
+disturbance is restricted to a few common genera (e.g. \emph{Cecropia}
+spp., \emph{Vismia} spp.) \citep{Guitet2018}. The taxonomic richness
+trajectories following intense disturbance then decreased, all the more
+so that the disturbance intensity was high \citep{Molino2001}. At the
+same time, pioneers became persistently dominant and prevented a return
+towards the initial evenness and the recovery of hyper-dominant
+shade-tolerant species. The intensity threshold then marked two distinct
+post-disturbance trajectories, and a change in community response to
+disturbance. In accordance with the IDH, an intense disturbance below
+the intensity threshold would then maximise community richness and
+evenness. Such intermediate disturbance would increase the availability
+of resources, creating opportunities for pioneers, while allowing a high
+diversity among surviving trees and the maintenance of shade-tolerant
+species \citep{Bongers2009}.
 
-Conversely the IDH was disproved regarding the disturbance impact on
-community functional richness and evenness. Irrespective of the
-disturbance intensity, both community functional richness and evenness
-increased following the recruitment of pioneers, functionally highly
-different from the composition of pre-disturbance community.
+Regarding community functional response in contrast, the IDH seemed
+disproved. Whenever the disturbance intensity, both community functional
+richness and evenness increased, with the recruitment of pioneers that
+were functionally highly different from the species of the
+pre-disturbance community \citep{Denslow1980, Molino2001}. Above the
+intensity threshold however, for the most intense disturbance, community
+functional richness and evenness started to decrease after 15 to 20
+years. The vacated environmental space was then occupied by short-lived
+species which prevented the establishment of other species. The decline
+of these short-lived pioneers decreased the functional richness and
+evenness, but we suggest that the establishment of long-lasting pioneers
+will follow and the taxonomic and functional trajectories will catch up
+with those observed for intermediate disturbance \citep{Walker2009}.
 
-Along time, taxonomic richness trajectories of all disturbed communities
-similarly dropped at first place, following the species loss due to
-disturbance, and then displayed a species gain depending on the
-disturbance intensity. Up to an intensity threshold, the species gain
-was all the more significant that the disturbance intensity increased,
-with the establishment of long-lived pioneers enhancing community
-taxonomic richness and evenness in the long term. These long-lived
-pioneers, functionally quite different from the functional composition,
-entailed as well a progressive and long-lasting increase of the
-functional richness and evenness \citep{Denslow1980, Molino2001}. Beyond
-an intensity threshold, though, a few short-lived pioneers occupied the
-vacated environmental space and prevented the establishment of other
-species. These short-lived pioneers were functionally very different
-from the pre-disturbance community and entailed a rapid and significant
-increase of functional richness and evenness. Already after 10 years,
-though, short-lived pioneers started to decline and the functional
-richness and evenness decreased. Likely this decrease will be followed
-by the establishment of long-lasting pioneers, and by the time they
-recruit we expect the taxonomic and functional trajectories to catch up
-with those observed after intermediate disturbance \citep{Walker2009}.
+The IDH then translated into community post-disturbance taxonomic
+trajectories, with marked differences below and above an intensity
+threshold. It was however not tangible regarding the functional
+trajectories, because of their decoupling with taxonomic trajectories.
 
@@
 
-For 15 years the species loss during disturbance, determined by the
-disturbance intensity, significantly decreased the functional redundancy
-within the pre-disturbance functional space. The redundancy decrease was
-not compensated for in the first place as the first recruited pioneers
-were functionally different from the pre-disturbance functional
-composition. Progressively though, first established species were
-replaced by more competitive long-lived pioneers or late-successional
-species resembling more the pre-disturbance functional composition and
-restoring the functional redundancy. This replacement was stochastic and
-followed the lottery recruitment rules, implying a recruitment eased for
-the first recruited species but then increasingly hampered by the
-emergence of interspecific competition \citep{Busing2002}. Along time
-the recovery of infrequent species was increasingly slow, so that the
-time to full recovery of the functional redundancy, in some communities
-just initiated after 30 years, was extremely difficult to estimate
+The decoupling between taxonomic and functional trajectories was
+explained by a decrease in the functional redundancy within the
+pre-disturbance functional space, due to the loss of species following
+disturbance. The recruitment of pioneers, functionally different from
+the pre-disturbance functional composition, did not compensate the
+decrease of functional redundancy in the first place. Progressively
+though, the functional redundancy was restored through the replacement
+of the first established species by more competitive, long-lived
+pioneers or late-successional species that were functionally closer to
+the pre-disturbance community. This replacement followed the lottery
+recruitment rules, implying a recruitment easy for the first recruited
+species but increasingly hampered by the emergence of interspecific
+competition \citep{Busing2002}. The functional redundancy then relied
+upon stochastic processes and its recovery, depending on the recruitment
+of infrequent species constrained by the increasing competition, was
+increasingly slow and difficult to anticipate
 \citep{Elmqvist2003, Diaz2005}.
 
-The mid-term impact of disturbance on community functional redundancy
-meant a lower resilience of the pre-disturbance communities, with higher
-chances to see the persistence of disturbance-specific species at the
-expense of late-successional ones \citep{Haddad2008}. Besides, the
-mid-term recovery of infrequent species increases the risks to loose
-keystone species, with unexpected ecological consequences
+The impact of disturbance on community functional redundancy meant a
+lower resilience of the pre-disturbance communities, with higher chances
+to see the persistence of disturbance-specific species at the expense of
+late-successional ones \citep{Haddad2008}. Besides, the stochastic
+recovery of infrequent species increases the risks to lose keystone
+species, with unexpected ecological consequences
 \citep{Jones1994, Chazdon2003a, Diaz2005}. Apart from the functional
@@
 
-Our study revealed community recovery through the combination of
-deterministic processes driving their convergence in functional
-composition, and dispersal limitation maintaining their divergence in
-taxonomic composition. The IDH was validated for community taxonomic
-richness and, to some extent, taxonomic evenness but disproved regarding
-community functional richness and evenness which were enhanced for any
-disturbance intensity by the high functional differences of pioneers
-compared to late-successional functional composition. The IDH was
-translated in time by the recruitment, beyond an intensity threshold, of
-short-lived pioneers that prevented in the first years after disturbance
-the establishment of more diverse long-lived pioneers, recruited
-otherwise below the intensity threshold. The resilience of tropical
-forests, defined in terms of recovery to pre-disturbance state, proved
-tangible but requiring several decades. Still, the disturbance impact on
-community redundancy cautioned against the risks of infrequent species
-loss and the persistence of disturbance-specific communities
-\citep{Herault2018}.
+Our study highlighted the combination of deterministic and stochastic
+processes determining the succession of recruited species and shaping
+community response to disturbance. The deterministic recruitment of
+pioneers shaped functional trajectories convergent in the functional
+space, while dispersal limitations shaped stochastic taxonomic
+trajectories maintaining community divergence in taxonomic composition.
+In accordance with the IDH, community taxonomic diversity was maximise
+for intermediate disturbance intensity, before reaching an intensity
+threshold beyond which the taxonomic richness decreased and the
+taxonomic evenness remained persistently higher. Conversely, functional
+trajectories proved decoupled from the taxonomic ones and did not follow
+the hypothesis of the IDH as no distinct trajectories emerged with
+increasing disturbance intensity. The decoupling between taxonomic and
+functional disturbance was explained by a decrease of community
+functional redundancy mitigating the functoinal impact of disturbance.
+The resilience of tropical forests, defined in terms of recovery to
+pre-disturbance state, proved tangible but requiring several decades.
+Still, the disturbance impact on community redundancy cautioned against
+the risks of infrequent species loss and the persistence of
+disturbance-specific communities \citep{Herault2018}.
 

### 2019-01-28 WholePlotTrajectories.Rmd
intro

@@
  The huge diversity of tropical forests is often assumed to be regularly reshaped by natural disturbance yielding a diversity peak at intermediate intensity.
-This intermediate disturbance hypothesis (IDH), though, remains debated and the controversy questions the extent of community resilience regarding their taxonomic and functional facets.
- To disentangle the ecological processes driving community response to disturbance, we analyzed the tree community trajectories over 30 years following a disturbance gradient in a Neotropical forest.
-Specifically, we examined community functional and taxonomic trajectories with regards to diversity, composition and redundancy.
-Functional trajectories were drawn based on 7 leaf, stem and life-history traits.
+This intermediate disturbance hypothesis (IDH), though, remains debated and the resilience of community taxonomic and functional diversity and composition remains controversial.
+ To disentangle the ecological processes driving community response to disturbance, we analyzed tree community trajectories following a disturbance gradient in a Neotropical forest.
+We examined community diversity, composition and redundancy trajectories over 30 years, considering both taxonomic and functional trajectories based on 7 leaf, stem and life-history traits.
 We highlighted the cyclic recovery of community taxonomic and functional composition.
@@
 
-Tropical forests cover large areas worldwide and represent crucial environmental, economic and social values. 
-They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, determine the cycles of carbon, water and nutrient, and ensure cultural and human well-being.
-However, the growing demand in forests products together with current global changes increases the pressure on remaining natural forests [@Morales-Hidalgo2015].
+The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
+They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, the carbon, water and nutrient cycles, and ensure cultural and human well-being.
+The growing demand in forests products together with current global changes, however, increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
 This context specifically threatens the maintenance of community structure, composition and functioning [@Anderson-Teixeira2013].
-In tropical forests indeed, tree communities are shaped and maintained by a natural disturbance regime, specifically tree fall gaps[@Schnitzer2001].
-This disturbance regime is affected by the current changing context and anthropic activities [@Sist2015], so clarify community responses to disturbance and the underlying ecological processes is crucial to anticipate the fate of tropical forests [@Chazdon2003a].
+In tropical forests indeed, tree communities biodiversity and composition rely on a natural disturbance regime, specifically through tree fall gaps [@Schnitzer2001].
+This disturbance regime is affected by the current changing context and anthropic activities [@Sist2015], so clarify community response to disturbance and the underlying ecological processes is crucial to anticipate the fate of tropical forests [@Chazdon2003].
 
-Disturbances change both biotic and abiotic environment, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and modifications in the interactions among species [@Chesson2000].
-Community response has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
+Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and modifications in species interactions, mainly through the reduction of species density [@Chesson2000].
+For now, community response to disturbance has been largely studied through forest structural parameters, rapid and convenient to measure, such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016].
 Thereafter, models based on the observed trajectories after disturbance gave important insights into community recovery and resilience [@Herault2018]. 
-Regarding community diversity and composition, however, post-disturbance trajectories are not as thoroughly understood.
-It is recognized that a determined succession, mainly driven by the recruitment of pioneer and heliophilous species, shape community post-disturbance trajectories [@Carreno-Rocabado2012;@Poorter2016].
-Deterministic processes would then be major processes of community response to disturbance.
-This would make post-disturbance trajectories deterministic, and then predictable according to the disturbance intensity and to the characteristics of the pre-disturbance community.
-Yet, observed trajectories in diversity and composition often deviate from predictions.
-Although the deterministic ecological processes at stake are identified, their interplay with stochastic processes and their relative importance remains debated [@Norden2015].
-This debate translates into the validation and the extent of the intermediate disturbance hypothesis (IDH).
-The IDH assumes the predominance of deterministic processes, and predicts a maximum species diversity after moderate disturbance intensity [@Molino2001;@Kariuki2006a].
-The IDH has been scarcely validated in the long-term and often rely on the analysis of taxonomic richness [@Molino2001]. 
-Taxonomic richness alone, however, gives limited or misleading information on forests recovery and functioning [@Chaudhary2016]. 
-More ecological-meaningful analysis would couple richness with species evenness, which reveals the changes in the species abundance distribution, and (ii)species composition, which is crucial for conservation issues [@Lavorel2002;@Bellwood2006].
-Furthermore, a functional approach accounting for species biological attributes would directly link community diversity, composition and redundancy to ecosystem functioning and to its environmental constraints [@Violle2007b;@Baraloto2012a].
-In that respect, the functional trait-based approach that focus on major traits related to species ecology and mediate species performance in a given environment was successfully adopted [@Diaz2005].
-For instance, the functional approach revealed in tropical rainforests the deterministic processes entailing, after disturbance, a functional shift from a dominance of "conservative" slow-growing species dealing with scarce resources to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Reich2014;@Herault2011].
-This shift is translated into the trajectories of key functional traits related to resource acquisition (leaf and stem traits) and life-history traits (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
-Eventually, a complete overview of community response to disturbance would encompass the changes in functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
-The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines community resilience after disturbance [@Elmqvist2003; @Diaz2005].
+Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood.
+A leading idea on how disturbance affect community diversity is the Intermediate Disturbance Hypothesis (IDH) [@Connell1978] predicting biodiversity peaks at intermediate value of disturbance intensity and at intermediate time after disturbance.
+The IDH is founded on the emergence after disturbance of deterministic processes that prevent competitive exclusion through the fluctuations of dominant competitors along time due to disturbance environmental impacts [@Fox2013].
+The IDH assumes predictable post-disturbance trajectories depending on the intensity of disturbance, that define the importance of disturbance impacts [@Willig2018].
+It also assumes, without additional disturbance, a recovery of the pre-disturbance state defined by the dominance of late-successional species [@Rees2001;@Carreno-Rocabado2012;@Poorter2016].
+In highly diverse tropical forests, however, empirical validation of the IDH remains debated [@Fox2013;@Sheil2013] and observed post-disturbance trajectories are often blurry compared to theoretical expectations [@Hugues2007].
 
-Improve the understanding of diversity and composition response to disturbance would then require to combine taxonomic and functional diversity and composition metrics that is not often provided [@Carreno-Rocabado2012,@Mouillot2013].
-Long term monitoring provided trajectories over several decades would allow to put in perspectives the existing studies that often correspond to chronosequence studies [@Chazdon2007] or cover 10 or 15 years after disturbance [@Carreno-Rocabado2012].
+Still, empirical evidence often relied on the analysis of a single diversity metric, often the taxonomic richness [@Molino2001], and this might give limited or misleading information on forests recovery and functioning.
+More ecological-meaningful analysis would first couple taxonomic richness and evenness, which reveals the changes in species abundance distribution and hence in the underlying ecological processes [@Chaudhary2016].
+Analysis would as well include taxonomic composition, which is crucial for conservation issues [@Lavorel2002;@Bellwood2006]. 
+Eventually, including a functional approach accounting for species biological attributes would be highly revealing of ecological process involved in community response to disturbance.
+The functional trait-based approach indeed directly links community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
+In that respect, a vast literature allowed recognizing major traits related to species ecology and mediating species performance in a given environment [@Diaz2005]. 
+Specifically in tropical forests, the functional approach revealed the post-disturbance deterministic processes entailing a shift from a dominance of “conservative” slow-growing species dealing with scarce resources, to a dominance of “acquisitive” fast-growing species with rapid and efficient use of abundant resources [@Reich2014;@Herault2011].
+This shift is translated into the trajectories of average community value of key functional traits related to resource acquisition (leaf and stem traits) and life-history strategy (seed mass, maximum size)  [@Wright2004;@Diaz2005;@TerSteege2006;@Westoby2006a;@Chave2009b].
+The functional approach would eventually be completed with the analysis of functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
+The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Elmqvist2003; @Diaz2005].
 
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-From post-disturbance trajectories (i) we determined the taxonomic and functional translation post-disturbance succession at community scale.
-(ii) we validated the importance of deterministic processes in community response to disturbance.
-We defined th interplay of deterministic and stochastic processes along time, which allowed discussing the IDH.
-(iii) examined community recovery time.
+Specifically, we (i) elucidated community taxonomic and functional recovery and the underlying ecological processes, (ii) clarified the validity of the IDH in the long term for tropical forest and its translation into different trajectories in time, and (iii) examined community recovery time.
+
+@Rees2001].
+Yet, observed trajectories proved to often deviate from predictions and from the recovery of initial state and thus evidenced the interplay of stochastic processes [@Hubbel2001;@Norden2015; @Norden2017].
+The debate underlying the study of community response to disturbance is now on the relative importance of stochastic and deterministic processes, and on the nature of ecological rules and mechanisms they generate.
+
+Specifically, the determinism of post-disturbance trajectories is the base of the intermediate disturbance hypothesis (IDH) which is assumed to be a determinant of tropical forest biodiversity [@Connell1978].
 

### 2019-01-28 WholePlotTrajectories.tex
intro

@@
 intensity. This intermediate disturbance hypothesis (IDH), though,
-remains debated and the controversy questions the extent of community
-resilience regarding their taxonomic and functional facets. To
-disentangle the ecological processes driving community response to
-disturbance, we analyzed the tree community trajectories over 30 years
-following a disturbance gradient in a Neotropical forest. Specifically,
-we examined community functional and taxonomic trajectories with regards
-to diversity, composition and redundancy. Functional trajectories were
-drawn based on 7 leaf, stem and life-history traits. We highlighted the
-cyclic recovery of community taxonomic and functional composition. While
-pre-disturbance taxonomic differences were maintained over time,
-functional composition trajectories were quite similar among
+remains debated and the resilience of community taxonomic and functional
+diversity and composition remains controversial. To disentangle the
+ecological processes driving community response to disturbance, we
+analyzed tree community trajectories following a disturbance gradient in
+a Neotropical forest. We examined community diversity, composition and
+redundancy trajectories over 30 years, considering both taxonomic and
+functional trajectories based on 7 leaf, stem and life-history traits.
+We highlighted the cyclic recovery of community taxonomic and functional
+composition. While pre-disturbance taxonomic differences were maintained
+over time, functional composition trajectories were quite similar among
 communities. The IDH did predict community taxonomic diversity response
@@
 
-Tropical forests cover large areas worldwide and represent crucial
+The large areas covered with tropical forests worldwide hold crucial
 environmental, economic and social values. They provide wood and
 multiple non-timber forest products, shelter a diversified fauna,
-regulate the local and regional climates, determine the cycles of
-carbon, water and nutrient, and ensure cultural and human well-being.
-However, the growing demand in forests products together with current
-global changes increases the pressure on remaining natural forests
-\citep{Morales-Hidalgo2015}. This context specifically threatens the
-maintenance of community structure, composition and functioning
-\citep{Anderson-Teixeira2013}. In tropical forests indeed, tree
-communities are shaped and maintained by a natural disturbance regime,
-specifically tree fall gaps\citep{Schnitzer2001}. This disturbance
-regime is affected by the current changing context and anthropic
-activities \citep{Sist2015}, so clarify community responses to
-disturbance and the underlying ecological processes is crucial to
-anticipate the fate of tropical forests \citep{Chazdon2003a}.
+regulate the local and regional climates, the carbon, water and nutrient
+cycles, and ensure cultural and human well-being. The growing demand in
+forests products together with current global changes, however, increase
+the pressure on remaining natural forests \citep{Morales-Hidalgo2015}.
+This context specifically threatens the maintenance of community
+structure, composition and functioning \citep{Anderson-Teixeira2013}. In
+tropical forests indeed, tree communities biodiversity and composition
+rely on a natural disturbance regime, specifically through tree fall
+gaps \citep{Schnitzer2001}. This disturbance regime is affected by the
+current changing context and anthropic activities \citep{Sist2015}, so
+clarify community response to disturbance and the underlying ecological
+processes is crucial to anticipate the fate of tropical forests
+\citep{Chazdon2003}.
 
-Disturbances change both biotic and abiotic environment, through
-modifications in the fluxes of light, heat and water
-\citep{Goulamoussene2017}, and modifications in the interactions among
-species \citep{Chesson2000}. Community response has been largely studied
-through forest structural parameters such as aboveground biomass, tree
-height or stem density \citep{Piponiot2016, Rutishauser2016}.
-Thereafter, models based on the observed trajectories after disturbance
-gave important insights into community recovery and resilience
-\citep{Herault2018}. Regarding community diversity and composition,
-however, post-disturbance trajectories are not as thoroughly understood.
-It is recognized that a determined succession, mainly driven by the
-recruitment of pioneer and heliophilous species, shape community
-post-disturbance trajectories \citep{Carreno-Rocabado2012, Poorter2016}.
-Deterministic processes would then be major processes of community
-response to disturbance. This would make post-disturbance trajectories
-deterministic, and then predictable according to the disturbance
-intensity and to the characteristics of the pre-disturbance community.
-Yet, observed trajectories in diversity and composition often deviate
-from predictions. Although the deterministic ecological processes at
-stake are identified, their interplay with stochastic processes and
-their relative importance remains debated \citep{Norden2015}. This
-debate translates into the validation and the extent of the intermediate
-disturbance hypothesis (IDH). The IDH assumes the predominance of
-deterministic processes, and predicts a maximum species diversity after
-moderate disturbance intensity \citep{Molino2001, Kariuki2006a}. The IDH
-has been scarcely validated in the long-term and often rely on the
-analysis of taxonomic richness \citep{Molino2001}. Taxonomic richness
-alone, however, gives limited or misleading information on forests
-recovery and functioning \citep{Chaudhary2016}. More
-ecological-meaningful analysis would couple richness with species
-evenness, which reveals the changes in the species abundance
-distribution, and (ii)species composition, which is crucial for
-conservation issues \citep{Lavorel2002, Bellwood2006}. Furthermore, a
-functional approach accounting for species biological attributes would
-directly link community diversity, composition and redundancy to
-ecosystem functioning and to its environmental constraints
-\citep{Violle2007b, Baraloto2012a}. In that respect, the functional
-trait-based approach that focus on major traits related to species
-ecology and mediate species performance in a given environment was
-successfully adopted \citep{Diaz2005}. For instance, the functional
-approach revealed in tropical rainforests the deterministic processes
-entailing, after disturbance, a functional shift from a dominance of
-``conservative'' slow-growing species dealing with scarce resources to
+Disturbance change both community abiotic and biotic environments,
+through modifications in the fluxes of light, heat and water
+\citep{Goulamoussene2017}, and modifications in species interactions,
+mainly through the reduction of species density \citep{Chesson2000}. For
+now, community response to disturbance has been largely studied through
+forest structural parameters, rapid and convenient to measure, such as
+aboveground biomass, tree height or stem density
+\citep{Piponiot2016, Rutishauser2016}. Thereafter, models based on the
+observed trajectories after disturbance gave important insights into
+community recovery and resilience \citep{Herault2018}. Regarding
+community diversity and composition, however, post-disturbance
+trajectories have not been as thoroughly understood. A leading idea on
+how disturbance affect community diversity is the Intermediate
+Disturbance Hypothesis (IDH) \citep{Connell1978} predicting biodiversity
+peaks at intermediate value of disturbance intensity and at intermediate
+time after disturbance. The IDH is founded on the emergence after
+disturbance of deterministic processes that prevent competitive
+exclusion through the fluctuations of dominant competitors along time
+due to disturbance environmental impacts \citep{Fox2013}. The IDH
+assumes predictable post-disturbance trajectories depending on the
+intensity of disturbance, that define the importance of disturbance
+impacts \citep{Willig2018}. It also assumes, without additional
+disturbance, a recovery of the pre-disturbance state defined by the
+dominance of late-successional species
+\citep{Rees2001, Carreno-Rocabado2012, Poorter2016}. In highly diverse
+tropical forests, however, empirical validation of the IDH remains
+debated \citep{Fox2013, Sheil2013} and observed post-disturbance
+trajectories are often blurry compared to theoretical expectations
+\citep{Hugues2007}.
+
+Still, empirical evidence often relied on the analysis of a single
+diversity metric, often the taxonomic richness \citep{Molino2001}, and
+this might give limited or misleading information on forests recovery
+and functioning. More ecological-meaningful analysis would first couple
+taxonomic richness and evenness, which reveals the changes in species
+abundance distribution and hence in the underlying ecological processes
+\citep{Chaudhary2016}. Analysis would as well include taxonomic
+composition, which is crucial for conservation issues
+\citep{Lavorel2002, Bellwood2006}. Eventually, including a functional
+approach accounting for species biological attributes would be highly
+revealing of ecological process involved in community response to
+disturbance. The functional trait-based approach indeed directly links
+community diversity, composition and redundancy to ecosystem functioning
+and environmental constraints \citep{Violle2007b, Baraloto2012a}. In
+that respect, a vast literature allowed recognizing major traits related
+to species ecology and mediating species performance in a given
+environment \citep{Diaz2005}. Specifically in tropical forests, the
+functional approach revealed the post-disturbance deterministic
+processes entailing a shift from a dominance of ``conservative''
+slow-growing species dealing with scarce resources, to a dominance of
 ``acquisitive'' fast-growing species with rapid and efficient use of
 abundant resources \citep{Reich2014, Herault2011}. This shift is
-translated into the trajectories of key functional traits related to
-resource acquisition (leaf and stem traits) and life-history traits
-(seed mass, maximum size)
-\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Eventually,
-a complete overview of community response to disturbance would encompass
-the changes in functional redundancy, that quantifies the amount of
-shared trait values among species \citep{Carmona2016}. The high
-functional redundancy of hyper-diverse tropical forests
-\citep{Bellwood2006} mitigates the impacts of species removal on
-ecosystem functioning and determines community resilience after
-disturbance \citep{Elmqvist2003, Diaz2005}.
-
-Improve the understanding of diversity and composition response to
-disturbance would then require to combine taxonomic and functional
-diversity and composition metrics that is not often provided
-\citep[\citet{Mouillot2013}]{Carreno-Rocabado2012}. Long term monitoring
-provided trajectories over several decades would allow to put in
-perspectives the existing studies that often correspond to
-chronosequence studies \citep{Chazdon2007} or cover 10 or 15 years after
-disturbance \citep{Carreno-Rocabado2012}.
+translated into the trajectories of average community value of key
+functional traits related to resource acquisition (leaf and stem traits)
+and life-history strategy (seed mass, maximum size)
+\citep{Wright2004, Diaz2005, TerSteege2006, Westoby2006a, Chave2009b}.
+The functional approach would eventually be completed with the analysis
+of functional redundancy that quantifies the amount of shared trait
+values among species \citep{Carmona2016}. The high functional redundancy
+of hyper-diverse tropical forests \citep{Bellwood2006} mitigates the
+impacts of species removal on ecosystem functioning and determines the
+resilience of communities after disturbance
+\citep{Elmqvist2003, Diaz2005}.
 
@@
 trajectories in terms of richness, evenness, composition and redundancy.
-From post-disturbance trajectories (i) we determined the taxonomic and
-functional translation post-disturbance succession at community scale.
-(ii) we validated the importance of deterministic processes in community
-response to disturbance. We defined th interplay of deterministic and
-stochastic processes along time, which allowed discussing the IDH. (iii)
-examined community recovery time.
+Specifically, we (i) elucidated community taxonomic and functional
+recovery and the underlying ecological processes, (ii) clarified the
+validity of the IDH in the long term for tropical forest and its
+translation into different trajectories in time, and (iii) examined
+community recovery time.
+
+\citet{Rees2001}{]}. Yet, observed trajectories proved to often deviate
+from predictions and from the recovery of initial state and thus
+evidenced the interplay of stochastic processes
+\citep{Hubbel2001, Norden2015, Norden2017}. The debate underlying the
+study of community response to disturbance is now on the relative
+importance of stochastic and deterministic processes, and on the nature
+of ecological rules and mechanisms they generate.
+
+Specifically, the determinism of post-disturbance trajectories is the
+base of the intermediate disturbance hypothesis (IDH) which is assumed
+to be a determinant of tropical forest biodiversity \citep{Connell1978}.
 

### 2019-01-29 WholePlotTrajectories.Rmd
intro

@@
 The growing demand in forests products together with current global changes, however, increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
-This context specifically threatens the maintenance of community structure, composition and functioning [@Anderson-Teixeira2013].
-In tropical forests indeed, tree communities biodiversity and composition rely on a natural disturbance regime, specifically through tree fall gaps [@Schnitzer2001].
-This disturbance regime is affected by the current changing context and anthropic activities [@Sist2015], so clarify community response to disturbance and the underlying ecological processes is crucial to anticipate the fate of tropical forests [@Chazdon2003].
+This changing context alters the natural disturbance regime that shape and maintain tree community structure, composition and functioning [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015], so anticipate the fate of tropical forests requires to understand community response to disturbance.
+Disturbance initiate ecological succession comprising a suit of changes in community structure and functioning over time, which rate, endpoint an duration depend on the disturbance intensity [@Chazdon2003;@Willig2018].
+Regarding tropical forests, community response to disturbance would be secondary succession, starting with an intact soil and a consistent seed bank, from which we aim would be to highlight predictable trajectories and their underlying ecological processes.
 
-Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and modifications in species interactions, mainly through the reduction of species density [@Chesson2000].
-For now, community response to disturbance has been largely studied through forest structural parameters, rapid and convenient to measure, such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016].
+Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and modifications in species interactions, mainly through modifications of species density [@Chesson2000].
+For now, post-disturbance response has been largely studied through forest structural parameters, rapid and convenient to measure, such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016].
 Thereafter, models based on the observed trajectories after disturbance gave important insights into community recovery and resilience [@Herault2018]. 
 Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood.
-A leading idea on how disturbance affect community diversity is the Intermediate Disturbance Hypothesis (IDH) [@Connell1978] predicting biodiversity peaks at intermediate value of disturbance intensity and at intermediate time after disturbance.
-The IDH is founded on the emergence after disturbance of deterministic processes that prevent competitive exclusion through the fluctuations of dominant competitors along time due to disturbance environmental impacts [@Fox2013].
-The IDH assumes predictable post-disturbance trajectories depending on the intensity of disturbance, that define the importance of disturbance impacts [@Willig2018].
-It also assumes, without additional disturbance, a recovery of the pre-disturbance state defined by the dominance of late-successional species [@Rees2001;@Carreno-Rocabado2012;@Poorter2016].
+A leading idea on how the impacts of disturbance on community diversity is the Intermediate Disturbance Hypothesis (IDH) [@Connell1978].
+The IDH assumes biodiversity changes in relation to disturbance intensity and frequency, with diversity peaks for intermediate intensity and at intermediate time after disturbance. 
+At intermediate level of disturbance, the fluctuation of environmental conditions entails deterministic processes favoring both competitively superior species and fast colonizers, hence preventing competitive exclusion [@Shea2004;@Pulsford2016].
+
+The IDH then assumes both predictable trajectories depending on the disturbance intensity, and, without additional disturbance, the recovery of pre-disturbance state defined by the dominance of late-successional species [@Rees2001;@Sheil2003;@Willig2018].
 In highly diverse tropical forests, however, empirical validation of the IDH remains debated [@Fox2013;@Sheil2013] and observed post-disturbance trajectories are often blurry compared to theoretical expectations [@Hugues2007].

### 2019-01-29 WholePlotTrajectories.tex
intro

@@
 the pressure on remaining natural forests \citep{Morales-Hidalgo2015}.
-This context specifically threatens the maintenance of community
-structure, composition and functioning \citep{Anderson-Teixeira2013}. In
-tropical forests indeed, tree communities biodiversity and composition
-rely on a natural disturbance regime, specifically through tree fall
-gaps \citep{Schnitzer2001}. This disturbance regime is affected by the
-current changing context and anthropic activities \citep{Sist2015}, so
-clarify community response to disturbance and the underlying ecological
-processes is crucial to anticipate the fate of tropical forests
-\citep{Chazdon2003}.
+This changing context alters the natural disturbance regime that shape
+and maintain tree community structure, composition and functioning
+\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}, so anticipate
+the fate of tropical forests requires to understand community response
+to disturbance. Disturbance initiate ecological succession comprising a
+suit of changes in community structure and functioning over time, which
+rate, endpoint an duration depend on the disturbance intensity
+\citep{Chazdon2003, Willig2018}. Regarding tropical forests, community
+response to disturbance would be secondary succession, starting with an
+intact soil and a consistent seed bank, from which we aim would be to
+highlight predictable trajectories and their underlying ecological
+processes.
 
@@
 \citep{Goulamoussene2017}, and modifications in species interactions,
-mainly through the reduction of species density \citep{Chesson2000}. For
-now, community response to disturbance has been largely studied through
-forest structural parameters, rapid and convenient to measure, such as
+mainly through modifications of species density \citep{Chesson2000}. For
+now, post-disturbance response has been largely studied through forest
+structural parameters, rapid and convenient to measure, such as
 aboveground biomass, tree height or stem density
@@
 trajectories have not been as thoroughly understood. A leading idea on
-how disturbance affect community diversity is the Intermediate
-Disturbance Hypothesis (IDH) \citep{Connell1978} predicting biodiversity
-peaks at intermediate value of disturbance intensity and at intermediate
-time after disturbance. The IDH is founded on the emergence after
-disturbance of deterministic processes that prevent competitive
-exclusion through the fluctuations of dominant competitors along time
-due to disturbance environmental impacts \citep{Fox2013}. The IDH
-assumes predictable post-disturbance trajectories depending on the
-intensity of disturbance, that define the importance of disturbance
-impacts \citep{Willig2018}. It also assumes, without additional
-disturbance, a recovery of the pre-disturbance state defined by the
-dominance of late-successional species
-\citep{Rees2001, Carreno-Rocabado2012, Poorter2016}. In highly diverse
+how the impacts of disturbance on community diversity is the
+Intermediate Disturbance Hypothesis (IDH) \citep{Connell1978}. The IDH
+assumes biodiversity changes in relation to disturbance intensity and
+frequency, with diversity peaks for intermediate intensity and at
+intermediate time after disturbance. At intermediate level of
+disturbance, the fluctuation of environmental conditions entails
+deterministic processes favoring both competitively superior species and
+fast colonizers, hence preventing competitive exclusion
+\citep{Shea2004, Pulsford2016}.
+
+The IDH then assumes both predictable trajectories depending on the
+disturbance intensity, and, without additional disturbance, the recovery
+of pre-disturbance state defined by the dominance of late-successional
+species \citep{Rees2001, Sheil2003, Willig2018}. In highly diverse
 tropical forests, however, empirical validation of the IDH remains

### 2019-01-30 WholePlotTrajectories.Rmd
Discussion IDH

@@
 The IDH then assumes both predictable trajectories depending on the disturbance intensity, and, without additional disturbance, the recovery of pre-disturbance state defined by the dominance of late-successional species [@Rees2001;@Sheil2003;@Willig2018].
-In highly diverse tropical forests, however, empirical validation of the IDH remains debated [@Fox2013;@Sheil2013] and observed post-disturbance trajectories are often blurry compared to theoretical expectations [@Hugues2007].
-
-Still, empirical evidence often relied on the analysis of a single diversity metric, often the taxonomic richness [@Molino2001], and this might give limited or misleading information on forests recovery and functioning.
+In highly diverse tropical forests, however, the validation of the IDH remains controversial [@Fox2013;@Sheil2013] and observed post-disturbance trajectories are often blurry compared to theoretical expectations [@Hugues2007].
+Empirical evidence generally rely on the analysis of a single diversity metric, often the taxonomic richness [@Molino2001], which might give limited or misleading information on community disturbance response.
+The IDH pattern of diversity peak under an intermediate disturbance regime can indeed result from various ecological processes [@Sheil2003].
+To disentangle these various processes and grasp the different facets of community response to disturbance it is crucial to combine a suitable range of approaches [@Shea2004].
 More ecological-meaningful analysis would first couple taxonomic richness and evenness, which reveals the changes in species abundance distribution and hence in the underlying ecological processes [@Chaudhary2016].
@@
 
-Community taxonomic richness and evenness trajectories both proved to depend on the intensity of disturbance and to markedly change above an intensity threshold, thus supporting the IDH.
+Community taxonomic richness and evenness trajectories were determined by the disturbance intensity, ranging from a limited and temporary impact to significant and persistant alterations of community diversity.
+The taxonomic trajectories markedly changed above an intensity threshold for which community response corresponded to a maximized richness and a resilience of the evenness. 
 The disturbance intensity determined the balance in the community between pre-disturbance surviving trees, and trees recruited afterward. 
-Below an intensity of 20%-25% AGB loss, the trees surviving after disturbance remained numerous enough to maintain the high taxonomic richness of the pre-disturbance community.
-The recruitment of pionners, infrequent or absent before disturbance, then increased the community taxonomic richness positively correlated with the disturbance intensity [@Martin2015; @Chaudhary2016].
-Besides, the pioneers became increasingly dominant and temporarily balanced the usual hyper-dominance of a few species [@Baraloto2012a]. 
-The taxonomic evenness then followed hump-shaped trajectoriesas already observed in the Guiana shield [@Baraloto2012a] and in Borenan tropical forests [@Cannon1998], with first an increase positively correlated to the disturbance intensity, and then an return towards initial values.
+Below a 20%-25% AGB loss, the trees surviving after disturbance remained numerous enough to maintain the high taxonomic richness of the pre-disturbance community [@Bongers2009].
+The recruitment of pionners, infrequent or absent before disturbance, then increased the taxonomic richness all the more so that the disturbance was intense [@Martin2015; @Chaudhary2016].
+As these pioneers became more dominant they besides temporarily balanced the usual hyper-dominance of tropical forests, hence temporarily increasing the taxonomic evenness [@Baraloto2012a].
 Above the intensity threshold, the taxonomic richness of surviving trees was too low to be offset by the recruitment of pioneers. 
 In the Guiana shield indeed, the pool of true pioneers specifically recruited after disturbance is restricted to a few common genera (e.g. *Cecropia* spp., *Vismia* spp.) [@Guitet2018].
-The taxonomic richness trajectories following intense disturbance then decreased, all the more so that the disturbance intensity was high [@Molino2001].
+The richness trajectories following intense disturbance then decreased, all the more so that the disturbance intensity was high [@Molino2001]. 
 At the same time, pioneers became persistently dominant and prevented a return towards the initial evenness and the recovery of hyper-dominant shade-tolerant species.
-The intensity threshold then marked two distinct post-disturbance trajectories, and a change in community response to disturbance. 
-In accordance with the IDH, an intense disturbance below the intensity threshold would then maximise community richness and evenness.
-Such intermediate disturbance would increase the availability of resources, creating opportunities for pioneers, while allowing a high diversity among surviving trees and the maintenance of shade-tolerant species [@Bongers2009].
+
+Observed trajectories supported the IDH in highlighting a threshold disturbance intensity, expressed as percentage of AGB removed, marking a change in post-disturbance trajectories.
+Taxonomic trajectories besides highlighted, below the intensity threshold, an intermediate time after when the taxonomic evenness was maximized.
+This time corresponded to the intermediate between the generation time of pionneers, first to dominate the post-disturbance recruitment, and the time for competitive exclusion, leading to the dominance of late-successional species.
+As already observed in the Guiana shield [@Baraloto2012a] and in Borenan tropical forests [@Cannon1998],
+taxonomic evenness followed humped-shaped trajectories with a peak when the recruitment was balanced between pioneers and late-successional species.
 

### 2019-01-30 WholePlotTrajectories.tex
Discussion IDH

@@
 species \citep{Rees2001, Sheil2003, Willig2018}. In highly diverse
-tropical forests, however, empirical validation of the IDH remains
-debated \citep{Fox2013, Sheil2013} and observed post-disturbance
+tropical forests, however, the validation of the IDH remains
+controversial \citep{Fox2013, Sheil2013} and observed post-disturbance
 trajectories are often blurry compared to theoretical expectations
-\citep{Hugues2007}.
-
-Still, empirical evidence often relied on the analysis of a single
-diversity metric, often the taxonomic richness \citep{Molino2001}, and
-this might give limited or misleading information on forests recovery
-and functioning. More ecological-meaningful analysis would first couple
+\citep{Hugues2007}. Empirical evidence generally rely on the analysis of
+a single diversity metric, often the taxonomic richness
+\citep{Molino2001}, which might give limited or misleading information
+on community disturbance response. The IDH pattern of diversity peak
+under an intermediate disturbance regime can indeed result from various
+ecological processes \citep{Sheil2003}. To disentangle these various
+processes and grasp the different facets of community response to
+disturbance it is crucial to combine a suitable range of approaches
+\citep{Shea2004}. More ecological-meaningful analysis would first couple
 taxonomic richness and evenness, which reveals the changes in species
@@
 
-Community taxonomic richness and evenness trajectories both proved to
-depend on the intensity of disturbance and to markedly change above an
-intensity threshold, thus supporting the IDH. The disturbance intensity
-determined the balance in the community between pre-disturbance
-surviving trees, and trees recruited afterward. Below an intensity of
-20\%-25\% AGB loss, the trees surviving after disturbance remained
-numerous enough to maintain the high taxonomic richness of the
-pre-disturbance community. The recruitment of pionners, infrequent or
-absent before disturbance, then increased the community taxonomic
-richness positively correlated with the disturbance intensity
-\citep{Martin2015, Chaudhary2016}. Besides, the pioneers became
-increasingly dominant and temporarily balanced the usual hyper-dominance
-of a few species \citep{Baraloto2012a}. The taxonomic evenness then
-followed hump-shaped trajectoriesas already observed in the Guiana
-shield \citep{Baraloto2012a} and in Borenan tropical forests
-\citep{Cannon1998}, with first an increase positively correlated to the
-disturbance intensity, and then an return towards initial values. Above
-the intensity threshold, the taxonomic richness of surviving trees was
-too low to be offset by the recruitment of pioneers. In the Guiana
-shield indeed, the pool of true pioneers specifically recruited after
+Community taxonomic richness and evenness trajectories were determined
+by the disturbance intensity, ranging from a limited and temporary
+impact to significant and persistant alterations of community diversity.
+The taxonomic trajectories markedly changed above an intensity threshold
+for which community response corresponded to a maximized richness and a
+resilience of the evenness. The disturbance intensity determined the
+balance in the community between pre-disturbance surviving trees, and
+trees recruited afterward. Below a 20\%-25\% AGB loss, the trees
+surviving after disturbance remained numerous enough to maintain the
+high taxonomic richness of the pre-disturbance community
+\citep{Bongers2009}. The recruitment of pionners, infrequent or absent
+before disturbance, then increased the taxonomic richness all the more
+so that the disturbance was intense \citep{Martin2015, Chaudhary2016}.
+As these pioneers became more dominant they besides temporarily balanced
+the usual hyper-dominance of tropical forests, hence temporarily
+increasing the taxonomic evenness \citep{Baraloto2012a}. Above the
+intensity threshold, the taxonomic richness of surviving trees was too
+low to be offset by the recruitment of pioneers. In the Guiana shield
+indeed, the pool of true pioneers specifically recruited after
 disturbance is restricted to a few common genera (e.g. \emph{Cecropia}
-spp., \emph{Vismia} spp.) \citep{Guitet2018}. The taxonomic richness
-trajectories following intense disturbance then decreased, all the more
-so that the disturbance intensity was high \citep{Molino2001}. At the
-same time, pioneers became persistently dominant and prevented a return
-towards the initial evenness and the recovery of hyper-dominant
-shade-tolerant species. The intensity threshold then marked two distinct
-post-disturbance trajectories, and a change in community response to
-disturbance. In accordance with the IDH, an intense disturbance below
-the intensity threshold would then maximise community richness and
-evenness. Such intermediate disturbance would increase the availability
-of resources, creating opportunities for pioneers, while allowing a high
-diversity among surviving trees and the maintenance of shade-tolerant
-species \citep{Bongers2009}.
+spp., \emph{Vismia} spp.) \citep{Guitet2018}. The richness trajectories
+following intense disturbance then decreased, all the more so that the
+disturbance intensity was high \citep{Molino2001}. At the same time,
+pioneers became persistently dominant and prevented a return towards the
+initial evenness and the recovery of hyper-dominant shade-tolerant
+species.
+
+Observed trajectories supported the IDH in highlighting a threshold
+disturbance intensity, expressed as percentage of AGB removed, marking a
+change in post-disturbance trajectories. Taxonomic trajectories besides
+highlighted, below the intensity threshold, an intermediate time after
+when the taxonomic evenness was maximized. This time corresponded to the
+intermediate between the generation time of pionneers, first to dominate
+the post-disturbance recruitment, and the time for competitive
+exclusion, leading to the dominance of late-successional species. As
+already observed in the Guiana shield \citep{Baraloto2012a} and in
+Borenan tropical forests \citep{Cannon1998}, taxonomic evenness followed
+humped-shaped trajectories with a peak when the recruitment was balanced
+between pioneers and late-successional species.
 

### 2019-01-31 WholePlotTrajectories.Rmd
intro

@@
 The IDH then assumes both predictable trajectories depending on the disturbance intensity, and, without additional disturbance, the recovery of pre-disturbance state defined by the dominance of late-successional species [@Rees2001;@Sheil2003;@Willig2018].
-In highly diverse tropical forests, however, the validation of the IDH remains controversial [@Fox2013;@Sheil2013] and observed post-disturbance trajectories are often blurry compared to theoretical expectations [@Hugues2007].
+In highly diverse tropical forests, however, the validation of the IDH remains controversial [@Fox2013;@Sheil2013] and observed post-disturbance trajectories are often blurry compared to theoretical expectations [@Hubbel2001;@Hugues2007;@Norden2015;@Norden2017].
 Empirical evidence generally rely on the analysis of a single diversity metric, often the taxonomic richness [@Molino2001], which might give limited or misleading information on community disturbance response.
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, we (i) elucidated community taxonomic and functional recovery and the underlying ecological processes, (ii) clarified the validity of the IDH in the long term for tropical forest and its translation into different trajectories in time, and (iii) examined community recovery time.
-
-@Rees2001].
-Yet, observed trajectories proved to often deviate from predictions and from the recovery of initial state and thus evidenced the interplay of stochastic processes [@Hubbel2001;@Norden2015; @Norden2017].
-The debate underlying the study of community response to disturbance is now on the relative importance of stochastic and deterministic processes, and on the nature of ecological rules and mechanisms they generate.
-
-Specifically, the determinism of post-disturbance trajectories is the base of the intermediate disturbance hypothesis (IDH) which is assumed to be a determinant of tropical forest biodiversity [@Connell1978].
+Specifically, we elucidateed (i) the community taxonomic and functional recovery and (ii) the ecological processes underlying community post-disturbance trajectories and discussed the validity and the extent of the IDH in the long term for tropical forest, and (iii) examined community recovery time.
 
@@
 The deterministic recruitment of pioneers shaped functional trajectories convergent in the functional space, while dispersal limitations shaped stochastic taxonomic trajectories maintaining community divergence in taxonomic composition.
-In accordance with the IDH, community taxonomic diversity was maximise for intermediate disturbance intensity, before reaching an intensity threshold beyond which the taxonomic richness decreased and the taxonomic evenness remained persistently higher.
+In accordance with the IDH, community taxonomic diversity was maximised for intermediate disturbance intensity, before reaching an intensity threshold beyond which the taxonomic richness decreased and the taxonomic evenness remained persistently higher.
 Conversely, functional trajectories proved decoupled from the taxonomic ones and did not follow the hypothesis of the IDH as no distinct trajectories emerged with increasing disturbance intensity. 

### 2019-01-31 WholePlotTrajectories.tex
intro

@@
 trajectories are often blurry compared to theoretical expectations
-\citep{Hugues2007}. Empirical evidence generally rely on the analysis of
-a single diversity metric, often the taxonomic richness
-\citep{Molino2001}, which might give limited or misleading information
-on community disturbance response. The IDH pattern of diversity peak
-under an intermediate disturbance regime can indeed result from various
-ecological processes \citep{Sheil2003}. To disentangle these various
-processes and grasp the different facets of community response to
-disturbance it is crucial to combine a suitable range of approaches
-\citep{Shea2004}. More ecological-meaningful analysis would first couple
-taxonomic richness and evenness, which reveals the changes in species
-abundance distribution and hence in the underlying ecological processes
-\citep{Chaudhary2016}. Analysis would as well include taxonomic
-composition, which is crucial for conservation issues
-\citep{Lavorel2002, Bellwood2006}. Eventually, including a functional
-approach accounting for species biological attributes would be highly
-revealing of ecological process involved in community response to
-disturbance. The functional trait-based approach indeed directly links
-community diversity, composition and redundancy to ecosystem functioning
-and environmental constraints \citep{Violle2007b, Baraloto2012a}. In
-that respect, a vast literature allowed recognizing major traits related
-to species ecology and mediating species performance in a given
-environment \citep{Diaz2005}. Specifically in tropical forests, the
-functional approach revealed the post-disturbance deterministic
-processes entailing a shift from a dominance of ``conservative''
-slow-growing species dealing with scarce resources, to a dominance of
-``acquisitive'' fast-growing species with rapid and efficient use of
-abundant resources \citep{Reich2014, Herault2011}. This shift is
-translated into the trajectories of average community value of key
-functional traits related to resource acquisition (leaf and stem traits)
-and life-history strategy (seed mass, maximum size)
+\citep{Hubbel2001, Hugues2007, Norden2015, Norden2017}. Empirical
+evidence generally rely on the analysis of a single diversity metric,
+often the taxonomic richness \citep{Molino2001}, which might give
+limited or misleading information on community disturbance response. The
+IDH pattern of diversity peak under an intermediate disturbance regime
+can indeed result from various ecological processes \citep{Sheil2003}.
+To disentangle these various processes and grasp the different facets of
+community response to disturbance it is crucial to combine a suitable
+range of approaches \citep{Shea2004}. More ecological-meaningful
+analysis would first couple taxonomic richness and evenness, which
+reveals the changes in species abundance distribution and hence in the
+underlying ecological processes \citep{Chaudhary2016}. Analysis would as
+well include taxonomic composition, which is crucial for conservation
+issues \citep{Lavorel2002, Bellwood2006}. Eventually, including a
+functional approach accounting for species biological attributes would
+be highly revealing of ecological process involved in community response
+to disturbance. The functional trait-based approach indeed directly
+links community diversity, composition and redundancy to ecosystem
+functioning and environmental constraints
+\citep{Violle2007b, Baraloto2012a}. In that respect, a vast literature
+allowed recognizing major traits related to species ecology and
+mediating species performance in a given environment \citep{Diaz2005}.
+Specifically in tropical forests, the functional approach revealed the
+post-disturbance deterministic processes entailing a shift from a
+dominance of ``conservative'' slow-growing species dealing with scarce
+resources, to a dominance of ``acquisitive'' fast-growing species with
+rapid and efficient use of abundant resources
+\citep{Reich2014, Herault2011}. This shift is translated into the
+trajectories of average community value of key functional traits related
+to resource acquisition (leaf and stem traits) and life-history strategy
+(seed mass, maximum size)
 \citep{Wright2004, Diaz2005, TerSteege2006, Westoby2006a, Chave2009b}.
@@
 community recovery time.
-
-\citet{Rees2001}{]}. Yet, observed trajectories proved to often deviate
-from predictions and from the recovery of initial state and thus
-evidenced the interplay of stochastic processes
-\citep{Hubbel2001, Norden2015, Norden2017}. The debate underlying the
-study of community response to disturbance is now on the relative
-importance of stochastic and deterministic processes, and on the nature
-of ecological rules and mechanisms they generate.
-
-Specifically, the determinism of post-disturbance trajectories is the
-base of the intermediate disturbance hypothesis (IDH) which is assumed
-to be a determinant of tropical forest biodiversity \citep{Connell1978}.
 
@@
 trajectories maintaining community divergence in taxonomic composition.
-In accordance with the IDH, community taxonomic diversity was maximise
+In accordance with the IDH, community taxonomic diversity was maximised
 for intermediate disturbance intensity, before reaching an intensity

### 2019-01-31 WholePlotTrajectories.Rmd
relecture

@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, we elucidateed (i) the community taxonomic and functional recovery and (ii) the ecological processes underlying community post-disturbance trajectories and discussed the validity and the extent of the IDH in the long term for tropical forest, and (iii) examined community recovery time.
+Specifically, we elucidateed (i) the community taxonomic and functional recovery and (ii) the ecological processes underlying community post-disturbance trajectories and discussed the validity and the extent of the IDH, and (iii) examined community recovery time.
 
@@
 
-Regarding community functional response in contrast, the IDH seemed disproved.
-Whenever the disturbance intensity, both community functional richness and evenness increased, with the recruitment of pioneers that were functionally highly different from the species of the pre-disturbance community [@Denslow1980;@Molino2001].
+Regarding community functional trajectories in contrast, there was neither intermediate disturbance intensity nor time to disturbance, which dismissed the IDH.
+The functional response according to disturbance intensity did not display the characteristic hump-shaped trajectories.
+Whenever the disturbance intensity, both community functional richness and evenness increased with the recruitment of pioneers that were functionally highly different from the species of the pre-disturbance community [@Denslow1980;@Molino2001].
 Above the intensity threshold however, for the most intense disturbance, community functional richness and evenness started to decrease after 15 to 20 years. 
@@
 
-The IDH then translated into community post-disturbance taxonomic trajectories, with marked differences below and above an intensity threshold. 
-It was however not tangible regarding the functional trajectories, because of their decoupling with taxonomic trajectories.
+The IDH then translated into community taxonomic response, with markedly different trajectories below and above an intensity threshold, and with an intermediate time to disturbance when the intensity remained below the threshold.
+It was however not tangible regarding the functional response, with functional trajectories decoupled from taxonmic ones.
 

### 2019-01-31 WholePlotTrajectories.tex
relecture

@@
 trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, we (i) elucidated community taxonomic and functional
-recovery and the underlying ecological processes, (ii) clarified the
-validity of the IDH in the long term for tropical forest and its
-translation into different trajectories in time, and (iii) examined
-community recovery time.
+Specifically, we elucidateed (i) the community taxonomic and functional
+recovery and (ii) the ecological processes underlying community
+post-disturbance trajectories and discussed the validity and the extent
+of the IDH, and (iii) examined community recovery time.
 
@@
 
-Regarding community functional response in contrast, the IDH seemed
-disproved. Whenever the disturbance intensity, both community functional
-richness and evenness increased, with the recruitment of pioneers that
-were functionally highly different from the species of the
+Regarding community functional trajectories in contrast, there was
+neither intermediate disturbance intensity nor time to disturbance,
+which dismissed the IDH. The functional response according to
+disturbance intensity did not display the characteristic hump-shaped
+trajectories. Whenever the disturbance intensity, both community
+functional richness and evenness increased with the recruitment of
+pioneers that were functionally highly different from the species of the
 pre-disturbance community \citep{Denslow1980, Molino2001}. Above the
@@
 
-The IDH then translated into community post-disturbance taxonomic
-trajectories, with marked differences below and above an intensity
-threshold. It was however not tangible regarding the functional
-trajectories, because of their decoupling with taxonomic trajectories.
+The IDH then translated into community taxonomic response, with markedly
+different trajectories below and above an intensity threshold, and with
+an intermediate time to disturbance when the intensity remained below
+the threshold. It was however not tangible regarding the functional
+response, with functional trajectories decoupled from taxonmic ones.
 

### 2019-02-05 WholePlotTrajectories.Rmd
intro

@@
 The growing demand in forests products together with current global changes, however, increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
-This changing context alters the natural disturbance regime that shape and maintain tree community structure, composition and functioning [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015], so anticipate the fate of tropical forests requires to understand community response to disturbance.
-Disturbance initiate ecological succession comprising a suit of changes in community structure and functioning over time, which rate, endpoint an duration depend on the disturbance intensity [@Chazdon2003;@Willig2018].
-Regarding tropical forests, community response to disturbance would be secondary succession, starting with an intact soil and a consistent seed bank, from which we aim would be to highlight predictable trajectories and their underlying ecological processes.
+This changing context alters the natural disturbance regime that shape and maintain tree community structure, composition and functioning [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
+Anticipate the fate of tropical forests then requires to understand the determinants of community response to disturbance and the underlying processes.
+In tropical forests, disturbance is followed by secondary succession that starts with a maintained forest environment.
+Post-disturbance successional trajectories follow environmental variability and comprises a suit of changes in community structure and functioning.
+This succession is driven by various processes involving the different species responses to disturbance (as colonisation, facilitation, coexistence, etc) [@Lindenmayer2012;@Garcia_florez2017].
+The rate, endpoint an duration of these trajectories depend on the intensity of the disturbance: a current aim of tropical ecology is to highlight the linkage between biodiversity and disturbance and infer predictable trajectories [@Chazdon2003a;@Willig2018].
 
-Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017], and modifications in species interactions, mainly through modifications of species density [@Chesson2000].
-For now, post-disturbance response has been largely studied through forest structural parameters, rapid and convenient to measure, such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016].
-Thereafter, models based on the observed trajectories after disturbance gave important insights into community recovery and resilience [@Herault2018]. 
-Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood.
-A leading idea on how the impacts of disturbance on community diversity is the Intermediate Disturbance Hypothesis (IDH) [@Connell1978].
-The IDH assumes biodiversity changes in relation to disturbance intensity and frequency, with diversity peaks for intermediate intensity and at intermediate time after disturbance. 
-At intermediate level of disturbance, the fluctuation of environmental conditions entails deterministic processes favoring both competitively superior species and fast colonizers, hence preventing competitive exclusion [@Shea2004;@Pulsford2016].
+Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
+For now, post-disturbance trajectories has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
+Models based on observed trajectories then assessed community time to recovery and highlighted the role of forest structure, forest composition, environmental parameters and disturbance characteristics for the determination of post-disturbance trajectories [@Herault2018].
+Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
+There is a need to further analyse community post-disturbance trajectories to assess community recovery and linkage between biodiversity and disturbance.
 
-The IDH then assumes both predictable trajectories depending on the disturbance intensity, and, without additional disturbance, the recovery of pre-disturbance state defined by the dominance of late-successional species [@Rees2001;@Sheil2003;@Willig2018].
-In highly diverse tropical forests, however, the validation of the IDH remains controversial [@Fox2013;@Sheil2013] and observed post-disturbance trajectories are often blurry compared to theoretical expectations [@Hubbel2001;@Hugues2007;@Norden2015;@Norden2017].
-Empirical evidence generally rely on the analysis of a single diversity metric, often the taxonomic richness [@Molino2001], which might give limited or misleading information on community disturbance response.
-The IDH pattern of diversity peak under an intermediate disturbance regime can indeed result from various ecological processes [@Sheil2003].
-To disentangle these various processes and grasp the different facets of community response to disturbance it is crucial to combine a suitable range of approaches [@Shea2004].
-More ecological-meaningful analysis would first couple taxonomic richness and evenness, which reveals the changes in species abundance distribution and hence in the underlying ecological processes [@Chaudhary2016].
-Analysis would as well include taxonomic composition, which is crucial for conservation issues [@Lavorel2002;@Bellwood2006]. 
-Eventually, including a functional approach accounting for species biological attributes would be highly revealing of ecological process involved in community response to disturbance.
-The functional trait-based approach indeed directly links community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
-In that respect, a vast literature allowed recognizing major traits related to species ecology and mediating species performance in a given environment [@Diaz2005]. 
-Specifically in tropical forests, the functional approach revealed the post-disturbance deterministic processes entailing a shift from a dominance of “conservative” slow-growing species dealing with scarce resources, to a dominance of “acquisitive” fast-growing species with rapid and efficient use of abundant resources [@Reich2014;@Herault2011].
-This shift is translated into the trajectories of average community value of key functional traits related to resource acquisition (leaf and stem traits) and life-history strategy (seed mass, maximum size)  [@Wright2004;@Diaz2005;@TerSteege2006;@Westoby2006a;@Chave2009b].
-The functional approach would eventually be completed with the analysis of functional redundancy that quantifies the amount of shared trait values among species [@Carmona2016].
+An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
+The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978], and states that community diversity peaks at intermediate level of disturbance.
+Post-disturbance fluctuations of environmental conditions are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
+In highly diverse tropical forests, however, the validation of the IDH remains controversial [@Hubbell2001;@Fox2013;@Sheil2013] and observations are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017].
+The IDH pattern of peaked diversity under intermediate disturbance can result from various processes that should be identified to grasp the conditions for the IDH application and the impact on communities, and to anticipate disturbance outcomes [@Sheil2003;@Shea2004].
+
+Empirical evidence of the IDH generally rely on the analysis of a single diversity metric, often the taxonomic richness [@Molino2001], which is limited as neither community changes in species abundance nor in functional characteristics are accounted for [@Mayfield2010;@Villeger2011].
+The post-disturbance processes involved, though, can only be disentangled in combining different approaches [@Sheil2003;@Shea2004].
+More ecological-meaningful analysis would couple taxonomic richness, evenness, and composition, to highlight the changes in species composition and abundance distribution, and hence the changes in fostered species [@Chaudhary2016].
+Analysis would as well include a functional approach, that accounts for species biological attributes, would include as the trait-based approach indeed directly links community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
+
+In that respect, a vast literature allowed recognizing major traits that represent species ecological strategy and determine how they respond to changing conditions [@Diaz2005]. 
+Specifically in tropical forests, the functional approach revealed the post-disturbance deterministic processes entailing a shift from a dominance of “conservative” slow-growing species dealing with scarce resources, to a dominance of “acquisitive” fast-growing species with rapid and efficient use of abundant resources [@Rees2001;@Reich2014;@Herault2011].
+This shift is translated into the trajectories of average community value of key functional traits related to resource acquisition (leaf and stem traits) and life-history strategy (seed mass, maximum size)  [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
+The functional approach also encompasses the analysis of functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
 The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Elmqvist2003; @Diaz2005].
 
-In this study, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem above-ground biomass (AGB) loss.
+In this study, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of above-ground biomass (AGB) loss.
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.

### 2019-02-05 WholePlotTrajectories.tex
intro

@@
 and maintain tree community structure, composition and functioning
-\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}, so anticipate
-the fate of tropical forests requires to understand community response
-to disturbance. Disturbance initiate ecological succession comprising a
-suit of changes in community structure and functioning over time, which
-rate, endpoint an duration depend on the disturbance intensity
-\citep{Chazdon2003, Willig2018}. Regarding tropical forests, community
-response to disturbance would be secondary succession, starting with an
-intact soil and a consistent seed bank, from which we aim would be to
-highlight predictable trajectories and their underlying ecological
-processes.
+\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. Anticipate the
+fate of tropical forests then requires to understand the determinants of
+community response to disturbance and the underlying processes. In
+tropical forests, disturbance is followed by secondary succession that
+starts with a maintained forest environment. Post-disturbance
+successional trajectories follow environmental variability and comprises
+a suit of changes in community structure and functioning. This
+succession is driven by various processes involving the different
+species responses to disturbance (as colonisation, facilitation,
+coexistence, etc) \citep{Lindenmayer2012, Garcia_florez2017}. The rate,
+endpoint an duration of these trajectories depend on the intensity of
+the disturbance: a current aim of tropical ecology is to highlight the
+linkage between biodiversity and disturbance and infer predictable
+trajectories \citep{Chazdon2003a, Willig2018}.
 
@@
 through modifications in the fluxes of light, heat and water
-\citep{Goulamoussene2017}, and modifications in species interactions,
-mainly through modifications of species density \citep{Chesson2000}. For
-now, post-disturbance response has been largely studied through forest
-structural parameters, rapid and convenient to measure, such as
+\citep{Goulamoussene2017}. For now, post-disturbance trajectories has
+been largely studied through forest structural parameters such as
 aboveground biomass, tree height or stem density
-\citep{Piponiot2016, Rutishauser2016}. Thereafter, models based on the
-observed trajectories after disturbance gave important insights into
-community recovery and resilience \citep{Herault2018}. Regarding
-community diversity and composition, however, post-disturbance
-trajectories have not been as thoroughly understood. A leading idea on
-how the impacts of disturbance on community diversity is the
-Intermediate Disturbance Hypothesis (IDH) \citep{Connell1978}. The IDH
-assumes biodiversity changes in relation to disturbance intensity and
-frequency, with diversity peaks for intermediate intensity and at
-intermediate time after disturbance. At intermediate level of
-disturbance, the fluctuation of environmental conditions entails
-deterministic processes favoring both competitively superior species and
-fast colonizers, hence preventing competitive exclusion
-\citep{Shea2004, Pulsford2016}.
+\citep{Piponiot2016, Rutishauser2016}. Models based on observed
+trajectories then assessed community time to recovery and highlighted
+the role of forest structure, forest composition, environmental
+parameters and disturbance characteristics for the determination of
+post-disturbance trajectories \citep{Herault2018}. Regarding community
+diversity and composition, however, post-disturbance trajectories have
+not been as thoroughly understood \citep{Guitet2018, Molino2001}. There
+is a need to further analyse community post-disturbance trajectories to
+assess community recovery and linkage between biodiversity and
+disturbance.
 
-The IDH then assumes both predictable trajectories depending on the
-disturbance intensity, and, without additional disturbance, the recovery
-of pre-disturbance state defined by the dominance of late-successional
-species \citep{Rees2001, Sheil2003, Willig2018}. In highly diverse
-tropical forests, however, the validation of the IDH remains
-controversial \citep{Fox2013, Sheil2013} and observed post-disturbance
-trajectories are often blurry compared to theoretical expectations
-\citep{Hubbel2001, Hugues2007, Norden2015, Norden2017}. Empirical
-evidence generally rely on the analysis of a single diversity metric,
-often the taxonomic richness \citep{Molino2001}, which might give
-limited or misleading information on community disturbance response. The
-IDH pattern of diversity peak under an intermediate disturbance regime
-can indeed result from various ecological processes \citep{Sheil2003}.
-To disentangle these various processes and grasp the different facets of
-community response to disturbance it is crucial to combine a suitable
-range of approaches \citep{Shea2004}. More ecological-meaningful
-analysis would first couple taxonomic richness and evenness, which
-reveals the changes in species abundance distribution and hence in the
-underlying ecological processes \citep{Chaudhary2016}. Analysis would as
-well include taxonomic composition, which is crucial for conservation
-issues \citep{Lavorel2002, Bellwood2006}. Eventually, including a
-functional approach accounting for species biological attributes would
-be highly revealing of ecological process involved in community response
-to disturbance. The functional trait-based approach indeed directly
-links community diversity, composition and redundancy to ecosystem
-functioning and environmental constraints
-\citep{Violle2007b, Baraloto2012a}. In that respect, a vast literature
-allowed recognizing major traits related to species ecology and
-mediating species performance in a given environment \citep{Diaz2005}.
-Specifically in tropical forests, the functional approach revealed the
-post-disturbance deterministic processes entailing a shift from a
-dominance of ``conservative'' slow-growing species dealing with scarce
-resources, to a dominance of ``acquisitive'' fast-growing species with
-rapid and efficient use of abundant resources
-\citep{Reich2014, Herault2011}. This shift is translated into the
-trajectories of average community value of key functional traits related
-to resource acquisition (leaf and stem traits) and life-history strategy
-(seed mass, maximum size)
-\citep{Wright2004, Diaz2005, TerSteege2006, Westoby2006a, Chave2009b}.
-The functional approach would eventually be completed with the analysis
-of functional redundancy that quantifies the amount of shared trait
-values among species \citep{Carmona2016}. The high functional redundancy
-of hyper-diverse tropical forests \citep{Bellwood2006} mitigates the
+An early conceptual basis of the linkage between biodiversity and
+disturbance is the Intermediate Disturbance Hypothesis (IDH). The IDH
+assumes a relationship between community diversity and the intensity and
+frequency of disturbance events \citep{Connell1978}, and states that
+community diversity peaks at intermediate level of disturbance.
+Post-disturbance fluctuations of environmental conditions are assumed to
+prevent competitive exclusion in fostering both competitively superior
+species and fast colonizers \citep{Shea2004, Pulsford2016}. In highly
+diverse tropical forests, however, the validation of the IDH remains
+controversial \citep{Hubbell2001, Fox2013, Sheil2013} and observations
+are often blurry compared to theoretical expectations
+\citep{Hugues2007, Sheil2003, Norden2017}. The IDH pattern of peaked
+diversity under intermediate disturbance can result from various
+processes that should be identified to grasp the conditions for the IDH
+application and the impact on communities, and to anticipate disturbance
+outcomes \citep{Sheil2003, Shea2004}.
+
+Empirical evidence of the IDH generally rely on the analysis of a single
+diversity metric, often the taxonomic richness \citep{Molino2001}, which
+is limited as neither community changes in species abundance nor in
+functional characteristics are accounted for
+\citep{Mayfield2010, Villeger2011}. The post-disturbance processes
+involved, though, can only be disentangled in combining different
+approaches \citep{Sheil2003, Shea2004}. More ecological-meaningful
+analysis would couple taxonomic richness, evenness, and composition, to
+highlight the changes in species composition and abundance distribution,
+and hence the changes in fostered species \citep{Chaudhary2016}.
+Analysis would as well include a functional approach, that accounts for
+species biological attributes, would include as the trait-based approach
+indeed directly links community diversity, composition and redundancy to
+ecosystem functioning and environmental constraints
+\citep{Violle2007b, Baraloto2012a}.
+
+In that respect, a vast literature allowed recognizing major traits that
+represent species ecological strategy and determine how they respond to
+changing conditions \citep{Diaz2005}. Specifically in tropical forests,
+the functional approach revealed the post-disturbance deterministic
+processes entailing a shift from a dominance of ``conservative''
+slow-growing species dealing with scarce resources, to a dominance of
+``acquisitive'' fast-growing species with rapid and efficient use of
+abundant resources \citep{Rees2001, Reich2014, Herault2011}. This shift
+is translated into the trajectories of average community value of key
+functional traits related to resource acquisition (leaf and stem traits)
+and life-history strategy (seed mass, maximum size)
+\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. The
+functional approach also encompasses the analysis of functional
+redundancy, that quantifies the amount of shared trait values among
+species \citep{Carmona2016}. The high functional redundancy of
+hyper-diverse tropical forests \citep{Bellwood2006} mitigates the
 impacts of species removal on ecosystem functioning and determines the
@@
 Neotropical forest plots set up on a gradient of disturbance intensity,
-from 10 to 60\% of ecosystem above-ground biomass (AGB) loss. We made
-use of a large functional traits database encompassing major leaf, stem
-and life-history traits in order to draw the taxonomic and functional
+from 10 to 60\% of above-ground biomass (AGB) loss. We made use of a
+large functional traits database encompassing major leaf, stem and
+life-history traits in order to draw the taxonomic and functional
 trajectories in terms of richness, evenness, composition and redundancy.

### 2019-02-07 WholePlotTrajectories.Rmd
intro

@@
 The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
-They provide wood and multiple non-timber forest products, shelter a diversified fauna, regulate the local and regional climates, the carbon, water and nutrient cycles, and ensure cultural and human well-being.
-The growing demand in forests products together with current global changes, however, increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
-This changing context alters the natural disturbance regime that shape and maintain tree community structure, composition and functioning [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
-Anticipate the fate of tropical forests then requires to understand the determinants of community response to disturbance and the underlying processes.
-In tropical forests, disturbance is followed by secondary succession that starts with a maintained forest environment.
-Post-disturbance successional trajectories follow environmental variability and comprises a suit of changes in community structure and functioning.
-This succession is driven by various processes involving the different species responses to disturbance (as colonisation, facilitation, coexistence, etc) [@Lindenmayer2012;@Garcia_florez2017].
-The rate, endpoint an duration of these trajectories depend on the intensity of the disturbance: a current aim of tropical ecology is to highlight the linkage between biodiversity and disturbance and infer predictable trajectories [@Chazdon2003a;@Willig2018].
-
+They provide wood and multiple non-timber forest products, shelter a diversified fauna, and ensure cultural and human well-being.
+They regulate as well the local and regional climates, and the carbon, water and nutrient cycles.
+However, the growing demand in forests products together with current global changes increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
+This growing pressure affects the natural disturbance regime that shape and maintain the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
+It is then urgent to understand the processes underlying community response to disturbance in order to anticipate the fate of tropical forests in the current context.
+In tropical forests, disturbance maintains a forest environment and is followed by secondary succession comprising a suit of changes in the structure and functioning of communities.
 Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
 For now, post-disturbance trajectories has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
-Models based on observed trajectories then assessed community time to recovery and highlighted the role of forest structure, forest composition, environmental parameters and disturbance characteristics for the determination of post-disturbance trajectories [@Herault2018].
+Models based on observed trajectories then assessed community time to recovery and highlighted the determinants of post-disturbance trajectories, as forest structure, forest composition, environmental parameters and disturbance characteristics [@Herault2018].
 Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
-There is a need to further analyse community post-disturbance trajectories to assess community recovery and linkage between biodiversity and disturbance.
+Biodiversity trajectories result from the differences in species response to disturbance which entail a variety of processes like colonisation, facilitation, coexistence [@Lindenmayer2012;@Garcia_florez2017].
+The rate, endpoint and duration of post-disturbance trajectories depend on the intensity and frequency of disturbance: a current objective of tropical ecology is to highlight this linkage betwenn biodiversity and disturbance, and anticipate community response [@Chazdon2003a;@Willig2018].
 
@@
 The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978], and states that community diversity peaks at intermediate level of disturbance.
-Post-disturbance fluctuations of environmental conditions are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
-In highly diverse tropical forests, however, the validation of the IDH remains controversial [@Hubbell2001;@Fox2013;@Sheil2013] and observations are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017].
-The IDH pattern of peaked diversity under intermediate disturbance can result from various processes that should be identified to grasp the conditions for the IDH application and the impact on communities, and to anticipate disturbance outcomes [@Sheil2003;@Shea2004].
+The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion as they foster both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
+In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017] the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
+As previously highlighted, there is a variety of ecological processes emerging after disturbance, and the pattern of the IDH (the diversity peak under intermediate disturbance) can result from various processes.
+Disentangle these processes in the context of tropical forests would assess the conditions for the IDH application and clarify the disturbance impact on communit diversity [@Sheil2003;@Shea2004].
 

### 2019-02-07 WholePlotTrajectories.tex
intro

@@
 environmental, economic and social values. They provide wood and
-multiple non-timber forest products, shelter a diversified fauna,
-regulate the local and regional climates, the carbon, water and nutrient
-cycles, and ensure cultural and human well-being. The growing demand in
-forests products together with current global changes, however, increase
-the pressure on remaining natural forests \citep{Morales-Hidalgo2015}.
-This changing context alters the natural disturbance regime that shape
-and maintain tree community structure, composition and functioning
-\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. Anticipate the
-fate of tropical forests then requires to understand the determinants of
-community response to disturbance and the underlying processes. In
-tropical forests, disturbance is followed by secondary succession that
-starts with a maintained forest environment. Post-disturbance
-successional trajectories follow environmental variability and comprises
-a suit of changes in community structure and functioning. This
-succession is driven by various processes involving the different
-species responses to disturbance (as colonisation, facilitation,
-coexistence, etc) \citep{Lindenmayer2012, Garcia_florez2017}. The rate,
-endpoint an duration of these trajectories depend on the intensity of
-the disturbance: a current aim of tropical ecology is to highlight the
-linkage between biodiversity and disturbance and infer predictable
-trajectories \citep{Chazdon2003a, Willig2018}.
-
-Disturbance change both community abiotic and biotic environments,
-through modifications in the fluxes of light, heat and water
+multiple non-timber forest products, shelter a diversified fauna, and
+ensure cultural and human well-being. They regulate as well the local
+and regional climates, and the carbon, water and nutrient cycles.
+However, the growing demand in forests products together with current
+global changes increase the pressure on remaining natural forests
+\citep{Morales-Hidalgo2015}. This growing pressure affects the natural
+disturbance regime that shape and maintain the structure, composition
+and functioning of tree communities
+\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. It is then
+urgent to understand the processes underlying community response to
+disturbance in order to anticipate the fate of tropical forests in the
+current context. In tropical forests, disturbance maintains a forest
+environment and is followed by secondary succession comprising a suit of
+changes in the structure and functioning of communities. Disturbance
+change both community abiotic and biotic environments, through
+modifications in the fluxes of light, heat and water
 \citep{Goulamoussene2017}. For now, post-disturbance trajectories has
@@
 trajectories then assessed community time to recovery and highlighted
-the role of forest structure, forest composition, environmental
-parameters and disturbance characteristics for the determination of
-post-disturbance trajectories \citep{Herault2018}. Regarding community
-diversity and composition, however, post-disturbance trajectories have
-not been as thoroughly understood \citep{Guitet2018, Molino2001}. There
-is a need to further analyse community post-disturbance trajectories to
-assess community recovery and linkage between biodiversity and
-disturbance.
+the determinants of post-disturbance trajectories, as forest structure,
+forest composition, environmental parameters and disturbance
+characteristics \citep{Herault2018}. Regarding community diversity and
+composition, however, post-disturbance trajectories have not been as
+thoroughly understood \citep{Guitet2018, Molino2001}. Biodiversity
+trajectories result from the differences in species response to
+disturbance which entail a variety of processes like colonisation,
+facilitation, coexistence \citep{Lindenmayer2012, Garcia_florez2017}.
+The rate, endpoint and duration of post-disturbance trajectories depend
+on the intensity and frequency of disturbance: a current objective of
+tropical ecology is to highlight this linkage betwenn biodiversity and
+disturbance, and anticipate community response
+\citep{Chazdon2003a, Willig2018}.
 
@@
 frequency of disturbance events \citep{Connell1978}, and states that
-community diversity peaks at intermediate level of disturbance.
-Post-disturbance fluctuations of environmental conditions are assumed to
-prevent competitive exclusion in fostering both competitively superior
-species and fast colonizers \citep{Shea2004, Pulsford2016}. In highly
-diverse tropical forests, however, the validation of the IDH remains
-controversial \citep{Hubbell2001, Fox2013, Sheil2013} and observations
-are often blurry compared to theoretical expectations
-\citep{Hugues2007, Sheil2003, Norden2017}. The IDH pattern of peaked
-diversity under intermediate disturbance can result from various
-processes that should be identified to grasp the conditions for the IDH
-application and the impact on communities, and to anticipate disturbance
-outcomes \citep{Sheil2003, Shea2004}.
+community diversity peaks at intermediate level of disturbance. The
+fluctuations of environmental conditions after disturbance are assumed
+to prevent competitive exclusion as they foster both competitively
+superior species and fast colonizers \citep{Shea2004, Pulsford2016}. In
+highly diverse tropical forests, however, observations of the IDH are
+often blurry compared to theoretical expectations
+\citep{Hugues2007, Sheil2003, Norden2017} the validation of the theory
+remains controversial \citep{Hubbell2001, Fox2013, Sheil2013}. As
+previously highlighted, there is a variety of ecological processes
+emerging after disturbance, and the pattern of the IDH (the diversity
+peak under intermediate disturbance) can result from various processes.
+Disentangle these processes in the context of tropical forests would
+assess the conditions for the IDH application and clarify the
+disturbance impact on communit diversity \citep{Sheil2003, Shea2004}.
 

### 2019-02-09 WholePlotTrajectories.Rmd
intro

@@
 However, the growing demand in forests products together with current global changes increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
-This growing pressure affects the natural disturbance regime that shape and maintain the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
-It is then urgent to understand the processes underlying community response to disturbance in order to anticipate the fate of tropical forests in the current context.
-In tropical forests, disturbance maintains a forest environment and is followed by secondary succession comprising a suit of changes in the structure and functioning of communities.
+These threats affect the natural disturbance regime that shape and maintain the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
+It is now urgent to understand the processes underlying community response to disturbance, and thus to be able to anticipate the fate of tropical forests in the current context.
+In tropical forests, disturbance maintains a forest environment and initiates a suit of changes in the structure and functioning of communities consituting a secondary succession.
 Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
 For now, post-disturbance trajectories has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
-Models based on observed trajectories then assessed community time to recovery and highlighted the determinants of post-disturbance trajectories, as forest structure, forest composition, environmental parameters and disturbance characteristics [@Herault2018].
+Models based on observed trajectories then assessed community time to recovery and identified some of the determinants of post-disturbance trajectories, like forest structure, forest composition, environmental parameters and disturbance characteristics [@Herault2018].
 Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
-Biodiversity trajectories result from the differences in species response to disturbance which entail a variety of processes like colonisation, facilitation, coexistence [@Lindenmayer2012;@Garcia_florez2017].
-The rate, endpoint and duration of post-disturbance trajectories depend on the intensity and frequency of disturbance: a current objective of tropical ecology is to highlight this linkage betwenn biodiversity and disturbance, and anticipate community response [@Chazdon2003a;@Willig2018].
+The variety of species response to disturbance entail different processes like colonisation, facilitation, coexistence, etc, which succession shapes biodiversity post-disturbance trajectories [@Lindenmayer2012;@Garcia_florez2017].
+The rate, endpoint and duration of the trajectories depend on the intensity and frequency of disturbance: a current objective of tropical ecology is to highlight the link betwenn biodiversity and disturbance and anticipate community trajectories [@Chazdon2003a;@Willig2018].
 
 An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
-The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978], and states that community diversity peaks at intermediate level of disturbance.
-The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion as they foster both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
-In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017] the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
-As previously highlighted, there is a variety of ecological processes emerging after disturbance, and the pattern of the IDH (the diversity peak under intermediate disturbance) can result from various processes.
-Disentangle these processes in the context of tropical forests would assess the conditions for the IDH application and clarify the disturbance impact on communit diversity [@Sheil2003;@Shea2004].
+The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978].
+The theory and states that community diversity peaks at intermediate level of disturbance.
+The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
+In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
+As previously highlighted, there is a variety of ecological processes emerging after disturbance, and the pattern of the IDH (the diversity peak under intermediate disturbance) can result from several such processes.
+Disentangle the processes involved and their succession in time would allow discussing the conditions for the application of the IDH, its scope in time, and its implication for the fate of forests [@Sheil2003;@Shea2004].
 
-Empirical evidence of the IDH generally rely on the analysis of a single diversity metric, often the taxonomic richness [@Molino2001], which is limited as neither community changes in species abundance nor in functional characteristics are accounted for [@Mayfield2010;@Villeger2011].
-The post-disturbance processes involved, though, can only be disentangled in combining different approaches [@Sheil2003;@Shea2004].
-More ecological-meaningful analysis would couple taxonomic richness, evenness, and composition, to highlight the changes in species composition and abundance distribution, and hence the changes in fostered species [@Chaudhary2016].
-Analysis would as well include a functional approach, that accounts for species biological attributes, would include as the trait-based approach indeed directly links community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
-
+Examine post-disturbance trajectories requires a set of diversity metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
+Specifically, diversity metrics should assess taxonomic richness, evenness and composition, in order to highlight the changes the community abundance distribution and in the species fostered by the disturbance.
+Analysis would as well include a functional approach accounting for species biological attributes and ecological strategy.
+The trait-based approach directly links community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a]. 
 In that respect, a vast literature allowed recognizing major traits that represent species ecological strategy and determine how they respond to changing conditions [@Diaz2005]. 
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, we elucidateed (i) the community taxonomic and functional recovery and (ii) the ecological processes underlying community post-disturbance trajectories and discussed the validity and the extent of the IDH, and (iii) examined community recovery time.
+Specifically, we (i) elucidated community taxonomic and functional resilience, (ii) discussed the scope of the IDH and the underlying ecological processes, and (iii) examined community recovery time.
 

### 2019-02-09 WholePlotTrajectories.tex
intro

@@
 global changes increase the pressure on remaining natural forests
-\citep{Morales-Hidalgo2015}. This growing pressure affects the natural
+\citep{Morales-Hidalgo2015}. These threats affect the natural
 disturbance regime that shape and maintain the structure, composition
 and functioning of tree communities
-\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. It is then
-urgent to understand the processes underlying community response to
-disturbance in order to anticipate the fate of tropical forests in the
-current context. In tropical forests, disturbance maintains a forest
-environment and is followed by secondary succession comprising a suit of
-changes in the structure and functioning of communities. Disturbance
-change both community abiotic and biotic environments, through
-modifications in the fluxes of light, heat and water
-\citep{Goulamoussene2017}. For now, post-disturbance trajectories has
-been largely studied through forest structural parameters such as
+\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. It is now urgent
+to understand the processes underlying community response to
+disturbance, and thus to be able to anticipate the fate of tropical
+forests in the current context. In tropical forests, disturbance
+maintains a forest environment and initiates a suit of changes in the
+structure and functioning of communities consituting a secondary
+succession. Disturbance change both community abiotic and biotic
+environments, through modifications in the fluxes of light, heat and
+water \citep{Goulamoussene2017}. For now, post-disturbance trajectories
+has been largely studied through forest structural parameters such as
 aboveground biomass, tree height or stem density
 \citep{Piponiot2016, Rutishauser2016}. Models based on observed
-trajectories then assessed community time to recovery and highlighted
-the determinants of post-disturbance trajectories, as forest structure,
-forest composition, environmental parameters and disturbance
+trajectories then assessed community time to recovery and identified
+some of the determinants of post-disturbance trajectories, like forest
+structure, forest composition, environmental parameters and disturbance
 characteristics \citep{Herault2018}. Regarding community diversity and
 composition, however, post-disturbance trajectories have not been as
-thoroughly understood \citep{Guitet2018, Molino2001}. Biodiversity
-trajectories result from the differences in species response to
-disturbance which entail a variety of processes like colonisation,
-facilitation, coexistence \citep{Lindenmayer2012, Garcia_florez2017}.
-The rate, endpoint and duration of post-disturbance trajectories depend
-on the intensity and frequency of disturbance: a current objective of
-tropical ecology is to highlight this linkage betwenn biodiversity and
-disturbance, and anticipate community response
-\citep{Chazdon2003a, Willig2018}.
+thoroughly understood \citep{Guitet2018, Molino2001}. The variety of
+species response to disturbance entail different processes like
+colonisation, facilitation, coexistence, etc, which succession shapes
+biodiversity post-disturbance trajectories
+\citep{Lindenmayer2012, Garcia_florez2017}. The rate, endpoint and
+duration of the trajectories depend on the intensity and frequency of
+disturbance: a current objective of tropical ecology is to highlight the
+link betwenn biodiversity and disturbance and anticipate community
+trajectories \citep{Chazdon2003a, Willig2018}.
 
@@
 assumes a relationship between community diversity and the intensity and
-frequency of disturbance events \citep{Connell1978}, and states that
-community diversity peaks at intermediate level of disturbance. The
-fluctuations of environmental conditions after disturbance are assumed
-to prevent competitive exclusion as they foster both competitively
-superior species and fast colonizers \citep{Shea2004, Pulsford2016}. In
-highly diverse tropical forests, however, observations of the IDH are
-often blurry compared to theoretical expectations
-\citep{Hugues2007, Sheil2003, Norden2017} the validation of the theory
-remains controversial \citep{Hubbell2001, Fox2013, Sheil2013}. As
-previously highlighted, there is a variety of ecological processes
-emerging after disturbance, and the pattern of the IDH (the diversity
-peak under intermediate disturbance) can result from various processes.
-Disentangle these processes in the context of tropical forests would
-assess the conditions for the IDH application and clarify the
-disturbance impact on communit diversity \citep{Sheil2003, Shea2004}.
+frequency of disturbance events \citep{Connell1978}. The theory and
+states that community diversity peaks at intermediate level of
+disturbance. The fluctuations of environmental conditions after
+disturbance are assumed to prevent competitive exclusion in fostering
+both competitively superior species and fast colonizers
+\citep{Shea2004, Pulsford2016}. In highly diverse tropical forests,
+however, observations of the IDH are often blurry compared to
+theoretical expectations \citep{Hugues2007, Sheil2003, Norden2017}, and
+the validation of the theory remains controversial
+\citep{Hubbell2001, Fox2013, Sheil2013}. As previously highlighted,
+there is a variety of ecological processes emerging after disturbance,
+and the pattern of the IDH (the diversity peak under intermediate
+disturbance) can result from several such processes. Disentangle the
+processes involved and their succession in time would allow discussing
+the conditions for the application of the IDH, its scope in time, and
+its implication for the fate of forests \citep{Sheil2003, Shea2004}.
 
-Empirical evidence of the IDH generally rely on the analysis of a single
-diversity metric, often the taxonomic richness \citep{Molino2001}, which
-is limited as neither community changes in species abundance nor in
-functional characteristics are accounted for
-\citep{Mayfield2010, Villeger2011}. The post-disturbance processes
-involved, though, can only be disentangled in combining different
-approaches \citep{Sheil2003, Shea2004}. More ecological-meaningful
-analysis would couple taxonomic richness, evenness, and composition, to
-highlight the changes in species composition and abundance distribution,
-and hence the changes in fostered species \citep{Chaudhary2016}.
-Analysis would as well include a functional approach, that accounts for
-species biological attributes, would include as the trait-based approach
-indeed directly links community diversity, composition and redundancy to
-ecosystem functioning and environmental constraints
-\citep{Violle2007b, Baraloto2012a}.
-
-In that respect, a vast literature allowed recognizing major traits that
-represent species ecological strategy and determine how they respond to
-changing conditions \citep{Diaz2005}. Specifically in tropical forests,
-the functional approach revealed the post-disturbance deterministic
-processes entailing a shift from a dominance of ``conservative''
-slow-growing species dealing with scarce resources, to a dominance of
-``acquisitive'' fast-growing species with rapid and efficient use of
-abundant resources \citep{Rees2001, Reich2014, Herault2011}. This shift
-is translated into the trajectories of average community value of key
-functional traits related to resource acquisition (leaf and stem traits)
-and life-history strategy (seed mass, maximum size)
+Examine post-disturbance trajectories requires a set of diversity
+metrics, to grasp all aspects of community changes
+\citep{Sheil2003, Shea2004, Mayfield2010}. Specifically, diversity
+metrics should assess taxonomic richness, evenness and composition, in
+order to highlight the changes the community abundance distribution and
+in the species fostered by the disturbance. Analysis would as well
+include a functional approach accounting for species biological
+attributes and ecological strategy. The trait-based approach directly
+links community diversity, composition and redundancy to ecosystem
+functioning and environmental constraints
+\citep{Violle2007b, Baraloto2012a}. In that respect, a vast literature
+allowed recognizing major traits that represent species ecological
+strategy and determine how they respond to changing conditions
+\citep{Diaz2005}. Specifically in tropical forests, the functional
+approach revealed the post-disturbance deterministic processes entailing
+a shift from a dominance of ``conservative'' slow-growing species
+dealing with scarce resources, to a dominance of ``acquisitive''
+fast-growing species with rapid and efficient use of abundant resources
+\citep{Rees2001, Reich2014, Herault2011}. This shift is translated into
+the trajectories of average community value of key functional traits
+related to resource acquisition (leaf and stem traits) and life-history
+strategy (seed mass, maximum size)
 \citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. The
@@
 trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, we elucidateed (i) the community taxonomic and functional
-recovery and (ii) the ecological processes underlying community
-post-disturbance trajectories and discussed the validity and the extent
-of the IDH, and (iii) examined community recovery time.
+Specifically, we (i) elucidated community taxonomic and functional
+resilience, (ii) discussed the scope of the IDH and the underlying
+ecological processes, and (iii) examined community recovery time.
 

### 2019-02-11 WholePlotTrajectories.Rmd
intro, dicussion

@@
 
-Examine post-disturbance trajectories requires a set of diversity metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
-Specifically, diversity metrics should assess taxonomic richness, evenness and composition, in order to highlight the changes the community abundance distribution and in the species fostered by the disturbance.
-Analysis would as well include a functional approach accounting for species biological attributes and ecological strategy.
-The trait-based approach directly links community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a]. 
+Fully analyse post-disturbance trajectories requires a set of metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
+The analysis should first consider community composition, which is crucial for conservation issues [@Lavorel2002;@Bellwood2006] and assesses the different species fostered by the disturbance.
+Diversity metrics encompassing taxonomic richness and evenness would assess the changes community abundance distribution. 
+A functional approach accounting for species biological attributes and ecological strategy, would directly link community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
 In that respect, a vast literature allowed recognizing major traits that represent species ecological strategy and determine how they respond to changing conditions [@Diaz2005]. 
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, we (i) elucidated community taxonomic and functional resilience, (ii) discussed the scope of the IDH and the underlying ecological processes, and (iii) examined community recovery time.
+Specifically, (i) from community trajectories we examined their taxonomic and functional resilience, (ii) discussed the scope of the IDH and the underlying ecological processes, and (iii) analyzed community time to recovery.
 
@@
 
-## A cyclic recovery of community composition
+## Decoupled taxonomic and functional trajectories
 
-The hump-shaped trajectories of community composition returned towards pre-disturbance state after 30 years, highlighting community resilience in terms of taxonomic and functional composition.
+Before disturbance, the different communities had different taxonomic compositions, visualized by their distinct starting points on the NMDS axis 2.
+These initial differences were maintained all along the 30 years following disturbance, with post-disturbance trajectories corresponding to a displacement on the NMDS axis 1.
+The taxonomic compositional changes following disturbance were similar among communities, which may correspond to the recruitment of a group of pioneers common to all communities whenever their initial differences and the disturbance intensity [@Denslow2000;@Bongers2009].
+Taxonomic trajectories continued with a recovery of the initial composition which, althought not fully achieved after 30 years, suggested the community taxonomic resilience and the maintenance of community initial differences [@Folke2006].
+The  return towards initial taxonomic composition showed that species not belonging to the pre-disturbance community were rarely recruited, probably because of the common dispersal limitations among tropical tree species [@Svenning2005].
 
-Before disturbance, there were already taxonomic differences among local communities, as revealed by the distinct starting points on the NMDS axis 2.
-These initial differences were maintained throughout the trajectories following disturbance, so community initial state influenced the trajectories following disturbance more than commonly thought.
-The initial composition indeed partly determined the pool of species recruited, and hence shaped the composition trajectories and drove them towards the recovery of the initial composition.
-This taxonomic recovery, although it was not fully achieved after 30 years, suggested the resilience of community taxonomic composition and the maintenance of community initial differences [@Folke2006].
-Species not belonging to the pre-disturbance community were then rarely recruited, probably because of the common dispersal limitations among tropical tree species [@Svenning2005].
-
-Functional composition trajectories were, conversely, highly similar among disturbed communities.
-As the composition of pre-disturbance surviving trees is representative of the initial community [@Herault2018], changes in functional composition relied upon the recruitment of species or functional types that were infrequent or absent before disturbance.
-Competitive pioneers became dominant in filling the environmental niches vacated by the disturbance, with a high availability of light, space and nutrients.
-This recruitment of pioneers similarly changed the functional composition of all communities, whenever the disturbance intensity.
-The functional composition went towards more resource-acquisitive strategies, translated in the functional plan (Fig.
-\@ref(fig:NMDSplans)) by a displacement on the right along the first axis [@Westoby1998; @Wright2004; @Reich2014].
-Thereafter, the pioneers first recruited were progressively excluded by long-lived, more resistant and shade-tolerant species.
-The commpunity functional composition then returned towards more resource-conservative strategies, suggesting the recovery of the initial community composition and translated in the functional plan by a displacement left along the first axis and upward along the second axis (Fig.
+Regarding functional composition, initial communities were similar and trajectories were confounded in the functional space.
+Following disturbance, the functional composition of pre-disturbance surviving trees is belived to be the same as this of the initial community [@Herault2018].
+Post-disturbance changes in functional composition then relied upon the recruitment of species or functional types previously infrequent or absent.
+These species were probably competitive pioneers becoming dominant as the most efficient to benefit from the light, space and nutrients made available by disturbance.
+This recruitment of pioneers drove functional composition trajectories similar for all plots and disturbance intensity.
+The composition first went towards more resource-acquisitive strategies, which was translated by a displacement on the right along the first axis in the functional plan [@Westoby1998; @Wright2004; @Reich2014].
+Thereafter, the pioneers recruited primarily were progressively excluded by long-lived, more resistant and shade-tolerant species.
+The community functional composition then returned towards more resource-conservative strategies, suggesting the recovery of the initial community composition and translated in the functional plan by a displacement left along the first axis and upward along the second axis (Fig.
 \@ref(fig:NMDSplans)).
 
-In the light of these taxonomic and functional composition trajectories, community response to disturbance proved simultaneously deterministic and stochastic at different levels of community organization.
-Taxonomic trajectories were shaped by dispersal limitations maintaining the stochastic differences among communities due to historical contingency. 
-Functional trajectories were shaped by trait-based processes resulting the the convergence of community functional composition [@Fukami2005].
+Both taxonomic and functional composition trajectories initiated a return towards pre-disturbance state after 30 years, which highlighted community taxonomic and functional resilience. 
+There was however a decoupling between taxonomic and functional composition trajectories as taxonomic trajectories maintained the initial differences among communities, while functional trajectories where similar and convergent in the functional space [@Fukami2005].
 
@@
 Community taxonomic richness and evenness trajectories were determined by the disturbance intensity, ranging from a limited and temporary impact to significant and persistant alterations of community diversity.
-The taxonomic trajectories markedly changed above an intensity threshold for which community response corresponded to a maximized richness and a resilience of the evenness. 
+The taxonomic trajectories markedly changed above an intensity threshold for which the taxonomic richness was maximized and the taxonomic evenness was resilient. 
 The disturbance intensity determined the balance in the community between pre-disturbance surviving trees, and trees recruited afterward. 
@@
 The recruitment of pionners, infrequent or absent before disturbance, then increased the taxonomic richness all the more so that the disturbance was intense [@Martin2015; @Chaudhary2016].
-As these pioneers became more dominant they besides temporarily balanced the usual hyper-dominance of tropical forests, hence temporarily increasing the taxonomic evenness [@Baraloto2012a].
+As these pioneers became more dominant they balanced the usual hyper-dominance of tropical forests, hence temporarily increasing the taxonomic evenness [@Baraloto2012a].
 Above the intensity threshold, the taxonomic richness of surviving trees was too low to be offset by the recruitment of pioneers. 
 In the Guiana shield indeed, the pool of true pioneers specifically recruited after disturbance is restricted to a few common genera (e.g. *Cecropia* spp., *Vismia* spp.) [@Guitet2018].
-The richness trajectories following intense disturbance then decreased, all the more so that the disturbance intensity was high [@Molino2001]. 
+The taxonomic richness then decreased following intense disturbance, all the more so that the disturbance intensity was high [@Molino2001]. 
 At the same time, pioneers became persistently dominant and prevented a return towards the initial evenness and the recovery of hyper-dominant shade-tolerant species.
 
-Observed trajectories supported the IDH in highlighting a threshold disturbance intensity, expressed as percentage of AGB removed, marking a change in post-disturbance trajectories.
-Taxonomic trajectories besides highlighted, below the intensity threshold, an intermediate time after when the taxonomic evenness was maximized.
-This time corresponded to the intermediate between the generation time of pionneers, first to dominate the post-disturbance recruitment, and the time for competitive exclusion, leading to the dominance of late-successional species.
+Taxonomic trajectories following disturbance of intermediate intensity besides highlighted an intermediate time after disturbance for which the taxonomic evenness was maximized.
 As already observed in the Guiana shield [@Baraloto2012a] and in Borenan tropical forests [@Cannon1998],
-taxonomic evenness followed humped-shaped trajectories with a peak when the recruitment was balanced between pioneers and late-successional species.
+taxonomic evenness followed hump-shaped trajectories with a peak when the recruitment was balanced between pioneers and late-successional species. 
+This time then corresponded to the match between the generation time of pionneers, first to dominate the post-disturbance recruitment, and the time for competitive exclusion to emerge, which leads to the dominance of late-successional species.
 
-Regarding community functional trajectories in contrast, there was neither intermediate disturbance intensity nor time to disturbance, which dismissed the IDH.
-The functional response according to disturbance intensity did not display the characteristic hump-shaped trajectories.
-Whenever the disturbance intensity, both community functional richness and evenness increased with the recruitment of pioneers that were functionally highly different from the species of the pre-disturbance community [@Denslow1980;@Molino2001].
+Regarding community functional trajectories in contrast, there was neither intermediate disturbance intensity nor intermediate time after disturbance, which dismissed the IDH.
+Neither functional richness nor functional evenness displayed a humped-shaped pattern, but both rather increased with the recruitment of pioneers that were functionally highly different from the pre-disturbance community [@Denslow1980;@Molino2001].
 Above the intensity threshold however, for the most intense disturbance, community functional richness and evenness started to decrease after 15 to 20 years. 
-The vacated environmental space was then occupied by short-lived species which prevented the establishment of other species. 
-The decline of these short-lived pioneers decreased the functional richness and evenness, but we suggest that the establishment of long-lasting pioneers will follow and the taxonomic and functional trajectories will catch up with those observed for intermediate disturbance [@Walker2009].
+The right after disturbance, short-lived species benefited from the resources made available and prevented the establishment of other species. 
+The decline of these short-lived pioneers decreased the functional richness and evenness, but we suggest that the establishment of long-lasting pioneers will follow and that the trajectories will catch up with those observed for intermediate disturbance [@Walker2009].
 
-The IDH then translated into community taxonomic response, with markedly different trajectories below and above an intensity threshold, and with an intermediate time to disturbance when the intensity remained below the threshold.
-It was however not tangible regarding the functional response, with functional trajectories decoupled from taxonmic ones.
+The IDH then translated into community taxonomic response, with post-disturbance trajectories markedly different below and above an intensity threshold.
+The IDH was also valid at an intermediate time after disturbance, when the recruitment of early- and late-successional species were balanced. 
+Functional trajectories, however, remained decoupled from taxonomic ones and dismissed the IDH.
 
@@
 Progressively though, the functional redundancy was restored through the replacement of the first established species by more competitive, long-lived pioneers or late-successional species that were functionally closer to the pre-disturbance community.
-This replacement followed the lottery recruitment rules, implying a recruitment easy for the first recruited species but increasingly hampered by the emergence of interspecific competition [@Busing2002].
-The functional redundancy then relied upon stochastic processes and its recovery, depending on the recruitment of infrequent species constrained by the increasing competition, was increasingly slow and difficult to anticipate [@Elmqvist2003;@Diaz2005].
+This replacement followed the lottery recruitment rules, implying an recruitment easy for the first species but becoming increasingly hampered by the emergence of interspecific competition [@Busing2002].
+The recovery of the functional redundancy then relied upon the random processe of species recruitment and was increasingly slow and difficult to anticipate [@Elmqvist2003;@Diaz2005].
 
-The impact of disturbance on community functional redundancy meant a lower resilience of the pre-disturbance communities, with higher chances to see the persistence of disturbance-specific species at the expense of late-successional ones [@Haddad2008].
-Besides, the stochastic recovery of infrequent species increases the risks to lose keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
-Apart from the functional characteristics considered here, infrequent species might indeed have unique functions in the ecosystem or be a key for some fauna [@Schleuning2016].
+The impact of disturbance on community functional redundancy meant a lower resilience of the disturbed communities, with higher chances to see the settltement of pioneers and early-successional species at the expense of late-successional ones [@Haddad2008].
+Besides, the random recovery of infrequent species increases the risks to lose keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
+Infrequent species might indeed have unique functional characteristics, apart from those considered here, in the ecosystem or be a key for some fauna [@Schleuning2016].
 

### 2019-02-11 WholePlotTrajectories.tex
intro, dicussion

@@
 
-Examine post-disturbance trajectories requires a set of diversity
-metrics, to grasp all aspects of community changes
-\citep{Sheil2003, Shea2004, Mayfield2010}. Specifically, diversity
-metrics should assess taxonomic richness, evenness and composition, in
-order to highlight the changes the community abundance distribution and
-in the species fostered by the disturbance. Analysis would as well
-include a functional approach accounting for species biological
-attributes and ecological strategy. The trait-based approach directly
-links community diversity, composition and redundancy to ecosystem
-functioning and environmental constraints
-\citep{Violle2007b, Baraloto2012a}. In that respect, a vast literature
-allowed recognizing major traits that represent species ecological
-strategy and determine how they respond to changing conditions
-\citep{Diaz2005}. Specifically in tropical forests, the functional
-approach revealed the post-disturbance deterministic processes entailing
-a shift from a dominance of ``conservative'' slow-growing species
-dealing with scarce resources, to a dominance of ``acquisitive''
-fast-growing species with rapid and efficient use of abundant resources
-\citep{Rees2001, Reich2014, Herault2011}. This shift is translated into
-the trajectories of average community value of key functional traits
-related to resource acquisition (leaf and stem traits) and life-history
-strategy (seed mass, maximum size)
+Fully analyse post-disturbance trajectories requires a set of metrics,
+to grasp all aspects of community changes
+\citep{Sheil2003, Shea2004, Mayfield2010}. The analysis should first
+consider community composition, which is crucial for conservation issues
+\citep{Lavorel2002, Bellwood2006} and assesses the different species
+fostered by the disturbance. Diversity metrics encompassing taxonomic
+richness and evenness would assess the changes community abundance
+distribution. A functional approach accounting for species biological
+attributes and ecological strategy, would directly link community
+diversity, composition and redundancy to ecosystem functioning and
+environmental constraints \citep{Violle2007b, Baraloto2012a}. In that
+respect, a vast literature allowed recognizing major traits that
+represent species ecological strategy and determine how they respond to
+changing conditions \citep{Diaz2005}. Specifically in tropical forests,
+the functional approach revealed the post-disturbance deterministic
+processes entailing a shift from a dominance of ``conservative''
+slow-growing species dealing with scarce resources, to a dominance of
+``acquisitive'' fast-growing species with rapid and efficient use of
+abundant resources \citep{Rees2001, Reich2014, Herault2011}. This shift
+is translated into the trajectories of average community value of key
+functional traits related to resource acquisition (leaf and stem traits)
+and life-history strategy (seed mass, maximum size)
 \citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. The
@@
 trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, we (i) elucidated community taxonomic and functional
-resilience, (ii) discussed the scope of the IDH and the underlying
-ecological processes, and (iii) examined community recovery time.
+Specifically, (i) from community trajectories we examined their
+taxonomic and functional resilience, (ii) discussed the scope of the IDH
+and the underlying ecological processes, and (iii) analyzed community
+time to recovery.
 
@@
 
-\subsection{A cyclic recovery of community
-composition}\label{a-cyclic-recovery-of-community-composition}
+\subsection{Decoupled taxonomic and functional
+trajectories}\label{decoupled-taxonomic-and-functional-trajectories}
 
-The hump-shaped trajectories of community composition returned towards
-pre-disturbance state after 30 years, highlighting community resilience
-in terms of taxonomic and functional composition.
+Before disturbance, the different communities had different taxonomic
+compositions, visualized by their distinct starting points on the NMDS
+axis 2. These initial differences were maintained all along the 30 years
+following disturbance, with post-disturbance trajectories corresponding
+to a displacement on the NMDS axis 1. The taxonomic compositional
+changes following disturbance were similar among communities, which may
+correspond to the recruitment of a group of pioneers common to all
+communities whenever their initial differences and the disturbance
+intensity \citep{Denslow2000, Bongers2009}. Taxonomic trajectories
+continued with a recovery of the initial composition which, althought
+not fully achieved after 30 years, suggested the community taxonomic
+resilience and the maintenance of community initial differences
+\citep{Folke2006}. The return towards initial taxonomic composition
+showed that species not belonging to the pre-disturbance community were
+rarely recruited, probably because of the common dispersal limitations
+among tropical tree species \citep{Svenning2005}.
 
-Before disturbance, there were already taxonomic differences among local
-communities, as revealed by the distinct starting points on the NMDS
-axis 2. These initial differences were maintained throughout the
-trajectories following disturbance, so community initial state
-influenced the trajectories following disturbance more than commonly
-thought. The initial composition indeed partly determined the pool of
-species recruited, and hence shaped the composition trajectories and
-drove them towards the recovery of the initial composition. This
-taxonomic recovery, although it was not fully achieved after 30 years,
-suggested the resilience of community taxonomic composition and the
-maintenance of community initial differences \citep{Folke2006}. Species
-not belonging to the pre-disturbance community were then rarely
-recruited, probably because of the common dispersal limitations among
-tropical tree species \citep{Svenning2005}.
+Regarding functional composition, initial communities were similar and
+trajectories were confounded in the functional space. Following
+disturbance, the functional composition of pre-disturbance surviving
+trees is belived to be the same as this of the initial community
+\citep{Herault2018}. Post-disturbance changes in functional composition
+then relied upon the recruitment of species or functional types
+previously infrequent or absent. These species were probably competitive
+pioneers becoming dominant as the most efficient to benefit from the
+light, space and nutrients made available by disturbance. This
+recruitment of pioneers drove functional composition trajectories
+similar for all plots and disturbance intensity. The composition first
+went towards more resource-acquisitive strategies, which was translated
+by a displacement on the right along the first axis in the functional
+plan \citep{Westoby1998, Wright2004, Reich2014}. Thereafter, the
+pioneers recruited primarily were progressively excluded by long-lived,
+more resistant and shade-tolerant species. The community functional
+composition then returned towards more resource-conservative strategies,
+suggesting the recovery of the initial community composition and
+translated in the functional plan by a displacement left along the first
+axis and upward along the second axis (Fig. \ref{fig:NMDSplans}).
 
-Functional composition trajectories were, conversely, highly similar
-among disturbed communities. As the composition of pre-disturbance
-surviving trees is representative of the initial community
-\citep{Herault2018}, changes in functional composition relied upon the
-recruitment of species or functional types that were infrequent or
-absent before disturbance. Competitive pioneers became dominant in
-filling the environmental niches vacated by the disturbance, with a high
-availability of light, space and nutrients. This recruitment of pioneers
-similarly changed the functional composition of all communities,
-whenever the disturbance intensity. The functional composition went
-towards more resource-acquisitive strategies, translated in the
-functional plan (Fig. \ref{fig:NMDSplans}) by a displacement on the
-right along the first axis \citep{Westoby1998, Wright2004, Reich2014}.
-Thereafter, the pioneers first recruited were progressively excluded by
-long-lived, more resistant and shade-tolerant species. The commpunity
-functional composition then returned towards more resource-conservative
-strategies, suggesting the recovery of the initial community composition
-and translated in the functional plan by a displacement left along the
-first axis and upward along the second axis (Fig. \ref{fig:NMDSplans}).
-
-In the light of these taxonomic and functional composition trajectories,
-community response to disturbance proved simultaneously deterministic
-and stochastic at different levels of community organization. Taxonomic
-trajectories were shaped by dispersal limitations maintaining the
-stochastic differences among communities due to historical contingency.
-Functional trajectories were shaped by trait-based processes resulting
-the the convergence of community functional composition
-\citep{Fukami2005}.
+Both taxonomic and functional composition trajectories initiated a
+return towards pre-disturbance state after 30 years, which highlighted
+community taxonomic and functional resilience. There was however a
+decoupling between taxonomic and functional composition trajectories as
+taxonomic trajectories maintained the initial differences among
+communities, while functional trajectories where similar and convergent
+in the functional space \citep{Fukami2005}.
 
@@
 The taxonomic trajectories markedly changed above an intensity threshold
-for which community response corresponded to a maximized richness and a
-resilience of the evenness. The disturbance intensity determined the
-balance in the community between pre-disturbance surviving trees, and
-trees recruited afterward. Below a 20\%-25\% AGB loss, the trees
-surviving after disturbance remained numerous enough to maintain the
-high taxonomic richness of the pre-disturbance community
-\citep{Bongers2009}. The recruitment of pionners, infrequent or absent
-before disturbance, then increased the taxonomic richness all the more
-so that the disturbance was intense \citep{Martin2015, Chaudhary2016}.
-As these pioneers became more dominant they besides temporarily balanced
-the usual hyper-dominance of tropical forests, hence temporarily
-increasing the taxonomic evenness \citep{Baraloto2012a}. Above the
-intensity threshold, the taxonomic richness of surviving trees was too
-low to be offset by the recruitment of pioneers. In the Guiana shield
-indeed, the pool of true pioneers specifically recruited after
-disturbance is restricted to a few common genera (e.g. \emph{Cecropia}
-spp., \emph{Vismia} spp.) \citep{Guitet2018}. The richness trajectories
-following intense disturbance then decreased, all the more so that the
-disturbance intensity was high \citep{Molino2001}. At the same time,
-pioneers became persistently dominant and prevented a return towards the
-initial evenness and the recovery of hyper-dominant shade-tolerant
-species.
+for which the taxonomic richness was maximized and the taxonomic
+evenness was resilient. The disturbance intensity determined the balance
+in the community between pre-disturbance surviving trees, and trees
+recruited afterward. Below a 20\%-25\% AGB loss, the trees surviving
+after disturbance remained numerous enough to maintain the high
+taxonomic richness of the pre-disturbance community \citep{Bongers2009}.
+The recruitment of pionners, infrequent or absent before disturbance,
+then increased the taxonomic richness all the more so that the
+disturbance was intense \citep{Martin2015, Chaudhary2016}. As these
+pioneers became more dominant they balanced the usual hyper-dominance of
+tropical forests, hence temporarily increasing the taxonomic evenness
+\citep{Baraloto2012a}. Above the intensity threshold, the taxonomic
+richness of surviving trees was too low to be offset by the recruitment
+of pioneers. In the Guiana shield indeed, the pool of true pioneers
+specifically recruited after disturbance is restricted to a few common
+genera (e.g. \emph{Cecropia} spp., \emph{Vismia} spp.)
+\citep{Guitet2018}. The taxonomic richness then decreased following
+intense disturbance, all the more so that the disturbance intensity was
+high \citep{Molino2001}. At the same time, pioneers became persistently
+dominant and prevented a return towards the initial evenness and the
+recovery of hyper-dominant shade-tolerant species.
 
-Observed trajectories supported the IDH in highlighting a threshold
-disturbance intensity, expressed as percentage of AGB removed, marking a
-change in post-disturbance trajectories. Taxonomic trajectories besides
-highlighted, below the intensity threshold, an intermediate time after
-when the taxonomic evenness was maximized. This time corresponded to the
-intermediate between the generation time of pionneers, first to dominate
-the post-disturbance recruitment, and the time for competitive
-exclusion, leading to the dominance of late-successional species. As
-already observed in the Guiana shield \citep{Baraloto2012a} and in
-Borenan tropical forests \citep{Cannon1998}, taxonomic evenness followed
-humped-shaped trajectories with a peak when the recruitment was balanced
-between pioneers and late-successional species.
+Taxonomic trajectories following disturbance of intermediate intensity
+besides highlighted an intermediate time after disturbance for which the
+taxonomic evenness was maximized. As already observed in the Guiana
+shield \citep{Baraloto2012a} and in Borenan tropical forests
+\citep{Cannon1998}, taxonomic evenness followed hump-shaped trajectories
+with a peak when the recruitment was balanced between pioneers and
+late-successional species. This time then corresponded to the match
+between the generation time of pionneers, first to dominate the
+post-disturbance recruitment, and the time for competitive exclusion to
+emerge, which leads to the dominance of late-successional species.
 
 Regarding community functional trajectories in contrast, there was
-neither intermediate disturbance intensity nor time to disturbance,
-which dismissed the IDH. The functional response according to
-disturbance intensity did not display the characteristic hump-shaped
-trajectories. Whenever the disturbance intensity, both community
-functional richness and evenness increased with the recruitment of
-pioneers that were functionally highly different from the species of the
-pre-disturbance community \citep{Denslow1980, Molino2001}. Above the
-intensity threshold however, for the most intense disturbance, community
-functional richness and evenness started to decrease after 15 to 20
-years. The vacated environmental space was then occupied by short-lived
-species which prevented the establishment of other species. The decline
+neither intermediate disturbance intensity nor intermediate time after
+disturbance, which dismissed the IDH. Neither functional richness nor
+functional evenness displayed a humped-shaped pattern, but both rather
+increased with the recruitment of pioneers that were functionally highly
+different from the pre-disturbance community
+\citep{Denslow1980, Molino2001}. Above the intensity threshold however,
+for the most intense disturbance, community functional richness and
+evenness started to decrease after 15 to 20 years. The right after
+disturbance, short-lived species benefited from the resources made
+available and prevented the establishment of other species. The decline
 of these short-lived pioneers decreased the functional richness and
 evenness, but we suggest that the establishment of long-lasting pioneers
-will follow and the taxonomic and functional trajectories will catch up
-with those observed for intermediate disturbance \citep{Walker2009}.
+will follow and that the trajectories will catch up with those observed
+for intermediate disturbance \citep{Walker2009}.
 
-The IDH then translated into community taxonomic response, with markedly
-different trajectories below and above an intensity threshold, and with
-an intermediate time to disturbance when the intensity remained below
-the threshold. It was however not tangible regarding the functional
-response, with functional trajectories decoupled from taxonmic ones.
+The IDH then translated into community taxonomic response, with
+post-disturbance trajectories markedly different below and above an
+intensity threshold. The IDH was also valid at an intermediate time
+after disturbance, when the recruitment of early- and late-successional
+species were balanced. Functional trajectories, however, remained
+decoupled from taxonomic ones and dismissed the IDH.
 
@@
 the pre-disturbance community. This replacement followed the lottery
-recruitment rules, implying a recruitment easy for the first recruited
-species but increasingly hampered by the emergence of interspecific
-competition \citep{Busing2002}. The functional redundancy then relied
-upon stochastic processes and its recovery, depending on the recruitment
-of infrequent species constrained by the increasing competition, was
-increasingly slow and difficult to anticipate
+recruitment rules, implying an recruitment easy for the first species
+but becoming increasingly hampered by the emergence of interspecific
+competition \citep{Busing2002}. The recovery of the functional
+redundancy then relied upon the random processe of species recruitment
+and was increasingly slow and difficult to anticipate
 \citep{Elmqvist2003, Diaz2005}.
@@
 The impact of disturbance on community functional redundancy meant a
-lower resilience of the pre-disturbance communities, with higher chances
-to see the persistence of disturbance-specific species at the expense of
-late-successional ones \citep{Haddad2008}. Besides, the stochastic
-recovery of infrequent species increases the risks to lose keystone
-species, with unexpected ecological consequences
-\citep{Jones1994, Chazdon2003a, Diaz2005}. Apart from the functional
-characteristics considered here, infrequent species might indeed have
-unique functions in the ecosystem or be a key for some fauna
+lower resilience of the disturbed communities, with higher chances to
+see the settltement of pioneers and early-successional species at the
+expense of late-successional ones \citep{Haddad2008}. Besides, the
+random recovery of infrequent species increases the risks to lose
+keystone species, with unexpected ecological consequences
+\citep{Jones1994, Chazdon2003a, Diaz2005}. Infrequent species might
+indeed have unique functional characteristics, apart from those
+considered here, in the ecosystem or be a key for some fauna
 \citep{Schleuning2016}.

### 2019-02-13 WholePlotTrajectories.Rmd
intro

@@
 Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
-The variety of species response to disturbance entail different processes like colonisation, facilitation, coexistence, etc, which succession shapes biodiversity post-disturbance trajectories [@Lindenmayer2012;@Garcia_florez2017].
-The rate, endpoint and duration of the trajectories depend on the intensity and frequency of disturbance: a current objective of tropical ecology is to highlight the link betwenn biodiversity and disturbance and anticipate community trajectories [@Chazdon2003a;@Willig2018].
+Given the variety of species response to disturbance and the huge diversity of tropical forests, biodiversity trajectories following disturbance might be very diverse [@Lindenmayer2012;@Garcia_florez2017].
+We assumed that the rate, endpoint and duration of the trajectories depend on the intensity and frequency of disturbance: we then aimed to highlight the link between biodiversity and disturbance and thus anticipate community post-disturbance trajectories [@Chazdon2003a;@Willig2018].
 
@@
 The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
-In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
-As previously highlighted, there is a variety of ecological processes emerging after disturbance, and the pattern of the IDH (the diversity peak under intermediate disturbance) can result from several such processes.
-Disentangle the processes involved and their succession in time would allow discussing the conditions for the application of the IDH, its scope in time, and its implication for the fate of forests [@Sheil2003;@Shea2004].
+In tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
+The IDH may result from several processes and in the highly diverse tree communities its pattern might differ from those recognized otherwise.
+Most of the studies rely on a single diversity metric, often the taxonomic richness [@Molino2001], which is limited to grasp all facets of the diversity-disturbance relationship.
+In this study we aimed as well to clarify the scope of the IDH and discuss its implication for the fate of tropical forests [@Sheil2003;@Shea2004].
 
-Fully analyse post-disturbance trajectories requires a set of metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
+The analyse of community response to disturbance requires a set of metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
 The analysis should first consider community composition, which is crucial for conservation issues [@Lavorel2002;@Bellwood2006] and assesses the different species fostered by the disturbance.
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, (i) from community trajectories we examined their taxonomic and functional resilience, (ii) discussed the scope of the IDH and the underlying ecological processes, and (iii) analyzed community time to recovery.
+Specifically, (i) from post-disturbance trajectories we examined community taxonomic and functional response, (ii) then we discussed the scope of the IDH in taxonomic and functional facets and in time after disturbance, and (iii) we analyzed community resilience and time to recovery.
 
@@
 
-## Decoupled taxonomic and functional trajectories
+## Decoupled taxonomic and functional composition trajectories
 
@@
 
-## A new perspective on the intermediate disturbance hypothesis
+## The scope of the intermediate disturbance hypothesis
 
 Community taxonomic richness and evenness trajectories were determined by the disturbance intensity, ranging from a limited and temporary impact to significant and persistant alterations of community diversity.
-The taxonomic trajectories markedly changed above an intensity threshold for which the taxonomic richness was maximized and the taxonomic evenness was resilient. 
+The taxonomic trajectories markedly changed above an intensity threshold for which the taxonomic richness was maximized and the taxonomic evenness remained resilient. 
 The disturbance intensity determined the balance in the community between pre-disturbance surviving trees, and trees recruited afterward. 
@@
 
-Taxonomic trajectories following disturbance of intermediate intensity besides highlighted an intermediate time after disturbance for which the taxonomic evenness was maximized.
+Taxonomic trajectories following disturbance of intermediate intensity besides highlighted an intermediate time after disturbance, around 15-20 years, for which the taxonomic evenness was maximized.
 As already observed in the Guiana shield [@Baraloto2012a] and in Borenan tropical forests [@Cannon1998],
 taxonomic evenness followed hump-shaped trajectories with a peak when the recruitment was balanced between pioneers and late-successional species. 
-This time then corresponded to the match between the generation time of pionneers, first to dominate the post-disturbance recruitment, and the time for competitive exclusion to emerge, which leads to the dominance of late-successional species.
+This intermediate time corresponded to the match between the generation time of pionneers, first to dominate the post-disturbance recruitment, and the time for competitive exclusion to emerge, which leads to the dominance of late-successional species.
 

### 2019-02-13 WholePlotTrajectories.tex
intro

@@
 composition, however, post-disturbance trajectories have not been as
-thoroughly understood \citep{Guitet2018, Molino2001}. The variety of
-species response to disturbance entail different processes like
-colonisation, facilitation, coexistence, etc, which succession shapes
-biodiversity post-disturbance trajectories
-\citep{Lindenmayer2012, Garcia_florez2017}. The rate, endpoint and
-duration of the trajectories depend on the intensity and frequency of
-disturbance: a current objective of tropical ecology is to highlight the
-link betwenn biodiversity and disturbance and anticipate community
-trajectories \citep{Chazdon2003a, Willig2018}.
+thoroughly understood \citep{Guitet2018, Molino2001}. Given the variety
+of species response to disturbance and the huge diversity of tropical
+forests, biodiversity trajectories following disturbance might be very
+diverse \citep{Lindenmayer2012, Garcia_florez2017}. We assumed that the
+rate, endpoint and duration of the trajectories depend on the intensity
+and frequency of disturbance: we then aimed to highlight the link
+between biodiversity and disturbance and thus anticipate community
+post-disturbance trajectories \citep{Chazdon2003a, Willig2018}.
 
@@
 both competitively superior species and fast colonizers
-\citep{Shea2004, Pulsford2016}. In highly diverse tropical forests,
-however, observations of the IDH are often blurry compared to
-theoretical expectations \citep{Hugues2007, Sheil2003, Norden2017}, and
-the validation of the theory remains controversial
-\citep{Hubbell2001, Fox2013, Sheil2013}. As previously highlighted,
-there is a variety of ecological processes emerging after disturbance,
-and the pattern of the IDH (the diversity peak under intermediate
-disturbance) can result from several such processes. Disentangle the
-processes involved and their succession in time would allow discussing
-the conditions for the application of the IDH, its scope in time, and
-its implication for the fate of forests \citep{Sheil2003, Shea2004}.
+\citep{Shea2004, Pulsford2016}. In tropical forests, however,
+observations of the IDH are often blurry compared to theoretical
+expectations \citep{Hugues2007, Sheil2003, Norden2017}, and the
+validation of the theory remains controversial
+\citep{Hubbell2001, Fox2013, Sheil2013}. The IDH may result from several
+processes and in the highly diverse tree communities its pattern might
+differ from those recognized otherwise. Most of the studies rely on a
+single diversity metric, often the taxonomic richness
+\citep{Molino2001}, which is limited to grasp all facets of the
+diversity-disturbance relationship. In this study we aimed as well to
+clarify the scope of the IDH and discuss its implication for the fate of
+tropical forests \citep{Sheil2003, Shea2004}.
 
-Fully analyse post-disturbance trajectories requires a set of metrics,
-to grasp all aspects of community changes
+The analyse of community response to disturbance requires a set of
+metrics, to grasp all aspects of community changes
 \citep{Sheil2003, Shea2004, Mayfield2010}. The analysis should first
@@
 trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, (i) from community trajectories we examined their
-taxonomic and functional resilience, (ii) discussed the scope of the IDH
-and the underlying ecological processes, and (iii) analyzed community
-time to recovery.
+Specifically, (i) from post-disturbance trajectories we examined
+community taxonomic and functional response, (ii) then we discussed the
+scope of the IDH in taxonomic and functional facets and in time after
+disturbance, and (iii) we analyzed community resilience and time to
+recovery.
 
@@
 
-\subsection{Decoupled taxonomic and functional
-trajectories}\label{decoupled-taxonomic-and-functional-trajectories}
+\subsection{Decoupled taxonomic and functional composition
+trajectories}\label{decoupled-taxonomic-and-functional-composition-trajectories}
 
@@
 
-\subsection{A new perspective on the intermediate disturbance
-hypothesis}\label{a-new-perspective-on-the-intermediate-disturbance-hypothesis}
+\subsection{The scope of the intermediate disturbance
+hypothesis}\label{the-scope-of-the-intermediate-disturbance-hypothesis}
 
@@
 for which the taxonomic richness was maximized and the taxonomic
-evenness was resilient. The disturbance intensity determined the balance
-in the community between pre-disturbance surviving trees, and trees
-recruited afterward. Below a 20\%-25\% AGB loss, the trees surviving
-after disturbance remained numerous enough to maintain the high
-taxonomic richness of the pre-disturbance community \citep{Bongers2009}.
-The recruitment of pionners, infrequent or absent before disturbance,
-then increased the taxonomic richness all the more so that the
-disturbance was intense \citep{Martin2015, Chaudhary2016}. As these
-pioneers became more dominant they balanced the usual hyper-dominance of
-tropical forests, hence temporarily increasing the taxonomic evenness
-\citep{Baraloto2012a}. Above the intensity threshold, the taxonomic
-richness of surviving trees was too low to be offset by the recruitment
-of pioneers. In the Guiana shield indeed, the pool of true pioneers
-specifically recruited after disturbance is restricted to a few common
-genera (e.g. \emph{Cecropia} spp., \emph{Vismia} spp.)
+evenness remained resilient. The disturbance intensity determined the
+balance in the community between pre-disturbance surviving trees, and
+trees recruited afterward. Below a 20\%-25\% AGB loss, the trees
+surviving after disturbance remained numerous enough to maintain the
+high taxonomic richness of the pre-disturbance community
+\citep{Bongers2009}. The recruitment of pionners, infrequent or absent
+before disturbance, then increased the taxonomic richness all the more
+so that the disturbance was intense \citep{Martin2015, Chaudhary2016}.
+As these pioneers became more dominant they balanced the usual
+hyper-dominance of tropical forests, hence temporarily increasing the
+taxonomic evenness \citep{Baraloto2012a}. Above the intensity threshold,
+the taxonomic richness of surviving trees was too low to be offset by
+the recruitment of pioneers. In the Guiana shield indeed, the pool of
+true pioneers specifically recruited after disturbance is restricted to
+a few common genera (e.g. \emph{Cecropia} spp., \emph{Vismia} spp.)
 \citep{Guitet2018}. The taxonomic richness then decreased following
@@
 Taxonomic trajectories following disturbance of intermediate intensity
-besides highlighted an intermediate time after disturbance for which the
-taxonomic evenness was maximized. As already observed in the Guiana
-shield \citep{Baraloto2012a} and in Borenan tropical forests
-\citep{Cannon1998}, taxonomic evenness followed hump-shaped trajectories
-with a peak when the recruitment was balanced between pioneers and
-late-successional species. This time then corresponded to the match
-between the generation time of pionneers, first to dominate the
-post-disturbance recruitment, and the time for competitive exclusion to
-emerge, which leads to the dominance of late-successional species.
+besides highlighted an intermediate time after disturbance, around 15-20
+years, for which the taxonomic evenness was maximized. As already
+observed in the Guiana shield \citep{Baraloto2012a} and in Borenan
+tropical forests \citep{Cannon1998}, taxonomic evenness followed
+hump-shaped trajectories with a peak when the recruitment was balanced
+between pioneers and late-successional species. This intermediate time
+corresponded to the match between the generation time of pionneers,
+first to dominate the post-disturbance recruitment, and the time for
+competitive exclusion to emerge, which leads to the dominance of
+late-successional species.
 

### 2019-02-13 WholePlotTrajectories.Rmd
intro

@@
 The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
-In tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
-The IDH may result from several processes and in the highly diverse tree communities its pattern might differ from those recognized otherwise.
-Most of the studies rely on a single diversity metric, often the taxonomic richness [@Molino2001], which is limited to grasp all facets of the diversity-disturbance relationship.
+In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
+The scope of the IDH should then be clarified regarding the different facets of community richness, abundance distribution and functioning.
+The underlying processes of the IDH are besides part of the post-disturbance succession and the question is on the time after disturbance validating the hypothesis.
 In this study we aimed as well to clarify the scope of the IDH and discuss its implication for the fate of tropical forests [@Sheil2003;@Shea2004].
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, (i) from post-disturbance trajectories we examined community taxonomic and functional response, (ii) then we discussed the scope of the IDH in taxonomic and functional facets and in time after disturbance, and (iii) we analyzed community resilience and time to recovery.
+Specifically, (i) from post-disturbance trajectories we examined community taxonomic and functional response, (ii) then we discussed the scope of the IDH in its aspect of taxonomic and functional diversity and time after disturbance, and (iii) we analyzed community resilience and time to recovery.
 

### 2019-02-13 WholePlotTrajectories.tex
intro

@@
 both competitively superior species and fast colonizers
-\citep{Shea2004, Pulsford2016}. In tropical forests, however,
-observations of the IDH are often blurry compared to theoretical
-expectations \citep{Hugues2007, Sheil2003, Norden2017}, and the
-validation of the theory remains controversial
-\citep{Hubbell2001, Fox2013, Sheil2013}. The IDH may result from several
-processes and in the highly diverse tree communities its pattern might
-differ from those recognized otherwise. Most of the studies rely on a
-single diversity metric, often the taxonomic richness
-\citep{Molino2001}, which is limited to grasp all facets of the
-diversity-disturbance relationship. In this study we aimed as well to
-clarify the scope of the IDH and discuss its implication for the fate of
-tropical forests \citep{Sheil2003, Shea2004}.
+\citep{Shea2004, Pulsford2016}. In highly diverse tropical forests,
+however, observations of the IDH are often blurry compared to
+theoretical expectations \citep{Hugues2007, Sheil2003, Norden2017}, and
+the validation of the theory remains controversial
+\citep{Hubbell2001, Fox2013, Sheil2013}. The scope of the IDH should
+then be clarified regarding the different facets of community richness,
+abundance distribution and functioning. The underlying processes of the
+IDH are besides part of the post-disturbance succession and the question
+is on the time after disturbance validating the hypothesis. In this
+study we aimed as well to clarify the scope of the IDH and discuss its
+implication for the fate of tropical forests
+\citep{Sheil2003, Shea2004}.
 
@@
 community taxonomic and functional response, (ii) then we discussed the
-scope of the IDH in taxonomic and functional facets and in time after
-disturbance, and (iii) we analyzed community resilience and time to
-recovery.
+scope of the IDH in its aspect of taxonomic and functional diversity and
+time after disturbance, and (iii) we analyzed community resilience and
+time to recovery.
 

### 2019-02-14 WholePlotTrajectories.Rmd
files

@@
 These threats affect the natural disturbance regime that shape and maintain the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
-It is now urgent to understand the processes underlying community response to disturbance, and thus to be able to anticipate the fate of tropical forests in the current context.
-In tropical forests, disturbance maintains a forest environment and initiates a suit of changes in the structure and functioning of communities consituting a secondary succession.
+It is urgent to understand community response to disturbance and the processes underlying to anticipate the fate of tropical forests in the current context.
 Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
-For now, post-disturbance trajectories has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
+In tropical forests, the forest environment is maintained after disturbance and there is a suit of changes in the structure and functioning of communities that consitutes a secondary succession.
+For now, this succession has been largely studied through post-disturbance trajectories of forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016].
 Models based on observed trajectories then assessed community time to recovery and identified some of the determinants of post-disturbance trajectories, like forest structure, forest composition, environmental parameters and disturbance characteristics [@Herault2018].
 Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
-Given the variety of species response to disturbance and the huge diversity of tropical forests, biodiversity trajectories following disturbance might be very diverse [@Lindenmayer2012;@Garcia_florez2017].
-We assumed that the rate, endpoint and duration of the trajectories depend on the intensity and frequency of disturbance: we then aimed to highlight the link between biodiversity and disturbance and thus anticipate community post-disturbance trajectories [@Chazdon2003a;@Willig2018].
+Given the variety of species response to disturbance and the huge diversity of tropical forests, post-disturbance biodiversity trajectories might be very diverse [@Lindenmayer2012;@Garcia_florez2017].
+It is essential to anticipate these trajectories, and hence to highlight the link between biodiversity and disturbance, specifically how disturbance intensity and frequency determine the rate, endpoint and duration of the trajectories [@Chazdon2003a;@Willig2018].
 
@@
 In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
-The scope of the IDH should then be clarified regarding the different facets of community richness, abundance distribution and functioning.
+Given the huge diversity of tropical forests and the variety of ecological processes that may underlie the IDH The scope of the IDH need be clarified regarding the different facets of community richness, abundance distribution and functioning.
 The underlying processes of the IDH are besides part of the post-disturbance succession and the question is on the time after disturbance validating the hypothesis.

### 2019-02-14 WholePlotTrajectories.Rmd
intro

@@
 The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
-In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
-Given the huge diversity of tropical forests and the variety of ecological processes that may underlie the IDH The scope of the IDH need be clarified regarding the different facets of community richness, abundance distribution and functioning.
-The underlying processes of the IDH are besides part of the post-disturbance succession and the question is on the time after disturbance validating the hypothesis.
-In this study we aimed as well to clarify the scope of the IDH and discuss its implication for the fate of tropical forests [@Sheil2003;@Shea2004].
+In tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
+The link between biodiversity and disturbance might be complicated by the huge diversity of tropical communities and by the variety of ecological processes that may underlie the IDH (colonisation, facilitation, coexistence,etc)[@Lindenmayer2012;@Garcia_florez2017].
+The scope of the IDH need be clarified regarding the different facets of community diversity, like taxonomic richness, abundance distribution and functioning.
+Besides, the question is highlighting the moment when emerge the conditions for the IDH, which would be an intermediate time after disturbance. 
+In this study we then aimed to clarify the scope of the IDH in time and regarding the different facets of community diversity [@Sheil2003;@Shea2004].
 

### 2019-02-14 WholePlotTrajectories.tex
intro

@@
 and functioning of tree communities
-\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. It is now urgent
-to understand the processes underlying community response to
-disturbance, and thus to be able to anticipate the fate of tropical
-forests in the current context. In tropical forests, disturbance
-maintains a forest environment and initiates a suit of changes in the
-structure and functioning of communities consituting a secondary
-succession. Disturbance change both community abiotic and biotic
+\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. It is urgent to
+understand community response to disturbance and the processes
+underlying to anticipate the fate of tropical forests in the current
+context. Disturbance change both community abiotic and biotic
 environments, through modifications in the fluxes of light, heat and
-water \citep{Goulamoussene2017}. For now, post-disturbance trajectories
-has been largely studied through forest structural parameters such as
-aboveground biomass, tree height or stem density
+water \citep{Goulamoussene2017}. In tropical forests, the forest
+environment is maintained after disturbance and there is a suit of
+changes in the structure and functioning of communities that consitutes
+a secondary succession. For now, this succession has been largely
+studied through post-disturbance trajectories of forest structural
+parameters such as aboveground biomass, tree height or stem density
 \citep{Piponiot2016, Rutishauser2016}. Models based on observed
@@
 of species response to disturbance and the huge diversity of tropical
-forests, biodiversity trajectories following disturbance might be very
-diverse \citep{Lindenmayer2012, Garcia_florez2017}. We assumed that the
-rate, endpoint and duration of the trajectories depend on the intensity
-and frequency of disturbance: we then aimed to highlight the link
-between biodiversity and disturbance and thus anticipate community
-post-disturbance trajectories \citep{Chazdon2003a, Willig2018}.
+forests, post-disturbance biodiversity trajectories might be very
+diverse \citep{Lindenmayer2012, Garcia_florez2017}. It is essential to
+anticipate these trajectories, and hence to highlight the link between
+biodiversity and disturbance, specifically how disturbance intensity and
+frequency determine the rate, endpoint and duration of the trajectories
+\citep{Chazdon2003a, Willig2018}.
 
@@
 both competitively superior species and fast colonizers
-\citep{Shea2004, Pulsford2016}. In highly diverse tropical forests,
-however, observations of the IDH are often blurry compared to
-theoretical expectations \citep{Hugues2007, Sheil2003, Norden2017}, and
-the validation of the theory remains controversial
-\citep{Hubbell2001, Fox2013, Sheil2013}. The scope of the IDH should
-then be clarified regarding the different facets of community richness,
-abundance distribution and functioning. The underlying processes of the
-IDH are besides part of the post-disturbance succession and the question
-is on the time after disturbance validating the hypothesis. In this
-study we aimed as well to clarify the scope of the IDH and discuss its
-implication for the fate of tropical forests
-\citep{Sheil2003, Shea2004}.
+\citep{Shea2004, Pulsford2016}. In tropical forests, however,
+observations of the IDH are often blurry compared to theoretical
+expectations \citep{Hugues2007, Sheil2003, Norden2017}, and the
+validation of the theory remains controversial
+\citep{Hubbell2001, Fox2013, Sheil2013}. The link between biodiversity
+and disturbance might be complicated by the huge diversity of tropical
+communities and by the variety of ecological processes that may underlie
+the IDH (colonisation, facilitation,
+coexistence,etc)\citep{Lindenmayer2012, Garcia_florez2017}. The scope of
+the IDH need be clarified regarding the different facets of community
+diversity, like taxonomic richness, abundance distribution and
+functioning. Besides, the question is highlighting the moment when
+emerge the conditions for the IDH, which would be an intermediate time
+after disturbance. In this study we then aimed to clarify the scope of
+the IDH in time and regarding the different facets of community
+diversity \citep{Sheil2003, Shea2004}.
 

### 2019-02-14 WholePlotTrajectories.Rmd
conclusions

@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, (i) from post-disturbance trajectories we examined community taxonomic and functional response, (ii) then we discussed the scope of the IDH in its aspect of taxonomic and functional diversity and time after disturbance, and (iii) we analyzed community resilience and time to recovery.
+Specifically, (i) we defined taxonomic and functional post-disturbance trajectories and examined the underlying ecological process, (ii) we discussed the scope of the IDH regarding the different facets of community diversity and regarding time after disturbance, and (iii) we analyzed community resilience and time to recovery.
 
@@
 
-# Conclusions
+# Conclusion
 
-Our study highlighted the combination of deterministic and stochastic processes determining the succession of recruited species and shaping community response to disturbance. 
-The deterministic recruitment of pioneers shaped functional trajectories convergent in the functional space, while dispersal limitations shaped stochastic taxonomic trajectories maintaining community divergence in taxonomic composition.
-In accordance with the IDH, community taxonomic diversity was maximised for intermediate disturbance intensity, before reaching an intensity threshold beyond which the taxonomic richness decreased and the taxonomic evenness remained persistently higher.
-Conversely, functional trajectories proved decoupled from the taxonomic ones and did not follow the hypothesis of the IDH as no distinct trajectories emerged with increasing disturbance intensity. 
-The decoupling between taxonomic and functional disturbance was explained by a decrease of community functional redundancy mitigating the functoinal impact of disturbance.
-The resilience of tropical forests, defined in terms of recovery to pre-disturbance state, proved tangible but requiring several decades. 
-Still, the disturbance impact on community redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].
+Community compositon and diversity post-disturbance trajectories were shaped by the recruitment of pioneers, identical for all communities and disturbance intensity.
+However, taxonomic and functional composition trajectories were decoupled: while functional trajectories were similar in the functional space and converged towards the recovery of the same initial state, taxonomic trajectories displayed initial composition differences maintained along time.
+This decoupling was explained by a decrease of the functional redundancy which mitigated the functional impact of disturbance.
+The disturbance impact was contrasted as well between functoinal and taxonomic diversity.
+While the functional trajectories remained similar whenever the disturbance intensity, taxonomic trajectories were markedly different above a threshold of 20-25% AGB removed for which taxonomic richness was maximized and taxonomic eveness remained resilient.
+The Intermediate Disturbance Hypothesis was then validated regarding community taxonomic response but inferred regarding the functional one.
+Taxonomic diversity was maximized following a disturbance of intermediate intensity, that was the threshold of 25% AGB removed, and at an intermediate time after disturbance when the recruitment of early- and late-successional species were balanced, that was around 25 years after disturbance.
+Whenever the disturbance intensity, community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied on the lottery, random recruitment of rare species.
+Although resilient, the functional redundancy was lower for more than 30 years after disturbance which cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].
 

### 2019-02-14 WholePlotTrajectories.tex
conclusions

@@
 trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, (i) from post-disturbance trajectories we examined
-community taxonomic and functional response, (ii) then we discussed the
-scope of the IDH in its aspect of taxonomic and functional diversity and
-time after disturbance, and (iii) we analyzed community resilience and
-time to recovery.
+Specifically, (i) we defined taxonomic and functional post-disturbance
+trajectories and examined the underlying ecological process, (ii) we
+discussed the scope of the IDH regarding the different facets of
+community diversity and regarding time after disturbance, and (iii) we
+analyzed community resilience and time to recovery.
 
@@
 
-\section{Conclusions}\label{conclusions}
+\section{Conclusion}\label{conclusion}
 
-Our study highlighted the combination of deterministic and stochastic
-processes determining the succession of recruited species and shaping
-community response to disturbance. The deterministic recruitment of
-pioneers shaped functional trajectories convergent in the functional
-space, while dispersal limitations shaped stochastic taxonomic
-trajectories maintaining community divergence in taxonomic composition.
-In accordance with the IDH, community taxonomic diversity was maximised
-for intermediate disturbance intensity, before reaching an intensity
-threshold beyond which the taxonomic richness decreased and the
-taxonomic evenness remained persistently higher. Conversely, functional
-trajectories proved decoupled from the taxonomic ones and did not follow
-the hypothesis of the IDH as no distinct trajectories emerged with
-increasing disturbance intensity. The decoupling between taxonomic and
-functional disturbance was explained by a decrease of community
-functional redundancy mitigating the functoinal impact of disturbance.
-The resilience of tropical forests, defined in terms of recovery to
-pre-disturbance state, proved tangible but requiring several decades.
-Still, the disturbance impact on community redundancy cautioned against
-the risks of infrequent species loss and the persistence of
-disturbance-specific communities \citep{Herault2018}.
+Community compositon and diversity post-disturbance trajectories were
+shaped by the recruitment of pioneers, identical for all communities and
+disturbance intensity. However, taxonomic and functional composition
+trajectories were decoupled: while functional trajectories were similar
+in the functional space and converged towards the recovery of the same
+initial state, taxonomic trajectories displayed initial composition
+differences maintained along time. This decoupling was explained by a
+decrease of the functional redundancy which mitigated the functional
+impact of disturbance. The disturbance impact was contrasted as well
+between functoinal and taxonomic diversity. While the functional
+trajectories remained similar whenever the disturbance intensity,
+taxonomic trajectories were markedly different above a threshold of
+20-25\% AGB removed for which taxonomic richness was maximized and
+taxonomic eveness remained resilient. The Intermediate Disturbance
+Hypothesis was then validated regarding community taxonomic response but
+inferred regarding the functional one. Taxonomic diversity was maximized
+following a disturbance of intermediate intensity, that was the
+threshold of 25\% AGB removed, and at an intermediate time after
+disturbance when the recruitment of early- and late-successional species
+were balanced, that was around 25 years after disturbance. Whenever the
+disturbance intensity, community resilience (in terms of recovery of the
+pre-disturbance state) was tangible but required several decades and
+relied on the lottery, random recruitment of rare species. Although
+resilient, the functional redundancy was lower for more than 30 years
+after disturbance which cautioned against the risks of infrequent
+species loss and the persistence of disturbance-specific communities
+\citep{Herault2018}.
 

### 2019-02-15 WholePlotTrajectories.Rmd
abstract

@@
 abstract: 
- Understanding the ecological rules underlying the maintenance of tropical forests biodiversity, structure, functioning and dynamics is urgent to anticipate their fate in the global change context.
- The huge diversity of tropical forests is often assumed to be regularly reshaped by natural disturbance yielding a diversity peak at intermediate intensity.
-This intermediate disturbance hypothesis (IDH), though, remains debated and the resilience of community taxonomic and functional diversity and composition remains controversial.
- To disentangle the ecological processes driving community response to disturbance, we analyzed tree community trajectories following a disturbance gradient in a Neotropical forest.
-We examined community diversity, composition and redundancy trajectories over 30 years, considering both taxonomic and functional trajectories based on 7 leaf, stem and life-history traits.
-We highlighted the cyclic recovery of community taxonomic and functional composition.
-While pre-disturbance taxonomic differences were maintained over time, functional composition trajectories were quite similar among communities.
-The IDH did predict community taxonomic diversity response while functional diversity was enhanced whatever the disturbance intensity.
+ Understand biodiversity trajectories of tropical forests following disturbance and clarify the link between biodiversity and sitruabnce is urgent to anticipate the impact of the global change context.
+ The taxonomic and functional facets of tree communities trajectories following disturbance and the underlying processes remain to be clarified.
+The explicit impact of disturbance intensity, often formalized with the Intermediate Disturbance Hypothesis (IDH), is controversial as well in tropical forests.
+ To clarify post-disturbance trajectories and examine the scope of the IDH in tropical forests, we analyzed tree community response to a disturbance gradient in a Neotropical forest.
+We examined community diversity, composition and redundancy trajectories over 30 years.
+We considered taxonomic and functional trajectories, based on 7 leaf, stem and life-history traits.
+We highlighted a decoupling between taxonomic trajectories, different between communities, and functional trajectories that were identical, which we explained by the variations in community functional redundancy.
+We besides clarified the scope of the IDH, which did not apply to the functional diversity but corroborated the hump-shaped taxonomic diversity pattern.
  Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years.
@@
 
-Community compositon and diversity post-disturbance trajectories were shaped by the recruitment of pioneers, identical for all communities and disturbance intensity.
-However, taxonomic and functional composition trajectories were decoupled: while functional trajectories were similar in the functional space and converged towards the recovery of the same initial state, taxonomic trajectories displayed initial composition differences maintained along time.
+Community post-disturbance trajectories in compositon and diversity were shaped by the recruitment of a determined pool of pioneers identical among communities and disturbance intensity.
+Nevertheless, taxonomic and functional composition trajectories were decoupled: while functional trajectories remained similar in the functional space and converged towards the recovery of a comparable initial state, taxonomic trajectories shiwed initial differences in community composition that were maintained along time.
 This decoupling was explained by a decrease of the functional redundancy which mitigated the functional impact of disturbance.
-The disturbance impact was contrasted as well between functoinal and taxonomic diversity.
+Community functional and taxonomic diversity response to disturbance was contrasted as well.
 While the functional trajectories remained similar whenever the disturbance intensity, taxonomic trajectories were markedly different above a threshold of 20-25% AGB removed for which taxonomic richness was maximized and taxonomic eveness remained resilient.
-The Intermediate Disturbance Hypothesis was then validated regarding community taxonomic response but inferred regarding the functional one.
-Taxonomic diversity was maximized following a disturbance of intermediate intensity, that was the threshold of 25% AGB removed, and at an intermediate time after disturbance when the recruitment of early- and late-successional species were balanced, that was around 25 years after disturbance.
-Whenever the disturbance intensity, community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied on the lottery, random recruitment of rare species.
+The Intermediate Disturbance Hypothesis applied to the linkage of disturbance with the taxonomic diversity, but not with the functional diversity.
+Taxonomic diversity was maximized following a disturbance of intermediate intensity, that was the threshold of 25% AGB removed, and at an intermediate time after disturbance, that was around 25 years after disturbance when the recruitment of early- and late-successional species were balanced.
+Whenever the disturbance intensity, community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied upon the random, lottery recruitment of rare species.
 Although resilient, the functional redundancy was lower for more than 30 years after disturbance which cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].

### 2019-02-15 WholePlotTrajectories.tex
abstract

@@
 \Abstract{
-Understanding the ecological rules underlying the maintenance of
-tropical forests biodiversity, structure, functioning and dynamics is
-urgent to anticipate their fate in the global change context. The huge
-diversity of tropical forests is often assumed to be regularly reshaped
-by natural disturbance yielding a diversity peak at intermediate
-intensity. This intermediate disturbance hypothesis (IDH), though,
-remains debated and the resilience of community taxonomic and functional
-diversity and composition remains controversial. To disentangle the
-ecological processes driving community response to disturbance, we
-analyzed tree community trajectories following a disturbance gradient in
-a Neotropical forest. We examined community diversity, composition and
-redundancy trajectories over 30 years, considering both taxonomic and
-functional trajectories based on 7 leaf, stem and life-history traits.
-We highlighted the cyclic recovery of community taxonomic and functional
-composition. While pre-disturbance taxonomic differences were maintained
-over time, functional composition trajectories were quite similar among
-communities. The IDH did predict community taxonomic diversity response
-while functional diversity was enhanced whatever the disturbance
-intensity. Although consistent, the recovery of community composition,
-diversity and redundancy remained unachieved after 30 years. This
-acknowledged the need of decades-long cycles with no disturbance to
-ensure a complete recovery, and questioned tropical forest community
-resilience after repeated disturbances.
+Understand biodiversity trajectories of tropical forests following
+disturbance and clarify the link between biodiversity and sitruabnce is
+urgent to anticipate the impact of the global change context. The
+taxonomic and functional facets of tree communities trajectories
+following disturbance and the underlying processes remain to be
+clarified. The explicit impact of disturbance intensity, often
+formalized with the Intermediate Disturbance Hypothesis (IDH), is
+controversial as well in tropical forests. To clarify post-disturbance
+trajectories and examine the scope of the IDH in tropical forests, we
+analyzed tree community response to a disturbance gradient in a
+Neotropical forest. We examined community diversity, composition and
+redundancy trajectories over 30 years. We considered taxonomic and
+functional trajectories, based on 7 leaf, stem and life-history traits.
+We highlighted a decoupling between taxonomic trajectories, different
+between communities, and functional trajectories that were identical,
+which we explained by the variations in community functional redundancy.
+We besides clarified the scope of the IDH, which did not apply to the
+functional diversity but corroborated the hump-shaped taxonomic
+diversity pattern. Although consistent, the recovery of community
+composition, diversity and redundancy remained unachieved after 30
+years. This acknowledged the need of decades-long cycles with no
+disturbance to ensure a complete recovery, and questioned tropical
+forest community resilience after repeated disturbances.
 }
@@
 
-Community compositon and diversity post-disturbance trajectories were
-shaped by the recruitment of pioneers, identical for all communities and
-disturbance intensity. However, taxonomic and functional composition
-trajectories were decoupled: while functional trajectories were similar
-in the functional space and converged towards the recovery of the same
-initial state, taxonomic trajectories displayed initial composition
-differences maintained along time. This decoupling was explained by a
-decrease of the functional redundancy which mitigated the functional
-impact of disturbance. The disturbance impact was contrasted as well
-between functoinal and taxonomic diversity. While the functional
-trajectories remained similar whenever the disturbance intensity,
-taxonomic trajectories were markedly different above a threshold of
-20-25\% AGB removed for which taxonomic richness was maximized and
-taxonomic eveness remained resilient. The Intermediate Disturbance
-Hypothesis was then validated regarding community taxonomic response but
-inferred regarding the functional one. Taxonomic diversity was maximized
-following a disturbance of intermediate intensity, that was the
-threshold of 25\% AGB removed, and at an intermediate time after
-disturbance when the recruitment of early- and late-successional species
-were balanced, that was around 25 years after disturbance. Whenever the
-disturbance intensity, community resilience (in terms of recovery of the
+Community post-disturbance trajectories in compositon and diversity were
+shaped by the recruitment of a determined pool of pioneers identical
+among communities and disturbance intensity. Nevertheless, taxonomic and
+functional composition trajectories were decoupled: while functional
+trajectories remained similar in the functional space and converged
+towards the recovery of a comparable initial state, taxonomic
+trajectories shiwed initial differences in community composition that
+were maintained along time. This decoupling was explained by a decrease
+of the functional redundancy which mitigated the functional impact of
+disturbance. Community functional and taxonomic diversity response to
+disturbance was contrasted as well. While the functional trajectories
+remained similar whenever the disturbance intensity, taxonomic
+trajectories were markedly different above a threshold of 20-25\% AGB
+removed for which taxonomic richness was maximized and taxonomic eveness
+remained resilient. The Intermediate Disturbance Hypothesis applied to
+the linkage of disturbance with the taxonomic diversity, but not with
+the functional diversity. Taxonomic diversity was maximized following a
+disturbance of intermediate intensity, that was the threshold of 25\%
+AGB removed, and at an intermediate time after disturbance, that was
+around 25 years after disturbance when the recruitment of early- and
+late-successional species were balanced. Whenever the disturbance
+intensity, community resilience (in terms of recovery of the
 pre-disturbance state) was tangible but required several decades and
-relied on the lottery, random recruitment of rare species. Although
+relied upon the random, lottery recruitment of rare species. Although
 resilient, the functional redundancy was lower for more than 30 years

### 2019-02-19 WholePlotTrajectories.Rmd
all

@@
 abstract: 
- Understand biodiversity trajectories of tropical forests following disturbance and clarify the link between biodiversity and sitruabnce is urgent to anticipate the impact of the global change context.
- The taxonomic and functional facets of tree communities trajectories following disturbance and the underlying processes remain to be clarified.
-The explicit impact of disturbance intensity, often formalized with the Intermediate Disturbance Hypothesis (IDH), is controversial as well in tropical forests.
- To clarify post-disturbance trajectories and examine the scope of the IDH in tropical forests, we analyzed tree community response to a disturbance gradient in a Neotropical forest.
-We examined community diversity, composition and redundancy trajectories over 30 years.
-We considered taxonomic and functional trajectories, based on 7 leaf, stem and life-history traits.
-We highlighted a decoupling between taxonomic trajectories, different between communities, and functional trajectories that were identical, which we explained by the variations in community functional redundancy.
-We besides clarified the scope of the IDH, which did not apply to the functional diversity but corroborated the hump-shaped taxonomic diversity pattern.
+ Understand biodiversity trajectories of tropical forests following disturbance and clarify the link between biodiversity and disturbance is urgent to anticipate the impact of the global change context.
+ It is urgent to understand the taxonomic and functional facets of tree community trajectories and the underlying processes, and to clarify the link between biodiversity and disturbance, specifically the scope of the Intermediate Disturbance Hypothesis (IDH) that is debated tropical forests.
+ In this study, we analyzed tree community response to a disturbance gradient in a Neotropical forest through diversity, composition and redundancy post-disturbance trajectories over 30 years.
+We considered taxonomic trajectories and functional trajectories based on 7 leaf, stem and life-history traits.
+While taxonomic trajectories differed among communities, functional trajectories remained similar and convergent, and this decoupling was explained by the variations in community functional redundancy mitigating the functional impact of disturbance.
+We validated the IDH, based on an intensity threshold of 20-25% of AGB removed, and we clarified its scope, showing that it applied to taxonomic but not functional diversity, and applied around 20 years after disturbance when the recruitment of pioneers and late-successionals were balanced.
  Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years.
-This acknowledged the need of decades-long cycles with no disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbances.
+This acknowledged the need of decades-long cycles without disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbances.
 keywords: [Community Ecology, Community Diversity Determinants, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
@@
 These threats affect the natural disturbance regime that shape and maintain the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
-It is urgent to understand community response to disturbance and the processes underlying to anticipate the fate of tropical forests in the current context.
+It is urgent to understand community response to disturbance and the underlying processes to anticipate the fate of tropical forests in the current context.
 Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
@@
 Given the variety of species response to disturbance and the huge diversity of tropical forests, post-disturbance biodiversity trajectories might be very diverse [@Lindenmayer2012;@Garcia_florez2017].
-It is essential to anticipate these trajectories, and hence to highlight the link between biodiversity and disturbance, specifically how disturbance intensity and frequency determine the rate, endpoint and duration of the trajectories [@Chazdon2003a;@Willig2018].
+It is essential to anticipate these trajectories and highlight the link between biodiversity and disturbance, and specifically how disturbance intensity and frequency determine the rate, endpoint and duration of post-disturbance trajectories [@Chazdon2003a;@Willig2018].
 
@@
 In tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
-The link between biodiversity and disturbance might be complicated by the huge diversity of tropical communities and by the variety of ecological processes that may underlie the IDH (colonisation, facilitation, coexistence,etc)[@Lindenmayer2012;@Garcia_florez2017].
-The scope of the IDH need be clarified regarding the different facets of community diversity, like taxonomic richness, abundance distribution and functioning.
-Besides, the question is highlighting the moment when emerge the conditions for the IDH, which would be an intermediate time after disturbance. 
-In this study we then aimed to clarify the scope of the IDH in time and regarding the different facets of community diversity [@Sheil2003;@Shea2004].
+The link between biodiversity and disturbance might be complicated by the huge diversity of tropical communities and a variety of ecological processes may underlie the IDH (colonisation, facilitation, coexistence,etc)[@Lindenmayer2012;@Garcia_florez2017].
+There is a need to clarify the scope of the IDH in terms of the different facets of community diversity, that are the taxonomic richness, the species abundance distribution and the community functioning.
+Besides, there is a need to delineate the scope of the IDH in time, and highlight the moment when the conditions for the IDH emerge [@Sheil2003;@Shea2004].
 
-The analyse of community response to disturbance requires a set of metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
-The analysis should first consider community composition, which is crucial for conservation issues [@Lavorel2002;@Bellwood2006] and assesses the different species fostered by the disturbance.
-Diversity metrics encompassing taxonomic richness and evenness would assess the changes community abundance distribution. 
-A functional approach accounting for species biological attributes and ecological strategy, would directly link community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
+Analysing community response to disturbance requires a set of metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
+The analysis should first consider community composition, which is crucial for conservation issues and reveals the species fostered by the disturbance [@Lavorel2002;@Bellwood2006].
+Then, diversity metrics encompassing taxonomic richness and evenness should be considered to assess the changes in community abundance distribution. 
+A functional approach, that account for species biological attributes and ecological strategy, would directly link community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
 In that respect, a vast literature allowed recognizing major traits that represent species ecological strategy and determine how they respond to changing conditions [@Diaz2005]. 
@@
 The functional approach also encompasses the analysis of functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016].
-The high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines the resilience of communities after disturbance [@Elmqvist2003; @Diaz2005].
+The typical high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines communities resilience after disturbance [@Elmqvist2003; @Diaz2005].
 
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, (i) we defined taxonomic and functional post-disturbance trajectories and examined the underlying ecological process, (ii) we discussed the scope of the IDH regarding the different facets of community diversity and regarding time after disturbance, and (iii) we analyzed community resilience and time to recovery.
+Specifically, (i) we defined taxonomic and functional post-disturbance trajectories and examined the underlying ecological process, (ii) we discussed the scope of the IDH regarding the different facets of community diversity and the time after disturbance, and (iii) we analyzed community resilience and time to recovery.
 
@@
 
-## Decoupled taxonomic and functional composition trajectories
+## Decoupled taxonomic and functional trajectories
 
@@
 
-Both taxonomic and functional composition trajectories initiated a return towards pre-disturbance state after 30 years, which highlighted community taxonomic and functional resilience. 
-There was however a decoupling between taxonomic and functional composition trajectories as taxonomic trajectories maintained the initial differences among communities, while functional trajectories where similar and convergent in the functional space [@Fukami2005].
+Both taxonomic and functional composition trajectories initiated a return towards pre-disturbance state after 30 years, which highlighted the taxonomic and functional resilience of communities. 
+There was however a decoupling between taxonomic and functional composition trajectories: while as taxonomic trajectories maintained the initial differences among communities, the functional trajectories where similar and convergent in the functional space [@Fukami2005].
 

### 2019-02-19 WholePlotTrajectories.tex
all

@@
 Understand biodiversity trajectories of tropical forests following
-disturbance and clarify the link between biodiversity and sitruabnce is
-urgent to anticipate the impact of the global change context. The
-taxonomic and functional facets of tree communities trajectories
-following disturbance and the underlying processes remain to be
-clarified. The explicit impact of disturbance intensity, often
-formalized with the Intermediate Disturbance Hypothesis (IDH), is
-controversial as well in tropical forests. To clarify post-disturbance
-trajectories and examine the scope of the IDH in tropical forests, we
-analyzed tree community response to a disturbance gradient in a
-Neotropical forest. We examined community diversity, composition and
-redundancy trajectories over 30 years. We considered taxonomic and
-functional trajectories, based on 7 leaf, stem and life-history traits.
-We highlighted a decoupling between taxonomic trajectories, different
-between communities, and functional trajectories that were identical,
-which we explained by the variations in community functional redundancy.
-We besides clarified the scope of the IDH, which did not apply to the
-functional diversity but corroborated the hump-shaped taxonomic
-diversity pattern. Although consistent, the recovery of community
-composition, diversity and redundancy remained unachieved after 30
-years. This acknowledged the need of decades-long cycles with no
-disturbance to ensure a complete recovery, and questioned tropical
-forest community resilience after repeated disturbances.
+disturbance and clarify the link between biodiversity and disturbance is
+urgent to anticipate the impact of the global change context. It is
+urgent to understand the taxonomic and functional facets of tree
+community trajectories and the underlying processes, and to clarify the
+link between biodiversity and disturbance, specifically the scope of the
+Intermediate Disturbance Hypothesis (IDH) that is debated tropical
+forests. In this study, we analyzed tree community response to a
+disturbance gradient in a Neotropical forest through diversity,
+composition and redundancy post-disturbance trajectories over 30 years.
+We considered taxonomic trajectories and functional trajectories based
+on 7 leaf, stem and life-history traits. While taxonomic trajectories
+differed among communities, functional trajectories remained similar and
+convergent, and this decoupling was explained by the variations in
+community functional redundancy mitigating the functional impact of
+disturbance. We validated the IDH, based on an intensity threshold of
+20-25\% of AGB removed, and we clarified its scope, showing that it
+applied to taxonomic but not functional diversity, and applied around 20
+years after disturbance when the recruitment of pioneers and
+late-successionals were balanced. Although consistent, the recovery of
+community composition, diversity and redundancy remained unachieved
+after 30 years. This acknowledged the need of decades-long cycles
+without disturbance to ensure a complete recovery, and questioned
+tropical forest community resilience after repeated disturbances.
 }
@@
 \citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. It is urgent to
-understand community response to disturbance and the processes
-underlying to anticipate the fate of tropical forests in the current
+understand community response to disturbance and the underlying
+processes to anticipate the fate of tropical forests in the current
 context. Disturbance change both community abiotic and biotic
@@
 diverse \citep{Lindenmayer2012, Garcia_florez2017}. It is essential to
-anticipate these trajectories, and hence to highlight the link between
-biodiversity and disturbance, specifically how disturbance intensity and
-frequency determine the rate, endpoint and duration of the trajectories
-\citep{Chazdon2003a, Willig2018}.
+anticipate these trajectories and highlight the link between
+biodiversity and disturbance, and specifically how disturbance intensity
+and frequency determine the rate, endpoint and duration of
+post-disturbance trajectories \citep{Chazdon2003a, Willig2018}.
 
@@
 and disturbance might be complicated by the huge diversity of tropical
-communities and by the variety of ecological processes that may underlie
-the IDH (colonisation, facilitation,
-coexistence,etc)\citep{Lindenmayer2012, Garcia_florez2017}. The scope of
-the IDH need be clarified regarding the different facets of community
-diversity, like taxonomic richness, abundance distribution and
-functioning. Besides, the question is highlighting the moment when
-emerge the conditions for the IDH, which would be an intermediate time
-after disturbance. In this study we then aimed to clarify the scope of
-the IDH in time and regarding the different facets of community
-diversity \citep{Sheil2003, Shea2004}.
+communities and a variety of ecological processes may underlie the IDH
+(colonisation, facilitation,
+coexistence,etc)\citep{Lindenmayer2012, Garcia_florez2017}. There is a
+need to clarify the scope of the IDH in terms of the different facets of
+community diversity, that are the taxonomic richness, the species
+abundance distribution and the community functioning. Besides, there is
+a need to delineate the scope of the IDH in time, and highlight the
+moment when the conditions for the IDH emerge
+\citep{Sheil2003, Shea2004}.
 
-The analyse of community response to disturbance requires a set of
-metrics, to grasp all aspects of community changes
+Analysing community response to disturbance requires a set of metrics,
+to grasp all aspects of community changes
 \citep{Sheil2003, Shea2004, Mayfield2010}. The analysis should first
 consider community composition, which is crucial for conservation issues
-\citep{Lavorel2002, Bellwood2006} and assesses the different species
-fostered by the disturbance. Diversity metrics encompassing taxonomic
-richness and evenness would assess the changes community abundance
-distribution. A functional approach accounting for species biological
-attributes and ecological strategy, would directly link community
-diversity, composition and redundancy to ecosystem functioning and
-environmental constraints \citep{Violle2007b, Baraloto2012a}. In that
-respect, a vast literature allowed recognizing major traits that
-represent species ecological strategy and determine how they respond to
-changing conditions \citep{Diaz2005}. Specifically in tropical forests,
-the functional approach revealed the post-disturbance deterministic
-processes entailing a shift from a dominance of ``conservative''
-slow-growing species dealing with scarce resources, to a dominance of
-``acquisitive'' fast-growing species with rapid and efficient use of
-abundant resources \citep{Rees2001, Reich2014, Herault2011}. This shift
-is translated into the trajectories of average community value of key
-functional traits related to resource acquisition (leaf and stem traits)
-and life-history strategy (seed mass, maximum size)
+and reveals the species fostered by the disturbance
+\citep{Lavorel2002, Bellwood2006}. Then, diversity metrics encompassing
+taxonomic richness and evenness should be considered to assess the
+changes in community abundance distribution. A functional approach, that
+account for species biological attributes and ecological strategy, would
+directly link community diversity, composition and redundancy to
+ecosystem functioning and environmental constraints
+\citep{Violle2007b, Baraloto2012a}. In that respect, a vast literature
+allowed recognizing major traits that represent species ecological
+strategy and determine how they respond to changing conditions
+\citep{Diaz2005}. Specifically in tropical forests, the functional
+approach revealed the post-disturbance deterministic processes entailing
+a shift from a dominance of ``conservative'' slow-growing species
+dealing with scarce resources, to a dominance of ``acquisitive''
+fast-growing species with rapid and efficient use of abundant resources
+\citep{Rees2001, Reich2014, Herault2011}. This shift is translated into
+the trajectories of average community value of key functional traits
+related to resource acquisition (leaf and stem traits) and life-history
+strategy (seed mass, maximum size)
 \citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. The
@@
 redundancy, that quantifies the amount of shared trait values among
-species \citep{Carmona2016}. The high functional redundancy of
+species \citep{Carmona2016}. The typical high functional redundancy of
 hyper-diverse tropical forests \citep{Bellwood2006} mitigates the
-impacts of species removal on ecosystem functioning and determines the
-resilience of communities after disturbance
-\citep{Elmqvist2003, Diaz2005}.
+impacts of species removal on ecosystem functioning and determines
+communities resilience after disturbance \citep{Elmqvist2003, Diaz2005}.
 
@@
 discussed the scope of the IDH regarding the different facets of
-community diversity and regarding time after disturbance, and (iii) we
+community diversity and the time after disturbance, and (iii) we
 analyzed community resilience and time to recovery.
@@
 
-\subsection{Decoupled taxonomic and functional composition
-trajectories}\label{decoupled-taxonomic-and-functional-composition-trajectories}
+\subsection{Decoupled taxonomic and functional
+trajectories}\label{decoupled-taxonomic-and-functional-trajectories}
 
@@
 return towards pre-disturbance state after 30 years, which highlighted
-community taxonomic and functional resilience. There was however a
-decoupling between taxonomic and functional composition trajectories as
-taxonomic trajectories maintained the initial differences among
-communities, while functional trajectories where similar and convergent
-in the functional space \citep{Fukami2005}.
+the taxonomic and functional resilience of communities. There was
+however a decoupling between taxonomic and functional composition
+trajectories: while as taxonomic trajectories maintained the initial
+differences among communities, the functional trajectories where similar
+and convergent in the functional space \citep{Fukami2005}.
 

### 2019-02-19 WholePlotTrajectories.Rmd
all

@@
  Understand biodiversity trajectories of tropical forests following disturbance and clarify the link between biodiversity and disturbance is urgent to anticipate the impact of the global change context.
- It is urgent to understand the taxonomic and functional facets of tree community trajectories and the underlying processes, and to clarify the link between biodiversity and disturbance, specifically the scope of the Intermediate Disturbance Hypothesis (IDH) that is debated tropical forests.
+ This calls to highlight the taxonomic and functional facets of tree community trajectories and the underlying processes, and to clarify the link between biodiversity and disturbance, specifically the scope of the Intermediate Disturbance Hypothesis (IDH) that is debated tropical forests.
  In this study, we analyzed tree community response to a disturbance gradient in a Neotropical forest through diversity, composition and redundancy post-disturbance trajectories over 30 years.
@@
 This acknowledged the need of decades-long cycles without disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbances.
-keywords: [Community Ecology, Community Diversity Determinants, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
+keywords: [Community Ecology, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
 preamble: >

### 2019-02-19 WholePlotTrajectories.tex
all

@@
 
-\Keywords{Community Ecology, Community Diversity Determinants, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity} % Keywords - if you don't want any simply remove all the text between the curly brackets
+\Keywords{Community Ecology, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity} % Keywords - if you don't want any simply remove all the text between the curly brackets
 \newcommand{\keywordname}{Keywords} % Defines the keywords heading name
@@
 disturbance and clarify the link between biodiversity and disturbance is
-urgent to anticipate the impact of the global change context. It is
-urgent to understand the taxonomic and functional facets of tree
-community trajectories and the underlying processes, and to clarify the
-link between biodiversity and disturbance, specifically the scope of the
+urgent to anticipate the impact of the global change context. This calls
+to highlight the taxonomic and functional facets of tree community
+trajectories and the underlying processes, and to clarify the link
+between biodiversity and disturbance, specifically the scope of the
 Intermediate Disturbance Hypothesis (IDH) that is debated tropical

### 2019-03-24 WholePlotTrajectories.Rmd
Correction EM

@@
 abstract: 
- Understand biodiversity trajectories of tropical forests following disturbance and clarify the link between biodiversity and disturbance is urgent to anticipate the impact of the global change context.
- This calls to highlight the taxonomic and functional facets of tree community trajectories and the underlying processes, and to clarify the link between biodiversity and disturbance, specifically the scope of the Intermediate Disturbance Hypothesis (IDH) that is debated tropical forests.
- In this study, we analyzed tree community response to a disturbance gradient in a Neotropical forest through diversity, composition and redundancy post-disturbance trajectories over 30 years.
+ Understanding biodiversity trajectories of tropical forests following disturbance and clarifying the link between biodiversity and disturbance is urgent to anticipate the impact of the global change context.
+ This calls to highlight the taxonomic and functional facets of tree community trajectories and the underlying processes, and to clarify the link between biodiversity and disturbance, specifically the scope of the Intermediate Disturbance Hypothesis (IDH) that is debated in tropical forests.
+ In this study, we analyzed the tree community response to a disturbance gradient in a Neotropical forest through diversity, composition and redundancy post-disturbance trajectories over 30 years.
 We considered taxonomic trajectories and functional trajectories based on 7 leaf, stem and life-history traits.
@@
 Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
-In tropical forests, the forest environment is maintained after disturbance and there is a suit of changes in the structure and functioning of communities that consitutes a secondary succession.
+In tropical forests, the forest environment is maintained after disturbance and there is a succession of changes in the structure and functioning of communities that consitutes a secondary succession.
 For now, this succession has been largely studied through post-disturbance trajectories of forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016].
@@
 The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978].
-The theory and states that community diversity peaks at intermediate level of disturbance.
+The theory states that community diversity peaks at intermediate level of disturbance.
 The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
@@
 The link between biodiversity and disturbance might be complicated by the huge diversity of tropical communities and a variety of ecological processes may underlie the IDH (colonisation, facilitation, coexistence,etc)[@Lindenmayer2012;@Garcia_florez2017].
-There is a need to clarify the scope of the IDH in terms of the different facets of community diversity, that are the taxonomic richness, the species abundance distribution and the community functioning.
+There is a need to clarify the scope of the IDH in terms of the different facets of community diversity that are taxonomic richness, species abundance distribution and community functioning.
 Besides, there is a need to delineate the scope of the IDH in time, and highlight the moment when the conditions for the IDH emerge [@Sheil2003;@Shea2004].
@@
 Then, diversity metrics encompassing taxonomic richness and evenness should be considered to assess the changes in community abundance distribution. 
-A functional approach, that account for species biological attributes and ecological strategy, would directly link community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
+A functional approach, that accounts for species biological attributes and ecological strategy, would directly link community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
 In that respect, a vast literature allowed recognizing major traits that represent species ecological strategy and determine how they respond to changing conditions [@Diaz2005]. 
@@
 
-Because of the variability in the precision of botanical identification efforts, we were constraint to conduct the taxonomic composition and diversity analysis at the genus level. 
+Because of the variability in the precision of botanical identification efforts, we were constrained to conduct the taxonomic composition and diversity analysis at the genus level. 
 Taxonomic and functional trajectories of community composition were followed in a two-dimensional NMDS ordination plan.
@@
 Third, the functional space was divided into a 100 x 100 grid and the number of species with a positive TDP was counted in each cell.
-The average count across cells minus 1 returnedthe Community Functional Redundancy, which was the average number of species in the community that share the same trait values.
+The average count across cells minus 1 returned the Community Functional Redundancy, which was the average number of species in the community that share the same trait values.
 
@@
 
-The second-degree polynomial regressions between (i) the percentage AGB loss and (ii) taxonomic and functional diversity after 10, 20 and 30 years best predicted the hump-shaped curve of the disturbance impact along the disturbance intensity gradient \@ref(fig:IDHplot).
+The second-degree polynomial regressions between (i) the percentage AGB loss and (ii) taxonomic and functional diversity after 10, 20 and 30 years best predicted the hump-shaped curve of the disturbance impact along the disturbance intensity gradient (Fig.
+\@ref(fig:IDHplot)).
 The relationship between the disturbance impact and its intensity was more markedly hump-shaped for the taxonomic richness than for the Simpson diversity.
@@
 Regarding functional composition, initial communities were similar and trajectories were confounded in the functional space.
-Following disturbance, the functional composition of pre-disturbance surviving trees is belived to be the same as this of the initial community [@Herault2018].
+Following disturbance, the functional composition of pre-disturbance surviving trees is believed to be the same as this of the initial community [@Herault2018].
 Post-disturbance changes in functional composition then relied upon the recruitment of species or functional types previously infrequent or absent.
@@
 This recruitment of pioneers drove functional composition trajectories similar for all plots and disturbance intensity.
-The composition first went towards more resource-acquisitive strategies, which was translated by a displacement on the right along the first axis in the functional plan [@Westoby1998; @Wright2004; @Reich2014].
+The composition first went towards more resource-acquisitive strategies, which was translated by a displacement to the right along the first axis in the functional plan [@Westoby1998; @Wright2004; @Reich2014].
 Thereafter, the pioneers recruited primarily were progressively excluded by long-lived, more resistant and shade-tolerant species.
@@
 Below a 20%-25% AGB loss, the trees surviving after disturbance remained numerous enough to maintain the high taxonomic richness of the pre-disturbance community [@Bongers2009].
-The recruitment of pionners, infrequent or absent before disturbance, then increased the taxonomic richness all the more so that the disturbance was intense [@Martin2015; @Chaudhary2016].
+The recruitment of pioneers, infrequent or absent before disturbance, then increased the taxonomic richness all the more so that the disturbance was intense [@Martin2015; @Chaudhary2016].
 As these pioneers became more dominant they balanced the usual hyper-dominance of tropical forests, hence temporarily increasing the taxonomic evenness [@Baraloto2012a].
@@
 Taxonomic trajectories following disturbance of intermediate intensity besides highlighted an intermediate time after disturbance, around 15-20 years, for which the taxonomic evenness was maximized.
-As already observed in the Guiana shield [@Baraloto2012a] and in Borenan tropical forests [@Cannon1998],
+As already observed in the Guiana shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998],
 taxonomic evenness followed hump-shaped trajectories with a peak when the recruitment was balanced between pioneers and late-successional species. 
@@
 Progressively though, the functional redundancy was restored through the replacement of the first established species by more competitive, long-lived pioneers or late-successional species that were functionally closer to the pre-disturbance community.
-This replacement followed the lottery recruitment rules, implying an recruitment easy for the first species but becoming increasingly hampered by the emergence of interspecific competition [@Busing2002].
+This replacement followed the lottery recruitment rules, implying an easy recruitment for the first species but becoming increasingly hampered by the emergence of interspecific competition [@Busing2002].
 The recovery of the functional redundancy then relied upon the random processe of species recruitment and was increasingly slow and difficult to anticipate [@Elmqvist2003;@Diaz2005].
@@
 
-AM, EM & BH designed the study, developed the analysis framework, interpreted the results and wrote the manuscript. All authors gave final approval for publication.
+AM, EM & BH designed the study, developed the analysis framework and interpreted the results.
+AM wrote the manuscript with contributions by EM & BH.
+All authors gave final approval for publication.
 
@@
 This article is based upon the dataset of the Paracou station, which is part of the Guyafor permanent plot network in French Guiana (Cirad-CNRS-ONF). 
-The dataset is available upon request to the scientific director (https://paracou.cirad.
-fr).
+The dataset is available upon request to the scientific director (https://paracou.cirad.fr).
 

### 2019-03-24 WholePlotTrajectories.Rmd
Plan -> plane

@@
 Because of the variability in the precision of botanical identification efforts, we were constrained to conduct the taxonomic composition and diversity analysis at the genus level. 
-Taxonomic and functional trajectories of community composition were followed in a two-dimensional NMDS ordination plan.
+Taxonomic and functional trajectories of community composition were followed in a two-dimensional NMDS ordination plane.
 Two NMDS using abundance-based (Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the latter based on the seven leaf, stem and life history traits (without seed mass classes).
@@
 Functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
-First, the individuals of the trait database were mapped in the plane of the two first axes from a PCA analysis.
+First, the individuals of the trait database were mapped in the plane of the first two axes from a PCA analysis. 
 Then, for each species, the traits probability density (TPD) were computed through two-dimension kernel density estimators. 
-Second, for each community, the TDB weigthed by species abundance were summed accross the functional space.
-Third, the functional space was divided into a 100 x 100 grid and the number of species with a positive TDP was counted in each cell.
+Second, for each community, the TPD weigthed by species abundance were summed accross the functional space. 
+Third, the functional space was divided into a 100 x 100 grid and the number of species with a positive TPD was counted in each cell. 
 The average count across cells minus 1 returned the Community Functional Redundancy, which was the average number of species in the community that share the same trait values.
@@
 
-```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of taxonomic composition ((ref:Bolda) and (ref:Boldc)) and functional composition ((ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS plan.
+```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of taxonomic composition ((ref:Bolda) and (ref:Boldc)) and functional composition ((ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS plane.
 Lower panels ((ref:Boldc) and (ref:Boldd)) represent the Euclidean distance to initial condition along the 30 sampled years.
@@
 This recruitment of pioneers drove functional composition trajectories similar for all plots and disturbance intensity.
-The composition first went towards more resource-acquisitive strategies, which was translated by a displacement to the right along the first axis in the functional plan [@Westoby1998; @Wright2004; @Reich2014].
+The composition first went towards more resource-acquisitive strategies, which was translated by a displacement to the right along the first axis in the functional plane [@Westoby1998; @Wright2004; @Reich2014].
 Thereafter, the pioneers recruited primarily were progressively excluded by long-lived, more resistant and shade-tolerant species.
-The community functional composition then returned towards more resource-conservative strategies, suggesting the recovery of the initial community composition and translated in the functional plan by a displacement left along the first axis and upward along the second axis (Fig.
+The community functional composition then returned towards more resource-conservative strategies, suggesting the recovery of the initial community composition and translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig.
 \@ref(fig:NMDSplans)).

### 2019-03-25 WholePlotTrajectories.Rmd
unchanged

--- 2019-02-19 WholePlotTrajectories.Rmd
@@
 Paracou station in French Guiana (5\textdegree 18'N and 52\textdegree 53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26\textdegree C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
-Elevation ranges between 5 and 50 m and soils correspond to thin acrisols over a layer of transformed saprolite with low permeability which produces lateral drainage during heavy rains.
+Elevation ranges from 5 to 50 m and mean annual temperature is 26°C. 
+Soils are thin ferralitic acrisols over a layer of transformed saprolite.
+The ferralitic surface is thin in the whole area, with low permeability and lateral drainage, except in the highest hilltops where it is thick and with a free vertical drainage [@Gourlet-Fleury2004].
 

### 2019-03-25 WholePlotTrajectories.tex
unchanged

@@
 mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
-one-month dry season in March \citep{Wagner2011}. Elevation ranges
-between 5 and 50 m and soils correspond to thin acrisols over a layer of
-transformed saprolite with low permeability which produces lateral
-drainage during heavy rains.
+one-month dry season in March \citep{Wagner2011}. Elevation ranges from
+5 to 50 m and mean annual temperature is 26°C. Soils are thin ferralitic
+acrisols over a layer of transformed saprolite. The ferralitic surface
+is thin in the whole area, with low permeability and lateral drainage,
+except in the highest hilltops where it is thick and with a free
+vertical drainage \citep{Gourlet-Fleury2004}.
 

### 2019-03-25 WholePlotTrajectories.Rmd
relecture

--- 2019-03-24 WholePlotTrajectories.Rmd
@@
 However, the growing demand in forests products together with current global changes increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
-These threats affect the natural disturbance regime that shape and maintain the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
-It is urgent to understand community response to disturbance and the underlying processes to anticipate the fate of tropical forests in the current context.
-Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
-In tropical forests, the forest environment is maintained after disturbance and there is a succession of changes in the structure and functioning of communities that consitutes a secondary succession.
-For now, this succession has been largely studied through post-disturbance trajectories of forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016].
-Models based on observed trajectories then assessed community time to recovery and identified some of the determinants of post-disturbance trajectories, like forest structure, forest composition, environmental parameters and disturbance characteristics [@Herault2018].
-Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
-Given the variety of species response to disturbance and the huge diversity of tropical forests, post-disturbance biodiversity trajectories might be very diverse [@Lindenmayer2012;@Garcia_florez2017].
-It is essential to anticipate these trajectories and highlight the link between biodiversity and disturbance, and specifically how disturbance intensity and frequency determine the rate, endpoint and duration of post-disturbance trajectories [@Chazdon2003a;@Willig2018].
+These threats affect the disturbance regime that naturally shapes and maintains the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
+To anticipate the fate of tropical forests in the current context it is then urgent to understand the response of communities to disturbance and the underlying processes.
+Disturbance impact forest communities in changing both abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
+Although in the tropical context the forest environment is maintained after disturbance, there is a succession of changes in the structure and functioning of communities constituting a post-disturbance secondary succession.
+For now, this succession has been largely studied through trajectories of forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
+Models based on observed trajectories assessed community time to recovery and identified some of the determinants of post-disturbance trajectories, like forest structure, forest composition, environmental parameters and disturbance characteristics [@Herault2018].
+Regarding community diversity and composition however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
+Given the variety of species response to disturbance and the huge diversity of tropical forests, manifold post-disturbance biodiversity trajectories might emerge [@Lindenmayer2012;@Garcia_florez2017].
+It is then essential to anticipate post-disturbance trajectories and highlight the link between biodiversity and disturbance, specifically the link between  disturbance intensity and frequency and the rate, endpoint and duration of the trajectories [@Chazdon2003a;@Willig2018].
 
 An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
-The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978].
-The theory states that community diversity peaks at intermediate level of disturbance.
-The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
-In tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
-The link between biodiversity and disturbance might be complicated by the huge diversity of tropical communities and a variety of ecological processes may underlie the IDH (colonisation, facilitation, coexistence,etc)[@Lindenmayer2012;@Garcia_florez2017].
+The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events, stating that community diversity peaks at intermediate level of disturbance [@Connell1978].
+The theory assumes that the environmental fluctuations following disturbance foster both competitively superior species and fast colonizers and thus prevent competitive exclusion [@Shea2004;@Pulsford2016].
+In tropical forests, however, observations of the IDH often diverge from theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
+The link between biodiversity and disturbance might be complicated by the huge diversity of tropical communities and a variety of ecological processes may underlie the IDH (colonisation, facilitation, coexistence, etc)[@Lindenmayer2012;@Garcia_florez2017].
 There is a need to clarify the scope of the IDH in terms of the different facets of community diversity that are taxonomic richness, species abundance distribution and community functioning.
@@
 Analysing community response to disturbance requires a set of metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
-The analysis should first consider community composition, which is crucial for conservation issues and reveals the species fostered by the disturbance [@Lavorel2002;@Bellwood2006].
+The analysis should first consider community composition, which is crucial for conservation issues and which reveals the species fostered by the disturbance [@Lavorel2002;@Bellwood2006].
 Then, diversity metrics encompassing taxonomic richness and evenness should be considered to assess the changes in community abundance distribution. 
-A functional approach, that accounts for species biological attributes and ecological strategy, would directly link community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
+Besides, a functional approach that accounts for species biological attributes and ecological strategy would directly link community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
 In that respect, a vast literature allowed recognizing major traits that represent species ecological strategy and determine how they respond to changing conditions [@Diaz2005]. 
@@
 Paracou station in French Guiana (5\textdegree 18'N and 52\textdegree 53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26\textdegree C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
-Elevation ranges between 5 and 50 m and soils correspond to thin acrisols over a layer of transformed saprolite with low permeability which produces lateral drainage during heavy rains.
+The mean annual temperature is 26\textdegree C. 
+Elevation ranges from 5 to 50 m.
+Across all plots the topography mainly corresponds to hilltops or hillsides, while bottomlands cover less than 1 % of the area. 
+Plots are shallow ferralitic acrisols over a layer of transformed saprolite with low permeability and lateral drainage.
+Soil conditions are homogeneous, to the exception of the highest hilltops where the thick surface allows a free vertical drainage [@Gourlet-Fleury2004].
 
-The experiment is a network of 12 6.25 ha plots that underwent a disturbance gradient of three logging, thinning and fuelwood cutting treatments (Table \@ref(tab:Tab1)) according to a randomized plot design with three replicate blocks of four plots.
-The disturbance corresponds to an intensity gradient.
-For treatment 1 (T1) 10 trees of commercial species (of a diameter at 1.3 m height (DBH) equal or above 50 cm) were felled per hectare.
-For treatment 2 (T2) 10 trees/ha of commercial species (DBH  $\geq$ 50 cm) were felled and 30 trees/ha of non-valuable species (DBH $\geq$ 40 cm) were removed by poison girdling.
-For treatment 3 (T3) 10 trees/ha of commercial species (DBH $\geq$ 50 cm) were felled and 30 trees/ha of non-valuable species (15 with DBH $\geq$ 50 cm and 15 with DBH $\geq$ 40 cm) were removed by poison girdling [@Schmitt1990].
-Disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Biomass2018].
+The experiment is a network of twelve 6.25 ha plots (Table \@ref(tab:Tab1)) that underwent three disturbance treatments in 1987 according to a randomized plot design [@Gourlet-Fleury2004].  
+The experiment comprised three replicates of three sylvicultural treatments (hereafter plots T1, T2 and T3) and three control plots (T0).
+All treatments T1, T2 and T3 comprised the logging of 10 trees/ha with 50 cm minimum DBH that belonged to a set of 58 commercial species [@Gourlet-Fleury2004]. 
+Treatment T2 additionally comprised a thinning treatment by poison-girdling of non-commercial, randomly selected species with an average 30 trees/ha with 40 cm minimum DBH.
+Treatment T3 additionally comprised the logging of 15 trees /ha with 40 cm minimum DBH and the poison-gidling of 20 trees/ha with a 50 cm minimum DBH, all belonging to non-commercial species. 
+Considering the silvicultural treatments and the following damage, disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Biomass2018].
+The three treatments constituted a disturbance intensity gradient with increasing of above-ground biomass (AGB) lost and surface disturbed.
 
@@
 Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007]. 
-Species seed mass were given in 5 mass classes.
-Seed mass trajectories were therefore reported as the proportion of each class in the inventories (Supplementary materials).
+Species seed mass were given in 5 mass classes, and seed mass trajectories were reported as the proportion of each class in the inventories (Supplementary materials).
 
 The taxonomic diversity was reported through species richness and the Hill number translation of the Simpson index [@Hill1973].
-These metrics allowed assessing the taxonomic richness as well as evenness, through the comparison between richness and Simpson diversity.
-R.esults will thus be discussed directly in terms of taxonomic richness and evenness.
+These metrics allowed assessing the taxonomic richness as well as evenness, through the comparison between these two metrics.: results will be discussed directly in terms of taxonomic richness and evenness.
 Both indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recommended for diversity studies [@Marcon2015b]. 
@@
 Functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
-First, the individuals of the trait database were mapped in the plane of the first two axes from a PCA analysis. 
-Then, for each species, the traits probability density (TPD) were computed through two-dimension kernel density estimators. 
+First, the individuals of the trait database were mapped in the plane of the first two axes from a PCA analysis, which lowered the weight of correlations among traits as the PCA axes were combinations of most decoupled traits.
+For each species, the traits probability density (TPD) were computed from the mapping of individuals through two-dimension kernel density estimators. 
 Second, for each community, the TPD weigthed by species abundance were summed accross the functional space. 
@@
 In disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high mean WSG to high mean SLA and chlorophyll content (see appendix I).
-For functional composition, this translated into cyclic compositional changes with an incomplete recovery of the initial composition (Fig.
-\@ref(fig:NMDSplans)).
+For functional composition, this translated into cyclic compositional changes with an incomplete recovery of the initial composition (Fig. \@ref(fig:NMDSplans)). 
 The maximum dissimilarity with the initial state was positively correlated with the disturbance intensity for both taxonomic and functional composition ($\rho_{Spearman}^{Taxonomic}=0.87$ and $\rho_{Spearman}^{Functional}=0.90$ respectively).
@@
 Negative correlation was observed for Leaf toughness, WSG and Hmax ($\rho_{Spearman}^{Leaf toughness}=-0.53$, $\rho_{Spearman}^{WSG}=-0.75$, $\rho_{Spearman}^{Hmax}=-0.40$)
-The proportions of the three lightest seed mass classes increased in all disturbed plots, and decreased after 30 years for the lightest class while it stabilized for the two other (Supp.
-Mat.
-- Fig.
-S2).
+The proportions of the three lightest seed mass classes increased in all disturbed plots.
+After 30 years the proportion of lightest seed mass class decreased while it stabilized for the two other (Supp. Mat. - Fig. S2).
 

### 2019-03-25 WholePlotTrajectories.tex
relecture

--- 2019-02-19 WholePlotTrajectories.tex
@@
 \Abstract{
-Understand biodiversity trajectories of tropical forests following
-disturbance and clarify the link between biodiversity and disturbance is
-urgent to anticipate the impact of the global change context. This calls
-to highlight the taxonomic and functional facets of tree community
+Understanding biodiversity trajectories of tropical forests following
+disturbance and clarifying the link between biodiversity and disturbance
+is urgent to anticipate the impact of the global change context. This
+calls to highlight the taxonomic and functional facets of tree community
 trajectories and the underlying processes, and to clarify the link
 between biodiversity and disturbance, specifically the scope of the
-Intermediate Disturbance Hypothesis (IDH) that is debated tropical
-forests. In this study, we analyzed tree community response to a
+Intermediate Disturbance Hypothesis (IDH) that is debated in tropical
+forests. In this study, we analyzed the tree community response to a
 disturbance gradient in a Neotropical forest through diversity,
@@
 global changes increase the pressure on remaining natural forests
-\citep{Morales-Hidalgo2015}. These threats affect the natural
-disturbance regime that shape and maintain the structure, composition
-and functioning of tree communities
-\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. It is urgent to
-understand community response to disturbance and the underlying
-processes to anticipate the fate of tropical forests in the current
-context. Disturbance change both community abiotic and biotic
-environments, through modifications in the fluxes of light, heat and
-water \citep{Goulamoussene2017}. In tropical forests, the forest
-environment is maintained after disturbance and there is a suit of
-changes in the structure and functioning of communities that consitutes
-a secondary succession. For now, this succession has been largely
-studied through post-disturbance trajectories of forest structural
-parameters such as aboveground biomass, tree height or stem density
-\citep{Piponiot2016, Rutishauser2016}. Models based on observed
-trajectories then assessed community time to recovery and identified
+\citep{Morales-Hidalgo2015}. These threats affect the disturbance regime
+that naturally shapes and maintains the structure, composition and
+functioning of tree communities
+\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. To anticipate
+the fate of tropical forests in the current context it is then urgent to
+understand the response of communities to disturbance and the underlying
+processes. Disturbance impact forest communities in changing both
+abiotic and biotic environments, through modifications in the fluxes of
+light, heat and water \citep{Goulamoussene2017}. Although in the
+tropical context the forest environment is maintained after disturbance,
+there is a succession of changes in the structure and functioning of
+communities constituting a post-disturbance secondary succession. For
+now, this succession has been largely studied through trajectories of
+forest structural parameters such as aboveground biomass, tree height or
+stem density \citep{Piponiot2016, Rutishauser2016}. Models based on
+observed trajectories assessed community time to recovery and identified
 some of the determinants of post-disturbance trajectories, like forest
@@
 characteristics \citep{Herault2018}. Regarding community diversity and
-composition, however, post-disturbance trajectories have not been as
+composition however, post-disturbance trajectories have not been as
 thoroughly understood \citep{Guitet2018, Molino2001}. Given the variety
 of species response to disturbance and the huge diversity of tropical
-forests, post-disturbance biodiversity trajectories might be very
-diverse \citep{Lindenmayer2012, Garcia_florez2017}. It is essential to
-anticipate these trajectories and highlight the link between
-biodiversity and disturbance, and specifically how disturbance intensity
-and frequency determine the rate, endpoint and duration of
-post-disturbance trajectories \citep{Chazdon2003a, Willig2018}.
+forests, manifold post-disturbance biodiversity trajectories might
+emerge \citep{Lindenmayer2012, Garcia_florez2017}. It is then essential
+to anticipate post-disturbance trajectories and highlight the link
+between biodiversity and disturbance, specifically the link between
+disturbance intensity and frequency and the rate, endpoint and duration
+of the trajectories \citep{Chazdon2003a, Willig2018}.
 
@@
 assumes a relationship between community diversity and the intensity and
-frequency of disturbance events \citep{Connell1978}. The theory and
-states that community diversity peaks at intermediate level of
-disturbance. The fluctuations of environmental conditions after
-disturbance are assumed to prevent competitive exclusion in fostering
-both competitively superior species and fast colonizers
-\citep{Shea2004, Pulsford2016}. In tropical forests, however,
-observations of the IDH are often blurry compared to theoretical
+frequency of disturbance events, stating that community diversity peaks
+at intermediate level of disturbance \citep{Connell1978}. The theory
+assumes that the environmental fluctuations following disturbance foster
+both competitively superior species and fast colonizers and thus prevent
+competitive exclusion \citep{Shea2004, Pulsford2016}. In tropical
+forests, however, observations of the IDH often diverge from theoretical
 expectations \citep{Hugues2007, Sheil2003, Norden2017}, and the
@@
 communities and a variety of ecological processes may underlie the IDH
-(colonisation, facilitation,
-coexistence,etc)\citep{Lindenmayer2012, Garcia_florez2017}. There is a
-need to clarify the scope of the IDH in terms of the different facets of
-community diversity, that are the taxonomic richness, the species
-abundance distribution and the community functioning. Besides, there is
-a need to delineate the scope of the IDH in time, and highlight the
-moment when the conditions for the IDH emerge
-\citep{Sheil2003, Shea2004}.
+(colonisation, facilitation, coexistence,
+etc)\citep{Lindenmayer2012, Garcia_florez2017}. There is a need to
+clarify the scope of the IDH in terms of the different facets of
+community diversity that are taxonomic richness, species abundance
+distribution and community functioning. Besides, there is a need to
+delineate the scope of the IDH in time, and highlight the moment when
+the conditions for the IDH emerge \citep{Sheil2003, Shea2004}.
 
@@
 consider community composition, which is crucial for conservation issues
-and reveals the species fostered by the disturbance
+and which reveals the species fostered by the disturbance
 \citep{Lavorel2002, Bellwood2006}. Then, diversity metrics encompassing
 taxonomic richness and evenness should be considered to assess the
-changes in community abundance distribution. A functional approach, that
-account for species biological attributes and ecological strategy, would
-directly link community diversity, composition and redundancy to
-ecosystem functioning and environmental constraints
+changes in community abundance distribution. Besides, a functional
+approach that accounts for species biological attributes and ecological
+strategy would directly link community diversity, composition and
+redundancy to ecosystem functioning and environmental constraints
 \citep{Violle2007b, Baraloto2012a}. In that respect, a vast literature
@@
 mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
-one-month dry season in March \citep{Wagner2011}. Elevation ranges
-between 5 and 50 m and soils correspond to thin acrisols over a layer of
-transformed saprolite with low permeability which produces lateral
-drainage during heavy rains.
+one-month dry season in March \citep{Wagner2011}. The mean annual
+temperature is 26\textdegree C. Elevation ranges from 5 to 50 m. Across
+all plots the topography mainly corresponds to hilltops or hillsides,
+while bottomlands cover less than 1 \% of the area. Plots are shallow
+ferralitic acrisols over a layer of transformed saprolite with low
+permeability and lateral drainage. Soil conditions are homogeneous, to
+the exception of the highest hilltops where the thick surface allows a
+free vertical drainage \citep{Gourlet-Fleury2004}.
 
-The experiment is a network of 12 6.25 ha plots that underwent a
-disturbance gradient of three logging, thinning and fuelwood cutting
-treatments (Table \ref{tab:Tab1}) according to a randomized plot design
-with three replicate blocks of four plots. The disturbance corresponds
-to an intensity gradient. For treatment 1 (T1) 10 trees of commercial
-species (of a diameter at 1.3 m height (DBH) equal or above 50 cm) were
-felled per hectare. For treatment 2 (T2) 10 trees/ha of commercial
-species (DBH \(\geq\) 50 cm) were felled and 30 trees/ha of non-valuable
-species (DBH \(\geq\) 40 cm) were removed by poison girdling. For
-treatment 3 (T3) 10 trees/ha of commercial species (DBH \(\geq\) 50 cm)
-were felled and 30 trees/ha of non-valuable species (15 with DBH
-\(\geq\) 50 cm and 15 with DBH \(\geq\) 40 cm) were removed by poison
-girdling \citep{Schmitt1990}. Disturbance intensity was measured as the
-percentage of aboveground biomass (\%AGB) lost between the first
-inventory in 1984 and five years after disturbance \citep{Piponiot2016}
-estimated with the BIOMASS R package \citep{Biomass2018}.
+The experiment is a network of twelve 6.25 ha plots (Table
+\ref{tab:Tab1}) that underwent three disturbance treatments in 1987
+according to a randomized plot design \citep{Gourlet-Fleury2004}.\\
+The experiment comprised three replicates of three sylvicultural
+treatments (hereafter plots T1, T2 and T3) and three control plots (T0).
+All treatments T1, T2 and T3 comprised the logging of 10 trees/ha with
+50 cm minimum DBH that belonged to a set of 58 commercial species
+\citep{Gourlet-Fleury2004}. Treatment T2 additionally comprised a
+thinning treatment by poison-girdling of non-commercial, randomly
+selected species with an average 30 trees/ha with 40 cm minimum DBH.
+Treatment T3 additionally comprised the logging of 15 trees /ha with 40
+cm minimum DBH and the poison-gidling of 20 trees/ha with a 50 cm
+minimum DBH, all belonging to non-commercial species. Considering the
+silvicultural treatments and the following damage, disturbance intensity
+was measured as the percentage of aboveground biomass (\%AGB) lost
+between the first inventory in 1984 and five years after disturbance
+\citep{Piponiot2016} estimated with the BIOMASS R package
+\citep{Biomass2018}. The three treatments constituted a disturbance
+intensity gradient with increasing of above-ground biomass (AGB) lost
+and surface disturbed.
 
@@
 Because of the variability in the precision of botanical identification
-efforts, we were constraint to conduct the taxonomic composition and
+efforts, we were constrained to conduct the taxonomic composition and
 diversity analysis at the genus level. Taxonomic and functional
 trajectories of community composition were followed in a two-dimensional
-NMDS ordination plan. Two NMDS using abundance-based (Bray-Curtis)
+NMDS ordination plane. Two NMDS using abundance-based (Bray-Curtis)
 dissimilarity measures were conducted to map either taxonomic or
@@
 means (CWM) \citep{Diaz2007}. Species seed mass were given in 5 mass
-classes. Seed mass trajectories were therefore reported as the
-proportion of each class in the inventories (Supplementary materials).
+classes, and seed mass trajectories were reported as the proportion of
+each class in the inventories (Supplementary materials).
 
@@
 metrics allowed assessing the taxonomic richness as well as evenness,
-through the comparison between richness and Simpson diversity. R.esults
-will thus be discussed directly in terms of taxonomic richness and
-evenness. Both indices belong to the set of HCDT or generalized entropy,
-respectively corresponding to the 0 and 2 order of diversity (q),
-recommended for diversity studies \citep{Marcon2015b}. The functional
-diversity was reported using the functional richness and functional
-evenness, \emph{i.e} the Rao index of quadratic entropy. The Rao index
-combines species abundance distribution, and the average pairwise
-functional dissimilarity between species computed by the gower distance.
+through the comparison between these two metrics.: results will be
+discussed directly in terms of taxonomic richness and evenness. Both
+indices belong to the set of HCDT or generalized entropy, respectively
+corresponding to the 0 and 2 order of diversity (q), recommended for
+diversity studies \citep{Marcon2015b}. The functional diversity was
+reported using the functional richness and functional evenness,
+\emph{i.e} the Rao index of quadratic entropy. The Rao index combines
+species abundance distribution, and the average pairwise functional
+dissimilarity between species computed by the gower distance.
 
@@
 community functional space \citep{Carmona2016}. First, the individuals
-of the trait database were mapped in the plane of the two first axes
-from a PCA analysis. Then, for each species, the traits probability
-density (TPD) were computed through two-dimension kernel density
-estimators. Second, for each community, the TDB weigthed by species
+of the trait database were mapped in the plane of the first two axes
+from a PCA analysis, which lowered the weight of correlations among
+traits as the PCA axes were combinations of most decoupled traits. For
+each species, the traits probability density (TPD) were computed from
+the mapping of individuals through two-dimension kernel density
+estimators. Second, for each community, the TPD weigthed by species
 abundance were summed accross the functional space. Third, the
 functional space was divided into a 100 x 100 grid and the number of
-species with a positive TDP was counted in each cell. The average count
-across cells minus 1 returnedthe Community Functional Redundancy, which
+species with a positive TPD was counted in each cell. The average count
+across cells minus 1 returned the Community Functional Redundancy, which
 was the average number of species in the community that share the same
@@
 
-\caption{Plot trajectories in terms of taxonomic composition (\textbf{(a)} and \textbf{(c)}) and functional composition (\textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS plan.
+\caption{Plot trajectories in terms of taxonomic composition (\textbf{(a)} and \textbf{(c)}) and functional composition (\textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS plane.
 Lower panels (\textbf{(c)} and \textbf{(d)}) represent the Euclidean distance to initial condition along the 30 sampled years.
@@
 proportions of the three lightest seed mass classes increased in all
-disturbed plots, and decreased after 30 years for the lightest class
-while it stabilized for the two other (Supp. Mat. - Fig. S2).
+disturbed plots. After 30 years the proportion of lightest seed mass
+class decreased while it stabilized for the two other (Supp. Mat. - Fig.
+S2).
 
@@
 years best predicted the hump-shaped curve of the disturbance impact
-along the disturbance intensity gradient \ref{fig:IDHplot}. The
+along the disturbance intensity gradient (Fig. \ref{fig:IDHplot}). The
 relationship between the disturbance impact and its intensity was more
@@
 disturbance, the functional composition of pre-disturbance surviving
-trees is belived to be the same as this of the initial community
+trees is believed to be the same as this of the initial community
 \citep{Herault2018}. Post-disturbance changes in functional composition
@@
 went towards more resource-acquisitive strategies, which was translated
-by a displacement on the right along the first axis in the functional
-plan \citep{Westoby1998, Wright2004, Reich2014}. Thereafter, the
+by a displacement to the right along the first axis in the functional
+plane \citep{Westoby1998, Wright2004, Reich2014}. Thereafter, the
 pioneers recruited primarily were progressively excluded by long-lived,
@@
 suggesting the recovery of the initial community composition and
-translated in the functional plan by a displacement left along the first
-axis and upward along the second axis (Fig. \ref{fig:NMDSplans}).
+translated in the functional plane by a displacement left along the
+first axis and upward along the second axis (Fig. \ref{fig:NMDSplans}).
 
@@
 high taxonomic richness of the pre-disturbance community
-\citep{Bongers2009}. The recruitment of pionners, infrequent or absent
+\citep{Bongers2009}. The recruitment of pioneers, infrequent or absent
 before disturbance, then increased the taxonomic richness all the more
@@
 years, for which the taxonomic evenness was maximized. As already
-observed in the Guiana shield \citep{Baraloto2012a} and in Borenan
+observed in the Guiana shield \citep{Baraloto2012a} and in Bornean
 tropical forests \citep{Cannon1998}, taxonomic evenness followed
@@
 the pre-disturbance community. This replacement followed the lottery
-recruitment rules, implying an recruitment easy for the first species
+recruitment rules, implying an easy recruitment for the first species
 but becoming increasingly hampered by the emergence of interspecific
@@
 
-AM, EM \& BH designed the study, developed the analysis framework,
-interpreted the results and wrote the manuscript. All authors gave final
-approval for publication.
+AM, EM \& BH designed the study, developed the analysis framework and
+interpreted the results. AM wrote the manuscript with contributions by
+EM \& BH. All authors gave final approval for publication.
 
@@
 (Cirad-CNRS-ONF). The dataset is available upon request to the
-scientific director (https://paracou.cirad. fr).
+scientific director (https://paracou.cirad.fr).
 

### 2019-03-26 WholePlotTrajectories.Rmd
relecture

@@
 Library("kableExtra")
-
-## abstract: 204 words / text without legends : 3659 words / references: 68
 
@@
 \@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-\break disturbance time.
-In disturbed communities, these compositional changes corresponded to shifts towards species with more acquisitive functional strategies, from communities with high mean WSG to high mean SLA and chlorophyll content (see appendix I).
-For functional composition, this translated into cyclic compositional changes with an incomplete recovery of the initial composition (Fig. \@ref(fig:NMDSplans)). 
+In disturbed communities the taxonomic composition changed towards more species with acquisitive functional strategies.
+This translated into a switch from high mean WSG in the community to high mean SLA and chlorophyll content (see appendix I). 
+The functional composition of disturbed communities corresponded to cyclic compositional changes with an incomplete recovery of the initial composition (Fig. \@ref(fig:NMDSplans)).
+
 The maximum dissimilarity with the initial state was positively correlated with the disturbance intensity for both taxonomic and functional composition ($\rho_{Spearman}^{Taxonomic}=0.87$ and $\rho_{Spearman}^{Functional}=0.90$ respectively).
-The maximum dissimilarity with the initial was reached for taxonomic composition between 15 to 25 years, for most of the plots, and around 22 years for functional composition.
+The maximum dissimilarity with the initial state was reached for taxonomic composition between 15 to 25 years, for most of the plots, and around 22 years for functional composition.
 
@@
 
-Community CWM average value of Maximum height at adult stage (*Hmax*), leaf toughness and wood specific gravity (*WSG*) first decreased and then slightly increased but remained significantly lower than their initial value (Fig.
+Community CWM average value of Maximum height at adult stage (*Hmax*), leaf toughness and wood specific gravity (*WSG*) decreased and then slightly increased, but remained significantly lower than their initial value (Fig.
 \@ref(fig:CWM)).
-On the other side, bark thickness and specific leaf area (*SLA*) increased and while bark thickness remained substantially high after 30 years, *SLA* had almost recovered to its initial value. 
+On the other side, bark thickness and specific leaf area (*SLA*) both increased.
+Bark thickness remained substantially high after 30 years, and *SLA* had almost recovered to its initial value. 
 For all traits, the maximum difference to initial value was correlated to the disturbance intensity.
@@
 The proportions of the three lightest seed mass classes increased in all disturbed plots.
-After 30 years the proportion of lightest seed mass class decreased while it stabilized for the two other (Supp. Mat. - Fig. S2).
+After 30 years the proportion of lightest seed mass class decreased while it stabilized for the two other lightest seed mass classes (Supp. Mat. - Fig. S2).
 
@@
 The maximum richness loss or gain after disturbance was positively correlated with the disturbance intensity ($\rho_{Spearman}^{Richness}=0.50$).
+
 In all disturbed plots the Simpson diversity first increased until a maximum reached after around 20 years. 
@@
 In disturbed plots, both trajectories depended on the disturbance intensity, with their maximum being positively correlated to %AGB loss $\rho_{Spearman}^{Richness}=0.76$ and $\rho_{Spearman}^{Rao}=0.60$.
-Functional richness and Rao diversity displayed for low disturbance intensity a low but long-lasting increase up to a maximum reached after 20-25 years, and for high intensity, a fast but short increase followed after 10 years by a slow decrease towards the initial values.
+Functional richness and Rao diversity displayed for low disturbance intensity a low but long-lasting increase up to a maximum reached after 20-25 years.
+Following high intensity, they displayed a fast but short increase followed after 10 years by a slow decrease towards the initial values.
 
@@
 Before disturbance, the different communities had different taxonomic compositions, visualized by their distinct starting points on the NMDS axis 2.
-These initial differences were maintained all along the 30 years following disturbance, with post-disturbance trajectories corresponding to a displacement on the NMDS axis 1.
-The taxonomic compositional changes following disturbance were similar among communities, which may correspond to the recruitment of a group of pioneers common to all communities whenever their initial differences and the disturbance intensity [@Denslow2000;@Bongers2009].
-Taxonomic trajectories continued with a recovery of the initial composition which, althought not fully achieved after 30 years, suggested the community taxonomic resilience and the maintenance of community initial differences [@Folke2006].
+These initial differences were maintained all along the 30 years following disturbance, with the disturbance leading a displacement on the NMDS axis 1.
+The taxonomic compositional changes following disturbance were similar among communities, which may correspond to the recruitment of a group of pioneers common to all communities whenever their initial differences and the intensity of disturbance [@Denslow2000;@Bongers2009].
+Taxonomic trajectories continued with a recovery of the initial composition which, althought not fully achieved after 30 years, suggested the resilience of the taxonomic composition resilience and the maintenance of the initial differences among communities [@Folke2006].
 The  return towards initial taxonomic composition showed that species not belonging to the pre-disturbance community were rarely recruited, probably because of the common dispersal limitations among tropical tree species [@Svenning2005].
@@
 Following disturbance, the functional composition of pre-disturbance surviving trees is believed to be the same as this of the initial community [@Herault2018].
-Post-disturbance changes in functional composition then relied upon the recruitment of species or functional types previously infrequent or absent.
-These species were probably competitive pioneers becoming dominant as the most efficient to benefit from the light, space and nutrients made available by disturbance.
-This recruitment of pioneers drove functional composition trajectories similar for all plots and disturbance intensity.
+The changes in functional composition following disturbance then relied upon the recruitment of species or functional types previously infrequent or absent.
+These species were probably competitive pioneers becoming dominant as the most efficient to benefit from the light, space and nutrients made available by the disturbance.
+This recruitment of pioneers led similar functional composition trajectories for all plots, whenever the disturbance intensity.
 The composition first went towards more resource-acquisitive strategies, which was translated by a displacement to the right along the first axis in the functional plane [@Westoby1998; @Wright2004; @Reich2014].
 Thereafter, the pioneers recruited primarily were progressively excluded by long-lived, more resistant and shade-tolerant species.
-The community functional composition then returned towards more resource-conservative strategies, suggesting the recovery of the initial community composition and translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig.
-\@ref(fig:NMDSplans)).
+The community functional composition then returned towards more resource-conservative strategies, suggesting the recovery of the initial functional composition.
+This recovery translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig. \@ref(fig:NMDSplans)).
 
 Both taxonomic and functional composition trajectories initiated a return towards pre-disturbance state after 30 years, which highlighted the taxonomic and functional resilience of communities. 
-There was however a decoupling between taxonomic and functional composition trajectories: while as taxonomic trajectories maintained the initial differences among communities, the functional trajectories where similar and convergent in the functional space [@Fukami2005].
+Taxonomicand functional trajectories however appeared decoupled: while taxonomic trajectories maintained the initial differences among communities, the functional trajectories where similar and convergent in the functional space [@Fukami2005].
 
@@
 Community taxonomic richness and evenness trajectories were determined by the disturbance intensity, ranging from a limited and temporary impact to significant and persistant alterations of community diversity.
-The taxonomic trajectories markedly changed above an intensity threshold for which the taxonomic richness was maximized and the taxonomic evenness remained resilient. 
-The disturbance intensity determined the balance in the community between pre-disturbance surviving trees, and trees recruited afterward. 
-Below a 20%-25% AGB loss, the trees surviving after disturbance remained numerous enough to maintain the high taxonomic richness of the pre-disturbance community [@Bongers2009].
+The taxonomic trajectories markedly changed above an intensity threshold at which the taxonomic richness was maximized and the taxonomic evenness remained resilient. 
+The disturbance intensity impacted the balance in the community between pre-disturbance surviving trees, and trees recruited afterward. 
+On the one hand, below a 20%-25% AGB loss, the trees surviving after disturbance remained numerous enough to maintain the high taxonomic richness of the pre-disturbance community [@Bongers2009].
 The recruitment of pioneers, infrequent or absent before disturbance, then increased the taxonomic richness all the more so that the disturbance was intense [@Martin2015; @Chaudhary2016].
 As these pioneers became more dominant they balanced the usual hyper-dominance of tropical forests, hence temporarily increasing the taxonomic evenness [@Baraloto2012a].
-Above the intensity threshold, the taxonomic richness of surviving trees was too low to be offset by the recruitment of pioneers. 
+On the other hand, above the intensity threshold, the taxonomic richness of surviving trees was too low to be offset by the recruitment of pioneers. 
 In the Guiana shield indeed, the pool of true pioneers specifically recruited after disturbance is restricted to a few common genera (e.g. *Cecropia* spp., *Vismia* spp.) [@Guitet2018].
@@
 taxonomic evenness followed hump-shaped trajectories with a peak when the recruitment was balanced between pioneers and late-successional species. 
-This intermediate time corresponded to the match between the generation time of pionneers, first to dominate the post-disturbance recruitment, and the time for competitive exclusion to emerge, which leads to the dominance of late-successional species.
+This intermediate time corresponded to the match between the generation time of pionneers, dominating first the post-disturbance recruitment, and the time for competitive exclusion to emerge, leading afterwards to the dominance of late-successional species.
 
@@
 Above the intensity threshold however, for the most intense disturbance, community functional richness and evenness started to decrease after 15 to 20 years. 
-The right after disturbance, short-lived species benefited from the resources made available and prevented the establishment of other species. 
+Right after disturbance, short-lived species benefited from the resources made available and prevented the establishment of other species. 
 The decline of these short-lived pioneers decreased the functional richness and evenness, but we suggest that the establishment of long-lasting pioneers will follow and that the trajectories will catch up with those observed for intermediate disturbance [@Walker2009].
 
-The IDH then translated into community taxonomic response, with post-disturbance trajectories markedly different below and above an intensity threshold.
+The IDH then translated into community taxonomic response, with post-disturbance trajectories changing markedly below and above an intensity threshold of 25 % AGB loss.
 The IDH was also valid at an intermediate time after disturbance, when the recruitment of early- and late-successional species were balanced. 
@@
 This replacement followed the lottery recruitment rules, implying an easy recruitment for the first species but becoming increasingly hampered by the emergence of interspecific competition [@Busing2002].
-The recovery of the functional redundancy then relied upon the random processe of species recruitment and was increasingly slow and difficult to anticipate [@Elmqvist2003;@Diaz2005].
+The recovery of the functional redundancy then relied upon the random process of species recruitment and was increasingly slow and difficult to anticipate [@Elmqvist2003;@Diaz2005].
 
@@
 Community post-disturbance trajectories in compositon and diversity were shaped by the recruitment of a determined pool of pioneers identical among communities and disturbance intensity.
-Nevertheless, taxonomic and functional composition trajectories were decoupled: while functional trajectories remained similar in the functional space and converged towards the recovery of a comparable initial state, taxonomic trajectories shiwed initial differences in community composition that were maintained along time.
-This decoupling was explained by a decrease of the functional redundancy which mitigated the functional impact of disturbance.
+Nevertheless, taxonomic and functional composition trajectories appeared decoupled.
+While functional trajectories remained similar in the functional space and converged towards the recovery of a comparable initial state, taxonomic trajectories showed initial differences in community composition that were maintained along time.
+This decoupling was explained by a decrease of the functional redundancy that mitigated the functional impact of disturbance.
 Community functional and taxonomic diversity response to disturbance was contrasted as well.
@@
 Taxonomic diversity was maximized following a disturbance of intermediate intensity, that was the threshold of 25% AGB removed, and at an intermediate time after disturbance, that was around 25 years after disturbance when the recruitment of early- and late-successional species were balanced.
-Whenever the disturbance intensity, community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied upon the random, lottery recruitment of rare species.
+Whenever the disturbance intensity, community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied upon the random lottery recruitment of rare species.
 Although resilient, the functional redundancy was lower for more than 30 years after disturbance which cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].

### 2019-03-26 WholePlotTrajectories.tex
relecture

@@
 and consistent trajectories over post-\break disturbance time. In
-disturbed communities, these compositional changes corresponded to
-shifts towards species with more acquisitive functional strategies, from
-communities with high mean WSG to high mean SLA and chlorophyll content
-(see appendix I). For functional composition, this translated into
-cyclic compositional changes with an incomplete recovery of the initial
-composition (Fig. \ref{fig:NMDSplans}). The maximum dissimilarity with
-the initial state was positively correlated with the disturbance
-intensity for both taxonomic and functional composition
-(\(\rho_{Spearman}^{Taxonomic}=0.87\) and
+disturbed communities the taxonomic composition changed towards more
+species with acquisitive functional strategies. This translated into a
+switch from high mean WSG in the community to high mean SLA and
+chlorophyll content (see appendix I). The functional composition of
+disturbed communities corresponded to cyclic compositional changes with
+an incomplete recovery of the initial composition (Fig.
+\ref{fig:NMDSplans}).
+
+The maximum dissimilarity with the initial state was positively
+correlated with the disturbance intensity for both taxonomic and
+functional composition (\(\rho_{Spearman}^{Taxonomic}=0.87\) and
 \(\rho_{Spearman}^{Functional}=0.90\) respectively). The maximum
-dissimilarity with the initial was reached for taxonomic composition
-between 15 to 25 years, for most of the plots, and around 22 years for
-functional composition.
+dissimilarity with the initial state was reached for taxonomic
+composition between 15 to 25 years, for most of the plots, and around 22
+years for functional composition.
 
@@
 (\emph{Hmax}), leaf toughness and wood specific gravity (\emph{WSG})
-first decreased and then slightly increased but remained significantly
-lower than their initial value (Fig. \ref{fig:CWM}). On the other side,
-bark thickness and specific leaf area (\emph{SLA}) increased and while
-bark thickness remained substantially high after 30 years, \emph{SLA}
-had almost recovered to its initial value. For all traits, the maximum
+decreased and then slightly increased, but remained significantly lower
+than their initial value (Fig. \ref{fig:CWM}). On the other side, bark
+thickness and specific leaf area (\emph{SLA}) both increased. Bark
+thickness remained substantially high after 30 years, and \emph{SLA} had
+almost recovered to its initial value. For all traits, the maximum
 difference to initial value was correlated to the disturbance intensity.
@@
 disturbed plots. After 30 years the proportion of lightest seed mass
-class decreased while it stabilized for the two other (Supp. Mat. - Fig.
-S2).
+class decreased while it stabilized for the two other lightest seed mass
+classes (Supp. Mat. - Fig. S2).
 
@@
 disturbance was positively correlated with the disturbance intensity
-(\(\rho_{Spearman}^{Richness}=0.50\)). In all disturbed plots the
-Simpson diversity first increased until a maximum reached after around
-20 years. This maximum was positively correlated with the disturbance
-intensity (\(\rho_{Spearman}^{Simpson}=0.77\)). The Simpson diversity
-then stabilized except for two T3 plots (plots 8 and 12) for which it
-kept increasing (Fig. \ref{fig:DivTaxo}).
+(\(\rho_{Spearman}^{Richness}=0.50\)).
+
+In all disturbed plots the Simpson diversity first increased until a
+maximum reached after around 20 years. This maximum was positively
+correlated with the disturbance intensity
+(\(\rho_{Spearman}^{Simpson}=0.77\)). The Simpson diversity then
+stabilized except for two T3 plots (plots 8 and 12) for which it kept
+increasing (Fig. \ref{fig:DivTaxo}).
 
@@
 intensity a low but long-lasting increase up to a maximum reached after
-20-25 years, and for high intensity, a fast but short increase followed
-after 10 years by a slow decrease towards the initial values.
+20-25 years. Following high intensity, they displayed a fast but short
+increase followed after 10 years by a slow decrease towards the initial
+values.
 
@@
 axis 2. These initial differences were maintained all along the 30 years
-following disturbance, with post-disturbance trajectories corresponding
-to a displacement on the NMDS axis 1. The taxonomic compositional
-changes following disturbance were similar among communities, which may
-correspond to the recruitment of a group of pioneers common to all
-communities whenever their initial differences and the disturbance
-intensity \citep{Denslow2000, Bongers2009}. Taxonomic trajectories
-continued with a recovery of the initial composition which, althought
-not fully achieved after 30 years, suggested the community taxonomic
-resilience and the maintenance of community initial differences
-\citep{Folke2006}. The return towards initial taxonomic composition
-showed that species not belonging to the pre-disturbance community were
-rarely recruited, probably because of the common dispersal limitations
-among tropical tree species \citep{Svenning2005}.
+following disturbance, with the disturbance leading a displacement on
+the NMDS axis 1. The taxonomic compositional changes following
+disturbance were similar among communities, which may correspond to the
+recruitment of a group of pioneers common to all communities whenever
+their initial differences and the intensity of disturbance
+\citep{Denslow2000, Bongers2009}. Taxonomic trajectories continued with
+a recovery of the initial composition which, althought not fully
+achieved after 30 years, suggested the resilience of the taxonomic
+composition resilience and the maintenance of the initial differences
+among communities \citep{Folke2006}. The return towards initial
+taxonomic composition showed that species not belonging to the
+pre-disturbance community were rarely recruited, probably because of the
+common dispersal limitations among tropical tree species
+\citep{Svenning2005}.
 
@@
 trees is believed to be the same as this of the initial community
-\citep{Herault2018}. Post-disturbance changes in functional composition
-then relied upon the recruitment of species or functional types
-previously infrequent or absent. These species were probably competitive
-pioneers becoming dominant as the most efficient to benefit from the
-light, space and nutrients made available by disturbance. This
-recruitment of pioneers drove functional composition trajectories
-similar for all plots and disturbance intensity. The composition first
-went towards more resource-acquisitive strategies, which was translated
-by a displacement to the right along the first axis in the functional
-plane \citep{Westoby1998, Wright2004, Reich2014}. Thereafter, the
-pioneers recruited primarily were progressively excluded by long-lived,
-more resistant and shade-tolerant species. The community functional
-composition then returned towards more resource-conservative strategies,
-suggesting the recovery of the initial community composition and
-translated in the functional plane by a displacement left along the
-first axis and upward along the second axis (Fig. \ref{fig:NMDSplans}).
+\citep{Herault2018}. The changes in functional composition following
+disturbance then relied upon the recruitment of species or functional
+types previously infrequent or absent. These species were probably
+competitive pioneers becoming dominant as the most efficient to benefit
+from the light, space and nutrients made available by the disturbance.
+This recruitment of pioneers led similar functional composition
+trajectories for all plots, whenever the disturbance intensity. The
+composition first went towards more resource-acquisitive strategies,
+which was translated by a displacement to the right along the first axis
+in the functional plane \citep{Westoby1998, Wright2004, Reich2014}.
+Thereafter, the pioneers recruited primarily were progressively excluded
+by long-lived, more resistant and shade-tolerant species. The community
+functional composition then returned towards more resource-conservative
+strategies, suggesting the recovery of the initial functional
+composition. This recovery translated in the functional plane by a
+displacement left along the first axis and upward along the second axis
+(Fig. \ref{fig:NMDSplans}).
 
@@
 return towards pre-disturbance state after 30 years, which highlighted
-the taxonomic and functional resilience of communities. There was
-however a decoupling between taxonomic and functional composition
-trajectories: while as taxonomic trajectories maintained the initial
-differences among communities, the functional trajectories where similar
-and convergent in the functional space \citep{Fukami2005}.
+the taxonomic and functional resilience of communities. Taxonomicand
+functional trajectories however appeared decoupled: while taxonomic
+trajectories maintained the initial differences among communities, the
+functional trajectories where similar and convergent in the functional
+space \citep{Fukami2005}.
 
@@
 The taxonomic trajectories markedly changed above an intensity threshold
-for which the taxonomic richness was maximized and the taxonomic
-evenness remained resilient. The disturbance intensity determined the
-balance in the community between pre-disturbance surviving trees, and
-trees recruited afterward. Below a 20\%-25\% AGB loss, the trees
-surviving after disturbance remained numerous enough to maintain the
-high taxonomic richness of the pre-disturbance community
+at which the taxonomic richness was maximized and the taxonomic evenness
+remained resilient. The disturbance intensity impacted the balance in
+the community between pre-disturbance surviving trees, and trees
+recruited afterward. On the one hand, below a 20\%-25\% AGB loss, the
+trees surviving after disturbance remained numerous enough to maintain
+the high taxonomic richness of the pre-disturbance community
 \citep{Bongers2009}. The recruitment of pioneers, infrequent or absent
@@
 hyper-dominance of tropical forests, hence temporarily increasing the
-taxonomic evenness \citep{Baraloto2012a}. Above the intensity threshold,
-the taxonomic richness of surviving trees was too low to be offset by
-the recruitment of pioneers. In the Guiana shield indeed, the pool of
-true pioneers specifically recruited after disturbance is restricted to
-a few common genera (e.g. \emph{Cecropia} spp., \emph{Vismia} spp.)
-\citep{Guitet2018}. The taxonomic richness then decreased following
-intense disturbance, all the more so that the disturbance intensity was
-high \citep{Molino2001}. At the same time, pioneers became persistently
-dominant and prevented a return towards the initial evenness and the
-recovery of hyper-dominant shade-tolerant species.
+taxonomic evenness \citep{Baraloto2012a}. On the other hand, above the
+intensity threshold, the taxonomic richness of surviving trees was too
+low to be offset by the recruitment of pioneers. In the Guiana shield
+indeed, the pool of true pioneers specifically recruited after
+disturbance is restricted to a few common genera (e.g. \emph{Cecropia}
+spp., \emph{Vismia} spp.) \citep{Guitet2018}. The taxonomic richness
+then decreased following intense disturbance, all the more so that the
+disturbance intensity was high \citep{Molino2001}. At the same time,
+pioneers became persistently dominant and prevented a return towards the
+initial evenness and the recovery of hyper-dominant shade-tolerant
+species.
 
@@
 corresponded to the match between the generation time of pionneers,
-first to dominate the post-disturbance recruitment, and the time for
-competitive exclusion to emerge, which leads to the dominance of
+dominating first the post-disturbance recruitment, and the time for
+competitive exclusion to emerge, leading afterwards to the dominance of
 late-successional species.
@@
 for the most intense disturbance, community functional richness and
-evenness started to decrease after 15 to 20 years. The right after
+evenness started to decrease after 15 to 20 years. Right after
 disturbance, short-lived species benefited from the resources made
@@
 The IDH then translated into community taxonomic response, with
-post-disturbance trajectories markedly different below and above an
-intensity threshold. The IDH was also valid at an intermediate time
-after disturbance, when the recruitment of early- and late-successional
-species were balanced. Functional trajectories, however, remained
-decoupled from taxonomic ones and dismissed the IDH.
+post-disturbance trajectories changing markedly below and above an
+intensity threshold of 25 \% AGB loss. The IDH was also valid at an
+intermediate time after disturbance, when the recruitment of early- and
+late-successional species were balanced. Functional trajectories,
+however, remained decoupled from taxonomic ones and dismissed the IDH.
 
@@
 competition \citep{Busing2002}. The recovery of the functional
-redundancy then relied upon the random processe of species recruitment
+redundancy then relied upon the random process of species recruitment
 and was increasingly slow and difficult to anticipate
@@
 among communities and disturbance intensity. Nevertheless, taxonomic and
-functional composition trajectories were decoupled: while functional
+functional composition trajectories appeared decoupled. While functional
 trajectories remained similar in the functional space and converged
 towards the recovery of a comparable initial state, taxonomic
-trajectories shiwed initial differences in community composition that
+trajectories showed initial differences in community composition that
 were maintained along time. This decoupling was explained by a decrease
-of the functional redundancy which mitigated the functional impact of
+of the functional redundancy that mitigated the functional impact of
 disturbance. Community functional and taxonomic diversity response to
@@
 pre-disturbance state) was tangible but required several decades and
-relied upon the random, lottery recruitment of rare species. Although
+relied upon the random lottery recruitment of rare species. Although
 resilient, the functional redundancy was lower for more than 30 years

### 2019-03-26 WholePlotTrajectories.Rmd
abstract

@@
 abstract: 
- Understanding biodiversity trajectories of tropical forests following disturbance and clarifying the link between biodiversity and disturbance is urgent to anticipate the impact of the global change context.
- This calls to highlight the taxonomic and functional facets of tree community trajectories and the underlying processes, and to clarify the link between biodiversity and disturbance, specifically the scope of the Intermediate Disturbance Hypothesis (IDH) that is debated in tropical forests.
- In this study, we analyzed the tree community response to a disturbance gradient in a Neotropical forest through diversity, composition and redundancy post-disturbance trajectories over 30 years.
-We considered taxonomic trajectories and functional trajectories based on 7 leaf, stem and life-history traits.
-While taxonomic trajectories differed among communities, functional trajectories remained similar and convergent, and this decoupling was explained by the variations in community functional redundancy mitigating the functional impact of disturbance.
-We validated the IDH, based on an intensity threshold of 20-25% of AGB removed, and we clarified its scope, showing that it applied to taxonomic but not functional diversity, and applied around 20 years after disturbance when the recruitment of pioneers and late-successionals were balanced.
+ Anticipate the impact of the global change context on tropical forests is now urgent.
+ This calls to highlight the taxonomic and functional facets of tree community post-disturbance trajectories and the underlying processes. 
+ It also calls to clarify the link between biodiversity and disturbance, specifically the scope of the Intermediate Disturbance Hypothesis (IDH) that is debated in tropical forests.
+ In this study, we analyzed the tree community trajectories following a disturbance gradient in a Neotropical forest over 30 years. 
+ We considered community taxonomic and functional diversity, composition and redundancy trajectories.
+We based on the annual botanical inventories of 75 ha of a Neotropical forest and on large trait datasets comprising 7 leaf, stem and life-history traits.
+ We highlighted a decoupling between taxonomic trajectories, that differed among communities, and functional trajectories, remained similar and convergent.
+We explained this decoupling by the variations in community functional redundancy that mitigated the functional impact of disturbance.
+Post-disturbance trajectories clarified the scope of the Intermediate Disturbance Hypothesis that was validated regarding taxonomic diversity, with an intensity threshold of 20-25%, but not regarding functional diversity.The IDH applied around 20 years after disturbance when the recruitment of pioneers and late-successionals were balanced.
  Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years.

### 2019-03-26 WholePlotTrajectories.tex
abstract

@@
 \Abstract{
-Understanding biodiversity trajectories of tropical forests following
-disturbance and clarifying the link between biodiversity and disturbance
-is urgent to anticipate the impact of the global change context. This
-calls to highlight the taxonomic and functional facets of tree community
-trajectories and the underlying processes, and to clarify the link
-between biodiversity and disturbance, specifically the scope of the
-Intermediate Disturbance Hypothesis (IDH) that is debated in tropical
-forests. In this study, we analyzed the tree community response to a
-disturbance gradient in a Neotropical forest through diversity,
-composition and redundancy post-disturbance trajectories over 30 years.
-We considered taxonomic trajectories and functional trajectories based
-on 7 leaf, stem and life-history traits. While taxonomic trajectories
-differed among communities, functional trajectories remained similar and
-convergent, and this decoupling was explained by the variations in
-community functional redundancy mitigating the functional impact of
-disturbance. We validated the IDH, based on an intensity threshold of
-20-25\% of AGB removed, and we clarified its scope, showing that it
-applied to taxonomic but not functional diversity, and applied around 20
-years after disturbance when the recruitment of pioneers and
-late-successionals were balanced. Although consistent, the recovery of
-community composition, diversity and redundancy remained unachieved
-after 30 years. This acknowledged the need of decades-long cycles
-without disturbance to ensure a complete recovery, and questioned
-tropical forest community resilience after repeated disturbances.
+Anticipate the impact of the global change context on tropical forests
+is now urgent. This calls to highlight the taxonomic and functional
+facets of tree community post-disturbance trajectories and the
+underlying processes. It also calls to clarify the link between
+biodiversity and disturbance, specifically the scope of the Intermediate
+Disturbance Hypothesis (IDH) that is debated in tropical forests. In
+this study, we analyzed the tree community trajectories following a
+disturbance gradient in a Neotropical forest over 30 years. We
+considered community taxonomic and functional diversity, composition and
+redundancy trajectories. We based on the annual botanical inventories of
+75 ha of a Neotropical forest and on large trait datasets comprising 7
+leaf, stem and life-history traits. We highlighted a decoupling between
+taxonomic trajectories, that differed among communities, and functional
+trajectories, remained similar and convergent. We explained this
+decoupling by the variations in community functional redundancy that
+mitigated the functional impact of disturbance. Post-disturbance
+trajectories clarified the scope of the Intermediate Disturbance
+Hypothesis that was validated regarding taxonomic diversity, with an
+intensity threshold of 20-25\%, but not regarding functional
+diversity.The IDH applied around 20 years after disturbance when the
+recruitment of pioneers and late-successionals were balanced. Although
+consistent, the recovery of community composition, diversity and
+redundancy remained unachieved after 30 years. This acknowledged the
+need of decades-long cycles without disturbance to ensure a complete
+recovery, and questioned tropical forest community resilience after
+repeated disturbances.
 }

### 2019-04-08 WholePlotTrajectories.Rmd
review Bruno

@@
 corrauthor:
-    email: ariane.mirabel@ecofog.gf
+    email: ariane.mirabel@gmail.com
     url: https://github.com/ArianeMirabel
@@
 Treatment T3 additionally comprised the logging of 15 trees /ha with 40 cm minimum DBH and the poison-gidling of 20 trees/ha with a 50 cm minimum DBH, all belonging to non-commercial species. 
-Considering the silvicultural treatments and the following damage, disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Biomass2018].
+Considering the silvicultural treatments and the following damage, disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Rejou2017].
 The three treatments constituted a disturbance intensity gradient with increasing of above-ground biomass (AGB) lost and surface disturbed.
@@
 names(Intervention)[5] <- "\\%AGB lost"
-knitr::kable(Intervention, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.", longtable = FALSE, booktabs = TRUE, escape = FALSE,format="latex") %>%
+knitr::kable(Intervention, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
+Treatment intensities are defined by the minimum logging DBH (Diameter at Breast Height), the type of logged species (commercial or not), the density of logged trees, and the total AGB (Above Ground Biomass) lost after treatment.", longtable = FALSE, booktabs = TRUE, escape = FALSE,format="latex") %>%
   kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
@@
 
-Six functional traits representing leaf economics (leaf thickness, toughness, total chlorophyll content and specific leaf area) and stem economics (wood specific gravity and bark thickness), and life-history traits (maximum specific height and seed mass) came from the BRIDGE project [^1].
+Six functional traits representing leaf economics (leaf thickness, toughness, total chlorophyll content and specific leaf area) and stem economics (wood specific gravity and bark thickness) came from the BRIDGE project [^1].
 Trait values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 species pertaining to 157 genera. 
-Missing trait values (10%) were filled using multivariate imputation by chained equation [@Mice2011].
-Imputations were restricted within genus or family when samples were too scarce, in order to account for the phylogenetic signal. 
+Whenever a species was in the dataset but missed some trait values (10% of the species), missing values were filled using multivariate imputation by chained equation [@Mice2011].
+To account for the phylogenetic signal in the filling process, imputations based on samples from the same genus or from the same family. 
 Whenever a species was not in the dataset, it was attributed a set of trait values randomly sampled among species of the next higher taxonomic level (same genus or family). 
-As seed mass information was classified into classes, no data filling process was applied and analyses were restricted to the 414 botanical species recorded.
+Two life-history traits (maximum specific height and seed mass) came from the Mariwenn database [^2]. 
+The database compiles information from a vast literature on the flora of French Guiana [@Ollivier2007] and comprises 362 species pertaining to 188 genera.
+As seed mass information was classified into classes, no data filling process was applied and analyses were restricted to the botanical species recorded.
 
@@
 [^1]: http://www.ecofog.gf/Bridge/ 
+[^2]: https://www.ecofog.gf/mariwenn/
 
@@
 Two NMDS using abundance-based (Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the latter based on the seven leaf, stem and life history traits (without seed mass classes).
-Trajectories along time were reported through the Euclidean distance between the target inventories and the reference inventories in 1989, *i.e.* 2 years after disturbance. 
+Trajectories along time were reported through the Euclidean distance between the target inventories and the reference pre-disturbance inventories of 1984.
 Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007]. 
@@
 
+The taxonomic and functional trajectories were analysed from the reference pre-disturbance inventories of 1984.
 The taxonomic diversity was reported through species richness and the Hill number translation of the Simpson index [@Hill1973].
-These metrics allowed assessing the taxonomic richness as well as evenness, through the comparison between these two metrics.: results will be discussed directly in terms of taxonomic richness and evenness.
-Both indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recommended for diversity studies [@Marcon2015b]. 
+These metrics allowed assessing the taxonomic richness as well as evenness, through the comparison between these two metrics: results will be discussed directly in terms of taxonomic richness and evenness.
+Both indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recommended for diversity studies [@Marcon2015]. 
 The functional diversity was reported using the functional richness and functional evenness, *i.e* the Rao index of quadratic entropy.
@@
 
-From 1989 (2 years after disturbance) to 2015 (28 years after disturbance), 828 388 individual trees and 591 botanical species pertaining to 223 genera and 64 families were recorded.
+From 1984, the first pre-disturbance inventory, to 2015 (28 years after disturbance), 828 388 individual trees and 591 botanical species pertaining to 223 genera and 64 families were recorded.
 
@@
 EuclidDist(TaxoEuclid)
-mtext("Euclidean distance\nfrom 1985 inventory",side=2,padj=0,line=2,cex=0.8)
+mtext("Euclidean distance\nfrom 1984 inventory",side=2,padj=0,line=2,cex=0.8)
 mtext("(c)",side=3,adj=0,line=0.5)
@@
 
-```{r DivTaxo, out.width = '100%', echo=FALSE, fig.cap="Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic richness (ref:Bolda), Simpson diversity (ref:Boldb), functional richness (ref:Boldc), and Rao diversity (ref:Boldd).
+```{r DivTaxo, out.width = '100%', echo=FALSE, fig.cap="Trajectories of community taxonomic richness (ref:Bolda), Simpson diversity (ref:Boldb), functional richness (ref:Boldc), and Rao diversity (ref:Boldd).
+Values correspond to the difference over 30 years of community diversity with the values of reference 1984 pre-disturbance inventories.
 Shaded areas are the credibility intervals ", fig.env="figure*"}
@@
 
+Our analysis highlighted the decoupling between functional and taxonomic  trajectories.
+While initial differences in community taxonomic composition were maintained after disturbance, their trajectories in functional composition were similar.
+In parallel, while the different taxonomic diversity trajectories displayed marked differences according to the disturbance intensity, the functional diversity trajectories were homogeneous whenever the disturbance intensity.
+This decoupling was accompanied by a decrease in community functional redundancy which appeared to be the key element of community post-disturbance recovery.
+
 ## Decoupled taxonomic and functional trajectories
@@
 These initial differences were maintained all along the 30 years following disturbance, with the disturbance leading a displacement on the NMDS axis 1.
-The taxonomic compositional changes following disturbance were similar among communities, which may correspond to the recruitment of a group of pioneers common to all communities whenever their initial differences and the intensity of disturbance [@Denslow2000;@Bongers2009].
+The taxonomic compositional changes following disturbance were similar among communities, which may correspond to the recruitment of a group of pioneers, as *Cecropia* spp., *Vismia* spp., common to all communities whenever their initial differences and the intensity of disturbance [@Denslow2000;@Bongers2009].
 Taxonomic trajectories continued with a recovery of the initial composition which, althought not fully achieved after 30 years, suggested the resilience of the taxonomic composition resilience and the maintenance of the initial differences among communities [@Folke2006].
@@
 
-Community taxonomic richness and evenness trajectories were determined by the disturbance intensity, ranging from a limited and temporary impact to significant and persistant alterations of community diversity.
-The taxonomic trajectories markedly changed above an intensity threshold at which the taxonomic richness was maximized and the taxonomic evenness remained resilient. 
+Comparing community diversity along time after disturbance revealed constrasting trajectories according to the distrubance intensity.
+Below a disturbance intensity threshold, in the case of treatments 1 and 2, the community taxonomic richness increased.
+In parallell, after a first increase, the taxonomic eveness stabilized or strated to decrease after 30 years.
+Below the intensity threshold 
+Contrastingly, above an intensity threshold, in the case of treatment 3, the taxonomic richness did not exceeded the initial value and the taxonomic evenness continiously increased for 30 years without sign of return towards the initial value.
+Below and above a threshold, corresponding to a disturbance intensity between treatments 2 and 3, community taxonomic richness and evenness changed markedly.
+A the disturbance intensity threshold, the taxonomic richness was maximized and the taxonomic evenness remained resilient. 
 The disturbance intensity impacted the balance in the community between pre-disturbance surviving trees, and trees recruited afterward. 
@@
 Taxonomic trajectories following disturbance of intermediate intensity besides highlighted an intermediate time after disturbance, around 15-20 years, for which the taxonomic evenness was maximized.
-As already observed in the Guiana shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998],
-taxonomic evenness followed hump-shaped trajectories with a peak when the recruitment was balanced between pioneers and late-successional species. 
+As already observed in the Guiana shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998], taxonomic evenness followed hump-shaped trajectories with a peak when the recruitment was balanced between pioneers and late-successional species.
 This intermediate time corresponded to the match between the generation time of pionneers, dominating first the post-disturbance recruitment, and the time for competitive exclusion to emerge, leading afterwards to the dominance of late-successional species.
@@
 
-The decoupling between taxonomic and functional trajectories was explained by a decrease in the functional redundancy within the pre-disturbance functional space, due to the loss of species following disturbance.
-The recruitment of pioneers, functionally different from the pre-disturbance functional composition, did not compensate the decrease of functional redundancy in the first place.
-Progressively though, the functional redundancy was restored through the replacement of the first established species by more competitive, long-lived pioneers or late-successional species that were functionally closer to the pre-disturbance community.
+The decoupling between taxonomic and functional trajectories was accompanied by a decrease in the functional redundancy within the pre-disturbance functional space.
+This decrease was explained by a loss of species following disturbance and then a lower redundancy of their functional characteristics.
+This decrease of the functional redundancy was not compensated by the recruitment of pioneers, as those were functionally different from the pre-disturbance community.
+Progressively though, more competitive,long-lived pioneers or late-successional species functionally closer to the pre-disturbance community progressively replaced  the first established species and restored the functional redundancy.
 This replacement followed the lottery recruitment rules, implying an easy recruitment for the first species but becoming increasingly hampered by the emergence of interspecific competition [@Busing2002].
@@
 While functional trajectories remained similar in the functional space and converged towards the recovery of a comparable initial state, taxonomic trajectories showed initial differences in community composition that were maintained along time.
-This decoupling was explained by a decrease of the functional redundancy that mitigated the functional impact of disturbance.
-Community functional and taxonomic diversity response to disturbance was contrasted as well.
+Community high functional redundancy could explain this decoupling as the loss of a species does not necessarily the loss of its functional charactristics in the community.
+In this sense, the functional redundancy would mitigate the functional impact of disturbance.
+Community functional and taxonomic diversity response to disturbance were contrasted as well.
 While the functional trajectories remained similar whenever the disturbance intensity, taxonomic trajectories were markedly different above a threshold of 20-25% AGB removed for which taxonomic richness was maximized and taxonomic eveness remained resilient.
@@
 Although resilient, the functional redundancy was lower for more than 30 years after disturbance which cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].
+Our conclusions first highlighted risks to lose rare species with visible impacts only in the long term, and led us to advice paying particular attention to rare and undangered species and to conduct regular inventories during exploitation times.
+Second, post-disturbance trajectories seemed to greatly rely on post-disturbance recruitment dynamics that would be key to model post-disturbance trajectories.
 

### 2019-04-08 WholePlotTrajectories.tex
review Bruno

@@
 }
-\affiliation{*\textbf{Corresponding author}: ariane.mirabel@ecofog.gf, https://github.com/ArianeMirabel} % Corresponding author
+\affiliation{*\textbf{Corresponding author}: ariane.mirabel@gmail.com, https://github.com/ArianeMirabel} % Corresponding author
 
@@
 \citep{Piponiot2016} estimated with the BIOMASS R package
-\citep{Biomass2018}. The three treatments constituted a disturbance
+\citep{Rejou2017}. The three treatments constituted a disturbance
 intensity gradient with increasing of above-ground biomass (AGB) lost
@@
 
-\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.}
+\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
+Treatment intensities are defined by the minimum logging DBH (Diameter at Breast Height), the type of logged species (commercial or not), the density of logged trees, and the total AGB (Above Ground Biomass) lost after treatment.}
 \centering
@@
 toughness, total chlorophyll content and specific leaf area) and stem
-economics (wood specific gravity and bark thickness), and life-history
-traits (maximum specific height and seed mass) came from the BRIDGE
-project \footnote{http://www.ecofog.gf/Bridge/}. Trait values were
-assessed from a selection of individuals located in nine permanent plots
-in French Guiana, including two in Paracou, and comprised 294 species
-pertaining to 157 genera. Missing trait values (10\%) were filled using
-multivariate imputation by chained equation \citep{Mice2011}.
-Imputations were restricted within genus or family when samples were too
-scarce, in order to account for the phylogenetic signal. Whenever a
-species was not in the dataset, it was attributed a set of trait values
-randomly sampled among species of the next higher taxonomic level (same
-genus or family). As seed mass information was classified into classes,
-no data filling process was applied and analyses were restricted to the
-414 botanical species recorded.
+economics (wood specific gravity and bark thickness) came from the
+BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}. Trait values
+were assessed from a selection of individuals located in nine permanent
+plots in French Guiana, including two in Paracou, and comprised 294
+species pertaining to 157 genera. Whenever a species was in the dataset
+but missed some trait values (10\% of the species), missing values were
+filled using multivariate imputation by chained equation
+\citep{Mice2011}. To account for the phylogenetic signal in the filling
+process, imputations based on samples from the same genus or from the
+same family. Whenever a species was not in the dataset, it was
+attributed a set of trait values randomly sampled among species of the
+next higher taxonomic level (same genus or family). Two life-history
+traits (maximum specific height and seed mass) came from the Mariwenn
+database \footnote{https://www.ecofog.gf/mariwenn/}. The database
+compiles information from a vast literature on the flora of French
+Guiana \citep{Ollivier2007} and comprises 362 species pertaining to 188
+genera. As seed mass information was classified into classes, no data
+filling process was applied and analyses were restricted to the
+botanical species recorded.
 
@@
 were reported through the Euclidean distance between the target
-inventories and the reference inventories in 1989, \emph{i.e.} 2 years
-after disturbance. Univariate trajectories of the leaf, stem and
-life-history traits were also visualized with the community weighted
-means (CWM) \citep{Diaz2007}. Species seed mass were given in 5 mass
-classes, and seed mass trajectories were reported as the proportion of
-each class in the inventories (Supplementary materials).
+inventories and the reference pre-disturbance inventories of 1984.
+Univariate trajectories of the leaf, stem and life-history traits were
+also visualized with the community weighted means (CWM)
+\citep{Diaz2007}. Species seed mass were given in 5 mass classes, and
+seed mass trajectories were reported as the proportion of each class in
+the inventories (Supplementary materials).
 
-The taxonomic diversity was reported through species richness and the
-Hill number translation of the Simpson index \citep{Hill1973}. These
-metrics allowed assessing the taxonomic richness as well as evenness,
-through the comparison between these two metrics.: results will be
-discussed directly in terms of taxonomic richness and evenness. Both
-indices belong to the set of HCDT or generalized entropy, respectively
-corresponding to the 0 and 2 order of diversity (q), recommended for
-diversity studies \citep{Marcon2015b}. The functional diversity was
-reported using the functional richness and functional evenness,
-\emph{i.e} the Rao index of quadratic entropy. The Rao index combines
-species abundance distribution, and the average pairwise functional
-dissimilarity between species computed by the gower distance.
+The taxonomic and functional trajectories were analysed from the
+reference pre-disturbance inventories of 1984. The taxonomic diversity
+was reported through species richness and the Hill number translation of
+the Simpson index \citep{Hill1973}. These metrics allowed assessing the
+taxonomic richness as well as evenness, through the comparison between
+these two metrics: results will be discussed directly in terms of
+taxonomic richness and evenness. Both indices belong to the set of HCDT
+or generalized entropy, respectively corresponding to the 0 and 2 order
+of diversity (q), recommended for diversity studies \citep{Marcon2015}.
+The functional diversity was reported using the functional richness and
+functional evenness, \emph{i.e} the Rao index of quadratic entropy. The
+Rao index combines species abundance distribution, and the average
+pairwise functional dissimilarity between species computed by the gower
+distance.
 
@@
 
-From 1989 (2 years after disturbance) to 2015 (28 years after
+From 1984, the first pre-disturbance inventory, to 2015 (28 years after
 disturbance), 828 388 individual trees and 591 botanical species
@@
 
-\caption{Trajectories over 30 years of the difference with the 1989 inventory (2 years after disturbance) of community taxonomic richness \textbf{(a)}, Simpson diversity \textbf{(b)}, functional richness \textbf{(c)}, and Rao diversity \textbf{(d)}.
+\caption{Trajectories of community taxonomic richness \textbf{(a)}, Simpson diversity \textbf{(b)}, functional richness \textbf{(c)}, and Rao diversity \textbf{(d)}.
+Values correspond to the difference over 30 years of community diversity with the values of reference 1984 pre-disturbance inventories.
 Shaded areas are the credibility intervals }\label{fig:DivTaxo}
@@
 
+Our analysis highlighted the decoupling between functional and taxonomic
+trajectories. While initial differences in community taxonomic
+composition were maintained after disturbance, their trajectories in
+functional composition were similar. In parallel, while the different
+taxonomic diversity trajectories displayed marked differences according
+to the disturbance intensity, the functional diversity trajectories were
+homogeneous whenever the disturbance intensity. This decoupling was
+accompanied by a decrease in community functional redundancy which
+appeared to be the key element of community post-disturbance recovery.
+
 \subsection{Decoupled taxonomic and functional
@@
 disturbance were similar among communities, which may correspond to the
-recruitment of a group of pioneers common to all communities whenever
-their initial differences and the intensity of disturbance
+recruitment of a group of pioneers, as \emph{Cecropia} spp.,
+\emph{Vismia} spp., common to all communities whenever their initial
+differences and the intensity of disturbance
 \citep{Denslow2000, Bongers2009}. Taxonomic trajectories continued with
@@
 
-Community taxonomic richness and evenness trajectories were determined
-by the disturbance intensity, ranging from a limited and temporary
-impact to significant and persistant alterations of community diversity.
-The taxonomic trajectories markedly changed above an intensity threshold
-at which the taxonomic richness was maximized and the taxonomic evenness
-remained resilient. The disturbance intensity impacted the balance in
-the community between pre-disturbance surviving trees, and trees
-recruited afterward. On the one hand, below a 20\%-25\% AGB loss, the
-trees surviving after disturbance remained numerous enough to maintain
-the high taxonomic richness of the pre-disturbance community
+Comparing community diversity along time after disturbance revealed
+constrasting trajectories according to the distrubance intensity. Below
+a disturbance intensity threshold, in the case of treatments 1 and 2,
+the community taxonomic richness increased. In parallell, after a first
+increase, the taxonomic eveness stabilized or strated to decrease after
+30 years. Below the intensity threshold Contrastingly, above an
+intensity threshold, in the case of treatment 3, the taxonomic richness
+did not exceeded the initial value and the taxonomic evenness
+continiously increased for 30 years without sign of return towards the
+initial value. Below and above a threshold, corresponding to a
+disturbance intensity between treatments 2 and 3, community taxonomic
+richness and evenness changed markedly. A the disturbance intensity
+threshold, the taxonomic richness was maximized and the taxonomic
+evenness remained resilient. The disturbance intensity impacted the
+balance in the community between pre-disturbance surviving trees, and
+trees recruited afterward. On the one hand, below a 20\%-25\% AGB loss,
+the trees surviving after disturbance remained numerous enough to
+maintain the high taxonomic richness of the pre-disturbance community
 \citep{Bongers2009}. The recruitment of pioneers, infrequent or absent
@@
 The decoupling between taxonomic and functional trajectories was
-explained by a decrease in the functional redundancy within the
-pre-disturbance functional space, due to the loss of species following
-disturbance. The recruitment of pioneers, functionally different from
-the pre-disturbance functional composition, did not compensate the
-decrease of functional redundancy in the first place. Progressively
-though, the functional redundancy was restored through the replacement
-of the first established species by more competitive, long-lived
-pioneers or late-successional species that were functionally closer to
-the pre-disturbance community. This replacement followed the lottery
-recruitment rules, implying an easy recruitment for the first species
-but becoming increasingly hampered by the emergence of interspecific
-competition \citep{Busing2002}. The recovery of the functional
-redundancy then relied upon the random process of species recruitment
-and was increasingly slow and difficult to anticipate
+accompanied by a decrease in the functional redundancy within the
+pre-disturbance functional space. This decrease was explained by a loss
+of species following disturbance and then a lower redundancy of their
+functional characteristics. This decrease of the functional redundancy
+was not compensated by the recruitment of pioneers, as those were
+functionally different from the pre-disturbance community. Progressively
+though, more competitive,long-lived pioneers or late-successional
+species functionally closer to the pre-disturbance community
+progressively replaced the first established species and restored the
+functional redundancy. This replacement followed the lottery recruitment
+rules, implying an easy recruitment for the first species but becoming
+increasingly hampered by the emergence of interspecific competition
+\citep{Busing2002}. The recovery of the functional redundancy then
+relied upon the random process of species recruitment and was
+increasingly slow and difficult to anticipate
 \citep{Elmqvist2003, Diaz2005}.
@@
 trajectories showed initial differences in community composition that
-were maintained along time. This decoupling was explained by a decrease
-of the functional redundancy that mitigated the functional impact of
+were maintained along time. Community high functional redundancy could
+explain this decoupling as the loss of a species does not necessarily
+the loss of its functional charactristics in the community. In this
+sense, the functional redundancy would mitigate the functional impact of
 disturbance. Community functional and taxonomic diversity response to
-disturbance was contrasted as well. While the functional trajectories
+disturbance were contrasted as well. While the functional trajectories
 remained similar whenever the disturbance intensity, taxonomic
@@
 species loss and the persistence of disturbance-specific communities
-\citep{Herault2018}.
+\citep{Herault2018}. Our conclusions first highlighted risks to lose
+rare species with visible impacts only in the long term, and led us to
+advice paying particular attention to rare and undangered species and to
+conduct regular inventories during exploitation times. Second,
+post-disturbance trajectories seemed to greatly rely on post-disturbance
+recruitment dynamics that would be key to model post-disturbance
+trajectories.
 

### 2019-04-09 WholePlotTrajectories.Rmd
relecture

@@
 They regulate as well the local and regional climates, and the carbon, water and nutrient cycles.
-However, the growing demand in forests products together with current global changes increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
+However, the growing demand in forests products together with current global changes increase the pressure on remaining natural undisturbed forests [@Morales-Hidalgo2015].
 These threats affect the disturbance regime that naturally shapes and maintains the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
-To anticipate the fate of tropical forests in the current context it is then urgent to understand the response of communities to disturbance and the underlying processes.
-Disturbance impact forest communities in changing both abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
-Although in the tropical context the forest environment is maintained after disturbance, there is a succession of changes in the structure and functioning of communities constituting a post-disturbance secondary succession.
-For now, this succession has been largely studied through trajectories of forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
-Models based on observed trajectories assessed community time to recovery and identified some of the determinants of post-disturbance trajectories, like forest structure, forest composition, environmental parameters and disturbance characteristics [@Herault2018].
-Regarding community diversity and composition however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
-Given the variety of species response to disturbance and the huge diversity of tropical forests, manifold post-disturbance biodiversity trajectories might emerge [@Lindenmayer2012;@Garcia_florez2017].
-It is then essential to anticipate post-disturbance trajectories and highlight the link between biodiversity and disturbance, specifically the link between  disturbance intensity and frequency and the rate, endpoint and duration of the trajectories [@Chazdon2003a;@Willig2018].
+To anticipate the fate of tropical forests in this context, understanding the response of communities to disturbance and deciphering the underlying ecological processes is of utmost importance.
+Disturbance impacts forest communities in changing both abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
+Although the forest cover is generally maintained after disturbance, there is a succession of changes in the structure and functioning of ecological communities that defines a post-disturbance ecosystem trajectory.
+For now, these trajectories have been largely studied through trajectories of forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
+Recent works have identified some of the determinants of post-disturbance biomass trajectories, like pre-disturbance forest structure and composition or post-disturbance environmental parameters [@Herault2018].
+Regarding tree community diversity and composition however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001]. 
+Given the variety of species response to disturbance and the huge diversity of tropical forests, manifold post-disturbance biodiversity trajectories might
+emerge [@Lindenmayer2012;@Garcia_florez2017].
 
@@
 In tropical forests, however, observations of the IDH often diverge from theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
-The link between biodiversity and disturbance might be complicated by the huge diversity of tropical communities and a variety of ecological processes may underlie the IDH (colonisation, facilitation, coexistence, etc)[@Lindenmayer2012;@Garcia_florez2017].
-There is a need to clarify the scope of the IDH in terms of the different facets of community diversity that are taxonomic richness, species abundance distribution and community functioning.
-Besides, there is a need to delineate the scope of the IDH in time, and highlight the moment when the conditions for the IDH emerge [@Sheil2003;@Shea2004].
+
+Moreover, we argue that there is a need to delineate the scope of the IDH in time, using ecosystem time-explicit trajectories, and highlight the moment when the conditions for the IDH emerge [@Sheil2003;@Shea2004].
+Indeed, the link between biodiversity, IDH
+and post-disturbance disturbance trajectories might be complicated by the huge diversity found in tropical
+tree communities and the variety of ecological processes that may underlie the IDH
+(colonization, facilitation, coexistence,
+etc)[@Lindenmayer2012;@Garcia_florez2017].
 
 Analysing community response to disturbance requires a set of metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
-The analysis should first consider community composition, which is crucial for conservation issues and which reveals the species fostered by the disturbance [@Lavorel2002;@Bellwood2006].
-Then, diversity metrics encompassing taxonomic richness and evenness should be considered to assess the changes in community abundance distribution. 
-Besides, a functional approach that accounts for species biological attributes and ecological strategy would directly link community diversity, composition and redundancy to ecosystem functioning and environmental constraints [@Violle2007b;@Baraloto2012a].
+The analysis should first consider community composition, which is crucial for conservation issues and which reveals the pool of species fostered (or hampered) by disturbance [@Lavorel2002;@Bellwood2006].
+Then, diversity metrics encompassing both richness and evenness should also be considered to assess the changes in community abundance distribution. 
+Besides, functional approaches have been shown to usefully complement pure taxonomic approaches because they shed light on the species biological attributes that directly link community diversity, composition and redundancy to ecosystem functioning [@Violle2007b;@Baraloto2012a].
 In that respect, a vast literature allowed recognizing major traits that represent species ecological strategy and determine how they respond to changing conditions [@Diaz2005]. 
-Specifically in tropical forests, the functional approach revealed the post-disturbance deterministic processes entailing a shift from a dominance of “conservative” slow-growing species dealing with scarce resources, to a dominance of “acquisitive” fast-growing species with rapid and efficient use of abundant resources [@Rees2001;@Reich2014;@Herault2011].
+Specifically in tropical forests, the functional approach revealed the post-disturbance deterministic processes entailing a shift from a dominance of "conservative" slow-growing species dealing with scarce resources, to a dominance of "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Rees2001;@Reich2014;@Herault2011].
 This shift is translated into the trajectories of average community value of key functional traits related to resource acquisition (leaf and stem traits) and life-history strategy (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
+
 The functional approach also encompasses the analysis of functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016]. 
@@
 
-In this study, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60% of above-ground biomass (AGB) loss.
+In this study, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60\% of above-ground biomass (AGB) loss.
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, (i) we defined taxonomic and functional post-disturbance trajectories and examined the underlying ecological process, (ii) we discussed the scope of the IDH regarding the different facets of community diversity and the time after disturbance, and (iii) we analyzed community resilience and time to recovery.
+Specifically, (i) we draw taxonomic and functional post-disturbance trajectories and examine the underlying ecological process, (ii) we discuss the scope of the IDH regarding the different facets of community diversity and the time after disturbance, and (iii) we analyze community resilience and time to recovery.
 
@@
 Paracou station in French Guiana (5\textdegree 18'N and 52\textdegree 53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26\textdegree C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
-The mean annual temperature is 26\textdegree C. 
 Elevation ranges from 5 to 50 m.
@@
 Considering the silvicultural treatments and the following damage, disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Rejou2017].
-The three treatments constituted a disturbance intensity gradient with increasing of above-ground biomass (AGB) lost and surface disturbed.
+The three treatments were then transformed into a continuous disturbance intensity gradient with increasing of above-ground biomass (AGB) lost.
 
@@
 In the 12 experimental plots, all trees above 10 cm DBH have been mapped and measured annually since 1984. 
-Trees are first identified with a vernacular name assigned by the forest worker team, and afterward with a scientific name assigned by botanists during regular botanical campaigns. 
+Trees are first identified with a vernacular name assigned by the forest worker team, and afterward with a
+scientific name assigned by botanists during regular botanical campaigns. 
 In 1984, specific vernacular names were given to 62 commercial or common species whereas more infrequent ones were identified under general identifiers only distinguishing trees and palms. 
@@
 Both indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recommended for diversity studies [@Marcon2015]. 
-The functional diversity was reported using the functional richness and functional evenness, *i.e* the Rao index of quadratic entropy.
-The Rao index combines species abundance distribution, and the average pairwise functional dissimilarity between species computed by the gower distance.
+The functional diversity was reported using the functional richness and functional evenness, through the Rao index of quadratic entropy.
+The Rao index combines species abundance distribution, and the average pairwise functional dissimilarity between species computed by the Gower distance.
 
-The impacts of initial disturbance were tested with the Spearman rank correlation between the extremes of taxonomic and functional metrics reached over the 30 years and the initial %AGB loss. 
-They were besides analyzed through polynomial regression between (i) taxonomic and functional richness and evenness and (ii) the initial %AGB loss at 10, 20 and 30 years after disturbance.
+The impacts of the initial disturbance levels on the maximum gain or loss in richness and evenness were tested with the Spearman rank correlation tests. 
+Richness and evenness trajectories were besides analyzed through polynomial regression between (i) taxonomic and
+functional richness and evenness and (ii) the initial \%AGB loss at 10, 20 and 30 years after disturbance.
 
-Functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
+Finally, functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
  First, the individuals of the trait database were mapped in the plane of the first two axes from a PCA analysis, which lowered the weight of correlations among traits as the PCA axes were combinations of most decoupled traits.
 For each species, the traits probability density (TPD) were computed from the mapping of individuals through two-dimension kernel density estimators.
-Second, for each community, the TPD weigthed by species abundance were summed accross the functional space. 
-Third, the functional space was divided into a 100 x 100 grid and the number of species with a positive TPD was counted in each cell. 
-The average count across cells minus 1 returned the Community Functional Redundancy, which was the average number of species in the community that share the same trait values.
+Second, for each community, the TPD weigthed by species
+abundance were summed accross the functional space. Third, the functional space was divided into a 100 x 100 grid and the number of species with a positive TPD was counted in each cell. 
+The average count across cells minus 1 returned the Community Functional Redundancy, which was the average number of species in the community that share the same
+trait values.
 
@@
 
-While both taxonomic and functional composition remained stable in undisturbed communities (Fig.
-\@ref(fig:NMDSplans)), they followed marked and consistent trajectories over post-\break disturbance time.
-In disturbed communities the taxonomic composition changed towards more species with acquisitive functional strategies.
-This translated into a switch from high mean WSG in the community to high mean SLA and chlorophyll content (see appendix I). 
-The functional composition of disturbed communities corresponded to cyclic compositional changes with an incomplete recovery of the initial composition (Fig. \@ref(fig:NMDSplans)).
+While both taxonomic and functional composition remained stable in undisturbed communities (Fig. \@ref(fig:NMDSplans)), they followed marked and consistent trajectories in disturbed plots over time.
 
+The functional composition trajectory in disturbed communities resembled to cyclic compositional changes with an incomplete recovery to the initial composition (Fig.\ref{fig:NMDSplans}). 
 The maximum dissimilarity with the initial state was positively correlated with the disturbance intensity for both taxonomic and functional composition ($\rho_{Spearman}^{Taxonomic}=0.87$ and $\rho_{Spearman}^{Functional}=0.90$ respectively).
-The maximum dissimilarity with the initial state was reached for taxonomic composition between 15 to 25 years, for most of the plots, and around 22 years for functional composition.
+The maximum dissimilarity with the initial state was reached for taxonomic composition between 15 to 25 years and around 20 to 25 years for functional composition.
 
@@
 
-Community CWM average value of all traits and seed mass proportions followed unimodal trajectories, either stabilizing or returning towards their initial values, to the exception of leaf chlorophyll content, which continued to increase for some T3 and T2 plots 30 years after disturbance.
-
-Community CWM average value of Maximum height at adult stage (*Hmax*), leaf toughness and wood specific gravity (*WSG*) decreased and then slightly increased, but remained significantly lower than their initial value (Fig.
-\@ref(fig:CWM)).
-On the other side, bark thickness and specific leaf area (*SLA*) both increased.
-Bark thickness remained substantially high after 30 years, and *SLA* had almost recovered to its initial value. 
-For all traits, the maximum difference to initial value was correlated to the disturbance intensity.
-Positive correlations were observed for Leaf thickness, chlorophyll content, SLA and bark thickness ($\rho_{Spearman}^{Leaf thickness}=0.76$, $\rho_{Spearman}^{Chlorophyll content}=0.60$, $\rho_{Spearman}^{SLA}=0.93$, $\rho_{Spearman}^{Bark thickness}=0.71$).
-Negative correlation was observed for Leaf toughness, WSG and Hmax ($\rho_{Spearman}^{Leaf toughness}=-0.53$, $\rho_{Spearman}^{WSG}=-0.75$, $\rho_{Spearman}^{Hmax}=-0.40$)
+In control plots, Community Weighted Means (CWM) of functional traits remain stable in time.  
+In disturbed plots, they mostly followed unimodal trajectories, either stabilizing or returning towards
+their initial values, to the exception of leaf chlorophyll content, which continued to increase 30 years after disturbance for some highly disturbed plots.
+Maximum height at adult stage (*Hmax*), leaf toughness and wood specific gravity (*WSG*) decreased in time and then slightly increased, but remained significantly lower than their initial value (Fig.
+\ref{fig:CWM}).
+Bark thickness and specific leaf area (*SLA*) both increased in time. 
+Bark thickness remained substantially high after 30 years, and *SLA* had almost recovered to its initial value. Whatever the functional traits, the maximum
+difference to initial value was highly correlated to the disturbance intensity.
+Positive correlations were observed for Leaf thickness, chlorophyll content, SLA and bark thickness ($\rho_{Spearman}^{Leaf thickness}=0.76$, $\rho_{Spearman}^{Chlorophyll content}=0.60$,
+$\rho_{Spearman}^{SLA}=0.93$, $\rho_{Spearman}^{Bark thickness}=0.71$). 
+Negative correlation was observed for Leaf toughness, WSG and Hmax ($\rho_{Spearman}^{Leaf toughness}=-0.53$,
+$\rho_{Spearman}^{WSG}=-0.75$, $\rho_{Spearman}^{Hmax}=-0.40$)
 The proportions of the three lightest seed mass classes increased in all disturbed plots. 
@@
 
-For undisturbed plots, taxonomic richness and Simpson diversity remained stable over the 30 years of monitoring.
-In disturbed communities, after low disturbance intensity the taxonomic richness increased, reaching a maximum gain of 14 botanical genera (plot 3 from treatment 2). 
+Taxonomic richness and evenness remained stable in control plots over the 30 years of monitoring. 
+In disturbed communities, after low disturbance intensity the taxonomic richness increased, reaching a maximum gain of 14 botanical genera. 
 After intense disturbance the taxonomic richness followed a more complex trajectory, decreasing for ten years after disturbance before recovering to pre-disturbance values.
@@
 
-In all disturbed plots the Simpson diversity first increased until a maximum reached after around 20 years. 
+In all disturbed plots the evenness first increased until a maximum reached after around 20 years. 
 This maximum was positively correlated with the disturbance intensity ($\rho_{Spearman}^{Simpson}=0.77$).
-The Simpson diversity then stabilized except for two T3 plots (plots 8 and 12) for which it kept increasing (Fig. \@ref(fig:DivTaxo)).
+The evenness then stabilized except for two intensively-disturbed plots (number 8 and 12) for which it kept increasing \@ref(fig:DivTaxo)).
 
@@
 
-The plot 7 from treatment 1 displayed constantly outlying functional richness and Rao diversity and was removed from the graphical representation for better readability.
-In undisturbed plots both functional richness and Rao diversity remained stable along the 30 years. 
-In disturbed plots, both trajectories depended on the disturbance intensity, with their maximum being positively correlated to %AGB loss $\rho_{Spearman}^{Richness}=0.76$ and $\rho_{Spearman}^{Rao}=0.60$.
-Functional richness and Rao diversity displayed for low disturbance intensity a low but long-lasting increase up to a maximum reached after 20-25 years.
-Following high intensity, they displayed a fast but short increase followed after 10 years by a slow decrease towards the initial values.
+Functional richness and evenness remained stable in control plots over the 30 years of monitoring. 
+In disturbed communities, both trajectories depended on the disturbance intensity, with their maximum values in time being positively correlated to \%AGB loss $\rho_{Spearman}^{Richness}=0.76$ and $\rho_{Spearman}^{Rao}=0.60$.
+Functional richness and evenness displayed for low disturbance intensity a low but long-lasting increase up to a maximum reached after 20-25 years. 
+For high disturbance intensity, they generally displayed a fast but short increase followed after 10 years by a slow decrease towards the initial values.
 
-The second-degree polynomial regressions between (i) the percentage AGB loss and (ii) taxonomic and functional diversity after 10, 20 and 30 years best predicted the hump-shaped curve of the disturbance impact along the disturbance intensity gradient (Fig.
-\@ref(fig:IDHplot)).
-The relationship between the disturbance impact and its intensity was more markedly hump-shaped for the taxonomic richness than for the Simpson diversity. 
-For both functional richness and Rao diversity the relationship was almost linear.
-The regression model better predicted the functional richness and Rao diversity ($0.55<R^2_{Functional Richness}<0.72$, and $0.60<R^2_{Functional Rao}<0.81$) than the taxonomic richness and evenness ($0.21<R^2_{Taxonomic Richness}<0.4$, and $-0.15<R^2_{Taxonomic Simpson}<0.43$ respectively).
+The second-degree polynomial regressions between (i) the percentage AGB loss and (ii) the taxonomic and functional diversity showed various shape depending on the diversity indices and on the time since disturbance  (Fig.
+\ref{fig:IDHplot}).
+Regarding taxonomic diversity, the relationship between disturbance intensity and diversity was more markedly hump-shaped for richness than for evenness and peaked at 20\% of initial AGB loss.
+Regarding functional diversity, the relationship was almost linear and similar between richness and evenness.
+Generally, all relationships were stronger 20 or 30 years after disturbance than observed for 10 years.
 
@@
 
-All disturbed plots had lower functional redundancy than control plots and followed similar hump-shaped trajectories (\@ref(fig:RedFunRest)). 
-The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{Spearman}=0.47$) and the recovery had not attained initial values for any disturbed communities after 30 years.
+Control plots displayed stable functional redundancy over time. 
+All disturbed plots had lower functional redundancy than control plots and followed similar hump-shaped trajectories (\ref{fig:RedFunRest}).
+The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{Spearman}=0.47$) and the recovery trajectory had not attained initial values for any disturbed communities after 30 years.
 
@@
 
-Our analysis highlighted the decoupling between functional and taxonomic  trajectories.
-While initial differences in community taxonomic composition were maintained after disturbance, their trajectories in functional composition were similar.
-In parallel, while the different taxonomic diversity trajectories displayed marked differences according to the disturbance intensity, the functional diversity trajectories were homogeneous whenever the disturbance intensity.
-This decoupling was accompanied by a decrease in community functional redundancy which appeared to be the key element of community post-disturbance recovery.
+Our analysis highlighted the decoupling between functional and taxonomic  trajectories, with taxonomic trajectories maintaining initial differences among plots while functional trajectories remained similar.
+We revealed that if the IDH explained the taxonomic response to disturbance, it was not validated regarding the functional response.
+The decoupling between taxonomic and functional response was linked to the disturbance impact on the functional redundancy, that proved determining in community recovery.
 
@@
 
-Before disturbance, the different communities had different taxonomic compositions, visualized by their distinct starting points on the NMDS axis 2.
-These initial differences were maintained all along the 30 years following disturbance, with the disturbance leading a displacement on the NMDS axis 1.
-The taxonomic compositional changes following disturbance were similar among communities, which may correspond to the recruitment of a group of pioneers, as *Cecropia* spp., *Vismia* spp., common to all communities whenever their initial differences and the intensity of disturbance [@Denslow2000;@Bongers2009].
-Taxonomic trajectories continued with a recovery of the initial composition which, althought not fully achieved after 30 years, suggested the resilience of the taxonomic composition resilience and the maintenance of the initial differences among communities [@Folke2006].
-The  return towards initial taxonomic composition showed that species not belonging to the pre-disturbance community were rarely recruited, probably because of the common dispersal limitations among tropical tree species [@Svenning2005].
+Regarding taxonomic composition, the different tree communities were substantially different before disturbance, and this is visualized by their distinct starting points on the NMDS axis 2. 
+These initial differences were maintained all along the 30 years following disturbance, with the disturbance leading a displacement on the NMDS axis 1 only. 
+Changes were thus similar and may correspond to the recruitment of a group of pioneers, e.g.
+*Cecropia* spp., *Vismia* spp.,  shared by all plots and whatever their initial taxonomic differences and the intensity of disturbance [@Denslow2000;@Bongers2009].
+Taxonomic trajectories seems to start a recovery to the initial composition which, although far from fully achieved after 30 years, suggested that taxonomic composition is resilient and that the initial differences will be maintained in time [@Folke2006].
+This would show that species not belonging to the pre-disturbance community were rarely recruited in the long-term, probably because of the common dispersal limitations among tropical tree species [@Svenning2005].
 
-Regarding functional composition, initial communities were similar and trajectories were confounded in the functional space.
-Following disturbance, the functional composition of pre-disturbance surviving trees is believed to be the same as this of the initial community [@Herault2018].
-The changes in functional composition following disturbance then relied upon the recruitment of species or functional types previously infrequent or absent.
-These species were probably competitive pioneers becoming dominant as the most efficient to benefit from the light, space and nutrients made available by the disturbance.
-This recruitment of pioneers led similar functional composition trajectories for all plots, whenever the disturbance intensity.
-The composition first went towards more resource-acquisitive strategies, which was translated by a displacement to the right along the first axis in the functional plane [@Westoby1998; @Wright2004; @Reich2014].
-Thereafter, the pioneers recruited primarily were progressively excluded by long-lived, more resistant and shade-tolerant species.
-The community functional composition then returned towards more resource-conservative strategies, suggesting the recovery of the initial functional composition.
-This recovery translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig. \@ref(fig:NMDSplans)).
+Regarding functional composition, initial communities had similar starting points and this contrasts to taxonomic composition. 
+This means that, despite differences in species composition, the functional signature of the tree communities were quite similar. 
+Following disturbance, functional trajectories were consistent in the NMDS plane with displacement intensities linked to disturbance intensities and thus relied upon the recruitment of a shared pool of functional types previously infrequent or absent.
+This common pool is composed of pioneer resource-acquisitive (low leaf toughness, wood specific gravity, maximum height and high Specific Leaf area) strategies that translated by a displacement to the right along the first NMDS axis [@Westoby1998;@Wright2004;@Reich2014].
+Thereafter, the pioneers recruited primarily were progressively excluded by long-lived, more competitive and shade-tolerant species. 
+The community functional composition then quickly returned towards more resource-conservative strategies, suggesting the recovery of the initial functional composition. 
+This recovery translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig. \ref{fig:NMDSplans}).
 
 Both taxonomic and functional composition trajectories initiated a return towards pre-disturbance state after 30 years, which highlighted the taxonomic and functional resilience of communities. 
-Taxonomicand functional trajectories however appeared decoupled: while taxonomic trajectories maintained the initial differences among communities, the functional trajectories where similar and convergent in the functional space [@Fukami2005].
+Taxonomic and functional trajectories however appeared decoupled: while taxonomic trajectories maintained the initial differences among communities, the functional trajectories where similar and convergent in the functional space [@Fukami2005] exemplifying the simultaneous operation of trait-based assembly rules and species-level priority effects that shapes tree community assembly in this forest, making it both deterministic in functional space and historically contingent in taxonomic space.
 
@@
 
-Comparing community diversity along time after disturbance revealed constrasting trajectories according to the distrubance intensity.
-Below a disturbance intensity threshold, in the case of treatments 1 and 2, the community taxonomic richness increased.
-In parallell, after a first increase, the taxonomic eveness stabilized or strated to decrease after 30 years.
-Below the intensity threshold 
-Contrastingly, above an intensity threshold, in the case of treatment 3, the taxonomic richness did not exceeded the initial value and the taxonomic evenness continiously increased for 30 years without sign of return towards the initial value.
-Below and above a threshold, corresponding to a disturbance intensity between treatments 2 and 3, community taxonomic richness and evenness changed markedly.
-A the disturbance intensity threshold, the taxonomic richness was maximized and the taxonomic evenness remained resilient. 
-The disturbance intensity impacted the balance in the community between pre-disturbance surviving trees, and trees recruited afterward. 
-On the one hand, below a 20%-25% AGB loss, the trees surviving after disturbance remained numerous enough to maintain the high taxonomic richness of the pre-disturbance community [@Bongers2009].
-The recruitment of pioneers, infrequent or absent before disturbance, then increased the taxonomic richness all the more so that the disturbance was intense [@Martin2015; @Chaudhary2016].
-As these pioneers became more dominant they balanced the usual hyper-dominance of tropical forests, hence temporarily increasing the taxonomic evenness [@Baraloto2012a].
-On the other hand, above the intensity threshold, the taxonomic richness of surviving trees was too low to be offset by the recruitment of pioneers. 
+Community trajectories depended on the disturbance intensity, with a threshold of 20-25% of AGB loss for which the taxonomic richness was maximized and the taxonomic evenness remained resilient.
+Below a disturbance intensity threshold (treatments 1 and 2), the taxonomic richness increased and the taxonomic eveness increased first before stabilizing or starting to decrease.
+In those cases, the trees surviving after disturbance remained numerous enough to maintain the high taxonomic richness of the pre-disturbance community [@Bongers2009].
+In parallel, the recruitment of pioneers, infrequent or absent before disturbance, increased the taxonomic richness all the more so that the disturbance was intense [@Martin2015;@Chaudhary2016].
+As these pioneers became more dominant they balanced the usual hyper-dominance of tropical forests and increased the taxonomic evenness [@Baraloto2012a].
+Contrastingly, above the intensity threshold (treatment 3), the taxonomic richness did not exceed the initial value and the taxonomic evenness continuously increased without sign of reversal.
+Above the intensity threshold, the taxonomic richness of surviving trees was too low to be offset by the recruitment of pioneers. 
 In the Guiana shield indeed, the pool of true pioneers specifically recruited after disturbance is restricted to a few common genera (e.g. *Cecropia* spp., *Vismia* spp.) [@Guitet2018].
-The taxonomic richness then decreased following intense disturbance, all the more so that the disturbance intensity was high [@Molino2001]. 
 At the same time, pioneers became persistently dominant and prevented a return towards the initial evenness and the recovery of hyper-dominant shade-tolerant species.
 
-Taxonomic trajectories following disturbance of intermediate intensity besides highlighted an intermediate time after disturbance, around 15-20 years, for which the taxonomic evenness was maximized.
-As already observed in the Guiana shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998], taxonomic evenness followed hump-shaped trajectories with a peak when the recruitment was balanced between pioneers and late-successional species.
-This intermediate time corresponded to the match between the generation time of pionneers, dominating first the post-disturbance recruitment, and the time for competitive exclusion to emerge, leading afterwards to the dominance of late-successional species.
+After disturbance of intermediate intensity, taxonomic evenness trajectories were hump-shaped, as already observed in the Guiana shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
+Trajectories peaked at an intermediate time after disturbance, around 15-20 years.
+This time corresponded to a balance between first established pioneers and the more competitve late-successional species.
 
@@
 
-The IDH then translated into community taxonomic response, with post-disturbance trajectories changing markedly below and above an intensity threshold of 25 % AGB loss.
-The IDH was also valid at an intermediate time after disturbance, when the recruitment of early- and late-successional species were balanced. 
-Functional trajectories, however, remained decoupled from taxonomic ones and dismissed the IDH.
-
 ## The functional redundancy, key of community resilience
 
-The decoupling between taxonomic and functional trajectories was accompanied by a decrease in the functional redundancy within the pre-disturbance functional space.
-This decrease was explained by a loss of species following disturbance and then a lower redundancy of their functional characteristics.
-This decrease of the functional redundancy was not compensated by the recruitment of pioneers, as those were functionally different from the pre-disturbance community.
-Progressively though, more competitive,long-lived pioneers or late-successional species functionally closer to the pre-disturbance community progressively replaced  the first established species and restored the functional redundancy.
-This replacement followed the lottery recruitment rules, implying an easy recruitment for the first species but becoming increasingly hampered by the emergence of interspecific competition [@Busing2002].
-The recovery of the functional redundancy then relied upon the random process of species recruitment and was increasingly slow and difficult to anticipate [@Elmqvist2003;@Diaz2005].
-
-The impact of disturbance on community functional redundancy meant a lower resilience of the disturbed communities, with higher chances to see the settltement of pioneers and early-successional species at the expense of late-successional ones [@Haddad2008].
-Besides, the random recovery of infrequent species increases the risks to lose keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
-Infrequent species might indeed have unique functional characteristics, apart from those considered here, in the ecosystem or be a key for some fauna [@Schleuning2016].
+The loss of species following disturbance decreased the functional redundancy during the first 15 years. 
+Progressively though, the functional redundancy was restored through the replacement of the resource-acquisitive strategies species by more late-successional species that were functionally closer to the pre-disturbance community.
+This replacement followed the lottery recruitment rules, implying an easy recruitment for the first species but becoming increasingly hampered by the emergence of interspecific competition [@Busing2002] and this explains why the slope of the recovery trajectory became less and less significant 20 years after disturbance.
+The recovery of the functional redundancy then relied upon the random process of species recruitment and was increasingly slow and difficult to anticipate
+[@Elmqvist2003;@Diaz2005]. 
+This suggests a low resilience of the functional redundancy with the random recovery of infrequent species increasing the risks to lose keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
+Infrequent species might indeed have unique functional characteristics, apart from those considered here, in the ecosystem or be a key resource for some fauna [@Schleuning2016].
 
@@
 
-Community post-disturbance trajectories in compositon and diversity were shaped by the recruitment of a determined pool of pioneers identical among communities and disturbance intensity.
-Nevertheless, taxonomic and functional composition trajectories appeared decoupled.
+Post-disturbance trajectories of tree community composition and diversity were both shaped by the recruitment of a determined pool of pioneers identical
+among local communities and independent of the disturbance intensity.
+(i) Composition trajectories in taxonomic and functional space nevertheless appeared decoupled.
 While functional trajectories remained similar in the functional space and converged towards the recovery of a comparable initial state, taxonomic trajectories showed initial differences in community composition that were maintained along time.
-Community high functional redundancy could explain this decoupling as the loss of a species does not necessarily the loss of its functional charactristics in the community.
-In this sense, the functional redundancy would mitigate the functional impact of disturbance.
-Community functional and taxonomic diversity response to disturbance were contrasted as well.
-While the functional trajectories remained similar whenever the disturbance intensity, taxonomic trajectories were markedly different above a threshold of 20-25% AGB removed for which taxonomic richness was maximized and taxonomic eveness remained resilient.
-The Intermediate Disturbance Hypothesis applied to the linkage of disturbance with the taxonomic diversity, but not with the functional diversity.
-Taxonomic diversity was maximized following a disturbance of intermediate intensity, that was the threshold of 25% AGB removed, and at an intermediate time after disturbance, that was around 25 years after disturbance when the recruitment of early- and late-successional species were balanced.
+Community high functional redundancy mediated this decoupling, as the loss of a species does not necessarily entails the loss of its functional charactristics.
+(ii) Diversity trajectories were contrasted as well.
+While the functional trajectories remained similar whatever the disturbance intensity, taxonomic trajectories were markedly different with a threshold at 20-25\% AGB removed for which taxonomic richness was maximized.
+The Intermediate Disturbance Hypothesis applied well to taxonomic diversity, but not to the functional diversity.
+Taxonomic diversity was maximized at an intermediate time after disturbance, that was around 25 years after disturbance when the recruitment of early- and late-successional species were balanced.
 Whenever the disturbance intensity, community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied upon the random lottery recruitment of rare species.
-Although resilient, the functional redundancy was lower for more than 30 years after disturbance which cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Herault2018].
-Our conclusions first highlighted risks to lose rare species with visible impacts only in the long term, and led us to advice paying particular attention to rare and undangered species and to conduct regular inventories during exploitation times.
-Second, post-disturbance trajectories seemed to greatly rely on post-disturbance recruitment dynamics that would be key to model post-disturbance trajectories.
+Given the long-term impacts of disturbance observed, we suggest that 30 years is not enough time for tropical communities to recover, even after relatively low intensity disturbance.
+Much of community response to disturbance rely on species recruitment: refined understanding of post-disturbance trajectories would be given by a closer analysis of recruitment processes.
 

### 2019-04-09 WholePlotTrajectories.tex
relecture

@@
 However, the growing demand in forests products together with current
-global changes increase the pressure on remaining natural forests
-\citep{Morales-Hidalgo2015}. These threats affect the disturbance regime
-that naturally shapes and maintains the structure, composition and
-functioning of tree communities
+global changes increase the pressure on remaining natural undisturbed
+forests \citep{Morales-Hidalgo2015}. These threats affect the
+disturbance regime that naturally shapes and maintains the structure,
+composition and functioning of tree communities
 \citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. To anticipate
-the fate of tropical forests in the current context it is then urgent to
-understand the response of communities to disturbance and the underlying
-processes. Disturbance impact forest communities in changing both
-abiotic and biotic environments, through modifications in the fluxes of
-light, heat and water \citep{Goulamoussene2017}. Although in the
-tropical context the forest environment is maintained after disturbance,
-there is a succession of changes in the structure and functioning of
-communities constituting a post-disturbance secondary succession. For
-now, this succession has been largely studied through trajectories of
-forest structural parameters such as aboveground biomass, tree height or
-stem density \citep{Piponiot2016, Rutishauser2016}. Models based on
-observed trajectories assessed community time to recovery and identified
-some of the determinants of post-disturbance trajectories, like forest
-structure, forest composition, environmental parameters and disturbance
-characteristics \citep{Herault2018}. Regarding community diversity and
-composition however, post-disturbance trajectories have not been as
-thoroughly understood \citep{Guitet2018, Molino2001}. Given the variety
-of species response to disturbance and the huge diversity of tropical
-forests, manifold post-disturbance biodiversity trajectories might
-emerge \citep{Lindenmayer2012, Garcia_florez2017}. It is then essential
-to anticipate post-disturbance trajectories and highlight the link
-between biodiversity and disturbance, specifically the link between
-disturbance intensity and frequency and the rate, endpoint and duration
-of the trajectories \citep{Chazdon2003a, Willig2018}.
+the fate of tropical forests in this context, understanding the response
+of communities to disturbance and deciphering the underlying ecological
+processes is of utmost importance. Disturbance impacts forest
+communities in changing both abiotic and biotic environments, through
+modifications in the fluxes of light, heat and water
+\citep{Goulamoussene2017}. Although the forest cover is generally
+maintained after disturbance, there is a succession of changes in the
+structure and functioning of ecological communities that defines a
+post-disturbance ecosystem trajectory. For now, these trajectories have
+been largely studied through trajectories of forest structural
+parameters such as aboveground biomass, tree height or stem density
+\citep{Piponiot2016, Rutishauser2016}. Recent works have identified some
+of the determinants of post-disturbance biomass trajectories, like
+pre-disturbance forest structure and composition or post-disturbance
+environmental parameters \citep{Herault2018}. Regarding tree community
+diversity and composition however, post-disturbance trajectories have
+not been as thoroughly understood \citep{Guitet2018, Molino2001}. Given
+the variety of species response to disturbance and the huge diversity of
+tropical forests, manifold post-disturbance biodiversity trajectories
+might emerge \citep{Lindenmayer2012, Garcia_florez2017}.
 
@@
 validation of the theory remains controversial
-\citep{Hubbell2001, Fox2013, Sheil2013}. The link between biodiversity
-and disturbance might be complicated by the huge diversity of tropical
-communities and a variety of ecological processes may underlie the IDH
-(colonisation, facilitation, coexistence,
-etc)\citep{Lindenmayer2012, Garcia_florez2017}. There is a need to
-clarify the scope of the IDH in terms of the different facets of
-community diversity that are taxonomic richness, species abundance
-distribution and community functioning. Besides, there is a need to
-delineate the scope of the IDH in time, and highlight the moment when
-the conditions for the IDH emerge \citep{Sheil2003, Shea2004}.
+\citep{Hubbell2001, Fox2013, Sheil2013}.
+
+Moreover, we argue that there is a need to delineate the scope of the
+IDH in time, using ecosystem time-explicit trajectories, and highlight
+the moment when the conditions for the IDH emerge
+\citep{Sheil2003, Shea2004}. Indeed, the link between biodiversity, IDH
+and post-disturbance disturbance trajectories might be complicated by
+the huge diversity found in tropical tree communities and the variety of
+ecological processes that may underlie the IDH (colonization,
+facilitation, coexistence,
+etc)\citep{Lindenmayer2012, Garcia_florez2017}.
 
@@
 consider community composition, which is crucial for conservation issues
-and which reveals the species fostered by the disturbance
-\citep{Lavorel2002, Bellwood2006}. Then, diversity metrics encompassing
-taxonomic richness and evenness should be considered to assess the
-changes in community abundance distribution. Besides, a functional
-approach that accounts for species biological attributes and ecological
-strategy would directly link community diversity, composition and
-redundancy to ecosystem functioning and environmental constraints
-\citep{Violle2007b, Baraloto2012a}. In that respect, a vast literature
-allowed recognizing major traits that represent species ecological
-strategy and determine how they respond to changing conditions
-\citep{Diaz2005}. Specifically in tropical forests, the functional
-approach revealed the post-disturbance deterministic processes entailing
-a shift from a dominance of ``conservative'' slow-growing species
-dealing with scarce resources, to a dominance of ``acquisitive''
-fast-growing species with rapid and efficient use of abundant resources
-\citep{Rees2001, Reich2014, Herault2011}. This shift is translated into
-the trajectories of average community value of key functional traits
-related to resource acquisition (leaf and stem traits) and life-history
-strategy (seed mass, maximum size)
-\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. The
-functional approach also encompasses the analysis of functional
+and which reveals the pool of species fostered (or hampered) by
+disturbance \citep{Lavorel2002, Bellwood2006}. Then, diversity metrics
+encompassing both richness and evenness should also be considered to
+assess the changes in community abundance distribution. Besides,
+functional approaches have been shown to usefully complement pure
+taxonomic approaches because they shed light on the species biological
+attributes that directly link community diversity, composition and
+redundancy to ecosystem functioning \citep{Violle2007b, Baraloto2012a}.
+In that respect, a vast literature allowed recognizing major traits that
+represent species ecological strategy and determine how they respond to
+changing conditions \citep{Diaz2005}. Specifically in tropical forests,
+the functional approach revealed the post-disturbance deterministic
+processes entailing a shift from a dominance of ``conservative''
+slow-growing species dealing with scarce resources, to a dominance of
+``acquisitive'' fast-growing species with rapid and efficient use of
+abundant resources \citep{Rees2001, Reich2014, Herault2011}. This shift
+is translated into the trajectories of average community value of key
+functional traits related to resource acquisition (leaf and stem traits)
+and life-history strategy (seed mass, maximum size)
+\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}.
+
+The functional approach also encompasses the analysis of functional
 redundancy, that quantifies the amount of shared trait values among
@@
 trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, (i) we defined taxonomic and functional post-disturbance
-trajectories and examined the underlying ecological process, (ii) we
-discussed the scope of the IDH regarding the different facets of
-community diversity and the time after disturbance, and (iii) we
-analyzed community resilience and time to recovery.
+Specifically, (i) we draw taxonomic and functional post-disturbance
+trajectories and examine the underlying ecological process, (ii) we
+discuss the scope of the IDH regarding the different facets of community
+diversity and the time after disturbance, and (iii) we analyze community
+resilience and time to recovery.
 
@@
 mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
-one-month dry season in March \citep{Wagner2011}. The mean annual
-temperature is 26\textdegree C. Elevation ranges from 5 to 50 m. Across
-all plots the topography mainly corresponds to hilltops or hillsides,
-while bottomlands cover less than 1 \% of the area. Plots are shallow
-ferralitic acrisols over a layer of transformed saprolite with low
-permeability and lateral drainage. Soil conditions are homogeneous, to
-the exception of the highest hilltops where the thick surface allows a
-free vertical drainage \citep{Gourlet-Fleury2004}.
+one-month dry season in March \citep{Wagner2011}. Elevation ranges from
+5 to 50 m. Across all plots the topography mainly corresponds to
+hilltops or hillsides, while bottomlands cover less than 1 \% of the
+area. Plots are shallow ferralitic acrisols over a layer of transformed
+saprolite with low permeability and lateral drainage. Soil conditions
+are homogeneous, to the exception of the highest hilltops where the
+thick surface allows a free vertical drainage
+\citep{Gourlet-Fleury2004}.
 
@@
 \citep{Piponiot2016} estimated with the BIOMASS R package
-\citep{Rejou2017}. The three treatments constituted a disturbance
-intensity gradient with increasing of above-ground biomass (AGB) lost
-and surface disturbed.
+\citep{Rejou2017}. The three treatments were then transformed into a
+continuous disturbance intensity gradient with increasing of
+above-ground biomass (AGB) lost.
 
@@
 The functional diversity was reported using the functional richness and
-functional evenness, \emph{i.e} the Rao index of quadratic entropy. The
-Rao index combines species abundance distribution, and the average
-pairwise functional dissimilarity between species computed by the gower
-distance.
+functional evenness, through the Rao index of quadratic entropy. The Rao
+index combines species abundance distribution, and the average pairwise
+functional dissimilarity between species computed by the Gower distance.
 
-The impacts of initial disturbance were tested with the Spearman rank
-correlation between the extremes of taxonomic and functional metrics
-reached over the 30 years and the initial \%AGB loss. They were besides
+The impacts of the initial disturbance levels on the maximum gain or
+loss in richness and evenness were tested with the Spearman rank
+correlation tests. Richness and evenness trajectories were besides
 analyzed through polynomial regression between (i) taxonomic and
@@
 
-Functional redundancy was measured as the overlap among species in
-community functional space \citep{Carmona2016}. First, the individuals
-of the trait database were mapped in the plane of the first two axes
-from a PCA analysis, which lowered the weight of correlations among
-traits as the PCA axes were combinations of most decoupled traits. For
-each species, the traits probability density (TPD) were computed from
-the mapping of individuals through two-dimension kernel density
+Finally, functional redundancy was measured as the overlap among species
+in community functional space \citep{Carmona2016}. First, the
+individuals of the trait database were mapped in the plane of the first
+two axes from a PCA analysis, which lowered the weight of correlations
+among traits as the PCA axes were combinations of most decoupled traits.
+For each species, the traits probability density (TPD) were computed
+from the mapping of individuals through two-dimension kernel density
 estimators. Second, for each community, the TPD weigthed by species
@@
 undisturbed communities (Fig. \ref{fig:NMDSplans}), they followed marked
-and consistent trajectories over post-\break disturbance time. In
-disturbed communities the taxonomic composition changed towards more
-species with acquisitive functional strategies. This translated into a
-switch from high mean WSG in the community to high mean SLA and
-chlorophyll content (see appendix I). The functional composition of
-disturbed communities corresponded to cyclic compositional changes with
-an incomplete recovery of the initial composition (Fig.
-\ref{fig:NMDSplans}).
+and consistent trajectories in disturbed plots over time.
 
-The maximum dissimilarity with the initial state was positively
-correlated with the disturbance intensity for both taxonomic and
-functional composition (\(\rho_{Spearman}^{Taxonomic}=0.87\) and
+The functional composition trajectory in disturbed communities resembled
+to cyclic compositional changes with an incomplete recovery to the
+initial composition (Fig.\ref{fig:NMDSplans}). The maximum dissimilarity
+with the initial state was positively correlated with the disturbance
+intensity for both taxonomic and functional composition
+(\(\rho_{Spearman}^{Taxonomic}=0.87\) and
 \(\rho_{Spearman}^{Functional}=0.90\) respectively). The maximum
 dissimilarity with the initial state was reached for taxonomic
-composition between 15 to 25 years, for most of the plots, and around 22
-years for functional composition.
+composition between 15 to 25 years and around 20 to 25 years for
+functional composition.
 
@@
 
-Community CWM average value of all traits and seed mass proportions
-followed unimodal trajectories, either stabilizing or returning towards
-their initial values, to the exception of leaf chlorophyll content,
-which continued to increase for some T3 and T2 plots 30 years after
-disturbance.
-
-Community CWM average value of Maximum height at adult stage
-(\emph{Hmax}), leaf toughness and wood specific gravity (\emph{WSG})
-decreased and then slightly increased, but remained significantly lower
-than their initial value (Fig. \ref{fig:CWM}). On the other side, bark
-thickness and specific leaf area (\emph{SLA}) both increased. Bark
-thickness remained substantially high after 30 years, and \emph{SLA} had
-almost recovered to its initial value. For all traits, the maximum
-difference to initial value was correlated to the disturbance intensity.
-Positive correlations were observed for Leaf thickness, chlorophyll
-content, SLA and bark thickness
+In control plots, Community Weighted Means (CWM) of functional traits
+remain stable in time.\\
+In disturbed plots, they mostly followed unimodal trajectories, either
+stabilizing or returning towards their initial values, to the exception
+of leaf chlorophyll content, which continued to increase 30 years after
+disturbance for some highly disturbed plots. Maximum height at adult
+stage (\emph{Hmax}), leaf toughness and wood specific gravity
+(\emph{WSG}) decreased in time and then slightly increased, but remained
+significantly lower than their initial value (Fig. \ref{fig:CWM}). Bark
+thickness and specific leaf area (\emph{SLA}) both increased in time.
+Bark thickness remained substantially high after 30 years, and
+\emph{SLA} had almost recovered to its initial value. Whatever the
+functional traits, the maximum difference to initial value was highly
+correlated to the disturbance intensity. Positive correlations were
+observed for Leaf thickness, chlorophyll content, SLA and bark thickness
 (\(\rho_{Spearman}^{Leaf thickness}=0.76\),
@@
 
-For undisturbed plots, taxonomic richness and Simpson diversity remained
-stable over the 30 years of monitoring. In disturbed communities, after
-low disturbance intensity the taxonomic richness increased, reaching a
-maximum gain of 14 botanical genera (plot 3 from treatment 2). After
-intense disturbance the taxonomic richness followed a more complex
-trajectory, decreasing for ten years after disturbance before recovering
-to pre-disturbance values. The maximum richness loss or gain after
-disturbance was positively correlated with the disturbance intensity
+Taxonomic richness and evenness remained stable in control plots over
+the 30 years of monitoring. In disturbed communities, after low
+disturbance intensity the taxonomic richness increased, reaching a
+maximum gain of 14 botanical genera. After intense disturbance the
+taxonomic richness followed a more complex trajectory, decreasing for
+ten years after disturbance before recovering to pre-disturbance values.
+The maximum richness loss or gain after disturbance was positively
+correlated with the disturbance intensity
 (\(\rho_{Spearman}^{Richness}=0.50\)).
 
-In all disturbed plots the Simpson diversity first increased until a
-maximum reached after around 20 years. This maximum was positively
-correlated with the disturbance intensity
-(\(\rho_{Spearman}^{Simpson}=0.77\)). The Simpson diversity then
-stabilized except for two T3 plots (plots 8 and 12) for which it kept
-increasing (Fig. \ref{fig:DivTaxo}).
+In all disturbed plots the evenness first increased until a maximum
+reached after around 20 years. This maximum was positively correlated
+with the disturbance intensity (\(\rho_{Spearman}^{Simpson}=0.77\)). The
+evenness then stabilized except for two intensively-disturbed plots
+(number 8 and 12) for which it kept increasing \ref{fig:DivTaxo}).
 
@@
 
-The plot 7 from treatment 1 displayed constantly outlying functional
-richness and Rao diversity and was removed from the graphical
-representation for better readability. In undisturbed plots both
-functional richness and Rao diversity remained stable along the 30
-years. In disturbed plots, both trajectories depended on the disturbance
-intensity, with their maximum being positively correlated to \%AGB loss
+Functional richness and evenness remained stable in control plots over
+the 30 years of monitoring. In disturbed communities, both trajectories
+depended on the disturbance intensity, with their maximum values in time
+being positively correlated to \%AGB loss
 \(\rho_{Spearman}^{Richness}=0.76\) and \(\rho_{Spearman}^{Rao}=0.60\).
-Functional richness and Rao diversity displayed for low disturbance
-intensity a low but long-lasting increase up to a maximum reached after
-20-25 years. Following high intensity, they displayed a fast but short
-increase followed after 10 years by a slow decrease towards the initial
-values.
+Functional richness and evenness displayed for low disturbance intensity
+a low but long-lasting increase up to a maximum reached after 20-25
+years. For high disturbance intensity, they generally displayed a fast
+but short increase followed after 10 years by a slow decrease towards
+the initial values.
 
 The second-degree polynomial regressions between (i) the percentage AGB
-loss and (ii) taxonomic and functional diversity after 10, 20 and 30
-years best predicted the hump-shaped curve of the disturbance impact
-along the disturbance intensity gradient (Fig. \ref{fig:IDHplot}). The
-relationship between the disturbance impact and its intensity was more
-markedly hump-shaped for the taxonomic richness than for the Simpson
-diversity. For both functional richness and Rao diversity the
-relationship was almost linear. The regression model better predicted
-the functional richness and Rao diversity
-(\(0.55<R^2_{Functional Richness}<0.72\), and
-\(0.60<R^2_{Functional Rao}<0.81\)) than the taxonomic richness and
-evenness (\(0.21<R^2_{Taxonomic Richness}<0.4\), and
-\(-0.15<R^2_{Taxonomic Simpson}<0.43\) respectively).
+loss and (ii) the taxonomic and functional diversity showed various
+shape depending on the diversity indices and on the time since
+disturbance (Fig. \ref{fig:IDHplot}). Regarding taxonomic diversity, the
+relationship between disturbance intensity and diversity was more
+markedly hump-shaped for richness than for evenness and peaked at 20\%
+of initial AGB loss. Regarding functional diversity, the relationship
+was almost linear and similar between richness and evenness. Generally,
+all relationships were stronger 20 or 30 years after disturbance than
+observed for 10 years.
 
@@
 
-All disturbed plots had lower functional redundancy than control plots
-and followed similar hump-shaped trajectories (\ref{fig:RedFunRest}).
-The maximum redundancy loss was positively correlated with the
-disturbance intensity (\(\rho_{Spearman}=0.47\)) and the recovery had
-not attained initial values for any disturbed communities after 30
-years.
+Control plots displayed stable functional redundancy over time. All
+disturbed plots had lower functional redundancy than control plots and
+followed similar hump-shaped trajectories (\ref{fig:RedFunRest}). The
+maximum redundancy loss was positively correlated with the disturbance
+intensity (\(\rho_{Spearman}=0.47\)) and the recovery trajectory had not
+attained initial values for any disturbed communities after 30 years.
 
@@
 Our analysis highlighted the decoupling between functional and taxonomic
-trajectories. While initial differences in community taxonomic
-composition were maintained after disturbance, their trajectories in
-functional composition were similar. In parallel, while the different
-taxonomic diversity trajectories displayed marked differences according
-to the disturbance intensity, the functional diversity trajectories were
-homogeneous whenever the disturbance intensity. This decoupling was
-accompanied by a decrease in community functional redundancy which
-appeared to be the key element of community post-disturbance recovery.
+trajectories, with taxonomic trajectories maintaining initial
+differences among plots while functional trajectories remained similar.
+We revealed that if the IDH explained the taxonomic response to
+disturbance, it was not validated regarding the functional response. The
+decoupling between taxonomic and functional response was linked to the
+disturbance impact on the functional redundancy, that proved determining
+in community recovery.
 
@@
 
-Before disturbance, the different communities had different taxonomic
-compositions, visualized by their distinct starting points on the NMDS
-axis 2. These initial differences were maintained all along the 30 years
-following disturbance, with the disturbance leading a displacement on
-the NMDS axis 1. The taxonomic compositional changes following
-disturbance were similar among communities, which may correspond to the
-recruitment of a group of pioneers, as \emph{Cecropia} spp.,
-\emph{Vismia} spp., common to all communities whenever their initial
-differences and the intensity of disturbance
-\citep{Denslow2000, Bongers2009}. Taxonomic trajectories continued with
-a recovery of the initial composition which, althought not fully
-achieved after 30 years, suggested the resilience of the taxonomic
-composition resilience and the maintenance of the initial differences
-among communities \citep{Folke2006}. The return towards initial
-taxonomic composition showed that species not belonging to the
-pre-disturbance community were rarely recruited, probably because of the
-common dispersal limitations among tropical tree species
-\citep{Svenning2005}.
+Regarding taxonomic composition, the different tree communities were
+substantially different before disturbance, and this is visualized by
+their distinct starting points on the NMDS axis 2. These initial
+differences were maintained all along the 30 years following
+disturbance, with the disturbance leading a displacement on the NMDS
+axis 1 only. Changes were thus similar and may correspond to the
+recruitment of a group of pioneers, e.g. \emph{Cecropia} spp.,
+\emph{Vismia} spp., shared by all plots and whatever their initial
+taxonomic differences and the intensity of disturbance
+\citep{Denslow2000, Bongers2009}. Taxonomic trajectories seems to start
+a recovery to the initial composition which, although far from fully
+achieved after 30 years, suggested that taxonomic composition is
+resilient and that the initial differences will be maintained in time
+\citep{Folke2006}. This would show that species not belonging to the
+pre-disturbance community were rarely recruited in the long-term,
+probably because of the common dispersal limitations among tropical tree
+species \citep{Svenning2005}.
 
-Regarding functional composition, initial communities were similar and
-trajectories were confounded in the functional space. Following
-disturbance, the functional composition of pre-disturbance surviving
-trees is believed to be the same as this of the initial community
-\citep{Herault2018}. The changes in functional composition following
-disturbance then relied upon the recruitment of species or functional
-types previously infrequent or absent. These species were probably
-competitive pioneers becoming dominant as the most efficient to benefit
-from the light, space and nutrients made available by the disturbance.
-This recruitment of pioneers led similar functional composition
-trajectories for all plots, whenever the disturbance intensity. The
-composition first went towards more resource-acquisitive strategies,
-which was translated by a displacement to the right along the first axis
-in the functional plane \citep{Westoby1998, Wright2004, Reich2014}.
-Thereafter, the pioneers recruited primarily were progressively excluded
-by long-lived, more resistant and shade-tolerant species. The community
-functional composition then returned towards more resource-conservative
+Regarding functional composition, initial communities had similar
+starting points and this contrasts to taxonomic composition. This means
+that, despite differences in species composition, the functional
+signature of the tree communities were quite similar. Following
+disturbance, functional trajectories were consistent in the NMDS plane
+with displacement intensities linked to disturbance intensities and thus
+relied upon the recruitment of a shared pool of functional types
+previously infrequent or absent. This common pool is composed of pioneer
+resource-acquisitive (low leaf toughness, wood specific gravity, maximum
+height and high Specific Leaf area) strategies that translated by a
+displacement to the right along the first NMDS axis
+\citep{Westoby1998, Wright2004, Reich2014}. Thereafter, the pioneers
+recruited primarily were progressively excluded by long-lived, more
+competitive and shade-tolerant species. The community functional
+composition then quickly returned towards more resource-conservative
 strategies, suggesting the recovery of the initial functional
@@
 return towards pre-disturbance state after 30 years, which highlighted
-the taxonomic and functional resilience of communities. Taxonomicand
+the taxonomic and functional resilience of communities. Taxonomic and
 functional trajectories however appeared decoupled: while taxonomic
@@
 functional trajectories where similar and convergent in the functional
-space \citep{Fukami2005}.
+space \citep{Fukami2005} exemplifying the simultaneous operation of
+trait-based assembly rules and species-level priority effects that
+shapes tree community assembly in this forest, making it both
+deterministic in functional space and historically contingent in
+taxonomic space.
 
@@
 
-Comparing community diversity along time after disturbance revealed
-constrasting trajectories according to the distrubance intensity. Below
-a disturbance intensity threshold, in the case of treatments 1 and 2,
-the community taxonomic richness increased. In parallell, after a first
-increase, the taxonomic eveness stabilized or strated to decrease after
-30 years. Below the intensity threshold Contrastingly, above an
-intensity threshold, in the case of treatment 3, the taxonomic richness
-did not exceeded the initial value and the taxonomic evenness
-continiously increased for 30 years without sign of return towards the
-initial value. Below and above a threshold, corresponding to a
-disturbance intensity between treatments 2 and 3, community taxonomic
-richness and evenness changed markedly. A the disturbance intensity
-threshold, the taxonomic richness was maximized and the taxonomic
-evenness remained resilient. The disturbance intensity impacted the
-balance in the community between pre-disturbance surviving trees, and
-trees recruited afterward. On the one hand, below a 20\%-25\% AGB loss,
-the trees surviving after disturbance remained numerous enough to
-maintain the high taxonomic richness of the pre-disturbance community
-\citep{Bongers2009}. The recruitment of pioneers, infrequent or absent
-before disturbance, then increased the taxonomic richness all the more
-so that the disturbance was intense \citep{Martin2015, Chaudhary2016}.
-As these pioneers became more dominant they balanced the usual
-hyper-dominance of tropical forests, hence temporarily increasing the
-taxonomic evenness \citep{Baraloto2012a}. On the other hand, above the
-intensity threshold, the taxonomic richness of surviving trees was too
-low to be offset by the recruitment of pioneers. In the Guiana shield
-indeed, the pool of true pioneers specifically recruited after
-disturbance is restricted to a few common genera (e.g. \emph{Cecropia}
-spp., \emph{Vismia} spp.) \citep{Guitet2018}. The taxonomic richness
-then decreased following intense disturbance, all the more so that the
-disturbance intensity was high \citep{Molino2001}. At the same time,
-pioneers became persistently dominant and prevented a return towards the
-initial evenness and the recovery of hyper-dominant shade-tolerant
-species.
+Community trajectories depended on the disturbance intensity, with a
+threshold of 20-25\% of AGB loss for which the taxonomic richness was
+maximized and the taxonomic evenness remained resilient. Below a
+disturbance intensity threshold (treatments 1 and 2), the taxonomic
+richness increased and the taxonomic eveness increased first before
+stabilizing or starting to decrease. In those cases, the trees surviving
+after disturbance remained numerous enough to maintain the high
+taxonomic richness of the pre-disturbance community \citep{Bongers2009}.
+In parallel, the recruitment of pioneers, infrequent or absent before
+disturbance, increased the taxonomic richness all the more so that the
+disturbance was intense \citep{Martin2015, Chaudhary2016}. As these
+pioneers became more dominant they balanced the usual hyper-dominance of
+tropical forests and increased the taxonomic evenness
+\citep{Baraloto2012a}. Contrastingly, above the intensity threshold
+(treatment 3), the taxonomic richness did not exceed the initial value
+and the taxonomic evenness continuously increased without sign of
+reversal. Above the intensity threshold, the taxonomic richness of
+surviving trees was too low to be offset by the recruitment of pioneers.
+In the Guiana shield indeed, the pool of true pioneers specifically
+recruited after disturbance is restricted to a few common genera (e.g.
+\emph{Cecropia} spp., \emph{Vismia} spp.) \citep{Guitet2018}. At the
+same time, pioneers became persistently dominant and prevented a return
+towards the initial evenness and the recovery of hyper-dominant
+shade-tolerant species.
 
-Taxonomic trajectories following disturbance of intermediate intensity
-besides highlighted an intermediate time after disturbance, around 15-20
-years, for which the taxonomic evenness was maximized. As already
-observed in the Guiana shield \citep{Baraloto2012a} and in Bornean
-tropical forests \citep{Cannon1998}, taxonomic evenness followed
-hump-shaped trajectories with a peak when the recruitment was balanced
-between pioneers and late-successional species. This intermediate time
-corresponded to the match between the generation time of pionneers,
-dominating first the post-disturbance recruitment, and the time for
-competitive exclusion to emerge, leading afterwards to the dominance of
+After disturbance of intermediate intensity, taxonomic evenness
+trajectories were hump-shaped, as already observed in the Guiana shield
+\citep{Baraloto2012a} and in Bornean tropical forests
+\citep{Cannon1998}. Trajectories peaked at an intermediate time after
+disturbance, around 15-20 years. This time corresponded to a balance
+between first established pioneers and the more competitve
 late-successional species.
@@
 
-The IDH then translated into community taxonomic response, with
-post-disturbance trajectories changing markedly below and above an
-intensity threshold of 25 \% AGB loss. The IDH was also valid at an
-intermediate time after disturbance, when the recruitment of early- and
-late-successional species were balanced. Functional trajectories,
-however, remained decoupled from taxonomic ones and dismissed the IDH.
-
 \subsection{The functional redundancy, key of community
@@
 
-The decoupling between taxonomic and functional trajectories was
-accompanied by a decrease in the functional redundancy within the
-pre-disturbance functional space. This decrease was explained by a loss
-of species following disturbance and then a lower redundancy of their
-functional characteristics. This decrease of the functional redundancy
-was not compensated by the recruitment of pioneers, as those were
-functionally different from the pre-disturbance community. Progressively
-though, more competitive,long-lived pioneers or late-successional
-species functionally closer to the pre-disturbance community
-progressively replaced the first established species and restored the
-functional redundancy. This replacement followed the lottery recruitment
-rules, implying an easy recruitment for the first species but becoming
-increasingly hampered by the emergence of interspecific competition
-\citep{Busing2002}. The recovery of the functional redundancy then
-relied upon the random process of species recruitment and was
-increasingly slow and difficult to anticipate
-\citep{Elmqvist2003, Diaz2005}.
-
-The impact of disturbance on community functional redundancy meant a
-lower resilience of the disturbed communities, with higher chances to
-see the settltement of pioneers and early-successional species at the
-expense of late-successional ones \citep{Haddad2008}. Besides, the
-random recovery of infrequent species increases the risks to lose
-keystone species, with unexpected ecological consequences
+The loss of species following disturbance decreased the functional
+redundancy during the first 15 years. Progressively though, the
+functional redundancy was restored through the replacement of the
+resource-acquisitive strategies species by more late-successional
+species that were functionally closer to the pre-disturbance community.
+This replacement followed the lottery recruitment rules, implying an
+easy recruitment for the first species but becoming increasingly
+hampered by the emergence of interspecific competition
+\citep{Busing2002} and this explains why the slope of the recovery
+trajectory became less and less significant 20 years after disturbance.
+The recovery of the functional redundancy then relied upon the random
+process of species recruitment and was increasingly slow and difficult
+to anticipate \citep{Elmqvist2003, Diaz2005}. This suggests a low
+resilience of the functional redundancy with the random recovery of
+infrequent species increasing the risks to lose keystone species, with
+unexpected ecological consequences
 \citep{Jones1994, Chazdon2003a, Diaz2005}. Infrequent species might
 indeed have unique functional characteristics, apart from those
-considered here, in the ecosystem or be a key for some fauna
+considered here, in the ecosystem or be a key resource for some fauna
 \citep{Schleuning2016}.
@@
 
-Community post-disturbance trajectories in compositon and diversity were
-shaped by the recruitment of a determined pool of pioneers identical
-among communities and disturbance intensity. Nevertheless, taxonomic and
-functional composition trajectories appeared decoupled. While functional
+Post-disturbance trajectories of tree community composition and
+diversity were both shaped by the recruitment of a determined pool of
+pioneers identical among local communities and independent of the
+disturbance intensity. (i) Composition trajectories in taxonomic and
+functional space nevertheless appeared decoupled. While functional
 trajectories remained similar in the functional space and converged
@@
 trajectories showed initial differences in community composition that
-were maintained along time. Community high functional redundancy could
-explain this decoupling as the loss of a species does not necessarily
-the loss of its functional charactristics in the community. In this
-sense, the functional redundancy would mitigate the functional impact of
-disturbance. Community functional and taxonomic diversity response to
-disturbance were contrasted as well. While the functional trajectories
-remained similar whenever the disturbance intensity, taxonomic
-trajectories were markedly different above a threshold of 20-25\% AGB
-removed for which taxonomic richness was maximized and taxonomic eveness
-remained resilient. The Intermediate Disturbance Hypothesis applied to
-the linkage of disturbance with the taxonomic diversity, but not with
-the functional diversity. Taxonomic diversity was maximized following a
-disturbance of intermediate intensity, that was the threshold of 25\%
-AGB removed, and at an intermediate time after disturbance, that was
-around 25 years after disturbance when the recruitment of early- and
-late-successional species were balanced. Whenever the disturbance
-intensity, community resilience (in terms of recovery of the
-pre-disturbance state) was tangible but required several decades and
-relied upon the random lottery recruitment of rare species. Although
-resilient, the functional redundancy was lower for more than 30 years
-after disturbance which cautioned against the risks of infrequent
-species loss and the persistence of disturbance-specific communities
-\citep{Herault2018}. Our conclusions first highlighted risks to lose
-rare species with visible impacts only in the long term, and led us to
-advice paying particular attention to rare and undangered species and to
-conduct regular inventories during exploitation times. Second,
-post-disturbance trajectories seemed to greatly rely on post-disturbance
-recruitment dynamics that would be key to model post-disturbance
-trajectories.
+were maintained along time. Community high functional redundancy
+mediated this decoupling, as the loss of a species does not necessarily
+entails the loss of its functional charactristics. (ii) Diversity
+trajectories were contrasted as well. While the functional trajectories
+remained similar whatever the disturbance intensity, taxonomic
+trajectories were markedly different with a threshold at 20-25\% AGB
+removed for which taxonomic richness was maximized. The Intermediate
+Disturbance Hypothesis applied well to taxonomic diversity, but not to
+the functional diversity. Taxonomic diversity was maximized at an
+intermediate time after disturbance, that was around 25 years after
+disturbance when the recruitment of early- and late-successional species
+were balanced. Whenever the disturbance intensity, community resilience
+(in terms of recovery of the pre-disturbance state) was tangible but
+required several decades and relied upon the random lottery recruitment
+of rare species. Given the long-term impacts of disturbance observed, we
+suggest that 30 years is not enough time for tropical communities to
+recover, even after relatively low intensity disturbance. Much of
+community response to disturbance rely on species recruitment: refined
+understanding of post-disturbance trajectories would be given by a
+closer analysis of recruitment processes.
 

### 2019-04-10 WholePlotTrajectories.Rmd
details

@@
  We highlighted a decoupling between taxonomic trajectories, that differed among communities, and functional trajectories, remained similar and convergent.
-We explained this decoupling by the variations in community functional redundancy that mitigated the functional impact of disturbance.
-Post-disturbance trajectories clarified the scope of the Intermediate Disturbance Hypothesis that was validated regarding taxonomic diversity, with an intensity threshold of 20-25%, but not regarding functional diversity.The IDH applied around 20 years after disturbance when the recruitment of pioneers and late-successionals were balanced.
+We explained this decoupling by the variations in community functional redundancy mitigating the functional impact of disturbance.
+The humped-shaped taxonomic trajectories which amplitude varied with disturbance intensity validated the IDH (Intermediate Disturbance Hypothesis).
+The IDH was however dismissed for functional diversity.
  Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years.
@@
 
-Our analysis highlighted the decoupling between functional and taxonomic  trajectories, with taxonomic trajectories maintaining initial differences among plots while functional trajectories remained similar.
-We revealed that if the IDH explained the taxonomic response to disturbance, it was not validated regarding the functional response.
+Our analysis highlighted the decoupling between functional and taxonomic  trajectories, with taxonomic trajectories maintaining initial differences among plots while functional trajectories converged in the functional space.
+
+The humped-shaped taxonomic trajectories which amplitude varied with disturbance intensity validated the IDH.
+Taxonomic diversity was maximized at an intensity threshold of 20-25% AGB loss and 20 years after disturbance, when the recruitment of pioneers and late-successionals were balanced.
+The IDH however did not apply to the functional diversity.
 The decoupling between taxonomic and functional response was linked to the disturbance impact on the functional redundancy, that proved determining in community recovery.

### 2019-04-10 WholePlotTrajectories.tex
details

@@
 trajectories, remained similar and convergent. We explained this
-decoupling by the variations in community functional redundancy that
-mitigated the functional impact of disturbance. Post-disturbance
-trajectories clarified the scope of the Intermediate Disturbance
-Hypothesis that was validated regarding taxonomic diversity, with an
-intensity threshold of 20-25\%, but not regarding functional
-diversity.The IDH applied around 20 years after disturbance when the
-recruitment of pioneers and late-successionals were balanced. Although
-consistent, the recovery of community composition, diversity and
-redundancy remained unachieved after 30 years. This acknowledged the
-need of decades-long cycles without disturbance to ensure a complete
-recovery, and questioned tropical forest community resilience after
-repeated disturbances.
+decoupling by the variations in community functional redundancy
+mitigating the functional impact of disturbance. The humped-shaped
+taxonomic trajectories which amplitude varied with disturbance intensity
+validated the IDH (Intermediate Disturbance Hypothesis). The IDH was
+however dismissed for functional diversity. Although consistent, the
+recovery of community composition, diversity and redundancy remained
+unachieved after 30 years. This acknowledged the need of decades-long
+cycles without disturbance to ensure a complete recovery, and questioned
+tropical forest community resilience after repeated disturbances.
 }
@@
 trajectories, with taxonomic trajectories maintaining initial
-differences among plots while functional trajectories remained similar.
-We revealed that if the IDH explained the taxonomic response to
-disturbance, it was not validated regarding the functional response. The
-decoupling between taxonomic and functional response was linked to the
-disturbance impact on the functional redundancy, that proved determining
-in community recovery.
+differences among plots while functional trajectories converged in the
+functional space.
+
+The humped-shaped taxonomic trajectories which amplitude varied with
+disturbance intensity validated the IDH. Taxonomic diversity was
+maximized at an intensity threshold of 20-25\% AGB loss and 20 years
+after disturbance, when the recruitment of pioneers and
+late-successionals were balanced. The IDH however did not apply to the
+functional diversity. The decoupling between taxonomic and functional
+response was linked to the disturbance impact on the functional
+redundancy, that proved determining in community recovery.
 

### 2019-04-12 WholePlotTrajectories.Rmd
relecture

@@
 In tropical forests, however, observations of the IDH often diverge from theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
+In tropical forests the processes underlying the IDH might indeed be complicated by the huge diversity found in tropical tree communities and the variety of ecological processes that may underlie the IDH
+(colonization, facilitation, coexistence, etc) [@Lindenmayer2012;@Garcia_florez2017].
+It is then crucial to test the IDH in this context and determine its scope in time, using ecosystem time-explicit trajectories [@Sheil2003;@Shea2004].
 
-Moreover, we argue that there is a need to delineate the scope of the IDH in time, using ecosystem time-explicit trajectories, and highlight the moment when the conditions for the IDH emerge [@Sheil2003;@Shea2004].
-Indeed, the link between biodiversity, IDH
-and post-disturbance disturbance trajectories might be complicated by the huge diversity found in tropical
-tree communities and the variety of ecological processes that may underlie the IDH
-(colonization, facilitation, coexistence,
-etc)[@Lindenmayer2012;@Garcia_florez2017].
-
-Analysing community response to disturbance requires a set of metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
+Analysing community response to disturbance requiers a set of metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
 The analysis should first consider community composition, which is crucial for conservation issues and which reveals the pool of species fostered (or hampered) by disturbance [@Lavorel2002;@Bellwood2006].
@@
 
-Community trajectories depended on the disturbance intensity, with a threshold of 20-25% of AGB loss for which the taxonomic richness was maximized and the taxonomic evenness remained resilient.
-Below a disturbance intensity threshold (treatments 1 and 2), the taxonomic richness increased and the taxonomic eveness increased first before stabilizing or starting to decrease.
-In those cases, the trees surviving after disturbance remained numerous enough to maintain the high taxonomic richness of the pre-disturbance community [@Bongers2009].
-In parallel, the recruitment of pioneers, infrequent or absent before disturbance, increased the taxonomic richness all the more so that the disturbance was intense [@Martin2015;@Chaudhary2016].
-As these pioneers became more dominant they balanced the usual hyper-dominance of tropical forests and increased the taxonomic evenness [@Baraloto2012a].
-Contrastingly, above the intensity threshold (treatment 3), the taxonomic richness did not exceed the initial value and the taxonomic evenness continuously increased without sign of reversal.
-Above the intensity threshold, the taxonomic richness of surviving trees was too low to be offset by the recruitment of pioneers. 
+Trajectories of taxonomic richness and evenness were markedly different below and above an intensity threshold (\ref{fig:DivTaxo}).
+Below the intensity threshold, for disturbances in the order of treatments 1 and 2, both taxonomic richness and evenness increased according to the disturbance intensity [@Martin2015;@Chaudhary2016].
+This suggested that while trees surviving after disturbance remained numerous enough to maintain the richness of the pre-disturbance community [@Bongers2009], the recruitment of pioneers previously infrequent or absent allowed increasing the taxonomic richness [@Martin2015;@Chaudhary2016].
+As these pioneers became more dominant, they balanced the usual hyper-dominance of tropical forests and increased the taxonomic evenness [@Baraloto2012a].
+Above the intensity threshold, like for treatment 3, the taxonomic richness did not exceed the initial value.
+The richness of surviving trees was then too low to be offset by the recruitment of pioneers. 
 In the Guiana shield indeed, the pool of true pioneers specifically recruited after disturbance is restricted to a few common genera (e.g. *Cecropia* spp., *Vismia* spp.) [@Guitet2018].
-At the same time, pioneers became persistently dominant and prevented a return towards the initial evenness and the recovery of hyper-dominant shade-tolerant species.
 
-After disturbance of intermediate intensity, taxonomic evenness trajectories were hump-shaped, as already observed in the Guiana shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].
-Trajectories peaked at an intermediate time after disturbance, around 15-20 years.
-This time corresponded to a balance between first established pioneers and the more competitve late-successional species.
+Considering different moments of the post-disturbance trajectories, there was always a humped-shaped pattern linking taxonomic richness or evenness and the intensity of disturbance (\ref{fig:IDHplot}).
+Both taxonomic richness and evenness were maximized at an intermediate intensity, around 20-25% of AGB lost, and this along the 30 years following disturbance.
 
-Regarding community functional trajectories in contrast, there was neither intermediate disturbance intensity nor intermediate time after disturbance, which dismissed the IDH.
-Neither functional richness nor functional evenness displayed a humped-shaped pattern, but both rather increased with the recruitment of pioneers that were functionally highly different from the pre-disturbance community [@Denslow1980;@Molino2001].
-Above the intensity threshold however, for the most intense disturbance, community functional richness and evenness started to decrease after 15 to 20 years. 
-Right after disturbance, short-lived species benefited from the resources made available and prevented the establishment of other species. 
-The decline of these short-lived pioneers decreased the functional richness and evenness, but we suggest that the establishment of long-lasting pioneers will follow and that the trajectories will catch up with those observed for intermediate disturbance [@Walker2009].
+Regarding community functional trajectories, no marked differences were observed among the post-disturbance trajectories (\ref{fig:DivTaxo}).
+Whenever the disturbance intensity, functional richness and evenness first increased with the recruitment of pioneers that were functionally highly different from the pre-disturbance community [@Denslow1980;@Molino2001].
+Then, after 15 to 20 years community functional richness and evenness started to decrease, with the decline of the first established species.
+Right after disturbance, short-lived pioneers benefited from the resources made available and prevented the establishment of other species. 
+These pioneers then declined, which decreased the functional richness and evenness [@Walker2009].
+
+Considering different times after disturbance, both functional richness and evenness kept increasing with the disturbance intensity and did not display humped-shaped trajectories (\ref{fig:IDHplot}).
 

### 2019-04-12 WholePlotTrajectories.tex
relecture

@@
 validation of the theory remains controversial
-\citep{Hubbell2001, Fox2013, Sheil2013}.
+\citep{Hubbell2001, Fox2013, Sheil2013}. In tropical forests the
+processes underlying the IDH might indeed be complicated by the huge
+diversity found in tropical tree communities and the variety of
+ecological processes that may underlie the IDH (colonization,
+facilitation, coexistence, etc)
+\citep{Lindenmayer2012, Garcia_florez2017}. It is then crucial to test
+the IDH in this context and determine its scope in time, using ecosystem
+time-explicit trajectories \citep{Sheil2003, Shea2004}.
 
-Moreover, we argue that there is a need to delineate the scope of the
-IDH in time, using ecosystem time-explicit trajectories, and highlight
-the moment when the conditions for the IDH emerge
-\citep{Sheil2003, Shea2004}. Indeed, the link between biodiversity, IDH
-and post-disturbance disturbance trajectories might be complicated by
-the huge diversity found in tropical tree communities and the variety of
-ecological processes that may underlie the IDH (colonization,
-facilitation, coexistence,
-etc)\citep{Lindenmayer2012, Garcia_florez2017}.
-
-Analysing community response to disturbance requires a set of metrics,
+Analysing community response to disturbance requiers a set of metrics,
 to grasp all aspects of community changes
@@
 
-Community trajectories depended on the disturbance intensity, with a
-threshold of 20-25\% of AGB loss for which the taxonomic richness was
-maximized and the taxonomic evenness remained resilient. Below a
-disturbance intensity threshold (treatments 1 and 2), the taxonomic
-richness increased and the taxonomic eveness increased first before
-stabilizing or starting to decrease. In those cases, the trees surviving
-after disturbance remained numerous enough to maintain the high
-taxonomic richness of the pre-disturbance community \citep{Bongers2009}.
-In parallel, the recruitment of pioneers, infrequent or absent before
-disturbance, increased the taxonomic richness all the more so that the
-disturbance was intense \citep{Martin2015, Chaudhary2016}. As these
-pioneers became more dominant they balanced the usual hyper-dominance of
-tropical forests and increased the taxonomic evenness
-\citep{Baraloto2012a}. Contrastingly, above the intensity threshold
-(treatment 3), the taxonomic richness did not exceed the initial value
-and the taxonomic evenness continuously increased without sign of
-reversal. Above the intensity threshold, the taxonomic richness of
-surviving trees was too low to be offset by the recruitment of pioneers.
-In the Guiana shield indeed, the pool of true pioneers specifically
-recruited after disturbance is restricted to a few common genera (e.g.
-\emph{Cecropia} spp., \emph{Vismia} spp.) \citep{Guitet2018}. At the
-same time, pioneers became persistently dominant and prevented a return
-towards the initial evenness and the recovery of hyper-dominant
-shade-tolerant species.
+Trajectories of taxonomic richness and evenness were markedly different
+below and above an intensity threshold (\ref{fig:DivTaxo}). Below the
+intensity threshold, for disturbances in the order of treatments 1 and
+2, both taxonomic richness and evenness increased according to the
+disturbance intensity \citep{Martin2015, Chaudhary2016}. This suggested
+that while trees surviving after disturbance remained numerous enough to
+maintain the richness of the pre-disturbance community
+\citep{Bongers2009}, the recruitment of pioneers previously infrequent
+or absent allowed increasing the taxonomic richness
+\citep{Martin2015, Chaudhary2016}. As these pioneers became more
+dominant, they balanced the usual hyper-dominance of tropical forests
+and increased the taxonomic evenness \citep{Baraloto2012a}. Above the
+intensity threshold, like for treatment 3, the taxonomic richness did
+not exceed the initial value. The richness of surviving trees was then
+too low to be offset by the recruitment of pioneers. In the Guiana
+shield indeed, the pool of true pioneers specifically recruited after
+disturbance is restricted to a few common genera (e.g. \emph{Cecropia}
+spp., \emph{Vismia} spp.) \citep{Guitet2018}.
 
-After disturbance of intermediate intensity, taxonomic evenness
-trajectories were hump-shaped, as already observed in the Guiana shield
-\citep{Baraloto2012a} and in Bornean tropical forests
-\citep{Cannon1998}. Trajectories peaked at an intermediate time after
-disturbance, around 15-20 years. This time corresponded to a balance
-between first established pioneers and the more competitve
-late-successional species.
+Considering different moments of the post-disturbance trajectories,
+there was always a humped-shaped pattern linking taxonomic richness or
+evenness and the intensity of disturbance (\ref{fig:IDHplot}). Both
+taxonomic richness and evenness were maximized at an intermediate
+intensity, around 20-25\% of AGB lost, and this along the 30 years
+following disturbance.
 
-Regarding community functional trajectories in contrast, there was
-neither intermediate disturbance intensity nor intermediate time after
-disturbance, which dismissed the IDH. Neither functional richness nor
-functional evenness displayed a humped-shaped pattern, but both rather
-increased with the recruitment of pioneers that were functionally highly
-different from the pre-disturbance community
-\citep{Denslow1980, Molino2001}. Above the intensity threshold however,
-for the most intense disturbance, community functional richness and
-evenness started to decrease after 15 to 20 years. Right after
-disturbance, short-lived species benefited from the resources made
-available and prevented the establishment of other species. The decline
-of these short-lived pioneers decreased the functional richness and
-evenness, but we suggest that the establishment of long-lasting pioneers
-will follow and that the trajectories will catch up with those observed
-for intermediate disturbance \citep{Walker2009}.
+Regarding community functional trajectories, no marked differences were
+observed among the post-disturbance trajectories (\ref{fig:DivTaxo}).
+Whenever the disturbance intensity, functional richness and evenness
+first increased with the recruitment of pioneers that were functionally
+highly different from the pre-disturbance community
+\citep{Denslow1980, Molino2001}. Then, after 15 to 20 years community
+functional richness and evenness started to decrease, with the decline
+of the first established species. Right after disturbance, short-lived
+pioneers benefited from the resources made available and prevented the
+establishment of other species. These pioneers then declined, which
+decreased the functional richness and evenness \citep{Walker2009}.
+
+Considering different times after disturbance, both functional richness
+and evenness kept increasing with the disturbance intensity and did not
+display humped-shaped trajectories (\ref{fig:IDHplot}).
 

### 2019-07-18 WholePlotTrajectories.Rmd
Final_V1

@@
 abstract: 
- Anticipate the impact of the global change context on tropical forests is now urgent.
- This calls to highlight the taxonomic and functional facets of tree community post-disturbance trajectories and the underlying processes. 
- It also calls to clarify the link between biodiversity and disturbance, specifically the scope of the Intermediate Disturbance Hypothesis (IDH) that is debated in tropical forests.
- In this study, we analyzed the tree community trajectories following a disturbance gradient in a Neotropical forest over 30 years. 
- We considered community taxonomic and functional diversity, composition and redundancy trajectories.
+ In the current global change context, it is urgent to anticipate the fate of tropical forests.
+ This means highlighting tree community response to disturbance and understand the underlying processes.
+ In that respect we aim here to clarify taxonomic and functional post-disturbance  trajectories and determine the scope of the Intermediate Disturbance Hypothesis (IDH) that remains debated in tropical forests.
+ We analyzed community trajectories following a disturbance gradient in a Neotropical forest over 30 years. 
+ We considered trajectories along time of community taxonomic and functional diversity and composition, and functional redundancy.
 We based on the annual botanical inventories of 75 ha of a Neotropical forest and on large trait datasets comprising 7 leaf, stem and life-history traits.
- We highlighted a decoupling between taxonomic trajectories, that differed among communities, and functional trajectories, remained similar and convergent.
-We explained this decoupling by the variations in community functional redundancy mitigating the functional impact of disturbance.
-The humped-shaped taxonomic trajectories which amplitude varied with disturbance intensity validated the IDH (Intermediate Disturbance Hypothesis).
-The IDH was however dismissed for functional diversity.
+ We highlighted a decoupling between taxonomic composition that differed among communities, and functional composition, that remained similar and convergent. 
+ The taxonomic diversity followed humped-shaped trajectories depending on the disturbance intensity that validated the IDH (Intermediate Disturbance Hypothesis), while functional diversity trajectories homogeneous among all plots dismissed the IDH.
+ We explained this decoupling by the variations in community functional redundancy that mitigates the functional impact of disturbance. 
  Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years.
@@
   \usepackage{textcomp}
-  \DeclareUnicodeCharacter{B0}{\textdegree}
   \usepackage{tabu}
@@
 They provide wood and multiple non-timber forest products, shelter a diversified fauna, and ensure cultural and human well-being.
-They regulate as well the local and regional climates, and the carbon, water and nutrient cycles.
-However, the growing demand in forests products together with current global changes increase the pressure on remaining natural undisturbed forests [@Morales-Hidalgo2015].
-These threats affect the disturbance regime that naturally shapes and maintains the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
-To anticipate the fate of tropical forests in this context, understanding the response of communities to disturbance and deciphering the underlying ecological processes is of utmost importance.
-Disturbance impacts forest communities in changing both abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
-Although the forest cover is generally maintained after disturbance, there is a succession of changes in the structure and functioning of ecological communities that defines a post-disturbance ecosystem trajectory.
-For now, these trajectories have been largely studied through trajectories of forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
-Recent works have identified some of the determinants of post-disturbance biomass trajectories, like pre-disturbance forest structure and composition or post-disturbance environmental parameters [@Herault2018].
-Regarding tree community diversity and composition however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001]. 
-Given the variety of species response to disturbance and the huge diversity of tropical forests, manifold post-disturbance biodiversity trajectories might
-emerge [@Lindenmayer2012;@Garcia_florez2017].
+They regulate the local and regional climates, as well as the carbon, water and nutrient cycles.
+However, the growing demand in forests products together with current global changes increase the pressure on remaining undisturbed forests [@Morales-Hidalgo2015].
+These threats affect the natural disturbance regime that defines and maintains the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
+To anticipate the fate of tropical forests, it is urgent to understand the response to disturbance of tree communities and the underlying ecological processes.
+The forest cover is generally maintained following disturbance, but modifications in the fluxes of light, heat and water [@Goulamoussene2017] change community abiotic and biotic environments.
+These changes translate into post-disturbance community trajectories that have been largely studied through trajectories of forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016].
+Recent works have identified some of the determinants of post-disturbance biomass trajectories, like the structure and composition of the pre-disturbance community or the post-disturbance environmental parameters [@Herault2018].
+Community diversity and composition trajectories, however, have not been as thoroughly understood [@Guitet2018;@Molino2001], and manifold biodiversity trajectories might emerge given the variety of species response to disturbance and the huge diversity of tropical forests [@Lindenmayer2012;@Garcia_florez2017].
 
 An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
-The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events, stating that community diversity peaks at intermediate level of disturbance [@Connell1978].
-The theory assumes that the environmental fluctuations following disturbance foster both competitively superior species and fast colonizers and thus prevent competitive exclusion [@Shea2004;@Pulsford2016].
-In tropical forests, however, observations of the IDH often diverge from theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
-In tropical forests the processes underlying the IDH might indeed be complicated by the huge diversity found in tropical tree communities and the variety of ecological processes that may underlie the IDH
-(colonization, facilitation, coexistence, etc) [@Lindenmayer2012;@Garcia_florez2017].
-It is then crucial to test the IDH in this context and determine its scope in time, using ecosystem time-explicit trajectories [@Sheil2003;@Shea2004].
+The IDH states a relationship between the community diversity and the intensity and frequency of disturbance events, and a diversity peak at intermediate level of disturbance [@Connell1978].
+This bases on the fluctuations of community environment following disturbance that foster both competitively superior species and fast colonizers, and prevents competitive exclusion [@Shea2004;@Pulsford2016].
+In tropical forests, however, observations of the IDH often diverge from theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the underlying processes might be complicated by the huge diversity of tropical tree communities [@Lindenmayer2012;@Garcia_florez2017].
+In this context, it is urgent to validate the IDH that remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
 
-Analysing community response to disturbance requiers a set of metrics, to grasp all aspects of community changes [@Sheil2003;@Shea2004;@Mayfield2010].
-The analysis should first consider community composition, which is crucial for conservation issues and which reveals the pool of species fostered (or hampered) by disturbance [@Lavorel2002;@Bellwood2006].
-Then, diversity metrics encompassing both richness and evenness should also be considered to assess the changes in community abundance distribution. 
-Besides, functional approaches have been shown to usefully complement pure taxonomic approaches because they shed light on the species biological attributes that directly link community diversity, composition and redundancy to ecosystem functioning [@Violle2007b;@Baraloto2012a].
-In that respect, a vast literature allowed recognizing major traits that represent species ecological strategy and determine how they respond to changing conditions [@Diaz2005]. 
-Specifically in tropical forests, the functional approach revealed the post-disturbance deterministic processes entailing a shift from a dominance of "conservative" slow-growing species dealing with scarce resources, to a dominance of "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Rees2001;@Reich2014;@Herault2011].
+Analysing community response to disturbance and grasp all aspects of community changes requires a whole set of metrics [@Sheil2003;@Shea2004;@Mayfield2010].
+The analysis should first consider community composition, which is crucial for conservation issues and reveals the pool of species fostered (or hampered) by disturbance [@Lavorel2002;@Bellwood2006].
+This should be completed with diversity metrics encompassing both richness and evenness to assess the changes in community abundance distribution. 
+Besides, functional approaches have been shown to usefully complement pure taxonomic approaches as they shed light on the species biological attributes directly linking community diversity, composition and redundancy to ecosystem functioning [@Violle2007b;@Baraloto2012a].
+In that respect, a vast literature allowed recognizing major traits representing species ecological strategy and determining how they respond to changing conditions [@Diaz2005]. 
+Specifically in tropical forests, the functional approach revealed the deterministic processes entailing after disturbance a shift from a dominance of "conservative" slow-growing species dealing with scarce resources, to a dominance of "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Rees2001;@Reich2014;@Herault2011].
 This shift is translated into the trajectories of average community value of key functional traits related to resource acquisition (leaf and stem traits) and life-history strategy (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
@@
 We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, (i) we draw taxonomic and functional post-disturbance trajectories and examine the underlying ecological process, (ii) we discuss the scope of the IDH regarding the different facets of community diversity and the time after disturbance, and (iii) we analyze community resilience and time to recovery.
+Specifically, (i) we drew taxonomic and functional post-disturbance trajectories and examined the underlying ecological processes, (ii) we discussed the scope of the IDH regarding taxonomic and functional facets of community diversity, and (iii) we analyzed community resilience and time to recovery.
 
@@
 
-Paracou station in French Guiana (5\textdegree 18'N and 52\textdegree 53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26\textdegree C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
+Paracou station in French Guiana (5° 18'N and 52° 53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26° C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
 Elevation ranges from 5 to 50 m.
@@
 Treatment T2 additionally comprised a thinning treatment by poison-girdling of non-commercial, randomly selected species with an average 30 trees/ha with 40 cm minimum DBH.
-Treatment T3 additionally comprised the logging of 15 trees /ha with 40 cm minimum DBH and the poison-gidling of 20 trees/ha with a 50 cm minimum DBH, all belonging to non-commercial species. 
+Treatment T3 additionally comprised the logging of 15 trees/ha with 40 cm minimum DBH and the poison-gidling of 20 trees/ha with a 50 cm minimum DBH, all belonging to non-commercial species. 
 Considering the silvicultural treatments and the following damage, disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Rejou2017].
@@
 
-This variability of protocols in time raised methodological issues as vernacular names usually correspond to different botanical species. 
-This resulted in significant taxonomic uncertainty that had to be propagated to composition and diversity metrics. 
-The uncertainty propagation was done through a Bayesian framework reconstituting complete inventories at genus level from real incomplete ones on the basis of vernacular/botanical names association. 
+This variability of protocols in time raised methodological issues as vernacular names usually correspond to various botanical species. 
+This resulted in significant taxonomic uncertainty that had to be accounted for in the measure of composition and diversity metrics. 
+The uncertainty propagation based on a Bayesian framework using vernacular/botanical names associations to reconstitute complete inventories at genus level from real incomplete ones. 
 Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,..., \alpha_V\big]$ observed across all inventories between each vernacular name *v* and all species $\big[s_1, s_2,..., s_N\big]$:
@@
 Whenever a species was in the dataset but missed some trait values (10% of the species), missing values were filled using multivariate imputation by chained equation [@Mice2011].
-To account for the phylogenetic signal in the filling process, imputations based on samples from the same genus or from the same family. 
+To account for the phylogenetic signal in the filling process, imputations based on samples of species from the same genus or from the same family. 
 Whenever a species was not in the dataset, it was attributed a set of trait values randomly sampled among species of the next higher taxonomic level (same genus or family). 
@@
 In disturbed communities, both trajectories depended on the disturbance intensity, with their maximum values in time being positively correlated to \%AGB loss $\rho_{Spearman}^{Richness}=0.76$ and $\rho_{Spearman}^{Rao}=0.60$.
-Functional richness and evenness displayed for low disturbance intensity a low but long-lasting increase up to a maximum reached after 20-25 years. 
+For low disturbance intensity, functional richness and evenness displayed a low but long-lasting increase up to a maximum reached after 20-25 years. 
 For high disturbance intensity, they generally displayed a fast but short increase followed after 10 years by a slow decrease towards the initial values.
@@
 
-Our analysis highlighted the decoupling between functional and taxonomic  trajectories, with taxonomic trajectories maintaining initial differences among plots while functional trajectories converged in the functional space.
-
-The humped-shaped taxonomic trajectories which amplitude varied with disturbance intensity validated the IDH.
-Taxonomic diversity was maximized at an intensity threshold of 20-25% AGB loss and 20 years after disturbance, when the recruitment of pioneers and late-successionals were balanced.
-The IDH however did not apply to the functional diversity.
-The decoupling between taxonomic and functional response was linked to the disturbance impact on the functional redundancy, that proved determining in community recovery.
+Our analysis highlighted the decoupling between functional and taxonomic trajectories.
+While the initial differences in taxonomic composition among plots were maintained, the functionalcomposition trajectories converged in the functional space.
+In terms of diversity, only the taxonomic trajectories validated the IDH that explained humped-shaped post-disturbance trajectories whith an amplitude depending on the disturbance intensity.
+The decoupling between taxonomic and functional response was explained by variations of community functional redundancy that mitigated the functional impact of disturbance, and appeared as a determinant of community recovery.
 
@@
 
-Regarding taxonomic composition, the different tree communities were substantially different before disturbance, and this is visualized by their distinct starting points on the NMDS axis 2. 
+Community taxonomic composition substantially differed before disturbance, as materialized by their distinct location on the NMDS axis 2. 
 These initial differences were maintained all along the 30 years following disturbance, with the disturbance leading a displacement on the NMDS axis 1 only. 
-Changes were thus similar and may correspond to the recruitment of a group of pioneers, e.g.
-*Cecropia* spp., *Vismia* spp.,  shared by all plots and whatever their initial taxonomic differences and the intensity of disturbance [@Denslow2000;@Bongers2009].
-Taxonomic trajectories seems to start a recovery to the initial composition which, although far from fully achieved after 30 years, suggested that taxonomic composition is resilient and that the initial differences will be maintained in time [@Folke2006].
-This would show that species not belonging to the pre-disturbance community were rarely recruited in the long-term, probably because of the common dispersal limitations among tropical tree species [@Svenning2005].
+Taxonomic composition changes were similar among plot and may correspond to the recruitment of a group of pioneers, e.g.
+*Cecropia* spp., *Vismia* spp.,  shared by all plots whatever their initial taxonomic differences and the intensity of disturbance [@Denslow2000;@Bongers2009].
+Taxonomic trajectories initiated a recovery towards the initial composition.
+This recovery, although far from being achieved after 30 years, suggested the resilience of community taxonomic composition and the maintenance of initial composition differences [@Folke2006]. 
+This suggested that species not belonging to the pre-disturbance community were rarely recruited in the long-term, probably because of species dispersal limitation that is common among tropical species [@Svenning2005].
 
-Regarding functional composition, initial communities had similar starting points and this contrasts to taxonomic composition. 
-This means that, despite differences in species composition, the functional signature of the tree communities were quite similar. 
-Following disturbance, functional trajectories were consistent in the NMDS plane with displacement intensities linked to disturbance intensities and thus relied upon the recruitment of a shared pool of functional types previously infrequent or absent.
-This common pool is composed of pioneer resource-acquisitive (low leaf toughness, wood specific gravity, maximum height and high Specific Leaf area) strategies that translated by a displacement to the right along the first NMDS axis [@Westoby1998;@Wright2004;@Reich2014].
-Thereafter, the pioneers recruited primarily were progressively excluded by long-lived, more competitive and shade-tolerant species. 
-The community functional composition then quickly returned towards more resource-conservative strategies, suggesting the recovery of the initial functional composition. 
-This recovery translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig. \ref{fig:NMDSplans}).
+Community functional composition trajectories, in contrast, were similar in the functional space [@Fukami2005].
+The amplitude of the compositional changes depended on the disturbance intensity.
+Functional trajectories were probably driven by the recruitment of species infrequent or absent before disturbance and belonging to a pool common to all plots. 
+This common pool was composed of pioneers with resource-acquisitive strategies (low leaf toughness, wood specific gravity, maximum height and high Specific Leaf area) which recruitment drove a displacement to the right along the first NMDS axis [@Westoby1998;@Wright2004;@Reich2014].
+Thereafter, the first recruited pioneers were progressively excluded by long-lived, more competitive and shade-tolerant species, which drove the recovery of the initial functional composition. 
+This recovery corresponded to a return of the composition towards resource-conservative strategies that translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig.
+\ref{fig:NMDSplans}).
 
-Both taxonomic and functional composition trajectories initiated a return towards pre-disturbance state after 30 years, which highlighted the taxonomic and functional resilience of communities. 
-Taxonomic and functional trajectories however appeared decoupled: while taxonomic trajectories maintained the initial differences among communities, the functional trajectories where similar and convergent in the functional space [@Fukami2005] exemplifying the simultaneous operation of trait-based assembly rules and species-level priority effects that shapes tree community assembly in this forest, making it both deterministic in functional space and historically contingent in taxonomic space.
+The decoupling between taxonomic and functional trajectories suggested the simultaneous operation of trait-based assembly rules and species-level priority effects to shape tree community assembly in this forest.
+Tree community assembly would then be both deterministic in the functional space and historically contingent in the taxonomic space.
 
@@
 
-Trajectories of taxonomic richness and evenness were markedly different below and above an intensity threshold (\ref{fig:DivTaxo}).
-Below the intensity threshold, for disturbances in the order of treatments 1 and 2, both taxonomic richness and evenness increased according to the disturbance intensity [@Martin2015;@Chaudhary2016].
-This suggested that while trees surviving after disturbance remained numerous enough to maintain the richness of the pre-disturbance community [@Bongers2009], the recruitment of pioneers previously infrequent or absent allowed increasing the taxonomic richness [@Martin2015;@Chaudhary2016].
-As these pioneers became more dominant, they balanced the usual hyper-dominance of tropical forests and increased the taxonomic evenness [@Baraloto2012a].
-Above the intensity threshold, like for treatment 3, the taxonomic richness did not exceed the initial value.
-The richness of surviving trees was then too low to be offset by the recruitment of pioneers. 
+Trajectories of taxonomic richness and evenness differed markedly below and above an intensity threshold (\ref{fig:DivTaxo}).
+For low and intermediate disturbances, both taxonomic richness and evenness increased according to the disturbance intensity [@Martin2015;@Chaudhary2016].
+This suggested that the recruitment of pioneers previously infrequent or absent increased the taxonomic richness [@Martin2015;@Chaudhary2016], and that trees surviving after disturbance remained numerous enough to maintain the richness of the pre-disturbance community [@Bongers2009].
+The pioneers thus recruited became abundant, which balanced the usual hyper-dominance of tropical forests and increased the taxonomic evenness [@Baraloto2012a].
+Above the intensity threshold, like for treatment 3 (high intensity), the taxonomic richness did not exceed the initial value in the first years following disturbance.
+This suggested that the richness of surviving trees was lower than this of the pre-disturbance community, and that this decrease was not offset by the recruitment of pioneers. 
 In the Guiana shield indeed, the pool of true pioneers specifically recruited after disturbance is restricted to a few common genera (e.g. *Cecropia* spp., *Vismia* spp.) [@Guitet2018].
+Considering different time after disturbance, there was always a humped-shaped pattern linking the disturbance intensity with the post-disturbance increase in taxonomic richness and evenness (\ref{fig:IDHplot}).
+Both taxonomic richness and evenness were maximized at an intermediate intensity, around 20-25% of AGB lost.
 
-Considering different moments of the post-disturbance trajectories, there was always a humped-shaped pattern linking taxonomic richness or evenness and the intensity of disturbance (\ref{fig:IDHplot}).
-Both taxonomic richness and evenness were maximized at an intermediate intensity, around 20-25% of AGB lost, and this along the 30 years following disturbance.
-
-Regarding community functional trajectories, no marked differences were observed among the post-disturbance trajectories (\ref{fig:DivTaxo}).
-Whenever the disturbance intensity, functional richness and evenness first increased with the recruitment of pioneers that were functionally highly different from the pre-disturbance community [@Denslow1980;@Molino2001].
-Then, after 15 to 20 years community functional richness and evenness started to decrease, with the decline of the first established species.
-Right after disturbance, short-lived pioneers benefited from the resources made available and prevented the establishment of other species. 
-These pioneers then declined, which decreased the functional richness and evenness [@Walker2009].
-
-Considering different times after disturbance, both functional richness and evenness kept increasing with the disturbance intensity and did not display humped-shaped trajectories (\ref{fig:IDHplot}).
+Regarding community functional trajectories, however, no marked differences were observed among post-disturbance trajectories (\ref{fig:DivTaxo}).
+Whenever the disturbance intensity, the recruitment of pioneers functionally highly different from the pre-disturbance community, first increased both community functional richness and evenness [@Denslow1980;@Molino2001].
+These recruitment hampered the establishment of other species in the first place.
+After 15 to 20 years, the first recruits declined and were replaced by species functionally more similar to the pre-disturbance community, which decreased the functional richness and evenness [@Walker2009].
 
@@
 The loss of species following disturbance decreased the functional redundancy during the first 15 years. 
-Progressively though, the functional redundancy was restored through the replacement of the resource-acquisitive strategies species by more late-successional species that were functionally closer to the pre-disturbance community.
-This replacement followed the lottery recruitment rules, implying an easy recruitment for the first species but becoming increasingly hampered by the emergence of interspecific competition [@Busing2002] and this explains why the slope of the recovery trajectory became less and less significant 20 years after disturbance.
-The recovery of the functional redundancy then relied upon the random process of species recruitment and was increasingly slow and difficult to anticipate
-[@Elmqvist2003;@Diaz2005]. 
-This suggests a low resilience of the functional redundancy with the random recovery of infrequent species increasing the risks to lose keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
+Progressively though, the functional redundancy was restored through the replacement of the resource-acquisitive strategies species by more late-successional species, functionally closer to the pre-disturbance community.
+This replacement followed the lottery recruitment rules, implying an easy recruitment for the first species but becoming increasingly hampered by the emergence of interspecific competition [@Busing2002].
+This explained why the recovery trajectory slowed down 20 years after disturbance. 
+The recovery of the functional redundancy then relied upon the random process of species recruitment and was increasingly slow and difficult to anticipate [@Elmqvist2003;@Diaz2005]. 
+This suggested a low resilience of the functional redundancy with the random recovery of infrequent species increasing the risks to lose keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
 Infrequent species might indeed have unique functional characteristics, apart from those considered here, in the ecosystem or be a key resource for some fauna [@Schleuning2016].
@@
 
-Post-disturbance trajectories of tree community composition and diversity were both shaped by the recruitment of a determined pool of pioneers identical
-among local communities and independent of the disturbance intensity.
-(i) Composition trajectories in taxonomic and functional space nevertheless appeared decoupled.
-While functional trajectories remained similar in the functional space and converged towards the recovery of a comparable initial state, taxonomic trajectories showed initial differences in community composition that were maintained along time.
-Community high functional redundancy mediated this decoupling, as the loss of a species does not necessarily entails the loss of its functional charactristics.
-(ii) Diversity trajectories were contrasted as well.
-While the functional trajectories remained similar whatever the disturbance intensity, taxonomic trajectories were markedly different with a threshold at 20-25\% AGB removed for which taxonomic richness was maximized.
+Post-disturbance trajectories of tree community composition and diversity were driven by the recruitment of a determined pool of pioneers, identical among local communities and independent of the disturbance intensity.
+The taxonomic composition trajectories maintained the initial differences among communities, while the functional trajectories were similar and converged in the functional space towards the recovery of the initial composition.
+The diversity trajectories were contrasted as well. 
+While the functional trajectories remained similar whatever the disturbance intensity, taxonomic trajectories were markedly different from a threshold of 20-25\% AGB removed that maximized the taxonomic richness.
 The Intermediate Disturbance Hypothesis applied well to taxonomic diversity, but not to the functional diversity.
-Taxonomic diversity was maximized at an intermediate time after disturbance, that was around 25 years after disturbance when the recruitment of early- and late-successional species were balanced.
-Whenever the disturbance intensity, community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied upon the random lottery recruitment of rare species.
+The decoupling between taxonomic and functional trajectories was mediated by the variations in functional redundancy, as the loss of a species does not necessarily entails the loss of its functional characteristics.
+Community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied upon the random lottery recruitment of rare species.
 Given the long-term impacts of disturbance observed, we suggest that 30 years is not enough time for tropical communities to recover, even after relatively low intensity disturbance.

### 2019-07-18 WholePlotTrajectories2.tex
Final_V1

--- 2019-04-12 WholePlotTrajectories.tex
@@
 % User-adder preamble
-\usepackage{textcomp} \DeclareUnicodeCharacter{B0}{\textdegree}
-\usepackage{tabu} \usepackage{caption}
+\usepackage{textcomp} \usepackage{tabu} \usepackage{caption}
 \captionsetup{justification = justified}
@@
 leaf, stem and life-history traits. We highlighted a decoupling between
-taxonomic trajectories, that differed among communities, and functional
-trajectories, remained similar and convergent. We explained this
-decoupling by the variations in community functional redundancy
-mitigating the functional impact of disturbance. The humped-shaped
-taxonomic trajectories which amplitude varied with disturbance intensity
-validated the IDH (Intermediate Disturbance Hypothesis). The IDH was
-however dismissed for functional diversity. Although consistent, the
-recovery of community composition, diversity and redundancy remained
-unachieved after 30 years. This acknowledged the need of decades-long
-cycles without disturbance to ensure a complete recovery, and questioned
+taxonomic composition trajectories, that differed among communities, and
+functional trajectories, that remained similar and convergent. The
+taxonomic diversity followed humped-shaped trajectories depending on the
+disturbance intensity that validated the IDH (Intermediate Disturbance
+Hypothesis), while functional diversity trajectories homogeneous among
+all plots dismissed the IDH. We explained this decoupling by the
+variations in community functional redundancy that mitigates the
+functional impact of disturbance. Although consistent, the recovery of
+community composition, diversity and redundancy remained unachieved
+after 30 years. This acknowledged the need of decades-long cycles
+without disturbance to ensure a complete recovery, and questioned
 tropical forest community resilience after repeated disturbances.
@@
 assumes that the environmental fluctuations following disturbance foster
-both competitively superior species and fast colonizers and thus prevent
+both competitively superior species and fast colonizers, thus preventing
 competitive exclusion \citep{Shea2004, Pulsford2016}. In tropical
@@
 validation of the theory remains controversial
-\citep{Hubbell2001, Fox2013, Sheil2013}. In tropical forests the
-processes underlying the IDH might indeed be complicated by the huge
-diversity found in tropical tree communities and the variety of
-ecological processes that may underlie the IDH (colonization,
-facilitation, coexistence, etc)
-\citep{Lindenmayer2012, Garcia_florez2017}. It is then crucial to test
-the IDH in this context and determine its scope in time, using ecosystem
-time-explicit trajectories \citep{Sheil2003, Shea2004}.
+\citep{Hubbell2001, Fox2013, Sheil2013}. In this context it is urgent to
+test the IDH and determine its scope, as the underlying processes might
+be complicated by the huge diversity of tropical tree communities
+\citep{Lindenmayer2012, Garcia_florez2017}.
 
@@
 consider community composition, which is crucial for conservation issues
-and which reveals the pool of species fostered (or hampered) by
-disturbance \citep{Lavorel2002, Bellwood2006}. Then, diversity metrics
-encompassing both richness and evenness should also be considered to
-assess the changes in community abundance distribution. Besides,
-functional approaches have been shown to usefully complement pure
-taxonomic approaches because they shed light on the species biological
-attributes that directly link community diversity, composition and
-redundancy to ecosystem functioning \citep{Violle2007b, Baraloto2012a}.
-In that respect, a vast literature allowed recognizing major traits that
-represent species ecological strategy and determine how they respond to
-changing conditions \citep{Diaz2005}. Specifically in tropical forests,
-the functional approach revealed the post-disturbance deterministic
+and reveals the pool of species fostered (or hampered) by disturbance
+\citep{Lavorel2002, Bellwood2006}. Then, diversity metrics encompassing
+both richness and evenness should also be considered to assess the
+changes in community abundance distribution. Besides, functional
+approaches have been shown to usefully complement pure taxonomic
+approaches as they shed light on the species biological attributes that
+directly link community diversity, composition and redundancy to
+ecosystem functioning \citep{Violle7b, Baraloto2012a}. In that respect,
+a vast literature allowed recognizing major traits that represent
+species ecological strategy and determine how they respond to changing
+conditions \citep{Diaz2005}. Specifically in tropical forests, the
+functional approach revealed the post-disturbance deterministic
 processes entailing a shift from a dominance of ``conservative''
@@
 trajectories and examine the underlying ecological process, (ii) we
-discuss the scope of the IDH regarding the different facets of community
-diversity and the time after disturbance, and (iii) we analyze community
-resilience and time to recovery.
+discuss the scope of the IDH regarding taxonomic and functional facets
+of community diversity, and (iii) we analyze community resilience and
+time to recovery.
 
@@
 
-Paracou station in French Guiana (5\textdegree 18'N and
-52\textdegree 53'W) is located in a lowland tropical rain forest in a
-tropical wet climate with mean annual temperature of 26\textdegree C,
-mean annual precipitation averaging 2980 mm.y\textsuperscript{-1} (30-y
-period) and a 3-month dry season (\textless{} 100
-mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
-one-month dry season in March \citep{Wagner2011}. Elevation ranges from
-5 to 50 m. Across all plots the topography mainly corresponds to
-hilltops or hillsides, while bottomlands cover less than 1 \% of the
-area. Plots are shallow ferralitic acrisols over a layer of transformed
-saprolite with low permeability and lateral drainage. Soil conditions
-are homogeneous, to the exception of the highest hilltops where the
-thick surface allows a free vertical drainage
+Paracou station in French Guiana (5° 18'N and 52° 53'W) is located in a
+lowland tropical rain forest in a tropical wet climate with mean annual
+temperature of 26° C, mean annual precipitation averaging 2980
+mm.y\textsuperscript{-1} (30-y period) and a 3-month dry season
+(\textless{} 100 mm.month\textsuperscript{-1}) from mid-August to
+mid-November, and a one-month dry season in March \citep{Wagner2011}.
+Elevation ranges from 5 to 50 m. Across all plots the topography mainly
+corresponds to hilltops or hillsides, while bottomlands cover less than
+1 \% of the area. Plots are shallow ferralitic acrisols over a layer of
+transformed saprolite with low permeability and lateral drainage. Soil
+conditions are homogeneous, to the exception of the highest hilltops
+where the thick surface allows a free vertical drainage
 \citep{Gourlet-Fleury2004}.
@@
 Our analysis highlighted the decoupling between functional and taxonomic
-trajectories, with taxonomic trajectories maintaining initial
-differences among plots while functional trajectories converged in the
-functional space.
-
-The humped-shaped taxonomic trajectories which amplitude varied with
-disturbance intensity validated the IDH. Taxonomic diversity was
-maximized at an intensity threshold of 20-25\% AGB loss and 20 years
-after disturbance, when the recruitment of pioneers and
-late-successionals were balanced. The IDH however did not apply to the
-functional diversity. The decoupling between taxonomic and functional
-response was linked to the disturbance impact on the functional
-redundancy, that proved determining in community recovery.
+trajectories. In terms of composition, while taxonomic trajectories
+maintained the initial differences among plots, the functional
+trajectories converged in the functional space. In terms of diversity,
+only the taxonomic trajectories validated the IDH that explained the
+humped-shaped taxonomic trajectories which amplitude depended on the
+disturbance intensity. The decoupling between taxonomic and functional
+response was explained by variations of community functional redundancy,
+that mitigated the functional impact of disturbance and appeared as a
+determinant of community recovery.
 
@@
 
-Regarding taxonomic composition, the different tree communities were
-substantially different before disturbance, and this is visualized by
-their distinct starting points on the NMDS axis 2. These initial
-differences were maintained all along the 30 years following
-disturbance, with the disturbance leading a displacement on the NMDS
-axis 1 only. Changes were thus similar and may correspond to the
-recruitment of a group of pioneers, e.g. \emph{Cecropia} spp.,
-\emph{Vismia} spp., shared by all plots and whatever their initial
-taxonomic differences and the intensity of disturbance
-\citep{Denslow2000, Bongers2009}. Taxonomic trajectories seems to start
-a recovery to the initial composition which, although far from fully
-achieved after 30 years, suggested that taxonomic composition is
-resilient and that the initial differences will be maintained in time
-\citep{Folke2006}. This would show that species not belonging to the
+Community taxonomic composition substantially differed before
+disturbance, as materialized by their distinct location on the NMDS axis
+2. These initial differences were maintained all along the 30 years
+following disturbance, with the disturbance leading a displacement on
+the NMDS axis 1 only. Taxonomic composition changes were similar among
+plot and may correspond to the recruitment of a group of pioneers, e.g.
+\emph{Cecropia} spp., \emph{Vismia} spp., shared by all plots whatever
+their initial taxonomic differences and the intensity of disturbance
+\citep{Denslow2000, Bongers2009}. Taxonomic trajectories initiated a
+recovery towards the initial composition which, although far from being
+achieved after 30 years, suggested the resilience of community taxonomic
+composition and the maintenance of initial composition differences along
+time \citep{Folke2006}. This suggested that species not belonging to the
 pre-disturbance community were rarely recruited in the long-term,
@@
 
-Regarding functional composition, initial communities had similar
-starting points and this contrasts to taxonomic composition. This means
-that, despite differences in species composition, the functional
-signature of the tree communities were quite similar. Following
-disturbance, functional trajectories were consistent in the NMDS plane
-with displacement intensities linked to disturbance intensities and thus
-relied upon the recruitment of a shared pool of functional types
-previously infrequent or absent. This common pool is composed of pioneer
-resource-acquisitive (low leaf toughness, wood specific gravity, maximum
-height and high Specific Leaf area) strategies that translated by a
-displacement to the right along the first NMDS axis
-\citep{Westoby1998, Wright2004, Reich2014}. Thereafter, the pioneers
-recruited primarily were progressively excluded by long-lived, more
-competitive and shade-tolerant species. The community functional
-composition then quickly returned towards more resource-conservative
-strategies, suggesting the recovery of the initial functional
-composition. This recovery translated in the functional plane by a
-displacement left along the first axis and upward along the second axis
-(Fig. \ref{fig:NMDSplans}).
+Community functional composition in contrast followed similar
+trajectories in the functional space and the amplitude of the
+compositional changes depended on the disturbance intensity. Functional
+trajectories were probably driven by the recruitment of a pool of
+species common to all plots that were infrequent or absent before
+disturbance. This common pool was composed of pioneers with
+resource-acquisitive strategies (low leaf toughness, wood specific
+gravity, maximum height and high Specific Leaf area) which recruitment
+drove a displacement to the right along the first NMDS axis
+\citep{Westoby1998, Wright2004, Reich2014}. Thereafter, the first
+recruited pioneers were progressively excluded by long-lived, more
+competitive and shade-tolerant species, which suggested the recovery of
+the initial functional composition. This recovery translated into a
+return towards resource-conservative strategies and in the functional
+plane by a displacement left along the first axis and upward along the
+second axis (Fig. \ref{fig:NMDSplans}).
 
-Both taxonomic and functional composition trajectories initiated a
-return towards pre-disturbance state after 30 years, which highlighted
-the taxonomic and functional resilience of communities. Taxonomic and
-functional trajectories however appeared decoupled: while taxonomic
-trajectories maintained the initial differences among communities, the
-functional trajectories where similar and convergent in the functional
-space \citep{Fukami2005} exemplifying the simultaneous operation of
-trait-based assembly rules and species-level priority effects that
-shapes tree community assembly in this forest, making it both
-deterministic in functional space and historically contingent in
-taxonomic space.
+The return after 30 years towards pre-disturbance state of composition
+trajectories highlighted the taxonomic and functional resilience of
+communities. Taxonomic and functional trajectories however appeared
+decoupled: while taxonomic trajectories maintained the initial
+differences among communities, the functional trajectories where similar
+and convergent in the functional space \citep{Fukami2005}. This
+decoupling exemplified the simultaneous operation of trait-based
+assembly rules and species-level priority effects that shapes tree
+community assembly in this forest, making it both deterministic in
+functional space and historically contingent in taxonomic space.
 
@@
 
-Trajectories of taxonomic richness and evenness were markedly different
-below and above an intensity threshold (\ref{fig:DivTaxo}). Below the
-intensity threshold, for disturbances in the order of treatments 1 and
-2, both taxonomic richness and evenness increased according to the
-disturbance intensity \citep{Martin2015, Chaudhary2016}. This suggested
-that while trees surviving after disturbance remained numerous enough to
-maintain the richness of the pre-disturbance community
-\citep{Bongers2009}, the recruitment of pioneers previously infrequent
-or absent allowed increasing the taxonomic richness
-\citep{Martin2015, Chaudhary2016}. As these pioneers became more
-dominant, they balanced the usual hyper-dominance of tropical forests
-and increased the taxonomic evenness \citep{Baraloto2012a}. Above the
-intensity threshold, like for treatment 3, the taxonomic richness did
-not exceed the initial value. The richness of surviving trees was then
-too low to be offset by the recruitment of pioneers. In the Guiana
-shield indeed, the pool of true pioneers specifically recruited after
-disturbance is restricted to a few common genera (e.g. \emph{Cecropia}
-spp., \emph{Vismia} spp.) \citep{Guitet2018}.
+Trajectories of taxonomic richness and evenness differed markedly below
+and above an intensity threshold (\ref{fig:DivTaxo}). For disturbances
+in the order of treatments 1 and 2, below the intensity threshold, both
+taxonomic richness and evenness increased according to the disturbance
+intensity \citep{Martin2015, Chaudhary2016}. This suggested that trees
+surviving after disturbance remained numerous enough to maintain the
+richness of the pre-disturbance community \citep{Bongers2009}, and that
+the recruitment of pioneers previously infrequent or absent increased
+the taxonomic richness \citep{Martin2015, Chaudhary2016}. The pioneers
+thus recruited became more dominant, which balanced the usual
+hyper-dominance of tropical forests and increased the taxonomic evenness
+\citep{Baraloto2012a}. Above the intensity threshold, like for treatment
+3, the taxonomic richness did not exceed the initial value. This
+suggested that the richness of surviving trees was lower than this of
+the pre-disturbance community,and the difference was not offset by the
+recruitment of pioneers. In the Guiana shield indeed, the pool of true
+pioneers specifically recruited after disturbance is restricted to a few
+common genera (e.g. \emph{Cecropia} spp., \emph{Vismia} spp.)
+\citep{Guitet2018}.
 
-Considering different moments of the post-disturbance trajectories,
-there was always a humped-shaped pattern linking taxonomic richness or
-evenness and the intensity of disturbance (\ref{fig:IDHplot}). Both
-taxonomic richness and evenness were maximized at an intermediate
-intensity, around 20-25\% of AGB lost, and this along the 30 years
-following disturbance.
+A humped-shaped pattern always linked both taxonomic richness and
+evenness with the disturbance intensity, at different time after
+disturbance (\ref{fig:IDHplot}). Both taxonomic richness and evenness
+were maximized at an intermediate intensity, around 20-25\% of AGB lost.
 
 Regarding community functional trajectories, no marked differences were
-observed among the post-disturbance trajectories (\ref{fig:DivTaxo}).
-Whenever the disturbance intensity, functional richness and evenness
-first increased with the recruitment of pioneers that were functionally
-highly different from the pre-disturbance community
-\citep{Denslow1980, Molino2001}. Then, after 15 to 20 years community
-functional richness and evenness started to decrease, with the decline
-of the first established species. Right after disturbance, short-lived
-pioneers benefited from the resources made available and prevented the
-establishment of other species. These pioneers then declined, which
-decreased the functional richness and evenness \citep{Walker2009}.
-
-Considering different times after disturbance, both functional richness
-and evenness kept increasing with the disturbance intensity and did not
-display humped-shaped trajectories (\ref{fig:IDHplot}).
+observed among post-disturbance trajectories (\ref{fig:DivTaxo}). First,
+whenever the disturbance intensity, the recruitment of pioneers that
+were functionally highly different from the pre-disturbance community
+increased both community functional richness and evenness
+\citep{Denslow1980, Molino2001} and hampered the establishment of other
+species. Then, after 15 to 20 years, the first recruits declined and
+were replaced by species functionally more similar to the
+pre-disturbance community, which decreased the functional richness and
+evenness \citep{Walker2009}.
 
@@
 Post-disturbance trajectories of tree community composition and
-diversity were both shaped by the recruitment of a determined pool of
-pioneers identical among local communities and independent of the
-disturbance intensity. (i) Composition trajectories in taxonomic and
-functional space nevertheless appeared decoupled. While functional
-trajectories remained similar in the functional space and converged
-towards the recovery of a comparable initial state, taxonomic
-trajectories showed initial differences in community composition that
-were maintained along time. Community high functional redundancy
-mediated this decoupling, as the loss of a species does not necessarily
-entails the loss of its functional charactristics. (ii) Diversity
-trajectories were contrasted as well. While the functional trajectories
-remained similar whatever the disturbance intensity, taxonomic
-trajectories were markedly different with a threshold at 20-25\% AGB
-removed for which taxonomic richness was maximized. The Intermediate
-Disturbance Hypothesis applied well to taxonomic diversity, but not to
-the functional diversity. Taxonomic diversity was maximized at an
-intermediate time after disturbance, that was around 25 years after
-disturbance when the recruitment of early- and late-successional species
-were balanced. Whenever the disturbance intensity, community resilience
-(in terms of recovery of the pre-disturbance state) was tangible but
-required several decades and relied upon the random lottery recruitment
-of rare species. Given the long-term impacts of disturbance observed, we
-suggest that 30 years is not enough time for tropical communities to
-recover, even after relatively low intensity disturbance. Much of
-community response to disturbance rely on species recruitment: refined
-understanding of post-disturbance trajectories would be given by a
-closer analysis of recruitment processes.
+diversity were driven by the recruitment of a determined pool of
+pioneers, identical among local communities and independent of the
+disturbance intensity. The taxonomic composition trajectories maintained
+the initial differences among communities, while the functional
+trajectories were similar and converged in the functoinal space towards
+the recovery of the initial composition. The diversity trajectories were
+contrasted as well. While the functional trajectories remained similar
+whatever the disturbance intensity, taxonomic trajectories were markedly
+different from a threshold of 20-25\% AGB removed that maximized the
+taxonomic richness. The Intermediate Disturbance Hypothesis applied well
+to taxonomic diversity, but not to the functional diversity. The
+decoupling between taxonomic and functional trajectories was mediated by
+the variations in functional redundancy, as the loss of a species does
+not necessarily entails the loss of its functional charactristics.
+Community resilience (in terms of recovery of the pre-disturbance state)
+was tangible but required several decades and relied upon the random
+lottery recruitment of rare species. Given the long-term impacts of
+disturbance observed, we suggest that 30 years is not enough time for
+tropical communities to recover, even after relatively low intensity
+disturbance. Much of community response to disturbance rely on species
+recruitment: refined understanding of post-disturbance trajectories
+would be given by a closer analysis of recruitment processes.
 

### 2019-08-16 WholePlotTrajectories.Rmd
manuscript final

@@
  In the current global change context, it is urgent to anticipate the fate of tropical forests.
- This means highlighting tree community response to disturbance and understand the underlying processes.
- In that respect we aim here to clarify taxonomic and functional post-disturbance  trajectories and determine the scope of the Intermediate Disturbance Hypothesis (IDH) that remains debated in tropical forests.
+ This means understanding tree community response to disturbance and the underlying processes.
+ In that respect, we aim here to clarify taxonomic and functional post-disturbance  trajectories, and determine the scope of the Intermediate Disturbance Hypothesis (IDH) that remains debated in tropical forests.
  We analyzed community trajectories following a disturbance gradient in a Neotropical forest over 30 years. 
- We considered trajectories along time of community taxonomic and functional diversity and composition, and functional redundancy.
-We based on the annual botanical inventories of 75 ha of a Neotropical forest and on large trait datasets comprising 7 leaf, stem and life-history traits.
- We highlighted a decoupling between taxonomic composition that differed among communities, and functional composition, that remained similar and convergent. 
- The taxonomic diversity followed humped-shaped trajectories depending on the disturbance intensity that validated the IDH (Intermediate Disturbance Hypothesis), while functional diversity trajectories homogeneous among all plots dismissed the IDH.
- We explained this decoupling by the variations in community functional redundancy that mitigates the functional impact of disturbance. 
- Although consistent, the recovery of community composition, diversity and redundancy remained unachieved after 30 years.
-This acknowledged the need of decades-long cycles without disturbance to ensure a complete recovery, and questioned tropical forest community resilience after repeated disturbances.
+ We considered trajectories along time of community taxonomic and functional trajectories in terms of richness, evenness, composition, and redundancy.
+ We based on the annual botanical inventories of 75 ha of a Neotropical forest and on large trait datasets comprising seven leaf, stem, and life-history traits. 
+ We identified a decoupling between taxonomic composition, differing among communities, and functional composition, remaining similar and convergent. 
+ The taxonomic diversity followed humped-shaped trajectories depending on the disturbance intensity, which validated the IDH (Intermediate Disturbance Hypothesis).
+ The functional diversity trajectories,however, were homogeneous among all plots and dismissed the IDH.
+ We explained this decoupling by the variations in community functional redundancy that mitigated the functional impact of disturbance. 
+ Although consistent, the recovery of community composition, diversity, and redundancy remained unachieved after 30 years.
+These results acknowledged the need of decades-long cycles without disturbance to ensure community complete recovery, and questioned community resilience after repeated disturbances.
 keywords: [Community Ecology, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
@@
 
-The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
+The large areas covered with tropical forests worldwide hold crucial environmental, economic, and social values. 
 They provide wood and multiple non-timber forest products, shelter a diversified fauna, and ensure cultural and human well-being.
-They regulate the local and regional climates, as well as the carbon, water and nutrient cycles.
+They regulate local and regional climates, as well as carbon, water and nutrient cycles.
 However, the growing demand in forests products together with current global changes increase the pressure on remaining undisturbed forests [@Morales-Hidalgo2015].
-These threats affect the natural disturbance regime that defines and maintains the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
-To anticipate the fate of tropical forests, it is urgent to understand the response to disturbance of tree communities and the underlying ecological processes.
-The forest cover is generally maintained following disturbance, but modifications in the fluxes of light, heat and water [@Goulamoussene2017] change community abiotic and biotic environments.
+These threats affect the natural disturbance regime that defines and maintains the structure, composition, and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
+To anticipate the fate of tropical forests, it is urgent to understand tree community response to disturbance, and the underlying ecological processes.
+The forest cover is generally maintained following disturbance, but modifications in the fluxes of light, heat, and water [@Goulamoussene2017] change community abiotic and biotic environments.
 These changes translate into post-disturbance community trajectories that have been largely studied through trajectories of forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016].
-Recent works have identified some of the determinants of post-disturbance biomass trajectories, like the structure and composition of the pre-disturbance community or the post-disturbance environmental parameters [@Herault2018].
-Community diversity and composition trajectories, however, have not been as thoroughly understood [@Guitet2018;@Molino2001], and manifold biodiversity trajectories might emerge given the variety of species response to disturbance and the huge diversity of tropical forests [@Lindenmayer2012;@Garcia_florez2017].
+Some of the determinants of post-disturbance biomass trajectories are already identified, like the structure and composition of the pre-disturbance community, or the post-disturbance environmental parameters [@Herault2018].
+Community diversity and composition trajectories, however, have not been as thoroughly understood [@Guitet2018;@Molino2001], and manifold biodiversity trajectories might emerge given the variety of species response to disturbance and the diversity of tropical forests [@Lindenmayer2012;@Garcia_florez2017].
 
 An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
-The IDH states a relationship between the community diversity and the intensity and frequency of disturbance events, and a diversity peak at intermediate level of disturbance [@Connell1978].
-This bases on the fluctuations of community environment following disturbance that foster both competitively superior species and fast colonizers, and prevents competitive exclusion [@Shea2004;@Pulsford2016].
-In tropical forests, however, observations of the IDH often diverge from theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the underlying processes might be complicated by the huge diversity of tropical tree communities [@Lindenmayer2012;@Garcia_florez2017].
-In this context, it is urgent to validate the IDH that remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
+The IDH states a relationship between the community diversity and the intensity and frequency of disturbance events, and postulates a diversity peak at intermediate level of disturbance [@Connell1978].
+This is based on the fluctuations of community environment following disturbance that foster both competitively superior species and fast colonizers, and prevents competitive exclusion [@Shea2004;@Pulsford2016].
+In tropical forests, however, observations of the IDH often diverge from theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the underlying processes might be complicated by the diversity of tropical tree communities [@Lindenmayer2012;@Garcia_florez2017].
+In this context, the IDH is controversial in tropical forests and remains to be tested [@Hubbell2001;@Fox2013;@Sheil2013].
 
-Analysing community response to disturbance and grasp all aspects of community changes requires a whole set of metrics [@Sheil2003;@Shea2004;@Mayfield2010].
-The analysis should first consider community composition, which is crucial for conservation issues and reveals the pool of species fostered (or hampered) by disturbance [@Lavorel2002;@Bellwood2006].
-This should be completed with diversity metrics encompassing both richness and evenness to assess the changes in community abundance distribution. 
-Besides, functional approaches have been shown to usefully complement pure taxonomic approaches as they shed light on the species biological attributes directly linking community diversity, composition and redundancy to ecosystem functioning [@Violle2007b;@Baraloto2012a].
+Analysing community response to disturbance and grasp all aspects of community changes requires a wide array of metrics [@Sheil2003;@Shea2004;@Mayfield2010].
+The analysis should first consider community composition, which is crucial for conservation issues and reveals the pool of species fostered or hampered by disturbance [@Lavorel2002;@Bellwood2006].
+This should be completed with diversity metrics encompassing both community richness and evenness to assess the changes in community abundance distribution. 
+Besides, functional approaches have been shown to usefully complement pure taxonomic approaches as they shed light on the species biological attributes directly linking community diversity, composition, and redundancy to ecosystem functioning [@Violle2007b;@Baraloto2012a].
 In that respect, a vast literature allowed recognizing major traits representing species ecological strategy and determining how they respond to changing conditions [@Diaz2005]. 
-Specifically in tropical forests, the functional approach revealed the deterministic processes entailing after disturbance a shift from a dominance of "conservative" slow-growing species dealing with scarce resources, to a dominance of "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Rees2001;@Reich2014;@Herault2011].
-This shift is translated into the trajectories of average community value of key functional traits related to resource acquisition (leaf and stem traits) and life-history strategy (seed mass, maximum size) [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
+Specifically, in tropical forests, the functional approach revealed the emergence of deterministic processes following disturbance. 
+Such deterministic processes entailed a shift from a dominance of "conservative" slow-growing species dealing with scarce resources, to a dominance of "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@Rees2001;@Reich2014;@Herault2011].
+This shift is translated into the trajectories of average community value of key functional traits related to resource acquisition, as leaf and stem traits, and life-history strategy, as seed mass and maximum size [@Wright2004;@TerSteege2006;@Westoby2006a;@Chave2009b].
 
 The functional approach also encompasses the analysis of functional redundancy, that quantifies the amount of shared trait values among species [@Carmona2016]. 
-The typical high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning and determines communities resilience after disturbance [@Elmqvist2003; @Diaz2005].
+The typical high functional redundancy of hyper-diverse tropical forests [@Bellwood2006] mitigates the impacts of species removal on ecosystem functioning, and determines community resilience after disturbance [@Elmqvist2003; @Diaz2005].
 
-In this study, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60\% of above-ground biomass (AGB) loss.
-We made use of a large functional traits database encompassing major leaf, stem and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition and redundancy.
+Here, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60\% of above-ground biomass (AGB) loss.
+We made use of a large functional traits database encompassing major leaf, stem, and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition, and redundancy.
 Specifically, (i) we drew taxonomic and functional post-disturbance trajectories and examined the underlying ecological processes, (ii) we discussed the scope of the IDH regarding taxonomic and functional facets of community diversity, and (iii) we analyzed community resilience and time to recovery.
@@
 
-Paracou station in French Guiana (5° 18'N and 52° 53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26° C, mean annual precipitation averaging 2980 mm.y^-1^ (30-y period) and a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
-Elevation ranges from 5 to 50 m.
-Across all plots the topography mainly corresponds to hilltops or hillsides, while bottomlands cover less than 1 % of the area. 
+Paracou station in French Guiana (5° 18'N and 52° 53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26° C, and mean annual precipitation averaging 2980 mm.y^-1^ (30-y period).
+The climate comprises a 3-month dry season (< 100 mm.month^-1^) from mid-August to mid-November, and a one-month dry season in March [@Wagner2011].
+Across all plots, elevation ranges from 5 to 50 m, and the topography mainly corresponds to hilltops or hillsides, while bottomlands cover less than 1 % of the area. 
 Plots are shallow ferralitic acrisols over a layer of transformed saprolite with low permeability and lateral drainage.
@@
 The experiment is a network of twelve 6.25 ha plots (Table \@ref(tab:Tab1)) that underwent three disturbance treatments in 1987 according to a randomized plot design [@Gourlet-Fleury2004].  
-The experiment comprised three replicates of three sylvicultural treatments (hereafter plots T1, T2 and T3) and three control plots (T0).
-All treatments T1, T2 and T3 comprised the logging of 10 trees/ha with 50 cm minimum DBH that belonged to a set of 58 commercial species [@Gourlet-Fleury2004]. 
-Treatment T2 additionally comprised a thinning treatment by poison-girdling of non-commercial, randomly selected species with an average 30 trees/ha with 40 cm minimum DBH.
-Treatment T3 additionally comprised the logging of 15 trees/ha with 40 cm minimum DBH and the poison-gidling of 20 trees/ha with a 50 cm minimum DBH, all belonging to non-commercial species. 
+The experiment comprised three replicates of three sylvicultural treatments (hereafter plots T1, T2, and T3), and three control plots (T0).
+All treatments T1, T2, and T3 comprised the logging of 10 trees/ha with 50 cm minimum DBH that belonged to a set of 58 commercially exploited species [@Gourlet-Fleury2004]. 
+Treatment T2 additionally comprised a thinning treatment by poison-girdling of non-commercially exploited species randomly selected with an average of 30 trees/ha with 40 cm minimum DBH.
+Treatment T3 additionally comprised the logging of 15 trees/ha with 40 cm minimum DBH, and the poison-girdling of 20 trees/ha with a 50 cm minimum DBH, all belonging to non-commercially exploited species.
 Considering the silvicultural treatments and the following damage, disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Rejou2017].
-The three treatments were then transformed into a continuous disturbance intensity gradient with increasing of above-ground biomass (AGB) lost.
+The three treatments were then transformed into a continuous disturbance intensity gradient with increasing above-ground biomass (AGB) loss.
 
@@
   Timber=c("-", 
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10   $trees.ha^{-1}$", 
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10  $trees.ha^{-1}$",
-           "DBH $\\geq$ 50 cm, commercial species, $\\approx$ 10  $trees.ha^{-1}$"),
+           "DBH $\\geq$ 50 cm, commercially exploited species, $\\approx$ 10   $trees.ha^{-1}$", 
+           "DBH $\\geq$ 50 cm, commercially exploited species, $\\approx$ 10  $trees.ha^{-1}$",
+           "DBH $\\geq$ 50 cm, commercially exploited species, $\\approx$ 10  $trees.ha^{-1}$"),
   Thinning=c("-", 
              "-", 
-             "DBH $\\geq$ 40 cm, non-valuable species, $\\approx$ 30   $trees.ha^{-1}$", 
-             "DBH $\\geq$ 50 cm, non-valuable species, $\\approx$ 15  $trees.ha^{-1}$  "),
+             "DBH $\\geq$ 40 cm, non-commercially exploited species, $\\approx$ 30   $trees.ha^{-1}$", 
+             "DBH $\\geq$ 50 cm, non-commercially exploited species, $\\approx$ 15  $trees.ha^{-1}$  "),
   Fuelwood=c("-", 
@@
              "-", 
-             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-valuable species,\\ $\\approx$ 15 $trees.ha^{-1}$"),
+             "40 cm $\\leq$ DBH $\\leq$ 50 cm, non-commercially exploited species,\\ $\\approx$ 15 $trees.ha^{-1}$"),
   AGBLost=c(0, "$[12-33]$", "$[33-56]$", "$[35-56]$")
@@
 knitr::kable(Intervention, caption="Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
-Treatment intensities are defined by the minimum logging DBH (Diameter at Breast Height), the type of logged species (commercial or not), the density of logged trees, and the total AGB (Above Ground Biomass) lost after treatment.", longtable = FALSE, booktabs = TRUE, escape = FALSE,format="latex") %>%
+Treatment intensities are defined by the minimum logging DBH (Diameter at Breast Height), the type of logged species (commercially exploited or not), the density of logged trees, and the total AGB (Above Ground Biomass) lost after treatment.", longtable = FALSE, booktabs = TRUE, escape = FALSE,format="latex") %>%
   kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
@@
 
-The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of *Fabaceae*, *Chrysobalanaceae*, *Lecythidaceae* and *Sapotaceae*. 
+The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of *Fabaceae*, *Chrysobalanaceae*, *Lecythidaceae*, and *Sapotaceae*. 
 In the 12 experimental plots, all trees above 10 cm DBH have been mapped and measured annually since 1984. 
-Trees are first identified with a vernacular name assigned by the forest worker team, and afterward with a
-scientific name assigned by botanists during regular botanical campaigns. 
-In 1984, specific vernacular names were given to 62 commercial or common species whereas more infrequent ones were identified under general identifiers only distinguishing trees and palms. 
-From 2003, botanical campaigns have been conducted every 5 to 6 years to identify all trees at the species level but identification levels still varied among plots and campaigns.
+Trees are first identified with a vernacular name assigned by the forest worker team, and afterward with a scientific name assigned by botanists during regular botanical campaigns. 
+In 1984, specific vernacular names were given to 62 commmon or commercially exploited species.
+More infrequent species were identified under general identifiers only distinguishing trees and palms. 
+From 2003, botanical campaigns have been conducted every 5 to 6 years to identify all trees at the species level.
+In 2015, however, identification levels still varied among plots and campaigns.
 
@@
 This resulted in significant taxonomic uncertainty that had to be accounted for in the measure of composition and diversity metrics. 
-The uncertainty propagation based on a Bayesian framework using vernacular/botanical names associations to reconstitute complete inventories at genus level from real incomplete ones. 
+Uncertainty propagation was implented withing a Bayesian framework using vernacular/botanical names associations to reconstitute complete inventories at genus level from real incomplete ones. 
 Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,..., \alpha_V\big]$ observed across all inventories between each vernacular name *v* and all species $\big[s_1, s_2,..., s_N\big]$:
@@
 
-Six functional traits representing leaf economics (leaf thickness, toughness, total chlorophyll content and specific leaf area) and stem economics (wood specific gravity and bark thickness) came from the BRIDGE project [^1].
-Trait values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 species pertaining to 157 genera. 
+Six functional traits representing the leaf economics (*i.e.* leaf thickness, toughness, total chlorophyll content, and specific leaf area), and the stem economics (*i.e.* wood specific gravity and bark thickness) were obtained from the BRIDGE project [^1].
+Trait values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 species belonging to 157 genera. 
 Whenever a species was in the dataset but missed some trait values (10% of the species), missing values were filled using multivariate imputation by chained equation [@Mice2011].
-To account for the phylogenetic signal in the filling process, imputations based on samples of species from the same genus or from the same family. 
+To account for the phylogenetic signal in the filling process, imputations were based on samples of species from the same genus or from the same family. 
 Whenever a species was not in the dataset, it was attributed a set of trait values randomly sampled among species of the next higher taxonomic level (same genus or family). 
-Two life-history traits (maximum specific height and seed mass) came from the Mariwenn database [^2]. 
-The database compiles information from a vast literature on the flora of French Guiana [@Ollivier2007] and comprises 362 species pertaining to 188 genera.
+Two life-history traits, maximum specific height and seed mass, came from the Mariwenn database [^2]. 
+The database compiles information from a vast literature on the flora of French Guiana [@Ollivier2007] and comprises 362 species belonging to 188 genera.
 As seed mass information was classified into classes, no data filling process was applied and analyses were restricted to the botanical species recorded.
@@
 
-Because of the variability in the precision of botanical identification efforts, we were constrained to conduct the taxonomic composition and diversity analysis at the genus level. 
-Taxonomic and functional trajectories of community composition were followed in a two-dimensional NMDS ordination plane.
-Two NMDS using abundance-based (Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the latter based on the seven leaf, stem and life history traits (without seed mass classes).
-Trajectories along time were reported through the Euclidean distance between the target inventories and the reference pre-disturbance inventories of 1984.
-Univariate trajectories of the leaf, stem and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007]. 
+Because of the variable precision of botanical identification efforts, we had to conduct the taxonomic composition and diversity analysis at the genus level. 
+Taxonomic and functional trajectories of community composition were drawn in a two-dimensional NMDS ordination plane.
+Two NMDS using abundance-based (*i.e.* Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the latter based on the seven leaf, stem, and life history traits (without seed mass classes).
+Trajectories along time were reported through the Euclidean distance between the target inventories and the 1984 pre-disturbance inventories of reference.
+Univariate trajectories of the leaf, stem, and life-history traits were also visualized with the community weighted means (CWM) [@Diaz2007]. 
 Species seed mass were given in 5 mass classes, and seed mass trajectories were reported as the proportion of each class in the inventories (Supplementary materials).
 
-The taxonomic and functional trajectories were analysed from the reference pre-disturbance inventories of 1984.
+The taxonomic and functional trajectories were analysed from the 1984 pre-disturbance inventories of reference.
 The taxonomic diversity was reported through species richness and the Hill number translation of the Simpson index [@Hill1973].
-These metrics allowed assessing the taxonomic richness as well as evenness, through the comparison between these two metrics: results will be discussed directly in terms of taxonomic richness and evenness.
-Both indices belong to the set of HCDT or generalized entropy, respectively corresponding to the 0 and 2 order of diversity (q), recommended for diversity studies [@Marcon2015]. 
+The comparison between these two metrics assesses community taxonomic richness and evenness: thereafter, results will be discussed directly in terms of taxonomic richness and evenness.
+Both indices are recommended for diversity studies [@Marcon2015], and belong to the set of HCDT or generalized entropy corresponding, respectively, to the 0 and 2 order of diversity (q). 
 The functional diversity was reported using the functional richness and functional evenness, through the Rao index of quadratic entropy.
@@
 Richness and evenness trajectories were besides analyzed through polynomial regression between (i) taxonomic and
-functional richness and evenness and (ii) the initial \%AGB loss at 10, 20 and 30 years after disturbance.
+functional richness and evenness, and (ii) the initial \%AGB loss at 10, 20, and 30 years after disturbance.
 
 Finally, functional redundancy was measured as the overlap among species in community functional space [@Carmona2016]. 
-First, the individuals of the trait database were mapped in the plane of the first two axes from a PCA analysis, which lowered the weight of correlations among traits as the PCA axes were combinations of most decoupled traits.
+First, the individuals of the trait database were mapped in the plane of the first two axes from a PCA analysis.
+The PCA analysis lowered the weight of correlations among traits as the axes are combinations of the most decoupled traits. 
 For each species, the traits probability density (TPD) were computed from the mapping of individuals through two-dimension kernel density estimators. 
 Second, for each community, the TPD weigthed by species
-abundance were summed accross the functional space. Third, the functional space was divided into a 100 x 100 grid and the number of species with a positive TPD was counted in each cell. 
-The average count across cells minus 1 returned the Community Functional Redundancy, which was the average number of species in the community that share the same
-trait values.
+abundance were summed accross the functional space. Third, the functional space was divided into a 100 x 100 grid, and the number of species with a positive TPD was counted in each cell. 
+The average count across cells minus 1 returned the Community Functional Redundancy, which was the average number of species in the community that share the same trait values.
 
@@
 
-From 1984, the first pre-disturbance inventory, to 2015 (28 years after disturbance), 828 388 individual trees and 591 botanical species pertaining to 223 genera and 64 families were recorded.
+From 1984, the first pre-disturbance inventory, to 2015, 28 years after disturbance, 828 388 individual trees and 591 botanical species spanning 223 genera and 64 families were recorded.
 
-While both taxonomic and functional composition remained stable in undisturbed communities (Fig. \@ref(fig:NMDSplans)), they followed marked and consistent trajectories in disturbed plots over time.
+In undisturbed plots, both taxonomic and functional composition remained stable (Fig. \@ref(fig:NMDSplans)).
+In disturbed plots, both trajectories followed marked and consistent trajectories over time.
 
-The functional composition trajectory in disturbed communities resembled to cyclic compositional changes with an incomplete recovery to the initial composition (Fig.\ref{fig:NMDSplans}). 
-The maximum dissimilarity with the initial state was positively correlated with the disturbance intensity for both taxonomic and functional composition ($\rho_{Spearman}^{Taxonomic}=0.87$ and $\rho_{Spearman}^{Functional}=0.90$ respectively).
-The maximum dissimilarity with the initial state was reached for taxonomic composition between 15 to 25 years and around 20 to 25 years for functional composition.
+The functional composition trajectory resembled, in disturbed communities, to cyclic compositional changes with an incomplete recovery of the initial composition (Fig.\ref{fig:NMDSplans}). 
+The maximum dissimilarity with the initial state was positively correlated with the disturbance intensity for both taxonomic and functional composition ($\rho_{Spearman}^{Taxonomic}=0.87$ and $\rho_{Spearman}^{Functional}=0.90$, respectively).
+The maximum dissimilarity with the initial state was reached for taxonomic composition between 15 to 25 years, and between 20 to 25 years for functional composition.
 
-```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of taxonomic composition ((ref:Bolda) and (ref:Boldc)) and functional composition ((ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS plane.
+```{r NMDSplans, out.width = '100%', echo=FALSE, fig.cap="Plot trajectories in terms of taxonomic composition ((ref:Bolda) and (ref:Boldc)), and functional composition ((ref:Boldb) and (ref:Boldd)) in a two-dimensional NMDS plane.
 Lower panels ((ref:Boldc) and (ref:Boldd)) represent the Euclidean distance to initial condition along the 30 sampled years.
@@
 
-In control plots, Community Weighted Means (CWM) of functional traits remain stable in time.  
+In control plots, Community Weighted Means (CWM) of functional traits remained stable in time.  
 In disturbed plots, they mostly followed unimodal trajectories, either stabilizing or returning towards
-their initial values, to the exception of leaf chlorophyll content, which continued to increase 30 years after disturbance for some highly disturbed plots.
-Maximum height at adult stage (*Hmax*), leaf toughness and wood specific gravity (*WSG*) decreased in time and then slightly increased, but remained significantly lower than their initial value (Fig.
+their initial values, to the exception of leaf chlorophyll content, which continued to increase 30 years after disturbance for 4 out of 6 highly disturbed plots.
+Maximum height at adult stage (*Hmax*), leaf toughness, and wood specific gravity (*WSG*) decreased in time and then slightly increased, but remained significantly lower than their initial value (Fig.
 \ref{fig:CWM}).
-Bark thickness and specific leaf area (*SLA*) both increased in time. 
+Bark thickness and specific leaf area (*SLA*) both increased along time. 
 Bark thickness remained substantially high after 30 years, and *SLA* had almost recovered to its initial value. Whatever the functional traits, the maximum
 difference to initial value was highly correlated to the disturbance intensity.
-Positive correlations were observed for Leaf thickness, chlorophyll content, SLA and bark thickness ($\rho_{Spearman}^{Leaf thickness}=0.76$, $\rho_{Spearman}^{Chlorophyll content}=0.60$,
+Positive correlations were observed for leaf thickness, chlorophyll content, SLA and bark thickness ($\rho_{Spearman}^{Leaf thickness}=0.76$, $\rho_{Spearman}^{Chlorophyll content}=0.60$,
 $\rho_{Spearman}^{SLA}=0.93$, $\rho_{Spearman}^{Bark thickness}=0.71$). 
-Negative correlation was observed for Leaf toughness, WSG and Hmax ($\rho_{Spearman}^{Leaf toughness}=-0.53$,
+Negative correlation was observed for leaf toughness, WSG, and Hmax ($\rho_{Spearman}^{Leaf toughness}=-0.53$,
 $\rho_{Spearman}^{WSG}=-0.75$, $\rho_{Spearman}^{Hmax}=-0.40$)
@@
 
-```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of community weighted means over 30 years after disturbance of four leaf traits (Leaf thickness, chlorophyll content, toughness, and specific area), two stem traits (wood specific gravity, and bark thickness) and one life history trait (Specific maximum height at adult stage).
+```{r CWM, out.width = '100%', echo=FALSE, fig.cap="Trajectories of community weighted means over 30 years after disturbance of four leaf traits (leaf thickness, chlorophyll content, toughness, and specific area), two stem traits (wood specific gravity and bark thickness), and one life history trait (species maximum height at adult stage).
 ", fig.env="figure*"}
@@
 This maximum was positively correlated with the disturbance intensity ($\rho_{Spearman}^{Simpson}=0.77$).
-The evenness then stabilized except for two intensively-disturbed plots (number 8 and 12) for which it kept increasing \@ref(fig:DivTaxo)).
+The evenness then stabilized except for two intensively-disturbed plots (number 8 and 12) for which it kept increasing (Fig. \@ref(fig:DivTaxo)).
 
 ```{r DivTaxo, out.width = '100%', echo=FALSE, fig.cap="Trajectories of community taxonomic richness (ref:Bolda), Simpson diversity (ref:Boldb), functional richness (ref:Boldc), and Rao diversity (ref:Boldd).
-Values correspond to the difference over 30 years of community diversity with the values of reference 1984 pre-disturbance inventories.
+Values correspond to the difference over 30 years of community diversity with the values of 1984 pre-disturbance inventories of reference.
 Shaded areas are the credibility intervals ", fig.env="figure*"}
@@
 Functional richness and evenness remained stable in control plots over the 30 years of monitoring. 
-In disturbed communities, both trajectories depended on the disturbance intensity, with their maximum values in time being positively correlated to \%AGB loss $\rho_{Spearman}^{Richness}=0.76$ and $\rho_{Spearman}^{Rao}=0.60$.
+In disturbed communities, both trajectories depended on the disturbance intensity, with their maximum values in time being positively correlated to \%AGB loss $\rho_{Spearman}^{Richness} = 0.76$ and
+$\rho_{Spearman}^{Rao} = 0.60$.
 For low disturbance intensity, functional richness and evenness displayed a low but long-lasting increase up to a maximum reached after 20-25 years. 
-For high disturbance intensity, they generally displayed a fast but short increase followed after 10 years by a slow decrease towards the initial values.
+For high disturbance intensity, they generally displayed a fast but short increase followed, after 10 years, by a slow decrease towards the initial values.
 
-The second-degree polynomial regressions between (i) the percentage AGB loss and (ii) the taxonomic and functional diversity showed various shape depending on the diversity indices and on the time since disturbance  (Fig.
+The second-degree polynomial regressions between (i) the percentage AGB loss, and (ii) the taxonomic and functional diversity showed various shapes depending on the diversity indices and on the time since disturbance  (Fig.
 \ref{fig:IDHplot}).
-Regarding taxonomic diversity, the relationship between disturbance intensity and diversity was more markedly hump-shaped for richness than for evenness and peaked at 20\% of initial AGB loss.
-Regarding functional diversity, the relationship was almost linear and similar between richness and evenness.
-Generally, all relationships were stronger 20 or 30 years after disturbance than observed for 10 years.
+Regarding taxonomic diversity, the relationship between disturbance intensity and diversity was more markedly hump-shaped for richness than for evenness, and peaked at 20\% of initial AGB loss.
+Regarding functional diversity, the relationship was almost linear, and was similar between richness and evenness. 
+All the relationships between disturbance intensity and diversity were stronger 20 or 30 years after disturbance than they were 10 years after disturbance.
 
-```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/4,fig.cap="Relationship between the initial \\%AGB loss and community taxonomic richness (ref:Bolda), taxonomic evenness (ref:Boldb), functional richness (ref:Boldc),and functional evenness (ref:Boldd) at 10, 20 and 30 years after disturbance", fig.env="figure*"}
+```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/4,fig.cap="Relationship between the initial \\%AGB loss and community taxonomic richness (ref:Bolda), taxonomic evenness (ref:Boldb), functional richness (ref:Boldc), and functional evenness (ref:Boldd) at 10, 20, and 30 years after disturbance.", fig.env="figure*"}
 load("DB/RichnessIDH");load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB");load("DB/FunRichIDH")
@@
 Control plots displayed stable functional redundancy over time. 
-All disturbed plots had lower functional redundancy than control plots and followed similar hump-shaped trajectories (\ref{fig:RedFunRest}).
+All disturbed plots had lower functional redundancy than control plots and followed similar hump-shaped trajectories (Fig. \ref{fig:RedFunRest}).
 The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{Spearman}=0.47$) and the recovery trajectory had not attained initial values for any disturbed communities after 30 years.
@@
 
-Our analysis highlighted the decoupling between functional and taxonomic trajectories.
-While the initial differences in taxonomic composition among plots were maintained, the functionalcomposition trajectories converged in the functional space.
+Our analysis revealed the decoupling between functional and taxonomic trajectories.
+While the initial differences in taxonomic composition among plots were maintained, the functional composition trajectories converged in the functional space.
 In terms of diversity, only the taxonomic trajectories validated the IDH that explained humped-shaped post-disturbance trajectories whith an amplitude depending on the disturbance intensity.
@@
 Community taxonomic composition substantially differed before disturbance, as materialized by their distinct location on the NMDS axis 2. 
-These initial differences were maintained all along the 30 years following disturbance, with the disturbance leading a displacement on the NMDS axis 1 only. 
-Taxonomic composition changes were similar among plot and may correspond to the recruitment of a group of pioneers, e.g.
-*Cecropia* spp., *Vismia* spp.,  shared by all plots whatever their initial taxonomic differences and the intensity of disturbance [@Denslow2000;@Bongers2009].
+These initial differences were maintained along the 30 years following disturbance, with the disturbance leading a displacement on the NMDS axis 1 only. 
+Taxonomic composition changes were similar among plots and may correspond to the recruitment of a group of pioneers, like *Cecropia spp.* or *Miconia spp.*,  common to all plots, whatever their initial taxonomic differences and the intensity of disturbance [@Denslow2000;@Bongers2009].
 Taxonomic trajectories initiated a recovery towards the initial composition.
 This recovery, although far from being achieved after 30 years, suggested the resilience of community taxonomic composition and the maintenance of initial composition differences [@Folke2006]. 
-This suggested that species not belonging to the pre-disturbance community were rarely recruited in the long-term, probably because of species dispersal limitation that is common among tropical species [@Svenning2005].
+Such resilience suggested that species not belonging to the pre-disturbance community were rarely recruited in the long-term, probably because of species dispersal limitation that is common among tropical species [@Svenning2005].
 
@@
 The amplitude of the compositional changes depended on the disturbance intensity.
-Functional trajectories were probably driven by the recruitment of species infrequent or absent before disturbance and belonging to a pool common to all plots. 
-This common pool was composed of pioneers with resource-acquisitive strategies (low leaf toughness, wood specific gravity, maximum height and high Specific Leaf area) which recruitment drove a displacement to the right along the first NMDS axis [@Westoby1998;@Wright2004;@Reich2014].
-Thereafter, the first recruited pioneers were progressively excluded by long-lived, more competitive and shade-tolerant species, which drove the recovery of the initial functional composition. 
-This recovery corresponded to a return of the composition towards resource-conservative strategies that translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig.
-\ref{fig:NMDSplans}).
+Functional trajectories were probably driven by the recruitment of species infrequent or absent before disturbance, and belonging to a pool of species common to all plots. 
+This common pool was composed of pioneers with "resource-acquisitive"" strategies displaying low leaf toughness, wood specific gravity, maximum height, and high specific leaf area.
+The recruitment of these pioneers drove a displacement to the right along the first NMDS axis [@Westoby1998;@Wright2004;@Reich2014]. 
+Thereafter, the first recruited pioneers were progressively excluded by long-lived, more competitive, and shade-tolerant species.
+The recruitment of these late-successionals marked the recovery of the initial functional composition with more "resource-conservative"" strategies .
+This recovery translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig. \ref{fig:NMDSplans}).
 
-The decoupling between taxonomic and functional trajectories suggested the simultaneous operation of trait-based assembly rules and species-level priority effects to shape tree community assembly in this forest.
-Tree community assembly would then be both deterministic in the functional space and historically contingent in the taxonomic space.
+The decoupling between taxonomic and functional trajectories suggested that simultaneous operation of trait-based assembly rules and species-level priority effects shaped tree community assembly in Paracou forest.
+Tree community assembly would then be both deterministic in the functional space, and historically contingent in the taxonomic space.
 
@@
 
-Trajectories of taxonomic richness and evenness differed markedly below and above an intensity threshold (\ref{fig:DivTaxo}).
-For low and intermediate disturbances, both taxonomic richness and evenness increased according to the disturbance intensity [@Martin2015;@Chaudhary2016].
-This suggested that the recruitment of pioneers previously infrequent or absent increased the taxonomic richness [@Martin2015;@Chaudhary2016], and that trees surviving after disturbance remained numerous enough to maintain the richness of the pre-disturbance community [@Bongers2009].
+Trajectories of taxonomic richness and evenness differed markedly below and above an intensity threshold (Fig. \ref{fig:DivTaxo}).
+For low and intermediate disturbance, both taxonomic richness and evenness increased according to the disturbance intensity [@Martin2015;@Chaudhary2016].
+This suggested that the recruitment of pioneers, previously infrequent or absent, increased the taxonomic richness [@Martin2015;@Chaudhary2016], and that trees surviving after disturbance remained numerous enough to maintain the richness of the pre-disturbance community [@Bongers2009].
 The pioneers thus recruited became abundant, which balanced the usual hyper-dominance of tropical forests and increased the taxonomic evenness [@Baraloto2012a].
 Above the intensity threshold, like for treatment 3 (high intensity), the taxonomic richness did not exceed the initial value in the first years following disturbance.
-This suggested that the richness of surviving trees was lower than this of the pre-disturbance community, and that this decrease was not offset by the recruitment of pioneers. 
-In the Guiana shield indeed, the pool of true pioneers specifically recruited after disturbance is restricted to a few common genera (e.g. *Cecropia* spp., *Vismia* spp.) [@Guitet2018].
-Considering different time after disturbance, there was always a humped-shaped pattern linking the disturbance intensity with the post-disturbance increase in taxonomic richness and evenness (\ref{fig:IDHplot}).
+No increase of the taxonomic richness suggested that the richness of surviving trees was lower than this of the pre-disturbance community, and that this difference was not offset by the recruitment of pioneers.
+In the Guiana shield, the pool of true pioneers specifically recruited after disturbance is restricted to a few common genera (e.g. *Cecropia spp.*, *Miconia spp.*, *Tapirira spp.*) [@Guitet2018].
+Considering the different times after disturbance, there was always a humped-shaped pattern linking the disturbance intensity with the post-disturbance increase in taxonomic richness and evenness (Fig.
+\ref{fig:IDHplot}).
 Both taxonomic richness and evenness were maximized at an intermediate intensity, around 20-25% of AGB lost.
 
-Regarding community functional trajectories, however, no marked differences were observed among post-disturbance trajectories (\ref{fig:DivTaxo}).
-Whenever the disturbance intensity, the recruitment of pioneers functionally highly different from the pre-disturbance community, first increased both community functional richness and evenness [@Denslow1980;@Molino2001].
-These recruitment hampered the establishment of other species in the first place.
-After 15 to 20 years, the first recruits declined and were replaced by species functionally more similar to the pre-disturbance community, which decreased the functional richness and evenness [@Walker2009].
+Regarding community functional trajectories, however, no marked differences were observed among post-disturbance trajectories (Fig. \ref{fig:DivTaxo}).
+Whenever the disturbance intensity, there was first an increase of both functional richness and evenness.
+Such increase suggested the recruitment of pioneers that were functionally highly different from the pre-disturbance community [@Denslow1980;@Molino2001].
+The recruitment of pioneers hampered the recruitment of other species in the first place.
+After 15 to 20 years, the first pioneer recruits declined and were replaced by species functionally more similar to the pre-disturbance community, which decreased the functional richness and evenness [@Walker2009].
 
@@
 
-The loss of species following disturbance decreased the functional redundancy during the first 15 years. 
-Progressively though, the functional redundancy was restored through the replacement of the resource-acquisitive strategies species by more late-successional species, functionally closer to the pre-disturbance community.
-This replacement followed the lottery recruitment rules, implying an easy recruitment for the first species but becoming increasingly hampered by the emergence of interspecific competition [@Busing2002].
-This explained why the recovery trajectory slowed down 20 years after disturbance. 
-The recovery of the functional redundancy then relied upon the random process of species recruitment and was increasingly slow and difficult to anticipate [@Elmqvist2003;@Diaz2005]. 
-This suggested a low resilience of the functional redundancy with the random recovery of infrequent species increasing the risks to lose keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
-Infrequent species might indeed have unique functional characteristics, apart from those considered here, in the ecosystem or be a key resource for some fauna [@Schleuning2016].
+The loss of species following disturbance decreased community functional redundancy during the first 15 years. 
+Progressively though, the functional redundancy was restored through the replacement of the species with "resource-acquisitive"" strategies by more late-successional species, functionally closer to the pre-disturbance community.
+The replacement of pioneers by late-successionals followed the lottery recruitment rules.
+This rules mean an easy recruitment of the first species that becomes increasingly difficult, as the following species are hampered by the emergence of interspecific competition [@Busing2002].
+The increasing competition among species explained why the recovery trajectory slowed down 20 years after disturbance. 
+The recovery of the functional redundancy then relied upon the random process of species recruitment, and was increasingly slow and difficult to anticipate [@Elmqvist2003;@Diaz2005]. 
+This suggested a low resilience of the functional redundancy with the random recovery of infrequent species increasing the risks of losing keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
+Infrequent species might indeed have unique functional characteristics in the ecosystem, apart from the ones considered here, in the ecosystem or be a key resource for some of the fauna [@Schleuning2016].
 
@@
 
-Post-disturbance trajectories of tree community composition and diversity were driven by the recruitment of a determined pool of pioneers, identical among local communities and independent of the disturbance intensity.
-The taxonomic composition trajectories maintained the initial differences among communities, while the functional trajectories were similar and converged in the functional space towards the recovery of the initial composition.
+Post-disturbance trajectories of tree community composition and diversity were driven by the recruitment of a determined pool of pioneers, identical among local communities, and independent of the disturbance intensity.
+The taxonomic composition trajectories maintained the initial differences among communities, while the functional trajectories were similar, and converged in the functional space towards the recovery of the initial composition.
 The diversity trajectories were contrasted as well. 
-While the functional trajectories remained similar whatever the disturbance intensity, taxonomic trajectories were markedly different from a threshold of 20-25\% AGB removed that maximized the taxonomic richness.
-The Intermediate Disturbance Hypothesis applied well to taxonomic diversity, but not to the functional diversity.
+While the functional trajectories remained similar whatever the disturbance intensity, taxonomic trajectories were markedly different from a threshold of 20-25\% AGB lost that maximized the taxonomic richness.
+The Intermediate Disturbance Hypothesis applied well to taxonomic diversity, but not to functional diversity.
 The decoupling between taxonomic and functional trajectories was mediated by the variations in functional redundancy, as the loss of a species does not necessarily entails the loss of its functional characteristics.
-Community resilience (in terms of recovery of the pre-disturbance state) was tangible but required several decades and relied upon the random lottery recruitment of rare species.
+Community resilience, in terms of recovery of the pre-disturbance state, was tangible but required several decades, and relied upon the random lottery recruitment of rare species.
 Given the long-term impacts of disturbance observed, we suggest that 30 years is not enough time for tropical communities to recover, even after relatively low intensity disturbance.
-Much of community response to disturbance rely on species recruitment: refined understanding of post-disturbance trajectories would be given by a closer analysis of recruitment processes.
+Much of community response to disturbance rely on the processes of species recruitment.
+A more refined understanding of the post-disturbance trajectories would be gained by a closer analysis of the recruitment process.
 
@@
 
-We are in debt with all technicians and colleagues who helped setting up the plots and collecting data over years. 
+We are in debt with all Paracou station technicians and colleagues who helped setting up the plots and collecting data over years. 
 Without their precious work, this study would have not been possible and they may be warmly thanked here.
+Our work benefited from an “Investissement d’Avenir” grant managed by the Agence
+Nationale de la Recherche (LABEX CEBA, ref ANR-10-LBX-25).
+We thank Niklas Tysklind for the usefull help with english proofing.
 
@@
 
-AM, EM & BH designed the study, developed the analysis framework and interpreted the results.
+AM, EM & BH designed the study, developed the analysis framework, and interpreted the results.
 AM wrote the manuscript with contributions by EM & BH.

### 2019-08-16 WholePlotTrajectories.tex
manuscript final

@@
 \Abstract{
-Anticipate the impact of the global change context on tropical forests
-is now urgent. This calls to highlight the taxonomic and functional
-facets of tree community post-disturbance trajectories and the
-underlying processes. It also calls to clarify the link between
-biodiversity and disturbance, specifically the scope of the Intermediate
-Disturbance Hypothesis (IDH) that is debated in tropical forests. In
-this study, we analyzed the tree community trajectories following a
-disturbance gradient in a Neotropical forest over 30 years. We
-considered community taxonomic and functional diversity, composition and
-redundancy trajectories. We based on the annual botanical inventories of
-75 ha of a Neotropical forest and on large trait datasets comprising 7
-leaf, stem and life-history traits. We highlighted a decoupling between
-taxonomic trajectories, that differed among communities, and functional
-trajectories, remained similar and convergent. We explained this
-decoupling by the variations in community functional redundancy
-mitigating the functional impact of disturbance. The humped-shaped
-taxonomic trajectories which amplitude varied with disturbance intensity
-validated the IDH (Intermediate Disturbance Hypothesis). The IDH was
-however dismissed for functional diversity. Although consistent, the
-recovery of community composition, diversity and redundancy remained
-unachieved after 30 years. This acknowledged the need of decades-long
-cycles without disturbance to ensure a complete recovery, and questioned
-tropical forest community resilience after repeated disturbances.
+In the current global change context, it is urgent to anticipate the
+fate of tropical forests. This means understanding tree community
+response to disturbance and the underlying processes. In that respect,
+we aim here to clarify taxonomic and functional post-disturbance
+trajectories, and determine the scope of the Intermediate Disturbance
+Hypothesis (IDH) that remains debated in tropical forests. We analyzed
+community trajectories following a disturbance gradient in a Neotropical
+forest over 30 years. We considered trajectories along time of community
+taxonomic and functional trajectories in terms of richness, evenness,
+composition, and redundancy. We based on the annual botanical
+inventories of 75 ha of a Neotropical forest and on large trait datasets
+comprising seven leaf, stem, and life-history traits. We identified a
+decoupling between taxonomic composition, differing among communities,
+and functional composition, remaining similar and convergent. The
+taxonomic diversity followed humped-shaped trajectories depending on the
+disturbance intensity, which validated the IDH (Intermediate Disturbance
+Hypothesis). The functional diversity trajectories,however, were
+homogeneous among all plots and dismissed the IDH. We explained this
+decoupling by the variations in community functional redundancy that
+mitigated the functional impact of disturbance. Although consistent, the
+recovery of community composition, diversity, and redundancy remained
+unachieved after 30 years. These results acknowledged the need of
+decades-long cycles without disturbance to ensure community complete
+recovery, and questioned community resilience after repeated
+disturbances.
 }
@@
 
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}{Lemma}[section]
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{definition}
+\newtheorem{example}{Example}[section]
+\theoremstyle{definition}
+\newtheorem{exercise}{Exercise}[section]
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
+\newtheorem*{solution}{Solution}
 \begin{document}
@@
 The large areas covered with tropical forests worldwide hold crucial
-environmental, economic and social values. They provide wood and
+environmental, economic, and social values. They provide wood and
 multiple non-timber forest products, shelter a diversified fauna, and
-ensure cultural and human well-being. They regulate as well the local
-and regional climates, and the carbon, water and nutrient cycles.
-However, the growing demand in forests products together with current
-global changes increase the pressure on remaining natural undisturbed
-forests \citep{Morales-Hidalgo2015}. These threats affect the
-disturbance regime that naturally shapes and maintains the structure,
-composition and functioning of tree communities
+ensure cultural and human well-being. They regulate local and regional
+climates, as well as carbon, water and nutrient cycles. However, the
+growing demand in forests products together with current global changes
+increase the pressure on remaining undisturbed forests
+\citep{Morales-Hidalgo2015}. These threats affect the natural
+disturbance regime that defines and maintains the structure,
+composition, and functioning of tree communities
 \citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. To anticipate
-the fate of tropical forests in this context, understanding the response
-of communities to disturbance and deciphering the underlying ecological
-processes is of utmost importance. Disturbance impacts forest
-communities in changing both abiotic and biotic environments, through
-modifications in the fluxes of light, heat and water
-\citep{Goulamoussene2017}. Although the forest cover is generally
-maintained after disturbance, there is a succession of changes in the
-structure and functioning of ecological communities that defines a
-post-disturbance ecosystem trajectory. For now, these trajectories have
-been largely studied through trajectories of forest structural
-parameters such as aboveground biomass, tree height or stem density
-\citep{Piponiot2016, Rutishauser2016}. Recent works have identified some
-of the determinants of post-disturbance biomass trajectories, like
-pre-disturbance forest structure and composition or post-disturbance
-environmental parameters \citep{Herault2018}. Regarding tree community
-diversity and composition however, post-disturbance trajectories have
-not been as thoroughly understood \citep{Guitet2018, Molino2001}. Given
-the variety of species response to disturbance and the huge diversity of
-tropical forests, manifold post-disturbance biodiversity trajectories
-might emerge \citep{Lindenmayer2012, Garcia_florez2017}.
+the fate of tropical forests, it is urgent to understand tree community
+response to disturbance, and the underlying ecological processes. The
+forest cover is generally maintained following disturbance, but
+modifications in the fluxes of light, heat, and water
+\citep{Goulamoussene2017} change community abiotic and biotic
+environments. These changes translate into post-disturbance community
+trajectories that have been largely studied through trajectories of
+forest structural parameters such as aboveground biomass, tree height or
+stem density \citep{Piponiot2016, Rutishauser2016}. Some of the
+determinants of post-disturbance biomass trajectories are already
+identified, like the structure and composition of the pre-disturbance
+community, or the post-disturbance environmental parameters
+\citep{Herault2018}. Community diversity and composition trajectories,
+however, have not been as thoroughly understood
+\citep{Guitet2018, Molino2001}, and manifold biodiversity trajectories
+might emerge given the variety of species response to disturbance and
+the diversity of tropical forests
+\citep{Lindenmayer2012, Garcia_florez2017}.
 
@@
 disturbance is the Intermediate Disturbance Hypothesis (IDH). The IDH
-assumes a relationship between community diversity and the intensity and
-frequency of disturbance events, stating that community diversity peaks
-at intermediate level of disturbance \citep{Connell1978}. The theory
-assumes that the environmental fluctuations following disturbance foster
-both competitively superior species and fast colonizers and thus prevent
-competitive exclusion \citep{Shea2004, Pulsford2016}. In tropical
-forests, however, observations of the IDH often diverge from theoretical
-expectations \citep{Hugues2007, Sheil2003, Norden2017}, and the
-validation of the theory remains controversial
-\citep{Hubbell2001, Fox2013, Sheil2013}. In tropical forests the
-processes underlying the IDH might indeed be complicated by the huge
-diversity found in tropical tree communities and the variety of
-ecological processes that may underlie the IDH (colonization,
-facilitation, coexistence, etc)
-\citep{Lindenmayer2012, Garcia_florez2017}. It is then crucial to test
-the IDH in this context and determine its scope in time, using ecosystem
-time-explicit trajectories \citep{Sheil2003, Shea2004}.
+states a relationship between the community diversity and the intensity
+and frequency of disturbance events, and postulates a diversity peak at
+intermediate level of disturbance \citep{Connell1978}. This is based on
+the fluctuations of community environment following disturbance that
+foster both competitively superior species and fast colonizers, and
+prevents competitive exclusion \citep{Shea2004, Pulsford2016}. In
+tropical forests, however, observations of the IDH often diverge from
+theoretical expectations \citep{Hugues2007, Sheil2003, Norden2017}, and
+the underlying processes might be complicated by the diversity of
+tropical tree communities \citep{Lindenmayer2012, Garcia_florez2017}. In
+this context, the IDH is controversial in tropical forests and remains
+to be tested \citep{Hubbell2001, Fox2013, Sheil2013}.
 
-Analysing community response to disturbance requiers a set of metrics,
-to grasp all aspects of community changes
+Analysing community response to disturbance and grasp all aspects of
+community changes requires a wide array of metrics
 \citep{Sheil2003, Shea2004, Mayfield2010}. The analysis should first
 consider community composition, which is crucial for conservation issues
-and which reveals the pool of species fostered (or hampered) by
-disturbance \citep{Lavorel2002, Bellwood2006}. Then, diversity metrics
-encompassing both richness and evenness should also be considered to
+and reveals the pool of species fostered or hampered by disturbance
+\citep{Lavorel2002, Bellwood2006}. This should be completed with
+diversity metrics encompassing both community richness and evenness to
 assess the changes in community abundance distribution. Besides,
 functional approaches have been shown to usefully complement pure
-taxonomic approaches because they shed light on the species biological
-attributes that directly link community diversity, composition and
+taxonomic approaches as they shed light on the species biological
+attributes directly linking community diversity, composition, and
 redundancy to ecosystem functioning \citep{Violle2007b, Baraloto2012a}.
-In that respect, a vast literature allowed recognizing major traits that
-represent species ecological strategy and determine how they respond to
-changing conditions \citep{Diaz2005}. Specifically in tropical forests,
-the functional approach revealed the post-disturbance deterministic
-processes entailing a shift from a dominance of ``conservative''
+In that respect, a vast literature allowed recognizing major traits
+representing species ecological strategy and determining how they
+respond to changing conditions \citep{Diaz2005}. Specifically, in
+tropical forests, the functional approach revealed the emergence of
+deterministic processes following disturbance. Such deterministic
+processes entailed a shift from a dominance of ``conservative''
 slow-growing species dealing with scarce resources, to a dominance of
@@
 is translated into the trajectories of average community value of key
-functional traits related to resource acquisition (leaf and stem traits)
-and life-history strategy (seed mass, maximum size)
+functional traits related to resource acquisition, as leaf and stem
+traits, and life-history strategy, as seed mass and maximum size
 \citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}.
@@
 hyper-diverse tropical forests \citep{Bellwood2006} mitigates the
-impacts of species removal on ecosystem functioning and determines
-communities resilience after disturbance \citep{Elmqvist2003, Diaz2005}.
+impacts of species removal on ecosystem functioning, and determines
+community resilience after disturbance \citep{Elmqvist2003, Diaz2005}.
 
-In this study, we monitored over 30 years the response of 75 ha of
-Neotropical forest plots set up on a gradient of disturbance intensity,
-from 10 to 60\% of above-ground biomass (AGB) loss. We made use of a
-large functional traits database encompassing major leaf, stem and
+Here, we monitored over 30 years the response of 75 ha of Neotropical
+forest plots set up on a gradient of disturbance intensity, from 10 to
+60\% of above-ground biomass (AGB) loss. We made use of a large
+functional traits database encompassing major leaf, stem, and
 life-history traits in order to draw the taxonomic and functional
-trajectories in terms of richness, evenness, composition and redundancy.
-Specifically, (i) we draw taxonomic and functional post-disturbance
-trajectories and examine the underlying ecological process, (ii) we
-discuss the scope of the IDH regarding the different facets of community
-diversity and the time after disturbance, and (iii) we analyze community
-resilience and time to recovery.
+trajectories in terms of richness, evenness, composition, and
+redundancy. Specifically, (i) we drew taxonomic and functional
+post-disturbance trajectories and examined the underlying ecological
+processes, (ii) we discussed the scope of the IDH regarding taxonomic
+and functional facets of community diversity, and (iii) we analyzed
+community resilience and time to recovery.
 
@@
 
-Paracou station in French Guiana (5\textdegree 18'N and
-52\textdegree 53'W) is located in a lowland tropical rain forest in a
-tropical wet climate with mean annual temperature of 26\textdegree C,
-mean annual precipitation averaging 2980 mm.y\textsuperscript{-1} (30-y
-period) and a 3-month dry season (\textless{} 100
-mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
-one-month dry season in March \citep{Wagner2011}. Elevation ranges from
-5 to 50 m. Across all plots the topography mainly corresponds to
-hilltops or hillsides, while bottomlands cover less than 1 \% of the
-area. Plots are shallow ferralitic acrisols over a layer of transformed
-saprolite with low permeability and lateral drainage. Soil conditions
-are homogeneous, to the exception of the highest hilltops where the
-thick surface allows a free vertical drainage
-\citep{Gourlet-Fleury2004}.
+Paracou station in French Guiana (5° 18'N and 52° 53'W) is located in a
+lowland tropical rain forest in a tropical wet climate with mean annual
+temperature of 26° C, and mean annual precipitation averaging 2980
+mm.y\textsuperscript{-1} (30-y period). The climate comprises a 3-month
+dry season (\textless{} 100 mm.month\textsuperscript{-1}) from
+mid-August to mid-November, and a one-month dry season in March
+\citep{Wagner2011}. Across all plots, elevation ranges from 5 to 50 m,
+and the topography mainly corresponds to hilltops or hillsides, while
+bottomlands cover less than 1 \% of the area. Plots are shallow
+ferralitic acrisols over a layer of transformed saprolite with low
+permeability and lateral drainage. Soil conditions are homogeneous, to
+the exception of the highest hilltops where the thick surface allows a
+free vertical drainage \citep{Gourlet-Fleury2004}.
 
@@
 The experiment comprised three replicates of three sylvicultural
-treatments (hereafter plots T1, T2 and T3) and three control plots (T0).
-All treatments T1, T2 and T3 comprised the logging of 10 trees/ha with
-50 cm minimum DBH that belonged to a set of 58 commercial species
-\citep{Gourlet-Fleury2004}. Treatment T2 additionally comprised a
-thinning treatment by poison-girdling of non-commercial, randomly
-selected species with an average 30 trees/ha with 40 cm minimum DBH.
-Treatment T3 additionally comprised the logging of 15 trees /ha with 40
-cm minimum DBH and the poison-gidling of 20 trees/ha with a 50 cm
-minimum DBH, all belonging to non-commercial species. Considering the
-silvicultural treatments and the following damage, disturbance intensity
-was measured as the percentage of aboveground biomass (\%AGB) lost
-between the first inventory in 1984 and five years after disturbance
-\citep{Piponiot2016} estimated with the BIOMASS R package
-\citep{Rejou2017}. The three treatments were then transformed into a
-continuous disturbance intensity gradient with increasing of
-above-ground biomass (AGB) lost.
+treatments (hereafter plots T1, T2, and T3), and three control plots
+(T0). All treatments T1, T2, and T3 comprised the logging of 10 trees/ha
+with 50 cm minimum DBH that belonged to a set of 58 commercially
+exploited species \citep{Gourlet-Fleury2004}. Treatment T2 additionally
+comprised a thinning treatment by poison-girdling of non-commercially
+exploited species randomly selected with an average of 30 trees/ha with
+40 cm minimum DBH. Treatment T3 additionally comprised the logging of 15
+trees/ha with 40 cm minimum DBH, and the poison-girdling of 20 trees/ha
+with a 50 cm minimum DBH, all belonging to non-commercially exploited
+species. Considering the silvicultural treatments and the following
+damage, disturbance intensity was measured as the percentage of
+aboveground biomass (\%AGB) lost between the first inventory in 1984 and
+five years after disturbance \citep{Piponiot2016} estimated with the
+BIOMASS R package \citep{Rejou2017}. The three treatments were then
+transformed into a continuous disturbance intensity gradient with
+increasing above-ground biomass (AGB) loss.
 
@@
 \caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.
-Treatment intensities are defined by the minimum logging DBH (Diameter at Breast Height), the type of logged species (commercial or not), the density of logged trees, and the total AGB (Above Ground Biomass) lost after treatment.}
+Treatment intensities are defined by the minimum logging DBH (Diameter at Breast Height), the type of logged species (commercially exploited or not), the density of logged trees, and the total AGB (Above Ground Biomass) lost after treatment.}
 \centering
@@
 Control & - & - & - & 0\\
-T1, low & DBH $\geq$ 50 cm, commercial species, $\approx$ 10   $trees.ha^{-1}$ & - & - & $[12-33]$\\
-T2, intermediate & DBH $\geq$ 50 cm, commercial species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30   $trees.ha^{-1}$ & - & $[33-56]$\\
-T3, high & DBH $\geq$ 50 cm, commercial species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15  $trees.ha^{-1}$ & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species,\ $\approx$ 15 $trees.ha^{-1}$ & $[35-56]$\\
+T1, low & DBH $\geq$ 50 cm, commercially exploited species, $\approx$ 10   $trees.ha^{-1}$ & - & - & $[12-33]$\\
+T2, intermediate & DBH $\geq$ 50 cm, commercially exploited species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 40 cm, non-commercially exploited species, $\approx$ 30   $trees.ha^{-1}$ & - & $[33-56]$\\
+T3, high & DBH $\geq$ 50 cm, commercially exploited species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 50 cm, non-commercially exploited species, $\approx$ 15  $trees.ha^{-1}$ & 40 cm $\leq$ DBH $\leq$ 50 cm, non-commercially exploited species,\ $\approx$ 15 $trees.ha^{-1}$ & $[35-56]$\\
 \bottomrule
@@
 Guiana Shield with a dominance of \emph{Fabaceae},
-\emph{Chrysobalanaceae}, \emph{Lecythidaceae} and \emph{Sapotaceae}. In
+\emph{Chrysobalanaceae}, \emph{Lecythidaceae}, and \emph{Sapotaceae}. In
 the 12 experimental plots, all trees above 10 cm DBH have been mapped
@@
 scientific name assigned by botanists during regular botanical
-campaigns. In 1984, specific vernacular names were given to 62
-commercial or common species whereas more infrequent ones were
+campaigns. In 1984, specific vernacular names were given to 62 commmon
+or commercially exploited species. More infrequent species were
 identified under general identifiers only distinguishing trees and
 palms. From 2003, botanical campaigns have been conducted every 5 to 6
-years to identify all trees at the species level but identification
-levels still varied among plots and campaigns.
+years to identify all trees at the species level. In 2015, however,
+identification levels still varied among plots and campaigns.
 
 This variability of protocols in time raised methodological issues as
-vernacular names usually correspond to different botanical species. This
-resulted in significant taxonomic uncertainty that had to be propagated
-to composition and diversity metrics. The uncertainty propagation was
-done through a Bayesian framework reconstituting complete inventories at
-genus level from real incomplete ones on the basis of
-vernacular/botanical names association. Vernacular names were replaced
-through multinomial trials based on the association probability
-\(\big[\alpha_1, \alpha_2,..., \alpha_V\big]\) observed across all
-inventories between each vernacular name \emph{v} and all species
-\(\big[s_1, s_2,..., s_N\big]\):
+vernacular names usually correspond to various botanical species. This
+resulted in significant taxonomic uncertainty that had to be accounted
+for in the measure of composition and diversity metrics. Uncertainty
+propagation was implented withing a Bayesian framework using
+vernacular/botanical names associations to reconstitute complete
+inventories at genus level from real incomplete ones. Vernacular names
+were replaced through multinomial trials based on the association
+probability \(\big[\alpha_1, \alpha_2,..., \alpha_V\big]\) observed
+across all inventories between each vernacular name \emph{v} and all
+species \(\big[s_1, s_2,..., s_N\big]\):
 
@@
 
-Six functional traits representing leaf economics (leaf thickness,
-toughness, total chlorophyll content and specific leaf area) and stem
-economics (wood specific gravity and bark thickness) came from the
-BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}. Trait values
-were assessed from a selection of individuals located in nine permanent
-plots in French Guiana, including two in Paracou, and comprised 294
-species pertaining to 157 genera. Whenever a species was in the dataset
-but missed some trait values (10\% of the species), missing values were
-filled using multivariate imputation by chained equation
+Six functional traits representing the leaf economics (\emph{i.e.} leaf
+thickness, toughness, total chlorophyll content, and specific leaf
+area), and the stem economics (\emph{i.e.} wood specific gravity and
+bark thickness) were obtained from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}.
+Trait values were assessed from a selection of individuals located in
+nine permanent plots in French Guiana, including two in Paracou, and
+comprised 294 species belonging to 157 genera. Whenever a species was in
+the dataset but missed some trait values (10\% of the species), missing
+values were filled using multivariate imputation by chained equation
 \citep{Mice2011}. To account for the phylogenetic signal in the filling
-process, imputations based on samples from the same genus or from the
-same family. Whenever a species was not in the dataset, it was
-attributed a set of trait values randomly sampled among species of the
-next higher taxonomic level (same genus or family). Two life-history
-traits (maximum specific height and seed mass) came from the Mariwenn
-database \footnote{https://www.ecofog.gf/mariwenn/}. The database
-compiles information from a vast literature on the flora of French
-Guiana \citep{Ollivier2007} and comprises 362 species pertaining to 188
-genera. As seed mass information was classified into classes, no data
-filling process was applied and analyses were restricted to the
+process, imputations were based on samples of species from the same
+genus or from the same family. Whenever a species was not in the
+dataset, it was attributed a set of trait values randomly sampled among
+species of the next higher taxonomic level (same genus or family). Two
+life-history traits, maximum specific height and seed mass, came from
+the Mariwenn database \footnote{https://www.ecofog.gf/mariwenn/}. The
+database compiles information from a vast literature on the flora of
+French Guiana \citep{Ollivier2007} and comprises 362 species belonging
+to 188 genera. As seed mass information was classified into classes, no
+data filling process was applied and analyses were restricted to the
 botanical species recorded.
@@
 
-Because of the variability in the precision of botanical identification
-efforts, we were constrained to conduct the taxonomic composition and
-diversity analysis at the genus level. Taxonomic and functional
-trajectories of community composition were followed in a two-dimensional
-NMDS ordination plane. Two NMDS using abundance-based (Bray-Curtis)
-dissimilarity measures were conducted to map either taxonomic or
-functional composition, the latter based on the seven leaf, stem and
-life history traits (without seed mass classes). Trajectories along time
-were reported through the Euclidean distance between the target
-inventories and the reference pre-disturbance inventories of 1984.
-Univariate trajectories of the leaf, stem and life-history traits were
-also visualized with the community weighted means (CWM)
-\citep{Diaz2007}. Species seed mass were given in 5 mass classes, and
-seed mass trajectories were reported as the proportion of each class in
-the inventories (Supplementary materials).
+Because of the variable precision of botanical identification efforts,
+we had to conduct the taxonomic composition and diversity analysis at
+the genus level. Taxonomic and functional trajectories of community
+composition were drawn in a two-dimensional NMDS ordination plane. Two
+NMDS using abundance-based (\emph{i.e.} Bray-Curtis) dissimilarity
+measures were conducted to map either taxonomic or functional
+composition, the latter based on the seven leaf, stem, and life history
+traits (without seed mass classes). Trajectories along time were
+reported through the Euclidean distance between the target inventories
+and the 1984 pre-disturbance inventories of reference. Univariate
+trajectories of the leaf, stem, and life-history traits were also
+visualized with the community weighted means (CWM) \citep{Diaz2007}.
+Species seed mass were given in 5 mass classes, and seed mass
+trajectories were reported as the proportion of each class in the
+inventories (Supplementary materials).
 
-The taxonomic and functional trajectories were analysed from the
-reference pre-disturbance inventories of 1984. The taxonomic diversity
-was reported through species richness and the Hill number translation of
-the Simpson index \citep{Hill1973}. These metrics allowed assessing the
-taxonomic richness as well as evenness, through the comparison between
-these two metrics: results will be discussed directly in terms of
-taxonomic richness and evenness. Both indices belong to the set of HCDT
-or generalized entropy, respectively corresponding to the 0 and 2 order
-of diversity (q), recommended for diversity studies \citep{Marcon2015}.
-The functional diversity was reported using the functional richness and
-functional evenness, through the Rao index of quadratic entropy. The Rao
-index combines species abundance distribution, and the average pairwise
+The taxonomic and functional trajectories were analysed from the 1984
+pre-disturbance inventories of reference. The taxonomic diversity was
+reported through species richness and the Hill number translation of the
+Simpson index \citep{Hill1973}. The comparison between these two metrics
+assesses community taxonomic richness and evenness: thereafter, results
+will be discussed directly in terms of taxonomic richness and evenness.
+Both indices are recommended for diversity studies \citep{Marcon2015},
+and belong to the set of HCDT or generalized entropy corresponding,
+respectively, to the 0 and 2 order of diversity (q). The functional
+diversity was reported using the functional richness and functional
+evenness, through the Rao index of quadratic entropy. The Rao index
+combines species abundance distribution, and the average pairwise
 functional dissimilarity between species computed by the Gower distance.
@@
 analyzed through polynomial regression between (i) taxonomic and
-functional richness and evenness and (ii) the initial \%AGB loss at 10,
-20 and 30 years after disturbance.
+functional richness and evenness, and (ii) the initial \%AGB loss at 10,
+20, and 30 years after disturbance.
 
@@
 individuals of the trait database were mapped in the plane of the first
-two axes from a PCA analysis, which lowered the weight of correlations
-among traits as the PCA axes were combinations of most decoupled traits.
-For each species, the traits probability density (TPD) were computed
-from the mapping of individuals through two-dimension kernel density
-estimators. Second, for each community, the TPD weigthed by species
-abundance were summed accross the functional space. Third, the
-functional space was divided into a 100 x 100 grid and the number of
-species with a positive TPD was counted in each cell. The average count
-across cells minus 1 returned the Community Functional Redundancy, which
-was the average number of species in the community that share the same
-trait values.
+two axes from a PCA analysis. The PCA analysis lowered the weight of
+correlations among traits as the axes are combinations of the most
+decoupled traits. For each species, the traits probability density (TPD)
+were computed from the mapping of individuals through two-dimension
+kernel density estimators. Second, for each community, the TPD weigthed
+by species abundance were summed accross the functional space. Third,
+the functional space was divided into a 100 x 100 grid, and the number
+of species with a positive TPD was counted in each cell. The average
+count across cells minus 1 returned the Community Functional Redundancy,
+which was the average number of species in the community that share the
+same trait values.
 
@@
 
-From 1984, the first pre-disturbance inventory, to 2015 (28 years after
-disturbance), 828 388 individual trees and 591 botanical species
-pertaining to 223 genera and 64 families were recorded.
+From 1984, the first pre-disturbance inventory, to 2015, 28 years after
+disturbance, 828 388 individual trees and 591 botanical species spanning
+223 genera and 64 families were recorded.
 
-While both taxonomic and functional composition remained stable in
-undisturbed communities (Fig. \ref{fig:NMDSplans}), they followed marked
-and consistent trajectories in disturbed plots over time.
+In undisturbed plots, both taxonomic and functional composition remained
+stable (Fig. \ref{fig:NMDSplans}). In disturbed plots, both trajectories
+followed marked and consistent trajectories over time.
 
-The functional composition trajectory in disturbed communities resembled
-to cyclic compositional changes with an incomplete recovery to the
-initial composition (Fig.\ref{fig:NMDSplans}). The maximum dissimilarity
-with the initial state was positively correlated with the disturbance
-intensity for both taxonomic and functional composition
+The functional composition trajectory resembled, in disturbed
+communities, to cyclic compositional changes with an incomplete recovery
+of the initial composition (Fig.\ref{fig:NMDSplans}). The maximum
+dissimilarity with the initial state was positively correlated with the
+disturbance intensity for both taxonomic and functional composition
 (\(\rho_{Spearman}^{Taxonomic}=0.87\) and
-\(\rho_{Spearman}^{Functional}=0.90\) respectively). The maximum
+\(\rho_{Spearman}^{Functional}=0.90\), respectively). The maximum
 dissimilarity with the initial state was reached for taxonomic
-composition between 15 to 25 years and around 20 to 25 years for
+composition between 15 to 25 years, and between 20 to 25 years for
 functional composition.
@@
 
-\caption{Plot trajectories in terms of taxonomic composition (\textbf{(a)} and \textbf{(c)}) and functional composition (\textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS plane.
+\caption{Plot trajectories in terms of taxonomic composition (\textbf{(a)} and \textbf{(c)}), and functional composition (\textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS plane.
 Lower panels (\textbf{(c)} and \textbf{(d)}) represent the Euclidean distance to initial condition along the 30 sampled years.
@@
 In control plots, Community Weighted Means (CWM) of functional traits
-remain stable in time.\\
+remained stable in time.\\
 In disturbed plots, they mostly followed unimodal trajectories, either
@@
 of leaf chlorophyll content, which continued to increase 30 years after
-disturbance for some highly disturbed plots. Maximum height at adult
-stage (\emph{Hmax}), leaf toughness and wood specific gravity
+disturbance for 4 out of 6 highly disturbed plots. Maximum height at
+adult stage (\emph{Hmax}), leaf toughness, and wood specific gravity
 (\emph{WSG}) decreased in time and then slightly increased, but remained
 significantly lower than their initial value (Fig. \ref{fig:CWM}). Bark
-thickness and specific leaf area (\emph{SLA}) both increased in time.
+thickness and specific leaf area (\emph{SLA}) both increased along time.
 Bark thickness remained substantially high after 30 years, and
@@
 correlated to the disturbance intensity. Positive correlations were
-observed for Leaf thickness, chlorophyll content, SLA and bark thickness
+observed for leaf thickness, chlorophyll content, SLA and bark thickness
 (\(\rho_{Spearman}^{Leaf thickness}=0.76\),
@@
 \(\rho_{Spearman}^{Bark thickness}=0.71\)). Negative correlation was
-observed for Leaf toughness, WSG and Hmax
+observed for leaf toughness, WSG, and Hmax
 (\(\rho_{Spearman}^{Leaf toughness}=-0.53\),
@@
 
-\caption{Trajectories of community weighted means over 30 years after disturbance of four leaf traits (Leaf thickness, chlorophyll content, toughness, and specific area), two stem traits (wood specific gravity, and bark thickness) and one life history trait (Specific maximum height at adult stage).
+\caption{Trajectories of community weighted means over 30 years after disturbance of four leaf traits (leaf thickness, chlorophyll content, toughness, and specific area), two stem traits (wood specific gravity and bark thickness), and one life history trait (species maximum height at adult stage).
 }\label{fig:CWM}
@@
 evenness then stabilized except for two intensively-disturbed plots
-(number 8 and 12) for which it kept increasing \ref{fig:DivTaxo}).
+(number 8 and 12) for which it kept increasing (Fig. \ref{fig:DivTaxo}).
 
@@
 \caption{Trajectories of community taxonomic richness \textbf{(a)}, Simpson diversity \textbf{(b)}, functional richness \textbf{(c)}, and Rao diversity \textbf{(d)}.
-Values correspond to the difference over 30 years of community diversity with the values of reference 1984 pre-disturbance inventories.
+Values correspond to the difference over 30 years of community diversity with the values of 1984 pre-disturbance inventories of reference.
 Shaded areas are the credibility intervals }\label{fig:DivTaxo}
@@
 being positively correlated to \%AGB loss
-\(\rho_{Spearman}^{Richness}=0.76\) and \(\rho_{Spearman}^{Rao}=0.60\).
-Functional richness and evenness displayed for low disturbance intensity
-a low but long-lasting increase up to a maximum reached after 20-25
-years. For high disturbance intensity, they generally displayed a fast
-but short increase followed after 10 years by a slow decrease towards
-the initial values.
+\(\rho_{Spearman}^{Richness} = 0.76\) and\\
+\(\rho_{Spearman}^{Rao} = 0.60\). For low disturbance intensity,
+functional richness and evenness displayed a low but long-lasting
+increase up to a maximum reached after 20-25 years. For high disturbance
+intensity, they generally displayed a fast but short increase followed,
+after 10 years, by a slow decrease towards the initial values.
 
 The second-degree polynomial regressions between (i) the percentage AGB
-loss and (ii) the taxonomic and functional diversity showed various
-shape depending on the diversity indices and on the time since
+loss, and (ii) the taxonomic and functional diversity showed various
+shapes depending on the diversity indices and on the time since
 disturbance (Fig. \ref{fig:IDHplot}). Regarding taxonomic diversity, the
 relationship between disturbance intensity and diversity was more
-markedly hump-shaped for richness than for evenness and peaked at 20\%
+markedly hump-shaped for richness than for evenness, and peaked at 20\%
 of initial AGB loss. Regarding functional diversity, the relationship
-was almost linear and similar between richness and evenness. Generally,
-all relationships were stronger 20 or 30 years after disturbance than
-observed for 10 years.
+was almost linear, and was similar between richness and evenness. All
+the relationships between disturbance intensity and diversity were
+stronger 20 or 30 years after disturbance than they were 10 years after
+disturbance.
 
@@
 
-\caption{Relationship between the initial \%AGB loss and community taxonomic richness \textbf{(a)}, taxonomic evenness \textbf{(b)}, functional richness \textbf{(c)},and functional evenness \textbf{(d)} at 10, 20 and 30 years after disturbance}\label{fig:IDHplot}
+\caption{Relationship between the initial \%AGB loss and community taxonomic richness \textbf{(a)}, taxonomic evenness \textbf{(b)}, functional richness \textbf{(c)}, and functional evenness \textbf{(d)} at 10, 20, and 30 years after disturbance.}\label{fig:IDHplot}
 \end{figure*}
@@
 disturbed plots had lower functional redundancy than control plots and
-followed similar hump-shaped trajectories (\ref{fig:RedFunRest}). The
-maximum redundancy loss was positively correlated with the disturbance
-intensity (\(\rho_{Spearman}=0.47\)) and the recovery trajectory had not
-attained initial values for any disturbed communities after 30 years.
+followed similar hump-shaped trajectories (Fig. \ref{fig:RedFunRest}).
+The maximum redundancy loss was positively correlated with the
+disturbance intensity (\(\rho_{Spearman}=0.47\)) and the recovery
+trajectory had not attained initial values for any disturbed communities
+after 30 years.
 
@@
 
-Our analysis highlighted the decoupling between functional and taxonomic
-trajectories, with taxonomic trajectories maintaining initial
-differences among plots while functional trajectories converged in the
-functional space.
-
-The humped-shaped taxonomic trajectories which amplitude varied with
-disturbance intensity validated the IDH. Taxonomic diversity was
-maximized at an intensity threshold of 20-25\% AGB loss and 20 years
-after disturbance, when the recruitment of pioneers and
-late-successionals were balanced. The IDH however did not apply to the
-functional diversity. The decoupling between taxonomic and functional
-response was linked to the disturbance impact on the functional
-redundancy, that proved determining in community recovery.
+Our analysis revealed the decoupling between functional and taxonomic
+trajectories. While the initial differences in taxonomic composition
+among plots were maintained, the functional composition trajectories
+converged in the functional space. In terms of diversity, only the
+taxonomic trajectories validated the IDH that explained humped-shaped
+post-disturbance trajectories whith an amplitude depending on the
+disturbance intensity. The decoupling between taxonomic and functional
+response was explained by variations of community functional redundancy
+that mitigated the functional impact of disturbance, and appeared as a
+determinant of community recovery.
 
@@
 
-Regarding taxonomic composition, the different tree communities were
-substantially different before disturbance, and this is visualized by
-their distinct starting points on the NMDS axis 2. These initial
-differences were maintained all along the 30 years following
-disturbance, with the disturbance leading a displacement on the NMDS
-axis 1 only. Changes were thus similar and may correspond to the
-recruitment of a group of pioneers, e.g. \emph{Cecropia} spp.,
-\emph{Vismia} spp., shared by all plots and whatever their initial
-taxonomic differences and the intensity of disturbance
-\citep{Denslow2000, Bongers2009}. Taxonomic trajectories seems to start
-a recovery to the initial composition which, although far from fully
-achieved after 30 years, suggested that taxonomic composition is
-resilient and that the initial differences will be maintained in time
-\citep{Folke2006}. This would show that species not belonging to the
-pre-disturbance community were rarely recruited in the long-term,
-probably because of the common dispersal limitations among tropical tree
-species \citep{Svenning2005}.
+Community taxonomic composition substantially differed before
+disturbance, as materialized by their distinct location on the NMDS axis
+2. These initial differences were maintained along the 30 years
+following disturbance, with the disturbance leading a displacement on
+the NMDS axis 1 only. Taxonomic composition changes were similar among
+plots and may correspond to the recruitment of a group of pioneers, like
+\emph{Cecropia spp.} or \emph{Miconia spp.}, common to all plots,
+whatever their initial taxonomic differences and the intensity of
+disturbance \citep{Denslow2000, Bongers2009}. Taxonomic trajectories
+initiated a recovery towards the initial composition. This recovery,
+although far from being achieved after 30 years, suggested the
+resilience of community taxonomic composition and the maintenance of
+initial composition differences \citep{Folke2006}. Such resilience
+suggested that species not belonging to the pre-disturbance community
+were rarely recruited in the long-term, probably because of species
+dispersal limitation that is common among tropical species
+\citep{Svenning2005}.
 
-Regarding functional composition, initial communities had similar
-starting points and this contrasts to taxonomic composition. This means
-that, despite differences in species composition, the functional
-signature of the tree communities were quite similar. Following
-disturbance, functional trajectories were consistent in the NMDS plane
-with displacement intensities linked to disturbance intensities and thus
-relied upon the recruitment of a shared pool of functional types
-previously infrequent or absent. This common pool is composed of pioneer
-resource-acquisitive (low leaf toughness, wood specific gravity, maximum
-height and high Specific Leaf area) strategies that translated by a
-displacement to the right along the first NMDS axis
-\citep{Westoby1998, Wright2004, Reich2014}. Thereafter, the pioneers
-recruited primarily were progressively excluded by long-lived, more
-competitive and shade-tolerant species. The community functional
-composition then quickly returned towards more resource-conservative
-strategies, suggesting the recovery of the initial functional
-composition. This recovery translated in the functional plane by a
+Community functional composition trajectories, in contrast, were similar
+in the functional space \citep{Fukami2005}. The amplitude of the
+compositional changes depended on the disturbance intensity. Functional
+trajectories were probably driven by the recruitment of species
+infrequent or absent before disturbance, and belonging to a pool of
+species common to all plots. This common pool was composed of pioneers
+with ``resource-acquisitive''" strategies displaying low leaf toughness,
+wood specific gravity, maximum height, and high specific leaf area. The
+recruitment of these pioneers drove a displacement to the right along
+the first NMDS axis \citep{Westoby1998, Wright2004, Reich2014}.
+Thereafter, the first recruited pioneers were progressively excluded by
+long-lived, more competitive, and shade-tolerant species. The
+recruitment of these late-successionals marked the recovery of the
+initial functional composition with more ``resource-conservative''"
+strategies . This recovery translated in the functional plane by a
 displacement left along the first axis and upward along the second axis
@@
 
-Both taxonomic and functional composition trajectories initiated a
-return towards pre-disturbance state after 30 years, which highlighted
-the taxonomic and functional resilience of communities. Taxonomic and
-functional trajectories however appeared decoupled: while taxonomic
-trajectories maintained the initial differences among communities, the
-functional trajectories where similar and convergent in the functional
-space \citep{Fukami2005} exemplifying the simultaneous operation of
-trait-based assembly rules and species-level priority effects that
-shapes tree community assembly in this forest, making it both
-deterministic in functional space and historically contingent in
-taxonomic space.
+The decoupling between taxonomic and functional trajectories suggested
+that simultaneous operation of trait-based assembly rules and
+species-level priority effects shaped tree community assembly in Paracou
+forest. Tree community assembly would then be both deterministic in the
+functional space, and historically contingent in the taxonomic space.
 
@@
 
-Trajectories of taxonomic richness and evenness were markedly different
-below and above an intensity threshold (\ref{fig:DivTaxo}). Below the
-intensity threshold, for disturbances in the order of treatments 1 and
-2, both taxonomic richness and evenness increased according to the
-disturbance intensity \citep{Martin2015, Chaudhary2016}. This suggested
-that while trees surviving after disturbance remained numerous enough to
-maintain the richness of the pre-disturbance community
-\citep{Bongers2009}, the recruitment of pioneers previously infrequent
-or absent allowed increasing the taxonomic richness
-\citep{Martin2015, Chaudhary2016}. As these pioneers became more
-dominant, they balanced the usual hyper-dominance of tropical forests
-and increased the taxonomic evenness \citep{Baraloto2012a}. Above the
-intensity threshold, like for treatment 3, the taxonomic richness did
-not exceed the initial value. The richness of surviving trees was then
-too low to be offset by the recruitment of pioneers. In the Guiana
-shield indeed, the pool of true pioneers specifically recruited after
-disturbance is restricted to a few common genera (e.g. \emph{Cecropia}
-spp., \emph{Vismia} spp.) \citep{Guitet2018}.
+Trajectories of taxonomic richness and evenness differed markedly below
+and above an intensity threshold (Fig. \ref{fig:DivTaxo}). For low and
+intermediate disturbance, both taxonomic richness and evenness increased
+according to the disturbance intensity
+\citep{Martin2015, Chaudhary2016}. This suggested that the recruitment
+of pioneers, previously infrequent or absent, increased the taxonomic
+richness \citep{Martin2015, Chaudhary2016}, and that trees surviving
+after disturbance remained numerous enough to maintain the richness of
+the pre-disturbance community \citep{Bongers2009}. The pioneers thus
+recruited became abundant, which balanced the usual hyper-dominance of
+tropical forests and increased the taxonomic evenness
+\citep{Baraloto2012a}. Above the intensity threshold, like for treatment
+3 (high intensity), the taxonomic richness did not exceed the initial
+value in the first years following disturbance. No increase of the
+taxonomic richness suggested that the richness of surviving trees was
+lower than this of the pre-disturbance community, and that this
+difference was not offset by the recruitment of pioneers. In the Guiana
+shield, the pool of true pioneers specifically recruited after
+disturbance is restricted to a few common genera (e.g. \emph{Cecropia
+spp.}, \emph{Miconia spp.}, \emph{Tapirira spp.}) \citep{Guitet2018}.
+Considering the different times after disturbance, there was always a
+humped-shaped pattern linking the disturbance intensity with the
+post-disturbance increase in taxonomic richness and evenness (Fig.
+\ref{fig:IDHplot}). Both taxonomic richness and evenness were maximized
+at an intermediate intensity, around 20-25\% of AGB lost.
 
-Considering different moments of the post-disturbance trajectories,
-there was always a humped-shaped pattern linking taxonomic richness or
-evenness and the intensity of disturbance (\ref{fig:IDHplot}). Both
-taxonomic richness and evenness were maximized at an intermediate
-intensity, around 20-25\% of AGB lost, and this along the 30 years
-following disturbance.
-
-Regarding community functional trajectories, no marked differences were
-observed among the post-disturbance trajectories (\ref{fig:DivTaxo}).
-Whenever the disturbance intensity, functional richness and evenness
-first increased with the recruitment of pioneers that were functionally
-highly different from the pre-disturbance community
-\citep{Denslow1980, Molino2001}. Then, after 15 to 20 years community
-functional richness and evenness started to decrease, with the decline
-of the first established species. Right after disturbance, short-lived
-pioneers benefited from the resources made available and prevented the
-establishment of other species. These pioneers then declined, which
+Regarding community functional trajectories, however, no marked
+differences were observed among post-disturbance trajectories (Fig.
+\ref{fig:DivTaxo}). Whenever the disturbance intensity, there was first
+an increase of both functional richness and evenness. Such increase
+suggested the recruitment of pioneers that were functionally highly
+different from the pre-disturbance community
+\citep{Denslow1980, Molino2001}. The recruitment of pioneers hampered
+the recruitment of other species in the first place. After 15 to 20
+years, the first pioneer recruits declined and were replaced by species
+functionally more similar to the pre-disturbance community, which
 decreased the functional richness and evenness \citep{Walker2009}.
-
-Considering different times after disturbance, both functional richness
-and evenness kept increasing with the disturbance intensity and did not
-display humped-shaped trajectories (\ref{fig:IDHplot}).
 
@@
 
-The loss of species following disturbance decreased the functional
+The loss of species following disturbance decreased community functional
 redundancy during the first 15 years. Progressively though, the
 functional redundancy was restored through the replacement of the
-resource-acquisitive strategies species by more late-successional
-species that were functionally closer to the pre-disturbance community.
-This replacement followed the lottery recruitment rules, implying an
-easy recruitment for the first species but becoming increasingly
-hampered by the emergence of interspecific competition
-\citep{Busing2002} and this explains why the slope of the recovery
-trajectory became less and less significant 20 years after disturbance.
-The recovery of the functional redundancy then relied upon the random
-process of species recruitment and was increasingly slow and difficult
-to anticipate \citep{Elmqvist2003, Diaz2005}. This suggests a low
+species with ``resource-acquisitive''" strategies by more
+late-successional species, functionally closer to the pre-disturbance
+community. The replacement of pioneers by late-successionals followed
+the lottery recruitment rules. This rules mean an easy recruitment of
+the first species that becomes increasingly difficult, as the following
+species are hampered by the emergence of interspecific competition
+\citep{Busing2002}. The increasing competition among species explained
+why the recovery trajectory slowed down 20 years after disturbance. The
+recovery of the functional redundancy then relied upon the random
+process of species recruitment, and was increasingly slow and difficult
+to anticipate \citep{Elmqvist2003, Diaz2005}. This suggested a low
 resilience of the functional redundancy with the random recovery of
-infrequent species increasing the risks to lose keystone species, with
+infrequent species increasing the risks of losing keystone species, with
 unexpected ecological consequences
 \citep{Jones1994, Chazdon2003a, Diaz2005}. Infrequent species might
-indeed have unique functional characteristics, apart from those
-considered here, in the ecosystem or be a key resource for some fauna
-\citep{Schleuning2016}.
+indeed have unique functional characteristics in the ecosystem, apart
+from the ones considered here, in the ecosystem or be a key resource for
+some of the fauna \citep{Schleuning2016}.
 
@@
 Post-disturbance trajectories of tree community composition and
-diversity were both shaped by the recruitment of a determined pool of
-pioneers identical among local communities and independent of the
-disturbance intensity. (i) Composition trajectories in taxonomic and
-functional space nevertheless appeared decoupled. While functional
-trajectories remained similar in the functional space and converged
-towards the recovery of a comparable initial state, taxonomic
-trajectories showed initial differences in community composition that
-were maintained along time. Community high functional redundancy
-mediated this decoupling, as the loss of a species does not necessarily
-entails the loss of its functional charactristics. (ii) Diversity
-trajectories were contrasted as well. While the functional trajectories
-remained similar whatever the disturbance intensity, taxonomic
-trajectories were markedly different with a threshold at 20-25\% AGB
-removed for which taxonomic richness was maximized. The Intermediate
-Disturbance Hypothesis applied well to taxonomic diversity, but not to
-the functional diversity. Taxonomic diversity was maximized at an
-intermediate time after disturbance, that was around 25 years after
-disturbance when the recruitment of early- and late-successional species
-were balanced. Whenever the disturbance intensity, community resilience
-(in terms of recovery of the pre-disturbance state) was tangible but
-required several decades and relied upon the random lottery recruitment
-of rare species. Given the long-term impacts of disturbance observed, we
-suggest that 30 years is not enough time for tropical communities to
-recover, even after relatively low intensity disturbance. Much of
-community response to disturbance rely on species recruitment: refined
-understanding of post-disturbance trajectories would be given by a
-closer analysis of recruitment processes.
+diversity were driven by the recruitment of a determined pool of
+pioneers, identical among local communities, and independent of the
+disturbance intensity. The taxonomic composition trajectories maintained
+the initial differences among communities, while the functional
+trajectories were similar, and converged in the functional space towards
+the recovery of the initial composition. The diversity trajectories were
+contrasted as well. While the functional trajectories remained similar
+whatever the disturbance intensity, taxonomic trajectories were markedly
+different from a threshold of 20-25\% AGB lost that maximized the
+taxonomic richness. The Intermediate Disturbance Hypothesis applied well
+to taxonomic diversity, but not to functional diversity. The decoupling
+between taxonomic and functional trajectories was mediated by the
+variations in functional redundancy, as the loss of a species does not
+necessarily entails the loss of its functional characteristics.
+Community resilience, in terms of recovery of the pre-disturbance state,
+was tangible but required several decades, and relied upon the random
+lottery recruitment of rare species. Given the long-term impacts of
+disturbance observed, we suggest that 30 years is not enough time for
+tropical communities to recover, even after relatively low intensity
+disturbance. Much of community response to disturbance rely on the
+processes of species recruitment. A more refined understanding of the
+post-disturbance trajectories would be gained by a closer analysis of
+the recruitment process.
 
@@
 
-We are in debt with all technicians and colleagues who helped setting up
-the plots and collecting data over years. Without their precious work,
-this study would have not been possible and they may be warmly thanked
-here.
+We are in debt with all Paracou station technicians and colleagues who
+helped setting up the plots and collecting data over years. Without
+their precious work, this study would have not been possible and they
+may be warmly thanked here. Our work benefited from an ``Investissement
+d'Avenir'' grant managed by the Agence Nationale de la Recherche (LABEX
+CEBA, ref ANR-10-LBX-25). We thank Niklas Tysklind for the usefull help
+with english proofing.
 
@@
 
-AM, EM \& BH designed the study, developed the analysis framework and
+AM, EM \& BH designed the study, developed the analysis framework, and
 interpreted the results. AM wrote the manuscript with contributions by

### 2019-09-17 WholePlotTrajectories.tex
final text

@@
 \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
+\usepackage[left]{lineno} \linenumbers
 
@@
 Control & - & - & - & 0\\
-T1, low & DBH $\geq$ 50 cm, commercially exploited species, $\approx$ 10   $trees.ha^{-1}$ & - & - & $[12-33]$\\
+T1, low & DBH $\geq$ 50 cm, commercially exploited species, <br/> $\approx$ 10   $trees.ha^{-1}$ & - & - & $[12-33]$\\
 T2, intermediate & DBH $\geq$ 50 cm, commercially exploited species, $\approx$ 10  $trees.ha^{-1}$ & DBH $\geq$ 40 cm, non-commercially exploited species, $\approx$ 30   $trees.ha^{-1}$ & - & $[33-56]$\\
@@
 
-See Supplementary Materials -Fig. S1 and \citet{Aubry-Kientz2013} for
+See Supplementary Materials -Fig. A.1 and \citet{Aubry-Kientz2013} for
 the detailed methodology.
@@
 class decreased while it stabilized for the two other lightest seed mass
-classes (Supp. Mat. - Fig. S2).
+classes (Supp. Mat. - Fig. A.2).
 

### 2019-09-17 WholePlotTrajectories_SuppMat.Rmd
mise en page
final text

@@
   - affsuperscript: 2
-    dptuniv: INPHB (Institut National Ploytechnique Félix Houphoüet Boigny) 
+    dptuniv: INPHB (Institut National Polytechnique Félix Houphoüet Boigny) 
     address: Yamoussoukro, Ivory Coast

### 2019-09-17 Scripts/GraphicalAbstract.R

@@ -1,0 +1,67 @@
+library("shape")
+library("plotrix")
+
+treatments<-list(c(1,6,11),c(2,7,9),c(3,5,10),c(4,8,12))
+names(treatments)<-c("Control","Low","Intermediate","High")
+ColorsTr<-c("darkolivegreen2","gold","orangered","darkred")
+colyear<-c("deepskyblue","cornflowerblue","darkslateblue")
+time<-c("1995","2005","2015")
+
+smooth<-function(mat,larg){return(do.call(cbind,lapply(1:ncol(mat),function(step){
+  range<-max(1,step-larg):min(ncol(mat),step+larg)
+  rowSums(mat[,range])/length(range)})))}
+
+
+load("DB/RichnessIDH");load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB");load("DB/FunRichIDH")
+  
+windows()
+  
+Data<-Rich
+AgbLoss<-AGBloss
+
+Ylim<-c(min(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))),
+          max(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))))
+
+plot(AgbLoss[,"AGB"],AgbLoss[,"AGB"],type="n",xlab="",ylab="",
+       ylim=Ylim,xaxt='n',cex.axis=1.5)
+Axis(side=1, labels=FALSE)
+  
+Ti<-2
+
+toplot<-unlist(lapply(Data, function(tr){return(tr[,time[Ti],"0.5"])}))
+
+abs<-AgbLoss[which(AgbLoss[,"plot"]%in%names(toplot)),"AGB"]
+points(abs,toplot,pch=20)
+Lm<-lm(toplot~abs)
+Lm2<-lm(toplot~abs+I(abs^2))
+
+abs_pred<-seq(min(abs),max(abs),length.out=100)
+lines(sort(abs_pred),predict(Lm2,newdata=data.frame(abs=abs_pred)),lwd=5)
+
+
+
+
+
+windows()
+
+Data<-FunRich
+AgbLoss<-AGBloss
+
+Ylim<-c(min(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))),
+        max(unlist(lapply(Data, function(tr){return(tr[,,"0.5"])}))))
+
[…]

### 2019-09-25 WholePlotTrajectories.Rmd
files

@@
   \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
+  \usepackage[left]{lineno}
+  \linenumbers
 bibliography: references.bib
@@
 output:
+  bookdown::word_document2: default
   bookdown::pdf_book:
@@
     keep_tex: TRUE
-  bookdown::word_document2: default
   bookdown::gitbook: default
@@
   Timber=c("-", 
-           "DBH $\\geq$ 50 cm, commercially exploited species, $\\approx$ 10   $trees.ha^{-1}$", 
+           "DBH $\\geq$ 50 cm, commercially exploited species, <br/> $\\approx$ 10   $trees.ha^{-1}$", 
            "DBH $\\geq$ 50 cm, commercially exploited species, $\\approx$ 10  $trees.ha^{-1}$",
@@
 
-See Supplementary Materials -Fig. S1 and @Aubry-Kientz2013 for the detailed methodology.
+See Supplementary Materials -Fig. A.1 and @Aubry-Kientz2013 for the detailed methodology.
 
@@
 The proportions of the three lightest seed mass classes increased in all disturbed plots. 
-After 30 years the proportion of lightest seed mass class decreased while it stabilized for the two other lightest seed mass classes (Supp. Mat. - Fig. S2).
+After 30 years the proportion of lightest seed mass class decreased while it stabilized for the two other lightest seed mass classes (Supp. Mat. - Fig. A.2).
 

### 2019-09-25 WholePlotTrajectories_SuppMat.Rmd
files

@@
   \hyphenation{bio-di-ver-si-ty sap-lings}
+  \captionsetup{labelformat=empty}
 lang: english
@@
 
+```{r, include=F}
+library(captioner)
+figures <- captioner(prefix = "Figure S", auto_space = FALSE)
+
+figures("FigBayesianScheme", "Scheme of the Bayesian process propagating the taxonomic uncertainty", display=FALSE)
+figures("FigSeedMassClasses", "Trajectories of seed mass classes proportions over 30 years after disturbance.
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).", display=FALSE)
+figures("PCAgraphs", "Summarizing plots of the functional PCA analysis defining communities functional space.
+(ref:Bolda) Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
+(ref:Boldb) bow plot of the PCA eignevalues in percentage.
+(ref:Boldc) Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).", display=FALSE)
+```
+
 \onecolumn
+
+#Supplementary Material {-}
+
+\pagenumbering{arabic}
 
@@
 
-```{r FigBayesianScheme,out.width='100%',fig.cap="Scheme of the Bayesian process propagating the taxonomic uncertainty",echo=F, fig.env='figure',fig.pos='h'}
+```{r FigBayesianScheme,out.width='100%',fig.cap=figures("FigBayesianScheme"),echo=F, fig.env='figure',fig.pos='h'}
 knitr::include_graphics("ExternalFig/BayesianSheme.png")
 ```
+
+\pagebreak
 
@@
 
-```{r FigSeedMassClasses, echo=FALSE, out.width='80%', fig.asp=2/5,fig.cap="Trajectories of seed mass classes proportions over 30 years after disturbance.
-Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).", fig.env='figure',fig.pos='h'}
+```{r FigSeedMassClasses, echo=FALSE, out.width='80%', fig.asp=2/5,fig.cap=figures("FigSeedMassClasses"), fig.env='figure',fig.pos='h'}
 load("DB/Smass_ClassProp")
@@
 
+\pagebreak
+
 **Figure S3**: Functional PCA
 
-```{r PCA_graphs, out.width = '100%', echo=FALSE, fig.cap="Summarizing plots of the functional PCA analysis defining communities functional space.
-(ref:Bolda) Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
-(ref:Boldb) bow plot of the PCA eignevalues in percentage.
-(ref:Boldc) Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).", fig.env='figure',fig.pos='h'}
+```{r PCAgraphs, echo=FALSE, out.width = '100%', fig.cap=figures("PCAgraphs"), fig.env='figure',fig.pos='h'}
 load("DB/FunctionalPCA")

### 2019-09-25 WholePlotTrajectories_SuppMat.tex
final text
files

@@
 \hyphenation{bio-di-ver-si-ty sap-lings}
+\captionsetup{labelformat=empty}
 
@@
 \textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles,
-Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}INPHB (Institut National Ploytechnique Félix Houphoüet Boigny)\\ \hspace{1em} Yamoussoukro, Ivory Coast
+Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}INPHB (Institut National Polytechnique Félix Houphoüet Boigny)\\ \hspace{1em} Yamoussoukro, Ivory Coast
 }
@@
 
-The following information are available for this article:
+\section*{Supplementary Material}\label{supplementary-material}
+\addcontentsline{toc}{section}{Supplementary Material}
+
+\pagenumbering{arabic}
 
@@
 
-\caption{Scheme of the Bayesian process propagating the taxonomic uncertainty}\label{fig:FigBayesianScheme}
+\caption{Figure S1: Scheme of the Bayesian process propagating the taxonomic uncertainty}\label{fig:FigBayesianScheme}
 \end{figure}
+
+\pagebreak
 
@@
 
-\caption{Trajectories of seed mass classes proportions over 30 years after disturbance. Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).}\label{fig:FigSeedMassClasses}
+\caption{Figure S2: Trajectories of seed mass classes proportions over 30 years after disturbance.
+Colors are treatments: green (control), yellow (T1), orange (T2), red (T3).}\label{fig:FigSeedMassClasses}
 \end{figure}
+
+\pagebreak
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_SuppMat_files/figure-latex/PCA_graphs-1}
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_SuppMat_files/figure-latex/PCAgraphs-1}
 
@@
 
-\caption{Summarizing plots of the functional PCA analysis defining communities functional space.
+\caption{Figure S3: Summarizing plots of the functional PCA analysis defining communities functional space.
 \textbf{(a)} Plot of the samples from the BRIDGE functional database in the two dimensional main plan.
 \textbf{(b)} bow plot of the PCA eignevalues in percentage.
-\textbf{(c)} Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).}(\#fig:PCA_graphs)
+\textbf{(c)} Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).}\label{fig:PCAgraphs}
 \end{figure}

### 2019-09-25 WholePlotTrajectories.Rmd
Rm

@@
 output:
-  bookdown::word_document2: default
   bookdown::pdf_book:
@@
     keep_tex: TRUE
+  bookdown::word_document2: default
   bookdown::gitbook: default

### 2020-03-30 WholePlotTrajectories.Rmd
STOTEN article version

--- 2019-08-16 WholePlotTrajectories.Rmd
@@
 ---
-title: "Post-Disturbance Tree Community Trajectories in a Neotropical Forest"
+title: "Diverging taxonomic and functional trajectories following disturbance in a Neotropical forest"
 author:
@@
  In that respect, we aim here to clarify taxonomic and functional post-disturbance  trajectories, and determine the scope of the Intermediate Disturbance Hypothesis (IDH) that remains debated in tropical forests.
- We analyzed community trajectories following a disturbance gradient in a Neotropical forest over 30 years. 
+ \color{black}We analyzed community trajectories following a disturbance gradient from 10 to 60\% of above-ground biomass loss in a Neotropical forest over 30 years.\color{red} 
  We considered trajectories along time of community taxonomic and functional trajectories in terms of richness, evenness, composition, and redundancy.
  We based on the annual botanical inventories of 75 ha of a Neotropical forest and on large trait datasets comprising seven leaf, stem, and life-history traits. 
- We identified a decoupling between taxonomic composition, differing among communities, and functional composition, remaining similar and convergent. 
- The taxonomic diversity followed humped-shaped trajectories depending on the disturbance intensity, which validated the IDH (Intermediate Disturbance Hypothesis).
- The functional diversity trajectories,however, were homogeneous among all plots and dismissed the IDH.
+ We identified a decoupling between taxonomic composition, differing among communities, and functional composition, \color{red} similar among communities and convergent in the functional space.\color{black}
+ \color{red} The taxonomic diversity followed humped-shaped trajectories along time after disturbance depending on the initial disturbance intensity, which validated the IDH (Intermediate Disturbance Hypothesis).\color{black}
+ The functional diversity trajectories,however, were homogeneous among plots and dismissed the IDH.
  We explained this decoupling by the variations in community functional redundancy that mitigated the functional impact of disturbance. 
  Although consistent, the recovery of community composition, diversity, and redundancy remained unachieved after 30 years.
-These results acknowledged the need of decades-long cycles without disturbance to ensure community complete recovery, and questioned community resilience after repeated disturbances.
+\color{red} These results acknowledged the need of decades-long cycles without disturbance to ensure community complete recovery.\color{black}
 keywords: [Community Ecology, Disturbance Trajectories, Intermediate Disturbance Hypothesis, Mid-term Resilience, Neotropical Forests, Taxonomic and Functional Biodiversity]
@@
   \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
+  \usepackage[left]{lineno}
+  \linenumbers
+  \usepackage{bookmark}
 bibliography: references.bib
@@
   bookdown::word_document2: default
-  bookdown::gitbook: default
 ---
@@
 However, the growing demand in forests products together with current global changes increase the pressure on remaining undisturbed forests [@Morales-Hidalgo2015].
-These threats affect the natural disturbance regime that defines and maintains the structure, composition, and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
+\color{red}These threats may change the frequency and magnitude of the natural disturbance regime that defines and maintains the structure, composition, and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
+\color{black}
 To anticipate the fate of tropical forests, it is urgent to understand tree community response to disturbance, and the underlying ecological processes.
@@
 This is based on the fluctuations of community environment following disturbance that foster both competitively superior species and fast colonizers, and prevents competitive exclusion [@Shea2004;@Pulsford2016].
-In tropical forests, however, observations of the IDH often diverge from theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the underlying processes might be complicated by the diversity of tropical tree communities [@Lindenmayer2012;@Garcia_florez2017].
-In this context, the IDH is controversial in tropical forests and remains to be tested [@Hubbell2001;@Fox2013;@Sheil2013].
+\color{red}
+In tropical forests, some studies advocate community deterministic response to disturbance, as for example in the dry and wet forests in Ghana [@Bongers2009] or in the guianese rainforests [@Molino2001], where intermediate disturbance proved to increase the community richness by increasing the proportion of pioneers or heliophilous species in the community while maintaining the richness in old-growth forest species.
+In other cases, however, observations of the IDH diverge from theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017].
+Some studies refute the role of disturbance intensity, as observed in forest gaps in Panama where no difference in species richness was found between forest gaps and undisturbed areas  [@Hubbell2001], or in costa-rican forests where forests that did not follow deterministic convergence display stochastic trajectories [@Norden2015].
+The high diversity of tropical forests may foster the emergence of numerous facilitation, adaptation, and inter- and intra-specific competition following disturbance [@Garcia_florez2017;@Bongers2009].These interactions can result in miscellaneous responses to disturbance [@Lindenmayer2012], which question the validity of the IDH in tropical forests [@Hubbell2001;@Fox2013;@Sheil2013].
+\color{black}
 
@@
 Specifically, (i) we drew taxonomic and functional post-disturbance trajectories and examined the underlying ecological processes, (ii) we discussed the scope of the IDH regarding taxonomic and functional facets of community diversity, and (iii) we analyzed community resilience and time to recovery.
+\color{red}
+We hypothesized that community functional and taxonomic trajectories might not track each other, given the high functional redundancy in tropical forests.
+While functional diversity would be enhanced by the environmental changes following disturbance, specifically the increase in light availability, taxonomic trajectories would only increase until a disturbance intensity threshold, as pioneers recruited following disturbance are not as diverse as late-successional species.
+Community taxonomic and functional would hence follow different rates of recovery.
+\color{black}
 
@@
 
-The experiment is a network of twelve 6.25 ha plots (Table \@ref(tab:Tab1)) that underwent three disturbance treatments in 1987 according to a randomized plot design [@Gourlet-Fleury2004].  
+The experiment is a network of twelve 6.25 ha plots\color{red}, distributed over a 400ha area \color{black}(Table \@ref(tab:Tab1), Supp.
+Mat.
+Fig.
+S1), that undergone three disturbance treatments in 1987 according to a randomized plot design [@Gourlet-Fleury2004].
+\color{red}The experiment intended to simulate logging intensities and distinguished the treatments for non-commercial commercial species, logged at 40cm DBH to simulate thinning operations.\color{black}
 The experiment comprised three replicates of three sylvicultural treatments (hereafter plots T1, T2, and T3), and three control plots (T0).
+\color{red}Control plots are set in mature forest stand that did not undergone any logging nor intense anthropic activities activities for the last 500 years\color{black}.
 All treatments T1, T2, and T3 comprised the logging of 10 trees/ha with 50 cm minimum DBH that belonged to a set of 58 commercially exploited species [@Gourlet-Fleury2004]. 
@@
 Considering the silvicultural treatments and the following damage, disturbance intensity was measured as the percentage of aboveground biomass (%AGB) lost between the first inventory in 1984 and five years after disturbance [@Piponiot2016] estimated with the BIOMASS R package [@Rejou2017].
+\color{red}
+Tree diameter are provided in the experimental dataset, wood density is extracted from global wood density (GWD) database [@Zanne2009], and tree height is retrieved from a generic H–D model based on a single bioclimatic predictor (eqn 6a in [@Chave2014]).
+The AGB is measured from the generalized allometric model eqn 4 of [@Chave2014].
+\color{black}
 The three treatments were then transformed into a continuous disturbance intensity gradient with increasing above-ground biomass (AGB) loss.
@@
 
-The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of *Fabaceae*, *Chrysobalanaceae*, *Lecythidaceae*, and *Sapotaceae*. 
+The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae, and Sapotaceae. 
 In the 12 experimental plots, all trees above 10 cm DBH have been mapped and measured annually since 1984. 
@@
 This resulted in significant taxonomic uncertainty that had to be accounted for in the measure of composition and diversity metrics. 
-Uncertainty propagation was implented withing a Bayesian framework using vernacular/botanical names associations to reconstitute complete inventories at genus level from real incomplete ones. 
+Uncertainty propagation was implented within a Bayesian framework using vernacular/botanical names associations to reconstitute complete inventories at genus level from real incomplete ones. 
 Vernacular names were replaced through multinomial trials based on the association probability $\big[\alpha_1, \alpha_2,..., \alpha_V\big]$ observed across all inventories between each vernacular name *v* and all species $\big[s_1, s_2,..., s_N\big]$:
@@
 
-See Supplementary Materials -Fig. S1 and @Aubry-Kientz2013 for the detailed methodology.
+See Supplementary Materials -Fig. S2 and @Aubry-Kientz2013 for the detailed methodology.
 
@@
 Whenever a species was in the dataset but missed some trait values (10% of the species), missing values were filled using multivariate imputation by chained equation [@Mice2011].
-To account for the phylogenetic signal in the filling process, imputations were based on samples of species from the same genus or from the same family. 
+To account for the phylogenetic signal in the imputation process, imputations were based on samples of species from the same genus or from the same family. 
 Whenever a species was not in the dataset, it was attributed a set of trait values randomly sampled among species of the next higher taxonomic level (same genus or family). 
@@
 
-All composition and diversity metrics were obtained after 50 iterations of the uncertainty propagation framework.
+\color{red}Confidence intervals of the composition and diversity metrics were computed after 50 iterations of the uncertainty propagation framework.\color{black}
 
@@
 Because of the variable precision of botanical identification efforts, we had to conduct the taxonomic composition and diversity analysis at the genus level. 
+\color{red}Functional composition, diversity, and redundancy measures were conducted at species level.\color{black}
 Taxonomic and functional trajectories of community composition were drawn in a two-dimensional NMDS ordination plane.
@@
 The comparison between these two metrics assesses community taxonomic richness and evenness: thereafter, results will be discussed directly in terms of taxonomic richness and evenness.
-Both indices are recommended for diversity studies [@Marcon2015], and belong to the set of HCDT or generalized entropy corresponding, respectively, to the 0 and 2 order of diversity (q). 
-The functional diversity was reported using the functional richness and functional evenness, through the Rao index of quadratic entropy.
-The Rao index combines species abundance distribution, and the average pairwise functional dissimilarity between species computed by the Gower distance.
+Both indices are recommended for diversity studies [@Marcon2015], and belong to the set of \color{red} Tsallis or generalized entropy \color{black} corresponding, respectively, to the 0 and 2 order of diversity (q).
+\color{red}Diversity metrics were measured using the Hqz function from the entropart package [@Marcon2015].\color{black}
+The functional diversity was reported using the functional richness and functional evenness.
+\color{red}Both functional richness and evenness were measured through a distance-based dissimilarity matrix built with the daisy function of the cluster R package [@Cluster2019], using the Gower distance between species, similar to @Herault2007a, @Herault2007b.\color{black}
+The functional diversity was measured through the Rao index of quadratic entropy.
+The Rao index combines the community abundance distribution\color{red}, from species relative abundance, \color{black} and the average pairwise functional dissimilarity between species computed by the Gower distance.
+\color{red}The comparison between these two metrics assesses community functional richness and evenness: thereafter, results will be discussed directly in terms of functional richness and evenness.\color{black}
 
 The impacts of the initial disturbance levels on the maximum gain or loss in richness and evenness were tested with the Spearman rank correlation tests. 
-Richness and evenness trajectories were besides analyzed through polynomial regression between (i) taxonomic and
-functional richness and evenness, and (ii) the initial \%AGB loss at 10, 20, and 30 years after disturbance.
+Richness and evenness trajectories were besides analyzed through polynomial regression \color{red}or simple linear models \color{black} between (i) taxonomic and functional richness and evenness, and (ii) the initial \%AGB loss at 10, 20, and 30 years after disturbance.
+\color{red}The best model fitting, between linear or quadratic model, was chosen based on their respective AIC criterion.
+The explanatory power of the fit were measured to adjusted r-square values.\color{black}
 
@@
 The PCA analysis lowered the weight of correlations among traits as the axes are combinations of the most decoupled traits. 
+\color{red}The two main PCA axis respectively hold 26% and 19% of the trait dataset variance.\color{black}
 For each species, the traits probability density (TPD) were computed from the mapping of individuals through two-dimension kernel density estimators. 
-Second, for each community, the TPD weigthed by species
-abundance were summed accross the functional space. Third, the functional space was divided into a 100 x 100 grid, and the number of species with a positive TPD was counted in each cell. 
+\color{red}
+We used the kde2d function of the MASS package [@MASS2002] that is a gaussian density estimator for a 2-dimensional space.
+\color{black}
+Second, for each community, the TPD weigthed by species abundance were summed across the functional space.
+Third, the functional space was divided into a 100 x 100 grid, and the number of species with a positive TPD was counted in each cell.
 The average count across cells minus 1 returned the Community Functional Redundancy, which was the average number of species in the community that share the same trait values.
@@
 The functional composition trajectory resembled, in disturbed communities, to cyclic compositional changes with an incomplete recovery of the initial composition (Fig.\ref{fig:NMDSplans}). 
-The maximum dissimilarity with the initial state was positively correlated with the disturbance intensity for both taxonomic and functional composition ($\rho_{Spearman}^{Taxonomic}=0.87$ and $\rho_{Spearman}^{Functional}=0.90$, respectively).
+The maximum dissimilarity with the initial state was positively correlated with the disturbance intensity for both taxonomic and functional composition ($\rho_{Spearman}^{Taxonomic}=0.94$ and $\rho_{Spearman}^{Functional}=0.90$, respectively.
+Fig.
+\@ref(fig:SumUp)).
 The maximum dissimilarity with the initial state was reached for taxonomic composition between 15 to 25 years, and between 20 to 25 years for functional composition.
@@
 their initial values, to the exception of leaf chlorophyll content, which continued to increase 30 years after disturbance for 4 out of 6 highly disturbed plots. 
-Maximum height at adult stage (*Hmax*), leaf toughness, and wood specific gravity (*WSG*) decreased in time and then slightly increased, but remained significantly lower than their initial value (Fig.
+ \color{red}Main tendencies for disturbed plots were a decrease followed by a slight increase of maximum height at adult stage (*Hmax*), leaf toughness, and wood specific gravity (*WSG*), but trait values remained significantly lower than their initial value (Fig.
 \ref{fig:CWM}).
+\color{black}
 Bark thickness and specific leaf area (*SLA*) both increased along time. 
-Bark thickness remained substantially high after 30 years, and *SLA* had almost recovered to its initial value. Whatever the functional traits, the maximum
-difference to initial value was highly correlated to the disturbance intensity.
-Positive correlations were observed for leaf thickness, chlorophyll content, SLA and bark thickness ($\rho_{Spearman}^{Leaf thickness}=0.76$, $\rho_{Spearman}^{Chlorophyll content}=0.60$,
-$\rho_{Spearman}^{SLA}=0.93$, $\rho_{Spearman}^{Bark thickness}=0.71$). 
-Negative correlation was observed for leaf toughness, WSG, and Hmax ($\rho_{Spearman}^{Leaf toughness}=-0.53$,
-$\rho_{Spearman}^{WSG}=-0.75$, $\rho_{Spearman}^{Hmax}=-0.40$)
+Bark thickness remained substantially high after 30 years, and *SLA* had almost recovered to its initial value.
+Whatever the functional traits, the maximum difference to initial value was highly correlated to the disturbance intensity.
+\color{red}Positive correlations were observed for leaf thickness, chlorophyll content, SLA and bark thickness ($\rho_{Spearman}^{Leaf thickness}=0.79$, $\rho_{Spearman}^{Chlorophyll content}=0.61$,
+$\rho_{Spearman}^{SLA}=0.99$, $\rho_{Spearman}^{Bark thickness}=0.68$). 
+Maxima were observed between 5 and 15 years following disturbance for leaf thickness and SLA while chlorophyll content and bark thickness kept increasing after 30 years for leaf chlorophyll content and bark thickness.
+Negative correlation was observed for leaf toughness, WSG, and Hmax ($\rho_{Spearman}^{Leaf toughness}=-0.47$,
+$\rho_{Spearman}^{WSG}=-0.80$, $\rho_{Spearman}^{Hmax}=-0.28$).
+Minima were observed between 5 and 15 years following disturbance for Hmax, 10 and 20 years for WSG, and between 15 and 25 years following disturbance for leaf toughness.\color{black}
 The proportions of the three lightest seed mass classes increased in all disturbed plots. 
-After 30 years the proportion of lightest seed mass class decreased while it stabilized for the two other lightest seed mass classes (Supp. Mat. - Fig. S2).
+After 30 years the proportion of lightest seed mass class decreased while it stabilized for the two other lightest seed mass classes (Supp. Mat. - Fig. S3).
 
@@
 After intense disturbance the taxonomic richness followed a more complex trajectory, decreasing for ten years after disturbance before recovering to pre-disturbance values.
-The maximum richness loss or gain after disturbance was positively correlated with the disturbance intensity ($\rho_{Spearman}^{Richness}=0.50$).
+\color{red}The maximum rate of change compared to initial richness was positively correlated with the disturbance intensity ($\rho_{Spearman}^{Richness}=0.73$).\color{black}
 
-In all disturbed plots the evenness first increased until a maximum reached after around 20 years. 
-This maximum was positively correlated with the disturbance intensity ($\rho_{Spearman}^{Simpson}=0.77$).
+In all disturbed plots \color{red}, to the exception of plots 5 (T2) and 12 (T3) that showed a small initial drop, \color{black} the evenness first increased until a maximum reached after around 20 years.
+This maximum was positively correlated with the disturbance intensity ($\rho_{Spearman}^{Evenness}=0.82$, \color{red} Fig. \@ref(fig:SumUp))\color{black}.
 The evenness then stabilized except for two intensively-disturbed plots (number 8 and 12) for which it kept increasing (Fig. \@ref(fig:DivTaxo)).
 
-```{r DivTaxo, out.width = '100%', echo=FALSE, fig.cap="Trajectories of community taxonomic richness (ref:Bolda), Simpson diversity (ref:Boldb), functional richness (ref:Boldc), and Rao diversity (ref:Boldd).
-Values correspond to the difference over 30 years of community diversity with the values of 1984 pre-disturbance inventories of reference.
+```{r DivTaxo, out.width = '100%', echo=FALSE, fig.cap="Trajectories of community taxonomic richness (ref:Bolda), taxonomic evenness, (ref:Boldb), functional richness (ref:Boldc), and functional evenness (ref:Boldd).
+Values correspond to the difference over 30 years of community diversity with the values of 1989 inventories of reference, 5 years after disturbance.
 Shaded areas are the credibility intervals ", fig.env="figure*"}
@@
 Functional richness and evenness remained stable in control plots over the 30 years of monitoring. 
-In disturbed communities, both trajectories depended on the disturbance intensity, with their maximum values in time being positively correlated to \%AGB loss $\rho_{Spearman}^{Richness} = 0.76$ and
-$\rho_{Spearman}^{Rao} = 0.60$.
-For low disturbance intensity, functional richness and evenness displayed a low but long-lasting increase up to a maximum reached after 20-25 years. 
+\color{red}In disturbed communities, both functional richness and evenness trajectories increased following disturbance, with maxima positively correlated to \%AGB loss $\rho_{Spearman}^{Richness} = 0.76$ and $\rho_{Spearman}^{Evenness} = 0.60$ (Fig.
+\@ref(fig:SumUp)).
+\color{black}
+For low \color{red}and intermediate \color{black}disturbance intensity, functional richness and evenness displayed a low but long-lasting increase up to a maximum reached after 20-25 years. 
 For high disturbance intensity, they generally displayed a fast but short increase followed, after 10 years, by a slow decrease towards the initial values.
@@
 \ref{fig:IDHplot}).
-Regarding taxonomic diversity, the relationship between disturbance intensity and diversity was more markedly hump-shaped for richness than for evenness, and peaked at 20\% of initial AGB loss.
-Regarding functional diversity, the relationship was almost linear, and was similar between richness and evenness. 
-All the relationships between disturbance intensity and diversity were stronger 20 or 30 years after disturbance than they were 10 years after disturbance.
+\color{red}Regarding taxonomic diversity, the taxonomic richness showed a humped-shaped trajectory with the disturbance intensity, and peaked at 20\% of initial AGB loss.
+The relationship between disturbance intensity and taxonomic evenness, however, was humped-shaped only 20 years after disturbance and linear for 10 and 30 years after disturbance times.\color{black}
+Regarding functional diversity, \color{red} the relationship was linear for both taxonomic richness and evenness, except for the functional diversity after 30 years. \color{black}
 
@@
 par(mfrow=c(1,4),mar=c(2,2.5,3,0),oma=c(1,1,1,4),no.readonly=TRUE)
-plotIDH(Data=Rich,AgbLoss=AGBloss)
+plotIDH(Data=Rich,AgbLoss=AGBloss_cor)
 mtext("(a) Taxonomic\nRichness",line=1,side=3,adj=0,cex=0.85)
 mtext("Equivalent diversity",side=2,line=2.2,adj=1,cex=0.8)
-plotIDH(Data=Sim,AgbLoss=AGBloss)
+plotIDH(Data=Sim,AgbLoss=AGBloss_cor)
 mtext("(b) Taxonomic\nEvenness",line=1,side=3,adj=0,cex=0.85)
-plotIDH(Data=Rao,AgbLoss=AGBloss)
+plotIDH(Data=Rao,AgbLoss=AGBloss_cor)
 mtext("(c) Functional\nRichness",line=1,side=3,adj=0,cex=0.85)
-plotIDH(Data=FunRich,AgbLoss=AGBloss)
+plotIDH(Data=FunRich,AgbLoss=AGBloss_cor)
 mtext("(d) Functional\nEvenness",line=1,side=3,adj=0,cex=0.85)
 mtext("initial AGB loss (%)",side=1,line=2.2,adj=1,cex=0.8)
-legend("right",inset=-0.5,xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
+legend("right",inset=-0.4,xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")
 
@@
 Control plots displayed stable functional redundancy over time. 
-All disturbed plots had lower functional redundancy than control plots and followed similar hump-shaped trajectories (Fig. \ref{fig:RedFunRest}).
-The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{Spearman}=0.47$) and the recovery trajectory had not attained initial values for any disturbed communities after 30 years.
+All disturbed plots had lower functional redundancy than control plots and followed similar humped-shaped trajectories (Fig. \ref{fig:RedFunRest}).
+The maximum redundancy loss was positively correlated with the disturbance intensity ($\rho_{Spearman}=0.31$, (Fig.
+\@ref(fig:SumUp))) and the recovery trajectory had not attained initial values for any disturbed communities after 30 years.
 
-```{r RedFunRest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
+```{r RedFunRest, out.width = '40%', echo=FALSE, fig.cap="Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
+The redundancy is measured in a two-dimensional plan summarizing 45\\% of species variance in the functional space of the 7 leaf, stem, and life history traits.
+Redundancy value represent the average number of species in the community that share the same trait values.
 Shaded areas are the credibility intervals.", fig.env="figure"}
@@
 
+```{r SumUp, out.width = '100%', echo=FALSE, fig.cap="Maximum changes in taxonomic and functional diversity and composition against disturbance intensity, expressed as percentage of initial disturbance.
+Upper panels display taxonomic richness and evenness (ref:Bolda), and functional richness and evenness (ref:Boldb) expressed in percentage of change compared to initial value.
+Lower panels display taxonomic and functional euclidean distance from pre-disturbance inventory, respectively in the spaces of species and functional traits (ref:Boldc), and functional redundancy expressed as number of redundant species in the community (ref:Boldd).", fig.env="figure"}
+load("DB/IndexChanges")
+
+ChangesDraw(Tbl)
+
+```
+
 # Discussion
 
-Our analysis revealed the decoupling between functional and taxonomic trajectories.
-While the initial differences in taxonomic composition among plots were maintained, the functional composition trajectories converged in the functional space.
+\color{red}Our analysis revealed that community post-disturbance trajectories in taxonomic and functional diversity and composition were decoupled.
+The taxonomic composition specificities between communities before disturbance were maintained following disturbance, while their functional composition converged in the functional space.
+\color{black}
 In terms of diversity, only the taxonomic trajectories validated the IDH that explained humped-shaped post-disturbance trajectories whith an amplitude depending on the disturbance intensity.
-The decoupling between taxonomic and functional response was explained by variations of community functional redundancy that mitigated the functional impact of disturbance, and appeared as a determinant of community recovery.
+The decoupling between taxonomic and functional response \color{red} could be \color{black}explained by variations in community functional redundancy.
+\color{red}The loss of species or the changes in their abundance would impact the taxonomic but not the functional diversity and composition, as redundant species with the same functional characteristics remain in the community.
+The trajectory of the functional redundancy hence appeared determinant for community recovery.\color{black}
 
@@
 
-Community taxonomic composition substantially differed before disturbance, as materialized by their distinct location on the NMDS axis 2. 
-These initial differences were maintained along the 30 years following disturbance, with the disturbance leading a displacement on the NMDS axis 1 only. 
-Taxonomic composition changes were similar among plots and may correspond to the recruitment of a group of pioneers, like *Cecropia spp.* or *Miconia spp.*,  common to all plots, whatever their initial taxonomic differences and the intensity of disturbance [@Denslow2000;@Bongers2009].
+\color{red}From pre-disturbance to 30 years after disturbance, communities showed different location along NMDS axis 2 that hardly changed along time.
+Specificities in community taxonomic composition, materialized by the distinct location on the NMDS axis 2, existed before disturbance.
+The disturbance led a displacement on the NMDS axis 1 only.\color{black}
+Taxonomic \color{red} post-disturbance \color{black} changes in composition were similar among plots and may correspond to the recruitment of a group of pioneers, like *Cecropia spp.* or *Miconia spp.*, common to all plots, whatever their initial taxonomic \color{red}specificities \color{black} and the intensity of the disturbance [@Denslow2000;@Bongers2009].
 Taxonomic trajectories initiated a recovery towards the initial composition.
-This recovery, although far from being achieved after 30 years, suggested the resilience of community taxonomic composition and the maintenance of initial composition differences [@Folke2006]. 
-Such resilience suggested that species not belonging to the pre-disturbance community were rarely recruited in the long-term, probably because of species dispersal limitation that is common among tropical species [@Svenning2005].
+This recovery, although far from being achieved after 30 years, suggested the resilience of community taxonomic composition.
+\color{red}Although the composition dramatically changed following disturbance, community specificities were maintained [@Folke2006]. 
+This maintenance suggested that the recruitment came from (i) a common set of pioneer and light-demanding species in all plots that shape the similar trajectories on NMDS axis 1 and (ii) a very local set of late-successional species that signs (and maintains in time) the initial position of each plot on NMDS axis 2.
+This can be related to species dispersal limitation that is common among tropical species [@Svenning2005].\color{black}
 
-Community functional composition trajectories, in contrast, were similar in the functional space [@Fukami2005].
+Community functional composition trajectories, in contrast, were similar in the functional space.
+\color{red}Such functional convergence contrasting with community taxonomic divergence was already observed in plant communities [@Fukami2005].\color{black}
 The amplitude of the compositional changes depended on the disturbance intensity.
@@
 The recruitment of these pioneers drove a displacement to the right along the first NMDS axis [@Westoby1998;@Wright2004;@Reich2014]. 
+\color{red}This recruitment went along with an increase of the mean SLA, and a decrease in the mean WSG in disturbed communities  (\ref{fig:CWM}).
+These fast and significant changes in mean functional traits closely linked to species light acquisition [@Wright2004;@Chave2009b] suggested major changes in the light environment following disturbance that shape species assemblage in tropical forests [@Pena2008;@Carreno-Rocabado2012].
 Thereafter, the first recruited pioneers were progressively excluded by long-lived, more competitive, and shade-tolerant species.
-The recruitment of these late-successionals marked the recovery of the initial functional composition with more "resource-conservative"" strategies .
+The recruitment of these late-successional species marked the recovery of the initial functional composition with more "resource-conservative"" strategies, corresponding to a fast decrease in community mean SLA and a stabilisation of community WSG, suggesting the progressive closing of forest canopy.
+\color{black}
 This recovery translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig. \ref{fig:NMDSplans}).
@@
 
+\color{red}
+We must temper these results with the low number of plots (3) for each treatment.
+However, the size of each plot is 6.25ha that has been regularly censused over 25 years so that the  trajectories have been drawn using 75ha of permanent plot information, a spatial hold very rarely achieved in such studies.
+Given the relative homogeneity of environmental characteristics of terra firme forests over the Guiana Shield and even the amazonian basin, it seems reasonable that similar trajectories would apply to neighboring areas [@Guitet2015;@Guitet2018].
+\color{black}
+
 ## The scope of the intermediate disturbance hypothesis
@@
 For low and intermediate disturbance, both taxonomic richness and evenness increased according to the disturbance intensity [@Martin2015;@Chaudhary2016].
-This suggested that the recruitment of pioneers, previously infrequent or absent, increased the taxonomic richness [@Martin2015;@Chaudhary2016], and that trees surviving after disturbance remained numerous enough to maintain the richness of the pre-disturbance community [@Bongers2009].
+This suggested that the recruitment of pioneers, previously infrequent or absent, increased the taxonomic richness, and that trees surviving after disturbance remained numerous enough to maintain the richness of the pre-disturbance community [@Bongers2009].
 The pioneers thus recruited became abundant, which balanced the usual hyper-dominance of tropical forests and increased the taxonomic evenness [@Baraloto2012a].
@@
 No increase of the taxonomic richness suggested that the richness of surviving trees was lower than this of the pre-disturbance community, and that this difference was not offset by the recruitment of pioneers.
-In the Guiana shield, the pool of true pioneers specifically recruited after disturbance is restricted to a few common genera (e.g. *Cecropia spp.*, *Miconia spp.*, *Tapirira spp.*) [@Guitet2018].
-Considering the different times after disturbance, there was always a humped-shaped pattern linking the disturbance intensity with the post-disturbance increase in taxonomic richness and evenness (Fig.
-\ref{fig:IDHplot}).
-Both taxonomic richness and evenness were maximized at an intermediate intensity, around 20-25% of AGB lost.
+In the Guiana shield, the pool of \color{red}hard \color{black} pioneers specifically recruited after disturbance is restricted to a few common genera (e.g.
+*Cecropia spp.*, *Miconia spp.*, *Tapirira spp.*) [@Guitet2018].
+\color{red}The disturbance intensity showed a humped-shaped relationship with the post-disturbance increase in taxonomic richness at all the different times after disturbance, and with the increase in taxonomic evenness 10 years after disturbance (Fig.
+\ref{fig:IDHplot}).\color{black}
+Taxonomic richness, and to some extend taxonomic evenness, were maximized at an intermediate intensity, around 20-25% of AGB lost.
 
-Regarding community functional trajectories, however, no marked differences were observed among post-disturbance trajectories (Fig. \ref{fig:DivTaxo}).
-Whenever the disturbance intensity, there was first an increase of both functional richness and evenness.
-Such increase suggested the recruitment of pioneers that were functionally highly different from the pre-disturbance community [@Denslow1980;@Molino2001].
-The recruitment of pioneers hampered the recruitment of other species in the first place.
-After 15 to 20 years, the first pioneer recruits declined and were replaced by species functionally more similar to the pre-disturbance community, which decreased the functional richness and evenness [@Walker2009].
+## The functional redundancy explaining the taxonomic-functional decoupling
 
-## The functional redundancy, key of community resilience
+\color{red}Regarding community functional trajectories (Fig.
+\ref{fig:IDHplot}), there was no intensity threshold above which the diversity-disturbance relationship changed.
+The functional diversity kept increasing along with the disturbance intensity, hence the IDH would not apply.
+Surprisingly, although some species were lost after disturbance, the functional diversity did not decrease in the first place.
+On the contrary, functional diversity increased, probably following the rapid recruitment of pioneers that were functionally highly different from the pre-disturbance community [@Denslow1980;@Molino2001].
 
-The loss of species following disturbance decreased community functional redundancy during the first 15 years. 
-Progressively though, the functional redundancy was restored through the replacement of the species with "resource-acquisitive"" strategies by more late-successional species, functionally closer to the pre-disturbance community.
-The replacement of pioneers by late-successionals followed the lottery recruitment rules.
-This rules mean an easy recruitment of the first species that becomes increasingly difficult, as the following species are hampered by the emergence of interspecific competition [@Busing2002].
-The increasing competition among species explained why the recovery trajectory slowed down 20 years after disturbance. 
-The recovery of the functional redundancy then relied upon the random process of species recruitment, and was increasingly slow and difficult to anticipate [@Elmqvist2003;@Diaz2005]. 
-This suggested a low resilience of the functional redundancy with the random recovery of infrequent species increasing the risks of losing keystone species, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005].
-Infrequent species might indeed have unique functional characteristics in the ecosystem, apart from the ones considered here, in the ecosystem or be a key resource for some of the fauna [@Schleuning2016].
+The loss of species following disturbance, however, decreased the functional redundancy, all the more so that disturbance was intense.
+All in all, because functional diversity was not lowered by disturbance while functional redundancy was, this means that the species that were lost in disturbance are, on the whole, functionally equivalent to the species that survive the disturbance.
+This makes functional redundancy a key to understand community dynamics at a functional level.
+In other words, the high redundancy of tropical forest mean that several species occupy the same functional space, so changes in taxonomic diversity or composition do not necessarily result in changes in community functional characteristics because species are commutable.
+The loss of a redundant species doesn’t change anything for the community functional characteristics, so functional trajectories do not necessarily track taxonomic ones.
+This commutability explains the taxonomic-functional decoupling and the fact that the taxonomic diversity-disturbance relationship supports the IDH while it functional diversity do not.
+Following disturbance, the redundancy was progressively restored through the replacement of "resource-acquisitive" species by more late-successional "resource-conservative" species functionally closer to the pre-disturbance community.
+The time for the functional redundancy to recover would hence be important to assess community recovery.
+\color{black}
 
@@
 
-Post-disturbance trajectories of tree community composition and diversity were driven by the recruitment of a determined pool of pioneers, identical among local communities, and independent of the disturbance intensity.
+Post-disturbance trajectories of tree community composition and diversity would be driven by the recruitment of a determined pool of pioneers, identical among local communities, and independent of the disturbance intensity.
 The taxonomic composition trajectories maintained the initial differences among communities, while the functional trajectories were similar, and converged in the functional space towards the recovery of the initial composition.
 The diversity trajectories were contrasted as well. 
-While the functional trajectories remained similar whatever the disturbance intensity, taxonomic trajectories were markedly different from a threshold of 20-25\% AGB lost that maximized the taxonomic richness.
-The Intermediate Disturbance Hypothesis applied well to taxonomic diversity, but not to functional diversity.
+While the functional trajectories \color{red} displayed the same tendency with a linear increase following all disturbance intensity, \color{black} taxonomic trajectories were markedly different after a threshold of 20-25\% AGB lost that maximized the taxonomic richness.
+The Intermediate Disturbance Hypothesis would apply well to taxonomic diversity, but not to functional diversity.
 The decoupling between taxonomic and functional trajectories was mediated by the variations in functional redundancy, as the loss of a species does not necessarily entails the loss of its functional characteristics.

### 2020-03-30 WholePlotTrajectories.tex
STOTEN article version

--- 2019-08-16 WholePlotTrajectories.tex
@@
 \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
+\usepackage[left]{lineno} \linenumbers \usepackage{bookmark}
 
@@
 
-\PaperTitle{Post-Disturbance Tree Community Trajectories in a Neotropical Forest} % Article title
+\PaperTitle{Diverging taxonomic and functional trajectories following disturbance in
+a Neotropical forest} % Article title
 
@@
 trajectories, and determine the scope of the Intermediate Disturbance
-Hypothesis (IDH) that remains debated in tropical forests. We analyzed
-community trajectories following a disturbance gradient in a Neotropical
-forest over 30 years. We considered trajectories along time of community
-taxonomic and functional trajectories in terms of richness, evenness,
-composition, and redundancy. We based on the annual botanical
+Hypothesis (IDH) that remains debated in tropical forests.
+\color{black}We analyzed community trajectories following a disturbance
+gradient from 10 to 60\% of above-ground biomass loss in a Neotropical
+forest over 30 years.\color{red} We considered trajectories along time
+of community taxonomic and functional trajectories in terms of richness,
+evenness, composition, and redundancy. We based on the annual botanical
 inventories of 75 ha of a Neotropical forest and on large trait datasets
@@
 decoupling between taxonomic composition, differing among communities,
-and functional composition, remaining similar and convergent. The
-taxonomic diversity followed humped-shaped trajectories depending on the
-disturbance intensity, which validated the IDH (Intermediate Disturbance
-Hypothesis). The functional diversity trajectories,however, were
-homogeneous among all plots and dismissed the IDH. We explained this
-decoupling by the variations in community functional redundancy that
-mitigated the functional impact of disturbance. Although consistent, the
-recovery of community composition, diversity, and redundancy remained
-unachieved after 30 years. These results acknowledged the need of
+and functional composition, \color{red} similar among communities and
+convergent in the functional space.\color{black} \color{red} The
+taxonomic diversity followed humped-shaped trajectories along time after
+disturbance depending on the initial disturbance intensity, which
+validated the IDH (Intermediate Disturbance Hypothesis).\color{black}
+The functional diversity trajectories,however, were homogeneous among
+plots and dismissed the IDH. We explained this decoupling by the
+variations in community functional redundancy that mitigated the
+functional impact of disturbance. Although consistent, the recovery of
+community composition, diversity, and redundancy remained unachieved
+after 30 years. \color{red} These results acknowledged the need of
 decades-long cycles without disturbance to ensure community complete
-recovery, and questioned community resilience after repeated
-disturbances.
+recovery.\color{black}
 }
@@
 increase the pressure on remaining undisturbed forests
-\citep{Morales-Hidalgo2015}. These threats affect the natural
-disturbance regime that defines and maintains the structure,
-composition, and functioning of tree communities
-\citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}. To anticipate
-the fate of tropical forests, it is urgent to understand tree community
-response to disturbance, and the underlying ecological processes. The
-forest cover is generally maintained following disturbance, but
-modifications in the fluxes of light, heat, and water
+\citep{Morales-Hidalgo2015}. \color{red}These threats may change the
+frequency and magnitude of the natural disturbance regime that defines
+and maintains the structure, composition, and functioning of tree
+communities \citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}.
+\color{black} To anticipate the fate of tropical forests, it is urgent
+to understand tree community response to disturbance, and the underlying
+ecological processes. The forest cover is generally maintained following
+disturbance, but modifications in the fluxes of light, heat, and water
 \citep{Goulamoussene2017} change community abiotic and biotic
@@
 foster both competitively superior species and fast colonizers, and
-prevents competitive exclusion \citep{Shea2004, Pulsford2016}. In
-tropical forests, however, observations of the IDH often diverge from
-theoretical expectations \citep{Hugues2007, Sheil2003, Norden2017}, and
-the underlying processes might be complicated by the diversity of
-tropical tree communities \citep{Lindenmayer2012, Garcia_florez2017}. In
-this context, the IDH is controversial in tropical forests and remains
-to be tested \citep{Hubbell2001, Fox2013, Sheil2013}.
+prevents competitive exclusion \citep{Shea2004, Pulsford2016}.
+\color{red} In tropical forests, some studies advocate community
+deterministic response to disturbance, as for example in the dry and wet
+forests in Ghana \citep{Bongers2009} or in the guianese rainforests
+\citep{Molino2001}, where intermediate disturbance proved to increase
+the community richness by increasing the proportion of pioneers or
+heliophilous species in the community while maintaining the richness in
+old-growth forest species. In other cases, however, observations of the
+IDH diverge from theoretical expectations
+\citep{Hugues2007, Sheil2003, Norden2017}. Some studies refute the role
+of disturbance intensity, as observed in forest gaps in Panama where no
+difference in species richness was found between forest gaps and
+undisturbed areas \citep{Hubbell2001}, or in costa-rican forests where
+forests that did not follow deterministic convergence display stochastic
+trajectories \citep{Norden2015}.\\
+The high diversity of tropical forests may foster the emergence of
+numerous facilitation, adaptation, and inter- and intra-specific
+competition following disturbance
+\citep{Garcia_florez2017, Bongers2009}.These interactions can result in
+miscellaneous responses to disturbance \citep{Lindenmayer2012}, which
+question the validity of the IDH in tropical forests
+\citep{Hubbell2001, Fox2013, Sheil2013}. \color{black}
 
@@
 and functional facets of community diversity, and (iii) we analyzed
-community resilience and time to recovery.
+community resilience and time to recovery. \color{red} We hypothesized
+that community functional and taxonomic trajectories might not track
+each other, given the high functional redundancy in tropical forests.
+While functional diversity would be enhanced by the environmental
+changes following disturbance, specifically the increase in light
+availability, taxonomic trajectories would only increase until a
+disturbance intensity threshold, as pioneers recruited following
+disturbance are not as diverse as late-successional species. Community
+taxonomic and functional would hence follow different rates of recovery.
+\color{black}
 
@@
 
-The experiment is a network of twelve 6.25 ha plots (Table
-\ref{tab:Tab1}) that underwent three disturbance treatments in 1987
+The experiment is a network of twelve 6.25 ha plots\color{red},
+distributed over a 400ha area \color{black}(Table \ref{tab:Tab1}, Supp.
+Mat. Fig. S1), that undergone three disturbance treatments in 1987
 according to a randomized plot design \citep{Gourlet-Fleury2004}.\\
-The experiment comprised three replicates of three sylvicultural
-treatments (hereafter plots T1, T2, and T3), and three control plots
-(T0). All treatments T1, T2, and T3 comprised the logging of 10 trees/ha
-with 50 cm minimum DBH that belonged to a set of 58 commercially
-exploited species \citep{Gourlet-Fleury2004}. Treatment T2 additionally
-comprised a thinning treatment by poison-girdling of non-commercially
-exploited species randomly selected with an average of 30 trees/ha with
-40 cm minimum DBH. Treatment T3 additionally comprised the logging of 15
-trees/ha with 40 cm minimum DBH, and the poison-girdling of 20 trees/ha
-with a 50 cm minimum DBH, all belonging to non-commercially exploited
-species. Considering the silvicultural treatments and the following
-damage, disturbance intensity was measured as the percentage of
-aboveground biomass (\%AGB) lost between the first inventory in 1984 and
-five years after disturbance \citep{Piponiot2016} estimated with the
-BIOMASS R package \citep{Rejou2017}. The three treatments were then
-transformed into a continuous disturbance intensity gradient with
-increasing above-ground biomass (AGB) loss.
+\color{red}The experiment intended to simulate logging intensities and
+distinguished the treatments for non-commercial commercial species,
+logged at 40cm DBH to simulate thinning operations.\color{black} The
+experiment comprised three replicates of three sylvicultural treatments
+(hereafter plots T1, T2, and T3), and three control plots (T0).
+\color{red}Control plots are set in mature forest stand that did not
+undergone any logging nor intense anthropic activities activities for
+the last 500 years\color{black}. All treatments T1, T2, and T3 comprised
+the logging of 10 trees/ha with 50 cm minimum DBH that belonged to a set
+of 58 commercially exploited species \citep{Gourlet-Fleury2004}.
+Treatment T2 additionally comprised a thinning treatment by
+poison-girdling of non-commercially exploited species randomly selected
+with an average of 30 trees/ha with 40 cm minimum DBH. Treatment T3
+additionally comprised the logging of 15 trees/ha with 40 cm minimum
+DBH, and the poison-girdling of 20 trees/ha with a 50 cm minimum DBH,
+all belonging to non-commercially exploited species. Considering the
+silvicultural treatments and the following damage, disturbance intensity
+was measured as the percentage of aboveground biomass (\%AGB) lost
+between the first inventory in 1984 and five years after disturbance
+\citep{Piponiot2016} estimated with the BIOMASS R package
+\citep{Rejou2017}. \color{red} Tree diameter are provided in the
+experimental dataset, wood density is extracted from global wood density
+(GWD) database \citep{Zanne2009}, and tree height is retrieved from a
+generic H--D model based on a single bioclimatic predictor (eqn 6a in
+\citep{Chave2014}). The AGB is measured from the generalized allometric
+model eqn 4 of \citep{Chave2014}. \color{black} The three treatments
+were then transformed into a continuous disturbance intensity gradient
+with increasing above-ground biomass (AGB) loss.
 
@@
 The study site corresponds to a tropical rainforest typical of the
-Guiana Shield with a dominance of \emph{Fabaceae},
-\emph{Chrysobalanaceae}, \emph{Lecythidaceae}, and \emph{Sapotaceae}. In
-the 12 experimental plots, all trees above 10 cm DBH have been mapped
-and measured annually since 1984. Trees are first identified with a
-vernacular name assigned by the forest worker team, and afterward with a
-scientific name assigned by botanists during regular botanical
-campaigns. In 1984, specific vernacular names were given to 62 commmon
-or commercially exploited species. More infrequent species were
-identified under general identifiers only distinguishing trees and
-palms. From 2003, botanical campaigns have been conducted every 5 to 6
-years to identify all trees at the species level. In 2015, however,
-identification levels still varied among plots and campaigns.
+Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae,
+Lecythidaceae, and Sapotaceae. In the 12 experimental plots, all trees
+above 10 cm DBH have been mapped and measured annually since 1984. Trees
+are first identified with a vernacular name assigned by the forest
+worker team, and afterward with a scientific name assigned by botanists
+during regular botanical campaigns. In 1984, specific vernacular names
+were given to 62 commmon or commercially exploited species. More
+infrequent species were identified under general identifiers only
+distinguishing trees and palms. From 2003, botanical campaigns have been
+conducted every 5 to 6 years to identify all trees at the species level.
+In 2015, however, identification levels still varied among plots and
+campaigns.
 
@@
 for in the measure of composition and diversity metrics. Uncertainty
-propagation was implented withing a Bayesian framework using
+propagation was implented within a Bayesian framework using
 vernacular/botanical names associations to reconstitute complete
@@
 
-See Supplementary Materials -Fig. S1 and \citet{Aubry-Kientz2013} for
+See Supplementary Materials -Fig. S2 and \citet{Aubry-Kientz2013} for
 the detailed methodology.
@@
 values were filled using multivariate imputation by chained equation
-\citep{Mice2011}. To account for the phylogenetic signal in the filling
-process, imputations were based on samples of species from the same
-genus or from the same family. Whenever a species was not in the
-dataset, it was attributed a set of trait values randomly sampled among
-species of the next higher taxonomic level (same genus or family). Two
-life-history traits, maximum specific height and seed mass, came from
-the Mariwenn database \footnote{https://www.ecofog.gf/mariwenn/}. The
-database compiles information from a vast literature on the flora of
+\citep{Mice2011}. To account for the phylogenetic signal in the
+imputation process, imputations were based on samples of species from
+the same genus or from the same family. Whenever a species was not in
+the dataset, it was attributed a set of trait values randomly sampled
+among species of the next higher taxonomic level (same genus or family).
+Two life-history traits, maximum specific height and seed mass, came
+from the Mariwenn database \footnote{https://www.ecofog.gf/mariwenn/}.
+The database compiles information from a vast literature on the flora of
 French Guiana \citep{Ollivier2007} and comprises 362 species belonging
@@
 
-All composition and diversity metrics were obtained after 50 iterations
-of the uncertainty propagation framework.
+\color{red}Confidence intervals of the composition and diversity metrics
+were computed after 50 iterations of the uncertainty propagation
+framework.\color{black}
 
@@
 we had to conduct the taxonomic composition and diversity analysis at
-the genus level. Taxonomic and functional trajectories of community
-composition were drawn in a two-dimensional NMDS ordination plane. Two
-NMDS using abundance-based (\emph{i.e.} Bray-Curtis) dissimilarity
-measures were conducted to map either taxonomic or functional
-composition, the latter based on the seven leaf, stem, and life history
-traits (without seed mass classes). Trajectories along time were
-reported through the Euclidean distance between the target inventories
-and the 1984 pre-disturbance inventories of reference. Univariate
-trajectories of the leaf, stem, and life-history traits were also
-visualized with the community weighted means (CWM) \citep{Diaz2007}.
-Species seed mass were given in 5 mass classes, and seed mass
-trajectories were reported as the proportion of each class in the
-inventories (Supplementary materials).
+the genus level. \color{red}Functional composition, diversity, and
+redundancy measures were conducted at species level.\color{black}
+Taxonomic and functional trajectories of community composition were
+drawn in a two-dimensional NMDS ordination plane. Two NMDS using
+abundance-based (\emph{i.e.} Bray-Curtis) dissimilarity measures were
+conducted to map either taxonomic or functional composition, the latter
+based on the seven leaf, stem, and life history traits (without seed
+mass classes). Trajectories along time were reported through the
+Euclidean distance between the target inventories and the 1984
+pre-disturbance inventories of reference. Univariate trajectories of the
+leaf, stem, and life-history traits were also visualized with the
+community weighted means (CWM) \citep{Diaz2007}. Species seed mass were
+given in 5 mass classes, and seed mass trajectories were reported as the
+proportion of each class in the inventories (Supplementary materials).
 
@@
 Both indices are recommended for diversity studies \citep{Marcon2015},
-and belong to the set of HCDT or generalized entropy corresponding,
-respectively, to the 0 and 2 order of diversity (q). The functional
-diversity was reported using the functional richness and functional
-evenness, through the Rao index of quadratic entropy. The Rao index
-combines species abundance distribution, and the average pairwise
-functional dissimilarity between species computed by the Gower distance.
+and belong to the set of \color{red} Tsallis or generalized entropy
+\color{black} corresponding, respectively, to the 0 and 2 order of
+diversity (q). \color{red}Diversity metrics were measured using the Hqz
+function from the entropart package \citep{Marcon2015}.\color{black} The
+functional diversity was reported using the functional richness and
+functional evenness. \color{red}Both functional richness and evenness
+were measured through a distance-based dissimilarity matrix built with
+the daisy function of the cluster R package \citep{Cluster2019}, using
+the Gower distance between species, similar to \citet{Herault2007a},
+\citet{Herault2007b}.\color{black} The functional diversity was measured
+through the Rao index of quadratic entropy. The Rao index combines the
+community abundance distribution\color{red}, from species relative
+abundance, \color{black} and the average pairwise functional
+dissimilarity between species computed by the Gower distance.
+\color{red}The comparison between these two metrics assesses community
+functional richness and evenness: thereafter, results will be discussed
+directly in terms of functional richness and evenness.\color{black}
 
@@
 correlation tests. Richness and evenness trajectories were besides
-analyzed through polynomial regression between (i) taxonomic and
-functional richness and evenness, and (ii) the initial \%AGB loss at 10,
-20, and 30 years after disturbance.
+analyzed through polynomial regression \color{red}or simple linear
+models \color{black} between (i) taxonomic and functional richness and
+evenness, and (ii) the initial \%AGB loss at 10, 20, and 30 years after
+disturbance. \color{red}The best model fitting, between linear or
+quadratic model, was chosen based on their respective AIC criterion. The
+explanatory power of the fit were measured to adjusted r-square
+values.\color{black}
 
@@
 correlations among traits as the axes are combinations of the most
-decoupled traits. For each species, the traits probability density (TPD)
-were computed from the mapping of individuals through two-dimension
-kernel density estimators. Second, for each community, the TPD weigthed
-by species abundance were summed accross the functional space. Third,
-the functional space was divided into a 100 x 100 grid, and the number
-of species with a positive TPD was counted in each cell. The average
-count across cells minus 1 returned the Community Functional Redundancy,
-which was the average number of species in the community that share the
-same trait values.
+decoupled traits. \color{red}The two main PCA axis respectively hold
+26\% and 19\% of the trait dataset variance.\color{black} For each
+species, the traits probability density (TPD) were computed from the
+mapping of individuals through two-dimension kernel density estimators.
+\color{red} We used the kde2d function of the MASS package
+\citep{MASS2002} that is a gaussian density estimator for a
+2-dimensional space. \color{black} Second, for each community, the TPD
+weigthed by species abundance were summed across the functional space.
+Third, the functional space was divided into a 100 x 100 grid, and the
+number of species with a positive TPD was counted in each cell. The
+average count across cells minus 1 returned the Community Functional
+Redundancy, which was the average number of species in the community
+that share the same trait values.
 
@@
 disturbance intensity for both taxonomic and functional composition
-(\(\rho_{Spearman}^{Taxonomic}=0.87\) and
-\(\rho_{Spearman}^{Functional}=0.90\), respectively). The maximum
-dissimilarity with the initial state was reached for taxonomic
-composition between 15 to 25 years, and between 20 to 25 years for
-functional composition.
+(\(\rho_{Spearman}^{Taxonomic}=0.94\) and
+\(\rho_{Spearman}^{Functional}=0.90\), respectively. Fig.
+\ref{fig:SumUp}). The maximum dissimilarity with the initial state was
+reached for taxonomic composition between 15 to 25 years, and between 20
+to 25 years for functional composition.
 
@@
 of leaf chlorophyll content, which continued to increase 30 years after
-disturbance for 4 out of 6 highly disturbed plots. Maximum height at
-adult stage (\emph{Hmax}), leaf toughness, and wood specific gravity
-(\emph{WSG}) decreased in time and then slightly increased, but remained
-significantly lower than their initial value (Fig. \ref{fig:CWM}). Bark
-thickness and specific leaf area (\emph{SLA}) both increased along time.
-Bark thickness remained substantially high after 30 years, and
-\emph{SLA} had almost recovered to its initial value. Whatever the
-functional traits, the maximum difference to initial value was highly
-correlated to the disturbance intensity. Positive correlations were
-observed for leaf thickness, chlorophyll content, SLA and bark thickness
-(\(\rho_{Spearman}^{Leaf thickness}=0.76\),
-\(\rho_{Spearman}^{Chlorophyll content}=0.60\),
-\(\rho_{Spearman}^{SLA}=0.93\),
-\(\rho_{Spearman}^{Bark thickness}=0.71\)). Negative correlation was
-observed for leaf toughness, WSG, and Hmax
-(\(\rho_{Spearman}^{Leaf toughness}=-0.53\),
-\(\rho_{Spearman}^{WSG}=-0.75\), \(\rho_{Spearman}^{Hmax}=-0.40\)) The
-proportions of the three lightest seed mass classes increased in all
-disturbed plots. After 30 years the proportion of lightest seed mass
-class decreased while it stabilized for the two other lightest seed mass
-classes (Supp. Mat. - Fig. S2).
+disturbance for 4 out of 6 highly disturbed plots. \color{red}Main
+tendencies for disturbed plots were a decrease followed by a slight
+increase of maximum height at adult stage (\emph{Hmax}), leaf toughness,
+and wood specific gravity (\emph{WSG}), but trait values remained
+significantly lower than their initial value (Fig. \ref{fig:CWM}).
+\color{black} Bark thickness and specific leaf area (\emph{SLA}) both
+increased along time. Bark thickness remained substantially high after
+30 years, and \emph{SLA} had almost recovered to its initial value.
+Whatever the functional traits, the maximum difference to initial value
+was highly correlated to the disturbance intensity. \color{red}Positive
+correlations were observed for leaf thickness, chlorophyll content, SLA
+and bark thickness (\(\rho_{Spearman}^{Leaf thickness}=0.79\),
+\(\rho_{Spearman}^{Chlorophyll content}=0.61\),
+\(\rho_{Spearman}^{SLA}=0.99\),
+\(\rho_{Spearman}^{Bark thickness}=0.68\)). Maxima were observed between
+5 and 15 years following disturbance for leaf thickness and SLA while
+chlorophyll content and bark thickness kept increasing after 30 years
+for leaf chlorophyll content and bark thickness. Negative correlation
+was observed for leaf toughness, WSG, and Hmax
+(\(\rho_{Spearman}^{Leaf toughness}=-0.47\),
+\(\rho_{Spearman}^{WSG}=-0.80\), \(\rho_{Spearman}^{Hmax}=-0.28\)).
+Minima were observed between 5 and 15 years following disturbance for
+Hmax, 10 and 20 years for WSG, and between 15 and 25 years following
+disturbance for leaf toughness.\color{black} The proportions of the
+three lightest seed mass classes increased in all disturbed plots. After
+30 years the proportion of lightest seed mass class decreased while it
+stabilized for the two other lightest seed mass classes (Supp. Mat. -
+Fig. S3).
 
@@
 ten years after disturbance before recovering to pre-disturbance values.
-The maximum richness loss or gain after disturbance was positively
-correlated with the disturbance intensity
-(\(\rho_{Spearman}^{Richness}=0.50\)).
+\color{red}The maximum rate of change compared to initial richness was
+positively correlated with the disturbance intensity
+(\(\rho_{Spearman}^{Richness}=0.73\)).\color{black}
 
-In all disturbed plots the evenness first increased until a maximum
-reached after around 20 years. This maximum was positively correlated
-with the disturbance intensity (\(\rho_{Spearman}^{Simpson}=0.77\)). The
-evenness then stabilized except for two intensively-disturbed plots
-(number 8 and 12) for which it kept increasing (Fig. \ref{fig:DivTaxo}).
+In all disturbed plots \color{red}, to the exception of plots 5 (T2) and
+12 (T3) that showed a small initial drop, \color{black} the evenness
+first increased until a maximum reached after around 20 years. This
+maximum was positively correlated with the disturbance intensity
+(\(\rho_{Spearman}^{Evenness}=0.82\), \color{red} Fig.
+\ref{fig:SumUp})\color{black}. The evenness then stabilized except for
+two intensively-disturbed plots (number 8 and 12) for which it kept
+increasing (Fig. \ref{fig:DivTaxo}).
 
@@
 
-\caption{Trajectories of community taxonomic richness \textbf{(a)}, Simpson diversity \textbf{(b)}, functional richness \textbf{(c)}, and Rao diversity \textbf{(d)}.
-Values correspond to the difference over 30 years of community diversity with the values of 1984 pre-disturbance inventories of reference.
+\caption{Trajectories of community taxonomic richness \textbf{(a)}, taxonomic evenness, \textbf{(b)}, functional richness \textbf{(c)}, and functional evenness \textbf{(d)}.
+Values correspond to the difference over 30 years of community diversity with the values of 1989 inventories of reference, 5 years after disturbance.
 Shaded areas are the credibility intervals }\label{fig:DivTaxo}
@@
 Functional richness and evenness remained stable in control plots over
-the 30 years of monitoring. In disturbed communities, both trajectories
-depended on the disturbance intensity, with their maximum values in time
-being positively correlated to \%AGB loss
-\(\rho_{Spearman}^{Richness} = 0.76\) and\\
-\(\rho_{Spearman}^{Rao} = 0.60\). For low disturbance intensity,
-functional richness and evenness displayed a low but long-lasting
-increase up to a maximum reached after 20-25 years. For high disturbance
-intensity, they generally displayed a fast but short increase followed,
-after 10 years, by a slow decrease towards the initial values.
+the 30 years of monitoring. \color{red}In disturbed communities, both
+functional richness and evenness trajectories increased following
+disturbance, with maxima positively correlated to \%AGB loss
+\(\rho_{Spearman}^{Richness} = 0.76\) and
+\(\rho_{Spearman}^{Evenness} = 0.60\) (Fig. \ref{fig:SumUp}).
+\color{black} For low \color{red}and intermediate
+\color{black}disturbance intensity, functional richness and evenness
+displayed a low but long-lasting increase up to a maximum reached after
+20-25 years. For high disturbance intensity, they generally displayed a
+fast but short increase followed, after 10 years, by a slow decrease
+towards the initial values.
 
@@
 shapes depending on the diversity indices and on the time since
-disturbance (Fig. \ref{fig:IDHplot}). Regarding taxonomic diversity, the
-relationship between disturbance intensity and diversity was more
-markedly hump-shaped for richness than for evenness, and peaked at 20\%
-of initial AGB loss. Regarding functional diversity, the relationship
-was almost linear, and was similar between richness and evenness. All
-the relationships between disturbance intensity and diversity were
-stronger 20 or 30 years after disturbance than they were 10 years after
-disturbance.
+disturbance (Fig. \ref{fig:IDHplot}). \color{red}Regarding taxonomic
+diversity, the taxonomic richness showed a humped-shaped trajectory with
+the disturbance intensity, and peaked at 20\% of initial AGB loss. The
+relationship between disturbance intensity and taxonomic evenness,
+however, was humped-shaped only 20 years after disturbance and linear
+for 10 and 30 years after disturbance times.\color{black} Regarding
+functional diversity, \color{red} the relationship was linear for both
+taxonomic richness and evenness, except for the functional diversity
+after 30 years. \color{black}
 
@@
 disturbed plots had lower functional redundancy than control plots and
-followed similar hump-shaped trajectories (Fig. \ref{fig:RedFunRest}).
+followed similar humped-shaped trajectories (Fig. \ref{fig:RedFunRest}).
 The maximum redundancy loss was positively correlated with the
-disturbance intensity (\(\rho_{Spearman}=0.47\)) and the recovery
-trajectory had not attained initial values for any disturbed communities
-after 30 years.
+disturbance intensity (\(\rho_{Spearman}=0.31\), (Fig. \ref{fig:SumUp}))
+and the recovery trajectory had not attained initial values for any
+disturbed communities after 30 years.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/RedFunRest-1}
+{\centering \includegraphics[width=0.4\linewidth]{WholePlotTrajectories_files/figure-latex/RedFunRest-1}
 
@@
 
-\caption{Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance. Shaded areas are the credibility intervals.}\label{fig:RedFunRest}
+\caption{Trajectories of the functional redundancy within the initial functional space over 30 years after disturbance.
+The redundancy is measured in a two-dimensional plan summarizing 45\% of species variance in the functional space of the 7 leaf, stem, and life history traits.
+Redundancy value represent the average number of species in the community that share the same trait values.
+Shaded areas are the credibility intervals.}\label{fig:RedFunRest}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{WholePlotTrajectories_files/figure-latex/SumUp-1}
+
+}
+
+\caption{Maximum changes in taxonomic and functional diversity and composition against disturbance intensity, expressed as percentage of initial disturbance.
+Upper panels display taxonomic richness and evenness \textbf{(a)}, and functional richness and evenness \textbf{(b)} expressed in percentage of change compared to initial value.
+Lower panels display taxonomic and functional euclidean distance from pre-disturbance inventory, respectively in the spaces of species and functional traits \textbf{(c)}, and functional redundancy expressed as number of redundant species in the community \textbf{(d)}.}\label{fig:SumUp}
 \end{figure}
@@
 
-Our analysis revealed the decoupling between functional and taxonomic
-trajectories. While the initial differences in taxonomic composition
-among plots were maintained, the functional composition trajectories
-converged in the functional space. In terms of diversity, only the
-taxonomic trajectories validated the IDH that explained humped-shaped
-post-disturbance trajectories whith an amplitude depending on the
-disturbance intensity. The decoupling between taxonomic and functional
-response was explained by variations of community functional redundancy
-that mitigated the functional impact of disturbance, and appeared as a
-determinant of community recovery.
+\color{red}Our analysis revealed that community post-disturbance
+trajectories in taxonomic and functional diversity and composition were
+decoupled. The taxonomic composition specificities between communities
+before disturbance were maintained following disturbance, while their
+functional composition converged in the functional space. \color{black}
+In terms of diversity, only the taxonomic trajectories validated the IDH
+that explained humped-shaped post-disturbance trajectories whith an
+amplitude depending on the disturbance intensity. The decoupling between
+taxonomic and functional response \color{red} could be
+\color{black}explained by variations in community functional redundancy.
+\color{red}The loss of species or the changes in their abundance would
+impact the taxonomic but not the functional diversity and composition,
+as redundant species with the same functional characteristics remain in
+the community. The trajectory of the functional redundancy hence
+appeared determinant for community recovery.\color{black}
 
@@
 
-Community taxonomic composition substantially differed before
-disturbance, as materialized by their distinct location on the NMDS axis
-2. These initial differences were maintained along the 30 years
-following disturbance, with the disturbance leading a displacement on
-the NMDS axis 1 only. Taxonomic composition changes were similar among
-plots and may correspond to the recruitment of a group of pioneers, like
-\emph{Cecropia spp.} or \emph{Miconia spp.}, common to all plots,
-whatever their initial taxonomic differences and the intensity of
-disturbance \citep{Denslow2000, Bongers2009}. Taxonomic trajectories
+\color{red}From pre-disturbance to 30 years after disturbance,
+communities showed different location along NMDS axis 2 that hardly
+changed along time. Specificities in community taxonomic composition,
+materialized by the distinct location on the NMDS axis 2, existed before
+disturbance. The disturbance led a displacement on the NMDS axis 1
+only.\color{black} Taxonomic \color{red} post-disturbance \color{black}
+changes in composition were similar among plots and may correspond to
+the recruitment of a group of pioneers, like \emph{Cecropia spp.} or
+\emph{Miconia spp.}, common to all plots, whatever their initial
+taxonomic \color{red}specificities \color{black} and the intensity of
+the disturbance \citep{Denslow2000, Bongers2009}. Taxonomic trajectories
 initiated a recovery towards the initial composition. This recovery,
 although far from being achieved after 30 years, suggested the
-resilience of community taxonomic composition and the maintenance of
-initial composition differences \citep{Folke2006}. Such resilience
-suggested that species not belonging to the pre-disturbance community
-were rarely recruited in the long-term, probably because of species
-dispersal limitation that is common among tropical species
-\citep{Svenning2005}.
+resilience of community taxonomic composition. \color{red}Although the
+composition dramatically changed following disturbance, community
+specificities were maintained \citep{Folke2006}. This maintenance
+suggested that the recruitment came from (i) a common set of pioneer and
+light-demanding species in all plots that shape the similar trajectories
+on NMDS axis 1 and (ii) a very local set of late-successional species
+that signs (and maintains in time) the initial position of each plot on
+NMDS axis 2. This can be related to species dispersal limitation that is
+common among tropical species \citep{Svenning2005}.\color{black}
 
 Community functional composition trajectories, in contrast, were similar
-in the functional space \citep{Fukami2005}. The amplitude of the
+in the functional space. \color{red}Such functional convergence
+contrasting with community taxonomic divergence was already observed in
+plant communities \citep{Fukami2005}.\color{black} The amplitude of the
 compositional changes depended on the disturbance intensity. Functional
@@
 the first NMDS axis \citep{Westoby1998, Wright2004, Reich2014}.
-Thereafter, the first recruited pioneers were progressively excluded by
-long-lived, more competitive, and shade-tolerant species. The
-recruitment of these late-successionals marked the recovery of the
-initial functional composition with more ``resource-conservative''"
-strategies . This recovery translated in the functional plane by a
-displacement left along the first axis and upward along the second axis
-(Fig. \ref{fig:NMDSplans}).
+\color{red}This recruitment went along with an increase of the mean SLA,
+and a decrease in the mean WSG in disturbed communities (\ref{fig:CWM}).
+These fast and significant changes in mean functional traits closely
+linked to species light acquisition \citep{Wright2004, Chave2009b}
+suggested major changes in the light environment following disturbance
+that shape species assemblage in tropical forests
+\citep{Pena2008, Carreno-Rocabado2012}. Thereafter, the first recruited
+pioneers were progressively excluded by long-lived, more competitive,
+and shade-tolerant species. The recruitment of these late-successional
+species marked the recovery of the initial functional composition with
+more ``resource-conservative''" strategies, corresponding to a fast
+decrease in community mean SLA and a stabilisation of community WSG,
+suggesting the progressive closing of forest canopy. \color{black} This
+recovery translated in the functional plane by a displacement left along
+the first axis and upward along the second axis (Fig.
+\ref{fig:NMDSplans}).
 
@@
 functional space, and historically contingent in the taxonomic space.
+
+\color{red} We must temper these results with the low number of plots
+(3) for each treatment. However, the size of each plot is 6.25ha that
+has been regularly censused over 25 years so that the trajectories have
+been drawn using 75ha of permanent plot information, a spatial hold very
+rarely achieved in such studies. Given the relative homogeneity of
+environmental characteristics of terra firme forests over the Guiana
+Shield and even the amazonian basin, it seems reasonable that similar
+trajectories would apply to neighboring areas
+\citep{Guitet2015, Guitet2018}. \color{black}
 
@@
 of pioneers, previously infrequent or absent, increased the taxonomic
-richness \citep{Martin2015, Chaudhary2016}, and that trees surviving
-after disturbance remained numerous enough to maintain the richness of
-the pre-disturbance community \citep{Bongers2009}. The pioneers thus
-recruited became abundant, which balanced the usual hyper-dominance of
-tropical forests and increased the taxonomic evenness
-\citep{Baraloto2012a}. Above the intensity threshold, like for treatment
-3 (high intensity), the taxonomic richness did not exceed the initial
-value in the first years following disturbance. No increase of the
-taxonomic richness suggested that the richness of surviving trees was
-lower than this of the pre-disturbance community, and that this
-difference was not offset by the recruitment of pioneers. In the Guiana
-shield, the pool of true pioneers specifically recruited after
-disturbance is restricted to a few common genera (e.g. \emph{Cecropia
-spp.}, \emph{Miconia spp.}, \emph{Tapirira spp.}) \citep{Guitet2018}.
-Considering the different times after disturbance, there was always a
-humped-shaped pattern linking the disturbance intensity with the
-post-disturbance increase in taxonomic richness and evenness (Fig.
-\ref{fig:IDHplot}). Both taxonomic richness and evenness were maximized
-at an intermediate intensity, around 20-25\% of AGB lost.
+richness, and that trees surviving after disturbance remained numerous
+enough to maintain the richness of the pre-disturbance community
+\citep{Bongers2009}. The pioneers thus recruited became abundant, which
+balanced the usual hyper-dominance of tropical forests and increased the
+taxonomic evenness \citep{Baraloto2012a}. Above the intensity threshold,
+like for treatment 3 (high intensity), the taxonomic richness did not
+exceed the initial value in the first years following disturbance. No
+increase of the taxonomic richness suggested that the richness of
+surviving trees was lower than this of the pre-disturbance community,
+and that this difference was not offset by the recruitment of pioneers.
+In the Guiana shield, the pool of \color{red}hard \color{black} pioneers
+specifically recruited after disturbance is restricted to a few common
+genera (e.g. \emph{Cecropia spp.}, \emph{Miconia spp.}, \emph{Tapirira
+spp.}) \citep{Guitet2018}. \color{red}The disturbance intensity showed a
+humped-shaped relationship with the post-disturbance increase in
+taxonomic richness at all the different times after disturbance, and
+with the increase in taxonomic evenness 10 years after disturbance (Fig.
+\ref{fig:IDHplot}).\color{black} Taxonomic richness, and to some extend
+taxonomic evenness, were maximized at an intermediate intensity, around
+20-25\% of AGB lost.
 
-Regarding community functional trajectories, however, no marked
-differences were observed among post-disturbance trajectories (Fig.
-\ref{fig:DivTaxo}). Whenever the disturbance intensity, there was first
-an increase of both functional richness and evenness. Such increase
-suggested the recruitment of pioneers that were functionally highly
-different from the pre-disturbance community
-\citep{Denslow1980, Molino2001}. The recruitment of pioneers hampered
-the recruitment of other species in the first place. After 15 to 20
-years, the first pioneer recruits declined and were replaced by species
-functionally more similar to the pre-disturbance community, which
-decreased the functional richness and evenness \citep{Walker2009}.
+\subsection{The functional redundancy explaining the
+taxonomic-functional
+decoupling}\label{the-functional-redundancy-explaining-the-taxonomic-functional-decoupling}
 
-\subsection{The functional redundancy, key of community
-resilience}\label{the-functional-redundancy-key-of-community-resilience}
+\color{red}Regarding community functional trajectories (Fig.
+\ref{fig:IDHplot}), there was no intensity threshold above which the
+diversity-disturbance relationship changed. The functional diversity
+kept increasing along with the disturbance intensity, hence the IDH
+would not apply. Surprisingly, although some species were lost after
+disturbance, the functional diversity did not decrease in the first
+place. On the contrary, functional diversity increased, probably
+following the rapid recruitment of pioneers that were functionally
+highly different from the pre-disturbance community
+\citep{Denslow1980, Molino2001}.
 
-The loss of species following disturbance decreased community functional
-redundancy during the first 15 years. Progressively though, the
-functional redundancy was restored through the replacement of the
-species with ``resource-acquisitive''" strategies by more
-late-successional species, functionally closer to the pre-disturbance
-community. The replacement of pioneers by late-successionals followed
-the lottery recruitment rules. This rules mean an easy recruitment of
-the first species that becomes increasingly difficult, as the following
-species are hampered by the emergence of interspecific competition
-\citep{Busing2002}. The increasing competition among species explained
-why the recovery trajectory slowed down 20 years after disturbance. The
-recovery of the functional redundancy then relied upon the random
-process of species recruitment, and was increasingly slow and difficult
-to anticipate \citep{Elmqvist2003, Diaz2005}. This suggested a low
-resilience of the functional redundancy with the random recovery of
-infrequent species increasing the risks of losing keystone species, with
-unexpected ecological consequences
-\citep{Jones1994, Chazdon2003a, Diaz2005}. Infrequent species might
-indeed have unique functional characteristics in the ecosystem, apart
-from the ones considered here, in the ecosystem or be a key resource for
-some of the fauna \citep{Schleuning2016}.
+The loss of species following disturbance, however, decreased the
+functional redundancy, all the more so that disturbance was intense. All
+in all, because functional diversity was not lowered by disturbance
+while functional redundancy was, this means that the species that were
+lost in disturbance are, on the whole, functionally equivalent to the
+species that survive the disturbance. This makes functional redundancy a
+key to understand community dynamics at a functional level. In other
+words, the high redundancy of tropical forest mean that several species
+occupy the same functional space, so changes in taxonomic diversity or
+composition do not necessarily result in changes in community functional
+characteristics because species are commutable. The loss of a redundant
+species doesn't change anything for the community functional
+characteristics, so functional trajectories do not necessarily track
+taxonomic ones. This commutability explains the taxonomic-functional
+decoupling and the fact that the taxonomic diversity-disturbance
+relationship supports the IDH while it functional diversity do not.\\
+Following disturbance, the redundancy was progressively restored through
+the replacement of ``resource-acquisitive'' species by more
+late-successional ``resource-conservative'' species functionally closer
+to the pre-disturbance community. The time for the functional redundancy
+to recover would hence be important to assess community recovery.
+\color{black}
 
@@
 Post-disturbance trajectories of tree community composition and
-diversity were driven by the recruitment of a determined pool of
+diversity would be driven by the recruitment of a determined pool of
 pioneers, identical among local communities, and independent of the
@@
 the recovery of the initial composition. The diversity trajectories were
-contrasted as well. While the functional trajectories remained similar
-whatever the disturbance intensity, taxonomic trajectories were markedly
-different from a threshold of 20-25\% AGB lost that maximized the
-taxonomic richness. The Intermediate Disturbance Hypothesis applied well
-to taxonomic diversity, but not to functional diversity. The decoupling
-between taxonomic and functional trajectories was mediated by the
-variations in functional redundancy, as the loss of a species does not
-necessarily entails the loss of its functional characteristics.
+contrasted as well. While the functional trajectories \color{red}
+displayed the same tendency with a linear increase following all
+disturbance intensity, \color{black} taxonomic trajectories were
+markedly different after a threshold of 20-25\% AGB lost that maximized
+the taxonomic richness. The Intermediate Disturbance Hypothesis would
+apply well to taxonomic diversity, but not to functional diversity. The
+decoupling between taxonomic and functional trajectories was mediated by
+the variations in functional redundancy, as the loss of a species does
+not necessarily entails the loss of its functional characteristics.
 Community resilience, in terms of recovery of the pre-disturbance state,
@@
 \makeatletter
-% The filename has .bib extension the must be eliminated
+% The filename has .bib extension that must be eliminated
 \filename@parse{references.bib}

### 2020-03-30 WholePlotTrajectories_SuppMat.Rmd
STOTEN article version

@@
 output: 
+  bookdown::word_document2: default
+  bookdown::gitbook: default
   bookdown::pdf_book:
@@
 (ref:Boldc) Plot of the 7 functional traits sed in the PCA analysis (Leaf and stem traits, and maximum species height).", display=FALSE)
+figures("ParacouMap","Map of the Paracou experimental site with the 12 studied plots. Square colors indicate the silvicultural treatment applied in 1984.")
 ```
@@
 
-**Figure S1**: Whenever the family or genus were recorded for a vernacular name, the pool of species sampled to replace the vernacular name was restricted to those pertaining to the same higher taxonomic level (i.e species of the same genus when the species is missing, species of the same family when the genus is missing) \@ref(fig:FigBayesianScheme).
+**Figure S1**: Map of the experimental design
+
+```{r ParacouMap,out.width='100%',fig.cap=figures("ParacouMap"),echo=F, fig.env='figure',fig.pos='h'}
+knitr::include_graphics("ExternalFig/ParacouMap.png")
+```
+
+**Figure S2**: Whenever the family or genus were recorded for a vernacular name, the pool of species sampled to replace the vernacular name was restricted to those pertaining to the same higher taxonomic level (i.e species of the same genus when the species is missing, species of the same family when the genus is missing) \@ref(fig:FigBayesianScheme).
 
@@
 
-**Figure S2**: Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
+**Figure S3**: Trajectories of seed mass classes proportions over 30 years after disturbance, corresponding to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation.
 No gap filling process was applied in this case.
@@
 
-**Figure S3**: Functional PCA
+**Figure S4**: Functional PCA
 

### 2020-03-30 WholePlotTrajectories_SuppMat.tex
STOTEN article version

@@
 
-\textbf{Figure S1}: Whenever the family or genus were recorded for a
+\textbf{Figure S1}: Map of the experimental design
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\linewidth]{ExternalFig/ParacouMap}
+
+}
+
+\caption{Figure S4: Map of the Paracou experimental site with the 12 studied plots. Square colors indicate the silvicultural treatment applied in 1984.}\label{fig:ParacouMap}
+\end{figure}
+
+\textbf{Figure S2}: Whenever the family or genus were recorded for a
 vernacular name, the pool of species sampled to replace the vernacular
@@
 
-\textbf{Figure S2}: Trajectories of seed mass classes proportions over
+\textbf{Figure S3}: Trajectories of seed mass classes proportions over
 30 years after disturbance, corresponding to the median (solid line) and
@@
 
-\textbf{Figure S3}: Functional PCA
+\textbf{Figure S4}: Functional PCA
 
@@
 \makeatletter
-% The filename has .bib extension the must be eliminated
+% The filename has .bib extension that must be eliminated
 \filename@parse{}

### 2020-03-30 Scripts/AGBloss.R

@@ -31,6 +31,11 @@ names(Carr)<-paste(plo,unique(Plo[,"n_carre"]),sep=".")
 return(Carr)
 }))
 
+AGBi<-unlist(lapply(1:12,function(plo){
+  Plo<-init[which(init[,"n_parcelle"]==plo),]
+    return(sum(computeAGB(D=Plo[,"circonf"],WD=rep(wd,nrow(Plo)),coord=c(mean(Plo[,"Lon"]),mean(Plo[,"Lat"])))))
+  }))
+
 AGBp<-unlist(lapply(1:12,function(plo){
   Plo<-logg[which(logg[,"n_parcelle"]==plo),]
   Carr<-unlist(lapply(unique(Plo[,"n_carre"]),function(car){
@@ -48,5 +53,13 @@ c(1,6,11,2,7,9,3,5,10,4,8,12)
 AGBloss<-cbind(rep(1:12,rep(4,12)),rep(1:4,12),AGBloss)
 colnames(AGBloss)<-c("plot","carre","AGBloss")
 
-save(AGBloss,file="DB/LostAGBcarre")
+save(AGBloss,file="DB/LostAGB")
+
+load("DB/LostAGB")
+AGBloss_cor<-cbind(c(AGBloss[1:3,"AGB"],24.9,18.6,20.8,52.8,41,47.8,51,58.8,48.4),AGBloss[,"plot"])
+colnames(AGBloss_cor)<-c("AGB","plot")
+save(AGBloss_cor,file="DB/LostAGB")
+
+
+
 

### 2020-03-30 Scripts/DivTaxo_trajectories.R

@@ -82,3 +82,43 @@ CompleteTaxo<-lapply(1:4,function(t){
 
 save(CompleteTaxo,file = "DB/ReplacementTraj_ForGraphs")
 
+
+
+
+EuclidDist<-function(Distances){
+  colnames(Distances)<-as.numeric(colnames(Distances))-1986
+  ret<-lapply(c(0.025,0.5,0.975),function(quant){
+    return(apply(Distances,c(1,2),function(col){
+      return(quantile(col,probs=quant))}))})
+  names(ret)<-c(0.025,0.5,0.975)
+  
+  plot(colnames(ret[[2]]),ret[[2]][1,],type="n",xlab="",ylab="",xaxt="n",
+       ylim=c(min(unlist(ret)),max(unlist(ret))),cex.axis=0.7)
+  axis(1,at=as.numeric(colnames(ret[[1]])),labels=TRUE)  
+  invisible(lapply(1:length(treatments),function(tr){
+    Toplot<-lapply(treatments[[tr]],function(plo){return(do.call(rbind,lapply(ret,function(quant){return(quant[plo,])})))})
+    invisible(lapply(Toplot,function(plo){
+      lines(colnames(plo),plo["0.5",],col=ColorsTr[[tr]],lwd=2)
+      polygon(c(colnames(plo),rev(colnames(plo))),c(plo["0.025",],rev(plo["0.975",])),
+              col=rgb(0,0,0,alpha=0.1),border=NA)
+    }))
+  }))
+}
+
+load("DB/FunctionalComposition_forGraphs");load("DB/TaxoComposition_forGraphs")
+load("DB/FunDistance_ForGraphs");load("DB/TaxoDistance_ForGraphs")
+
+windows()
+par(mfrow=c(2,2),mar=c(2,2.5,2.5,1),oma=c(1,2,1,4),no.readonly = T)
+TaxoCompo(MatrepTaxo)
+mtext("Taxonomic composition",side=3,line=1.8,cex=1.05)
+FunCompo(MatrepFun)
+mtext("Functional composition",side=3,line=1.8,cex=1.05)
+EuclidDist(TaxoEuclid)
+mtext("Euclidean distance\nfrom 1984 inventory",side=2,padj=0,line=2,cex=0.8)
+mtext("(c)",side=3,adj=0,line=0.5)
+EuclidDist(FunEuclid)
+mtext("(d)",side=3,adj=0,line=0.5)
[…]

### 2020-03-30 Scripts/DrawingTools.R

@@ -18,6 +18,7 @@ matT$plot<-substr(rownames(matT),start=1,stop=regexpr("_",rownames(matT))-1)
 matT$year<-substr(rownames(matT),start=regexpr("_",rownames(matT))+1,
                   stop=nchar(rownames(matT)))
 colnames(matT)<-c("X","Y","plot","year")
+matT<-subset(matT, as.numeric(matT$year) >= 1989 )
 plot(matT$X,matT$Y,type="n",xlab="",ylab="",cex.axis=0.7)
 invisible(lapply(1:length(treatments),function(tr){
   plots<-treatments[[tr]]
@@ -26,7 +27,7 @@ invisible(lapply(1:length(treatments),function(tr){
     points(toplot[1,"X"],toplot[1,"Y"],col=ColorsTr[[tr]],pch=16,cex=1.7)
     points(toplot[nrow(toplot),"X"],toplot[nrow(toplot),"Y"],col=ColorsTr[[tr]],pch=1,cex=1.7,lwd=2)
     
-    lines(toplot$X,toplot$Y,col=ColorsTr[[tr]],lty=2)
+    lines(toplot$X,toplot$Y,col=ColorsTr[[tr]])#,lty=2)
     invisible(lapply(round(seq(2,nrow(toplot),length.out=5)),function(step){
       Arrows(x0=toplot[step,"X"],y0=toplot[step,"Y"],
              x1=toplot[step-1,"X"],y1=toplot[step-1,"Y"],
@@ -53,7 +54,7 @@ invisible(lapply(1:length(treatments),function(tr){
     points(toplot[1,"X"],toplot[1,"Y"],col=ColorsTr[[tr]],pch=16,cex=1.7)
     points(toplot[nrow(toplot),"X"],toplot[nrow(toplot),"Y"],col=ColorsTr[[tr]],pch=1,cex=1.7,lwd=2)
     
-    lines(toplot$X,toplot$Y,col=ColorsTr[[tr]],lty=2)
+    lines(toplot$X,toplot$Y,col=ColorsTr[[tr]])#,lty=2)
     invisible(lapply(round(seq(2,nrow(toplot),length.out=5)),function(step){
       Arrows(x0=toplot[step,"X"],y0=toplot[step,"Y"],
              x1=toplot[step-1,"X"],y1=toplot[step-1,"Y"],
@@ -66,23 +67,23 @@ mtext("NMDS 2",side=2,padj=0,line=2,cex=0.8)
 }
 
[…]

### 2020-03-30 Scripts/Functionalredundancy.R

@@ -41,7 +41,7 @@ RedundancyTraj<-lapply(1:Nrep,function(rep){
   
   traits_filled<-Traits_filling(Traits1,Traits2,InventorySp)
   
-  ACP<-dudi.pca(scale(traits_filled[,TraitsName]),scannf=FALSE,nf=2)
+  ACP<-dudi.pca(scale(traits_filled[,TraitsName]),scannf=T,nf=2)
   Bigacp<-merge(ACP$li,traits_filled[,c("Family","Genus","name" )],by="row.names")[,c("Family","Genus","name","Axis1","Axis2")]
 
   xgen<-c(min(Bigacp[,"Axis1"]),max(Bigacp[,"Axis1"]))

### 2020-03-30 Scripts/Stats.R

@@ -1,3 +1,5 @@
+library(dplyr)
+
 load("DB/ReplacementTraj_ForGraphs")
 load("DB/FunctionalTraj_ForGraphs")
 load("DB/FunctionalRichnessTraj_ForGraphs")
@@ -86,11 +88,11 @@ plotIDH(FunRich,AGBloss)
 ## Rho Spearman, Taxo
 load("DB/FunDistance_ForGraphs");load("DB/TaxoDistance_ForGraphs");load("DB/LostAGB")
 
-Dist<-TaxoEuclid
+Dist<-FunEuclid
 ret<-apply(apply(Dist,c(1,2),median),1,function(li){return(max(abs(li)))})
 ret<-data.frame(cbind(ret,names(ret)))
 colnames(ret)<-c("Dist","plot")
-ret<-merge(ret,AGBloss,by="plot")
+ret<-merge(ret,AGBloss_cor,by="plot")
 ret<-apply(ret,2,as.numeric)
 cor(ret[,"Dist"],ret[,"AGB"],method="spearman")
 
@@ -131,7 +133,7 @@ Maxred<-unlist(lapply(1:12,function(pl){
 treats<-cbind(c(1,6,11,2,7,9,3,5,10,4,8,12),rep(0:3,each=3))
 Maxred<-cbind(Maxred,1:12)
 colnames(Maxred)<-c("Max","plot")
-Maxred<-merge(Maxred,AGBloss,by="plot")
+Maxred<-merge(Maxred,AGBloss_cor,by="plot")
 cor(Maxred[,"Max"],Maxred[,"AGB"],method="spearman")
 
 
@@ -144,7 +146,7 @@ Max<-do.call(rbind,lapply(CWM,function(pl){ return(apply(pl[,,"0.5"],2,max))}))
 treats<-cbind(c(1,6,11,2,7,9,3,5,10,4,8,12),rep(0:3,each=3))
 Max<-apply(cbind(Max,rownames(Max)),2,as.numeric)#,treats[order(treats[,1]),2])
 colnames(Max)<-c(colnames(CWM[[1]]),"plot")
-Max<-merge(Max,AGBloss,by="plot")
+Max<-merge(Max,AGBloss_cor,by="plot")
 apply(Max[,colnames(CWM[[1]])],2,function(x){return(cor(x,Max[,"AGB"],method="spearman"))})
 
 
@@ -187,7 +189,7 @@ for(q in 1:3){
[…]

### 2020-03-30 Scripts/Stats_traits.R

@@ -1,0 +1,62 @@
+load("DB/Paracou_R_Subdivided_ok")
+Traits1<-read.csv("DB/BridgeOK.csv",sep=";",na.strings="")
+Traits2<-read.csv("DB/DataLifeTraits.csv",sep=";",na.strings="")
+TraitsName<-c("Hmax","L_thickness","L_chloro","L_toughness","L_DryMass","SLA","WD","Bark_thick")
+InventorySp<-do.call(rbind,lapply(LivingStand_all,function(yr){
+  +   ret<-do.call(rbind,yr)[,c("Famille","Genre","name")]
+  +   ret<-ret[which(!grepl("Indet.",ret[,"name"])),]
+  +   return(ret[which(!duplicated(ret)),])}))
+InventorySp<-InventorySp[which(!duplicated(InventorySp)),]
+
+Traits1<-Traits1[which(apply(Traits1,1,function(li){return(any(!is.na(li)))})),]
+Traits2<-Traits2[which(apply(Traits2,1,function(li){return(any(!is.na(li)))})),]
+
+#Ajout colonne nom genre_espece pour avoir un seul identifiant par espèce
+Traits1["name"]<-as.factor(paste(Traits1$Genus,"_",Traits1$species,sep=""))
+Traits2["name"]<-sub(" ","_",Traits2[,"Name"])
+
+#Renommer les colonnes,purement pratique
+traits1<-Traits1[,c("Family","Genus","species","name","bar_code","thickness","SPAD","toughness","dry_mass","traits_surf_area","ind_surf_area",
+                    "sapwood_dens","moisture","bark_thick")]
+colnames(traits1)<-c("Family","Genus","species","name","bar_code","L_thickness","L_chloro","L_toughness","L_DryMass","LA","SLA","WD",
+                     "moisture","Bark_thick")
+traits2<-Traits2[,c("name","Hauteur","Masse")]
+colnames(traits2)<-c("name","Hmax","S_mass")
+traits2[which(traits2[,"name"]=="Sterculia_speciosa"),"Hmax"]<-43    # Anecdotique, pb d'estimation de la hauteur pour cette espèce
+
+# Réunir les deux tables de traits en 1
+traits<-merge(traits2,traits1,by="name")
+
+traits[which(!is.na(traits[,"LA"])),"SLA"]<-
+  traits[which(!is.na(traits[,"LA"])),"LA"]/traits[which(!is.na(traits[,"LA"])),"L_DryMass"]
+
[…]

### 2020-03-30 Scripts/TaxonomicNMDS.R

@@ -69,7 +69,9 @@ save(MatrepTaxo,file="DB/TaxoComposition_ForGraphs")
 #####################################################
 ### Taxonomic Euclidean distance
 
-Nrep<-10
+dates<-dates[which(as.numeric(dates)>=1989)]
+
+Nrep<-50
 
 TaxoEuclid<-lapply(1:Nrep,function(rep){
   Mat<-lapply(1:12,function(p){
@@ -90,7 +92,7 @@ TaxoEuclid<-lapply(1:Nrep,function(rep){
     Ok<-names(which(unlist(lapply(2:length(Trajnmds),function(step){return(Ntrees[step]-Ntrees[step-1])}))<=-90))
     Trajnmds<-Trajnmds[which(!names(Trajnmds)%in%Ok)]
     namesOk<-names(Trajnmds)
-    ref<-as.data.frame(Trajnmds[["1985"]])
+    ref<-as.data.frame(Trajnmds[["1989"]])
     Trajnmds<-do.call(cbind,lapply(Trajnmds,function(yr){
       All<-as.data.frame(AllInv[, "Genre"]);colnames(All)<-"Genre"
       ret<-merge(All,yr,by.x="Genre",by.y="row.names",all.x=TRUE)
@@ -106,6 +108,6 @@ TaxoEuclid<-lapply(1:Nrep,function(rep){
 TaxoEuclid<-array(unlist(Matrep),dim=c(nrow(Matrep[[1]]),ncol(Matrep[[1]]),length(Matrep)),
            dimnames=list(1:nrow(Matrep[[1]]),colnames(Matrep[[1]]),1:length(Matrep)))
 
-save(TaxoEuclid,file="DB/TaxoDistance_ForGraphs")
+save(TaxoEuclid,file="DB/TaxoDistance_ForGraphs_rect")
 
 

### 2020-03-30 Scripts/script_trajectories_paracou.R

@@ -1,0 +1,113 @@
+###########################################################
+### Trajectories of recovery in Paracou
+###########################################################
+
+# trajectories of the 12 plots of disturbance experiment
+
+
+library(data.table)
+library(ggplot2)
+library(BIOMASS)
+library(EcoFoG)
+
+load("DB/AGBlostDB")
+DataParacou <- AllYears 
+
+###### get data and tidying up ######
+######  
+DataParacou <- Guyafor2df(UID="Geraldine.Derroire",PWD="geraldine973+", WHERE="Forest='Paracou'")
+# DataParacou <- Guyafor2df(UID="Geraldine.Derroire",PWD="geraldine973+", WHERE="Forest='Paracou'AND Plot='16'")
+# Data <- as.data.table(read.csv("190310_DataParacou.csv", header=TRUE))
+Data <- as.data.table(DataParacou)
+# calculate DBH
+Data$DBH <- Data$CircCorr / pi
+# keep only DBH >=0 (a few tree have a circ-corr < 0, mostly Bois Cathédrale, cf code correction Camilia)
+Data <- Data[DBH >= 10]
+
+# keep only plot 1 to 12
+# Data <- Data[Plot %in% 1:12]
+# table(Data$Plot, Data$CensusYear)
+
+# add treatment
+Data$Treat <- as.factor(NA)
+Data[n_parcelle %in% c("1", "6", "11"), Treat := "Control"]
+Data[n_parcelle %in% c("2", "7", "9"), Treat := "T1"]
+Data[n_parcelle %in% c("3", "5", "10"), Treat := "T2"]
+Data[n_parcelle %in% c("4", "8", "12"), Treat := "T3"]
+Data$Treat <- as.factor(Data$Treat)
+# table(Data$Plot, Data$Treat)
+
+
+###### set appearance of graphs ######
+######  
+# colour of treatments
+MyColTreat <-c("Control" = "yellowgreen" ,"T1" = "gold",
+               "T2" = "orangered2", "T3" = "tomato4") 
+
+
+###### trajectories of AGB ######
+######  
+# get WD 
+Taxo=correctTaxo(genus=as.character(Data$Genre), species=as.character(Data$Espece)) # correct the species names
[…]

### 2020-03-30 WholePlotTrajectories.tex
cleanup

@@
 \usepackage{textcomp} \usepackage{tabu} \usepackage{caption}
-\captionsetup{justification = justified}
-\renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend}
-\hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
+
+\captionsetup{justification = justified} \renewenvironment{table}{\begin{table*}}{\end{table*}\ignorespacesafterend} \hyphenation{bio-di-ver-si-ty sap-lings re-or-gan-i-za-tion post-dis-tur-bance dis-tur-bance}
+
 \usepackage[left]{lineno} \linenumbers \usepackage{bookmark}
@@
 
-\PaperTitle{Diverging taxonomic and functional trajectories following disturbance in
-a Neotropical forest} % Article title
+\PaperTitle{Diverging taxonomic and functional trajectories following disturbance in a Neotropical forest} % Article title
 
@@
 \affiliation{
-\textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles,
-Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}INPHB, Institut National Polytechnique Félix Houphoüet-Boigny\\ \hspace{1em} Yamoussoukro, Ivory Coast.
+\textsuperscript{1}UMR EcoFoG, AgroParistech, CNRS, Cirad, INRA, Université des Antilles, Université de Guyane.\\ \hspace{1em} Campus Agronomique, 97310 Kourou, France.\\\textsuperscript{2}INPHB, Institut National Polytechnique Félix Houphoüet-Boigny\\ \hspace{1em} Yamoussoukro, Ivory Coast.
 }
@@
 \Abstract{
-In the current global change context, it is urgent to anticipate the
-fate of tropical forests. This means understanding tree community
-response to disturbance and the underlying processes. In that respect,
-we aim here to clarify taxonomic and functional post-disturbance
-trajectories, and determine the scope of the Intermediate Disturbance
-Hypothesis (IDH) that remains debated in tropical forests.
-\color{black}We analyzed community trajectories following a disturbance
-gradient from 10 to 60\% of above-ground biomass loss in a Neotropical
-forest over 30 years.\color{red} We considered trajectories along time
-of community taxonomic and functional trajectories in terms of richness,
-evenness, composition, and redundancy. We based on the annual botanical
-inventories of 75 ha of a Neotropical forest and on large trait datasets
-comprising seven leaf, stem, and life-history traits. We identified a
-decoupling between taxonomic composition, differing among communities,
-and functional composition, \color{red} similar among communities and
-convergent in the functional space.\color{black} \color{red} The
-taxonomic diversity followed humped-shaped trajectories along time after
-disturbance depending on the initial disturbance intensity, which
-validated the IDH (Intermediate Disturbance Hypothesis).\color{black}
-The functional diversity trajectories,however, were homogeneous among
-plots and dismissed the IDH. We explained this decoupling by the
-variations in community functional redundancy that mitigated the
-functional impact of disturbance. Although consistent, the recovery of
-community composition, diversity, and redundancy remained unachieved
-after 30 years. \color{red} These results acknowledged the need of
-decades-long cycles without disturbance to ensure community complete
-recovery.\color{black}
+In the current global change context, it is urgent to anticipate the fate of tropical forests.
+This means understanding tree community response to disturbance and the underlying processes.
+In that respect, we aim here to clarify taxonomic and functional post-disturbance trajectories, and determine the scope of the Intermediate Disturbance Hypothesis (IDH) that remains debated in tropical forests.
+\color{black}We analyzed community trajectories following a disturbance gradient from 10 to 60\% of above-ground biomass loss in a Neotropical forest over 30 years.\color{red} We considered trajectories along time of community taxonomic and functional trajectories in terms of richness, evenness, composition, and redundancy.
+We based on the annual botanical inventories of 75 ha of a Neotropical forest and on large trait datasets comprising seven leaf, stem, and life-history traits.
+We identified a decoupling between taxonomic composition, differing among communities, and functional composition, \color{red} similar among communities and convergent in the functional space.\color{black} \color{red} The taxonomic diversity followed humped-shaped trajectories along time after disturbance depending on the initial disturbance intensity, which validated the IDH (Intermediate Disturbance Hypothesis).\color{black} The functional diversity trajectories,however, were homogeneous among plots and dismissed the IDH.
+We explained this decoupling by the variations in community functional redundancy that mitigated the functional impact of disturbance.
+Although consistent, the recovery of community composition, diversity, and redundancy remained unachieved after 30 years.
+\color{red} These results acknowledged the need of decades-long cycles without disturbance to ensure community complete recovery.\color{black}
 }
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-The large areas covered with tropical forests worldwide hold crucial
-environmental, economic, and social values. They provide wood and
-multiple non-timber forest products, shelter a diversified fauna, and
-ensure cultural and human well-being. They regulate local and regional
-climates, as well as carbon, water and nutrient cycles. However, the
-growing demand in forests products together with current global changes
-increase the pressure on remaining undisturbed forests
-\citep{Morales-Hidalgo2015}. \color{red}These threats may change the
-frequency and magnitude of the natural disturbance regime that defines
-and maintains the structure, composition, and functioning of tree
-communities \citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}.
-\color{black} To anticipate the fate of tropical forests, it is urgent
-to understand tree community response to disturbance, and the underlying
-ecological processes. The forest cover is generally maintained following
-disturbance, but modifications in the fluxes of light, heat, and water
-\citep{Goulamoussene2017} change community abiotic and biotic
-environments. These changes translate into post-disturbance community
-trajectories that have been largely studied through trajectories of
-forest structural parameters such as aboveground biomass, tree height or
-stem density \citep{Piponiot2016, Rutishauser2016}. Some of the
-determinants of post-disturbance biomass trajectories are already
-identified, like the structure and composition of the pre-disturbance
-community, or the post-disturbance environmental parameters
-\citep{Herault2018}. Community diversity and composition trajectories,
-however, have not been as thoroughly understood
-\citep{Guitet2018, Molino2001}, and manifold biodiversity trajectories
-might emerge given the variety of species response to disturbance and
-the diversity of tropical forests
-\citep{Lindenmayer2012, Garcia_florez2017}.
+The large areas covered with tropical forests worldwide hold crucial environmental, economic, and social values.
+They provide wood and multiple non-timber forest products, shelter a diversified fauna, and ensure cultural and human well-being.
+They regulate local and regional climates, as well as carbon, water and nutrient cycles.
+However, the growing demand in forests products together with current global changes increase the pressure on remaining undisturbed forests \citep{Morales-Hidalgo2015}.
+\color{red}These threats may change the frequency and magnitude of the natural disturbance regime that defines and maintains the structure, composition, and functioning of tree communities \citep{Schnitzer2001, Anderson-Teixeira2013, Sist2015}.
+\color{black}
+To anticipate the fate of tropical forests, it is urgent to understand tree community response to disturbance, and the underlying ecological processes.
+The forest cover is generally maintained following disturbance, but modifications in the fluxes of light, heat, and water \citep{Goulamoussene2017} change community abiotic and biotic environments.
+These changes translate into post-disturbance community trajectories that have been largely studied through trajectories of forest structural parameters such as aboveground biomass, tree height or stem density \citep{Piponiot2016, Rutishauser2016}.
+Some of the determinants of post-disturbance biomass trajectories are already identified, like the structure and composition of the pre-disturbance community, or the post-disturbance environmental parameters \citep{Herault2018}.
+Community diversity and composition trajectories, however, have not been as thoroughly understood \citep{Guitet2018, Molino2001}, and manifold biodiversity trajectories might emerge given the variety of species response to disturbance and the diversity of tropical forests \citep{Lindenmayer2012, Garcia_florez2017}.
 
-An early conceptual basis of the linkage between biodiversity and
-disturbance is the Intermediate Disturbance Hypothesis (IDH). The IDH
-states a relationship between the community diversity and the intensity
-and frequency of disturbance events, and postulates a diversity peak at
-intermediate level of disturbance \citep{Connell1978}. This is based on
-the fluctuations of community environment following disturbance that
-foster both competitively superior species and fast colonizers, and
-prevents competitive exclusion \citep{Shea2004, Pulsford2016}.
-\color{red} In tropical forests, some studies advocate community
-deterministic response to disturbance, as for example in the dry and wet
-forests in Ghana \citep{Bongers2009} or in the guianese rainforests
-\citep{Molino2001}, where intermediate disturbance proved to increase
-the community richness by increasing the proportion of pioneers or
-heliophilous species in the community while maintaining the richness in
-old-growth forest species. In other cases, however, observations of the
-IDH diverge from theoretical expectations
-\citep{Hugues2007, Sheil2003, Norden2017}. Some studies refute the role
-of disturbance intensity, as observed in forest gaps in Panama where no
-difference in species richness was found between forest gaps and
-undisturbed areas \citep{Hubbell2001}, or in costa-rican forests where
-forests that did not follow deterministic convergence display stochastic
-trajectories \citep{Norden2015}.\\
-The high diversity of tropical forests may foster the emergence of
-numerous facilitation, adaptation, and inter- and intra-specific
-competition following disturbance
-\citep{Garcia_florez2017, Bongers2009}.These interactions can result in
-miscellaneous responses to disturbance \citep{Lindenmayer2012}, which
-question the validity of the IDH in tropical forests
-\citep{Hubbell2001, Fox2013, Sheil2013}. \color{black}
-
-Analysing community response to disturbance and grasp all aspects of
-community changes requires a wide array of metrics
-\citep{Sheil2003, Shea2004, Mayfield2010}. The analysis should first
-consider community composition, which is crucial for conservation issues
-and reveals the pool of species fostered or hampered by disturbance
-\citep{Lavorel2002, Bellwood2006}. This should be completed with
-diversity metrics encompassing both community richness and evenness to
-assess the changes in community abundance distribution. Besides,
-functional approaches have been shown to usefully complement pure
-taxonomic approaches as they shed light on the species biological
-attributes directly linking community diversity, composition, and
-redundancy to ecosystem functioning \citep{Violle2007b, Baraloto2012a}.
-In that respect, a vast literature allowed recognizing major traits
-representing species ecological strategy and determining how they
-respond to changing conditions \citep{Diaz2005}. Specifically, in
-tropical forests, the functional approach revealed the emergence of
-deterministic processes following disturbance. Such deterministic
-processes entailed a shift from a dominance of ``conservative''
-slow-growing species dealing with scarce resources, to a dominance of
-``acquisitive'' fast-growing species with rapid and efficient use of
-abundant resources \citep{Rees2001, Reich2014, Herault2011}. This shift
-is translated into the trajectories of average community value of key
-functional traits related to resource acquisition, as leaf and stem
-traits, and life-history strategy, as seed mass and maximum size
-\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}.
-
-The functional approach also encompasses the analysis of functional
-redundancy, that quantifies the amount of shared trait values among
-species \citep{Carmona2016}. The typical high functional redundancy of
-hyper-diverse tropical forests \citep{Bellwood2006} mitigates the
-impacts of species removal on ecosystem functioning, and determines
-community resilience after disturbance \citep{Elmqvist2003, Diaz2005}.
-
-Here, we monitored over 30 years the response of 75 ha of Neotropical
-forest plots set up on a gradient of disturbance intensity, from 10 to
-60\% of above-ground biomass (AGB) loss. We made use of a large
-functional traits database encompassing major leaf, stem, and
-life-history traits in order to draw the taxonomic and functional
-trajectories in terms of richness, evenness, composition, and
-redundancy. Specifically, (i) we drew taxonomic and functional
-post-disturbance trajectories and examined the underlying ecological
-processes, (ii) we discussed the scope of the IDH regarding taxonomic
-and functional facets of community diversity, and (iii) we analyzed
-community resilience and time to recovery. \color{red} We hypothesized
-that community functional and taxonomic trajectories might not track
-each other, given the high functional redundancy in tropical forests.
-While functional diversity would be enhanced by the environmental
-changes following disturbance, specifically the increase in light
-availability, taxonomic trajectories would only increase until a
-disturbance intensity threshold, as pioneers recruited following
-disturbance are not as diverse as late-successional species. Community
-taxonomic and functional would hence follow different rates of recovery.
+An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
+The IDH states a relationship between the community diversity and the intensity and frequency of disturbance events, and postulates a diversity peak at intermediate level of disturbance \citep{Connell1978}.
+This is based on the fluctuations of community environment following disturbance that foster both competitively superior species and fast colonizers, and prevents competitive exclusion \citep{Shea2004, Pulsford2016}.
+\color{red}
+In tropical forests, some studies advocate community deterministic response to disturbance, as for example in the dry and wet forests in Ghana \citep{Bongers2009} or in the guianese rainforests \citep{Molino2001}, where intermediate disturbance proved to increase the community richness by increasing the proportion of pioneers or heliophilous species in the community while maintaining the richness in old-growth forest species.
+In other cases, however, observations of the IDH diverge from theoretical expectations \citep{Hugues2007, Sheil2003, Norden2017}.
+Some studies refute the role of disturbance intensity, as observed in forest gaps in Panama where no difference in species richness was found between forest gaps and undisturbed areas \citep{Hubbell2001}, or in costa-rican forests where forests that did not follow deterministic convergence display stochastic trajectories \citep{Norden2015}.\\
+The high diversity of tropical forests may foster the emergence of numerous facilitation, adaptation, and inter- and intra-specific competition following disturbance \citep{Garcia_florez2017, Bongers2009}.These interactions can result in miscellaneous responses to disturbance \citep{Lindenmayer2012}, which question the validity of the IDH in tropical forests \citep{Hubbell2001, Fox2013, Sheil2013}.
 \color{black}
 
-\section{Material and Methods}\label{material-and-methods}
+Analysing community response to disturbance and grasp all aspects of community changes requires a wide array of metrics \citep{Sheil2003, Shea2004, Mayfield2010}.
+The analysis should first consider community composition, which is crucial for conservation issues and reveals the pool of species fostered or hampered by disturbance \citep{Lavorel2002, Bellwood2006}.
+This should be completed with diversity metrics encompassing both community richness and evenness to assess the changes in community abundance distribution.
+Besides, functional approaches have been shown to usefully complement pure taxonomic approaches as they shed light on the species biological attributes directly linking community diversity, composition, and redundancy to ecosystem functioning \citep{Violle2007b, Baraloto2012a}.
+In that respect, a vast literature allowed recognizing major traits representing species ecological strategy and determining how they respond to changing conditions \citep{Diaz2005}.
+Specifically, in tropical forests, the functional approach revealed the emergence of deterministic processes following disturbance.
+Such deterministic processes entailed a shift from a dominance of ``conservative'' slow-growing species dealing with scarce resources, to a dominance of ``acquisitive'' fast-growing species with rapid and efficient use of abundant resources \citep{Rees2001, Reich2014, Herault2011}.
+This shift is translated into the trajectories of average community value of key functional traits related to resource acquisition, as leaf and stem traits, and life-history strategy, as seed mass and maximum size \citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}.
 
-\subsection{Study site}\label{study-site}
+The functional approach also encompasses the analysis of functional redundancy, that quantifies the amount of shared trait values among species \citep{Carmona2016}.
+The typical high functional redundancy of hyper-diverse tropical forests \citep{Bellwood2006} mitigates the impacts of species removal on ecosystem functioning, and determines community resilience after disturbance \citep{Elmqvist2003, Diaz2005}.
 
-Paracou station in French Guiana (5° 18'N and 52° 53'W) is located in a
-lowland tropical rain forest in a tropical wet climate with mean annual
-temperature of 26° C, and mean annual precipitation averaging 2980
-mm.y\textsuperscript{-1} (30-y period). The climate comprises a 3-month
-dry season (\textless{} 100 mm.month\textsuperscript{-1}) from
-mid-August to mid-November, and a one-month dry season in March
-\citep{Wagner2011}. Across all plots, elevation ranges from 5 to 50 m,
-and the topography mainly corresponds to hilltops or hillsides, while
-bottomlands cover less than 1 \% of the area. Plots are shallow
-ferralitic acrisols over a layer of transformed saprolite with low
-permeability and lateral drainage. Soil conditions are homogeneous, to
-the exception of the highest hilltops where the thick surface allows a
-free vertical drainage \citep{Gourlet-Fleury2004}.
+Here, we monitored over 30 years the response of 75 ha of Neotropical forest plots set up on a gradient of disturbance intensity, from 10 to 60\% of above-ground biomass (AGB) loss.
+We made use of a large functional traits database encompassing major leaf, stem, and life-history traits in order to draw the taxonomic and functional trajectories in terms of richness, evenness, composition, and redundancy.
+Specifically, (i) we drew taxonomic and functional post-disturbance trajectories and examined the underlying ecological processes, (ii) we discussed the scope of the IDH regarding taxonomic and functional facets of community diversity, and (iii) we analyzed community resilience and time to recovery.
+\color{red}
+We hypothesized that community functional and taxonomic trajectories might not track each other, given the high functional redundancy in tropical forests.
+While functional diversity would be enhanced by the environmental changes following disturbance, specifically the increase in light availability, taxonomic trajectories would only increase until a disturbance intensity threshold, as pioneers recruited following disturbance are not as diverse as late-successional species.
+Community taxonomic and functional would hence follow different rates of recovery.
+\color{black}
 
-The experiment is a network of twelve 6.25 ha plots\color{red},
-distributed over a 400ha area \color{black}(Table \ref{tab:Tab1}, Supp.
-Mat. Fig. S1), that undergone three disturbance treatments in 1987
-according to a randomized plot design \citep{Gourlet-Fleury2004}.\\
-\color{red}The experiment intended to simulate logging intensities and
-distinguished the treatments for non-commercial commercial species,
-logged at 40cm DBH to simulate thinning operations.\color{black} The
-experiment comprised three replicates of three sylvicultural treatments
-(hereafter plots T1, T2, and T3), and three control plots (T0).
-\color{red}Control plots are set in mature forest stand that did not
-undergone any logging nor intense anthropic activities activities for
-the last 500 years\color{black}. All treatments T1, T2, and T3 comprised
-the logging of 10 trees/ha with 50 cm minimum DBH that belonged to a set
-of 58 commercially exploited species \citep{Gourlet-Fleury2004}.
-Treatment T2 additionally comprised a thinning treatment by
-poison-girdling of non-commercially exploited species randomly selected
-with an average of 30 trees/ha with 40 cm minimum DBH. Treatment T3
-additionally comprised the logging of 15 trees/ha with 40 cm minimum
-DBH, and the poison-girdling of 20 trees/ha with a 50 cm minimum DBH,
-all belonging to non-commercially exploited species. Considering the
-silvicultural treatments and the following damage, disturbance intensity
-was measured as the percentage of aboveground biomass (\%AGB) lost
-between the first inventory in 1984 and five years after disturbance
-\citep{Piponiot2016} estimated with the BIOMASS R package
-\citep{Rejou2017}. \color{red} Tree diameter are provided in the
-experimental dataset, wood density is extracted from global wood density
-(GWD) database \citep{Zanne2009}, and tree height is retrieved from a
-generic H--D model based on a single bioclimatic predictor (eqn 6a in
-\citep{Chave2014}). The AGB is measured from the generalized allometric
-model eqn 4 of \citep{Chave2014}. \color{black} The three treatments
-were then transformed into a continuous disturbance intensity gradient
-with increasing above-ground biomass (AGB) loss.
+\hypertarget{material-and-methods}{%
+\section{Material and Methods}\label{material-and-methods}}
 
-\begin{table}
+\hypertarget{study-site}{%
+\subsection{Study site}\label{study-site}}
+
+Paracou station in French Guiana (5° 18'N and 52° 53'W) is located in a lowland tropical rain forest in a tropical wet climate with mean annual temperature of 26° C, and mean annual precipitation averaging 2980 mm.y\textsuperscript{-1} (30-y period).
+The climate comprises a 3-month dry season (\textless{} 100 mm.month\textsuperscript{-1}) from mid-August to mid-November, and a one-month dry season in March \citep{Wagner2011}.
+Across all plots, elevation ranges from 5 to 50 m, and the topography mainly corresponds to hilltops or hillsides, while bottomlands cover less than 1 \% of the area.
+Plots are shallow ferralitic acrisols over a layer of transformed saprolite with low permeability and lateral drainage.
+Soil conditions are homogeneous, to the exception of the highest hilltops where the thick surface allows a free vertical drainage \citep{Gourlet-Fleury2004}.
+
+The experiment is a network of twelve 6.25 ha plots\color{red}, distributed over a 400ha area \color{black}(Table \ref{tab:Tab1}, Supp.
+Mat.
+Fig.
+S1), that undergone three disturbance treatments in 1987 according to a randomized plot design \citep{Gourlet-Fleury2004}.\\
+\color{red}The experiment intended to simulate logging intensities and distinguished the treatments for non-commercial commercial species, logged at 40cm DBH to simulate thinning operations.\color{black}
+The experiment comprised three replicates of three sylvicultural treatments (hereafter plots T1, T2, and T3), and three control plots (T0).
+\color{red}Control plots are set in mature forest stand that did not undergone any logging nor intense anthropic activities activities for the last 500 years\color{black}.
+All treatments T1, T2, and T3 comprised the logging of 10 trees/ha with 50 cm minimum DBH that belonged to a set of 58 commercially exploited species \citep{Gourlet-Fleury2004}.
+Treatment T2 additionally comprised a thinning treatment by poison-girdling of non-commercially exploited species randomly selected with an average of 30 trees/ha with 40 cm minimum DBH.
+Treatment T3 additionally comprised the logging of 15 trees/ha with 40 cm minimum DBH, and the poison-girdling of 20 trees/ha with a 50 cm minimum DBH, all belonging to non-commercially exploited species.
+Considering the silvicultural treatments and the following damage, disturbance intensity was measured as the percentage of aboveground biomass (\%AGB) lost between the first inventory in 1984 and five years after disturbance \citep{Piponiot2016} estimated with the BIOMASS R package \citep{Rejou2017}.
+\color{red}
+Tree diameter are provided in the experimental dataset, wood density is extracted from global wood density (GWD) database \citep{Zanne2009}, and tree height is retrieved from a generic H--D model based on a single bioclimatic predictor (eqn 6a in \citep{Chave2014}).
+The AGB is measured from the generalized allometric model eqn 4 of \citep{Chave2014}.
+\color{black}
+The three treatments were then transformed into a continuous disturbance intensity gradient with increasing above-ground biomass (AGB) loss.
+
+\begin{table*}[t]
 
@@
 \end{tabu}
-\end{table}
+\end{table*}\ignorespacesafterend
 
-\subsection{Inventories protocol and dataset
-collection}\label{inventories-protocol-and-dataset-collection}
+\hypertarget{inventories-protocol-and-dataset-collection}{%
+\subsection{Inventories protocol and dataset collection}\label{inventories-protocol-and-dataset-collection}}
 
-The study site corresponds to a tropical rainforest typical of the
-Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae,
-Lecythidaceae, and Sapotaceae. In the 12 experimental plots, all trees
-above 10 cm DBH have been mapped and measured annually since 1984. Trees
-are first identified with a vernacular name assigned by the forest
-worker team, and afterward with a scientific name assigned by botanists
-during regular botanical campaigns. In 1984, specific vernacular names
-were given to 62 commmon or commercially exploited species. More
-infrequent species were identified under general identifiers only
-distinguishing trees and palms. From 2003, botanical campaigns have been
-conducted every 5 to 6 years to identify all trees at the species level.
-In 2015, however, identification levels still varied among plots and
-campaigns.
+The study site corresponds to a tropical rainforest typical of the Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae, Lecythidaceae, and Sapotaceae.
+In the 12 experimental plots, all trees above 10 cm DBH have been mapped and measured annually since 1984.
+Trees are first identified with a vernacular name assigned by the forest worker team, and afterward with a scientific name assigned by botanists during regular botanical campaigns.
+In 1984, specific vernacular names were given to 62 commmon or commercially exploited species.
+More infrequent species were identified under general identifiers only distinguishing trees and palms.
+From 2003, botanical campaigns have been conducted every 5 to 6 years to identify all trees at the species level.
+In 2015, however, identification levels still varied among plots and campaigns.
 
-This variability of protocols in time raised methodological issues as
-vernacular names usually correspond to various botanical species. This
-resulted in significant taxonomic uncertainty that had to be accounted
-for in the measure of composition and diversity metrics. Uncertainty
-propagation was implented within a Bayesian framework using
-vernacular/botanical names associations to reconstitute complete
-inventories at genus level from real incomplete ones. Vernacular names
-were replaced through multinomial trials based on the association
-probability \(\big[\alpha_1, \alpha_2,..., \alpha_V\big]\) observed
-across all inventories between each vernacular name \emph{v} and all
-species \(\big[s_1, s_2,..., s_N\big]\):
+This variability of protocols in time raised methodological issues as vernacular names usually correspond to various botanical species.
+This resulted in significant taxonomic uncertainty that had to be accounted for in the measure of composition and diversity metrics.
+Uncertainty propagation was implented within a Bayesian framework using vernacular/botanical names associations to reconstitute complete inventories at genus level from real incomplete ones.
+Vernacular names were replaced through multinomial trials based on the association probability \(\big[\alpha_1, \alpha_2,..., \alpha_V\big]\) observed across all inventories between each vernacular name \emph{v} and all species \(\big[s_1, s_2,..., s_N\big]\):
 
@@
 
-See Supplementary Materials -Fig. S2 and \citet{Aubry-Kientz2013} for
-the detailed methodology.
+See Supplementary Materials -Fig. S2 and \citet{Aubry-Kientz2013} for the detailed methodology.
 
-Six functional traits representing the leaf economics (\emph{i.e.} leaf
-thickness, toughness, total chlorophyll content, and specific leaf
-area), and the stem economics (\emph{i.e.} wood specific gravity and
-bark thickness) were obtained from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}.
-Trait values were assessed from a selection of individuals located in
-nine permanent plots in French Guiana, including two in Paracou, and
-comprised 294 species belonging to 157 genera. Whenever a species was in
-the dataset but missed some trait values (10\% of the species), missing
-values were filled using multivariate imputation by chained equation
-\citep{Mice2011}. To account for the phylogenetic signal in the
-imputation process, imputations were based on samples of species from
-the same genus or from the same family. Whenever a species was not in
-the dataset, it was attributed a set of trait values randomly sampled
-among species of the next higher taxonomic level (same genus or family).
-Two life-history traits, maximum specific height and seed mass, came
-from the Mariwenn database \footnote{https://www.ecofog.gf/mariwenn/}.
-The database compiles information from a vast literature on the flora of
-French Guiana \citep{Ollivier2007} and comprises 362 species belonging
-to 188 genera. As seed mass information was classified into classes, no
-data filling process was applied and analyses were restricted to the
-botanical species recorded.
+Six functional traits representing the leaf economics (\emph{i.e.} leaf thickness, toughness, total chlorophyll content, and specific leaf area), and the stem economics (\emph{i.e.} wood specific gravity and bark thickness) were obtained from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}.
+Trait values were assessed from a selection of individuals located in nine permanent plots in French Guiana, including two in Paracou, and comprised 294 species belonging to 157 genera.
+Whenever a species was in the dataset but missed some trait values (10\% of the species), missing values were filled using multivariate imputation by chained equation \citep{Mice2011}.
+To account for the phylogenetic signal in the imputation process, imputations were based on samples of species from the same genus or from the same family.
+Whenever a species was not in the dataset, it was attributed a set of trait values randomly sampled among species of the next higher taxonomic level (same genus or family).
+Two life-history traits, maximum specific height and seed mass, came from the Mariwenn database \footnote{https://www.ecofog.gf/mariwenn/}.
+The database compiles information from a vast literature on the flora of French Guiana \citep{Ollivier2007} and comprises 362 species belonging to 188 genera.
+As seed mass information was classified into classes, no data filling process was applied and analyses were restricted to the botanical species recorded.
 
-\color{red}Confidence intervals of the composition and diversity metrics
-were computed after 50 iterations of the uncertainty propagation
-framework.\color{black}
+\color{red}Confidence intervals of the composition and diversity metrics were computed after 50 iterations of the uncertainty propagation framework.\color{black}
 
-\subsection{Composition and diversity
-metrics}\label{composition-and-diversity-metrics}
+\hypertarget{composition-and-diversity-metrics}{%
+\subsection{Composition and diversity metrics}\label{composition-and-diversity-metrics}}
 
-Because of the variable precision of botanical identification efforts,
-we had to conduct the taxonomic composition and diversity analysis at
-the genus level. \color{red}Functional composition, diversity, and
-redundancy measures were conducted at species level.\color{black}
-Taxonomic and functional trajectories of community composition were
-drawn in a two-dimensional NMDS ordination plane. Two NMDS using
-abundance-based (\emph{i.e.} Bray-Curtis) dissimilarity measures were
-conducted to map either taxonomic or functional composition, the latter
-based on the seven leaf, stem, and life history traits (without seed
-mass classes). Trajectories along time were reported through the
-Euclidean distance between the target inventories and the 1984
-pre-disturbance inventories of reference. Univariate trajectories of the
-leaf, stem, and life-history traits were also visualized with the
-community weighted means (CWM) \citep{Diaz2007}. Species seed mass were
-given in 5 mass classes, and seed mass trajectories were reported as the
-proportion of each class in the inventories (Supplementary materials).
+Because of the variable precision of botanical identification efforts, we had to conduct the taxonomic composition and diversity analysis at the genus level.
+\color{red}Functional composition, diversity, and redundancy measures were conducted at species level.\color{black}
+Taxonomic and functional trajectories of community composition were drawn in a two-dimensional NMDS ordination plane.
+Two NMDS using abundance-based (\emph{i.e.} Bray-Curtis) dissimilarity measures were conducted to map either taxonomic or functional composition, the latter based on the seven leaf, stem, and life history traits (without seed mass classes).
+Trajectories along time were reported through the Euclidean distance between the target inventories and the 1984 pre-disturbance inventories of reference.
+Univariate trajectories of the leaf, stem, and life-history traits were also visualized with the community weighted means (CWM) \citep{Diaz2007}.
+Species seed mass were given in 5 mass classes, and seed mass trajectories were reported as the proportion of each class in the inventories (Supplementary materials).
 
-The taxonomic and functional trajectories were analysed from the 1984
-pre-disturbance inventories of reference. The taxonomic diversity was
-reported through species richness and the Hill number translation of the
-Simpson index \citep{Hill1973}. The comparison between these two metrics
-assesses community taxonomic richness and evenness: thereafter, results
-will be discussed directly in terms of taxonomic richness and evenness.
-Both indices are recommended for diversity studies \citep{Marcon2015},
-and belong to the set of \color{red} Tsallis or generalized entropy
-\color{black} corresponding, respectively, to the 0 and 2 order of
-diversity (q). \color{red}Diversity metrics were measured using the Hqz
-function from the entropart package \citep{Marcon2015}.\color{black} The
-functional diversity was reported using the functional richness and
-functional evenness. \color{red}Both functional richness and evenness
-were measured through a distance-based dissimilarity matrix built with
-the daisy function of the cluster R package \citep{Cluster2019}, using
-the Gower distance between species, similar to \citet{Herault2007a},
-\citet{Herault2007b}.\color{black} The functional diversity was measured
-through the Rao index of quadratic entropy. The Rao index combines the
-community abundance distribution\color{red}, from species relative
-abundance, \color{black} and the average pairwise functional
-dissimilarity between species computed by the Gower distance.
-\color{red}The comparison between these two metrics assesses community
-functional richness and evenness: thereafter, results will be discussed
-directly in terms of functional richness and evenness.\color{black}
+The taxonomic and functional trajectories were analysed from the 1984 pre-disturbance inventories of reference.
+The taxonomic diversity was reported through species richness and the Hill number translation of the Simpson index \citep{Hill1973}.
+The comparison between these two metrics assesses community taxonomic richness and evenness: thereafter, results will be discussed directly in terms of taxonomic richness and evenness.
+Both indices are recommended for diversity studies \citep{Marcon2015}, and belong to the set of \color{red} Tsallis or generalized entropy \color{black} corresponding, respectively, to the 0 and 2 order of diversity (q).
+\color{red}Diversity metrics were measured using the Hqz function from the entropart package \citep{Marcon2015}.\color{black}
+The functional diversity was reported using the functional richness and functional evenness.
+\color{red}Both functional richness and evenness were measured through a distance-based dissimilarity matrix built with the daisy function of the cluster R package \citep{Cluster2019}, using the Gower distance between species, similar to \citet{Herault2007a}, \citet{Herault2007b}.\color{black}
+The functional diversity was measured through the Rao index of quadratic entropy.
+The Rao index combines the community abundance distribution\color{red}, from species relative abundance, \color{black} and the average pairwise functional dissimilarity between species computed by the Gower distance.
+\color{red}The comparison between these two metrics assesses community functional richness and evenness: thereafter, results will be discussed directly in terms of functional richness and evenness.\color{black}
 
-The impacts of the initial disturbance levels on the maximum gain or
-loss in richness and evenness were tested with the Spearman rank
-correlation tests. Richness and evenness trajectories were besides
-analyzed through polynomial regression \color{red}or simple linear
-models \color{black} between (i) taxonomic and functional richness and
-evenness, and (ii) the initial \%AGB loss at 10, 20, and 30 years after
-disturbance. \color{red}The best model fitting, between linear or
-quadratic model, was chosen based on their respective AIC criterion. The
-explanatory power of the fit were measured to adjusted r-square
-values.\color{black}
+The impacts of the initial disturbance levels on the maximum gain or loss in richness and evenness were tested with the Spearman rank correlation tests.
+Richness and evenness trajectories were besides analyzed through polynomial regression \color{red}or simple linear models \color{black} between (i) taxonomic and functional richness and evenness, and (ii) the initial \%AGB loss at 10, 20, and 30 years after disturbance.
+\color{red}The best model fitting, between linear or quadratic model, was chosen based on their respective AIC criterion.
+The explanatory power of the fit were measured to adjusted r-square values.\color{black}
 
-Finally, functional redundancy was measured as the overlap among species
-in community functional space \citep{Carmona2016}. First, the
-individuals of the trait database were mapped in the plane of the first
-two axes from a PCA analysis. The PCA analysis lowered the weight of
-correlations among traits as the axes are combinations of the most
-decoupled traits. \color{red}The two main PCA axis respectively hold
-26\% and 19\% of the trait dataset variance.\color{black} For each
-species, the traits probability density (TPD) were computed from the
-mapping of individuals through two-dimension kernel density estimators.
-\color{red} We used the kde2d function of the MASS package
-\citep{MASS2002} that is a gaussian density estimator for a
-2-dimensional space. \color{black} Second, for each community, the TPD
-weigthed by species abundance were summed across the functional space.
-Third, the functional space was divided into a 100 x 100 grid, and the
-number of species with a positive TPD was counted in each cell. The
-average count across cells minus 1 returned the Community Functional
-Redundancy, which was the average number of species in the community
-that share the same trait values.
+Finally, functional redundancy was measured as the overlap among species in community functional space \citep{Carmona2016}.
+First, the individuals of the trait database were mapped in the plane of the first two axes from a PCA analysis.
+The PCA analysis lowered the weight of correlations among traits as the axes are combinations of the most decoupled traits.
+\color{red}The two main PCA axis respectively hold 26\% and 19\% of the trait dataset variance.\color{black}
+For each species, the traits probability density (TPD) were computed from the mapping of individuals through two-dimension kernel density estimators.
+\color{red}
+We used the kde2d function of the MASS package \citep{MASS2002} that is a gaussian density estimator for a 2-dimensional space.
+\color{black}
+Second, for each community, the TPD weigthed by species abundance were summed across the functional space.
+Third, the functional space was divided into a 100 x 100 grid, and the number of species with a positive TPD was counted in each cell.
+The average count across cells minus 1 returned the Community Functional Redundancy, which was the average number of species in the community that share the same trait values.
 
-\section{Results}\label{results}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-\subsection{Community Composition}\label{community-composition}
+\hypertarget{community-composition}{%
+\subsection{Community Composition}\label{community-composition}}
 
-From 1984, the first pre-disturbance inventory, to 2015, 28 years after
-disturbance, 828 388 individual trees and 591 botanical species spanning
-223 genera and 64 families were recorded.
+From 1984, the first pre-disturbance inventory, to 2015, 28 years after disturbance, 828 388 individual trees and 591 botanical species spanning 223 genera and 64 families were recorded.
 
-In undisturbed plots, both taxonomic and functional composition remained
-stable (Fig. \ref{fig:NMDSplans}). In disturbed plots, both trajectories
-followed marked and consistent trajectories over time.
+In undisturbed plots, both taxonomic and functional composition remained stable (Fig. \ref{fig:NMDSplans}).
+In disturbed plots, both trajectories followed marked and consistent trajectories over time.
 
-The functional composition trajectory resembled, in disturbed
-communities, to cyclic compositional changes with an incomplete recovery
-of the initial composition (Fig.\ref{fig:NMDSplans}). The maximum
-dissimilarity with the initial state was positively correlated with the
-disturbance intensity for both taxonomic and functional composition
-(\(\rho_{Spearman}^{Taxonomic}=0.94\) and
-\(\rho_{Spearman}^{Functional}=0.90\), respectively. Fig.
-\ref{fig:SumUp}). The maximum dissimilarity with the initial state was
-reached for taxonomic composition between 15 to 25 years, and between 20
-to 25 years for functional composition.
+The functional composition trajectory resembled, in disturbed communities, to cyclic compositional changes with an incomplete recovery of the initial composition (Fig.\ref{fig:NMDSplans}).
+The maximum dissimilarity with the initial state was positively correlated with the disturbance intensity for both taxonomic and functional composition (\(\rho_{Spearman}^{Taxonomic}=0.94\) and \(\rho_{Spearman}^{Functional}=0.90\), respectively.
+Fig.
+\ref{fig:SumUp}).
+The maximum dissimilarity with the initial state was reached for taxonomic composition between 15 to 25 years, and between 20 to 25 years for functional composition.
 
@@
 
-In control plots, Community Weighted Means (CWM) of functional traits
-remained stable in time.\\
-In disturbed plots, they mostly followed unimodal trajectories, either
-stabilizing or returning towards their initial values, to the exception
-of leaf chlorophyll content, which continued to increase 30 years after
-disturbance for 4 out of 6 highly disturbed plots. \color{red}Main
-tendencies for disturbed plots were a decrease followed by a slight
-increase of maximum height at adult stage (\emph{Hmax}), leaf toughness,
-and wood specific gravity (\emph{WSG}), but trait values remained
-significantly lower than their initial value (Fig. \ref{fig:CWM}).
-\color{black} Bark thickness and specific leaf area (\emph{SLA}) both
-increased along time. Bark thickness remained substantially high after
-30 years, and \emph{SLA} had almost recovered to its initial value.
-Whatever the functional traits, the maximum difference to initial value
-was highly correlated to the disturbance intensity. \color{red}Positive
-correlations were observed for leaf thickness, chlorophyll content, SLA
-and bark thickness (\(\rho_{Spearman}^{Leaf thickness}=0.79\),
-\(\rho_{Spearman}^{Chlorophyll content}=0.61\),
-\(\rho_{Spearman}^{SLA}=0.99\),
-\(\rho_{Spearman}^{Bark thickness}=0.68\)). Maxima were observed between
-5 and 15 years following disturbance for leaf thickness and SLA while
-chlorophyll content and bark thickness kept increasing after 30 years
-for leaf chlorophyll content and bark thickness. Negative correlation
-was observed for leaf toughness, WSG, and Hmax
-(\(\rho_{Spearman}^{Leaf toughness}=-0.47\),
+In control plots, Community Weighted Means (CWM) of functional traits remained stable in time.\\
+In disturbed plots, they mostly followed unimodal trajectories, either stabilizing or returning towards
+their initial values, to the exception of leaf chlorophyll content, which continued to increase 30 years after disturbance for 4 out of 6 highly disturbed plots.
+\color{red}Main tendencies for disturbed plots were a decrease followed by a slight increase of maximum height at adult stage (\emph{Hmax}), leaf toughness, and wood specific gravity (\emph{WSG}), but trait values remained significantly lower than their initial value (Fig.
+\ref{fig:CWM}).
+\color{black}
+Bark thickness and specific leaf area (\emph{SLA}) both increased along time.
+Bark thickness remained substantially high after 30 years, and \emph{SLA} had almost recovered to its initial value.
+Whatever the functional traits, the maximum difference to initial value was highly correlated to the disturbance intensity.
+\color{red}Positive correlations were observed for leaf thickness, chlorophyll content, SLA and bark thickness (\(\rho_{Spearman}^{Leaf thickness}=0.79\), \(\rho_{Spearman}^{Chlorophyll content}=0.61\),
+\(\rho_{Spearman}^{SLA}=0.99\), \(\rho_{Spearman}^{Bark thickness}=0.68\)).
+Maxima were observed between 5 and 15 years following disturbance for leaf thickness and SLA while chlorophyll content and bark thickness kept increasing after 30 years for leaf chlorophyll content and bark thickness.
+Negative correlation was observed for leaf toughness, WSG, and Hmax (\(\rho_{Spearman}^{Leaf toughness}=-0.47\),
 \(\rho_{Spearman}^{WSG}=-0.80\), \(\rho_{Spearman}^{Hmax}=-0.28\)).
-Minima were observed between 5 and 15 years following disturbance for
-Hmax, 10 and 20 years for WSG, and between 15 and 25 years following
-disturbance for leaf toughness.\color{black} The proportions of the
-three lightest seed mass classes increased in all disturbed plots. After
-30 years the proportion of lightest seed mass class decreased while it
-stabilized for the two other lightest seed mass classes (Supp. Mat. -
-Fig. S3).
+Minima were observed between 5 and 15 years following disturbance for Hmax, 10 and 20 years for WSG, and between 15 and 25 years following disturbance for leaf toughness.\color{black}
+The proportions of the three lightest seed mass classes increased in all disturbed plots.
+After 30 years the proportion of lightest seed mass class decreased while it stabilized for the two other lightest seed mass classes (Supp. Mat. - Fig. S3).
 
@@
 
-\subsection{Community taxonomic and functional
-diversity}\label{community-taxonomic-and-functional-diversity}
+\hypertarget{community-taxonomic-and-functional-diversity}{%
+\subsection{Community taxonomic and functional diversity}\label{community-taxonomic-and-functional-diversity}}
 
-Taxonomic richness and evenness remained stable in control plots over
-the 30 years of monitoring. In disturbed communities, after low
-disturbance intensity the taxonomic richness increased, reaching a
-maximum gain of 14 botanical genera. After intense disturbance the
-taxonomic richness followed a more complex trajectory, decreasing for
-ten years after disturbance before recovering to pre-disturbance values.
-\color{red}The maximum rate of change compared to initial richness was
-positively correlated with the disturbance intensity
-(\(\rho_{Spearman}^{Richness}=0.73\)).\color{black}
+Taxonomic richness and evenness remained stable in control plots over the 30 years of monitoring.
+In disturbed communities, after low disturbance intensity the taxonomic richness increased, reaching a maximum gain of 14 botanical genera.
+After intense disturbance the taxonomic richness followed a more complex trajectory, decreasing for ten years after disturbance before recovering to pre-disturbance values.
+\color{red}The maximum rate of change compared to initial richness was positively correlated with the disturbance intensity (\(\rho_{Spearman}^{Richness}=0.73\)).\color{black}
 
-In all disturbed plots \color{red}, to the exception of plots 5 (T2) and
-12 (T3) that showed a small initial drop, \color{black} the evenness
-first increased until a maximum reached after around 20 years. This
-maximum was positively correlated with the disturbance intensity
-(\(\rho_{Spearman}^{Evenness}=0.82\), \color{red} Fig.
-\ref{fig:SumUp})\color{black}. The evenness then stabilized except for
-two intensively-disturbed plots (number 8 and 12) for which it kept
-increasing (Fig. \ref{fig:DivTaxo}).
+In all disturbed plots \color{red}, to the exception of plots 5 (T2) and 12 (T3) that showed a small initial drop, \color{black} the evenness first increased until a maximum reached after around 20 years.
+This maximum was positively correlated with the disturbance intensity (\(\rho_{Spearman}^{Evenness}=0.82\), \color{red} Fig. \ref{fig:SumUp})\color{black}.
+The evenness then stabilized except for two intensively-disturbed plots (number 8 and 12) for which it kept increasing (Fig. \ref{fig:DivTaxo}).
 
@@
 
-Functional richness and evenness remained stable in control plots over
-the 30 years of monitoring. \color{red}In disturbed communities, both
-functional richness and evenness trajectories increased following
-disturbance, with maxima positively correlated to \%AGB loss
-\(\rho_{Spearman}^{Richness} = 0.76\) and
-\(\rho_{Spearman}^{Evenness} = 0.60\) (Fig. \ref{fig:SumUp}).
-\color{black} For low \color{red}and intermediate
-\color{black}disturbance intensity, functional richness and evenness
-displayed a low but long-lasting increase up to a maximum reached after
-20-25 years. For high disturbance intensity, they generally displayed a
-fast but short increase followed, after 10 years, by a slow decrease
-towards the initial values.
+Functional richness and evenness remained stable in control plots over the 30 years of monitoring.
+\color{red}In disturbed communities, both functional richness and evenness trajectories increased following disturbance, with maxima positively correlated to \%AGB loss \(\rho_{Spearman}^{Richness} = 0.76\) and \(\rho_{Spearman}^{Evenness} = 0.60\) (Fig.
+\ref{fig:SumUp}).
+\color{black}
+For low \color{red}and intermediate \color{black}disturbance intensity, functional richness and evenness displayed a low but long-lasting increase up to a maximum reached after 20-25 years.
+For high disturbance intensity, they generally displayed a fast but short increase followed, after 10 years, by a slow decrease towards the initial values.
 
-The second-degree polynomial regressions between (i) the percentage AGB
-loss, and (ii) the taxonomic and functional diversity showed various
-shapes depending on the diversity indices and on the time since
-disturbance (Fig. \ref{fig:IDHplot}). \color{red}Regarding taxonomic
-diversity, the taxonomic richness showed a humped-shaped trajectory with
-the disturbance intensity, and peaked at 20\% of initial AGB loss. The
-relationship between disturbance intensity and taxonomic evenness,
-however, was humped-shaped only 20 years after disturbance and linear
-for 10 and 30 years after disturbance times.\color{black} Regarding
-functional diversity, \color{red} the relationship was linear for both
-taxonomic richness and evenness, except for the functional diversity
-after 30 years. \color{black}
+The second-degree polynomial regressions between (i) the percentage AGB loss, and (ii) the taxonomic and functional diversity showed various shapes depending on the diversity indices and on the time since disturbance (Fig.
+\ref{fig:IDHplot}).
+\color{red}Regarding taxonomic diversity, the taxonomic richness showed a humped-shaped trajectory with the disturbance intensity, and peaked at 20\% of initial AGB loss.
+The relationship between disturbance intensity and taxonomic evenness, however, was humped-shaped only 20 years after disturbance and linear for 10 and 30 years after disturbance times.\color{black}
+Regarding functional diversity, \color{red} the relationship was linear for both taxonomic richness and evenness, except for the functional diversity after 30 years. \color{black}
 
@@
 
-\subsection{Functional redundancy}\label{functional-redundancy}
+\hypertarget{functional-redundancy}{%
+\subsection{Functional redundancy}\label{functional-redundancy}}
 
-Control plots displayed stable functional redundancy over time. All
-disturbed plots had lower functional redundancy than control plots and
-followed similar humped-shaped trajectories (Fig. \ref{fig:RedFunRest}).
-The maximum redundancy loss was positively correlated with the
-disturbance intensity (\(\rho_{Spearman}=0.31\), (Fig. \ref{fig:SumUp}))
-and the recovery trajectory had not attained initial values for any
-disturbed communities after 30 years.
+Control plots displayed stable functional redundancy over time.
+All disturbed plots had lower functional redundancy than control plots and followed similar humped-shaped trajectories (Fig. \ref{fig:RedFunRest}).
+The maximum redundancy loss was positively correlated with the disturbance intensity (\(\rho_{Spearman}=0.31\), (Fig.
+\ref{fig:SumUp})) and the recovery trajectory had not attained initial values for any disturbed communities after 30 years.
 
@@
 
-\section{Discussion}\label{discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-\color{red}Our analysis revealed that community post-disturbance
-trajectories in taxonomic and functional diversity and composition were
-decoupled. The taxonomic composition specificities between communities
-before disturbance were maintained following disturbance, while their
-functional composition converged in the functional space. \color{black}
-In terms of diversity, only the taxonomic trajectories validated the IDH
-that explained humped-shaped post-disturbance trajectories whith an
-amplitude depending on the disturbance intensity. The decoupling between
-taxonomic and functional response \color{red} could be
-\color{black}explained by variations in community functional redundancy.
-\color{red}The loss of species or the changes in their abundance would
-impact the taxonomic but not the functional diversity and composition,
-as redundant species with the same functional characteristics remain in
-the community. The trajectory of the functional redundancy hence
-appeared determinant for community recovery.\color{black}
+\color{red}Our analysis revealed that community post-disturbance trajectories in taxonomic and functional diversity and composition were decoupled.
+The taxonomic composition specificities between communities before disturbance were maintained following disturbance, while their functional composition converged in the functional space.
+\color{black}
+In terms of diversity, only the taxonomic trajectories validated the IDH that explained humped-shaped post-disturbance trajectories whith an amplitude depending on the disturbance intensity.
+The decoupling between taxonomic and functional response \color{red} could be \color{black}explained by variations in community functional redundancy.
+\color{red}The loss of species or the changes in their abundance would impact the taxonomic but not the functional diversity and composition, as redundant species with the same functional characteristics remain in the community.
+The trajectory of the functional redundancy hence appeared determinant for community recovery.\color{black}
 
-\subsection{Decoupled taxonomic and functional
-trajectories}\label{decoupled-taxonomic-and-functional-trajectories}
+\hypertarget{decoupled-taxonomic-and-functional-trajectories}{%
+\subsection{Decoupled taxonomic and functional trajectories}\label{decoupled-taxonomic-and-functional-trajectories}}
 
-\color{red}From pre-disturbance to 30 years after disturbance,
-communities showed different location along NMDS axis 2 that hardly
-changed along time. Specificities in community taxonomic composition,
-materialized by the distinct location on the NMDS axis 2, existed before
-disturbance. The disturbance led a displacement on the NMDS axis 1
-only.\color{black} Taxonomic \color{red} post-disturbance \color{black}
-changes in composition were similar among plots and may correspond to
-the recruitment of a group of pioneers, like \emph{Cecropia spp.} or
-\emph{Miconia spp.}, common to all plots, whatever their initial
-taxonomic \color{red}specificities \color{black} and the intensity of
-the disturbance \citep{Denslow2000, Bongers2009}. Taxonomic trajectories
-initiated a recovery towards the initial composition. This recovery,
-although far from being achieved after 30 years, suggested the
-resilience of community taxonomic composition. \color{red}Although the
-composition dramatically changed following disturbance, community
-specificities were maintained \citep{Folke2006}. This maintenance
-suggested that the recruitment came from (i) a common set of pioneer and
-light-demanding species in all plots that shape the similar trajectories
-on NMDS axis 1 and (ii) a very local set of late-successional species
-that signs (and maintains in time) the initial position of each plot on
-NMDS axis 2. This can be related to species dispersal limitation that is
-common among tropical species \citep{Svenning2005}.\color{black}
+\color{red}From pre-disturbance to 30 years after disturbance, communities showed different location along NMDS axis 2 that hardly changed along time.
+Specificities in community taxonomic composition, materialized by the distinct location on the NMDS axis 2, existed before disturbance.
+The disturbance led a displacement on the NMDS axis 1 only.\color{black}
+Taxonomic \color{red} post-disturbance \color{black} changes in composition were similar among plots and may correspond to the recruitment of a group of pioneers, like \emph{Cecropia spp.} or \emph{Miconia spp.}, common to all plots, whatever their initial taxonomic \color{red}specificities \color{black} and the intensity of the disturbance \citep{Denslow2000, Bongers2009}.
+Taxonomic trajectories initiated a recovery towards the initial composition.
+This recovery, although far from being achieved after 30 years, suggested the resilience of community taxonomic composition.
+\color{red}Although the composition dramatically changed following disturbance, community specificities were maintained \citep{Folke2006}.
+This maintenance suggested that the recruitment came from (i) a common set of pioneer and light-demanding species in all plots that shape the similar trajectories on NMDS axis 1 and (ii) a very local set of late-successional species that signs (and maintains in time) the initial position of each plot on NMDS axis 2.
+This can be related to species dispersal limitation that is common among tropical species \citep{Svenning2005}.\color{black}
 
-Community functional composition trajectories, in contrast, were similar
-in the functional space. \color{red}Such functional convergence
-contrasting with community taxonomic divergence was already observed in
-plant communities \citep{Fukami2005}.\color{black} The amplitude of the
-compositional changes depended on the disturbance intensity. Functional
-trajectories were probably driven by the recruitment of species
-infrequent or absent before disturbance, and belonging to a pool of
-species common to all plots. This common pool was composed of pioneers
-with ``resource-acquisitive''" strategies displaying low leaf toughness,
-wood specific gravity, maximum height, and high specific leaf area. The
-recruitment of these pioneers drove a displacement to the right along
-the first NMDS axis \citep{Westoby1998, Wright2004, Reich2014}.
-\color{red}This recruitment went along with an increase of the mean SLA,
-and a decrease in the mean WSG in disturbed communities (\ref{fig:CWM}).
-These fast and significant changes in mean functional traits closely
-linked to species light acquisition \citep{Wright2004, Chave2009b}
-suggested major changes in the light environment following disturbance
-that shape species assemblage in tropical forests
-\citep{Pena2008, Carreno-Rocabado2012}. Thereafter, the first recruited
-pioneers were progressively excluded by long-lived, more competitive,
-and shade-tolerant species. The recruitment of these late-successional
-species marked the recovery of the initial functional composition with
-more ``resource-conservative''" strategies, corresponding to a fast
-decrease in community mean SLA and a stabilisation of community WSG,
-suggesting the progressive closing of forest canopy. \color{black} This
-recovery translated in the functional plane by a displacement left along
-the first axis and upward along the second axis (Fig.
-\ref{fig:NMDSplans}).
+Community functional composition trajectories, in contrast, were similar in the functional space.
+\color{red}Such functional convergence contrasting with community taxonomic divergence was already observed in plant communities \citep{Fukami2005}.\color{black}
+The amplitude of the compositional changes depended on the disturbance intensity.
+Functional trajectories were probably driven by the recruitment of species infrequent or absent before disturbance, and belonging to a pool of species common to all plots.
+This common pool was composed of pioneers with ``resource-acquisitive''" strategies displaying low leaf toughness, wood specific gravity, maximum height, and high specific leaf area.
+The recruitment of these pioneers drove a displacement to the right along the first NMDS axis \citep{Westoby1998, Wright2004, Reich2014}.
+\color{red}This recruitment went along with an increase of the mean SLA, and a decrease in the mean WSG in disturbed communities (\ref{fig:CWM}).
+These fast and significant changes in mean functional traits closely linked to species light acquisition \citep{Wright2004, Chave2009b} suggested major changes in the light environment following disturbance that shape species assemblage in tropical forests \citep{Pena2008, Carreno-Rocabado2012}.
+Thereafter, the first recruited pioneers were progressively excluded by long-lived, more competitive, and shade-tolerant species.
+The recruitment of these late-successional species marked the recovery of the initial functional composition with more ``resource-conservative''" strategies, corresponding to a fast decrease in community mean SLA and a stabilisation of community WSG, suggesting the progressive closing of forest canopy.
+\color{black}
+This recovery translated in the functional plane by a displacement left along the first axis and upward along the second axis (Fig. \ref{fig:NMDSplans}).
 
-The decoupling between taxonomic and functional trajectories suggested
-that simultaneous operation of trait-based assembly rules and
-species-level priority effects shaped tree community assembly in Paracou
-forest. Tree community assembly would then be both deterministic in the
-functional space, and historically contingent in the taxonomic space.
+The decoupling between taxonomic and functional trajectories suggested that simultaneous operation of trait-based assembly rules and species-level priority effects shaped tree community assembly in Paracou forest.
+Tree community assembly would then be both deterministic in the functional space, and historically contingent in the taxonomic space.
 
-\color{red} We must temper these results with the low number of plots
-(3) for each treatment. However, the size of each plot is 6.25ha that
-has been regularly censused over 25 years so that the trajectories have
-been drawn using 75ha of permanent plot information, a spatial hold very
-rarely achieved in such studies. Given the relative homogeneity of
-environmental characteristics of terra firme forests over the Guiana
-Shield and even the amazonian basin, it seems reasonable that similar
-trajectories would apply to neighboring areas
-\citep{Guitet2015, Guitet2018}. \color{black}
+\color{red}
 
-\subsection{The scope of the intermediate disturbance
-hypothesis}\label{the-scope-of-the-intermediate-disturbance-hypothesis}
+We must temper these results with the low number of plots (3) for each treatment.
+However, the size of each plot is 6.25ha that has been regularly censused over 25 years so that the trajectories have been drawn using 75ha of permanent plot information, a spatial hold very rarely achieved in such studies.
+Given the relative homogeneity of environmental characteristics of terra firme forests over the Guiana Shield and even the amazonian basin, it seems reasonable that similar trajectories would apply to neighboring areas \citep{Guitet2015, Guitet2018}.
+\color{black}
 
-Trajectories of taxonomic richness and evenness differed markedly below
-and above an intensity threshold (Fig. \ref{fig:DivTaxo}). For low and
-intermediate disturbance, both taxonomic richness and evenness increased
-according to the disturbance intensity
-\citep{Martin2015, Chaudhary2016}. This suggested that the recruitment
-of pioneers, previously infrequent or absent, increased the taxonomic
-richness, and that trees surviving after disturbance remained numerous
-enough to maintain the richness of the pre-disturbance community
-\citep{Bongers2009}. The pioneers thus recruited became abundant, which
-balanced the usual hyper-dominance of tropical forests and increased the
-taxonomic evenness \citep{Baraloto2012a}. Above the intensity threshold,
-like for treatment 3 (high intensity), the taxonomic richness did not
-exceed the initial value in the first years following disturbance. No
-increase of the taxonomic richness suggested that the richness of
-surviving trees was lower than this of the pre-disturbance community,
-and that this difference was not offset by the recruitment of pioneers.
-In the Guiana shield, the pool of \color{red}hard \color{black} pioneers
-specifically recruited after disturbance is restricted to a few common
-genera (e.g. \emph{Cecropia spp.}, \emph{Miconia spp.}, \emph{Tapirira
-spp.}) \citep{Guitet2018}. \color{red}The disturbance intensity showed a
-humped-shaped relationship with the post-disturbance increase in
-taxonomic richness at all the different times after disturbance, and
-with the increase in taxonomic evenness 10 years after disturbance (Fig.
-\ref{fig:IDHplot}).\color{black} Taxonomic richness, and to some extend
-taxonomic evenness, were maximized at an intermediate intensity, around
-20-25\% of AGB lost.
+\hypertarget{the-scope-of-the-intermediate-disturbance-hypothesis}{%
+\subsection{The scope of the intermediate disturbance hypothesis}\label{the-scope-of-the-intermediate-disturbance-hypothesis}}
 
-\subsection{The functional redundancy explaining the
-taxonomic-functional
-decoupling}\label{the-functional-redundancy-explaining-the-taxonomic-functional-decoupling}
+Trajectories of taxonomic richness and evenness differed markedly below and above an intensity threshold (Fig. \ref{fig:DivTaxo}).
+For low and intermediate disturbance, both taxonomic richness and evenness increased according to the disturbance intensity \citep{Martin2015, Chaudhary2016}.
+This suggested that the recruitment of pioneers, previously infrequent or absent, increased the taxonomic richness, and that trees surviving after disturbance remained numerous enough to maintain the richness of the pre-disturbance community \citep{Bongers2009}.
+The pioneers thus recruited became abundant, which balanced the usual hyper-dominance of tropical forests and increased the taxonomic evenness \citep{Baraloto2012a}.
+Above the intensity threshold, like for treatment 3 (high intensity), the taxonomic richness did not exceed the initial value in the first years following disturbance.
+No increase of the taxonomic richness suggested that the richness of surviving trees was lower than this of the pre-disturbance community, and that this difference was not offset by the recruitment of pioneers.
+In the Guiana shield, the pool of \color{red}hard \color{black} pioneers specifically recruited after disturbance is restricted to a few common genera (e.g.~\emph{Cecropia spp.}, \emph{Miconia spp.}, \emph{Tapirira spp.}) \citep{Guitet2018}.
+\color{red}The disturbance intensity showed a humped-shaped relationship with the post-disturbance increase in taxonomic richness at all the different times after disturbance, and with the increase in taxonomic evenness 10 years after disturbance (Fig.
+\ref{fig:IDHplot}).\color{black}
+Taxonomic richness, and to some extend taxonomic evenness, were maximized at an intermediate intensity, around 20-25\% of AGB lost.
+
+\hypertarget{the-functional-redundancy-explaining-the-taxonomic-functional-decoupling}{%
+\subsection{The functional redundancy explaining the taxonomic-functional decoupling}\label{the-functional-redundancy-explaining-the-taxonomic-functional-decoupling}}
 
 \color{red}Regarding community functional trajectories (Fig.
-\ref{fig:IDHplot}), there was no intensity threshold above which the
-diversity-disturbance relationship changed. The functional diversity
-kept increasing along with the disturbance intensity, hence the IDH
-would not apply. Surprisingly, although some species were lost after
-disturbance, the functional diversity did not decrease in the first
-place. On the contrary, functional diversity increased, probably
-following the rapid recruitment of pioneers that were functionally
-highly different from the pre-disturbance community
-\citep{Denslow1980, Molino2001}.
+\ref{fig:IDHplot}), there was no intensity threshold above which the diversity-disturbance relationship changed.
+The functional diversity kept increasing along with the disturbance intensity, hence the IDH would not apply.
+Surprisingly, although some species were lost after disturbance, the functional diversity did not decrease in the first place.
+On the contrary, functional diversity increased, probably following the rapid recruitment of pioneers that were functionally highly different from the pre-disturbance community \citep{Denslow1980, Molino2001}.
 
-The loss of species following disturbance, however, decreased the
-functional redundancy, all the more so that disturbance was intense. All
-in all, because functional diversity was not lowered by disturbance
-while functional redundancy was, this means that the species that were
-lost in disturbance are, on the whole, functionally equivalent to the
-species that survive the disturbance. This makes functional redundancy a
-key to understand community dynamics at a functional level. In other
-words, the high redundancy of tropical forest mean that several species
-occupy the same functional space, so changes in taxonomic diversity or
-composition do not necessarily result in changes in community functional
-characteristics because species are commutable. The loss of a redundant
-species doesn't change anything for the community functional
-characteristics, so functional trajectories do not necessarily track
-taxonomic ones. This commutability explains the taxonomic-functional
-decoupling and the fact that the taxonomic diversity-disturbance
-relationship supports the IDH while it functional diversity do not.\\
-Following disturbance, the redundancy was progressively restored through
-the replacement of ``resource-acquisitive'' species by more
-late-successional ``resource-conservative'' species functionally closer
-to the pre-disturbance community. The time for the functional redundancy
-to recover would hence be important to assess community recovery.
+The loss of species following disturbance, however, decreased the functional redundancy, all the more so that disturbance was intense.
+All in all, because functional diversity was not lowered by disturbance while functional redundancy was, this means that the species that were lost in disturbance are, on the whole, functionally equivalent to the species that survive the disturbance.
+This makes functional redundancy a key to understand community dynamics at a functional level.
+In other words, the high redundancy of tropical forest mean that several species occupy the same functional space, so changes in taxonomic diversity or composition do not necessarily result in changes in community functional characteristics because species are commutable.
+The loss of a redundant species doesn't change anything for the community functional characteristics, so functional trajectories do not necessarily track taxonomic ones.
+This commutability explains the taxonomic-functional decoupling and the fact that the taxonomic diversity-disturbance relationship supports the IDH while it functional diversity do not.\\
+Following disturbance, the redundancy was progressively restored through the replacement of ``resource-acquisitive'' species by more late-successional ``resource-conservative'' species functionally closer to the pre-disturbance community.
+The time for the functional redundancy to recover would hence be important to assess community recovery.
 \color{black}
 
-\section{Conclusion}\label{conclusion}
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 
-Post-disturbance trajectories of tree community composition and
-diversity would be driven by the recruitment of a determined pool of
-pioneers, identical among local communities, and independent of the
-disturbance intensity. The taxonomic composition trajectories maintained
-the initial differences among communities, while the functional
-trajectories were similar, and converged in the functional space towards
-the recovery of the initial composition. The diversity trajectories were
-contrasted as well. While the functional trajectories \color{red}
-displayed the same tendency with a linear increase following all
-disturbance intensity, \color{black} taxonomic trajectories were
-markedly different after a threshold of 20-25\% AGB lost that maximized
-the taxonomic richness. The Intermediate Disturbance Hypothesis would
-apply well to taxonomic diversity, but not to functional diversity. The
-decoupling between taxonomic and functional trajectories was mediated by
-the variations in functional redundancy, as the loss of a species does
-not necessarily entails the loss of its functional characteristics.
-Community resilience, in terms of recovery of the pre-disturbance state,
-was tangible but required several decades, and relied upon the random
-lottery recruitment of rare species. Given the long-term impacts of
-disturbance observed, we suggest that 30 years is not enough time for
-tropical communities to recover, even after relatively low intensity
-disturbance. Much of community response to disturbance rely on the
-processes of species recruitment. A more refined understanding of the
-post-disturbance trajectories would be gained by a closer analysis of
-the recruitment process.
+Post-disturbance trajectories of tree community composition and diversity would be driven by the recruitment of a determined pool of pioneers, identical among local communities, and independent of the disturbance intensity.
+The taxonomic composition trajectories maintained the initial differences among communities, while the functional trajectories were similar, and converged in the functional space towards the recovery of the initial composition.
+The diversity trajectories were contrasted as well.
+While the functional trajectories \color{red} displayed the same tendency with a linear increase following all disturbance intensity, \color{black} taxonomic trajectories were markedly different after a threshold of 20-25\% AGB lost that maximized the taxonomic richness.
+The Intermediate Disturbance Hypothesis would apply well to taxonomic diversity, but not to functional diversity.
+The decoupling between taxonomic and functional trajectories was mediated by the variations in functional redundancy, as the loss of a species does not necessarily entails the loss of its functional characteristics.
+Community resilience, in terms of recovery of the pre-disturbance state, was tangible but required several decades, and relied upon the random lottery recruitment of rare species.
+Given the long-term impacts of disturbance observed, we suggest that 30 years is not enough time for tropical communities to recover, even after relatively low intensity disturbance.
+Much of community response to disturbance rely on the processes of species recruitment.
+A more refined understanding of the post-disturbance trajectories would be gained by a closer analysis of the recruitment process.
 
-\section{Acknowledgement}\label{acknowledgement}
+\hypertarget{acknowledgement}{%
+\section{Acknowledgement}\label{acknowledgement}}
 
-We are in debt with all Paracou station technicians and colleagues who
-helped setting up the plots and collecting data over years. Without
-their precious work, this study would have not been possible and they
-may be warmly thanked here. Our work benefited from an ``Investissement
-d'Avenir'' grant managed by the Agence Nationale de la Recherche (LABEX
-CEBA, ref ANR-10-LBX-25). We thank Niklas Tysklind for the usefull help
-with english proofing.
+We are in debt with all Paracou station technicians and colleagues who helped setting up the plots and collecting data over years.
+Without their precious work, this study would have not been possible and they may be warmly thanked here.
+Our work benefited from an ``Investissement d'Avenir'' grant managed by the Agence
+Nationale de la Recherche (LABEX CEBA, ref ANR-10-LBX-25).
+We thank Niklas Tysklind for the usefull help with english proofing.
 
-\section{Author's contributions}\label{authors-contributions}
+\hypertarget{authors-contributions}{%
+\section{Author's contributions}\label{authors-contributions}}
 
-AM, EM \& BH designed the study, developed the analysis framework, and
-interpreted the results. AM wrote the manuscript with contributions by
-EM \& BH. All authors gave final approval for publication.
+AM, EM \& BH designed the study, developed the analysis framework, and interpreted the results.
+AM wrote the manuscript with contributions by EM \& BH.
+All authors gave final approval for publication.
 
-\section{Data availability}\label{data-availability}
+\hypertarget{data-availability}{%
+\section{Data availability}\label{data-availability}}
 
-This article is based upon the dataset of the Paracou station, which is
-part of the Guyafor permanent plot network in French Guiana
-(Cirad-CNRS-ONF). The dataset is available upon request to the
-scientific director (https://paracou.cirad.fr).
+This article is based upon the dataset of the Paracou station, which is part of the Guyafor permanent plot network in French Guiana (Cirad-CNRS-ONF).
+The dataset is available upon request to the scientific director (https://paracou.cirad.fr).
 
@@
 \makeatletter
-% The filename has .bib extension that must be eliminated
+% The filename has .bib extension the must be eliminated
 \filename@parse{references.bib}
